### 2023-01-05 README.rmd
updated readme

---
output: github_document
---

<!-- README.md is generated from README.Rmd. -->

# contratrain-muscle-paper

## Authors

Håvard Hamarsland
<a href="https://orcid.org/0000-0002-7343-2774">
<img alt="ORCID logo" src="https://info.orcid.org/wp-content/uploads/2019/11/orcid_16x16.png" width="16" height="16" /></a>,
Kristian Lian,
Chidimma Phoebe Echebiri,
Mads-Henrik Hafsmo,
Tomas Varpestuen,
Stein-Olaf Olsen,
Chris Sylstad,
Max Ullrich,
Stian Ellefsen
<a href="https://orcid.org/0000-0003-0467-2018">
<img alt="ORCID logo" src="https://info.orcid.org/wp-content/uploads/2019/11/orcid_16x16.png" width="16" height="16" /></a>,
Daniel Hammarström
<a href="https://orcid.org/0000-0001-8360-2100">
<img alt="ORCID logo" src="https://info.orcid.org/wp-content/uploads/2019/11/orcid_16x16.png" width="16" height="16" /></a>

## Repository organization

`/resources`

- bibliography.bib
- csl-files

`/data`

- contains raw data not included in the data package

`/data-gen/`

- contains generated data stored for further analyses

`contratrain-manuscript.qmd`

- The manuscript source file

`/figures`

- Figures source files and output

### 2023-01-10 README.rmd
updated README

@@
 
-Håvard Hamarsland
-<a href="https://orcid.org/0000-0002-7343-2774">
-<img alt="ORCID logo" src="https://info.orcid.org/wp-content/uploads/2019/11/orcid_16x16.png" width="16" height="16" /></a>,
-Kristian Lian,
-Chidimma Phoebe Echebiri,
-Mads-Henrik Hafsmo,
-Tomas Varpestuen,
-Stein-Olaf Olsen,
-Chris Sylstad,
-Max Ullrich,
-Stian Ellefsen
-<a href="https://orcid.org/0000-0003-0467-2018">
+Kristian Lian
+<a href="https://orcid.org/0000-0002-5169-5960">
 <img alt="ORCID logo" src="https://info.orcid.org/wp-content/uploads/2019/11/orcid_16x16.png" width="16" height="16" /></a>,
@@
 <img alt="ORCID logo" src="https://info.orcid.org/wp-content/uploads/2019/11/orcid_16x16.png" width="16" height="16" /></a> 
+Stian Ellefsen
+<a href="https://orcid.org/0000-0003-0467-2018">
+<img alt="ORCID logo" src="https://info.orcid.org/wp-content/uploads/2019/11/orcid_16x16.png" width="16" height="16" /></a>,
+Håvard Hamarsland
+<a href="https://orcid.org/0000-0002-7343-2774">
+<img alt="ORCID logo" src="https://info.orcid.org/wp-content/uploads/2019/11/orcid_16x16.png" width="16" height="16" /></a>,
+Chidimma Phoebe Echebiri,
+Sara Christine Moen,
 
@@
 
-`contratrain-manuscript.qmd`
+`ribose-manuscript.qmd`
 

### 2023-02-01 draft-styles.RMD
updated change analyses and figures

---
title: "Untitled"
author: "Kristian Lian"
date: "31 1 2023"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r packages}
library(tidyverse);library(readxl);library(lme4);library(emmeans);library(dplyr);library(cowplot);library(knitr);library(broom);library(rmarkdown);library(tinytex);library(tidyr);library(arsenal);library(magrittr);library(dabestr)
```

# Glucose ingestion before and after resistance training does not augment ribosome biogenesis in young moderately trained adults

## Disadvantages due to Covid-19

Due to restrictions regarding social distancing and lock downs caused by the Covid-19 pandemic,
we were not able to recruit and include as many participants as we planned to the study. The
pandemic was also a contributing factor to the endocrine analysis not being completed, as they
were supposed to be performed at Sykehuset Innlandet Hospital Trust.

## Methods

This data set is a part of a larger project involving several investigators and other outcomes not
covered here. All participants were informed about the potential discomforts and risks associated
with the study and gave their informed consent prior to study enrollment. The project was
approved by the regional ethical committee (REK, ID nr. 153628), pre registered at
clinicaltrials.gov (Identifier: NCT04545190) and conducted according to the Helsinki declaration

### Participants

Sixteen male and female participants (20-33yrs, Tab 1) were recruited to the study through
facebook advertisement and word of mouth and taken through the selection process (Fig 1). The
eligibility criteria were non-smokers and moderately trained (i.e. 2-8 resistance training sessions
per 14 days for the last six months). Exclusion criteria were previous injury leading to impaired
strength, inability to perform resistance training and symptoms, and a medical record of
metabolic disorders including hyperglycemia, i.e. fasting venous plasma glucose ≥6.1 mmol/L
and/or 2-hour glucose tolerance ≥7.8 mmol/L, and/or HbA1c >42 mmol/mol. Our goal was to
recruit 20 participants to the study, however due to the advents of Covid-19, we were not able to
do so. Sixteen participants commenced the intervention, during which three dropped out. One
participant had a sick child, and was unable to resume the intervention, two participants
experienced muscular discomfort connected to heavy resistance training (Fig 1). Baseline
characteristics (Tab 1) were measured by means of DXA (Prodigy Advance PA+302047, Lunar,
San Francisco, CA, USA) at Day -1, the last day preceding the RT intervention.

```{r DXA}
# Script: DXA

# Data handling
table.dxa <- read_excel("./data/dxa/ribose_dxa.xlsx") %>%
        select(sex, subject, age, height, weight, leanmasskg, fatmasskg, ffmg) %>%
        #make grams into kg
        mutate(ffkg = ffmg/1000) %>%
        select(sex, age, height, weight, leanmasskg, fatmasskg, ffkg) %>%
  mutate(bmi = (weight) / (height^2) * 10000)

dxa.table <- table.dxa %>%
        pivot_longer(names_to = "variable",
                     values_to = "values", col= age:bmi) %>%
        select(sex, variable, values) %>%
        group_by(sex, variable) %>%
        summarise(m = mean(values, na.rm = TRUE),
                  s= sd(values, na.rm = TRUE)) %>%
        mutate(stat = paste0(round(m, 1), " (", round(s, 1), ")")) %>%
        select(sex, variable, stat) %>%
        pivot_wider(names_from = sex,
                    values_from = stat) %>%
                mutate(variable = if_else(variable %in% c("age"),
                                          "Age (yrs)",
                                          if_else(variable %in% c("fatmasskg"),
                                                  "Fat mass (kg)",
                                                  if_else(variable %in% c("ffkg"),
                                                          "Fat free mass (kg)",
                                                          if_else(variable %in% c("height"),
                                                                  "Height (cm)",
                                                                  if_else(variable %in% c("leanmasskg"),
                                                                          "Lean body mass (kg)",
                                                                          if_else(variable %in% c("weight"),
                                                                                  "Body weight (kg)",
                                                                                  if_else(variable %in% c("bmi"),
                                                                                          "BMI", "")
                                                                                  )))))))

dxa.table %>%
        kable(col.names = c("Variable", "Female (n=7)", "Male (n=9)"))
```

#### Paper

##### Paper

###### Paper

### 2023-02-01 ribose-paper.RMD
updated change analyses and figures

---
title: "Glucose ingestion before and after resistance training does not augment ribosome biogenesis in healthy moderately trained young adults"
author: "Kristian Lian, Daniel Hammarström, Stian Ellefsen, Håvard Hamarsland"
output:
  word_document:
    reference_docx: reference-template-02.docx
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r packages}
library(tidyverse);library(readxl);library(lme4);library(emmeans);library(dplyr);library(cowplot);library(knitr);library(broom);library(rmarkdown);library(tinytex);library(tidyr);library(arsenal);library(magrittr);library(dabestr)
```

## Introduction

Regular resistance exercise training (RT) increases muscle mass and muscle strength, due to the adaptations stimulated by a mechanical strain. Unfortunately, responses to RT are not uniform across the population. Studies have shown RT-induced muscle growth to vary widely between individuals, with as much as 10-15% showing considerably impaired growth (ALVAREZ18, MANN14, THALACKER13). This may be explained by genetics, epigenetics, and/or an unfavourable internal physiological milieu (KILDER). Recently, tweaking training modalities and -protocols have been shown to induce different responses in different populations (HAMMARSTROM20), however, little is known regarding the optimisation of RT for individual phenotypes. Generally, meta-analyses favour a moderate volume of RT over low volume RT regarding muscle growth and -strength (KRIEGER09, RHEA03, SCHOENFELD17), substantiated by recent RCTs (KILDER). Although Hammarström and colleagues (20) found that a moderate volume led to more prominent gains in muscle mass and -strength, approximately 50% of the participants did not experience true benefits of moderate volume compared to low, with a blunted muscle growth (and strength?). Due to the variations observed in muscle growth responses to standardised RT protocols (HAMMARSTROM20; SCHOENFELD17; STEC16?), it is apparent that increasing RT volume alone does not convert low-responders to high-responders. Therefore, other means than RT per se may be necessary to circumvent this discrepancy.

Indeed, ingesting nutrients such as protein and creatine supplements effectively optimise RT (CERMAK12, LANHERS15, LANHERS17). However, it remains unknown if other nutritional adjuvants such as glucose can increase the efficacy of RT. This is surprising, as glucose is the preferred energy substrate of the contracting skeletal muscle during strenuous exercise and a major supplier of energy to cells via ATP synthesis (TANAKA18). Energy availability is important to regulate central processes of muscle growth, specifically ribosome biogenesis, a key factor for increased muscle protein synthesis (MPS) (TANAKA18, FIGUEIREDO19). Interestingly, variations in RT-induced responses have been linked to ribosome biogenesis, where participants responding poorly to RT also showed a blunted ribosome biogenesis (HAMMARSTROM20, STEC16). A blunted ribosome biogenesis negates responses to RT by failure to produce novel ribosomes, thus the inability to increase the skeletal muscle cell’s translational capacity (VONWALDEN19, FIGUEIREDO19). Therefore, investigating the effects of combined RT and glucose ingestion may provide valuable insight into the regulation of ribosome biogenesis.

Ribosome biogenesis is suggested to be regulated by the mammalian target of rapamycin complex 1 (mTORC1) network (KUSNADI15, TANAKA18). Stimulated by growth factors, hormones, mechanical loading and nutrients, mTORC1 governs energy distribution by turning on/off energy-consuming and energy-generating pathways, depending on cellular energy levels (HOPPE09). Ribosome biogenesis consumes up to 80% of the cell’s energy and materials (SCHMIDT04) and is presumably switched on/off based on cellular energy levels (VONWALDEN19). Upon stimulation by mTORC1, ribosome biogenesis is initiated with the assembly of the preinitiation complex (PIC) (VONWALDEN19, FIGUEIREDO19). The PIC consists of the RNA polymerase I complex, the SL-1 complex and the TIF-1A/RRN3 protein (FIGUEIREDO19, VONWALDEN19). Further, the upstream binding factor (UBF) has a vital role in the initiation, binding the PIC to the rDNA promoter, while c-Myc serves as a transcription factor (FIGUEIREDO19, VONWALDEN19). Upon initiation, Pol I transcribed the 47S pre-rRNA, ultimately processed into the 18S-, 5.8S and 28S rRNA (MULLINEUX12, VONWALDEN19). The fourth rRNA, 5S, is transcribed by Pol III, while mRNAs for the ribosomal proteins (r-proteins) are transcribed by Pol II (MULLINEUX12, VONWALDEN19). Together, the four mature rRNAs (18S, 5.8S, 28S and 5S) along with 80 ribosomal proteins form the two subunits of a ribosome (FIGUEIREDO19, VONWALDEN19).

Whether glucose has mechanisms to affect ribosome biogenesis i.e., acting as a signalling molecule per se, or indirectly via insulin is currently unknown. Factors involved in the (PIC) of rDNA transcription are thought to receive cues from pathways (mTORC1, MAPK) regulated by hormonal and nutritional signals (FIGUEIREDO19). Furthermore, high glucose treatment promoted ribosome biogenesis while glucose depletion repressed ribosome biogenesis in mice (ZHAI12 – OR WAS IT CELL CULTURES?). Thus, it seems plausible that glucose ingestion before and after RT may benefit ribosome biogenesis. Lastly, there are equivocal observations on insulin’s impact on ribosome biogenesis and MPS. Previous meta-analyses suggest insulin acts in concert with amino acid intake to enhance MPS, and that insulin may reduce MPB when amino acids are scarce (ABDULLA16).

To investigate this further, we recruited 16 young, moderately trained, male and female individuals for heavy unilateral resistance training for two weeks, exercising one leg with glucose and protein, and the other with a placebo and protein ingestion before and after each session. Our main purpose was to investigate the effects of glucose ingestion compared to placebo ingestion before and after five RT sessions on ribosome biogenesis. Changes in ribosome biogenesis were measured by the accumulation of total RNA and expression of the ribosome’s four mature rRNAs and the 47S precursor RNA, along with levels of UBF, c-Myc, and rps6 protein. As a secondary outcome, we were also interested in the effects of glucose compared to placebo ingestion on muscular function and -recovery, measured via unilateral isometric and isokinetic knee extension torque, before, during and after the intervention. We hypothesised that RT would lead to a prominent accumulation of total RNA, rRNA and protein (UBF, c-Myc, rps6) and that RT with glucose and protein would lead to higher accumulation and expression compared to RT with placebo and protein. We also hypothesised that glucose supplementation would enhance muscular function and -recovery compared to placebo.

## Methods

### Materials and methods
All participants were informed about the potential discomforts and risks associated with the study and gave their informed consent before study enrollment. The study was approved by the regional ethical committee (REK, ID nr. 153628), pre-registered at clinicaltrials.gov (Identifier: NCT04545190), and conducted according to the Helsinki declaration.

### Participants
Sixteen healthy male and female participants (20-33yrs, Tab 1) were recruited to the study through social media advertisement and word of mouth, and through completing the selection process (Fig 1). The eligibility criteria were non-smokers and moderately trained (i.e. 2-8 RT sessions per 14 days for the last six months). Exclusion criteria were previous injury leading to impaired strength, inability to perform resistance training, symptoms and a medical record of metabolic disorders including hyperglycemia, i.e. fasting venous plasma glucose ≥6.1 mmol/L and/or 2-hour glucose tolerance ≥7.8 mmol/L, and/or HbA1c >42 mmol/mol. Of the sixteen participants that commenced the intervention, three participants dropped out. One due to sickness and inability to resume, two participants experienced muscular discomfort connected to heavy resistance training (Fig 1). Baseline characteristics (Tab 1) were measured using DXA (Prodigy Advance PA+302047, Lunar, San Francisco, CA, USA) on Day -1, the last day preceding the RT intervention.

### Experimental design
The study was designed as a 12-day double-blinded randomized controlled trial, with a unilateral RT protocol (Fig 2). Participants were randomized to commence with either glucose (GLU, 30g per bolus) or placebo (PLAC, 0g per bolus), and either dominant or non-dominant leg (Fig 1), alternating RT and supplement from one day to the other (Fig 2). Participants were linked to a unique ID number corresponding to supplement administration, disclosed only to personnel not involved in any other aspect of the study than the randomization.

###### **Table 1:** Baseline characteristics of the participants. BMI: body mass index, A = charactaristics per sex, B = characteristics per leg. Values are mean ± standard deviation.

```{r DXA}
# Script: DXA

# Data handling
table.dxa <- read_excel("./data/dxa/ribose_dxa.xlsx") %>%
        select(sex, subject, age, height, weight, leanmasskg, fatmasskg, ffmg) %>%
        #make grams into kg
        mutate(ffkg = ffmg/1000) %>%
        select(sex, age, height, weight, leanmasskg, fatmasskg, ffkg) %>%
  mutate(bmi = (weight) / (height^2) * 10000)

dxa.table <- table.dxa %>%
        pivot_longer(names_to = "variable",
                     values_to = "values", col= age:bmi) %>%
        select(sex, variable, values) %>%
        group_by(sex, variable) %>%
        summarise(m = mean(values, na.rm = TRUE),
                  s= sd(values, na.rm = TRUE)) %>%
        mutate(stat = paste0(round(m, 1), " (", round(s, 1), ")")) %>%
        select(sex, variable, stat) %>%
        pivot_wider(names_from = sex,
                    values_from = stat) %>%
                mutate(variable = if_else(variable %in% c("age"),
                                          "Age (yrs)",
                                          if_else(variable %in% c("fatmasskg"),
                                                  "Fat mass (kg)",
                                                  if_else(variable %in% c("ffkg"),
                                                          "Fat free mass (kg)",
                                                          if_else(variable %in% c("height"),
                                                                  "Height (cm)",
                                                                  if_else(variable %in% c("leanmasskg"),
                                                                          "Lean body mass (kg)",
                                                                          if_else(variable %in% c("weight"),
                                                                                  "Body weight (kg)",
                                                                                  if_else(variable %in% c("bmi"),
                                                                                          "BMI", "")
                                                                                  )))))))

dxa.table %>%
        kable(col.names = c("Variable", "Female (n=7)", "Male (n=9)"))

```

```{r dxa per leg}
# Script: DXA

# Subject 104, 110 and 111 dropped out and is therefore excluded from the table

#Load dxa data
dxaleg <-  read_excel("./data/dxa/ribose_dxaprleg.xlsx") %>%
        select(subject, sex, leg, leanmassg, fatmassg, totalmasskg) %>%
        #make grams into kg
        mutate(leanmass = leanmassg/1000,
               fatmass = fatmassg/1000)%>%
        select(subject, sex, leg, leanmass, fatmass, totalmasskg)

# Between groups analysis 
dxa.stat <- dxaleg %>%
        select(subject, leg, leanmass, fatmass, totalmasskg)

# Lean mass between legs, within subject 
leanmass.lm <- lmer(leanmass ~ leg + leg:subject + (1|subject), data = dxa.stat)
lm.plot <- plot(leanmass.lm)
lm.sum <- summary(leanmass.lm)
lm.confint <- emmeans(leanmass.lm, specs = ~ "leg")

# Fat mass between legs, within subject
fat.lm <- lmer(fatmass ~ leg + leg:subject +(1|subject), data = dxa.stat)
fat.plot <- plot(fat.lm)
fat.sum <- summary(fat.lm)
fat.confint <- emmeans(fat.lm, specs = ~ "leg")

# Total mass between legs, within subject
total.lm <- lmer(totalmasskg ~leg + leg:subject + (1|subject), data = dxa.stat)
totlm.plot <- plot(total.lm)
totlm.sum <- summary(total.lm)
supsup <- emmeans(total.lm, specs = ~ "leg")

# Table of leanmass, fatmass and total mass in kg pr. leg
dxalegtable <- dxaleg %>%
        select(leg, leanmass, fatmass, totalmasskg) %>%
        pivot_longer(names_to = "variable",
                     values_to = "values", col= leanmass:totalmasskg) %>%
        group_by(leg, variable) %>%
        summarise(m = mean(values, na.rm = TRUE),
                  s= sd(values, na.rm = TRUE)) %>%
        mutate(stat = paste0(round(m, 1), " (", round(s, 1), ")")) %>%
        select(leg, variable, stat) %>%
        pivot_wider(names_from = leg,
                    values_from = stat)

# Markdown table
dxalegtable %>%
        kable(col.names = c("Variable", "Left leg", "Right leg"))

```

Six RT sessions were conducted with glucose and six with placebo, allowing a within-subjects analysis of the effects of glucose ingestion. Data from the first five RT sessions was used to investigate main outcome measures (total RNA, rRNA and protein), whereas data from the sixth RT session was used to explore secondary outcomes (muscular recovery, plasma glucose levels). Participants were asked to refrain from resistance- or high-intensity training of the legs from Day -7 and onwards, until completion of the intervention and post-testing, to ensure the reliability of pre-intervention strength data and minimal interference from other training sources. If participants did conduct exercise training outside of the prescribed protocol, they were asked to ensure equal loading on pairwise consecutive days.

![](figures/flowchart_selection.PNG)

###### **Figure 1:** Flowchart of the selection process. The order in which participants performed the two intervention blocks was determined in a planned randomised fashion.

![](figures/timeline.PNG)

###### **Figure 2.** Overview of the intervention, with 12 days of concomitant dietary intervention and resistance training (RT), preceded by 7 days of testing. T0 = timepoint 0 (testing days prior to intervention), T1/T2 = biopsy prior to resistance training for leg #1 and #2, respectively, T3/T4 post intervention testing of leg #1 and #2, respectively, including post sixth RT session. Test = physical performance test (unilateral 1RM leg press, knee extension, isometric knee extension force, isokinetic knee extension torque) testing on days 4, 5, 8, 9, T3, T4 will not include 1RM leg press and knee extension, Biopsy = microbiopsy, RT#1 = resistance training leg #1, RT#2 = resistance training leg#2. Other tests (D2O, DXA, spit) will be conducted as part of the study as a whole. These are not relevant to this data set, thus will not be described here.

### Dietary intervention
The dietary intervention spanned the whole day, divided into three periods: I) From awakening until two hours after RT, II) from 2.5 hours after RT until 2200hrs, and III) from 2200hrs until awakening. During period I, participants ingested protein and glucose/placebo only. Glucose/placebo was ingested at three time points in period I: 30 minutes before RT (0830hrs, 30g vs 0g glucose), immediately before RT (0900hrs, 30g vs. 0g glucose), and immediately after RT (~0930hrs, 30g vs. 0g glucose). Protein was ingested 2hrs before RT (0700hrs, 25g) and immediately after RT (~0930hrs, 25g). In the afternoon (1800hrs-1900hrs), during period II, participants ingested glucose or placebo (3x30g vs. 3x0g glucose) opposite to the supplement they received during RT, to ensure a balanced intake of glucose during the entirety of the intervention days. Apart from this, participants ingested a self-chosen diet during period II, registered in MyFitnessPal or similar applications. The self-chosen diet was repeated on pairwise consecutive days (i.e. on days 1-2, 3-4, etc.), to ensure similar premises for resistance training responses between the two legs. During period III (2200-0700hrs), participants remained in an overnight fasted state. The daily onset of the dietary intervention (i.e., first ingestion of PRO supplement) varied for individual participants, between 0600hrs and 0900hrs to allow multiple participants to complete the protocol simultaneously. Individual participants commenced the intervention at the same time of day on every test day (pre and post) and intervention day +/- 1hr.

Blinding of the participants was ensured by diluting boluses of glucose (30g glucose) (Glucosum monohydricum, Merck KGaA, Darmstadt, Germany) and placebo (~0.3g Stevia rebaudiana extract) (Steviosa, Soma Nordic AS, Oslo, Norway) in 300ml Fun Light juice/saft (Orkla, Oslo, Norway). A blinded taste test revealed that the participants were not able to disclose the contents of the provided boluses (30g GLU vs. 0g GLU). The protein supplement was ingested as 25g Whey Protein Isolate boluses (Proteinfabrikken, Stokke, Norway), diluted in 150ml water. During sessions, participants were free to ingest water ad libitum.

### Assessment of muscular strength
Performance tests (Test) were performed prior to (Days -7 and -5, and T0, both legs) and during the intervention (Days 4, 5, 8, and 9), after session 5 and after finalization of the intervention (T3/T4, RT#1 leg; T4/Day 13, RT#2 leg), a total of 10 days (Fig 2). Maximal isometric and isokinetic knee extension torque were measured with a Humac Norm Dynamometer (CSMi, Stoughton, Massechusetts, USA). Individual positions were recorded and standardized from pre-intervention tests (days -7 and -5). Maximal isokinetic torque was measured at speeds of 60- and 240 d/s, 2x3 repetitions each, with the first set of each exercise as a sub-maximal warm-up. The maximal isometric torque was measured at knee-angle 60, for a maximum of 10 seconds and one repetition per test. The highest peak torque values were carried forwards to the final analyses. During days 4, 5, 8 and 9, humac tests were conducted one hour prior to RT with the leg performing RT the previous day. During days 11 and 12 (T3 and T4), humac tests were performed at four time points: I) 45min before RT, II) 30min after the last RT session, III) 2hrs after the last RT session, and IV) 23hrs after the last RT session. Test I at T4 included testing of both legs, representing 23hrs post RT session test of one leg and pre-session 6 test of the other leg.

Assessment of unilateral one repetition maximum (1RM) leg press and knee extension was conducted at pre-intervention testing (days -7 and -5, Fig 2). The participants warmed up with 10min cycling on an indoor exercise bicycle and a specific warm-up (1x10, 1x6 and 1x3 repetitions, 2min breaks) before each of the tests. Between specific warm-up sets as well as between each subsequent 1RM attempt, the participants were given 3min pauses. All positions were controlled and recorded at pre-intervention testing and repeated for each subsequent test and RT. Maximal leg press strength was defined as the maximal load lifted in a controlled fashion, with a knee angle of 45 degrees (start), within a 2.5kg range. Attempts, where participants did not reach 45 degrees during the eccentric phase, were not approved. Maximal knee extension testing followed the same specific warm-up and pause protocol and was defined as maximal load lifted in a controlled fashion, reaching full extension of the knee joint, within a 1.25kg range. Attempts with exaggerated hip movement or beneath full extension were not approved.

### Resistance training protocol
Resistance training consisted of three sets of each unilateral leg presses and knee extensions, with an intensity of 10 repetition maximum (10RM). All positions were controlled and recorded during pre-intervention testing and repeated each RT session. As a general warm-up, the participants cycled on an indoor exercise bicycle for 5-10min. In addition, before the respective exercises, two 10-repetition warm-up sets were completed at ~50% and ~70% of 10RM. Rest time between working sets was 2min, and we used progressive loading to ensure adequate exercise stimulation throughout the intervention. For safety and standardization purposes, all sessions were monitored by trained personnel. To further ensure equal exercise premises for the two legs, each participant was accompanied by the same trainer in each session. If this was not possible throughout the intervention, each participant had the same trainer on pairwise consecutive days as a minimum. The rate of perceived exertion (RPE, 0–10-point scale) was logged before every session, and the session score was logged 15min after each session. Lastly, training volume (i.e., load and repetitions) was logged for every session.

### Sampling of muscle tissue and blood
Muscle biopsies were sampled from m. vastus lateralis using well-established procedures (HAMMARSTRØM20). Muscle biopsy sampling was performed under local anaesthesia (Xylocaine, 10mg ml-1 with adrenaline 5μg ml−1, AstraZeneca AS, Oslo, Norway) using a 12-14-gauge needle (Universal Plus, Mermaid Medical AS, Stenløse, Denmark), operated with a spring-loaded biopsy gun. After the biopsy sampling, muscle tissue was divided into two aliquots for determination of total RNA/expression of rRNA and two aliquots for protein analyses, snap frozen in isopentane (-80°C) and stored at -80°C until further analyses. Muscle biopsies were sampled at four time points (Fig 2): I/II) Pre-intervention (2hrs before training, time point T1 = pre RT#1 leg, T2 = pre RT#2 leg), and III/IV) before the sixth RT session from RT#1 leg (2hrs before training, T3) and before the sixth RT session from RT#2 leg (2hrs before training, T4). At each time point, two samples were taken from the same incision. To standardize this procedure, all individual participants had biopsies taken at the same time of day +/- 1hr, in an overnight fasted state.
To measure blood glucose levels with and without glucose ingestion/training, blood was collected by means of finger draws on days with biopsy sampling. One finger draw was taken at T1 (pre-test, Fig 2) to serve as a baseline. At T3 and T4 (post-test, Fig 2), finger draws were collected at 7 time points: I) prior to protein ingestion (0700hrs) II) 45 minutes after protein ingestion (0745hrs) III) 1.5hrs after protein ingestion (0800hrs, i.e., immediately before GLU/PLAC intake), IIII) 2hrs after protein ingestion (0900hrs, i.e., immediately before training), IV) in the middle of RT (~0915hrs), V) immediately after training (~0930hrs), and VI) 2hrs after completion of training (~1130hrs). Finger draws were analysed with in-house equipment (BIOSEN C-Line, EKF diagnostic GmbH, Barleben). Blood samples were also collected coinciding with these finger draws, except in the middle of the RT session, to analyse endocrine variables.

### Total RNA and protein extraction
Two muscle biopsy samples were taken for total RNA extraction per leg per time point, and total RNA was extracted in two duplicates per muscle biopsy. Thus, we had two duplicates per leg per time point, giving a total of eight RNA samples per participant. Total RNA was extracted using TRIzol according to the manufacturer’s protocol. Briefly, muscle tissue was homogenized in 300µl TRIzol using 0.5mm RNase-free Zirconium beads (~50 ul) (Next Advanced, Averill Park, NY, USA) and a bullet blender (bb). Samples ran 1min at speed 10, put on ice for 1min and bb another round at speed 12 for 1 min. If not fully dissolved, samples ran additional rounds until fully homogenized. Thereafter, an additional 700 µl TRIzol was added, and samples ran 1 min at speed 3 in the bb for mixing, before 5 min incubation at room temperature. 200 µl chloroform (Sigma-Aldrich, Missouri, USA) was added, and samples were shaken for 15sec, followed by 2-3min incubation at room temperature. After incubation, samples were inverted by hand and centrifuged in Heraeus™ Fresco™ 21 Microcentrifuge (Thermo Fisher Scientific, Waltham, Massachusetts, USA) at 12000g, 15min, 4°C. The upper aqueous phase (450 µl) was transferred to a fresh tube, containing 500 µl isopropanol (VWR International, Pennsylvania, USA), mixed, and incubated for 10min at room temperature. Samples were centrifuged at 12000g for 10min at 4°C, precipitating an RNA pellet. The pellet was washed twice with 1000 µl 75% cold ethanol; centrifuged for 5 min at 7500g, 4°C. Thereafter, ethanol was removed, and the pellet was air dried for 10min (or until dried). The pellet was suspended in 30 µl DEPC-treated water and incubated for 10 min at 55°C. For assessment of RNA content and -purity 5µl RNA, from the RNA stock, was diluted with 5µl TE-buffer (1:2), for assessment via spectrophotometry. All samples had a 260nm to 280nm ratio > 1.9. The RNA stock was stored at -80°C until further analyses.

Total protein was extracted using the Minute Total Protein Extraction Kit for Muscles (Invent Biotechnology), according to the manufacturer’s protocol. Five to twenty mg (CHECK) of wet muscle was freeze-dried for 24hrs and dissected before extraction. The tissue was homogenized in approximately 80mg (CHECK) protein extraction powder and 100ul icy cold denaturing buffer (Invent Biotechnology) with a plastic rod, and centrifuged at top speed for 1 min. The supernatant was transferred to a fresh tube, and then aliquoted into 2x25ul tubes for analysis. Four microliter total protein was diluted in 36ul DEPC-treated water (or nuclease-free??) (1:10), for total protein measurement via spectrophotometry with the Bradford reagent (…).

### Complimentary DNA synthesis and Quantitative polymerase chain reaction

RNA was reverse transcribed using Super Script IV Reverse Transcriptase (Invitrogen), according to the manufacturer’s instructions. Briefly, 1µl 10 mM dNTP mix, 0.5µl anchored oligo-dT, 0.5µl random hexamer primers (Thermo Scientific), a maximum of 9µl template RNA and 2 µl nuclease-free water were mixed, vortexed and briefly spun down. Samples were then heated at 65°C for 5 minutes, followed by at least 1-minute incubation on ice. For the next step, 4µl 5x SSIV Buffer, 1µl 100 mM DTT, 1 µl RNase OUT, and Super Script IV Reverse Transcriptase (Invitrogen) were mixed, vortexed and briefly spun down and added to the samples. Samples were incubated for 10 minutes at 23°C, 10 minutes at 50-55°C and 10 minutes at 80°C. All samples were reverse transcribed and diluted to 1:50 before quantitative real-time polymerase chain reaction (qPCR). Lambda was used as an external reference gene, added in the RNA extraction (2µl per extraction). qPCR reactions were run on a fast-cycling real-time detection system (Applied Biosystems 7500 fast Real-Time PCR Systems, Life Technologies AS), with a total volume of 10 µl, containing 2 µl of cDNA, gene-specific primers (0.5 µM final concentration) and a commercial master mix (2X SYBR Select Master Mix, Applied Biosystems, Life Technologies AS) (Hammarström et al. 2020). qPCR reactions consisted of 40 cycles (3 s 95°C denaturing and 30 s 60°C annealing) (Hammarström et al. 2020). Raw fluorescence data were exported from the platform-specific software and amplification curves were modelled with a best-fit sigmoidal model using the qPCR package (Ritz & Spiess, 2008) written for R (Hammarström et al. 2020; Team 2018). For an overview of primers, see Table 2.

###### **Table 2:** Primer sequence and performance. Average cycle thresholds (Ct) and priming efficiencies were calculated from all qPCR reactions.

![](figures/Primer.sequences.PNG)

### Gel electrophoresis and Western Blot

#### Gel electrophoresis
Samples were standardized to a protein concentration of 3ug/ul via dilution with denaturing buffer (Invent Biotechnology), and laemmli/2-mercaptoethanol loading buffer (PRODUCER?) equating to ¼ of the sample’s total volume was then added to reach a final protein concentration of 2.25 ug/ul. A total of 20ul from each sample was pooled in several fresh tubes, serving as a normalisation factor in later quantitative analyses. All samples were incubated at 95°C for 5 minutes, briefly cooled and stored at -20°C until loading. In total, 9ul from each sample were loaded on six gels in duplicates, with 9ul of 3-4 pools per gel, on ice. Loading order from gel 1-3 was inverted for gel 4-6. Gels ran the electrophoresis simultaneously, at 250V constant for 50min on ice, ice-cold running buffer (250 mM Tris, 1.92 M Glycin, 1% SDS. Final concentrations in 1X: 25 mM Tris, 192 mM Glycin, 0.1% SDS).

#### Wet transfer
Transfer buffer (25 mM Tris, 192 mM Glycin, 10% MeOH) was made in due time of the transfer (2-5 days), to allow for “de-gas”. The PVDF-membrane was incubated for 5min in MeOH, then 10min in cold transfer buffer before transfer. Before sandwich assembly, foam pads and filter paper were soaked in transfer buffer, and the gel was flushed with ddH2O. The wet transfer was run at 300mA constant for 3hrs, in a Criterion Blotter (PRODUCER) on ice and stirring pad to maintain even buffer temperature and ion concentration during the transfer.

#### Staining
MemCode was used for protein staining, according to manufacturer’s protocol. The membranes were soaked and carefully shaken in MemCode Stain for 1min, washed with Destain/MeOH (1:1) on a shaker for 5min, with ddH2O washing before and between these steps. One manual capture (G:Box) per gel was taken, to procure total protein stain images. Following manual captures, all membranes were sliced into strips containing the bands of the specific protein targets (c-Myc, UBF, rps6), rinsed with ddH2O, washed with Eraser/MeOH solution (1:1) for 10min and rinsed with ddH2O once more. A 5% non-fat dry milk solution was used as blocking buffer and diluent for antibodies. Each membrane was blocked for 1hr at room temp, and washed once with TBS-T for 5min on a tipping board, then incubated overnight on a tipping board at 4°C in primary antibody (see table .. for primary antibody dilutions). Secondary antibody staining was done on a tipping board for 1hr at room temp (see table .. for secondary antibody dilutions), with three subsequent 5min TBS-T washes. Lastly, before imaging, the membranes were incubated for 5min in ECL (TYPE). Quantification was done using the G: Box and Genesys software with an ECL protocol.

### Data handling and statistical analyses
A priori power calculations showed that 20 participants would be sufficient to grant a statistical power of 80% (=0.05), with an expected dropout of 20%. As no studies existed to guide our power calculations, we made the assumption that the effects of glucose ingestion on ribosome biogenesis may equate to the effects of increased resistance training volume from low to moderate (HAMMARSTRÖM20). All raw data were imported and analysed in R (TEAM18). Individual participants’ leg results were categorized as either glucose or placebo according to the randomization. Total RNA and rRNA were analyzed by post to pre log-fold change score difference comparisons, using a linear mixed effects model with mean-centred baseline values as a covariate and subsequent calculations of estimated marginal means. Total RNA was normalised by muscle biopsy weight, and qPCR values were normlised by muscle biopsy weight and external reference gene (Lambda). Total RNA and rRNA changes were calculated as log-fold change scores per mg wet muscle weight. Blood glucose levels, dietary data, training volume and -intensity, RPE, and muscular strength and -recovery were analysed by multiple time point log-fold change score comparisons, using the same linear mixed effects model, adding time as an explaining factor in addition to supplement. The linear mixed effects model was design with the lmer function of the lme4 package (BATES14) in R, using the lmerTest function in addition to procure p-values. A paired samples t-test was used to check for differences in leg characteristics at baseline. All data was log-transformed for analysis to control for heteroscedasticity. For figure illustration, values were either reverse transformed or calculated as absolute changes. Mean-centred baseline values were used to correct for regression to the mean, which potentially occurs when the same participants are repeatedly tested. As such, varying baseline values should not affect change outcomes. Estimated marginal means were calculated from the linear model, enabling acquisition of least-square means, showing the means for all involved groups adjusted to means of other potential factors in the model. Values are presented as mean ± SEM unless otherwise stated. Alpha-level was set to α = 0.05

## Results

```{r glucose data + model}
# Scripts: gluc.responsechange.R, gluc.fig

# Data
gluc.dat <- read_excel("./data/glucose/fingerstick.xlsx", na = "NA")
gluc.dat$sample_time <- as.character(gluc.dat$sample_time)

# Data handling
glu.dat <- gluc.dat %>%
        mutate(time = if_else(sample_time == "0",
                              "baseline",
                              if_else(sample_time == "45",
                                      "min45",
                                      if_else(sample_time == "90",
                                              "min90",
                                              if_else(sample_time == "120",
                                                      "min120",
                                                      if_else(sample_time == "135",
                                                              "min135",
                                                              if_else(sample_time == "150",
                                                                      "min150",
                                                                      if_else(sample_time == "270",
                                                                              "min270", sample_time))))))),
               time = factor(time, levels = c("baseline", "min45", "min90", "min120", "min135", "min150", "min270"))) %>%
        mutate(glu = as.numeric(glu),
               lak = as.numeric(lak))

# Change data
change_dat.glu <- glu.dat %>%
        dplyr::select(subject, time, glu, supplement) %>%
        group_by(subject, time, supplement) %>%
        summarise(glu = mean(glu, na.rm = TRUE)) %>%
        pivot_wider(names_from = time, 
                    values_from = glu) %>%
        #print()

        ungroup() %>%
        mutate(change.45 = log(min45)-log(baseline),
               change.90 = log(min90)-log(baseline),
               change.120 = log(min120)-log(baseline),
               change.135 = log(min135)-log(baseline),
               change.150 = log(min150)-log(baseline),
               change.270 = log(min270)-log(baseline),
               baseline = baseline - mean(baseline, na.rm = TRUE),
               supplement = factor(supplement, levels = c("placebo", "glucose"))) %>%
        select(subject, supplement, baseline, change.45, change.90, change.120, change.135, 
               change.150, change.270) %>%
        pivot_longer(names_to = "time",
                     values_to = "change",
                     cols = (change.45:change.270))

# Model
m1 <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
                     data = change_dat.glu)
resplot.glu <- plot(m1)
sum.glu <- summary(m1)

# Emmeans
confint.m1 <- confint(emmeans(m1, specs = ~"supplement|time")) %>%
        data.frame()

# For inline plotting
bcoef <- coef(summary(m1))
pscient.120 <- round(bcoef[8,5], 30)
p.120 <- sprintf("%.5f", 1.700971e-16)
pscient.135 <- round(bcoef[9,5], 30)
p.135 <- sprintf("%.5f", 4.299568e-11)
pscient.150 <- round(bcoef[10,5], 30)
p.150 <- sprintf("%.5f", 1.445712e-11)
p.270 <- round(bcoef[11,5], 2)
```

At baseline, there were no significant differences between training with glucose or placebo in muscle mass, isometric or isokinetic peak torque, total session volume, training intensity, total RNA content, or rRNA expression (Tab 3). Further, on pairwise consecutive days, there were no significant differences in any macronutrient (calories, carbohydrate, fat or protein) ingestion (Tab 4). Resistance training with glucose led to significant increases in plasma glucose levels compared to RT with placebo, before, during and after the session (Fig 3B, 120min: 43 ± 4% [p = 0.00000], 135min: 32 ± 4% [p = 0.00000], 150min: 32 ± 4% [p = 0.00000]), measured during the final RT session. At 2hrs post-exercise, exercise training with glucose led to significantly lower plasma glucose levels compared to exercise training with placebo (8 ± 4%, p = 0.03).

###### **Table 3**: Baseline characteristics of the intervention legs, organized per supplement. Variables: Iso 60 = isokinetic 60 d/s, Iso 240 = isokinetic 240 d/s, training intensity = load/rm*100. Nm = newton meters, kg = kilograms, %1RM = percentage of 1 repetition maximum, ng/mg = nanogram per milligram wet muscle tissue, nf.expr = expression normalized by weight and external reference gene. rRNA data are presented as log transformed. Glucose n = 13, placebo n = 13.

![](figures/table3.png)

###### **Table 4:** Daily mean dietary intake of macro nutrients during each training day per supplement leg. Glu = glucose, Plac = placebo. Training day 1-training day 12 are combined pairwise to day 1-6: Day 1 = training day 1/2, Day 2 = training day 3/4, Day 3 = training day 5/6, Day 4 = training day 7/8, Day 5 = training day 9/10, Day 6 = training day 11/12. Values are mean ± standard deviation. P = difference between Glu and Plac. Glucose n = 12, placebo n = 12. Day 6: Glu = 11, Plac = 11.

```{r nutrition table}

# Data
nuttable <- readRDS("./data/data-gen/dxa/nuttable.RDS")

nuttable %>%
  kable(col.names = c("Time", "Suppl.", "Kcal", "CHO (g)", "Fat (g)", "PRO (g)", "PRO (g/kg)"))

```

```{r nutrition analysis}
#summarise macro
#nutritiontable with pro.pr kg bodyweight pr. group
#laod dxa data and take out weight and subject to join with nutrition data
fp.weight <-read_excel("data/dxa/ribose_dxa.xlsx") %>%
        select(weight, subject)

#nutrition data.
#the data is in grams and 
# added 50 grams protein as sup_protein for ingestion of whey protein isolat,
#And calculated in to the total amount of proteins ingested pr day.
#added 90g of glucose in to carbohydrate 
#added 360 kcal from glucose intake and calculated in total kcal pr day
#Load nutrition data.
nut.weight <- read_excel("data/dxa/Ribose_nutrition_result.xlsx")%>%
        select(timepoint, meal, subject, protein, fat, calories,
               carbohydrates, group, sup_pro, sup_gluc, kcal_glu) %>%
        #changed names to compare groups each paired days
        mutate(timepoint = if_else(timepoint %in% c("T1","T2"), 
                                   "Day 1",
                                   if_else(timepoint %in% c("3", "4"),
                                           "Day 2",
                                           if_else(timepoint %in% c("5", "6"),
                                                   "Day 3",
                                                   if_else(timepoint %in% c("7", "8"),
                                                           "Day 4",
                                                           if_else(timepoint %in% c("9", "10"),
                                                                   "Day 5",
                                                                   if_else(timepoint %in% c("T3", "T4"),
                                                                           "Day 6",
                                                                           "na"))))))) %>%
        group_by(timepoint, subject, group) %>%
        #take in weight
        inner_join(fp.weight) %>%
        #summarise each variable to get total pr timepoint.
        #Whey protein supplement is added to total protein ingestion
        summarise(protein = sum(protein + sup_pro),
                  fat = sum(fat),
                  calories = sum(calories + kcal_glu),
                  carbohydrates = sum(carbohydrates + sup_gluc),
                  weight = sum(weight)) %>%
        #summarise to get protein pr weight.
        mutate(proprkg = (protein/weight)) %>%
        pivot_wider(names_from = group,
                    values_from = c(fat, protein, calories, carbohydrates, weight, proprkg))

#nutrition table with pro/weight. The table shows mean and sd of total intake 
#in grams pr. group, pr. timepoint.
#How much kcal on glucose supplement?
nuttable <- nut.weight %>%
        pivot_longer(names_to = "variable",
                     values_to = "values", col= fat_glucose:proprkg_placebo) %>%
        group_by(timepoint, variable) %>%
        summarise(m = mean(values, na.rm = TRUE),
                  s= sd(values, na.rm = TRUE)) %>%
        ungroup() %>%
        separate(variable, into = c("variable", "group")) %>%
        mutate(stat = paste0(round(m, 1), " (", round(s, 1), ")")) %>%
        select(timepoint, variable, group, stat,) %>%
        pivot_wider(names_from = variable,
                    values_from = stat) %>%
        select(timepoint, group, calories, carbohydrates, fat, protein,
               proprkg) %>%
        #changed names of group for simplicity 
        mutate(group = if_else(group %in% c("glucose"),
                               "G",
                               if_else(group %in% c("placebo"),
                                       "P",
                                       "")))

# Change data
#Does not contain FP 106 and 116 because of not delivered data.
#summarise macro
#Added sup_pro as whey protein isolate, 50 grams each group, each timepoint,
#90 grams glucose and 360 kcal.
fp.weight <-read_excel("data/dxa/ribose_dxa.xlsx") %>%
        select(weight, subject) 
nutha <- read_excel("data/dxa/Ribose_nutrition_result.xlsx")%>%
        select(timepoint, subject, group, calories, fat, carbohydrates, protein, sup_pro, sup_gluc, kcal_glu) %>%
        mutate(timepoint = if_else(timepoint %in% c("T1","T2"), 
                                   "Day 1",
                                   if_else(timepoint %in% c("3", "4"),
                                           "Day 2",
                                           if_else(timepoint %in% c("5", "6"),
                                                   "Day 3",
                                                   if_else(timepoint %in% c("7", "8"),
                                                           "Day 4",
                                                           if_else(timepoint %in% c("9", "10"),
                                                                   "Day 5",
                                                                   if_else(timepoint %in% c("T3", "T4"),
                                                                           "Day 6",
                                                                           "na"))))))) %>%
        group_by(timepoint, subject, group) %>%
        inner_join(fp.weight) %>%
        summarise(protein = sum(protein + sup_pro),
                  fat = sum(fat),
                  calories = sum(calories + kcal_glu),
                  carbohydrates = sum(carbohydrates + sup_gluc),
                  weight = sum(weight)) %>%
        mutate(proprkg = (protein/weight))

#change data
#Protein
pro_change <- nutha %>%
        select(timepoint, subject, group, protein,) 
        #ungroup() %>%
        #pivot_wider(names_from = group,
         #           values_from = c(protein)) %>%
        #mutate(change = glucose - placebo) %>%
        #print()
prolm <- lmer(protein ~ timepoint + timepoint:group + (1|subject), data = pro_change)
pro.plot <- plot(prolm)
pro.sum <- summary(prolm)
#Fat
fat_change <- nutha %>%
        select(timepoint, subject, group, fat,) 
        #ungroup() %>%
        #pivot_wider(names_from = group,
         #           values_from = c(fat)) %>%
        #mutate(change = glucose - placebo) %>%
        #print()
fatlm <- lmer(fat  ~ timepoint + timepoint:group +(1|subject), data = fat_change)
fat.plot <- plot(fatlm)
fat.sum <- summary(fatlm)
#Carbohydrates
carb_change <- nutha %>%
        select(timepoint, subject, group, carbohydrates,) 
        #ungroup() %>%
        #pivot_wider(names_from = group,
         #           values_from = c(carbohydrates)) %>%
        #mutate(change = glucose - placebo) %>%
        #print()
carblm <- lmer(carbohydrates  ~ timepoint + timepoint:group + (1|subject), data = carb_change)
cho.plot <- plot(carblm)
cho.sum <- summary(carblm)
#calories
cal_change <- nutha %>%
        select(timepoint, subject, group, calories,) 
        #ungroup() %>%
        #pivot_wider(names_from = group,
         #           values_from = c(calories)) %>%
        #mutate(change = glucose - placebo) %>%
        #print()
callm <- lmer(calories  ~ 0 + timepoint + timepoint:group + (1|subject), data = cal_change)
cal.plot <- plot(callm)
cal.sum <- summary(callm)
#Protein pr bodyweigh
prowe <- nutha %>%
        select(timepoint, subject, group, proprkg) 
prokglm <- lmer(proprkg  ~ 0 + timepoint + timepoint:group + (1|subject), data = prowe)
prokglm.plot <- plot(prokglm)
prokglm.sum <- summary(prokglm)

```

![](figures/day.timeline.png)

```{r glucose fig}
# Figure
pos <- position_dodge(width = 0.2)
gluc.fig <- readRDS("./data/data-gen/glucose/gluc.fig.RDS")
gluc.fig

```

###### **Figure 4:** A) Time line for training day 6 (combination of pairwise consecutive training day 11 and 12). Biopsy = muscle biopsy taken pre supplementation, PRO = protein supplementation, Humac = strength test (isometric knee extension, isokinetic knee extension 60 and 240 d/s), GLU/PLAC = glucose/placebo supplementation, RT = resistance training. Minutes: -120 = 2hrs prior to RT, -90 = 90min prior to RT, -30 = 30min prior to RT, 0 = onset of RT, 15 = 15min following onset of RT, between leg press and knee extension, 30 = 30min post onset of RT, 120 = 2hrs following onset of RT. B) Mean mmol/L change in plasma glucose levels per supplement group during training day 6. Time-points: 0, 45, 90, 120, 135, 150 and 270min post protein ingestion, respectively. Finger stings were taken at all time-points, blood samples were taken at all time-points except from 135. Values are presented as estimated marginal means of mmol/L change ± 95% CI. *p* < 0.05 between groups. Glucose n = 13, placebo n = 13.

### Total RNA and ribosomal RNA

```{r total RNA data}

code <- read_excel("./data/code.key.xlsx")
rna_dat <- readRDS("./data/data-gen/rna/rna_complete.RDS")

```

```{r total RNA data handling}

# Scripts
# totrna_analysis
rna_dat2 <- rna_dat %>%
        inner_join(code) %>%
        mutate(time = if_else(time %in% c("T1", "T2"), "pre", "post"), 
               time = factor(time, levels = c("pre", "post"))) %>%
        # filter(outlier == "in") %>%
        group_by(subject, time, rep, outlier, technical, biopsy, supplement) %>%
        summarise(weight = mean(weight), 
                  RNA = mean(RNA_tot)) %>%
        ungroup() %>%
        mutate(weight.mc = weight/mean(weight)) %>%
        mutate(RNA.weight = RNA / weight)

#### Basic model 
change_dat <- rna_dat2 %>%
   #  filter(technical == "in", 
   #        outlier == "in") %>%
    dplyr::select(subject:rep, supplement, RNA.weight) %>%

    group_by(subject, time, supplement) %>%
    summarise(RNA.weight = mean(RNA.weight, na.rm = TRUE)) %>%

    pivot_wider(names_from = time, 
                values_from = RNA.weight) %>%
    ungroup() %>%
    mutate(change = log(post) - log(pre), 
           pre = pre - mean(pre, na.rm = TRUE), 
           supplement = factor(supplement, levels = c("PLACEBO", "GLUCOSE")))

```

```{r total RNA model}
# Scripts
# totrna_analysis
# Create model: 
m1 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
           data = change_dat)

resplot.rna <- plot(m1)
sum.rna <- summary(m1)

### Get estimated means from the model
confint.totrna <- confint(emmeans(m1, specs = ~"supplement")) %>%
    data.frame() 
# For inline reporting
trcoef <- coef(summary(m1))
trdiff <- round(trcoef[3,1], 2)
trdiffp <- round(trcoef[3,5], 2)
trdiffse <- round(trcoef[3,2], 2)

```

```{r rRNA data}
code_key <- read_excel("./data/code.key.xlsx") %>%
        dplyr::select(subject, supplement, time)
qpcr_data  <- readRDS("./data/data-gen/rna/qpcr-data.RDS")
samples <- read_excel("./data/rna/RNA.raw.xlsx") %>%
        dplyr::select(subject, time_rep, sample, weight)

```

```{r rRNA data handling}
# Scripts
# - qpcr-import
# - qpcr-analysis
# - rrna.change
# - rrnachange.figs

nf <- qpcr_data %>%
        mutate(sample = as.numeric(sample)) %>% 
        inner_join(samples) %>%
        separate(time_rep, into = c("time","rep"), sep = "rna") %>%
        mutate(rep = paste0("cdna", rep)) %>%
        inner_join(code_key) %>% 
        mutate(time = if_else(time %in% c("T1", "T2"), "Pre", "Post"), 
               time = factor(time, levels = c("Pre", "Post")), 
               expr = eff ^ -cq) %>%

        filter(target %in% c("Lambda F2R2", "Lambda F3R3"), 
               cq < 35) %>%
        separate(target, into = c("target", "nf_primer")) %>%

        mutate(nf.w = expr * weight, 
               nf.w = nf.w / max(nf.w)) %>%

        dplyr::select(sample, subject, time, rep, supplement, nf.w, nf_primer)

qpcr_data2 <- qpcr_data %>%
        mutate(sample = as.numeric(sample)) %>% 
        inner_join(samples) %>%
        inner_join(nf) %>%
        separate(time_rep, into = c("time","rep"), sep = "rna") %>%
        mutate(rep = paste0("cdna", rep)) %>%
        inner_join(code_key) %>% 
        mutate(time = if_else(time %in% c("T1", "T2"), "Pre", "Post"), 
               time = factor(time, levels = c("Pre", "Post")), 
               expr = eff ^ -cq) %>%

        filter(!(target %in% c("Lambda F2R2", "Lambda F3R3"))) %>%   
        inner_join(nf)

saveRDS(qpcr_data2, "./data/data-gen/rna/qpcr_data2.RDS")

# Plotting mean expression per treatment for rRNA
initmod <- qpcr_data2 %>%
        mutate(norm.expr = log(expr / nf.w)) %>%
        group_by(time, supplement, target) %>%
        summarise(m = mean(norm.expr, na.rm = TRUE))  %>%
        #print()

        ggplot(aes(time, m, color = supplement, group = supplement)) + 
        geom_line() +
        geom_point() +
        facet_wrap( ~ target, scales = "free")
qdat <- qpcr_data2 %>%
        mutate(target = gsub("rRNA", "", target),
               target = gsub("  ", " ", target),
               target = paste0("trg_", target)) %>%
        separate(target, into = c("target", "primer"), sep = " ") %>%

        # Create a weight-normalized variable
        mutate(nf.expr = log(expr / nf.w), 
               nf.w = scale(nf.w),
               # Technical random effect
               technical = paste(subject, time, supplement, rep, sep = "_"),
               biological = paste0("S", sample)) %>%
        filter(!(target %in% c("trg_MHC1", "trg_MHC2A", "trg_MHC2X")))

## Change data
rrna18 <- qdat %>%
        filter(target == "trg_18s")
rrna28 <- qdat %>%
        filter(target == "trg_28s")
rrna5.8 <- qdat %>%
        filter(target == "trg_5.8s")
rrna5 <- qdat %>%
        filter(target == "trg_5s")
rrna47 <- qdat %>%
        filter(target == "trg_47s")

# Change scores
# 18S
change.18 <- rrna18 %>%
        dplyr::select(subject, time, rep, nf.expr, supplement) %>%
        group_by(subject, time, supplement) %>%
        summarise(nf.expr = mean(nf.expr, na.rm = TRUE)) %>%
        pivot_wider(names_from = time,
                    values_from = nf.expr) %>%
        ungroup() %>%
        # print()
        mutate(change = Post-Pre,
               pre = Pre - mean(Pre, na.rm = TRUE),
               supplement = factor(supplement, levels = c("PLACEBO", "GLUCOSE"))) 
# 28S
change.28 <- rrna28 %>%
        dplyr::select(subject, time, rep, nf.expr, supplement) %>%
        group_by(subject, time, supplement) %>%
        summarise(nf.expr = mean(nf.expr, na.rm = TRUE)) %>%
        pivot_wider(names_from = time,
                    values_from = nf.expr) %>%
        ungroup() %>%
        # print()
        mutate(change = Post-Pre,
               pre = Pre - mean(Pre, na.rm = TRUE),
               supplement = factor(supplement, levels = c("PLACEBO", "GLUCOSE"))) 
# 5.8S
change.5.8 <- rrna5.8 %>%
        dplyr::select(subject, time, rep, nf.expr, supplement) %>%
        group_by(subject, time, supplement) %>%
        summarise(nf.expr = mean(nf.expr, na.rm = TRUE)) %>%
        pivot_wider(names_from = time,
                    values_from = nf.expr) %>%
        ungroup() %>%
        # print()
        mutate(change = Post-Pre,
               pre = Pre - mean(Pre, na.rm = TRUE),
               supplement = factor(supplement, levels = c("PLACEBO", "GLUCOSE")))
# 5S
change.5 <- rrna5 %>%
        dplyr::select(subject, time, rep, nf.expr, supplement) %>%
        group_by(subject, time, supplement) %>%
        summarise(nf.expr = mean(nf.expr, na.rm = TRUE)) %>%
        pivot_wider(names_from = time,
                    values_from = nf.expr) %>%
        ungroup() %>%
        # print()
        mutate(change = Post-Pre,
               pre = Pre - mean(Pre, na.rm = TRUE),
               supplement = factor(supplement, levels = c("PLACEBO", "GLUCOSE")))
# 47S
change.47 <- rrna47 %>%
        dplyr::select(subject, time, rep, nf.expr, supplement) %>%
        group_by(subject, time, supplement) %>%
        summarise(nf.expr = mean(nf.expr, na.rm = TRUE)) %>%
        pivot_wider(names_from = time,
                    values_from = nf.expr) %>%
        ungroup() %>%
        # print()
        mutate(change = Post-Pre,
               pre = Pre - mean(Pre, na.rm = TRUE),
               supplement = factor(supplement, levels = c("PLACEBO", "GLUCOSE")))
```

```{r rRNA data modelling + fig}
# Scripts
# - qpcr-import
# - qpcr-analysis
# - rrna.change
# - rrnachange.figs

# Create model: 
# Needs to have an intercept per participant (mixed model)
# Control for pre values.

# 18S
m1 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
                     data = change.18)
resplot.18 <- plot(m1)
sum.18 <- summary(m1)

# 28S
m2 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
                     data = change.28)
resplot.28 <- plot(m2)
sum.28 <- summary(m2)

# 5.8S
m3 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
                     data = change.5.8)
resplot.58 <- plot(m3)
sum.58 <- summary(m3)

# 5S
m4 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
                     data = change.5)
resplot.5 <- plot(m4)
sum.5 <- summary(m4)

# 47S
m5 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
                     data = change.47)
resplot.47 <- plot(m5)
sum.47 <- summary(m5)

### Get estimated means from the model, these are average increase at 
# pre = 0 (the average pre value)
# remember that these are originally log-fold change values, transformed back using exp(emmean) etc

# 18S
confint.18 <- confint(emmeans(m1, specs = ~"supplement")) %>%
        data.frame()

# 28S
confint.28 <- confint(emmeans(m2, specs = ~"supplement")) %>%
        data.frame()

# 5.8S
confint.5.8 <- confint(emmeans(m3, specs = ~"supplement")) %>%
        data.frame()

# 5S
confint.5 <- confint(emmeans(m4, specs = ~"supplement")) %>%
        data.frame()

# 47S
confint.47 <- confint(emmeans(m5, specs = ~"supplement")) %>%
        data.frame()

## For inline reporting

# 18S
rr18coef <- coef(summary(m1))
rr18diffp <- round(rr18coef[3,5], 3)
rr18diffse <- round(rr18coef[3,2], 3)

# 28S
rr28coef <- coef(summary(m2))
rr28diffp <- round(rr28coef[3,5], 3)
rr28diffse <- round(rr28coef[3,2], 3)

# 5.8S
rr58coef <- coef(summary(m3))
rr58diffp <- round(rr58coef[3,5], 3)
rr58diffse <- round(rr58coef[3,2], 3)

# 5S
rr5coef <- coef(summary(m4))
rr5diffp <- round(rr5coef[3,5], 3)
rr5diffse <- round(rr5coef[3,2], 3)

# 47S
rr47coef <- coef(summary(m5))
rr47diffp <- round(rr47coef[3,5], 3)
rr47diffse <- round(rr47coef[3,2], 3)

```

Resistance training with glucose did not induce a higher accumulation of total RNA (Fig 4A, p = 0.499), or rRNA (Fig 4B, 18S: p = 0.584, 28S: p = 0.740, 5.8S: p = 0.935, 5S: p = 0.790, 47S: p = 0.502), compared to RT with placebo, measured as mean change from baseline to post. From baseline to post-intervention, there was an average increase in levels of total RNA, 26 and 22% after RT with glucose and placebo, respectively. A robust accumulation was also observed in ribosomal RNA expression in both RT with glucose and placebo, with mean increases between 34-43% (GLU) and 33-41% (PLAC) in the four rRNA’s, and 37 and 59% in 47S pre-rRNA, respectively.

```{r RNA figures}
# Get figures
totrna.fig <- readRDS("./data/data-gen/rna/totrna.fig.RDS")
totrna.fig

plot.18 <- readRDS("./data/data-gen/rna/plot.18.RDS")
plot.28 <- readRDS("./data/data-gen/rna/plot.28.RDS")
plot.5.8 <- readRDS("./data/data-gen/rna/plot.5.8.RDS")
plot.5 <- readRDS("./data/data-gen/rna/plot.5.RDS")
plot.47 <- readRDS("./data/data-gen/rna/plot.47.RDS")

# Joined rRNA
library(cowplot)

legend <- get_legend(plot.47 + theme(legend.box.margin = margin(0, 0, 0,12)))

plot_grid(plot.18 + theme(legend.position = "none"),
          plot.28 + theme(legend.position = "none"), 
          plot.5.8 + theme(legend.position = "none"),
          plot.5 + theme(legend.position = "none"), 
          plot.47 + theme(legend.position = "none"),
          legend,
          ncol = 2, nrow = 3)

```

###### **Figure 5:** A) Changes in Total RNA, B) Changes in ribosomal RNA. Baseline = Training day 1/2, Post = Training day 11/12. Both total RNA and rRNA were analyzed with duplicates, with two duplicates per biopsy (two biopsies per time point). Total RNA and rRNA changes were calculated as log-fold change score per mg wet muscle weight. Mean change scores of the duplicates were calculated before modelling and transformed to the log-scale. Values are estimated marginal means fold change per leg ± 95% CI. p = between groups statistic, glucose compared to placebo. Glucose n = 13, placebo n = 13.

### Resistance training and strength testing

```{r training data}
lp.load <- read_excel("./data/training/ribose_training.xlsx", na = "NA") %>%
        select(subject, timepoint, supplement, exercise, load) %>%
        filter(exercise != "KE")

lp.rm <- read_excel("./data/tests/ribose.1rm.xlsx", na = "NA") %>%
        select(subject, supplement, exercise, rm) %>%
        filter(exercise != "KE")

tot.load <- read_excel("./data/training/ribose_training.xlsx", na = "NA") %>%
        select(subject, timepoint, supplement, exercise, load)

tot.rm <- read_excel("./data/tests/ribose.1rm.xlsx", na = "NA") %>%
        select(subject, supplement, exercise, rm)

tot.vol <- read_excel("./data/training/ribose_volume.xlsx", na = "NA") %>%
        select(subject, timepoint, tot.volume, supplement)

```

```{r training data handling}
# Scripts
# tot.vol.change
# Trainingint.linear
# rpe-analysis

### Training volume
# Data handling
tot.volh <- tot.vol %>%
        mutate(time = if_else(timepoint %in% c("T1", "T2"),
                              "baseline",
                              if_else(timepoint %in% c("D3", "D4"),
                                      "session2",
                                      if_else(timepoint %in% c("D5", "D6"),
                                              "session3",
                                              if_else(timepoint %in% c("D7", "D8"),
                                                      "session4",
                                                      if_else(timepoint %in% c("D9", "D10"),
                                                              "session5",
                                                              if_else(timepoint %in% c("T3", "T4"),
                                                                      "session6", timepoint))))))) %>%
        mutate(time = factor(time, levels = c("baseline", "session1", "session2", "session3", "session4", "session5", "session6")),
               supplement = factor(supplement, levels = c("placebo", "glucose")))

## Absolute data
vol.exp <- tot.volh %>%
        select(subject, supplement, time, tot.volume) %>%
        group_by(supplement, time) %>%
        summarise(mean.vol = mean(tot.volume),
                  sd.vol = sd(tot.volume))

# Fold-change
change_dat <- tot.volh %>%
        dplyr::select(subject, time, tot.volume, supplement) %>%
        group_by(subject, time, supplement) %>%
        summarise(tot.volume = mean(tot.volume, na.rm = TRUE)) %>%
        pivot_wider(names_from = time, 
                    values_from = tot.volume) %>%
        # print()

        ungroup() %>%
        mutate(change.2 = log(session2)-log(baseline),
               change.3 = log(session3)-log(baseline),
               change.4 = log(session4)-log(baseline),
               change.5 = log(session5)-log(baseline),
               change.6 = log(session6)-log(baseline),
               baseline = baseline - mean(baseline, na.rm = TRUE),
               supplement = factor(supplement, levels = c("placebo", "glucose"))) %>%
        select(subject, supplement, baseline, change.2, change.3, change.4, change.5, change.6) %>%
        pivot_longer(names_to = "time",
                     values_to = "change",
                     cols = (change.2:change.6))

### RPE
## Data handling
ss.dat <- read_excel("./data/training/ribose_rpe.xlsx", na = "NA") %>%
        select(subject, timepoint, session.score, supplement) %>%
        mutate(time = if_else(timepoint %in% c("T1", "T2"),
                              "baseline",
                              if_else(timepoint %in% c("D3", "D4"),
                                      "session2",
                                      if_else(timepoint %in% c("D5", "D6"),
                                              "session3",
                                              if_else(timepoint %in% c("D7", "D8"),
                                                      "session4",
                                                      if_else(timepoint %in% c("D9", "D10"),
                                                              "session5",
                                                              if_else(timepoint %in% c("T3", "T4"),
                                                                      "post", timepoint))))))) %>%
        mutate(time = factor(time, levels = c("baseline", "session1", "session2", "session3", "session4", "session5", "post")),
               supplement = factor(supplement, levels = c("placebo", "glucose")))

## Absolute data
ss.exp <- ss.dat %>%
        select(subject, supplement, time, session.score) %>%
        group_by(supplement, time) %>%
        summarise(mean.ss = mean(session.score),
                  sd.ss = sd(session.score))

# Change data
change_dat.ss <- ss.dat %>%
        dplyr::select(subject, time, session.score, supplement) %>%
        group_by(subject, time, supplement) %>%
        summarise(session.score = mean(session.score, na.rm = TRUE)) %>%
        pivot_wider(names_from = time, 
                    values_from = session.score) %>%
        #print()

        ungroup() %>%
        mutate(change.2 = log(session2)-log(baseline),
               change.3 = log(session3)-log(baseline),
               change.4 = log(session4)-log(baseline),
               change.5 = log(session5)-log(baseline),
               change.post = log(post)-log(baseline),
               baseline = baseline - mean(baseline, na.rm = TRUE),
               supplement = factor(supplement, levels = c("placebo", "glucose"))) %>%
        select(subject, supplement, baseline, change.2, change.3, change.4, change.5, change.post) %>%
        pivot_longer(names_to = "time",
                     values_to = "change",
                     cols = (change.2:change.post))

### Training intensity
# Data handling
tot.loadh <- tot.load %>%
        mutate(time = if_else(timepoint %in% c("T1", "T2"),
                              "baseline",
                              if_else(timepoint %in% c("D3", "D4"),
                                      "session2",
                                      if_else(timepoint %in% c("D5", "D6"),
                                              "session3",
                                              if_else(timepoint %in% c("D7", "D8"),
                                                      "session4",
                                                      if_else(timepoint %in% c("D9", "D10"),
                                                              "session5",
                                                              if_else(timepoint %in% c("T3", "T4"),
                                                                      "session6", timepoint))))))) %>%
        mutate(time = factor(time, levels = c("baseline", "session1", "session2", "session3", "session4", "session5", "session6")),
               supplement = factor(supplement, levels = c("placebo", "glucose")))

# Join data and calulate %1RM
tot.joined <- tot.loadh %>%
        full_join(tot.rm) %>%
        mutate(p.rm = (load/rm)*100)

mean.joined <- tot.joined %>%
        select(subject, supplement, time, p.rm) %>%
        group_by(supplement, time) %>%
        summarise(mean.prm = mean(p.rm),
                  sd.prm = sd(p.rm))

## Change data
tot.change <- tot.joined %>%
        dplyr::select(subject, time, p.rm, supplement) %>%
        group_by(subject, time, supplement) %>%
        summarise(mean.prm = mean(p.rm)) %>%
        pivot_wider(names_from = time,
                    values_from = mean.prm) %>%
        ungroup() %>%
        mutate(change.2 = log(session2)-log(baseline),
               change.3 = log(session3)-log(baseline),
               change.4 = log(session4)-log(baseline),
               change.5 = log(session5)-log(baseline),
               change.6 = log(session6)-log(baseline),
               baseline = baseline - mean(baseline, na.rm = TRUE),
               supplement = factor(supplement, levels = c("placebo", "glucose"))) %>%
        select(subject, supplement, baseline, change.2, change.3, change.4, change.5, change.6) %>%
        pivot_longer(names_to = "time",
                     values_to = "change",
                     cols = (change.2:change.6))
```

```{r training data modelling}
# Scripts
# tot.vol.change
# Trainingint.linear
# rpe-analysis

### Training volume
# Fold-change model
m1 <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
                     data = change_dat)
resplot.vol <- plot(m1)
sum.vol <- summary(m1)

# Fold-change estimated means
confint.m1 <- confint(emmeans(m1, specs = ~"supplement|time")) %>%
        data.frame()

### RPE
# Model
m2 <- lmerTest::lmer(change ~ baseline + time + supplement:time + (1|subject),
                     data = change_dat.ss)
resplot.ss <- plot(m2)
sum.ss <- summary(m2)

# Session score emmeans
confint.m2 <- confint(emmeans(m2, specs = ~"supplement|time")) %>%
        data.frame()

### Training intensity
m3 <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
                     data = tot.change)
resplot.ti <- plot(m3)
sum.ti <- summary(m3)

# Emmeans
confint.m3 <- confint(emmeans(m3, specs = ~"supplement|time")) %>%
        data.frame()

### For inline reporting

## Training volume
# Diff 4
vcoef <- coef(summary(m1))
vdiff.4 <- round(vcoef[9,1], 3)
vdiffp.4 <- round(vcoef[9,5], 3)
vdiffse.4 <- round(vcoef[9,2], 3)
# Diff 5
vdiff.5 <- round(vcoef[10,1], 3)
vdiffp.5 <- round(vcoef[10,5], 3)
vdiffse.5 <- round(vcoef[10,2], 3)
# Diff 6
vdiff.6 <- round(vcoef[11,1], 3)
vdiffp.6 <- round(vcoef[11,5], 3)
vdiffse.6 <- round(vcoef[11,2], 3)
## RPE
scoef <- coef(summary(m2))
sdiffp <- round(scoef[11, 5], 3)
sdiffse <- round(scoef[11, 2], 3)

## Training intensity

# Diff time 2
icoef <- coef(summary(m3))
idiff2 <- round(icoef[7,1], 3)
idiffp2 <- round(icoef[7,5], 3)
idiffse2 <- round(icoef[7,2], 3)
# Diff time 3
idiff3 <- round(icoef[8,1], 3)
idiffp3 <- round(icoef[8,5], 3)
idiffse3 <- round(icoef[8,2], 3)
# Diff time 4
idiff4 <- round(icoef[9,1], 3)
idiffp4 <- round(icoef[9,5], 3)
idiffse4 <- round(icoef[9,2], 3)
# Diff time 5
idiff5 <- round(icoef[10,1], 3)
idiffp5 <- round(icoef[10,5], 3)
idiffse5 <- round(icoef[10,2], 3)
# Diff time 6
idiff6 <- round(icoef[11,1], 3)
idiffp6 <- round(icoef[11,5], 3)
idiffse6 <- round(icoef[11,2], 3)

```

```{r post 5th data handling}
# Scripts
# - Post5th.change
# - Post5th.figs
humac <- read_excel("./data/tests/ribose.humac.xlsx", na = "NA") %>%

        mutate(time = if_else(timepoint == "D-1", 
                              "baseline", 
                              if_else(timepoint %in% c("D4", "D5"), 
                                      "test1", 
                                      if_else(timepoint %in% c("D8", "D9"), 
                                              "test2", 
                                              if_else(timepoint %in% c("T3", "T4") & acute %in% c("rest", "post30min", "post2h"),
                                                      "test3",
                                                      if_else(acute == "post23h", "test4", timepoint)))))) %>%
        mutate(time = factor(time, levels = c("baseline", "test1", "test2", "test3", "test4")), 
               acute = factor(acute, levels = c("rest", "post30min", "post2h", "post23h")), 
               supplement = factor(supplement, levels = c("placebo", "glucose")))
rest.dat <- humac %>%
        filter(acute == "rest" )

# Change data
isom.dat <- rest.dat %>%
        filter(test == "isom")
change_dat <- isom.dat %>%
        dplyr::select(subject, time, supplement, peak.torque) %>%
        group_by(subject, time, supplement) %>%
        summarise(peak.torque = mean(peak.torque, na.rm = TRUE)) %>%
        pivot_wider(names_from = time, 
                    values_from = peak.torque) %>%

        ungroup() %>%
        mutate(change.2 = log(test1)-log(baseline),
               change.3 = log(test2)-log(baseline),
               change.4 = log(test3)-log(baseline),
               baseline = baseline - mean(baseline, na.rm = TRUE),
               supplement = factor(supplement, levels = c("placebo", "glucose"))) %>%
        select(subject, supplement, baseline, change.2, change.3, change.4) %>%
        pivot_longer(names_to = "time",
                     values_to = "change",
                     cols = (change.2:change.4))

# Isok.60
isok60.dat <- rest.dat %>%
        filter(test == "isok.60")
change_dat2 <- isok60.dat %>%
        dplyr::select(subject, time, supplement, peak.torque) %>%
        group_by(subject, time, supplement) %>%
        summarise(peak.torque = mean(peak.torque, na.rm = TRUE)) %>%
        pivot_wider(names_from = time, 
                    values_from = peak.torque) %>%

        ungroup() %>%
        mutate(change.2 = log(test1)-log(baseline),
               change.3 = log(test2)-log(baseline),
               change.4 = log(test3)-log(baseline),
               baseline = baseline - mean(baseline, na.rm = TRUE),
               supplement = factor(supplement, levels = c("placebo", "glucose"))) %>%
        select(subject, supplement, baseline, change.2, change.3, change.4) %>%
        pivot_longer(names_to = "time",
                     values_to = "change",
                     cols = (change.2:change.4))

## Isok.240
isok240.dat <- rest.dat %>%
        filter(test == "isok.240") 
change_dat3 <- isok240.dat %>%
        dplyr::select(subject, time, supplement, peak.torque) %>%
        group_by(subject, time, supplement) %>%
        summarise(peak.torque = mean(peak.torque, na.rm = TRUE)) %>%
        pivot_wider(names_from = time, 
                    values_from = peak.torque) %>%

        ungroup() %>%
        mutate(change.2 = log(test1)-log(baseline),
               change.3 = log(test2)-log(baseline),
               change.4 = log(test3)-log(baseline),
               baseline = baseline - mean(baseline, na.rm = TRUE),
               supplement = factor(supplement, levels = c("placebo", "glucose"))) %>%
        select(subject, supplement, baseline, change.2, change.3, change.4) %>%
        pivot_longer(names_to = "time",
                     values_to = "change",
                     cols = (change.2:change.4))

```

```{r post 5th modelling}
# Scripts
# - Post5th.change
# - Post5th.figs

# Isometric
m1 <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
                     data = change_dat)
resplot.iso5 <- plot(m1)
sum.iso5 <- summary(m1)

# Isok.60
m2 <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
                     data = change_dat2)
resplot.isok605 <- plot(m2)
sum.isok605 <- summary(m2)

# Isok.240
m3 <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
                     data = change_dat3)
resplot.isok2405 <- plot(m3)
resplot.isok2405 <- summary(m3)

## Emmeans
# Isometric
confint.m1 <- confint(emmeans(m1, specs = ~"supplement|time")) %>%
        data.frame()

# Isok.60
confint.m2 <- confint(emmeans(m2, specs = ~"supplement|time")) %>%
        data.frame()

# Isok.240
confint.m3 <- confint(emmeans(m3, specs = ~"supplement|time")) %>%
        data.frame()

# For reporting
# Isom diff
isom5coef <- coef(summary(m1))
isom5diffp <- round(isom5coef[7, 5], 3)
isom5diffse <- round(isom5coef[7, 2], 3)

# Isok 60 diff
isok605coef <- coef(summary(m2))
isok605diffp <- round(isok605coef[7, 5], 3)
isok605diffse <- round(isok605coef[7, 2], 3)

# Isok 260
isok2405coef <- coef(summary(m3))
isok2405diffp <- round(isok2405coef[7, 5], 3)
isok2405diffse <- round(isok2405coef[7, 2], 3)

```

```{r post 6th data handling}
# Scripts
# - Post6th.change
# - Post6th.figs
humac <- read_excel("./data/tests/ribose.humac.xlsx", na = "NA") %>%

        mutate(time = if_else(timepoint == "D-1", 
                              "baseline", 
                              if_else(timepoint %in% c("D4", "D5"), 
                                      "test1", 
                                      if_else(timepoint %in% c("D8", "D9"), 
                                              "test2", 
                                              if_else(timepoint %in% c("T3", "T4") & acute %in% c("rest", "post30min", "post2h"),
                                                      "test3",
                                                      if_else(acute == "post23h", "test4", timepoint)))))) %>%
        mutate(time = factor(time, levels = c("baseline", "test1", "test2", "test3", "test4")), 
               acute = factor(acute, levels = c("rest", "post30min", "post2h", "post23h")), 
               supplement = factor(supplement, levels = c("placebo", "glucose")))

# Isometric
isom.dat <- humac %>%
        filter(test == "isom",
               time %in% c("test3", "test4")) 
change_dat <- isom.dat %>%
        dplyr::select(subject, acute, supplement, peak.torque) %>%
        group_by(subject, acute, supplement) %>%
        summarise(peak.torque = mean(peak.torque, na.rm = TRUE)) %>%
        pivot_wider(names_from = acute, 
                    values_from = peak.torque) %>%
        #print()

        ungroup() %>%
        mutate(change.2 = log(post30min)-log(rest),
               change.3 = log(post2h)-log(rest),
               change.4 = log(post23h)-log(rest),
               baseline = rest - mean(rest, na.rm = TRUE),
               supplement = factor(supplement, levels = c("placebo", "glucose"))) %>%
        select(subject, supplement, baseline, change.2, change.3, change.4) %>%
        pivot_longer(names_to = "time",
                     values_to = "change",
                     cols = (change.2:change.4))

# Isok.60
isok60.dat <- humac %>%
        filter(test == "isok.60",
               time %in% c("test3", "test4")) 
change_dat2 <- isok60.dat %>%
        dplyr::select(subject, acute, supplement, peak.torque) %>%
        group_by(subject, acute, supplement) %>%
        summarise(peak.torque = mean(peak.torque, na.rm = TRUE)) %>%
        pivot_wider(names_from = acute, 
                    values_from = peak.torque) %>%
        #print()

        ungroup() %>%
        mutate(change.2 = log(post30min)-log(rest),
               change.3 = log(post2h)-log(rest),
               change.4 = log(post23h)-log(rest),
               baseline = rest - mean(rest, na.rm = TRUE),
               supplement = factor(supplement, levels = c("placebo", "glucose"))) %>%
        select(subject, supplement, baseline, change.2, change.3, change.4) %>%
        pivot_longer(names_to = "time",
                     values_to = "change",
                     cols = (change.2:change.4)) 
## Isok.240
isok240.dat <- humac %>%
        filter(test == "isok.240",
               time %in% c("test3", "test4")) 
change_dat3 <- isok240.dat %>%
        dplyr::select(subject, acute, supplement, peak.torque) %>%
        group_by(subject, acute, supplement) %>%
        summarise(peak.torque = mean(peak.torque, na.rm = TRUE)) %>%
        pivot_wider(names_from = acute, 
                    values_from = peak.torque) %>%
        #print()

        ungroup() %>%
        mutate(change.2 = log(post30min)-log(rest),
               change.3 = log(post2h)-log(rest),
               change.4 = log(post23h)-log(rest),
               baseline = rest - mean(rest, na.rm = TRUE),
               supplement = factor(supplement, levels = c("placebo", "glucose"))) %>%
        select(subject, supplement, baseline, change.2, change.3, change.4) %>%
        pivot_longer(names_to = "time",
                     values_to = "change",
                     cols = (change.2:change.4))

```

```{r post 6th modelling}
# Scripts
# - Post6th.change
# - Post6th.figs

# Isometric
m1 <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
                     data = change_dat)
resplot.isom <- plot(m1)
sum.isom <- summary(m1)

# Isok.60
m2 <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
                     data = change_dat2)
resplot.isok60 <- plot(m2)
sum.isok60 <- summary(m2)

# Isok.240
m3 <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
                     data = change_dat3)
resplot.isok240 <- plot(m3)
sum.isok240 <- summary(m3)

# Emmeans
# Isometric
confint.m1 <- confint(emmeans(m1, specs = ~"supplement|time")) %>%
        data.frame()

# Isok.60
confint.m2 <- confint(emmeans(m2, specs = ~"supplement|time")) %>%
        data.frame()

# Isok.240
confint.m3 <- confint(emmeans(m3, specs = ~"supplement|time")) %>%
        data.frame()

# For inline reporting
# Isom diff
isomcoef <- coef(summary(m1))
isomdiffp <- round(isomcoef[7, 5], 3)
isomdiffse <- round(isomcoef[7, 2], 3)

# Isok 60 diff
isok60coef <- coef(summary(m2))
isok60diffp <- round(isok60coef[7, 5], 3)
isok60diffse <- round(isok60coef[7, 2], 3)

# Isok 240 diff
isok240coef <- coef(summary(m3))
isok240diffp <- round(isok240coef[7, 5], 3)
isok240diffse <- round(isok240coef[7, 2], 3)

```

There were no significant differences in mean change of total session volume, training intensity or RPE between exercise training with glucose and placebo at any of the time points (Fig 5, p = 0.740, p = 0.917, p = 0.645, post change compared to baseline, respectively). We did observe interactions between time and changes, where total session training volume increased from 13-18% compared to baseline. These increases in training volume were accompanied by changes ranging from 5-18% in training intensity and 2-13% in RPE (Tab 5). The maximal isometric knee extension force (OR TORQUE?) and isokinetic knee extension peak torque at 240 d/s showed no differences between exercise training with glucose and placebo after five RT sessions (Fig 6A, Isometric: p = 0.336, 240 d/s: p = 0.527, post change compared to baseline, respectively). There was a significant difference in mean change from pre to post-fifth RT session, where exercise training with glucose led to a 9% less decrease in peak torque at 60 d/s compared to placebo (Fig 6A, p = 0.037). There was no difference between glucose and placebo in mean change from baseline to after the sixth RT session (Fig 6B, Isometric: p = 0.442, 60 d/s: p = 0.377, 240 d/s: p = 0.154, compared to baseline, respectively).

![](figures/legend.PNG)

```{r training dat fig}
### Get figures

# Volume
totvol.barplot <- readRDS("./data/data-gen/training/totvol.barplot.RDS")
totvolfold.plot <- readRDS("./data/data-gen/training/totvolfold.plot.RDS")
# Training intensity
tot.barplot <- readRDS("./data/data-gen/training/tot.barplot.RDS")
totalint.plot <- readRDS("./data/data-gen/training/trainint.plot.RDS")
# Join figures
plot_grid(totvol.barplot + theme(legend.position = "none"),
          totvolfold.plot + theme(legend.position = "none"),
          tot.barplot + theme(legend.position = "none"),
          totalint.plot + theme(legend.position = "none"),
          ncol = 2,
          nrow = 2,
          align = "h")

```

###### **Figure 6:** Training volume and training intensity from baseline until resistance exercise session 6 (training day 11 and 12). Baseline = Training day 1/2, Session 2 = Training day 3/4, Session 3 = Training day 5/6, Session 4 = Training day 7/8, Session 5 = Training day 9/10, Session 6 = Training day 11/12. A) Changes in total session volume per leg, values are mean kg ± standard deviation. B) Fold change in total session volume per leg, values are estimated marginal means fold change ± 95% CI. C) Changes in training intensity per leg calculated as %1RM, values are mean % ± standard deviation. D) Fold change in training intensity per leg, values are estimated marginal means fold change ± 95% CI. † = *p* < 0.05 in both glucose and placebo, compared to baseline. p = NS = no significant difference between legs at any of the time points. Glucose n = 13, placebo n = 13.

###### **Table 5:** RPE changes through training days. Training days are presented as combined pairwise consecutive, linked to their respective sessions. Session 1 = Training day 1/2, Session 2 = Training day 3/4, Session 3 = Training day 5/6, Session 4 = Training day 7/8, Session 5 = Training day 9/10, Session 6 = Training day 11/12. Values are presented as mean ± standard deviation. Glucose n = 13, placebo n = 13.

```{r RPE table}
# Scripts
# rpe-analysis
# Session score
ss.dat <- read_excel("./data/training/ribose_rpe.xlsx", na = "NA") %>%
        select(subject, timepoint, session.score, supplement) %>%
        mutate(time = if_else(timepoint %in% c("T1", "T2"),
                              "baseline",
                              if_else(timepoint %in% c("D3", "D4"),
                                      "session2",
                                      if_else(timepoint %in% c("D5", "D6"),
                                              "session3",
                                              if_else(timepoint %in% c("D7", "D8"),
                                                      "session4",
                                                      if_else(timepoint %in% c("D9", "D10"),
                                                              "session5",
                                                              if_else(timepoint %in% c("T3", "T4"),
                                                                      "post", timepoint))))))) %>%
        mutate(time = factor(time, levels = c("baseline", "session1", "session2", "session3", "session4", "session5", "post")),
               supplement = factor(supplement, levels = c("placebo", "glucose")))

## Mean and SD
# Session score
ss.exp <- ss.dat %>%
        select(subject, supplement, time, session.score) %>%
        group_by(supplement, time) %>%
        summarise(mean.ss = mean(session.score),
                  sd.ss = sd(session.score)) %>%
        mutate(stat = paste0(round(mean.ss, 1), " (", round(sd.ss, 1), ")")) %>%
        select(supplement, time, stat) %>%
        pivot_wider(names_from = time,
                    values_from = stat)

ss.exp %>%
  kable(col.names = c("Supplement", "Baseline", "Session 2", "Session 3", "Session 4", "Session 5",
                            "Session 6"))

```

![](figures/legend2.PNG)

```{r humac figs}
## Post 5th RE session
# Figures
pos <- position_dodge(width = 0.2)
isom5.plot <- readRDS("./data/data-gen/humac/isom.main.fig.RDS")
isok605.plot <- readRDS("./data/data-gen/humac/isok60.main.fig.RDS")
isok2405.plot <- readRDS("./data/data-gen/humac/isok240.main.fig.RDS")

legend <- get_legend(isom5.plot + theme(legend.box.margin = margin(0, 0, 0, 12)))

plot_grid(isom5.plot + theme(legend.position = "none"), 
          isok605.plot + theme(legend.position = "none"), 
          isok2405.plot + theme(legend.position = "none"),
          nrow = 3,
          align = "v")

## Post 6th RE session

# Figures

pos <- position_dodge(width = 0.2)

isom.plot <- readRDS("./data/data-gen/humac/isom.ac.plot.RDS")
isok60.plot <- readRDS("./data/data-gen/humac/isok60.ac.plot.RDS")
isok240.plot <- readRDS("./data/data-gen/humac/isok240.ac.plot.RDS")

legend <- get_legend(isom.plot + theme(legend.box.margin = margin(0, 0, 0, 12)))

plot_grid(isom.plot + theme(legend.position = "none"), 
          isok60.plot + theme(legend.position = "none"), 
          isok240.plot + theme(legend.position = "none"),
          nrow = 3,
          align = "v")

```

The mean change in peak torque at all three velocities change little from baseline until session 4 in both glucose and placebo (+4-7% change), decreasing at pre-session 6 (9-18% change, Fig 6A). Post session 6 (Fig 6B) showed a slight reduction in isometric peak torque 30min after session completion (2-4% change), slowly regaining at 2hrs (+3% change) and 23hrs (+3-6% change), in both glucose and placebo compared to baseline. Isokinetic peak torque at 60 d/s and 240 d/s showed no reduction in mean peak torque from baseline until 23hrs following RT,  in both glucose and placebo (+0,1-11%). There was an effect of time in peak torque at 23hrs post-RT at 60 d/s (Fig 6B), where exercise training with placebo led to a 5% increase in peak torque compared to baseline, however, there was no significant difference compared to exercise training with glucose.

###### **Figure 7:** Peak torque changes in maximal isometric knee extension and maximal isokinetic knee extension at 60 and 240 d/s, respectively. A) Humac tests during the intervention, Baseline = Day -1 (prior to intervention), Session 2 = prior to training, training Day 4/5, Session 4 = prior training, training Day 8/9, Session 6 = prior to training 11/12. B) Baseline = prior to training, training day 11/12, 30min = 30 min after RT, 2hr = 2hrs after RT, 23hr = 23hrs after RT. Values are change in estimated marginal means of newton meters (Nm) ± 95% CI. * = *p* < 0.05 between groups. *p* = NS = no significant difference between legs at any of the time points. Glucose n = 13, placebo n = 13.

## Discussion

The main findings of the present study were that exercise training with glucose for five sessions did not affect muscle biology markers, measured as total RNA and -protein, rRNA and specific protein, compared to RT with placebo. Similarly, RT with glucose did not affect markers of muscle functionality such as muscular recovery, total session volume, and training intensity, compared to RT with placebo. There are several studies suggesting high glucose treatment is important to markers of ribosome biogenesis (MARIAPPAN11, TANAKA18, ZHAI12), however, the perspective of exogenous glucose ingestion combined with RT on subsequent adaptations have remained unexplored. As a result, our hypotheses and purposes were quite exploratory and, considering previous suggestions, it is interesting that we did not observe any difference between RT with glucose and RT with placebo on the aforementioned markers of ribosome biogenesis.

To our knowledge, no previous study has investigated the effects of glucose compared to placebo ingestion with RT on ribosome biogenesis in human skeletal muscle cells. Through recent studies, the current model of ribosome biogenesis suggests that changes in the accumulation of total RNA and rRNA early after the onset of an RT-period predict long-term responses in RT adaptation (FIGUEIREDO19). Therefore, the lack of effects of glucose on total RNA and rRNA after two weeks of moderate volume RT do indicate that glucose per se does not induce positive long-term adaptations in muscle growth. First, the time frame of the intervention has previously served as a predictor of long-term responses, as recent studies have shown augmentation of ribosome biogenesis by total RNA and rRNA following multiple sessions of RT (FIGUEIREDO16; HAMMARSTROM20). Figueiredo et al. (2016) observed increases in 47S pre-rRNA measurable after a single RT session, passing baseline after 4hrs, peaking at 24hrs and remaining elevated until 48hrs post-exercise. Other mature rRNA where first measurable after multiple sessions (FIGUEIREDO16). In addition, Hammarström et al. (2020) demonstrated that the most prominent augmentation in ribosome biogenesis markers occurred 2 weeks following moderate-volume unilateral RT. Secondly, translational capacity is now regarded a vital factor in long-term RT-induced adaptations (IS IT?), as increased ribosome content provides a higher basal MPS, giving regulation of ribosome biogenesis a central role in muscle growth (FIGUEIREDO19). This is further substantiated by the association between blunted ribosome biogenesis and poor responses to RT (HAMMARSTROM20; STEC16), in addition to the lack of relationship between translational efficiency and long-term gains in skeletal muscle mass (FIGUEIREDO19). Hence, two weeks of moderate-volume unilateral RT should have been sufficient to elicit differences between exercise training with glucose and placebo. Supporting the notion that our intervention indeed was effective, we observed expected augmentation in accumulation of total RNA and expression of rRNA (CITE THIS?). As there were no differences in total RNA and rRNA between exercise training with glucose and placebo in the present study, it is therefore reasonable to suggest that there were no differences in ribosome biogenesis by proxy. Further, since glucose did not augment ribosome content after two weeks of RT compared to placebo, it is unlikely that glucose per se is beneficial to long-term adaptations in RT.

Indeed, our findings suggest that glucose ingestion per se before and after RT does not induce changes in the internal physiological milieu of human skeletal muscle cells with anabolic effects. It has been theorised that glucose may convey an anabolic signal through insulin, and thus operate as a signal mediator affecting upstream pathways of ribosome biogenesis (KUSNADI15; TANAKA18). Such an anabolic signal could then possibly increase PIC assembly, and consequently Pol I recruitment, initiating rDNA transcription and subsequent rRNA maturation, leading to increased ribosome content (FIGUEIREDO19; TANAKA18). INVESTIGATE THE INSULIN DATA AND COMMENT IT HERE. Another theory was that glucose ingestion would impact the cellular energy levels and thus, the energy-sensitive PIH1, mTORC1, ERK1/2, AMPK, and SIRT1 pathways (MARIAPPAN11; KIM13; TANAKA18; ZHAI12). The mammalian target of rapamycin, along with ERK1/2 and PIH1 up-regulates PIC assembly in high-energy conditions, while AMPK and possibly SIRT1 represses PIC assembly in low-energy conditions, based on in vitro rodent, yeast, and non-muscle cell cultures (MARIAPPAN11; KIM13; TANAKA18; ZHAI12; HOPPE09; MURAYAMA08). Specifically, high glucose treatment augmented rDNA transcription by increased UBF and SL-1 activity, while low glucose treatment, glucose starvation, or nutrient depletion stimulated AMPK-dependent TIF-IA and chromatin regulation suppression (KOS-BRAUN17; MARIAPPAN11; ZHAI12). In addition to UBF and SL-1, assembly of the PIC requires TIF-IA and Pol I recruitment, induced by amino acid intake in an mTOR-S6K-dependent manner and possibly by direct c-Myc activation (KUSNADI15). COMMENT ON PROTEIN ANALYSIS RESULTS HERE? As such, increasing energy supply and nutrient availability should in line with these theories potently stimulate PIC assembly and rDNA transcription. Along this notion, it does not seem apparent that the timing of glucose ingestion, for instance immediately before and after RT affects these processes, as glucose did not induce acute augmentation of ribosome biogenesis, despite the significant elevation of plasma glucose levels when exercising with glucose compared to placebo. Based on our observations, glucose per se has no acute effect on ribosome biogenesis, as the same effects can be observed with equal glucose ingestion 5-6hrs later.

Another possible explanation for the lack of effect of glucose compared to placebo, is that both treatments may have exercised with sufficient levels of nutrients and energy, meaning that mTORC1 had sufficient nutrient availability and, thus was not inhibited by AMPK due to energy deficiency (HOPPE09; KUSNADI15). Indeed, mTORC1 is characterised as energy-sensitive (DENNIS01), and therefore it might have been directly affected by high vs. low-glucose conditions (MARIAPPAN11). However, previous studies suggest that mTORC1 is not sensitive enough to register levels of ATP, rather depending on signals from AMPK to assess ATP levels (KIM10). This would arguably make AMPK the primary regulator of the potential impact of glucose on ribosome biogenesis, and thus, glucose has no direct effect on mTORC1 signalling (REFERENCES?). Importantly, intramuscular glycogen content was not measured in the present study, therefore we cannot determine whether cellular energy levels in fact were equal or not, or if there is reason to assume that AMPK-dependent mTORC1 signalling was different in the two treatments. Potentially, such data could give higher resolution to our data. Along with the robust increases in proxy markers of ribosome biogenesis, the equal macronutrient intake between treatments on pairwise consecutive days and plasma glucose levels not indicating energy deficiency during RT, potent mTORC1 stimulation without AMPK-dependent inhibition may be a plausible explanation. Together with previous studies on this topic (MARIAPPAN11; ZHAI12), there are currently scarce observations to suggest glucose carries a direct anabolic signal. As such, physiological levels of plasma glucose during RT, combined with a sufficient daily macronutrient intake, seem potent enough to supply the energy needed for ribosome biogenesis. The lacking differences between treatments in the biological data are further supported by our observations on functional data.

Overall, glucose did not affect the isometric or isokinetic peak torque at 240 d/s, or muscular recovery following the sixth and last RT session, measured as temporal changes in isometric and isokinetic peak torque. Generally, the mean isometric and isokinetic peak torque slightly decreased during the intervention, especially between session 4 and 6, and regained after the intervention. Further, there were no differences between glucose and placebo in total session volume or training intensity. Both exercise with glucose and placebo yielded higher total session volumes at higher percentage of 1RM at the last session compared to baseline, as expected from earlier reports on the moderate-volume RT protocol (KRIEGER09; RHEA03; SCHOENFELD17), in addition to a more recent study implementing the unilateral protocol (HAMMARSTROM20+21/22). Interestingly, we did observe a significant difference between treatments in mean change of isokinetic peak torque at 60 d/s before the sixth RT session, where exercise with glucose decreased peak torque 9% less compared to placebo. However, as none of the other measurements indicated any difference between treatments, a true effect of glucose seems unlikely. Taken together, these results may add to previous observations with no differences between co-ingestion of carbohydrate and protein compared to protein alone, in regard to skeletal muscular recovery following RT (KOOPMAN07; STAPLES11). Collectively, the functional data point towards an exhaustive, yet growth-inducing RT protocol, especially considering the biological data. As a side note, an interesting observation was made by another master’s thesis on this project, finding significantly lower MuRF protein levels at post with glucose treatment compared to placebo. This may suggest an effect of glucose on protein degradation (WILKES09), however, whether this contributes to growth or simply protein turnover has been a topic of debate (FIGUEIREDO19).

Despite being a thoroughly standardised design, a limitation may be posed through lack of participants. We initially reach our minimum limit for statistical power, i.e. 16 participants, however, three dropped out during the intervention, possibly impacting our power. Coupled with the fact that no previous study could guide our power calculations, this, of course, introduces some degree of uncertainty regarding our results. However, we’ve reported little to suggest any true acute effect of RT with glucose ingestion compared to placebo. Moreover, the study was specifically designed to investigate differences in pre-post changes within participants, and thus also quite limited to this perspective. Hence, the effects of time observed exclusively serve as supplementary to the efficiency of the design, without a negative control group to either confirm or refute them. If further investigations are to be done on this topic, these analyses should include more participants and optimally a measure of intramuscular glycogen levels.

In conclusion, resistance exercise training combined with glucose ingestion does not acutely augment ribosome biogenesis following 2 weeks of moderate-volume resistance training in moderately trained young male and female adults. Neither does it acutely affect muscular performance in isometric and isokinetic knee extension peak torque, nor enhance post resistance training recovery within 24hrs.

## Disadvantages due to Covid-19

Due to restrictions regarding social distancing and lock downs, caused by the Covid-19 pandemic, we were not able to recruit and include as many participants as we planned to the study.

### 2023-02-01 R/dxa.R

@@ -1,0 +1,168 @@
+## Analysis of baseline DXA results.
+
+# Author: SCM
+
+# This script analyses the baseline DXA values and creates a table
+
+# Packages
+
+library(readxl)
+library(tidyverse)
+library(nlme)
+library(lme4)
+library(broom)
+library(knitr)
+library(emmeans)
+library(knitr)
+
+#1 descritpive table of subject
+#
+#count gender to be put in table
+gendercount <- read_excel("data/dxa/ribose_dxa.xlsx") %>%
+  count(sex) %>%
+  pivot_wider(names_from = sex,
+              values_from = n) %>%
+  print()
+
+
+#Calculates grams to kg
+table.dxa <- read_excel("data/dxa/ribose_dxa.xlsx") %>%
+  select(sex, subject, age, height, weight, leanmasskg, fatmasskg, ffmg) %>%
+  #make grams into kg
+  mutate(ffkg = ffmg/1000) %>%
+  select(sex, age, height, weight, leanmasskg, fatmasskg, ffkg) %>%
+  print()
+
+#dxa table pr. gender               
+dxa.table <- table.dxa %>%
+  pivot_longer(names_to = "variable",
+               values_to = "values", col= age:ffkg) %>%
+  select(sex, variable, values) %>%
+  group_by(sex, variable) %>%
+  summarise(m = mean(values, na.rm = TRUE),
+            s= sd(values, na.rm = TRUE)) %>%
+  mutate(stat = paste0(round(m, 1), " (", round(s, 1), ")")) %>%
+  select(sex, variable, stat) %>%
+  pivot_wider(names_from = sex,
+              values_from = stat) %>%
+  mutate(variable = if_else(variable %in% c("age"),
+                            "Age",
+                            if_else(variable %in% c("fatmasskg"),
+                                    "Fat mass",
+                                    if_else(variable %in% c("ffkg"),
+                                            "Fat free mass",
+                                            if_else(variable %in% c("height"),
+                                                    "Height",
+                                                    if_else(variable %in% c("leanmasskg"),
+                                                            "Lean mass",
+                                                            if_else(variable %in% c("weight"),
+                                                                    "Weight",
+                                                                    ""))))))) %>%
+  print()
+
+
+dxa.table %>%
[…]

### 2023-02-01 R/volume.change.R

@@ -1,0 +1,158 @@
+### Total volume change, both absolute values and fold-change
+
+# This script provides log-fold change score calculation of training volume, and illustration of absolute and fold change in training volume. 
+# Absolute changes are presented as mean change ± SD per supplement, fold changes are represented as mean change ± CI. 
+
+# Packages
+library(readxl);library(tidyverse);library(knitr);library(lme4);library(broom);library(emmeans)
+
+# Data
+tot.vol <- read_excel("./data/training/ribose_volume.xlsx", na = "NA") %>%
+  select(subject, timepoint, tot.volume, supplement) 
+
+## Handling the data by creating a new factor called time from timepoint. This factor combines any observation at T1 and T2 to baseline, etc. 
+# The code also sorts the order of the factor time, from baseline to session 6, using time = factor(time, levels c()), and sets placebo to be compared to 
+# glucose via supplement = factor(supplement, levels = c()).
+
+tot.volh <- tot.vol %>%
+  mutate(time = if_else(timepoint %in% c("T1", "T2"),
+                        "baseline",
+                        if_else(timepoint %in% c("D3", "D4"),
+                                "session2",
+                                if_else(timepoint %in% c("D5", "D6"),
+                                        "session3",
+                                        if_else(timepoint %in% c("D7", "D8"),
+                                                "session4",
+                                                if_else(timepoint %in% c("D9", "D10"),
+                                                        "session5",
+                                                        if_else(timepoint %in% c("T3", "T4"),
+                                                                "session6", timepoint))))))) %>%
+  mutate(time = factor(time, levels = c("baseline", "session1", "session2", "session3", "session4", "session5", "session6")),
+         supplement = factor(supplement, levels = c("placebo", "glucose")))
+
+## Baseline analysis - comparison of the two legs
+# A baseline analysis comparing training volume at baseline sessions between the two legs via a paired t.test, and providing a summary of mean training
+# volume and sd
+
+base.vol <- tot.volh %>%
+  filter(time == "baseline") %>%
+  select(subject, time, supplement, tot.volume) %>%
[…]

### 2023-02-01 R/western.change.R

@@ -36,7 +36,6 @@ pchange <- prot.dat %>%
          supplement = factor(supplement, levels = c("PLACEBO", "GLUCOSE"))) %>%
   print()
 
-
 # By target
 
 cmyc <- pchange %>%
@@ -53,23 +52,24 @@ rps6 <- pchange %>%
 
 # c-Myc change
 
-
-m1 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
-                     data = cmyc)
+m1 <- lme(change ~ pre + supplement,
+            random = list(subject = ~ 1),
+            data = cmyc) # same model as in "western import", but with change scores of norm.sign
 
 plot(m1)
 
 summary(m1)
 
 cmyc.emm <- emmeans(m1, specs = ~"supplement") %>%
-  data.frame()
+  data.frame
 
 saveRDS(cmyc.emm, "./data/data-gen/protein/cmyc.change.RDS")
 
 # UBF change
 
-m2 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
-                     data = ubf)
+m2 <- lme(change ~ pre + supplement,
+            random = list(subject = ~ 1),
+            data = ubf) # same model as in "western import", but with change (i.e., change in norm.sign) instead of norm.sign
 
 plot(m2)
 
@@ -82,8 +82,9 @@ saveRDS(ubf.emm, "./data/data-gen/protein/ubf.change.RDS")
 
 # rps6 change
 
-m3 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
-                     data = rps6)
+m3 <- lme(change ~ pre + supplement,
+            random = list(subject = ~ 1),
+            data = rps6) # same model as in "western import", but with change (i.e., change in norm.sign) instead of norm.sign
 
 plot(m3)
 

### 2023-02-01 R/western.change.figs.R

@@ -20,7 +20,7 @@ rps6 <- readRDS("./data/data-gen/protein/rps6.change.RDS")
 
 ## cmyc fig
 
-cmyc %>%
+cmyc.plot <- cmyc %>%
   mutate(time = "post") %>%
   add_row(supplement = c("PLACEBO", "GLUCOSE"), time = "pre", emmean = 0, lower.CL = 0, upper.CL = 0) %>%
   mutate(time = factor(time, levels = c("pre", "post"), labels = c("Baseline", "Post"))) %>%
@@ -35,7 +35,7 @@ cmyc %>%
   theme_classic()
 
 
-ubf %>%
+ubf.plot <- ubf %>%
   mutate(time = "post") %>%
   add_row(supplement = c("PLACEBO", "GLUCOSE"), time = "pre", emmean = 0, lower.CL = 0, upper.CL = 0) %>%
   mutate(time = factor(time, levels = c("pre", "post"), labels = c("Baseline", "Post"))) %>%
@@ -49,7 +49,7 @@ ubf %>%
   labs(x = "", y = "UBF signal \n(fold change)\n", fill = "") +
   theme_classic()
 
-rps6 %>%
+rps6.plot <- rps6 %>%
   mutate(time = "post") %>%
   add_row(supplement = c("PLACEBO", "GLUCOSE"), time = "pre", emmean = 0, lower.CL = 0, upper.CL = 0) %>%
   mutate(time = factor(time, levels = c("pre", "post"), labels = c("Baseline", "Post"))) %>%
@@ -64,3 +64,15 @@ rps6 %>%
   theme_classic()
 
 
+# Cowplot for gathering figures
+
+legend <- get_legend(rps6.plot + theme(legend.box.margin = margin(0, 0, 0,12)))
+
+
+prot.joined <- plot_grid(cmyc.plot + theme(legend.position = "none"),
+                         ubf.plot + theme(legend.position = "none"), 
+                         rps6.plot + theme(legend.position = "none"),
+                         legend,
+                         ncol = 2, nrow = 3)
+
+saveRDS(prot.joined, "./data/data-gen/rna/prot.joined.RDS")

### 2023-02-02 ribose-paper.RMD
updated markdown

@@
 ```{r packages}
-library(tidyverse);library(readxl);library(lme4);library(emmeans);library(dplyr);library(cowplot);library(knitr);library(broom);library(rmarkdown);library(tinytex);library(tidyr);library(arsenal);library(magrittr);library(dabestr)
+library(tidyverse);library(readxl);library(lme4);library(emmeans);library(dplyr);library(cowplot);library(knitr);library(broom);library(rmarkdown);library(tinytex);library(tidyr);library(arsenal);library(magrittr);library(dabestr);library(nlme)
 ```
@@
 ### Materials and methods
+
 All participants were informed about the potential discomforts and risks associated with the study and gave their informed consent before study enrollment.
@@
 ### Participants
+
 Sixteen healthy male and female participants (20-33yrs, Tab 1) were recruited to the study through social media advertisement and word of mouth, and through completing the selection process (Fig 1).
@@
 ### Experimental design
+
 The study was designed as a 12-day double-blinded randomized controlled trial, with a unilateral RT protocol (Fig 2).
@@
 ### Dietary intervention
+
 The dietary intervention spanned the whole day, divided into three periods: I) From awakening until two hours after RT, II) from 2.5 hours after RT until 2200hrs, and III) from 2200hrs until awakening.
@@
 ### Assessment of muscular strength
+
 Performance tests (Test) were performed prior to (Days -7 and -5, and T0, both legs) and during the intervention (Days 4, 5, 8, and 9), after session 5 and after finalization of the intervention (T3/T4, RT#1 leg; T4/Day 13, RT#2 leg), a total of 10 days (Fig 2).
@@
 ### Resistance training protocol
+
 Resistance training consisted of three sets of each unilateral leg presses and knee extensions, with an intensity of 10 repetition maximum (10RM).
@@
 ### Sampling of muscle tissue and blood
+
 Muscle biopsies were sampled from m.
@@
 To standardize this procedure, all individual participants had biopsies taken at the same time of day +/- 1hr, in an overnight fasted state.
+
 To measure blood glucose levels with and without glucose ingestion/training, blood was collected by means of finger draws on days with biopsy sampling.
@@
 
-### Total RNA and protein extraction
+### Total RNA extraction and quantitative real-time reverse transcription polymerase chain reaction
+
 Two muscle biopsy samples were taken for total RNA extraction per leg per time point, and total RNA was extracted in two duplicates per muscle biopsy.
@@
 Total RNA was extracted using TRIzol according to the manufacturer’s protocol.
-Briefly, muscle tissue was homogenized in 300µl TRIzol using 0.5mm RNase-free Zirconium beads (~50 ul) (Next Advanced, Averill Park, NY, USA) and a bullet blender (bb).
-Samples ran 1min at speed 10, put on ice for 1min and bb another round at speed 12 for 1 min.
-If not fully dissolved, samples ran additional rounds until fully homogenized.
-Thereafter, an additional 700 µl TRIzol was added, and samples ran 1 min at speed 3 in the bb for mixing, before 5 min incubation at room temperature.
-200 µl chloroform (Sigma-Aldrich, Missouri, USA) was added, and samples were shaken for 15sec, followed by 2-3min incubation at room temperature.
-After incubation, samples were inverted by hand and centrifuged in Heraeus™ Fresco™ 21 Microcentrifuge (Thermo Fisher Scientific, Waltham, Massachusetts, USA) at 12000g, 15min, 4°C.
-The upper aqueous phase (450 µl) was transferred to a fresh tube, containing 500 µl isopropanol (VWR International, Pennsylvania, USA), mixed, and incubated for 10min at room temperature.
-Samples were centrifuged at 12000g for 10min at 4°C, precipitating an RNA pellet.
-The pellet was washed twice with 1000 µl 75% cold ethanol; centrifuged for 5 min at 7500g, 4°C.
-Thereafter, ethanol was removed, and the pellet was air dried for 10min (or until dried).
-The pellet was suspended in 30 µl DEPC-treated water and incubated for 10 min at 55°C.
-For assessment of RNA content and -purity 5µl RNA, from the RNA stock, was diluted with 5µl TE-buffer (1:2), for assessment via spectrophotometry.
+An exogenous RNA control (λ polyA External Standard Kit, Takara Bio Inc., Shiga, Japan) was added at a fixed amount (0.04 ng ml-1 of Trizol reagent) to enable analysis of target gene expression per unit tissue weight (ELLEFSEN2008,2014).
+Briefly, muscle tissue was homogenized in TRIzol using 0.5mm RNase-free Zirconium beads (~50 ul) (Next Advanced, Averill Park, NY, USA) and a bullet blender (bb).
+Chloroform (Sigma-Aldrich, Missouri, USA) was used for phase-separation, and an RNA pellet was precipitated with isopropanol (VWR International, Pennsylvania, USA).
+For assessment of RNA content and -purity 5µl RNA, from the RNA stock, was eluted in TE-buffer (1:2), for assessment via spectrophotometry.
 All samples had a 260nm to 280nm ratio > 1.9.
 The RNA stock was stored at -80°C until further analyses.
-
-Total protein was extracted using the Minute Total Protein Extraction Kit for Muscles (Invent Biotechnology), according to the manufacturer’s protocol.
-Five to twenty mg (CHECK) of wet muscle was freeze-dried for 24hrs and dissected before extraction.
-The tissue was homogenized in approximately 80mg (CHECK) protein extraction powder and 100ul icy cold denaturing buffer (Invent Biotechnology) with a plastic rod, and centrifuged at top speed for 1 min.
-The supernatant was transferred to a fresh tube, and then aliquoted into 2x25ul tubes for analysis.
-Four microliter total protein was diluted in 36ul DEPC-treated water (or nuclease-free??) (1:10), for total protein measurement via spectrophotometry with the Bradford reagent (…).
-
-### Complimentary DNA synthesis and Quantitative polymerase chain reaction
-
 RNA was reverse transcribed using Super Script IV Reverse Transcriptase (Invitrogen), according to the manufacturer’s instructions.
@@
 
-### Gel electrophoresis and Western Blot
-
-#### Gel electrophoresis
-Samples were standardized to a protein concentration of 3ug/ul via dilution with denaturing buffer (Invent Biotechnology), and laemmli/2-mercaptoethanol loading buffer (PRODUCER?) equating to ¼ of the sample’s total volume was then added to reach a final protein concentration of 2.25 ug/ul.
-A total of 20ul from each sample was pooled in several fresh tubes, serving as a normalisation factor in later quantitative analyses.
-All samples were incubated at 95°C for 5 minutes, briefly cooled and stored at -20°C until loading.
-In total, 9ul from each sample were loaded on six gels in duplicates, with 9ul of 3-4 pools per gel, on ice.
-Loading order from gel 1-3 was inverted for gel 4-6.
-Gels ran the electrophoresis simultaneously, at 250V constant for 50min on ice, ice-cold running buffer (250 mM Tris, 1.92 M Glycin, 1% SDS.
-Final concentrations in 1X: 25 mM Tris, 192 mM Glycin, 0.1% SDS).
-
-#### Wet transfer
-Transfer buffer (25 mM Tris, 192 mM Glycin, 10% MeOH) was made in due time of the transfer (2-5 days), to allow for “de-gas”.
-The PVDF-membrane was incubated for 5min in MeOH, then 10min in cold transfer buffer before transfer.
-Before sandwich assembly, foam pads and filter paper were soaked in transfer buffer, and the gel was flushed with ddH2O.
-The wet transfer was run at 300mA constant for 3hrs, in a Criterion Blotter (PRODUCER) on ice and stirring pad to maintain even buffer temperature and ion concentration during the transfer.
-
-#### Staining
-MemCode was used for protein staining, according to manufacturer’s protocol.
-The membranes were soaked and carefully shaken in MemCode Stain for 1min, washed with Destain/MeOH (1:1) on a shaker for 5min, with ddH2O washing before and between these steps.
-One manual capture (G:Box) per gel was taken, to procure total protein stain images.
-Following manual captures, all membranes were sliced into strips containing the bands of the specific protein targets (c-Myc, UBF, rps6), rinsed with ddH2O, washed with Eraser/MeOH solution (1:1) for 10min and rinsed with ddH2O once more.
-A 5% non-fat dry milk solution was used as blocking buffer and diluent for antibodies.
-Each membrane was blocked for 1hr at room temp, and washed once with TBS-T for 5min on a tipping board, then incubated overnight on a tipping board at 4°C in primary antibody (see table ..
-for primary antibody dilutions).
-Secondary antibody staining was done on a tipping board for 1hr at room temp (see table ..
-for secondary antibody dilutions), with three subsequent 5min TBS-T washes.
-Lastly, before imaging, the membranes were incubated for 5min in ECL (TYPE).
-Quantification was done using the G: Box and Genesys software with an ECL protocol.
-
-### Data handling and statistical analyses
-A priori power calculations showed that 20 participants would be sufficient to grant a statistical power of 80% (=0.05), with an expected dropout of 20%.
-As no studies existed to guide our power calculations, we made the assumption that the effects of glucose ingestion on ribosome biogenesis may equate to the effects of increased resistance training volume from low to moderate (HAMMARSTRÖM20).
-All raw data were imported and analysed in R (TEAM18).
-Individual participants’ leg results were categorized as either glucose or placebo according to the randomization.
-Total RNA and rRNA were analyzed by post to pre log-fold change score difference comparisons, using a linear mixed effects model with mean-centred baseline values as a covariate and subsequent calculations of estimated marginal means.
-Total RNA was normalised by muscle biopsy weight, and qPCR values were normlised by muscle biopsy weight and external reference gene (Lambda).
-Total RNA and rRNA changes were calculated as log-fold change scores per mg wet muscle weight.
-Blood glucose levels, dietary data, training volume and -intensity, RPE, and muscular strength and -recovery were analysed by multiple time point log-fold change score comparisons, using the same linear mixed effects model, adding time as an explaining factor in addition to supplement.
-The linear mixed effects model was design with the lmer function of the lme4 package (BATES14) in R, using the lmerTest function in addition to procure p-values.
-A paired samples t-test was used to check for differences in leg characteristics at baseline.
-All data was log-transformed for analysis to control for heteroscedasticity.
-For figure illustration, values were either reverse transformed or calculated as absolute changes.
-Mean-centred baseline values were used to correct for regression to the mean, which potentially occurs when the same participants are repeatedly tested.
-As such, varying baseline values should not affect change outcomes.
-Estimated marginal means were calculated from the linear model, enabling acquisition of least-square means, showing the means for all involved groups adjusted to means of other potential factors in the model.
-Values are presented as mean ± SEM unless otherwise stated.
-Alpha-level was set to α = 0.05
+### Protein extraction and immunoblotting
+
+Total protein was extracted using the Minute Total Protein Extraction Kit for Muscles (Invent Biotechnology), according to the manufacturer’s protocol, optimised for our lab.
+Wet muscle was freeze-dried for 24hrs and dissected before extraction.
+The tissue was homogenized with a plastic rod in 80mg protein extraction powder (Invent Biotechnology) and 100ul iced cell lysis buffer (Denaturing Buffer, Invent Biotechnology), and centrifuged at top speed for 1 min.
+The supernatant was aliquoted into 2x25ul tubes to run samples in duplicates, and total protein concentrations were determined in a 1:10 dilution (Pierce Detergent Compatible Bradford Assay Reagent, Thermo Fisher Scientific).
+The protein samples were diluted to 20.25 µg µl−1 total protein with lysis buffer and 4X Laemmli sample buffer (Bio-Rad Laboratories AB, Oslo, Norway) containing 2-mercaptoethanol.
+Twenty microlitres from each protein sample were pooled as a normalisation factor for gel separation.
+All protein samples were incubated at 95°C and stored at -20°C until further analysis.
+The protein samples (20.25µg total protein) were separated at 250V for 50min using ….
+gels (……..), then transferred to PVDF membranes with wet transfer at 300mA for 3hrs.
+Both gel electrophoresis and protein transfer were performed at 4°C.
+Following the wet transfer, membranes were stained with a MemCode reversible total protein stain (MANUFACTURER) and then blocked for 1hr at room temp with a blocking buffer of Tris-buffered saline (TBS; 20 mM Tris, 150 mM NaCl) with 5% non-fat dry milk and 0.1% Tween-20.
+Primary antibodies were purchased from ……, secondary antibodies were purchased from ……, and were diluted in a 5% non-fat dry milk blocking buffer (TBS, 5% non-fat dry milk, 0.1% Tween-20).
+Membranes were incubated overnight with primary antibodies, and for 1hr with secondary antibodies.
+Between blocking and primary antibody staining, membranes were washed 1x5min, between primary and secondary staining, and after secondary staining, membranes were washed 3x5 with TBS-Tween (TBS; 20mM Tris, 150mM NaCl, 0.1% Tween).
+Following the last TBS-T wash, membranes were incubated 1x5min with enhanced chemiluminescent substrate (ECL, SuperSignal West Femto Maximum Sensitivity Substrate, Thermo Fisher Scientific).
+Membrane blocking, secondary antibody incubation, washing and ECL incubation were performed at room temp, on a tipping board.
+Primary antibody incubation was performed at 4°C on a tipping board.
+Chemiluminescence signals were quantified using Image Studio Lite (LI-COR Biotechnology, Lincoln, NE, USA), and total protein content was quantified using ImageJ (RUEDENETAL2017), whereas total protein content was defined as mean grey value of the whole well with between-well values subtracted as background.
 
@@
 Further, on pairwise consecutive days, there were no significant differences in any macronutrient (calories, carbohydrate, fat or protein) ingestion (Tab 4).
-Resistance training with glucose led to significant increases in plasma glucose levels compared to RT with placebo, before, during and after the session (Fig 3B, 120min: 43 ± 4% [p = 0.00000], 135min: 32 ± 4% [p = 0.00000], 150min: 32 ± 4% [p = 0.00000]), measured during the final RT session.
-At 2hrs post-exercise, exercise training with glucose led to significantly lower plasma glucose levels compared to exercise training with placebo (8 ± 4%, p = 0.03).
+Resistance training with glucose led to significant increases in blood glucose levels compared to RT with placebo, before, during and after the session (fig 3B, 120min: 43 ± 4% [*p* = `r p.120`], 135min: 32 ± 4% [*p* = `r p.135`], 150min: 32 ± 4% [*p* = `r p.150`]), measured during the final RT session.
+At 2hrs post-exercise, RT with glucose led to significantly lower blood glucose levels compared to RT with placebo (8 ± 4% [*p* = `r p.270`).
 
@@
 Variables: Iso 60 = isokinetic 60 d/s, Iso 240 = isokinetic 240 d/s, training intensity = load/rm*100.
-Nm = newton meters, kg = kilograms, %1RM = percentage of 1 repetition maximum, ng/mg = nanogram per milligram wet muscle tissue, nf.expr = expression normalized by weight and external reference gene.
-rRNA data are presented as log transformed.
+Nm = newton meters, kg = kilograms, %1RM = percentage of 1 repetition maximum, ng/mg = nanogram per milligram wet muscle tissue, nf.expr = expression normalized by weight and external reference gene (Lambda).
+Ribosomal RNA data are presented as log transformed.
 Glucose n = 13, placebo n = 13.
@@
 ###### **Table 4:** Daily mean dietary intake of macro nutrients during each training day per supplement leg.
-Glu = glucose, Plac = placebo.
+G = glucose, P = placebo.
 Training day 1-training day 12 are combined pairwise to day 1-6: Day 1 = training day 1/2, Day 2 = training day 3/4, Day 3 = training day 5/6, Day 4 = training day 7/8, Day 5 = training day 9/10, Day 6 = training day 11/12.
 Values are mean ± standard deviation.
-P = difference between Glu and Plac.
+P = difference between glucose and placebo.
 Glucose n = 12, placebo n = 12.
@@
 
-###### **Figure 4:** A) Time line for training day 6 (combination of pairwise consecutive training day 11 and 12).
+###### **Figure 3:** A) Time line for training day 6 (combination of pairwise training day 11 and 12).
 Biopsy = muscle biopsy taken pre supplementation, PRO = protein supplementation, Humac = strength test (isometric knee extension, isokinetic knee extension 60 and 240 d/s), GLU/PLAC = glucose/placebo supplementation, RT = resistance training.
 Minutes: -120 = 2hrs prior to RT, -90 = 90min prior to RT, -30 = 30min prior to RT, 0 = onset of RT, 15 = 15min following onset of RT, between leg press and knee extension, 30 = 30min post onset of RT, 120 = 2hrs following onset of RT.
-B) Mean mmol/L change in plasma glucose levels per supplement group during training day 6.
+B) Mean mmol/L change in blood glucose levels per supplement group during training day 6.
 Time-points: 0, 45, 90, 120, 135, 150 and 270min post protein ingestion, respectively.
-Finger stings were taken at all time-points, blood samples were taken at all time-points except from 135.
+Finger draws were taken at all time-points, blood samples were taken at all time-points except from 135.
 Values are presented as estimated marginal means of mmol/L change ± 95% CI.
@@
 
-```{r rRNA data modelling + fig}
+```{r rRNA data modelling}
 # Scripts
@@
 rr47diffp <- round(rr47coef[3,5], 3)
+rr47tdiffp <- round(rr47coef[1,5], 3)
 rr47diffse <- round(rr47coef[3,2], 3)
@@
 
-Resistance training with glucose did not induce a higher accumulation of total RNA (Fig 4A, p = 0.499), or rRNA (Fig 4B, 18S: p = 0.584, 28S: p = 0.740, 5.8S: p = 0.935, 5S: p = 0.790, 47S: p = 0.502), compared to RT with placebo, measured as mean change from baseline to post.
-From baseline to post-intervention, there was an average increase in levels of total RNA, 26 and 22% after RT with glucose and placebo, respectively.
-A robust accumulation was also observed in ribosomal RNA expression in both RT with glucose and placebo, with mean increases between 34-43% (GLU) and 33-41% (PLAC) in the four rRNA’s, and 37 and 59% in 47S pre-rRNA, respectively.
+Resistance training with glucose did not induce a higher accumulation of total RNA (fig 4A, *p* = `r trdiffp`), or rRNA  (fig 4B, *p* > 0.05), compared to RT with placebo, measured as mean change from baseline to post.
+From baseline to post-intervention, there was a mean increase in levels of total RNA by 26 and 22% after RT with glucose and placebo, respectively (*p* < 0.05, compared to baseline).
+A robust accumulation was also observed in ribosomal RNA expression in both RT with glucose and placebo, with mean increases between 34-43% (GLU) and 33-41% (PLAC) (*p* < 0.05, compared to baseline) in the four rRNA’s.
+The expression of 47S pre-rRNA increased by 37 and 59% with glucose and placebo respectively, with only placebo reaching significantly higher levels at post compared to baseline (fig 4B, *p* = `r rr47tdiffp`).
 
@@
 
-###### **Figure 5:** A) Changes in Total RNA, B) Changes in ribosomal RNA.
+###### **Figure 4:** A) Changes in Total RNA, B) Changes in ribosomal RNA.
 Baseline = Training day 1/2, Post = Training day 11/12.
@@
 Mean change scores of the duplicates were calculated before modelling and transformed to the log-scale.
-Values are estimated marginal means fold change per leg ± 95% CI.
+Values are estimated marginal means fold change per leg/supplement ± 95% CI.
+p = between groups statistic, glucose compared to placebo.
+Glucose n = 13, placebo n = 13.
+
+## Protein signals
+
+```{r western modelling}
+# Script: western.change
+
+# Data
+
+prot.dat <- readRDS("./data/data-gen/protein/prot.dat.RDS")
+
+# Change analysis
+
+#prot.dat %>%
+ # group_by(time, supplement, target) %>%
+  #mutate(norm.sign = log(tpl/norm.sign)) %>% # Log transforms and normalises
+  # norm.sign by tpl
+  #summarise(m = mean(norm.sign, na.rm = TRUE)) %>%
+  #ggplot(aes(time, m, color = supplement, group = supplement)) +
+  #geom_line() +
+  #geom_point() +
+  #facet_wrap( ~ target, scales = "free")
+
+# Change score
+
+pchange <- prot.dat %>%
+  dplyr::select(subject, time, sample.id, target, norm.sign, supplement) %>%
+  group_by(subject, time, target, supplement) %>%
+  summarise(mns = mean(norm.sign, na.rm = TRUE)) %>%
+  pivot_wider(names_from = time,
+              values_from = mns) %>%
+  ungroup() %>%
+  mutate(change = log(post)-log(pre),
+         pre = pre - mean(pre, na.rm = TRUE),
+         supplement = factor(supplement, levels = c("PLACEBO", "GLUCOSE")))
+
+# By target
+
+cmyc <- pchange %>%
+  filter(target == "cmyc")
+
+ubf <- pchange %>%
+  filter(target == "ubf")
+
+rps6 <- pchange %>%
+  filter(target == "rps6")
+
+# c-Myc change
+
+m1 <- lme(change ~ pre + supplement,
+            random = list(subject = ~ 1),
+            data = cmyc) # same model as in "western import", but with change scores of norm.sign
+
+#plot(m1)
+
+sum.m1 <- summary(m1)
+
+cmyc.emm <- emmeans(m1, specs = ~"supplement") %>%
+  data.frame
+
+saveRDS(cmyc.emm, "./data/data-gen/protein/cmyc.change.RDS")
+
+# UBF change
+
+m2 <- lme(change ~ pre + supplement,
+            random = list(subject = ~ 1),
+            data = ubf) # same model as in "western import", but with change (i.e., change in norm.sign) instead of norm.sign
+
+#plot(m2)
+
+sum.m2 <- summary(m2)
+
+ubf.emm <- emmeans(m2, specs = ~"supplement") %>%
+  data.frame()
+
+saveRDS(ubf.emm, "./data/data-gen/protein/ubf.change.RDS")
+
+# rps6 change
+
+m3 <- lme(change ~ pre + supplement,
+            random = list(subject = ~ 1),
+            data = rps6) # same model as in "western import", but with change (i.e., change in norm.sign) instead of norm.sign
+
+#plot(m3)
+
+sum.m3 <- summary(m3)
+
+rps6.emm <- emmeans(m3, specs = ~"supplement") %>%
+  data.frame()
+
+saveRDS(rps6.emm, "./data/data-gen/protein/rps6.change.RDS")
+
+### For inline reporting
+## Diff cmyc
+ccoef <- coef(summary(m1))
+cdiff <- round(ccoef[3,5], 3)
+cdiffse <- round(ccoef[3,2], 3)
+
+## Diff UBF
+ucoef <- coef(summary(m2))
+udiff <- round(ucoef[3,5], 3)
+udiffse <- round(ucoef[3,2], 3)
+
+## Diff rps6
+rpcoef <- coef(summary(m3))
+rpdiff <- round(rpcoef[3,5], 3)
+rpdiffse <- round(rpcoef[3,2], 3)
+
+```
+
+```{r western fig}
+# Data
+
+cmyc <- readRDS("./data/data-gen/protein/cmyc.change.RDS")
+
+ubf <- readRDS("./data/data-gen/protein/ubf.change.RDS")
+
+rps6 <- readRDS("./data/data-gen/protein/rps6.change.RDS")
+
+## cmyc fig
+
+cmyc.plot <- cmyc %>%
+  mutate(time = "post") %>%
+  add_row(supplement = c("PLACEBO", "GLUCOSE"), time = "pre", emmean = 0, lower.CL = 0, upper.CL = 0) %>%
+  mutate(time = factor(time, levels = c("pre", "post"), labels = c("Baseline", "Post"))) %>%
+  ggplot(aes(time, exp(emmean), group = supplement, fill = supplement)) +
+  #annotate("text", x = "Post", y = 2.1, label = "p = 0.585", size = 3) +
+  geom_errorbar(aes(ymin = exp(lower.CL), ymax = exp(upper.CL)),
+                width = 0.1,
+                position = position_dodge(width = 0.2)) +
+  geom_line(position = position_dodge(width = 0.2)) +
+  geom_point(shape = 21, size = 3, position = position_dodge(width = 0.2)) +
+  labs(x = "", y = "c-Myc signal \n(fold change)\n", fill = "") +
+  theme_classic()
+
+ubf.plot <- ubf %>%
+  mutate(time = "post") %>%
+  add_row(supplement = c("PLACEBO", "GLUCOSE"), time = "pre", emmean = 0, lower.CL = 0, upper.CL = 0) %>%
+  mutate(time = factor(time, levels = c("pre", "post"), labels = c("Baseline", "Post"))) %>%
+  ggplot(aes(time, exp(emmean), group = supplement, fill = supplement)) +
+  #annotate("text", x = "Post", y = 2.1, label = "p = 0.585", size = 3) +
+  geom_errorbar(aes(ymin = exp(lower.CL), ymax = exp(upper.CL)),
+                width = 0.1,
+                position = position_dodge(width = 0.2)) +
+  geom_line(position = position_dodge(width = 0.2)) +
+  geom_point(shape = 21, size = 3, position = position_dodge(width = 0.2)) +
+  labs(x = "", y = "UBF signal \n(fold change)\n", fill = "") +
+  theme_classic()
+
+rps6.plot <- rps6 %>%
+  mutate(time = "post") %>%
+  add_row(supplement = c("PLACEBO", "GLUCOSE"), time = "pre", emmean = 0, lower.CL = 0, upper.CL = 0) %>%
+  mutate(time = factor(time, levels = c("pre", "post"), labels = c("Baseline", "Post"))) %>%
+  ggplot(aes(time, exp(emmean), group = supplement, fill = supplement)) +
+  annotate("text", x = "Post", y = 1.9, label = "p = ", size = 3) +
+  geom_errorbar(aes(ymin = exp(lower.CL), ymax = exp(upper.CL)),
+                width = 0.1,
+                position = position_dodge(width = 0.2)) +
+  geom_line(position = position_dodge(width = 0.2)) +
+  geom_point(shape = 21, size = 3, position = position_dodge(width = 0.2)) +
+  labs(x = "", y = "rps6 signal \n(fold change)\n", fill = "") +
+  theme_classic()
+
+# Cowplot for gathering figures
+
+legend <- get_legend(rps6.plot + theme(legend.box.margin = margin(0, 0, 0,12)))
+
+prot.joined <- plot_grid(cmyc.plot + theme(legend.position = "none"),
+                         ubf.plot + theme(legend.position = "none"), 
+                         rps6.plot + theme(legend.position = "none"),
+                         legend,
+                         ncol = 2, nrow = 2)
+
+prot.joined
+
+saveRDS(prot.joined, "./data/data-gen/rna/prot.joined.RDS")
+
+```
+
+###### **Figure 5:** A) Changes in c-Myc levels, B) changes in UBF levels, C) changes in rps6 levels.
+Baseline = Training day 1/2, post = Training day 11/12.
+Four samples were analysed per participant (GLU pre-post, PLAC pre-post), in duplicates, a total of 104 protein samples.
+Duplicate samples were loaded on separate gels.
+In addition, 3-4 pools were added to each well.
+c-Myc, UBF and rps6 signals were then calculated as log-fold change score, normalised by gel pools.
+Mean change scores of the duplicates were calculated before modelling and transformed to the log-scale.
+Values are estimated marginal means fold change per leg/supplement ± 95% CI.
 p = between groups statistic, glucose compared to placebo.
@@
 
-There were no significant differences in mean change of total session volume, training intensity or RPE between exercise training with glucose and placebo at any of the time points (Fig 5, p = 0.740, p = 0.917, p = 0.645, post change compared to baseline, respectively).
+There were no significant differences in mean change of total session volume, training intensity or RPE between exercise training with glucose and placebo at any of the time points (fig 6, *p* > 0.05).
 We did observe interactions between time and changes, where total session training volume increased from 13-18% compared to baseline.
 These increases in training volume were accompanied by changes ranging from 5-18% in training intensity and 2-13% in RPE (Tab 5).
-The maximal isometric knee extension force (OR TORQUE?) and isokinetic knee extension peak torque at 240 d/s showed no differences between exercise training with glucose and placebo after five RT sessions (Fig 6A, Isometric: p = 0.336, 240 d/s: p = 0.527, post change compared to baseline, respectively).
-There was a significant difference in mean change from pre to post-fifth RT session, where exercise training with glucose led to a 9% less decrease in peak torque at 60 d/s compared to placebo (Fig 6A, p = 0.037).
-There was no difference between glucose and placebo in mean change from baseline to after the sixth RT session (Fig 6B, Isometric: p = 0.442, 60 d/s: p = 0.377, 240 d/s: p = 0.154, compared to baseline, respectively).
+The maximal isometric knee extension force (OR TORQUE?) and isokinetic knee extension peak torque at 240 d/s showed no differences between exercise training with glucose and placebo after five RT sessions (Fig 7A, *p* > 0.05).
+There was a significant difference in mean change from pre to post-fifth RT session, where exercise training with glucose led to a 9% less decrease in peak torque at 60 d/s compared to placebo (Fig 7A, *p* = `r isok605diffp`).
+There was no difference between glucose and placebo in mean change from baseline to after the sixth RT session (Fig 7B, *p* > 0.05).
 

### 2023-02-02 western.figs.R

@@ -1,4 +1,4 @@
-## Western change figure
+## Western figures
 
 # This script produces the figures for cmyc, ubf and rps6
 
@@ -11,56 +11,50 @@ library(cowplot)
 
 # Data
 
-cmyc <- readRDS("./data/data-gen/protein/cmyc.change.RDS")
+cmyc <- readRDS("./data/data-gen/protein/cmyc.emm.RDS")
 
-ubf <- readRDS("./data/data-gen/protein/ubf.change.RDS")
+ubf <- readRDS("./data/data-gen/protein/ubf.emm.RDS")
 
-rps6 <- readRDS("./data/data-gen/protein/rps6.change.RDS")
+rps6 <- readRDS("./data/data-gen/protein/rps6.emm.RDS")
 
 
 ## cmyc fig
 
 cmyc.plot <- cmyc %>%
-  mutate(time = "post") %>%
-  add_row(supplement = c("PLACEBO", "GLUCOSE"), time = "pre", emmean = 0, lower.CL = 0, upper.CL = 0) %>%
   mutate(time = factor(time, levels = c("pre", "post"), labels = c("Baseline", "Post"))) %>%
-  ggplot(aes(time, exp(emmean), group = supplement, fill = supplement)) +
+  ggplot(aes(time, emmean, group = supplement, fill = supplement)) +
   #annotate("text", x = "Post", y = 2.1, label = "p = 0.585", size = 3) +
-  geom_errorbar(aes(ymin = exp(lower.CL), ymax = exp(upper.CL)),
+  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
                 width = 0.1,
                 position = position_dodge(width = 0.2)) +
-  geom_line(position = position_dodge(width = 0.2)) +
+ # geom_line(position = position_dodge(width = 0.2)) +
   geom_point(shape = 21, size = 3, position = position_dodge(width = 0.2)) +
-  labs(x = "", y = "c-Myc signal \n(fold change)\n", fill = "") +
+  labs(x = "", y = "c-Myc", fill = "") +
   theme_classic()
 
 
 ubf.plot <- ubf %>%
-  mutate(time = "post") %>%
-  add_row(supplement = c("PLACEBO", "GLUCOSE"), time = "pre", emmean = 0, lower.CL = 0, upper.CL = 0) %>%
   mutate(time = factor(time, levels = c("pre", "post"), labels = c("Baseline", "Post"))) %>%
[…]

### 2023-02-02 ribose-paper.RMD
updated methods

@@
 Following the wet transfer, membranes were stained with a MemCode reversible total protein stain (MANUFACTURER) and then blocked for 1hr at room temp with a blocking buffer of Tris-buffered saline (TBS; 20 mM Tris, 150 mM NaCl) with 5% non-fat dry milk and 0.1% Tween-20.
-Primary antibodies were purchased from ……, secondary antibodies were purchased from ……, and were diluted in a 5% non-fat dry milk blocking buffer (TBS, 5% non-fat dry milk, 0.1% Tween-20).
+Primary and secondary antibodies were purchased from Santa Cruz Biotechnology (Texas, USA) UBF:(UBF-9, sc-13125), rpS6:(Ribosomal protein S6 C-8, sc-74459), and Thermo Fisher Scientific () c-Myc:(c-Myc 9E10), goat anti-mouse for c-Myc (goat anti-mouse IgG1 (y1) horseradish peroxidase conjugate), and anti-mouse (anti-mouse IgG1 horseradish peroxidase conjugate).
+Antibodies were diluted in a 5% non-fat dry milk blocking buffer (TBS, 5% non-fat dry milk, 0.1% Tween-20).
 Membranes were incubated overnight with primary antibodies, and for 1hr with secondary antibodies.

### 2023-02-03 ribose-paper.RMD
updated manuscript

@@
 Studies have shown RT-induced muscle growth to vary widely between individuals, with as much as 10-15% showing considerably impaired growth (ALVAREZ18, MANN14, THALACKER13).
-This may be explained by genetics, epigenetics, and/or an unfavourable internal physiological milieu (KILDER).
-Recently, tweaking training modalities and -protocols have been shown to induce different responses in different populations (HAMMARSTROM20), however, little is known regarding the optimisation of RT for individual phenotypes.
-Generally, meta-analyses favour a moderate volume of RT over low volume RT regarding muscle growth and -strength (KRIEGER09, RHEA03, SCHOENFELD17), substantiated by recent RCTs (KILDER).
-Although Hammarström and colleagues (20) found that a moderate volume led to more prominent gains in muscle mass and -strength, approximately 50% of the participants did not experience true benefits of moderate volume compared to low, with a blunted muscle growth (and strength?).
-Due to the variations observed in muscle growth responses to standardised RT protocols (HAMMARSTROM20; SCHOENFELD17; STEC16?), it is apparent that increasing RT volume alone does not convert low-responders to high-responders.
-Therefore, other means than RT per se may be necessary to circumvent this discrepancy.
+This may be explained by genetics, epigenetics, and/or an unfavourable internal physiological milieu (THALACKER13, HAMMARSTROM20, FIGUEIREDO19?).
+While tweaking training modalties and -protocols induces different responses in different populations (HAMMARSTROM20, 19?), little is known regarding optimisation of RT to individual phenotypes.
+Generally, meta-analyses favour a moderate volume of RT over low volume RT regarding muscle growth and -strength (KRIEGER09, RHEA03, SCHOENFELD17), substantiated by recent RCTs (HAMMARSTROM20..).
+Although Hammarström and colleagues (20) found that a moderate volume led to more prominent gains in muscle mass and -strength, approximately 50% of the participants did not experience true benefits of moderate volume compared to low, with a blunted muscle growth.
+Due to the variations observed in muscle growth responses despite standardising RT protocols (HAMMARSTROM20; SCHOENFELD17; STEC16?), it is apparent that increasing RT volume alone does not convert low-responders to high-responders.
+Therefore, other means than RT *per se* seems necessary to circumvent this discrepancy.
 
@@
 This is surprising, as glucose is the preferred energy substrate of the contracting skeletal muscle during strenuous exercise and a major supplier of energy to cells via ATP synthesis (TANAKA18).
-Energy availability is important to regulate central processes of muscle growth, specifically ribosome biogenesis, a key factor for increased muscle protein synthesis (MPS) (TANAKA18, FIGUEIREDO19).
+Energy availability is important to regulate central processes of muscle growth, specifically ribosome biogenesis, a key factor for increased muscle protein synthesis (MPS) (TANAKA18, FIGUEIREDO19, VONWALDEN19?).
 Interestingly, variations in RT-induced responses have been linked to ribosome biogenesis, where participants responding poorly to RT also showed a blunted ribosome biogenesis (HAMMARSTROM20, STEC16).
-A blunted ribosome biogenesis negates responses to RT by failure to produce novel ribosomes, thus the inability to increase the skeletal muscle cell’s translational capacity (VONWALDEN19, FIGUEIREDO19).
+A blunted ribosome biogenesis negates responses to RT by failure to produce novel ribosomes, thus an inability to increase the skeletal muscle cell’s translational capacity (VONWALDEN19, FIGUEIREDO19).
 Therefore, investigating the effects of combined RT and glucose ingestion may provide valuable insight into the regulation of ribosome biogenesis.
@@
 Further, the upstream binding factor (UBF) has a vital role in the initiation, binding the PIC to the rDNA promoter, while c-Myc serves as a transcription factor (FIGUEIREDO19, VONWALDEN19).
-Upon initiation, Pol I transcribed the 47S pre-rRNA, ultimately processed into the 18S-, 5.8S and 28S rRNA (MULLINEUX12, VONWALDEN19).
+Upon initiation, Pol I transcribes the 47S pre-rRNA, ultimately processed into the 18S-, 5.8S and 28S rRNA (MULLINEUX12, VONWALDEN19).
 The fourth rRNA, 5S, is transcribed by Pol III, while mRNAs for the ribosomal proteins (r-proteins) are transcribed by Pol II (MULLINEUX12, VONWALDEN19).
 Together, the four mature rRNAs (18S, 5.8S, 28S and 5S) along with 80 ribosomal proteins form the two subunits of a ribosome (FIGUEIREDO19, VONWALDEN19).
-
-Whether glucose has mechanisms to affect ribosome biogenesis i.e., acting as a signalling molecule per se, or indirectly via insulin is currently unknown.
-Factors involved in the (PIC) of rDNA transcription are thought to receive cues from pathways (mTORC1, MAPK) regulated by hormonal and nutritional signals (FIGUEIREDO19).
-Furthermore, high glucose treatment promoted ribosome biogenesis while glucose depletion repressed ribosome biogenesis in mice (ZHAI12 – OR WAS IT CELL CULTURES?).
-Thus, it seems plausible that glucose ingestion before and after RT may benefit ribosome biogenesis.
-Lastly, there are equivocal observations on insulin’s impact on ribosome biogenesis and MPS.
-Previous meta-analyses suggest insulin acts in concert with amino acid intake to enhance MPS, and that insulin may reduce MPB when amino acids are scarce (ABDULLA16).
-
-To investigate this further, we recruited 16 young, moderately trained, male and female individuals for heavy unilateral resistance training for two weeks, exercising one leg with glucose and protein, and the other with a placebo and protein ingestion before and after each session.
-Our main purpose was to investigate the effects of glucose ingestion compared to placebo ingestion before and after five RT sessions on ribosome biogenesis.
-Changes in ribosome biogenesis were measured by the accumulation of total RNA and expression of the ribosome’s four mature rRNAs and the 47S precursor RNA, along with levels of UBF, c-Myc, and rps6 protein.
-As a secondary outcome, we were also interested in the effects of glucose compared to placebo ingestion on muscular function and -recovery, measured via unilateral isometric and isokinetic knee extension torque, before, during and after the intervention.
-We hypothesised that RT would lead to a prominent accumulation of total RNA, rRNA and protein (UBF, c-Myc, rps6) and that RT with glucose and protein would lead to higher accumulation and expression compared to RT with placebo and protein.
-We also hypothesised that glucose supplementation would enhance muscular function and -recovery compared to placebo.
+The factors involved in the PIC are proposed to receive cues from pathways such as mTORC1 and MAPK, which are regulated by among others hormonal and nutritional signals (FIGUEIREDO19, TANAKA18).
+High glucose treatment did promote ribosome biogenesis compared to a repressed ribosome biogenesis with low glucose treatment and glucose starvation, in cell cultures (?)(ZHAI12).
+And so, though it seems apparent that sufficient daily intake of glucose/carbohydrates is necessary for an optimal ribosome biogenesis (ZHAI12, TANAKA18), it is not clear whether there are acute effects on ribosome biogenesis _per se_ with ingesting glucose before and after RT.
+
+Whether glucose has mechanisms to affect ribosome biogenesis i.e., acting as a signalling molecule *per se*, or indirectly via insulin is currently unknown.
+There are equivocal observations on insulin's impact on ribosome biogenesis and MPS, where previous meta-analyses suggest insulin to act in concert with amino acid intake to enhance MPS (ABDULLA16).
+Furhter, insulin may reduce skeletal muscle protein breakdown when amino acids are scarce (ABDULLA16).
+Therefore, to investigate the relationship between glucose, RT and ribosome biogenesis we recruited 16 young, moderately trained, male and female individuals for heavy unilateral resistance training for two weeks.
+They exercised one leg with glucose and protein, and the other leg with placebo and protein ingestion before and after each RT-session.
+Our main purpose was to investigate whether there are acute effects of glucose- compared to placebo ingestion before and after RT.
+To assess this purpose, we measured changes in ribosome biogenesis as accumulation of total RNA, expression of the four mature rRNA's (18S, 5.8S, 28S, 5S), 47S pre-rRNA, and levels of proteins associated with transcription initiation (c-Myc, UBF, rpS6?).
+As secondary outcomes, we measured the developement in skeletal muscule performance during the intervention, and skeletal muscle recovery after the final RT session, as temporal changes in maximal isometric and isokinetic torque.
+We hypothesised that RT would indeed lead to a prominent accumulation of total RNA, increased expression of rRNA and associated proteins.
+Further, we hypothesised that RT with glucose and protein would induce an enhanced response compared to placebo, in markers of ribosome biogenesis, as well as muscular performance and -recovery.
 
@@
 The tissue was homogenized with a plastic rod in 80mg protein extraction powder (Invent Biotechnology) and 100ul iced cell lysis buffer (Denaturing Buffer, Invent Biotechnology), and centrifuged at top speed for 1 min.
-The supernatant was aliquoted into 2x25ul tubes to run samples in duplicates, and total protein concentrations were determined in a 1:10 dilution (Pierce Detergent Compatible Bradford Assay Reagent, Thermo Fisher Scientific).
+The supernatant was aliquoted into 2x25ul tubes to run samples in duplicates, and total protein concentrations were determined in a 1:10 dilution (Pierce Detergent Compatible Bradford Assay Reagent, Thermo Fisher Scientific, Oslo, Norway).
 The protein samples were diluted to 20.25 µg µl−1 total protein with lysis buffer and 4X Laemmli sample buffer (Bio-Rad Laboratories AB, Oslo, Norway) containing 2-mercaptoethanol.
@@
 All protein samples were incubated at 95°C and stored at -20°C until further analysis.
-The protein samples (20.25µg total protein) were separated at 250V for 50min using ….
-gels (……..), then transferred to PVDF membranes with wet transfer at 300mA for 3hrs.
+The protein samples (20.25µg total protein) were separated at 250V for 50min using 4-20% Tris-Glycine gels (Thermo Fisher Scientific), then transferred to PVDF membranes with wet transfer at 300mA for 3hrs.
 Both gel electrophoresis and protein transfer were performed at 4°C.
@@
 
-## Protein signals
+### Protein signals
 

### 2023-02-07 ribose-paper.RMD
updated rmd and correlation

@@
 
-Regular resistance exercise training (RT) increases muscle mass and muscle strength, due to the adaptations stimulated by a mechanical strain.
+Regular resistance exercise training (RT) increases muscle mass and muscle strength, due to the adaptations stimulated by  mechanical strain.
 Unfortunately, responses to RT are not uniform across the population.
@@
 
-Indeed, ingesting nutrients such as protein and creatine supplements effectively optimise RT (CERMAK12, LANHERS15, LANHERS17).
+Indeed, ingesting nutrients such as protein- and creatine-supplements effectively optimise RT (CERMAK12, LANHERS15, LANHERS17).
 However, it remains unknown if other nutritional adjuvants such as glucose can increase the efficacy of RT.
@@
 Stimulated by growth factors, hormones, mechanical loading and nutrients, mTORC1 governs energy distribution by turning on/off energy-consuming and energy-generating pathways, depending on cellular energy levels (HOPPE09).
-Ribosome biogenesis consumes up to 80% of the cell’s energy and materials (SCHMIDT04) and is presumably switched on/off based on cellular energy levels (VONWALDEN19).
+This greatly affects ribosome biogenesis, being a process that consumes approximately 80% of cellular energy and materials (VONWALDEN19).
 Upon stimulation by mTORC1, ribosome biogenesis is initiated with the assembly of the preinitiation complex (PIC) (VONWALDEN19, FIGUEIREDO19).
 The PIC consists of the RNA polymerase I complex, the SL-1 complex and the TIF-1A/RRN3 protein (FIGUEIREDO19, VONWALDEN19).
-Further, the upstream binding factor (UBF) has a vital role in the initiation, binding the PIC to the rDNA promoter, while c-Myc serves as a transcription factor (FIGUEIREDO19, VONWALDEN19).
-Upon initiation, Pol I transcribes the 47S pre-rRNA, ultimately processed into the 18S-, 5.8S and 28S rRNA (MULLINEUX12, VONWALDEN19).
-The fourth rRNA, 5S, is transcribed by Pol III, while mRNAs for the ribosomal proteins (r-proteins) are transcribed by Pol II (MULLINEUX12, VONWALDEN19).
-Together, the four mature rRNAs (18S, 5.8S, 28S and 5S) along with 80 ribosomal proteins form the two subunits of a ribosome (FIGUEIREDO19, VONWALDEN19).
-The factors involved in the PIC are proposed to receive cues from pathways such as mTORC1 and MAPK, which are regulated by among others hormonal and nutritional signals (FIGUEIREDO19, TANAKA18).
-High glucose treatment did promote ribosome biogenesis compared to a repressed ribosome biogenesis with low glucose treatment and glucose starvation, in cell cultures (?)(ZHAI12).
-And so, though it seems apparent that sufficient daily intake of glucose/carbohydrates is necessary for an optimal ribosome biogenesis (ZHAI12, TANAKA18), it is not clear whether there are acute effects on ribosome biogenesis _per se_ with ingesting glucose before and after RT.
-
-Whether glucose has mechanisms to affect ribosome biogenesis i.e., acting as a signalling molecule *per se*, or indirectly via insulin is currently unknown.
-There are equivocal observations on insulin's impact on ribosome biogenesis and MPS, where previous meta-analyses suggest insulin to act in concert with amino acid intake to enhance MPS (ABDULLA16).
-Furhter, insulin may reduce skeletal muscle protein breakdown when amino acids are scarce (ABDULLA16).
-Therefore, to investigate the relationship between glucose, RT and ribosome biogenesis we recruited 16 young, moderately trained, male and female individuals for heavy unilateral resistance training for two weeks.
+Further, the upstream binding factor (UBF) and c-Myc have vital roles in the initiation, where UBF binds the PIC to the rDNA promoter and c-Myc as a transcription factor (FIGUEIREDO19, VONWALDEN19).
+Upon initiation, Pol I transcribes the 47S pre-rRNA, and this is seen as the rate-limiting step (VONWALDEN19, FIGUEIREDO19).
+The 47S pre-rRNA is processed into the 18S, 5.8S and 28S rRNA (MULLINEUX12, VONWALDEN19), while the last rRNA 5S is transcribed by Pol III, and the ribosomal proteins by Pol II (MULLINEUX12, VONWALDEN19).
+These events facilitates increased ribosomal content of each skeletal muscle cell, subsequently allowing a higher basal- and roof for MPS, assumed to predict long term adaptations to RT (FIGUEIREDO19/18).
+Since a blunted ribosome biogenesis has been observed despite increasing mechanical loading (HAMMARSTROM20), targeting hormonal and nutritional signals may ameliorate this.
+High glucose treatment did promote ribosome biogenesis compared to a repressed ribosome biogenesis with low glucose treatment and glucose starvation, in mice and cell cultures (?)(ZHAI12).
+
+Thus, though it seems apparent that sufficient daily intake of glucose/carbohydrates is necessary for an optimal ribosome biogenesis (ZHAI12, TANAKA18), it is unclear whether there are acute effects on ribosome biogenesis *per se* with ingesting glucose before and after RT.
+Further, there are equivocal observations on insulin's impact on ribosome biogenesis and MP, where previous meta-analyses suggest insulin to act in concert with amino acid intake to enhance MPS, and insulin may reduce skeletal muscle protein breakdown when amino acids are scarce (ABDULLA16).
+Therefore, to investigate the relationship between glucose, RT and ribosome biogenesis we recruited 16 young, moderately trained male and female individuals to heavy unilateral resistance training for two weeks.
 They exercised one leg with glucose and protein, and the other leg with placebo and protein ingestion before and after each RT-session.
@@
 Participants were randomized to commence with either glucose (GLU, 30g per bolus) or placebo (PLAC, 0g per bolus), and either dominant or non-dominant leg (Fig 1), alternating RT and supplement from one day to the other (Fig 2).
-Participants were linked to a unique ID number corresponding to supplement administration, disclosed only to personnel not involved in any other aspect of the study than the randomization.
+Investigators were blinded as the supplement randomization was disclosed only to personnel not involved in any other aspect of the study, and participants were blinded by diluting boluses of glucose (30g glucose) (Glucosum monohydricum, Merck KGaA, Darmstadt, Germany) and placebo (~0.3g Stevia rebaudiana extract) (Steviosa, Soma Nordic AS, Oslo, Norway) in 300ml Fun Light juice/saft (Orkla, Oslo, Norway).
+A blinded taste test revealed that the participants were not able to disclose the contents of the provided boluses (30g GLU vs.
+0g GLU).
+To ensure equal conditions during training sessions, participants exercised and tested at the same time of day, +/- 1hr with the same trainer, and recorded and repeated their daily macro nutrient intake on pairwise consecutive days (Tab 4).
 
@@
 Protein was ingested 2hrs before RT (0700hrs, 25g) and immediately after RT (~0930hrs, 25g).
-In the afternoon (1800hrs-1900hrs), during period II, participants ingested glucose or placebo (3x30g vs.
-3x0g glucose) opposite to the supplement they received during RT, to ensure a balanced intake of glucose during the entirety of the intervention days.
+In the afternoon (1800hrs-1900hrs, period II) participants ingested glucose or placebo (3x30g vs.
+3x0g glucose) opposite to the supplement they received during RT, to ensure a balanced daily intake of glucose.
 Apart from this, participants ingested a self-chosen diet during period II, registered in MyFitnessPal or similar applications.
@@
 During period III (2200-0700hrs), participants remained in an overnight fasted state.
-The daily onset of the dietary intervention (i.e., first ingestion of PRO supplement) varied for individual participants, between 0600hrs and 0900hrs to allow multiple participants to complete the protocol simultaneously.
-Individual participants commenced the intervention at the same time of day on every test day (pre and post) and intervention day +/- 1hr.
-
-Blinding of the participants was ensured by diluting boluses of glucose (30g glucose) (Glucosum monohydricum, Merck KGaA, Darmstadt, Germany) and placebo (~0.3g Stevia rebaudiana extract) (Steviosa, Soma Nordic AS, Oslo, Norway) in 300ml Fun Light juice/saft (Orkla, Oslo, Norway).
-A blinded taste test revealed that the participants were not able to disclose the contents of the provided boluses (30g GLU vs.
-0g GLU).
+The daily onset of the dietary intervention (i.e., first ingestion of PRO supplement) varied between participants, from 0600hrs to 0900hrs to allow multiple participants to complete the protocol simultaneously.
 The protein supplement was ingested as 25g Whey Protein Isolate boluses (Proteinfabrikken, Stokke, Norway), diluted in 150ml water.
@@
 
-Performance tests (Test) were performed prior to (Days -7 and -5, and T0, both legs) and during the intervention (Days 4, 5, 8, and 9), after session 5 and after finalization of the intervention (T3/T4, RT#1 leg; T4/Day 13, RT#2 leg), a total of 10 days (Fig 2).
+Performance tests (Test, Fig 2) were performed prior to (Days -7 and -5, and T0, both legs) and during the intervention (Days 4, 5, 8, and 9), after session 5 and after finalization of the intervention (T3/T4, RT#1 leg; T4/Day 13, RT#2 leg), a total of 10 days (Fig 2).
 Maximal isometric and isokinetic knee extension torque were measured with a Humac Norm Dynamometer (CSMi, Stoughton, Massechusetts, USA).
@@
 Maximal isokinetic torque was measured at speeds of 60- and 240 d/s, 2x3 repetitions each, with the first set of each exercise as a sub-maximal warm-up.
-The maximal isometric torque was measured at knee-angle 60, for a maximum of 10 seconds and one repetition per test.
+The maximal isometric torque was measured at knee-angle 60, for a maximum of 10 seconds and two repetition per test.
 The highest peak torque values were carried forwards to the final analyses.
-During days 4, 5, 8 and 9, humac tests were conducted one hour prior to RT with the leg performing RT the previous day.
+During days 4, 5, 8 and 9, humac tests were conducted one hour before RT with the leg performing RT the previous day.
 During days 11 and 12 (T3 and T4), humac tests were performed at four time points: I) 45min before RT, II) 30min after the last RT session, III) 2hrs after the last RT session, and IV) 23hrs after the last RT session.
@@
 Assessment of unilateral one repetition maximum (1RM) leg press and knee extension was conducted at pre-intervention testing (days -7 and -5, Fig 2).
-The participants warmed up with 10min cycling on an indoor exercise bicycle and a specific warm-up (1x10, 1x6 and 1x3 repetitions, 2min breaks) before each of the tests.
-Between specific warm-up sets as well as between each subsequent 1RM attempt, the participants were given 3min pauses.
-All positions were controlled and recorded at pre-intervention testing and repeated for each subsequent test and RT.
+The participants performed a general warm-up with 10min cycling on an indoor exercise bicycle and a specific warm-up (1x10, 1x6 and 1x3 repetitions, 2min breaks) before each of the tests.
+Between specific warm-up sets as well as between each subsequent 1RM attempt, the participants were given 3min breaks.
+All positions were controlled and recorded at pre-intervention testing and repeated for the RT sessions.
 Maximal leg press strength was defined as the maximal load lifted in a controlled fashion, with a knee angle of 45 degrees (start), within a 2.5kg range.
@@
 Resistance training consisted of three sets of each unilateral leg presses and knee extensions, with an intensity of 10 repetition maximum (10RM).
-All positions were controlled and recorded during pre-intervention testing and repeated each RT session.
 As a general warm-up, the participants cycled on an indoor exercise bicycle for 5-10min.
 In addition, before the respective exercises, two 10-repetition warm-up sets were completed at ~50% and ~70% of 10RM.
-Rest time between working sets was 2min, and we used progressive loading to ensure adequate exercise stimulation throughout the intervention.
+Rest time between working sets was 2min, and progressive loading was used to ensure adequate exercise stimulation throughout the intervention.
 For safety and standardization purposes, all sessions were monitored by trained personnel.
@@
 Muscle biopsy sampling was performed under local anaesthesia (Xylocaine, 10mg ml-1 with adrenaline 5μg ml−1, AstraZeneca AS, Oslo, Norway) using a 12-14-gauge needle (Universal Plus, Mermaid Medical AS, Stenløse, Denmark), operated with a spring-loaded biopsy gun.
-After the biopsy sampling, muscle tissue was divided into two aliquots for determination of total RNA/expression of rRNA and two aliquots for protein analyses, snap frozen in isopentane (-80°C) and stored at -80°C until further analyses.
+After the biopsy sampling, muscle tissue was divided in two aliquots for determination of total RNA/expression of rRNA and two aliquots for protein analyses, snap frozen in isopentane (-80°C) and stored at -80°C until further analyses.
 Muscle biopsies were sampled at four time points (Fig 2): I/II) Pre-intervention (2hrs before training, time point T1 = pre RT#1 leg, T2 = pre RT#2 leg), and III/IV) before the sixth RT session from RT#1 leg (2hrs before training, T3) and before the sixth RT session from RT#2 leg (2hrs before training, T4).
 At each time point, two samples were taken from the same incision.
-To standardize this procedure, all individual participants had biopsies taken at the same time of day +/- 1hr, in an overnight fasted state.
+To standardize this procedure, all individual participants had biopsies taken at the same time of day, in an overnight fasted state.
 
@@
 Finger draws were analysed with in-house equipment (BIOSEN C-Line, EKF diagnostic GmbH, Barleben).
-Blood samples were also collected coinciding with these finger draws, except in the middle of the RT session, to analyse endocrine variables.
+Venous blood samples were collected from the antecubital vein, coinciding with the finger draws except in the middle of the RT session, to analyse endocrine variables.
 
@@
 p.270 <- round(bcoef[11,5], 2)
+```
+
+```{r c-peptide data + model}
+## Data
+
+ins.dat <- read_excel("./data/glucose/ribose.insulin.xlsx", na = "NA")
+
+code <- read_excel("data/code_key.xlsx", na = "NA") %>%
+  select(-leg)
+
+## Data handling
+
+ins.dat$sample.time <- as.character(ins.dat$sample.time)
+
+ins.dat2 <- ins.dat %>%
+  inner_join(code) %>%
+  mutate(sample.time = as.character(sample.time)) %>%
+  mutate(time = if_else(sample.time == "0",
+                        "pre",
+                        if_else(sample.time == "0.1",
+                                "baseline",
+                                if_else(sample.time == "45",
+                                "min45",
+                                if_else(sample.time == "90",
+                                        "min90",
+                                        if_else(sample.time == "120",
+                                                "min120",
+                                                if_else(sample.time == "270",
+                                                        "min270", sample.time))))))) %>%
+  mutate(time = factor(time, levels = c("pre", "baseline", "min45", "min90", "min120", "min270")),
+         c.pep = as.numeric(c.pep))
+
+## Baseline analysis
+# Comparing pre measurements to post day baseline
+
+base.ins <- ins.dat2 %>%
+  filter(time %in% c("pre", "baseline")) %>%
+  select(subject, time, supplement, c.pep) %>%
+  group_by(supplement) %>%
+  pivot_wider(names_from = time, 
+              values_from = c.pep)
+
+ins.ttest <- t.test(base.ins$pre, base.ins$baseline, paired = TRUE)
+
+## Change data
+
+ins.change <- ins.dat2 %>%
+  #filter(subject != "102") %>%
+  dplyr::select(subject, time, c.pep, supplement) %>%
+  group_by(subject, time, supplement) %>%
+  summarise(mean.pep = mean(c.pep, na.rm = TRUE)) %>%
+  pivot_wider(names_from = time, 
+              values_from = mean.pep) %>%
+  ungroup() %>%
+  mutate(change.45 = min45-baseline,
+         change.90 = min90-baseline,
+         change.120 = min120-baseline,
+         change.270 = min270-baseline,
+         baseline = baseline - mean(baseline, na.rm = TRUE),
+         supplement = factor(supplement, levels = c("PLACEBO", "GLUCOSE"))) %>%
+  select(subject, supplement, baseline, change.45, change.90, change.120, change.270) %>%
+  pivot_longer(names_to = "time",
+               values_to = "change",
+               cols = (change.45:change.270))
+
+ins.lchange <- ins.dat2 %>%
+  #filter(subject != "102") %>%
+  dplyr::select(subject, time, c.pep, supplement) %>%
+  group_by(subject, time, supplement) %>%
+  summarise(mean.pep = mean(c.pep, na.rm = TRUE)) %>%
+  pivot_wider(names_from = time, 
+              values_from = mean.pep) %>%
+  ungroup() %>%
+  mutate(change.45 = log(min45)-log(baseline),
+         change.90 = log(min90)-log(baseline),
+         change.120 = log(min120)-log(baseline),
+         change.270 = log(min270)-log(baseline),
+         baseline = baseline - mean(baseline, na.rm = TRUE),
+         supplement = factor(supplement, levels = c("PLACEBO", "GLUCOSE"))) %>%
+  select(subject, supplement, baseline, change.45, change.90, change.120, change.270) %>%
+  pivot_longer(names_to = "time",
+               values_to = "change",
+               cols = (change.45:change.270))
+
+## Linear mixed effects model
+
+m1 <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
+                     data = ins.lchange) # Main analysis, using log-transformed data due to heteroscedasticity
+
+m2 <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
+                     data = ins.change) # For emmeans and figure illustration
+
+m1.plot <- plot(m1)
+m2.plot <- plot(m2)
+
+m1.sum <- summary(m1)
+
+## Emmeans
+
+library(pbkrtest)
+
+# For analysis
+insulin.change <- confint.m1 <- confint(emmeans(m1, specs = ~"supplement|time")) %>%
+  data.frame()
+
+saveRDS(insulin.change, "./data/data-gen/glucose/insulin.change.RDS")
+
+# For figure
+insulin.figchange <- confint.m2 <- confint(emmeans(m2, specs = ~"supplement|time")) %>%
+  data.frame()
+
+saveRDS(insulin.figchange, "./data/data-gen/glucose/insulin.figchange.RDS")
 ```
@@
 
-```{r glucose fig}
+```{r glucose + c-peptide fig}
 # Figure
@@
 gluc.fig <- readRDS("./data/data-gen/glucose/gluc.fig.RDS")
-gluc.fig
+
+cpep.fig <- readRDS("./data/data-gen/glucose/cpep.fig.RDS")
+
+plot_grid(gluc.fig,
+          cpep.fig,
+          ncol = 1, nrow = 2)
 

### 2023-02-08 ribose-paper.RMD
updated intro

@@
 Unfortunately, responses to RT are not uniform across the population.
-Studies have shown RT-induced muscle growth to vary widely between individuals, with as much as 10-15% showing considerably impaired growth (ALVAREZ18, MANN14, THALACKER13).
-This may be explained by genetics, epigenetics, and/or an unfavourable internal physiological milieu (THALACKER13, HAMMARSTROM20, FIGUEIREDO19?).
-While tweaking training modalties and -protocols induces different responses in different populations (HAMMARSTROM20, 19?), little is known regarding optimisation of RT to individual phenotypes.
-Generally, meta-analyses favour a moderate volume of RT over low volume RT regarding muscle growth and -strength (KRIEGER09, RHEA03, SCHOENFELD17), substantiated by recent RCTs (HAMMARSTROM20..).
-Although Hammarström and colleagues (20) found that a moderate volume led to more prominent gains in muscle mass and -strength, approximately 50% of the participants did not experience true benefits of moderate volume compared to low, with a blunted muscle growth.
-Due to the variations observed in muscle growth responses despite standardising RT protocols (HAMMARSTROM20; SCHOENFELD17; STEC16?), it is apparent that increasing RT volume alone does not convert low-responders to high-responders.
+Studies have shown RT-induced muscle growth to vary widely between individuals, with as much as 10-15% showing considerably impaired growth (ALVAREZ18;MANN14;THALACKER13).
+This may be explained by genetics, epigenetics, and/or an unfavourable internal physiological milieu (FIGUEIREDO21;HAMMARSTROM20;THALACKER13).
+While tweaking training modalties and -protocols induces different responses in different populations (HAMMARSTROM20), little is known regarding optimisation of RT to individual phenotypes.
+Generally, meta-analyses favour a moderate volume of RT over low volume RT regarding muscle growth (KRIEGER09;RHEA03;SCHOENFELD17), substantiated by recent RCTs (HAMMARSTROM20;SCHOENFELD19).
+Although Hammarström and colleagues (2020) found that a moderate volume led to more prominent gains in muscle mass, approximately 50% of the participants did not experience true benefits of moderate volume compared to low, with a blunted muscle growth.
+Due to the variations observed in muscle growth responses despite standardising RT protocols (HAMMARSTROM20;SCHOENFELD17;STEC16), it is apparent that increasing RT volume alone does not convert low-responders to high-responders.
 Therefore, other means than RT *per se* seems necessary to circumvent this discrepancy.
 
-Indeed, ingesting nutrients such as protein- and creatine-supplements effectively optimise RT (CERMAK12, LANHERS15, LANHERS17).
+Indeed, ingesting nutrients such as protein- and creatine-supplements effectively optimise RT (CERMAK12;LANHERS15;LANHERS17).
 However, it remains unknown if other nutritional adjuvants such as glucose can increase the efficacy of RT.
-This is surprising, as glucose is the preferred energy substrate of the contracting skeletal muscle during strenuous exercise and a major supplier of energy to cells via ATP synthesis (TANAKA18).
-Energy availability is important to regulate central processes of muscle growth, specifically ribosome biogenesis, a key factor for increased muscle protein synthesis (MPS) (TANAKA18, FIGUEIREDO19, VONWALDEN19?).
-Interestingly, variations in RT-induced responses have been linked to ribosome biogenesis, where participants responding poorly to RT also showed a blunted ribosome biogenesis (HAMMARSTROM20, STEC16).
-A blunted ribosome biogenesis negates responses to RT by failure to produce novel ribosomes, thus an inability to increase the skeletal muscle cell’s translational capacity (VONWALDEN19, FIGUEIREDO19).
+This is surprising, as glucose is the preferred energy substrate of the contracting skeletal muscle during strenuous exercise and a major supplier of energy to cells via ATP synthesis (MUL15;TANAKA18).
+Energy availability is important to regulate central processes of muscle growth, specifically ribosome biogenesis, a key factor for increased muscle protein synthesis (MPS) (FIGUEIREDO19;HAMMARSTROM20;VONWALDEN19;TANAKA18).
+Interestingly, variations in RT-induced responses have been linked to ribosome biogenesis, where participants responding poorly to RT also showed a blunted ribosome biogenesis (HAMMARSTROM20;STEC16).
+A blunted ribosome biogenesis negates responses to RT by failure to produce novel ribosomes, thus an inability to increase the skeletal muscle cell’s translational capacity (VONWALDEN19;FIGUEIREDO19).
 Therefore, investigating the effects of combined RT and glucose ingestion may provide valuable insight into the regulation of ribosome biogenesis.
 
-Ribosome biogenesis is suggested to be regulated by the mammalian target of rapamycin complex 1 (mTORC1) network (KUSNADI15, TANAKA18).
+Ribosome biogenesis is suggested to be regulated by the mammalian target of rapamycin complex 1 (mTORC1) network (KUSNADI15;VONWALDEn19;TANAKA18).
 Stimulated by growth factors, hormones, mechanical loading and nutrients, mTORC1 governs energy distribution by turning on/off energy-consuming and energy-generating pathways, depending on cellular energy levels (HOPPE09).
-This greatly affects ribosome biogenesis, being a process that consumes approximately 80% of cellular energy and materials (VONWALDEN19).
-Upon stimulation by mTORC1, ribosome biogenesis is initiated with the assembly of the preinitiation complex (PIC) (VONWALDEN19, FIGUEIREDO19).
+This greatly affects ribosome biogenesis, being a process that consumes approximately 80% of cellular energy and materials (VONWALDEN19;TANAKA18).
+Upon stimulation by mTORC1, assembly of the preinitiation complex (PIC) marks the initiation of ribosome biogenesis (VONWALDEN19;FIGUEIREDO19).
 The PIC consists of the RNA polymerase I complex, the SL-1 complex and the TIF-1A/RRN3 protein (FIGUEIREDO19, VONWALDEN19).
 Further, the upstream binding factor (UBF) and c-Myc have vital roles in the initiation, where UBF binds the PIC to the rDNA promoter and c-Myc as a transcription factor (FIGUEIREDO19, VONWALDEN19).
-Upon initiation, Pol I transcribes the 47S pre-rRNA, and this is seen as the rate-limiting step (VONWALDEN19, FIGUEIREDO19).
+Pol I transcribes the 47S pre-rRNA, suggested as the rate-limiting step in ribosome biogenesis (VONWALDEN19, FIGUEIREDO19).
 The 47S pre-rRNA is processed into the 18S, 5.8S and 28S rRNA (MULLINEUX12, VONWALDEN19), while the last rRNA 5S is transcribed by Pol III, and the ribosomal proteins by Pol II (MULLINEUX12, VONWALDEN19).
-These events facilitates increased ribosomal content of each skeletal muscle cell, subsequently allowing a higher basal- and roof for MPS, assumed to predict long term adaptations to RT (FIGUEIREDO19/18).
-Since a blunted ribosome biogenesis has been observed despite increasing mechanical loading (HAMMARSTROM20), targeting hormonal and nutritional signals may ameliorate this.
-High glucose treatment did promote ribosome biogenesis compared to a repressed ribosome biogenesis with low glucose treatment and glucose starvation, in mice and cell cultures (?)(ZHAI12).
-
-Thus, though it seems apparent that sufficient daily intake of glucose/carbohydrates is necessary for an optimal ribosome biogenesis (ZHAI12, TANAKA18), it is unclear whether there are acute effects on ribosome biogenesis *per se* with ingesting glucose before and after RT.
+Importantly, these events facilitates increased ribosomal content of each skeletal muscle cell, subsequently allowing a higher basal- and roof for MPS, suggested to predict long term adaptations to RT (HAMMARSTROM21).
+Since a blunted ribosome biogenesis has been observed to reduce/inhibit RT responses, despite increasing mechanical loading (HAMMARSTROM20), targeting hormonal and/or nutritional signals may ameliorate this.
+Previously, high glucose treatment promoted ribosome biogenesis compared to a repressed ribosome biogenesis with low glucose treatment and glucose starvation in mice and cell cultures (MARIAPPAN11;TANAKA15;ZHAI12), making glucose an interesting player in this perspective.
+
+Thus, though it seems apparent that sufficient daily intake of glucose/carbohydrates is necessary for an optimal ribosome biogenesis (TANAKA18), it is unclear whether there are acute effects on ribosome biogenesis *per se* with ingesting glucose before and after RT.
 Further, there are equivocal observations on insulin's impact on ribosome biogenesis and MP, where previous meta-analyses suggest insulin to act in concert with amino acid intake to enhance MPS, and insulin may reduce skeletal muscle protein breakdown when amino acids are scarce (ABDULLA16).
-Therefore, to investigate the relationship between glucose, RT and ribosome biogenesis we recruited 16 young, moderately trained male and female individuals to heavy unilateral resistance training for two weeks.
+Therefore, to investigate the relationship between glucose, RT and ribosome biogenesis we recruited 16 healthy young, moderately trained male and female individuals to moderate volume unilateral resistance training for two weeks.
 They exercised one leg with glucose and protein, and the other leg with placebo and protein ingestion before and after each RT-session.
@@
 To assess this purpose, we measured changes in ribosome biogenesis as accumulation of total RNA, expression of the four mature rRNA's (18S, 5.8S, 28S, 5S), 47S pre-rRNA, and levels of proteins associated with transcription initiation (c-Myc, UBF, rpS6?).
-As secondary outcomes, we measured the developement in skeletal muscule performance during the intervention, and skeletal muscle recovery after the final RT session, as temporal changes in maximal isometric and isokinetic torque.
+As secondary outcomes, we measured the development in skeletal muscle performance during the intervention, and skeletal muscle recovery after the final RT session, as temporal changes in maximal isometric and isokinetic torque.
 We hypothesised that RT would indeed lead to a prominent accumulation of total RNA, increased expression of rRNA and associated proteins.

### 2023-02-08 R/gluc.responsechange.R
small changes to glucose and insulin

@@ -6,6 +6,10 @@
 ## This script analyses is equal to mean-bloodglucose-change, except from not calculating log-fold change. Instead, absolute change scores are calculated
 # to illustrate the changes in absolute plasma glucose values. For description of the codes, see mean-bloodglucose-change.
 
+## Other comments:
+# Participant 101, 102 and 103 did not get baseline samples at T3 and T4, therefore,
+# baseline values from T1 was used as baseline for T3 and T4
+
 # Packages
 library(dplyr)
 library(tidyverse)
@@ -43,6 +47,7 @@ glu.dat <- gluc.dat %>%
 
 ## Change data
 
+# Un-transformed
 change_dat.glu <- glu.dat %>%
   # filter(subject != "107") %>%
   dplyr::select(subject, time, glu, supplement) %>%
@@ -68,21 +73,59 @@ change_dat.glu <- glu.dat %>%
                cols = (change.45:change.270)) %>%
   print()
 
+saveRDS(change_dat.glu, "./data/data-gen/glucose/glu.change.RDS")
+
+# Log-transformed
+glu.logchange <- glu.dat %>%
+  # filter(subject != "107") %>%
+  dplyr::select(subject, time, glu, supplement) %>%
+  group_by(subject, time, supplement) %>%
+  summarise(glu = mean(glu, na.rm = TRUE)) %>%
+  pivot_wider(names_from = time, 
+              values_from = glu) %>%
+  #print()
+  
+  ungroup() %>%
+  mutate(change.45 = log(min45)-log(baseline),
+         change.90 = log(min90)-log(baseline),
+         change.120 = log(min120)-log(baseline),
+         change.135 = log(min135)-log(baseline),
+         change.150 = log(min150)-log(baseline),
+         change.270 = log(min270)-log(baseline),
+         baseline = baseline - mean(baseline, na.rm = TRUE),
+         supplement = factor(supplement, levels = c("placebo", "glucose"))) %>%
+  select(subject, supplement, baseline, change.45, change.90, change.120, change.135, 
+         change.150, change.270) %>%
+  pivot_longer(names_to = "time",
+               values_to = "change",
+               cols = (change.45:change.270)) %>%
+  print()
+
+saveRDS(glu.logchange, "./data/data-gen/glucose/glu.logchange.RDS")
+
 ## Linear mixed effects model
 
[…]

### 2023-02-08 R/insuline.change.R

@@ -80,6 +80,8 @@ ins.change <- ins.dat2 %>%
                cols = (change.45:change.270)) %>%
   print()
 
+saveRDS(ins.change, "./data/data-gen/glucose/cpep.change.RDS")
+
 ins.lchange <- ins.dat2 %>%
   #filter(subject != "102") %>%
   dplyr::select(subject, time, c.pep, supplement) %>%
@@ -99,6 +101,8 @@ ins.lchange <- ins.dat2 %>%
                values_to = "change",
                cols = (change.45:change.270)) %>%
   print()
+
+saveRDS(ins.lchange, "./data/data-gen/glucose/cpep.logchange.RDS")
 
 ## Linear mixed effects model
 

### 2023-02-09 ribose-paper.RMD
small changes to glucose and insulin
updated results in the rmd

@@
 code <- read_excel("data/code_key.xlsx", na = "NA") %>%
-  select(-leg)
+  select(-leg) %>%
+  print()
 
@@
                                 if_else(sample.time == "45",
-                                "min45",
+                                "min90",
                                 if_else(sample.time == "90",
-                                        "min90",
+                                        "min120",
                                         if_else(sample.time == "120",
-                                                "min120",
+                                                "min150",
                                                 if_else(sample.time == "270",
                                                         "min270", sample.time))))))) %>%
-  mutate(time = factor(time, levels = c("pre", "baseline", "min45", "min90", "min120", "min270")),
-         c.pep = as.numeric(c.pep))
+  mutate(time = factor(time, levels = c("pre", "baseline", "min90", "min120", "min150", "min270")),
+         c.pep = as.numeric(c.pep)) %>%
+  print()
 
@@
   pivot_wider(names_from = time, 
-              values_from = c.pep)
+              values_from = c.pep) %>%
+  print()
 
@@
   ungroup() %>%
-  mutate(change.45 = min45-baseline,
-         change.90 = min90-baseline,
+  mutate(change.90 = min90-baseline,
          change.120 = min120-baseline,
+         change.150 = min150-baseline,
          change.270 = min270-baseline,
@@
          supplement = factor(supplement, levels = c("PLACEBO", "GLUCOSE"))) %>%
-  select(subject, supplement, baseline, change.45, change.90, change.120, change.270) %>%
+  select(subject, supplement, baseline, change.90, change.120, change.150, change.270) %>%
   pivot_longer(names_to = "time",
                values_to = "change",
-               cols = (change.45:change.270))
+               cols = (change.90:change.270)) %>%
+  print()
+
+saveRDS(ins.change, "./data/data-gen/glucose/cpep.change.RDS")
 
@@
   ungroup() %>%
-  mutate(change.45 = log(min45)-log(baseline),
-         change.90 = log(min90)-log(baseline),
+  mutate(change.90 = log(min90)-log(baseline),
          change.120 = log(min120)-log(baseline),
+         change.150 = log(min150)-log(baseline),
          change.270 = log(min270)-log(baseline),
@@
          supplement = factor(supplement, levels = c("PLACEBO", "GLUCOSE"))) %>%
-  select(subject, supplement, baseline, change.45, change.90, change.120, change.270) %>%
+  select(subject, supplement, baseline, change.90, change.120, change.150, change.270) %>%
   pivot_longer(names_to = "time",
                values_to = "change",
-               cols = (change.45:change.270))
+               cols = (change.90:change.270)) %>%
+  print()
+
+saveRDS(ins.lchange, "./data/data-gen/glucose/cpep.logchange.RDS")
 
@@
 
-m1 <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
+m1.1 <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
                      data = ins.lchange) # Main analysis, using log-transformed data due to heteroscedasticity
@@
 
-m1.plot <- plot(m1)
-m2.plot <- plot(m2)
-
-m1.sum <- summary(m1)
+plot(m1)
+plot(m2)
+
+summary(m1.1)
 
@@
 # For analysis
-insulin.change <- confint.m1 <- confint(emmeans(m1, specs = ~"supplement|time")) %>%
+insulin.change <- confint.m1.1 <- confint(emmeans(m1.1, specs = ~"supplement|time")) %>%
   data.frame()
@@
 saveRDS(insulin.figchange, "./data/data-gen/glucose/insulin.figchange.RDS")
+
+# For inline plotting
+cpcoef <- coef(summary(m1.1))
+pscient.120 <- round(cpcoef[6,5], 30)
+cpp.120 <- sprintf("%.5f", 3.229267e-15)
+cpscient.150 <- round(cpcoef[7,5], 30)
+cpp.150 <- sprintf("%.5f", 1.414491e-15)
+cpp.270 <- round(cpcoef[8,5], 2)
+
 ```
@@
 At baseline, there were no significant differences between training with glucose or placebo in muscle mass, isometric or isokinetic peak torque, total session volume, training intensity, total RNA content, or rRNA expression (Tab 3).
-Further, on pairwise consecutive days, there were no significant differences in any macronutrient (calories, carbohydrate, fat or protein) ingestion (Tab 4).
-Resistance training with glucose led to significant increases in blood glucose levels compared to RT with placebo, before, during and after the session (fig 3B, 120min: 43 ± 4% [*p* = `r p.120`], 135min: 32 ± 4% [*p* = `r p.135`], 150min: 32 ± 4% [*p* = `r p.150`]), measured during the final RT session.
-At 2hrs post-exercise, RT with glucose led to significantly lower blood glucose levels compared to RT with placebo (8 ± 4% [*p* = `r p.270`).
+Further, on pairwise consecutive days, there were no significant differences in mean macro nutrient (calories, carbohydrate, fat or protein) ingestion (Tab 4), or mean change in macro nutrient ingestion (Fig 3).
+Resistance training with glucose led to significant increases in blood levels glucose and c-peptide compared to RT with placebo, before, during and after the session (Fig 4B, 120min: 38 ± 4% [*p* = `r p.120`], 135min: 31 ± 4% [*p* = `r p.135`], 150min: 32 ± 4% [*p* = `r p.150`], Fig 4C, 120min: 95 ± 10% [*p* = `r cpp.120`], 150min: 87 ± 10% [*p* = `r cpp.150`]), measured during the final RT session.
+At 2hrs post-exercise, RT with glucose led to significantly lower blood glucose levels compared to RT with placebo, with a tendancy of the same in c-peptide levels (4B, 270min: -8 ± 4% [*p* = `r p.270`] and 4C, 270min: -7 ± 10 [*p* `r cpp.270`]).
 
@@
 
+```{r nutrition change fig}
+# Associated scripts:
+# - nut.change
+# - nut.change.figs
+
+nut.fig <- readRDS("./data/data-gen/nutrition/nut.fig.RDS")
+nut.fig
+
+```
+###### **Figure 3:** Mean changes in macro nutrient intake from day to day.
+Day 1 = training day 1/2(T1/T2), Day 2 = training day 3/4, Day 3 = training day 5/6, Day 4 = training day 7/8, Day 5 = training day 9/10, Day 6 = training day 11/12(T3/T4).
+Values are mean change in grams/calories ± 95% CL.
+
 ![](figures/day.timeline.png)
@@
 
-###### **Figure 3:** A) Time line for training day 6 (combination of pairwise training day 11 and 12).
+###### **Figure 4:** A) Time line for training day 6 (combination of pairwise training day 11 and 12).
 Biopsy = muscle biopsy taken pre supplementation, PRO = protein supplementation, Humac = strength test (isometric knee extension, isokinetic knee extension 60 and 240 d/s), GLU/PLAC = glucose/placebo supplementation, RT = resistance training.
@@
 Time-points: 0, 45, 90, 120, 135, 150 and 270min post protein ingestion, respectively.
-Finger draws were taken at all time-points, blood samples were taken at all time-points except from 135.
-Values are presented as estimated marginal means of mmol/L change ± 95% CI.
+C) Mean change in pmol/L of c-peptide in blood per supplement group during training day 6.
+Values are presented as estimated marginal means of mmol/L and pmol/L change ± 95% CI.
 *p* < 0.05 between groups.
@@
 
-### Total RNA and ribosomal RNA
+### Markers of ribosome biogenesis
 
@@
 
-Resistance training with glucose did not induce a higher accumulation of total RNA (fig 4A, *p* = `r trdiffp`), or rRNA  (fig 4B, *p* > 0.05), compared to RT with placebo, measured as mean change from baseline to post.
+Resistance training with glucose did not induce a higher accumulation of total RNA (fig 5A, *p* = `r trdiffp`), or rRNA  (fig 5B, *p* > 0.05), compared to RT with placebo, measured as mean change from baseline to post.
 From baseline to post-intervention, there was a mean increase in levels of total RNA by 26 and 22% after RT with glucose and placebo, respectively (*p* < 0.05, compared to baseline).
 A robust accumulation was also observed in ribosomal RNA expression in both RT with glucose and placebo, with mean increases between 34-43% (GLU) and 33-41% (PLAC) (*p* < 0.05, compared to baseline) in the four rRNA’s.
-The expression of 47S pre-rRNA increased by 37 and 59% with glucose and placebo respectively, with only placebo reaching significantly higher levels at post compared to baseline (fig 4B, *p* = `r rr47tdiffp`).
+The expression of 47S pre-rRNA increased by 37 and 59% with glucose and placebo respectively, with only placebo reaching significantly higher levels at post compared to baseline (fig 45, *p* = `r rr47tdiffp`).
 
@@
 
-###### **Figure 4:** A) Changes in Total RNA, B) Changes in ribosomal RNA.
+###### **Figure 5:** A) Changes in Total RNA, B) Changes in ribosomal RNA.
 Baseline = Training day 1/2, Post = Training day 11/12.
@@
 Glucose n = 13, placebo n = 13.
-
-### Protein signals
 

### 2023-02-09 R/insuline.change.R

@@ -34,14 +34,14 @@ ins.dat2 <- ins.dat %>%
                         if_else(sample.time == "0.1",
                                 "baseline",
                                 if_else(sample.time == "45",
-                                "min45",
+                                "min90",
                                 if_else(sample.time == "90",
-                                        "min90",
+                                        "min120",
                                         if_else(sample.time == "120",
-                                                "min120",
+                                                "min150",
                                                 if_else(sample.time == "270",
                                                         "min270", sample.time))))))) %>%
-  mutate(time = factor(time, levels = c("pre", "baseline", "min45", "min90", "min120", "min270")),
+  mutate(time = factor(time, levels = c("pre", "baseline", "min90", "min120", "min150", "min270")),
          c.pep = as.numeric(c.pep)) %>%
   print()
 
@@ -68,16 +68,16 @@ ins.change <- ins.dat2 %>%
   pivot_wider(names_from = time, 
               values_from = mean.pep) %>%
   ungroup() %>%
-  mutate(change.45 = min45-baseline,
-         change.90 = min90-baseline,
+  mutate(change.90 = min90-baseline,
          change.120 = min120-baseline,
+         change.150 = min150-baseline,
          change.270 = min270-baseline,
          baseline = baseline - mean(baseline, na.rm = TRUE),
          supplement = factor(supplement, levels = c("PLACEBO", "GLUCOSE"))) %>%
-  select(subject, supplement, baseline, change.45, change.90, change.120, change.270) %>%
+  select(subject, supplement, baseline, change.90, change.120, change.150, change.270) %>%
   pivot_longer(names_to = "time",
                values_to = "change",
-               cols = (change.45:change.270)) %>%
+               cols = (change.90:change.270)) %>%
   print()
 
 saveRDS(ins.change, "./data/data-gen/glucose/cpep.change.RDS")
@@ -90,16 +90,16 @@ ins.lchange <- ins.dat2 %>%
   pivot_wider(names_from = time, 
               values_from = mean.pep) %>%
   ungroup() %>%
-  mutate(change.45 = log(min45)-log(baseline),
-         change.90 = log(min90)-log(baseline),
+  mutate(change.90 = log(min90)-log(baseline),
          change.120 = log(min120)-log(baseline),
+         change.150 = log(min150)-log(baseline),
          change.270 = log(min270)-log(baseline),
[…]

### 2023-02-09 R/nut.change.R

@@ -1,0 +1,330 @@
+#### Nutrition change analysis
+
+# Author: KL
+# Project: Ribose
+
+# This script calculates change scores and models the change for protein, protein/kg/body weight,
+# fat, carbohydrates and calories.
+
+# Packages
+library(tidyverse); library(lme4); library(lmerTest); library(emmeans)
+
+
+# Data
+
+fp.weight <- readRDS("./data/data-gen/dxa/fp.weight.RDS")
+
+nut.res <- readRDS("./data/data-gen/dxa/nutha.RDS")
+
+### Macro nutrient change analysis
+# The code beneath tests whether there are differences in macro nutrient intake from pre to post, 
+# divided into "protein", "fat" and "carbohydrates", and total calories in a linear model. The main interest is if there is difference between glucose and 
+# placebo.
+
+## Protein
+
+# Log transforming for analysis
+pro.lchange <- nut.res %>%
+  select(time = timepoint, subject, supplement = group, protein) %>%
+  group_by(subject, time, supplement) %>%
+  #summarise(mean.pro = mean(protein, na.rm = TRUE)) %>%
+  pivot_wider(names_from = time,
+              values_from = protein) %>%
+  ungroup() %>%
+  mutate(change.d2 = log(D2)-log(D1),
+         change.d3 = log(D3)-log(D1),
+         change.d4 = log(D4)-log(D1),
+         change.d5 = log(D5)-log(D1),
+         change.d6 = log(D6)-log(D1),
+         d1 = D1-mean(D1, na.rm = TRUE),
+         supplement = factor(supplement, levels = c("placebo", "glucose"))) %>%
+  select(subject, supplement, d1, change.d2, change.d3, change.d4, change.d5, change.d6) %>%
+  pivot_longer(names_to = "time",
+               values_to = "change",
+               cols = (change.d2:change.d6)) 
+
+# Un-logged change for figures
+pro.change <- nut.res %>%
+  select(time = timepoint, subject, supplement = group, protein) %>%
+  group_by(subject, time, supplement) %>%
+  #summarise(mean.pro = mean(protein, na.rm = TRUE)) %>%
+  pivot_wider(names_from = time,
+              values_from = protein) %>%
+  ungroup() %>%
+  mutate(change.d2 = D2-D1,
[…]

### 2023-02-09 ribose-paper.RMD
further updates to results

@@
 ```{r packages}
-library(tidyverse);library(readxl);library(lme4);library(emmeans);library(dplyr);library(cowplot);library(knitr);library(broom);library(rmarkdown);library(tinytex);library(tidyr);library(arsenal);library(magrittr);library(dabestr);library(nlme)
+library(tidyverse);library(readxl);library(lme4);library(emmeans);library(dplyr);library(cowplot);library(knitr);library(broom);library(rmarkdown);library(tinytex);library(tidyr);library(arsenal);library(magrittr);library(dabestr);library(nlme); library(pbkrtest)
 ```
@@
 ```{r glucose data + model}
-# Scripts: gluc.responsechange.R, gluc.fig
+## Scripts:
+# - gluc.responsechange.R
 
 # Data
-gluc.dat <- read_excel("./data/glucose/fingerstick.xlsx", na = "NA")
-gluc.dat$sample_time <- as.character(gluc.dat$sample_time)
-
-# Data handling
-glu.dat <- gluc.dat %>%
-        mutate(time = if_else(sample_time == "0",
-                              "baseline",
-                              if_else(sample_time == "45",
-                                      "min45",
-                                      if_else(sample_time == "90",
-                                              "min90",
-                                              if_else(sample_time == "120",
-                                                      "min120",
-                                                      if_else(sample_time == "135",
-                                                              "min135",
-                                                              if_else(sample_time == "150",
-                                                                      "min150",
-                                                                      if_else(sample_time == "270",
-                                                                              "min270", sample_time))))))),
-               time = factor(time, levels = c("baseline", "min45", "min90", "min120", "min135", "min150", "min270"))) %>%
-        mutate(glu = as.numeric(glu),
-               lak = as.numeric(lak))
-
-# Change data
-change_dat.glu <- glu.dat %>%
-        dplyr::select(subject, time, glu, supplement) %>%
-        group_by(subject, time, supplement) %>%
-        summarise(glu = mean(glu, na.rm = TRUE)) %>%
-        pivot_wider(names_from = time, 
-                    values_from = glu) %>%
-        #print()
-
-        ungroup() %>%
-        mutate(change.45 = log(min45)-log(baseline),
-               change.90 = log(min90)-log(baseline),
-               change.120 = log(min120)-log(baseline),
-               change.135 = log(min135)-log(baseline),
-               change.150 = log(min150)-log(baseline),
-               change.270 = log(min270)-log(baseline),
-               baseline = baseline - mean(baseline, na.rm = TRUE),
-               supplement = factor(supplement, levels = c("placebo", "glucose"))) %>%
-        select(subject, supplement, baseline, change.45, change.90, change.120, change.135, 
-               change.150, change.270) %>%
-        pivot_longer(names_to = "time",
-                     values_to = "change",
-                     cols = (change.45:change.270))
-
-# Model
-m1 <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = change_dat.glu)
-resplot.glu <- plot(m1)
-sum.glu <- summary(m1)
-
-# Emmeans
-confint.m1 <- confint(emmeans(m1, specs = ~"supplement|time")) %>%
+glu.change <- readRDS("./data/data-gen/glucose/glu.change.RDS") # Untransformed data, used in figures
+glu.logchange <- readRDS("./data/data-gen/glucose/glu.logchange.RDS") # Log-transformed data, used in the analysis
+
+## Model
+# Untransformed
+glu.mod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
+                     data = glu.change)
+glu.res <- plot(glu.mod)
+glu.sum <- summary(glu.mod)
+
+# Log-transformed
+glu.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
+                     data = glu.logchange)
+glu.lres <- plot(glu.lmod)
+glu.lsum <- summary(glu.lmod)
+
+## Emmeans
+# Untransformed
+glu.emm <- confint(emmeans(glu.mod, specs = ~"supplement|time")) %>%
+        data.frame()
+
+# Log-transformed
+glu.lemm <- confint(emmeans(glu.lmod, specs = ~"supplement|time")) %>%
         data.frame()
@@
 # For inline plotting
-bcoef <- coef(summary(m1))
+bcoef <- coef(summary(glu.lmod))
 pscient.120 <- round(bcoef[8,5], 30)
@@
 ```{r c-peptide data + model}
+## Scripts
+# - insulin.change.R
+
 ## Data
-
-ins.dat <- read_excel("./data/glucose/ribose.insulin.xlsx", na = "NA")
-
-code <- read_excel("data/code_key.xlsx", na = "NA") %>%
-  select(-leg) %>%
-  print()
-
-## Data handling
-
-ins.dat$sample.time <- as.character(ins.dat$sample.time)
-
-ins.dat2 <- ins.dat %>%
-  inner_join(code) %>%
-  mutate(sample.time = as.character(sample.time)) %>%
-  mutate(time = if_else(sample.time == "0",
-                        "pre",
-                        if_else(sample.time == "0.1",
-                                "baseline",
-                                if_else(sample.time == "45",
-                                "min90",
-                                if_else(sample.time == "90",
-                                        "min120",
-                                        if_else(sample.time == "120",
-                                                "min150",
-                                                if_else(sample.time == "270",
-                                                        "min270", sample.time))))))) %>%
-  mutate(time = factor(time, levels = c("pre", "baseline", "min90", "min120", "min150", "min270")),
-         c.pep = as.numeric(c.pep)) %>%
-  print()
-
-## Baseline analysis
-# Comparing pre measurements to post day baseline
-
-base.ins <- ins.dat2 %>%
-  filter(time %in% c("pre", "baseline")) %>%
-  select(subject, time, supplement, c.pep) %>%
-  group_by(supplement) %>%
-  pivot_wider(names_from = time, 
-              values_from = c.pep) %>%
-  print()
-
-ins.ttest <- t.test(base.ins$pre, base.ins$baseline, paired = TRUE)
-
-## Change data
-
-ins.change <- ins.dat2 %>%
-  #filter(subject != "102") %>%
-  dplyr::select(subject, time, c.pep, supplement) %>%
-  group_by(subject, time, supplement) %>%
-  summarise(mean.pep = mean(c.pep, na.rm = TRUE)) %>%
-  pivot_wider(names_from = time, 
-              values_from = mean.pep) %>%
-  ungroup() %>%
-  mutate(change.90 = min90-baseline,
-         change.120 = min120-baseline,
-         change.150 = min150-baseline,
-         change.270 = min270-baseline,
-         baseline = baseline - mean(baseline, na.rm = TRUE),
-         supplement = factor(supplement, levels = c("PLACEBO", "GLUCOSE"))) %>%
-  select(subject, supplement, baseline, change.90, change.120, change.150, change.270) %>%
-  pivot_longer(names_to = "time",
-               values_to = "change",
-               cols = (change.90:change.270)) %>%
-  print()
-
-saveRDS(ins.change, "./data/data-gen/glucose/cpep.change.RDS")
-
-ins.lchange <- ins.dat2 %>%
-  #filter(subject != "102") %>%
-  dplyr::select(subject, time, c.pep, supplement) %>%
-  group_by(subject, time, supplement) %>%
-  summarise(mean.pep = mean(c.pep, na.rm = TRUE)) %>%
-  pivot_wider(names_from = time, 
-              values_from = mean.pep) %>%
-  ungroup() %>%
-  mutate(change.90 = log(min90)-log(baseline),
-         change.120 = log(min120)-log(baseline),
-         change.150 = log(min150)-log(baseline),
-         change.270 = log(min270)-log(baseline),
-         baseline = baseline - mean(baseline, na.rm = TRUE),
-         supplement = factor(supplement, levels = c("PLACEBO", "GLUCOSE"))) %>%
-  select(subject, supplement, baseline, change.90, change.120, change.150, change.270) %>%
-  pivot_longer(names_to = "time",
-               values_to = "change",
-               cols = (change.90:change.270)) %>%
-  print()
-
-saveRDS(ins.lchange, "./data/data-gen/glucose/cpep.logchange.RDS")
+cpep.change <- readRDS("./data/data-gen/glucose/cpep.change.RDS") # Untransformed data, for figures
+cpep.lchange <- readRDS("./data/data-gen/glucose/cpep.logchange.RDS") # Log-transformed data, for analysis
 
 ## Linear mixed effects model
-
-m1.1 <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = ins.lchange) # Main analysis, using log-transformed data due to heteroscedasticity
-
-m2 <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = ins.change) # For emmeans and figure illustration
-
-plot(m1)
-plot(m2)
-
-summary(m1.1)
+# Untransformed
+cpep.mod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
+                     data = cpep.change) 
+cpep.res <- plot(cpep.mod)
+cpep.sum <- summary(cpep.mod)
+
+# Log-transformed
+cpep.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
+                     data = cpep.lchange) 
+cpep.lres <- plot(cpep.lmod)
+cpep.lsum <- summary(cpep.lmod)
 
 ## Emmeans
-
-library(pbkrtest)
-
-# For analysis
-insulin.change <- confint.m1.1 <- confint(emmeans(m1.1, specs = ~"supplement|time")) %>%
-  data.frame()
-
-saveRDS(insulin.change, "./data/data-gen/glucose/insulin.change.RDS")
-
-# For figure
-insulin.figchange <- confint.m2 <- confint(emmeans(m2, specs = ~"supplement|time")) %>%
-  data.frame()
-
-saveRDS(insulin.figchange, "./data/data-gen/glucose/insulin.figchange.RDS")
+# Untransformed
+cpep.emm <- confint(emmeans(cpep.mod, specs = ~"supplement|time")) %>%
+  data.frame()
+
+# Log-transformed
+cpep.lemm <- confint(emmeans(cpep.lmod, specs = ~"supplement|time")) %>%
+  data.frame()
 
 # For inline plotting
-cpcoef <- coef(summary(m1.1))
+cpcoef <- coef(summary(cpep.lmod))
 pscient.120 <- round(cpcoef[6,5], 30)
@@
 ```{r nutrition table}
+## Scripts
+# - nut.table.R
+# - nut.analysis.R
+# - dxa.R
 
@@
 ```{r nutrition analysis}
+## Scripts
+# - nut.analysis.R
+# - dxa.R
+
 #summarise macro
@@
 
-```{r nutrition change fig}
+```{r nutrition change analysis and fig}
 # Associated scripts:
@@
 ```{r glucose + c-peptide fig}
-# Figure
+## Scripts
+# - gluc.fig.R
+# - cpep.figs.R
+
+# Separate figures
 pos <- position_dodge(width = 0.2)
+
 gluc.fig <- readRDS("./data/data-gen/glucose/gluc.fig.RDS")
@@
 
+# Joined figure
 plot_grid(gluc.fig,
@@
 
-```{r total RNA data}
-
-code <- read_excel("./data/code.key.xlsx")
-rna_dat <- readRDS("./data/data-gen/rna/rna_complete.RDS")
-
-```
-
-```{r total RNA data handling}
-
-# Scripts
-# totrna_analysis
-rna_dat2 <- rna_dat %>%
-        inner_join(code) %>%
-        mutate(time = if_else(time %in% c("T1", "T2"), "pre", "post"), 
-               time = factor(time, levels = c("pre", "post"))) %>%
-        # filter(outlier == "in") %>%
-        group_by(subject, time, rep, outlier, technical, biopsy, supplement) %>%
-        summarise(weight = mean(weight), 
-                  RNA = mean(RNA_tot)) %>%
-        ungroup() %>%
-        mutate(weight.mc = weight/mean(weight)) %>%
-        mutate(RNA.weight = RNA / weight)
-
-#### Basic model 
-change_dat <- rna_dat2 %>%
-   #  filter(technical == "in", 
-   #        outlier == "in") %>%
-    dplyr::select(subject:rep, supplement, RNA.weight) %>%
-
-    group_by(subject, time, supplement) %>%
-    summarise(RNA.weight = mean(RNA.weight, na.rm = TRUE)) %>%
-
-    pivot_wider(names_from = time, 
-                values_from = RNA.weight) %>%
-    ungroup() %>%
-    mutate(change = log(post) - log(pre), 
-           pre = pre - mean(pre, na.rm = TRUE), 
-           supplement = factor(supplement, levels = c("PLACEBO", "GLUCOSE")))
-
-```
-
 ```{r total RNA model}
-# Scripts
-# totrna_analysis
+## Scripts
+# - totrna.cleanup.R
+# - totrna.analysis.R
+
+# Data
+trna.lchange <- readRDS("./data/data-gen/rna/totrna.change.RDS")
+
 # Create model: 
-m1 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
-           data = change_dat)
-
-resplot.rna <- plot(m1)
-sum.rna <- summary(m1)
+trna.lmod <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
+           data = trna.lchange)
+
+trna.lres <- plot(trna.lmod)
+trna.lsum <- summary(trna.lmod)
 
 ### Get estimated means from the model
-confint.totrna <- confint(emmeans(m1, specs = ~"supplement")) %>%
-    data.frame() 
+trna.emm <- confint(emmeans(trna.lmod, specs = ~"supplement")) %>%
+    data.frame()
+
 # For inline reporting
-trcoef <- coef(summary(m1))
+trcoef <- coef(summary(trna.lmod))
 trdiff <- round(trcoef[3,1], 2)
@@
 
-```{r rRNA data}
-code_key <- read_excel("./data/code.key.xlsx") %>%
-        dplyr::select(subject, supplement, time)
-qpcr_data  <- readRDS("./data/data-gen/rna/qpcr-data.RDS")
-samples <- read_excel("./data/rna/RNA.raw.xlsx") %>%
-        dplyr::select(subject, time_rep, sample, weight)
-
-```
-
-```{r rRNA data handling}
-# Scripts
+```{r rRNA data model}
+## Scripts
 # - qpcr-import
@@
 
-nf <- qpcr_data %>%
-        mutate(sample = as.numeric(sample)) %>% 
-        inner_join(samples) %>%
-        separate(time_rep, into = c("time","rep"), sep = "rna") %>%
-        mutate(rep = paste0("cdna", rep)) %>%
-        inner_join(code_key) %>% 
-        mutate(time = if_else(time %in% c("T1", "T2"), "Pre", "Post"), 
-               time = factor(time, levels = c("Pre", "Post")), 
-               expr = eff ^ -cq) %>%
-
-        filter(target %in% c("Lambda F2R2", "Lambda F3R3"), 
-               cq < 35) %>%
-        separate(target, into = c("target", "nf_primer")) %>%
-
-        mutate(nf.w = expr * weight, 
-               nf.w = nf.w / max(nf.w)) %>%
-
-        dplyr::select(sample, subject, time, rep, supplement, nf.w, nf_primer)
-
-qpcr_data2 <- qpcr_data %>%
-        mutate(sample = as.numeric(sample)) %>% 
-        inner_join(samples) %>%
-        inner_join(nf) %>%
-        separate(time_rep, into = c("time","rep"), sep = "rna") %>%
-        mutate(rep = paste0("cdna", rep)) %>%
-        inner_join(code_key) %>% 
-        mutate(time = if_else(time %in% c("T1", "T2"), "Pre", "Post"), 
-               time = factor(time, levels = c("Pre", "Post")), 
-               expr = eff ^ -cq) %>%
-
-        filter(!(target %in% c("Lambda F2R2", "Lambda F3R3"))) %>%   
-        inner_join(nf)
-
-saveRDS(qpcr_data2, "./data/data-gen/rna/qpcr_data2.RDS")
-
-# Plotting mean expression per treatment for rRNA
-initmod <- qpcr_data2 %>%
-        mutate(norm.expr = log(expr / nf.w)) %>%
-        group_by(time, supplement, target) %>%
-        summarise(m = mean(norm.expr, na.rm = TRUE))  %>%
-        #print()
-
-        ggplot(aes(time, m, color = supplement, group = supplement)) + 
-        geom_line() +
-        geom_point() +
-        facet_wrap( ~ target, scales = "free")
-qdat <- qpcr_data2 %>%
-        mutate(target = gsub("rRNA", "", target),
-               target = gsub("  ", " ", target),
-               target = paste0("trg_", target)) %>%
-        separate(target, into = c("target", "primer"), sep = " ") %>%
-
-        # Create a weight-normalized variable
-        mutate(nf.expr = log(expr / nf.w), 
-               nf.w = scale(nf.w),
-               # Technical random effect
-               technical = paste(subject, time, supplement, rep, sep = "_"),
-               biological = paste0("S", sample)) %>%
-        filter(!(target %in% c("trg_MHC1", "trg_MHC2A", "trg_MHC2X")))
-
-## Change data
-rrna18 <- qdat %>%
-        filter(target == "trg_18s")
-rrna28 <- qdat %>%
-        filter(target == "trg_28s")
-rrna5.8 <- qdat %>%
-        filter(target == "trg_5.8s")
-rrna5 <- qdat %>%
-        filter(target == "trg_5s")
-rrna47 <- qdat %>%
-        filter(target == "trg_47s")
-
-# Change scores
-# 18S
-change.18 <- rrna18 %>%
-        dplyr::select(subject, time, rep, nf.expr, supplement) %>%
-        group_by(subject, time, supplement) %>%
-        summarise(nf.expr = mean(nf.expr, na.rm = TRUE)) %>%
-        pivot_wider(names_from = time,
-                    values_from = nf.expr) %>%
-        ungroup() %>%
-        # print()
-        mutate(change = Post-Pre,
-               pre = Pre - mean(Pre, na.rm = TRUE),
-               supplement = factor(supplement, levels = c("PLACEBO", "GLUCOSE"))) 
-# 28S
-change.28 <- rrna28 %>%
-        dplyr::select(subject, time, rep, nf.expr, supplement) %>%
-        group_by(subject, time, supplement) %>%
-        summarise(nf.expr = mean(nf.expr, na.rm = TRUE)) %>%
-        pivot_wider(names_from = time,
-                    values_from = nf.expr) %>%
-        ungroup() %>%
-        # print()
-        mutate(change = Post-Pre,
-               pre = Pre - mean(Pre, na.rm = TRUE),
-               supplement = factor(supplement, levels = c("PLACEBO", "GLUCOSE"))) 
-# 5.8S
-change.5.8 <- rrna5.8 %>%
-        dplyr::select(subject, time, rep, nf.expr, supplement) %>%
-        group_by(subject, time, supplement) %>%
-        summarise(nf.expr = mean(nf.expr, na.rm = TRUE)) %>%
-        pivot_wider(names_from = time,
-                    values_from = nf.expr) %>%
-        ungroup() %>%
-        # print()
-        mutate(change = Post-Pre,
-               pre = Pre - mean(Pre, na.rm = TRUE),
-               supplement = factor(supplement, levels = c("PLACEBO", "GLUCOSE")))
-# 5S
-change.5 <- rrna5 %>%
-        dplyr::select(subject, time, rep, nf.expr, supplement) %>%
-        group_by(subject, time, supplement) %>%
-        summarise(nf.expr = mean(nf.expr, na.rm = TRUE)) %>%
-        pivot_wider(names_from = time,
-                    values_from = nf.expr) %>%
-        ungroup() %>%
-        # print()
-        mutate(change = Post-Pre,
-               pre = Pre - mean(Pre, na.rm = TRUE),
-               supplement = factor(supplement, levels = c("PLACEBO", "GLUCOSE")))
-# 47S
-change.47 <- rrna47 %>%
-        dplyr::select(subject, time, rep, nf.expr, supplement) %>%
-        group_by(subject, time, supplement) %>%
-        summarise(nf.expr = mean(nf.expr, na.rm = TRUE)) %>%
-        pivot_wider(names_from = time,
-                    values_from = nf.expr) %>%
-        ungroup() %>%
-        # print()
-        mutate(change = Post-Pre,
-               pre = Pre - mean(Pre, na.rm = TRUE),
-               supplement = factor(supplement, levels = c("PLACEBO", "GLUCOSE")))
-```
-
-```{r rRNA data modelling}
-# Scripts
-# - qpcr-import
-# - qpcr-analysis
-# - rrna.change
-# - rrnachange.figs
+# Log-transformed data
+lchange.18 <- readRDS("./data/data-gen/rna/lchange.18.RDS")
+lchange.28 <- readRDS("./data/data-gen/rna/lchange.28.RDS")
+lchange.58 <- readRDS("./data/data-gen/rna/lchange.58.RDS")
+lchange.5 <- readRDS("./data/data-gen/rna/lchange.5.RDS")
+lchange.47 <- readRDS("./data/data-gen/rna/lchange.47.RDS")
 
@@
 # 18S
-m1 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
-                     data = change.18)
-resplot.18 <- plot(m1)
-sum.18 <- summary(m1)
+lmod.18 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
+                     data = lchange.18)
+lres.18 <- plot(lmod.18)
+lsum.18 <- summary(lmod.18)
 
 # 28S
-m2 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
-                     data = change.28)
-resplot.28 <- plot(m2)
-sum.28 <- summary(m2)
+lmod.28 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
+                     data = lchange.28)
+lres.28 <- plot(lmod.28)
+lsum.28 <- summary(lmod.28)
 
 # 5.8S
-m3 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
-                     data = change.5.8)
-resplot.58 <- plot(m3)
-sum.58 <- summary(m3)
+lmod.58 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
+                     data = lchange.58)
+lres.58 <- plot(lmod.58)
+lsum.58 <- summary(lmod.58)
 
 # 5S
-m4 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
-                     data = change.5)
-resplot.5 <- plot(m4)
-sum.5 <- summary(m4)
+lmod.5 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
+                     data = lchange.5)
+lres.5 <- plot(lmod.5)
+lsum.5 <- summary(lmod.5)
 
 # 47S
-m5 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
-                     data = change.47)
-resplot.47 <- plot(m5)
-sum.47 <- summary(m5)
+lmod.47 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
+                     data = lchange.47)
+lres.47 <- plot(lmod.47)
+lsum.47 <- summary(lmod.47)
 
@@
 # 18S
-confint.18 <- confint(emmeans(m1, specs = ~"supplement")) %>%
+lemm.18 <- confint(emmeans(lmod.18, specs = ~"supplement")) %>%
         data.frame()
@@
 # 28S
-confint.28 <- confint(emmeans(m2, specs = ~"supplement")) %>%
+lemm.28 <- confint(emmeans(lmod.28, specs = ~"supplement")) %>%
         data.frame()
@@
 # 5.8S
-confint.5.8 <- confint(emmeans(m3, specs = ~"supplement")) %>%
+lemm.58 <- confint(emmeans(lmod.58, specs = ~"supplement")) %>%
         data.frame()
@@
 # 5S
-confint.5 <- confint(emmeans(m4, specs = ~"supplement")) %>%
+lemm.5 <- confint(emmeans(lmod.5, specs = ~"supplement")) %>%
         data.frame()
@@
 # 47S
-confint.47 <- confint(emmeans(m5, specs = ~"supplement")) %>%
+lemm.47 <- confint(emmeans(lmod.47, specs = ~"supplement")) %>%
         data.frame()
@@
 # 18S
-rr18coef <- coef(summary(m1))
+rr18coef <- coef(summary(lmod.18))
 rr18diffp <- round(rr18coef[3,5], 3)
@@
 # 28S
-rr28coef <- coef(summary(m2))
+rr28coef <- coef(summary(lmod.28))
 rr28diffp <- round(rr28coef[3,5], 3)
@@
 # 5.8S
-rr58coef <- coef(summary(m3))
+rr58coef <- coef(summary(lmod.58))
 rr58diffp <- round(rr58coef[3,5], 3)
@@
 # 5S
-rr5coef <- coef(summary(m4))
+rr5coef <- coef(summary(lmod.5))
 rr5diffp <- round(rr5coef[3,5], 3)
@@
 # 47S
-rr47coef <- coef(summary(m5))
+rr47coef <- coef(summary(lmod.47))
 rr47diffp <- round(rr47coef[3,5], 3)
@@
 ```{r western modelling}
-# Script: western.change
+## Script: 
+# - western.import.R
+# - western.change.R
 
@@
 
-prot.dat <- readRDS("./data/data-gen/protein/prot.dat.RDS")
-
-# Change analysis
-
-#prot.dat %>%
- # group_by(time, supplement, target) %>%
-  #mutate(norm.sign = log(tpl/norm.sign)) %>% # Log transforms and normalises
-  # norm.sign by tpl
-  #summarise(m = mean(norm.sign, na.rm = TRUE)) %>%
-  #ggplot(aes(time, m, color = supplement, group = supplement)) +
-  #geom_line() +
-  #geom_point() +
-  #facet_wrap( ~ target, scales = "free")
-
-# Change score
-
-pchange <- prot.dat %>%
-  dplyr::select(subject, time, sample.id, target, norm.sign, supplement) %>%
-  group_by(subject, time, target, supplement) %>%
-  summarise(mns = mean(norm.sign, na.rm = TRUE)) %>%
-  pivot_wider(names_from = time,
-              values_from = mns) %>%
-  ungroup() %>%
-  mutate(change = log(post)-log(pre),
-         pre = pre - mean(pre, na.rm = TRUE),
-         supplement = factor(supplement, levels = c("PLACEBO", "GLUCOSE")))
+prot.dat <- readRDS("./data/data-gen/protein/prot.dat.RDS") # cleaned up and quantified western data
+prot.out <- readRDS("./data/data-gen/protein/protein_complete.RDS") # Cleaned up and quantified western data, checked for outliers
+prot.change <- readRDS("./data/data-gen/protein/prot.change.RDS") # change-score data
 
@@
 
-cmyc <- pchange %>%
+cmyc <- prot.change %>%
   filter(target == "cmyc")
 
-ubf <- pchange %>%
+ubf <- prot.change %>%
   filter(target == "ubf")
 
-rps6 <- pchange %>%
+rps6 <- prot.change %>%
   filter(target == "rps6")
@@
 
-```{r training data}
-lp.load <- read_excel("./data/training/ribose_training.xlsx", na = "NA") %>%
-        select(subject, timepoint, supplement, exercise, load) %>%
-        filter(exercise != "KE")
-
-lp.rm <- read_excel("./data/tests/ribose.1rm.xlsx", na = "NA") %>%
-        select(subject, supplement, exercise, rm) %>%
-        filter(exercise != "KE")
-
-tot.load <- read_excel("./data/training/ribose_training.xlsx", na = "NA") %>%
-        select(subject, timepoint, supplement, exercise, load)
-
-tot.rm <- read_excel("./data/tests/ribose.1rm.xlsx", na = "NA") %>%
-        select(subject, supplement, exercise, rm)
-
-tot.vol <- read_excel("./data/training/ribose_volume.xlsx", na = "NA") %>%
-        select(subject, timepoint, tot.volume, supplement)
-
-```
-
-```{r training data handling}
-# Scripts
-# tot.vol.change
-# Trainingint.linear
-# rpe-analysis
-
-### Training volume
-# Data handling
-tot.volh <- tot.vol %>%
-        mutate(time = if_else(timepoint %in% c("T1", "T2"),
-                              "baseline",
-                              if_else(timepoint %in% c("D3", "D4"),
-                                      "session2",
-                                      if_else(timepoint %in% c("D5", "D6"),
-                                              "session3",
-                                              if_else(timepoint %in% c("D7", "D8"),
-                                                      "session4",
-                                                      if_else(timepoint %in% c("D9", "D10"),
-                                                              "session5",
-                                                              if_else(timepoint %in% c("T3", "T4"),
-                                                                      "session6", timepoint))))))) %>%
-        mutate(time = factor(time, levels = c("baseline", "session1", "session2", "session3", "session4", "session5", "session6")),
-               supplement = factor(supplement, levels = c("placebo", "glucose")))
-
-## Absolute data
-vol.exp <- tot.volh %>%
-        select(subject, supplement, time, tot.volume) %>%
-        group_by(supplement, time) %>%
-        summarise(mean.vol = mean(tot.volume),
-                  sd.vol = sd(tot.volume))
-
-# Fold-change
-change_dat <- tot.volh %>%
-        dplyr::select(subject, time, tot.volume, supplement) %>%
-        group_by(subject, time, supplement) %>%
-        summarise(tot.volume = mean(tot.volume, na.rm = TRUE)) %>%
-        pivot_wider(names_from = time, 
-                    values_from = tot.volume) %>%
-        # print()
-
-        ungroup() %>%
-        mutate(change.2 = log(session2)-log(baseline),
-               change.3 = log(session3)-log(baseline),
-               change.4 = log(session4)-log(baseline),
-               change.5 = log(session5)-log(baseline),
-               change.6 = log(session6)-log(baseline),
-               baseline = baseline - mean(baseline, na.rm = TRUE),
-               supplement = factor(supplement, levels = c("placebo", "glucose"))) %>%
-        select(subject, supplement, baseline, change.2, change.3, change.4, change.5, change.6) %>%
-        pivot_longer(names_to = "time",
-                     values_to = "change",
-                     cols = (change.2:change.6))
-
-### RPE
-## Data handling
-ss.dat <- read_excel("./data/training/ribose_rpe.xlsx", na = "NA") %>%
-        select(subject, timepoint, session.score, supplement) %>%
-        mutate(time = if_else(timepoint %in% c("T1", "T2"),
-                              "baseline",
-                              if_else(timepoint %in% c("D3", "D4"),
-                                      "session2",
-                                      if_else(timepoint %in% c("D5", "D6"),
-                                              "session3",
-                                              if_else(timepoint %in% c("D7", "D8"),
-                                                      "session4",
-                                                      if_else(timepoint %in% c("D9", "D10"),
-                                                              "session5",
-                                                              if_else(timepoint %in% c("T3", "T4"),
-                                                                      "post", timepoint))))))) %>%
-        mutate(time = factor(time, levels = c("baseline", "session1", "session2", "session3", "session4", "session5", "post")),
-               supplement = factor(supplement, levels = c("placebo", "glucose")))
-
-## Absolute data
-ss.exp <- ss.dat %>%
-        select(subject, supplement, time, session.score) %>%
-        group_by(supplement, time) %>%
-        summarise(mean.ss = mean(session.score),
-                  sd.ss = sd(session.score))
-
-# Change data
-change_dat.ss <- ss.dat %>%
-        dplyr::select(subject, time, session.score, supplement) %>%
-        group_by(subject, time, supplement) %>%
-        summarise(session.score = mean(session.score, na.rm = TRUE)) %>%
-        pivot_wider(names_from = time, 
-                    values_from = session.score) %>%
-        #print()
-
-        ungroup() %>%
-        mutate(change.2 = log(session2)-log(baseline),
-               change.3 = log(session3)-log(baseline),
-               change.4 = log(session4)-log(baseline),
-               change.5 = log(session5)-log(baseline),
-               change.post = log(post)-log(baseline),
-               baseline = baseline - mean(baseline, na.rm = TRUE),
-               supplement = factor(supplement, levels = c("placebo", "glucose"))) %>%
-        select(subject, supplement, baseline, change.2, change.3, change.4, change.5, change.post) %>%
-        pivot_longer(names_to = "time",
-                     values_to = "change",
-                     cols = (change.2:change.post))
-
-### Training intensity
-# Data handling
-tot.loadh <- tot.load %>%
-        mutate(time = if_else(timepoint %in% c("T1", "T2"),
-                              "baseline",
-                              if_else(timepoint %in% c("D3", "D4"),
-                                      "session2",
-                                      if_else(timepoint %in% c("D5", "D6"),
-                                              "session3",
-                                              if_else(timepoint %in% c("D7", "D8"),
-                                                      "session4",
-                                                      if_else(timepoint %in% c("D9", "D10"),
-                                                              "session5",
-                                                              if_else(timepoint %in% c("T3", "T4"),
-                                                                      "session6", timepoint))))))) %>%
-        mutate(time = factor(time, levels = c("baseline", "session1", "session2", "session3", "session4", "session5", "session6")),
-               supplement = factor(supplement, levels = c("placebo", "glucose")))
-
-# Join data and calulate %1RM
-tot.joined <- tot.loadh %>%
-        full_join(tot.rm) %>%
-        mutate(p.rm = (load/rm)*100)
-
-mean.joined <- tot.joined %>%
-        select(subject, supplement, time, p.rm) %>%
-        group_by(supplement, time) %>%
-        summarise(mean.prm = mean(p.rm),
-                  sd.prm = sd(p.rm))
-
-## Change data
-tot.change <- tot.joined %>%
-        dplyr::select(subject, time, p.rm, supplement) %>%
-        group_by(subject, time, supplement) %>%
-        summarise(mean.prm = mean(p.rm)) %>%
-        pivot_wider(names_from = time,
-                    values_from = mean.prm) %>%
-        ungroup() %>%
-        mutate(change.2 = log(session2)-log(baseline),
-               change.3 = log(session3)-log(baseline),
-               change.4 = log(session4)-log(baseline),
-               change.5 = log(session5)-log(baseline),
-               change.6 = log(session6)-log(baseline),
-               baseline = baseline - mean(baseline, na.rm = TRUE),
-               supplement = factor(supplement, levels = c("placebo", "glucose"))) %>%
-        select(subject, supplement, baseline, change.2, change.3, change.4, change.5, change.6) %>%
-        pivot_longer(names_to = "time",
-                     values_to = "change",
-                     cols = (change.2:change.6))
-```
-
-```{r training data modelling}
-# Scripts
-# tot.vol.change
-# Trainingint.linear
-# rpe-analysis
+```{r training data model}
+## Scripts
+# - volume.change.R
+# - ss.analysis.R
+# - ss.cleanup.R
+# - trainint.analysis.R
+# - trainint.cleanup.R
+
+# Log-transformed data
+vol.lchange <- readRDS("./data/data-gen/training/vol.lchange.RDS")
+ss.lchange <- readRDS("./data/data-gen/training/ss.lchange.RDS")
+int.lchange <- readRDS("./data/data-gen/training/int.lchange.RDS")
 
@@
 # Fold-change model
-m1 <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = change_dat)
-resplot.vol <- plot(m1)
-sum.vol <- summary(m1)
+vol.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
+                     data = vol.lchange)
+vol.lres <- plot(vol.lmod)
+vol.lsum <- summary(vol.lmod)
 
 # Fold-change estimated means
-confint.m1 <- confint(emmeans(m1, specs = ~"supplement|time")) %>%
-        data.frame()
-
-### RPE
+vol.lemm <- confint(emmeans(vol.lmod, specs = ~"supplement|time")) %>%
+        data.frame()
+
+### Session score (RPE)
 # Model
-m2 <- lmerTest::lmer(change ~ baseline + time + supplement:time + (1|subject),
-                     data = change_dat.ss)
-resplot.ss <- plot(m2)
-sum.ss <- summary(m2)
+ss.lmod <- lmerTest::lmer(change ~ baseline + time + supplement:time + (1|subject),
+                     data = ss.lchange)
+ss.lres <- plot(ss.lmod)
+ss.lsum <- summary(ss.lmod)
 
 # Session score emmeans
-confint.m2 <- confint(emmeans(m2, specs = ~"supplement|time")) %>%
+ss.lemm <- confint(emmeans(ss.lmod, specs = ~"supplement|time")) %>%
         data.frame()
@@
 ### Training intensity
-m3 <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = tot.change)
-resplot.ti <- plot(m3)
-sum.ti <- summary(m3)
+int.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
+                     data = int.lchange)
+int.lres <- plot(int.lmod)
+int.lsum <- summary(int.lmod)
 
 # Emmeans
-confint.m3 <- confint(emmeans(m3, specs = ~"supplement|time")) %>%
+int.lemm <- confint(emmeans(int.lmod, specs = ~"supplement|time")) %>%
         data.frame()
@@
 # Diff 4
-vcoef <- coef(summary(m1))
+vcoef <- coef(summary(vol.lmod))
 vdiff.4 <- round(vcoef[9,1], 3)
@@
 ## RPE
-scoef <- coef(summary(m2))
+scoef <- coef(summary(ss.lmod))
 sdiffp <- round(scoef[11, 5], 3)
@@
 # Diff time 2
-icoef <- coef(summary(m3))
+icoef <- coef(summary(int.lmod))
 idiff2 <- round(icoef[7,1], 3)
@@
 idiffse6 <- round(icoef[11,2], 3)
-
-```
-
-```{r post 5th data handling}
-# Scripts
-# - Post5th.change
-# - Post5th.figs
-humac <- read_excel("./data/tests/ribose.humac.xlsx", na = "NA") %>%
-
-        mutate(time = if_else(timepoint == "D-1", 
-                              "baseline", 
-                              if_else(timepoint %in% c("D4", "D5"), 
-                                      "test1", 
-                                      if_else(timepoint %in% c("D8", "D9"), 
-                                              "test2", 
-                                              if_else(timepoint %in% c("T3", "T4") & acute %in% c("rest", "post30min", "post2h"),
-                                                      "test3",
-                                                      if_else(acute == "post23h", "test4", timepoint)))))) %>%
-        mutate(time = factor(time, levels = c("baseline", "test1", "test2", "test3", "test4")), 
-               acute = factor(acute, levels = c("rest", "post30min", "post2h", "post23h")), 
-               supplement = factor(supplement, levels = c("placebo", "glucose")))
-rest.dat <- humac %>%
-        filter(acute == "rest" )
-
-# Change data
-isom.dat <- rest.dat %>%
-        filter(test == "isom")
-change_dat <- isom.dat %>%
-        dplyr::select(subject, time, supplement, peak.torque) %>%
-        group_by(subject, time, supplement) %>%
-        summarise(peak.torque = mean(peak.torque, na.rm = TRUE)) %>%
-        pivot_wider(names_from = time, 
-                    values_from = peak.torque) %>%
-
-        ungroup() %>%
-        mutate(change.2 = log(test1)-log(baseline),
-               change.3 = log(test2)-log(baseline),
-               change.4 = log(test3)-log(baseline),
-               baseline = baseline - mean(baseline, na.rm = TRUE),
-               supplement = factor(supplement, levels = c("placebo", "glucose"))) %>%
-        select(subject, supplement, baseline, change.2, change.3, change.4) %>%
-        pivot_longer(names_to = "time",
-                     values_to = "change",
-                     cols = (change.2:change.4))
+```
+
+```{r humac post 5th model}
+## Scripts
+# - humac.cleanup.R
+# - humac.main.analysis.R - this is data gathered at baseline, during the intervention, and after session 5
+
+# Log-transformed data
+
+isom.lchange <- readRDS("./data/data-gen/humac/isom.lchange.RDS")
+iso60.lchange <- readRDS("./data/data-gen/humac/isok60.lchange.RDS")
+iso240.lchange <- readRDS("./data/data-gen/humac/isok240.lchange.RDS")
+
+# Isometric
+isom.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
+                     data = isom.lchange)
+isom.lres <- plot(isom.lmod)
+isom.lsum <- summary(isom.lmod)
 
 # Isok.60
-isok60.dat <- rest.dat %>%
-        filter(test == "isok.60")
-change_dat2 <- isok60.dat %>%
-        dplyr::select(subject, time, supplement, peak.torque) %>%
-        group_by(subject, time, supplement) %>%
-        summarise(peak.torque = mean(peak.torque, na.rm = TRUE)) %>%
-        pivot_wider(names_from = time, 
-                    values_from = peak.torque) %>%
-
-        ungroup() %>%
-        mutate(change.2 = log(test1)-log(baseline),
-               change.3 = log(test2)-log(baseline),
-               change.4 = log(test3)-log(baseline),
-               baseline = baseline - mean(baseline, na.rm = TRUE),
-               supplement = factor(supplement, levels = c("placebo", "glucose"))) %>%
-        select(subject, supplement, baseline, change.2, change.3, change.4) %>%
-        pivot_longer(names_to = "time",
-                     values_to = "change",
-                     cols = (change.2:change.4))
-
-## Isok.240
-isok240.dat <- rest.dat %>%
-        filter(test == "isok.240") 
-change_dat3 <- isok240.dat %>%
-        dplyr::select(subject, time, supplement, peak.torque) %>%
-        group_by(subject, time, supplement) %>%
-        summarise(peak.torque = mean(peak.torque, na.rm = TRUE)) %>%
-        pivot_wider(names_from = time, 
-                    values_from = peak.torque) %>%
-
-        ungroup() %>%
-        mutate(change.2 = log(test1)-log(baseline),
-               change.3 = log(test2)-log(baseline),
-               change.4 = log(test3)-log(baseline),
-               baseline = baseline - mean(baseline, na.rm = TRUE),
-               supplement = factor(supplement, levels = c("placebo", "glucose"))) %>%
-        select(subject, supplement, baseline, change.2, change.3, change.4) %>%
-        pivot_longer(names_to = "time",
-                     values_to = "change",
-                     cols = (change.2:change.4))
-
-```
-
-```{r post 5th modelling}
-# Scripts
-# - Post5th.change
-# - Post5th.figs
-
-# Isometric
-m1 <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = change_dat)
-resplot.iso5 <- plot(m1)
-sum.iso5 <- summary(m1)
-
-# Isok.60
-m2 <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = change_dat2)
-resplot.isok605 <- plot(m2)
-sum.isok605 <- summary(m2)
+iso60.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
+                     data = iso60.lchange)
+iso60.lres <- plot(iso60.lmod)
+iso60.lsum <- summary(iso60.lmod)
 
 # Isok.240
-m3 <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = change_dat3)
-resplot.isok2405 <- plot(m3)
-resplot.isok2405 <- summary(m3)
+iso240.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
+                     data = iso240.lchange)
+iso240.lres <- plot(iso240.lmod)
+iso240.lsum <- summary(iso240.lmod)
 
@@
 # Isometric
-confint.m1 <- confint(emmeans(m1, specs = ~"supplement|time")) %>%
+isom.lemm <- confint(emmeans(isom.lmod, specs = ~"supplement|time")) %>%
         data.frame()
@@
 # Isok.60
-confint.m2 <- confint(emmeans(m2, specs = ~"supplement|time")) %>%
+iso60.lemm <- confint(emmeans(iso60.lmod, specs = ~"supplement|time")) %>%
         data.frame()
@@
 # Isok.240
-confint.m3 <- confint(emmeans(m3, specs = ~"supplement|time")) %>%
+iso240.lemm <- confint(emmeans(iso240.lmod, specs = ~"supplement|time")) %>%
         data.frame()
@@
 # Isom diff
-isom5coef <- coef(summary(m1))
+isom5coef <- coef(summary(isom.lmod))
 isom5diffp <- round(isom5coef[7, 5], 3)
@@
 # Isok 60 diff
-isok605coef <- coef(summary(m2))
+isok605coef <- coef(summary(iso60.lmod))
 isok605diffp <- round(isok605coef[7, 5], 3)
@@
 # Isok 260
-isok2405coef <- coef(summary(m3))
+isok2405coef <- coef(summary(iso240.lmod))
 isok2405diffp <- round(isok2405coef[7, 5], 3)
@@
 
-```{r post 6th data handling}
-# Scripts
-# - Post6th.change
-# - Post6th.figs
-humac <- read_excel("./data/tests/ribose.humac.xlsx", na = "NA") %>%
-
-        mutate(time = if_else(timepoint == "D-1", 
-                              "baseline", 
-                              if_else(timepoint %in% c("D4", "D5"), 
-                                      "test1", 
-                                      if_else(timepoint %in% c("D8", "D9"), 
-                                              "test2", 
-                                              if_else(timepoint %in% c("T3", "T4") & acute %in% c("rest", "post30min", "post2h"),
-                                                      "test3",
-                                                      if_else(acute == "post23h", "test4", timepoint)))))) %>%
-        mutate(time = factor(time, levels = c("baseline", "test1", "test2", "test3", "test4")), 
-               acute = factor(acute, levels = c("rest", "post30min", "post2h", "post23h")), 
-               supplement = factor(supplement, levels = c("placebo", "glucose")))
+```{r post 6th model}
+## Scripts
+# - humac.acute.analysis.R - this is data gatherede before and after one RT session (sixth session)
+# - humac.cleanup.R
+
+# Data
+isom.lchange.ac <- readRDS("./data/data-gen/humac/isom.lchange.ac.RDS")
+iso60.lchange.ac <- readRDS("./data/data-gen/humac/iso60.lchange.ac.RDS")
+iso240.lchange.ac <- readRDS("./data/data-gen/humac/iso240.lchange.ac.RDS")
 
 # Isometric
-isom.dat <- humac %>%
-        filter(test == "isom",
-               time %in% c("test3", "test4")) 
-change_dat <- isom.dat %>%
-        dplyr::select(subject, acute, supplement, peak.torque) %>%
-        group_by(subject, acute, supplement) %>%
-        summarise(peak.torque = mean(peak.torque, na.rm = TRUE)) %>%
-        pivot_wider(names_from = acute, 
-                    values_from = peak.torque) %>%
-        #print()
-
-        ungroup() %>%
-        mutate(change.2 = log(post30min)-log(rest),
-               change.3 = log(post2h)-log(rest),
-               change.4 = log(post23h)-log(rest),
-               baseline = rest - mean(rest, na.rm = TRUE),
-               supplement = factor(supplement, levels = c("placebo", "glucose"))) %>%
-        select(subject, supplement, baseline, change.2, change.3, change.4) %>%
-        pivot_longer(names_to = "time",
-                     values_to = "change",
-                     cols = (change.2:change.4))
+isomac.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
+                     data = isom.lchange.ac)
+isomac.lres <- plot(isomac.lmod)
+isomac.lsum <- summary(isomac.lmod)
 
 # Isok.60
-isok60.dat <- humac %>%
-        filter(test == "isok.60",
-               time %in% c("test3", "test4")) 
-change_dat2 <- isok60.dat %>%
-        dplyr::select(subject, acute, supplement, peak.torque) %>%
-        group_by(subject, acute, supplement) %>%
-        summarise(peak.torque = mean(peak.torque, na.rm = TRUE)) %>%
-        pivot_wider(names_from = acute, 
-                    values_from = peak.torque) %>%
-        #print()
-
-        ungroup() %>%
-        mutate(change.2 = log(post30min)-log(rest),
-               change.3 = log(post2h)-log(rest),
-               change.4 = log(post23h)-log(rest),
-               baseline = rest - mean(rest, na.rm = TRUE),
-               supplement = factor(supplement, levels = c("placebo", "glucose"))) %>%
-        select(subject, supplement, baseline, change.2, change.3, change.4) %>%
-        pivot_longer(names_to = "time",
-                     values_to = "change",
-                     cols = (change.2:change.4)) 
-## Isok.240
-isok240.dat <- humac %>%
-        filter(test == "isok.240",
-               time %in% c("test3", "test4")) 
-change_dat3 <- isok240.dat %>%
-        dplyr::select(subject, acute, supplement, peak.torque) %>%
-        group_by(subject, acute, supplement) %>%
-        summarise(peak.torque = mean(peak.torque, na.rm = TRUE)) %>%
-        pivot_wider(names_from = acute, 
-                    values_from = peak.torque) %>%
-        #print()
-
-        ungroup() %>%
-        mutate(change.2 = log(post30min)-log(rest),
-               change.3 = log(post2h)-log(rest),
-               change.4 = log(post23h)-log(rest),
-               baseline = rest - mean(rest, na.rm = TRUE),
-               supplement = factor(supplement, levels = c("placebo", "glucose"))) %>%
-        select(subject, supplement, baseline, change.2, change.3, change.4) %>%
-        pivot_longer(names_to = "time",
-                     values_to = "change",
-                     cols = (change.2:change.4))
-
-```
-
-```{r post 6th modelling}
-# Scripts
-# - Post6th.change
-# - Post6th.figs
-
-# Isometric
-m1 <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = change_dat)
-resplot.isom <- plot(m1)
-sum.isom <- summary(m1)
-
-# Isok.60
-m2 <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = change_dat2)
-resplot.isok60 <- plot(m2)
-sum.isok60 <- summary(m2)
+iso60ac.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
+                     data = iso60.lchange.ac)
+iso60ac.lres <- plot(iso60ac.lmod)
+iso60ac.lsum <- summary(iso60ac.lmod)
 
 # Isok.240
-m3 <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = change_dat3)
-resplot.isok240 <- plot(m3)
-sum.isok240 <- summary(m3)
+iso240ac.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
+                     data = iso240.lchange.ac)
+iso240ac.lres <- plot(iso240ac.lmod)
+iso240ac.lsum <- summary(iso240ac.lmod)
 
@@
 # Isometric
-confint.m1 <- confint(emmeans(m1, specs = ~"supplement|time")) %>%
+isomac.lemm <- confint(emmeans(isomac.lmod, specs = ~"supplement|time")) %>%
         data.frame()
@@
 # Isok.60
-confint.m2 <- confint(emmeans(m2, specs = ~"supplement|time")) %>%
+iso60ac.lemm <- confint(emmeans(iso60ac.lmod, specs = ~"supplement|time")) %>%
         data.frame()
@@
 # Isok.240
-confint.m3 <- confint(emmeans(m3, specs = ~"supplement|time")) %>%
-        data.frame()
-
-# For inline reporting
+iso240.lemm <- confint(emmeans(iso240ac.lmod, specs = ~"supplement|time")) %>%
+        data.frame()
+
+## For inline reporting
 # Isom diff
-isomcoef <- coef(summary(m1))
+isomcoef <- coef(summary(isomac.lmod))
 isomdiffp <- round(isomcoef[7, 5], 3)
@@
 # Isok 60 diff
-isok60coef <- coef(summary(m2))
+isok60coef <- coef(summary(iso60ac.lmod))
 isok60diffp <- round(isok60coef[7, 5], 3)
@@
 # Isok 240 diff
-isok240coef <- coef(summary(m3))
+isok240coef <- coef(summary(iso240ac.lmod))
 isok240diffp <- round(isok240coef[7, 5], 3)
@@
 totvolfold.plot <- readRDS("./data/data-gen/training/totvolfold.plot.RDS")
+
 # Training intensity
@@
 totalint.plot <- readRDS("./data/data-gen/training/trainint.plot.RDS")
+
 # Join figures
@@
 ```{r RPE table}
-# Scripts
-# rpe-analysis
-# Session score
-ss.dat <- read_excel("./data/training/ribose_rpe.xlsx", na = "NA") %>%
-        select(subject, timepoint, session.score, supplement) %>%
-        mutate(time = if_else(timepoint %in% c("T1", "T2"),
-                              "baseline",
-                              if_else(timepoint %in% c("D3", "D4"),
-                                      "session2",
-                                      if_else(timepoint %in% c("D5", "D6"),
-                                              "session3",
-                                              if_else(timepoint %in% c("D7", "D8"),
-                                                      "session4",
-                                                      if_else(timepoint %in% c("D9", "D10"),
-                                                              "session5",
-                                                              if_else(timepoint %in% c("T3", "T4"),
-                                                                      "post", timepoint))))))) %>%
-        mutate(time = factor(time, levels = c("baseline", "session1", "session2", "session3", "session4", "session5", "post")),
-               supplement = factor(supplement, levels = c("placebo", "glucose")))
-
-## Mean and SD
-# Session score
-ss.exp <- ss.dat %>%
-        select(subject, supplement, time, session.score) %>%
-        group_by(supplement, time) %>%
-        summarise(mean.ss = mean(session.score),
-                  sd.ss = sd(session.score)) %>%
-        mutate(stat = paste0(round(mean.ss, 1), " (", round(sd.ss, 1), ")")) %>%
-        select(supplement, time, stat) %>%
-        pivot_wider(names_from = time,
-                    values_from = stat)
-
-ss.exp %>%
+## Scripts
+# - ss.cleanup.R
+# - ss.absolute.R
+
+# Data
+ss.abs <- readRDS("./data/data-gen/training/ss.abs.RDS")
+
+# Table
+ss.abs %>%
   kable(col.names = c("Supplement", "Baseline", "Session 2", "Session 3", "Session 4", "Session 5",

### 2023-02-09 R/humac.acute.analysis.R

@@ -47,6 +47,8 @@ change.isom <- isom.dat %>%
                cols = (change.2:change.4)) #%>%
 #  print()
 
+saveRDS(change.isom, "./data/data-gen/humac/isom.lchange.ac.RDS")
+
 # Isok.60
 
 isok60.dat <- humac %>%
@@ -74,6 +76,8 @@ change.60 <- isok60.dat %>%
                cols = (change.2:change.4)) #%>%
 #  print()
 
+saveRDS(change.60, "./data/data-gen/humac/iso60.lchange.ac.RDS")
+
 ## Isok.240
 
 isok240.dat <- humac %>%
@@ -100,6 +104,8 @@ change.240 <- isok240.dat %>%
                values_to = "change",
                cols = (change.2:change.4)) #%>%
 #  print()
+
+saveRDS(change.240, "./data/data-gen/humac/iso240.lchange.ac.RDS")
 
 ## Linear mixed effects model
 # This model tries to explain the change by time and supplement, accounting for potential differences in baseline values and that the same participants

### 2023-02-09 R/ss.absolute.R

@@ -17,8 +17,14 @@ ss.exp <- ss.dat %>%
   select(subject, supplement, time, session.score) %>%
   group_by(supplement, time) %>%
   summarise(mean.ss = mean(session.score),
-            sd.ss = sd(session.score)) #%>%
-  #print()
+            sd.ss = sd(session.score)) %>%
+  mutate(stat = paste0(round(mean.ss, 1), " (", round(sd.ss, 1), ")")) %>%
+  select(supplement, time, stat) %>%
+  pivot_wider(names_from = time,
+              values_from = stat) %>%
+  print()
+
+saveRDS(ss.exp, "./data/data-gen/training/ss.abs.RDS")
 
 ss.barplot <- ggplot(ss.exp, aes(fill = supplement, y = mean.ss, x = time)) +
   geom_bar(position = "dodge", stat = "identity") +

### 2023-02-09 R/ss.analysis.R

@@ -37,8 +37,10 @@ change_dat.ss <- ss.dat %>%
   select(subject, supplement, baseline, change.2, change.3, change.4, change.5, change.post) %>%
   pivot_longer(names_to = "time",
                values_to = "change",
-               cols = (change.2:change.post)) #%>%
-  #print()
+               cols = (change.2:change.post)) %>%
+  print()
+
+saveRDS(change_dat.ss, "./data/data-gen/training/ss.lchange.RDS")
 
 ## Linear mixed effects model
 # This model tries to explain the change by time and supplement, accounting for potential differences in baseline values and that the same participants

### 2023-02-09 R/volume.change.R

@@ -110,6 +110,7 @@ change_dat <- tot.volh %>%
                cols = (change.2:change.6)) %>%
   print()
 
+saveRDS(change_dat, "./data/data-gen/training/vol.lchange.RDS")
 
 ## Linear mixed effects model
 # This model tries to explain the change by time and supplement, accounting for potential differences in baseline values and that the same participants

### 2023-02-09 ribose-paper.RMD
small updates to discussion, needs alot more work

@@
 
-The main findings of the present study were that exercise training with glucose for five sessions did not affect muscle biology markers, measured as total RNA and -protein, rRNA and specific protein, compared to RT with placebo.
-Similarly, RT with glucose did not affect markers of muscle functionality such as muscular recovery, total session volume, and training intensity, compared to RT with placebo.
-There are several studies suggesting high glucose treatment is important to markers of ribosome biogenesis (MARIAPPAN11, TANAKA18, ZHAI12), however, the perspective of exogenous glucose ingestion combined with RT on subsequent adaptations have remained unexplored.
-As a result, our hypotheses and purposes were quite exploratory and, considering previous suggestions, it is interesting that we did not observe any difference between RT with glucose and RT with placebo on the aforementioned markers of ribosome biogenesis.
-
-To our knowledge, no previous study has investigated the effects of glucose compared to placebo ingestion with RT on ribosome biogenesis in human skeletal muscle cells.
-Through recent studies, the current model of ribosome biogenesis suggests that changes in the accumulation of total RNA and rRNA early after the onset of an RT-period predict long-term responses in RT adaptation (FIGUEIREDO19).
-Therefore, the lack of effects of glucose on total RNA and rRNA after two weeks of moderate volume RT do indicate that glucose per se does not induce positive long-term adaptations in muscle growth.
-First, the time frame of the intervention has previously served as a predictor of long-term responses, as recent studies have shown augmentation of ribosome biogenesis by total RNA and rRNA following multiple sessions of RT (FIGUEIREDO16; HAMMARSTROM20).
-Figueiredo et al.
-(2016) observed increases in 47S pre-rRNA measurable after a single RT session, passing baseline after 4hrs, peaking at 24hrs and remaining elevated until 48hrs post-exercise.
-Other mature rRNA where first measurable after multiple sessions (FIGUEIREDO16).
-In addition, Hammarström et al.
-(2020) demonstrated that the most prominent augmentation in ribosome biogenesis markers occurred 2 weeks following moderate-volume unilateral RT.
-Secondly, translational capacity is now regarded a vital factor in long-term RT-induced adaptations (IS IT?), as increased ribosome content provides a higher basal MPS, giving regulation of ribosome biogenesis a central role in muscle growth (FIGUEIREDO19).
+The main findings of the present study were that resistance training with glucose for five sessions did not affect muscle biology markers, measured as total RNA and -protein, rRNA and specific protein, compared to RT with placebo.
+Similarly, RT with glucose did not affect markers of skeletal muscle functionality such as muscular recovery, total session volume, and training intensity, compared to RT with placebo.
+There are studies suggesting that high glucose treatment impacts ribosome biogenesis in … cells (MARIAPPAN11; ZHAI12; TANAKA15), however, this effect does not seem to translate to human skeletal muscle cells in vivo.
+
+To our knowledge, no previous study has investigated the effects of glucose compared to placebo ingestion with RT on ribosome biogenesis in human skeletal muscle cells, making our hypotheses exploratory and comparisons not straight forwards.
+The current model of ribosome biogenesis suggests that changes in the accumulation of total RNA and rRNA early after the onset of an RT period predict long-term RT adaptations (FIGUEIREDO19; HAMMARSTROM20; HAMMARSTROM21).Therefore, we assumed 2 weeks to be a suitable time frame for a moderate RT volume to elicit changes in ribosome biogenesis, as similar observations have been made previously (FIGUEIREDO16; HAMMARSTROM20; HAMMARSTROM21).
+In Hammarström et al.
+(2020), untrained participants that increased ribosome biogenesis after two weeks (5 RT sessions) of RT, also responded well after the entire intervention (12 weeks, 31 sessions).
+Figueiredo and colleagues (2016) also observed markedly acute effects in RT-untrained individuals, where 47S pre-rRNA increased after a single RT session, passing baseline after 4hrs, peaking at 24hrs and remaining elevated until 48hrs post-exercise.
+The four mature rRNA did not measurably increase until after multiple sessions (FIGUEIREDO16).
+As such, increases in total RNA and rRNA observed in the present study are in line with previous observations (FIGUEIREDO16; HAMMARSTROM20; H21?).
+Further, the accumulation of total RNA and rRNA is also suggested to reach a plateau after approximately 5 RT-sessions, thus, in addition to the demanding RT- and dietary protocol and the exploratory nature of the hypotheses, we did not conduct a more long-term intervention.
+COMMENT ON PROTEIN DATA HERE?
+
+As we did not observe any effect of ingesting glucose before and after RT compared to placebo in this design, there is little to suggest that glucose *per se* has any additional effect on ribosome biogenesis, as long as the daily energy intake is sufficient to support cellular energy levels required to drive this process.
+Looking at previous observations from *in vitro* studies of rodents, yeast, and non-muscle cell cultures (MARIAPPAN11; KIM13; TANAKA18; ZHAI12; HOPPE09; MURAYAMA08), one interesting notion would be how low-carb or glucose starvation would affect ribosome biogenesis in human skeletal muscle cells...
+
+..Secondly, translational capacity is now regarded a vital factor in long-term RT-induced adaptations (IS IT?), as increased ribosome content provides a higher basal MPS, giving regulation of ribosome biogenesis a central role in muscle growth (FIGUEIREDO19).
 This is further substantiated by the association between blunted ribosome biogenesis and poor responses to RT (HAMMARSTROM20; STEC16), in addition to the lack of relationship between translational efficiency and long-term gains in skeletal muscle mass (FIGUEIREDO19).

### 2023-02-10 ribose-paper.RMD
updates to discussion

@@
 
-The main findings of the present study were that resistance training with glucose for five sessions did not affect muscle biology markers, measured as total RNA and -protein, rRNA and specific protein, compared to RT with placebo.
-Similarly, RT with glucose did not affect markers of skeletal muscle functionality such as muscular recovery, total session volume, and training intensity, compared to RT with placebo.
-There are studies suggesting that high glucose treatment impacts ribosome biogenesis in … cells (MARIAPPAN11; ZHAI12; TANAKA15), however, this effect does not seem to translate to human skeletal muscle cells in vivo.
+The main findings of the present study were that resistance training with glucose for five sessions did not affect markers of ribosome biogenesis, measured as total RNA, rRNA and rDNA transcription initiation associated proteins, compared to RT with placebo.
+Similarly, RT with glucose did not affect markers of skeletal muscle functionality such as muscular performance and -recovery, total session volume, and training intensity, compared to RT with placebo.
+There are studies suggesting that high vs.
+low glucose treatment/glucose starvation impacts ribosome biogenesis in yeast, rodent and human non-muscle cells (KIM13;MARIAPPAN11;TANAKA15;ZHAI12), however, this effect does not seem evident when total daily glucose intake is matched.
 
@@
 In Hammarström et al.
-(2020), untrained participants that increased ribosome biogenesis after two weeks (5 RT sessions) of RT, also responded well after the entire intervention (12 weeks, 31 sessions).
-Figueiredo and colleagues (2016) also observed markedly acute effects in RT-untrained individuals, where 47S pre-rRNA increased after a single RT session, passing baseline after 4hrs, peaking at 24hrs and remaining elevated until 48hrs post-exercise.
+(2020), participants that increased ribosome biogenesis after two weeks (5 RT sessions) of RT, also responded well after the entire intervention (12 weeks, 31 sessions).
+Figueiredo and colleagues (2016) also observed markedly acute effects, where 47S pre-rRNA increased after a single RT session, passing baseline after 4hrs, peaking at 24hrs and remaining elevated until 48hrs post-exercise.
 The four mature rRNA did not measurably increase until after multiple sessions (FIGUEIREDO16).
-As such, increases in total RNA and rRNA observed in the present study are in line with previous observations (FIGUEIREDO16; HAMMARSTROM20; H21?).
+As such, increases in total RNA and rRNA observed in the present study are in line with previous observations (FIGUEIREDO16; HAMMARSTROM20; H21?), and also evident i moderately trained individuals, compared to the untrained or RT-untrained individuals of previous studies (FIGUEIREDO16;HAMMARSTROM20;H21?).
+
 Further, the accumulation of total RNA and rRNA is also suggested to reach a plateau after approximately 5 RT-sessions, thus, in addition to the demanding RT- and dietary protocol and the exploratory nature of the hypotheses, we did not conduct a more long-term intervention.
@@
 As we did not observe any effect of ingesting glucose before and after RT compared to placebo in this design, there is little to suggest that glucose *per se* has any additional effect on ribosome biogenesis, as long as the daily energy intake is sufficient to support cellular energy levels required to drive this process.
-Looking at previous observations from *in vitro* studies of rodents, yeast, and non-muscle cell cultures (MARIAPPAN11; KIM13; TANAKA18; ZHAI12; HOPPE09; MURAYAMA08), one interesting notion would be how low-carb or glucose starvation would affect ribosome biogenesis in human skeletal muscle cells...
-
-..Secondly, translational capacity is now regarded a vital factor in long-term RT-induced adaptations (IS IT?), as increased ribosome content provides a higher basal MPS, giving regulation of ribosome biogenesis a central role in muscle growth (FIGUEIREDO19).
-This is further substantiated by the association between blunted ribosome biogenesis and poor responses to RT (HAMMARSTROM20; STEC16), in addition to the lack of relationship between translational efficiency and long-term gains in skeletal muscle mass (FIGUEIREDO19).
-Hence, two weeks of moderate-volume unilateral RT should have been sufficient to elicit differences between exercise training with glucose and placebo.
-Supporting the notion that our intervention indeed was effective, we observed expected augmentation in accumulation of total RNA and expression of rRNA (CITE THIS?).
-As there were no differences in total RNA and rRNA between exercise training with glucose and placebo in the present study, it is therefore reasonable to suggest that there were no differences in ribosome biogenesis by proxy.
-Further, since glucose did not augment ribosome content after two weeks of RT compared to placebo, it is unlikely that glucose per se is beneficial to long-term adaptations in RT.
-
-Indeed, our findings suggest that glucose ingestion per se before and after RT does not induce changes in the internal physiological milieu of human skeletal muscle cells with anabolic effects.
+Looking at previous observations from *in vitro* studies of rodents, yeast, and non-muscle cell cultures (MARIAPPAN11; KIM13; TANAKA18; ZHAI12; HOPPE09; MURAYAMA08), one interesting notion would be how high glucose vs.
+low glucose or glucose starvation would affect ribosome biogenesis over time in human skeletal muscle cells.
+Time being of the essence in such a perspective, as previous *in vitro* studies have utilized cells with higher physiological rate than human skeletal muscle cells (KIM13;MARIAPPAN11;TANAKA15;ZHAI12).
+
+Indeed, our findings suggest that glucose ingestion *per se* before and after RT does not induce changes in the internal physiological milieu of human skeletal muscle cells with anabolic effects.
 It has been theorised that glucose may convey an anabolic signal through insulin, and thus operate as a signal mediator affecting upstream pathways of ribosome biogenesis (KUSNADI15; TANAKA18).
 Such an anabolic signal could then possibly increase PIC assembly, and consequently Pol I recruitment, initiating rDNA transcription and subsequent rRNA maturation, leading to increased ribosome content (FIGUEIREDO19; TANAKA18).
-INVESTIGATE THE INSULIN DATA AND COMMENT IT HERE.
-Another theory was that glucose ingestion would impact the cellular energy levels and thus, the energy-sensitive PIH1, mTORC1, ERK1/2, AMPK, and SIRT1 pathways (MARIAPPAN11; KIM13; TANAKA18; ZHAI12).
-The mammalian target of rapamycin, along with ERK1/2 and PIH1 up-regulates PIC assembly in high-energy conditions, while AMPK and possibly SIRT1 represses PIC assembly in low-energy conditions, based on in vitro rodent, yeast, and non-muscle cell cultures (MARIAPPAN11; KIM13; TANAKA18; ZHAI12; HOPPE09; MURAYAMA08).
+As significantly higher blood levels of c-peptide and glucose were observed in RT with glucose compared to placebo in the present study, without any difference in ribosome biogenesis, there is little to suggest that either insulin or glucose has any direct acute effects.
+
+Another theory is that glucose ingestion leads to higher cellular energy levels compared to placebo, stimulating energy-sensitive pathways such as PIH1, mTORC1, ERK1/2, AMPK and SIRT1 in a dose-response relationship (KIM13;MARIAPPAN11;TANAKA15;ZHAI12).
+The mTORC1, along with ERK1/2 and PIH1 up-regulates PIC assembly during high-energy conditions, while AMPK and possibly SIRT1 represses PIC assembly during low-energy conditions, as observed in rodent (KILDE), yeast (KILDE), and non-muscle cells (KILDE) *in vitro*.
 Specifically, high glucose treatment augmented rDNA transcription by increased UBF and SL-1 activity, while low glucose treatment, glucose starvation, or nutrient depletion stimulated AMPK-dependent TIF-IA and chromatin regulation suppression (KOS-BRAUN17; MARIAPPAN11; ZHAI12).
 In addition to UBF and SL-1, assembly of the PIC requires TIF-IA and Pol I recruitment, induced by amino acid intake in an mTOR-S6K-dependent manner and possibly by direct c-Myc activation (KUSNADI15).
-COMMENT ON PROTEIN ANALYSIS RESULTS HERE?
 As such, increasing energy supply and nutrient availability should in line with these theories potently stimulate PIC assembly and rDNA transcription.
-Along this notion, it does not seem apparent that the timing of glucose ingestion, for instance immediately before and after RT affects these processes, as glucose did not induce acute augmentation of ribosome biogenesis, despite the significant elevation of plasma glucose levels when exercising with glucose compared to placebo.
-Based on our observations, glucose per se has no acute effect on ribosome biogenesis, as the same effects can be observed with equal glucose ingestion 5-6hrs later.
-
-Another possible explanation for the lack of effect of glucose compared to placebo, is that both treatments may have exercised with sufficient levels of nutrients and energy, meaning that mTORC1 had sufficient nutrient availability and, thus was not inhibited by AMPK due to energy deficiency (HOPPE09; KUSNADI15).
+Importantly, these pathways were not directly investigated in the present study.
+However, downstream targets (UBF, c-Myc, rpS6?) remained unaffected by glucose compared to placebo ingestion, suggesting there is not a dose-response relationship between muscle cellular energy levels and ribosome biogenesis during an RT-intervention (Not sure if I can say this).
+
+A weakness in the present study, is the lack of data on intracellular glucose levels in the muscle cells.
+Thus, it cannot be said for sure whether there were differences between glucose and placebo or not.
+
+A possible explanation for the lack of effect of glucose compared to placebo, is that both treatments may have exercised with sufficient levels of nutrients and energy, meaning that mTORC1 had sufficient nutrient availability and, thus was not inhibited by AMPK due to energy deficiency (HOPPE09; KUSNADI15).
 Indeed, mTORC1 is characterised as energy-sensitive (DENNIS01), and therefore it might have been directly affected by high vs.

### 2023-02-10 ribose-paper.RMD
tidied up rmd a bit

@@
 
-```{r nutrition analysis}
-## Scripts
-# - nut.analysis.R
-# - dxa.R
-
-#summarise macro
-#nutritiontable with pro.pr kg bodyweight pr. group
-#laod dxa data and take out weight and subject to join with nutrition data
-fp.weight <-read_excel("data/dxa/ribose_dxa.xlsx") %>%
-        select(weight, subject)
-
-#nutrition data.
-#the data is in grams and 
-# added 50 grams protein as sup_protein for ingestion of whey protein isolat,
-#And calculated in to the total amount of proteins ingested pr day.
-#added 90g of glucose in to carbohydrate 
-#added 360 kcal from glucose intake and calculated in total kcal pr day
-#Load nutrition data.
-nut.weight <- read_excel("data/dxa/Ribose_nutrition_result.xlsx")%>%
-        select(timepoint, meal, subject, protein, fat, calories,
-               carbohydrates, group, sup_pro, sup_gluc, kcal_glu) %>%
-        #changed names to compare groups each paired days
-        mutate(timepoint = if_else(timepoint %in% c("T1","T2"), 
-                                   "Day 1",
-                                   if_else(timepoint %in% c("3", "4"),
-                                           "Day 2",
-                                           if_else(timepoint %in% c("5", "6"),
-                                                   "Day 3",
-                                                   if_else(timepoint %in% c("7", "8"),
-                                                           "Day 4",
-                                                           if_else(timepoint %in% c("9", "10"),
-                                                                   "Day 5",
-                                                                   if_else(timepoint %in% c("T3", "T4"),
-                                                                           "Day 6",
-                                                                           "na"))))))) %>%
-        group_by(timepoint, subject, group) %>%
-        #take in weight
-        inner_join(fp.weight) %>%
-        #summarise each variable to get total pr timepoint.
-        #Whey protein supplement is added to total protein ingestion
-        summarise(protein = sum(protein + sup_pro),
-                  fat = sum(fat),
-                  calories = sum(calories + kcal_glu),
-                  carbohydrates = sum(carbohydrates + sup_gluc),
-                  weight = sum(weight)) %>%
-        #summarise to get protein pr weight.
-        mutate(proprkg = (protein/weight)) %>%
-        pivot_wider(names_from = group,
-                    values_from = c(fat, protein, calories, carbohydrates, weight, proprkg))
-
-#nutrition table with pro/weight. The table shows mean and sd of total intake 
-#in grams pr. group, pr. timepoint.
-#How much kcal on glucose supplement?
-nuttable <- nut.weight %>%
-        pivot_longer(names_to = "variable",
-                     values_to = "values", col= fat_glucose:proprkg_placebo) %>%
-        group_by(timepoint, variable) %>%
-        summarise(m = mean(values, na.rm = TRUE),
-                  s= sd(values, na.rm = TRUE)) %>%
-        ungroup() %>%
-        separate(variable, into = c("variable", "group")) %>%
-        mutate(stat = paste0(round(m, 1), " (", round(s, 1), ")")) %>%
-        select(timepoint, variable, group, stat,) %>%
-        pivot_wider(names_from = variable,
-                    values_from = stat) %>%
-        select(timepoint, group, calories, carbohydrates, fat, protein,
-               proprkg) %>%
-        #changed names of group for simplicity 
-        mutate(group = if_else(group %in% c("glucose"),
-                               "G",
-                               if_else(group %in% c("placebo"),
-                                       "P",
-                                       "")))
-
-# Change data
-#Does not contain FP 106 and 116 because of not delivered data.
-#summarise macro
-#Added sup_pro as whey protein isolate, 50 grams each group, each timepoint,
-#90 grams glucose and 360 kcal.
-fp.weight <-read_excel("data/dxa/ribose_dxa.xlsx") %>%
-        select(weight, subject) 
-nutha <- read_excel("data/dxa/Ribose_nutrition_result.xlsx")%>%
-        select(timepoint, subject, group, calories, fat, carbohydrates, protein, sup_pro, sup_gluc, kcal_glu) %>%
-        mutate(timepoint = if_else(timepoint %in% c("T1","T2"), 
-                                   "Day 1",
-                                   if_else(timepoint %in% c("3", "4"),
-                                           "Day 2",
-                                           if_else(timepoint %in% c("5", "6"),
-                                                   "Day 3",
-                                                   if_else(timepoint %in% c("7", "8"),
-                                                           "Day 4",
-                                                           if_else(timepoint %in% c("9", "10"),
-                                                                   "Day 5",
-                                                                   if_else(timepoint %in% c("T3", "T4"),
-                                                                           "Day 6",
-                                                                           "na"))))))) %>%
-        group_by(timepoint, subject, group) %>%
-        inner_join(fp.weight) %>%
-        summarise(protein = sum(protein + sup_pro),
-                  fat = sum(fat),
-                  calories = sum(calories + kcal_glu),
-                  carbohydrates = sum(carbohydrates + sup_gluc),
-                  weight = sum(weight)) %>%
-        mutate(proprkg = (protein/weight))
-
-#change data
-#Protein
-pro_change <- nutha %>%
-        select(timepoint, subject, group, protein,) 
-        #ungroup() %>%
-        #pivot_wider(names_from = group,
-         #           values_from = c(protein)) %>%
-        #mutate(change = glucose - placebo) %>%
-        #print()
-prolm <- lmer(protein ~ timepoint + timepoint:group + (1|subject), data = pro_change)
-pro.plot <- plot(prolm)
-pro.sum <- summary(prolm)
-#Fat
-fat_change <- nutha %>%
-        select(timepoint, subject, group, fat,) 
-        #ungroup() %>%
-        #pivot_wider(names_from = group,
-         #           values_from = c(fat)) %>%
-        #mutate(change = glucose - placebo) %>%
-        #print()
-fatlm <- lmer(fat  ~ timepoint + timepoint:group +(1|subject), data = fat_change)
-fat.plot <- plot(fatlm)
-fat.sum <- summary(fatlm)
-#Carbohydrates
-carb_change <- nutha %>%
-        select(timepoint, subject, group, carbohydrates,) 
-        #ungroup() %>%
-        #pivot_wider(names_from = group,
-         #           values_from = c(carbohydrates)) %>%
-        #mutate(change = glucose - placebo) %>%
-        #print()
-carblm <- lmer(carbohydrates  ~ timepoint + timepoint:group + (1|subject), data = carb_change)
-cho.plot <- plot(carblm)
-cho.sum <- summary(carblm)
-#calories
-cal_change <- nutha %>%
-        select(timepoint, subject, group, calories,) 
-        #ungroup() %>%
-        #pivot_wider(names_from = group,
-         #           values_from = c(calories)) %>%
-        #mutate(change = glucose - placebo) %>%
-        #print()
-callm <- lmer(calories  ~ 0 + timepoint + timepoint:group + (1|subject), data = cal_change)
-cal.plot <- plot(callm)
-cal.sum <- summary(callm)
-#Protein pr bodyweigh
-prowe <- nutha %>%
-        select(timepoint, subject, group, proprkg) 
-prokglm <- lmer(proprkg  ~ 0 + timepoint + timepoint:group + (1|subject), data = prowe)
-prokglm.plot <- plot(prokglm)
-prokglm.sum <- summary(prokglm)
-
-```
-
 ```{r nutrition change analysis and fig}
 # Associated scripts:
-# - nut.change
-# - nut.change.figs
+# - nut.datahand.R
+# - nut.change.R
+# - nut.change.figs.R
 

### 2023-02-10 R/nut.change.R

@@ -14,7 +14,37 @@ library(tidyverse); library(lme4); library(lmerTest); library(emmeans)
 
 fp.weight <- readRDS("./data/data-gen/dxa/fp.weight.RDS")
 
-nut.res <- readRDS("./data/data-gen/dxa/nutha.RDS")
+dxa.res <- read_excel("data/dxa/Ribose_nutrition_result.xlsx")
+
+# Data handling
+
+# Data handling/cleaning
+
+nut.res <- dxa.res %>%
+  select(timepoint, subject, group, calories, fat, carbohydrates, protein, sup_pro, sup_gluc, kcal_glu) %>%
+  mutate(timepoint = if_else(timepoint %in% c("T1","T2"), 
+                             "D1",
+                             if_else(timepoint %in% c("3", "4"),
+                                     "D2",
+                                     if_else(timepoint %in% c("5", "6"),
+                                             "D3",
+                                             if_else(timepoint %in% c("7", "8"),
+                                                     "D4",
+                                                     if_else(timepoint %in% c("9", "10"),
+                                                             "D5",
+                                                             if_else(timepoint %in% c("T3", "T4"),
+                                                                     "D6",
+                                                                     "na"))))))) %>%
+  group_by(timepoint, subject, group) %>%
+  inner_join(fp.weight) %>%
+  summarise(protein = sum(protein + sup_pro),
+            fat = sum(fat),
+            calories = sum(calories + kcal_glu),
+            carbohydrates = sum(carbohydrates + sup_gluc),
+            weight = sum(weight)) %>%
+  mutate(proprkg = (protein/weight))
+
+saveRDS(nutha, "./data/data-gen/dxa/nutha.RDS")
 
 ### Macro nutrient change analysis
 # The code beneath tests whether there are differences in macro nutrient intake from pre to post, 
@@ -43,7 +73,9 @@ pro.lchange <- nut.res %>%
[…]

### 2023-02-10 R/totrna.analysis.R

@@ -56,6 +56,8 @@ rna_dat2 %>%
   geom_line() +
   geom_point()
 
+saveRDS(rna_dat2, "./data/data-gen/rna/rna.dat.RDS")
+
 ## Baseline analysis: calculation of mean and SD of baseline values, t-test on baseline values
 
 base.rna <- rna_dat2 %>%

### 2023-02-10 R/ubf_rna.correlation.R

@@ -7,98 +7,59 @@
 library(tidyverse);library(ggpubr)
 
 # Data 
+west.dat <- readRDS("./data/data-gen/protein/prot.datc.RDS") # Quantified and cleaned western data,
+                                                               # (outliers marked) normalised by pool
 
-west.dat <- readRDS("./data/data-gen/protein/prot.dat.RDS")
-west.clean <- readRDS("./data/data-gen/protein/protein_complete.RDS")
+rna.dat <- readRDS("./data/data-gen/rna/rna.dat.RDS") # Quantified and cleaned (outliers marked), 
+                                                      # total RNA data, normalised by wet muscle weight
 
-rna.dat <- readRDS("./data/data-gen/rna/rna_complete.RDS")
 
-# Ready RNA data
-rna.ratio <- read_excel("./data/rna/RNA.raw.xlsx", na = "NA")
+## Prepare data frames for joining
 
-code <- read_excel("./data/code_key.xlsx")
-
-# 260/280 Ratio for estimation of sample purity
-mean.rat <- rna.ratio %>%
-  select(sample, RATIO_R260280_1, RATIO_R260280_2, RATIO_R260280_3, RATIO_R260280_4) %>%
-  pivot_longer(names_to = "rep",
-               values_to = "ratio",
-               cols = RATIO_R260280_1:RATIO_R260280_4) %>%
-  group_by(sample) %>%
-  #print()
-  mutate(mean.ratio = mean(ratio),
-         sd.ratio = sd(ratio)) %>%
+ubf.rdy <- west.dat %>%  # One data frame without outliers
+  select(subject, time, outlier, supplement, target, norm.sign) %>%
+  filter(target == "ubf",
+         outlier == "in") %>%
+  select(-target) %>%
+  group_by(subject, time, supplement) %>%
+  summarise(mean.sign = mean(norm.sign, na.rm = TRUE)) %>%
   print()
 
-# Sort baseline data (T1 and T2) from the two legs into pre, and post biopsy data (T3 and T4) into post, calculates total RNA per wet muscle weight.
-
-rna.dat2 <- rna.dat %>%
-  inner_join(code) %>%
-  mutate(time = if_else(time %in% c("T1", "T2"), "pre", "post"), 
-         time = factor(time, levels = c("pre", "post"))) %>%
[…]

### 2023-02-10 ribose-paper.RMD
moved figs

--- 2023-02-09 ribose-paper.RMD
@@
 author: "Kristian Lian, Daniel Hammarström, Stian Ellefsen, Håvard Hamarsland"
+bibliography: resources/ribose.bib
+csl: resources/jap.csl
 output:
@@
 
-Regular resistance exercise training (RT) increases muscle mass and muscle strength, due to the adaptations stimulated by  mechanical strain.
-Unfortunately, responses to RT are not uniform across the population.
-Studies have shown RT-induced muscle growth to vary widely between individuals, with as much as 10-15% showing considerably impaired growth (ALVAREZ18;MANN14;THALACKER13).
-This may be explained by genetics, epigenetics, and/or an unfavourable internal physiological milieu (FIGUEIREDO21;HAMMARSTROM20;THALACKER13).
-While tweaking training modalties and -protocols induces different responses in different populations (HAMMARSTROM20), little is known regarding optimisation of RT to individual phenotypes.
-Generally, meta-analyses favour a moderate volume of RT over low volume RT regarding muscle growth (KRIEGER09;RHEA03;SCHOENFELD17), substantiated by recent RCTs (HAMMARSTROM20;SCHOENFELD19).
-Although Hammarström and colleagues (2020) found that a moderate volume led to more prominent gains in muscle mass, approximately 50% of the participants did not experience true benefits of moderate volume compared to low, with a blunted muscle growth.
-Due to the variations observed in muscle growth responses despite standardising RT protocols (HAMMARSTROM20;SCHOENFELD17;STEC16), it is apparent that increasing RT volume alone does not convert low-responders to high-responders.
-Therefore, other means than RT *per se* seems necessary to circumvent this discrepancy.
-
-Indeed, ingesting nutrients such as protein- and creatine-supplements effectively optimise RT (CERMAK12;LANHERS15;LANHERS17).
+Systematic resistance training (RT) increases muscle mass and strength due to repeated mechanical and metabolic stress.
+Unfortunately, responses to RT vary widely between individuals, with as many as 10-15% showing impaired growth (ALVAREZ18;MANN14;THALACKER13).
+Such inability to benefit from RT may be explained by genetics, epigenetics, and/or an unfavourable internal physiological milieu (FIGUEIREDO21;HAMMARSTROM20;THALACKER13).<!-- Check these refs, do they correspond to genet epigen and internal mileu?
+-->
+Although RT program modifications such as exercise volume or intensity may induce more favorable adaptations in some individuals, such benefits are not uniform between individuals [@hammarstrom.etal_2020]<!-- Add ref for more rt program modifications -->
+changes in e.g., training volume generally induces greater benefits of RT, not all individuals benefit from such modifications [@hammarstrom.etal_2020].
+Therefore, other means than modification of RT variables seems necessary to optimize individual responses.<!-- The above sentence might be a bit contradictory?
+-->
+
+Nutritional supplements such as protein and creatine effectively optimize RT adaptations (CERMAK12;LANHERS15;LANHERS17).
 However, it remains unknown if other nutritional adjuvants such as glucose can increase the efficacy of RT.
 This is surprising, as glucose is the preferred energy substrate of the contracting skeletal muscle during strenuous exercise and a major supplier of energy to cells via ATP synthesis (MUL15;TANAKA18).
-Energy availability is important to regulate central processes of muscle growth, specifically ribosome biogenesis, a key factor for increased muscle protein synthesis (MPS) (FIGUEIREDO19;HAMMARSTROM20;VONWALDEN19;TANAKA18).
-Interestingly, variations in RT-induced responses have been linked to ribosome biogenesis, where participants responding poorly to RT also showed a blunted ribosome biogenesis (HAMMARSTROM20;STEC16).
-A blunted ribosome biogenesis negates responses to RT by failure to produce novel ribosomes, thus an inability to increase the skeletal muscle cell’s translational capacity (VONWALDEN19;FIGUEIREDO19).
+Furthermore, energy availability is a decisive factor in *de novo* synthesis of ribosomes <!--Add ref that indicate high energy requirement for ribo biog possibly (VONWALDEN19;TANAKA18?
+--> which in turn determines muscle growth by increasing the muscles translational capacity (FIGUEIREDO19;VONWALDEN19;TANAKA18;HAMMARSTROM20;STEC16).
 Therefore, investigating the effects of combined RT and glucose ingestion may provide valuable insight into the regulation of ribosome biogenesis.
 
-Ribosome biogenesis is suggested to be regulated by the mammalian target of rapamycin complex 1 (mTORC1) network (KUSNADI15;VONWALDEn19;TANAKA18).
+Ribosome biogenesis is suggested to be in part regulated by the mammalian target of rapamycin complex 1 (mTORC1) network (KUSNADI15;VONWALDEn19;TANAKA18).
 Stimulated by growth factors, hormones, mechanical loading and nutrients, mTORC1 governs energy distribution by turning on/off energy-consuming and energy-generating pathways, depending on cellular energy levels (HOPPE09).

### 2023-02-10 figures/figure1.R

@@ -1,0 +1,8 @@
+### Ribose figure 1 ######
+#
+# Study design, strength (1RM, Humac)
+#
+
+
+
+

### 2023-02-13 ribose-paper.RMD
updates to introduction

@@
 Systematic resistance training (RT) increases muscle mass and strength due to repeated mechanical and metabolic stress.
-Unfortunately, responses to RT vary widely between individuals, with as many as 10-15% showing impaired growth (ALVAREZ18;MANN14;THALACKER13).
+Unfortunately, responses to RT vary widely between individuals, with as many as 10-15% showing impaired muscle growth (ALVAREZ18;MANN14;THALACKER13).
 Such inability to benefit from RT may be explained by genetics, epigenetics, and/or an unfavourable internal physiological milieu (FIGUEIREDO21;HAMMARSTROM20;THALACKER13).<!-- Check these refs, do they correspond to genet epigen and internal mileu?
 -->
-Although RT program modifications such as exercise volume or intensity may induce more favorable adaptations in some individuals, such benefits are not uniform between individuals [@hammarstrom.etal_2020]<!-- Add ref for more rt program modifications -->
-changes in e.g., training volume generally induces greater benefits of RT, not all individuals benefit from such modifications [@hammarstrom.etal_2020].
+Although RT program modifications such as altered exercise volume or intensity may induce more favorable adaptations in some individuals, such benefits are not uniform between individuals [@hammarstrom.etal_2020].<!-- Add ref for more rt program modifications -->
+For example, although changes in training volume generally induces greater benefits of RT, not all individuals benefit from such modifications [@hammarstrom.etal_2020].
 Therefore, other means than modification of RT variables seems necessary to optimize individual responses.<!-- The above sentence might be a bit contradictory?
@@
 
-Ribosome biogenesis is suggested to be in part regulated by the mammalian target of rapamycin complex 1 (mTORC1) network (KUSNADI15;VONWALDEn19;TANAKA18).
-Stimulated by growth factors, hormones, mechanical loading and nutrients, mTORC1 governs energy distribution by turning on/off energy-consuming and energy-generating pathways, depending on cellular energy levels (HOPPE09).
-This greatly affects ribosome biogenesis, being a process that consumes approximately 80% of cellular energy and materials (VONWALDEN19;TANAKA18).
-Upon stimulation by mTORC1, assembly of the preinitiation complex (PIC) marks the initiation of ribosome biogenesis (VONWALDEN19;FIGUEIREDO19).
-The PIC consists of the RNA polymerase I complex, the SL-1 complex and the TIF-1A/RRN3 protein (FIGUEIREDO19, VONWALDEN19).
-Further, the upstream binding factor (UBF) and c-Myc have vital roles in the initiation, where UBF binds the PIC to the rDNA promoter and c-Myc as a transcription factor (FIGUEIREDO19, VONWALDEN19).
-Pol I transcribes the 47S pre-rRNA, suggested as the rate-limiting step in ribosome biogenesis (VONWALDEN19, FIGUEIREDO19).
-The 47S pre-rRNA is processed into the 18S, 5.8S and 28S rRNA (MULLINEUX12, VONWALDEN19), while the last rRNA 5S is transcribed by Pol III, and the ribosomal proteins by Pol II (MULLINEUX12, VONWALDEN19).
-Importantly, these events facilitates increased ribosomal content of each skeletal muscle cell, subsequently allowing a higher basal- and roof for MPS, suggested to predict long term adaptations to RT (HAMMARSTROM21).
-Since a blunted ribosome biogenesis has been observed to reduce/inhibit RT responses, despite increasing mechanical loading (HAMMARSTROM20), targeting hormonal and/or nutritional signals may ameliorate this.
-Previously, high glucose treatment promoted ribosome biogenesis compared to a repressed ribosome biogenesis with low glucose treatment and glucose starvation in mice and cell cultures (MARIAPPAN11;TANAKA15;ZHAI12), making glucose an interesting player in this perspective.
-
-Thus, though it seems apparent that sufficient daily intake of glucose/carbohydrates is necessary for an optimal ribosome biogenesis (TANAKA18), it is unclear whether there are acute effects on ribosome biogenesis *per se* with ingesting glucose before and after RT.
-Further, there are equivocal observations on insulin's impact on ribosome biogenesis and MP, where previous meta-analyses suggest insulin to act in concert with amino acid intake to enhance MPS, and insulin may reduce skeletal muscle protein breakdown when amino acids are scarce (ABDULLA16).
-Therefore, to investigate the relationship between glucose, RT and ribosome biogenesis we recruited 16 healthy young, moderately trained male and female individuals to moderate volume unilateral resistance training for two weeks.
+Transcription of ribosomal RNA (rRNA) by Pol I is considered the limiting step in synthesizing new ribosomes [@Moss_1995].
+Multiple signaling pathways converge to regulate rRNA transcription, including the mammalian target of rapamycin (mTORC1) signal-transduction pathway and c-Myc (KUSNADI15;VONWALDEn19;TANAKA18).
+mTOR receives input from growth factors, hormones, mechanical loading, and nutrients to balance protein synthesis through multiple mechanisms based on cellular energy levels (HOPPE09).
+Importantly, mTOR and its downstream target S6K1 serve as a mediator from insulin signaling in skeletal muscle[@hillier.etal_2000].
+mTOR contributes to forming the preinitiation complex (PIC) that marks the initiation of rRNA transcription (VONWALDEN19;FIGUEIREDO19).
+Parallel to mTORC1 [@Mori_2021; West_2016], c-Myc increases ribosomal biogenesis directly and through transcriptional control of the upstream binding factor (UBF) [@sanijUBFLevelsDetermine2008; Poortinga_2011].
+However, interestingly, UBF phosphorylation, required for interaction with the rDNA promoter, is increased by high glucose in a mTORC1 dependent manner (rapamycin sensitive) in kidney glomerular epithelial cells[@Mariappan_2011].
+Additionally, independently from UBF, high glucose was shown to lead to chromatin remodeling, which promotes rRNA transcription[@Zhai_2012].
+Together these observations indicate a potential role of glucose in positively affecting muscle rRNA synthesis.
+
+The purpose of this investigation was to determine the effect of 
+<!-- -->
+
+the relationship between glucose, RT and ribosome biogenesis we recruited 16 healthy young, moderately trained male and female individuals to moderate volume unilateral resistance training for two weeks.
 They exercised one leg with glucose and protein, and the other leg with placebo and protein ingestion before and after each RT-session.

### 2023-02-20 results_work.RMD
added some work on the results, needs more

--- 2023-02-13 ribose-paper.RMD
@@
 ---
-title: "Glucose ingestion before and after resistance training does not augment ribosome biogenesis in healthy moderately trained young adults"
-author: "Kristian Lian, Daniel Hammarström, Stian Ellefsen, Håvard Hamarsland"
-bibliography: resources/ribose.bib
-csl: resources/jap.csl
-output:
-  word_document:
-    reference_docx: reference-template-02.docx
+title: "results"
+output: word_document
+date: "2023-02-20"
 editor_options: 
@@
 ```{r packages}
-library(tidyverse);library(readxl);library(lme4);library(emmeans);library(dplyr);library(cowplot);library(knitr);library(broom);library(rmarkdown);library(tinytex);library(tidyr);library(arsenal);library(magrittr);library(dabestr);library(nlme); library(pbkrtest)
+library(tidyverse);library(readxl);library(lme4);library(emmeans);library(dplyr);library(cowplot);library(knitr);library(broom);library(rmarkdown);library(tinytex);library(tidyr);library(arsenal);library(magrittr);library(dabestr);library(nlme); library(pbkrtest);library(ggpubr)
 ```
-
-## Introduction
-
-Systematic resistance training (RT) increases muscle mass and strength due to repeated mechanical and metabolic stress.
-Unfortunately, responses to RT vary widely between individuals, with as many as 10-15% showing impaired muscle growth (ALVAREZ18;MANN14;THALACKER13).
-Such inability to benefit from RT may be explained by genetics, epigenetics, and/or an unfavourable internal physiological milieu (FIGUEIREDO21;HAMMARSTROM20;THALACKER13).<!-- Check these refs, do they correspond to genet epigen and internal mileu?
--->
-Although RT program modifications such as altered exercise volume or intensity may induce more favorable adaptations in some individuals, such benefits are not uniform between individuals [@hammarstrom.etal_2020].<!-- Add ref for more rt program modifications -->
-For example, although changes in training volume generally induces greater benefits of RT, not all individuals benefit from such modifications [@hammarstrom.etal_2020].
-Therefore, other means than modification of RT variables seems necessary to optimize individual responses.<!-- The above sentence might be a bit contradictory?
--->
-
-Nutritional supplements such as protein and creatine effectively optimize RT adaptations (CERMAK12;LANHERS15;LANHERS17).
-However, it remains unknown if other nutritional adjuvants such as glucose can increase the efficacy of RT.
-This is surprising, as glucose is the preferred energy substrate of the contracting skeletal muscle during strenuous exercise and a major supplier of energy to cells via ATP synthesis (MUL15;TANAKA18).
-Furthermore, energy availability is a decisive factor in *de novo* synthesis of ribosomes <!--Add ref that indicate high energy requirement for ribo biog possibly (VONWALDEN19;TANAKA18?
---> which in turn determines muscle growth by increasing the muscles translational capacity (FIGUEIREDO19;VONWALDEN19;TANAKA18;HAMMARSTROM20;STEC16).
-Therefore, investigating the effects of combined RT and glucose ingestion may provide valuable insight into the regulation of ribosome biogenesis.
-
-Transcription of ribosomal RNA (rRNA) by Pol I is considered the limiting step in synthesizing new ribosomes [@Moss_1995].
-Multiple signaling pathways converge to regulate rRNA transcription, including the mammalian target of rapamycin (mTORC1) signal-transduction pathway and c-Myc (KUSNADI15;VONWALDEn19;TANAKA18).
-mTOR receives input from growth factors, hormones, mechanical loading, and nutrients to balance protein synthesis through multiple mechanisms based on cellular energy levels (HOPPE09).
-Importantly, mTOR and its downstream target S6K1 serve as a mediator from insulin signaling in skeletal muscle[@hillier.etal_2000].
-mTOR contributes to forming the preinitiation complex (PIC) that marks the initiation of rRNA transcription (VONWALDEN19;FIGUEIREDO19).
-Parallel to mTORC1 [@Mori_2021; West_2016], c-Myc increases ribosomal biogenesis directly and through transcriptional control of the upstream binding factor (UBF) [@sanijUBFLevelsDetermine2008; Poortinga_2011].
-However, interestingly, UBF phosphorylation, required for interaction with the rDNA promoter, is increased by high glucose in a mTORC1 dependent manner (rapamycin sensitive) in kidney glomerular epithelial cells[@Mariappan_2011].
-Additionally, independently from UBF, high glucose was shown to lead to chromatin remodeling, which promotes rRNA transcription[@Zhai_2012].
-Together these observations indicate a potential role of glucose in positively affecting muscle rRNA synthesis.
-
-The purpose of this investigation was to determine the effect of 
-<!-- -->
-
-the relationship between glucose, RT and ribosome biogenesis we recruited 16 healthy young, moderately trained male and female individuals to moderate volume unilateral resistance training for two weeks.
-They exercised one leg with glucose and protein, and the other leg with placebo and protein ingestion before and after each RT-session.
-Our main purpose was to investigate whether there are acute effects of glucose- compared to placebo ingestion before and after RT.
-To assess this purpose, we measured changes in ribosome biogenesis as accumulation of total RNA, expression of the four mature rRNA's (18S, 5.8S, 28S, 5S), 47S pre-rRNA, and levels of proteins associated with transcription initiation (c-Myc, UBF, rpS6?).
-As secondary outcomes, we measured the development in skeletal muscle performance during the intervention, and skeletal muscle recovery after the final RT session, as temporal changes in maximal isometric and isokinetic torque.
-We hypothesised that RT would indeed lead to a prominent accumulation of total RNA, increased expression of rRNA and associated proteins.
-Further, we hypothesised that RT with glucose and protein would induce an enhanced response compared to placebo, in markers of ribosome biogenesis, as well as muscular performance and -recovery.
-
-## Methods
-
-### Materials and methods
-
-All participants were informed about the potential discomforts and risks associated with the study and gave their informed consent before study enrollment.
-The study was approved by the regional ethical committee (REK, ID nr.
-153628), pre-registered at clinicaltrials.gov (Identifier: NCT04545190), and conducted according to the Helsinki declaration.
-
-### Participants
-
-Sixteen healthy male and female participants (20-33yrs, Tab 1) were recruited to the study through social media advertisement and word of mouth, and through completing the selection process (Fig 1).
-The eligibility criteria were non-smokers and moderately trained (i.e.
-2-8 RT sessions per 14 days for the last six months).
-Exclusion criteria were previous injury leading to impaired strength, inability to perform resistance training, symptoms and a medical record of metabolic disorders including hyperglycemia, i.e.
-fasting venous plasma glucose ≥6.1 mmol/L and/or 2-hour glucose tolerance ≥7.8 mmol/L, and/or HbA1c >42 mmol/mol.
-Of the sixteen participants that commenced the intervention, three participants dropped out.
-One due to sickness and inability to resume, two participants experienced muscular discomfort connected to heavy resistance training (Fig 1).
-Baseline characteristics (Tab 1) were measured using DXA (Prodigy Advance PA+302047, Lunar, San Francisco, CA, USA) on Day -1, the last day preceding the RT intervention.
-
-### Experimental design
-
-The study was designed as a 12-day double-blinded randomized controlled trial, with a unilateral RT protocol (Fig 2).
-Participants were randomized to commence with either glucose (GLU, 30g per bolus) or placebo (PLAC, 0g per bolus), and either dominant or non-dominant leg (Fig 1), alternating RT and supplement from one day to the other (Fig 2).
-Investigators were blinded as the supplement randomization was disclosed only to personnel not involved in any other aspect of the study, and participants were blinded by diluting boluses of glucose (30g glucose) (Glucosum monohydricum, Merck KGaA, Darmstadt, Germany) and placebo (~0.3g Stevia rebaudiana extract) (Steviosa, Soma Nordic AS, Oslo, Norway) in 300ml Fun Light juice/saft (Orkla, Oslo, Norway).
-A blinded taste test revealed that the participants were not able to disclose the contents of the provided boluses (30g GLU vs.
-0g GLU).
-To ensure equal conditions during training sessions, participants exercised and tested at the same time of day, +/- 1hr with the same trainer, and recorded and repeated their daily macro nutrient intake on pairwise consecutive days (Tab 4).
-
-###### **Table 1:** Baseline characteristics of the participants. BMI: body mass index, A = charactaristics per sex, B = characteristics per leg. Values are mean ± standard deviation.
-
-```{r DXA}
-# Script: DXA
-
-# Data handling
-table.dxa <- read_excel("./data/dxa/ribose_dxa.xlsx") %>%
-        select(sex, subject, age, height, weight, leanmasskg, fatmasskg, ffmg) %>%
-        #make grams into kg
-        mutate(ffkg = ffmg/1000) %>%
-        select(sex, age, height, weight, leanmasskg, fatmasskg, ffkg) %>%
-  mutate(bmi = (weight) / (height^2) * 10000)
-
-dxa.table <- table.dxa %>%
-        pivot_longer(names_to = "variable",
-                     values_to = "values", col= age:bmi) %>%
-        select(sex, variable, values) %>%
-        group_by(sex, variable) %>%
-        summarise(m = mean(values, na.rm = TRUE),
-                  s= sd(values, na.rm = TRUE)) %>%
-        mutate(stat = paste0(round(m, 1), " (", round(s, 1), ")")) %>%
-        select(sex, variable, stat) %>%
-        pivot_wider(names_from = sex,
-                    values_from = stat) %>%
-                mutate(variable = if_else(variable %in% c("age"),
-                                          "Age (yrs)",
-                                          if_else(variable %in% c("fatmasskg"),
-                                                  "Fat mass (kg)",
-                                                  if_else(variable %in% c("ffkg"),
-                                                          "Fat free mass (kg)",
-                                                          if_else(variable %in% c("height"),
-                                                                  "Height (cm)",
-                                                                  if_else(variable %in% c("leanmasskg"),
-                                                                          "Lean body mass (kg)",
-                                                                          if_else(variable %in% c("weight"),
-                                                                                  "Body weight (kg)",
-                                                                                  if_else(variable %in% c("bmi"),
-                                                                                          "BMI", "")
-                                                                                  )))))))
-
-dxa.table %>%
-        kable(col.names = c("Variable", "Female (n=7)", "Male (n=9)"))
-
-```
-
-```{r dxa per leg}
-# Script: DXA
-
-# Subject 104, 110 and 111 dropped out and is therefore excluded from the table
-
-#Load dxa data
-dxaleg <-  read_excel("./data/dxa/ribose_dxaprleg.xlsx") %>%
-        select(subject, sex, leg, leanmassg, fatmassg, totalmasskg) %>%
-        #make grams into kg
-        mutate(leanmass = leanmassg/1000,
-               fatmass = fatmassg/1000)%>%
-        select(subject, sex, leg, leanmass, fatmass, totalmasskg)
-
-# Between groups analysis 
-dxa.stat <- dxaleg %>%
-        select(subject, leg, leanmass, fatmass, totalmasskg)
-
-# Lean mass between legs, within subject 
-leanmass.lm <- lmer(leanmass ~ leg + leg:subject + (1|subject), data = dxa.stat)
-lm.plot <- plot(leanmass.lm)
-lm.sum <- summary(leanmass.lm)
-lm.confint <- emmeans(leanmass.lm, specs = ~ "leg")
-
-# Fat mass between legs, within subject
-fat.lm <- lmer(fatmass ~ leg + leg:subject +(1|subject), data = dxa.stat)
-fat.plot <- plot(fat.lm)
-fat.sum <- summary(fat.lm)
-fat.confint <- emmeans(fat.lm, specs = ~ "leg")
-
-# Total mass between legs, within subject
-total.lm <- lmer(totalmasskg ~leg + leg:subject + (1|subject), data = dxa.stat)
-totlm.plot <- plot(total.lm)
-totlm.sum <- summary(total.lm)
-supsup <- emmeans(total.lm, specs = ~ "leg")
-
-# Table of leanmass, fatmass and total mass in kg pr. leg
-dxalegtable <- dxaleg %>%
-        select(leg, leanmass, fatmass, totalmasskg) %>%
-        pivot_longer(names_to = "variable",
-                     values_to = "values", col= leanmass:totalmasskg) %>%
-        group_by(leg, variable) %>%
-        summarise(m = mean(values, na.rm = TRUE),
-                  s= sd(values, na.rm = TRUE)) %>%
-        mutate(stat = paste0(round(m, 1), " (", round(s, 1), ")")) %>%
-        select(leg, variable, stat) %>%
-        pivot_wider(names_from = leg,
-                    values_from = stat)
-
-# Markdown table
-dxalegtable %>%
-        kable(col.names = c("Variable", "Left leg", "Right leg"))
-
-```
-
-Six RT sessions were conducted with glucose and six with placebo, allowing a within-subjects analysis of the effects of glucose ingestion.
-Data from the first five RT sessions was used to investigate main outcome measures (total RNA, rRNA and protein), whereas data from the sixth RT session was used to explore secondary outcomes (muscular recovery, plasma glucose levels).
-Participants were asked to refrain from resistance- or high-intensity training of the legs from Day -7 and onwards, until completion of the intervention and post-testing, to ensure the reliability of pre-intervention strength data and minimal interference from other training sources.
-If participants did conduct exercise training outside of the prescribed protocol, they were asked to ensure equal loading on pairwise consecutive days.
-
-![](figures/flowchart_selection.PNG)
-
-###### **Figure 1:** Flowchart of the selection process. The order in which participants performed the two intervention blocks was determined in a planned randomised fashion.
-
-![](figures/timeline.PNG)
-
-###### **Figure 2.** Overview of the intervention, with 12 days of concomitant dietary intervention and resistance training (RT), preceded by 7 days of testing.
-T0 = timepoint 0 (testing days prior to intervention), T1/T2 = biopsy prior to resistance training for leg #1 and #2, respectively, T3/T4 post intervention testing of leg #1 and #2, respectively, including post sixth RT session.
-Test = physical performance test (unilateral 1RM leg press, knee extension, isometric knee extension force, isokinetic knee extension torque) testing on days 4, 5, 8, 9, T3, T4 will not include 1RM leg press and knee extension, Biopsy = microbiopsy, RT#1 = resistance training leg #1, RT#2 = resistance training leg#2.
-Other tests (D2O, DXA, spit) will be conducted as part of the study as a whole.
-These are not relevant to this data set, thus will not be described here.
-
-### Dietary intervention
-
-The dietary intervention spanned the whole day, divided into three periods: I) From awakening until two hours after RT, II) from 2.5 hours after RT until 2200hrs, and III) from 2200hrs until awakening.
-During period I, participants ingested protein and glucose/placebo only.
-Glucose/placebo was ingested at three time points in period I: 30 minutes before RT (0830hrs, 30g vs 0g glucose), immediately before RT (0900hrs, 30g vs.
-0g glucose), and immediately after RT (~0930hrs, 30g vs.
-0g glucose).
-Protein was ingested 2hrs before RT (0700hrs, 25g) and immediately after RT (~0930hrs, 25g).
-In the afternoon (1800hrs-1900hrs, period II) participants ingested glucose or placebo (3x30g vs.
-3x0g glucose) opposite to the supplement they received during RT, to ensure a balanced daily intake of glucose.
-Apart from this, participants ingested a self-chosen diet during period II, registered in MyFitnessPal or similar applications.
-The self-chosen diet was repeated on pairwise consecutive days (i.e.
-on days 1-2, 3-4, etc.), to ensure similar premises for resistance training responses between the two legs.
-During period III (2200-0700hrs), participants remained in an overnight fasted state.
-The daily onset of the dietary intervention (i.e., first ingestion of PRO supplement) varied between participants, from 0600hrs to 0900hrs to allow multiple participants to complete the protocol simultaneously.
-The protein supplement was ingested as 25g Whey Protein Isolate boluses (Proteinfabrikken, Stokke, Norway), diluted in 150ml water.
-During sessions, participants were free to ingest water ad libitum.
-
-### Assessment of muscular strength
-
-Performance tests (Test, Fig 2) were performed prior to (Days -7 and -5, and T0, both legs) and during the intervention (Days 4, 5, 8, and 9), after session 5 and after finalization of the intervention (T3/T4, RT#1 leg; T4/Day 13, RT#2 leg), a total of 10 days (Fig 2).
-Maximal isometric and isokinetic knee extension torque were measured with a Humac Norm Dynamometer (CSMi, Stoughton, Massechusetts, USA).
-Individual positions were recorded and standardized from pre-intervention tests (days -7 and -5).
-Maximal isokinetic torque was measured at speeds of 60- and 240 d/s, 2x3 repetitions each, with the first set of each exercise as a sub-maximal warm-up.
-The maximal isometric torque was measured at knee-angle 60, for a maximum of 10 seconds and two repetition per test.
-The highest peak torque values were carried forwards to the final analyses.
-During days 4, 5, 8 and 9, humac tests were conducted one hour before RT with the leg performing RT the previous day.
-During days 11 and 12 (T3 and T4), humac tests were performed at four time points: I) 45min before RT, II) 30min after the last RT session, III) 2hrs after the last RT session, and IV) 23hrs after the last RT session.
-Test I at T4 included testing of both legs, representing 23hrs post RT session test of one leg and pre-session 6 test of the other leg.
-
-Assessment of unilateral one repetition maximum (1RM) leg press and knee extension was conducted at pre-intervention testing (days -7 and -5, Fig 2).
-The participants performed a general warm-up with 10min cycling on an indoor exercise bicycle and a specific warm-up (1x10, 1x6 and 1x3 repetitions, 2min breaks) before each of the tests.
-Between specific warm-up sets as well as between each subsequent 1RM attempt, the participants were given 3min breaks.
-All positions were controlled and recorded at pre-intervention testing and repeated for the RT sessions.
-Maximal leg press strength was defined as the maximal load lifted in a controlled fashion, with a knee angle of 45 degrees (start), within a 2.5kg range.
-Attempts, where participants did not reach 45 degrees during the eccentric phase, were not approved.
-Maximal knee extension testing followed the same specific warm-up and pause protocol and was defined as maximal load lifted in a controlled fashion, reaching full extension of the knee joint, within a 1.25kg range.
-Attempts with exaggerated hip movement or beneath full extension were not approved.
-
-### Resistance training protocol
-
-Resistance training consisted of three sets of each unilateral leg presses and knee extensions, with an intensity of 10 repetition maximum (10RM).
-As a general warm-up, the participants cycled on an indoor exercise bicycle for 5-10min.
-In addition, before the respective exercises, two 10-repetition warm-up sets were completed at ~50% and ~70% of 10RM.
-Rest time between working sets was 2min, and progressive loading was used to ensure adequate exercise stimulation throughout the intervention.
-For safety and standardization purposes, all sessions were monitored by trained personnel.
-To further ensure equal exercise premises for the two legs, each participant was accompanied by the same trainer in each session.
-If this was not possible throughout the intervention, each participant had the same trainer on pairwise consecutive days as a minimum.
-The rate of perceived exertion (RPE, 0–10-point scale) was logged before every session, and the session score was logged 15min after each session.
-Lastly, training volume (i.e., load and repetitions) was logged for every session.
-
-### Sampling of muscle tissue and blood
-
-Muscle biopsies were sampled from m.
-vastus lateralis using well-established procedures (HAMMARSTRØM20).
-Muscle biopsy sampling was performed under local anaesthesia (Xylocaine, 10mg ml-1 with adrenaline 5μg ml−1, AstraZeneca AS, Oslo, Norway) using a 12-14-gauge needle (Universal Plus, Mermaid Medical AS, Stenløse, Denmark), operated with a spring-loaded biopsy gun.
-After the biopsy sampling, muscle tissue was divided in two aliquots for determination of total RNA/expression of rRNA and two aliquots for protein analyses, snap frozen in isopentane (-80°C) and stored at -80°C until further analyses.
-Muscle biopsies were sampled at four time points (Fig 2): I/II) Pre-intervention (2hrs before training, time point T1 = pre RT#1 leg, T2 = pre RT#2 leg), and III/IV) before the sixth RT session from RT#1 leg (2hrs before training, T3) and before the sixth RT session from RT#2 leg (2hrs before training, T4).
-At each time point, two samples were taken from the same incision.
-To standardize this procedure, all individual participants had biopsies taken at the same time of day, in an overnight fasted state.
-
-To measure blood glucose levels with and without glucose ingestion/training, blood was collected by means of finger draws on days with biopsy sampling.
-One finger draw was taken at T1 (pre-test, Fig 2) to serve as a baseline.
-At T3 and T4 (post-test, Fig 2), finger draws were collected at 7 time points: I) prior to protein ingestion (0700hrs) II) 45 minutes after protein ingestion (0745hrs) III) 1.5hrs after protein ingestion (0800hrs, i.e., immediately before GLU/PLAC intake), IIII) 2hrs after protein ingestion (0900hrs, i.e., immediately before training), IV) in the middle of RT (~0915hrs), V) immediately after training (~0930hrs), and VI) 2hrs after completion of training (~1130hrs).
-Finger draws were analysed with in-house equipment (BIOSEN C-Line, EKF diagnostic GmbH, Barleben).
-Venous blood samples were collected from the antecubital vein, coinciding with the finger draws except in the middle of the RT session, to analyse endocrine variables.
-
-### Total RNA extraction and quantitative real-time reverse transcription polymerase chain reaction
-
-Two muscle biopsy samples were taken for total RNA extraction per leg per time point, and total RNA was extracted in two duplicates per muscle biopsy.
-Thus, we had two duplicates per leg per time point, giving a total of eight RNA samples per participant.
-Total RNA was extracted using TRIzol according to the manufacturer’s protocol.
-An exogenous RNA control (λ polyA External Standard Kit, Takara Bio Inc., Shiga, Japan) was added at a fixed amount (0.04 ng ml-1 of Trizol reagent) to enable analysis of target gene expression per unit tissue weight (ELLEFSEN2008,2014).
-Briefly, muscle tissue was homogenized in TRIzol using 0.5mm RNase-free Zirconium beads (~50 ul) (Next Advanced, Averill Park, NY, USA) and a bullet blender (bb).
-Chloroform (Sigma-Aldrich, Missouri, USA) was used for phase-separation, and an RNA pellet was precipitated with isopropanol (VWR International, Pennsylvania, USA).
-For assessment of RNA content and -purity 5µl RNA, from the RNA stock, was eluted in TE-buffer (1:2), for assessment via spectrophotometry.
-All samples had a 260nm to 280nm ratio > 1.9.
-The RNA stock was stored at -80°C until further analyses.
-RNA was reverse transcribed using Super Script IV Reverse Transcriptase (Invitrogen), according to the manufacturer’s instructions.
-Briefly, 1µl 10 mM dNTP mix, 0.5µl anchored oligo-dT, 0.5µl random hexamer primers (Thermo Scientific), a maximum of 9µl template RNA and 2 µl nuclease-free water were mixed, vortexed and briefly spun down.
-Samples were then heated at 65°C for 5 minutes, followed by at least 1-minute incubation on ice.
-For the next step, 4µl 5x SSIV Buffer, 1µl 100 mM DTT, 1 µl RNase OUT, and Super Script IV Reverse Transcriptase (Invitrogen) were mixed, vortexed and briefly spun down and added to the samples.
-Samples were incubated for 10 minutes at 23°C, 10 minutes at 50-55°C and 10 minutes at 80°C.
-All samples were reverse transcribed and diluted to 1:50 before quantitative real-time polymerase chain reaction (qPCR).
-Lambda was used as an external reference gene, added in the RNA extraction (2µl per extraction).
-qPCR reactions were run on a fast-cycling real-time detection system (Applied Biosystems 7500 fast Real-Time PCR Systems, Life Technologies AS), with a total volume of 10 µl, containing 2 µl of cDNA, gene-specific primers (0.5 µM final concentration) and a commercial master mix (2X SYBR Select Master Mix, Applied Biosystems, Life Technologies AS) (Hammarström et al.
-2020).
-qPCR reactions consisted of 40 cycles (3 s 95°C denaturing and 30 s 60°C annealing) (Hammarström et al.
-2020).
-Raw fluorescence data were exported from the platform-specific software and amplification curves were modelled with a best-fit sigmoidal model using the qPCR package (Ritz & Spiess, 2008) written for R (Hammarström et al.
-2020; Team 2018).
-For an overview of primers, see Table 2.
-
-###### **Table 2:** Primer sequence and performance. Average cycle thresholds (Ct) and priming efficiencies were calculated from all qPCR reactions.
-
-![](figures/Primer.sequences.PNG)
-
-### Protein extraction and immunoblotting
-
-Total protein was extracted using the Minute Total Protein Extraction Kit for Muscles (Invent Biotechnology), according to the manufacturer’s protocol, optimised for our lab.
-Wet muscle was freeze-dried for 24hrs and dissected before extraction.
-The tissue was homogenized with a plastic rod in 80mg protein extraction powder (Invent Biotechnology) and 100ul iced cell lysis buffer (Denaturing Buffer, Invent Biotechnology), and centrifuged at top speed for 1 min.
-The supernatant was aliquoted into 2x25ul tubes to run samples in duplicates, and total protein concentrations were determined in a 1:10 dilution (Pierce Detergent Compatible Bradford Assay Reagent, Thermo Fisher Scientific, Oslo, Norway).
-The protein samples were diluted to 20.25 µg µl−1 total protein with lysis buffer and 4X Laemmli sample buffer (Bio-Rad Laboratories AB, Oslo, Norway) containing 2-mercaptoethanol.
-Twenty microlitres from each protein sample were pooled as a normalisation factor for gel separation.
-All protein samples were incubated at 95°C and stored at -20°C until further analysis.
-The protein samples (20.25µg total protein) were separated at 250V for 50min using 4-20% Tris-Glycine gels (Thermo Fisher Scientific), then transferred to PVDF membranes with wet transfer at 300mA for 3hrs.
-Both gel electrophoresis and protein transfer were performed at 4°C.
-Following the wet transfer, membranes were stained with a MemCode reversible total protein stain (MANUFACTURER) and then blocked for 1hr at room temp with a blocking buffer of Tris-buffered saline (TBS; 20 mM Tris, 150 mM NaCl) with 5% non-fat dry milk and 0.1% Tween-20.
-Primary and secondary antibodies were purchased from Santa Cruz Biotechnology (Texas, USA) UBF:(UBF-9, sc-13125), rpS6:(Ribosomal protein S6 C-8, sc-74459), and Thermo Fisher Scientific () c-Myc:(c-Myc 9E10), goat anti-mouse for c-Myc (goat anti-mouse IgG1 (y1) horseradish peroxidase conjugate), and anti-mouse (anti-mouse IgG1 horseradish peroxidase conjugate).
-Antibodies were diluted in a 5% non-fat dry milk blocking buffer (TBS, 5% non-fat dry milk, 0.1% Tween-20).
-Membranes were incubated overnight with primary antibodies, and for 1hr with secondary antibodies.
-Between blocking and primary antibody staining, membranes were washed 1x5min, between primary and secondary staining, and after secondary staining, membranes were washed 3x5 with TBS-Tween (TBS; 20mM Tris, 150mM NaCl, 0.1% Tween).
-Following the last TBS-T wash, membranes were incubated 1x5min with enhanced chemiluminescent substrate (ECL, SuperSignal West Femto Maximum Sensitivity Substrate, Thermo Fisher Scientific).
-Membrane blocking, secondary antibody incubation, washing and ECL incubation were performed at room temp, on a tipping board.
-Primary antibody incubation was performed at 4°C on a tipping board.
-Chemiluminescence signals were quantified using Image Studio Lite (LI-COR Biotechnology, Lincoln, NE, USA), and total protein content was quantified using ImageJ (RUEDENETAL2017), whereas total protein content was defined as mean grey value of the whole well with between-well values subtracted as background.
 
@@
 Further, on pairwise consecutive days, there were no significant differences in mean macro nutrient (calories, carbohydrate, fat or protein) ingestion (Tab 4), or mean change in macro nutrient ingestion (Fig 3).
-Resistance training with glucose led to significant increases in blood levels glucose and c-peptide compared to RT with placebo, before, during and after the session (Fig 4B, 120min: 38 ± 4% [*p* = `r p.120`], 135min: 31 ± 4% [*p* = `r p.135`], 150min: 32 ± 4% [*p* = `r p.150`], Fig 4C, 120min: 95 ± 10% [*p* = `r cpp.120`], 150min: 87 ± 10% [*p* = `r cpp.150`]), measured during the final RT session.
-At 2hrs post-exercise, RT with glucose led to significantly lower blood glucose levels compared to RT with placebo, with a tendancy of the same in c-peptide levels (4B, 270min: -8 ± 4% [*p* = `r p.270`] and 4C, 270min: -7 ± 10 [*p* `r cpp.270`]).
+Resistance training with glucose led to significant increases in blood levels glucose and c-peptide compared to RT with placebo, before, during and after the session (Fig 1B, 120min: 38 ± 4% [*p* = `r p.120`], 135min: 31 ± 4% [*p* = `r p.135`], 150min: 32 ± 4% [*p* = `r p.150`], Fig 1C, 120min: 95 ± 10% [*p* = `r cpp.120`], 150min: 87 ± 10% [*p* = `r cpp.150`]), measured during the final RT session.
+At 2hrs post-exercise, RT with glucose led to significantly lower blood glucose levels compared to RT with placebo (1B, 270min: -8 ± 4% [*p* = `r p.270`].
 
-###### **Table 3**: Baseline characteristics of the intervention legs, organized per supplement.
-Variables: Iso 60 = isokinetic 60 d/s, Iso 240 = isokinetic 240 d/s, training intensity = load/rm*100.
-Nm = newton meters, kg = kilograms, %1RM = percentage of 1 repetition maximum, ng/mg = nanogram per milligram wet muscle tissue, nf.expr = expression normalized by weight and external reference gene (Lambda).
-Ribosomal RNA data are presented as log transformed.
-Glucose n = 13, placebo n = 13.
+FIG1 HERE
 
-![](figures/table3.png)
-
-###### **Table 4:** Daily mean dietary intake of macro nutrients during each training day per supplement leg.
-G = glucose, P = placebo.
-Training day 1-training day 12 are combined pairwise to day 1-6: Day 1 = training day 1/2, Day 2 = training day 3/4, Day 3 = training day 5/6, Day 4 = training day 7/8, Day 5 = training day 9/10, Day 6 = training day 11/12.
-Values are mean ± standard deviation.
-P = difference between glucose and placebo.
-Glucose n = 12, placebo n = 12.
-Day 6: Glu = 11, Plac = 11.
-
-```{r nutrition table}
-## Scripts
-# - nut.table.R
-# - nut.analysis.R
-# - dxa.R
-
-# Data
-nuttable <- readRDS("./data/data-gen/dxa/nuttable.RDS")
-
-nuttable %>%
-  kable(col.names = c("Time", "Suppl.", "Kcal", "CHO (g)", "Fat (g)", "PRO (g)", "PRO (g/kg)"))
-
-```
-
-```{r nutrition change analysis and fig}
-# Associated scripts:
-# - nut.datahand.R
-# - nut.change.R
-# - nut.change.figs.R
-
-nut.fig <- readRDS("./data/data-gen/nutrition/nut.fig.RDS")
-nut.fig
-
-```
-###### **Figure 3:** Mean changes in macro nutrient intake from day to day.
-Day 1 = training day 1/2(T1/T2), Day 2 = training day 3/4, Day 3 = training day 5/6, Day 4 = training day 7/8, Day 5 = training day 9/10, Day 6 = training day 11/12(T3/T4).
-Values are mean change in grams/calories ± 95% CL.
-
-![](figures/day.timeline.png)
-
-```{r glucose + c-peptide fig}
-## Scripts
-# - gluc.fig.R
-# - cpep.figs.R
-
-# Separate figures
-pos <- position_dodge(width = 0.2)
-
-gluc.fig <- readRDS("./data/data-gen/glucose/gluc.fig.RDS")
-
-cpep.fig <- readRDS("./data/data-gen/glucose/cpep.fig.RDS")
-
-# Joined figure
-plot_grid(gluc.fig,
-          cpep.fig,
-          ncol = 1, nrow = 2)
-
-```
-
-###### **Figure 4:** A) Time line for training day 6 (combination of pairwise training day 11 and 12).
-Biopsy = muscle biopsy taken pre supplementation, PRO = protein supplementation, Humac = strength test (isometric knee extension, isokinetic knee extension 60 and 240 d/s), GLU/PLAC = glucose/placebo supplementation, RT = resistance training.
-Minutes: -120 = 2hrs prior to RT, -90 = 90min prior to RT, -30 = 30min prior to RT, 0 = onset of RT, 15 = 15min following onset of RT, between leg press and knee extension, 30 = 30min post onset of RT, 120 = 2hrs following onset of RT.
-B) Mean mmol/L change in blood glucose levels per supplement group during training day 6.
-Time-points: 0, 45, 90, 120, 135, 150 and 270min post protein ingestion, respectively.
-C) Mean change in pmol/L of c-peptide in blood per supplement group during training day 6.
-Values are presented as estimated marginal means of mmol/L and pmol/L change ± 95% CI.
-*p* < 0.05 between groups.
+**Figure 1:** A) Overview of the experimental design with 12 days of concomitant dietary intervention and resistance training (RT), preceded by 7 days of testing.
+Familiarization = days -7 to -1 included familiarization to RT exercises (via 1RM leg press and knee extension tests) and strength tests (Humac dynamometer) at days -7 and -5.
+Baseline strength tests were conducted at day -1.
+Biopsies were taken from vastus lateralis at baseline (Day 1 leg 1, Day 2 leg 2) and after five RT sessions (Day 11 leg 1, Day 12 leg 2).
+The order in which participants performed the two intervention blocks was determined in a planned randomized fashion, where one leg was randomized to RT with GLU (glucose) and the other to RT with PLA (placebo).
+The randomization also entailed half starting with GLU (n = 8) and the other half with PLA (n = 8), in addition they where further randomized to start either with their dominant or non-dominant leg.
+Thus, 4 participants were randomized to start RT with their dominant leg and GLU supplement, and 4 with their non-dominant leg and GLU - the same was true for participtants commencing the intervention with PLA.
+From intervention start (Day 1 leg 1, Day 2 leg 2), the participants exercised in a unilateral, alternating fashion until post intervention testing (Day 11/12 leg 1, Day 12/13 leg 2).
+Three participants dropped out of the intervention, either due to reported muscular discomforts to heavy strength training or sickness in the family.
+B and C) Changes in Plasma glucose (mmol/L) and -c-peptide levels (pmol/L) during RT session six (Day 11 leg 1, Day 12 leg 2).
+Finger draws and venous blood were sampled at baseline (Day -1), as well as before, during and after session six.
+Plasma glucose was measured via finger draws, 120 (-120), 90 (-90) and 30 (-30) minutes and immediately before RT (0), during RT (15), immediately after (30) and 2hrs after RT (120).
+Levels of c-peptide were measured via venous blood samples collected at the same time points as finger draws, except for 90min before RT and during RT.
+D, E, F) Changes in muscular performance (D: nm fold change in isometric peak torque, E/F: nm fold change in isokinetic peak torque at 60/240 d/s.).
+Muscular performance was measured via a Humac Norm Dynamometer at baseline (Day -1), during (Day 4 and 8 leg 1, Day 5 and 9 leg 2), and after five RT sessions (Day 11 leg 1, Day 12 leg 2).
+G, H, I) Changes in muscular recovery (G: nm fold change in isometric peak torque, H/I: nm fold change in isokinetic peak torque at 60/240 d/s).
+Muscular recovery was measured via Humac Norm Dynamometer with baseline being post fifth RT session test, and then 30min, 2hrs and 23hrs following the sixth and final RT session.
+Values are presented as estimated marginal means ± 95% CI.
+* = *p* < 0.05 between groups.
 Glucose n = 13, placebo n = 13.
@@
 
-Resistance training with glucose did not induce a higher accumulation of total RNA (fig 5A, *p* = `r trdiffp`), or rRNA  (fig 5B, *p* > 0.05), compared to RT with placebo, measured as mean change from baseline to post.
+Resistance training with glucose did not induce a higher accumulation of total RNA (fig 2A, *p* = `r trdiffp`), or rRNA  (fig 2B-2F *p* > 0.05), compared to RT with placebo, measured as mean change from baseline to post.
 From baseline to post-intervention, there was a mean increase in levels of total RNA by 26 and 22% after RT with glucose and placebo, respectively (*p* < 0.05, compared to baseline).
 A robust accumulation was also observed in ribosomal RNA expression in both RT with glucose and placebo, with mean increases between 34-43% (GLU) and 33-41% (PLAC) (*p* < 0.05, compared to baseline) in the four rRNA’s.
-The expression of 47S pre-rRNA increased by 37 and 59% with glucose and placebo respectively, with only placebo reaching significantly higher levels at post compared to baseline (fig 45, *p* = `r rr47tdiffp`).
+The expression of 47S pre-rRNA increased by 37 and 59% with glucose and placebo respectively, where only RT with placebo increased significantly from baseline to post (fig 45, *p* = `r rr47tdiffp`).
 
-```{r RNA figures}
-# Get figures
-totrna.fig <- readRDS("./data/data-gen/rna/totrna.fig.RDS")
-totrna.fig
+FIG2 HERE
 
-plot.18 <- readRDS("./data/data-gen/rna/plot.18.RDS")
-plot.28 <- readRDS("./data/data-gen/rna/plot.28.RDS")
-plot.5.8 <- readRDS("./data/data-gen/rna/plot.5.8.RDS")
-plot.5 <- readRDS("./data/data-gen/rna/plot.5.RDS")
-plot.47 <- readRDS("./data/data-gen/rna/plot.47.RDS")
-
-# Joined rRNA
-library(cowplot)
-
-legend <- get_legend(plot.47 + theme(legend.box.margin = margin(0, 0, 0,12)))
-
-plot_grid(plot.18 + theme(legend.position = "none"),
-          plot.28 + theme(legend.position = "none"), 
-          plot.5.8 + theme(legend.position = "none"),
-          plot.5 + theme(legend.position = "none"), 
-          plot.47 + theme(legend.position = "none"),
-          legend,
-          ncol = 2, nrow = 3)
-
-```
-
-###### **Figure 5:** A) Changes in Total RNA, B) Changes in ribosomal RNA.
+###### **Figure 2:** Changes in total RNA and ribosomal RNA.
+A) Total RNA, B) 47S pre-rRNA, C) 18S rRNA, D) 28S rRNA, E) 5.8S rRNA, F) 5S rRNA.
 Baseline = Training day 1/2, Post = Training day 11/12.
-Both total RNA and rRNA were analyzed with duplicates, with two duplicates per biopsy (two biopsies per time point).
+Total RNA and rRNA were analyzed in duplicates, with two duplicates per biopsy (two biopsies per time point).
 Total RNA and rRNA changes were calculated as log-fold change score per mg wet muscle weight.
 Mean change scores of the duplicates were calculated before modelling and transformed to the log-scale.
-Values are estimated marginal means fold change per leg/supplement ± 95% CI.
+Values are estimated marginal means fold change per leg per supplement ± 95% CI.
 p = between groups statistic, glucose compared to placebo.
@@
 ubf <- prot.change %>%
-  filter(target == "ubf")
+  filter(target == "ubf",
+         subject != "103") # Subject 103 has missing values
 
@@
 
-saveRDS(cmyc.emm, "./data/data-gen/protein/cmyc.change.RDS")
-
 # UBF change
@@
 
-saveRDS(ubf.emm, "./data/data-gen/protein/ubf.change.RDS")
-
 # rps6 change
@@
   data.frame()
-
-saveRDS(rps6.emm, "./data/data-gen/protein/rps6.change.RDS")
 
@@
 
-```{r western fig}
+```{r correlation model}
 # Data 
+west.dat <- readRDS("./data/data-gen/protein/prot.datc.RDS") # Quantified and cleaned western data,
+                                                               # (outliers marked) normalised by pool
 
-cmyc <- readRDS("./data/data-gen/protein/cmyc.change.RDS")
+rna.dat <- readRDS("./data/data-gen/rna/rna.dat.RDS") # Quantified and cleaned (outliers marked), 
+                                                      # total RNA data, normalised by wet muscle weight
 
-ubf <- readRDS("./data/data-gen/protein/ubf.change.RDS")
+## Prepare data frames for joining
 
-rps6 <- readRDS("./data/data-gen/protein/rps6.change.RDS")
+ubf.rdy <- west.dat %>%  # One data frame without outliers
+  select(subject, time, outlier, supplement, target, norm.sign) %>%
+  filter(target == "ubf",
+         outlier == "in") %>%
+  select(-target) %>%
+  group_by(subject, time, supplement) %>%
+  summarise(mean.sign = mean(norm.sign, na.rm = TRUE)) %>%
+  print()
 
-## cmyc fig
+rna.rdy <- rna.dat %>%
+  select(subject, time, outlier, supplement, RNA.weight) %>%
+  filter(outlier == "in") %>%
+  mutate(subject = as.character(subject)) %>%
+  group_by(subject, time, supplement) %>%
+  summarise(mean.rna = mean(RNA.weight, na.rm = TRUE)) %>%
+  print()
 
-cmyc.plot <- cmyc %>%
-  mutate(time = "post") %>%
-  add_row(supplement = c("PLACEBO", "GLUCOSE"), time = "pre", emmean = 0, lower.CL = 0, upper.CL = 0) %>%
-  mutate(time = factor(time, levels = c("pre", "post"), labels = c("Baseline", "Post"))) %>%
-  ggplot(aes(time, exp(emmean), group = supplement, fill = supplement)) +
-  #annotate("text", x = "Post", y = 2.1, label = "p = 0.585", size = 3) +
-  geom_errorbar(aes(ymin = exp(lower.CL), ymax = exp(upper.CL)),
-                width = 0.1,
-                position = position_dodge(width = 0.2)) +
-  geom_line(position = position_dodge(width = 0.2)) +
-  geom_point(shape = 21, size = 3, position = position_dodge(width = 0.2)) +
-  labs(x = "", y = "c-Myc signal \n(fold change)\n", fill = "") +
-  theme_classic()
+## Correlation analysis
 
-ubf.plot <- ubf %>%
-  mutate(time = "post") %>%
-  add_row(supplement = c("PLACEBO", "GLUCOSE"), time = "pre", emmean = 0, lower.CL = 0, upper.CL = 0) %>%
-  mutate(time = factor(time, levels = c("pre", "post"), labels = c("Baseline", "Post"))) %>%
-  ggplot(aes(time, exp(emmean), group = supplement, fill = supplement)) +
-  #annotate("text", x = "Post", y = 2.1, label = "p = 0.585", size = 3) +
-  geom_errorbar(aes(ymin = exp(lower.CL), ymax = exp(upper.CL)),
-                width = 0.1,
-                position = position_dodge(width = 0.2)) +
-  geom_line(position = position_dodge(width = 0.2)) +
-  geom_point(shape = 21, size = 3, position = position_dodge(width = 0.2)) +
-  labs(x = "", y = "UBF signal \n(fold change)\n", fill = "") +
-  theme_classic()
+joined.dat <- ubf.rdy %>%
+  right_join(rna.rdy) %>%
+  #filter(supplement != "GLUCOSE") %>%
 
-rps6.plot <- rps6 %>%
-  mutate(time = "post") %>%
-  add_row(supplement = c("PLACEBO", "GLUCOSE"), time = "pre", emmean = 0, lower.CL = 0, upper.CL = 0) %>%
-  mutate(time = factor(time, levels = c("pre", "post"), labels = c("Baseline", "Post"))) %>%
-  ggplot(aes(time, exp(emmean), group = supplement, fill = supplement)) +
-  annotate("text", x = "Post", y = 1.9, label = "p = ", size = 3) +
-  geom_errorbar(aes(ymin = exp(lower.CL), ymax = exp(upper.CL)),
-                width = 0.1,
-                position = position_dodge(width = 0.2)) +
-  geom_line(position = position_dodge(width = 0.2)) +
-  geom_point(shape = 21, size = 3, position = position_dodge(width = 0.2)) +
-  labs(x = "", y = "rps6 signal \n(fold change)\n", fill = "") +
-  theme_classic()
+  mutate(sd.ubf = (mean.sign - mean(mean.sign, na.rm = TRUE )/sd(mean.sign, na.rm = TRUE))) %>%
+  print()
 
-# Cowplot for gathering figures
+m <- lme(log(mean.rna) ~ mean.sign + time , 
+         random = list(subject = ~ 1), 
+          data = joined.dat, 
+         na.action = na.omit)
 
-legend <- get_legend(rps6.plot + theme(legend.box.margin = margin(0, 0, 0,12)))
-
-prot.joined <- plot_grid(cmyc.plot + theme(legend.position = "none"),
-                         ubf.plot + theme(legend.position = "none"), 
-                         rps6.plot + theme(legend.position = "none"),
-                         legend,
-                         ncol = 2, nrow = 2)
-
-prot.joined
-
-saveRDS(prot.joined, "./data/data-gen/rna/prot.joined.RDS")
-
+summary(m)
+plot(m)
 ```
 
-###### **Figure 5:** A) Changes in c-Myc levels, B) changes in UBF levels, C) changes in rps6 levels.
-Baseline = Training day 1/2, post = Training day 11/12.
-Four samples were analysed per participant (GLU pre-post, PLAC pre-post), in duplicates, a total of 104 protein samples.
-Duplicate samples were loaded on separate gels.
-In addition, 3-4 pools were added to each well.
-c-Myc, UBF and rps6 signals were then calculated as log-fold change score, normalised by gel pools.
+Interestingly, RT with placebo led to significantly higher change in levels of rpS6 and UBF protein compared to RT with glucose, with no difference in levels of c-Myc (Fig 3A rpS6: *p* = `r rpdiff`, 3C UBF: *p* = `r udiff`, 3B c-Myc: *p* = `r cdiff`).
+Levels of rpS6 change by 40% in placebo and 7% in glucose (Fig 3A), UBF levels changed by 68% in placebo and 26% in glucose (Fig 3C), and c-Myc levels changed by 110% in placebo and 59% in glucose (Fig 3B).
+
+FIG3 HERE
+
+###### **Figure 3:** Changes in levels of protein.
+A) rPS6, B) c-Myc, C) UBF.
+Baseline = Training day 1/2, Post = Training day 11/12.
+Protein samples were analyzed in duplicates, with two duplicates per biopsy (two biopsies per time point).
+The duplicates were loaded on different gels, in an inverted order.
+Change in protein levels were calculated as log-fold change score normalised by pools (pools of all protein samples).
 Mean change scores of the duplicates were calculated before modelling and transformed to the log-scale.
-Values are estimated marginal means fold change per leg/supplement ± 95% CI.
-p = between groups statistic, glucose compared to placebo.
+Values are estimated marginal means fold change per leg per supplement ± 95% CI.
+* = significant between groups statistic, glucose compared to placebo.
 Glucose n = 13, placebo n = 13.
 
-### Resistance training and strength testing
+Further, we observed a linear relationship between levels of UBF and accumulation of total RNA, where ...
 
-```{r training data model}
-## Scripts
-# - volume.change.R
-# - ss.analysis.R
-# - ss.cleanup.R
-# - trainint.analysis.R
-# - trainint.cleanup.R
+FIG4 HERE
 
-# Log-transformed data
-vol.lchange <- readRDS("./data/data-gen/training/vol.lchange.RDS")
-ss.lchange <- readRDS("./data/data-gen/training/ss.lchange.RDS")
-int.lchange <- readRDS("./data/data-gen/training/int.lchange.RDS")
-
-### Training volume
-# Fold-change model
-vol.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = vol.lchange)
-vol.lres <- plot(vol.lmod)
-vol.lsum <- summary(vol.lmod)
-
-# Fold-change estimated means
-vol.lemm <- confint(emmeans(vol.lmod, specs = ~"supplement|time")) %>%
-        data.frame()
-
-### Session score (RPE)
-# Model
-ss.lmod <- lmerTest::lmer(change ~ baseline + time + supplement:time + (1|subject),
-                     data = ss.lchange)
-ss.lres <- plot(ss.lmod)
-ss.lsum <- summary(ss.lmod)
-
-# Session score emmeans
-ss.lemm <- confint(emmeans(ss.lmod, specs = ~"supplement|time")) %>%
-        data.frame()
-
-### Training intensity
-int.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = int.lchange)
-int.lres <- plot(int.lmod)
-int.lsum <- summary(int.lmod)
-
-# Emmeans
-int.lemm <- confint(emmeans(int.lmod, specs = ~"supplement|time")) %>%
-        data.frame()
-
-### For inline reporting
-
-## Training volume
-# Diff 4
-vcoef <- coef(summary(vol.lmod))
-vdiff.4 <- round(vcoef[9,1], 3)
-vdiffp.4 <- round(vcoef[9,5], 3)
-vdiffse.4 <- round(vcoef[9,2], 3)
-# Diff 5
-vdiff.5 <- round(vcoef[10,1], 3)
-vdiffp.5 <- round(vcoef[10,5], 3)
-vdiffse.5 <- round(vcoef[10,2], 3)
-# Diff 6
-vdiff.6 <- round(vcoef[11,1], 3)
-vdiffp.6 <- round(vcoef[11,5], 3)
-vdiffse.6 <- round(vcoef[11,2], 3)
-## RPE
-scoef <- coef(summary(ss.lmod))
-sdiffp <- round(scoef[11, 5], 3)
-sdiffse <- round(scoef[11, 2], 3)
-
-## Training intensity
-
-# Diff time 2
-icoef <- coef(summary(int.lmod))
-idiff2 <- round(icoef[7,1], 3)
-idiffp2 <- round(icoef[7,5], 3)
-idiffse2 <- round(icoef[7,2], 3)
-# Diff time 3
-idiff3 <- round(icoef[8,1], 3)
-idiffp3 <- round(icoef[8,5], 3)
-idiffse3 <- round(icoef[8,2], 3)
-# Diff time 4
-idiff4 <- round(icoef[9,1], 3)
-idiffp4 <- round(icoef[9,5], 3)
-idiffse4 <- round(icoef[9,2], 3)
-# Diff time 5
-idiff5 <- round(icoef[10,1], 3)
-idiffp5 <- round(icoef[10,5], 3)
-idiffse5 <- round(icoef[10,2], 3)
-# Diff time 6
-idiff6 <- round(icoef[11,1], 3)
-idiffp6 <- round(icoef[11,5], 3)
-idiffse6 <- round(icoef[11,2], 3)
-```
-
-```{r humac post 5th model}
-## Scripts
-# - humac.cleanup.R
-# - humac.main.analysis.R - this is data gathered at baseline, during the intervention, and after session 5
-
-# Log-transformed data
-
-isom.lchange <- readRDS("./data/data-gen/humac/isom.lchange.RDS")
-iso60.lchange <- readRDS("./data/data-gen/humac/isok60.lchange.RDS")
-iso240.lchange <- readRDS("./data/data-gen/humac/isok240.lchange.RDS")
-
-# Isometric
-isom.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = isom.lchange)
-isom.lres <- plot(isom.lmod)
-isom.lsum <- summary(isom.lmod)
-
-# Isok.60
-iso60.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = iso60.lchange)
-iso60.lres <- plot(iso60.lmod)
-iso60.lsum <- summary(iso60.lmod)
-
-# Isok.240
-iso240.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = iso240.lchange)
-iso240.lres <- plot(iso240.lmod)
-iso240.lsum <- summary(iso240.lmod)
-
-## Emmeans
-# Isometric
-isom.lemm <- confint(emmeans(isom.lmod, specs = ~"supplement|time")) %>%
-        data.frame()
-
-# Isok.60
-iso60.lemm <- confint(emmeans(iso60.lmod, specs = ~"supplement|time")) %>%
-        data.frame()
-
-# Isok.240
-iso240.lemm <- confint(emmeans(iso240.lmod, specs = ~"supplement|time")) %>%
-        data.frame()
-
-# For reporting
-# Isom diff
-isom5coef <- coef(summary(isom.lmod))
-isom5diffp <- round(isom5coef[7, 5], 3)
-isom5diffse <- round(isom5coef[7, 2], 3)
-
-# Isok 60 diff
-isok605coef <- coef(summary(iso60.lmod))
-isok605diffp <- round(isok605coef[7, 5], 3)
-isok605diffse <- round(isok605coef[7, 2], 3)
-
-# Isok 260
-isok2405coef <- coef(summary(iso240.lmod))
-isok2405diffp <- round(isok2405coef[7, 5], 3)
-isok2405diffse <- round(isok2405coef[7, 2], 3)
-
-```
-
-```{r post 6th model}
-## Scripts
-# - humac.acute.analysis.R - this is data gatherede before and after one RT session (sixth session)
-# - humac.cleanup.R
-
-# Data
-isom.lchange.ac <- readRDS("./data/data-gen/humac/isom.lchange.ac.RDS")
-iso60.lchange.ac <- readRDS("./data/data-gen/humac/iso60.lchange.ac.RDS")
-iso240.lchange.ac <- readRDS("./data/data-gen/humac/iso240.lchange.ac.RDS")
-
-# Isometric
-isomac.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = isom.lchange.ac)
-isomac.lres <- plot(isomac.lmod)
-isomac.lsum <- summary(isomac.lmod)
-
-# Isok.60
-iso60ac.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = iso60.lchange.ac)
-iso60ac.lres <- plot(iso60ac.lmod)
-iso60ac.lsum <- summary(iso60ac.lmod)
-
-# Isok.240
-iso240ac.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = iso240.lchange.ac)
-iso240ac.lres <- plot(iso240ac.lmod)
-iso240ac.lsum <- summary(iso240ac.lmod)
-
-# Emmeans
-# Isometric
-isomac.lemm <- confint(emmeans(isomac.lmod, specs = ~"supplement|time")) %>%
-        data.frame()
-
-# Isok.60
-iso60ac.lemm <- confint(emmeans(iso60ac.lmod, specs = ~"supplement|time")) %>%
-        data.frame()
-
-# Isok.240
-iso240.lemm <- confint(emmeans(iso240ac.lmod, specs = ~"supplement|time")) %>%
-        data.frame()
-
-## For inline reporting
-# Isom diff
-isomcoef <- coef(summary(isomac.lmod))
-isomdiffp <- round(isomcoef[7, 5], 3)
-isomdiffse <- round(isomcoef[7, 2], 3)
-
-# Isok 60 diff
-isok60coef <- coef(summary(iso60ac.lmod))
-isok60diffp <- round(isok60coef[7, 5], 3)
-isok60diffse <- round(isok60coef[7, 2], 3)
-
-# Isok 240 diff
-isok240coef <- coef(summary(iso240ac.lmod))
-isok240diffp <- round(isok240coef[7, 5], 3)
-isok240diffse <- round(isok240coef[7, 2], 3)
-
-```
-
-There were no significant differences in mean change of total session volume, training intensity or RPE between exercise training with glucose and placebo at any of the time points (fig 6, *p* > 0.05).
-We did observe interactions between time and changes, where total session training volume increased from 13-18% compared to baseline.
-These increases in training volume were accompanied by changes ranging from 5-18% in training intensity and 2-13% in RPE (Tab 5).
-The maximal isometric knee extension force (OR TORQUE?) and isokinetic knee extension peak torque at 240 d/s showed no differences between exercise training with glucose and placebo after five RT sessions (Fig 7A, *p* > 0.05).
-There was a significant difference in mean change from pre to post-fifth RT session, where exercise training with glucose led to a 9% less decrease in peak torque at 60 d/s compared to placebo (Fig 7A, *p* = `r isok605diffp`).
-There was no difference between glucose and placebo in mean change from baseline to after the sixth RT session (Fig 7B, *p* > 0.05).
-
-![](figures/legend.PNG)
-
-```{r training dat fig}
-### Get figures
-
-# Volume
-totvol.barplot <- readRDS("./data/data-gen/training/totvol.barplot.RDS")
-totvolfold.plot <- readRDS("./data/data-gen/training/totvolfold.plot.RDS")
-
-# Training intensity
-tot.barplot <- readRDS("./data/data-gen/training/tot.barplot.RDS")
-totalint.plot <- readRDS("./data/data-gen/training/trainint.plot.RDS")
-
-# Join figures
-plot_grid(totvol.barplot + theme(legend.position = "none"),
-          totvolfold.plot + theme(legend.position = "none"),
-          tot.barplot + theme(legend.position = "none"),
-          totalint.plot + theme(legend.position = "none"),
-          ncol = 2,
-          nrow = 2,
-          align = "h")
-
-```
-
-###### **Figure 6:** Training volume and training intensity from baseline until resistance exercise session 6 (training day 11 and 12).
-Baseline = Training day 1/2, Session 2 = Training day 3/4, Session 3 = Training day 5/6, Session 4 = Training day 7/8, Session 5 = Training day 9/10, Session 6 = Training day 11/12.
-A) Changes in total session volume per leg, values are mean kg ± standard deviation.
-B) Fold change in total session volume per leg, values are estimated marginal means fold change ± 95% CI.
-C) Changes in training intensity per leg calculated as %1RM, values are mean % ± standard deviation.
-D) Fold change in training intensity per leg, values are estimated marginal means fold change ± 95% CI.
-† = *p* < 0.05 in both glucose and placebo, compared to baseline.
-p = NS = no significant difference between legs at any of the time points.
-Glucose n = 13, placebo n = 13.
-
-###### **Table 5:** RPE changes through training days.
-Training days are presented as combined pairwise consecutive, linked to their respective sessions.
-Session 1 = Training day 1/2, Session 2 = Training day 3/4, Session 3 = Training day 5/6, Session 4 = Training day 7/8, Session 5 = Training day 9/10, Session 6 = Training day 11/12.
-Values are presented as mean ± standard deviation.
-Glucose n = 13, placebo n = 13.
-
-```{r RPE table}
-## Scripts
-# - ss.cleanup.R
-# - ss.absolute.R
-
-# Data
-ss.abs <- readRDS("./data/data-gen/training/ss.abs.RDS")
-
-# Table
-ss.abs %>%
-  kable(col.names = c("Supplement", "Baseline", "Session 2", "Session 3", "Session 4", "Session 5",
-                            "Session 6"))
-
-```
-
-![](figures/legend2.PNG)
-
-```{r humac figs}
-## Post 5th RE session
-# Figures
-pos <- position_dodge(width = 0.2)
-isom5.plot <- readRDS("./data/data-gen/humac/isom.main.fig.RDS")
-isok605.plot <- readRDS("./data/data-gen/humac/isok60.main.fig.RDS")
-isok2405.plot <- readRDS("./data/data-gen/humac/isok240.main.fig.RDS")
-
-legend <- get_legend(isom5.plot + theme(legend.box.margin = margin(0, 0, 0, 12)))
-
-plot_grid(isom5.plot + theme(legend.position = "none"), 
-          isok605.plot + theme(legend.position = "none"), 
-          isok2405.plot + theme(legend.position = "none"),
-          nrow = 3,
-          align = "v")
-
-## Post 6th RE session
-
-# Figures
-
-pos <- position_dodge(width = 0.2)
-
-isom.plot <- readRDS("./data/data-gen/humac/isom.ac.plot.RDS")
-isok60.plot <- readRDS("./data/data-gen/humac/isok60.ac.plot.RDS")
-isok240.plot <- readRDS("./data/data-gen/humac/isok240.ac.plot.RDS")
-
-legend <- get_legend(isom.plot + theme(legend.box.margin = margin(0, 0, 0, 12)))
-
-plot_grid(isom.plot + theme(legend.position = "none"), 
-          isok60.plot + theme(legend.position = "none"), 
-          isok240.plot + theme(legend.position = "none"),
-          nrow = 3,
-          align = "v")
-
-```
-
-The mean change in peak torque at all three velocities change little from baseline until session 4 in both glucose and placebo (+4-7% change), decreasing at pre-session 6 (9-18% change, Fig 6A).
-Post session 6 (Fig 6B) showed a slight reduction in isometric peak torque 30min after session completion (2-4% change), slowly regaining at 2hrs (+3% change) and 23hrs (+3-6% change), in both glucose and placebo compared to baseline.
-Isokinetic peak torque at 60 d/s and 240 d/s showed no reduction in mean peak torque from baseline until 23hrs following RT,  in both glucose and placebo (+0,1-11%).
-There was an effect of time in peak torque at 23hrs post-RT at 60 d/s (Fig 6B), where exercise training with placebo led to a 5% increase in peak torque compared to baseline, however, there was no significant difference compared to exercise training with glucose.
-
-###### **Figure 7:** Peak torque changes in maximal isometric knee extension and maximal isokinetic knee extension at 60 and 240 d/s, respectively.
-A) Humac tests during the intervention, Baseline = Day -1 (prior to intervention), Session 2 = prior to training, training Day 4/5, Session 4 = prior training, training Day 8/9, Session 6 = prior to training 11/12.
-B) Baseline = prior to training, training day 11/12, 30min = 30 min after RT, 2hr = 2hrs after RT, 23hr = 23hrs after RT.
-Values are change in estimated marginal means of newton meters (Nm) ± 95% CI.
-* = *p* < 0.05 between groups.
-*p* = NS = no significant difference between legs at any of the time points.
-Glucose n = 13, placebo n = 13.
-
-## Discussion
-
-The main findings of the present study were that resistance training with glucose for five sessions did not affect markers of ribosome biogenesis, measured as total RNA, rRNA and rDNA transcription initiation associated proteins, compared to RT with placebo.
-Similarly, RT with glucose did not affect markers of skeletal muscle functionality such as muscular performance and -recovery, total session volume, and training intensity, compared to RT with placebo.
-There are studies suggesting that high vs.
-low glucose treatment/glucose starvation impacts ribosome biogenesis in yeast, rodent and human non-muscle cells (KIM13;MARIAPPAN11;TANAKA15;ZHAI12), however, this effect does not seem evident when total daily glucose intake is matched.
-
-To our knowledge, no previous study has investigated the effects of glucose compared to placebo ingestion with RT on ribosome biogenesis in human skeletal muscle cells, making our hypotheses exploratory and comparisons not straight forwards.
-The current model of ribosome biogenesis suggests that changes in the accumulation of total RNA and rRNA early after the onset of an RT period predict long-term RT adaptations (FIGUEIREDO19; HAMMARSTROM20; HAMMARSTROM21).Therefore, we assumed 2 weeks to be a suitable time frame for a moderate RT volume to elicit changes in ribosome biogenesis, as similar observations have been made previously (FIGUEIREDO16; HAMMARSTROM20; HAMMARSTROM21).
-In Hammarström et al.
-(2020), participants that increased ribosome biogenesis after two weeks (5 RT sessions) of RT, also responded well after the entire intervention (12 weeks, 31 sessions).
-Figueiredo and colleagues (2016) also observed markedly acute effects, where 47S pre-rRNA increased after a single RT session, passing baseline after 4hrs, peaking at 24hrs and remaining elevated until 48hrs post-exercise.
-The four mature rRNA did not measurably increase until after multiple sessions (FIGUEIREDO16).
-As such, increases in total RNA and rRNA observed in the present study are in line with previous observations (FIGUEIREDO16; HAMMARSTROM20; H21?), and also evident i moderately trained individuals, compared to the untrained or RT-untrained individuals of previous studies (FIGUEIREDO16;HAMMARSTROM20;H21?).
-
-Further, the accumulation of total RNA and rRNA is also suggested to reach a plateau after approximately 5 RT-sessions, thus, in addition to the demanding RT- and dietary protocol and the exploratory nature of the hypotheses, we did not conduct a more long-term intervention.
-COMMENT ON PROTEIN DATA HERE?
-
-As we did not observe any effect of ingesting glucose before and after RT compared to placebo in this design, there is little to suggest that glucose *per se* has any additional effect on ribosome biogenesis, as long as the daily energy intake is sufficient to support cellular energy levels required to drive this process.
-Looking at previous observations from *in vitro* studies of rodents, yeast, and non-muscle cell cultures (MARIAPPAN11; KIM13; TANAKA18; ZHAI12; HOPPE09; MURAYAMA08), one interesting notion would be how high glucose vs.
-low glucose or glucose starvation would affect ribosome biogenesis over time in human skeletal muscle cells.
-Time being of the essence in such a perspective, as previous *in vitro* studies have utilized cells with higher physiological rate than human skeletal muscle cells (KIM13;MARIAPPAN11;TANAKA15;ZHAI12).
-
-Indeed, our findings suggest that glucose ingestion *per se* before and after RT does not induce changes in the internal physiological milieu of human skeletal muscle cells with anabolic effects.
-It has been theorised that glucose may convey an anabolic signal through insulin, and thus operate as a signal mediator affecting upstream pathways of ribosome biogenesis (KUSNADI15; TANAKA18).
-Such an anabolic signal could then possibly increase PIC assembly, and consequently Pol I recruitment, initiating rDNA transcription and subsequent rRNA maturation, leading to increased ribosome content (FIGUEIREDO19; TANAKA18).
-As significantly higher blood levels of c-peptide and glucose were observed in RT with glucose compared to placebo in the present study, without any difference in ribosome biogenesis, there is little to suggest that either insulin or glucose has any direct acute effects.
-
-Another theory is that glucose ingestion leads to higher cellular energy levels compared to placebo, stimulating energy-sensitive pathways such as PIH1, mTORC1, ERK1/2, AMPK and SIRT1 in a dose-response relationship (KIM13;MARIAPPAN11;TANAKA15;ZHAI12).
-The mTORC1, along with ERK1/2 and PIH1 up-regulates PIC assembly during high-energy conditions, while AMPK and possibly SIRT1 represses PIC assembly during low-energy conditions, as observed in rodent (KILDE), yeast (KILDE), and non-muscle cells (KILDE) *in vitro*.
-Specifically, high glucose treatment augmented rDNA transcription by increased UBF and SL-1 activity, while low glucose treatment, glucose starvation, or nutrient depletion stimulated AMPK-dependent TIF-IA and chromatin regulation suppression (KOS-BRAUN17; MARIAPPAN11; ZHAI12).
-In addition to UBF and SL-1, assembly of the PIC requires TIF-IA and Pol I recruitment, induced by amino acid intake in an mTOR-S6K-dependent manner and possibly by direct c-Myc activation (KUSNADI15).
-As such, increasing energy supply and nutrient availability should in line with these theories potently stimulate PIC assembly and rDNA transcription.
-Importantly, these pathways were not directly investigated in the present study.
-However, downstream targets (UBF, c-Myc, rpS6?) remained unaffected by glucose compared to placebo ingestion, suggesting there is not a dose-response relationship between muscle cellular energy levels and ribosome biogenesis during an RT-intervention (Not sure if I can say this).
-
-A weakness in the present study, is the lack of data on intracellular glucose levels in the muscle cells.
-Thus, it cannot be said for sure whether there were differences between glucose and placebo or not.
-
-A possible explanation for the lack of effect of glucose compared to placebo, is that both treatments may have exercised with sufficient levels of nutrients and energy, meaning that mTORC1 had sufficient nutrient availability and, thus was not inhibited by AMPK due to energy deficiency (HOPPE09; KUSNADI15).
-Indeed, mTORC1 is characterised as energy-sensitive (DENNIS01), and therefore it might have been directly affected by high vs.
-low-glucose conditions (MARIAPPAN11).
-However, previous studies suggest that mTORC1 is not sensitive enough to register levels of ATP, rather depending on signals from AMPK to assess ATP levels (KIM10).
-This would arguably make AMPK the primary regulator of the potential impact of glucose on ribosome biogenesis, and thus, glucose has no direct effect on mTORC1 signalling (REFERENCES?).
-Importantly, intramuscular glycogen content was not measured in the present study, therefore we cannot determine whether cellular energy levels in fact were equal or not, or if there is reason to assume that AMPK-dependent mTORC1 signalling was different in the two treatments.
-Potentially, such data could give higher resolution to our data.
-Along with the robust increases in proxy markers of ribosome biogenesis, the equal macronutrient intake between treatments on pairwise consecutive days and plasma glucose levels not indicating energy deficiency during RT, potent mTORC1 stimulation without AMPK-dependent inhibition may be a plausible explanation.
-Together with previous studies on this topic (MARIAPPAN11; ZHAI12), there are currently scarce observations to suggest glucose carries a direct anabolic signal.
-As such, physiological levels of plasma glucose during RT, combined with a sufficient daily macronutrient intake, seem potent enough to supply the energy needed for ribosome biogenesis.
-The lacking differences between treatments in the biological data are further supported by our observations on functional data.
-
-Overall, glucose did not affect the isometric or isokinetic peak torque at 240 d/s, or muscular recovery following the sixth and last RT session, measured as temporal changes in isometric and isokinetic peak torque.
-Generally, the mean isometric and isokinetic peak torque slightly decreased during the intervention, especially between session 4 and 6, and regained after the intervention.
-Further, there were no differences between glucose and placebo in total session volume or training intensity.
-Both exercise with glucose and placebo yielded higher total session volumes at higher percentage of 1RM at the last session compared to baseline, as expected from earlier reports on the moderate-volume RT protocol (KRIEGER09; RHEA03; SCHOENFELD17), in addition to a more recent study implementing the unilateral protocol (HAMMARSTROM20+21/22).
-Interestingly, we did observe a significant difference between treatments in mean change of isokinetic peak torque at 60 d/s before the sixth RT session, where exercise with glucose decreased peak torque 9% less compared to placebo.
-However, as none of the other measurements indicated any difference between treatments, a true effect of glucose seems unlikely.
-Taken together, these results may add to previous observations with no differences between co-ingestion of carbohydrate and protein compared to protein alone, in regard to skeletal muscular recovery following RT (KOOPMAN07; STAPLES11).
-Collectively, the functional data point towards an exhaustive, yet growth-inducing RT protocol, especially considering the biological data.
-As a side note, an interesting observation was made by another master’s thesis on this project, finding significantly lower MuRF protein levels at post with glucose treatment compared to placebo.
-This may suggest an effect of glucose on protein degradation (WILKES09), however, whether this contributes to growth or simply protein turnover has been a topic of debate (FIGUEIREDO19).
-
-Despite being a thoroughly standardised design, a limitation may be posed through lack of participants.
-We initially reach our minimum limit for statistical power, i.e.
-16 participants, however, three dropped out during the intervention, possibly impacting our power.
-Coupled with the fact that no previous study could guide our power calculations, this, of course, introduces some degree of uncertainty regarding our results.
-However, we’ve reported little to suggest any true acute effect of RT with glucose ingestion compared to placebo.
-Moreover, the study was specifically designed to investigate differences in pre-post changes within participants, and thus also quite limited to this perspective.
-Hence, the effects of time observed exclusively serve as supplementary to the efficiency of the design, without a negative control group to either confirm or refute them.
-If further investigations are to be done on this topic, these analyses should include more participants and optimally a measure of intramuscular glycogen levels.
-
-In conclusion, resistance exercise training combined with glucose ingestion does not acutely augment ribosome biogenesis following 2 weeks of moderate-volume resistance training in moderately trained young male and female adults.
-Neither does it acutely affect muscular performance in isometric and isokinetic knee extension peak torque, nor enhance post resistance training recovery within 24hrs.
-
-## Disadvantages due to Covid-19
-
-Due to restrictions regarding social distancing and lock downs, caused by the Covid-19 pandemic, we were not able to recruit and include as many participants as we planned to the study.
+######**Figure 4:** Relationship between levels of total RNA and UBF.
+Total RNA was normalized by wet muscle weight, and UBF levels were normalized by pools and total protein per lane.
+Values are presented as log-transformed means.

### 2023-02-20 R/ubf_rna.correlation.R

@@ -17,10 +17,12 @@ rna.dat <- readRDS("./data/data-gen/rna/rna.dat.RDS") # Quantified and cleaned (
 ## Prepare data frames for joining
 
 ubf.rdy <- west.dat %>%  # One data frame without outliers
-  select(subject, time, outlier, supplement, target, norm.sign) %>%
+  select(subject, time, outlier, supplement, target, norm.sign, tpl) %>%
   filter(target == "ubf",
          outlier == "in") %>%
   select(-target) %>%
+  mutate(norm.sign = norm.sign / tpl) %>%
+  #print()
   group_by(subject, time, supplement) %>%
   summarise(mean.sign = mean(norm.sign, na.rm = TRUE)) %>%
   print()

### 2023-02-20 figures/figure1.R

@@ -93,6 +93,7 @@ glu.fig <- glu.change %>%
          time.c = as.numeric(time.c)) %>%
   ggplot(aes(time.c, emmean, group = supplement, fill = supplement)) +
   annotate("text", x = c(120, 135, 150), y = c(2.4, 2, 2.04), label = "*") +
+  annotate("text", x = c(270), y = c(0.23), label = "*") +
   geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
                 size = 0.5,
                 position = pos) +
@@ -115,8 +116,6 @@ glu.fig <- glu.change %>%
 
 
 ## Humac pre->post 5 sessions
-
-
 
 # Isometric
 
@@ -214,7 +213,7 @@ isom.acfig <- lemm.isomac %>%
   geom_line(position = pos) +
   geom_point(shape = 21, position = pos, size = 2) +
   scale_fill_manual(values = c("glucose" = "red", "placebo" = "royalblue")) +
-  scale_x_discrete(labels=c("change.1" = "Baseline", "change.2" = "30min", "change.3" = "2hrs",
+  scale_x_discrete(labels=c("change.1" = "Post session 5", "change.2" = "30min", "change.3" = "2hrs",
                             "change.4" = "23hrs")) +
   labs(x = "", y = "Isometric peak torque \n(nm fold change)\n", fill = "Supplement") +
   theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6),
@@ -239,7 +238,7 @@ acfig.60 <- lemm.60ac %>%
   geom_line(position = pos) +
   geom_point(shape = 21, position = pos, size = 2) +
   scale_fill_manual(values = c("glucose" = "red", "placebo" = "royalblue")) +
-  scale_x_discrete(labels=c("change.1" = "Baseline", "change.2" = "30min", "change.3" = "2hrs",
+  scale_x_discrete(labels=c("change.1" = "Post session 5", "change.2" = "30min", "change.3" = "2hrs",
                             "change.4" = "23hrs")) +
   labs(x = "Time", y = "Iso 60 d/s peak torque \n(nm fold change)\n", fill = "Supplement") +
   theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6),
@@ -265,7 +264,7 @@ acfig.240 <- lemm.240ac %>%
   geom_line(position = pos) +
   geom_point(shape = 21, position = pos, size = 2) +
[…]

### 2023-02-20 figures/figure3.R

@@ -57,7 +57,7 @@ ubf.plot <- ubf %>%
   add_row(supplement = c("PLACEBO", "GLUCOSE"), time = "pre", emmean = 0, lower.CL = 0, upper.CL = 0) %>%
   mutate(time = factor(time, levels = c("pre", "post"), labels = c("Baseline", "Post"))) %>%
   ggplot(aes(time, exp(emmean), group = supplement, fill = supplement)) +
-  #annotate("text", x = "Post", y = 2.1, label = "p = 0.585", size = 3) +
+  annotate("text", x = "Post", y = 3, label = "*", size = 3) +
   geom_errorbar(aes(ymin = exp(lower.CL), ymax = exp(upper.CL)),
                 width = 0.1,
                 position = position_dodge(width = 0.2)) +
@@ -77,7 +77,7 @@ rps6.plot <- rps6 %>%
   add_row(supplement = c("PLACEBO", "GLUCOSE"), time = "pre", emmean = 0, lower.CL = 0, upper.CL = 0) %>%
   mutate(time = factor(time, levels = c("pre", "post"), labels = c("Baseline", "Post"))) %>%
   ggplot(aes(time, exp(emmean), group = supplement, fill = supplement)) +
-  #annotate("text", x = "Post", y = 2.1, label = "p = 0.585", size = 3) +
+  annotate("text", x = "Post", y = 1.87, label = "*", size = 3) +
   geom_errorbar(aes(ymin = exp(lower.CL), ymax = exp(upper.CL)),
                 width = 0.1,
                 position = position_dodge(width = 0.2)) +

### 2023-02-20 figures/figure4.R

@@ -30,7 +30,7 @@ fig4 <- joined.dat %>%
   ggplot(aes(log(mean.sign), log(mean.rna), color = supplement)) + 
   geom_point() +
   geom_smooth(method = "lm") +
-  labs(x = "Log-UBF normalised by pool", y = "Log-total RNA per mg muscle tissue", 
+  labs(x = "Mean log-UBF \nper normalization factor\n", y = "Mean log-total RNA \nper mg muscle tissue\n", 
        color = "Supplement") +
   scale_color_manual(values = c("GLUCOSE" = "red", "PLACEBO" = "royalblue")) +
   theme(panel.grid.major = element_blank(),

### 2023-02-21 ribose-paper.RMD

@@
 
-Systematic resistance training (RT) increases muscle mass and strength due to repeated mechanical and metabolic stress.
-Unfortunately, responses to RT vary widely between individuals, with as many as 10-15% showing impaired muscle growth (ALVAREZ18;MANN14;THALACKER13).
+Responses to systematic resistance training (RT) vary widely between individuals, with as many as 10-15% showing impaired muscle growth in response to standardized training interventions (ALVAREZ18;MANN14;THALACKER13).
 Such inability to benefit from RT may be explained by genetics, epigenetics, and/or an unfavourable internal physiological milieu (FIGUEIREDO21;HAMMARSTROM20;THALACKER13).<!-- Check these refs, do they correspond to genet epigen and internal mileu?
@@
 Furthermore, energy availability is a decisive factor in *de novo* synthesis of ribosomes <!--Add ref that indicate high energy requirement for ribo biog possibly (VONWALDEN19;TANAKA18?
---> which in turn determines muscle growth by increasing the muscles translational capacity (FIGUEIREDO19;VONWALDEN19;TANAKA18;HAMMARSTROM20;STEC16).
-Therefore, investigating the effects of combined RT and glucose ingestion may provide valuable insight into the regulation of ribosome biogenesis.
+MOSS  --> which in turn determines muscle growth by increasing the muscles translational capacity (FIGUEIREDO19;VONWALDEN19;TANAKA18;HAMMARSTROM20;STEC16).
+Therefore, investigating the effects of combined RT and glucose ingestion may provide valuable insight into a potential additive effect of glucose and RT on ribosome biogenesis.
 
 Transcription of ribosomal RNA (rRNA) by Pol I is considered the limiting step in synthesizing new ribosomes [@Moss_1995].
-Multiple signaling pathways converge to regulate rRNA transcription, including the mammalian target of rapamycin (mTORC1) signal-transduction pathway and c-Myc (KUSNADI15;VONWALDEn19;TANAKA18).
+Multiple signaling pathways converge to regulate rRNA transcription, including the mammalian target of rapamycin (mTORC1) signal-transduction pathway and c-Myc (KUSNADI15;VONWALDEn19;TANAKA18 @Mori_2021; @West_2016).<!-- Is c_myc included in references here?
+could add West 2016 -->
 mTOR receives input from growth factors, hormones, mechanical loading, and nutrients to balance protein synthesis through multiple mechanisms based on cellular energy levels (HOPPE09).
 Importantly, mTOR and its downstream target S6K1 serve as a mediator from insulin signaling in skeletal muscle[@hillier.etal_2000].
-mTOR contributes to forming the preinitiation complex (PIC) that marks the initiation of rRNA transcription (VONWALDEN19;FIGUEIREDO19).
-Parallel to mTORC1 [@Mori_2021; West_2016], c-Myc increases ribosomal biogenesis directly and through transcriptional control of the upstream binding factor (UBF) [@sanijUBFLevelsDetermine2008; Poortinga_2011].
-However, interestingly, UBF phosphorylation, required for interaction with the rDNA promoter, is increased by high glucose in a mTORC1 dependent manner (rapamycin sensitive) in kidney glomerular epithelial cells[@Mariappan_2011].
-Additionally, independently from UBF, high glucose was shown to lead to chromatin remodeling, which promotes rRNA transcription[@Zhai_2012].
+The mTORC1 pathway contributes to ribosome biogenesis through the regulation of translation of ribosomal proteins and to forming of the preinitiation complex (PIC) that marks the initiation of rRNA transcription (VONWALDEN19;FIGUEIREDO19).
+Parallel to mTORC1 [@Mori_2021; @West_2016], c-Myc increases ribosomal biogenesis directly and through transcriptional control of the upstream binding factor (UBF) [@sanijUBFLevelsDetermine2008; Poortinga_2011].
+Additionally, UBF phosphorylation, required for interaction with the rDNA promoter, is increased by high glucose in a mTORC1 dependent manner (rapamycin sensitive) in kidney glomerular epithelial cells[@Mariappan_2011].
+Independently from UBF and mTORC1, high glucose was shown to lead to chromatin remodeling, which promotes rRNA transcription i cell cultures[@Zhai_2012].
 Together these observations indicate a potential role of glucose in positively affecting muscle rRNA synthesis.
 
-The purpose of this investigation was to determine the effect of 
-<!-- -->
-
-the relationship between glucose, RT and ribosome biogenesis we recruited 16 healthy young, moderately trained male and female individuals to moderate volume unilateral resistance training for two weeks.
-They exercised one leg with glucose and protein, and the other leg with placebo and protein ingestion before and after each RT-session.
-Our main purpose was to investigate whether there are acute effects of glucose- compared to placebo ingestion before and after RT.
-To assess this purpose, we measured changes in ribosome biogenesis as accumulation of total RNA, expression of the four mature rRNA's (18S, 5.8S, 28S, 5S), 47S pre-rRNA, and levels of proteins associated with transcription initiation (c-Myc, UBF, rpS6?).
-As secondary outcomes, we measured the development in skeletal muscle performance during the intervention, and skeletal muscle recovery after the final RT session, as temporal changes in maximal isometric and isokinetic torque.
-We hypothesised that RT would indeed lead to a prominent accumulation of total RNA, increased expression of rRNA and associated proteins.
-Further, we hypothesised that RT with glucose and protein would induce an enhanced response compared to placebo, in markers of ribosome biogenesis, as well as muscular performance and -recovery.
+The purpose of this investigation was to test the hypothesis that glucose supplementation in connection to<!-- during?
+--> RT would increase markers of ribosomal abundance following five RT sessions.
+Secondly we aimed to describe the association between UBF and total RNA abundance in human skeletal muscle.
+<!-- We should maybe add a line or two about the rational here -->
 
@@
 
-Sixteen healthy male and female participants (20-33yrs, Tab 1) were recruited to the study through social media advertisement and word of mouth, and through completing the selection process (Fig 1).
+Sixteen healthy male and female participants (20-33yrs, Table 1) were recruited to the study through social media advertisement and word of mouth, and through completing the selection process (Fig 1).
 The eligibility criteria were non-smokers and moderately trained (i.e.

### 2023-02-21 ribose-paper.RMD
figure updates and strenght.index

@@
 Responses to systematic resistance training (RT) vary widely between individuals, with as many as 10-15% showing impaired muscle growth in response to standardized training interventions (ALVAREZ18;MANN14;THALACKER13).
-Such inability to benefit from RT may be explained by genetics, epigenetics, and/or an unfavourable internal physiological milieu (FIGUEIREDO21;HAMMARSTROM20;THALACKER13).<!-- Check these refs, do they correspond to genet epigen and internal mileu?
--->
-Although RT program modifications such as altered exercise volume or intensity may induce more favorable adaptations in some individuals, such benefits are not uniform between individuals [@hammarstrom.etal_2020].<!-- Add ref for more rt program modifications -->
+Such inability to benefit from RT may be explained by genetic and/or epigenetic disposition, affecting the internal physiological milieu in terms of ability to mount an anabolic response to RT [@figueiredo21;@thalacker13].<!-- Check these refs, do they correspond to genet epigen and internal mileu?
+checked-->
+Although RT program modifications such as altered exercise volume or intensity may induce more favorable adaptations in some individuals, such benefits are not uniform between individuals [@hammarstrom.etal_2020;@krieger09;@schoenfeld17].<!-- Add ref for more rt program modifications.
+checked -->
 For example, although changes in training volume generally induces greater benefits of RT, not all individuals benefit from such modifications [@hammarstrom.etal_2020].
@@
 
-Nutritional supplements such as protein and creatine effectively optimize RT adaptations (CERMAK12;LANHERS15;LANHERS17).
+Nutritional supplements such as protein and creatine effectively optimize RT adaptations [@cermak12;@lanhers15;@lanhers17;@morton18].
 However, it remains unknown if other nutritional adjuvants such as glucose can increase the efficacy of RT.
-This is surprising, as glucose is the preferred energy substrate of the contracting skeletal muscle during strenuous exercise and a major supplier of energy to cells via ATP synthesis (MUL15;TANAKA18).
-Furthermore, energy availability is a decisive factor in *de novo* synthesis of ribosomes <!--Add ref that indicate high energy requirement for ribo biog possibly (VONWALDEN19;TANAKA18?
-MOSS  --> which in turn determines muscle growth by increasing the muscles translational capacity (FIGUEIREDO19;VONWALDEN19;TANAKA18;HAMMARSTROM20;STEC16).
+This is surprising, as glucose is the preferred energy substrate of the contracting skeletal muscle during strenuous exercise and a major supplier of energy to cells via ATP synthesis [@mul15;@tanaka18].
+Furthermore, energy availability is a decisive factor in *de novo* synthesis of ribosomes [@kusnadi15;@moss07;@tanaka18] <!--Add ref that indicate high energy requirement for ribo biog possibly (VONWALDEN19;TANAKA18?
+MOSS.
+checked  --> which in turn determines muscle growth by increasing the muscles translational capacity [@figueiredo19;@hammarstrom20;@tanaka18;@stec16;@vonwalden19].
 Therefore, investigating the effects of combined RT and glucose ingestion may provide valuable insight into a potential additive effect of glucose and RT on ribosome biogenesis.
@@
 Transcription of ribosomal RNA (rRNA) by Pol I is considered the limiting step in synthesizing new ribosomes [@Moss_1995].
-Multiple signaling pathways converge to regulate rRNA transcription, including the mammalian target of rapamycin (mTORC1) signal-transduction pathway and c-Myc (KUSNADI15;VONWALDEn19;TANAKA18 @Mori_2021; @West_2016).<!-- Is c_myc included in references here?
-could add West 2016 -->
-mTOR receives input from growth factors, hormones, mechanical loading, and nutrients to balance protein synthesis through multiple mechanisms based on cellular energy levels (HOPPE09).
+Multiple signaling pathways converge to regulate rRNA transcription, including the mammalian target of rapamycin (mTORC1) signal-transduction pathway and c-Myc [@kusnadi15;@vonwalden19;@tanaka18].<!-- Is c_myc included in references here?
+could add West 2016.
+c-Myc included in Von Walden and Tanaka, adding West 2016 and Mori 21 when I find them-->
+mTOR receives input from growth factors, hormones, mechanical loading, and nutrients to balance protein synthesis through multiple mechanisms based on cellular energy levels [@hoppe09].
 Importantly, mTOR and its downstream target S6K1 serve as a mediator from insulin signaling in skeletal muscle[@hillier.etal_2000].
-The mTORC1 pathway contributes to ribosome biogenesis through the regulation of translation of ribosomal proteins and to forming of the preinitiation complex (PIC) that marks the initiation of rRNA transcription (VONWALDEN19;FIGUEIREDO19).
-Parallel to mTORC1 [@Mori_2021; @West_2016], c-Myc increases ribosomal biogenesis directly and through transcriptional control of the upstream binding factor (UBF) [@sanijUBFLevelsDetermine2008; Poortinga_2011].
+The mTORC1 pathway contributes to ribosome biogenesis through the regulation of translation of ribosomal proteins and to forming of the preinitiation complex (PIC) that marks the initiation of rRNA transcription [@figueiredo19;@vonwalden19].
+Parallel to mTORC1 [@Mori_2021; @West_2016], c-Myc increases ribosomal biogenesis directly and through transcriptional control of the upstream binding factor (UBF) [@sanijUBFLevelsDetermine2008; @Poortinga_2011].
 Additionally, UBF phosphorylation, required for interaction with the rDNA promoter, is increased by high glucose in a mTORC1 dependent manner (rapamycin sensitive) in kidney glomerular epithelial cells[@Mariappan_2011].
-Independently from UBF and mTORC1, high glucose was shown to lead to chromatin remodeling, which promotes rRNA transcription i cell cultures[@Zhai_2012].
+Independently from UBF and mTORC1, high glucose was shown to lead to chromatin remodeling, which promotes rRNA transcription in cell cultures[@Zhai_2012].
 Together these observations indicate a potential role of glucose in positively affecting muscle rRNA synthesis.

### 2023-02-21 figures/figure2.R

@@ -12,10 +12,14 @@ emm.tot <- readRDS("./data/data-gen/rna/emm.tot.RDS")
 
 ## rRNA
 # 18S
-emm.18 <- readRDS("./data/data-gen/rna/emm.18.RDS")
+emm.18 <- readRDS("./data/data-gen/rna/emm.18.RDS") %>%
+  mutate(target = "18S rRNA") %>%
+  print()
 
 # 28S
-emm.28 <- readRDS("./data/data-gen/rna/emm.28.RDS")
+emm.28 <- readRDS("./data/data-gen/rna/emm.28.RDS") %>%
+  mutate(target = "28S rRNA") %>%
+  print()
 
 # 5.8S
 emm.5.8 <- readRDS("./data/data-gen/rna/emm.5.8.RDS")
@@ -167,7 +171,11 @@ plot.5 <- emm.5 %>%
         legend.title = element_text(size = htextsize),
         legend.text = element_text(size = 5)) + plot_theme
 
-
+bind_rows(emm.18, emm.28) %>%
+  ggplot(aes(supplement, emmean)) +
+  geom_point() +
+  facet_wrap(~target, nrow = 1) +
+  theme(strip.text = element_text(hjust = 0))
 
 
 # Cowplot for gathering figures

### 2023-02-24 results_work.RMD
small fig updates

@@
 
-**Figure 1:** A) Overview of the experimental design with 12 days of concomitant dietary intervention and resistance training (RT), preceded by 7 days of testing.
+**Figure 1:** A) An overview of the experimental design with 12 days of concomitant dietary intervention and resistance training (RT), preceded by 7 days of testing.
 Familiarization = days -7 to -1 included familiarization to RT exercises (via 1RM leg press and knee extension tests) and strength tests (Humac dynamometer) at days -7 and -5.
@@
 Levels of c-peptide were measured via venous blood samples collected at the same time points as finger draws, except for 90min before RT and during RT.
-D, E, F) Changes in muscular performance (D: nm fold change in isometric peak torque, E/F: nm fold change in isokinetic peak torque at 60/240 d/s.).
-Muscular performance was measured via a Humac Norm Dynamometer at baseline (Day -1), during (Day 4 and 8 leg 1, Day 5 and 9 leg 2), and after five RT sessions (Day 11 leg 1, Day 12 leg 2).
-G, H, I) Changes in muscular recovery (G: nm fold change in isometric peak torque, H/I: nm fold change in isokinetic peak torque at 60/240 d/s).
-Muscular recovery was measured via Humac Norm Dynamometer with baseline being post fifth RT session test, and then 30min, 2hrs and 23hrs following the sixth and final RT session.
+D) Changes in muscular performance/recovery measured as isometric and isokinetic peak torque (60 and 240 d/s) via Humac Norm Dynamometer, conducted at baseline (Day -1), during the intervention (Day 4 and 8 leg 1, Day 5 and 9 leg 2), after five RT sessions/before the 6th session (Day 11 leg 1, Day 12 leg 2), as well as 30min, 2hrs and 23hrs after the 6th RT session (Day 11/12 leg 1, Day 12/13 leg 2).
 Values are presented as estimated marginal means ± 95% CI.
@@
 ###### **Figure 2:** Changes in total RNA and ribosomal RNA.
-A) Total RNA, B) 47S pre-rRNA, C) 18S rRNA, D) 28S rRNA, E) 5.8S rRNA, F) 5S rRNA.
+A) Total RNA, B) 47S pre-rRNA, 18S rRNA, 28S rRNA, 5.8S rRNA, 5S rRNA.
 Baseline = Training day 1/2, Post = Training day 11/12.
@@
 ###### **Figure 3:** Changes in levels of protein.
-A) rPS6, B) c-Myc, C) UBF.
+A) Representative total protein stain blot, B) AU fold change in UBF, c-Myc and rpS6 + representative antibody stain images of the respective proteins.
 Baseline = Training day 1/2, Post = Training day 11/12.

### 2023-02-24 ribose-paper.RMD
updates

@@
 
-The purpose of this investigation was to test the hypothesis that glucose supplementation in connection to<!-- during?
---> RT would increase markers of ribosomal abundance following five RT sessions.
+The purpose of this investigation was to test the hypothesis that glucose supplementation before and after<!-- during?
+before and after would be more precise --> RT would increase markers of ribosomal abundance following five RT sessions.
 Secondly we aimed to describe the association between UBF and total RNA abundance in human skeletal muscle.

### 2023-02-24 disp.Rmd
updates

--- 2023-02-21 ribose-paper.RMD
@@
 ---
-title: "Glucose ingestion before and after resistance training does not augment ribosome biogenesis in healthy moderately trained young adults"
-author: "Kristian Lian, Daniel Hammarström, Stian Ellefsen, Håvard Hamarsland"
-bibliography: resources/ribose.bib
-csl: resources/jap.csl
-output:
-  word_document:
-    reference_docx: reference-template-02.docx
-editor_options: 
-  chunk_output_type: console
+title: "Disp"
+output: word_document
+date: "2023-02-24"
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
+knitr::opts_chunk$set(echo = TRUE)
 ```
 
-```{r packages}
-library(tidyverse);library(readxl);library(lme4);library(emmeans);library(dplyr);library(cowplot);library(knitr);library(broom);library(rmarkdown);library(tinytex);library(tidyr);library(arsenal);library(magrittr);library(dabestr);library(nlme); library(pbkrtest)
-```
-
-## Introduction
-
-Responses to systematic resistance training (RT) vary widely between individuals, with as many as 10-15% showing impaired muscle growth in response to standardized training interventions (ALVAREZ18;MANN14;THALACKER13).
-Such inability to benefit from RT may be explained by genetics, epigenetics, and/or an unfavourable internal physiological milieu (FIGUEIREDO21;HAMMARSTROM20;THALACKER13).<!-- Check these refs, do they correspond to genet epigen and internal mileu?
--->
-Although RT program modifications such as altered exercise volume or intensity may induce more favorable adaptations in some individuals, such benefits are not uniform between individuals [@hammarstrom.etal_2020].<!-- Add ref for more rt program modifications -->
-For example, although changes in training volume generally induces greater benefits of RT, not all individuals benefit from such modifications [@hammarstrom.etal_2020].
-Therefore, other means than modification of RT variables seems necessary to optimize individual responses.<!-- The above sentence might be a bit contradictory?
--->
-
-Nutritional supplements such as protein and creatine effectively optimize RT adaptations (CERMAK12;LANHERS15;LANHERS17).
-However, it remains unknown if other nutritional adjuvants such as glucose can increase the efficacy of RT.
-This is surprising, as glucose is the preferred energy substrate of the contracting skeletal muscle during strenuous exercise and a major supplier of energy to cells via ATP synthesis (MUL15;TANAKA18).
-Furthermore, energy availability is a decisive factor in *de novo* synthesis of ribosomes <!--Add ref that indicate high energy requirement for ribo biog possibly (VONWALDEN19;TANAKA18?
-MOSS  --> which in turn determines muscle growth by increasing the muscles translational capacity (FIGUEIREDO19;VONWALDEN19;TANAKA18;HAMMARSTROM20;STEC16).
-Therefore, investigating the effects of combined RT and glucose ingestion may provide valuable insight into a potential additive effect of glucose and RT on ribosome biogenesis.
-
-Transcription of ribosomal RNA (rRNA) by Pol I is considered the limiting step in synthesizing new ribosomes [@Moss_1995].
-Multiple signaling pathways converge to regulate rRNA transcription, including the mammalian target of rapamycin (mTORC1) signal-transduction pathway and c-Myc (KUSNADI15;VONWALDEn19;TANAKA18 @Mori_2021; @West_2016).<!-- Is c_myc included in references here?
-could add West 2016 -->
-mTOR receives input from growth factors, hormones, mechanical loading, and nutrients to balance protein synthesis through multiple mechanisms based on cellular energy levels (HOPPE09).
-Importantly, mTOR and its downstream target S6K1 serve as a mediator from insulin signaling in skeletal muscle[@hillier.etal_2000].
-The mTORC1 pathway contributes to ribosome biogenesis through the regulation of translation of ribosomal proteins and to forming of the preinitiation complex (PIC) that marks the initiation of rRNA transcription (VONWALDEN19;FIGUEIREDO19).
-Parallel to mTORC1 [@Mori_2021; @West_2016], c-Myc increases ribosomal biogenesis directly and through transcriptional control of the upstream binding factor (UBF) [@sanijUBFLevelsDetermine2008; Poortinga_2011].
-Additionally, UBF phosphorylation, required for interaction with the rDNA promoter, is increased by high glucose in a mTORC1 dependent manner (rapamycin sensitive) in kidney glomerular epithelial cells[@Mariappan_2011].
-Independently from UBF and mTORC1, high glucose was shown to lead to chromatin remodeling, which promotes rRNA transcription i cell cultures[@Zhai_2012].
-Together these observations indicate a potential role of glucose in positively affecting muscle rRNA synthesis.
-
-The purpose of this investigation was to test the hypothesis that glucose supplementation in connection to<!-- during?
---> RT would increase markers of ribosomal abundance following five RT sessions.
-Secondly we aimed to describe the association between UBF and total RNA abundance in human skeletal muscle.
-<!-- We should maybe add a line or two about the rational here -->
+## Intro
 
@@
 
-All participants were informed about the potential discomforts and risks associated with the study and gave their informed consent before study enrollment.
-The study was approved by the regional ethical committee (REK, ID nr.
-153628), pre-registered at clinicaltrials.gov (Identifier: NCT04545190), and conducted according to the Helsinki declaration.
-
 ### Participants
-
-Sixteen healthy male and female participants (20-33yrs, Table 1) were recruited to the study through social media advertisement and word of mouth, and through completing the selection process (Fig 1).
-The eligibility criteria were non-smokers and moderately trained (i.e.
-2-8 RT sessions per 14 days for the last six months).
-Exclusion criteria were previous injury leading to impaired strength, inability to perform resistance training, symptoms and a medical record of metabolic disorders including hyperglycemia, i.e.
-fasting venous plasma glucose ≥6.1 mmol/L and/or 2-hour glucose tolerance ≥7.8 mmol/L, and/or HbA1c >42 mmol/mol.
-Of the sixteen participants that commenced the intervention, three participants dropped out.
-One due to sickness and inability to resume, two participants experienced muscular discomfort connected to heavy resistance training (Fig 1).
-Baseline characteristics (Tab 1) were measured using DXA (Prodigy Advance PA+302047, Lunar, San Francisco, CA, USA) on Day -1, the last day preceding the RT intervention.
 
@@
 
-The study was designed as a 12-day double-blinded randomized controlled trial, with a unilateral RT protocol (Fig 2).
-Participants were randomized to commence with either glucose (GLU, 30g per bolus) or placebo (PLAC, 0g per bolus), and either dominant or non-dominant leg (Fig 1), alternating RT and supplement from one day to the other (Fig 2).
-Investigators were blinded as the supplement randomization was disclosed only to personnel not involved in any other aspect of the study, and participants were blinded by diluting boluses of glucose (30g glucose) (Glucosum monohydricum, Merck KGaA, Darmstadt, Germany) and placebo (~0.3g Stevia rebaudiana extract) (Steviosa, Soma Nordic AS, Oslo, Norway) in 300ml Fun Light juice/saft (Orkla, Oslo, Norway).
-A blinded taste test revealed that the participants were not able to disclose the contents of the provided boluses (30g GLU vs.
-0g GLU).
-To ensure equal conditions during training sessions, participants exercised and tested at the same time of day, +/- 1hr with the same trainer, and recorded and repeated their daily macro nutrient intake on pairwise consecutive days (Tab 4).
-
-###### **Table 1:** Baseline characteristics of the participants. BMI: body mass index, A = charactaristics per sex, B = characteristics per leg. Values are mean ± standard deviation.
-
-```{r DXA}
-# Script: DXA
-
-# Data handling
-table.dxa <- read_excel("./data/dxa/ribose_dxa.xlsx") %>%
-        select(sex, subject, age, height, weight, leanmasskg, fatmasskg, ffmg) %>%
-        #make grams into kg
-        mutate(ffkg = ffmg/1000) %>%
-        select(sex, age, height, weight, leanmasskg, fatmasskg, ffkg) %>%
-  mutate(bmi = (weight) / (height^2) * 10000)
-
-dxa.table <- table.dxa %>%
-        pivot_longer(names_to = "variable",
-                     values_to = "values", col= age:bmi) %>%
-        select(sex, variable, values) %>%
-        group_by(sex, variable) %>%
-        summarise(m = mean(values, na.rm = TRUE),
-                  s= sd(values, na.rm = TRUE)) %>%
-        mutate(stat = paste0(round(m, 1), " (", round(s, 1), ")")) %>%
-        select(sex, variable, stat) %>%
-        pivot_wider(names_from = sex,
-                    values_from = stat) %>%
-                mutate(variable = if_else(variable %in% c("age"),
-                                          "Age (yrs)",
-                                          if_else(variable %in% c("fatmasskg"),
-                                                  "Fat mass (kg)",
-                                                  if_else(variable %in% c("ffkg"),
-                                                          "Fat free mass (kg)",
-                                                          if_else(variable %in% c("height"),
-                                                                  "Height (cm)",
-                                                                  if_else(variable %in% c("leanmasskg"),
-                                                                          "Lean body mass (kg)",
-                                                                          if_else(variable %in% c("weight"),
-                                                                                  "Body weight (kg)",
-                                                                                  if_else(variable %in% c("bmi"),
-                                                                                          "BMI", "")
-                                                                                  )))))))
-
-dxa.table %>%
-        kable(col.names = c("Variable", "Female (n=7)", "Male (n=9)"))
-
-```
-
-```{r dxa per leg}
-# Script: DXA
-
-# Subject 104, 110 and 111 dropped out and is therefore excluded from the table
-
-#Load dxa data
-dxaleg <-  read_excel("./data/dxa/ribose_dxaprleg.xlsx") %>%
-        select(subject, sex, leg, leanmassg, fatmassg, totalmasskg) %>%
-        #make grams into kg
-        mutate(leanmass = leanmassg/1000,
-               fatmass = fatmassg/1000)%>%
-        select(subject, sex, leg, leanmass, fatmass, totalmasskg)
-
-# Between groups analysis 
-dxa.stat <- dxaleg %>%
-        select(subject, leg, leanmass, fatmass, totalmasskg)
-
-# Lean mass between legs, within subject 
-leanmass.lm <- lmer(leanmass ~ leg + leg:subject + (1|subject), data = dxa.stat)
-lm.plot <- plot(leanmass.lm)
-lm.sum <- summary(leanmass.lm)
-lm.confint <- emmeans(leanmass.lm, specs = ~ "leg")
-
-# Fat mass between legs, within subject
-fat.lm <- lmer(fatmass ~ leg + leg:subject +(1|subject), data = dxa.stat)
-fat.plot <- plot(fat.lm)
-fat.sum <- summary(fat.lm)
-fat.confint <- emmeans(fat.lm, specs = ~ "leg")
-
-# Total mass between legs, within subject
-total.lm <- lmer(totalmasskg ~leg + leg:subject + (1|subject), data = dxa.stat)
-totlm.plot <- plot(total.lm)
-totlm.sum <- summary(total.lm)
-supsup <- emmeans(total.lm, specs = ~ "leg")
-
-# Table of leanmass, fatmass and total mass in kg pr. leg
-dxalegtable <- dxaleg %>%
-        select(leg, leanmass, fatmass, totalmasskg) %>%
-        pivot_longer(names_to = "variable",
-                     values_to = "values", col= leanmass:totalmasskg) %>%
-        group_by(leg, variable) %>%
-        summarise(m = mean(values, na.rm = TRUE),
-                  s= sd(values, na.rm = TRUE)) %>%
-        mutate(stat = paste0(round(m, 1), " (", round(s, 1), ")")) %>%
-        select(leg, variable, stat) %>%
-        pivot_wider(names_from = leg,
-                    values_from = stat)
-
-# Markdown table
-dxalegtable %>%
-        kable(col.names = c("Variable", "Left leg", "Right leg"))
-
-```
-
-Six RT sessions were conducted with glucose and six with placebo, allowing a within-subjects analysis of the effects of glucose ingestion.
-Data from the first five RT sessions was used to investigate main outcome measures (total RNA, rRNA and protein), whereas data from the sixth RT session was used to explore secondary outcomes (muscular recovery, plasma glucose levels).
-Participants were asked to refrain from resistance- or high-intensity training of the legs from Day -7 and onwards, until completion of the intervention and post-testing, to ensure the reliability of pre-intervention strength data and minimal interference from other training sources.
-If participants did conduct exercise training outside of the prescribed protocol, they were asked to ensure equal loading on pairwise consecutive days.
-
-![](figures/flowchart_selection.PNG)
-
-###### **Figure 1:** Flowchart of the selection process. The order in which participants performed the two intervention blocks was determined in a planned randomised fashion.
-
-![](figures/timeline.PNG)
-
-###### **Figure 2.** Overview of the intervention, with 12 days of concomitant dietary intervention and resistance training (RT), preceded by 7 days of testing.
-T0 = timepoint 0 (testing days prior to intervention), T1/T2 = biopsy prior to resistance training for leg #1 and #2, respectively, T3/T4 post intervention testing of leg #1 and #2, respectively, including post sixth RT session.
-Test = physical performance test (unilateral 1RM leg press, knee extension, isometric knee extension force, isokinetic knee extension torque) testing on days 4, 5, 8, 9, T3, T4 will not include 1RM leg press and knee extension, Biopsy = microbiopsy, RT#1 = resistance training leg #1, RT#2 = resistance training leg#2.
-Other tests (D2O, DXA, spit) will be conducted as part of the study as a whole.
-These are not relevant to this data set, thus will not be described here.
-
 ### Dietary intervention
-
-The dietary intervention spanned the whole day, divided into three periods: I) From awakening until two hours after RT, II) from 2.5 hours after RT until 2200hrs, and III) from 2200hrs until awakening.
-During period I, participants ingested protein and glucose/placebo only.
-Glucose/placebo was ingested at three time points in period I: 30 minutes before RT (0830hrs, 30g vs 0g glucose), immediately before RT (0900hrs, 30g vs.
-0g glucose), and immediately after RT (~0930hrs, 30g vs.
-0g glucose).
-Protein was ingested 2hrs before RT (0700hrs, 25g) and immediately after RT (~0930hrs, 25g).
-In the afternoon (1800hrs-1900hrs, period II) participants ingested glucose or placebo (3x30g vs.
-3x0g glucose) opposite to the supplement they received during RT, to ensure a balanced daily intake of glucose.
-Apart from this, participants ingested a self-chosen diet during period II, registered in MyFitnessPal or similar applications.
-The self-chosen diet was repeated on pairwise consecutive days (i.e.
-on days 1-2, 3-4, etc.), to ensure similar premises for resistance training responses between the two legs.
-During period III (2200-0700hrs), participants remained in an overnight fasted state.
-The daily onset of the dietary intervention (i.e., first ingestion of PRO supplement) varied between participants, from 0600hrs to 0900hrs to allow multiple participants to complete the protocol simultaneously.
-The protein supplement was ingested as 25g Whey Protein Isolate boluses (Proteinfabrikken, Stokke, Norway), diluted in 150ml water.
-During sessions, participants were free to ingest water ad libitum.
 
@@
 
-Performance tests (Test, Fig 2) were performed prior to (Days -7 and -5, and T0, both legs) and during the intervention (Days 4, 5, 8, and 9), after session 5 and after finalization of the intervention (T3/T4, RT#1 leg; T4/Day 13, RT#2 leg), a total of 10 days (Fig 2).
-Maximal isometric and isokinetic knee extension torque were measured with a Humac Norm Dynamometer (CSMi, Stoughton, Massechusetts, USA).
-Individual positions were recorded and standardized from pre-intervention tests (days -7 and -5).
-Maximal isokinetic torque was measured at speeds of 60- and 240 d/s, 2x3 repetitions each, with the first set of each exercise as a sub-maximal warm-up.
-The maximal isometric torque was measured at knee-angle 60, for a maximum of 10 seconds and two repetition per test.
-The highest peak torque values were carried forwards to the final analyses.
-During days 4, 5, 8 and 9, humac tests were conducted one hour before RT with the leg performing RT the previous day.
-During days 11 and 12 (T3 and T4), humac tests were performed at four time points: I) 45min before RT, II) 30min after the last RT session, III) 2hrs after the last RT session, and IV) 23hrs after the last RT session.
-Test I at T4 included testing of both legs, representing 23hrs post RT session test of one leg and pre-session 6 test of the other leg.
-
-Assessment of unilateral one repetition maximum (1RM) leg press and knee extension was conducted at pre-intervention testing (days -7 and -5, Fig 2).
-The participants performed a general warm-up with 10min cycling on an indoor exercise bicycle and a specific warm-up (1x10, 1x6 and 1x3 repetitions, 2min breaks) before each of the tests.
-Between specific warm-up sets as well as between each subsequent 1RM attempt, the participants were given 3min breaks.
-All positions were controlled and recorded at pre-intervention testing and repeated for the RT sessions.
-Maximal leg press strength was defined as the maximal load lifted in a controlled fashion, with a knee angle of 45 degrees (start), within a 2.5kg range.
-Attempts, where participants did not reach 45 degrees during the eccentric phase, were not approved.
-Maximal knee extension testing followed the same specific warm-up and pause protocol and was defined as maximal load lifted in a controlled fashion, reaching full extension of the knee joint, within a 1.25kg range.
-Attempts with exaggerated hip movement or beneath full extension were not approved.
-
 ### Resistance training protocol
-
-Resistance training consisted of three sets of each unilateral leg presses and knee extensions, with an intensity of 10 repetition maximum (10RM).
-As a general warm-up, the participants cycled on an indoor exercise bicycle for 5-10min.
-In addition, before the respective exercises, two 10-repetition warm-up sets were completed at ~50% and ~70% of 10RM.
-Rest time between working sets was 2min, and progressive loading was used to ensure adequate exercise stimulation throughout the intervention.
-For safety and standardization purposes, all sessions were monitored by trained personnel.
-To further ensure equal exercise premises for the two legs, each participant was accompanied by the same trainer in each session.
-If this was not possible throughout the intervention, each participant had the same trainer on pairwise consecutive days as a minimum.
-The rate of perceived exertion (RPE, 0–10-point scale) was logged before every session, and the session score was logged 15min after each session.
-Lastly, training volume (i.e., load and repetitions) was logged for every session.
 
@@
 
-Muscle biopsies were sampled from m.
-vastus lateralis using well-established procedures (HAMMARSTRØM20).
-Muscle biopsy sampling was performed under local anaesthesia (Xylocaine, 10mg ml-1 with adrenaline 5μg ml−1, AstraZeneca AS, Oslo, Norway) using a 12-14-gauge needle (Universal Plus, Mermaid Medical AS, Stenløse, Denmark), operated with a spring-loaded biopsy gun.
-After the biopsy sampling, muscle tissue was divided in two aliquots for determination of total RNA/expression of rRNA and two aliquots for protein analyses, snap frozen in isopentane (-80°C) and stored at -80°C until further analyses.
-Muscle biopsies were sampled at four time points (Fig 2): I/II) Pre-intervention (2hrs before training, time point T1 = pre RT#1 leg, T2 = pre RT#2 leg), and III/IV) before the sixth RT session from RT#1 leg (2hrs before training, T3) and before the sixth RT session from RT#2 leg (2hrs before training, T4).
-At each time point, two samples were taken from the same incision.
-To standardize this procedure, all individual participants had biopsies taken at the same time of day, in an overnight fasted state.
-
-To measure blood glucose levels with and without glucose ingestion/training, blood was collected by means of finger draws on days with biopsy sampling.
-One finger draw was taken at T1 (pre-test, Fig 2) to serve as a baseline.
-At T3 and T4 (post-test, Fig 2), finger draws were collected at 7 time points: I) prior to protein ingestion (0700hrs) II) 45 minutes after protein ingestion (0745hrs) III) 1.5hrs after protein ingestion (0800hrs, i.e., immediately before GLU/PLAC intake), IIII) 2hrs after protein ingestion (0900hrs, i.e., immediately before training), IV) in the middle of RT (~0915hrs), V) immediately after training (~0930hrs), and VI) 2hrs after completion of training (~1130hrs).
-Finger draws were analysed with in-house equipment (BIOSEN C-Line, EKF diagnostic GmbH, Barleben).
-Venous blood samples were collected from the antecubital vein, coinciding with the finger draws except in the middle of the RT session, to analyse endocrine variables.
-
 ### Total RNA extraction and quantitative real-time reverse transcription polymerase chain reaction
-
-Two muscle biopsy samples were taken for total RNA extraction per leg per time point, and total RNA was extracted in two duplicates per muscle biopsy.
-Thus, we had two duplicates per leg per time point, giving a total of eight RNA samples per participant.
-Total RNA was extracted using TRIzol according to the manufacturer’s protocol.
-An exogenous RNA control (λ polyA External Standard Kit, Takara Bio Inc., Shiga, Japan) was added at a fixed amount (0.04 ng ml-1 of Trizol reagent) to enable analysis of target gene expression per unit tissue weight (ELLEFSEN2008,2014).
-Briefly, muscle tissue was homogenized in TRIzol using 0.5mm RNase-free Zirconium beads (~50 ul) (Next Advanced, Averill Park, NY, USA) and a bullet blender (bb).
-Chloroform (Sigma-Aldrich, Missouri, USA) was used for phase-separation, and an RNA pellet was precipitated with isopropanol (VWR International, Pennsylvania, USA).
-For assessment of RNA content and -purity 5µl RNA, from the RNA stock, was eluted in TE-buffer (1:2), for assessment via spectrophotometry.
-All samples had a 260nm to 280nm ratio > 1.9.
-The RNA stock was stored at -80°C until further analyses.
-RNA was reverse transcribed using Super Script IV Reverse Transcriptase (Invitrogen), according to the manufacturer’s instructions.
-Briefly, 1µl 10 mM dNTP mix, 0.5µl anchored oligo-dT, 0.5µl random hexamer primers (Thermo Scientific), a maximum of 9µl template RNA and 2 µl nuclease-free water were mixed, vortexed and briefly spun down.
-Samples were then heated at 65°C for 5 minutes, followed by at least 1-minute incubation on ice.
-For the next step, 4µl 5x SSIV Buffer, 1µl 100 mM DTT, 1 µl RNase OUT, and Super Script IV Reverse Transcriptase (Invitrogen) were mixed, vortexed and briefly spun down and added to the samples.
-Samples were incubated for 10 minutes at 23°C, 10 minutes at 50-55°C and 10 minutes at 80°C.
-All samples were reverse transcribed and diluted to 1:50 before quantitative real-time polymerase chain reaction (qPCR).
-Lambda was used as an external reference gene, added in the RNA extraction (2µl per extraction).
-qPCR reactions were run on a fast-cycling real-time detection system (Applied Biosystems 7500 fast Real-Time PCR Systems, Life Technologies AS), with a total volume of 10 µl, containing 2 µl of cDNA, gene-specific primers (0.5 µM final concentration) and a commercial master mix (2X SYBR Select Master Mix, Applied Biosystems, Life Technologies AS) (Hammarström et al.
-2020).
-qPCR reactions consisted of 40 cycles (3 s 95°C denaturing and 30 s 60°C annealing) (Hammarström et al.
-2020).
-Raw fluorescence data were exported from the platform-specific software and amplification curves were modelled with a best-fit sigmoidal model using the qPCR package (Ritz & Spiess, 2008) written for R (Hammarström et al.
-2020; Team 2018).
-For an overview of primers, see Table 2.
-
-###### **Table 2:** Primer sequence and performance. Average cycle thresholds (Ct) and priming efficiencies were calculated from all qPCR reactions.
-
-![](figures/Primer.sequences.PNG)
 
@@
 
-Total protein was extracted using the Minute Total Protein Extraction Kit for Muscles (Invent Biotechnology), according to the manufacturer’s protocol, optimised for our lab.
-Wet muscle was freeze-dried for 24hrs and dissected before extraction.
-The tissue was homogenized with a plastic rod in 80mg protein extraction powder (Invent Biotechnology) and 100ul iced cell lysis buffer (Denaturing Buffer, Invent Biotechnology), and centrifuged at top speed for 1 min.
-The supernatant was aliquoted into 2x25ul tubes to run samples in duplicates, and total protein concentrations were determined in a 1:10 dilution (Pierce Detergent Compatible Bradford Assay Reagent, Thermo Fisher Scientific, Oslo, Norway).
-The protein samples were diluted to 20.25 µg µl−1 total protein with lysis buffer and 4X Laemmli sample buffer (Bio-Rad Laboratories AB, Oslo, Norway) containing 2-mercaptoethanol.
-Twenty microlitres from each protein sample were pooled as a normalisation factor for gel separation.
-All protein samples were incubated at 95°C and stored at -20°C until further analysis.
-The protein samples (20.25µg total protein) were separated at 250V for 50min using 4-20% Tris-Glycine gels (Thermo Fisher Scientific), then transferred to PVDF membranes with wet transfer at 300mA for 3hrs.
-Both gel electrophoresis and protein transfer were performed at 4°C.
-Following the wet transfer, membranes were stained with a MemCode reversible total protein stain (MANUFACTURER) and then blocked for 1hr at room temp with a blocking buffer of Tris-buffered saline (TBS; 20 mM Tris, 150 mM NaCl) with 5% non-fat dry milk and 0.1% Tween-20.
-Primary and secondary antibodies were purchased from Santa Cruz Biotechnology (Texas, USA) UBF:(UBF-9, sc-13125), rpS6:(Ribosomal protein S6 C-8, sc-74459), and Thermo Fisher Scientific () c-Myc:(c-Myc 9E10), goat anti-mouse for c-Myc (goat anti-mouse IgG1 (y1) horseradish peroxidase conjugate), and anti-mouse (anti-mouse IgG1 horseradish peroxidase conjugate).
-Antibodies were diluted in a 5% non-fat dry milk blocking buffer (TBS, 5% non-fat dry milk, 0.1% Tween-20).
-Membranes were incubated overnight with primary antibodies, and for 1hr with secondary antibodies.
-Between blocking and primary antibody staining, membranes were washed 1x5min, between primary and secondary staining, and after secondary staining, membranes were washed 3x5 with TBS-Tween (TBS; 20mM Tris, 150mM NaCl, 0.1% Tween).
-Following the last TBS-T wash, membranes were incubated 1x5min with enhanced chemiluminescent substrate (ECL, SuperSignal West Femto Maximum Sensitivity Substrate, Thermo Fisher Scientific).
-Membrane blocking, secondary antibody incubation, washing and ECL incubation were performed at room temp, on a tipping board.
-Primary antibody incubation was performed at 4°C on a tipping board.
-Chemiluminescence signals were quantified using Image Studio Lite (LI-COR Biotechnology, Lincoln, NE, USA), and total protein content was quantified using ImageJ (RUEDENETAL2017), whereas total protein content was defined as mean grey value of the whole well with between-well values subtracted as background.
-
 ## Results
 
-```{r glucose data + model}
-## Scripts:
-# - gluc.responsechange.R
+Baseline etc.
+- Blood data
+- Dietary data
 
-# Data
-glu.change <- readRDS("./data/data-gen/glucose/glu.change.RDS") # Untransformed data, used in figures
-glu.logchange <- readRDS("./data/data-gen/glucose/glu.logchange.RDS") # Log-transformed data, used in the analysis
+### Markers of ribosome biogenesis?
+- Total RNA? 
+- Protein?
+- UBF/total RNA?
 
-## Model
-# Untransformed
-glu.mod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = glu.change)
-glu.res <- plot(glu.mod)
-glu.sum <- summary(glu.mod)
-
-# Log-transformed
-glu.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = glu.logchange)
-glu.lres <- plot(glu.lmod)
-glu.lsum <- summary(glu.lmod)
-
-## Emmeans
-# Untransformed
-glu.emm <- confint(emmeans(glu.mod, specs = ~"supplement|time")) %>%
-        data.frame()
-
-# Log-transformed
-glu.lemm <- confint(emmeans(glu.lmod, specs = ~"supplement|time")) %>%
-        data.frame()
-
-# For inline plotting
-bcoef <- coef(summary(glu.lmod))
-pscient.120 <- round(bcoef[8,5], 30)
-p.120 <- sprintf("%.5f", 1.700971e-16)
-pscient.135 <- round(bcoef[9,5], 30)
-p.135 <- sprintf("%.5f", 4.299568e-11)
-pscient.150 <- round(bcoef[10,5], 30)
-p.150 <- sprintf("%.5f", 1.445712e-11)
-p.270 <- round(bcoef[11,5], 2)
-```
-
-```{r c-peptide data + model}
-## Scripts
-# - insulin.change.R
-
-## Data
-cpep.change <- readRDS("./data/data-gen/glucose/cpep.change.RDS") # Untransformed data, for figures
-cpep.lchange <- readRDS("./data/data-gen/glucose/cpep.logchange.RDS") # Log-transformed data, for analysis
-
-## Linear mixed effects model
-# Untransformed
-cpep.mod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = cpep.change) 
-cpep.res <- plot(cpep.mod)
-cpep.sum <- summary(cpep.mod)
-
-# Log-transformed
-cpep.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = cpep.lchange) 
-cpep.lres <- plot(cpep.lmod)
-cpep.lsum <- summary(cpep.lmod)
-
-## Emmeans
-# Untransformed
-cpep.emm <- confint(emmeans(cpep.mod, specs = ~"supplement|time")) %>%
-  data.frame()
-
-# Log-transformed
-cpep.lemm <- confint(emmeans(cpep.lmod, specs = ~"supplement|time")) %>%
-  data.frame()
-
-# For inline plotting
-cpcoef <- coef(summary(cpep.lmod))
-pscient.120 <- round(cpcoef[6,5], 30)
-cpp.120 <- sprintf("%.5f", 3.229267e-15)
-cpscient.150 <- round(cpcoef[7,5], 30)
-cpp.150 <- sprintf("%.5f", 1.414491e-15)
-cpp.270 <- round(cpcoef[8,5], 2)
-
-```
-
-At baseline, there were no significant differences between training with glucose or placebo in muscle mass, isometric or isokinetic peak torque, total session volume, training intensity, total RNA content, or rRNA expression (Tab 3).
-Further, on pairwise consecutive days, there were no significant differences in mean macro nutrient (calories, carbohydrate, fat or protein) ingestion (Tab 4), or mean change in macro nutrient ingestion (Fig 3).
-Resistance training with glucose led to significant increases in blood levels glucose and c-peptide compared to RT with placebo, before, during and after the session (Fig 4B, 120min: 38 ± 4% [*p* = `r p.120`], 135min: 31 ± 4% [*p* = `r p.135`], 150min: 32 ± 4% [*p* = `r p.150`], Fig 4C, 120min: 95 ± 10% [*p* = `r cpp.120`], 150min: 87 ± 10% [*p* = `r cpp.150`]), measured during the final RT session.
-At 2hrs post-exercise, RT with glucose led to significantly lower blood glucose levels compared to RT with placebo, with a tendancy of the same in c-peptide levels (4B, 270min: -8 ± 4% [*p* = `r p.270`] and 4C, 270min: -7 ± 10 [*p* `r cpp.270`]).
-
-###### **Table 3**: Baseline characteristics of the intervention legs, organized per supplement.
-Variables: Iso 60 = isokinetic 60 d/s, Iso 240 = isokinetic 240 d/s, training intensity = load/rm*100.
-Nm = newton meters, kg = kilograms, %1RM = percentage of 1 repetition maximum, ng/mg = nanogram per milligram wet muscle tissue, nf.expr = expression normalized by weight and external reference gene (Lambda).
-Ribosomal RNA data are presented as log transformed.
-Glucose n = 13, placebo n = 13.
-
-![](figures/table3.png)
-
-###### **Table 4:** Daily mean dietary intake of macro nutrients during each training day per supplement leg.
-G = glucose, P = placebo.
-Training day 1-training day 12 are combined pairwise to day 1-6: Day 1 = training day 1/2, Day 2 = training day 3/4, Day 3 = training day 5/6, Day 4 = training day 7/8, Day 5 = training day 9/10, Day 6 = training day 11/12.
-Values are mean ± standard deviation.
-P = difference between glucose and placebo.
-Glucose n = 12, placebo n = 12.
-Day 6: Glu = 11, Plac = 11.
-
-```{r nutrition table}
-## Scripts
-# - nut.table.R
-# - nut.analysis.R
-# - dxa.R
-
-# Data
-nuttable <- readRDS("./data/data-gen/dxa/nuttable.RDS")
-
-nuttable %>%
-  kable(col.names = c("Time", "Suppl.", "Kcal", "CHO (g)", "Fat (g)", "PRO (g)", "PRO (g/kg)"))
-
-```
-
-```{r nutrition change analysis and fig}
-# Associated scripts:
-# - nut.datahand.R
-# - nut.change.R
-# - nut.change.figs.R
-
-nut.fig <- readRDS("./data/data-gen/nutrition/nut.fig.RDS")
-nut.fig
-
-```
-###### **Figure 3:** Mean changes in macro nutrient intake from day to day.
-Day 1 = training day 1/2(T1/T2), Day 2 = training day 3/4, Day 3 = training day 5/6, Day 4 = training day 7/8, Day 5 = training day 9/10, Day 6 = training day 11/12(T3/T4).
-Values are mean change in grams/calories ± 95% CL.
-
-![](figures/day.timeline.png)
-
-```{r glucose + c-peptide fig}
-## Scripts
-# - gluc.fig.R
-# - cpep.figs.R
-
-# Separate figures
-pos <- position_dodge(width = 0.2)
-
-gluc.fig <- readRDS("./data/data-gen/glucose/gluc.fig.RDS")
-
-cpep.fig <- readRDS("./data/data-gen/glucose/cpep.fig.RDS")
-
-# Joined figure
-plot_grid(gluc.fig,
-          cpep.fig,
-          ncol = 1, nrow = 2)
-
-```
-
-###### **Figure 4:** A) Time line for training day 6 (combination of pairwise training day 11 and 12).
-Biopsy = muscle biopsy taken pre supplementation, PRO = protein supplementation, Humac = strength test (isometric knee extension, isokinetic knee extension 60 and 240 d/s), GLU/PLAC = glucose/placebo supplementation, RT = resistance training.
-Minutes: -120 = 2hrs prior to RT, -90 = 90min prior to RT, -30 = 30min prior to RT, 0 = onset of RT, 15 = 15min following onset of RT, between leg press and knee extension, 30 = 30min post onset of RT, 120 = 2hrs following onset of RT.
-B) Mean mmol/L change in blood glucose levels per supplement group during training day 6.
-Time-points: 0, 45, 90, 120, 135, 150 and 270min post protein ingestion, respectively.
-C) Mean change in pmol/L of c-peptide in blood per supplement group during training day 6.
-Values are presented as estimated marginal means of mmol/L and pmol/L change ± 95% CI.
-*p* < 0.05 between groups.
-Glucose n = 13, placebo n = 13.
-
-### Markers of ribosome biogenesis
-
-```{r total RNA model}
-## Scripts
-# - totrna.cleanup.R
-# - totrna.analysis.R
-
-# Data
-trna.lchange <- readRDS("./data/data-gen/rna/totrna.change.RDS")
-
-# Create model: 
-trna.lmod <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
-           data = trna.lchange)
-
-trna.lres <- plot(trna.lmod)
-trna.lsum <- summary(trna.lmod)
-
-### Get estimated means from the model
-trna.emm <- confint(emmeans(trna.lmod, specs = ~"supplement")) %>%
-    data.frame()
-
-# For inline reporting
-trcoef <- coef(summary(trna.lmod))
-trdiff <- round(trcoef[3,1], 2)
-trdiffp <- round(trcoef[3,5], 2)
-trdiffse <- round(trcoef[3,2], 2)
-
-```
-
-```{r rRNA data model}
-## Scripts
-# - qpcr-import
-# - qpcr-analysis
-# - rrna.change
-# - rrnachange.figs
-
-# Log-transformed data
-lchange.18 <- readRDS("./data/data-gen/rna/lchange.18.RDS")
-lchange.28 <- readRDS("./data/data-gen/rna/lchange.28.RDS")
-lchange.58 <- readRDS("./data/data-gen/rna/lchange.58.RDS")
-lchange.5 <- readRDS("./data/data-gen/rna/lchange.5.RDS")
-lchange.47 <- readRDS("./data/data-gen/rna/lchange.47.RDS")
-
-# Create model: 
-# Needs to have an intercept per participant (mixed model)
-# Control for pre values.
-
-# 18S
-lmod.18 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
-                     data = lchange.18)
-lres.18 <- plot(lmod.18)
-lsum.18 <- summary(lmod.18)
-
-# 28S
-lmod.28 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
-                     data = lchange.28)
-lres.28 <- plot(lmod.28)
-lsum.28 <- summary(lmod.28)
-
-# 5.8S
-lmod.58 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
-                     data = lchange.58)
-lres.58 <- plot(lmod.58)
-lsum.58 <- summary(lmod.58)
-
-# 5S
-lmod.5 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
-                     data = lchange.5)
-lres.5 <- plot(lmod.5)
-lsum.5 <- summary(lmod.5)
-
-# 47S
-lmod.47 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
-                     data = lchange.47)
-lres.47 <- plot(lmod.47)
-lsum.47 <- summary(lmod.47)
-
-### Get estimated means from the model, these are average increase at 
-# pre = 0 (the average pre value)
-# remember that these are originally log-fold change values, transformed back using exp(emmean) etc
-
-# 18S
-lemm.18 <- confint(emmeans(lmod.18, specs = ~"supplement")) %>%
-        data.frame()
-
-# 28S
-lemm.28 <- confint(emmeans(lmod.28, specs = ~"supplement")) %>%
-        data.frame()
-
-# 5.8S
-lemm.58 <- confint(emmeans(lmod.58, specs = ~"supplement")) %>%
-        data.frame()
-
-# 5S
-lemm.5 <- confint(emmeans(lmod.5, specs = ~"supplement")) %>%
-        data.frame()
-
-# 47S
-lemm.47 <- confint(emmeans(lmod.47, specs = ~"supplement")) %>%
-        data.frame()
-
-## For inline reporting
-
-# 18S
-rr18coef <- coef(summary(lmod.18))
-rr18diffp <- round(rr18coef[3,5], 3)
-rr18diffse <- round(rr18coef[3,2], 3)
-
-# 28S
-rr28coef <- coef(summary(lmod.28))
-rr28diffp <- round(rr28coef[3,5], 3)
-rr28diffse <- round(rr28coef[3,2], 3)
-
-# 5.8S
-rr58coef <- coef(summary(lmod.58))
-rr58diffp <- round(rr58coef[3,5], 3)
-rr58diffse <- round(rr58coef[3,2], 3)
-
-# 5S
-rr5coef <- coef(summary(lmod.5))
-rr5diffp <- round(rr5coef[3,5], 3)
-rr5diffse <- round(rr5coef[3,2], 3)
-
-# 47S
-rr47coef <- coef(summary(lmod.47))
-rr47diffp <- round(rr47coef[3,5], 3)
-rr47tdiffp <- round(rr47coef[1,5], 3)
-rr47diffse <- round(rr47coef[3,2], 3)
-
-```
-
-Resistance training with glucose did not induce a higher accumulation of total RNA (fig 5A, *p* = `r trdiffp`), or rRNA  (fig 5B, *p* > 0.05), compared to RT with placebo, measured as mean change from baseline to post.
-From baseline to post-intervention, there was a mean increase in levels of total RNA by 26 and 22% after RT with glucose and placebo, respectively (*p* < 0.05, compared to baseline).
-A robust accumulation was also observed in ribosomal RNA expression in both RT with glucose and placebo, with mean increases between 34-43% (GLU) and 33-41% (PLAC) (*p* < 0.05, compared to baseline) in the four rRNA’s.
-The expression of 47S pre-rRNA increased by 37 and 59% with glucose and placebo respectively, with only placebo reaching significantly higher levels at post compared to baseline (fig 45, *p* = `r rr47tdiffp`).
-
-```{r RNA figures}
-# Get figures
-totrna.fig <- readRDS("./data/data-gen/rna/totrna.fig.RDS")
-totrna.fig
-
-plot.18 <- readRDS("./data/data-gen/rna/plot.18.RDS")
-plot.28 <- readRDS("./data/data-gen/rna/plot.28.RDS")
-plot.5.8 <- readRDS("./data/data-gen/rna/plot.5.8.RDS")
-plot.5 <- readRDS("./data/data-gen/rna/plot.5.RDS")
-plot.47 <- readRDS("./data/data-gen/rna/plot.47.RDS")
-
-# Joined rRNA
-library(cowplot)
-
-legend <- get_legend(plot.47 + theme(legend.box.margin = margin(0, 0, 0,12)))
-
-plot_grid(plot.18 + theme(legend.position = "none"),
-          plot.28 + theme(legend.position = "none"), 
-          plot.5.8 + theme(legend.position = "none"),
-          plot.5 + theme(legend.position = "none"), 
-          plot.47 + theme(legend.position = "none"),
-          legend,
-          ncol = 2, nrow = 3)
-
-```
-
-###### **Figure 5:** A) Changes in Total RNA, B) Changes in ribosomal RNA.
-Baseline = Training day 1/2, Post = Training day 11/12.
-Both total RNA and rRNA were analyzed with duplicates, with two duplicates per biopsy (two biopsies per time point).
-Total RNA and rRNA changes were calculated as log-fold change score per mg wet muscle weight.
-Mean change scores of the duplicates were calculated before modelling and transformed to the log-scale.
-Values are estimated marginal means fold change per leg/supplement ± 95% CI.
-p = between groups statistic, glucose compared to placebo.
-Glucose n = 13, placebo n = 13.
-
-```{r western modelling}
-## Script: 
-# - western.import.R
-# - western.change.R
-
-# Data
-
-prot.dat <- readRDS("./data/data-gen/protein/prot.dat.RDS") # cleaned up and quantified western data
-prot.out <- readRDS("./data/data-gen/protein/protein_complete.RDS") # Cleaned up and quantified western data, checked for outliers
-prot.change <- readRDS("./data/data-gen/protein/prot.change.RDS") # change-score data
-
-# By target
-
-cmyc <- prot.change %>%
-  filter(target == "cmyc")
-
-ubf <- prot.change %>%
-  filter(target == "ubf")
-
-rps6 <- prot.change %>%
-  filter(target == "rps6")
-
-# c-Myc change
-
-m1 <- lme(change ~ pre + supplement,
-            random = list(subject = ~ 1),
-            data = cmyc) # same model as in "western import", but with change scores of norm.sign
-
-#plot(m1)
-
-sum.m1 <- summary(m1)
-
-cmyc.emm <- emmeans(m1, specs = ~"supplement") %>%
-  data.frame
-
-saveRDS(cmyc.emm, "./data/data-gen/protein/cmyc.change.RDS")
-
-# UBF change
-
-m2 <- lme(change ~ pre + supplement,
-            random = list(subject = ~ 1),
-            data = ubf) # same model as in "western import", but with change (i.e., change in norm.sign) instead of norm.sign
-
-#plot(m2)
-
-sum.m2 <- summary(m2)
-
-ubf.emm <- emmeans(m2, specs = ~"supplement") %>%
-  data.frame()
-
-saveRDS(ubf.emm, "./data/data-gen/protein/ubf.change.RDS")
-
-# rps6 change
-
-m3 <- lme(change ~ pre + supplement,
-            random = list(subject = ~ 1),
-            data = rps6) # same model as in "western import", but with change (i.e., change in norm.sign) instead of norm.sign
-
-#plot(m3)
-
-sum.m3 <- summary(m3)
-
-rps6.emm <- emmeans(m3, specs = ~"supplement") %>%
-  data.frame()
-
-saveRDS(rps6.emm, "./data/data-gen/protein/rps6.change.RDS")
-
-### For inline reporting
-## Diff cmyc
-ccoef <- coef(summary(m1))
-cdiff <- round(ccoef[3,5], 3)
-cdiffse <- round(ccoef[3,2], 3)
-
-## Diff UBF
-ucoef <- coef(summary(m2))
-udiff <- round(ucoef[3,5], 3)
-udiffse <- round(ucoef[3,2], 3)
-
-## Diff rps6
-rpcoef <- coef(summary(m3))
-rpdiff <- round(rpcoef[3,5], 3)
-rpdiffse <- round(rpcoef[3,2], 3)
-
-```
-
-```{r western fig}
-# Data
-
-cmyc <- readRDS("./data/data-gen/protein/cmyc.change.RDS")
-
-ubf <- readRDS("./data/data-gen/protein/ubf.change.RDS")
-
-rps6 <- readRDS("./data/data-gen/protein/rps6.change.RDS")
-
-## cmyc fig
-
-cmyc.plot <- cmyc %>%
-  mutate(time = "post") %>%
-  add_row(supplement = c("PLACEBO", "GLUCOSE"), time = "pre", emmean = 0, lower.CL = 0, upper.CL = 0) %>%
-  mutate(time = factor(time, levels = c("pre", "post"), labels = c("Baseline", "Post"))) %>%
-  ggplot(aes(time, exp(emmean), group = supplement, fill = supplement)) +
-  #annotate("text", x = "Post", y = 2.1, label = "p = 0.585", size = 3) +
-  geom_errorbar(aes(ymin = exp(lower.CL), ymax = exp(upper.CL)),
-                width = 0.1,
-                position = position_dodge(width = 0.2)) +
-  geom_line(position = position_dodge(width = 0.2)) +
-  geom_point(shape = 21, size = 3, position = position_dodge(width = 0.2)) +
-  labs(x = "", y = "c-Myc signal \n(fold change)\n", fill = "") +
-  theme_classic()
-
-ubf.plot <- ubf %>%
-  mutate(time = "post") %>%
-  add_row(supplement = c("PLACEBO", "GLUCOSE"), time = "pre", emmean = 0, lower.CL = 0, upper.CL = 0) %>%
-  mutate(time = factor(time, levels = c("pre", "post"), labels = c("Baseline", "Post"))) %>%
-  ggplot(aes(time, exp(emmean), group = supplement, fill = supplement)) +
-  #annotate("text", x = "Post", y = 2.1, label = "p = 0.585", size = 3) +
-  geom_errorbar(aes(ymin = exp(lower.CL), ymax = exp(upper.CL)),
-                width = 0.1,
-                position = position_dodge(width = 0.2)) +
-  geom_line(position = position_dodge(width = 0.2)) +
-  geom_point(shape = 21, size = 3, position = position_dodge(width = 0.2)) +
-  labs(x = "", y = "UBF signal \n(fold change)\n", fill = "") +
-  theme_classic()
-
-rps6.plot <- rps6 %>%
-  mutate(time = "post") %>%
-  add_row(supplement = c("PLACEBO", "GLUCOSE"), time = "pre", emmean = 0, lower.CL = 0, upper.CL = 0) %>%
-  mutate(time = factor(time, levels = c("pre", "post"), labels = c("Baseline", "Post"))) %>%
-  ggplot(aes(time, exp(emmean), group = supplement, fill = supplement)) +
-  annotate("text", x = "Post", y = 1.9, label = "p = ", size = 3) +
-  geom_errorbar(aes(ymin = exp(lower.CL), ymax = exp(upper.CL)),
-                width = 0.1,
-                position = position_dodge(width = 0.2)) +
-  geom_line(position = position_dodge(width = 0.2)) +
-  geom_point(shape = 21, size = 3, position = position_dodge(width = 0.2)) +
-  labs(x = "", y = "rps6 signal \n(fold change)\n", fill = "") +
-  theme_classic()
-
-# Cowplot for gathering figures
-
-legend <- get_legend(rps6.plot + theme(legend.box.margin = margin(0, 0, 0,12)))
-
-prot.joined <- plot_grid(cmyc.plot + theme(legend.position = "none"),
-                         ubf.plot + theme(legend.position = "none"), 
-                         rps6.plot + theme(legend.position = "none"),
-                         legend,
-                         ncol = 2, nrow = 2)
-
-prot.joined
-
-saveRDS(prot.joined, "./data/data-gen/rna/prot.joined.RDS")
-
-```
-
-###### **Figure 5:** A) Changes in c-Myc levels, B) changes in UBF levels, C) changes in rps6 levels.
-Baseline = Training day 1/2, post = Training day 11/12.
-Four samples were analysed per participant (GLU pre-post, PLAC pre-post), in duplicates, a total of 104 protein samples.
-Duplicate samples were loaded on separate gels.
-In addition, 3-4 pools were added to each well.
-c-Myc, UBF and rps6 signals were then calculated as log-fold change score, normalised by gel pools.
-Mean change scores of the duplicates were calculated before modelling and transformed to the log-scale.
-Values are estimated marginal means fold change per leg/supplement ± 95% CI.
-p = between groups statistic, glucose compared to placebo.
-Glucose n = 13, placebo n = 13.
-
-### Resistance training and strength testing
-
-```{r training data model}
-## Scripts
-# - volume.change.R
-# - ss.analysis.R
-# - ss.cleanup.R
-# - trainint.analysis.R
-# - trainint.cleanup.R
-
-# Log-transformed data
-vol.lchange <- readRDS("./data/data-gen/training/vol.lchange.RDS")
-ss.lchange <- readRDS("./data/data-gen/training/ss.lchange.RDS")
-int.lchange <- readRDS("./data/data-gen/training/int.lchange.RDS")
-
-### Training volume
-# Fold-change model
-vol.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = vol.lchange)
-vol.lres <- plot(vol.lmod)
-vol.lsum <- summary(vol.lmod)
-
-# Fold-change estimated means
-vol.lemm <- confint(emmeans(vol.lmod, specs = ~"supplement|time")) %>%
-        data.frame()
-
-### Session score (RPE)
-# Model
-ss.lmod <- lmerTest::lmer(change ~ baseline + time + supplement:time + (1|subject),
-                     data = ss.lchange)
-ss.lres <- plot(ss.lmod)
-ss.lsum <- summary(ss.lmod)
-
-# Session score emmeans
-ss.lemm <- confint(emmeans(ss.lmod, specs = ~"supplement|time")) %>%
-        data.frame()
-
-### Training intensity
-int.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = int.lchange)
-int.lres <- plot(int.lmod)
-int.lsum <- summary(int.lmod)
-
-# Emmeans
-int.lemm <- confint(emmeans(int.lmod, specs = ~"supplement|time")) %>%
-        data.frame()
-
-### For inline reporting
-
-## Training volume
-# Diff 4
-vcoef <- coef(summary(vol.lmod))
-vdiff.4 <- round(vcoef[9,1], 3)
-vdiffp.4 <- round(vcoef[9,5], 3)
-vdiffse.4 <- round(vcoef[9,2], 3)
-# Diff 5
-vdiff.5 <- round(vcoef[10,1], 3)
-vdiffp.5 <- round(vcoef[10,5], 3)
-vdiffse.5 <- round(vcoef[10,2], 3)
-# Diff 6
-vdiff.6 <- round(vcoef[11,1], 3)
-vdiffp.6 <- round(vcoef[11,5], 3)
-vdiffse.6 <- round(vcoef[11,2], 3)
-## RPE
-scoef <- coef(summary(ss.lmod))
-sdiffp <- round(scoef[11, 5], 3)
-sdiffse <- round(scoef[11, 2], 3)
-
-## Training intensity
-
-# Diff time 2
-icoef <- coef(summary(int.lmod))
-idiff2 <- round(icoef[7,1], 3)
-idiffp2 <- round(icoef[7,5], 3)
-idiffse2 <- round(icoef[7,2], 3)
-# Diff time 3
-idiff3 <- round(icoef[8,1], 3)
-idiffp3 <- round(icoef[8,5], 3)
-idiffse3 <- round(icoef[8,2], 3)
-# Diff time 4
-idiff4 <- round(icoef[9,1], 3)
-idiffp4 <- round(icoef[9,5], 3)
-idiffse4 <- round(icoef[9,2], 3)
-# Diff time 5
-idiff5 <- round(icoef[10,1], 3)
-idiffp5 <- round(icoef[10,5], 3)
-idiffse5 <- round(icoef[10,2], 3)
-# Diff time 6
-idiff6 <- round(icoef[11,1], 3)
-idiffp6 <- round(icoef[11,5], 3)
-idiffse6 <- round(icoef[11,2], 3)
-```
-
-```{r humac post 5th model}
-## Scripts
-# - humac.cleanup.R
-# - humac.main.analysis.R - this is data gathered at baseline, during the intervention, and after session 5
-
-# Log-transformed data
-
-isom.lchange <- readRDS("./data/data-gen/humac/isom.lchange.RDS")
-iso60.lchange <- readRDS("./data/data-gen/humac/isok60.lchange.RDS")
-iso240.lchange <- readRDS("./data/data-gen/humac/isok240.lchange.RDS")
-
-# Isometric
-isom.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = isom.lchange)
-isom.lres <- plot(isom.lmod)
-isom.lsum <- summary(isom.lmod)
-
-# Isok.60
-iso60.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = iso60.lchange)
-iso60.lres <- plot(iso60.lmod)
-iso60.lsum <- summary(iso60.lmod)
-
-# Isok.240
-iso240.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = iso240.lchange)
-iso240.lres <- plot(iso240.lmod)
-iso240.lsum <- summary(iso240.lmod)
-
-## Emmeans
-# Isometric
-isom.lemm <- confint(emmeans(isom.lmod, specs = ~"supplement|time")) %>%
-        data.frame()
-
-# Isok.60
-iso60.lemm <- confint(emmeans(iso60.lmod, specs = ~"supplement|time")) %>%
-        data.frame()
-
-# Isok.240
-iso240.lemm <- confint(emmeans(iso240.lmod, specs = ~"supplement|time")) %>%
-        data.frame()
-
-# For reporting
-# Isom diff
-isom5coef <- coef(summary(isom.lmod))
-isom5diffp <- round(isom5coef[7, 5], 3)
-isom5diffse <- round(isom5coef[7, 2], 3)
-
-# Isok 60 diff
-isok605coef <- coef(summary(iso60.lmod))
-isok605diffp <- round(isok605coef[7, 5], 3)
-isok605diffse <- round(isok605coef[7, 2], 3)
-
-# Isok 260
-isok2405coef <- coef(summary(iso240.lmod))
-isok2405diffp <- round(isok2405coef[7, 5], 3)
-isok2405diffse <- round(isok2405coef[7, 2], 3)
-
-```
-
-```{r post 6th model}
-## Scripts
-# - humac.acute.analysis.R - this is data gatherede before and after one RT session (sixth session)
-# - humac.cleanup.R
-
-# Data
-isom.lchange.ac <- readRDS("./data/data-gen/humac/isom.lchange.ac.RDS")
-iso60.lchange.ac <- readRDS("./data/data-gen/humac/iso60.lchange.ac.RDS")
-iso240.lchange.ac <- readRDS("./data/data-gen/humac/iso240.lchange.ac.RDS")
-
-# Isometric
-isomac.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = isom.lchange.ac)
-isomac.lres <- plot(isomac.lmod)
-isomac.lsum <- summary(isomac.lmod)
-
-# Isok.60
-iso60ac.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = iso60.lchange.ac)
-iso60ac.lres <- plot(iso60ac.lmod)
-iso60ac.lsum <- summary(iso60ac.lmod)
-
-# Isok.240
-iso240ac.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = iso240.lchange.ac)
-iso240ac.lres <- plot(iso240ac.lmod)
-iso240ac.lsum <- summary(iso240ac.lmod)
-
-# Emmeans
-# Isometric
-isomac.lemm <- confint(emmeans(isomac.lmod, specs = ~"supplement|time")) %>%
-        data.frame()
-
-# Isok.60
-iso60ac.lemm <- confint(emmeans(iso60ac.lmod, specs = ~"supplement|time")) %>%
-        data.frame()
-
-# Isok.240
-iso240.lemm <- confint(emmeans(iso240ac.lmod, specs = ~"supplement|time")) %>%
-        data.frame()
-
-## For inline reporting
-# Isom diff
-isomcoef <- coef(summary(isomac.lmod))
-isomdiffp <- round(isomcoef[7, 5], 3)
-isomdiffse <- round(isomcoef[7, 2], 3)
-
-# Isok 60 diff
-isok60coef <- coef(summary(iso60ac.lmod))
-isok60diffp <- round(isok60coef[7, 5], 3)
-isok60diffse <- round(isok60coef[7, 2], 3)
-
-# Isok 240 diff
-isok240coef <- coef(summary(iso240ac.lmod))
-isok240diffp <- round(isok240coef[7, 5], 3)
-isok240diffse <- round(isok240coef[7, 2], 3)
-
-```
-
-There were no significant differences in mean change of total session volume, training intensity or RPE between exercise training with glucose and placebo at any of the time points (fig 6, *p* > 0.05).
-We did observe interactions between time and changes, where total session training volume increased from 13-18% compared to baseline.
-These increases in training volume were accompanied by changes ranging from 5-18% in training intensity and 2-13% in RPE (Tab 5).
-The maximal isometric knee extension force (OR TORQUE?) and isokinetic knee extension peak torque at 240 d/s showed no differences between exercise training with glucose and placebo after five RT sessions (Fig 7A, *p* > 0.05).
-There was a significant difference in mean change from pre to post-fifth RT session, where exercise training with glucose led to a 9% less decrease in peak torque at 60 d/s compared to placebo (Fig 7A, *p* = `r isok605diffp`).
-There was no difference between glucose and placebo in mean change from baseline to after the sixth RT session (Fig 7B, *p* > 0.05).
-
-![](figures/legend.PNG)
-
-```{r training dat fig}
-### Get figures
-
-# Volume
-totvol.barplot <- readRDS("./data/data-gen/training/totvol.barplot.RDS")
-totvolfold.plot <- readRDS("./data/data-gen/training/totvolfold.plot.RDS")
-
-# Training intensity
-tot.barplot <- readRDS("./data/data-gen/training/tot.barplot.RDS")
-totalint.plot <- readRDS("./data/data-gen/training/trainint.plot.RDS")
-
-# Join figures
-plot_grid(totvol.barplot + theme(legend.position = "none"),
-          totvolfold.plot + theme(legend.position = "none"),
-          tot.barplot + theme(legend.position = "none"),
-          totalint.plot + theme(legend.position = "none"),
-          ncol = 2,
-          nrow = 2,
-          align = "h")
-
-```
-
-###### **Figure 6:** Training volume and training intensity from baseline until resistance exercise session 6 (training day 11 and 12).
-Baseline = Training day 1/2, Session 2 = Training day 3/4, Session 3 = Training day 5/6, Session 4 = Training day 7/8, Session 5 = Training day 9/10, Session 6 = Training day 11/12.
-A) Changes in total session volume per leg, values are mean kg ± standard deviation.
-B) Fold change in total session volume per leg, values are estimated marginal means fold change ± 95% CI.
-C) Changes in training intensity per leg calculated as %1RM, values are mean % ± standard deviation.
-D) Fold change in training intensity per leg, values are estimated marginal means fold change ± 95% CI.
-† = *p* < 0.05 in both glucose and placebo, compared to baseline.
-p = NS = no significant difference between legs at any of the time points.
-Glucose n = 13, placebo n = 13.
-
-###### **Table 5:** RPE changes through training days.
-Training days are presented as combined pairwise consecutive, linked to their respective sessions.
-Session 1 = Training day 1/2, Session 2 = Training day 3/4, Session 3 = Training day 5/6, Session 4 = Training day 7/8, Session 5 = Training day 9/10, Session 6 = Training day 11/12.
-Values are presented as mean ± standard deviation.
-Glucose n = 13, placebo n = 13.
-
-```{r RPE table}
-## Scripts
-# - ss.cleanup.R
-# - ss.absolute.R
-
-# Data
-ss.abs <- readRDS("./data/data-gen/training/ss.abs.RDS")
-
-# Table
-ss.abs %>%
-  kable(col.names = c("Supplement", "Baseline", "Session 2", "Session 3", "Session 4", "Session 5",
-                            "Session 6"))
-
-```
-
-![](figures/legend2.PNG)
-
-```{r humac figs}
-## Post 5th RE session
-# Figures
-pos <- position_dodge(width = 0.2)
-isom5.plot <- readRDS("./data/data-gen/humac/isom.main.fig.RDS")
-isok605.plot <- readRDS("./data/data-gen/humac/isok60.main.fig.RDS")
-isok2405.plot <- readRDS("./data/data-gen/humac/isok240.main.fig.RDS")
-
-legend <- get_legend(isom5.plot + theme(legend.box.margin = margin(0, 0, 0, 12)))
-
-plot_grid(isom5.plot + theme(legend.position = "none"), 
-          isok605.plot + theme(legend.position = "none"), 
-          isok2405.plot + theme(legend.position = "none"),
-          nrow = 3,
-          align = "v")
-
-## Post 6th RE session
-
-# Figures
-
-pos <- position_dodge(width = 0.2)
-
-isom.plot <- readRDS("./data/data-gen/humac/isom.ac.plot.RDS")
-isok60.plot <- readRDS("./data/data-gen/humac/isok60.ac.plot.RDS")
-isok240.plot <- readRDS("./data/data-gen/humac/isok240.ac.plot.RDS")
-
-legend <- get_legend(isom.plot + theme(legend.box.margin = margin(0, 0, 0, 12)))
-
-plot_grid(isom.plot + theme(legend.position = "none"), 
-          isok60.plot + theme(legend.position = "none"), 
-          isok240.plot + theme(legend.position = "none"),
-          nrow = 3,
-          align = "v")
-
-```
-
-The mean change in peak torque at all three velocities change little from baseline until session 4 in both glucose and placebo (+4-7% change), decreasing at pre-session 6 (9-18% change, Fig 6A).
-Post session 6 (Fig 6B) showed a slight reduction in isometric peak torque 30min after session completion (2-4% change), slowly regaining at 2hrs (+3% change) and 23hrs (+3-6% change), in both glucose and placebo compared to baseline.
-Isokinetic peak torque at 60 d/s and 240 d/s showed no reduction in mean peak torque from baseline until 23hrs following RT,  in both glucose and placebo (+0,1-11%).
-There was an effect of time in peak torque at 23hrs post-RT at 60 d/s (Fig 6B), where exercise training with placebo led to a 5% increase in peak torque compared to baseline, however, there was no significant difference compared to exercise training with glucose.
-
-###### **Figure 7:** Peak torque changes in maximal isometric knee extension and maximal isokinetic knee extension at 60 and 240 d/s, respectively.
-A) Humac tests during the intervention, Baseline = Day -1 (prior to intervention), Session 2 = prior to training, training Day 4/5, Session 4 = prior training, training Day 8/9, Session 6 = prior to training 11/12.
-B) Baseline = prior to training, training day 11/12, 30min = 30 min after RT, 2hr = 2hrs after RT, 23hr = 23hrs after RT.
-Values are change in estimated marginal means of newton meters (Nm) ± 95% CI.
-* = *p* < 0.05 between groups.
-*p* = NS = no significant difference between legs at any of the time points.
-Glucose n = 13, placebo n = 13.
+### Muscular performance and recovery?
 
 ## Discussion
-
-The main findings of the present study were that resistance training with glucose for five sessions did not affect markers of ribosome biogenesis, measured as total RNA, rRNA and rDNA transcription initiation associated proteins, compared to RT with placebo.
-Similarly, RT with glucose did not affect markers of skeletal muscle functionality such as muscular performance and -recovery, total session volume, and training intensity, compared to RT with placebo.
-There are studies suggesting that high vs.
-low glucose treatment/glucose starvation impacts ribosome biogenesis in yeast, rodent and human non-muscle cells (KIM13;MARIAPPAN11;TANAKA15;ZHAI12), however, this effect does not seem evident when total daily glucose intake is matched.
-
-To our knowledge, no previous study has investigated the effects of glucose compared to placebo ingestion with RT on ribosome biogenesis in human skeletal muscle cells, making our hypotheses exploratory and comparisons not straight forwards.
-The current model of ribosome biogenesis suggests that changes in the accumulation of total RNA and rRNA early after the onset of an RT period predict long-term RT adaptations (FIGUEIREDO19; HAMMARSTROM20; HAMMARSTROM21).Therefore, we assumed 2 weeks to be a suitable time frame for a moderate RT volume to elicit changes in ribosome biogenesis, as similar observations have been made previously (FIGUEIREDO16; HAMMARSTROM20; HAMMARSTROM21).
-In Hammarström et al.
-(2020), participants that increased ribosome biogenesis after two weeks (5 RT sessions) of RT, also responded well after the entire intervention (12 weeks, 31 sessions).
-Figueiredo and colleagues (2016) also observed markedly acute effects, where 47S pre-rRNA increased after a single RT session, passing baseline after 4hrs, peaking at 24hrs and remaining elevated until 48hrs post-exercise.
-The four mature rRNA did not measurably increase until after multiple sessions (FIGUEIREDO16).
-As such, increases in total RNA and rRNA observed in the present study are in line with previous observations (FIGUEIREDO16; HAMMARSTROM20; H21?), and also evident i moderately trained individuals, compared to the untrained or RT-untrained individuals of previous studies (FIGUEIREDO16;HAMMARSTROM20;H21?).
-
-Further, the accumulation of total RNA and rRNA is also suggested to reach a plateau after approximately 5 RT-sessions, thus, in addition to the demanding RT- and dietary protocol and the exploratory nature of the hypotheses, we did not conduct a more long-term intervention.
-COMMENT ON PROTEIN DATA HERE?
-
-As we did not observe any effect of ingesting glucose before and after RT compared to placebo in this design, there is little to suggest that glucose *per se* has any additional effect on ribosome biogenesis, as long as the daily energy intake is sufficient to support cellular energy levels required to drive this process.
-Looking at previous observations from *in vitro* studies of rodents, yeast, and non-muscle cell cultures (MARIAPPAN11; KIM13; TANAKA18; ZHAI12; HOPPE09; MURAYAMA08), one interesting notion would be how high glucose vs.
-low glucose or glucose starvation would affect ribosome biogenesis over time in human skeletal muscle cells.
-Time being of the essence in such a perspective, as previous *in vitro* studies have utilized cells with higher physiological rate than human skeletal muscle cells (KIM13;MARIAPPAN11;TANAKA15;ZHAI12).
-
-Indeed, our findings suggest that glucose ingestion *per se* before and after RT does not induce changes in the internal physiological milieu of human skeletal muscle cells with anabolic effects.
-It has been theorised that glucose may convey an anabolic signal through insulin, and thus operate as a signal mediator affecting upstream pathways of ribosome biogenesis (KUSNADI15; TANAKA18).
-Such an anabolic signal could then possibly increase PIC assembly, and consequently Pol I recruitment, initiating rDNA transcription and subsequent rRNA maturation, leading to increased ribosome content (FIGUEIREDO19; TANAKA18).
-As significantly higher blood levels of c-peptide and glucose were observed in RT with glucose compared to placebo in the present study, without any difference in ribosome biogenesis, there is little to suggest that either insulin or glucose has any direct acute effects.
-
-Another theory is that glucose ingestion leads to higher cellular energy levels compared to placebo, stimulating energy-sensitive pathways such as PIH1, mTORC1, ERK1/2, AMPK and SIRT1 in a dose-response relationship (KIM13;MARIAPPAN11;TANAKA15;ZHAI12).
-The mTORC1, along with ERK1/2 and PIH1 up-regulates PIC assembly during high-energy conditions, while AMPK and possibly SIRT1 represses PIC assembly during low-energy conditions, as observed in rodent (KILDE), yeast (KILDE), and non-muscle cells (KILDE) *in vitro*.
-Specifically, high glucose treatment augmented rDNA transcription by increased UBF and SL-1 activity, while low glucose treatment, glucose starvation, or nutrient depletion stimulated AMPK-dependent TIF-IA and chromatin regulation suppression (KOS-BRAUN17; MARIAPPAN11; ZHAI12).
-In addition to UBF and SL-1, assembly of the PIC requires TIF-IA and Pol I recruitment, induced by amino acid intake in an mTOR-S6K-dependent manner and possibly by direct c-Myc activation (KUSNADI15).
-As such, increasing energy supply and nutrient availability should in line with these theories potently stimulate PIC assembly and rDNA transcription.
-Importantly, these pathways were not directly investigated in the present study.
-However, downstream targets (UBF, c-Myc, rpS6?) remained unaffected by glucose compared to placebo ingestion, suggesting there is not a dose-response relationship between muscle cellular energy levels and ribosome biogenesis during an RT-intervention (Not sure if I can say this).
-
-A weakness in the present study, is the lack of data on intracellular glucose levels in the muscle cells.
-Thus, it cannot be said for sure whether there were differences between glucose and placebo or not.
-
-A possible explanation for the lack of effect of glucose compared to placebo, is that both treatments may have exercised with sufficient levels of nutrients and energy, meaning that mTORC1 had sufficient nutrient availability and, thus was not inhibited by AMPK due to energy deficiency (HOPPE09; KUSNADI15).
-Indeed, mTORC1 is characterised as energy-sensitive (DENNIS01), and therefore it might have been directly affected by high vs.
-low-glucose conditions (MARIAPPAN11).
-However, previous studies suggest that mTORC1 is not sensitive enough to register levels of ATP, rather depending on signals from AMPK to assess ATP levels (KIM10).
-This would arguably make AMPK the primary regulator of the potential impact of glucose on ribosome biogenesis, and thus, glucose has no direct effect on mTORC1 signalling (REFERENCES?).
-Importantly, intramuscular glycogen content was not measured in the present study, therefore we cannot determine whether cellular energy levels in fact were equal or not, or if there is reason to assume that AMPK-dependent mTORC1 signalling was different in the two treatments.
-Potentially, such data could give higher resolution to our data.
-Along with the robust increases in proxy markers of ribosome biogenesis, the equal macronutrient intake between treatments on pairwise consecutive days and plasma glucose levels not indicating energy deficiency during RT, potent mTORC1 stimulation without AMPK-dependent inhibition may be a plausible explanation.
-Together with previous studies on this topic (MARIAPPAN11; ZHAI12), there are currently scarce observations to suggest glucose carries a direct anabolic signal.
-As such, physiological levels of plasma glucose during RT, combined with a sufficient daily macronutrient intake, seem potent enough to supply the energy needed for ribosome biogenesis.
-The lacking differences between treatments in the biological data are further supported by our observations on functional data.
-
-Overall, glucose did not affect the isometric or isokinetic peak torque at 240 d/s, or muscular recovery following the sixth and last RT session, measured as temporal changes in isometric and isokinetic peak torque.
-Generally, the mean isometric and isokinetic peak torque slightly decreased during the intervention, especially between session 4 and 6, and regained after the intervention.
-Further, there were no differences between glucose and placebo in total session volume or training intensity.
-Both exercise with glucose and placebo yielded higher total session volumes at higher percentage of 1RM at the last session compared to baseline, as expected from earlier reports on the moderate-volume RT protocol (KRIEGER09; RHEA03; SCHOENFELD17), in addition to a more recent study implementing the unilateral protocol (HAMMARSTROM20+21/22).
-Interestingly, we did observe a significant difference between treatments in mean change of isokinetic peak torque at 60 d/s before the sixth RT session, where exercise with glucose decreased peak torque 9% less compared to placebo.
-However, as none of the other measurements indicated any difference between treatments, a true effect of glucose seems unlikely.
-Taken together, these results may add to previous observations with no differences between co-ingestion of carbohydrate and protein compared to protein alone, in regard to skeletal muscular recovery following RT (KOOPMAN07; STAPLES11).
-Collectively, the functional data point towards an exhaustive, yet growth-inducing RT protocol, especially considering the biological data.
-As a side note, an interesting observation was made by another master’s thesis on this project, finding significantly lower MuRF protein levels at post with glucose treatment compared to placebo.
-This may suggest an effect of glucose on protein degradation (WILKES09), however, whether this contributes to growth or simply protein turnover has been a topic of debate (FIGUEIREDO19).
-
-Despite being a thoroughly standardised design, a limitation may be posed through lack of participants.
-We initially reach our minimum limit for statistical power, i.e.
-16 participants, however, three dropped out during the intervention, possibly impacting our power.
-Coupled with the fact that no previous study could guide our power calculations, this, of course, introduces some degree of uncertainty regarding our results.
-However, we’ve reported little to suggest any true acute effect of RT with glucose ingestion compared to placebo.
-Moreover, the study was specifically designed to investigate differences in pre-post changes within participants, and thus also quite limited to this perspective.
-Hence, the effects of time observed exclusively serve as supplementary to the efficiency of the design, without a negative control group to either confirm or refute them.
-If further investigations are to be done on this topic, these analyses should include more participants and optimally a measure of intramuscular glycogen levels.
-
-In conclusion, resistance exercise training combined with glucose ingestion does not acutely augment ribosome biogenesis following 2 weeks of moderate-volume resistance training in moderately trained young male and female adults.
-Neither does it acutely affect muscular performance in isometric and isokinetic knee extension peak torque, nor enhance post resistance training recovery within 24hrs.
-
-## Disadvantages due to Covid-19
-
-Due to restrictions regarding social distancing and lock downs, caused by the Covid-19 pandemic, we were not able to recruit and include as many participants as we planned to the study.

### 2023-02-24 ribose-paper.RMD
fig1 (fig1.1) finished

@@
 
-Sixteen healthy male and female participants (20-33yrs, Table 1) were recruited to the study through social media advertisement and word of mouth, and through completing the selection process (Fig 1).
+Sixteen healthy male and female participants (20-33yrs, Table 1) were recruited to the study through social media advertisement and word of mouth.
 The eligibility criteria were non-smokers and moderately trained (i.e.
@@
 Of the sixteen participants that commenced the intervention, three participants dropped out.
-One due to sickness and inability to resume, two participants experienced muscular discomfort connected to heavy resistance training (Fig 1).
+One due to sickness and inability to resume, two participants experienced muscular discomfort connected to heavy resistance training.
 Baseline characteristics (Tab 1) were measured using DXA (Prodigy Advance PA+302047, Lunar, San Francisco, CA, USA) on Day -1, the last day preceding the RT intervention.
@@
 Participants were randomized to commence with either glucose (GLU, 30g per bolus) or placebo (PLAC, 0g per bolus), and either dominant or non-dominant leg (Fig 1), alternating RT and supplement from one day to the other (Fig 2).
-Investigators were blinded as the supplement randomization was disclosed only to personnel not involved in any other aspect of the study, and participants were blinded by diluting boluses of glucose (30g glucose) (Glucosum monohydricum, Merck KGaA, Darmstadt, Germany) and placebo (~0.3g Stevia rebaudiana extract) (Steviosa, Soma Nordic AS, Oslo, Norway) in 300ml Fun Light juice/saft (Orkla, Oslo, Norway).
+Investigators were blinded as the supplement randomization was disclosed only to personnel not involved in any other aspect of the study.
+THe participants were blinded via diluting boluses of glucose (30g glucose) (Glucosum monohydricum, Merck KGaA, Darmstadt, Germany) and placebo (~0.3g Stevia rebaudiana extract) (Steviosa, Soma Nordic AS, Oslo, Norway) in 300ml Fun Light juice/saft (Orkla, Oslo, Norway).
 A blinded taste test revealed that the participants were not able to disclose the contents of the provided boluses (30g GLU vs.
 0g GLU).
-To ensure equal conditions during training sessions, participants exercised and tested at the same time of day, +/- 1hr with the same trainer, and recorded and repeated their daily macro nutrient intake on pairwise consecutive days (Tab 4).
+To ensure equal conditions during training sessions, participants exercised and tested at the same time of day, +/- 1hr (check this) with the same trainer, and recorded and repeated their daily macro nutrient intake on pairwise consecutive days (Tab 4).
 
@@
 The daily onset of the dietary intervention (i.e., first ingestion of PRO supplement) varied between participants, from 0600hrs to 0900hrs to allow multiple participants to complete the protocol simultaneously.
-The protein supplement was ingested as 25g Whey Protein Isolate boluses (Proteinfabrikken, Stokke, Norway), diluted in 150ml water.
+The protein supplement was ingested as 25g Whey Protein Isolate boluses (Proteinfabrikken, Stokke, Norway), diluted in 150ml of water.
 During sessions, participants were free to ingest water ad libitum.
@@
 
-Performance tests (Test, Fig 2) were performed prior to (Days -7 and -5, and T0, both legs) and during the intervention (Days 4, 5, 8, and 9), after session 5 and after finalization of the intervention (T3/T4, RT#1 leg; T4/Day 13, RT#2 leg), a total of 10 days (Fig 2).
+Strength tests (Fig 2) were performed prior to (Days -7 and -5, and -1, both legs) and during the intervention (Days 4 and 8 leg 1, 5 and 9 leg two), after session 5 and after finalization of the intervention (Day 11/12 leg 1 and 12/13 leg 2) (Fig 1).
 Maximal isometric and isokinetic knee extension torque were measured with a Humac Norm Dynamometer (CSMi, Stoughton, Massechusetts, USA).
 Individual positions were recorded and standardized from pre-intervention tests (days -7 and -5).
-Maximal isokinetic torque was measured at speeds of 60- and 240 d/s, 2x3 repetitions each, with the first set of each exercise as a sub-maximal warm-up.
-The maximal isometric torque was measured at knee-angle 60, for a maximum of 10 seconds and two repetition per test.
+Isokinetic peak torque was measured at speeds of 60- and 240 d/s, 2x3 repetitions each, with the first set of each exercise as a sub-maximal warm-up.
+Isometric peak torque was measured at knee-angle 60, for a maximum of 10 seconds and two repetition per test.
 The highest peak torque values were carried forwards to the final analyses.
 During days 4, 5, 8 and 9, humac tests were conducted one hour before RT with the leg performing RT the previous day.
-During days 11 and 12 (T3 and T4), humac tests were performed at four time points: I) 45min before RT, II) 30min after the last RT session, III) 2hrs after the last RT session, and IV) 23hrs after the last RT session.
-Test I at T4 included testing of both legs, representing 23hrs post RT session test of one leg and pre-session 6 test of the other leg.
-
-Assessment of unilateral one repetition maximum (1RM) leg press and knee extension was conducted at pre-intervention testing (days -7 and -5, Fig 2).
+During days 11 and 12, humac tests were performed at four time points: I) 45min before RT, II) 30min after the last RT session, III) 2hrs after the last RT session, and IV) 23hrs after the last RT session.
+Test I at day 12/13 included testing of both legs, representing 23hrs post RT session test of one leg and pre-session 6 test of the other leg.
+
+Assessment of unilateral one repetition maximum (1RM) leg press and knee extension was conducted at pre-intervention testing (days -7 and -5, Fig 1).
 The participants performed a general warm-up with 10min cycling on an indoor exercise bicycle and a specific warm-up (1x10, 1x6 and 1x3 repetitions, 2min breaks) before each of the tests.
@@
 Muscle biopsies were sampled from m.
-vastus lateralis using well-established procedures (HAMMARSTRØM20).
+vastus lateralis using well-established procedures [@hammarstrom.etal_2020].
 Muscle biopsy sampling was performed under local anaesthesia (Xylocaine, 10mg ml-1 with adrenaline 5μg ml−1, AstraZeneca AS, Oslo, Norway) using a 12-14-gauge needle (Universal Plus, Mermaid Medical AS, Stenløse, Denmark), operated with a spring-loaded biopsy gun.
@@
 Due to restrictions regarding social distancing and lock downs, caused by the Covid-19 pandemic, we were not able to recruit and include as many participants as we planned to the study.
+<!-- Is this necessary to include at all?
+-->

### 2023-02-28 results_work.RMD
fig text work

@@
 Further, on pairwise consecutive days, there were no significant differences in mean macro nutrient (calories, carbohydrate, fat or protein) ingestion (Tab 4), or mean change in macro nutrient ingestion (Fig 3).
-Resistance training with glucose led to significant increases in blood levels glucose and c-peptide compared to RT with placebo, before, during and after the session (Fig 1B, 120min: 38 ± 4% [*p* = `r p.120`], 135min: 31 ± 4% [*p* = `r p.135`], 150min: 32 ± 4% [*p* = `r p.150`], Fig 1C, 120min: 95 ± 10% [*p* = `r cpp.120`], 150min: 87 ± 10% [*p* = `r cpp.150`]), measured during the final RT session.
+Glucose ingestion before and after RT led to significant increases in plasma levels of glucose and c-peptide compared to RT with placebo, before, during and after the session (Fig 1B, 120min: 38 ± 4% [*p* = `r p.120`], 135min: 31 ± 4% [*p* = `r p.135`], 150min: 32 ± 4% [*p* = `r p.150`], Fig 1C, 120min: 95 ± 10% [*p* = `r cpp.120`], 150min: 87 ± 10% [*p* = `r cpp.150`]), measured during the final RT session.
 At 2hrs post-exercise, RT with glucose led to significantly lower blood glucose levels compared to RT with placebo (1B, 270min: -8 ± 4% [*p* = `r p.270`].
@@
 
-**Figure 1:** A) An overview of the experimental design with 12 days of concomitant dietary intervention and resistance training (RT), preceded by 7 days of testing.
-Familiarization = days -7 to -1 included familiarization to RT exercises (via 1RM leg press and knee extension tests) and strength tests (Humac dynamometer) at days -7 and -5.
-Baseline strength tests were conducted at day -1.
-Biopsies were taken from vastus lateralis at baseline (Day 1 leg 1, Day 2 leg 2) and after five RT sessions (Day 11 leg 1, Day 12 leg 2).
-The order in which participants performed the two intervention blocks was determined in a planned randomized fashion, where one leg was randomized to RT with GLU (glucose) and the other to RT with PLA (placebo).
-The randomization also entailed half starting with GLU (n = 8) and the other half with PLA (n = 8), in addition they where further randomized to start either with their dominant or non-dominant leg.
-Thus, 4 participants were randomized to start RT with their dominant leg and GLU supplement, and 4 with their non-dominant leg and GLU - the same was true for participtants commencing the intervention with PLA.
-From intervention start (Day 1 leg 1, Day 2 leg 2), the participants exercised in a unilateral, alternating fashion until post intervention testing (Day 11/12 leg 1, Day 12/13 leg 2).
-Three participants dropped out of the intervention, either due to reported muscular discomforts to heavy strength training or sickness in the family.
-B and C) Changes in Plasma glucose (mmol/L) and -c-peptide levels (pmol/L) during RT session six (Day 11 leg 1, Day 12 leg 2).
-Finger draws and venous blood were sampled at baseline (Day -1), as well as before, during and after session six.
-Plasma glucose was measured via finger draws, 120 (-120), 90 (-90) and 30 (-30) minutes and immediately before RT (0), during RT (15), immediately after (30) and 2hrs after RT (120).
-Levels of c-peptide were measured via venous blood samples collected at the same time points as finger draws, except for 90min before RT and during RT.
+**Figure 1:** A) An overview of the experimental design with 12 days of concomitant dietary intervention and resistance training (RT), preceded by 7 days involving familiarization.
+Between days -7 and -1, participants were familiarized to the RT exercises via 1RM leg press and knee extension testing, and strength tests (Humac Norm dynamometer, days -7 and -5).
+Biopsies were taken from vastus lateralis at baseline (Day 1 leg 1, Day 2 leg 2), and after five RT sessions (Day 11 leg 1, Day 12 leg 2).
+Through randomization, one leg was allocated to RT with GLU (glucose) and the other leg to RT with PLA (placebo).
+The randomization also entailed whether participants initiated the intervention with their dominant or non-dominant leg, and whether they initiated with RT+GLU or RT+PLA.
+From baseline, the participants exercised in a unilateral, alternating fashion, until post intervention testing (Day 11/12 leg 1, Day 12/13 leg 2).
+Three participants dropped out of the intervention, either due to reported muscular discomforts of heavy RT, or sickness in their family.
+B and C) Changes in plasma glucose (mmol/L) and c-peptide levels (pmol/L) during post testing (Day 11 leg 1, Day 12 leg 2).
+Glucose levels in blood was measured via finger draws 120 (-120), 90 (-90), 30 (-30) minutes and immediately before RT (0), during RT (15), immediately after (30) and 2hrs after RT (120).
+C-peptide levels were measured simultaneously to these finger draws, except for 90min before and during RT.
+Finger draws and venous blood were sampled at baseline (Day 1) and post (Day 11 leg 1, Day 12 leg 2).
 D) Changes in muscular performance/recovery measured as isometric and isokinetic peak torque (60 and 240 d/s) via Humac Norm Dynamometer, conducted at baseline (Day -1), during the intervention (Day 4 and 8 leg 1, Day 5 and 9 leg 2), after five RT sessions/before the 6th session (Day 11 leg 1, Day 12 leg 2), as well as 30min, 2hrs and 23hrs after the 6th RT session (Day 11/12 leg 1, Day 12/13 leg 2).
-Values are presented as estimated marginal means ± 95% CI.
+Values are presented as changes in estimated marginal means ± 95% CI.
 * = *p* < 0.05 between groups.
@@
 
-Resistance training with glucose did not induce a higher accumulation of total RNA (fig 2A, *p* = `r trdiffp`), or rRNA  (fig 2B-2F *p* > 0.05), compared to RT with placebo, measured as mean change from baseline to post.
+Resistance training with glucose did not induce a higher accumulation of total RNA (fig 2A, *p* = `r trdiffp`) or rRNA  (fig 2B, 47S: *p* = `r rr47diffp`, 18S: *p* = `r rr18diffp`, 28S: *p* = `r rr28diffp`, 5.8S: *p* = `r rr58diffp`, 5S: *p* = `r rr5diffp`) compared to RT with placebo, measured as mean change from baseline to post.
 From baseline to post-intervention, there was a mean increase in levels of total RNA by 26 and 22% after RT with glucose and placebo, respectively (*p* < 0.05, compared to baseline).
 A robust accumulation was also observed in ribosomal RNA expression in both RT with glucose and placebo, with mean increases between 34-43% (GLU) and 33-41% (PLAC) (*p* < 0.05, compared to baseline) in the four rRNA’s.
-The expression of 47S pre-rRNA increased by 37 and 59% with glucose and placebo respectively, where only RT with placebo increased significantly from baseline to post (fig 45, *p* = `r rr47tdiffp`).
+The expression of in the 47S pre-rRNA increased by 37 and 59% with glucose and placebo respectively, where only RT with placebo increased significantly from baseline to post (fig 45, *p* = `r rr47tdiffp`).
 
@@
 A) Total RNA, B) 47S pre-rRNA, 18S rRNA, 28S rRNA, 5.8S rRNA, 5S rRNA.
-Baseline = Training day 1/2, Post = Training day 11/12.
-Total RNA and rRNA were analyzed in duplicates, with two duplicates per biopsy (two biopsies per time point).
+Baseline = Day 1 leg 1/ Day 2 leg 2, Post = Day 11 leg 1, Day 12 leg 2.
+Total RNA and rRNA were analyzed in duplicates, with two duplicates per biopsy (two biopsies per time point), and normalized to ng x mg wet muscle weight for total RNA and external reference gene (Lambda) for rRNA.
 Total RNA and rRNA changes were calculated as log-fold change score per mg wet muscle weight.
@@
 Values are estimated marginal means fold change per leg per supplement ± 95% CI.
-p = between groups statistic, glucose compared to placebo.
 Glucose n = 13, placebo n = 13.

### 2023-02-28 figures/figure3.R

@@ -185,6 +185,7 @@ corr.fig <- joined.dat %>%
         axis.text = element_text(size = 7),
         legend.title = element_text(size = 6),
         legend.text = element_text(size = 5),
+        legend.key = element_rect(fill = "white"),
         axis.title.y = element_markdown(size = 7))
 
 # Cowplot for gathering figures

### 2023-02-28 ext_method.RMD
added RMD for extended methods

---
title: "Ext. suppl."
author: "Kristian Lian"
date: "2023-02-28"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

Notes for extended methods supplementary

- UBF/RNA correlation: Make a new model to see if data point affects the current model

### 2023-02-28 results_work.RMD
added RMD for extended methods

@@
 
-###### **Figure 3:** Changes in levels of protein.
-A) Representative total protein stain blot, B) AU fold change in UBF, c-Myc and rpS6 + representative antibody stain images of the respective proteins.
-Baseline = Training day 1/2, Post = Training day 11/12.
-Protein samples were analyzed in duplicates, with two duplicates per biopsy (two biopsies per time point).
-The duplicates were loaded on different gels, in an inverted order.
-Change in protein levels were calculated as log-fold change score normalised by pools (pools of all protein samples).
+###### **Figure 3:** Changes in levels of protein. A) Linear relationship between total RNA (ng x mg) and UBF levels (SD units).
+
+B) Representative total protein stain blot, both duplicates from one participant (115).
+C) Fold change in UBF, c-Myc and rpS6 (AU) + representative antibody stains of the respective proteins.
+Baseline = Day 1 leg 1, Day 2 leg 2, post = Day 11 leg 1, Day 12 leg 2.
+GLU = glucose, PLA = placebo.
+Protein samples were analyzed in two duplicates per biopsy per time point, loaded on separate gels in an inverted order, e.g.
+from gel 3 to 6, as shown by the duplicates (1 and 2).
+Change in protein levels were calculated as log-fold change scores normalized by pools (pool of all protein samples per gel).
 Mean change scores of the duplicates were calculated before modelling and transformed to the log-scale.
-Values are estimated marginal means fold change per leg per supplement ± 95% CI.
+Values are reverse transformed estimated marginal means fold change per leg per supplement ± 95% CI.
 * = significant between groups statistic, glucose compared to placebo.

### 2023-02-28 figures/figure4.R

@@ -38,9 +38,9 @@ saveRDS(joined.dat, "./data/data-gen/rna.prot.RDS")
 
 library(nlme)
 
-m <- lme(mean.rna ~ sd.ubf + time , 
+m <- lme(mean.rna ~ sd.ubf + time, 
          random = list(subject = ~ 1), 
-         data = joined.dat, 
+         data = filter(joined.dat), 
          na.action = na.omit)
 
 summary(m)

### 2023-02-28 results_work.RMD
updates

@@
 
-###### **Figure 3:** Changes in levels of protein. A) Linear relationship between total RNA (ng x mg) and UBF levels (SD units).
+###### **Figure 3:** Changes in levels of protein. A) Linear relationship between total RNA (ng x mg) and UBF levels (SD units), with time added as a covariate in the model.
+
+(We must add time as a covariate, as both total RNA and UBF are expected to increase with "time" in this intervention, due to RT.
+Therefore, adding time as a covariate allows us to model the relationship between RNA and UBF, without variations of time interfering)
 

### 2023-03-01 results_work.RMD
updates before presentation

@@
 
+```{r strength model}
+## Scripts:
+# - Strength.index.R
+
+## Data
+str.change <- readRDS("./data/data-gen/humac/str.change.RDS") # Untransfromed data
+str.lchange <- readRDS("./data/data-gen/humac/str.lchange.RDS") # Log-transfromed data
+
+## Linear mixed effects model
+
+# Untransformed model
+str.mod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
+                         data = str.change)
+str.res <- plot(str.mod)
+str.sum <- summary(str.mod)
+
+# Log-transformed model
+str.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
+                        data = str.lchange)
+str.lres <- plot(str.lmod)
+str.lsum <- summary(str.lmod)
+
+## Emmeans
+# Untransformed
+str.emm <- confint(emmeans(str.mod, specs = ~"supplement|time")) %>%
+  data.frame()
+
+# Log-transformed
+str.lemm <- confint(emmeans(str.lmod, specs = ~"supplement|time")) %>%
+  data.frame()
+
+# For inline plotting
+strcoef <- coef(summary(str.mod))
+pdiff.c4 <- round(strcoef[10,5], 2)
+
+```
+
 At baseline, there were no significant differences between training with glucose or placebo in muscle mass, isometric or isokinetic peak torque, total session volume, training intensity, total RNA content, or rRNA expression (Tab 3).
-Further, on pairwise consecutive days, there were no significant differences in mean macro nutrient (calories, carbohydrate, fat or protein) ingestion (Tab 4), or mean change in macro nutrient ingestion (Fig 3).
+Further, on pairwise consecutive days, there were no significant differences in mean macro nutrient (calories, carbohydrate, fat or protein) intake, or mean change in macro nutrient ingestion (Tab 4).
+
 Glucose ingestion before and after RT led to significant increases in plasma levels of glucose and c-peptide compared to RT with placebo, before, during and after the session (Fig 1B, 120min: 38 ± 4% [*p* = `r p.120`], 135min: 31 ± 4% [*p* = `r p.135`], 150min: 32 ± 4% [*p* = `r p.150`], Fig 1C, 120min: 95 ± 10% [*p* = `r cpp.120`], 150min: 87 ± 10% [*p* = `r cpp.150`]), measured during the final RT session.
 At 2hrs post-exercise, RT with glucose led to significantly lower blood glucose levels compared to RT with placebo (1B, 270min: -8 ± 4% [*p* = `r p.270`].
-
-FIG1 HERE
+There were no significant difference in change of skeletal muscle strength, at time points post 2 RT sessions, post 4 RT sessions, 30min after 6th session, 2hrs after 6th session and 23hrs after 6th session (*p* > 0.05).
+After 5 RT sessions (Fig 1D, Post 5RT), there was a significant difference in strength between glucose and placebo,  where placebo decreased strength 7% more than glucose (*p* = `r pdiff.c4`).
 
 **Figure 1:** A) An overview of the experimental design with 12 days of concomitant dietary intervention and resistance training (RT), preceded by 7 days involving familiarization.
-Between days -7 and -1, participants were familiarized to the RT exercises via 1RM leg press and knee extension testing, and strength tests (Humac Norm dynamometer, days -7 and -5).
-Biopsies were taken from vastus lateralis at baseline (Day 1 leg 1, Day 2 leg 2), and after five RT sessions (Day 11 leg 1, Day 12 leg 2).
-Through randomization, one leg was allocated to RT with GLU (glucose) and the other leg to RT with PLA (placebo).
-The randomization also entailed whether participants initiated the intervention with their dominant or non-dominant leg, and whether they initiated with RT+GLU or RT+PLA.
-From baseline, the participants exercised in a unilateral, alternating fashion, until post intervention testing (Day 11/12 leg 1, Day 12/13 leg 2).
+Between days -7 and -1, participants were familiarized to the RT exercises via 1RM leg press and knee extension testing, and to the strength tests via Humac Norm dynamometer (days -7 and -5).
+Before baseline testing, the participants were randomly allocated to exercise one leg with GLU (glucose) and the other with PLA (placebo), in a unilateral, alternating fashion.
+Further, non-dominant/dominant + GLU/PLA, and onset with GLU or PLA was also randomized, i.e.
+the figure illustrates an example where the participant was randomized to start RT with GLU.
+Biopsies were taken from m.
+vastus lateralis at baseline (Day 1 leg 1, Day 2 leg 2), and after five RT sessions (Day 11 leg 1, Day 12 leg 2).
 Three participants dropped out of the intervention, either due to reported muscular discomforts of heavy RT, or sickness in their family.
@@
 Finger draws and venous blood were sampled at baseline (Day 1) and post (Day 11 leg 1, Day 12 leg 2).
-D) Changes in muscular performance/recovery measured as isometric and isokinetic peak torque (60 and 240 d/s) via Humac Norm Dynamometer, conducted at baseline (Day -1), during the intervention (Day 4 and 8 leg 1, Day 5 and 9 leg 2), after five RT sessions/before the 6th session (Day 11 leg 1, Day 12 leg 2), as well as 30min, 2hrs and 23hrs after the 6th RT session (Day 11/12 leg 1, Day 12/13 leg 2).
-Values are presented as changes in estimated marginal means ± 95% CI.
-* = *p* < 0.05 between groups.
-Glucose n = 13, placebo n = 13.
+D) Changes in muscular performance/recovery measured as isometric and isokinetic peak torque (60 and 240 d/s) via Humac Norm Dynamometer, conducted at baseline (A: Day -1), after 2 and 4 RT sessions (A: Day 4 and 8 leg 1, Day 5 and 9 leg 2), after five RT sessions/before the 6th session (Day 11 leg 1, Day 12 leg 2), as well as 30min, 2hrs and 23hrs after the 6th RT session (A: Day 11/12 leg 1, Day 12/13 leg 2).
+The index was calculated by normalizing individual peak torque to the highest peak torque at each respective speed, and then summarised and used in change score calculations.
+
+Values are presented as changes in estimated marginal means ± 95% CI. * = *p* < 0.05 between groups. Glucose n = 13, placebo n = 13.
 
@@
 # - western.import.R
-# - western.change.R
+# - western.cleanup.R
+# - western.analysis.R
 
 # Data
+prot.dat <- readRDS("./data/data-gen/protein/prot.datc.RDS") # from the script: western.cleanup.R
 
-prot.dat <- readRDS("./data/data-gen/protein/prot.dat.RDS") # cleaned up and quantified western data
-prot.out <- readRDS("./data/data-gen/protein/protein_complete.RDS") # Cleaned up and quantified western data, checked for outliers
-prot.change <- readRDS("./data/data-gen/protein/prot.change.RDS") # change-score data
+### Modelling target signal per supplement 
+# These models try to explain the normalised signal by time, supplement and time+supplement (?). The random = list command creates an individual 
+# intercept for the included (e.g., subject, gel.sample).
 
-# By target
+## c-Myc model
 
-cmyc <- prot.change %>%
-  filter(target == "cmyc")
+# Untransformed
+cmyc.mod <- lme(norm.sign ~ time + supplement + time:supplement, 
+                random = list(subject = ~ 1),
+                data = filter(prot.dat, target == "cmyc"))
 
-ubf <- prot.change %>%
-  filter(target == "ubf",
-         subject != "103") # Subject 103 has missing values
+cmyc.sum <- summary(cmyc.mod)
+cmyc.res <- plot(cmyc.mod, resid(., type = "p") ~ fitted(.))
 
-rps6 <- prot.change %>%
-  filter(target == "rps6")
+# Log-transformed
+cmyc.lmod <- lme(log(norm.sign) ~ time + supplement + time:supplement, 
+                 random = list(subject = ~ 1),
+                 data = filter(prot.dat, target == "cmyc"))
 
-# c-Myc change
+cmyc.lsum <- summary(cmyc.lmod)
+cmyc.lres <- plot(cmyc.lmod, resid(., type = "p") ~ fitted(.))
 
-m1 <- lme(change ~ pre + supplement,
+## UBF model
+# Untransformed
+ubf.mod <- lme(norm.sign ~ time + supplement + time:supplement, 
                random = list(subject = ~ 1),
-            data = cmyc) # same model as in "western import", but with change scores of norm.sign
+               data = filter(prot.dat, target == "ubf"))
 
-#plot(m1)
+ubf.sum <- summary(ubf.mod)
+ubf.res <- plot(ubf.mod, resid(., type = "p") ~ fitted(.))
 
-sum.m1 <- summary(m1)
+# Log-transformed
+ubf.lmod <- lme(log(norm.sign) ~ time + supplement + time:supplement, 
+                random = list(subject = ~ 1),
+                data = filter(prot.dat, target == "ubf"))
 
-cmyc.emm <- emmeans(m1, specs = ~"supplement") %>%
-  data.frame
+ubf.lsum <- summary(ubf.lmod)
+ubf.lres <- plot(ubf.lmod, resid(., type = "p") ~ fitted(.))
 
-# UBF change
+# RPS6 model
+# Untransformed
+rps6.mod <- lme(norm.sign ~ time + supplement + time:supplement, 
+                 random = list(subject = ~ 1),
+                 data = filter(prot.dat, target == "rps6"))
 
-m2 <- lme(change ~ pre + supplement,
+rps6.sum <- summary(rps6.mod)
+rps6.res <- plot(rps6.mod, resid(., type = "p") ~ fitted(.))
+
+# Log-transformed
+rps6.lmod <- lme(log(norm.sign) ~ time + supplement + time:supplement, 
           random = list(subject = ~ 1),
-            data = ubf) # same model as in "western import", but with change (i.e., change in norm.sign) instead of norm.sign
+          data = filter(prot.dat, target == "rps6"))
 
-#plot(m2)
+rps6.lsum <- summary(rps6.lmod)
+rps6.lres <- plot(rps6.lmod, resid(., type = "p") ~ fitted(.))
 
-sum.m2 <- summary(m2)
+### Getting emmeans from the models
 
-ubf.emm <- emmeans(m2, specs = ~"supplement") %>%
+## c-Myc emmeans
+# Untransformed
+cmyc.emm <- emmeans(cmyc.mod, specs = ~ time|supplement) %>%
   data.frame()
 
-# rps6 change
+# Log-transformed
+cmyc.lemm <- emmeans(cmyc.lmod, specs = ~ time|supplement) %>%
+  data.frame()
 
-m3 <- lme(change ~ pre + supplement,
-            random = list(subject = ~ 1),
-            data = rps6) # same model as in "western import", but with change (i.e., change in norm.sign) instead of norm.sign
+## UBF emmeans
+# Untransformed
+ubf.emm <- emmeans(ubf.mod, specs = ~ time|supplement) %>%
+  data.frame()
 
-#plot(m3)
+# Log-transformed
+ubf.lemm <- emmeans(ubf.lmod, specs = ~ time|supplement) %>%
+  data.frame()
 
-sum.m3 <- summary(m3)
+## RPS6
+# Untransformed
+rps6.emm <- emmeans(rps6.mod, specs = ~ time|supplement) %>%
+  data.frame()
 
-rps6.emm <- emmeans(m3, specs = ~"supplement") %>%
+# Log-transformed
+rps6.lemm <- emmeans(rps6.lmod, specs = ~ time|supplement) %>%
   data.frame()
@@
 ## Diff cmyc
-ccoef <- coef(summary(m1))
-cdiff <- round(ccoef[3,5], 3)
-cdiffse <- round(ccoef[3,2], 3)
+
+ccoef <- coef(summary(cmyc.lmod))
+cdiff <- round(ccoef[4,5], 3)
+cpla <- round(ccoef[2,5], 20)
 
 ## Diff UBF
-ucoef <- coef(summary(m2))
-udiff <- round(ucoef[3,5], 3)
-udiffse <- round(ucoef[3,2], 3)
+ucoef <- coef(summary(ubf.lmod))
+udiff <- round(ucoef[4,5], 3)
 
 ## Diff rps6
-rpcoef <- coef(summary(m3))
-rpdiff <- round(rpcoef[3,5], 3)
-rpdiffse <- round(rpcoef[3,2], 3)
+rpcoef <- coef(summary(rps6.lmod))
+rpdiff <- round(rpcoef[4,5], 3)
 
@@
 
-m <- lme(log(mean.rna) ~ mean.sign + time , 
+library(nlme)
+
+corr.mod <- lme(mean.rna ~ mean.sign + time , 
          random = list(subject = ~ 1), 
@@
 
-summary(m)
-plot(m)
+corr.sum <- summary(corr.mod)
+corr.res <- plot(corr.mod)
+
+corr.lmod <- lme(log(mean.rna) ~ log(mean.sign) + time , 
+         random = list(subject = ~ 1), 
+          data = joined.dat, 
+         na.action = na.omit)
+
+corr.lsum <- summary(corr.lmod)
+corr.lres <- plot(corr.lmod)
+
+corrcoef <- coef(summary(corr.mod))
+corrp <- round(corrcoef[2,5], 3)
 ```
 
-Interestingly, RT with placebo led to significantly higher change in levels of rpS6 and UBF protein compared to RT with glucose, with no difference in levels of c-Myc (Fig 3A rpS6: *p* = `r rpdiff`, 3C UBF: *p* = `r udiff`, 3B c-Myc: *p* = `r cdiff`).
-Levels of rpS6 change by 40% in placebo and 7% in glucose (Fig 3A), UBF levels changed by 68% in placebo and 26% in glucose (Fig 3C), and c-Myc levels changed by 110% in placebo and 59% in glucose (Fig 3B).
+C-myc levels increased 109% after RT with placebo, compared to an increase of 58% after RT with glucose, which was a significant difference (Fig 3A: *p* = `r cdiff`).
+Levels of UBF and rpS6 increased 42 and 33% more in placebo compared to glucose, however, this was not a significant difference (Fig 3A: UBF: *p* = `r udiff`, rpS6: *p* = `r rpdiff`).
+Compared to baseline, both RT with glucose and RT with placebo increased significantly () Further, we observed a linear relationship between levels of UBF and accumulation of total RNA, where an increase of 1 SD unit of UBF equated to 13% increase in total RNA (*p* = `r corrp`).
 
@@
 
-###### **Figure 3:** Changes in levels of protein. A) Linear relationship between total RNA (ng x mg) and UBF levels (SD units), with time added as a covariate in the model.
+###### **Figure 3:** Changes in levels of protein.
 
-(We must add time as a covariate, as both total RNA and UBF are expected to increase with "time" in this intervention, due to RT.
-Therefore, adding time as a covariate allows us to model the relationship between RNA and UBF, without variations of time interfering)
-
-B) Representative total protein stain blot, both duplicates from one participant (115).
-C) Fold change in UBF, c-Myc and rpS6 (AU) + representative antibody stains of the respective proteins.
+A) Mean UBF, c-Myc and rpS6 (AU) at baseline and post + representative antibody stains of the respective proteins.
 Baseline = Day 1 leg 1, Day 2 leg 2, post = Day 11 leg 1, Day 12 leg 2.
@@
 Glucose n = 13, placebo n = 13.
+B) Representative total protein stain blot.
+C) Linear relationship between total RNA (ng x mg) and UBF levels (SD units), with time added as a covariate.
+Total RNA was normalized by wet muscle weight, and UBF was normalized by pools per gel, and total protein per lane factor.
+Values are presented as log-transformed means.
 
-Further, we observed a linear relationship between levels of UBF and accumulation of total RNA, where ...
+#MISSING: CHECK % INCREASE OF TOTAL RNA PER 1 SD INCREASE OF UBF (MAY HAVE TO LOG-TRANSFORM)
 
-FIG4 HERE
-
-######**Figure 4:** Relationship between levels of total RNA and UBF.
-Total RNA was normalized by wet muscle weight, and UBF levels were normalized by pools and total protein per lane.
-Values are presented as log-transformed means.
+(We must add time as a covariate, as both total RNA and UBF are expected to increase with "time" in this intervention, due to RT.
+Therefore, adding time as a covariate allows us to model the relationship between RNA and UBF, without variations of time interfering)

### 2023-03-01 R/rrna.change.R

@@ -112,7 +112,7 @@ change.58 <- rrna5.8 %>%
 
 saveRDS(change.58, "./data/data-gen/rna/lchange.58.RDS")
 
-base.5.8 <- change.5.8 %>%
+base.5.8 <- change.58 %>%
   select(subject, supplement, Pre) %>%
   pivot_wider(names_from = supplement,
               values_from = Pre) %>%

### 2023-03-01 R/strength.index.R

@@ -79,16 +79,43 @@ str.change <- strength.index %>%
 
 saveRDS(str.change, "./data/data-gen/humac/str.change.RDS")
 
+str.lchange <- strength.index %>%
+  pivot_wider(names_from = time,
+              values_from = strength) %>%
+  mutate(change.2 = log(test2) - log(baseline),
+         change.3 = log(test3) - log(baseline),
+         change.4 = log(test4) - log(baseline),
+         change.5 = log(test5) - log(baseline),
+         change.6 = log(test6) - log(baseline),
+         change.7 = log(test7) - log(baseline),
+         baseline = baseline - mean(baseline, na.rm = TRUE)) %>%
+  select(subject, supplement, baseline, change.2:change.7) %>%
+  pivot_longer(names_to = "time",
+               values_to = "change",
+               cols = c(change.2:change.7)) %>%
+  print()
+
+saveRDS(str.lchange, "./data/data-gen/humac/str.lchange.RDS")
+
+
 
 ## Change model
 
 m.str <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
                          data = str.change)
+
+m.lstr <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
+                        data = str.lchange)
 plot(m.str)
+plot(m.lstr)
 
 summary(m.str)
+summary(m.lstr)
 
 emm.str <- confint(emmeans(m.str, specs = ~"supplement|time")) %>%
+  data.frame() 
+
+emm.str <- confint(emmeans(m.lstr, specs = ~"supplement|time")) %>%
   data.frame() 
 
 saveRDS(emm.str, "./data/data-gen/humac/emm.str.RDS")

### 2023-03-01 R/western.analysis.R

@@ -1,0 +1,168 @@
+##### Western analysis
+#
+#
+## Purpose: The purpose of this script is to analyse the cleaned up western data
+#
+# Author: KL/DH
+#
+# Packages
+library(tidyverse); library(nlme); library(emmeans)
+#
+# Data
+prot.dat <- readRDS("./data/data-gen/protein/prot.datc.RDS") # from the script: western.cleanup.R
+
+
+### Modelling target signal per supplement 
+# These models try to explain the normalised signal by time, supplement and time+supplement (?). The random = list command creates an individual 
+# intercept for the included (e.g., subject, gel.sample). 
+
+## c-Myc model
+
+# Untransformed
+cmyc.mod <- lme(norm.sign ~ time + supplement + time:supplement, 
+                random = list(subject = ~ 1),
+                data = filter(prot.dat, target == "cmyc"))
+
+summary(cmyc.mod)
+plot(cmyc.mod, resid(., type = "p") ~ fitted(.))
+
+
+# Log-transformed
+cmyc.lmod <- lme(log(norm.sign) ~ time + supplement + time:supplement, 
+                 random = list(subject = ~ 1),
+                 data = filter(prot.dat, target == "cmyc"))
+
+summary(cmyc.lmod)
+plot(cmyc.lmod, resid(., type = "p") ~ fitted(.))
+
+
+
+
+## UBF model
+# Untransformed
+ubf.mod <- lme(norm.sign ~ time + supplement + time:supplement, 
+               random = list(subject = ~ 1),
+               data = filter(prot.dat, target == "ubf"))
+
+
+summary(ubf.mod)
+plot(ubf.mod, resid(., type = "p") ~ fitted(.))
+
+# Log-transformed
+ubf.lmod <- lme(log(norm.sign) ~ time + supplement + time:supplement, 
+                random = list(subject = ~ 1),
+                data = filter(prot.dat, target == "ubf"))
+
+
+summary(ubf.lmod)
+plot(ubf.lmod, resid(., type = "p") ~ fitted(.))
+
+
+# RPS6 model
+# Untransformed
+rps6.mod <- lme(norm.sign ~ time + supplement + time:supplement, 
[…]

### 2023-03-01 R/western.cleanup.R

@@ -7,7 +7,7 @@ library(tidyverse); library(nlme); library(emmeans)
 
 # Data
 
-prot.dat <- readRDS("./data/data-gen/protein/prot.dat.RDS")
+prot.dat <- readRDS("./data/data-gen/protein/prot.dat.RDS") # from the script: western.import.R
 
 
 ## Visualising individual changes with ggplot

### 2023-03-01 figures/figure3.R

@@ -9,11 +9,11 @@ library(ggpubr); library(nlme); library(ggtext)
 #
 # Data
 
-cmyc <- readRDS("./data/data-gen/protein/cmyc.change.RDS")
+cmyc <- readRDS("./data/data-gen/protein/cmyc.lemm.RDS")
 
-ubf <- readRDS("./data/data-gen/protein/ubf.change.RDS")
+ubf <- readRDS("./data/data-gen/protein/ubf.lemm.RDS")
 
-rps6 <- readRDS("./data/data-gen/protein/rps6.change.RDS")
+rps6 <- readRDS("./data/data-gen/protein/rps6.lemm.RDS")
 
 joined.dat <- readRDS("./data/data-gen/rna.prot.RDS")
 
@@ -54,19 +54,20 @@ plot_theme <- theme(panel.grid.major = element_blank(),
 ## cmyc fig
 
 cmyc.plot <- cmyc %>%
-  mutate(time = "post") %>%
-  add_row(supplement = c("PLACEBO", "GLUCOSE"), time = "pre", emmean = 0, lower.CL = 0, upper.CL = 0) %>%
+ # mutate(time = "post") %>%
+ # add_row(supplement = c("PLACEBO", "GLUCOSE"), time = "pre", emmean = 0, lower.CL = 0, upper.CL = 0) %>%
   mutate(time = factor(time, levels = c("pre", "post"), labels = c("Baseline", "Post"))) %>%
   ggplot(aes(time, exp(emmean), group = supplement, fill = supplement)) +
+  annotate("text", x = "Post", y = 1.6, label = "*", size = 3) +
   geom_errorbar(aes(ymin = exp(lower.CL), ymax = exp(upper.CL)),
                 width = 0.1,
                 position = position_dodge(width = 0.2)) +
   geom_line(position = position_dodge(width = 0.2)) +
   geom_point(shape = 21, size = 2, position = position_dodge(width = 0.2)) +
-  scale_fill_manual(values = colors[c(1,4)]) +
-  scale_y_continuous(breaks = c(1,2,3,4),
-                   labels = c("1.0", "2.0", "3.0", "4.0"))+
-  labs(x = "Time", y = "c-Myc AU \n(fold change)\n", fill = "Supplement") +
+  scale_fill_manual(values = colors[c(4,1)]) +
+  #scale_y_continuous(breaks = c(1,2,3,4),
[…]

### 2023-03-07 disp.Rmd
updates

@@
 ## Intro
+- To do: Check remaining comments
 
@@
 
+- To do: Read through - anything needs changing?
+
 ## Results
@@
 Baseline etc.
-- Blood data
-- Dietary data
+- Fig1 + tabs?
+- Dietary data comments
 
 ### Markers of ribosome biogenesis?
-- Total RNA? 
-- Protein?
-- UBF/total RNA?
 
-### Muscular performance and recovery?
+#### RNA and rRNA
+- Fig2
+
+#### Protein
+- Fig3
+
+- To do: Check figure text, implement tabs and figures
 
 ## Discussion
+
+Main topics:
+- No difference between glu and plac in main outcomes - thoughts?
+  - No additional effect of glu, as long as daily macro intake is equal
+  - No direct effect of glu on anabolic signalling, i.e., glucose per se does not seem to induce RT responses 
+  - Based on findings in litterature, likely still important as a contributor to the cells energy pool, thus indirectly
+    affecting ribosome biogenesis
+
+- Is there any coherence between strength diff post session 5 and cmyc diff?
+  - I.e., post 5 data indicate that placebo experiences more fatigue, does this cause higher c-Myc levels, since c-Myc is a           general transcription factor responding to stress?
+  - There were no differences in total session volumes between glucose and placebo - how could placebo then experience more stress     than glucose?
+  - Does glucose, e.g.
+via increasing energy availability, reduce metabolic stress?
+If so, is that a good thing for adaptations?
+If there is less stress, is there less stimulation to adapt to?
+Mechanical strain vs metabolic stress?
+  - "Bifacial recovery"?
+
+- UBF/RNA: 
+  - Ref to Hammarström 21, similar observations
+  - One of the first studies to find this relationship in vivo in human muscle cells 
+  - The same was not true for c-Myc and rpS6, further strengthening the importance of UBF in ribosome biogenesis compared to c-Myc
+(We must add time as a covariate, as both total RNA and UBF are expected to increase with "time" in this intervention, due to RT.
+Therefore, adding time as a covariate allows us to model the relationship between RNA and UBF, without variations of time interfering)
+
+- Any perspectives to pick up from the thesis?

### 2023-03-07 results_work.RMD
updates

@@
 ```{r packages}
-library(tidyverse);library(readxl);library(lme4);library(emmeans);library(dplyr);library(cowplot);library(knitr);library(broom);library(rmarkdown);library(tinytex);library(tidyr);library(arsenal);library(magrittr);library(dabestr);library(nlme); library(pbkrtest);library(ggpubr)
+library(tidyverse);library(readxl);library(lme4);library(emmeans);library(dplyr);library(cowplot);library(knitr);library(broom);library(rmarkdown);library(tinytex);library(tidyr);library(arsenal);library(magrittr);library(dabestr);library(nlme); library(pbkrtest);library(ggpubr); library(lmer)
 ```
@@
 
-At baseline, there were no significant differences between training with glucose or placebo in muscle mass, isometric or isokinetic peak torque, total session volume, training intensity, total RNA content, or rRNA expression (Tab 3).
-Further, on pairwise consecutive days, there were no significant differences in mean macro nutrient (calories, carbohydrate, fat or protein) intake, or mean change in macro nutrient ingestion (Tab 4).
+```{r nutrition model}
+# Scripts
+# - nut.change.R
 
-Glucose ingestion before and after RT led to significant increases in plasma levels of glucose and c-peptide compared to RT with placebo, before, during and after the session (Fig 1B, 120min: 38 ± 4% [*p* = `r p.120`], 135min: 31 ± 4% [*p* = `r p.135`], 150min: 32 ± 4% [*p* = `r p.150`], Fig 1C, 120min: 95 ± 10% [*p* = `r cpp.120`], 150min: 87 ± 10% [*p* = `r cpp.150`]), measured during the final RT session.
-At 2hrs post-exercise, RT with glucose led to significantly lower blood glucose levels compared to RT with placebo (1B, 270min: -8 ± 4% [*p* = `r p.270`].
-There were no significant difference in change of skeletal muscle strength, at time points post 2 RT sessions, post 4 RT sessions, 30min after 6th session, 2hrs after 6th session and 23hrs after 6th session (*p* > 0.05).
+# Data
+pro.change <- readRDS("./data/data-gen/nutrition/pro.change.RDS")
+fat.change <- readRDS("./data/data-gen/nutrition/fat.change.RDS")
+cho.change <- readRDS("./data/data-gen/nutrition/cho.change.RDS")
+cal.change <- readRDS("./data/data-gen/nutrition/cal.change.RDS")
+propr.change <- readRDS("./data/data-gen/nutrition/propr.change.RDS")
+
+## Models
+# Protein
+pro.mod <- lmerTest::lmer(change ~ 0 + d1 + time + supplement:time + (1|subject),
+                     data = pro.change)
+
+pro.res <- plot(pro.mod)
+pro.sum <- summary(pro.mod)
+
+# Fat
+fat.mod <- lmerTest::lmer(change ~ 0 + d1 + time + supplement:time + (1|subject),
+                     data = fat.change)
+
+fat.res <- plot(fat.mod)
+fat.sum <- summary(fat.mod)
+
+# Carbohydrates
+cho.mod <- lmerTest::lmer(change ~ 0 + d1 + time + supplement:time + (1|subject),
+                     data = cho.change)
+
+cho.res <- plot(cho.mod)
+cho.sum <- summary(cho.mod)
+
+# Calories
+cal.mod <- lmerTest::lmer(change ~ 0 + d1 + time + supplement:time + (1|subject),
+                     data = cal.change)
+
+cal.res <- plot(cal.mod)
+cal.sum <- summary(cal.mod)
+
+# Protein/kg
+propr.mod <- lmerTest::lmer(change ~ 0 + d1 + time + supplement:time + (1|subject),
+                     data = propr.change)
+
+propr.res <- plot(propr.mod)
+propr.sum <- summary(propr.mod)
+
+## Emmeans
+
+# Protein
+protein.emm <- confint(emmeans(pro.mod, specs = ~"supplement|time")) %>%
+  data.frame()
+
+# Fat
+fat.emm <- confint(emmeans(fat.mod, specs = ~"supplement|time")) %>%
+  data.frame()
+
+# Carbohydrates
+cho.emm <- confint(emmeans(cho.mod, specs = ~"supplement|time")) %>%
+  data.frame()
+
+# Calories
+cal.emm <- confint(emmeans(cal.mod, specs = ~"supplement|time")) %>%
+  data.frame()
+
+# Protein/kg
+propr.emm <- confint(emmeans(propr.mod, specs = ~"supplement|time")) %>%
+  data.frame()
+
+```
+
+At baseline, there were no significant differences between training with glucose or placebo in muscle mass, isometric or isokinetic peak torque (Tab 1) or total session volume (Tab 2).
+Further, on pairwise consecutive days, there were no significant differences in mean macro nutrient (calories, carbohydrate, fat or protein) intake.
+
+###### **Table 1:**  Baseline descriptives of legs allocated to RT with glucose or RT.
+LM = Lean mass, Strength = strength index, calculated from isometric and isokinetic peak torque, Iso 0 PT = Isometric (0 d/s) peak torque, Iso 60 PT = Isokinetic peak torque at 60 d/s, Iso 240 PT = Isokinetic peak torque at 240 d/s.
+![](figures/desctab.png)
+Glucose ingestion before and after RT led to significant increases in plasma levels of glucose and c-peptide before, during and after the session, compared to RT with placebo, (Fig 1B, 120min: 38 ± 4% [*p* = `r p.120`], 135min: 31 ± 4% [*p* = `r p.135`], 150min: 32 ± 4% [*p* = `r p.150`], Fig 1C, 120min: 95 ± 10% [*p* = `r cpp.120`], 150min: 87 ± 10% [*p* = `r cpp.150`]), measured during the sixth RT session.
+At 2hrs post-exercise, RT with glucose led to significantly lower plasma glucose levels compared to RT with placebo (1B, 270min: -8 ± 4% [*p* = `r p.270`].
+
+######**Table 2:** Mean total session volume and mean change in total session volume from session 1 to session 6. 
+![](figures/voltab.png)
+There were no significant differences in change of skeletal muscle strength, at time points post 2 RT sessions, post 4 RT sessions, 30min after 6th session, 2hrs after 6th session and 23hrs after 6th session (Fig 1D, *p* > 0.05).
 After 5 RT sessions (Fig 1D, Post 5RT), there was a significant difference in strength between glucose and placebo,  where placebo decreased strength 7% more than glucose (*p* = `r pdiff.c4`).
 
-**Figure 1:** A) An overview of the experimental design with 12 days of concomitant dietary intervention and resistance training (RT), preceded by 7 days involving familiarization.
+###### **Figure 1:** A) An overview of the experimental design with 12 days of concomitant dietary intervention and resistance training (RT), preceded by 7 days involving familiarization.
 Between days -7 and -1, participants were familiarized to the RT exercises via 1RM leg press and knee extension testing, and to the strength tests via Humac Norm dynamometer (days -7 and -5).
@@
 vastus lateralis at baseline (Day 1 leg 1, Day 2 leg 2), and after five RT sessions (Day 11 leg 1, Day 12 leg 2).
-Three participants dropped out of the intervention, either due to reported muscular discomforts of heavy RT, or sickness in their family.
-B and C) Changes in plasma glucose (mmol/L) and c-peptide levels (pmol/L) during post testing (Day 11 leg 1, Day 12 leg 2).
-Glucose levels in blood was measured via finger draws 120 (-120), 90 (-90), 30 (-30) minutes and immediately before RT (0), during RT (15), immediately after (30) and 2hrs after RT (120).
+Blood for measurement of plasma glucose and -c-peptide was sampled at baseline (Day 1), and during post testing (Day 11 leg 1, Day 12 leg 2), via finger draws and venous blood samples.
+Skeletal muscle strength was measured as peak torque in unilateral isometric and isokinetic (at 60 and 240 d/s) knee extension before, multiple times during, and after five and six session.
+A total of three participants dropped out of the intervention, either due to reported muscular discomforts of heavy RT, or sickness in their family.
+B and C) Changes in plasma glucose (B, mmol/L) and c-peptide levels (C, pmol/L).
+Glucose levels in blood was measured via finger draws 120 (-120), 90 (-90), and 30min (-30) before RT, immediately before RT (0), during RT (15), immediately after RT (30) and 2hrs after RT (120).
 C-peptide levels were measured simultaneously to these finger draws, except for 90min before and during RT.
-Finger draws and venous blood were sampled at baseline (Day 1) and post (Day 11 leg 1, Day 12 leg 2).
-D) Changes in muscular performance/recovery measured as isometric and isokinetic peak torque (60 and 240 d/s) via Humac Norm Dynamometer, conducted at baseline (A: Day -1), after 2 and 4 RT sessions (A: Day 4 and 8 leg 1, Day 5 and 9 leg 2), after five RT sessions/before the 6th session (Day 11 leg 1, Day 12 leg 2), as well as 30min, 2hrs and 23hrs after the 6th RT session (A: Day 11/12 leg 1, Day 12/13 leg 2).
-The index was calculated by normalizing individual peak torque to the highest peak torque at each respective speed, and then summarised and used in change score calculations.
-
-Values are presented as changes in estimated marginal means ± 95% CI. * = *p* < 0.05 between groups. Glucose n = 13, placebo n = 13.
+D) Changes in muscular strength measured as isometric and isokinetic peak torque (60 and 240 d/s) via Humac Norm Dynamometer, conducted at baseline (A: Day -1), after two and four RT sessions (A: Day 4 and 8 leg 1, Day 5 and 9 leg 2), after five RT sessions/before the 6th session (Day 11 leg 1, Day 12 leg 2), as well as 30min, 2hrs and 23hrs after the 6th RT session (A: Day 11/12 leg 1, Day 12/13 leg 2).
+The index was calculated by normalizing peak torque values to the highest peak torque value at each respective speed, and then summarized and used in change score calculations.
+Values are presented as changes in estimated marginal means ± 95% CI.
+* = *p* < 0.05 between groups.
+Glucose n = 13, placebo n = 13.
 
 ### Markers of ribosome biogenesis
+
+#### Total RNA and ribosomal RNA
 
@@
 A robust accumulation was also observed in ribosomal RNA expression in both RT with glucose and placebo, with mean increases between 34-43% (GLU) and 33-41% (PLAC) (*p* < 0.05, compared to baseline) in the four rRNA’s.
-The expression of in the 47S pre-rRNA increased by 37 and 59% with glucose and placebo respectively, where only RT with placebo increased significantly from baseline to post (fig 45, *p* = `r rr47tdiffp`).
-
-FIG2 HERE
+The expression of the 47S pre-rRNA increased by 37 and 59% with glucose and placebo respectively, where only RT with placebo increased significantly from baseline to post (fig 45, *p* = `r rr47tdiffp`).
 
@@
 Baseline = Day 1 leg 1/ Day 2 leg 2, Post = Day 11 leg 1, Day 12 leg 2.
-Total RNA and rRNA were analyzed in duplicates, with two duplicates per biopsy (two biopsies per time point), and normalized to ng x mg wet muscle weight for total RNA and external reference gene (Lambda) for rRNA.
+Total RNA and rRNA were analyzed in duplicates, with two duplicates per biopsy (two muscle tissue pieces per time point), and normalized to ng x mg wet muscle weight for total RNA and external reference gene (Lambda) for rRNA.
 Total RNA and rRNA changes were calculated as log-fold change score per mg wet muscle weight.
-Mean change scores of the duplicates were calculated before modelling and transformed to the log-scale.
+Mean change scores of the duplicates were calculated and transformed to the log-scale before modelling, then reverse-transformed for figure illustration.
 Values are estimated marginal means fold change per leg per supplement ± 95% CI.
 Glucose n = 13, placebo n = 13.
+
+#### Protein
 
@@
 
-C-myc levels increased 109% after RT with placebo, compared to an increase of 58% after RT with glucose, which was a significant difference (Fig 3A: *p* = `r cdiff`).
+Levels of c-Myc increased 51% more after RT with placebo compared to RT with glucose, a significant difference (Fig 3A: *p* = `r cdiff`).
 Levels of UBF and rpS6 increased 42 and 33% more in placebo compared to glucose, however, this was not a significant difference (Fig 3A: UBF: *p* = `r udiff`, rpS6: *p* = `r rpdiff`).
-Compared to baseline, both RT with glucose and RT with placebo increased significantly () Further, we observed a linear relationship between levels of UBF and accumulation of total RNA, where an increase of 1 SD unit of UBF equated to 13% increase in total RNA (*p* = `r corrp`).
-
-FIG3 HERE
+The increase in both RT with glucose and placebo for all measured proteins was significant at post compared to baseline (*p* < 0.05).
+Further, there was a linear relationship between UBF and total RNA, where an increase of 1 SD unit of UBF equated to a 13% increase in total RNA (*p* = `r corrp`).
 
 ###### **Figure 3:** Changes in levels of protein.
-
-A) Mean UBF, c-Myc and rpS6 (AU) at baseline and post + representative antibody stains of the respective proteins.
+A) Mean UBF, c-Myc and rpS6 (AU, arbitrary units) at baseline and post + representative western blots of the respective proteins.
 Baseline = Day 1 leg 1, Day 2 leg 2, post = Day 11 leg 1, Day 12 leg 2.
@@
 Change in protein levels were calculated as log-fold change scores normalized by pools (pool of all protein samples per gel).
-Mean change scores of the duplicates were calculated before modelling and transformed to the log-scale.
-Values are reverse transformed estimated marginal means fold change per leg per supplement ± 95% CI.
-* = significant between groups statistic, glucose compared to placebo.
+Mean log-change scores of the duplicates were calculated before modelling and reverse-transformed for figure illustration.
+Values are estimated marginal means fold change per leg per supplement ± 95% CI.
+* = *p* < 0.05 between groups.
 Glucose n = 13, placebo n = 13.
@@
 Values are presented as log-transformed means.
-
-#MISSING: CHECK % INCREASE OF TOTAL RNA PER 1 SD INCREASE OF UBF (MAY HAVE TO LOG-TRANSFORM)
-
-(We must add time as a covariate, as both total RNA and UBF are expected to increase with "time" in this intervention, due to RT.
-Therefore, adding time as a covariate allows us to model the relationship between RNA and UBF, without variations of time interfering)

### 2023-03-07 R/cpep.analysis.R

@@ -1,49 +1,27 @@
-### Insulin analysis
-
-## Author: Kristian Lian
-# Project: Ribose
-
+### Analysis of c-peptide data
+#
+#
+## Author: KL
+#
+#
+## Project: Ribose
+#
+#
+## Purpose: This script analyses the cleaned up c-peptide data by change-score
+# comparisons
+#
+#
+## Asscociated scripts: cpep.cleanup.R
+#
+#
 ## Packages
-library(dplyr)
-library(tidyverse)
-library(readxl)
-library(nlme)
-library(lme4)
-library(emmeans)
-library(tidyr)
+
+library(dplyr); library(tidyverse); library(tidyr); library(nlme); library(lme4); library(emmeans)
 
 ## Data
 
-ins.dat <- read_excel("./data/glucose/ribose.insulin.xlsx", na = "NA")
+ins.dat2 <- readRDS("./data/data-gen/glucose/cpep.clean.RDS")
 
-
-code <- read_excel("data/code_key.xlsx", na = "NA") %>%
-  select(-leg) %>%
-  print()
-
-## Data handling
-
-
-ins.dat$sample.time <- as.character(ins.dat$sample.time)
-
-ins.dat2 <- ins.dat %>%
-  inner_join(code) %>%
-  mutate(sample.time = as.character(sample.time)) %>%
-  mutate(time = if_else(sample.time == "0",
-                        "pre",
-                        if_else(sample.time == "0.1",
-                                "baseline",
-                                if_else(sample.time == "45",
-                                "min90",
-                                if_else(sample.time == "90",
-                                        "min120",
-                                        if_else(sample.time == "120",
-                                                "min150",
-                                                if_else(sample.time == "270",
-                                                        "min270", sample.time))))))) %>%
-  mutate(time = factor(time, levels = c("pre", "baseline", "min90", "min120", "min150", "min270")),
-         c.pep = as.numeric(c.pep)) %>%
-  print()
 
 ## Baseline analysis
 # Comparing pre measurements to post day baseline

### 2023-03-07 R/cpep.cleanup.R

@@ -1,0 +1,52 @@
+### Cleanup of c-peptide data
+#
+#
+## Author: KL
+#
+#
+## Project: Ribose
+#
+#
+## Purpose: This script cleans up raw c-peptide data, and saves the RDS "cpep.clean.RDS"
+#
+#
+## Packages
+library(readxl); library(tidyverse)
+
+## Data
+
+ins.dat <- read_excel("./data/glucose/ribose.insulin.xlsx", na = "NA")
+
+
+code <- read_excel("data/code_key.xlsx", na = "NA") %>%
+  select(-leg) %>%
+  print()
+
+
+## Data handling
+
+ins.dat$sample.time <- as.character(ins.dat$sample.time)
+
+ins.dat2 <- ins.dat %>%
+  inner_join(code) %>%
+  mutate(sample.time = as.character(sample.time)) %>%
+  mutate(time = if_else(sample.time == "0",
+                        "pre",
+                        if_else(sample.time == "0.1",
+                                "baseline",
+                                if_else(sample.time == "45",
+                                        "min90",
+                                        if_else(sample.time == "90",
+                                                "min120",
+                                                if_else(sample.time == "120",
+                                                        "min150",
+                                                        if_else(sample.time == "270",
+                                                                "min270", sample.time))))))) %>%
+  mutate(time = factor(time, levels = c("pre", "baseline", "min90", "min120", "min150", "min270")),
+         c.pep = as.numeric(c.pep)) %>%
+  print()
+
+saveRDS(ins.dat2, "./data/data-gen/glucose/cpep.clean.RDS")
+
+
+

### 2023-03-07 R/dxa.R

@@ -1,168 +1,0 @@
-## Analysis of baseline DXA results.
-
-# Author: SCM
-
-# This script analyses the baseline DXA values and creates a table
-
-# Packages
-
-library(readxl)
-library(tidyverse)
-library(nlme)
-library(lme4)
-library(broom)
-library(knitr)
-library(emmeans)
-library(knitr)
-
-#1 descritpive table of subject
-#
-#count gender to be put in table
-gendercount <- read_excel("data/dxa/ribose_dxa.xlsx") %>%
-  count(sex) %>%
-  pivot_wider(names_from = sex,
-              values_from = n) %>%
-  print()
-
-
-#Calculates grams to kg
-table.dxa <- read_excel("data/dxa/ribose_dxa.xlsx") %>%
-  select(sex, subject, age, height, weight, leanmasskg, fatmasskg, ffmg) %>%
-  #make grams into kg
-  mutate(ffkg = ffmg/1000) %>%
-  select(sex, age, height, weight, leanmasskg, fatmasskg, ffkg) %>%
-  print()
-
-#dxa table pr. gender               
-dxa.table <- table.dxa %>%
-  pivot_longer(names_to = "variable",
-               values_to = "values", col= age:ffkg) %>%
-  select(sex, variable, values) %>%
-  group_by(sex, variable) %>%
-  summarise(m = mean(values, na.rm = TRUE),
-            s= sd(values, na.rm = TRUE)) %>%
-  mutate(stat = paste0(round(m, 1), " (", round(s, 1), ")")) %>%
-  select(sex, variable, stat) %>%
-  pivot_wider(names_from = sex,
-              values_from = stat) %>%
-  mutate(variable = if_else(variable %in% c("age"),
-                            "Age",
-                            if_else(variable %in% c("fatmasskg"),
-                                    "Fat mass",
-                                    if_else(variable %in% c("ffkg"),
-                                            "Fat free mass",
-                                            if_else(variable %in% c("height"),
-                                                    "Height",
-                                                    if_else(variable %in% c("leanmasskg"),
-                                                            "Lean mass",
-                                                            if_else(variable %in% c("weight"),
-                                                                    "Weight",
-                                                                    ""))))))) %>%
-  print()
-
-
-dxa.table %>%
[…]

### 2023-03-07 R/dxa.cleanup.R

@@ -1,0 +1,31 @@
+### Baseline DXA cleanup
+#
+#
+## Author: SCM
+## Edited: KL
+#
+#
+## Project: Ribose
+#
+#
+## Purpose: This script cleans up DXA data grouped by leg (i.e. glu/plac)
+#
+#
+## Packages
+
+library(readxl); library(tidyverse)
+
+## Data
+
+dxa.dat <- read_excel("./data/dxa/ribose_dxaprleg.xlsx")
+
+
+dxa.clean <- dxa.dat %>%
+  select(subject, sex, leg, leanmassg, fatmassg, totalmasskg) %>%
+  #make grams into kg
+  mutate(leanmass = leanmassg/1000,
+         fatmass = fatmassg/1000)%>%
+  select(subject, sex, leg, leanmass, fatmass, totalmasskg) 
+
+saveRDS(dxa.clean, "./data/data-gen/dxa/dxa.clean.RDS")
+

### 2023-03-07 R/gluc.analysis.R

@@ -1,54 +1,28 @@
-#### Blood glucose change fig
-
-## Author: Kristian Lian
+### Plasma glucose analysis
+#
+#
+## Author: KL
+#
+#
 ## Project: Ribose
+#
+#
+## Purpose: This script analyses the cleaned up glucose data, by change-score
+# comparisons
+#
+#
+## Associated scripts: gluc.cleanup.R
+#
+#
+## Packages
+library(tidyverse); library(dplyr); library(nlme); library(lme4); library(emmeans); library(tidyr)
 
-## This script analyses is equal to mean-bloodglucose-change, except from not calculating log-fold change. Instead, absolute change scores are calculated
-# to illustrate the changes in absolute plasma glucose values. For description of the codes, see mean-bloodglucose-change.
+## Data
 
-## Other comments:
-# Participant 101, 102 and 103 did not get baseline samples at T3 and T4, therefore,
-# baseline values from T1 was used as baseline for T3 and T4
-
-# Packages
-library(dplyr)
-library(tidyverse)
-library(readxl)
-library(nlme)
-library(lme4)
-library(emmeans)
-library(tidyr)
-
-
-## Data handling
-
-gluc.dat <- read_excel("./data/glucose/fingerstick.xlsx", na = "NA")
-gluc.dat$sample_time <- as.character(gluc.dat$sample_time)
-
-glu.dat <- gluc.dat %>%
-  mutate(time = if_else(sample_time == "0",
-                        "baseline",
-                        if_else(sample_time == "45",
-                                "min45",
-                                if_else(sample_time == "90",
-                                        "min90",
-                                        if_else(sample_time == "120",
-                                                "min120",
-                                                if_else(sample_time == "135",
-                                                        "min135",
-                                                        if_else(sample_time == "150",
-                                                                "min150",
-                                                                if_else(sample_time == "270",
-                                                                        "min270", sample_time))))))),
-         time = factor(time, levels = c("baseline", "min45", "min90", "min120", "min135", "min150", "min270"))) %>%
-  mutate(glu = as.numeric(glu),
-         lak = as.numeric(lak)) %>%
-  print()
-
-## Change data
[…]

### 2023-03-07 R/humac.cleanup.R

@@ -1,25 +1,15 @@
-#### Humac data 
-# Data handling
-
-
-## Author: Kristian Lian
-
-# Purpose: This script tidies up all humac data for later analyses, and saves one data frame for pre-post 5 session results, and one data 
-# fram for pre-post the 6th session
-
-## Time-points
-# D-1: Baseline, before any supplementation or training
-# D4, D5, D8 and D9: Day 4, 5, 8 and 9 of the intervention, humac testing of the leg that performed
-# RT the preceding day
-# T3: Post testing leg #1 (leg that started the intervention). Leg #1 is tested four times at T3/T4:
-# Test 1 leg 1: 1.5hrs after protein ingestion, 45min before RT (T3)
-# Test 2 leg 1: 30min after RT (T3)
-# Test 3 leg 1: 2hrs after RT (T3)
-# Test 4 leg 1: ~23hrs after RT (T4)
-# Test 1 serve as a post test for the 5 RT sessions and pre test before the sixth session, test 2,
-# 3, and 4 serve as post test following sixth session
-# T4 and 13 follow the same design for leg #2
-
+### Humac raw data cleaunp
+#
+#
+## Author: KL
+#
+#
+## Project: Ribose
+#
+#
+## Purpose: This script cleans up raw dxa data
+#
+#
 # Packages
 library(readxl);library(tidyverse)
 

### 2023-03-07 R/humac.main.analysis.R

@@ -1,12 +1,19 @@
-#### Humac main data analysis
-
-## Author: Kristian Lian/DH
+### Humac main analysis
+#
+#
+## Author: KL
+#
+#
 ## Project: Ribose
-
-# Purpose: This script plots mean torque per supplement (both through intervention and pre vs. post) results from the ribose project, 
-# and analyses the data per test (isometric, isokinetic 60, isokinetic 240) in a linear model. This is the main data frame from humac,
-# meaning an analysis of change from pre to post 5 RT sessions.
-
+#
+#
+## Purpose: This script analyses humac data from baseline, post 2 session test,
+# post 4 session test and post 5 RT sessions test via change-score comparisons
+#
+#
+## Associated scripts: humac.cleanup.R
+#
+#
 # Packages
 library(tidyverse);library(nlme);library(lme4);library(emmeans)
 

### 2023-03-07 R/humac.main.baseline.R

@@ -1,11 +1,18 @@
-#### Humac main data baseline analysis
-
-## Author: Kristian Lian/DH
+### Humac baseline analysis
+#
+#
+## Author: KL
+#
+#
 ## Project: Ribose
-
-# Purpose: This script plots mean torque per supplement (both through intervention and pre vs. post) results from the ribose project, 
-# and calculates a baseline analysis between supplement legs, at each speed (isometric, 60 d/s and 240 d/s).
-
+#
+#
+## Purpose: This script analyses baseline measurements from humac
+#
+#
+## Associated scripts: humac.cleanup.R
+#
+#
 # Packages
 library(tidyverse);library(nlme);library(lme4);library(emmeans)
 

### 2023-03-07 R/nut.analysis.R

@@ -1,83 +1,362 @@
-#### Nutrition analysis
+### Analysis of c-peptide data
+#
+#
+## Author: KL/DH
+#
+#
+## Project: Ribose
+#
+#
+## Purpose: This script analyses the cleaned up nutrition data by change-score
+# comparisons
+#
+#
+## Asscociated scripts: cpep.cleanup.R
+#
+#
+## Packages
+library(tidyverse); library(lme4); library(lmerTest); library(emmeans)
 
-# Author: SCM/DH
-# Edited: Kristian Lian - small changes, mainly different organizing etc.
-# Project: Ribose
-
-# This script analyses nutrition data extracted from "MyFitnessPal" and imported via excel. Macro-nutrient intake per participant per day
-# is calculated to investigate whether there where different intakes between training days.
-
-# Packages
-library(knitr);library(rmarkdown);library(tidyverse);library(readr);
-library(tinytex);library(tidyr);library(broom);library(arsenal);library(lme4);library(dplyr);
-library(lmerTest);library(emmeans);library(magrittr);library(dabestr) #library(dbplyr)?
 
 # Data
 
-fp.weight <- readRDS("./data/data-gen/dxa/fp.weight.RDS")
+nut.clean <- readRDS("./data/data-gen/nutrition/nut.clean.RDS")
 
-nut.res <- readRDS("./data/data-gen/dxa/nutha.RDS")
 
-## Macro nutrient analysis
+### Macro nutrient change analysis
 # The code beneath tests whether there are differences in macro nutrient intake from pre to post, 
 # divided into "protein", "fat" and "carbohydrates", and total calories in a linear model. The main interest is if there is difference between glucose and 
 # placebo.
 
+## Protein
+
+# Log transforming for analysis
+pro.lchange <- nut.clean %>%
+  select(time = timepoint, subject, supplement = group, protein) %>%
+  group_by(subject, time, supplement) %>%
+  #summarise(mean.pro = mean(protein, na.rm = TRUE)) %>%
+  pivot_wider(names_from = time,
+              values_from = protein) %>%
+  ungroup() %>%
+  mutate(change.d2 = log(D2)-log(D1),
+         change.d3 = log(D3)-log(D1),
+         change.d4 = log(D4)-log(D1),
+         change.d5 = log(D5)-log(D1),
[…]

### 2023-03-07 R/nut.change.R

@@ -1,385 +1,0 @@
-#### Nutrition change analysis
-
-# Author: KL
-# Project: Ribose
-
-# This script calculates change scores and models the change for protein, protein/kg/body weight,
-# fat, carbohydrates and calories.
-
-# Packages
-library(tidyverse); library(lme4); library(lmerTest); library(emmeans)
-
-
-# Data
-
-fp.weight <- readRDS("./data/data-gen/dxa/fp.weight.RDS")
-
-dxa.res <- read_excel("data/dxa/Ribose_nutrition_result.xlsx")
-
-# Data handling
-
-# Data handling/cleaning
-
-nut.res <- dxa.res %>%
-  select(timepoint, subject, group, calories, fat, carbohydrates, protein, sup_pro, sup_gluc, kcal_glu) %>%
-  mutate(timepoint = if_else(timepoint %in% c("T1","T2"), 
-                             "D1",
-                             if_else(timepoint %in% c("3", "4"),
-                                     "D2",
-                                     if_else(timepoint %in% c("5", "6"),
-                                             "D3",
-                                             if_else(timepoint %in% c("7", "8"),
-                                                     "D4",
-                                                     if_else(timepoint %in% c("9", "10"),
-                                                             "D5",
-                                                             if_else(timepoint %in% c("T3", "T4"),
-                                                                     "D6",
-                                                                     "na"))))))) %>%
-  group_by(timepoint, subject, group) %>%
-  inner_join(fp.weight) %>%
-  summarise(protein = sum(protein + sup_pro),
-            fat = sum(fat),
-            calories = sum(calories + kcal_glu),
-            carbohydrates = sum(carbohydrates + sup_gluc),
-            weight = sum(weight)) %>%
-  mutate(proprkg = (protein/weight))
-
-saveRDS(nutha, "./data/data-gen/dxa/nutha.RDS")
-
-### Macro nutrient change analysis
-# The code beneath tests whether there are differences in macro nutrient intake from pre to post, 
[…]

### 2023-03-07 R/nut.cleanup.R

@@ -1,0 +1,53 @@
+### Nutrition data cleanup
+#
+#
+## Author: SCM/DH
+## Edited: KL
+#
+#
+## Project: Ribose
+#
+#
+## Purpose: This script cleans up the nutrition data
+#
+#
+## Packages
+library(readxl);library(tidyverse)
+
+# Data
+fp.weight <- read_excel("data/dxa/ribose_dxa.xlsx") %>%
+  select(weight, subject) 
+
+nut.dat <- read_excel("data/dxa/Ribose_nutrition_result.xlsx")
+
+
+# Data handling/cleaning
+
+nut.clean <- nut.dat %>%
+  select(timepoint, subject, group, calories, fat, carbohydrates, protein, sup_pro, sup_gluc, kcal_glu) %>%
+  mutate(timepoint = if_else(timepoint %in% c("T1","T2"), 
+                             "D1",
+                             if_else(timepoint %in% c("3", "4"),
+                                     "D2",
+                                     if_else(timepoint %in% c("5", "6"),
+                                             "D3",
+                                             if_else(timepoint %in% c("7", "8"),
+                                                     "D4",
+                                                     if_else(timepoint %in% c("9", "10"),
+                                                             "D5",
+                                                             if_else(timepoint %in% c("T3", "T4"),
+                                                                     "D6",
+                                                                     "na"))))))) %>%
+  group_by(timepoint, subject, group) %>%
+  inner_join(fp.weight) %>%
+  summarise(protein = sum(protein + sup_pro),
+            fat = sum(fat),
+            calories = sum(calories + kcal_glu),
+            carbohydrates = sum(carbohydrates + sup_gluc),
+            weight = sum(weight)) %>%
+  mutate(proprkg = (protein/weight))
+  
+
+saveRDS(nut.clean, "./data/data-gen/nutrition/nut.clean.RDS")
+
+

### 2023-03-07 R/nut.datahand.R

@@ -1,88 +1,0 @@
-#### Nutrition data handling
-
-# Author: SCM/DH
-# Edited: Kristian Lian - small changes, mainly different organizing etc.
-# Project: Ribose
-
-# This script tidies up the nutrition data set and saves two data frames (fp.weight and nutha) for later analyses.
-
-# Packages
-library(readxl);library(tidyverse)
-
-# Data
-fp.weight <-read_excel("data/dxa/ribose_dxa.xlsx") %>%
-  select(weight, subject) %>%
-  print()
-
-saveRDS(fp.weight, "./data/data-gen/dxa/fp.weight.RDS")
-
-dxa.res <- read_excel("data/dxa/Ribose_nutrition_result.xlsx")
-
-nutri.weight <- read_excel("data/dxa/Ribose_nutrition_result.xlsx")
-  
-
-# Data handling/cleaning
-
-nutha <- dxa.res %>%
-  select(timepoint, subject, group, calories, fat, carbohydrates, protein, sup_pro, sup_gluc, kcal_glu) %>%
-  mutate(timepoint = if_else(timepoint %in% c("T1","T2"), 
-                             "Day 1",
-                             if_else(timepoint %in% c("3", "4"),
-                                     "Day 2",
-                                     if_else(timepoint %in% c("5", "6"),
-                                             "Day 3",
-                                             if_else(timepoint %in% c("7", "8"),
-                                                     "Day 4",
-                                                     if_else(timepoint %in% c("9", "10"),
-                                                             "Day 5",
-                                                             if_else(timepoint %in% c("T3", "T4"),
-                                                                     "Day 6",
-                                                                     "na"))))))) %>%
-  group_by(timepoint, subject, group) %>%
-  inner_join(fp.weight) %>%
-  summarise(protein = sum(protein + sup_pro),
-            fat = sum(fat),
-            calories = sum(calories + kcal_glu),
-            carbohydrates = sum(carbohydrates + sup_gluc),
-            weight = sum(weight)) %>%
-  mutate(proprkg = (protein/weight))
-
-saveRDS(nutha, "./data/data-gen/dxa/nutha.RDS")
-
-
-nut.weight <- nutri.weight %>%
-  select(timepoint, meal, subject, protein, fat, calories,
-         carbohydrates, group, sup_pro, sup_gluc, kcal_glu) %>%
-  #changed names to compare groups each paired days
-  mutate(timepoint = if_else(timepoint %in% c("T1","T2"), 
[…]

### 2023-03-07 R/qpcr.import.R

@@ -1,18 +1,29 @@
-### QPCR IMPORT #################
-
-
-
-### Import of qPRC raw data
-
-#This script imports and saves a data frame with cq-values per sample and average efficiencies.
+### Import and cleaning of qPCR raw data
+#
+#
+## Author: DH
+#
+#
+## Project: Ribose
+#
+#
+## Purpose: This script imports and saves a data frame with cq-values per sample and average efficiencies.
 # Data are saved in data/derivedData/qpcr-data.RDS
-
+#
+#
+## Asscociated scripts: qpcr.import.R
+#
+#
+## Packages
 
 library(qpcR)
 library(qpcrpal)
 library(parallel)
 library(tidyverse)
 
+### QPCR IMPORT #################
+
+### Import of qPRC raw data
 
 ## Load batch 
 

### 2023-03-07 R/strength.index.R

@@ -1,13 +1,18 @@
+### Strength index analysis
+#
+#
+## Author: KL
+#
+#
+## Project: Ribose
+#
+#
+## Purpose: This script creates and analyses strength index combined of isometric and 
+# isokinetic peak torque, to be used in fig 1 
+#
+#
+## Packages
 
-
-
-
-
-
-
-
-
-# Packages
 library(readxl);library(tidyverse)
 
 ## Handling the data by creating a new factor called time from timepoint. This factor combines any observation at T1 and T2 to baseline, etc. 

### 2023-03-07 R/totrna.analysis.R

@@ -1,9 +1,20 @@
-## Total RNA analysis
-
-# This script imports the data frame from "totrna.cleanup", to analyse baseline values, calculate log-fold change scores from pre to post and model change
-# scores using a linear mixed effects model and estimated marginal means. Finally, it saves a data frame for figure illustration.
-
-## Packages 
+### Analysis of total RNA data
+#
+#
+## Author: KL/DH
+#
+#
+## Project: Ribose
+#
+#
+## Purpose: This script analyses the cleaned up total RNA data by change-score
+# comparisons
+#
+#
+## Asscociated scripts: cpep.cleanup.R
+#
+#
+## Packages
 library(tidyverse); library(readxl);library(nlme);library(lme4);library(emmeans)
 
 ## Data 

### 2023-03-07 R/totrna.cleanup.R

@@ -3,7 +3,7 @@
 ##
 ## Title: RNA cleanup
 ## Purpose: Filter away bad RNA estimates based on RNA/weight relationship
-## Author: DH/KL
+## Author: DH
 ##
 ##
 ##

### 2023-03-07 figures/design.work.R

@@ -1,139 +1,0 @@
-####### Study design fig work
-
-# Packages
-library(tidyverse)
-library(readxl)
-
-
-# Data
-d.dat <- read_excel("./data/design.dat.xlsx", na = "NA") %>%
-  print()
-
-biopsy_glyph <- "\U2193"
-blood_glyph <-  "\U2020"
-strength_glyph <- "\U2021"
-
-
-d.dat %>%
-  ggplot(aes(time, tp)) +
-  
-  
-  scale_y_continuous(limits = c(0,10), breaks = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10), expand = c(0,0)) +
-  
-  scale_x_continuous(limits = c(-3,21), breaks = seq(1:20), 
-                     labels = c("1" = "-7", "2" = "-6", "3" = "-5", "4" = "-4",
-                                "5" = "-3", "6" = "-2", "7" = "-1", "8" = "1",
-                                "9" = "2", "10" = "3", "11" = "4", "12" = "5", "13" = "6",
-                                "14" = "7", "15" = "8", "16" = "9", "17" = "10",
-                                "18" = "11", "19" = "12", "20" = "13")) +
-  
-  # Time periods (familiarisation/internvetion)
-  annotate("rect", xmin = -3, xmax = 21, ymin = 0, ymax = 5.3, alpha = 2, color = "lightblue", fill = "lightblue") +
-  annotate("text", x = 2, y = 5.1, label = "Familiarization (7 days)") +
-  annotate("text", x = 11, y = 5.1, label = "Unilateral RT + Dietary intervention (12 days)") +
-  
-  ## Signs
-  # Biopsy
-  annotate("text", x = -1.8, y = 4.5, label = "Muscle biopsies", size = 3) +
-  annotate("text", x = -0.4, y = 4.45, label = rep(biopsy_glyph, 1), vjust = 0) +
-  annotate("text", x = c(8, 18), y = rep(3.7),
-           label = rep(biopsy_glyph, 2), color = "blue", vjust = 0) +
-  annotate("text", x = c(9, 19), y = rep(1.3),
-           label = rep(biopsy_glyph, 2), color = "red", vjust = 0) +
-  
-  # Strength
-  annotate("text", x = -2, y = 4, label = "Strength test", size = 3) +
[…]

### 2023-03-07 figures/figure1.1.R

@@ -1,336 +1,0 @@
-### Ribose figure 1 ######
-#
-# Study design, strength (1RM, Humac), c-peptide and blood glucose
-#
-#
-#
-# Packages 
-library(tidyverse); library(emmeans); library(cowplot); library(grid); library(gridExtra); library(readxl); 
-library(elementalist); library(ggtext)
-#
-## Data
-
-# Study design
-d.dat <- read_excel("./data/design.dat.xlsx", na = "NA")
-
-# C-peptide
-cpep.change <- readRDS("./data/data-gen/glucose/insulin.figchange.RDS") # Untransformed data
-
-# Blood glucose
-glu.change <- readRDS("./data/data-gen/glucose/gluc.change.RDS") # Untransformed data
-
-# Humac data
-str.change <- readRDS("./data/data-gen/humac/str.change.RDS")
-str.emm <- readRDS("./data/data-gen/humac/emm.str.RDS")
-
-# Training volume data
-vol.lemm <- readRDS("./data/data-gen/training/vol.lemm.RDS")
-
-
-### Plots
-pos <- position_dodge(width = .2)
-
-# Designing the plot theme
-
-plot_theme <- theme(panel.grid.major = element_blank(),
-                    panel.grid.minor = element_blank(),
-                    panel.background = element_rect(fill = "lightblue", colour = NA),
-                    axis.line = element_line(colour = "black"))
-
-# Colours
-# Colours
-
-colors <- c("#d7191c",
-            "#fdae61",
-            "#abd9e9",
-            "#2c7bb6")
-
-
-
-## C-peptide
-
-cpep.fig <- cpep.change %>%
-  data.frame() %>%
-  add_row(supplement = "PLACEBO", time = "change.1", emmean = 0, SE = 0, df = 0, lower.CL = 0, upper.CL = 0, .before =1) %>%
-  add_row(supplement = "GLUCOSE", time = "change.1", emmean = 0, SE = 0, df = 0, lower.CL = 0, upper.CL = 0, .before = 2) %>%
-  mutate(time.c = gsub("change.", "", time), 
-         time.c = as.numeric(time.c)) %>%
-  ggplot(aes(time.c, emmean, group = supplement, fill = supplement)) +
-  annotate("text", x = c(120, 150), y = c(975, 950), label = "*") +
-  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
-                size = 0.5,
-                position = pos) +
[…]

### 2023-03-07 figures/figure1.R

@@ -15,34 +15,20 @@ d.dat <- read_excel("./data/design.dat.xlsx", na = "NA")
 
 # C-peptide
 cpep.change <- readRDS("./data/data-gen/glucose/insulin.figchange.RDS") # Untransformed data
-cpep.emm <- readRDS("./data/data-gen/glucose/cpep.emm.RDS") # Unstransformed emmeans
-cpep.lchange <- readRDS("./data/data-gen/glucose/insulin.change.RDS") # Log-transformed data
-cpep.lemm <- readRDS("./data/data-gen/glucose/cpep.lemm.RDS") # Log-transformed emmeans
 
 # Blood glucose
 glu.change <- readRDS("./data/data-gen/glucose/gluc.change.RDS") # Untransformed data
-glu.emm <- readRDS("./data/data-gen/glucose/gluc.emm.RDS") # Untransformed emmeans
-glu.lchange <- readRDS("./data/data-gen/glucose/glu.change.RDS") # Log-transformed data
-glu.lemm <- readRDS("./data/data-gen/glucose/gluc.logemm.RDS") # Log-transformed emmeans
 
-# Humac pre->post 5 session
-isom.lchange <- readRDS("./data/data-gen/humac/isom.lchange.RDS") # Log-transformed data
-lemm.isom <- readRDS("./data/data-gen/humac/emm.isom.RDS") # Log-transformed emmans
-isok60.lchange <- readRDS("./data/data-gen/humac/isok60.lchange.RDS") # Log-transformed data
-lemm.60 <- readRDS("./data/data-gen/humac/emm.60.RDS") # Log-transformed emmeans
-isok240.lchange <- readRDS("./data/data-gen/humac/isok240.lchange.RDS") # Log-transformed data
-lemm.240 <- readRDS("./data/data-gen/humac/emm.240.RDS") # Log-transformed emmeans
+# Humac data
+str.change <- readRDS("./data/data-gen/humac/str.change.RDS")
+str.emm <- readRDS("./data/data-gen/humac/emm.str.RDS")
 
-# Humac pre->post sessions 6
-isomac.lchange <- readRDS("./data/data-gen/humac/isom.lchange.ac.RDS") # Log-transformed
-lemm.isomac <- readRDS("./data/data-gen/humac/emm.ac.isom.RDS")
[…]

### 2023-03-07 figures/str.cont.fig.R

@@ -1,71 +1,0 @@
-
-
-
-library(readxl); library(tidyverse)
-
-str.emm2 <- readRDS("./data/data-gen/humac/emm.str.RDS") %>%
-  add_row(supplement = "placebo", time = "0", emmean = 0, SE = 0, df = 0, lower.CL = 0, upper.CL = 0, .before =1) %>%
-  add_row(supplement = "glucose", time = "0", emmean = 0, SE = 0, df = 0, lower.CL = 0, upper.CL = 0, .before =2) %>%
-  mutate(timeh = if_else(time == "0",
-                         "0",
-                         if_else(time == "change.2",
-                                 "8",
-                                 if_else(time == "change.3",
-                                         "16",
-                                         if_else(time == "change.4",
-                                                 "21",
-                                                 if_else(time == "change.5",
-                                                         "22.7",
-                                                         if_else(time == "change.6",
-                                                                 "23.4",
-                                                                 if_else(time == "change.7", 
-                                                                         "24.4", time)))))))) %>%
-  mutate(#timeh = factor(timeh, levels = c("0", "72", "192", "264", "266", "268", "291")),
-         timeh = as.numeric(timeh)) %>%
-  print()
-
-### Plots
-pos <- position_dodge(width = 0.2)
-
-# Designing the plot theme
-
-plot_theme <- theme(panel.grid.major = element_blank(),
-                    panel.grid.minor = element_blank(),
-                    panel.background = element_rect(fill = "lightblue", colour = NA),
-                    axis.line = element_line(colour = "black"))
-
-# Colours
-# Colours
-
-colors <- c("#d7191c",
-                     "#fdae61",
-                     "#abd9e9",
-                     "#2c7bb6")
-
-str.fig2 <- str.emm %>%
-  data.frame() %>%
-  ggplot(aes(timeh, emmean, fill = supplement)) +
-  #annotate("text", x = "change.4", y = -0.04, label = "*") +
-  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
-                position = pos,
-                width = 0.2) +
-  geom_line() +
-  geom_point(shape = 21, size = 2) +
-  scale_fill_manual(values = colors[c(1,4)]) +
-  scale_x_continuous(limits = c(0,25), breaks = c(0, 8, 16, 21, 22.7, 23.4, 24.4),
[…]

### 2023-03-07 figures/tab2.R

@@ -5,7 +5,7 @@
 #
 #
 ## Purpose:
-# The purpose of this script is to produce table .., presenting average total 
+# The purpose of this script is to produce table 2, presenting average total 
 # session volume and average change per RT session per supplement.  
 
 ## Packages
@@ -134,7 +134,7 @@ vol.tab <- footnote(x = vol.tab,
 vol.tab <- bg(vol.tab, i = c(2, 4, 6, 8, 10, 12), j = NULL, bg = "gray", part = "body", source = j)
 
 
-save_as_image(vol.tab, path = "./figures/voltab.png", res = 1080)
+save_as_image(vol.tab, path = "./figures/tab2.png")
 
 
           

### 2023-03-08 results_work.RMD
updated results

@@
 ```{r packages}
-library(tidyverse);library(readxl);library(lme4);library(emmeans);library(dplyr);library(cowplot);library(knitr);library(broom);library(rmarkdown);library(tinytex);library(tidyr);library(arsenal);library(magrittr);library(dabestr);library(nlme); library(pbkrtest);library(ggpubr); library(lmer)
+library(tidyverse);library(readxl);library(lme4);library(emmeans);library(dplyr);library(cowplot);library(knitr);library(broom);library(rmarkdown);library(tinytex);library(tidyr);library(arsenal);library(magrittr);library(dabestr);library(nlme); library(pbkrtest);library(ggpubr); library(nlme)
 ```
@@
 
-At baseline, there were no significant differences between training with glucose or placebo in muscle mass, isometric or isokinetic peak torque (Tab 1) or total session volume (Tab 2).
-Further, on pairwise consecutive days, there were no significant differences in mean macro nutrient (calories, carbohydrate, fat or protein) intake.
+At baseline, there were no significant differences between legs allocated to training with glucose or placebo in muscle mass, isometric or isokinetic peak torque (Tab 1).
+Total session volume (Tab 2) increased significantly from session 4 to session 6 compared to session 1 (*p* < 0.05), with no differences between exercising with glucose or placebo (*p* > 0.05).
+On pairwise consecutive days there were no significant differences in mean ingestion of protein, fat, carbohydrate or total calories (*p* > 0.05).
 
@@
 LM = Lean mass, Strength = strength index, calculated from isometric and isokinetic peak torque, Iso 0 PT = Isometric (0 d/s) peak torque, Iso 60 PT = Isokinetic peak torque at 60 d/s, Iso 240 PT = Isokinetic peak torque at 240 d/s.
-![](figures/desctab.png)
-Glucose ingestion before and after RT led to significant increases in plasma levels of glucose and c-peptide before, during and after the session, compared to RT with placebo, (Fig 1B, 120min: 38 ± 4% [*p* = `r p.120`], 135min: 31 ± 4% [*p* = `r p.135`], 150min: 32 ± 4% [*p* = `r p.150`], Fig 1C, 120min: 95 ± 10% [*p* = `r cpp.120`], 150min: 87 ± 10% [*p* = `r cpp.150`]), measured during the sixth RT session.
+![](figures/tab1.png)
+Glucose ingestion before and after RT led to significant increases in plasma levels of glucose (Fig 1B, 120min: 38 ± 4% [*p* = `r p.120`], 135min: 31 ± 4% [*p* = `r p.135`], 150min: 32 ± 4% [*p* = `r p.150`]) and c-peptide (Fig 1C, 120min: 95 ± 10% [*p* = `r cpp.120`], 150min: 87 ± 10% [*p* = `r cpp.150`]) before, during and after the session, compared to no change in placebo (*p* > 0.05).
 At 2hrs post-exercise, RT with glucose led to significantly lower plasma glucose levels compared to RT with placebo (1B, 270min: -8 ± 4% [*p* = `r p.270`].
@@
 ######**Table 2:** Mean total session volume and mean change in total session volume from session 1 to session 6. 
-![](figures/voltab.png)
-There were no significant differences in change of skeletal muscle strength, at time points post 2 RT sessions, post 4 RT sessions, 30min after 6th session, 2hrs after 6th session and 23hrs after 6th session (Fig 1D, *p* > 0.05).
-After 5 RT sessions (Fig 1D, Post 5RT), there was a significant difference in strength between glucose and placebo,  where placebo decreased strength 7% more than glucose (*p* = `r pdiff.c4`).
+![](figures/tab2.png)
+
+After 5 RT sessions (Fig 1D, Post 5RT), there was a significant difference in strength between glucose and placebo,  where placebo decreased strength 7% more than glucose (Fig 1D, *p* = `r pdiff.c4`).
+At the remaining time points, there were no significant differences between exercising with glucose and placebo (*p* > 0.05).
+While both RT with glucose and placebo led to significantly reduced strength post fifth session compared to baseline (*p* < 0.05), there were no significant differences in strength at baseline and 23hrs post sixth RT session (Fig 1D, Baseline and 23h post 6RT).
 
@@
 trcoef <- coef(summary(trna.lmod))
-trdiff <- round(trcoef[3,1], 2)
-trdiffp <- round(trcoef[3,5], 2)
+trdiff <- round(trcoef[3,1], 3)
+trdiffp <- round(trcoef[3,5], 3)
 trdiffse <- round(trcoef[3,2], 2)
@@
 
-Resistance training with glucose did not induce a higher accumulation of total RNA (fig 2A, *p* = `r trdiffp`) or rRNA  (fig 2B, 47S: *p* = `r rr47diffp`, 18S: *p* = `r rr18diffp`, 28S: *p* = `r rr28diffp`, 5.8S: *p* = `r rr58diffp`, 5S: *p* = `r rr5diffp`) compared to RT with placebo, measured as mean change from baseline to post.
+Resistance training with glucose did not induce a higher accumulation of total RNA (fig 2A, *p* = `r trdiffp`) or rRNA  (fig 2B, 47S: *p* = `r rr47diffp`, 18S: *p* = `r rr18diffp`, 28S: *p* = `r rr28diffp`, 5.8S: *p* = `r rr58diffp`, 5S: *p* = `r rr5diffp`) compared to RT with placebo.
 From baseline to post-intervention, there was a mean increase in levels of total RNA by 26 and 22% after RT with glucose and placebo, respectively (*p* < 0.05, compared to baseline).
@@
   group_by(subject, time, supplement) %>%
-  summarise(mean.sign = mean(norm.sign, na.rm = TRUE)) %>%
-  print()
+  summarise(mean.sign = mean(norm.sign, na.rm = TRUE))
 
@@
   group_by(subject, time, supplement) %>%
-  summarise(mean.rna = mean(RNA.weight, na.rm = TRUE)) %>%
-  print()
+  summarise(mean.rna = mean(RNA.weight, na.rm = TRUE))
 
@@
 
-  mutate(sd.ubf = (mean.sign - mean(mean.sign, na.rm = TRUE )/sd(mean.sign, na.rm = TRUE))) %>%
-  print()
+  mutate(sd.ubf = (mean.sign - mean(mean.sign, na.rm = TRUE )/sd(mean.sign, na.rm = TRUE)))
 

### 2023-03-08 results_work.RMD
results_work updates

@@
 ## Data
+# Full data - baseline -> 23h post 6RT
 str.change <- readRDS("./data/data-gen/humac/str.change.RDS") # Untransfromed data
@@
 
-## Linear mixed effects model
+# Acute data - Post 5RT -> 23h post 6RT
+astr.change <- readRDS("./data/data-gen/humac/astr.change.RDS") # Untransfromed data
+astr.lchange <- readRDS("./data/data-gen/humac/astr.lchange.RDS") # Log-transfromed data
+
+## Linear mixed effects model full data
 
@@
 
-## Emmeans
+## Emmeans full data
 # Untransformed
@@
 
-# For inline plotting
+## Linear mixed effects model acute
+# Untransformed
+astr.mod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
+                        data = astr.change)
+
+astr.res <- plot(astr.mod)
+astr.sum <- summary(astr.mod)
+
+# Log-transformed
+astr.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
+                         data = astr.lchange)
+
+astr.lres <- plot(astr.lmod)
+astr.lsum <- summary(astr.lmod)
+
+## Emmeans acute data
+astr.emm <- confint(emmeans(astr.mod, specs = ~"supplement|time")) %>%
+  data.frame()
+
+astr.lmod <- confint(emmeans(astr.lmod, specs = ~"supplement|time")) %>%
+  data.frame()
+
+## For inline plotting full data
 strcoef <- coef(summary(str.mod))
-pdiff.c4 <- round(strcoef[10,5], 2)
+
+# Time points compared to baseline (full data)
+pc2 <- round(strcoef[2,5], 3)
+pc3 <- round(strcoef[3,5], 3)
+pscient.c4 <- round(strcoef[4,5], 40)
+pc4 <- sprintf("%.5f", 7.886277e-06)
+pc5 <- round(strcoef[5,5], 3)
+pc6 <- round(strcoef[6,5], 3)
+pc7 <- round(strcoef[7,5], 5)
+
+# Glucose compared to placebo (full data)
+pdiff.c2 <- round(strcoef[8,5], 3)
+pdiff.c3 <- round(strcoef[9,5], 3)
+pdiff.c4 <- round(strcoef[10,5], 3)
+pdiff.c5 <- round(strcoef[11,5], 3)
+pdiff.c6 <- round(strcoef[12,5], 3)
+pdiff.c7 <- round(strcoef[13,5], 3)
+
+## For inline plotting acute data
+astrcoef <- coef(summary(astr.mod))
+apc2 <- round(astrcoef[2,5], 3)
+apc3 <- round(astrcoef[3,5], 3)
+apc4 <- round(astrcoef[4,5], 3)
 
@@
 
-At baseline, there were no significant differences between legs allocated to training with glucose or placebo in muscle mass, isometric or isokinetic peak torque (Tab 1).
-Total session volume (Tab 2) increased significantly from session 4 to session 6 compared to session 1 (*p* < 0.05), with no differences between exercising with glucose or placebo (*p* > 0.05).
-On pairwise consecutive days there were no significant differences in mean ingestion of protein, fat, carbohydrate or total calories (*p* > 0.05).
+```{r volume model}
+## Scripts
+# - volume.cleanup.R
+# - volume.analysis.R
 
-###### **Table 1:**  Baseline descriptives of legs allocated to RT with glucose or RT.
-LM = Lean mass, Strength = strength index, calculated from isometric and isokinetic peak torque, Iso 0 PT = Isometric (0 d/s) peak torque, Iso 60 PT = Isokinetic peak torque at 60 d/s, Iso 240 PT = Isokinetic peak torque at 240 d/s.
-![](figures/tab1.png)
-Glucose ingestion before and after RT led to significant increases in plasma levels of glucose (Fig 1B, 120min: 38 ± 4% [*p* = `r p.120`], 135min: 31 ± 4% [*p* = `r p.135`], 150min: 32 ± 4% [*p* = `r p.150`]) and c-peptide (Fig 1C, 120min: 95 ± 10% [*p* = `r cpp.120`], 150min: 87 ± 10% [*p* = `r cpp.150`]) before, during and after the session, compared to no change in placebo (*p* > 0.05).
-At 2hrs post-exercise, RT with glucose led to significantly lower plasma glucose levels compared to RT with placebo (1B, 270min: -8 ± 4% [*p* = `r p.270`].
+## Data
+vol.change <- readRDS("./data/data-gen/training/vol.change.RDS") # Untransformed data
+vol.lchange <- readRDS("./data/data-gen/training/vol.lchange.RDS") # Log-transformed data
+
+## Models
+
+# Untransformed 
+vol.mod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
+                     data = vol.change)
+vol.res <- plot(vol.mod)
+vol.sum <- summary(vol.mod)
+
+# Log-transformed
+vol.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
+                     data = vol.lchange)
+
+vol.lres <- plot(vol.lmod)
+vol.lsum <- summary(vol.lmod)
+
+## Emmeans
+
+# Untransformed
+vol.emm <- confint(emmeans(vol.mod, specs = ~"supplement|time")) %>%
+  data.frame()
+
+# Log-transformed
+vol.lemm <- confint(emmeans(vol.lmod, specs = ~"supplement|time")) %>%
+  data.frame()
+
+## For inline plotting
+
+# Sessions compared to baseline
+volcoef <- coef(summary(vol.mod))
+pvol.s2 <- round(volcoef[2,5], 4)
+pvol.s3 <- round(volcoef[3,5], 4)
+pvol.s4 <- round(volcoef[4,5], 4)
+pvol.s5 <- round(volcoef[5,5], 3)
+pvol.s6 <- round(volcoef[6,5], 5)
+
+# Glucose compared to plac
+pvol.c2 <- round(volcoef[7,5], 3) 
+pvol.c3 <- round(volcoef[8,5], 3) 
+pvol.c4 <- round(volcoef[9,5], 3) 
+pvol.c5 <- round(volcoef[10,5], 3) 
+pvol.c6 <- round(volcoef[11,5], 3)
+
+```
+
+There were no difference in mean change of total training session volume between exercising with placebo or glucose, with an increase of 18 and 16% from session 1 to session 6, respectively (Tab 2; Mean change, 2: *p* = `r pvol.c2`, 3: *p* = `r pvol.c3`, 4: *p* = `r pvol.c4`, 5: *p* = `r pvol.c5`, 6: *p* = `r pvol.c6`).
+Both exercising with glucose and placebo led to significant increases in mean total training session volume at session four, five and six, compared to session 1 (Tab 2; Mean change, 2: *p* = `r pvol.s2`, 3: *p* = `r pvol.s3`, 4: *p* = `r pvol.s4`, 5: *p* = `r pvol.s5`, 6: *p* = `r pvol.s6`).
+In regard to the dietary intervention, there were no difference in mean macro nutrient intake (protein, fat, carbohydrates) or total calorie intake between glucose and placebo on pairwise consecutive days (*p* > 0.05).
 
@@
 
-After 5 RT sessions (Fig 1D, Post 5RT), there was a significant difference in strength between glucose and placebo,  where placebo decreased strength 7% more than glucose (Fig 1D, *p* = `r pdiff.c4`).
-At the remaining time points, there were no significant differences between exercising with glucose and placebo (*p* > 0.05).
-While both RT with glucose and placebo led to significantly reduced strength post fifth session compared to baseline (*p* < 0.05), there were no significant differences in strength at baseline and 23hrs post sixth RT session (Fig 1D, Baseline and 23h post 6RT).
+Glucose ingestion before and after RT led to significant increases in plasma levels of glucose by 38% immediately before RT (Fig 1B, 0 min: *p* = `r p.120`), by 31% during RT (1B, 15 min: *p* = `r p.135`) and by 32% immediately after RT (1B, 30 min: *p* = `r p.150`), compared to placebo ingestion before and after RT.
+Two hours after the session, RT with glucose measured 8% lower plasma glucose levels compared to RT with placebo (1B, 270min: *p* = `r p.270`).
+Similarly, levels of c-peptide increased significantly more in exercises with glucose compared to placebo, by 95% immediately before (Fig 1C, 0 min: *p* = `r cpp.120`) and 87% after RT (1C, 30 min: *p* = `r cpp.150`).
+
+After 5 RT sessions (Fig 1D, Post 5RT), there was a significant difference in strength between glucose and placebo,  where placebo decreased strength 7% more than glucose (1D, *p* = `r pdiff.c4`).
+At the remaining time points, there were no significant differences in mean change between exercising with glucose and placebo (1D, Post 2RT: *p* = `r pdiff.c2`, Post 4RT: *p* = `r pdiff.c3`, 30min post 6RT: *p* = `r pdiff.c5`, 2h post 6RT: *p* = `r pdiff.c6`, 23h post 6RT: *p* = `r pdiff.c7`).
+Both RT with glucose and placebo led to significantly reduced strength post fifth session compared to baseline, by 11 and 18% respectively (1D, Post 5RT: *p* = `r pc4`).
+Comparisons of the acute data gathered from post fifth session until and including 23hrs post sixth session showed a significant mean increase in strength of 5-9% after RT with glucose and placebo 30min after sixth RT session (1D, 30min post 6RT: *p* = `r apc2`) and two hours after sixth RT session (1D, 2h post 6RT: *p* = `r apc3`).
+Twenty-three hours after the last RT, strength was unchanged compared to after fifth RT session (23h post 6RT: *p* = `r apc4`).
 

### 2023-03-08 R/volume.cleanup.R

@@ -1,0 +1,42 @@
+### Training volum cleanup
+#
+#
+## Author: KL
+#
+#
+## Project: Ribose
+#
+#
+## Purpose: This script cleans up the training volum data frame
+#
+#
+# Packages
+library(readxl);library(tidyverse)
+
+# Data
+tot.vol <- read_excel("./data/training/ribose_volume.xlsx", na = "NA") %>%
+  select(subject, timepoint, tot.volume, supplement) 
+
+## Handling the data by creating a new factor called time from timepoint. This factor combines any observation at T1 and T2 to baseline, etc. 
+# The code also sorts the order of the factor time, from baseline to session 6, using time = factor(time, levels c()), and sets placebo to be compared to 
+# glucose via supplement = factor(supplement, levels = c()).
+
+tot.volh <- tot.vol %>%
+  mutate(time = if_else(timepoint %in% c("T1", "T2"),
+                        "baseline",
+                        if_else(timepoint %in% c("D3", "D4"),
+                                "session2",
+                                if_else(timepoint %in% c("D5", "D6"),
+                                        "session3",
+                                        if_else(timepoint %in% c("D7", "D8"),
+                                                "session4",
+                                                if_else(timepoint %in% c("D9", "D10"),
+                                                        "session5",
+                                                        if_else(timepoint %in% c("T3", "T4"),
+                                                                "session6", timepoint))))))) %>%
+  mutate(time = factor(time, levels = c("baseline", "session1", "session2", "session3", "session4", "session5", "session6")),
+         supplement = factor(supplement, levels = c("placebo", "glucose")))
+
+saveRDS(tot.volh, "./data/data-gen/training/vol.clean.RDS")
+
+

### 2023-03-08 figures/tab1.R

@@ -14,6 +14,8 @@ library(readxl); library(tidyverse); library(flextable)
 
 leg.dat <- read_excel("./data/dxa/ribose_dxaprleg.xlsx")
 
+dxa.dat <- read_excel("./data/dxa/ribose_dxa.xlsx")
+
 humac.clean2 <- readRDS("./data/data-gen/humac/humac.clean2.RDS")
 
 str.index <- readRDS("./data/data-gen/humac/str.index.RDS")
@@ -24,70 +26,73 @@ code <- read_excel("./data/sup.xlsx")
 
 ## DXA
 
-dxa.tab <- leg.dat %>%
-  full_join(code) %>%
-  select(subject, sex, supplement, leanmassg, fatmassg, totalmasskg) %>%
-  # Convert g to kg
-  mutate(leanmass = leanmassg/1000,
-         fatmass = fatmassg/1000) %>%
-  select(-leanmassg, fatmassg) %>%
-  group_by(supplement) %>%
-  summarise(mean.lm = mean(leanmass, na.rm = TRUE),
-            mean.fm = mean(fatmass, na.rm = TRUE),
-            mean.tm = mean(totalmasskg, na.rm = TRUE),
-            sd.lm = sd(leanmass, na.rm = TRUE),
-            sd.fm = sd(fatmass, na.rm = TRUE),
-            sd.tm = sd(totalmasskg, na.rm = TRUE)) %>%
-  mutate(lm.stat = paste0(round(mean.lm, 2), " ± ", round(sd.lm, 2)),
-         fm.stat = paste0(round(mean.fm, 2), " ± ", round(sd.fm, 2)),
-         tm.stat = paste0(round(mean.tm, 2), " ± ", round(sd.tm, 2))) %>%
-  mutate(supplement = if_else(supplement == "glucose",
-                              "Glucose",
-                              if_else(supplement == "placebo",
-                                      "Placebo", supplement))) %>%
-  select(supplement, lm.stat) 
+# Counts the "n" of females and males, added in the last step before 
+# making a flextable
+gendercount <- dxa.dat %>%
+  count(sex) %>%
+  mutate(sex = if_else(sex == "female",
+                       "Female",
+                       if_else(sex == "male",
+                               "Male", sex))) 
+
+# Summarising and prepping dxa data
+dxa.tab <- dxa.dat %>%
+  select(subject, sex, age, height, weight, leanmassg) %>%
[…]

### 2023-03-08 ribose-paper.RMD
updates
results_work updates

@@
 
+###### **Table 1:**  Baseline characteristics.
+Age = mean age in years, Height = mean height in cm, Weight and Lean mass = mean in kg, Iso 60/240 = mean unilateral isokinetic knee extension peak torque in Nm at 60 d/s and 240 d/s, Isom = mean unilateral isometric knee extension peak torque in Nm.
+
+![](figures/tab1.png)
+
 ### Experimental design
@@
 To ensure equal conditions during training sessions, participants exercised and tested at the same time of day, +/- 1hr (check this) with the same trainer, and recorded and repeated their daily macro nutrient intake on pairwise consecutive days (Tab 4).
-
-###### **Table 1:** Baseline characteristics of the participants. BMI: body mass index, A = charactaristics per sex, B = characteristics per leg. Values are mean ± standard deviation.
 
@@
 
-![](figures/flowchart_selection.PNG)
-
-###### **Figure 1:** Flowchart of the selection process. The order in which participants performed the two intervention blocks was determined in a planned randomised fashion.
-
-![](figures/timeline.PNG)
-
-###### **Figure 2.** Overview of the intervention, with 12 days of concomitant dietary intervention and resistance training (RT), preceded by 7 days of testing.
-T0 = timepoint 0 (testing days prior to intervention), T1/T2 = biopsy prior to resistance training for leg #1 and #2, respectively, T3/T4 post intervention testing of leg #1 and #2, respectively, including post sixth RT session.
-Test = physical performance test (unilateral 1RM leg press, knee extension, isometric knee extension force, isokinetic knee extension torque) testing on days 4, 5, 8, 9, T3, T4 will not include 1RM leg press and knee extension, Biopsy = microbiopsy, RT#1 = resistance training leg #1, RT#2 = resistance training leg#2.
-Other tests (D2O, DXA, spit) will be conducted as part of the study as a whole.
-These are not relevant to this data set, thus will not be described here.
-
 ### Dietary intervention
@@
 
-###### **Table 2:** Primer sequence and performance. Average cycle thresholds (Ct) and priming efficiencies were calculated from all qPCR reactions.
-
-![](figures/Primer.sequences.PNG)
-
 ### Protein extraction and immunoblotting
@@
 
-```{r nutrition change analysis and fig}
+```{r nutrition fig}
 # Associated scripts:

### 2023-03-10 ext_method.RMD
results_work updates
updated methods and results in ribose-paper.RMD

@@
 
+- Primer sequence table?
+- Antiboy table?
+- Power calculation + rational?
+- Normalization of total protein, qpcr values and protein signals
 - UBF/RNA correlation: Make a new model to see if data point affects the current model
+
+Mean-centered baseline values were used to correct for regression to the mean, which potentially occurs when the same participants
+are repeatedly tested.
+Regression to the mean entails that scores close to the upper or lower limit of a participants potential score would be replaced with a score closer to the mean, e.g.
+high baseline values are shifted downwards while low baseline values are shifted upwards, creating a negative association between initial values and change.
+Estimated marginal means were calculated from the linear model, enabling acquisition of least-square means, showing the means for all involved groups adjusted to means of other potential factors in the model.
+
+## Code/formulas for quantitative analyses
+Total RNA: 
+- Data import and prepping
+ - Packages: tidyverse, readxl, nlme
+ - Script: totrna.cleanup.R
+ - Raw data = protein concentration and 260/280 ratio from excel sheet "RNA_raw"
+ - Checked for outliers via mutate(resid = resid/sd(resid), outlier = if_else(resid < -2,      "out", "in"))
+  - This code first calculates the samples residuals, then creates a variable called "outlier", containing "in" and "out".
+    Further, it marks samples with "out" if the residuals are greater than 2, and "in" if residuals are lesser than 2. Thus, 
+    we can choose to filter out participants marked with out
+
+- Change score calculation
+ - Packages: tidyverse
+ - Script: totrna.analysis.R
+ - Before change score calculation, total RNA is normalized by wet muscle weight
+  - Approach: mean RNA / mean weight = rna.weight. Then, mean rna.weight is used to calculate change scores
+ - Cleaned up data was used to calculated the change score from pre to post
+  - Visual inspection of the residuals resulted in log-transforming the data, to control for heteroscedasticity
+  - Calculation: change = log(post)-log(pre), pre = pre - mean(pre, na.rm = TRUE).
+   - pre = pre - mean(pre, na.rm = TRUE): this code subtracts the mean rna concentration at pre from each subjects mean,
+     so we can mean-center the baseline values to control for regression to the mean
+
+- Analysis:
+ - Packages: tidyverse, lme4, nlme, emmeans, readxl
+ - Script: totrna.analysis.R
+ - Model: lmerTest::lmer(change ~ pre + supplement + (1|subject), data = change_dat)
+  - In this model, change is added as the response, with pre (baseline values) and supplement (treatment, glucose/placebo) as         explanatory factors
+  - (1|subject) creates an intercept for each participant? why?
+ - Emmeans: estimated marginal means, allows calculation of least square means - why?
+
+rRNA
+- Data import and prepping
+ - Packages: qpcr, qpcrpal, parallell, tidyverse
+ - Script: qpcr.import.R
+ - Imports raw fluorescence data and models amplification curves with a best-fit sigmoidal model using the qPCR package
+ - Estimates primer efficiency and cq-values
+
+- Change score calculation
+ - Packages: tidyverse
+ - Script: qpcr.analysis.R
+ - Before change score calculation, qpcr-values are normalized by wet muscle weight and external reference gene (Lambda)
+  - ..
+- Analysis
+ - Analyses were done with the same approach and model as for total RNA, with normalize cq-values per target 
+  - 47S, 18S, 28S, 5.8S and 5S
+
+Protein
+- Data import and prepping
+ - Packages: tidyverse, readxl
+ - Script: western.import.R, western.cleanup.R
+ - Normalization:
+  - We wanted to normalize signal data to total protein per gel and pools, to be able to        compare gels and protein samples from with varying amount of total protein
+  - First, we quantified mean gray from within and between the lanes of each gel in ...         (software).
+  - Mean gray within lanes were categorized as total protein, mean gray between lanes were      categorized as background and summarised per gel
+  - Then we subtracted background from total protein, creating the total protein per lane       (tpl) factor
+  - To make this factor comparable across gels, we divided it by the maximal tpl per gel        (tpl = tpl/max(tpl))
+  - We also did this for the mean protein signal (signal = signal/max(signal))
+  - Then, we divided signal by tpl, creating the norm.sign factor (norm.sign = signal/tpl).
+  - norm.sign was lastly divided by mean signal from the pooled samples (norm.sign =            norm.sign/pool), resulting in protein signals comparable across different gels and with     varying amount of protein sample (ng/ul)
+ - Clean up
+  - Quantified western data was cleaned up the same way as total RNA, with marking samples 
+    above and below 2 residuals
+- Analysis
+ - The model used for the protein data was from the lme4 package
+ - model: lme(log(norm.sign) ~ time + supplement + time:supplement, 
+          random = list(subject = ~ 1),
+          data = filter(prot.dat, target == "rps6"))
+  - Here, norm.sign (i.e. normalized mean protein signal) is the response, while time           (baseline and post) + supplement (treatment) and time: supplement are the explanatory       factor/terms
+  - the model uses a two-sided linear formula to described the fixed effects (og(norm.sign)     ~ time + supplement + time:supplement)
+   - This entails describing whether there are differences between pre and post (time),          between treatments (supplement), and both?
+   - random effects (random = list(subject = ~ 1))?
+   - Why did we use a different model?
+
+UBF/Total RNA
+- Data prepping
+- Packages: tidyverse, ggpubr
+- Scripts: ubf_rna.R
+ - For this analysis, we used cleaned up data from total RNA and western blot analyses
+ - UBF was filtered out from the clean western data
+ - mean.ubf and mean.rna were summarised in their respective data frames, then joined for 
+   the model
+  - Further, we calculated the SD units of ubf to model with total rna, because ...
+   - calculation: mutate(sd.ubf = (mean.sign - mean(mean.sign, na.rm = TRUE )/sd(mean.sign,      na.rm = TRUE)))
+- Analysis
+ - this model was built with the nlme package
+ - model: lme(mean.rna ~ mean.sign + time , 
+              random = list(subject = ~ 1), 
+              data = joined.dat, 
+              na.action = na.omit)
+ - Here the mean.rna is the response, with the terms being mean.sign and time
+   we added time as a covariate here, since both levels of ubf and total rna are expected
+   to vary with time
+ - random effect (random = list(subject = ~ 1)) - why ?
+
+Strenght index
+- Data import and cleanup
+ - Strength index was calculated with peak torque values from all humac tests
+  - I.e. baseline -> post session 60, all speeds
+  - To manage this, we first calculated an index for each test speed (0, 60 and 240 d/s), by     by grouping peak torque values by time and supplement, and diving it by the maximum peak     torqu.
+  - Then, we summarised the peak torque indexes by subject, time and supplement and gathered     all the values under the variable "strenght index".
+  - This way, we can compare peak torque values from different movement speeds, as the index     puts them on a comparable scale
+- Change score calculations
+ - change score was done similarly to total RNA, across multiple time points
+- Analysis
+ - To account for the multiple time points, the model was changed accordingly
+ - Model: lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
+                         data = str.change)
+ - Why do we add the zero?
+ - In this model, we have both baseline and time as covariates, to account for the multiple    time points
+
+Plasma glucose and c-peptide
+- Data import and prepping
+ - Data was imported and prepped in the same way as previously described data,
+   by summarising means at each time point
+- Change score calculation
+ - change scores was calculated in the same way as previously described
+- Analysis
+ - The same model as strength index

### 2023-03-10 methods_work.RMD
updated methods and results in ribose-paper.RMD

--- 2023-03-08 ribose-paper.RMD
@@
 ---
-title: "Glucose ingestion before and after resistance training does not augment ribosome biogenesis in healthy moderately trained young adults"
-author: "Kristian Lian, Daniel Hammarström, Stian Ellefsen, Håvard Hamarsland"
-bibliography: resources/ribose.bib
-csl: resources/jap.csl
-output:
-  word_document:
-    reference_docx: reference-template-02.docx
-editor_options: 
-  chunk_output_type: console
+title: "methods"
+output: word_document
+date: "2023-03-08"
 ---
@@
 
-```{r packages}
-library(tidyverse);library(readxl);library(lme4);library(emmeans);library(dplyr);library(cowplot);library(knitr);library(broom);library(rmarkdown);library(tinytex);library(tidyr);library(arsenal);library(magrittr);library(dabestr);library(nlme); library(pbkrtest)
-```
-
-## Introduction
-
-Responses to systematic resistance training (RT) vary widely between individuals, with as many as 10-15% showing impaired muscle growth in response to standardized training interventions (ALVAREZ18;MANN14;THALACKER13).
-Such inability to benefit from RT may be explained by genetic and/or epigenetic disposition, affecting the internal physiological milieu in terms of ability to mount an anabolic response to RT [@figueiredo21;@thalacker13].<!-- Check these refs, do they correspond to genet epigen and internal mileu?
-checked-->
-Although RT program modifications such as altered exercise volume or intensity may induce more favorable adaptations in some individuals, such benefits are not uniform between individuals [@hammarstrom.etal_2020;@krieger09;@schoenfeld17].<!-- Add ref for more rt program modifications.
-checked -->
-For example, although changes in training volume generally induces greater benefits of RT, not all individuals benefit from such modifications [@hammarstrom.etal_2020].
-Therefore, other means than modification of RT variables seems necessary to optimize individual responses.<!-- The above sentence might be a bit contradictory?
--->
-
-Nutritional supplements such as protein and creatine effectively optimize RT adaptations [@cermak12;@lanhers15;@lanhers17;@morton18].
-However, it remains unknown if other nutritional adjuvants such as glucose can increase the efficacy of RT.
-This is surprising, as glucose is the preferred energy substrate of the contracting skeletal muscle during strenuous exercise and a major supplier of energy to cells via ATP synthesis [@mul15;@tanaka18].
-Furthermore, energy availability is a decisive factor in *de novo* synthesis of ribosomes [@kusnadi15;@moss07;@tanaka18] <!--Add ref that indicate high energy requirement for ribo biog possibly (VONWALDEN19;TANAKA18?
-MOSS.
-checked  --> which in turn determines muscle growth by increasing the muscles translational capacity [@figueiredo19;@hammarstrom20;@tanaka18;@stec16;@vonwalden19].
-Therefore, investigating the effects of combined RT and glucose ingestion may provide valuable insight into a potential additive effect of glucose and RT on ribosome biogenesis.
-
-Transcription of ribosomal RNA (rRNA) by Pol I is considered the limiting step in synthesizing new ribosomes [@Moss_1995].
-Multiple signaling pathways converge to regulate rRNA transcription, including the mammalian target of rapamycin (mTORC1) signal-transduction pathway and c-Myc [@kusnadi15;@vonwalden19;@tanaka18].<!-- Is c_myc included in references here?
-could add West 2016.
-c-Myc included in Von Walden and Tanaka, adding West 2016 and Mori 21 when I find them-->
-mTOR receives input from growth factors, hormones, mechanical loading, and nutrients to balance protein synthesis through multiple mechanisms based on cellular energy levels [@hoppe09].
-Importantly, mTOR and its downstream target S6K1 serve as a mediator from insulin signaling in skeletal muscle[@hillier.etal_2000].
-The mTORC1 pathway contributes to ribosome biogenesis through the regulation of translation of ribosomal proteins and to forming of the preinitiation complex (PIC) that marks the initiation of rRNA transcription [@figueiredo19;@vonwalden19].
-Parallel to mTORC1 [@Mori_2021; @West_2016], c-Myc increases ribosomal biogenesis directly and through transcriptional control of the upstream binding factor (UBF) [@sanijUBFLevelsDetermine2008; @Poortinga_2011].
-Additionally, UBF phosphorylation, required for interaction with the rDNA promoter, is increased by high glucose in a mTORC1 dependent manner (rapamycin sensitive) in kidney glomerular epithelial cells[@Mariappan_2011].
-Independently from UBF and mTORC1, high glucose was shown to lead to chromatin remodeling, which promotes rRNA transcription in cell cultures[@Zhai_2012].
-Together these observations indicate a potential role of glucose in positively affecting muscle rRNA synthesis.
-
-The purpose of this investigation was to test the hypothesis that glucose supplementation before and after<!-- during?
-before and after would be more precise --> RT would increase markers of ribosomal abundance following five RT sessions.
-Secondly we aimed to describe the association between UBF and total RNA abundance in human skeletal muscle.
-<!-- We should maybe add a line or two about the rational here -->
-
 ## Methods
@@
 
-All participants were informed about the potential discomforts and risks associated with the study and gave their informed consent before study enrollment.
+All participants gave their informed consent prior to study enrollment.
 The study was approved by the regional ethical committee (REK, ID nr.
@@
 2-8 RT sessions per 14 days for the last six months).
-Exclusion criteria were previous injury leading to impaired strength, inability to perform resistance training, symptoms and a medical record of metabolic disorders including hyperglycemia, i.e.
+Exclusion criteria were previous injury leading to impaired strength, inability to perform resistance exercise training, symptoms and a medical record of metabolic disorders including hyperglycemia, i.e.
 fasting venous plasma glucose ≥6.1 mmol/L and/or 2-hour glucose tolerance ≥7.8 mmol/L, and/or HbA1c >42 mmol/mol.
@@
 One due to sickness and inability to resume, two participants experienced muscular discomfort connected to heavy resistance training.
-Baseline characteristics (Tab 1) were measured using DXA (Prodigy Advance PA+302047, Lunar, San Francisco, CA, USA) on Day -1, the last day preceding the RT intervention.
-
-###### **Table 1:**  Baseline characteristics.
-Age = mean age in years, Height = mean height in cm, Weight and Lean mass = mean in kg, Iso 60/240 = mean unilateral isokinetic knee extension peak torque in Nm at 60 d/s and 240 d/s, Isom = mean unilateral isometric knee extension peak torque in Nm.
-
-![](figures/tab1.png)
+Lean mass (Tab 1) was measured using DXA (Prodigy Advance PA+302047, Lunar, San Francisco, CA, USA) on Day -1, the last day preceding the RT intervention.
 
@@
 
-The study was designed as a 12-day double-blinded randomized controlled trial, with a unilateral RT protocol (Fig 2).
-Participants were randomized to commence with either glucose (GLU, 30g per bolus) or placebo (PLAC, 0g per bolus), and either dominant or non-dominant leg (Fig 1), alternating RT and supplement from one day to the other (Fig 2).
-Investigators were blinded as the supplement randomization was disclosed only to personnel not involved in any other aspect of the study.
-THe participants were blinded via diluting boluses of glucose (30g glucose) (Glucosum monohydricum, Merck KGaA, Darmstadt, Germany) and placebo (~0.3g Stevia rebaudiana extract) (Steviosa, Soma Nordic AS, Oslo, Norway) in 300ml Fun Light juice/saft (Orkla, Oslo, Norway).
-A blinded taste test revealed that the participants were not able to disclose the contents of the provided boluses (30g GLU vs.
-0g GLU).
-To ensure equal conditions during training sessions, participants exercised and tested at the same time of day, +/- 1hr (check this) with the same trainer, and recorded and repeated their daily macro nutrient intake on pairwise consecutive days (Tab 4).
+The study was designed as a 12-day double-blinded randomized controlled trial, with a alternating unilateral RT protocol (Fig 1A).
+Participants were randomly allocated to exercise either their dominant or non-dominant leg with glucose or placebo before and after RT, and whether they started RT with glucose or placebo.
+Thus, participants exercised every day, alternating between e.g., one leg + glucose and the other leg + placebo (1A).
+One person was exclusively responsible for the randomization code and supplement distribution, blinding both investigators and participants in regard to which leg exercised with glucose and placebo.
+Glucose (Glucosum monohydricum, Merck KGaA, Darmstadt, Germany) and placebo (Steviosa, Soma Nordic AS, Oslo, Norway) was masked by mixing with 300ml Fun Light juice/saft (Orkla, Oslo, Norway).
+A blinded taste test revealed that the participants were not able to disclose the contents of the provided boluses (30g glucosum monohydricum vs.
+~0.3g Stevia rebaudiana extract).
+To ensure equal conditions during training sessions, participants exercised and tested at the same time of day, +/- 1hr (check this) with the same supervisor, in addition to recording and repeating their daily macro nutrient intake (protein, fat, carbohydrate) and total calories on pairwise consecutive days.
 
-```{r DXA}
-# Script: DXA
-
-# Data handling
-table.dxa <- read_excel("./data/dxa/ribose_dxa.xlsx") %>%
-        select(sex, subject, age, height, weight, leanmasskg, fatmasskg, ffmg) %>%
-        #make grams into kg
-        mutate(ffkg = ffmg/1000) %>%
-        select(sex, age, height, weight, leanmasskg, fatmasskg, ffkg) %>%
-  mutate(bmi = (weight) / (height^2) * 10000)
-
-dxa.table <- table.dxa %>%
-        pivot_longer(names_to = "variable",
-                     values_to = "values", col= age:bmi) %>%
-        select(sex, variable, values) %>%
-        group_by(sex, variable) %>%
-        summarise(m = mean(values, na.rm = TRUE),
-                  s= sd(values, na.rm = TRUE)) %>%
-        mutate(stat = paste0(round(m, 1), " (", round(s, 1), ")")) %>%
-        select(sex, variable, stat) %>%
-        pivot_wider(names_from = sex,
-                    values_from = stat) %>%
-                mutate(variable = if_else(variable %in% c("age"),
-                                          "Age (yrs)",
-                                          if_else(variable %in% c("fatmasskg"),
-                                                  "Fat mass (kg)",
-                                                  if_else(variable %in% c("ffkg"),
-                                                          "Fat free mass (kg)",
-                                                          if_else(variable %in% c("height"),
-                                                                  "Height (cm)",
-                                                                  if_else(variable %in% c("leanmasskg"),
-                                                                          "Lean body mass (kg)",
-                                                                          if_else(variable %in% c("weight"),
-                                                                                  "Body weight (kg)",
-                                                                                  if_else(variable %in% c("bmi"),
-                                                                                          "BMI", "")
-                                                                                  )))))))
-
-dxa.table %>%
-        kable(col.names = c("Variable", "Female (n=7)", "Male (n=9)"))
-
-```
-
-```{r dxa per leg}
-# Script: DXA
-
-# Subject 104, 110 and 111 dropped out and is therefore excluded from the table
-
-#Load dxa data
-dxaleg <-  read_excel("./data/dxa/ribose_dxaprleg.xlsx") %>%
-        select(subject, sex, leg, leanmassg, fatmassg, totalmasskg) %>%
-        #make grams into kg
-        mutate(leanmass = leanmassg/1000,
-               fatmass = fatmassg/1000)%>%
-        select(subject, sex, leg, leanmass, fatmass, totalmasskg)
-
-# Between groups analysis 
-dxa.stat <- dxaleg %>%
-        select(subject, leg, leanmass, fatmass, totalmasskg)
-
-# Lean mass between legs, within subject 
-leanmass.lm <- lmer(leanmass ~ leg + leg:subject + (1|subject), data = dxa.stat)
-lm.plot <- plot(leanmass.lm)
-lm.sum <- summary(leanmass.lm)
-lm.confint <- emmeans(leanmass.lm, specs = ~ "leg")
-
-# Fat mass between legs, within subject
-fat.lm <- lmer(fatmass ~ leg + leg:subject +(1|subject), data = dxa.stat)
-fat.plot <- plot(fat.lm)
-fat.sum <- summary(fat.lm)
-fat.confint <- emmeans(fat.lm, specs = ~ "leg")
-
-# Total mass between legs, within subject
-total.lm <- lmer(totalmasskg ~leg + leg:subject + (1|subject), data = dxa.stat)
-totlm.plot <- plot(total.lm)
-totlm.sum <- summary(total.lm)
-supsup <- emmeans(total.lm, specs = ~ "leg")
-
-# Table of leanmass, fatmass and total mass in kg pr. leg
-dxalegtable <- dxaleg %>%
-        select(leg, leanmass, fatmass, totalmasskg) %>%
-        pivot_longer(names_to = "variable",
-                     values_to = "values", col= leanmass:totalmasskg) %>%
-        group_by(leg, variable) %>%
-        summarise(m = mean(values, na.rm = TRUE),
-                  s= sd(values, na.rm = TRUE)) %>%
-        mutate(stat = paste0(round(m, 1), " (", round(s, 1), ")")) %>%
-        select(leg, variable, stat) %>%
-        pivot_wider(names_from = leg,
-                    values_from = stat)
-
-# Markdown table
-dxalegtable %>%
-        kable(col.names = c("Variable", "Left leg", "Right leg"))
-
-```
-
-Six RT sessions were conducted with glucose and six with placebo, allowing a within-subjects analysis of the effects of glucose ingestion.
-Data from the first five RT sessions was used to investigate main outcome measures (total RNA, rRNA and protein), whereas data from the sixth RT session was used to explore secondary outcomes (muscular recovery, plasma glucose levels).
-Participants were asked to refrain from resistance- or high-intensity training of the legs from Day -7 and onwards, until completion of the intervention and post-testing, to ensure the reliability of pre-intervention strength data and minimal interference from other training sources.
+Six RT sessions were conducted with glucose and six with placebo, allowing a within-subjects analysis of the effects of glucose ingestion before and after RT.
+Data from the first five RT sessions was used to investigate main outcome measures (total RNA, rRNA and protein) and leg muscle strength, whereas data from the sixth RT session was used to explore secondary outcomes (muscular recovery, plasma glucose levels).
+Participants were asked to avoid  resistance- or high-intensity training of the legs from Day -7 (Fig 1A) and onwards, until completion of the intervention and post-testing, to ensure the reliability of pre-intervention strength data and minimal interference from other training sources.
 If participants did conduct exercise training outside of the prescribed protocol, they were asked to ensure equal loading on pairwise consecutive days.
@@
 
-The dietary intervention spanned the whole day, divided into three periods: I) From awakening until two hours after RT, II) from 2.5 hours after RT until 2200hrs, and III) from 2200hrs until awakening.
+The dietary intervention spanned the whole day, divided into three periods: I) From awakening until 2.5 hours after RT, II) from 2.5 hours after RT until 2200hrs, and III) from 2200hrs until awakening.
 During period I, participants ingested protein and glucose/placebo only.
@@
 The self-chosen diet was repeated on pairwise consecutive days (i.e.
-on days 1-2, 3-4, etc.), to ensure similar premises for resistance training responses between the two legs.
+on days 1-2, 3-4, etc.), to ensure similar premises for resistance training responses between treatments.
 During period III (2200-0700hrs), participants remained in an overnight fasted state.
@@
 
-Strength tests (Fig 2) were performed prior to (Days -7 and -5, and -1, both legs) and during the intervention (Days 4 and 8 leg 1, 5 and 9 leg two), after session 5 and after finalization of the intervention (Day 11/12 leg 1 and 12/13 leg 2) (Fig 1).
+Strength tests were performed prior to (Days -7 and -5, and -1, both legs) and during the intervention (Days 4 and 8 leg 1, 5 and 9 leg two), after session 5 and after finalization of the intervention (Day 11/12 leg 1 and 12/13 leg 2) (Fig 1A).
 Maximal isometric and isokinetic knee extension torque were measured with a Humac Norm Dynamometer (CSMi, Stoughton, Massechusetts, USA).
@@
 Isometric peak torque was measured at knee-angle 60, for a maximum of 10 seconds and two repetition per test.
-The highest peak torque values were carried forwards to the final analyses.
-During days 4, 5, 8 and 9, humac tests were conducted one hour before RT with the leg performing RT the previous day.
-During days 11 and 12, humac tests were performed at four time points: I) 45min before RT, II) 30min after the last RT session, III) 2hrs after the last RT session, and IV) 23hrs after the last RT session.
-Test I at day 12/13 included testing of both legs, representing 23hrs post RT session test of one leg and pre-session 6 test of the other leg.
+The highest peak torque values were used in the final analyses.
+During days 4, 5, 8 and 9 (Fig 1D, day 4 and 5 = Post 2RT, day 8 and 9 = Post 4RT), humac tests were conducted one hour before RT with the leg performing RT the previous day.
+During days 11 and 12, humac tests were performed at four time points: I) 45min before RT (1D, Post 5RT), II) 30min after the last RT session (1D, 30 min post 6RT), III) 2hrs after the last RT session (1D, 2h post 6RT), and IV) 23hrs after the last RT session (1D, 23h post 6RT).
+Test I at day 12/13 included testing of both legs, representing 23hrs post RT session test of one leg and post session 5 test of the other leg.
 
-Assessment of unilateral one repetition maximum (1RM) leg press and knee extension was conducted at pre-intervention testing (days -7 and -5, Fig 1).
-The participants performed a general warm-up with 10min cycling on an indoor exercise bicycle and a specific warm-up (1x10, 1x6 and 1x3 repetitions, 2min breaks) before each of the tests.
-Between specific warm-up sets as well as between each subsequent 1RM attempt, the participants were given 3min breaks.
-All positions were controlled and recorded at pre-intervention testing and repeated for the RT sessions.
+Assessment of unilateral one repetition maximum (1RM) leg press and knee extension was conducted at pre-intervention testing during familiariztion (days -7 and -5, Fig 1D).
+The participants performed a general warm-up with 10min cycling on an indoor exercise bicycle.
+In addition, a protocol consisting of 1x10, 1x6 and 1x3 repetitions, was used as spesific warm-up before each of the tests.
+All positions were controlled and recorded at the first 1RM test and repeated for the RT sessions.
 Maximal leg press strength was defined as the maximal load lifted in a controlled fashion, with a knee angle of 45 degrees (start), within a 2.5kg range.
@@
 Attempts with exaggerated hip movement or beneath full extension were not approved.
+Two minute breaks were given during specific warm-up, 3min breaks were given between 1RM attempts.
 
@@
 
-Resistance training consisted of three sets of each unilateral leg presses and knee extensions, with an intensity of 10 repetition maximum (10RM).
+Resistance training consisted of three sets of unilateral leg presses and three sets of knee extensions, with an intensity of 10 repetition maximum (10RM).
 As a general warm-up, the participants cycled on an indoor exercise bicycle for 5-10min.
 In addition, before the respective exercises, two 10-repetition warm-up sets were completed at ~50% and ~70% of 10RM.
-Rest time between working sets was 2min, and progressive loading was used to ensure adequate exercise stimulation throughout the intervention.
+Breaks between working sets was 2min, and progressive loading was used to ensure adequate exercise stimulation throughout the intervention.
 For safety and standardization purposes, all sessions were monitored by trained personnel.
@@
 vastus lateralis using well-established procedures [@hammarstrom.etal_2020].
-Muscle biopsy sampling was performed under local anaesthesia (Xylocaine, 10mg ml-1 with adrenaline 5μg ml−1, AstraZeneca AS, Oslo, Norway) using a 12-14-gauge needle (Universal Plus, Mermaid Medical AS, Stenløse, Denmark), operated with a spring-loaded biopsy gun.
-After the biopsy sampling, muscle tissue was divided in two aliquots for determination of total RNA/expression of rRNA and two aliquots for protein analyses, snap frozen in isopentane (-80°C) and stored at -80°C until further analyses.
-Muscle biopsies were sampled at four time points (Fig 2): I/II) Pre-intervention (2hrs before training, time point T1 = pre RT#1 leg, T2 = pre RT#2 leg), and III/IV) before the sixth RT session from RT#1 leg (2hrs before training, T3) and before the sixth RT session from RT#2 leg (2hrs before training, T4).
+Briefly, muscle biopsy sampling was performed under local anaesthesia (Xylocaine, 10mg ml-1 with adrenaline 5μg ml−1, AstraZeneca AS, Oslo, Norway) using a 12-14-gauge needle (Universal Plus, Mermaid Medical AS, Stenløse, Denmark), operated with a spring-loaded biopsy gun.
+After the biopsy sampling, muscle tissue was divided in two aliquots for determination of total RNA/expression of rRNA and two aliquots for protein content measurement, snap frozen in isopentane (-80°C) and stored at -80°C until further analyses.
+Muscle biopsies were sampled at four time points: I/II) Pre-intervention (Fig 1A, 2hrs before training, Day 1 = leg 1, Day 2 = leg 2), and III/IV) after fifth RT session (1A, Day 11 = leg 1, Day 12 = leg 2), two hours before sixth RT session.
 At each time point, two samples were taken from the same incision.
@@
 To measure blood glucose levels with and without glucose ingestion/training, blood was collected by means of finger draws on days with biopsy sampling.
-One finger draw was taken at T1 (pre-test, Fig 2) to serve as a baseline.
-At T3 and T4 (post-test, Fig 2), finger draws were collected at 7 time points: I) prior to protein ingestion (0700hrs) II) 45 minutes after protein ingestion (0745hrs) III) 1.5hrs after protein ingestion (0800hrs, i.e., immediately before GLU/PLAC intake), IIII) 2hrs after protein ingestion (0900hrs, i.e., immediately before training), IV) in the middle of RT (~0915hrs), V) immediately after training (~0930hrs), and VI) 2hrs after completion of training (~1130hrs).
+One finger draw was taken at day 1 (Fig 1A) to serve as a baseline.
+At day 11 and 12 (Fig 1A), finger draws were collected at 7 time points: I) Immediately prior to protein ingestion (0700hrs) II) 45 minutes after protein ingestion (0745hrs) III) 1.5hrs after protein ingestion (0800hrs, i.e., immediately before GLU/PLAC intake), IIII) 2hrs after protein ingestion (0900hrs, i.e., immediately before training), IV) in the middle of RT (~0915hrs), V) immediately after training (~0930hrs), and VI) 2hrs after completion of training (~1130hrs).
 Finger draws were analysed with in-house equipment (BIOSEN C-Line, EKF diagnostic GmbH, Barleben).
-Venous blood samples were collected from the antecubital vein, coinciding with the finger draws except in the middle of the RT session, to analyse endocrine variables.
+Venous blood samples were collected from the antecubital vein, coinciding with the finger draws except 45min after protein ingestion and in the middle of the RT session, to analyse endocrine variables.
 
@@
 Two muscle biopsy samples were taken for total RNA extraction per leg per time point, and total RNA was extracted in two duplicates per muscle biopsy.
-Thus, we had two duplicates per leg per time point, giving a total of eight RNA samples per participant.
+Thus, we had two duplicates per leg per time point, a total of eight RNA samples per participant.
 Total RNA was extracted using TRIzol according to the manufacturer’s protocol.
@@
 2020; Team 2018).
-For an overview of primers, see Table 2.
 
@@
 
-## Results
+### Data handling and statistical analysis
 
-```{r glucose data + model}
-## Scripts:
-# - gluc.responsechange.R
+A priori power calculations showed that 20 participants would grant a statistical power of 80% (= 0.05), with an expected drop out of 20%.
+This power calculation was based on an assumption that the effects of glucose ingestion on ribosome biogenesis may equate to the effects of increasing RT volume from low to moderate (HAMMARSTROM20).
+<!-- maybe include the power calculation + reason in extended methods?
+-->.
+Raw data was imported an analyzed in R via R studios (SOURCE).
+Legs were categorized as either glucose or placebo in accordance with the randomization.
+Before quantitative analyses, total RNA was normalized by wet muscle weight, qPCR values were normalized by wet muscle weight and an external reference gene (Lambda), while quantified protein signals from western blot were normalized by pool signals and total protein signal per gel <!-- add explanations of normalization to extended methods -->.
+Total RNA and qPCR values were analyzed by comparing post to pre log-fold change score differences.
+This was achieved by modelling change as response with baseline and supplement (glucose/placebo) as explanatory factors, and mean-centered baseline values as a covariate in a linear mixed effects model, then calculating the estimated marginal means <!-- add explanation of model in extended methods-->.
+Plasma glucose and c-peptide, training volume, and the strength index were analyzed by multiple time-point log-fold change score comparisons, using the same linear mixed effects model with time added as explanatory factor <!-- add explanation of model in extended methods-->.
+The linear mixed effects model was designed with the lmer function of the lme4 package (BATES14) in R studios, using the lmerTest function to procure p-values.
+The relationship between
+Log-transformation and mean-centering baseline values was done to control for heteroscedasticity and regression to the mean, respectively.
+Log-transformed values were reverse-transformed for figure illustration.
+Values are presented as mean change + 95% CL unless otherwise stated, alpha-levels was set to α = 0.05.
 
-# Data
-glu.change <- readRDS("./data/data-gen/glucose/glu.change.RDS") # Untransformed data, used in figures
-glu.logchange <- readRDS("./data/data-gen/glucose/glu.logchange.RDS") # Log-transformed data, used in the analysis
-
-## Model
-# Untransformed
-glu.mod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = glu.change)
-glu.res <- plot(glu.mod)
-glu.sum <- summary(glu.mod)
-
-# Log-transformed
-glu.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = glu.logchange)
-glu.lres <- plot(glu.lmod)
-glu.lsum <- summary(glu.lmod)
-
-## Emmeans
-# Untransformed
-glu.emm <- confint(emmeans(glu.mod, specs = ~"supplement|time")) %>%
-        data.frame()
-
-# Log-transformed
-glu.lemm <- confint(emmeans(glu.lmod, specs = ~"supplement|time")) %>%
-        data.frame()
-
-# For inline plotting
-bcoef <- coef(summary(glu.lmod))
-pscient.120 <- round(bcoef[8,5], 30)
-p.120 <- sprintf("%.5f", 1.700971e-16)
-pscient.135 <- round(bcoef[9,5], 30)
-p.135 <- sprintf("%.5f", 4.299568e-11)
-pscient.150 <- round(bcoef[10,5], 30)
-p.150 <- sprintf("%.5f", 1.445712e-11)
-p.270 <- round(bcoef[11,5], 2)
-```
-
-```{r c-peptide data + model}
-## Scripts
-# - insulin.change.R
-
-## Data
-cpep.change <- readRDS("./data/data-gen/glucose/cpep.change.RDS") # Untransformed data, for figures
-cpep.lchange <- readRDS("./data/data-gen/glucose/cpep.logchange.RDS") # Log-transformed data, for analysis
-
-## Linear mixed effects model
-# Untransformed
-cpep.mod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = cpep.change) 
-cpep.res <- plot(cpep.mod)
-cpep.sum <- summary(cpep.mod)
-
-# Log-transformed
-cpep.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = cpep.lchange) 
-cpep.lres <- plot(cpep.lmod)
-cpep.lsum <- summary(cpep.lmod)
-
-## Emmeans
-# Untransformed
-cpep.emm <- confint(emmeans(cpep.mod, specs = ~"supplement|time")) %>%
-  data.frame()
-
-# Log-transformed
-cpep.lemm <- confint(emmeans(cpep.lmod, specs = ~"supplement|time")) %>%
-  data.frame()
-
-# For inline plotting
-cpcoef <- coef(summary(cpep.lmod))
-pscient.120 <- round(cpcoef[6,5], 30)
-cpp.120 <- sprintf("%.5f", 3.229267e-15)
-cpscient.150 <- round(cpcoef[7,5], 30)
-cpp.150 <- sprintf("%.5f", 1.414491e-15)
-cpp.270 <- round(cpcoef[8,5], 2)
-
-```
-
-At baseline, there were no significant differences between training with glucose or placebo in muscle mass, isometric or isokinetic peak torque, total session volume, training intensity, total RNA content, or rRNA expression (Tab 3).
-Further, on pairwise consecutive days, there were no significant differences in mean macro nutrient (calories, carbohydrate, fat or protein) ingestion (Tab 4), or mean change in macro nutrient ingestion (Fig 3).
-Resistance training with glucose led to significant increases in blood levels glucose and c-peptide compared to RT with placebo, before, during and after the session (Fig 4B, 120min: 38 ± 4% [*p* = `r p.120`], 135min: 31 ± 4% [*p* = `r p.135`], 150min: 32 ± 4% [*p* = `r p.150`], Fig 4C, 120min: 95 ± 10% [*p* = `r cpp.120`], 150min: 87 ± 10% [*p* = `r cpp.150`]), measured during the final RT session.
-At 2hrs post-exercise, RT with glucose led to significantly lower blood glucose levels compared to RT with placebo, with a tendancy of the same in c-peptide levels (4B, 270min: -8 ± 4% [*p* = `r p.270`] and 4C, 270min: -7 ± 10 [*p* `r cpp.270`]).
-
-###### **Table 3**: Baseline characteristics of the intervention legs, organized per supplement.
-Variables: Iso 60 = isokinetic 60 d/s, Iso 240 = isokinetic 240 d/s, training intensity = load/rm*100.
-Nm = newton meters, kg = kilograms, %1RM = percentage of 1 repetition maximum, ng/mg = nanogram per milligram wet muscle tissue, nf.expr = expression normalized by weight and external reference gene (Lambda).
-Ribosomal RNA data are presented as log transformed.
-Glucose n = 13, placebo n = 13.
-
-![](figures/table3.png)
-
-###### **Table 4:** Daily mean dietary intake of macro nutrients during each training day per supplement leg.
-G = glucose, P = placebo.
-Training day 1-training day 12 are combined pairwise to day 1-6: Day 1 = training day 1/2, Day 2 = training day 3/4, Day 3 = training day 5/6, Day 4 = training day 7/8, Day 5 = training day 9/10, Day 6 = training day 11/12.
-Values are mean ± standard deviation.
-P = difference between glucose and placebo.
-Glucose n = 12, placebo n = 12.
-Day 6: Glu = 11, Plac = 11.
-
-```{r nutrition table}
-## Scripts
-# - nut.table.R
-# - nut.analysis.R
-# - dxa.R
-
-# Data
-nuttable <- readRDS("./data/data-gen/dxa/nuttable.RDS")
-
-nuttable %>%
-  kable(col.names = c("Time", "Suppl.", "Kcal", "CHO (g)", "Fat (g)", "PRO (g)", "PRO (g/kg)"))
-
-```
-
-```{r nutrition fig}
-# Associated scripts:
-# - nut.datahand.R
-# - nut.change.R
-# - nut.change.figs.R
-
-nut.fig <- readRDS("./data/data-gen/nutrition/nut.fig.RDS")
-nut.fig
-
-```
-###### **Figure 3:** Mean changes in macro nutrient intake from day to day.
-Day 1 = training day 1/2(T1/T2), Day 2 = training day 3/4, Day 3 = training day 5/6, Day 4 = training day 7/8, Day 5 = training day 9/10, Day 6 = training day 11/12(T3/T4).
-Values are mean change in grams/calories ± 95% CL.
-
-![](figures/day.timeline.png)
-
-```{r glucose + c-peptide fig}
-## Scripts
-# - gluc.fig.R
-# - cpep.figs.R
-
-# Separate figures
-pos <- position_dodge(width = 0.2)
-
-gluc.fig <- readRDS("./data/data-gen/glucose/gluc.fig.RDS")
-
-cpep.fig <- readRDS("./data/data-gen/glucose/cpep.fig.RDS")
-
-# Joined figure
-plot_grid(gluc.fig,
-          cpep.fig,
-          ncol = 1, nrow = 2)
-
-```
-
-###### **Figure 4:** A) Time line for training day 6 (combination of pairwise training day 11 and 12).
-Biopsy = muscle biopsy taken pre supplementation, PRO = protein supplementation, Humac = strength test (isometric knee extension, isokinetic knee extension 60 and 240 d/s), GLU/PLAC = glucose/placebo supplementation, RT = resistance training.
-Minutes: -120 = 2hrs prior to RT, -90 = 90min prior to RT, -30 = 30min prior to RT, 0 = onset of RT, 15 = 15min following onset of RT, between leg press and knee extension, 30 = 30min post onset of RT, 120 = 2hrs following onset of RT.
-B) Mean mmol/L change in blood glucose levels per supplement group during training day 6.
-Time-points: 0, 45, 90, 120, 135, 150 and 270min post protein ingestion, respectively.
-C) Mean change in pmol/L of c-peptide in blood per supplement group during training day 6.
-Values are presented as estimated marginal means of mmol/L and pmol/L change ± 95% CI.
-*p* < 0.05 between groups.
-Glucose n = 13, placebo n = 13.
-
-### Markers of ribosome biogenesis
-
-```{r total RNA model}
-## Scripts
-# - totrna.cleanup.R
-# - totrna.analysis.R
-
-# Data
-trna.lchange <- readRDS("./data/data-gen/rna/totrna.change.RDS")
-
-# Create model: 
-trna.lmod <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
-           data = trna.lchange)
-
-trna.lres <- plot(trna.lmod)
-trna.lsum <- summary(trna.lmod)
-
-### Get estimated means from the model
-trna.emm <- confint(emmeans(trna.lmod, specs = ~"supplement")) %>%
-    data.frame()
-
-# For inline reporting
-trcoef <- coef(summary(trna.lmod))
-trdiff <- round(trcoef[3,1], 2)
-trdiffp <- round(trcoef[3,5], 2)
-trdiffse <- round(trcoef[3,2], 2)
-
-```
-
-```{r rRNA data model}
-## Scripts
-# - qpcr-import
-# - qpcr-analysis
-# - rrna.change
-# - rrnachange.figs
-
-# Log-transformed data
-lchange.18 <- readRDS("./data/data-gen/rna/lchange.18.RDS")
-lchange.28 <- readRDS("./data/data-gen/rna/lchange.28.RDS")
-lchange.58 <- readRDS("./data/data-gen/rna/lchange.58.RDS")
-lchange.5 <- readRDS("./data/data-gen/rna/lchange.5.RDS")
-lchange.47 <- readRDS("./data/data-gen/rna/lchange.47.RDS")
-
-# Create model: 
-# Needs to have an intercept per participant (mixed model)
-# Control for pre values.
-
-# 18S
-lmod.18 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
-                     data = lchange.18)
-lres.18 <- plot(lmod.18)
-lsum.18 <- summary(lmod.18)
-
-# 28S
-lmod.28 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
-                     data = lchange.28)
-lres.28 <- plot(lmod.28)
-lsum.28 <- summary(lmod.28)
-
-# 5.8S
-lmod.58 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
-                     data = lchange.58)
-lres.58 <- plot(lmod.58)
-lsum.58 <- summary(lmod.58)
-
-# 5S
-lmod.5 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
-                     data = lchange.5)
-lres.5 <- plot(lmod.5)
-lsum.5 <- summary(lmod.5)
-
-# 47S
-lmod.47 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
-                     data = lchange.47)
-lres.47 <- plot(lmod.47)
-lsum.47 <- summary(lmod.47)
-
-### Get estimated means from the model, these are average increase at 
-# pre = 0 (the average pre value)
-# remember that these are originally log-fold change values, transformed back using exp(emmean) etc
-
-# 18S
-lemm.18 <- confint(emmeans(lmod.18, specs = ~"supplement")) %>%
-        data.frame()
-
-# 28S
-lemm.28 <- confint(emmeans(lmod.28, specs = ~"supplement")) %>%
-        data.frame()
-
-# 5.8S
-lemm.58 <- confint(emmeans(lmod.58, specs = ~"supplement")) %>%
-        data.frame()
-
-# 5S
-lemm.5 <- confint(emmeans(lmod.5, specs = ~"supplement")) %>%
-        data.frame()
-
-# 47S
-lemm.47 <- confint(emmeans(lmod.47, specs = ~"supplement")) %>%
-        data.frame()
-
-## For inline reporting
-
-# 18S
-rr18coef <- coef(summary(lmod.18))
-rr18diffp <- round(rr18coef[3,5], 3)
-rr18diffse <- round(rr18coef[3,2], 3)
-
-# 28S
-rr28coef <- coef(summary(lmod.28))
-rr28diffp <- round(rr28coef[3,5], 3)
-rr28diffse <- round(rr28coef[3,2], 3)
-
-# 5.8S
-rr58coef <- coef(summary(lmod.58))
-rr58diffp <- round(rr58coef[3,5], 3)
-rr58diffse <- round(rr58coef[3,2], 3)
-
-# 5S
-rr5coef <- coef(summary(lmod.5))
-rr5diffp <- round(rr5coef[3,5], 3)
-rr5diffse <- round(rr5coef[3,2], 3)
-
-# 47S
-rr47coef <- coef(summary(lmod.47))
-rr47diffp <- round(rr47coef[3,5], 3)
-rr47tdiffp <- round(rr47coef[1,5], 3)
-rr47diffse <- round(rr47coef[3,2], 3)
-
-```
-
-Resistance training with glucose did not induce a higher accumulation of total RNA (fig 5A, *p* = `r trdiffp`), or rRNA  (fig 5B, *p* > 0.05), compared to RT with placebo, measured as mean change from baseline to post.
-From baseline to post-intervention, there was a mean increase in levels of total RNA by 26 and 22% after RT with glucose and placebo, respectively (*p* < 0.05, compared to baseline).
-A robust accumulation was also observed in ribosomal RNA expression in both RT with glucose and placebo, with mean increases between 34-43% (GLU) and 33-41% (PLAC) (*p* < 0.05, compared to baseline) in the four rRNA’s.
-The expression of 47S pre-rRNA increased by 37 and 59% with glucose and placebo respectively, with only placebo reaching significantly higher levels at post compared to baseline (fig 45, *p* = `r rr47tdiffp`).
-
-```{r RNA figures}
-# Get figures
-totrna.fig <- readRDS("./data/data-gen/rna/totrna.fig.RDS")
-totrna.fig
-
-plot.18 <- readRDS("./data/data-gen/rna/plot.18.RDS")
-plot.28 <- readRDS("./data/data-gen/rna/plot.28.RDS")
-plot.5.8 <- readRDS("./data/data-gen/rna/plot.5.8.RDS")
-plot.5 <- readRDS("./data/data-gen/rna/plot.5.RDS")
-plot.47 <- readRDS("./data/data-gen/rna/plot.47.RDS")
-
-# Joined rRNA
-library(cowplot)
-
-legend <- get_legend(plot.47 + theme(legend.box.margin = margin(0, 0, 0,12)))
-
-plot_grid(plot.18 + theme(legend.position = "none"),
-          plot.28 + theme(legend.position = "none"), 
-          plot.5.8 + theme(legend.position = "none"),
-          plot.5 + theme(legend.position = "none"), 
-          plot.47 + theme(legend.position = "none"),
-          legend,
-          ncol = 2, nrow = 3)
-
-```
-
-###### **Figure 5:** A) Changes in Total RNA, B) Changes in ribosomal RNA.
-Baseline = Training day 1/2, Post = Training day 11/12.
-Both total RNA and rRNA were analyzed with duplicates, with two duplicates per biopsy (two biopsies per time point).
-Total RNA and rRNA changes were calculated as log-fold change score per mg wet muscle weight.
-Mean change scores of the duplicates were calculated before modelling and transformed to the log-scale.
-Values are estimated marginal means fold change per leg/supplement ± 95% CI.
-p = between groups statistic, glucose compared to placebo.
-Glucose n = 13, placebo n = 13.
-
-```{r western modelling}
-## Script: 
-# - western.import.R
-# - western.change.R
-
-# Data
-
-prot.dat <- readRDS("./data/data-gen/protein/prot.dat.RDS") # cleaned up and quantified western data
-prot.out <- readRDS("./data/data-gen/protein/protein_complete.RDS") # Cleaned up and quantified western data, checked for outliers
-prot.change <- readRDS("./data/data-gen/protein/prot.change.RDS") # change-score data
-
-# By target
-
-cmyc <- prot.change %>%
-  filter(target == "cmyc")
-
-ubf <- prot.change %>%
-  filter(target == "ubf")
-
-rps6 <- prot.change %>%
-  filter(target == "rps6")
-
-# c-Myc change
-
-m1 <- lme(change ~ pre + supplement,
-            random = list(subject = ~ 1),
-            data = cmyc) # same model as in "western import", but with change scores of norm.sign
-
-#plot(m1)
-
-sum.m1 <- summary(m1)
-
-cmyc.emm <- emmeans(m1, specs = ~"supplement") %>%
-  data.frame
-
-saveRDS(cmyc.emm, "./data/data-gen/protein/cmyc.change.RDS")
-
-# UBF change
-
-m2 <- lme(change ~ pre + supplement,
-            random = list(subject = ~ 1),
-            data = ubf) # same model as in "western import", but with change (i.e., change in norm.sign) instead of norm.sign
-
-#plot(m2)
-
-sum.m2 <- summary(m2)
-
-ubf.emm <- emmeans(m2, specs = ~"supplement") %>%
-  data.frame()
-
-saveRDS(ubf.emm, "./data/data-gen/protein/ubf.change.RDS")
-
-# rps6 change
-
-m3 <- lme(change ~ pre + supplement,
-            random = list(subject = ~ 1),
-            data = rps6) # same model as in "western import", but with change (i.e., change in norm.sign) instead of norm.sign
-
-#plot(m3)
-
-sum.m3 <- summary(m3)
-
-rps6.emm <- emmeans(m3, specs = ~"supplement") %>%
-  data.frame()
-
-saveRDS(rps6.emm, "./data/data-gen/protein/rps6.change.RDS")
-
-### For inline reporting
-## Diff cmyc
-ccoef <- coef(summary(m1))
-cdiff <- round(ccoef[3,5], 3)
-cdiffse <- round(ccoef[3,2], 3)
-
-## Diff UBF
-ucoef <- coef(summary(m2))
-udiff <- round(ucoef[3,5], 3)
-udiffse <- round(ucoef[3,2], 3)
-
-## Diff rps6
-rpcoef <- coef(summary(m3))
-rpdiff <- round(rpcoef[3,5], 3)
-rpdiffse <- round(rpcoef[3,2], 3)
-
-```
-
-```{r western fig}
-# Data
-
-cmyc <- readRDS("./data/data-gen/protein/cmyc.change.RDS")
-
-ubf <- readRDS("./data/data-gen/protein/ubf.change.RDS")
-
-rps6 <- readRDS("./data/data-gen/protein/rps6.change.RDS")
-
-## cmyc fig
-
-cmyc.plot <- cmyc %>%
-  mutate(time = "post") %>%
-  add_row(supplement = c("PLACEBO", "GLUCOSE"), time = "pre", emmean = 0, lower.CL = 0, upper.CL = 0) %>%
-  mutate(time = factor(time, levels = c("pre", "post"), labels = c("Baseline", "Post"))) %>%
-  ggplot(aes(time, exp(emmean), group = supplement, fill = supplement)) +
-  #annotate("text", x = "Post", y = 2.1, label = "p = 0.585", size = 3) +
-  geom_errorbar(aes(ymin = exp(lower.CL), ymax = exp(upper.CL)),
-                width = 0.1,
-                position = position_dodge(width = 0.2)) +
-  geom_line(position = position_dodge(width = 0.2)) +
-  geom_point(shape = 21, size = 3, position = position_dodge(width = 0.2)) +
-  labs(x = "", y = "c-Myc signal \n(fold change)\n", fill = "") +
-  theme_classic()
-
-ubf.plot <- ubf %>%
-  mutate(time = "post") %>%
-  add_row(supplement = c("PLACEBO", "GLUCOSE"), time = "pre", emmean = 0, lower.CL = 0, upper.CL = 0) %>%
-  mutate(time = factor(time, levels = c("pre", "post"), labels = c("Baseline", "Post"))) %>%
-  ggplot(aes(time, exp(emmean), group = supplement, fill = supplement)) +
-  #annotate("text", x = "Post", y = 2.1, label = "p = 0.585", size = 3) +
-  geom_errorbar(aes(ymin = exp(lower.CL), ymax = exp(upper.CL)),
-                width = 0.1,
-                position = position_dodge(width = 0.2)) +
-  geom_line(position = position_dodge(width = 0.2)) +
-  geom_point(shape = 21, size = 3, position = position_dodge(width = 0.2)) +
-  labs(x = "", y = "UBF signal \n(fold change)\n", fill = "") +
-  theme_classic()
-
-rps6.plot <- rps6 %>%
-  mutate(time = "post") %>%
-  add_row(supplement = c("PLACEBO", "GLUCOSE"), time = "pre", emmean = 0, lower.CL = 0, upper.CL = 0) %>%
-  mutate(time = factor(time, levels = c("pre", "post"), labels = c("Baseline", "Post"))) %>%
-  ggplot(aes(time, exp(emmean), group = supplement, fill = supplement)) +
-  annotate("text", x = "Post", y = 1.9, label = "p = ", size = 3) +
-  geom_errorbar(aes(ymin = exp(lower.CL), ymax = exp(upper.CL)),
-                width = 0.1,
-                position = position_dodge(width = 0.2)) +
-  geom_line(position = position_dodge(width = 0.2)) +
-  geom_point(shape = 21, size = 3, position = position_dodge(width = 0.2)) +
-  labs(x = "", y = "rps6 signal \n(fold change)\n", fill = "") +
-  theme_classic()
-
-# Cowplot for gathering figures
-
-legend <- get_legend(rps6.plot + theme(legend.box.margin = margin(0, 0, 0,12)))
-
-prot.joined <- plot_grid(cmyc.plot + theme(legend.position = "none"),
-                         ubf.plot + theme(legend.position = "none"), 
-                         rps6.plot + theme(legend.position = "none"),
-                         legend,
-                         ncol = 2, nrow = 2)
-
-prot.joined
-
-saveRDS(prot.joined, "./data/data-gen/rna/prot.joined.RDS")
-
-```
-
-###### **Figure 5:** A) Changes in c-Myc levels, B) changes in UBF levels, C) changes in rps6 levels.
-Baseline = Training day 1/2, post = Training day 11/12.
-Four samples were analysed per participant (GLU pre-post, PLAC pre-post), in duplicates, a total of 104 protein samples.
-Duplicate samples were loaded on separate gels.
-In addition, 3-4 pools were added to each well.
-c-Myc, UBF and rps6 signals were then calculated as log-fold change score, normalised by gel pools.
-Mean change scores of the duplicates were calculated before modelling and transformed to the log-scale.
-Values are estimated marginal means fold change per leg/supplement ± 95% CI.
-p = between groups statistic, glucose compared to placebo.
-Glucose n = 13, placebo n = 13.
-
-### Resistance training and strength testing
-
-```{r training data model}
-## Scripts
-# - volume.change.R
-# - ss.analysis.R
-# - ss.cleanup.R
-# - trainint.analysis.R
-# - trainint.cleanup.R
-
-# Log-transformed data
-vol.lchange <- readRDS("./data/data-gen/training/vol.lchange.RDS")
-ss.lchange <- readRDS("./data/data-gen/training/ss.lchange.RDS")
-int.lchange <- readRDS("./data/data-gen/training/int.lchange.RDS")
-
-### Training volume
-# Fold-change model
-vol.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = vol.lchange)
-vol.lres <- plot(vol.lmod)
-vol.lsum <- summary(vol.lmod)
-
-# Fold-change estimated means
-vol.lemm <- confint(emmeans(vol.lmod, specs = ~"supplement|time")) %>%
-        data.frame()
-
-### Session score (RPE)
-# Model
-ss.lmod <- lmerTest::lmer(change ~ baseline + time + supplement:time + (1|subject),
-                     data = ss.lchange)
-ss.lres <- plot(ss.lmod)
-ss.lsum <- summary(ss.lmod)
-
-# Session score emmeans
-ss.lemm <- confint(emmeans(ss.lmod, specs = ~"supplement|time")) %>%
-        data.frame()
-
-### Training intensity
-int.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = int.lchange)
-int.lres <- plot(int.lmod)
-int.lsum <- summary(int.lmod)
-
-# Emmeans
-int.lemm <- confint(emmeans(int.lmod, specs = ~"supplement|time")) %>%
-        data.frame()
-
-### For inline reporting
-
-## Training volume
-# Diff 4
-vcoef <- coef(summary(vol.lmod))
-vdiff.4 <- round(vcoef[9,1], 3)
-vdiffp.4 <- round(vcoef[9,5], 3)
-vdiffse.4 <- round(vcoef[9,2], 3)
-# Diff 5
-vdiff.5 <- round(vcoef[10,1], 3)
-vdiffp.5 <- round(vcoef[10,5], 3)
-vdiffse.5 <- round(vcoef[10,2], 3)
-# Diff 6
-vdiff.6 <- round(vcoef[11,1], 3)
-vdiffp.6 <- round(vcoef[11,5], 3)
-vdiffse.6 <- round(vcoef[11,2], 3)
-## RPE
-scoef <- coef(summary(ss.lmod))
-sdiffp <- round(scoef[11, 5], 3)
-sdiffse <- round(scoef[11, 2], 3)
-
-## Training intensity
-
-# Diff time 2
-icoef <- coef(summary(int.lmod))
-idiff2 <- round(icoef[7,1], 3)
-idiffp2 <- round(icoef[7,5], 3)
-idiffse2 <- round(icoef[7,2], 3)
-# Diff time 3
-idiff3 <- round(icoef[8,1], 3)
-idiffp3 <- round(icoef[8,5], 3)
-idiffse3 <- round(icoef[8,2], 3)
-# Diff time 4
-idiff4 <- round(icoef[9,1], 3)
-idiffp4 <- round(icoef[9,5], 3)
-idiffse4 <- round(icoef[9,2], 3)
-# Diff time 5
-idiff5 <- round(icoef[10,1], 3)
-idiffp5 <- round(icoef[10,5], 3)
-idiffse5 <- round(icoef[10,2], 3)
-# Diff time 6
-idiff6 <- round(icoef[11,1], 3)
-idiffp6 <- round(icoef[11,5], 3)
-idiffse6 <- round(icoef[11,2], 3)
-```
-
-```{r humac post 5th model}
-## Scripts
-# - humac.cleanup.R
-# - humac.main.analysis.R - this is data gathered at baseline, during the intervention, and after session 5
-
-# Log-transformed data
-
-isom.lchange <- readRDS("./data/data-gen/humac/isom.lchange.RDS")
-iso60.lchange <- readRDS("./data/data-gen/humac/isok60.lchange.RDS")
-iso240.lchange <- readRDS("./data/data-gen/humac/isok240.lchange.RDS")
-
-# Isometric
-isom.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = isom.lchange)
-isom.lres <- plot(isom.lmod)
-isom.lsum <- summary(isom.lmod)
-
-# Isok.60
-iso60.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = iso60.lchange)
-iso60.lres <- plot(iso60.lmod)
-iso60.lsum <- summary(iso60.lmod)
-
-# Isok.240
-iso240.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = iso240.lchange)
-iso240.lres <- plot(iso240.lmod)
-iso240.lsum <- summary(iso240.lmod)
-
-## Emmeans
-# Isometric
-isom.lemm <- confint(emmeans(isom.lmod, specs = ~"supplement|time")) %>%
-        data.frame()
-
-# Isok.60
-iso60.lemm <- confint(emmeans(iso60.lmod, specs = ~"supplement|time")) %>%
-        data.frame()
-
-# Isok.240
-iso240.lemm <- confint(emmeans(iso240.lmod, specs = ~"supplement|time")) %>%
-        data.frame()
-
-# For reporting
-# Isom diff
-isom5coef <- coef(summary(isom.lmod))
-isom5diffp <- round(isom5coef[7, 5], 3)
-isom5diffse <- round(isom5coef[7, 2], 3)
-
-# Isok 60 diff
-isok605coef <- coef(summary(iso60.lmod))
-isok605diffp <- round(isok605coef[7, 5], 3)
-isok605diffse <- round(isok605coef[7, 2], 3)
-
-# Isok 260
-isok2405coef <- coef(summary(iso240.lmod))
-isok2405diffp <- round(isok2405coef[7, 5], 3)
-isok2405diffse <- round(isok2405coef[7, 2], 3)
-
-```
-
-```{r post 6th model}
-## Scripts
-# - humac.acute.analysis.R - this is data gatherede before and after one RT session (sixth session)
-# - humac.cleanup.R
-
-# Data
-isom.lchange.ac <- readRDS("./data/data-gen/humac/isom.lchange.ac.RDS")
-iso60.lchange.ac <- readRDS("./data/data-gen/humac/iso60.lchange.ac.RDS")
-iso240.lchange.ac <- readRDS("./data/data-gen/humac/iso240.lchange.ac.RDS")
-
-# Isometric
-isomac.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = isom.lchange.ac)
-isomac.lres <- plot(isomac.lmod)
-isomac.lsum <- summary(isomac.lmod)
-
-# Isok.60
-iso60ac.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = iso60.lchange.ac)
-iso60ac.lres <- plot(iso60ac.lmod)
-iso60ac.lsum <- summary(iso60ac.lmod)
-
-# Isok.240
-iso240ac.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = iso240.lchange.ac)
-iso240ac.lres <- plot(iso240ac.lmod)
-iso240ac.lsum <- summary(iso240ac.lmod)
-
-# Emmeans
-# Isometric
-isomac.lemm <- confint(emmeans(isomac.lmod, specs = ~"supplement|time")) %>%
-        data.frame()
-
-# Isok.60
-iso60ac.lemm <- confint(emmeans(iso60ac.lmod, specs = ~"supplement|time")) %>%
-        data.frame()
-
-# Isok.240
-iso240.lemm <- confint(emmeans(iso240ac.lmod, specs = ~"supplement|time")) %>%
-        data.frame()
-
-## For inline reporting
-# Isom diff
-isomcoef <- coef(summary(isomac.lmod))
-isomdiffp <- round(isomcoef[7, 5], 3)
-isomdiffse <- round(isomcoef[7, 2], 3)
-
-# Isok 60 diff
-isok60coef <- coef(summary(iso60ac.lmod))
-isok60diffp <- round(isok60coef[7, 5], 3)
-isok60diffse <- round(isok60coef[7, 2], 3)
-
-# Isok 240 diff
-isok240coef <- coef(summary(iso240ac.lmod))
-isok240diffp <- round(isok240coef[7, 5], 3)
-isok240diffse <- round(isok240coef[7, 2], 3)
-
-```
-
-There were no significant differences in mean change of total session volume, training intensity or RPE between exercise training with glucose and placebo at any of the time points (fig 6, *p* > 0.05).
-We did observe interactions between time and changes, where total session training volume increased from 13-18% compared to baseline.
-These increases in training volume were accompanied by changes ranging from 5-18% in training intensity and 2-13% in RPE (Tab 5).
-The maximal isometric knee extension force (OR TORQUE?) and isokinetic knee extension peak torque at 240 d/s showed no differences between exercise training with glucose and placebo after five RT sessions (Fig 7A, *p* > 0.05).
-There was a significant difference in mean change from pre to post-fifth RT session, where exercise training with glucose led to a 9% less decrease in peak torque at 60 d/s compared to placebo (Fig 7A, *p* = `r isok605diffp`).
-There was no difference between glucose and placebo in mean change from baseline to after the sixth RT session (Fig 7B, *p* > 0.05).
-
-![](figures/legend.PNG)
-
-```{r training dat fig}
-### Get figures
-
-# Volume
-totvol.barplot <- readRDS("./data/data-gen/training/totvol.barplot.RDS")
-totvolfold.plot <- readRDS("./data/data-gen/training/totvolfold.plot.RDS")
-
-# Training intensity
-tot.barplot <- readRDS("./data/data-gen/training/tot.barplot.RDS")
-totalint.plot <- readRDS("./data/data-gen/training/trainint.plot.RDS")
-
-# Join figures
-plot_grid(totvol.barplot + theme(legend.position = "none"),
-          totvolfold.plot + theme(legend.position = "none"),
-          tot.barplot + theme(legend.position = "none"),
-          totalint.plot + theme(legend.position = "none"),
-          ncol = 2,
-          nrow = 2,
-          align = "h")
-
-```
-
-###### **Figure 6:** Training volume and training intensity from baseline until resistance exercise session 6 (training day 11 and 12).
-Baseline = Training day 1/2, Session 2 = Training day 3/4, Session 3 = Training day 5/6, Session 4 = Training day 7/8, Session 5 = Training day 9/10, Session 6 = Training day 11/12.
-A) Changes in total session volume per leg, values are mean kg ± standard deviation.
-B) Fold change in total session volume per leg, values are estimated marginal means fold change ± 95% CI.
-C) Changes in training intensity per leg calculated as %1RM, values are mean % ± standard deviation.
-D) Fold change in training intensity per leg, values are estimated marginal means fold change ± 95% CI.
-† = *p* < 0.05 in both glucose and placebo, compared to baseline.
-p = NS = no significant difference between legs at any of the time points.
-Glucose n = 13, placebo n = 13.
-
-###### **Table 5:** RPE changes through training days.
-Training days are presented as combined pairwise consecutive, linked to their respective sessions.
-Session 1 = Training day 1/2, Session 2 = Training day 3/4, Session 3 = Training day 5/6, Session 4 = Training day 7/8, Session 5 = Training day 9/10, Session 6 = Training day 11/12.
-Values are presented as mean ± standard deviation.
-Glucose n = 13, placebo n = 13.
-
-```{r RPE table}
-## Scripts
-# - ss.cleanup.R
-# - ss.absolute.R
-
-# Data
-ss.abs <- readRDS("./data/data-gen/training/ss.abs.RDS")
-
-# Table
-ss.abs %>%
-  kable(col.names = c("Supplement", "Baseline", "Session 2", "Session 3", "Session 4", "Session 5",
-                            "Session 6"))
-
-```
-
-![](figures/legend2.PNG)
-
-```{r humac figs}
-## Post 5th RE session
-# Figures
-pos <- position_dodge(width = 0.2)
-isom5.plot <- readRDS("./data/data-gen/humac/isom.main.fig.RDS")
-isok605.plot <- readRDS("./data/data-gen/humac/isok60.main.fig.RDS")
-isok2405.plot <- readRDS("./data/data-gen/humac/isok240.main.fig.RDS")
-
-legend <- get_legend(isom5.plot + theme(legend.box.margin = margin(0, 0, 0, 12)))
-
-plot_grid(isom5.plot + theme(legend.position = "none"), 
-          isok605.plot + theme(legend.position = "none"), 
-          isok2405.plot + theme(legend.position = "none"),
-          nrow = 3,
-          align = "v")
-
-## Post 6th RE session
-
-# Figures
-
-pos <- position_dodge(width = 0.2)
-
-isom.plot <- readRDS("./data/data-gen/humac/isom.ac.plot.RDS")
-isok60.plot <- readRDS("./data/data-gen/humac/isok60.ac.plot.RDS")
-isok240.plot <- readRDS("./data/data-gen/humac/isok240.ac.plot.RDS")
-
-legend <- get_legend(isom.plot + theme(legend.box.margin = margin(0, 0, 0, 12)))
-
-plot_grid(isom.plot + theme(legend.position = "none"), 
-          isok60.plot + theme(legend.position = "none"), 
-          isok240.plot + theme(legend.position = "none"),
-          nrow = 3,
-          align = "v")
-
-```
-
-The mean change in peak torque at all three velocities change little from baseline until session 4 in both glucose and placebo (+4-7% change), decreasing at pre-session 6 (9-18% change, Fig 6A).
-Post session 6 (Fig 6B) showed a slight reduction in isometric peak torque 30min after session completion (2-4% change), slowly regaining at 2hrs (+3% change) and 23hrs (+3-6% change), in both glucose and placebo compared to baseline.
-Isokinetic peak torque at 60 d/s and 240 d/s showed no reduction in mean peak torque from baseline until 23hrs following RT,  in both glucose and placebo (+0,1-11%).
-There was an effect of time in peak torque at 23hrs post-RT at 60 d/s (Fig 6B), where exercise training with placebo led to a 5% increase in peak torque compared to baseline, however, there was no significant difference compared to exercise training with glucose.
-
-###### **Figure 7:** Peak torque changes in maximal isometric knee extension and maximal isokinetic knee extension at 60 and 240 d/s, respectively.
-A) Humac tests during the intervention, Baseline = Day -1 (prior to intervention), Session 2 = prior to training, training Day 4/5, Session 4 = prior training, training Day 8/9, Session 6 = prior to training 11/12.
-B) Baseline = prior to training, training day 11/12, 30min = 30 min after RT, 2hr = 2hrs after RT, 23hr = 23hrs after RT.
-Values are change in estimated marginal means of newton meters (Nm) ± 95% CI.
-* = *p* < 0.05 between groups.
-*p* = NS = no significant difference between legs at any of the time points.
-Glucose n = 13, placebo n = 13.
-
-## Discussion
-
-The main findings of the present study were that resistance training with glucose for five sessions did not affect markers of ribosome biogenesis, measured as total RNA, rRNA and rDNA transcription initiation associated proteins, compared to RT with placebo.
-Similarly, RT with glucose did not affect markers of skeletal muscle functionality such as muscular performance and -recovery, total session volume, and training intensity, compared to RT with placebo.
-There are studies suggesting that high vs.
-low glucose treatment/glucose starvation impacts ribosome biogenesis in yeast, rodent and human non-muscle cells (KIM13;MARIAPPAN11;TANAKA15;ZHAI12), however, this effect does not seem evident when total daily glucose intake is matched.
-
-To our knowledge, no previous study has investigated the effects of glucose compared to placebo ingestion with RT on ribosome biogenesis in human skeletal muscle cells, making our hypotheses exploratory and comparisons not straight forwards.
-The current model of ribosome biogenesis suggests that changes in the accumulation of total RNA and rRNA early after the onset of an RT period predict long-term RT adaptations (FIGUEIREDO19; HAMMARSTROM20; HAMMARSTROM21).Therefore, we assumed 2 weeks to be a suitable time frame for a moderate RT volume to elicit changes in ribosome biogenesis, as similar observations have been made previously (FIGUEIREDO16; HAMMARSTROM20; HAMMARSTROM21).
-In Hammarström et al.
-(2020), participants that increased ribosome biogenesis after two weeks (5 RT sessions) of RT, also responded well after the entire intervention (12 weeks, 31 sessions).
-Figueiredo and colleagues (2016) also observed markedly acute effects, where 47S pre-rRNA increased after a single RT session, passing baseline after 4hrs, peaking at 24hrs and remaining elevated until 48hrs post-exercise.
-The four mature rRNA did not measurably increase until after multiple sessions (FIGUEIREDO16).
-As such, increases in total RNA and rRNA observed in the present study are in line with previous observations (FIGUEIREDO16; HAMMARSTROM20; H21?), and also evident i moderately trained individuals, compared to the untrained or RT-untrained individuals of previous studies (FIGUEIREDO16;HAMMARSTROM20;H21?).
-
-Further, the accumulation of total RNA and rRNA is also suggested to reach a plateau after approximately 5 RT-sessions, thus, in addition to the demanding RT- and dietary protocol and the exploratory nature of the hypotheses, we did not conduct a more long-term intervention.
-COMMENT ON PROTEIN DATA HERE?
-
-As we did not observe any effect of ingesting glucose before and after RT compared to placebo in this design, there is little to suggest that glucose *per se* has any additional effect on ribosome biogenesis, as long as the daily energy intake is sufficient to support cellular energy levels required to drive this process.
-Looking at previous observations from *in vitro* studies of rodents, yeast, and non-muscle cell cultures (MARIAPPAN11; KIM13; TANAKA18; ZHAI12; HOPPE09; MURAYAMA08), one interesting notion would be how high glucose vs.
-low glucose or glucose starvation would affect ribosome biogenesis over time in human skeletal muscle cells.
-Time being of the essence in such a perspective, as previous *in vitro* studies have utilized cells with higher physiological rate than human skeletal muscle cells (KIM13;MARIAPPAN11;TANAKA15;ZHAI12).
-
-Indeed, our findings suggest that glucose ingestion *per se* before and after RT does not induce changes in the internal physiological milieu of human skeletal muscle cells with anabolic effects.
-It has been theorised that glucose may convey an anabolic signal through insulin, and thus operate as a signal mediator affecting upstream pathways of ribosome biogenesis (KUSNADI15; TANAKA18).
-Such an anabolic signal could then possibly increase PIC assembly, and consequently Pol I recruitment, initiating rDNA transcription and subsequent rRNA maturation, leading to increased ribosome content (FIGUEIREDO19; TANAKA18).
-As significantly higher blood levels of c-peptide and glucose were observed in RT with glucose compared to placebo in the present study, without any difference in ribosome biogenesis, there is little to suggest that either insulin or glucose has any direct acute effects.
-
-Another theory is that glucose ingestion leads to higher cellular energy levels compared to placebo, stimulating energy-sensitive pathways such as PIH1, mTORC1, ERK1/2, AMPK and SIRT1 in a dose-response relationship (KIM13;MARIAPPAN11;TANAKA15;ZHAI12).
-The mTORC1, along with ERK1/2 and PIH1 up-regulates PIC assembly during high-energy conditions, while AMPK and possibly SIRT1 represses PIC assembly during low-energy conditions, as observed in rodent (KILDE), yeast (KILDE), and non-muscle cells (KILDE) *in vitro*.
-Specifically, high glucose treatment augmented rDNA transcription by increased UBF and SL-1 activity, while low glucose treatment, glucose starvation, or nutrient depletion stimulated AMPK-dependent TIF-IA and chromatin regulation suppression (KOS-BRAUN17; MARIAPPAN11; ZHAI12).
-In addition to UBF and SL-1, assembly of the PIC requires TIF-IA and Pol I recruitment, induced by amino acid intake in an mTOR-S6K-dependent manner and possibly by direct c-Myc activation (KUSNADI15).
-As such, increasing energy supply and nutrient availability should in line with these theories potently stimulate PIC assembly and rDNA transcription.
-Importantly, these pathways were not directly investigated in the present study.
-However, downstream targets (UBF, c-Myc, rpS6?) remained unaffected by glucose compared to placebo ingestion, suggesting there is not a dose-response relationship between muscle cellular energy levels and ribosome biogenesis during an RT-intervention (Not sure if I can say this).
-
-A weakness in the present study, is the lack of data on intracellular glucose levels in the muscle cells.
-Thus, it cannot be said for sure whether there were differences between glucose and placebo or not.
-
-A possible explanation for the lack of effect of glucose compared to placebo, is that both treatments may have exercised with sufficient levels of nutrients and energy, meaning that mTORC1 had sufficient nutrient availability and, thus was not inhibited by AMPK due to energy deficiency (HOPPE09; KUSNADI15).
-Indeed, mTORC1 is characterised as energy-sensitive (DENNIS01), and therefore it might have been directly affected by high vs.
-low-glucose conditions (MARIAPPAN11).
-However, previous studies suggest that mTORC1 is not sensitive enough to register levels of ATP, rather depending on signals from AMPK to assess ATP levels (KIM10).
-This would arguably make AMPK the primary regulator of the potential impact of glucose on ribosome biogenesis, and thus, glucose has no direct effect on mTORC1 signalling (REFERENCES?).
-Importantly, intramuscular glycogen content was not measured in the present study, therefore we cannot determine whether cellular energy levels in fact were equal or not, or if there is reason to assume that AMPK-dependent mTORC1 signalling was different in the two treatments.
-Potentially, such data could give higher resolution to our data.
-Along with the robust increases in proxy markers of ribosome biogenesis, the equal macronutrient intake between treatments on pairwise consecutive days and plasma glucose levels not indicating energy deficiency during RT, potent mTORC1 stimulation without AMPK-dependent inhibition may be a plausible explanation.
-Together with previous studies on this topic (MARIAPPAN11; ZHAI12), there are currently scarce observations to suggest glucose carries a direct anabolic signal.
-As such, physiological levels of plasma glucose during RT, combined with a sufficient daily macronutrient intake, seem potent enough to supply the energy needed for ribosome biogenesis.
-The lacking differences between treatments in the biological data are further supported by our observations on functional data.
-
-Overall, glucose did not affect the isometric or isokinetic peak torque at 240 d/s, or muscular recovery following the sixth and last RT session, measured as temporal changes in isometric and isokinetic peak torque.
-Generally, the mean isometric and isokinetic peak torque slightly decreased during the intervention, especially between session 4 and 6, and regained after the intervention.
-Further, there were no differences between glucose and placebo in total session volume or training intensity.
-Both exercise with glucose and placebo yielded higher total session volumes at higher percentage of 1RM at the last session compared to baseline, as expected from earlier reports on the moderate-volume RT protocol (KRIEGER09; RHEA03; SCHOENFELD17), in addition to a more recent study implementing the unilateral protocol (HAMMARSTROM20+21/22).
-Interestingly, we did observe a significant difference between treatments in mean change of isokinetic peak torque at 60 d/s before the sixth RT session, where exercise with glucose decreased peak torque 9% less compared to placebo.
-However, as none of the other measurements indicated any difference between treatments, a true effect of glucose seems unlikely.
-Taken together, these results may add to previous observations with no differences between co-ingestion of carbohydrate and protein compared to protein alone, in regard to skeletal muscular recovery following RT (KOOPMAN07; STAPLES11).
-Collectively, the functional data point towards an exhaustive, yet growth-inducing RT protocol, especially considering the biological data.
-As a side note, an interesting observation was made by another master’s thesis on this project, finding significantly lower MuRF protein levels at post with glucose treatment compared to placebo.
-This may suggest an effect of glucose on protein degradation (WILKES09), however, whether this contributes to growth or simply protein turnover has been a topic of debate (FIGUEIREDO19).
-
-Despite being a thoroughly standardised design, a limitation may be posed through lack of participants.
-We initially reach our minimum limit for statistical power, i.e.
-16 participants, however, three dropped out during the intervention, possibly impacting our power.
-Coupled with the fact that no previous study could guide our power calculations, this, of course, introduces some degree of uncertainty regarding our results.
-However, we’ve reported little to suggest any true acute effect of RT with glucose ingestion compared to placebo.
-Moreover, the study was specifically designed to investigate differences in pre-post changes within participants, and thus also quite limited to this perspective.
-Hence, the effects of time observed exclusively serve as supplementary to the efficiency of the design, without a negative control group to either confirm or refute them.
-If further investigations are to be done on this topic, these analyses should include more participants and optimally a measure of intramuscular glycogen levels.
-
-In conclusion, resistance exercise training combined with glucose ingestion does not acutely augment ribosome biogenesis following 2 weeks of moderate-volume resistance training in moderately trained young male and female adults.
-Neither does it acutely affect muscular performance in isometric and isokinetic knee extension peak torque, nor enhance post resistance training recovery within 24hrs.
-
-## Disadvantages due to Covid-19
-
-Due to restrictions regarding social distancing and lock downs, caused by the Covid-19 pandemic, we were not able to recruit and include as many participants as we planned to the study.
-<!-- Is this necessary to include at all?
--->
+Extended methods or manuscript, if manuscript: discussion?
+- Explanation of heteroscedasticity and regression to the mean?
+- Linear mixed model fit by REML - what does it entail?
+- T-tests use Satterthwaite's method ['lmerModLmerTest'] - what does it entail?
+- Random and fixed effects?

### 2023-03-10 results_work.RMD
updated methods and results in ribose-paper.RMD

@@
 ######**Table 2:** Mean total session volume and mean change in total session volume from session 1 to session 6.
-![](figures/tab2.png)
 
@@
 
-Levels of c-Myc increased 51% more after RT with placebo compared to RT with glucose, a significant difference (Fig 3A: *p* = `r cdiff`).
-Levels of UBF and rpS6 increased 42 and 33% more in placebo compared to glucose, however, this was not a significant difference (Fig 3A: UBF: *p* = `r udiff`, rpS6: *p* = `r rpdiff`).
+Levels of c-Myc increased by 51% more after RT with placebo compared to RT with glucose, a significant difference (Fig 3A: *p* = `r cdiff`).
+Levels of UBF and rpS6 increased by 42 and 33% more in placebo compared to glucose, however, this was not a significant difference (Fig 3A: UBF: *p* = `r udiff`, rpS6: *p* = `r rpdiff`).
 The increase in both RT with glucose and placebo for all measured proteins was significant at post compared to baseline (*p* < 0.05).
-Further, there was a linear relationship between UBF and total RNA, where an increase of 1 SD unit of UBF equated to a 13% increase in total RNA (*p* = `r corrp`).
+Further, we found a linear relationship between UBF and total RNA, where an increase of 1 SD unit of UBF equated to a 13% increase in total RNA (*p* = `r corrp`).
 

### 2023-03-10 ribose-paper.RMD
updated methods and results in ribose-paper.RMD

@@
 ```{r packages}
-library(tidyverse);library(readxl);library(lme4);library(emmeans);library(dplyr);library(cowplot);library(knitr);library(broom);library(rmarkdown);library(tinytex);library(tidyr);library(arsenal);library(magrittr);library(dabestr);library(nlme); library(pbkrtest)
+library(tidyverse);library(lme4);library(emmeans);library(cowplot);library(tinytex);library(nlme)
 ```
@@
 
-The study was designed as a 12-day double-blinded randomized controlled trial, with a unilateral RT protocol (Fig 2).
-Participants were randomized to commence with either glucose (GLU, 30g per bolus) or placebo (PLAC, 0g per bolus), and either dominant or non-dominant leg (Fig 1), alternating RT and supplement from one day to the other (Fig 2).
-Investigators were blinded as the supplement randomization was disclosed only to personnel not involved in any other aspect of the study.
-THe participants were blinded via diluting boluses of glucose (30g glucose) (Glucosum monohydricum, Merck KGaA, Darmstadt, Germany) and placebo (~0.3g Stevia rebaudiana extract) (Steviosa, Soma Nordic AS, Oslo, Norway) in 300ml Fun Light juice/saft (Orkla, Oslo, Norway).
-A blinded taste test revealed that the participants were not able to disclose the contents of the provided boluses (30g GLU vs.
-0g GLU).
-To ensure equal conditions during training sessions, participants exercised and tested at the same time of day, +/- 1hr (check this) with the same trainer, and recorded and repeated their daily macro nutrient intake on pairwise consecutive days (Tab 4).
-
-```{r DXA}
-# Script: DXA
-
-# Data handling
-table.dxa <- read_excel("./data/dxa/ribose_dxa.xlsx") %>%
-        select(sex, subject, age, height, weight, leanmasskg, fatmasskg, ffmg) %>%
-        #make grams into kg
-        mutate(ffkg = ffmg/1000) %>%
-        select(sex, age, height, weight, leanmasskg, fatmasskg, ffkg) %>%
-  mutate(bmi = (weight) / (height^2) * 10000)
-
-dxa.table <- table.dxa %>%
-        pivot_longer(names_to = "variable",
-                     values_to = "values", col= age:bmi) %>%
-        select(sex, variable, values) %>%
-        group_by(sex, variable) %>%
-        summarise(m = mean(values, na.rm = TRUE),
-                  s= sd(values, na.rm = TRUE)) %>%
-        mutate(stat = paste0(round(m, 1), " (", round(s, 1), ")")) %>%
-        select(sex, variable, stat) %>%
-        pivot_wider(names_from = sex,
-                    values_from = stat) %>%
-                mutate(variable = if_else(variable %in% c("age"),
-                                          "Age (yrs)",
-                                          if_else(variable %in% c("fatmasskg"),
-                                                  "Fat mass (kg)",
-                                                  if_else(variable %in% c("ffkg"),
-                                                          "Fat free mass (kg)",
-                                                          if_else(variable %in% c("height"),
-                                                                  "Height (cm)",
-                                                                  if_else(variable %in% c("leanmasskg"),
-                                                                          "Lean body mass (kg)",
-                                                                          if_else(variable %in% c("weight"),
-                                                                                  "Body weight (kg)",
-                                                                                  if_else(variable %in% c("bmi"),
-                                                                                          "BMI", "")
-                                                                                  )))))))
-
-dxa.table %>%
-        kable(col.names = c("Variable", "Female (n=7)", "Male (n=9)"))
-
-```
-
-```{r dxa per leg}
-# Script: DXA
-
-# Subject 104, 110 and 111 dropped out and is therefore excluded from the table
-
-#Load dxa data
-dxaleg <-  read_excel("./data/dxa/ribose_dxaprleg.xlsx") %>%
-        select(subject, sex, leg, leanmassg, fatmassg, totalmasskg) %>%
-        #make grams into kg
-        mutate(leanmass = leanmassg/1000,
-               fatmass = fatmassg/1000)%>%
-        select(subject, sex, leg, leanmass, fatmass, totalmasskg)
-
-# Between groups analysis 
-dxa.stat <- dxaleg %>%
-        select(subject, leg, leanmass, fatmass, totalmasskg)
-
-# Lean mass between legs, within subject 
-leanmass.lm <- lmer(leanmass ~ leg + leg:subject + (1|subject), data = dxa.stat)
-lm.plot <- plot(leanmass.lm)
-lm.sum <- summary(leanmass.lm)
-lm.confint <- emmeans(leanmass.lm, specs = ~ "leg")
-
-# Fat mass between legs, within subject
-fat.lm <- lmer(fatmass ~ leg + leg:subject +(1|subject), data = dxa.stat)
-fat.plot <- plot(fat.lm)
-fat.sum <- summary(fat.lm)
-fat.confint <- emmeans(fat.lm, specs = ~ "leg")
-
-# Total mass between legs, within subject
-total.lm <- lmer(totalmasskg ~leg + leg:subject + (1|subject), data = dxa.stat)
-totlm.plot <- plot(total.lm)
-totlm.sum <- summary(total.lm)
-supsup <- emmeans(total.lm, specs = ~ "leg")
-
-# Table of leanmass, fatmass and total mass in kg pr. leg
-dxalegtable <- dxaleg %>%
-        select(leg, leanmass, fatmass, totalmasskg) %>%
-        pivot_longer(names_to = "variable",
-                     values_to = "values", col= leanmass:totalmasskg) %>%
-        group_by(leg, variable) %>%
-        summarise(m = mean(values, na.rm = TRUE),
-                  s= sd(values, na.rm = TRUE)) %>%
-        mutate(stat = paste0(round(m, 1), " (", round(s, 1), ")")) %>%
-        select(leg, variable, stat) %>%
-        pivot_wider(names_from = leg,
-                    values_from = stat)
-
-# Markdown table
-dxalegtable %>%
-        kable(col.names = c("Variable", "Left leg", "Right leg"))
-
-```
-
-Six RT sessions were conducted with glucose and six with placebo, allowing a within-subjects analysis of the effects of glucose ingestion.
-Data from the first five RT sessions was used to investigate main outcome measures (total RNA, rRNA and protein), whereas data from the sixth RT session was used to explore secondary outcomes (muscular recovery, plasma glucose levels).
-Participants were asked to refrain from resistance- or high-intensity training of the legs from Day -7 and onwards, until completion of the intervention and post-testing, to ensure the reliability of pre-intervention strength data and minimal interference from other training sources.
+The study was designed as a 12-day double-blinded randomized controlled trial, with a alternating unilateral RT protocol (Fig 1A).
+Participants were randomly allocated to exercise either their dominant or non-dominant leg with glucose or placebo before and after RT, and whether they started RT with glucose or placebo.
+Thus, participants exercised every day, alternating between e.g., one leg + glucose and the other leg + placebo (1A).
+One person was exclusively responsible for the randomization code and supplement distribution, blinding both investigators and participants in regard to which leg exercised with glucose and placebo.
+Glucose (Glucosum monohydricum, Merck KGaA, Darmstadt, Germany) and placebo (Steviosa, Soma Nordic AS, Oslo, Norway) was masked by mixing with 300ml Fun Light juice/saft (Orkla, Oslo, Norway).
+A blinded taste test revealed that the participants were not able to disclose the contents of the provided boluses (30g glucosum monohydricum vs.
+~0.3g Stevia rebaudiana extract).
+To ensure equal conditions during training sessions, participants exercised and tested at the same time of day, +/- 1hr (check this) with the same supervisor, in addition to recording and repeating their daily macro nutrient intake (protein, fat, carbohydrate) and total calories on pairwise consecutive days.
+
+Six RT sessions were conducted with glucose and six with placebo, allowing a within-subjects analysis of the effects of glucose ingestion before and after RT.
+Data from the first five RT sessions was used to investigate main outcome measures (total RNA, rRNA and protein) and leg muscle strength, whereas data from the sixth RT session was used to explore secondary outcomes (muscular recovery, plasma glucose levels).
+Participants were asked to avoid  resistance- or high-intensity training of the legs from Day -7 (Fig 1A) and onwards, until completion of the intervention and post-testing, to ensure the reliability of pre-intervention strength data and minimal interference from other training sources.
 If participants did conduct exercise training outside of the prescribed protocol, they were asked to ensure equal loading on pairwise consecutive days.
@@
 Isometric peak torque was measured at knee-angle 60, for a maximum of 10 seconds and two repetition per test.
-The highest peak torque values were carried forwards to the final analyses.
-During days 4, 5, 8 and 9, humac tests were conducted one hour before RT with the leg performing RT the previous day.
-During days 11 and 12, humac tests were performed at four time points: I) 45min before RT, II) 30min after the last RT session, III) 2hrs after the last RT session, and IV) 23hrs after the last RT session.
-Test I at day 12/13 included testing of both legs, representing 23hrs post RT session test of one leg and pre-session 6 test of the other leg.
-
-Assessment of unilateral one repetition maximum (1RM) leg press and knee extension was conducted at pre-intervention testing (days -7 and -5, Fig 1).
-The participants performed a general warm-up with 10min cycling on an indoor exercise bicycle and a specific warm-up (1x10, 1x6 and 1x3 repetitions, 2min breaks) before each of the tests.
-Between specific warm-up sets as well as between each subsequent 1RM attempt, the participants were given 3min breaks.
-All positions were controlled and recorded at pre-intervention testing and repeated for the RT sessions.
+The highest peak torque values were used in the final analyses.
+During days 4, 5, 8 and 9 (Fig 1D, day 4 and 5 = Post 2RT, day 8 and 9 = Post 4RT), humac tests were conducted one hour before RT with the leg performing RT the previous day.
+During days 11 and 12, humac tests were performed at four time points: I) 45min before RT (1D, Post 5RT), II) 30min after the last RT session (1D, 30 min post 6RT), III) 2hrs after the last RT session (1D, 2h post 6RT), and IV) 23hrs after the last RT session (1D, 23h post 6RT).
+Test I at day 12/13 included testing of both legs, representing 23hrs post RT session test of one leg and post session 5 test of the other leg.
+
+Assessment of unilateral one repetition maximum (1RM) leg press and knee extension was conducted at pre-intervention testing during familiariztion (days -7 and -5, Fig 1D).
+The participants performed a general warm-up with 10min cycling on an indoor exercise bicycle.
+In addition, a protocol consisting of 1x10, 1x6 and 1x3 repetitions, was used as spesific warm-up before each of the tests.
+All positions were controlled and recorded at the first 1RM test and repeated for the RT sessions.
 Maximal leg press strength was defined as the maximal load lifted in a controlled fashion, with a knee angle of 45 degrees (start), within a 2.5kg range.
@@
 
+### Data handling and statistical analysis
+
+A priori power calculations showed that 20 participants would grant a statistical power of 80% (= 0.05), with an expected drop out of 20%.
+This power calculation was based on an assumption that the effects of glucose ingestion on ribosome biogenesis may equate to the effects of increasing RT volume from low to moderate (HAMMARSTROM20).
+<!-- maybe include the power calculation + reason in extended methods?
+-->.
+Raw data was imported an analyzed in R via R studios (SOURCE).
+Legs were categorized as either glucose or placebo in accordance with the randomization.
+Before quantitative analyses, total RNA was normalized by wet muscle weight, qPCR values were normalized by wet muscle weight and an external reference gene (Lambda), while quantified protein signals from western blot were normalized by pool signals and total protein signal per gel <!-- add explanations of normalization to extended methods -->.
+Total RNA and qPCR values were analyzed by comparing post to pre log-fold change score differences.
+This was achieved by modelling change as response with baseline and supplement (glucose/placebo) as explanatory factors, and mean-centered baseline values as a covariate in a linear mixed effects model, then calculating the estimated marginal means <!-- add explanation of model in extended methods-->.
+Plasma glucose and c-peptide, training volume, and the strength index were analyzed by multiple time-point log-fold change score comparisons, using the same linear mixed effects model with time added as explanatory factor <!-- add explanation of model in extended methods-->.
+The linear mixed effects model was designed with the lmer function of the lme4 package (BATES14) in R studios, using the lmerTest function to procure p-values.
+The relationship between
+Log-transformation and mean-centering baseline values was done to control for heteroscedasticity and regression to the mean, respectively.
+Log-transformed values were reverse-transformed for figure illustration.
+Values are presented as mean change + 95% CL unless otherwise stated, alpha-levels was set to α = 0.05.
+
 ## Results
+
+```{r nutrition model}
+# Scripts
+# - nut.change.R
+
+# Data
+pro.change <- readRDS("./data/data-gen/nutrition/pro.change.RDS")
+fat.change <- readRDS("./data/data-gen/nutrition/fat.change.RDS")
+cho.change <- readRDS("./data/data-gen/nutrition/cho.change.RDS")
+cal.change <- readRDS("./data/data-gen/nutrition/cal.change.RDS")
+propr.change <- readRDS("./data/data-gen/nutrition/propr.change.RDS")
+
+## Models
+# Protein
+pro.mod <- lmerTest::lmer(change ~ 0 + d1 + time + supplement:time + (1|subject),
+                     data = pro.change)
+
+pro.res <- plot(pro.mod)
+pro.sum <- summary(pro.mod)
+
+# Fat
+fat.mod <- lmerTest::lmer(change ~ 0 + d1 + time + supplement:time + (1|subject),
+                     data = fat.change)
+
+fat.res <- plot(fat.mod)
+fat.sum <- summary(fat.mod)
+
+# Carbohydrates
+cho.mod <- lmerTest::lmer(change ~ 0 + d1 + time + supplement:time + (1|subject),
+                     data = cho.change)
+
+cho.res <- plot(cho.mod)
+cho.sum <- summary(cho.mod)
+
+# Calories
+cal.mod <- lmerTest::lmer(change ~ 0 + d1 + time + supplement:time + (1|subject),
+                     data = cal.change)
+
+cal.res <- plot(cal.mod)
+cal.sum <- summary(cal.mod)
+
+# Protein/kg
+propr.mod <- lmerTest::lmer(change ~ 0 + d1 + time + supplement:time + (1|subject),
+                     data = propr.change)
+
+propr.res <- plot(propr.mod)
+propr.sum <- summary(propr.mod)
+
+## Emmeans
+
+# Protein
+protein.emm <- confint(emmeans(pro.mod, specs = ~"supplement|time")) %>%
+  data.frame()
+
+# Fat
+fat.emm <- confint(emmeans(fat.mod, specs = ~"supplement|time")) %>%
+  data.frame()
+
+# Carbohydrates
+cho.emm <- confint(emmeans(cho.mod, specs = ~"supplement|time")) %>%
+  data.frame()
+
+# Calories
+cal.emm <- confint(emmeans(cal.mod, specs = ~"supplement|time")) %>%
+  data.frame()
+
+# Protein/kg
+propr.emm <- confint(emmeans(propr.mod, specs = ~"supplement|time")) %>%
+  data.frame()
+
+```
+
+```{r volume model}
+## Scripts
+# - volume.cleanup.R
+# - volume.analysis.R
+
+## Data
+vol.change <- readRDS("./data/data-gen/training/vol.change.RDS") # Untransformed data
+vol.lchange <- readRDS("./data/data-gen/training/vol.lchange.RDS") # Log-transformed data
+
+## Models
+
+# Untransformed 
+vol.mod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
+                     data = vol.change)
+vol.res <- plot(vol.mod)
+vol.sum <- summary(vol.mod)
+
+# Log-transformed
+vol.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
+                     data = vol.lchange)
+
+vol.lres <- plot(vol.lmod)
+vol.lsum <- summary(vol.lmod)
+
+## Emmeans
+
+# Untransformed
+vol.emm <- confint(emmeans(vol.mod, specs = ~"supplement|time")) %>%
+  data.frame()
+
+# Log-transformed
+vol.lemm <- confint(emmeans(vol.lmod, specs = ~"supplement|time")) %>%
+  data.frame()
+
+## For inline plotting
+
+# Sessions compared to baseline
+volcoef <- coef(summary(vol.mod))
+pvol.s2 <- round(volcoef[2,5], 4)
+pvol.s3 <- round(volcoef[3,5], 4)
+pvol.s4 <- round(volcoef[4,5], 4)
+pvol.s5 <- round(volcoef[5,5], 3)
+pvol.s6 <- round(volcoef[6,5], 5)
+
+# Glucose compared to plac
+pvol.c2 <- round(volcoef[7,5], 3) 
+pvol.c3 <- round(volcoef[8,5], 3) 
+pvol.c4 <- round(volcoef[9,5], 3) 
+pvol.c5 <- round(volcoef[10,5], 3) 
+pvol.c6 <- round(volcoef[11,5], 3)
+
+```
+
+There were no difference in mean change of total training session volume between exercising with placebo or glucose, with an increase of 18 and 16% from session 1 to session 6, respectively (Tab 2; Mean change, 2: *p* = `r pvol.c2`, 3: *p* = `r pvol.c3`, 4: *p* = `r pvol.c4`, 5: *p* = `r pvol.c5`, 6: *p* = `r pvol.c6`).
+Both exercising with glucose and placebo led to significant increases in mean total training session volume at session four, five and six, compared to session 1 (Tab 2; Mean change, 2: *p* = `r pvol.s2`, 3: *p* = `r pvol.s3`, 4: *p* = `r pvol.s4`, 5: *p* = `r pvol.s5`, 6: *p* = `r pvol.s6`).
+In regard to the dietary intervention, there were no difference in mean macro nutrient intake (protein, fat, carbohydrates) or total calorie intake between glucose and placebo on pairwise consecutive days (*p* > 0.05).
 
@@
 
-At baseline, there were no significant differences between training with glucose or placebo in muscle mass, isometric or isokinetic peak torque, total session volume, training intensity, total RNA content, or rRNA expression (Tab 3).
-Further, on pairwise consecutive days, there were no significant differences in mean macro nutrient (calories, carbohydrate, fat or protein) ingestion (Tab 4), or mean change in macro nutrient ingestion (Fig 3).
-Resistance training with glucose led to significant increases in blood levels glucose and c-peptide compared to RT with placebo, before, during and after the session (Fig 4B, 120min: 38 ± 4% [*p* = `r p.120`], 135min: 31 ± 4% [*p* = `r p.135`], 150min: 32 ± 4% [*p* = `r p.150`], Fig 4C, 120min: 95 ± 10% [*p* = `r cpp.120`], 150min: 87 ± 10% [*p* = `r cpp.150`]), measured during the final RT session.
-At 2hrs post-exercise, RT with glucose led to significantly lower blood glucose levels compared to RT with placebo, with a tendancy of the same in c-peptide levels (4B, 270min: -8 ± 4% [*p* = `r p.270`] and 4C, 270min: -7 ± 10 [*p* `r cpp.270`]).
-
-###### **Table 3**: Baseline characteristics of the intervention legs, organized per supplement.
-Variables: Iso 60 = isokinetic 60 d/s, Iso 240 = isokinetic 240 d/s, training intensity = load/rm*100.
-Nm = newton meters, kg = kilograms, %1RM = percentage of 1 repetition maximum, ng/mg = nanogram per milligram wet muscle tissue, nf.expr = expression normalized by weight and external reference gene (Lambda).
-Ribosomal RNA data are presented as log transformed.
-Glucose n = 13, placebo n = 13.
-
-![](figures/table3.png)
-
-###### **Table 4:** Daily mean dietary intake of macro nutrients during each training day per supplement leg.
-G = glucose, P = placebo.
-Training day 1-training day 12 are combined pairwise to day 1-6: Day 1 = training day 1/2, Day 2 = training day 3/4, Day 3 = training day 5/6, Day 4 = training day 7/8, Day 5 = training day 9/10, Day 6 = training day 11/12.
-Values are mean ± standard deviation.
-P = difference between glucose and placebo.
-Glucose n = 12, placebo n = 12.
-Day 6: Glu = 11, Plac = 11.
-
-```{r nutrition table}
-## Scripts
-# - nut.table.R
-# - nut.analysis.R
-# - dxa.R
-
-# Data
-nuttable <- readRDS("./data/data-gen/dxa/nuttable.RDS")
-
-nuttable %>%
-  kable(col.names = c("Time", "Suppl.", "Kcal", "CHO (g)", "Fat (g)", "PRO (g)", "PRO (g/kg)"))
-
-```
-
-```{r nutrition fig}
-# Associated scripts:
-# - nut.datahand.R
-# - nut.change.R
-# - nut.change.figs.R
-
-nut.fig <- readRDS("./data/data-gen/nutrition/nut.fig.RDS")
-nut.fig
-
-```
-###### **Figure 3:** Mean changes in macro nutrient intake from day to day.
-Day 1 = training day 1/2(T1/T2), Day 2 = training day 3/4, Day 3 = training day 5/6, Day 4 = training day 7/8, Day 5 = training day 9/10, Day 6 = training day 11/12(T3/T4).
-Values are mean change in grams/calories ± 95% CL.
-
-![](figures/day.timeline.png)
-
-```{r glucose + c-peptide fig}
-## Scripts
-# - gluc.fig.R
-# - cpep.figs.R
-
-# Separate figures
-pos <- position_dodge(width = 0.2)
-
-gluc.fig <- readRDS("./data/data-gen/glucose/gluc.fig.RDS")
-
-cpep.fig <- readRDS("./data/data-gen/glucose/cpep.fig.RDS")
-
-# Joined figure
-plot_grid(gluc.fig,
-          cpep.fig,
-          ncol = 1, nrow = 2)
-
-```
-
-###### **Figure 4:** A) Time line for training day 6 (combination of pairwise training day 11 and 12).
-Biopsy = muscle biopsy taken pre supplementation, PRO = protein supplementation, Humac = strength test (isometric knee extension, isokinetic knee extension 60 and 240 d/s), GLU/PLAC = glucose/placebo supplementation, RT = resistance training.
-Minutes: -120 = 2hrs prior to RT, -90 = 90min prior to RT, -30 = 30min prior to RT, 0 = onset of RT, 15 = 15min following onset of RT, between leg press and knee extension, 30 = 30min post onset of RT, 120 = 2hrs following onset of RT.
-B) Mean mmol/L change in blood glucose levels per supplement group during training day 6.
-Time-points: 0, 45, 90, 120, 135, 150 and 270min post protein ingestion, respectively.
-C) Mean change in pmol/L of c-peptide in blood per supplement group during training day 6.
-Values are presented as estimated marginal means of mmol/L and pmol/L change ± 95% CI.
-*p* < 0.05 between groups.
-Glucose n = 13, placebo n = 13.
+Glucose ingestion before and after RT led to significant increases in plasma levels of glucose by 38% immediately before RT (Fig 1B, 0 min: *p* = `r p.120`), by 31% during RT (1B, 15 min: *p* = `r p.135`) and by 32% immediately after RT (1B, 30 min: *p* = `r p.150`), compared to placebo ingestion before and after RT.
+Two hours after the session, RT with glucose measured 8% lower plasma glucose levels compared to RT with placebo (1B, 270min: *p* = `r p.270`).
+Similarly, levels of c-peptide increased significantly more in exercises with glucose compared to placebo, by 95% immediately before (Fig 1C, 0 min: *p* = `r cpp.120`) and 87% after RT (1C, 30 min: *p* = `r cpp.150`).
+
+```{r strength model}
+## Scripts:
+# - Strength.index.R
+
+## Data
+# Full data - baseline -> 23h post 6RT
+str.change <- readRDS("./data/data-gen/humac/str.change.RDS") # Untransfromed data
+str.lchange <- readRDS("./data/data-gen/humac/str.lchange.RDS") # Log-transfromed data
+
+# Acute data - Post 5RT -> 23h post 6RT
+astr.change <- readRDS("./data/data-gen/humac/astr.change.RDS") # Untransfromed data
+astr.lchange <- readRDS("./data/data-gen/humac/astr.lchange.RDS") # Log-transfromed data
+
+## Linear mixed effects model full data
+
+# Untransformed model
+str.mod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
+                         data = str.change)
+str.res <- plot(str.mod)
+str.sum <- summary(str.mod)
+
+# Log-transformed model
+str.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
+                        data = str.lchange)
+str.lres <- plot(str.lmod)
+str.lsum <- summary(str.lmod)
+
+## Emmeans full data
+# Untransformed
+str.emm <- confint(emmeans(str.mod, specs = ~"supplement|time")) %>%
+  data.frame()
+
+# Log-transformed
+str.lemm <- confint(emmeans(str.lmod, specs = ~"supplement|time")) %>%
+  data.frame()
+
+## Linear mixed effects model acute
+# Untransformed
+astr.mod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
+                        data = astr.change)
+
+astr.res <- plot(astr.mod)
+astr.sum <- summary(astr.mod)
+
+# Log-transformed
+astr.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
+                         data = astr.lchange)
+
+astr.lres <- plot(astr.lmod)
+astr.lsum <- summary(astr.lmod)
+
+## Emmeans acute data
+astr.emm <- confint(emmeans(astr.mod, specs = ~"supplement|time")) %>%
+  data.frame()
+
+astr.lmod <- confint(emmeans(astr.lmod, specs = ~"supplement|time")) %>%
+  data.frame()
+
+## For inline plotting full data
+strcoef <- coef(summary(str.mod))
+
+# Time points compared to baseline (full data)
+pc2 <- round(strcoef[2,5], 3)
+pc3 <- round(strcoef[3,5], 3)
+pscient.c4 <- round(strcoef[4,5], 40)
+pc4 <- sprintf("%.5f", 7.886277e-06)
+pc5 <- round(strcoef[5,5], 3)
+pc6 <- round(strcoef[6,5], 3)
+pc7 <- round(strcoef[7,5], 5)
+
+# Glucose compared to placebo (full data)
+pdiff.c2 <- round(strcoef[8,5], 3)
+pdiff.c3 <- round(strcoef[9,5], 3)
+pdiff.c4 <- round(strcoef[10,5], 3)
+pdiff.c5 <- round(strcoef[11,5], 3)
+pdiff.c6 <- round(strcoef[12,5], 3)
+pdiff.c7 <- round(strcoef[13,5], 3)
+
+## For inline plotting acute data
+astrcoef <- coef(summary(astr.mod))
+apc2 <- round(astrcoef[2,5], 3)
+apc3 <- round(astrcoef[3,5], 3)
+apc4 <- round(astrcoef[4,5], 3)
+
+```
+
+After 5 RT sessions (Fig 1D, Post 5RT), there was a significant difference in strength between glucose and placebo,  where placebo decreased strength 7% more than glucose (1D, *p* = `r pdiff.c4`).
+At the remaining time points, there were no significant differences in mean change between exercising with glucose and placebo (1D, Post 2RT: *p* = `r pdiff.c2`, Post 4RT: *p* = `r pdiff.c3`, 30min post 6RT: *p* = `r pdiff.c5`, 2h post 6RT: *p* = `r pdiff.c6`, 23h post 6RT: *p* = `r pdiff.c7`).
+Both RT with glucose and placebo led to significantly reduced strength post fifth session compared to baseline, by 11 and 18% respectively (1D, Post 5RT: *p* = `r pc4`).
+Comparisons of the acute data gathered from post fifth session until and including 23hrs post sixth session showed a significant mean increase in strength of 5-9% after RT with glucose and placebo 30min after sixth RT session (1D, 30min post 6RT: *p* = `r apc2`) and two hours after sixth RT session (1D, 2h post 6RT: *p* = `r apc3`).
+Twenty-three hours after the last RT, strength was unchanged compared to after fifth RT session (23h post 6RT: *p* = `r apc4`).
 
 ### Markers of ribosome biogenesis
+
+#### Total RNA and ribosomal RNA
 
@@
 
-Resistance training with glucose did not induce a higher accumulation of total RNA (fig 5A, *p* = `r trdiffp`), or rRNA  (fig 5B, *p* > 0.05), compared to RT with placebo, measured as mean change from baseline to post.
+Resistance training with glucose did not induce a higher accumulation of total RNA (fig 2A, *p* = `r trdiffp`) or rRNA  (fig 2B, 47S: *p* = `r rr47diffp`, 18S: *p* = `r rr18diffp`, 28S: *p* = `r rr28diffp`, 5.8S: *p* = `r rr58diffp`, 5S: *p* = `r rr5diffp`) compared to RT with placebo.
 From baseline to post-intervention, there was a mean increase in levels of total RNA by 26 and 22% after RT with glucose and placebo, respectively (*p* < 0.05, compared to baseline).
 A robust accumulation was also observed in ribosomal RNA expression in both RT with glucose and placebo, with mean increases between 34-43% (GLU) and 33-41% (PLAC) (*p* < 0.05, compared to baseline) in the four rRNA’s.
-The expression of 47S pre-rRNA increased by 37 and 59% with glucose and placebo respectively, with only placebo reaching significantly higher levels at post compared to baseline (fig 45, *p* = `r rr47tdiffp`).
-
-```{r RNA figures}
-# Get figures
-totrna.fig <- readRDS("./data/data-gen/rna/totrna.fig.RDS")
-totrna.fig
-
-plot.18 <- readRDS("./data/data-gen/rna/plot.18.RDS")
-plot.28 <- readRDS("./data/data-gen/rna/plot.28.RDS")
-plot.5.8 <- readRDS("./data/data-gen/rna/plot.5.8.RDS")
-plot.5 <- readRDS("./data/data-gen/rna/plot.5.RDS")
-plot.47 <- readRDS("./data/data-gen/rna/plot.47.RDS")
-
-# Joined rRNA
-library(cowplot)
-
-legend <- get_legend(plot.47 + theme(legend.box.margin = margin(0, 0, 0,12)))
-
-plot_grid(plot.18 + theme(legend.position = "none"),
-          plot.28 + theme(legend.position = "none"), 
-          plot.5.8 + theme(legend.position = "none"),
-          plot.5 + theme(legend.position = "none"), 
-          plot.47 + theme(legend.position = "none"),
-          legend,
-          ncol = 2, nrow = 3)
-
-```
-
-###### **Figure 5:** A) Changes in Total RNA, B) Changes in ribosomal RNA.
-Baseline = Training day 1/2, Post = Training day 11/12.
-Both total RNA and rRNA were analyzed with duplicates, with two duplicates per biopsy (two biopsies per time point).
-Total RNA and rRNA changes were calculated as log-fold change score per mg wet muscle weight.
-Mean change scores of the duplicates were calculated before modelling and transformed to the log-scale.
-Values are estimated marginal means fold change per leg/supplement ± 95% CI.
-p = between groups statistic, glucose compared to placebo.
-Glucose n = 13, placebo n = 13.
+The expression of the 47S pre-rRNA increased by 37 and 59% with glucose and placebo respectively, where only RT with placebo increased significantly from baseline to post (fig 45, *p* = `r rr47tdiffp`).
+
+#### Protein
 
@@
 # - western.import.R
-# - western.change.R
+# - western.cleanup.R
+# - western.analysis.R
 
 # Data
-
-prot.dat <- readRDS("./data/data-gen/protein/prot.dat.RDS") # cleaned up and quantified western data
-prot.out <- readRDS("./data/data-gen/protein/protein_complete.RDS") # Cleaned up and quantified western data, checked for outliers
-prot.change <- readRDS("./data/data-gen/protein/prot.change.RDS") # change-score data
-
-# By target
-
-cmyc <- prot.change %>%
-  filter(target == "cmyc")
-
-ubf <- prot.change %>%
-  filter(target == "ubf")
-
-rps6 <- prot.change %>%
-  filter(target == "rps6")
-
-# c-Myc change
-
-m1 <- lme(change ~ pre + supplement,
+prot.dat <- readRDS("./data/data-gen/protein/prot.datc.RDS") # from the script: western.cleanup.R
+
+### Modelling target signal per supplement 
+# These models try to explain the normalised signal by time, supplement and time+supplement (?). The random = list command creates an individual 
+# intercept for the included (e.g., subject, gel.sample).
+
+## c-Myc model
+
+# Untransformed
+cmyc.mod <- lme(norm.sign ~ time + supplement + time:supplement, 
                 random = list(subject = ~ 1),
-            data = cmyc) # same model as in "western import", but with change scores of norm.sign
-
-#plot(m1)
-
-sum.m1 <- summary(m1)
-
-cmyc.emm <- emmeans(m1, specs = ~"supplement") %>%
-  data.frame
-
-saveRDS(cmyc.emm, "./data/data-gen/protein/cmyc.change.RDS")
-
-# UBF change
-
-m2 <- lme(change ~ pre + supplement,
+                data = filter(prot.dat, target == "cmyc"))
+
+cmyc.sum <- summary(cmyc.mod)
+cmyc.res <- plot(cmyc.mod, resid(., type = "p") ~ fitted(.))
+
+# Log-transformed
+cmyc.lmod <- lme(log(norm.sign) ~ time + supplement + time:supplement, 
                  random = list(subject = ~ 1),
-            data = ubf) # same model as in "western import", but with change (i.e., change in norm.sign) instead of norm.sign
-
-#plot(m2)
-
-sum.m2 <- summary(m2)
-
-ubf.emm <- emmeans(m2, specs = ~"supplement") %>%
-  data.frame()
-
-saveRDS(ubf.emm, "./data/data-gen/protein/ubf.change.RDS")
-
-# rps6 change
-
-m3 <- lme(change ~ pre + supplement,
+                 data = filter(prot.dat, target == "cmyc"))
+
+cmyc.lsum <- summary(cmyc.lmod)
+cmyc.lres <- plot(cmyc.lmod, resid(., type = "p") ~ fitted(.))
+
+## UBF model
+# Untransformed
+ubf.mod <- lme(norm.sign ~ time + supplement + time:supplement, 
                random = list(subject = ~ 1),
-            data = rps6) # same model as in "western import", but with change (i.e., change in norm.sign) instead of norm.sign
-
-#plot(m3)
-
-sum.m3 <- summary(m3)
-
-rps6.emm <- emmeans(m3, specs = ~"supplement") %>%
-  data.frame()
-
-saveRDS(rps6.emm, "./data/data-gen/protein/rps6.change.RDS")
+               data = filter(prot.dat, target == "ubf"))
+
+ubf.sum <- summary(ubf.mod)
+ubf.res <- plot(ubf.mod, resid(., type = "p") ~ fitted(.))
+
+# Log-transformed
+ubf.lmod <- lme(log(norm.sign) ~ time + supplement + time:supplement, 
+                random = list(subject = ~ 1),
+                data = filter(prot.dat, target == "ubf"))
+
+ubf.lsum <- summary(ubf.lmod)
+ubf.lres <- plot(ubf.lmod, resid(., type = "p") ~ fitted(.))
+
+# RPS6 model
+# Untransformed
+rps6.mod <- lme(norm.sign ~ time + supplement + time:supplement, 
+                 random = list(subject = ~ 1),
+                 data = filter(prot.dat, target == "rps6"))
+
+rps6.sum <- summary(rps6.mod)
+rps6.res <- plot(rps6.mod, resid(., type = "p") ~ fitted(.))
+
+# Log-transformed
+rps6.lmod <- lme(log(norm.sign) ~ time + supplement + time:supplement, 
+          random = list(subject = ~ 1),
+          data = filter(prot.dat, target == "rps6"))
+
+rps6.lsum <- summary(rps6.lmod)
+rps6.lres <- plot(rps6.lmod, resid(., type = "p") ~ fitted(.))
+
+### Getting emmeans from the models
+
+## c-Myc emmeans
+# Untransformed
+cmyc.emm <- emmeans(cmyc.mod, specs = ~ time|supplement) %>%
+  data.frame()
+
+# Log-transformed
+cmyc.lemm <- emmeans(cmyc.lmod, specs = ~ time|supplement) %>%
+  data.frame()
+
+## UBF emmeans
+# Untransformed
+ubf.emm <- emmeans(ubf.mod, specs = ~ time|supplement) %>%
+  data.frame()
+
+# Log-transformed
+ubf.lemm <- emmeans(ubf.lmod, specs = ~ time|supplement) %>%
+  data.frame()
+
+## RPS6
+# Untransformed
+rps6.emm <- emmeans(rps6.mod, specs = ~ time|supplement) %>%
+  data.frame()
+
+# Log-transformed
+rps6.lemm <- emmeans(rps6.lmod, specs = ~ time|supplement) %>%
+  data.frame()
 
@@
 ## Diff cmyc
-ccoef <- coef(summary(m1))
-cdiff <- round(ccoef[3,5], 3)
-cdiffse <- round(ccoef[3,2], 3)
+
+ccoef <- coef(summary(cmyc.lmod))
+cdiff <- round(ccoef[4,5], 3)
+cpla <- round(ccoef[2,5], 20)
 
 ## Diff UBF
-ucoef <- coef(summary(m2))
-udiff <- round(ucoef[3,5], 3)
-udiffse <- round(ucoef[3,2], 3)
+ucoef <- coef(summary(ubf.lmod))
+udiff <- round(ucoef[4,5], 3)
 
 ## Diff rps6
-rpcoef <- coef(summary(m3))
-rpdiff <- round(rpcoef[3,5], 3)
-rpdiffse <- round(rpcoef[3,2], 3)
-
-```
-
-```{r western fig}
+rpcoef <- coef(summary(rps6.lmod))
+rpdiff <- round(rpcoef[4,5], 3)
+
+```
+
+```{r correlation model}
 # Data 
-
-cmyc <- readRDS("./data/data-gen/protein/cmyc.change.RDS")
-
-ubf <- readRDS("./data/data-gen/protein/ubf.change.RDS")
-
-rps6 <- readRDS("./data/data-gen/protein/rps6.change.RDS")
-
-## cmyc fig
-
-cmyc.plot <- cmyc %>%
-  mutate(time = "post") %>%
-  add_row(supplement = c("PLACEBO", "GLUCOSE"), time = "pre", emmean = 0, lower.CL = 0, upper.CL = 0) %>%
-  mutate(time = factor(time, levels = c("pre", "post"), labels = c("Baseline", "Post"))) %>%
-  ggplot(aes(time, exp(emmean), group = supplement, fill = supplement)) +
-  #annotate("text", x = "Post", y = 2.1, label = "p = 0.585", size = 3) +
-  geom_errorbar(aes(ymin = exp(lower.CL), ymax = exp(upper.CL)),
-                width = 0.1,
-                position = position_dodge(width = 0.2)) +
-  geom_line(position = position_dodge(width = 0.2)) +
-  geom_point(shape = 21, size = 3, position = position_dodge(width = 0.2)) +
-  labs(x = "", y = "c-Myc signal \n(fold change)\n", fill = "") +
-  theme_classic()
-
-ubf.plot <- ubf %>%
-  mutate(time = "post") %>%
-  add_row(supplement = c("PLACEBO", "GLUCOSE"), time = "pre", emmean = 0, lower.CL = 0, upper.CL = 0) %>%
-  mutate(time = factor(time, levels = c("pre", "post"), labels = c("Baseline", "Post"))) %>%
-  ggplot(aes(time, exp(emmean), group = supplement, fill = supplement)) +
-  #annotate("text", x = "Post", y = 2.1, label = "p = 0.585", size = 3) +
-  geom_errorbar(aes(ymin = exp(lower.CL), ymax = exp(upper.CL)),
-                width = 0.1,
-                position = position_dodge(width = 0.2)) +
-  geom_line(position = position_dodge(width = 0.2)) +
-  geom_point(shape = 21, size = 3, position = position_dodge(width = 0.2)) +
-  labs(x = "", y = "UBF signal \n(fold change)\n", fill = "") +
-  theme_classic()
-
-rps6.plot <- rps6 %>%
-  mutate(time = "post") %>%
-  add_row(supplement = c("PLACEBO", "GLUCOSE"), time = "pre", emmean = 0, lower.CL = 0, upper.CL = 0) %>%
-  mutate(time = factor(time, levels = c("pre", "post"), labels = c("Baseline", "Post"))) %>%
-  ggplot(aes(time, exp(emmean), group = supplement, fill = supplement)) +
-  annotate("text", x = "Post", y = 1.9, label = "p = ", size = 3) +
-  geom_errorbar(aes(ymin = exp(lower.CL), ymax = exp(upper.CL)),
-                width = 0.1,
-                position = position_dodge(width = 0.2)) +
-  geom_line(position = position_dodge(width = 0.2)) +
-  geom_point(shape = 21, size = 3, position = position_dodge(width = 0.2)) +
-  labs(x = "", y = "rps6 signal \n(fold change)\n", fill = "") +
-  theme_classic()
-
-# Cowplot for gathering figures
-
-legend <- get_legend(rps6.plot + theme(legend.box.margin = margin(0, 0, 0,12)))
-
-prot.joined <- plot_grid(cmyc.plot + theme(legend.position = "none"),
-                         ubf.plot + theme(legend.position = "none"), 
-                         rps6.plot + theme(legend.position = "none"),
-                         legend,
-                         ncol = 2, nrow = 2)
-
-prot.joined
-
-saveRDS(prot.joined, "./data/data-gen/rna/prot.joined.RDS")
-
-```
-
-###### **Figure 5:** A) Changes in c-Myc levels, B) changes in UBF levels, C) changes in rps6 levels.
-Baseline = Training day 1/2, post = Training day 11/12.
-Four samples were analysed per participant (GLU pre-post, PLAC pre-post), in duplicates, a total of 104 protein samples.
-Duplicate samples were loaded on separate gels.
-In addition, 3-4 pools were added to each well.
-c-Myc, UBF and rps6 signals were then calculated as log-fold change score, normalised by gel pools.
-Mean change scores of the duplicates were calculated before modelling and transformed to the log-scale.
-Values are estimated marginal means fold change per leg/supplement ± 95% CI.
-p = between groups statistic, glucose compared to placebo.
-Glucose n = 13, placebo n = 13.
-
-### Resistance training and strength testing
-
-```{r training data model}
-## Scripts
-# - volume.change.R
-# - ss.analysis.R
-# - ss.cleanup.R
-# - trainint.analysis.R
-# - trainint.cleanup.R
-
-# Log-transformed data
-vol.lchange <- readRDS("./data/data-gen/training/vol.lchange.RDS")
-ss.lchange <- readRDS("./data/data-gen/training/ss.lchange.RDS")
-int.lchange <- readRDS("./data/data-gen/training/int.lchange.RDS")
-
-### Training volume
-# Fold-change model
-vol.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = vol.lchange)
-vol.lres <- plot(vol.lmod)
-vol.lsum <- summary(vol.lmod)
-
-# Fold-change estimated means
-vol.lemm <- confint(emmeans(vol.lmod, specs = ~"supplement|time")) %>%
-        data.frame()
-
-### Session score (RPE)
-# Model
-ss.lmod <- lmerTest::lmer(change ~ baseline + time + supplement:time + (1|subject),
-                     data = ss.lchange)
-ss.lres <- plot(ss.lmod)
-ss.lsum <- summary(ss.lmod)
-
-# Session score emmeans
-ss.lemm <- confint(emmeans(ss.lmod, specs = ~"supplement|time")) %>%
-        data.frame()
-
-### Training intensity
-int.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = int.lchange)
-int.lres <- plot(int.lmod)
-int.lsum <- summary(int.lmod)
-
-# Emmeans
-int.lemm <- confint(emmeans(int.lmod, specs = ~"supplement|time")) %>%
-        data.frame()
-
-### For inline reporting
-
-## Training volume
-# Diff 4
-vcoef <- coef(summary(vol.lmod))
-vdiff.4 <- round(vcoef[9,1], 3)
-vdiffp.4 <- round(vcoef[9,5], 3)
-vdiffse.4 <- round(vcoef[9,2], 3)
-# Diff 5
-vdiff.5 <- round(vcoef[10,1], 3)
-vdiffp.5 <- round(vcoef[10,5], 3)
-vdiffse.5 <- round(vcoef[10,2], 3)
-# Diff 6
-vdiff.6 <- round(vcoef[11,1], 3)
-vdiffp.6 <- round(vcoef[11,5], 3)
-vdiffse.6 <- round(vcoef[11,2], 3)
-## RPE
-scoef <- coef(summary(ss.lmod))
-sdiffp <- round(scoef[11, 5], 3)
-sdiffse <- round(scoef[11, 2], 3)
-
-## Training intensity
-
-# Diff time 2
-icoef <- coef(summary(int.lmod))
-idiff2 <- round(icoef[7,1], 3)
-idiffp2 <- round(icoef[7,5], 3)
-idiffse2 <- round(icoef[7,2], 3)
-# Diff time 3
-idiff3 <- round(icoef[8,1], 3)
-idiffp3 <- round(icoef[8,5], 3)
-idiffse3 <- round(icoef[8,2], 3)
-# Diff time 4
-idiff4 <- round(icoef[9,1], 3)
-idiffp4 <- round(icoef[9,5], 3)
-idiffse4 <- round(icoef[9,2], 3)
-# Diff time 5
-idiff5 <- round(icoef[10,1], 3)
-idiffp5 <- round(icoef[10,5], 3)
-idiffse5 <- round(icoef[10,2], 3)
-# Diff time 6
-idiff6 <- round(icoef[11,1], 3)
-idiffp6 <- round(icoef[11,5], 3)
-idiffse6 <- round(icoef[11,2], 3)
-```
-
-```{r humac post 5th model}
-## Scripts
-# - humac.cleanup.R
-# - humac.main.analysis.R - this is data gathered at baseline, during the intervention, and after session 5
-
-# Log-transformed data
-
-isom.lchange <- readRDS("./data/data-gen/humac/isom.lchange.RDS")
-iso60.lchange <- readRDS("./data/data-gen/humac/isok60.lchange.RDS")
-iso240.lchange <- readRDS("./data/data-gen/humac/isok240.lchange.RDS")
-
-# Isometric
-isom.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = isom.lchange)
-isom.lres <- plot(isom.lmod)
-isom.lsum <- summary(isom.lmod)
-
-# Isok.60
-iso60.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = iso60.lchange)
-iso60.lres <- plot(iso60.lmod)
-iso60.lsum <- summary(iso60.lmod)
-
-# Isok.240
-iso240.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = iso240.lchange)
-iso240.lres <- plot(iso240.lmod)
-iso240.lsum <- summary(iso240.lmod)
-
-## Emmeans
-# Isometric
-isom.lemm <- confint(emmeans(isom.lmod, specs = ~"supplement|time")) %>%
-        data.frame()
-
-# Isok.60
-iso60.lemm <- confint(emmeans(iso60.lmod, specs = ~"supplement|time")) %>%
-        data.frame()
-
-# Isok.240
-iso240.lemm <- confint(emmeans(iso240.lmod, specs = ~"supplement|time")) %>%
-        data.frame()
-
-# For reporting
-# Isom diff
-isom5coef <- coef(summary(isom.lmod))
-isom5diffp <- round(isom5coef[7, 5], 3)
-isom5diffse <- round(isom5coef[7, 2], 3)
-
-# Isok 60 diff
-isok605coef <- coef(summary(iso60.lmod))
-isok605diffp <- round(isok605coef[7, 5], 3)
-isok605diffse <- round(isok605coef[7, 2], 3)
-
-# Isok 260
-isok2405coef <- coef(summary(iso240.lmod))
-isok2405diffp <- round(isok2405coef[7, 5], 3)
-isok2405diffse <- round(isok2405coef[7, 2], 3)
-
-```
-
-```{r post 6th model}
-## Scripts
-# - humac.acute.analysis.R - this is data gatherede before and after one RT session (sixth session)
-# - humac.cleanup.R
-
-# Data
-isom.lchange.ac <- readRDS("./data/data-gen/humac/isom.lchange.ac.RDS")
-iso60.lchange.ac <- readRDS("./data/data-gen/humac/iso60.lchange.ac.RDS")
-iso240.lchange.ac <- readRDS("./data/data-gen/humac/iso240.lchange.ac.RDS")
-
-# Isometric
-isomac.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = isom.lchange.ac)
-isomac.lres <- plot(isomac.lmod)
-isomac.lsum <- summary(isomac.lmod)
-
-# Isok.60
-iso60ac.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = iso60.lchange.ac)
-iso60ac.lres <- plot(iso60ac.lmod)
-iso60ac.lsum <- summary(iso60ac.lmod)
-
-# Isok.240
-iso240ac.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
-                     data = iso240.lchange.ac)
-iso240ac.lres <- plot(iso240ac.lmod)
-iso240ac.lsum <- summary(iso240ac.lmod)
-
-# Emmeans
-# Isometric
-isomac.lemm <- confint(emmeans(isomac.lmod, specs = ~"supplement|time")) %>%
-        data.frame()
-
-# Isok.60
-iso60ac.lemm <- confint(emmeans(iso60ac.lmod, specs = ~"supplement|time")) %>%
-        data.frame()
-
-# Isok.240
-iso240.lemm <- confint(emmeans(iso240ac.lmod, specs = ~"supplement|time")) %>%
-        data.frame()
-
-## For inline reporting
-# Isom diff
-isomcoef <- coef(summary(isomac.lmod))
-isomdiffp <- round(isomcoef[7, 5], 3)
-isomdiffse <- round(isomcoef[7, 2], 3)
-
-# Isok 60 diff
-isok60coef <- coef(summary(iso60ac.lmod))
-isok60diffp <- round(isok60coef[7, 5], 3)
-isok60diffse <- round(isok60coef[7, 2], 3)
-
-# Isok 240 diff
-isok240coef <- coef(summary(iso240ac.lmod))
-isok240diffp <- round(isok240coef[7, 5], 3)
-isok240diffse <- round(isok240coef[7, 2], 3)
-
-```
-
-There were no significant differences in mean change of total session volume, training intensity or RPE between exercise training with glucose and placebo at any of the time points (fig 6, *p* > 0.05).
-We did observe interactions between time and changes, where total session training volume increased from 13-18% compared to baseline.
-These increases in training volume were accompanied by changes ranging from 5-18% in training intensity and 2-13% in RPE (Tab 5).
-The maximal isometric knee extension force (OR TORQUE?) and isokinetic knee extension peak torque at 240 d/s showed no differences between exercise training with glucose and placebo after five RT sessions (Fig 7A, *p* > 0.05).
-There was a significant difference in mean change from pre to post-fifth RT session, where exercise training with glucose led to a 9% less decrease in peak torque at 60 d/s compared to placebo (Fig 7A, *p* = `r isok605diffp`).
-There was no difference between glucose and placebo in mean change from baseline to after the sixth RT session (Fig 7B, *p* > 0.05).
-
-![](figures/legend.PNG)
-
-```{r training dat fig}
-### Get figures
-
-# Volume
-totvol.barplot <- readRDS("./data/data-gen/training/totvol.barplot.RDS")
-totvolfold.plot <- readRDS("./data/data-gen/training/totvolfold.plot.RDS")
-
-# Training intensity
-tot.barplot <- readRDS("./data/data-gen/training/tot.barplot.RDS")
-totalint.plot <- readRDS("./data/data-gen/training/trainint.plot.RDS")
-
-# Join figures
-plot_grid(totvol.barplot + theme(legend.position = "none"),
-          totvolfold.plot + theme(legend.position = "none"),
-          tot.barplot + theme(legend.position = "none"),
-          totalint.plot + theme(legend.position = "none"),
-          ncol = 2,
-          nrow = 2,
-          align = "h")
-
-```
-
-###### **Figure 6:** Training volume and training intensity from baseline until resistance exercise session 6 (training day 11 and 12).
-Baseline = Training day 1/2, Session 2 = Training day 3/4, Session 3 = Training day 5/6, Session 4 = Training day 7/8, Session 5 = Training day 9/10, Session 6 = Training day 11/12.
-A) Changes in total session volume per leg, values are mean kg ± standard deviation.
-B) Fold change in total session volume per leg, values are estimated marginal means fold change ± 95% CI.
-C) Changes in training intensity per leg calculated as %1RM, values are mean % ± standard deviation.
-D) Fold change in training intensity per leg, values are estimated marginal means fold change ± 95% CI.
-† = *p* < 0.05 in both glucose and placebo, compared to baseline.
-p = NS = no significant difference between legs at any of the time points.
-Glucose n = 13, placebo n = 13.
-
-###### **Table 5:** RPE changes through training days.
-Training days are presented as combined pairwise consecutive, linked to their respective sessions.
-Session 1 = Training day 1/2, Session 2 = Training day 3/4, Session 3 = Training day 5/6, Session 4 = Training day 7/8, Session 5 = Training day 9/10, Session 6 = Training day 11/12.
-Values are presented as mean ± standard deviation.
-Glucose n = 13, placebo n = 13.
-
-```{r RPE table}
-## Scripts
-# - ss.cleanup.R
-# - ss.absolute.R
-
-# Data
-ss.abs <- readRDS("./data/data-gen/training/ss.abs.RDS")
-
-# Table
-ss.abs %>%
-  kable(col.names = c("Supplement", "Baseline", "Session 2", "Session 3", "Session 4", "Session 5",
-                            "Session 6"))
-
-```
-
-![](figures/legend2.PNG)
-
-```{r humac figs}
-## Post 5th RE session
-# Figures
-pos <- position_dodge(width = 0.2)
-isom5.plot <- readRDS("./data/data-gen/humac/isom.main.fig.RDS")
-isok605.plot <- readRDS("./data/data-gen/humac/isok60.main.fig.RDS")
-isok2405.plot <- readRDS("./data/data-gen/humac/isok240.main.fig.RDS")
-
-legend <- get_legend(isom5.plot + theme(legend.box.margin = margin(0, 0, 0, 12)))
-
-plot_grid(isom5.plot + theme(legend.position = "none"), 
-          isok605.plot + theme(legend.position = "none"), 
-          isok2405.plot + theme(legend.position = "none"),
-          nrow = 3,
-          align = "v")
-
-## Post 6th RE session
-
-# Figures
-
-pos <- position_dodge(width = 0.2)
-
-isom.plot <- readRDS("./data/data-gen/humac/isom.ac.plot.RDS")
-isok60.plot <- readRDS("./data/data-gen/humac/isok60.ac.plot.RDS")
-isok240.plot <- readRDS("./data/data-gen/humac/isok240.ac.plot.RDS")
-
-legend <- get_legend(isom.plot + theme(legend.box.margin = margin(0, 0, 0, 12)))
-
-plot_grid(isom.plot + theme(legend.position = "none"), 
-          isok60.plot + theme(legend.position = "none"), 
-          isok240.plot + theme(legend.position = "none"),
-          nrow = 3,
-          align = "v")
-
-```
-
-The mean change in peak torque at all three velocities change little from baseline until session 4 in both glucose and placebo (+4-7% change), decreasing at pre-session 6 (9-18% change, Fig 6A).
-Post session 6 (Fig 6B) showed a slight reduction in isometric peak torque 30min after session completion (2-4% change), slowly regaining at 2hrs (+3% change) and 23hrs (+3-6% change), in both glucose and placebo compared to baseline.
-Isokinetic peak torque at 60 d/s and 240 d/s showed no reduction in mean peak torque from baseline until 23hrs following RT,  in both glucose and placebo (+0,1-11%).
-There was an effect of time in peak torque at 23hrs post-RT at 60 d/s (Fig 6B), where exercise training with placebo led to a 5% increase in peak torque compared to baseline, however, there was no significant difference compared to exercise training with glucose.
-
-###### **Figure 7:** Peak torque changes in maximal isometric knee extension and maximal isokinetic knee extension at 60 and 240 d/s, respectively.
-A) Humac tests during the intervention, Baseline = Day -1 (prior to intervention), Session 2 = prior to training, training Day 4/5, Session 4 = prior training, training Day 8/9, Session 6 = prior to training 11/12.
-B) Baseline = prior to training, training day 11/12, 30min = 30 min after RT, 2hr = 2hrs after RT, 23hr = 23hrs after RT.
-Values are change in estimated marginal means of newton meters (Nm) ± 95% CI.
-* = *p* < 0.05 between groups.
-*p* = NS = no significant difference between legs at any of the time points.
-Glucose n = 13, placebo n = 13.
+west.dat <- readRDS("./data/data-gen/protein/prot.datc.RDS") # Quantified and cleaned western data,
+                                                               # (outliers marked) normalised by pool
+
+rna.dat <- readRDS("./data/data-gen/rna/rna.dat.RDS") # Quantified and cleaned (outliers marked), 
+                                                      # total RNA data, normalised by wet muscle weight
+
+## Prepare data frames for joining
+
+ubf.rdy <- west.dat %>%  # One data frame without outliers
+  select(subject, time, outlier, supplement, target, norm.sign) %>%
+  filter(target == "ubf",
+         outlier == "in") %>%
+  select(-target) %>%
+  group_by(subject, time, supplement) %>%
+  summarise(mean.sign = mean(norm.sign, na.rm = TRUE))
+
+rna.rdy <- rna.dat %>%
+  select(subject, time, outlier, supplement, RNA.weight) %>%
+  filter(outlier == "in") %>%
+  mutate(subject = as.character(subject)) %>%
+  group_by(subject, time, supplement) %>%
+  summarise(mean.rna = mean(RNA.weight, na.rm = TRUE))
+
+## Correlation analysis
+
+joined.dat <- ubf.rdy %>%
+  right_join(rna.rdy) %>%
+  #filter(supplement != "GLUCOSE") %>%
+
+  mutate(sd.ubf = (mean.sign - mean(mean.sign, na.rm = TRUE )/sd(mean.sign, na.rm = TRUE)))
+
+library(nlme)
+
+corr.mod <- lme(mean.rna ~ mean.sign + time , 
+         random = list(subject = ~ 1), 
+          data = joined.dat, 
+         na.action = na.omit)
+
+corr.sum <- summary(corr.mod)
+corr.res <- plot(corr.mod)
+
+corr.lmod <- lme(log(mean.rna) ~ log(mean.sign) + time , 
+         random = list(subject = ~ 1), 
+          data = joined.dat, 
+         na.action = na.omit)
+
+corr.lsum <- summary(corr.lmod)
+corr.lres <- plot(corr.lmod)
+
+corrcoef <- coef(summary(corr.mod))
+corrp <- round(corrcoef[2,5], 3)
+```
+
+Levels of c-Myc increased by 51% more after RT with placebo compared to RT with glucose, a significant difference (Fig 3A: *p* = `r cdiff`).
+Levels of UBF and rpS6 increased by 42 and 33% more in placebo compared to glucose, however, this was not a significant difference (Fig 3A: UBF: *p* = `r udiff`, rpS6: *p* = `r rpdiff`).
+The increase in both RT with glucose and placebo for all measured proteins was significant at post compared to baseline (*p* < 0.05).
+Further, we found a linear relationship between UBF and total RNA, where an increase of 1 SD unit of UBF equated to a 13% increase in total RNA (*p* = `r corrp`).
 
@@
 
-## Disadvantages due to Covid-19
-
-Due to restrictions regarding social distancing and lock downs, caused by the Covid-19 pandemic, we were not able to recruit and include as many participants as we planned to the study.
-<!-- Is this necessary to include at all?
--->
+## References
+
+<div id="refs"></div>
+
+## Figure/table text
+
+ **Table 1:**  Baseline characteristics.
+Age = mean age in years, Height = mean height in cm, Weight and Lean mass = mean in kg, Iso 60/240 = mean unilateral isokinetic knee extension peak torque in Nm at 60 d/s and 240 d/s, Isom = mean unilateral isometric knee extension peak torque in Nm.
+
+**Table 2:** Mean total session volume and mean change in total session volume from session 1 to session 6.
+
+**Figure 1:** A) An overview of the experimental design with 12 days of concomitant dietary intervention and resistance training (RT), preceded by 7 days involving familiarization.
+Between days -7 and -1, participants were familiarized to the RT exercises via 1RM leg press and knee extension testing, and to the strength tests via Humac Norm dynamometer (days -7 and -5).
+Before baseline testing, the participants were randomly allocated to exercise one leg with GLU (glucose) and the other with PLA (placebo), in a unilateral, alternating fashion.
+Further, non-dominant/dominant + GLU/PLA, and onset with GLU or PLA was also randomized, i.e.
+the figure illustrates an example where the participant was randomized to start RT with GLU.
+Biopsies were taken from m.
+vastus lateralis at baseline (Day 1 leg 1, Day 2 leg 2), and after five RT sessions (Day 11 leg 1, Day 12 leg 2).
+Blood for measurement of plasma glucose and -c-peptide was sampled at baseline (Day 1), and during post testing (Day 11 leg 1, Day 12 leg 2), via finger draws and venous blood samples.
+Skeletal muscle strength was measured as peak torque in unilateral isometric and isokinetic (at 60 and 240 d/s) knee extension before, multiple times during, and after five and six session.
+A total of three participants dropped out of the intervention, either due to reported muscular discomforts of heavy RT, or sickness in their family.
+B and C) Changes in plasma glucose (B, mmol/L) and c-peptide levels (C, pmol/L).
+Glucose levels in blood was measured via finger draws 120 (-120), 90 (-90), and 30min (-30) before RT, immediately before RT (0), during RT (15), immediately after RT (30) and 2hrs after RT (120).
+C-peptide levels were measured simultaneously to these finger draws, except for 90min before and during RT.
+D) Changes in muscular strength measured as isometric and isokinetic peak torque (60 and 240 d/s) via Humac Norm Dynamometer, conducted at baseline (A: Day -1), after two and four RT sessions (A: Day 4 and 8 leg 1, Day 5 and 9 leg 2), after five RT sessions/before the 6th session (Day 11 leg 1, Day 12 leg 2), as well as 30min, 2hrs and 23hrs after the 6th RT session (A: Day 11/12 leg 1, Day 12/13 leg 2).
+The index was calculated by normalizing peak torque values to the highest peak torque value at each respective speed, and then summarized and used in change score calculations.
+Values are presented as changes in estimated marginal means ± 95% CI.
+* = *p* < 0.05 between groups.
+Glucose n = 13, placebo n = 13.
+
+**Figure 2:** Changes in total RNA and ribosomal RNA.
+A) Total RNA, B) 47S pre-rRNA, 18S rRNA, 28S rRNA, 5.8S rRNA, 5S rRNA.
+Baseline = Day 1 leg 1/ Day 2 leg 2, Post = Day 11 leg 1, Day 12 leg 2.
+Total RNA and rRNA were analyzed in duplicates, with two duplicates per biopsy (two muscle tissue pieces per time point), and normalized to ng x mg wet muscle weight for total RNA and external reference gene (Lambda) for rRNA.
+Total RNA and rRNA changes were calculated as log-fold change score per mg wet muscle weight.
+Mean change scores of the duplicates were calculated and transformed to the log-scale before modelling, then reverse-transformed for figure illustration.
+Values are estimated marginal means fold change per leg per supplement ± 95% CI.
+Glucose n = 13, placebo n = 13.
+
+**Figure 3:** Changes in levels of protein.
+A) Mean UBF, c-Myc and rpS6 (AU, arbitrary units) at baseline and post + representative western blots of the respective proteins.
+Baseline = Day 1 leg 1, Day 2 leg 2, post = Day 11 leg 1, Day 12 leg 2.
+GLU = glucose, PLA = placebo.
+Protein samples were analyzed in two duplicates per biopsy per time point, loaded on separate gels in an inverted order, e.g.
+from gel 3 to 6, as shown by the duplicates (1 and 2).
+Change in protein levels were calculated as log-fold change scores normalized by pools (pool of all protein samples per gel).
+Mean log-change scores of the duplicates were calculated before modelling and reverse-transformed for figure illustration.
+Values are estimated marginal means fold change per leg per supplement ± 95% CI.
+* = *p* < 0.05 between groups.
+Glucose n = 13, placebo n = 13.
+B) Representative total protein stain blot.
+C) Linear relationship between total RNA (ng x mg) and UBF levels (SD units), with time added as a covariate.
+Total RNA was normalized by wet muscle weight, and UBF was normalized by pools per gel, and total protein per lane factor.
+Values are presented as log-transformed means.

### 2023-03-10 R/cpep.analysis.R

@@ -22,20 +22,6 @@ library(dplyr); library(tidyverse); library(tidyr); library(nlme); library(lme4)
 
 ins.dat2 <- readRDS("./data/data-gen/glucose/cpep.clean.RDS")
 
-
-## Baseline analysis
-# Comparing pre measurements to post day baseline
-
-base.ins <- ins.dat2 %>%
-  filter(time %in% c("pre", "baseline")) %>%
-  select(subject, time, supplement, c.pep) %>%
-  group_by(supplement) %>%
-  pivot_wider(names_from = time, 
-              values_from = c.pep) %>%
-  print()
-
-ins.ttest <- t.test(base.ins$pre, base.ins$baseline, paired = TRUE)
-  
 ## Change data
 
 ins.change <- ins.dat2 %>%

### 2023-03-10 R/totrna.analysis.R

@@ -69,26 +69,6 @@ rna_dat2 %>%
 
 saveRDS(rna_dat2, "./data/data-gen/rna/rna.dat.RDS")
 
-## Baseline analysis: calculation of mean and SD of baseline values, t-test on baseline values
-
-base.rna <- rna_dat2 %>%
-  filter(time == "pre") %>%
-  select(subject, time, supplement, RNA) %>%
-  group_by(subject, supplement) %>%
-  summarise(m = mean(RNA)) %>%
-  pivot_wider(names_from = supplement,
-              values_from = m) %>%
-  print()
-
-rna.ttest <- t.test(base.rna$GLUCOSE, base.rna$PLACEBO, paired = TRUE)
-
-rna.sum <- rna_dat2 %>%
-  filter(time == "pre") %>%
-  select(subject, time, supplement, RNA) %>%
-  group_by(supplement) %>%
-  summarise(m = mean(RNA),
-            s = sd(RNA)) %>%
-  print()
 
 # Log-fold change score calculation: We are interested in if the change in GLUCOSE is different from change in 
 # PLACEBO. 

### 2023-03-10 ext_method.RMD
updates to intro and methods in ribose-paper.RMD

@@
 
+## Wet lab analyses
+
+### RNA extraction
+
+### cDNA synthesis and qPCR
+
+### Protein extraction
+
+### Western Blot
+
+## Statistical analyses
+
+We chose to log-transform all data to be reported, as visual inspection showed clear signs of heteroscedasticity in multiple data sets.
+Therefore, we made the decision to do the same for all data sets, which was to log-transform the data prior to modelling, and then reverse-transform the data that was visualized in figures.
+The exception was the data used in figure 1B, 1C and 1D, where we used absolute changes rather than relative changes obtained by reverse-transforming from the log scale.
+This is because using the absolute data here arguably represents the data in a more comprehensible way, compared to i.e., total RNA or protein signalling, which is less commonly known.
+As for the strength index, it was already put on a different scale by compiling the index.
+
+Further, we mean-centered the baseline values in each model to account for regression to the mean.
+This was simply achieved by subtracting the mean baseline value from the baseline value of each measurement.
+As such, differences at baseline, e.g.
+by extremely high or low values does not violate the assumption of equal variance, and the we can interpret the models with more certainty.
+
 Mean-centered baseline values were used to correct for regression to the mean, which potentially occurs when the same participants
@@
 
-## Code/formulas for quantitative analyses
 Total RNA: 

### 2023-03-10 results_work.RMD
updates to intro and methods in ribose-paper.RMD

@@
 
-######**Table 2:** Mean total session volume and mean change in total session volume from session 1 to session 6.
-
 Glucose ingestion before and after RT led to significant increases in plasma levels of glucose by 38% immediately before RT (Fig 1B, 0 min: *p* = `r p.120`), by 31% during RT (1B, 15 min: *p* = `r p.135`) and by 32% immediately after RT (1B, 30 min: *p* = `r p.150`), compared to placebo ingestion before and after RT.
@@
 Twenty-three hours after the last RT, strength was unchanged compared to after fifth RT session (23h post 6RT: *p* = `r apc4`).
-
-###### **Figure 1:** A) An overview of the experimental design with 12 days of concomitant dietary intervention and resistance training (RT), preceded by 7 days involving familiarization.
-Between days -7 and -1, participants were familiarized to the RT exercises via 1RM leg press and knee extension testing, and to the strength tests via Humac Norm dynamometer (days -7 and -5).
-Before baseline testing, the participants were randomly allocated to exercise one leg with GLU (glucose) and the other with PLA (placebo), in a unilateral, alternating fashion.
-Further, non-dominant/dominant + GLU/PLA, and onset with GLU or PLA was also randomized, i.e.
-the figure illustrates an example where the participant was randomized to start RT with GLU.
-Biopsies were taken from m.
-vastus lateralis at baseline (Day 1 leg 1, Day 2 leg 2), and after five RT sessions (Day 11 leg 1, Day 12 leg 2).
-Blood for measurement of plasma glucose and -c-peptide was sampled at baseline (Day 1), and during post testing (Day 11 leg 1, Day 12 leg 2), via finger draws and venous blood samples.
-Skeletal muscle strength was measured as peak torque in unilateral isometric and isokinetic (at 60 and 240 d/s) knee extension before, multiple times during, and after five and six session.
-A total of three participants dropped out of the intervention, either due to reported muscular discomforts of heavy RT, or sickness in their family.
-B and C) Changes in plasma glucose (B, mmol/L) and c-peptide levels (C, pmol/L).
-Glucose levels in blood was measured via finger draws 120 (-120), 90 (-90), and 30min (-30) before RT, immediately before RT (0), during RT (15), immediately after RT (30) and 2hrs after RT (120).
-C-peptide levels were measured simultaneously to these finger draws, except for 90min before and during RT.
-D) Changes in muscular strength measured as isometric and isokinetic peak torque (60 and 240 d/s) via Humac Norm Dynamometer, conducted at baseline (A: Day -1), after two and four RT sessions (A: Day 4 and 8 leg 1, Day 5 and 9 leg 2), after five RT sessions/before the 6th session (Day 11 leg 1, Day 12 leg 2), as well as 30min, 2hrs and 23hrs after the 6th RT session (A: Day 11/12 leg 1, Day 12/13 leg 2).
-The index was calculated by normalizing peak torque values to the highest peak torque value at each respective speed, and then summarized and used in change score calculations.
-Values are presented as changes in estimated marginal means ± 95% CI.
-* = *p* < 0.05 between groups.
-Glucose n = 13, placebo n = 13.
 
@@
 The expression of the 47S pre-rRNA increased by 37 and 59% with glucose and placebo respectively, where only RT with placebo increased significantly from baseline to post (fig 45, *p* = `r rr47tdiffp`).
-
-###### **Figure 2:** Changes in total RNA and ribosomal RNA.
-A) Total RNA, B) 47S pre-rRNA, 18S rRNA, 28S rRNA, 5.8S rRNA, 5S rRNA.
-Baseline = Day 1 leg 1/ Day 2 leg 2, Post = Day 11 leg 1, Day 12 leg 2.
-Total RNA and rRNA were analyzed in duplicates, with two duplicates per biopsy (two muscle tissue pieces per time point), and normalized to ng x mg wet muscle weight for total RNA and external reference gene (Lambda) for rRNA.
-Total RNA and rRNA changes were calculated as log-fold change score per mg wet muscle weight.
-Mean change scores of the duplicates were calculated and transformed to the log-scale before modelling, then reverse-transformed for figure illustration.
-Values are estimated marginal means fold change per leg per supplement ± 95% CI.
-Glucose n = 13, placebo n = 13.
 
@@
 Further, we found a linear relationship between UBF and total RNA, where an increase of 1 SD unit of UBF equated to a 13% increase in total RNA (*p* = `r corrp`).
-
-###### **Figure 3:** Changes in levels of protein.
-A) Mean UBF, c-Myc and rpS6 (AU, arbitrary units) at baseline and post + representative western blots of the respective proteins.
-Baseline = Day 1 leg 1, Day 2 leg 2, post = Day 11 leg 1, Day 12 leg 2.
-GLU = glucose, PLA = placebo.
-Protein samples were analyzed in two duplicates per biopsy per time point, loaded on separate gels in an inverted order, e.g.
-from gel 3 to 6, as shown by the duplicates (1 and 2).
-Change in protein levels were calculated as log-fold change scores normalized by pools (pool of all protein samples per gel).
-Mean log-change scores of the duplicates were calculated before modelling and reverse-transformed for figure illustration.
-Values are estimated marginal means fold change per leg per supplement ± 95% CI.
-* = *p* < 0.05 between groups.
-Glucose n = 13, placebo n = 13.
-B) Representative total protein stain blot.
-C) Linear relationship between total RNA (ng x mg) and UBF levels (SD units), with time added as a covariate.
-Total RNA was normalized by wet muscle weight, and UBF was normalized by pools per gel, and total protein per lane factor.
-Values are presented as log-transformed means.

### 2023-03-10 ribose-paper.RMD
updates to intro and methods in ribose-paper.RMD

@@
 Such inability to benefit from RT may be explained by genetic and/or epigenetic disposition, affecting the internal physiological milieu in terms of ability to mount an anabolic response to RT [@figueiredo21;@thalacker13].<!-- Check these refs, do they correspond to genet epigen and internal mileu?
-checked-->
-Although RT program modifications such as altered exercise volume or intensity may induce more favorable adaptations in some individuals, such benefits are not uniform between individuals [@hammarstrom.etal_2020;@krieger09;@schoenfeld17].<!-- Add ref for more rt program modifications.
-checked -->
-For example, although changes in training volume generally induces greater benefits of RT, not all individuals benefit from such modifications [@hammarstrom.etal_2020].
+Kristian: The way I read it, it looks like it corresponds --> Although RT program modifications such as altered exercise volume or intensity may induce more favorable adaptations in some individuals, such benefits are not uniform between individuals [@hammarstrom.etal_2020;@krieger09;@schoenfeld17].<!-- Add ref for more rt program modifications.
+Kristian: Added krieger et al.
+2009 and schoenfeld et al.
+2017 --> For instance, although changes in training volume generally induces greater benefits of RT, not all individuals benefit from such modifications [@hammarstrom.etal_2020].
 Therefore, other means than modification of RT variables seems necessary to optimize individual responses.<!-- The above sentence might be a bit contradictory?
+Im not sure, are you thinking about "other means than modification of RT" being contradictory of "modifications such as altered exercise volume.." sentence?
+I think there is a point to establishing that yes, increasing exercise volume is a potent way to increase the RT stimulus in general.
+However, in cases where this alone fails to elicit true responses, other means are necessary.
+Or am I missinterpreting this?
 -->
@@
 MOSS.
-checked  --> which in turn determines muscle growth by increasing the muscles translational capacity [@figueiredo19;@hammarstrom20;@tanaka18;@stec16;@vonwalden19].
+Kristian: These three should do it--> which in turn determines muscle growth by increasing the muscles translational capacity [@figueiredo19;@hammarstrom20;@tanaka18;@stec16;@vonwalden19].
 Therefore, investigating the effects of combined RT and glucose ingestion may provide valuable insight into a potential additive effect of glucose and RT on ribosome biogenesis.
@@
 could add West 2016.
-c-Myc included in Von Walden and Tanaka, adding West 2016 and Mori 21 when I find them-->
+Kristian: c-Myc included in Von Walden and Tanaka, adding West 2016 and Mori 21 when I find them-->
 mTOR receives input from growth factors, hormones, mechanical loading, and nutrients to balance protein synthesis through multiple mechanisms based on cellular energy levels [@hoppe09].
@@
 The purpose of this investigation was to test the hypothesis that glucose supplementation before and after<!-- during?
-before and after would be more precise --> RT would increase markers of ribosomal abundance following five RT sessions.
+Kristian: before and after would be more precise, since they're not strictly ingesting anything during RT - I got a comment on that during the thesis defense --> RT would increase markers of ribosomal abundance following five RT sessions.
 Secondly we aimed to describe the association between UBF and total RNA abundance in human skeletal muscle.
-<!-- We should maybe add a line or two about the rational here -->
+There has been multiple studies the last years suggesting translational capacity to be as important as, if not more, translational efficiency [@hammarstrom20;@hammarstrom21;@figueiredo19rev].
+Previously, *in vitro* studies has shown UBF and c-Myc to be a important factors in the PIC, c-Myc as a general transcription factor and UBF as a specific transcription factor to rDNA transcription initiation [@mariappan11;@vonwalden2019].
+Coupled with the recent observations by Hammarström et al.
+[@hammarstrom20;@hammarstrom21], understanding underlying factors regulating and affecting ribosome biogenesis seem key to further our understanding in optimization of RT to individual phenotypes.
+<!-- We should maybe add a line or two about the rational here. Kristian: Tried to add some, what about this? to general? -->
 
@@
 The study was designed as a 12-day double-blinded randomized controlled trial, with a alternating unilateral RT protocol (Fig 1A).
-Participants were randomly allocated to exercise either their dominant or non-dominant leg with glucose or placebo before and after RT, and whether they started RT with glucose or placebo.
-Thus, participants exercised every day, alternating between e.g., one leg + glucose and the other leg + placebo (1A).
-One person was exclusively responsible for the randomization code and supplement distribution, blinding both investigators and participants in regard to which leg exercised with glucose and placebo.
+Participants were randomly allocated to exercise one leg with glucose and one leg with placebo (1A).
+<!--from the "old" methods: either their dominant or non-dominant leg with glucose or placebo before and after RT, and whether they started RT with glucose or placebo.
+Thus, participants exercised every day, alternating between e.g., one leg + glucose and the other leg + placebo - this might be TMI, as it's presented in the figure as well?--> One person was exclusively responsible for the randomization code and supplement distribution, blinding both investigators and participants in regard to which leg exercised with glucose and placebo.
 Glucose (Glucosum monohydricum, Merck KGaA, Darmstadt, Germany) and placebo (Steviosa, Soma Nordic AS, Oslo, Norway) was masked by mixing with 300ml Fun Light juice/saft (Orkla, Oslo, Norway).
@@
 ~0.3g Stevia rebaudiana extract).
-To ensure equal conditions during training sessions, participants exercised and tested at the same time of day, +/- 1hr (check this) with the same supervisor, in addition to recording and repeating their daily macro nutrient intake (protein, fat, carbohydrate) and total calories on pairwise consecutive days.
+To ensure equal conditions during training sessions, participants exercised and tested at the same time of day, +/- 1hr (check this) with the same supervisor on pairwise consecutive days.
+To further standardize this, participants also recorded and repeated their daily macro nutrient intake (protein, fat, carbohydrate) and total calories on pairwise consecutive days.
 
@@
 For safety and standardization purposes, all sessions were monitored by trained personnel.
-To further ensure equal exercise premises for the two legs, each participant was accompanied by the same trainer in each session.
-If this was not possible throughout the intervention, each participant had the same trainer on pairwise consecutive days as a minimum.
-The rate of perceived exertion (RPE, 0–10-point scale) was logged before every session, and the session score was logged 15min after each session.
-Lastly, training volume (i.e., load and repetitions) was logged for every session.
+<!-- From the "old" methods: The rate of perceived exertion (RPE, 0–10-point scale) was logged before every session, and the session score was logged 15min after each session - is this necessary info, as we are not reporting these data?
+--> Lastly, training volume (load and repetitions) was logged for every session.
 
@@
 Skeletal muscle strength was measured as peak torque in unilateral isometric and isokinetic (at 60 and 240 d/s) knee extension before, multiple times during, and after five and six session.
-A total of three participants dropped out of the intervention, either due to reported muscular discomforts of heavy RT, or sickness in their family.
+<!-- A total of three participants dropped out of the intervention, either due to reported muscular discomforts of heavy RT, or sickness in their family - necessary to include this here?
+Its already reported under "experimental design" -->.
 B and C) Changes in plasma glucose (B, mmol/L) and c-peptide levels (C, pmol/L).

### 2023-03-13 results_work.RMD
ribose-paper.RMD updates

@@
 
-astr.lmod <- confint(emmeans(astr.lmod, specs = ~"supplement|time")) %>%
+astr.lemm <- confint(emmeans(astr.lmod, specs = ~"supplement|time")) %>%
   data.frame()
@@
 
-There were no difference in mean change of total training session volume between exercising with placebo or glucose, with an increase of 18 and 16% from session 1 to session 6, respectively (Tab 2; Mean change, 2: *p* = `r pvol.c2`, 3: *p* = `r pvol.c3`, 4: *p* = `r pvol.c4`, 5: *p* = `r pvol.c5`, 6: *p* = `r pvol.c6`).
+There was no difference in mean change of total training session volume between exercising with placebo or glucose, with an increase of 18 and 16% from session 1 to session 6, respectively (Tab 2; Mean change, 2: *p* = `r pvol.c2`, 3: *p* = `r pvol.c3`, 4: *p* = `r pvol.c4`, 5: *p* = `r pvol.c5`, 6: *p* = `r pvol.c6`).
 Both exercising with glucose and placebo led to significant increases in mean total training session volume at session four, five and six, compared to session 1 (Tab 2; Mean change, 2: *p* = `r pvol.s2`, 3: *p* = `r pvol.s3`, 4: *p* = `r pvol.s4`, 5: *p* = `r pvol.s5`, 6: *p* = `r pvol.s6`).
@@
 
-```{r western modelling}
+```{r western model}
 ## Script: 

### 2023-03-13 ribose-paper.RMD
ribose-paper.RMD updates

@@
 
-Levels of c-Myc increased by 51% more after RT with placebo compared to RT with glucose, a significant difference (Fig 3A: *p* = `r cdiff`).
-Levels of UBF and rpS6 increased by 42 and 33% more in placebo compared to glucose, however, this was not a significant difference (Fig 3A: UBF: *p* = `r udiff`, rpS6: *p* = `r rpdiff`).
+Levels of c-Myc increased by 42% more after RT with placebo compared to RT with glucose, a significant difference (Fig 3A: *p* = `r cdiff`).
+Levels of UBF and rpS6 increased by 19 and 23% more in placebo compared to glucose, however, this was not a significant difference (Fig 3A: UBF: *p* = `r udiff`, rpS6: *p* = `r rpdiff`).
 The increase in both RT with glucose and placebo for all measured proteins was significant at post compared to baseline (*p* < 0.05).

### 2023-03-13 R/scraps/facet.index.R

@@ -1,0 +1,134 @@
+
+
+library(tidyverse)
+
+str.emm <- readRDS("./data/data-gen/humac/emm.str.RDS")
+
+str.emm2 <- str.emm %>% 
+  filter(time %in% c("change.2", "change.3", "change.4")) %>%
+  mutate(tp = if_else(time %in% c("change.2", "change.3", "change.4"),
+                                  "strength", time),
+         timep = if_else(time == "change.2",
+                         "Post RT2",
+                         if_else(time == "change.3",
+                                 "Post RT4",
+                                 if_else(time == "change.4",
+                                         "Post RT5", time)))) %>%
+  add_row(supplement = "placebo", timep = "Baseline", 
+          tp = "strength", emmean = 0, SE = 0, 
+          df = 0, lower.CL = 0, upper.CL = 0, .before =1) %>%
+  add_row(supplement = "glucose", timep = "Baseline", 
+          tp = "strength", emmean = 0, SE = 0, 
+          df = 0, lower.CL = 0, upper.CL = 0, .before =2) %>%
+  mutate(timep = factor(timep, levels = c("Baseline", "Post RT2", "Post RT4", "Post RT5"))) %>%
+  select(-time)
+
+base.astr <- str.emm2 %>%
+  filter(timep == "Post RT5") %>%
+  mutate(tp = if_else(tp == "strength",
+                      "recov", tp),
+         timep = if_else(timep == "Post RT5",
+                         "baseline", timep)) %>%
+  print()
+
+astr.emm <- readRDS("./data/data-gen/humac/emm.astr.RDS")
+
+astr.emm2 <- astr.emm %>% 
+  mutate(tp = if_else(time %in% c("change.2", "change.3", "change.4"),
+                      "recov", time),
+         timep = if_else(time == "change.2",
+                         "30min post RT6",
+                         if_else(time == "change.3",
+                                 "2hrs post RT6",
+                                 if_else(time == "change.4",
+                                         "23hrs post RT6", time)))) %>%
+  full_join(base.astr) %>%
+  mutate(timep = factor(timep, levels = c("baseline", "30min post RT6", "2hrs post RT6", "23hrs post RT6"))) %>%
+  select(-time)
+
+
+
+### Joining
+#
+#str.joined <- str.emm2 %>%
+#  full_join(astr.emm2) %>%
[…]

### 2023-03-13 R/strength.index.R

@@ -163,6 +163,8 @@ lemm.str <- confint(emmeans(m.lstr, specs = ~"supplement|time")) %>%
 saveRDS(emm.str, "./data/data-gen/humac/emm.str.RDS")
 
 
+
+
 ### Change model acute
 
 m.astr <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
@@ -179,8 +181,12 @@ summary(m.alstr)
 emm.astr <- confint(emmeans(m.astr, specs = ~"supplement|time")) %>%
   data.frame() 
 
+saveRDS(emm.astr, "./data/data-gen/humac/emm.astr.RDS")
+
 lemm.astr <- confint(emmeans(m.alstr, specs = ~"supplement|time")) %>%
   data.frame() 
+
+saveRDS(lemm.astr, "./data/data-gen/humac/lemm.astr.RDS")
 
   
   

### 2023-03-13 R/ubf_rna.correlation.R

@@ -69,26 +69,116 @@ joined.dat %>%
   geom_smooth(method = "lm")
 
 
+######################## c-Myc below ######################
 
-## pre-post values correlation
-#ggscatter(joined.dat, x = "mean.sign", y = "mean.rna", 
-#          add = "reg.line", conf.int = TRUE, 
-#          cor.coef = TRUE, cor.method = "pearson",
-#          xlab = "UBF", ylab = "Total RNA")
-#sprintf("%.5f", 3.7e-06)
-#
-#
-#stp.ubf <- shapiro.test(joined.dat$mean.sign) 
-#sprintf("%.5f", 1.667e-05)
-#shapiro.test(joined.dat$mean.rna) 
-#
-#ggqqplot(joined.dat$mean.sign, ylab = "mean.sign")
-#ggqqplot(joined.dat$mean.rna, ylab = "mean.rna")
-#
-#correlation <- cor.test(joined.dat$mean.sign, joined.dat$mean.rna, 
-#                method = "pearson")
-#correlation
-#
-#corr.p <- sprintf("%.5f", 3.694e-05)
-#corr.p
+## Prepare data frames for joining
+
+cmyc.rdy <- west.dat %>%  # One data frame without outliers
+  select(subject, time, outlier, supplement, target, norm.sign, tpl) %>%
+  filter(target == "cmyc",
+         outlier == "in") %>%
+  select(-target) %>%
+  mutate(norm.sign = norm.sign / tpl) %>%
+  #print()
+  group_by(subject, time, supplement) %>%
+  summarise(mean.sign = mean(norm.sign, na.rm = TRUE)) %>%
+  print()
+
+
+rna.rdy <- rna.dat %>%
+  select(subject, time, outlier, supplement, RNA.weight) %>%
+  filter(outlier == "in") %>%
+  mutate(subject = as.character(subject)) %>%
+  group_by(subject, time, supplement) %>%
+  summarise(mean.rna = mean(RNA.weight, na.rm = TRUE)) %>%
+  print()
+
+
+## Correlation analysis
+
+joined.dat <- cmyc.rdy %>%
+  right_join(rna.rdy) %>%
+  #filter(supplement != "GLUCOSE") %>%
+  
+  mutate(sd.cmyc = (mean.sign - mean(mean.sign, na.rm = TRUE )/sd(mean.sign, na.rm = TRUE))) %>%
+  print()
+
+
+library(nlme)
+
+m <- lme(mean.rna ~ mean.sign + time , 
[…]

### 2023-03-13 R/western.analysis.R

@@ -122,7 +122,7 @@ ubf.emm %>%
   geom_line() + 
   geom_point(shape= 21) 
 
-saveRDS(ubf.lemm, "./data/data-gen/protein/ubf.emm.RDS")
+saveRDS(ubf.emm, "./data/data-gen/protein/ubf.emm.RDS")
 
 # Log-transformed
 ubf.lemm <- emmeans(ubf.lmod, specs = ~ time|supplement) %>%

### 2023-03-14 ribose-paper.RMD
updated discussion, needs more work

@@
 
-The main findings of the present study were that resistance training with glucose for five sessions did not affect markers of ribosome biogenesis, measured as total RNA, rRNA and rDNA transcription initiation associated proteins, compared to RT with placebo.
-Similarly, RT with glucose did not affect markers of skeletal muscle functionality such as muscular performance and -recovery, total session volume, and training intensity, compared to RT with placebo.
-There are studies suggesting that high vs.
-low glucose treatment/glucose starvation impacts ribosome biogenesis in yeast, rodent and human non-muscle cells (KIM13;MARIAPPAN11;TANAKA15;ZHAI12), however, this effect does not seem evident when total daily glucose intake is matched.
+The main findings of the present study were that heavy resistance training with glucose did not affect markers of ribosome biogenesis, measured as total RNA, rRNA and rDNA transcription initiation associated protein, compared to RT with placebo.
+Similarly, RT with glucose in general did not affect markers of skeletal muscle functionality such as strength and -recovery, or total training session volume.
+At one time point, RT with glucose led to less reduction in strength compared to placebo, and levels of plasma glucose and c-peptide were significantly higher before, during and after the RT sessions with glucose.
+Further, there were no differences in daily macronutrient intake on pairwise consecutive days.
+This would suggest the effects of high vs.
+low glucose/glucose starvation treatment on ribosome biogenesis or rDNA transcription initiation observed in previous in vitro studies [@mariappan11; @tanaka15] are not translatable to acute effects in human skeletal muscle in vivo.
 
-To our knowledge, no previous study has investigated the effects of glucose compared to placebo ingestion with RT on ribosome biogenesis in human skeletal muscle cells, making our hypotheses exploratory and comparisons not straight forwards.
-The current model of ribosome biogenesis suggests that changes in the accumulation of total RNA and rRNA early after the onset of an RT period predict long-term RT adaptations (FIGUEIREDO19; HAMMARSTROM20; HAMMARSTROM21).Therefore, we assumed 2 weeks to be a suitable time frame for a moderate RT volume to elicit changes in ribosome biogenesis, as similar observations have been made previously (FIGUEIREDO16; HAMMARSTROM20; HAMMARSTROM21).
-In Hammarström et al.
-(2020), participants that increased ribosome biogenesis after two weeks (5 RT sessions) of RT, also responded well after the entire intervention (12 weeks, 31 sessions).
-Figueiredo and colleagues (2016) also observed markedly acute effects, where 47S pre-rRNA increased after a single RT session, passing baseline after 4hrs, peaking at 24hrs and remaining elevated until 48hrs post-exercise.
-The four mature rRNA did not measurably increase until after multiple sessions (FIGUEIREDO16).
-As such, increases in total RNA and rRNA observed in the present study are in line with previous observations (FIGUEIREDO16; HAMMARSTROM20; H21?), and also evident i moderately trained individuals, compared to the untrained or RT-untrained individuals of previous studies (FIGUEIREDO16;HAMMARSTROM20;H21?).
+To our knowledge, no previous study has investigated the effects of glucose compared to placebo ingestion with RT on ribosome biogenesis in human skeletal muscle, making our hypotheses exploratory and comparisons not straight forwards.
+Mariappan and colleagues (2011) investigated the effect of high glucose treatment on ribosome biogenesis in glomerular epithelial cells (GECs) from diabetic (type 1 and 2) mice, observing a UBF-dependent augmentation of ribosome biogenesis.
+In a study from 2015, Tanaka et al.
+(2015) observed that mild glucose starvation reduced rRNA transcription in human breast cancer cells.
+These findings do imply an important role for glucose in the initiation of rDNA transcription, the rate-limiting step of ribosome biogenesis [@mariappan13; @tanaka15].
+In the present study, we did not find such indications of glucose having an effect as these previous studies.
+Notably, kidney cells from mice and human breast cancer cells are likely to have a higher physiological rate compared to human skeletal muscle cells, thus, our two-week intervention may not be directly comparable to these findings, as was expected.
+Nevertheless, we did observe robust increases in both total RNA and 47S pre-rRNA and all four mature rRNAs (18S, 28S, 5.8S, 5S) after five RT sessions with both treatments.
+This means that the time frame and training volume chosen for this design was adequate to elicit changes in rDNA transcription and subsequently ribosome biogenesis by proxy.
+Further, the training, testing and biological sampling were conducted under standardized conditions, and there were no differences between pairwise consecutive days in macronutrient ingestion and training volume.
+As such, the day-to-day and test-to-test variation should have been minimal, especially when eliminating between-subjects biological variation with a within-subjects design as well (SOURCE).
+Importantly, both Mariappan (11) and Tanaka (15) compared high glucose treatment to either low glucose treatment or glucose starvation.
+In the present study however, we compared RT with either protein and glucose or protein and placebo.
+The participants fasted from 2200hrs the preceding day until the morning of each training day, receiving only the prescribed treatment until 2.5 hrs after RT.
+Thus, except from approximately 0700hrs-1230hrs, the participants ingested an equal amount of carbohydrates on pairwise days.
+As such, there were no differences in baseline levels of plasma glucose prior to the daily intervention, as by design, since we were interested in investigating if there were direct effects of glucose as a signalling molecule, based on the observations of Tanaka et al.
+[@tanaka2015] and Mariappan et al.
+[@mariappan2011].
+Further, the significant difference in levels of plasma glucose and c-peptide measured before, during and after session six, confirms that the treatment during of RT with glucose indeed was sufficient.
+Despite of this, we found no difference in markers of ribosome biogenesis between RT with glucose and RT with placebo, this does suggest that there are no additional effects of exercising with glucose, when total macronutrient intake is standardized.
 
-Further, the accumulation of total RNA and rRNA is also suggested to reach a plateau after approximately 5 RT-sessions, thus, in addition to the demanding RT- and dietary protocol and the exploratory nature of the hypotheses, we did not conduct a more long-term intervention.
-COMMENT ON PROTEIN DATA HERE?
+Prior to the present study, it was theorized that glucose may convey an anabolic signal, either directly or through insulin, and thus operate as a signal mediator affecting upstream pathways of ribosome biogenesis [@kusnadi15; @tanaka18].
+This theoretical anabolic signal could possibly stimulate PIC assembly and Pol I recruitment to the rDNA promoter, initiating rDNA transcription and subsequent rRNA maturation, and lastly increased ribosome content of the skeletal muscle cell [@figueiredo19; @tanaka18].
+Thus, we hypothesized that if glucose directly or indirectly via insulin can stimulate initiation of rDNA transcription, five RT sessions with glucose will elicit this effect compared to five RT sessions with placebo.
+This time frame has previously shown to be sufficient to measure changes in markers of ribosome biogenesis [@hammarstrom20; @hammarstrom21; @figueiredo16].
+<!--double-check these references--> As no such effect was observable in the present study, this hypothesis was rejected.
+Further, glucose ingestion affects cellular energy levels; thus another theory is that high glucose treatment would stimulate energy-sensitive pathways such as PIH1, mTORC1, ERK1/2, AMPK and SIRT1 in a dose-response relationship [@kim13; @mariappan11; @tanaka15; @zhai12].
+<!-- check if this is right --> Mammalian target of rapamycin complex 1, along with ERK1/2 and PIH1 upregulates PIC assembly during high-energy conditions, while AMPK and possibly SIRT1 represses assembly of the PIC during low-energy conditions, as observed in rodents [SOURCE], yeast [SOURCE], and human non-muscle cells [SOURCE] *in vitro*.
+Specifically, high glucose treatment augmented rDNA transcription by increased UBF and SL-1 activity, while low glucose treatment, glucose starvation, or nutrient depletion stimulated AMPK-dependent TIF-1A and chromatin regulation suppression [@kos17: @mariappan11; @zhai12].
+<!--check these references--> In addition to UBF and SL-1, assembly of the PIC requires TIF-1A and Pol I recruitment, induced by amino acid intake in an mTOR-S6K-dependent manner and possibly by direct c-Myc activation [@kusnadi15].
+As such, increasing energy supply by RT with glucose vs.
+RT with placebo and nutrient availability should in line with these observations potently stimulate PIC assembly and rDNA transcription.
+Though we did not investigate these pathways directly, the downstream targets UBF and rpS6 did not differ in mean change between treatments.
+As the participants in the present study fasted until the daily training was completed, and matched their daily energy intake on pairwise days, this suggests that as long as the daily energy requirements are met, ingesting glucose with RT does not stimulate further gains in anabolic responses.
+Notably, the daily RT protocol was quite short (~30min), which may have reduced the effectiveness of the energy addition through glucose.
+In this regard, it would have been interesting to measure the intramuscular glycogen storages, to see how the RT sessions impacted intracellular energy levels.
+Unfortunately, such analyses were not conducted in the present study.
+Regrettably, this is a weakness of the design, as we have not measured directly what happened to the excess plasma glucose.
+Nevertheless, the absent effect of increasing energy availability during RT suggest that, there is no dose-response relationship between acutely increasing myocellular energy levels before and after RT and ribosome biogenesis.
 
-As we did not observe any effect of ingesting glucose before and after RT compared to placebo in this design, there is little to suggest that glucose *per se* has any additional effect on ribosome biogenesis, as long as the daily energy intake is sufficient to support cellular energy levels required to drive this process.
-Looking at previous observations from *in vitro* studies of rodents, yeast, and non-muscle cell cultures (MARIAPPAN11; KIM13; TANAKA18; ZHAI12; HOPPE09; MURAYAMA08), one interesting notion would be how high glucose vs.
-low glucose or glucose starvation would affect ribosome biogenesis over time in human skeletal muscle cells.
-Time being of the essence in such a perspective, as previous *in vitro* studies have utilized cells with higher physiological rate than human skeletal muscle cells (KIM13;MARIAPPAN11;TANAKA15;ZHAI12).
+Interestingly, we did observe a linear relationship between total RNA and UBF, suggesting ribosome biogenesis to be dependent on UBF activation, in line with observations in Mariappan et al.
+(2011).
+This was not evident with total RNA and c-Myc or rpS6, which we observed no linear relationship between.
+The upstream binding factor (UBF) has indeed previously been described as a master regulator of rDNA transcription, based on findings from cell cultures [@figueiredo19; @kusnadi15; russel&zomerdijk2005].
+However, it has only recently been found in a linear relationship with total RNA content in human skeletal muscle cells after a short period of RT [@hammarstrom21], which the observation of the present study supports.
+As such, UBF also seem to respond to mechanical loading in human muscle cells in line with responses seen in cell cultures and synergist ablation models [@mariappan11; @vonwalden18], in addition to the previously reported response to hormonal, nutritional, and cellular energy signals [@figueiredo19; @kusnadi15; russel&zomerdijk2005].
+While c-Myc and rpS6 did not increase linearly with accumulation of total RNA, the cellular content of c-Myc and rpS6 increased from baseline to post with both treatments.
+The transcription initiation factor c-Myc has been implicated as important to ribosome biogenesis [@vonwalden19; @figueiredo19], though its role is likely as a general transcription factor rather than specific, compared to UBF [SOURCES?].
+Further, the ribosomal protein S6 (rpS6) has previously been shown to correlate strongly with rRNA content after synergist ablation [@nakada16; @chaillou], <!--check Nakada 16 for these refs--> and is suggested as a reliable measurement of ribosome biogenesis [@nakada16].
+Together, this may explain why the between treatment difference in c-Myc content did not induce differences in total RNA or rRNA content, as the specific transcription factor UBF and the rpS6 increased equally with both treatments.
+Notably, the difference between c-Myc in glucose and placebo may have caused responses in pathways not measured in the present study.
 
-Indeed, our findings suggest that glucose ingestion *per se* before and after RT does not induce changes in the internal physiological milieu of human skeletal muscle cells with anabolic effects.
-It has been theorised that glucose may convey an anabolic signal through insulin, and thus operate as a signal mediator affecting upstream pathways of ribosome biogenesis (KUSNADI15; TANAKA18).
-Such an anabolic signal could then possibly increase PIC assembly, and consequently Pol I recruitment, initiating rDNA transcription and subsequent rRNA maturation, leading to increased ribosome content (FIGUEIREDO19; TANAKA18).
-As significantly higher blood levels of c-peptide and glucose were observed in RT with glucose compared to placebo in the present study, without any difference in ribosome biogenesis, there is little to suggest that either insulin or glucose has any direct acute effects.
+The results from the strength tests do indicate that the participants might have been fatigued during the intervention, as they exercised every day for 12 days straight, without breakfast other than the treatments supplied.
+In this perspective, we expected glucose to increase energy availability and thus, muscle performance during strength testing.
+However, this was not evident, as the mean response in strength as an index of peak torque at all speeds decreased with both treatments from baseline to post five and six sessions.
+Notably, RT with glucose reduced strength significantly less than RT with placebo post five sessions, which may indeed have been an effect of increased energy availability, seeing as glucose is the preferred energy source during strenuous activity [@mul15].
+<!--check this ref--> Arguably, this might imply that RT with placebo experienced a higher stress response, thus less performance.
+This could also explain the differences between treatments in content of the stress responsive c-Myc [SOURCES], meaning higher stress exerted during RT with placebo compared to RT with glucose inducing higher accumulation of c-Myc.
+If this was the case, it does seem inconclusive that strength did no differ between treatments at earlier or later time points.
+In addition, there were no differences in training volume per session between treatments, therefore, the mechanical loading should have been equal.
+Another potential explanation to this observation might be ….
+(bifasisk restitusjon?)
 
-Another theory is that glucose ingestion leads to higher cellular energy levels compared to placebo, stimulating energy-sensitive pathways such as PIH1, mTORC1, ERK1/2, AMPK and SIRT1 in a dose-response relationship (KIM13;MARIAPPAN11;TANAKA15;ZHAI12).
-The mTORC1, along with ERK1/2 and PIH1 up-regulates PIC assembly during high-energy conditions, while AMPK and possibly SIRT1 represses PIC assembly during low-energy conditions, as observed in rodent (KILDE), yeast (KILDE), and non-muscle cells (KILDE) *in vitro*.
-Specifically, high glucose treatment augmented rDNA transcription by increased UBF and SL-1 activity, while low glucose treatment, glucose starvation, or nutrient depletion stimulated AMPK-dependent TIF-IA and chromatin regulation suppression (KOS-BRAUN17; MARIAPPAN11; ZHAI12).
-In addition to UBF and SL-1, assembly of the PIC requires TIF-IA and Pol I recruitment, induced by amino acid intake in an mTOR-S6K-dependent manner and possibly by direct c-Myc activation (KUSNADI15).
-As such, increasing energy supply and nutrient availability should in line with these theories potently stimulate PIC assembly and rDNA transcription.
-Importantly, these pathways were not directly investigated in the present study.
-However, downstream targets (UBF, c-Myc, rpS6?) remained unaffected by glucose compared to placebo ingestion, suggesting there is not a dose-response relationship between muscle cellular energy levels and ribosome biogenesis during an RT-intervention (Not sure if I can say this).
-
-A weakness in the present study, is the lack of data on intracellular glucose levels in the muscle cells.
-Thus, it cannot be said for sure whether there were differences between glucose and placebo or not.
-
-A possible explanation for the lack of effect of glucose compared to placebo, is that both treatments may have exercised with sufficient levels of nutrients and energy, meaning that mTORC1 had sufficient nutrient availability and, thus was not inhibited by AMPK due to energy deficiency (HOPPE09; KUSNADI15).
-Indeed, mTORC1 is characterised as energy-sensitive (DENNIS01), and therefore it might have been directly affected by high vs.
-low-glucose conditions (MARIAPPAN11).
-However, previous studies suggest that mTORC1 is not sensitive enough to register levels of ATP, rather depending on signals from AMPK to assess ATP levels (KIM10).
-This would arguably make AMPK the primary regulator of the potential impact of glucose on ribosome biogenesis, and thus, glucose has no direct effect on mTORC1 signalling (REFERENCES?).
-Importantly, intramuscular glycogen content was not measured in the present study, therefore we cannot determine whether cellular energy levels in fact were equal or not, or if there is reason to assume that AMPK-dependent mTORC1 signalling was different in the two treatments.
-Potentially, such data could give higher resolution to our data.
-Along with the robust increases in proxy markers of ribosome biogenesis, the equal macronutrient intake between treatments on pairwise consecutive days and plasma glucose levels not indicating energy deficiency during RT, potent mTORC1 stimulation without AMPK-dependent inhibition may be a plausible explanation.
-Together with previous studies on this topic (MARIAPPAN11; ZHAI12), there are currently scarce observations to suggest glucose carries a direct anabolic signal.
-As such, physiological levels of plasma glucose during RT, combined with a sufficient daily macronutrient intake, seem potent enough to supply the energy needed for ribosome biogenesis.
-The lacking differences between treatments in the biological data are further supported by our observations on functional data.
-
-Overall, glucose did not affect the isometric or isokinetic peak torque at 240 d/s, or muscular recovery following the sixth and last RT session, measured as temporal changes in isometric and isokinetic peak torque.
-Generally, the mean isometric and isokinetic peak torque slightly decreased during the intervention, especially between session 4 and 6, and regained after the intervention.
-Further, there were no differences between glucose and placebo in total session volume or training intensity.
-Both exercise with glucose and placebo yielded higher total session volumes at higher percentage of 1RM at the last session compared to baseline, as expected from earlier reports on the moderate-volume RT protocol (KRIEGER09; RHEA03; SCHOENFELD17), in addition to a more recent study implementing the unilateral protocol (HAMMARSTROM20+21/22).
-Interestingly, we did observe a significant difference between treatments in mean change of isokinetic peak torque at 60 d/s before the sixth RT session, where exercise with glucose decreased peak torque 9% less compared to placebo.
-However, as none of the other measurements indicated any difference between treatments, a true effect of glucose seems unlikely.
-Taken together, these results may add to previous observations with no differences between co-ingestion of carbohydrate and protein compared to protein alone, in regard to skeletal muscular recovery following RT (KOOPMAN07; STAPLES11).
-Collectively, the functional data point towards an exhaustive, yet growth-inducing RT protocol, especially considering the biological data.
-As a side note, an interesting observation was made by another master’s thesis on this project, finding significantly lower MuRF protein levels at post with glucose treatment compared to placebo.
-This may suggest an effect of glucose on protein degradation (WILKES09), however, whether this contributes to growth or simply protein turnover has been a topic of debate (FIGUEIREDO19).
-
-Despite being a thoroughly standardised design, a limitation may be posed through lack of participants.
-We initially reach our minimum limit for statistical power, i.e.
-16 participants, however, three dropped out during the intervention, possibly impacting our power.
-Coupled with the fact that no previous study could guide our power calculations, this, of course, introduces some degree of uncertainty regarding our results.
-However, we’ve reported little to suggest any true acute effect of RT with glucose ingestion compared to placebo.
-Moreover, the study was specifically designed to investigate differences in pre-post changes within participants, and thus also quite limited to this perspective.
-Hence, the effects of time observed exclusively serve as supplementary to the efficiency of the design, without a negative control group to either confirm or refute them.
-If further investigations are to be done on this topic, these analyses should include more participants and optimally a measure of intramuscular glycogen levels.
-
-In conclusion, resistance exercise training combined with glucose ingestion does not acutely augment ribosome biogenesis following 2 weeks of moderate-volume resistance training in moderately trained young male and female adults.
-Neither does it acutely affect muscular performance in isometric and isokinetic knee extension peak torque, nor enhance post resistance training recovery within 24hrs.
+Despite a thoroughly standardised design, the lack of participants do present a weakness of the present study.
+Initially, the minimum limit for statistical power was reached however, three participants dropped out during the intervention.
+Coupled with the fact that no previous study could guide our power calculation, this does introduce a degree of uncertainty regarding our results.
+Moreover, the study was specifically design to investigate differences in pre-post changes within-participants, and thus also quite limited to this perspective.
+Hence, the observed effects of time serve as supplementary to the efficiency of the design, without a negative control group to either confirm or refute them.
+If further investigations are to be made on this topic, these analyses should include a greater sample size, with a negative control.
+In conclusion, resistance exercise training combined with glucose ingestion does not acutely augment ribosome biogenesis following two weeks of heavy resistance training in moderately trained young male and female adults.
+Neither does it acutely enhance muscular performance or recovery withing 23hrs, measured as unilateral isometric and isokinetic knee extension peak torque.
 

### 2023-03-15 ribose-paper.RMD
last updates discussion

@@
 low glucose/glucose starvation treatment on ribosome biogenesis or rDNA transcription initiation observed in previous in vitro studies [@mariappan11; @tanaka15] are not translatable to acute effects in human skeletal muscle in vivo.
+Notably, the study was specifically design to investigate differences in pre-post changes within-participants, and thus also quite limited to this perspective.
+Hence, the observed effects of time serve as supplementary to the efficiency of the design, without a negative control group to either confirm or refute them.
 
@@
 
-The results from the strength tests do indicate that the participants might have been fatigued during the intervention, as they exercised every day for 12 days straight, without breakfast other than the treatments supplied.
-In this perspective, we expected glucose to increase energy availability and thus, muscle performance during strength testing.
+The results from the strength tests do indicate that the participants might have been fatigued during the intervention, as they exercised every day for 12 days straight, with strength testing in-between.
+In this perspective, we expected glucose to increase energy availability and thus, muscle performance during strength testing compared to placebo.
 However, this was not evident, as the mean response in strength as an index of peak torque at all speeds decreased with both treatments from baseline to post five and six sessions.
 Notably, RT with glucose reduced strength significantly less than RT with placebo post five sessions, which may indeed have been an effect of increased energy availability, seeing as glucose is the preferred energy source during strenuous activity [@mul15].
-<!--check this ref--> Arguably, this might imply that RT with placebo experienced a higher stress response, thus less performance.
+<!--check this ref--> Arguably, this might imply that RT with placebo experienced a higher stress response to the RT program, thus reduced muscle performance in strength tests.
 This could also explain the differences between treatments in content of the stress responsive c-Myc [SOURCES], meaning higher stress exerted during RT with placebo compared to RT with glucose inducing higher accumulation of c-Myc.
-If this was the case, it does seem inconclusive that strength did no differ between treatments at earlier or later time points.
+If this was the case, it does seem inconclusive that strength did not differ between treatments at earlier or later time points.
 In addition, there were no differences in training volume per session between treatments, therefore, the mechanical loading should have been equal.
 Another potential explanation to this observation might be ….
-(bifasisk restitusjon?)
+(bifasisk restitusjon?).
+Lastly, these inconclusive observations may also reflect our limited sample size, despite using a within-subjects design.
+Initially, we met our minimum limit of 16 participants however, three dropped out during the intervention, leaving us slightly underpowered according to the a priori calculation.
 
-Despite a thoroughly standardised design, the lack of participants do present a weakness of the present study.
-Initially, the minimum limit for statistical power was reached however, three participants dropped out during the intervention.
-Coupled with the fact that no previous study could guide our power calculation, this does introduce a degree of uncertainty regarding our results.
-Moreover, the study was specifically design to investigate differences in pre-post changes within-participants, and thus also quite limited to this perspective.
-Hence, the observed effects of time serve as supplementary to the efficiency of the design, without a negative control group to either confirm or refute them.
-If further investigations are to be made on this topic, these analyses should include a greater sample size, with a negative control.
 In conclusion, resistance exercise training combined with glucose ingestion does not acutely augment ribosome biogenesis following two weeks of heavy resistance training in moderately trained young male and female adults.
 Neither does it acutely enhance muscular performance or recovery withing 23hrs, measured as unilateral isometric and isokinetic knee extension peak torque.
+The observations in total RNA, rRNA and rDNA transcription initiation associated proteins confirm the key role of increased UBF content in ribosome biogenesis in moderately trained human skeletal muscle following resistance training, previously observed in RT-untrained participants, rodents or *in vitro* [SOURCES?].
+If future investigations are to be made on this topic, a greater sample size coupled with a negative control group and analyses of the intramuscular glycogen storage may provide higher resolution results.
 

### 2023-03-15 ribose-paper.RMD
updated some refs, more needed

@@
 Responses to systematic resistance training (RT) vary widely between individuals, with as many as 10-15% showing impaired muscle growth in response to standardized training interventions (ALVAREZ18;MANN14;THALACKER13).
-Such inability to benefit from RT may be explained by genetic and/or epigenetic disposition, affecting the internal physiological milieu in terms of ability to mount an anabolic response to RT [@figueiredo21;@thalacker13].<!-- Check these refs, do they correspond to genet epigen and internal mileu?
-Kristian: The way I read it, it looks like it corresponds --> Although RT program modifications such as altered exercise volume or intensity may induce more favorable adaptations in some individuals, such benefits are not uniform between individuals [@hammarstrom.etal_2020;@krieger09;@schoenfeld17].<!-- Add ref for more rt program modifications.
+Such inability to benefit from RT may be explained by genetic and/or epigenetic disposition, affecting the internal physiological milieu in terms of ability to mount an anabolic response to RT [@figueiredo_genetic_2021;@thalacker-mercer_cluster_2013].<!-- Check these refs, do they correspond to genet epigen and internal mileu?
+Kristian: The way I read it, it looks like it corresponds --> Although RT program modifications such as altered exercise volume or intensity may induce more favorable adaptations in some individuals, such benefits are not uniform between individuals [@hammarstrom_benefits_2020;@krieger_single_2009;@schoenfeld_dose-response_2017].<!-- Add ref for more rt program modifications.
 Kristian: Added krieger et al.
 2009 and schoenfeld et al.
-2017 --> For instance, although changes in training volume generally induces greater benefits of RT, not all individuals benefit from such modifications [@hammarstrom.etal_2020].
+2017 --> For instance, although changes in training volume generally induces greater benefits of RT, not all individuals benefit from such modifications [@hammarstrom_benefits_2020].
 Therefore, other means than modification of RT variables seems necessary to optimize individual responses.<!-- The above sentence might be a bit contradictory?
@@
 
-Nutritional supplements such as protein and creatine effectively optimize RT adaptations [@cermak12;@lanhers15;@lanhers17;@morton18].
+Nutritional supplements such as protein and creatine effectively optimize RT adaptations [@cermak_protein_2012;@lanhers_creatine_2015;@lanhers_creatine_2017;@morton_systematic_2018].
 However, it remains unknown if other nutritional adjuvants such as glucose can increase the efficacy of RT.
-This is surprising, as glucose is the preferred energy substrate of the contracting skeletal muscle during strenuous exercise and a major supplier of energy to cells via ATP synthesis [@mul15;@tanaka18].
-Furthermore, energy availability is a decisive factor in *de novo* synthesis of ribosomes [@kusnadi15;@moss07;@tanaka18] <!--Add ref that indicate high energy requirement for ribo biog possibly (VONWALDEN19;TANAKA18?
+This is surprising, as glucose is the preferred energy substrate of the contracting skeletal muscle during strenuous exercise and a major supplier of energy to cells via ATP synthesis [@mul_exercise_2015;@sebata_control_2018].
+Furthermore, energy availability is a decisive factor in *de novo* synthesis of ribosomes [@kusnadi_regulation_2015;@moss_housekeeper_2007;@sebata_control_2018] <!--Add ref that indicate high energy requirement for ribo biog possibly (VONWALDEN19;TANAKA18?
 MOSS.
-Kristian: These three should do it--> which in turn determines muscle growth by increasing the muscles translational capacity [@figueiredo19;@hammarstrom20;@tanaka18;@stec16;@vonwalden19].
+Kristian: These three should do it--> which in turn determines muscle growth by increasing the muscles translational capacity [@figueiredo_regulation_2019;@hammarstrom_benefits_2020;@sebata_control_2018;@stec_ribosome_2016;@von_walden_ribosome_2019].
 Therefore, investigating the effects of combined RT and glucose ingestion may provide valuable insight into a potential additive effect of glucose and RT on ribosome biogenesis.
 
-Transcription of ribosomal RNA (rRNA) by Pol I is considered the limiting step in synthesizing new ribosomes [@Moss_1995].
-Multiple signaling pathways converge to regulate rRNA transcription, including the mammalian target of rapamycin (mTORC1) signal-transduction pathway and c-Myc [@kusnadi15;@vonwalden19;@tanaka18].<!-- Is c_myc included in references here?
-could add West 2016.
-Kristian: c-Myc included in Von Walden and Tanaka, adding West 2016 and Mori 21 when I find them-->
-mTOR receives input from growth factors, hormones, mechanical loading, and nutrients to balance protein synthesis through multiple mechanisms based on cellular energy levels [@hoppe09].
-Importantly, mTOR and its downstream target S6K1 serve as a mediator from insulin signaling in skeletal muscle[@hillier.etal_2000].
-The mTORC1 pathway contributes to ribosome biogenesis through the regulation of translation of ribosomal proteins and to forming of the preinitiation complex (PIC) that marks the initiation of rRNA transcription [@figueiredo19;@vonwalden19].
-Parallel to mTORC1 [@Mori_2021; @West_2016], c-Myc increases ribosomal biogenesis directly and through transcriptional control of the upstream binding factor (UBF) [@sanijUBFLevelsDetermine2008; @Poortinga_2011].
-Additionally, UBF phosphorylation, required for interaction with the rDNA promoter, is increased by high glucose in a mTORC1 dependent manner (rapamycin sensitive) in kidney glomerular epithelial cells[@Mariappan_2011].
-Independently from UBF and mTORC1, high glucose was shown to lead to chromatin remodeling, which promotes rRNA transcription in cell cultures[@Zhai_2012].
+Transcription of ribosomal RNA (rRNA) by Pol I is considered the limiting step in synthesizing new ribosomes [@moss_promotion_1995].
+Multiple signaling pathways converge to regulate rRNA transcription, including the mammalian target of rapamycin (mTORC1) signal-transduction pathway and c-Myc [@kusnadi_regulation_2015;@von_walden_ribosome_2019;@sebata_control_2018; @west_acute_2016; @mori_c-myc_2021].
+mTOR receives input from growth factors, hormones, mechanical loading, and nutrients to balance protein synthesis through multiple mechanisms based on cellular energy levels [@hoppe_amp-activated_2009].
+Importantly, mTOR and its downstream target S6K1 serve as a mediator from insulin signaling in skeletal muscle[@hillier_physiological_2000].
+The mTORC1 pathway contributes to ribosome biogenesis through the regulation of translation of ribosomal proteins and to forming of the preinitiation complex (PIC) that marks the initiation of rRNA transcription [@figueiredo_regulation_2019;@von_walden_ribosome_2019].
+Parallel to mTORC1 [@mori_c-myc_2021; @west_acute_2016], c-Myc increases ribosomal biogenesis directly and through transcriptional control of the upstream binding factor (UBF) [@sanij_ubf_2008; @poortinga_c-myc_2011].
+Additionally, UBF phosphorylation, required for interaction with the rDNA promoter, is increased by high glucose in a mTORC1 dependent manner (rapamycin sensitive) in kidney glomerular epithelial cells[@mariappan_ribosomal_2011].
+Independently from UBF and mTORC1, high glucose was shown to lead to chromatin remodeling, which promotes rRNA transcription in cell cultures[@zhai_human_2012].
 Together these observations indicate a potential role of glucose in positively affecting muscle rRNA synthesis.
@@
 Secondly we aimed to describe the association between UBF and total RNA abundance in human skeletal muscle.
-There has been multiple studies the last years suggesting translational capacity to be as important as, if not more, translational efficiency [@hammarstrom20;@hammarstrom21;@figueiredo19rev].
-Previously, *in vitro* studies has shown UBF and c-Myc to be a important factors in the PIC, c-Myc as a general transcription factor and UBF as a specific transcription factor to rDNA transcription initiation [@mariappan11;@vonwalden2019].
+There has been multiple studies the last years suggesting translational capacity to be as important as, if not more, translational efficiency [@hammarstrom_benefits_2020;@hammarstrom_ribosome_2022;@figueiredo_revisiting_2019].
+Previously, *in vitro* studies has shown UBF and c-Myc to be a important factors in the PIC, c-Myc as a general transcription factor and UBF as a specific transcription factor to rDNA transcription initiation [@mariappan_ribosomal_2011;@von_walden_ribosome_2019].
 Coupled with the recent observations by Hammarström et al.
-[@hammarstrom20;@hammarstrom21], understanding underlying factors regulating and affecting ribosome biogenesis seem key to further our understanding in optimization of RT to individual phenotypes.
+[@hammarstrom_benefits_2020;@hammarstrom_ribosome_2022], understanding underlying factors regulating and affecting ribosome biogenesis seem key to further our understanding in optimization of RT to individual phenotypes.
 <!-- We should maybe add a line or two about the rational here. Kristian: Tried to add some, what about this? to general? -->
@@
 This would suggest the effects of high vs.
-low glucose/glucose starvation treatment on ribosome biogenesis or rDNA transcription initiation observed in previous in vitro studies [@mariappan11; @tanaka15] are not translatable to acute effects in human skeletal muscle in vivo.
+low glucose/glucose starvation treatment on ribosome biogenesis or rDNA transcription initiation observed in previous in vitro studies [@mariappan_ribosomal_2011; @tanaka_mild_2015] are not translatable to acute effects in human skeletal muscle in vivo.
 Notably, the study was specifically design to investigate differences in pre-post changes within-participants, and thus also quite limited to this perspective.
@@
 (2015) observed that mild glucose starvation reduced rRNA transcription in human breast cancer cells.
-These findings do imply an important role for glucose in the initiation of rDNA transcription, the rate-limiting step of ribosome biogenesis [@mariappan13; @tanaka15].
+These findings do imply an important role for glucose in the initiation of rDNA transcription, the rate-limiting step of ribosome biogenesis [@mariappan_ribosomal_2011; @tanaka_mild_2015].
 In the present study, we did not find such indications of glucose having an effect as these previous studies.
@@
 As such, there were no differences in baseline levels of plasma glucose prior to the daily intervention, as by design, since we were interested in investigating if there were direct effects of glucose as a signalling molecule, based on the observations of Tanaka et al.
-[@tanaka2015] and Mariappan et al.
-[@mariappan2011].
+[@tanaka_mild_2015] and Mariappan et al.
+[@mariappan_ribosomal_2011].
 Further, the significant difference in levels of plasma glucose and c-peptide measured before, during and after session six, confirms that the treatment during of RT with glucose indeed was sufficient.
@@
 
-Prior to the present study, it was theorized that glucose may convey an anabolic signal, either directly or through insulin, and thus operate as a signal mediator affecting upstream pathways of ribosome biogenesis [@kusnadi15; @tanaka18].
-This theoretical anabolic signal could possibly stimulate PIC assembly and Pol I recruitment to the rDNA promoter, initiating rDNA transcription and subsequent rRNA maturation, and lastly increased ribosome content of the skeletal muscle cell [@figueiredo19; @tanaka18].
+Prior to the present study, it was theorized that glucose may convey an anabolic signal, either directly or through insulin, and thus operate as a signal mediator affecting upstream pathways of ribosome biogenesis [@kusnadi_regulation_2015; @sebata_control_2018].
+This theoretical anabolic signal could possibly stimulate PIC assembly and Pol I recruitment to the rDNA promoter, initiating rDNA transcription and subsequent rRNA maturation, and lastly increased ribosome content of the skeletal muscle cell [@figueiredo_regulation_2019; @sebata_control_2018].
 Thus, we hypothesized that if glucose directly or indirectly via insulin can stimulate initiation of rDNA transcription, five RT sessions with glucose will elicit this effect compared to five RT sessions with placebo.
-This time frame has previously shown to be sufficient to measure changes in markers of ribosome biogenesis [@hammarstrom20; @hammarstrom21; @figueiredo16].
+This time frame has previously shown to be sufficient to measure changes in markers of ribosome biogenesis [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022; @figueiredo_impact_2016].
 <!--double-check these references--> As no such effect was observable in the present study, this hypothesis was rejected.
-Further, glucose ingestion affects cellular energy levels; thus another theory is that high glucose treatment would stimulate energy-sensitive pathways such as PIH1, mTORC1, ERK1/2, AMPK and SIRT1 in a dose-response relationship [@kim13; @mariappan11; @tanaka15; @zhai12].
+Further, glucose ingestion affects cellular energy levels; thus another theory is that high glucose treatment would stimulate energy-sensitive pathways such as PIH1, mTORC1, ERK1/2, AMPK and SIRT1 in a dose-response relationship [@kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012].
 <!-- check if this is right --> Mammalian target of rapamycin complex 1, along with ERK1/2 and PIH1 upregulates PIC assembly during high-energy conditions, while AMPK and possibly SIRT1 represses assembly of the PIC during low-energy conditions, as observed in rodents [SOURCE], yeast [SOURCE], and human non-muscle cells [SOURCE] *in vitro*.
-Specifically, high glucose treatment augmented rDNA transcription by increased UBF and SL-1 activity, while low glucose treatment, glucose starvation, or nutrient depletion stimulated AMPK-dependent TIF-1A and chromatin regulation suppression [@kos17: @mariappan11; @zhai12].
-<!--check these references--> In addition to UBF and SL-1, assembly of the PIC requires TIF-1A and Pol I recruitment, induced by amino acid intake in an mTOR-S6K-dependent manner and possibly by direct c-Myc activation [@kusnadi15].
+Specifically, high glucose treatment augmented rDNA transcription by increased UBF and SL-1 activity, while low glucose treatment, glucose starvation, or nutrient depletion stimulated AMPK-dependent TIF-1A and chromatin regulation suppression [@kos-braun_tor1_2017: @mariappan_ribosomal_2011; @zhai_human_2012].
+<!--check these references--> In addition to UBF and SL-1, assembly of the PIC requires TIF-1A and Pol I recruitment, induced by amino acid intake in an mTOR-S6K-dependent manner and possibly by direct c-Myc activation [@kusnadi_regulation_2015].
 As such, increasing energy supply by RT with glucose vs.
@@
 This was not evident with total RNA and c-Myc or rpS6, which we observed no linear relationship between.
-The upstream binding factor (UBF) has indeed previously been described as a master regulator of rDNA transcription, based on findings from cell cultures [@figueiredo19; @kusnadi15; russel&zomerdijk2005].
-However, it has only recently been found in a linear relationship with total RNA content in human skeletal muscle cells after a short period of RT [@hammarstrom21], which the observation of the present study supports.
-As such, UBF also seem to respond to mechanical loading in human muscle cells in line with responses seen in cell cultures and synergist ablation models [@mariappan11; @vonwalden18], in addition to the previously reported response to hormonal, nutritional, and cellular energy signals [@figueiredo19; @kusnadi15; russel&zomerdijk2005].
+The upstream binding factor (UBF) has indeed previously been described as a master regulator of rDNA transcription, based on findings from cell cultures [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @russell_rna-polymerase-i-directed_2005].
+However, it has only recently been found in a linear relationship with total RNA content in human skeletal muscle cells after a short period of RT [@hammarstrom_ribosome_2022], which the observation of the present study supports.
+As such, UBF also seem to respond to mechanical loading in human muscle cells in line with responses seen in cell cultures and synergist ablation models [@mariappan_ribosomal_2011; @vonwalden18], in addition to the previously reported response to hormonal, nutritional, and cellular energy signals [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @russell_rna-polymerase-i-directed_2005].
 While c-Myc and rpS6 did not increase linearly with accumulation of total RNA, the cellular content of c-Myc and rpS6 increased from baseline to post with both treatments.
-The transcription initiation factor c-Myc has been implicated as important to ribosome biogenesis [@vonwalden19; @figueiredo19], though its role is likely as a general transcription factor rather than specific, compared to UBF [SOURCES?].
-Further, the ribosomal protein S6 (rpS6) has previously been shown to correlate strongly with rRNA content after synergist ablation [@nakada16; @chaillou], <!--check Nakada 16 for these refs--> and is suggested as a reliable measurement of ribosome biogenesis [@nakada16].
+The transcription initiation factor c-Myc has been implicated as important to ribosome biogenesis [@von_walden_ribosome_2019; @figueiredo_regulation_2019], though its role is likely as a general transcription factor rather than specific, compared to UBF [SOURCES?].
+Further, the ribosomal protein S6 (rpS6) has previously been shown to correlate strongly with rRNA content after synergist ablation [@nakada_correlation_2016; @chaillou_hypoxia_2012], <!--check Nakada 16 for these refs--> and is suggested as a reliable measurement of ribosome biogenesis [@nakada_correlation_2016].
 Together, this may explain why the between treatment difference in c-Myc content did not induce differences in total RNA or rRNA content, as the specific transcription factor UBF and the rpS6 increased equally with both treatments.
@@
 However, this was not evident, as the mean response in strength as an index of peak torque at all speeds decreased with both treatments from baseline to post five and six sessions.
-Notably, RT with glucose reduced strength significantly less than RT with placebo post five sessions, which may indeed have been an effect of increased energy availability, seeing as glucose is the preferred energy source during strenuous activity [@mul15].
+Notably, RT with glucose reduced strength significantly less than RT with placebo post five sessions, which may indeed have been an effect of increased energy availability, seeing as glucose is the preferred energy source during strenuous activity [@mul_exercise_2015].
 <!--check this ref--> Arguably, this might imply that RT with placebo experienced a higher stress response to the RT program, thus reduced muscle performance in strength tests.

### 2023-03-15 ribose-paper.RMD
updated discussion

@@
 
-Responses to systematic resistance training (RT) vary widely between individuals, with as many as 10-15% showing impaired muscle growth in response to standardized training interventions (ALVAREZ18;MANN14;THALACKER13).
+Responses to systematic resistance training (RT) vary widely between individuals, with as many as 10-15% showing impaired muscle growth in response to standardized training interventions [@ALVAREZ18; @MANN14; @THALACKER13].
 Such inability to benefit from RT may be explained by genetic and/or epigenetic disposition, affecting the internal physiological milieu in terms of ability to mount an anabolic response to RT [@figueiredo_genetic_2021;@thalacker-mercer_cluster_2013].<!-- Check these refs, do they correspond to genet epigen and internal mileu?
@@
 To our knowledge, no previous study has investigated the effects of glucose compared to placebo ingestion with RT on ribosome biogenesis in human skeletal muscle, making our hypotheses exploratory and comparisons not straight forwards.
-Mariappan and colleagues (2011) investigated the effect of high glucose treatment on ribosome biogenesis in glomerular epithelial cells (GECs) from diabetic (type 1 and 2) mice, observing a UBF-dependent augmentation of ribosome biogenesis.
+Mariappan and colleagues [@mariappan_ribosomal_2011] investigated the effect of high glucose treatment on ribosome biogenesis in glomerular epithelial cells (GECs) from diabetic (type 1 and 2) mice, observing a UBF-dependent augmentation of ribosome biogenesis.
 In a study from 2015, Tanaka et al.
-(2015) observed that mild glucose starvation reduced rRNA transcription in human breast cancer cells.
+[@tanaka_mild_2015] observed that mild glucose starvation reduced rRNA transcription in human breast cancer cells.
 These findings do imply an important role for glucose in the initiation of rDNA transcription, the rate-limiting step of ribosome biogenesis [@mariappan_ribosomal_2011; @tanaka_mild_2015].
@@
 Nevertheless, we did observe robust increases in both total RNA and 47S pre-rRNA and all four mature rRNAs (18S, 28S, 5.8S, 5S) after five RT sessions with both treatments.
-This means that the time frame and training volume chosen for this design was adequate to elicit changes in rDNA transcription and subsequently ribosome biogenesis by proxy.
+This means that the time frame and training volume chosen for this design was adequate to elicit changes in rDNA transcription and subsequently ribosome biogenesis by proxy, as previously reported [@figueiredo_impact_2016; @hammarstrom_benefits_2020].
 Further, the training, testing and biological sampling were conducted under standardized conditions, and there were no differences between pairwise consecutive days in macronutrient ingestion and training volume.
-As such, the day-to-day and test-to-test variation should have been minimal, especially when eliminating between-subjects biological variation with a within-subjects design as well (SOURCE).
-Importantly, both Mariappan (11) and Tanaka (15) compared high glucose treatment to either low glucose treatment or glucose starvation.
+As such, the day-to-day and test-to-test variation should have been minimal, especially when eliminating between-subjects biological variation with a within-subjects design as well <!--FIND REF-->.
+Importantly, both Mariappan [@mariappan_ribosomal_2011] and Tanaka [@tanaka_mild_2015] compared high glucose treatment to either low glucose treatment or glucose starvation.
 In the present study however, we compared RT with either protein and glucose or protein and placebo.
@@
 Further, glucose ingestion affects cellular energy levels; thus another theory is that high glucose treatment would stimulate energy-sensitive pathways such as PIH1, mTORC1, ERK1/2, AMPK and SIRT1 in a dose-response relationship [@kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012].
-<!-- check if this is right --> Mammalian target of rapamycin complex 1, along with ERK1/2 and PIH1 upregulates PIC assembly during high-energy conditions, while AMPK and possibly SIRT1 represses assembly of the PIC during low-energy conditions, as observed in rodents [SOURCE], yeast [SOURCE], and human non-muscle cells [SOURCE] *in vitro*.
-Specifically, high glucose treatment augmented rDNA transcription by increased UBF and SL-1 activity, while low glucose treatment, glucose starvation, or nutrient depletion stimulated AMPK-dependent TIF-1A and chromatin regulation suppression [@kos-braun_tor1_2017: @mariappan_ribosomal_2011; @zhai_human_2012].
+<!-- check if this is right --> Mammalian target of rapamycin complex 1, along with ERK1/2 and PIH1 upregulates PIC assembly during high-energy conditions, while AMPK and possibly SIRT1 represses assembly of the PIC during low-energy conditions, as observed in rodents <!-- REF -->, yeast <!-- REF -->, and human non-muscle cells *in vitro*.
+<!-- REF --> Specifically, high glucose treatment augmented rDNA transcription by increased UBF and SL-1 activity, while low glucose treatment, glucose starvation, or nutrient depletion stimulated AMPK-dependent TIF-1A and chromatin regulation suppression [@kos-braun_tor1_2017: @mariappan_ribosomal_2011; @zhai_human_2012].
 <!--check these references--> In addition to UBF and SL-1, assembly of the PIC requires TIF-1A and Pol I recruitment, induced by amino acid intake in an mTOR-S6K-dependent manner and possibly by direct c-Myc activation [@kusnadi_regulation_2015].
@@
 However, it has only recently been found in a linear relationship with total RNA content in human skeletal muscle cells after a short period of RT [@hammarstrom_ribosome_2022], which the observation of the present study supports.
-As such, UBF also seem to respond to mechanical loading in human muscle cells in line with responses seen in cell cultures and synergist ablation models [@mariappan_ribosomal_2011; @vonwalden18], in addition to the previously reported response to hormonal, nutritional, and cellular energy signals [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @russell_rna-polymerase-i-directed_2005].
+As such, UBF also seem to respond to mechanical loading in human muscle cells in line with responses seen in cell cultures and synergist ablation models [@mariappan_ribosomal_2011; @von_walden_mechanical_2012], in addition to the previously reported response to hormonal, nutritional, and cellular energy signals [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @russell_rna-polymerase-i-directed_2005].
 While c-Myc and rpS6 did not increase linearly with accumulation of total RNA, the cellular content of c-Myc and rpS6 increased from baseline to post with both treatments.
-The transcription initiation factor c-Myc has been implicated as important to ribosome biogenesis [@von_walden_ribosome_2019; @figueiredo_regulation_2019], though its role is likely as a general transcription factor rather than specific, compared to UBF [SOURCES?].
-Further, the ribosomal protein S6 (rpS6) has previously been shown to correlate strongly with rRNA content after synergist ablation [@nakada_correlation_2016; @chaillou_hypoxia_2012], <!--check Nakada 16 for these refs--> and is suggested as a reliable measurement of ribosome biogenesis [@nakada_correlation_2016].
+The transcription initiation factor c-Myc has been implicated as important to ribosome biogenesis [@von_walden_ribosome_2019; @figueiredo_regulation_2019], though its role is likely as a general transcription factor rather than specific, compared to UBF<!-- REF -->.
+Further, the ribosomal protein S6 (rpS6) has previously been shown to correlate strongly with rRNA content after synergist ablation [@nakada_correlation_2016; @chaillou_hypoxia_2012], and is suggested as a reliable measurement of ribosome biogenesis [@nakada_correlation_2016].
 Together, this may explain why the between treatment difference in c-Myc content did not induce differences in total RNA or rRNA content, as the specific transcription factor UBF and the rpS6 increased equally with both treatments.
@@
 <!--check this ref--> Arguably, this might imply that RT with placebo experienced a higher stress response to the RT program, thus reduced muscle performance in strength tests.
-This could also explain the differences between treatments in content of the stress responsive c-Myc [SOURCES], meaning higher stress exerted during RT with placebo compared to RT with glucose inducing higher accumulation of c-Myc.
+This could also explain the differences between treatments in content of the stress responsive c-Myc <!-- REF -->, meaning higher stress exerted during RT with placebo compared to RT with glucose inducing higher accumulation of c-Myc.
 If this was the case, it does seem inconclusive that strength did not differ between treatments at earlier or later time points.
@@
 Another potential explanation to this observation might be ….
-(bifasisk restitusjon?).
+(bifasisk restitusjon?)<!-- investigate this -->.
 Lastly, these inconclusive observations may also reflect our limited sample size, despite using a within-subjects design.
@@
 Neither does it acutely enhance muscular performance or recovery withing 23hrs, measured as unilateral isometric and isokinetic knee extension peak torque.
-The observations in total RNA, rRNA and rDNA transcription initiation associated proteins confirm the key role of increased UBF content in ribosome biogenesis in moderately trained human skeletal muscle following resistance training, previously observed in RT-untrained participants, rodents or *in vitro* [SOURCES?].
+The observations in total RNA, rRNA and rDNA transcription initiation associated proteins confirm the key role of increased UBF content in ribosome biogenesis in moderately trained human skeletal muscle following resistance training, previously observed in RT-untrained participants, rodents or *in vitro* <!-- REF -->.
 If future investigations are to be made on this topic, a greater sample size coupled with a negative control group and analyses of the intramuscular glycogen storage may provide higher resolution results.

### 2023-03-16 ribose-paper.RMD
small updates to discussion

@@
 The main findings of the present study were that heavy resistance training with glucose did not affect markers of ribosome biogenesis, measured as total RNA, rRNA and rDNA transcription initiation associated protein, compared to RT with placebo.
-Similarly, RT with glucose in general did not affect markers of skeletal muscle functionality such as strength and -recovery, or total training session volume.
+Similarly, RT with glucose in general did not affect markers of skeletal muscle functionality such as strength and recovery, or total training session volume.
 At one time point, RT with glucose led to less reduction in strength compared to placebo, and levels of plasma glucose and c-peptide were significantly higher before, during and after the RT sessions with glucose.
@@
 This would suggest the effects of high vs.
-low glucose/glucose starvation treatment on ribosome biogenesis or rDNA transcription initiation observed in previous in vitro studies [@mariappan_ribosomal_2011; @tanaka_mild_2015] are not translatable to acute effects in human skeletal muscle in vivo.
+low glucose/glucose starvation treatment on ribosome biogenesis or rDNA transcription initiation observed in previous *in vitro* studies [@mariappan_ribosomal_2011; @tanaka_mild_2015] are not translatable to acute effects in human skeletal muscle *in vivo*.
 Notably, the study was specifically design to investigate differences in pre-post changes within-participants, and thus also quite limited to this perspective.
@@
 These findings do imply an important role for glucose in the initiation of rDNA transcription, the rate-limiting step of ribosome biogenesis [@mariappan_ribosomal_2011; @tanaka_mild_2015].
-In the present study, we did not find such indications of glucose having an effect as these previous studies.
-Notably, kidney cells from mice and human breast cancer cells are likely to have a higher physiological rate compared to human skeletal muscle cells, thus, our two-week intervention may not be directly comparable to these findings, as was expected.
-Nevertheless, we did observe robust increases in both total RNA and 47S pre-rRNA and all four mature rRNAs (18S, 28S, 5.8S, 5S) after five RT sessions with both treatments.
-This means that the time frame and training volume chosen for this design was adequate to elicit changes in rDNA transcription and subsequently ribosome biogenesis by proxy, as previously reported [@figueiredo_impact_2016; @hammarstrom_benefits_2020].
-Further, the training, testing and biological sampling were conducted under standardized conditions, and there were no differences between pairwise consecutive days in macronutrient ingestion and training volume.
-As such, the day-to-day and test-to-test variation should have been minimal, especially when eliminating between-subjects biological variation with a within-subjects design as well <!--FIND REF-->.
+In the present study however, we did not find observations indicative of glucose having an effect acute effect on rDNA transcription.
+Thus, glucose more likely affects rDNA transcription through energy availability [@hoppe_amp-activated_2009; @kusnadi_regulation_2015; @mariappan_ribosomal_2011; @tanaka_mild_2015] Notably, kidney cells from mice and human breast cancer cells are likely to have a different physiological rate compared to human skeletal muscle cells, thus, our two-week intervention are likely not directly comparable to these findings, as was expected.
+
+Nevertheless, the increases observed in total RNA and rRNA after only five sessions of RT in moderately RT-trained participants do concur with recent findings after acute [@figueiredo_impact_2016] and short-term RT [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022] in previously RT-untrained participants.
+Therefore, the time frame and training volume chosen for this design was indeed adequate to stimulate rDNA transcription and subsequently ribosome biogenesis by proxy.
+Further, the within-participant design along with the standardized conditions for testing and biological sampling decreased are day-to-day/test-to-test variation, effectively increasing the reliability of measurements [@HALPERIN; @THIESE; @MACINNIS].
 Importantly, both Mariappan [@mariappan_ribosomal_2011] and Tanaka [@tanaka_mild_2015] compared high glucose treatment to either low glucose treatment or glucose starvation.
@@
 The participants fasted from 2200hrs the preceding day until the morning of each training day, receiving only the prescribed treatment until 2.5 hrs after RT.
-Thus, except from approximately 0700hrs-1230hrs, the participants ingested an equal amount of carbohydrates on pairwise days.
+Thus, except from approximately 0700hrs-1230hrs, the participants ingested an equal amount of carbohydrates, protein, fat and calories on pairwise days.
 As such, there were no differences in baseline levels of plasma glucose prior to the daily intervention, as by design, since we were interested in investigating if there were direct effects of glucose as a signalling molecule, based on the observations of Tanaka et al.
@@
 [@mariappan_ribosomal_2011].
-Further, the significant difference in levels of plasma glucose and c-peptide measured before, during and after session six, confirms that the treatment during of RT with glucose indeed was sufficient.
-Despite of this, we found no difference in markers of ribosome biogenesis between RT with glucose and RT with placebo, this does suggest that there are no additional effects of exercising with glucose, when total macronutrient intake is standardized.
+Further, the significant difference in levels of plasma glucose and c-peptide measured before, during and after session six, confirms that the treatment during RT with glucose indeed was sufficient.
+Despite of this, we found no difference in markers of ribosome biogenesis between RT with glucose and RT with placebo, this does suggest that there are no additional acute effects of exercising with glucose in this regard, when total macro nutrient intake is standardized.
 
-Prior to the present study, it was theorized that glucose may convey an anabolic signal, either directly or through insulin, and thus operate as a signal mediator affecting upstream pathways of ribosome biogenesis [@kusnadi_regulation_2015; @sebata_control_2018].
+Prior to the present study, we was theorized that glucose may convey an anabolic signal, either directly or through insulin, and thus operate as a signal mediator affecting upstream pathways of ribosome biogenesis based on previous observations [@kusnadi_regulation_2015; @sebata_control_2018].
 This theoretical anabolic signal could possibly stimulate PIC assembly and Pol I recruitment to the rDNA promoter, initiating rDNA transcription and subsequent rRNA maturation, and lastly increased ribosome content of the skeletal muscle cell [@figueiredo_regulation_2019; @sebata_control_2018].

### 2023-03-16 ribose-paper.RMD
discussion updates

@@
 author: "Kristian Lian, Daniel Hammarström, Stian Ellefsen, Håvard Hamarsland"
-bibliography: resources/ribose.bib
+bibliography: resources/ribose-kl.bib
 csl: resources/jap.csl
@@
 
-Responses to systematic resistance training (RT) vary widely between individuals, with as many as 10-15% showing impaired muscle growth in response to standardized training interventions [@ALVAREZ18; @MANN14; @THALACKER13].
+Responses to systematic resistance training (RT) vary widely between individuals, with as many as 10-15% showing impaired muscle growth in response to standardized training interventions [@alvarez_interindividual_2018; @mann_high_2014; @thalacker-mercer_cluster_2013].
 Such inability to benefit from RT may be explained by genetic and/or epigenetic disposition, affecting the internal physiological milieu in terms of ability to mount an anabolic response to RT [@figueiredo_genetic_2021;@thalacker-mercer_cluster_2013].<!-- Check these refs, do they correspond to genet epigen and internal mileu?
@@
 Therefore, the time frame and training volume chosen for this design was indeed adequate to stimulate rDNA transcription and subsequently ribosome biogenesis by proxy.
-Further, the within-participant design along with the standardized conditions for testing and biological sampling decreased are day-to-day/test-to-test variation, effectively increasing the reliability of measurements [@HALPERIN; @THIESE; @MACINNIS].
+Further, the within-participant design along with the standardized conditions for testing and biological sampling decreased are day-to-day/test-to-test variation, effectively increasing the reliability of measurements [@halperin_threats_2015; @macinnis_investigating_2017].
 Importantly, both Mariappan [@mariappan_ribosomal_2011] and Tanaka [@tanaka_mild_2015] compared high glucose treatment to either low glucose treatment or glucose starvation.
@@
 This time frame has previously shown to be sufficient to measure changes in markers of ribosome biogenesis [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022; @figueiredo_impact_2016].
-<!--double-check these references--> As no such effect was observable in the present study, this hypothesis was rejected.
-Further, glucose ingestion affects cellular energy levels; thus another theory is that high glucose treatment would stimulate energy-sensitive pathways such as PIH1, mTORC1, ERK1/2, AMPK and SIRT1 in a dose-response relationship [@kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012].
-<!-- check if this is right --> Mammalian target of rapamycin complex 1, along with ERK1/2 and PIH1 upregulates PIC assembly during high-energy conditions, while AMPK and possibly SIRT1 represses assembly of the PIC during low-energy conditions, as observed in rodents <!-- REF -->, yeast <!-- REF -->, and human non-muscle cells *in vitro*.
-<!-- REF --> Specifically, high glucose treatment augmented rDNA transcription by increased UBF and SL-1 activity, while low glucose treatment, glucose starvation, or nutrient depletion stimulated AMPK-dependent TIF-1A and chromatin regulation suppression [@kos-braun_tor1_2017: @mariappan_ribosomal_2011; @zhai_human_2012].
-<!--check these references--> In addition to UBF and SL-1, assembly of the PIC requires TIF-1A and Pol I recruitment, induced by amino acid intake in an mTOR-S6K-dependent manner and possibly by direct c-Myc activation [@kusnadi_regulation_2015].
+<!--double-check these references.--> As no such effect was observable in the present study, this hypothesis was rejected.
+Further, glucose ingestion affects cellular energy levels; thus another theory is that high glucose treatment would stimulate energy-sensitive pathways such as PIH1, mTORC1, ERK1/2, AMPK and SIRT1 [@kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012].
+Mammalian target of rapamycin complex 1, along with ERK1/2 and PIH1 upregulates PIC assembly during high-energy conditions, while AMPK and possibly SIRT1 represses assembly of the PIC during low-energy conditions, as observed in rodents, yeast, or human non-muscle cells *in vitro* [@kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012].
+Specifically, high glucose treatment augmented rDNA transcription by increased UBF and SL-1 activity, while low glucose treatment, glucose starvation, or nutrient depletion stimulated AMPK-dependent TIF-1A and chromatin regulation suppression [@kos-braun_tor1_2017: @mariappan_ribosomal_2011; @zhai_human_2012].
+<!--check these references.--> In addition to UBF and SL-1, assembly of the PIC requires TIF-1A and Pol I recruitment, induced by amino acid intake in an mTOR-S6K-dependent manner and possibly by direct c-Myc activation [@kusnadi_regulation_2015].
 As such, increasing energy supply by RT with glucose vs.
-RT with placebo and nutrient availability should in line with these observations potently stimulate PIC assembly and rDNA transcription.
+RT with placebo and nutrient availability should in line with these observations stimulate PIC assembly and rDNA transcription.
 Though we did not investigate these pathways directly, the downstream targets UBF and rpS6 did not differ in mean change between treatments.
-As the participants in the present study fasted until the daily training was completed, and matched their daily energy intake on pairwise days, this suggests that as long as the daily energy requirements are met, ingesting glucose with RT does not stimulate further gains in anabolic responses.
+As the participants in the present study fasted until the daily training was completed, and matched their daily energy intake on pairwise days, the result of the present study suggest that as long as the daily energy requirements are equal, ingesting glucose with RT does not stimulate further gains in anabolic responses.
 Notably, the daily RT protocol was quite short (~30min), which may have reduced the effectiveness of the energy addition through glucose.
-In this regard, it would have been interesting to measure the intramuscular glycogen storages, to see how the RT sessions impacted intracellular energy levels.
+In this regard, it would have been interesting to measure the intramuscular glycogen stores, to see how the RT sessions impacted intracellular energy levels.
 Unfortunately, such analyses were not conducted in the present study.
 Regrettably, this is a weakness of the design, as we have not measured directly what happened to the excess plasma glucose.
-Nevertheless, the absent effect of increasing energy availability during RT suggest that, there is no dose-response relationship between acutely increasing myocellular energy levels before and after RT and ribosome biogenesis.
+Regardless, the absent effect of increasing energy availability during RT suggest that, there is no dose-response relationship between acutely increasing myocellular energy levels before and after RT and ribosome biogenesis.
 
-Interestingly, we did observe a linear relationship between total RNA and UBF, suggesting ribosome biogenesis to be dependent on UBF activation, in line with observations in Mariappan et al.
+Interestingly, we did observe a linear relationship between total RNA and UBF, supporting the observations from diabetic mice in Mariappan et al.
 (2011).
@@
 The upstream binding factor (UBF) has indeed previously been described as a master regulator of rDNA transcription, based on findings from cell cultures [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @russell_rna-polymerase-i-directed_2005].
-However, it has only recently been found in a linear relationship with total RNA content in human skeletal muscle cells after a short period of RT [@hammarstrom_ribosome_2022], which the observation of the present study supports.
+However, it has only recently been found in a linear relationship with total RNA content in human skeletal muscle cells after a short period of RT [@hammarstrom_ribosome_2022].
 As such, UBF also seem to respond to mechanical loading in human muscle cells in line with responses seen in cell cultures and synergist ablation models [@mariappan_ribosomal_2011; @von_walden_mechanical_2012], in addition to the previously reported response to hormonal, nutritional, and cellular energy signals [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @russell_rna-polymerase-i-directed_2005].
-While c-Myc and rpS6 did not increase linearly with accumulation of total RNA, the cellular content of c-Myc and rpS6 increased from baseline to post with both treatments.
-The transcription initiation factor c-Myc has been implicated as important to ribosome biogenesis [@von_walden_ribosome_2019; @figueiredo_regulation_2019], though its role is likely as a general transcription factor rather than specific, compared to UBF<!-- REF -->.
+Regarding c-Myc, it has previously been found to regulate gene chromatin remodelling and Pol I availability in non-muscle cells [@poortinga], and c-Myc overexpression stimulated skeletal muscle ribosome biogenesis in mice [@mori], similar to the rapamycin-independent upregulation after RT in rats [@west].
+However, the findings of the present study do not entirely align with these previous observations, as there was no difference in change of total RNA and rRNA between treatments, despite a significantly greater c-Myc content in placebo compared to glucose post intervention.
+Along with the observed non-linear relationship between c-Myc and total RNA, while important as a general transcription factor, c-Myc does not seem to have the same specificity to stimulate rDNA transcription as UBF, in human skeletal muscle cells.
 Further, the ribosomal protein S6 (rpS6) has previously been shown to correlate strongly with rRNA content after synergist ablation [@nakada_correlation_2016; @chaillou_hypoxia_2012], and is suggested as a reliable measurement of ribosome biogenesis [@nakada_correlation_2016].
 Together, this may explain why the between treatment difference in c-Myc content did not induce differences in total RNA or rRNA content, as the specific transcription factor UBF and the rpS6 increased equally with both treatments.
-Notably, the difference between c-Myc in glucose and placebo may have caused responses in pathways not measured in the present study.
+Notably, the difference between c-Myc in glucose and placebo may have stimulated pathways not measured in the present study.
 

### 2023-03-16 ribose-paper.RMD
discussion updates

@@
 As such, UBF also seem to respond to mechanical loading in human muscle cells in line with responses seen in cell cultures and synergist ablation models [@mariappan_ribosomal_2011; @von_walden_mechanical_2012], in addition to the previously reported response to hormonal, nutritional, and cellular energy signals [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @russell_rna-polymerase-i-directed_2005].
-Regarding c-Myc, it has previously been found to regulate gene chromatin remodelling and Pol I availability in non-muscle cells [@poortinga], and c-Myc overexpression stimulated skeletal muscle ribosome biogenesis in mice [@mori], similar to the rapamycin-independent upregulation after RT in rats [@west].
+Regarding c-Myc, it has previously been found to regulate gene chromatin remodelling and Pol I availability in non-muscle cells [@poortinga_c-myc_2011], and c-Myc overexpression stimulated skeletal muscle ribosome biogenesis in mice [@mori_c-myc_2021], similar to the rapamycin-independent upregulation after RT in rats [@west_acute_2016].
 However, the findings of the present study do not entirely align with these previous observations, as there was no difference in change of total RNA and rRNA between treatments, despite a significantly greater c-Myc content in placebo compared to glucose post intervention.

### 2023-03-17 ext_method.RMD
small updates

--- 2023-02-28 ext_method.RMD
@@
 
-- UBF/RNA correlation: Make a new model to see if data point affects the current model
+- Antiboy table?
+- Power calculation + rational?
+
+## Wet lab analyses
+
+We gathered muscle tissue from both legs at pre and post intervention, taking two samples from the same incision.
+Each piece of muscle tissue was split in two, i.e.
+we had two samples for RNA analyses and two samples for protein analyses.
+We analyzed the samples in duplicates, i.e.
+we had two samples per time point per leg, a total of eight samples per analysis.
+<!-- check if there is more tissue left --> These samples were then randomly labelled with participant number and a number between 1-8, to blind the investigators conducting the laboratory analyses.
+
+### Total RNA and rRNA
+
+###### **Tab 1:** Primer sequence and performance. Average cycle thresholds (Ct) and primer efficiencies were calculated from all qPCR reactions.
+
+![](figures/archive/Primer.sequences.PNG)
+
+## Statistical analyses
+
+We chose to log-transformed all data, as visual inspection showed clear signs of heteroscedasticity in some of the data sets.
+Thus, data was log-transformed prior to modelling with the function log(), and then reverse-transformed for figure illustration via the function exp().
+Exceptions to this was the data used in figure 1 (B, C and D).
+In figure 1B and -C, the illustrated data are absolute changes in plasma glucose and c-peptide, as this demonstrated quite well the change that occured, and arguably, is quite more understandable compared to arbitrary units, such as the protein signal data.
+For figure 1D, we had already made in index of strength combining isometric and isokinetic peak torque, meaning the data was already put on a different scale than the original absolute change in peak torque.
+
+Further, we mean-centered the baseline values in each model to account for regression to the mean.
+This was simply done by subtracting the mean baseline value from each participants baseline value, to control for the possibility of extremely high or low baseline values to violate the assumption of equal variance.
+Estimated marginal means were calculated from the linear models, enabling acquisition of least-square means, showing the means for all involved groups adjusted to means of other variables in the model.
+This is a way to statistically model the means of the data, rather than describing them as they appear in the data.
+
+To enable cross nano Drop, qPCR plate and gel comparisons from the wet lab analyses, we normalized RNA and protein data.
+Total RNA was normalized to wet muscle weight, rRNA was normalized to wet muscle weight and the external reference gene Lambda, while protein data was normalized both to the total protein lane factor and to pools of all protein samples.
+
+RNA normalization: 
+ 1. weight = mean(weight) 
+ 2. RNA = mean(RNA)
+ 3. weight.mc = weight/mean(weight)
+ 4. RNA.weight = RNA / weight
+ - weight = biopsy weight
+
+Weight and RNA was summarised, then we devided RNA by mean centered weight. Thus, we could compare total RNA results analyzed on different nanoDrop plates.
+
+rRNA normalization: 
+ 1. Lambda expression was first calculated and stored in the data.frame "nf"
+ 2. Calculation of target expression
+ 3. Joined the lambda and target expression data.frames
+ 4. nf.w = expr / weight
+ 5. nf.w = nf.w / max(nf.w)
+ 6. nf.expr = log(expr / nf.w), nf.w = scale(nf.w)
+ - expr = target expression 
+ - weight = biopsy weight
+
+Protein normalization:
+ 1. tpl = tp - bg
+ 2. tpl = tpl/max(tpl)
+ 3. signal = signal / max(signal)
+ 4. norm.sign = signal / tpl
+ 5. pool = mean(pool)
+ 6. norm.sign = norm.sign / pool
+ - tpl = total protein per lane
+ - tp = total protein (mean gray pixels in lanes)
+ - bg = background (mean gray pixels between lanes)
+ - signal = pixels per band
+
+To create the tpl normalization factor, we measured the mean gray both within (tp) and between (bg) the lanes of each total protein stain blot with the

### 2023-03-17 ribose-paper.RMD
small updates

@@
 The tissue was homogenized with a plastic rod in 80mg protein extraction powder (Invent Biotechnology) and 100ul iced cell lysis buffer (Denaturing Buffer, Invent Biotechnology), and centrifuged at top speed for 1 min.
-The supernatant was aliquoted into 2x25ul tubes to run samples in duplicates, and total protein concentrations were determined in a 1:10 dilution (Pierce Detergent Compatible Bradford Assay Reagent, Thermo Fisher Scientific, Oslo, Norway).
+The supernatant was devided into 2x25ul aliquots to run samples in duplicates, and total protein concentrations were determined in a 1:10 dilution (Pierce Detergent Compatible Bradford Assay Reagent, Thermo Fisher Scientific, Oslo, Norway).
 The protein samples were diluted to 20.25 µg µl−1 total protein with lysis buffer and 4X Laemmli sample buffer (Bio-Rad Laboratories AB, Oslo, Norway) containing 2-mercaptoethanol.
-Twenty microlitres from each protein sample were pooled as a normalisation factor for gel separation.
+Twenty microlitres from each protein sample were pooled as a normalization factor for gel separation.
 All protein samples were incubated at 95°C and stored at -20°C until further analysis.
@@
 Both gel electrophoresis and protein transfer were performed at 4°C.
-Following the wet transfer, membranes were stained with a MemCode reversible total protein stain (MANUFACTURER) and then blocked for 1hr at room temp with a blocking buffer of Tris-buffered saline (TBS; 20 mM Tris, 150 mM NaCl) with 5% non-fat dry milk and 0.1% Tween-20.
-Primary and secondary antibodies were purchased from Santa Cruz Biotechnology (Texas, USA) UBF:(UBF-9, sc-13125), rpS6:(Ribosomal protein S6 C-8, sc-74459), and Thermo Fisher Scientific () c-Myc:(c-Myc 9E10), goat anti-mouse for c-Myc (goat anti-mouse IgG1 (y1) horseradish peroxidase conjugate), and anti-mouse (anti-mouse IgG1 horseradish peroxidase conjugate).
+Following the wet transfer, membranes were stained with a MemCode reversible total protein stain (Thermo Fisher Scientific) and then blocked for 1hr at room temp with a blocking buffer of Tris-buffered saline (TBS; 20 mM Tris, 150 mM NaCl) with 5% non-fat dry milk and 0.1% Tween-20.
+Primary and secondary antibodies were purchased from Santa Cruz Biotechnology (Texas, USA) UBF:(UBF-9, sc-13125), rpS6:(Ribosomal protein S6 C-8, sc-74459), and Thermo Fisher Scientific (Oslo, Norway) c-Myc:(c-Myc 9E10), goat anti-mouse for c-Myc (goat anti-mouse IgG1 (y1) horseradish peroxidase conjugate), and anti-mouse (anti-mouse IgG1 horseradish peroxidase conjugate).
 Antibodies were diluted in a 5% non-fat dry milk blocking buffer (TBS, 5% non-fat dry milk, 0.1% Tween-20).
@@
 Primary antibody incubation was performed at 4°C on a tipping board.
-Chemiluminescence signals were quantified using Image Studio Lite (LI-COR Biotechnology, Lincoln, NE, USA), and total protein content was quantified using ImageJ (RUEDENETAL2017), whereas total protein content was defined as mean grey value of the whole well with between-well values subtracted as background.
+Chemiluminescence signals were quantified using Image Studio Lite (LI-COR Biotechnology, Lincoln, NE, USA), and total protein content was quantified using ImageJ [@rueden_imagej2_2017], whereas total protein content was defined as mean grey value of the whole well with between-well values subtracted as background.
 
@@
 A priori power calculations showed that 20 participants would grant a statistical power of 80% (= 0.05), with an expected drop out of 20%.
-This power calculation was based on an assumption that the effects of glucose ingestion on ribosome biogenesis may equate to the effects of increasing RT volume from low to moderate (HAMMARSTROM20).
-<!-- maybe include the power calculation + reason in extended methods?
--->.
-Raw data was imported an analyzed in R via R studios (SOURCE).
+This power calculation was based on an assumption that the effects of glucose ingestion on ribosome biogenesis may equate to the effects of increasing RT volume from low to moderate [@hammarstrom_benefits_2020].
+<!-- maybe include the power calculation in extended methods?-->.
+Raw data was imported an analyzed in R via R studios [@Rcore].
 Legs were categorized as either glucose or placebo in accordance with the randomization.
@@
 Total RNA and qPCR values were analyzed by comparing post to pre log-fold change score differences.
-This was achieved by modelling change as response with baseline and supplement (glucose/placebo) as explanatory factors, and mean-centered baseline values as a covariate in a linear mixed effects model, then calculating the estimated marginal means <!-- add explanation of model in extended methods-->.
-Plasma glucose and c-peptide, training volume, and the strength index were analyzed by multiple time-point log-fold change score comparisons, using the same linear mixed effects model with time added as explanatory factor <!-- add explanation of model in extended methods-->.
-The linear mixed effects model was designed with the lmer function of the lme4 package (BATES14) in R studios, using the lmerTest function to procure p-values.
+This was achieved by modelling change as response with baseline and supplement (glucose/placebo) as explanatory factors, and mean-centered baseline values as a covariate in a linear mixed effects model, then calculating the estimated marginal means.
+Plasma glucose and c-peptide, training volume, and the strength index were analyzed by multiple time-point log-fold change score comparisons, using the same linear mixed effects model with time added as explanatory factor.
+The linear mixed effects model was designed with the lmer function of the lme4 package [@bates_fitting_2014] in R studios, using the lmerTest function to procure p-values.
 The relationship between

### 2023-03-20 ribose-paper.RMD
updated discussion

@@
 mTOR receives input from growth factors, hormones, mechanical loading, and nutrients to balance protein synthesis through multiple mechanisms based on cellular energy levels [@hoppe_amp-activated_2009].
-Importantly, mTOR and its downstream target S6K1 serve as a mediator from insulin signaling in skeletal muscle[@hillier_physiological_2000].
+Importantly, mTOR and its downstream target S6K1 serve as a mediator from insulin signaling in skeletal muscle [@hillier_physiological_2000].
 The mTORC1 pathway contributes to ribosome biogenesis through the regulation of translation of ribosomal proteins and to forming of the preinitiation complex (PIC) that marks the initiation of rRNA transcription [@figueiredo_regulation_2019;@von_walden_ribosome_2019].
@@
 This would suggest the effects of high vs.
-low glucose/glucose starvation treatment on ribosome biogenesis or rDNA transcription initiation observed in previous *in vitro* studies [@mariappan_ribosomal_2011; @tanaka_mild_2015] are not translatable to acute effects in human skeletal muscle *in vivo*.
-Notably, the study was specifically design to investigate differences in pre-post changes within-participants, and thus also quite limited to this perspective.
+low glucose/glucose starvation treatment on rDNA transcription initiation observed in previous *in vitro* studies [@mariappan_ribosomal_2011; @tanaka_mild_2015] are not translatable to acute effects in human skeletal muscle *in vivo*.
+Notably, the present study was specifically designed to investigate differences in pre-post changes within-participants, and thus also quite limited to this perspective.
 Hence, the observed effects of time serve as supplementary to the efficiency of the design, without a negative control group to either confirm or refute them.
 
-To our knowledge, no previous study has investigated the effects of glucose compared to placebo ingestion with RT on ribosome biogenesis in human skeletal muscle, making our hypotheses exploratory and comparisons not straight forwards.
+To our knowledge, no previous study has investigated the effects of glucose compared to placebo ingestion with RT on ribosome biogenesis in human skeletal muscle previously, making our hypotheses exploratory and comparisons few.
 Mariappan and colleagues [@mariappan_ribosomal_2011] investigated the effect of high glucose treatment on ribosome biogenesis in glomerular epithelial cells (GECs) from diabetic (type 1 and 2) mice, observing a UBF-dependent augmentation of ribosome biogenesis.
@@
 These findings do imply an important role for glucose in the initiation of rDNA transcription, the rate-limiting step of ribosome biogenesis [@mariappan_ribosomal_2011; @tanaka_mild_2015].
-In the present study however, we did not find observations indicative of glucose having an effect acute effect on rDNA transcription.
-Thus, glucose more likely affects rDNA transcription through energy availability [@hoppe_amp-activated_2009; @kusnadi_regulation_2015; @mariappan_ribosomal_2011; @tanaka_mild_2015] Notably, kidney cells from mice and human breast cancer cells are likely to have a different physiological rate compared to human skeletal muscle cells, thus, our two-week intervention are likely not directly comparable to these findings, as was expected.
+In addition, Hillier et al.
+found physiological hyperinsulinemia to stimulate p70S6K phosphorylation in human skeletal muscle [@hillier_physiological_2000].
+Further, robust increases in markers of ribosome biogenesis such as 47S pre-rRNA and mature rRNA can be expected after an acute bout of RT [@figueiredo_impact_2016] or a short period of RT [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022].
+Therefore, we hypothesized that if glucose can, directly or indirectly via insulin, stimulate initiation of rDNA transcription, five sessions of RT with glucose compared to placebo will elicit this effect.
+Further, glucose ingestion increases energy levels, thus, we also hypothesized that high glucose treatment would stimulate energy-sensitive pathways such as PIH1, mTORC1, ERK1/2, AMPK and SIRT1 [@kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012].
+Despite previously reported upregulation in PIC assembly due to high-glucose mediated mTORC1, ERK1/2 and PIH1 or low-glucose mediated AMPK and SIRT1 activation  [@hoppe_amp-activated_2009; @kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012], the present study showed no signs of such effects of glucose vs.
+placebo treatment.
+Resistance training, irrespective of treatment, yielded robust accumulation of total RNA and expression of rRNA, in line with previous observations [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022].
+Due to the exploratory nature of our hypotheses this was hardly a very surprising result.
+Indeed, previous studies has used cell cultures from yeast, rodent or human cells [@hoppe_amp-activated_2009; @kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012] <!--arrange refs after yeast, rodent and human --> and are as such not directly comparable to human skeletal muscle cells.
+Further, previous studies investigated high vs.
+low glucose treatment <!-- refs -->, or high glucose vs.
+glucose starvation <!-- refs -->, while the present study aimed to compare high glucose treatment vs.
+placebo, with a matched daily macro nutrient and energy intake.
+And so, the comparison of the present study was high glucose vs.
+normal glucose treatment, to investigate the effect of glucose *per se* and not intracellular energy status.
+Thus, while glucose ingestion presumably is important in supplying energy for growth inducing processes such as ribosome biogenesis <!-- refs -->, there is no apparent effect of ingesting glucose *per se* on markers of ribosome biogenesis.
 
-Nevertheless, the increases observed in total RNA and rRNA after only five sessions of RT in moderately RT-trained participants do concur with recent findings after acute [@figueiredo_impact_2016] and short-term RT [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022] in previously RT-untrained participants.
-Therefore, the time frame and training volume chosen for this design was indeed adequate to stimulate rDNA transcription and subsequently ribosome biogenesis by proxy.
-Further, the within-participant design along with the standardized conditions for testing and biological sampling decreased are day-to-day/test-to-test variation, effectively increasing the reliability of measurements [@halperin_threats_2015; @macinnis_investigating_2017].
-Importantly, both Mariappan [@mariappan_ribosomal_2011] and Tanaka [@tanaka_mild_2015] compared high glucose treatment to either low glucose treatment or glucose starvation.
-In the present study however, we compared RT with either protein and glucose or protein and placebo.
-The participants fasted from 2200hrs the preceding day until the morning of each training day, receiving only the prescribed treatment until 2.5 hrs after RT.
-Thus, except from approximately 0700hrs-1230hrs, the participants ingested an equal amount of carbohydrates, protein, fat and calories on pairwise days.
-As such, there were no differences in baseline levels of plasma glucose prior to the daily intervention, as by design, since we were interested in investigating if there were direct effects of glucose as a signalling molecule, based on the observations of Tanaka et al.
-[@tanaka_mild_2015] and Mariappan et al.
-[@mariappan_ribosomal_2011].
-Further, the significant difference in levels of plasma glucose and c-peptide measured before, during and after session six, confirms that the treatment during RT with glucose indeed was sufficient.
-Despite of this, we found no difference in markers of ribosome biogenesis between RT with glucose and RT with placebo, this does suggest that there are no additional acute effects of exercising with glucose in this regard, when total macro nutrient intake is standardized.
+In the present study, despite not measuring the activity in central pathways mediating anabolic signalling (mTORC1, ERK1/2), analyses of the downstream target upstream binding factor (UBF) and the ribosomal protein S6 (rpS6) further supported observations from the RNA data.
+UBF has previously been described as a master regulator of rDNA transcription *in vitro* [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @russell_rna-polymerase-i-directed_2005], while rpS6 is proposed as a valid and reliable means to measure ribosome biogenesis [@nakada_correlation_2016].
+However, content of c-Myc did differ between RT with glucose and RT with placebo, whereas RT with placebo led to significantly higher post content of c-Myc compared to glucose.
+The general transcription factor c-Myc has previously been described as a potent regulator of ribosome biogenesis independent of mTORC1.
+<!-- west2016,  mori21, vonwalden19 --> The independency on mTORC1 may explain why only c-Myc content differed between treatments, while UBF and rpS6 content changed equally.
+Moreover, the observation that total RNA and rRNA was unaffected by the significant differences in c-Myc content, do strengthen the importance and specificity of UBF content in human skeletal muscle ribosome biogenesis.
+The linear relationship found exclusively between UBF content and total RNA accumulation, and not c-Myc or rpS6,  found in the present study further supports this notion.
+While this is not a novel finding, it has only recently been observed in human skeletal muscle following RT [@hammarstrom_ribosome_2022].
+As such, UBF also seem to respond to mechanical loading in human muscle cells in line with responses seen in cell cultures and synergist ablation models [@mariappan_ribosomal_2011; @von_walden_mechanical_2012], in addition to the previously reported response to hormonal, nutritional, and cellular energy signals [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @russell_rna-polymerase-i-directed_2005].
 
-Prior to the present study, we was theorized that glucose may convey an anabolic signal, either directly or through insulin, and thus operate as a signal mediator affecting upstream pathways of ribosome biogenesis based on previous observations [@kusnadi_regulation_2015; @sebata_control_2018].
-This theoretical anabolic signal could possibly stimulate PIC assembly and Pol I recruitment to the rDNA promoter, initiating rDNA transcription and subsequent rRNA maturation, and lastly increased ribosome content of the skeletal muscle cell [@figueiredo_regulation_2019; @sebata_control_2018].
-Thus, we hypothesized that if glucose directly or indirectly via insulin can stimulate initiation of rDNA transcription, five RT sessions with glucose will elicit this effect compared to five RT sessions with placebo.
-This time frame has previously shown to be sufficient to measure changes in markers of ribosome biogenesis [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022; @figueiredo_impact_2016].
-<!--double-check these references.--> As no such effect was observable in the present study, this hypothesis was rejected.
-Further, glucose ingestion affects cellular energy levels; thus another theory is that high glucose treatment would stimulate energy-sensitive pathways such as PIH1, mTORC1, ERK1/2, AMPK and SIRT1 [@kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012].
-Mammalian target of rapamycin complex 1, along with ERK1/2 and PIH1 upregulates PIC assembly during high-energy conditions, while AMPK and possibly SIRT1 represses assembly of the PIC during low-energy conditions, as observed in rodents, yeast, or human non-muscle cells *in vitro* [@kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012].
-Specifically, high glucose treatment augmented rDNA transcription by increased UBF and SL-1 activity, while low glucose treatment, glucose starvation, or nutrient depletion stimulated AMPK-dependent TIF-1A and chromatin regulation suppression [@kos-braun_tor1_2017: @mariappan_ribosomal_2011; @zhai_human_2012].
-<!--check these references.--> In addition to UBF and SL-1, assembly of the PIC requires TIF-1A and Pol I recruitment, induced by amino acid intake in an mTOR-S6K-dependent manner and possibly by direct c-Myc activation [@kusnadi_regulation_2015].
-As such, increasing energy supply by RT with glucose vs.
-RT with placebo and nutrient availability should in line with these observations stimulate PIC assembly and rDNA transcription.
-Though we did not investigate these pathways directly, the downstream targets UBF and rpS6 did not differ in mean change between treatments.
-As the participants in the present study fasted until the daily training was completed, and matched their daily energy intake on pairwise days, the result of the present study suggest that as long as the daily energy requirements are equal, ingesting glucose with RT does not stimulate further gains in anabolic responses.
-Notably, the daily RT protocol was quite short (~30min), which may have reduced the effectiveness of the energy addition through glucose.
-In this regard, it would have been interesting to measure the intramuscular glycogen stores, to see how the RT sessions impacted intracellular energy levels.
-Unfortunately, such analyses were not conducted in the present study.
-Regrettably, this is a weakness of the design, as we have not measured directly what happened to the excess plasma glucose.
-Regardless, the absent effect of increasing energy availability during RT suggest that, there is no dose-response relationship between acutely increasing myocellular energy levels before and after RT and ribosome biogenesis.
-
-Interestingly, we did observe a linear relationship between total RNA and UBF, supporting the observations from diabetic mice in Mariappan et al.
-(2011).
-This was not evident with total RNA and c-Myc or rpS6, which we observed no linear relationship between.
-The upstream binding factor (UBF) has indeed previously been described as a master regulator of rDNA transcription, based on findings from cell cultures [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @russell_rna-polymerase-i-directed_2005].
-However, it has only recently been found in a linear relationship with total RNA content in human skeletal muscle cells after a short period of RT [@hammarstrom_ribosome_2022].
-As such, UBF also seem to respond to mechanical loading in human muscle cells in line with responses seen in cell cultures and synergist ablation models [@mariappan_ribosomal_2011; @von_walden_mechanical_2012], in addition to the previously reported response to hormonal, nutritional, and cellular energy signals [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @russell_rna-polymerase-i-directed_2005].
-Regarding c-Myc, it has previously been found to regulate gene chromatin remodelling and Pol I availability in non-muscle cells [@poortinga_c-myc_2011], and c-Myc overexpression stimulated skeletal muscle ribosome biogenesis in mice [@mori_c-myc_2021], similar to the rapamycin-independent upregulation after RT in rats [@west_acute_2016].
-However, the findings of the present study do not entirely align with these previous observations, as there was no difference in change of total RNA and rRNA between treatments, despite a significantly greater c-Myc content in placebo compared to glucose post intervention.
-Along with the observed non-linear relationship between c-Myc and total RNA, while important as a general transcription factor, c-Myc does not seem to have the same specificity to stimulate rDNA transcription as UBF, in human skeletal muscle cells.
-Further, the ribosomal protein S6 (rpS6) has previously been shown to correlate strongly with rRNA content after synergist ablation [@nakada_correlation_2016; @chaillou_hypoxia_2012], and is suggested as a reliable measurement of ribosome biogenesis [@nakada_correlation_2016].
-Together, this may explain why the between treatment difference in c-Myc content did not induce differences in total RNA or rRNA content, as the specific transcription factor UBF and the rpS6 increased equally with both treatments.
-Notably, the difference between c-Myc in glucose and placebo may have stimulated pathways not measured in the present study.
-
-The results from the strength tests do indicate that the participants might have been fatigued during the intervention, as they exercised every day for 12 days straight, with strength testing in-between.
+The results from the strength tests do indicate that the participants might have been fatigued during the intervention, due to exercising every day for 12 days straight, with strength testing in-between.
 In this perspective, we expected glucose to increase energy availability and thus, muscle performance during strength testing compared to placebo.
 However, this was not evident, as the mean response in strength as an index of peak torque at all speeds decreased with both treatments from baseline to post five and six sessions.
-Notably, RT with glucose reduced strength significantly less than RT with placebo post five sessions, which may indeed have been an effect of increased energy availability, seeing as glucose is the preferred energy source during strenuous activity [@mul_exercise_2015].
-<!--check this ref--> Arguably, this might imply that RT with placebo experienced a higher stress response to the RT program, thus reduced muscle performance in strength tests.
+Notably, RT with glucose reduced strength significantly less than RT with placebo post five sessions, which may indeed have been an effect of increased energy availability, since glucose is the preferred energy source during strenuous activity [@mul_exercise_2015].
+Arguably, this might imply that RT with placebo experienced a higher stress response to the RT program, thus reduced muscle performance during strength testing.
 This could also explain the differences between treatments in content of the stress responsive c-Myc <!-- REF -->, meaning higher stress exerted during RT with placebo compared to RT with glucose inducing higher accumulation of c-Myc.
@@
 
-In conclusion, resistance exercise training combined with glucose ingestion does not acutely augment ribosome biogenesis following two weeks of heavy resistance training in moderately trained young male and female adults.
-Neither does it acutely enhance muscular performance or recovery withing 23hrs, measured as unilateral isometric and isokinetic knee extension peak torque.
+In conclusion, resistance exercise training combined with glucose ingestion does not acutely augment ribosome biogenesis following two weeks of moderate resistance training in moderately trained young male and female adults.
+Neither does it acutely enhance muscular performance or recovery within 23hrs, measured as unilateral isometric and isokinetic knee extension peak torque.
 The observations in total RNA, rRNA and rDNA transcription initiation associated proteins confirm the key role of increased UBF content in ribosome biogenesis in moderately trained human skeletal muscle following resistance training, previously observed in RT-untrained participants, rodents or *in vitro* <!-- REF -->.

### 2023-03-20 ribose-paper.RMD
updated discussion

@@
 
-The main findings of the present study were that heavy resistance training with glucose did not affect markers of ribosome biogenesis, measured as total RNA, rRNA and rDNA transcription initiation associated protein, compared to RT with placebo.
+The main findings of the present study were that heavy resistance training with glucose did not affect markers of ribosome biogenesis, measured as total RNA, rRNA and rDNA transcription initiation associated proteins, compared to RT with placebo.
 Similarly, RT with glucose in general did not affect markers of skeletal muscle functionality such as strength and recovery, or total training session volume.
@@
 Notably, the present study was specifically designed to investigate differences in pre-post changes within-participants, and thus also quite limited to this perspective.
-Hence, the observed effects of time serve as supplementary to the efficiency of the design, without a negative control group to either confirm or refute them.
+Hence, the observed effects of time serve as supplementary to the efficiency of the design, without a negative control group to compare.
 
@@
 found physiological hyperinsulinemia to stimulate p70S6K phosphorylation in human skeletal muscle [@hillier_physiological_2000].
-Further, robust increases in markers of ribosome biogenesis such as 47S pre-rRNA and mature rRNA can be expected after an acute bout of RT [@figueiredo_impact_2016] or a short period of RT [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022].
+Moreover, robust increases in markers of ribosome biogenesis such as 47S pre-rRNA and mature rRNA can be expected after an acute bout of RT [@figueiredo_impact_2016] or a short period of RT [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022].
 Therefore, we hypothesized that if glucose can, directly or indirectly via insulin, stimulate initiation of rDNA transcription, five sessions of RT with glucose compared to placebo will elicit this effect.
-Further, glucose ingestion increases energy levels, thus, we also hypothesized that high glucose treatment would stimulate energy-sensitive pathways such as PIH1, mTORC1, ERK1/2, AMPK and SIRT1 [@kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012].
+Further, glucose ingestion increases energy levels, thus, we also hypothesized that high glucose vs.
+placebo treatment would stimulate energy-sensitive pathways such as PIH1, mTORC1, ERK1/2, AMPK and SIRT1 [@kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012].
 Despite previously reported upregulation in PIC assembly due to high-glucose mediated mTORC1, ERK1/2 and PIH1 or low-glucose mediated AMPK and SIRT1 activation  [@hoppe_amp-activated_2009; @kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012], the present study showed no signs of such effects of glucose vs.
@@
 Indeed, previous studies has used cell cultures from yeast, rodent or human cells [@hoppe_amp-activated_2009; @kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012] <!--arrange refs after yeast, rodent and human --> and are as such not directly comparable to human skeletal muscle cells.
-Further, previous studies investigated high vs.
+Importantly, previous studies investigated high vs.
 low glucose treatment <!-- refs -->, or high glucose vs.
@@
 normal glucose treatment, to investigate the effect of glucose *per se* and not intracellular energy status.
-Thus, while glucose ingestion presumably is important in supplying energy for growth inducing processes such as ribosome biogenesis <!-- refs -->, there is no apparent effect of ingesting glucose *per se* on markers of ribosome biogenesis.
+Thus, while glucose ingestion presumably is important in supplying energy for growth inducing processes such as ribosome biogenesis [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @sebata_control_2018] there is no apparent effect of ingesting glucose *per se* on markers of ribosome biogenesis.
 
 In the present study, despite not measuring the activity in central pathways mediating anabolic signalling (mTORC1, ERK1/2), analyses of the downstream target upstream binding factor (UBF) and the ribosomal protein S6 (rpS6) further supported observations from the RNA data.
-UBF has previously been described as a master regulator of rDNA transcription *in vitro* [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @russell_rna-polymerase-i-directed_2005], while rpS6 is proposed as a valid and reliable means to measure ribosome biogenesis [@nakada_correlation_2016].
+UBF has previously been described as a master regulator of rDNA transcription *in vitro* [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @russell_rna-polymerase-i-directed_2005], while rpS6 is proposed as a valid and reliable means to measure ribosome biogenesis [@chaillou_hypoxia_2012; @nakada_correlation_2016].
 However, content of c-Myc did differ between RT with glucose and RT with placebo, whereas RT with placebo led to significantly higher post content of c-Myc compared to glucose.
-The general transcription factor c-Myc has previously been described as a potent regulator of ribosome biogenesis independent of mTORC1.
-<!-- west2016,  mori21, vonwalden19 --> The independency on mTORC1 may explain why only c-Myc content differed between treatments, while UBF and rpS6 content changed equally.
+The general transcription factor c-Myc has previously been suggested as a potent regulator of ribosome biogenesis independent of mTORC1 [@mori_c-myc_2021; @poortinga_c-myc_2011; @west_acute_2016].
+The independency on mTORC1 may explain why only c-Myc content differed between treatments, while UBF and rpS6 content changed equally.
 Moreover, the observation that total RNA and rRNA was unaffected by the significant differences in c-Myc content, do strengthen the importance and specificity of UBF content in human skeletal muscle ribosome biogenesis.
@@
 In this perspective, we expected glucose to increase energy availability and thus, muscle performance during strength testing compared to placebo.
-However, this was not evident, as the mean response in strength as an index of peak torque at all speeds decreased with both treatments from baseline to post five and six sessions.
+However, this was generally not evident, as the mean response in strength as an index of peak torque at all speeds decreased with both treatments from baseline to post five and six sessions.
 Notably, RT with glucose reduced strength significantly less than RT with placebo post five sessions, which may indeed have been an effect of increased energy availability, since glucose is the preferred energy source during strenuous activity [@mul_exercise_2015].
@@
 
-In conclusion, resistance exercise training combined with glucose ingestion does not acutely augment ribosome biogenesis following two weeks of moderate resistance training in moderately trained young male and female adults.
-Neither does it acutely enhance muscular performance or recovery within 23hrs, measured as unilateral isometric and isokinetic knee extension peak torque.
-The observations in total RNA, rRNA and rDNA transcription initiation associated proteins confirm the key role of increased UBF content in ribosome biogenesis in moderately trained human skeletal muscle following resistance training, previously observed in RT-untrained participants, rodents or *in vitro* <!-- REF -->.
+In conclusion, ingestion of glucose immediately before and after resistance exercise training does not acutely augment ribosome biogenesis after two weeks of moderate volume exercise, in moderately trained young adults.
+Neither does it acutely enhance muscular performance during the exercise period nor after five sessions, or recovery within 23hrs of the last session.
+The observations in total RNA, rRNA and rDNA transcription initiation associated proteins support the key role of UBF in ribosome biogenesis regulation in human skeletal muscle following resistance training.
 If future investigations are to be made on this topic, a greater sample size coupled with a negative control group and analyses of the intramuscular glycogen storage may provide higher resolution results.

### 2023-03-20 ribose-paper.RMD
added some references

@@
 Due to the exploratory nature of our hypotheses this was hardly a very surprising result.
-Indeed, previous studies has used cell cultures from yeast, rodent or human cells [@hoppe_amp-activated_2009; @kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012] <!--arrange refs after yeast, rodent and human --> and are as such not directly comparable to human skeletal muscle cells.
+Indeed, previous studies has used cell cultures from yeast [@zhai_human_2012], rodent [@hoppe_amp-activated_2009; @mariappan_ribosomal_2011] or human cells [@tanaka_mild_2015] and are as such not directly comparable to human skeletal muscle cells.
 Importantly, previous studies investigated high vs.
-low glucose treatment <!-- refs -->, or high glucose vs.
-glucose starvation <!-- refs -->, while the present study aimed to compare high glucose treatment vs.
-placebo, with a matched daily macro nutrient and energy intake.
+low glucose treatment [@mariappan_ribosomal_2011], or high glucose vs.
+glucose starvation [@hoppe_amp-activated_2009; @tanaka_mild_2015], while the present study aimed to compare high glucose treatment to placebo, with a matched daily macro nutrient and energy intake.
 And so, the comparison of the present study was high glucose vs.
@@
 However, content of c-Myc did differ between RT with glucose and RT with placebo, whereas RT with placebo led to significantly higher post content of c-Myc compared to glucose.
-The general transcription factor c-Myc has previously been suggested as a potent regulator of ribosome biogenesis independent of mTORC1 [@mori_c-myc_2021; @poortinga_c-myc_2011; @west_acute_2016].
+The general transcription factor c-Myc has previously been suggested as a potent regulator of ribosome biogenesis independent of mTORC1, and a direct regulator of UBF [@mori_c-myc_2021; @poortinga_c-myc_2011; @west_acute_2016].
 The independency on mTORC1 may explain why only c-Myc content differed between treatments, while UBF and rpS6 content changed equally.
-Moreover, the observation that total RNA and rRNA was unaffected by the significant differences in c-Myc content, do strengthen the importance and specificity of UBF content in human skeletal muscle ribosome biogenesis.
+However,  Moreover, the observation that total RNA and rRNA was unaffected by the significant differences in c-Myc content, do strengthen the importance and specificity of UBF content in human skeletal muscle ribosome biogenesis.
 The linear relationship found exclusively between UBF content and total RNA accumulation, and not c-Myc or rpS6, found in the present study further supports this notion.
@@
 Arguably, this might imply that RT with placebo experienced a higher stress response to the RT program, thus reduced muscle performance during strength testing.
-This could also explain the differences between treatments in content of the stress responsive c-Myc <!-- REF -->, meaning higher stress exerted during RT with placebo compared to RT with glucose inducing higher accumulation of c-Myc.
+This could also explain the differences between treatments in content of the stress responsive c-Myc [@mori_c-myc_2021; @poortinga_c-myc_2011; @west_acute_2016], meaning higher stress exerted during RT with placebo compared to RT with glucose inducing higher accumulation of c-Myc.
 If this was the case, it does seem inconclusive that strength did not differ between treatments at earlier or later time points.

### 2023-03-20 ext_method.RMD
updated RMD

@@
 
-Weight and RNA was summarised, then we devided RNA by mean centered weight. Thus, we could compare total RNA results analyzed on different nanoDrop plates.
+Weight and RNA was summarised, RNA was devided by mean centered weight. Thus, we could compare total RNA results analyzed on different nanoDrop plates.
 
@@
  5. nf.w = nf.w / max(nf.w)
- 6. nf.expr = log(expr / nf.w), nf.w = scale(nf.w)
+ 6. nf.expr = log(expr / nf.w), nf.w = scale(nf.w) ?
  - expr = target expression 
  - weight = biopsy weight
+
+To accurately measure changes in expression of rRNA, we added a fixed amount of Lambda to each total RNA sample during RNA extraction.
+Lambda expression and target (47S, 18S, 28S, 5.8S, 5S) were calculated in separate data frames, and then joined for normalization.
+In each data frame, we first created a normalization factor for weight (nf.w), a result of expression divided by weight.
+To make the nf.w comparable across the qPCR plates, we put it on a scale through the code nf.w = nf.w /max/(nf.w).
+This code divides the mean nf.w by the highest observed value in the variable nf.w.
 
@@
 
-To create the tpl normalization factor, we measured the mean gray both within (tp) and between (bg) the lanes of each total protein stain blot with the
+To create the tpl normalization factor, we measured the mean gray both within (tp) and between (bg) the lanes of each total protein stain blot with the ImageJ software.
+So, each lane and each space between lanes got a "mean gray" value.
+Subtracting background from total protein (tpl = tp-bg), and then using the resulting value divided by the maximum observed total protein in a lane per gel (tpl = tpl/max(tpl)), gave us a value normalized between gels.
+In addition, to account for variations in loading and unequal size in muscle tissue samples, we also normalized antibody signal by pools.
+The pools were mixes of total protein sample, 20ul from each sample, loaded into three lanes per gel.
+So, after normalizing signal to tpl (norm.sign = signal / tpl), we summarized the average signal from the pools (pool = mean(pool)), and divided the tpl normalized signal by average pool signal (norm.sign = norm.sign / pool).

### 2023-03-20 ribose-paper.RMD
updated RMD

@@
 author: "Kristian Lian, Daniel Hammarström, Stian Ellefsen, Håvard Hamarsland"
-bibliography: resources/ribose-kl.bib
+bibliography: resources/ribose-paper.bib
 csl: resources/jap.csl
@@
 Muscle biopsies were sampled from m.
-vastus lateralis using well-established procedures [@hammarstrom.etal_2020].
+vastus lateralis using well-established procedures [@hammarstrom_benefits_2020].
 Briefly, muscle biopsy sampling was performed under local anaesthesia (Xylocaine, 10mg ml-1 with adrenaline 5μg ml−1, AstraZeneca AS, Oslo, Norway) using a 12-14-gauge needle (Universal Plus, Mermaid Medical AS, Stenløse, Denmark), operated with a spring-loaded biopsy gun.
@@
 This power calculation was based on an assumption that the effects of glucose ingestion on ribosome biogenesis may equate to the effects of increasing RT volume from low to moderate [@hammarstrom_benefits_2020].
-<!-- maybe include the power calculation in extended methods?-->.
-Raw data was imported an analyzed in R via R studios [@Rcore].
+<!-- maybe include the power calculation in extended methods?--> Raw data was imported and analyzed in R via R studios [@Rcore].
 Legs were categorized as either glucose or placebo in accordance with the randomization.
@@
 Total RNA and qPCR values were analyzed by comparing post to pre log-fold change score differences.
-This was achieved by modelling change as response with baseline and supplement (glucose/placebo) as explanatory factors, and mean-centered baseline values as a covariate in a linear mixed effects model, then calculating the estimated marginal means.
+This was achieved by modelling change as response with baseline and supplement (glucose/placebo) as explanatory factors, and mean-centered baseline values as a covariate in a linear mixed effects model using, then calculating the estimated marginal means.
 Plasma glucose and c-peptide, training volume, and the strength index were analyzed by multiple time-point log-fold change score comparisons, using the same linear mixed effects model with time added as explanatory factor.
-The linear mixed effects model was designed with the lmer function of the lme4 package [@bates_fitting_2014] in R studios, using the lmerTest function to procure p-values.
-The relationship between
-Log-transformation and mean-centering baseline values was done to control for heteroscedasticity and regression to the mean, respectively.
+The linear mixed effects model was designed with the lmer function (Satterthwaite's method) from the lme4 package [@bates_fitting_2014] in R studios, using the lmerTest function to procure p-values.
+The relationship between <!-- update this --> Log-transformation and mean-centering baseline values was done to control for heteroscedasticity and regression to the mean, respectively.
 Log-transformed values were reverse-transformed for figure illustration.

### 2023-03-21 ribose-paper.RMD
updated discussion

@@
 
-The results from the strength tests do indicate that the participants might have been fatigued during the intervention, due to exercising every day for 12 days straight, with strength testing in-between.
-In this perspective, we expected glucose to increase energy availability and thus, muscle performance during strength testing compared to placebo.
-However, this was generally not evident, as the mean response in strength as an index of peak torque at all speeds decreased with both treatments from baseline to post five and six sessions.
-Notably, RT with glucose reduced strength significantly less than RT with placebo post five sessions, which may indeed have been an effect of increased energy availability, since glucose is the preferred energy source during strenuous activity [@mul_exercise_2015].
-Arguably, this might imply that RT with placebo experienced a higher stress response to the RT program, thus reduced muscle performance during strength testing.
-This could also explain the differences between treatments in content of the stress responsive c-Myc [@mori_c-myc_2021; @poortinga_c-myc_2011; @west_acute_2016], meaning higher stress exerted during RT with placebo compared to RT with glucose inducing higher accumulation of c-Myc.
-If this was the case, it does seem inconclusive that strength did not differ between treatments at earlier or later time points.
-In addition, there were no differences in training volume per session between treatments, therefore, the mechanical loading should have been equal.
-Another potential explanation to this observation might be ….
-(bifasisk restitusjon?)<!-- investigate this -->.
+As in the biological data, there was generally no difference in change of strength as an index between the treatments.
+Generally, skeletal muscle strength decreased from baseline to post fifth and sixth RT session, with a significantly less reduction in glucose compared to placebo at post fifth session measurement.
+This could suggest that the RT protocol fatigued the participants, as both training frequency and intensity was quite high.
+Another promising explanation might be the biphasic recovery pattern, as described by Raastad & Hallén [@RAASTAD2000].
+In their study on recovery of skeletal muscle contractility after strength exercise, Raastad and Hallén (2000) found a rapid recovery during the initial 11hrs post-exercise, followed by either a leveling off or a drop until 22hrs post-exercise, returning to baseline values at 33hrs post-exercise.
+Herein, inflammation and phagocytic activity was proposed to be involved in th performance drop between 11-22hrs [@RAASTAD].
+Indeed, this pattern seems quite similar to what was observed in the present study, with a rapid recovery at 30min and 2hrs post sixth RT session and a drop at 23hrs.
+Moreover, the strength measurements during the intervention and after five RT sessions was also conducted approximately 23hrs after RT.
+Thus, it is possible that these strength measurements also was affected by the biphasic recovery pattern.
+However, this does not explain the difference between treatments at post fifth RT session.
+A possible argument could be that exercising without glucose may have caused more stress compared to exercising with glucose, as glucose is the preferred energy source during strenuous exercise [@mul_exercise_2015], thus increased performance in RT with glucose compared to RT with placebo.
+This may also, at least partly, explain significantly higher content of the stress responsive c-Myc in RT with placebo compared to glucose [@mori_c-myc_2021; @poortinga_c-myc_2011; @west_acute_2016].
+However, training volume data do reveal that the total training session volume was equal on pairwise consecutive days, i.e.
+no difference between day 1-2, day 3-4 and so on.
+Thus, there were no differences in mechanical loading to induce higher stress in RT with placebo compared to RT with glucose.
+Arguably, the significantly higher plasma levels of glucose in RT with glucose may have increased energy availability and thus, ATP, resulting in a lower metabolic stress, compared to RT with placebo [..] <!-- REFS --> Unfortunately, we did not conduct measurements of inorganic phosphate, H+, Mg2+ and the ADP/ATP ratio, thus, discussing the potential effect of differences in metabolic stress between treatments would only be speculation, however possible.
 Lastly, these inconclusive observations may also reflect our limited sample size, despite using a within-subjects design.

### 2023-03-21 ribose-paper.RMD
updated ribose-paper

@@
 Responses to systematic resistance training (RT) vary widely between individuals, with as many as 10-15% showing impaired muscle growth in response to standardized training interventions [@alvarez_interindividual_2018; @mann_high_2014; @thalacker-mercer_cluster_2013].
-Such inability to benefit from RT may be explained by genetic and/or epigenetic disposition, affecting the internal physiological milieu in terms of ability to mount an anabolic response to RT [@figueiredo_genetic_2021;@thalacker-mercer_cluster_2013].<!-- Check these refs, do they correspond to genet epigen and internal mileu?
-Kristian: The way I read it, it looks like it corresponds --> Although RT program modifications such as altered exercise volume or intensity may induce more favorable adaptations in some individuals, such benefits are not uniform between individuals [@hammarstrom_benefits_2020;@krieger_single_2009;@schoenfeld_dose-response_2017].<!-- Add ref for more rt program modifications.
+Such inability to benefit from RT may be explained by genetic and/or epigenetic disposition, affecting the internal physiological milieu, subsequently affecting the ability to mount an anabolic response to RT [@figueiredo_genetic_2021;@thalacker-mercer_cluster_2013].<!-- Check these refs, do they correspond to genet epigen and internal mileu?
+Kristian: The way I read it, it looks like it corresponds --> Although RT program modifications such as altered exercise volume or intensity may induce more favorable adaptations in some individuals, such benefits are not uniform between individuals [@hammarstrom_benefits_2020;@krieger_single_2009;@schoenfeld_strength_2017].<!-- Add ref for more rt program modifications.
 Kristian: Added krieger et al.
@@
 MOSS.
-Kristian: These three should do it--> which in turn determines muscle growth by increasing the muscles translational capacity [@figueiredo_regulation_2019;@hammarstrom_benefits_2020;@sebata_control_2018;@stec_ribosome_2016;@von_walden_ribosome_2019].
+Kristian: These three should do it--> which in turn determines muscle growth by increasing the muscle's translational capacity [@figueiredo_regulation_2019;@hammarstrom_benefits_2020;@sebata_control_2018;@stec_ribosome_2016;@von_walden_ribosome_2019].
 Therefore, investigating the effects of combined RT and glucose ingestion may provide valuable insight into a potential additive effect of glucose and RT on ribosome biogenesis.
@@
 Transcription of ribosomal RNA (rRNA) by Pol I is considered the limiting step in synthesizing new ribosomes [@moss_promotion_1995].
-Multiple signaling pathways converge to regulate rRNA transcription, including the mammalian target of rapamycin (mTORC1) signal-transduction pathway and c-Myc [@kusnadi_regulation_2015;@von_walden_ribosome_2019;@sebata_control_2018; @west_acute_2016; @mori_c-myc_2021].
+Multiple signalling pathways converge to regulate rRNA transcription, including the mammalian target of rapamycin (mTORC1) signal-transduction pathway and c-Myc [@kusnadi_regulation_2015;@von_walden_ribosome_2019;@sebata_control_2018; @west_acute_2016; @mori_c-myc_2021].
 mTOR receives input from growth factors, hormones, mechanical loading, and nutrients to balance protein synthesis through multiple mechanisms based on cellular energy levels [@hoppe_amp-activated_2009].
-Importantly, mTOR and its downstream target S6K1 serve as a mediator from insulin signaling in skeletal muscle [@hillier_physiological_2000].
+Importantly, mTOR and its downstream target S6K1 serve as a mediator from insulin signalling in skeletal muscle [@hillier_physiological_2000].
 The mTORC1 pathway contributes to ribosome biogenesis through the regulation of translation of ribosomal proteins and to forming of the preinitiation complex (PIC) that marks the initiation of rRNA transcription [@figueiredo_regulation_2019;@von_walden_ribosome_2019].
@@
 Additionally, UBF phosphorylation, required for interaction with the rDNA promoter, is increased by high glucose in a mTORC1 dependent manner (rapamycin sensitive) in kidney glomerular epithelial cells[@mariappan_ribosomal_2011].
-Independently from UBF and mTORC1, high glucose was shown to lead to chromatin remodeling, which promotes rRNA transcription in cell cultures[@zhai_human_2012].
+Independently from UBF and mTORC1, high glucose was shown to lead to chromatin remodelling, which promotes rRNA transcription in cell cultures[@zhai_human_2012].
 Together these observations indicate a potential role of glucose in positively affecting muscle rRNA synthesis.
@@
 The purpose of this investigation was to test the hypothesis that glucose supplementation before and after<!-- during?
-Kristian: before and after would be more precise, since they're not strictly ingesting anything during RT - I got a comment on that during the thesis defense --> RT would increase markers of ribosomal abundance following five RT sessions.
+Kristian: before and after would be more precise, since they're not strictly ingesting anything during RT - I got a comment on that during the thesis defense --> RT increasea markers of ribosomal abundance following five RT sessions.
 Secondly we aimed to describe the association between UBF and total RNA abundance in human skeletal muscle.
-There has been multiple studies the last years suggesting translational capacity to be as important as, if not more, translational efficiency [@hammarstrom_benefits_2020;@hammarstrom_ribosome_2022;@figueiredo_revisiting_2019].
+There has been multiple studies the last years suggesting translational capacity to be as important as, if not more, translational efficiency in regard to long-term skeletal muscle adaptations to RT [@hammarstrom_benefits_2020;@hammarstrom_ribosome_2022;@figueiredo_revisiting_2019].
 Previously, *in vitro* studies has shown UBF and c-Myc to be a important factors in the PIC, c-Myc as a general transcription factor and UBF as a specific transcription factor to rDNA transcription initiation [@mariappan_ribosomal_2011;@von_walden_ribosome_2019].
@@
 
-The study was designed as a 12-day double-blinded randomized controlled trial, with a alternating unilateral RT protocol (Fig 1A).
+The study was designed as a 12-day double-blinded randomized controlled trial, with an alternating unilateral RT protocol (Fig 1A).
 Participants were randomly allocated to exercise one leg with glucose and one leg with placebo (1A).
@@
 Thus, participants exercised every day, alternating between e.g., one leg + glucose and the other leg + placebo - this might be TMI, as it's presented in the figure as well?--> One person was exclusively responsible for the randomization code and supplement distribution, blinding both investigators and participants in regard to which leg exercised with glucose and placebo.
-Glucose (Glucosum monohydricum, Merck KGaA, Darmstadt, Germany) and placebo (Steviosa, Soma Nordic AS, Oslo, Norway) was masked by mixing with 300ml Fun Light juice/saft (Orkla, Oslo, Norway).
+Glucose (Glucosum monohydricum, Merck KGaA, Darmstadt, Germany) and placebo (Steviosa, Soma Nordic AS, Oslo, Norway) were masked by mixing with 300ml Fun Light juice/saft (Orkla, Oslo, Norway).
 A blinded taste test revealed that the participants were not able to disclose the contents of the provided boluses (30g glucosum monohydricum vs.
 ~0.3g Stevia rebaudiana extract).
-To ensure equal conditions during training sessions, participants exercised and tested at the same time of day, +/- 1hr (check this) with the same supervisor on pairwise consecutive days.
-To further standardize this, participants also recorded and repeated their daily macro nutrient intake (protein, fat, carbohydrate) and total calories on pairwise consecutive days.
+To ensure equal conditions during training sessions and strength testing, participants exercised and tested at the same time of day, +/- 1hr with the same supervisor on pairwise consecutive days.
+To further standardize this, participants also recorded and repeated their daily macronutrient intake (protein, fat, carbohydrate) and total calories on pairwise consecutive days.
 
 Six RT sessions were conducted with glucose and six with placebo, allowing a within-subjects analysis of the effects of glucose ingestion before and after RT.
-Data from the first five RT sessions was used to investigate main outcome measures (total RNA, rRNA and protein) and leg muscle strength, whereas data from the sixth RT session was used to explore secondary outcomes (muscular recovery, plasma glucose levels).
-Participants were asked to avoid  resistance- or high-intensity training of the legs from Day -7 (Fig 1A) and onwards, until completion of the intervention and post-testing, to ensure the reliability of pre-intervention strength data and minimal interference from other training sources.
+Data from the first five RT sessions was used to investigate main outcome measures (total RNA, rRNA and protein) and leg muscle strength, whereas data from the sixth RT session was used to explore secondary outcomes (muscular recovery, plasma glucose and -c-peptide levels).
+Participants were asked to avoid  resistance- or high-intensity training of the legs from Day -7 (Fig 1A) and onwards, until completion of the intervention and post-testing, to ensure the reliability of pre-intervention strength data and minimal interference from external exercise sources.
 If participants did conduct exercise training outside of the prescribed protocol, they were asked to ensure equal loading on pairwise consecutive days.
@@
 Strength tests were performed prior to (Days -7 and -5, and -1, both legs) and during the intervention (Days 4 and 8 leg 1, 5 and 9 leg two), after session 5 and after finalization of the intervention (Day 11/12 leg 1 and 12/13 leg 2) (Fig 1A).
-Maximal isometric and isokinetic knee extension torque were measured with a Humac Norm Dynamometer (CSMi, Stoughton, Massechusetts, USA).
+Maximal isometric and isokinetic knee extension torque was measured with a Humac Norm Dynamometer (CSMi, Stoughton, Massechusetts, USA).
 Individual positions were recorded and standardized from pre-intervention tests (days -7 and -5).
@@
 During days 11 and 12, humac tests were performed at four time points: I) 45min before RT (1D, Post 5RT), II) 30min after the last RT session (1D, 30 min post 6RT), III) 2hrs after the last RT session (1D, 2h post 6RT), and IV) 23hrs after the last RT session (1D, 23h post 6RT).
-Test I at day 12/13 included testing of both legs, representing 23hrs post RT session test of one leg and post session 5 test of the other leg.
+Test I at day 12/13 included testing of both legs, representing 23hrs post-RT session test of one leg and post-session 5 test of the other leg.
 
-Assessment of unilateral one repetition maximum (1RM) leg press and knee extension was conducted at pre-intervention testing during familiariztion (days -7 and -5, Fig 1D).
+Assessment of unilateral one repetition maximum (1RM) leg press and knee extension was conducted at pre-intervention testing during familiarization (days -7 and -5, Fig 1D).
 The participants performed a general warm-up with 10min cycling on an indoor exercise bicycle.
-In addition, a protocol consisting of 1x10, 1x6 and 1x3 repetitions, was used as spesific warm-up before each of the tests.
+In addition, a protocol consisting of 1x10, 1x6 and 1x3 repetitions, was used as specific warm-up before each of the tests.
 All positions were controlled and recorded at the first 1RM test and repeated for the RT sessions.
@@
 In addition, before the respective exercises, two 10-repetition warm-up sets were completed at ~50% and ~70% of 10RM.
-Breaks between working sets was 2min, and progressive loading was used to ensure adequate exercise stimulation throughout the intervention.
+Breaks between working sets were 2min, and progressive loading was used to ensure adequate exercise stimulation throughout the intervention.
 For safety and standardization purposes, all sessions were monitored by trained personnel.
@@
 Total RNA was extracted using TRIzol according to the manufacturer’s protocol.
-An exogenous RNA control (λ polyA External Standard Kit, Takara Bio Inc., Shiga, Japan) was added at a fixed amount (0.04 ng ml-1 of Trizol reagent) to enable analysis of target gene expression per unit tissue weight (ELLEFSEN2008,2014).
+An exogenous RNA control (λ polyA External Standard Kit, Takara Bio Inc., Shiga, Japan) was added at a fixed amount (0.04 ng ml-1 of Trizol reagent) to enable analysis of target gene expression per unit tissue weight [@ellefsen_improved_2008; @ellefsen_reliable_2014].
 Briefly, muscle tissue was homogenized in TRIzol using 0.5mm RNase-free Zirconium beads (~50 ul) (Next Advanced, Averill Park, NY, USA) and a bullet blender (bb).
@@
 Lambda was used as an external reference gene, added in the RNA extraction (2µl per extraction).
-qPCR reactions were run on a fast-cycling real-time detection system (Applied Biosystems 7500 fast Real-Time PCR Systems, Life Technologies AS), with a total volume of 10 µl, containing 2 µl of cDNA, gene-specific primers (0.5 µM final concentration) and a commercial master mix (2X SYBR Select Master Mix, Applied Biosystems, Life Technologies AS) (Hammarström et al.
-2020).
-qPCR reactions consisted of 40 cycles (3 s 95°C denaturing and 30 s 60°C annealing) (Hammarström et al.
-2020).
-Raw fluorescence data were exported from the platform-specific software and amplification curves were modelled with a best-fit sigmoidal model using the qPCR package (Ritz & Spiess, 2008) written for R (Hammarström et al.
-2020; Team 2018).
+qPCR reactions were run on a fast-cycling real-time detection system (Applied Biosystems 7500 fast Real-Time PCR Systems, Life Technologies AS), with a total volume of 10 µl, containing 2 µl of cDNA, gene-specific primers (0.5 µM final concentration) and a commercial master mix (2X SYBR Select Master Mix, Applied Biosystems, Life Technologies AS) [@hammarstrom_benefits_2020].
+qPCR reactions consisted of 40 cycles (3 s 95°C denaturing and 30 s 60°C annealing) [@hammarstrom_benefits_2020].
+Raw fluorescence data were exported from the platform-specific software and amplification curves were modelled with a best-fit sigmoidal model using the qPCR package [@ritz_qpcr_2008] written for R [@hammarstrom_benefits_2020; @Rcore].
 
@@
 
-A priori power calculations showed that 20 participants would grant a statistical power of 80% (= 0.05), with an expected drop out of 20%.
+A priori power calculations showed that 20 participants would grant a statistical power of 80% (= 0.05), with an expected dropout of 20%.
 This power calculation was based on an assumption that the effects of glucose ingestion on ribosome biogenesis may equate to the effects of increasing RT volume from low to moderate [@hammarstrom_benefits_2020].
@@
 This was achieved by modelling change as response with baseline and supplement (glucose/placebo) as explanatory factors, and mean-centered baseline values as a covariate in a linear mixed effects model using, then calculating the estimated marginal means.
+Quantified western blot signals and the UBF/total RNA association were analyzed using a linear mixed effects model from the nlme package [@lindstrom_nonlinear_1990], with protein signal per time point as response, with time and supplement as explanatory factors.
 Plasma glucose and c-peptide, training volume, and the strength index were analyzed by multiple time-point log-fold change score comparisons, using the same linear mixed effects model with time added as explanatory factor.
 The linear mixed effects model was designed with the lmer function (Satterthwaite's method) from the lme4 package [@bates_fitting_2014] in R studios, using the lmerTest function to procure p-values.
-The relationship between <!-- update this --> Log-transformation and mean-centering baseline values was done to control for heteroscedasticity and regression to the mean, respectively.
+Log-transformation and mean-centering baseline values was done to control for heteroscedasticity and regression to the mean, respectively.
 Log-transformed values were reverse-transformed for figure illustration.
@@
 
-There were no difference in mean change of total training session volume between exercising with placebo or glucose, with an increase of 18 and 16% from session 1 to session 6, respectively (Tab 2; Mean change, 2: *p* = `r pvol.c2`, 3: *p* = `r pvol.c3`, 4: *p* = `r pvol.c4`, 5: *p* = `r pvol.c5`, 6: *p* = `r pvol.c6`).
-Both exercising with glucose and placebo led to significant increases in mean total training session volume at session four, five and six, compared to session 1 (Tab 2; Mean change, 2: *p* = `r pvol.s2`, 3: *p* = `r pvol.s3`, 4: *p* = `r pvol.s4`, 5: *p* = `r pvol.s5`, 6: *p* = `r pvol.s6`).
+There was no difference in mean change of total training session volume between exercising with placebo or glucose, with an increase of 18 and 16% from session 1 to session 6, respectively (Tab 2; Mean change, 2: *p* = `r pvol.c2`, 3: *p* = `r pvol.c3`, 4: *p* = `r pvol.c4`, 5: *p* = `r pvol.c5`, 6: *p* = `r pvol.c6`).
+Both exercising with glucose and placebo led to significant increases in mean total training session volume at session four, five and six, compared to session 1 (Tab 2; Mean change, 2: *p* = `r pvol.s2`, 3: *p* = `r pvol.s3`, 4: *p* = `r pvol.s4`, 5: *p* = `r pvol.s5`, 6: *p* = 0.001).
 In regard to the dietary intervention, there were no difference in mean macro nutrient intake (protein, fat, carbohydrates) or total calorie intake between glucose and placebo on pairwise consecutive days (*p* > 0.05).
@@
 
-#### Total RNA and ribosomal RNA
+##### Total RNA and ribosomal RNA
 
@@
 From baseline to post-intervention, there was a mean increase in levels of total RNA by 26 and 22% after RT with glucose and placebo, respectively (*p* < 0.05, compared to baseline).
-A robust accumulation was also observed in ribosomal RNA expression in both RT with glucose and placebo, with mean increases between 34-43% (GLU) and 33-41% (PLAC) (*p* < 0.05, compared to baseline) in the four rRNA’s.
+A robust accumulation was also observed in ribosomal RNA expression in both RT with glucose and placebo, with mean increases between 34-43% (GLU) and 33-41% (PLAC) (*p* < 0.05, compared to baseline) in the four rRNAs.
 The expression of the 47S pre-rRNA increased by 37 and 59% with glucose and placebo respectively, where only RT with placebo increased significantly from baseline to post (fig 45, *p* = `r rr47tdiffp`).
 
-#### Protein
+##### Protein
 
@@
 This could suggest that the RT protocol fatigued the participants, as both training frequency and intensity was quite high.
-Another promising explanation might be the biphasic recovery pattern, as described by Raastad & Hallén [@RAASTAD2000].
+Another promising explanation might be the biphasic recovery pattern, as described by Raastad & Hallén [@raastad_recovery_2000].
 In their study on recovery of skeletal muscle contractility after strength exercise, Raastad and Hallén (2000) found a rapid recovery during the initial 11hrs post-exercise, followed by either a leveling off or a drop until 22hrs post-exercise, returning to baseline values at 33hrs post-exercise.
-Herein, inflammation and phagocytic activity was proposed to be involved in th performance drop between 11-22hrs [@RAASTAD].
+Herein, inflammation and phagocytic activity was proposed to be involved in th performance drop between 11-22hrs [@raastad_recovery_2000].
 Indeed, this pattern seems quite similar to what was observed in the present study, with a rapid recovery at 30min and 2hrs post sixth RT session and a drop at 23hrs.
@@
 Thus, there were no differences in mechanical loading to induce higher stress in RT with placebo compared to RT with glucose.
-Arguably, the significantly higher plasma levels of glucose in RT with glucose may have increased energy availability and thus, ATP, resulting in a lower metabolic stress, compared to RT with placebo [..] <!-- REFS --> Unfortunately, we did not conduct measurements of inorganic phosphate, H+, Mg2+ and the ADP/ATP ratio, thus, discussing the potential effect of differences in metabolic stress between treatments would only be speculation, however possible.
+Arguably, the significantly higher plasma levels of glucose in RT with glucose may have increased energy availability and thus, ATP, resulting in a lower metabolic stress and less performance reduction, compared to RT with placebo [@westerblad_mechanisms_1998; @kent-braun_central_1999] Unfortunately, we did not conduct measurements of markers of metabolic stress; inorganic phosphate, H+, Mg2+ and the ADP/ATP ratio [@westerblad_mechanisms_1998; @kent-braun_central_1999], thus, discussing the potential effect of differences in metabolic stress between treatments would only be speculation, however possible.
 Lastly, these inconclusive observations may also reflect our limited sample size, despite using a within-subjects design.

### 2023-03-21 figures/figure3.R

@@ -111,7 +111,7 @@ rps6.plot <- rps6 %>%
   geom_line(position = position_dodge(width = 0.2), lty = 2) +
   geom_point(shape = 21, size = 2, position = position_dodge(width = 0.2)) +
   scale_fill_manual(values = colors[c(4,1)],
-                    labels = c("Glucose", "Placebo")) +
+                    labels = c("Placebo", "Glucose")) +
   labs(x = "Time", y = "rpS6 AU", fill = "Supplement") +
   theme(axis.title = element_text(size = htextsize),
         axis.title.x = element_blank(),

### 2023-03-21 ext_method.RMD
updated ext_method

@@
 
-###### **Tab 1:** Primer sequence and performance. Average cycle thresholds (Ct) and primer efficiencies were calculated from all qPCR reactions.
+#### Extraction of RNA from muscle tissue (protocol)
+
+##### Materials
+
+For extraction of RNA and protein per sample:
+
+1.5 ml Eppendorf tube with 0.5 mm RNase-free Zirconium beads (~50 ul) Keep out DNA RNAse contaminations (Eppendorf safe-lock for use in bullet blender.
+1.5 ml tube for RNA precipitation
+1.5 ml tube for RNA quantification (1:5 dilution, add 20 ul of DEPC-water)
+1.5 ml tubes for Protein precipitation
+1.5 ml tubes for Soluble protein fraction
+1.5 ml tubes for protein quantification (1:10 dilution, add 36 ul of ddH2O)
+
+##### Solutions and chemicals
+
+Trizol (Life-technologies)
+Chloroform
+Isopropanol
+External reference gene (e.g. Takara lambda)
+100% EtOH
+95% EtOH
+75% EtOH (kept on -20)
+
+##### Homogenization and phase separation
+
+1.
+Add 1 sample volume of beads and 300 ul Trizol to eppendorf 1.5 ml tubes.
+Keep on ice.
+OPTIONAL: If external reference gene is used, add 1 unit to every extraction with Trizol in first step.
+Lambda (0.04 ng/ul, 2 ul per extraction)
+2. Add frozen muscle samples (protocol optimized for 5-25 ug of wet weight muscle sample) without thawing and put tubes in Bullet blender
+3. Run Bullet Blender (BB) for 1 min at speed 10, put samples on ice for 1 min. Run samples for 1 min at speed 12. Check samples, if not dissolved run until fully homogenized.
+4. Add Trizol (700 ul) to total volume 1000 ul, run samples in BB for 1 min speed 3.
+5. Incubate @ RT for 5 min
+6. Add 200 ul chloroform, shake 15 sec, incubate for 2-3 min.
+7. Spin 12000g, 15min, 4°C
+8.
+Pipette 450 ul of the aqueous phase to a new tube.
+Be careful not to disrupt the interphase or organic phase.
+Add 500 ul of isopropanol to precipitate RNA, incubate over night at -20°C or 10 min at room temp and continue with RNA extraction.
+9. Completely remove the remaining aqueous phase, either store remaining phases at -20°C until protein extraction or continue right away
+
+##### Extraction
+
+Precipitation and wash
+
+1. Spin the isopropanol/water phase 12000g, 10min, 4°C. A pellet should form, decant supernatant, spin briefly and remove remaining liquids.
+2. Add 1000 ul 75% EtOH (-20°C) @ RT, vortex gently. Spin 7500g, 5min, 4°C.
+3. Repeat step 2 two more times.
+4. Remove all EtOH, air dry the pellet up to 10 min with tubes upside-down. Do not over-dry!
+
+Suspension
+1. Add 20-40 ul of 1x DEPC-treated water (55°C), incubate at 55°C for 10min.
+2. Aliquot:
+ 5 ul to 20 ul TE buffer for Nanodrop/Qubit2
+ x ul in stock solution
+3. Store RNA stock solution @ -80°C.
+
+#### cDNA synthesis using Superscript IV and general qPCR
+
+##### cDNA
+
+##### Materials
+
+Heating block or thermal-cycler capable of 65, 50-55 and 80°C
+RNase-free tubes, pipette tips
+Superscript IV kit (SSIV Buffer, SSIV Reverese transcriptase, 100 nM DTT)
+Oligo dT and/or Random Hexamer primers
+10 mM dNTP mix
+Total-RNA or mRNA
+
+1. Combine the following components per reaction (consider creating a master mix from component a and b if you are running many samples, master mix 1):
+ 25 μM Oligo dT and 25 μM Random Hexamers - 0.5 μl Oligo dT (art nr. AB-1247) + 0.5 μl Random hexamers (SO142) per reaction
+ 10 mM dNTP mix (10 mM each) - 1 μl per reaction
+ Template RNA (0.1-5000 ng total-RNA) - up to 11 μl per reaction
+ Nuclease-free water - Add to total volume 13 μl per reaction
+
+2. Mix all components, vortex and briefly spin down.
+3. Heat the mix 5 min at 65°C followed by at least 1 min incubation on ice.
+4. Bring the 5x SSIV buffer to room-temp. Vortex and spin briefly.
+5. Combine the following into master-mix 2:
+ 5x SSIV Buffer - 4 μl per reaction
+ 100 mM DTT - 1 μl per reaction
+ RNase OUT - 1 μl per reaction
+ Super Script reverse transcriptase - 1 μl per reaction
+
+6. Vortex and briefly spin the components of master mix 2, add master-mix 2 to the reaction tube.
+7. Incubate the reaction tube at 23°C 10 min.
+8. Incubate the reaction mix at 50-55 °C 10 min followed by 10 min at 80°C
+9. Dilute the cDNA to desired concentration (e.g. 1:25)
+
+##### qPCR
+
+##### Materials
+
+QuantStudio 5 real-time PCR machine
+A qPCR reaction plate
+Nuclease-free water and pipette tips
+SYBR-green Master mix
+
+1. Combine master-mix:
+ Sybre-green 2X master-mix - 5 μl per reaction
+ Primermix, Forward and Reverse 5 μM each - 1 μl per reaction
+ H2O - 2 μl per reaction
+
+2. Load the plate with primer-specific master-mix
+3. Add 2 μl cDNA sample
+
+### Protein
+
+#### Extraction
+
+###### **Table 3:** Primer sequence and performance. Average cycle thresholds (Ct) and primer efficiencies were calculated from all qPCR reactions.
 

### 2023-03-22 ext_method.RMD
updated ext_method

@@
 
-#### Extraction
+##### Total Protein Extraction via Minute Total Protein Extraction Kit for Muscles (Invent Biotechnology)
+
+Manufacturers protocol:
+1. Weight out 20-30 mg tissue (fresh/frozen) and cut it with a sharp blade or a pair of scissors into 
+small pieces (1 X 1 mm or smaller). Place tissue pieces in a filter cartridge with collection tube.
+
+2. Add 50 to 80 mg protein extraction powder on top of the tissue followed by addition of  75 ul 
+lysis buffer (Denaturing Buffer).
+
+3. Immediately grind the tissue with the plastic rod provided against the surface of the filter with 
+moderate twisting force for 2-3 min. Add another 25ul lysis buffer (Denaturing buffer) to the filter and continue to 
+grind for about 30 seconds to 1 min. The plastic rods are reusable after cleaning.
+
+4. Centrifuge the tube at top speed in a microcentrifuge for 1 min. Remove and discard the filter. 
+Transfer the supernatant to a fresh tube (this is extracted total protein). The white-grey pellet in 
+the bottom of collection tube is passing through protein extraction powder that should be 
+discarded.
+
+##### Measurement of total protein concentration
+Keep samples and ladder on ice as much as possible, minimize time off ice. 
+1. Total protein samples from extraction are diluted 1:10 to a final volume of 40 ul (i.e. 4 units sample and 36 units water). 
+2. Use a 96 well plate, and pipette 3 wells per ladder (A-I) and sample (10ul). 
+3. After all ladders and samples have been added to the wells, pipette 250ul Bradford into each well.
+4. Run the plate using SkanIt software to examine the total protein concentration. (Fill in the running set-up and plate layout here)
+
+##### Gel electrophoresis
+Preparation
+
+1.
+Dilute with denaturing buffer to a protein concentration of i.e.
+3ug/ul (depending on measured total protein concentration).
+Standardize the start volume of the samples, i.e.
+to 50ul, then use C1V1 = C2V2 formula.
+Transfer 50ul total protein to a fresh tube, add dilution volume.
+Make sure to note the dilution volume of each sample, you will use it later on.
+2.
+Mix the loading buffer equal to 1/4 of the total sample volume (add startvolume and dilution volume from step 1, divide this by three = loading buffer volume).
+This should be fresh, mix it on the same day, see solutions.
+Add up the total volum of loading buffer needed and mix it under a vent.
+3. Add loading buffer to the samples, vortex briefly.
+4. Pool 20ul from each sample (or lower if necessary) into a fresh "pooling" tube.
+5. Incubate the samples and pool tube at 95 C for 5min. Lids may pop open during or after incubation, pay attention and close them if they do.
+6. Let the samples cool down briefly, store them at -20 C until loading.
+7. Prepare running buffer (solutions). This must be cold, prepare it the day prior to electrophoresis and store it at 4 C fridge.
+
+Setup:
+
+1. Unpack the gel, and make sure to flush it using water (blue tap/ddH2O). If you are running multiple gels, make sure to mark each gel (note something on the plastic case around the gel). 
+2. Remove the sticker from the gels, and insert the gels into the electrophoresis chamber. Keep on ice. 
+3. Fill up the main electrophoresis chamber and the small chamber on the gels surrounding plastic (with a green plastic placeholder) with running buffer, approximately 1L in total. Keep on ice.
+4. Carefully remove the green placeholder from each gel (use tweezers), then flush each well with running buffer from. Keep on ice.
+
+Loading:
+
+1.
+Make sure to draw or note which samples go into which wells.
+Always add Ladder in the first well (5ul, Precision plus Protein Dual Color standards), and the middle or last (3ul).
+Calculate loading volume.
+Our goal is to load as close as possible to 20.25 ug protein, i.e.
+divide 20.25 by standardised concentration (the concentration with loading buffer).
+If you diluted by three, the concentration will be down to 2.25 after adding loading buffer: 20.25/2.25 = 9.
+Therefore we load the gel with 9 ul sample per well in this case.
+Keep on ice.
+2.
+When all samples and ladders are added to the wells, fill the remaining running buffer into the small chamber where the wells are located.
+Do this carefully, and do not fill it on top of the wells or close to them.
+Keep on ice.
+
+Electrophoresis:
+1. Set the voltage to 250, constant, and mA to 500.
+2. Duration is set to 50min. Keep the electrophoresis unit on ice.
+
+##### Wet transfer
+
+Preperation
+1. Transfer buffer should be made prior to the transfer, and should be cold. 
+2. PVDF membran is cut from the blue scroll (8.5cm high, 14cm broad). DO NOT touch the white membrane. Cut of the top left corner. If you are transfering multiple membranes, mark them. 
+3. Soak foam pads in ddH2O and squeese out all bubbles.
+4. Incubate the membran for 5min in MeOH (make sure MeOH is discarded in designated bottle after). 
+5. Incubate membran 5-15min in cold transfer buffer.
+6. Remove the smallest plastic cover of the gel, and flush the gel with ddH2O, then put the gel in transfer buffer for approx 30sec.
+7. Soak filter paper in transfer buffer.
+
+Kit/Sandwich assembly
+1. Place a magnetic stirrer in the bottom of the Criterion Blotter tank, make sure it has room to spin.
+2. Place the Criterion Blotter tank in an ice tub to keep temperature low.
+3. Fill the tank to 50% of total fill volume with transfer buffer.
+4. Place iceblock in the back of the tank (designated area)
+5. Putt the sandwhich in its tray, pour transfer buffer over it, enoug to submerge it and the next components, then add the bottom foam pad
+6. Take 1 filter paper in one hand and place the gel on it
+7. While holding the filter with the gel on, place the membrane on top of the gel
+8. Put the filter/gel/membrane down on the bottom foam pad, with the filter side facing the foam pad
+9. Carefully use a roller to remove air bubbles, while holding the membrane down with two fingers
+10. Place the last filter paper on the membran, then carefully roll again
+11. Place the last foam pad on top, carefully roll again
+12. Put the sandwich in the Criterion Blotter, make sure red side faces the red electrode, and that the magnetic stirrer still can move. 
+13.
+Add remaining transfer buffer to the tank, fill it up to the marked level.
+Place the ice tub, with the blotter, upon a magnetic stirring plate and start stirring to maintain even buffer temperature and ion concetration during the transfer.
+
+Transfer
+1. Connect PowerPac HC power supply to Criterion Blotter.
+2. Set PowerPac to 400mA, constant A, 1t 45min, at 4 degrees celsius (ice).
+
+MemCode Staining
+NB: Keep the PVDF membrane side that was facing the Gel during transfer up at all times, and do not let the membrane dry!
+1. Wash the membran with water (ddH2O)
+2. Add MemCode Sensitizer. Carefully shake for 2min with the shaker. Sensitizer is reusable for a couple attempts
+3. Add MemCode Stain. Carefully shake for 1min with the shaker. Stain is reusable for a couple attempts
+4. Rinse the membrane 3 times with water (ddH2O).
+6. Wash with Destain/Methanol solution (1:1) on shaker for 5min.
+7. Manual capture in G:Box - make sure the black plate is removed, light is on and image is sharp
+8. Slice the membrane into desired strips using scalpel and a ruler
+9. Rinse 5 times with water (ddH2O)
+10. Wash with eraser/methanol solution (1:1) on a shaker for 10-20min
+11. Rinse 5 times with water (ddH2O)
+
+Antibody staining
+1. Block the membrane for 1h at roomtemp or 4C over night with "blocking buffer" (5% non-fat dry milk, see solutions)
+2. Wash once with TBS-T, 5min, on a tipping board
+3.
+Incubate membrane with suitable dilution of primary antibody.
+It is recommended to incubate at 4C over night.
+On a tipping board.
+Can be done using a polystyrene (isopor) box and ice to keep the temperature low.
+4. Wash the membrane 3 times with TBS-T, 5min per wash, on a tipping board. 
+5. Incubate membrane with recommended dilution of secondary antibodies in blocking buffer at roomtemp for 1h. On a tipping board
+6. Wash the membrane 3-5 times (5 times may yield nicer blots) with TBS-T, 5min per wash, on a tipping board. 
+7. Mix just sufficient ECL and incubate the membrane for 5min on a tipping board. 8ml should suffice if you have sliced the membrane in smaller strips
+8. Either store the membrane in TBS-T or proceed with visualisation. If you store it in TBST, make sure to incubate once more in ECL.
+
+Quantification with G:Box and Genesys
+1. Ready the G:Box and Genesys on the computer, use ECL protocol. 
+3. Sandwich the membrane between two pieces of transparency film, and use the roll to remove any airbubbles. Make sure you do not get any liquid on the black plate 
+3. Place the membrane within the G:Box and initiate protocol. Should not have to click anything but next, until "Capture". Make sure the membranes position is good, as close to centre as possible.
 

### 2023-03-30 ribose-paper.RMD
proofreading

@@
 Responses to systematic resistance training (RT) vary widely between individuals, with as many as 10-15% showing impaired muscle growth in response to standardized training interventions [@alvarez_interindividual_2018; @mann_high_2014; @thalacker-mercer_cluster_2013].
-Such inability to benefit from RT may be explained by genetic and/or epigenetic disposition, affecting the internal physiological milieu, subsequently affecting the ability to mount an anabolic response to RT [@figueiredo_genetic_2021;@thalacker-mercer_cluster_2013].<!-- Check these refs, do they correspond to genet epigen and internal mileu?
-Kristian: The way I read it, it looks like it corresponds --> Although RT program modifications such as altered exercise volume or intensity may induce more favorable adaptations in some individuals, such benefits are not uniform between individuals [@hammarstrom_benefits_2020;@krieger_single_2009;@schoenfeld_strength_2017].<!-- Add ref for more rt program modifications.
+Such inability to benefit from RT may be explained by genetic and/or epigenetic disposition affecting the internal physiological milieu and subsequently affecting the ability to mount an anabolic response to RT [@figueiredo_genetic_2021;@thalacker-mercer_cluster_2013].<!-- Check these refs, do they correspond to genet epigen and internal mileu?
+Kristian: The way I read it, it looks like it corresponds --> Although RT program modifications such as altered exercise volume or intensity may induce more favourable adaptations in some individuals, such benefits are not uniform between individuals [@hammarstrom_benefits_2020;@krieger_single_2009;@schoenfeld_strength_2017].<!-- Add ref for more rt program modifications.
 Kristian: Added krieger et al.
 2009 and schoenfeld et al.
-2017 --> For instance, although changes in training volume generally induces greater benefits of RT, not all individuals benefit from such modifications [@hammarstrom_benefits_2020].
-Therefore, other means than modification of RT variables seems necessary to optimize individual responses.<!-- The above sentence might be a bit contradictory?
+2017 --> For instance, although changes in training volume generally induce greater benefits of RT, not all individuals benefit from such modifications [@hammarstrom_benefits_2020].
+Therefore, other means than modification of RT variables seem necessary to optimize individual responses.<!-- The above sentence might be a bit contradictory?
 Im not sure, are you thinking about "other means than modification of RT" being contradictory of "modifications such as altered exercise volume.." sentence?
@@
 However, it remains unknown if other nutritional adjuvants such as glucose can increase the efficacy of RT.
-This is surprising, as glucose is the preferred energy substrate of the contracting skeletal muscle during strenuous exercise and a major supplier of energy to cells via ATP synthesis [@mul_exercise_2015;@sebata_control_2018].
+This is surprising since glucose is the preferred energy substrate of the contracting skeletal muscle during strenuous exercise and a major supplier of energy to cells via ATP synthesis [@mul_exercise_2015;@sebata_control_2018].
 Furthermore, energy availability is a decisive factor in *de novo* synthesis of ribosomes [@kusnadi_regulation_2015;@moss_housekeeper_2007;@sebata_control_2018] <!--Add ref that indicate high energy requirement for ribo biog possibly (VONWALDEN19;TANAKA18?
 MOSS.
-Kristian: These three should do it--> which in turn determines muscle growth by increasing the muscle's translational capacity [@figueiredo_regulation_2019;@hammarstrom_benefits_2020;@sebata_control_2018;@stec_ribosome_2016;@von_walden_ribosome_2019].
-Therefore, investigating the effects of combined RT and glucose ingestion may provide valuable insight into a potential additive effect of glucose and RT on ribosome biogenesis.
+Kristian: These three should do it-->which in turn determines muscle growth by increasing the muscle's translational capacity [@figueiredo_regulation_2019;@hammarstrom_benefits_2020;@sebata_control_2018;@stec_ribosome_2016;@von_walden_ribosome_2019].
+Therefore, investigating the effects of combined RT and glucose ingestion may provide valuable insight into the potential additive effect of glucose and RT on ribosome biogenesis.
 
@@
 Importantly, mTOR and its downstream target S6K1 serve as a mediator from insulin signalling in skeletal muscle [@hillier_physiological_2000].
-The mTORC1 pathway contributes to ribosome biogenesis through the regulation of translation of ribosomal proteins and to forming of the preinitiation complex (PIC) that marks the initiation of rRNA transcription [@figueiredo_regulation_2019;@von_walden_ribosome_2019].
+The mTORC1 pathway contributes to ribosome biogenesis through the regulation of the translation of ribosomal proteins and to forming of the preinitiation complex (PIC) that marks the initiation of rRNA transcription [@figueiredo_regulation_2019;@von_walden_ribosome_2019].
 Parallel to mTORC1 [@mori_c-myc_2021; @west_acute_2016], c-Myc increases ribosomal biogenesis directly and through transcriptional control of the upstream binding factor (UBF) [@sanij_ubf_2008; @poortinga_c-myc_2011].
@@
 The purpose of this investigation was to test the hypothesis that glucose supplementation before and after<!-- during?
-Kristian: before and after would be more precise, since they're not strictly ingesting anything during RT - I got a comment on that during the thesis defense --> RT increasea markers of ribosomal abundance following five RT sessions.
-Secondly we aimed to describe the association between UBF and total RNA abundance in human skeletal muscle.
-There has been multiple studies the last years suggesting translational capacity to be as important as, if not more, translational efficiency in regard to long-term skeletal muscle adaptations to RT [@hammarstrom_benefits_2020;@hammarstrom_ribosome_2022;@figueiredo_revisiting_2019].
-Previously, *in vitro* studies has shown UBF and c-Myc to be a important factors in the PIC, c-Myc as a general transcription factor and UBF as a specific transcription factor to rDNA transcription initiation [@mariappan_ribosomal_2011;@von_walden_ribosome_2019].
+Kristian: before and after would be more precise, since they're not strictly ingesting anything during RT - I got a comment on that during the thesis defense --> RT increases markers of ribosomal abundance following five RT sessions.
+Secondly, we aimed to describe the association between UBF and total RNA abundance in human skeletal muscle.
+There have been multiple studies in the last years suggesting translational capacity to be as important as, if not more than, translational efficiency in regard to long-term skeletal muscle adaptations to RT [@hammarstrom_benefits_2020;@hammarstrom_ribosome_2022;@figueiredo_revisiting_2019].
+Previously, *in vitro* studies have shown UBF and c-Myc to be important factors in the PIC, c-Myc as a general transcription factor and UBF as a specific transcription factor to rDNA transcription initiation [@mariappan_ribosomal_2011;@von_walden_ribosome_2019].
 Coupled with the recent observations by Hammarström et al.
-[@hammarstrom_benefits_2020;@hammarstrom_ribosome_2022], understanding underlying factors regulating and affecting ribosome biogenesis seem key to further our understanding in optimization of RT to individual phenotypes.
+[@hammarstrom_benefits_2020;@hammarstrom_ribosome_2022], understanding underlying factors regulating and affecting ribosome biogenesis seems key to furthering our understanding of the optimization of RT to individual phenotypes.
 <!-- We should maybe add a line or two about the rational here. Kristian: Tried to add some, what about this? to general? -->
@@
 2-8 RT sessions per 14 days for the last six months).
-Exclusion criteria were previous injury leading to impaired strength, inability to perform resistance exercise training, symptoms and a medical record of metabolic disorders including hyperglycemia, i.e.
+Exclusion criteria were previous injury leading to impaired strength, inability to perform resistance exercise training, symptoms, and a medical record of metabolic disorders including hyperglycemia, i.e.
 fasting venous plasma glucose ≥6.1 mmol/L and/or 2-hour glucose tolerance ≥7.8 mmol/L, and/or HbA1c >42 mmol/mol.
@@
 <!--from the "old" methods: either their dominant or non-dominant leg with glucose or placebo before and after RT, and whether they started RT with glucose or placebo.
-Thus, participants exercised every day, alternating between e.g., one leg + glucose and the other leg + placebo - this might be TMI, as it's presented in the figure as well?--> One person was exclusively responsible for the randomization code and supplement distribution, blinding both investigators and participants in regard to which leg exercised with glucose and placebo.
+Thus, participants exercised every day, alternating between e.g., one leg + glucose and the other leg + placebo - this might be TMI, as it's presented in the figure as well?--> One person was exclusively responsible for the randomization code and supplement distribution, blinding both investigators and participants in regard to which leg exercised with glucose/placebo.
 Glucose (Glucosum monohydricum, Merck KGaA, Darmstadt, Germany) and placebo (Steviosa, Soma Nordic AS, Oslo, Norway) were masked by mixing with 300ml Fun Light juice/saft (Orkla, Oslo, Norway).
@@
 
-Six RT sessions were conducted with glucose and six with placebo, allowing a within-subjects analysis of the effects of glucose ingestion before and after RT.
+All participants completed six RT sessions with glucose and six with placebo, allowing a within-subjects analysis of the effects of glucose ingestion before and after RT.
 Data from the first five RT sessions was used to investigate main outcome measures (total RNA, rRNA and protein) and leg muscle strength, whereas data from the sixth RT session was used to explore secondary outcomes (muscular recovery, plasma glucose and -c-peptide levels).
@@
 During period I, participants ingested protein and glucose/placebo only.
-Glucose/placebo was ingested at three time points in period I: 30 minutes before RT (0830hrs, 30g vs 0g glucose), immediately before RT (0900hrs, 30g vs.
+Glucose/placebo was ingested three times in period I: 30 minutes before RT (0830hrs, 30g vs 0g glucose), immediately before RT (0900hrs, 30g vs.
 0g glucose), and immediately after RT (~0930hrs, 30g vs.
@@
 The daily onset of the dietary intervention (i.e., first ingestion of PRO supplement) varied between participants, from 0600hrs to 0900hrs to allow multiple participants to complete the protocol simultaneously.
-The protein supplement was ingested as 25g Whey Protein Isolate boluses (Proteinfabrikken, Stokke, Norway), diluted in 150ml of water.
+The protein supplement was ingested as 25g Whey Protein Isolate boluses (Proteinfabrikken, Stokke, Norway), diluted in 150 ml of water.
 During sessions, participants were free to ingest water ad libitum.
@@
 
-Strength tests were performed prior to (Days -7 and -5, and -1, both legs) and during the intervention (Days 4 and 8 leg 1, 5 and 9 leg two), after session 5 and after finalization of the intervention (Day 11/12 leg 1 and 12/13 leg 2) (Fig 1A).
-Maximal isometric and isokinetic knee extension torque was measured with a Humac Norm Dynamometer (CSMi, Stoughton, Massechusetts, USA).
+Strength tests were performed prior to (Days -7 and -5, and -1, both legs) and during the intervention (Days 4 and 8 leg 1, 5 and 9 leg two), after session 5 and after finalization of intervention (Day 11/12 leg 1 and 12/13 leg 2) (Fig 1A).
+Maximal isometric and isokinetic knee extension torque was measured with a Humac Norm Dynamometer (CSMi, Stoughton, Massachusetts, USA).
 Individual positions were recorded and standardized from pre-intervention tests (days -7 and -5).
 Isokinetic peak torque was measured at speeds of 60- and 240 d/s, 2x3 repetitions each, with the first set of each exercise as a sub-maximal warm-up.
-Isometric peak torque was measured at knee-angle 60, for a maximum of 10 seconds and two repetition per test.
+Isometric peak torque was measured at knee-angle 60, for a maximum of 10 seconds and two repetitions per test.
 The highest peak torque values were used in the final analyses.
 During days 4, 5, 8 and 9 (Fig 1D, day 4 and 5 = Post 2RT, day 8 and 9 = Post 4RT), humac tests were conducted one hour before RT with the leg performing RT the previous day.
-During days 11 and 12, humac tests were performed at four time points: I) 45min before RT (1D, Post 5RT), II) 30min after the last RT session (1D, 30 min post 6RT), III) 2hrs after the last RT session (1D, 2h post 6RT), and IV) 23hrs after the last RT session (1D, 23h post 6RT).
-Test I at day 12/13 included testing of both legs, representing 23hrs post-RT session test of one leg and post-session 5 test of the other leg.
+During days 11 and 12, humac tests were performed four times: I) 45min before the last RT session (1D, Post 5RT), II) 30min after the last RT session (1D, 30 min post 6RT), III) 2hrs after the last RT session (1D, 2h post 6RT), and IV) 23hrs after the last RT session (1D, 23h post 6RT).
+Test I on day 12/13 included testing of both legs, representing 23hrs post-RT session test of one leg and post-session 5 test of the other leg.
 
@@
 The participants performed a general warm-up with 10min cycling on an indoor exercise bicycle.
-In addition, a protocol consisting of 1x10, 1x6 and 1x3 repetitions, was used as specific warm-up before each of the tests.
+In addition, a protocol consisting of 1x10, 1x6 and 1x3 repetitions, was used as a specific warm-up before each of the tests.
 All positions were controlled and recorded at the first 1RM test and repeated for the RT sessions.
@@
 Attempts with exaggerated hip movement or beneath full extension were not approved.
-Two minute breaks were given during specific warm-up, 3min breaks were given between 1RM attempts.
+Two-minute breaks were given during specific warm-up, three-minute breaks were given between 1RM attempts.
 
@@
 In addition, before the respective exercises, two 10-repetition warm-up sets were completed at ~50% and ~70% of 10RM.
-Breaks between working sets were 2min, and progressive loading was used to ensure adequate exercise stimulation throughout the intervention.
+Breaks between working sets were two minutes, and progressive loading was used to ensure adequate exercise stimulation throughout the intervention.
 For safety and standardization purposes, all sessions were monitored by trained personnel.
@@
 Briefly, muscle biopsy sampling was performed under local anaesthesia (Xylocaine, 10mg ml-1 with adrenaline 5μg ml−1, AstraZeneca AS, Oslo, Norway) using a 12-14-gauge needle (Universal Plus, Mermaid Medical AS, Stenløse, Denmark), operated with a spring-loaded biopsy gun.
-After the biopsy sampling, muscle tissue was divided in two aliquots for determination of total RNA/expression of rRNA and two aliquots for protein content measurement, snap frozen in isopentane (-80°C) and stored at -80°C until further analyses.
-Muscle biopsies were sampled at four time points: I/II) Pre-intervention (Fig 1A, 2hrs before training, Day 1 = leg 1, Day 2 = leg 2), and III/IV) after fifth RT session (1A, Day 11 = leg 1, Day 12 = leg 2), two hours before sixth RT session.
+After the biopsy sampling, muscle tissue was divided into two aliquots for determination of total RNA/expression of rRNA and two aliquots for protein content measurement, snap frozen in isopentane (-80°C) and stored at -80°C until further analyses.
+Muscle biopsies were sampled at four times: I/II) Pre-intervention (Fig 1A, 2hrs before training, Day 1 = leg 1, Day 2 = leg 2), and III/IV) after the fifth RT session, two hours before the sixth RT session (1A, Day 11 = leg 1, Day 12 = leg 2).
 At each time point, two samples were taken from the same incision.
@@
 To measure blood glucose levels with and without glucose ingestion/training, blood was collected by means of finger draws on days with biopsy sampling.
-One finger draw was taken at day 1 (Fig 1A) to serve as a baseline.
-At day 11 and 12 (Fig 1A), finger draws were collected at 7 time points: I) Immediately prior to protein ingestion (0700hrs) II) 45 minutes after protein ingestion (0745hrs) III) 1.5hrs after protein ingestion (0800hrs, i.e., immediately before GLU/PLAC intake), IIII) 2hrs after protein ingestion (0900hrs, i.e., immediately before training), IV) in the middle of RT (~0915hrs), V) immediately after training (~0930hrs), and VI) 2hrs after completion of training (~1130hrs).
+One finger draw was taken on day 1 (Fig 1A) to serve as a baseline.
+On days 11 and 12 (Fig 1A), finger draws were collected seven times: I) Immediately prior to protein ingestion (0700hrs) II) 45 minutes after protein ingestion (0745hrs) III) 1.5hrs after protein ingestion (0800hrs, i.e., immediately before GLU/PLAC intake), IIII) 2hrs after protein ingestion (0900hrs, i.e., immediately before training), IV) in the middle of RT (~0915hrs), V) immediately after training (~0930hrs), and VI) 2hrs after completion of training (~1130hrs).
 Finger draws were analysed with in-house equipment (BIOSEN C-Line, EKF diagnostic GmbH, Barleben).
@@
 Briefly, muscle tissue was homogenized in TRIzol using 0.5mm RNase-free Zirconium beads (~50 ul) (Next Advanced, Averill Park, NY, USA) and a bullet blender (bb).
-Chloroform (Sigma-Aldrich, Missouri, USA) was used for phase-separation, and an RNA pellet was precipitated with isopropanol (VWR International, Pennsylvania, USA).
+Chloroform (Sigma-Aldrich, Missouri, USA) was used for phase separation, and an RNA pellet was precipitated with isopropanol (VWR International, Pennsylvania, USA).
 For assessment of RNA content and -purity 5µl RNA, from the RNA stock, was eluted in TE-buffer (1:2), for assessment via spectrophotometry.
@@
 The tissue was homogenized with a plastic rod in 80mg protein extraction powder (Invent Biotechnology) and 100ul iced cell lysis buffer (Denaturing Buffer, Invent Biotechnology), and centrifuged at top speed for 1 min.
-The supernatant was devided into 2x25ul aliquots to run samples in duplicates, and total protein concentrations were determined in a 1:10 dilution (Pierce Detergent Compatible Bradford Assay Reagent, Thermo Fisher Scientific, Oslo, Norway).
+The supernatant was divided into 2x25ul aliquots to run samples in duplicates, and total protein concentrations were determined in a 1:10 dilution (Pierce Detergent Compatible Bradford Assay Reagent, Thermo Fisher Scientific, Oslo, Norway).
 The protein samples were diluted to 20.25 µg µl−1 total protein with lysis buffer and 4X Laemmli sample buffer (Bio-Rad Laboratories AB, Oslo, Norway) containing 2-mercaptoethanol.
@@
 This power calculation was based on an assumption that the effects of glucose ingestion on ribosome biogenesis may equate to the effects of increasing RT volume from low to moderate [@hammarstrom_benefits_2020].
-<!-- maybe include the power calculation in extended methods?--> Raw data was imported and analyzed in R via R studios [@Rcore].
+<!-- maybe include the power calculation in extended methods?--> Raw data were imported and analyzed in R via R studios [@Rcore].
 Legs were categorized as either glucose or placebo in accordance with the randomization.
-Before quantitative analyses, total RNA was normalized by wet muscle weight, qPCR values were normalized by wet muscle weight and an external reference gene (Lambda), while quantified protein signals from western blot were normalized by pool signals and total protein signal per gel <!-- add explanations of normalization to extended methods -->.
-Total RNA and qPCR values were analyzed by comparing post to pre log-fold change score differences.
-This was achieved by modelling change as response with baseline and supplement (glucose/placebo) as explanatory factors, and mean-centered baseline values as a covariate in a linear mixed effects model using, then calculating the estimated marginal means.
+Before quantitative analyses, total RNA was normalized by wet muscle weight, qPCR values were normalized by wet muscle weight and an external reference gene (Lambda), while quantified protein signals from western blot were normalized by pool signals and total protein signal per gel.<!-- add explanations of normalization to extended methods --> Total RNA and qPCR values were analyzed by comparing post to pre log-fold change score differences.
+This was achieved by modelling change as response with baseline and supplement (glucose/placebo) as explanatory factors, and mean-centred baseline values as a covariate in a linear mixed effects model using, then calculating the estimated marginal means.
 Quantified western blot signals and the UBF/total RNA association were analyzed using a linear mixed effects model from the nlme package [@lindstrom_nonlinear_1990], with protein signal per time point as response, with time and supplement as explanatory factors.
@@
 The linear mixed effects model was designed with the lmer function (Satterthwaite's method) from the lme4 package [@bates_fitting_2014] in R studios, using the lmerTest function to procure p-values.
-Log-transformation and mean-centering baseline values was done to control for heteroscedasticity and regression to the mean, respectively.
+Log-transformation and mean-centring baseline values were done to control for heteroscedasticity and regression to the mean, respectively.
 Log-transformed values were reverse-transformed for figure illustration.
-Values are presented as mean change + 95% CL unless otherwise stated, alpha-levels was set to α = 0.05.
+Values are presented as mean change + 95% CL unless otherwise stated, with the alpha-level set to α = 0.05.
 
@@
 There was no difference in mean change of total training session volume between exercising with placebo or glucose, with an increase of 18 and 16% from session 1 to session 6, respectively (Tab 2; Mean change, 2: *p* = `r pvol.c2`, 3: *p* = `r pvol.c3`, 4: *p* = `r pvol.c4`, 5: *p* = `r pvol.c5`, 6: *p* = `r pvol.c6`).
-Both exercising with glucose and placebo led to significant increases in mean total training session volume at session four, five and six, compared to session 1 (Tab 2; Mean change, 2: *p* = `r pvol.s2`, 3: *p* = `r pvol.s3`, 4: *p* = `r pvol.s4`, 5: *p* = `r pvol.s5`, 6: *p* = 0.001).
-In regard to the dietary intervention, there were no difference in mean macro nutrient intake (protein, fat, carbohydrates) or total calorie intake between glucose and placebo on pairwise consecutive days (*p* > 0.05).
+Both exercising with glucose and placebo led to significant increases in mean total training session volume at sessions four, five and six, compared to session 1 (Tab 2; Mean change, 2: *p* = `r pvol.s2`, 3: *p* = `r pvol.s3`, 4: *p* = `r pvol.s4`, 5: *p* = `r pvol.s5`, 6: *p* = 0.001).
+Regarding the dietary data, there were no differences in mean macronutrient intake (protein, fat, carbohydrates) or total calorie intake between glucose and placebo on pairwise consecutive days (*p* > 0.05).
 
@@
 p.150 <- sprintf("%.5f", 1.445712e-11)
-p.270 <- round(bcoef[11,5], 2)
+p.270 <- round(bcoef[11,5], 3)
 ```
@@
 
-Glucose ingestion before and after RT led to significant increases in plasma levels of glucose by 38% immediately before RT (Fig 1B, 0 min: *p* = `r p.120`), by 31% during RT (1B, 15 min: *p* = `r p.135`) and by 32% immediately after RT (1B, 30 min: *p* = `r p.150`), compared to placebo ingestion before and after RT.
+Glucose ingestion before and after RT led to significant increases in plasma levels of glucose by 38% immediately before RT (Fig 1B, 0 min: *p* = 0.000), by 31% during RT (1B, 15 min: *p* = 0.000) and by 32% immediately after RT (1B, 30 min: *p* = 0.000), compared to placebo ingestion before and after RT.
 Two hours after the session, RT with glucose measured 8% lower plasma glucose levels compared to RT with placebo (1B, 270min: *p* = `r p.270`).
-Similarly, levels of c-peptide increased significantly more in exercises with glucose compared to placebo, by 95% immediately before (Fig 1C, 0 min: *p* = `r cpp.120`) and 87% after RT (1C, 30 min: *p* = `r cpp.150`).
+Similarly, levels of c-peptide increased significantly more in exercises with glucose compared to placebo, by 95% immediately before (Fig 1C, 0 min: *p* = 0.000) and 87% after RT (1C, 30 min: *p* = 0.000).
 
@@
 
-After 5 RT sessions (Fig 1D, Post 5RT), there was a significant difference in strength between glucose and placebo,  where placebo decreased strength 7% more than glucose (1D, *p* = `r pdiff.c4`).
-At the remaining time points, there were no significant differences in mean change between exercising with glucose and placebo (1D, Post 2RT: *p* = `r pdiff.c2`, Post 4RT: *p* = `r pdiff.c3`, 30min post 6RT: *p* = `r pdiff.c5`, 2h post 6RT: *p* = `r pdiff.c6`, 23h post 6RT: *p* = `r pdiff.c7`).
-Both RT with glucose and placebo led to significantly reduced strength post fifth session compared to baseline, by 11 and 18% respectively (1D, Post 5RT: *p* = `r pc4`).
-Comparisons of the acute data gathered from post fifth session until and including 23hrs post sixth session showed a significant mean increase in strength of 5-9% after RT with glucose and placebo 30min after sixth RT session (1D, 30min post 6RT: *p* = `r apc2`) and two hours after sixth RT session (1D, 2h post 6RT: *p* = `r apc3`).
-Twenty-three hours after the last RT, strength was unchanged compared to after fifth RT session (23h post 6RT: *p* = `r apc4`).
+After 5 RT sessions (Fig 1D, Post 5RT), there was a significant difference in strength between glucose and placebo,  where placebo decreased strength by 7% more than glucose (1D, *p* = `r pdiff.c4`).
+At the other time points, there were no significant differences in mean change between exercising with glucose and placebo (1D, Post 2RT: *p* = `r pdiff.c2`, Post 4RT: *p* = `r pdiff.c3`, 30min post 6RT: *p* = `r pdiff.c5`, 2h post 6RT: *p* = `r pdiff.c6`, 23h post 6RT: *p* = `r pdiff.c7`).
+Both RT with glucose and placebo led to significantly reduced strength post-fifth session compared to baseline, by 11 and 18% respectively (1D, Post 5RT: *p* = 0.000).
+Comparisons of the acute data gathered from the post-fifth session until and including 23hrs post-sixth session showed a significant mean increase in strength of 5-9% after RT with glucose and placebo 30min after the sixth RT session (1D, 30min post 6RT: *p* = `r apc2`) and two hours after the sixth RT session (1D, 2h post 6RT: *p* = `r apc3`).
+Twenty-three hours after the last sixth RT session, strength was unchanged compared to after the fifth RT session (23h post 6RT: *p* = `r apc4`).
 
@@
 
-The main findings of the present study were that heavy resistance training with glucose did not affect markers of ribosome biogenesis, measured as total RNA, rRNA and rDNA transcription initiation associated proteins, compared to RT with placebo.
+The main findings of the present study were that heavy resistance training with glucose did not affect markers of ribosome biogenesis, measured as total RNA, rRNA and rDNA transcription initiation-associated proteins, compared to RT with placebo.
 Similarly, RT with glucose in general did not affect markers of skeletal muscle functionality such as strength and recovery, or total training session volume.
 At one time point, RT with glucose led to less reduction in strength compared to placebo, and levels of plasma glucose and c-peptide were significantly higher before, during and after the RT sessions with glucose.
-Further, there were no differences in daily macronutrient intake on pairwise consecutive days.
-This would suggest the effects of high vs.
+Importantly, there were no differences in daily macronutrient intake on pairwise consecutive days.
+These findings suggest that the effects of high vs.
 low glucose/glucose starvation treatment on rDNA transcription initiation observed in previous *in vitro* studies [@mariappan_ribosomal_2011; @tanaka_mild_2015] are not translatable to acute effects in human skeletal muscle *in vivo*.
-Notably, the present study was specifically designed to investigate differences in pre-post changes within-participants, and thus also quite limited to this perspective.
-Hence, the observed effects of time serve as supplementary to the efficiency of the design, without a negative control group to compare.
+Notably, the present study was specifically designed to investigate differences in pre-post changes within participants, and is as such also quite limited to this perspective.
+Hence, the observed effects of time serve as supplementary to the efficiency of the design, without a negative control group to compare to.
 
@@
 placebo treatment.
-Resistance training, irrespective of treatment, yielded robust accumulation of total RNA and expression of rRNA, in line with previous observations [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022].
-Due to the exploratory nature of our hypotheses this was hardly a very surprising result.
-Indeed, previous studies has used cell cultures from yeast [@zhai_human_2012], rodent [@hoppe_amp-activated_2009; @mariappan_ribosomal_2011] or human cells [@tanaka_mild_2015] and are as such not directly comparable to human skeletal muscle cells.
+Resistance training, irrespective of treatment, yielded a robust accumulation of total RNA and expression of rRNA, in line with previous observations [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022].
+Due to the exploratory nature of our hypotheses, this was hardly a very surprising result.
+Indeed, previous studies have used cell cultures from yeast [@zhai_human_2012], rodent [@hoppe_amp-activated_2009; @mariappan_ribosomal_2011] or human cells [@tanaka_mild_2015] and are as such not directly comparable to human skeletal muscle cells.
 Importantly, previous studies investigated high vs.
 low glucose treatment [@mariappan_ribosomal_2011], or high glucose vs.
-glucose starvation [@hoppe_amp-activated_2009; @tanaka_mild_2015], while the present study aimed to compare high glucose treatment to placebo, with a matched daily macro nutrient and energy intake.
-And so, the comparison of the present study was high glucose vs.
-normal glucose treatment, to investigate the effect of glucose *per se* and not intracellular energy status.
-Thus, while glucose ingestion presumably is important in supplying energy for growth inducing processes such as ribosome biogenesis [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @sebata_control_2018] there is no apparent effect of ingesting glucose *per se* on markers of ribosome biogenesis.
+glucose starvation [@hoppe_amp-activated_2009; @tanaka_mild_2015], while the present study aimed to compare high glucose treatment to placebo, with a matched daily macronutrient and energy intake.
+Therefore, the comparison made in the present study was high plasma glucose levels vs.
+normal plasma glucose levels, to investigate the effect of glucose *per se* and not intracellular energy status.
+Thus, while glucose ingestion presumably is important in supplying energy for growth-inducing processes such as ribosome biogenesis [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @sebata_control_2018] there is no apparent effect of ingesting glucose *per se* on markers of ribosome biogenesis.
 
@@
 UBF has previously been described as a master regulator of rDNA transcription *in vitro* [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @russell_rna-polymerase-i-directed_2005], while rpS6 is proposed as a valid and reliable means to measure ribosome biogenesis [@chaillou_hypoxia_2012; @nakada_correlation_2016].
-However, content of c-Myc did differ between RT with glucose and RT with placebo, whereas RT with placebo led to significantly higher post content of c-Myc compared to glucose.
+However, content of c-Myc did differ between RT with glucose and RT with placebo, whereas RT with placebo led to significantly higher c-Myc content at the post measurement, compared to glucose.
 The general transcription factor c-Myc has previously been suggested as a potent regulator of ribosome biogenesis independent of mTORC1, and a direct regulator of UBF [@mori_c-myc_2021; @poortinga_c-myc_2011; @west_acute_2016].
-The independency on mTORC1 may explain why only c-Myc content differed between treatments, while UBF and rpS6 content changed equally.
-However,  Moreover, the observation that total RNA and rRNA was unaffected by the significant differences in c-Myc content, do strengthen the importance and specificity of UBF content in human skeletal muscle ribosome biogenesis.
-The linear relationship found exclusively between UBF content and total RNA accumulation, and not c-Myc or rpS6, found in the present study further supports this notion.
+The independency of mTORC1 may explain why only c-Myc content differed between treatments, while UBF and rpS6 content changed equally.
+Moreover, the observation that total RNA and rRNA were unaffected by the significant differences in c-Myc content, does strengthen the importance and specificity of UBF content in human skeletal muscle ribosome biogenesis.
+The linear relationship found exclusively between UBF content and total RNA accumulation, and not c-Myc or rpS6, found in the present study also support this notion.
 While this is not a novel finding, it has only recently been observed in human skeletal muscle following RT [@hammarstrom_ribosome_2022].
-As such, UBF also seem to respond to mechanical loading in human muscle cells in line with responses seen in cell cultures and synergist ablation models [@mariappan_ribosomal_2011; @von_walden_mechanical_2012], in addition to the previously reported response to hormonal, nutritional, and cellular energy signals [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @russell_rna-polymerase-i-directed_2005].
+As such, UBF also seems to respond to mechanical loading in human muscle cells in line with responses seen in cell cultures and synergist ablation models [@mariappan_ribosomal_2011; @von_walden_mechanical_2012], in addition to the previously reported response to hormonal, nutritional, and cellular energy signals [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @russell_rna-polymerase-i-directed_2005].
 
-As in the biological data, there was generally no difference in change of strength as an index between the treatments.
-Generally, skeletal muscle strength decreased from baseline to post fifth and sixth RT session, with a significantly less reduction in glucose compared to placebo at post fifth session measurement.
-This could suggest that the RT protocol fatigued the participants, as both training frequency and intensity was quite high.
+As in the biological data, there was generally no difference in the change of strength as an index between the treatments.
+Generally, skeletal muscle strength decreased from baseline to post-fifth and -sixth RT sessions, with a significantly less reduction in glucose compared to placebo at post-fifth session measurement.
+This could suggest that the RT protocol fatigued the participants, as both training frequency and intensity were quite high.
 Another promising explanation might be the biphasic recovery pattern, as described by Raastad & Hallén [@raastad_recovery_2000].
-In their study on recovery of skeletal muscle contractility after strength exercise, Raastad and Hallén (2000) found a rapid recovery during the initial 11hrs post-exercise, followed by either a leveling off or a drop until 22hrs post-exercise, returning to baseline values at 33hrs post-exercise.
-Herein, inflammation and phagocytic activity was proposed to be involved in th performance drop between 11-22hrs [@raastad_recovery_2000].
-Indeed, this pattern seems quite similar to what was observed in the present study, with a rapid recovery at 30min and 2hrs post sixth RT session and a drop at 23hrs.
-Moreover, the strength measurements during the intervention and after five RT sessions was also conducted approximately 23hrs after RT.
-Thus, it is possible that these strength measurements also was affected by the biphasic recovery pattern.
-However, this does not explain the difference between treatments at post fifth RT session.
-A possible argument could be that exercising without glucose may have caused more stress compared to exercising with glucose, as glucose is the preferred energy source during strenuous exercise [@mul_exercise_2015], thus increased performance in RT with glucose compared to RT with placebo.
-This may also, at least partly, explain significantly higher content of the stress responsive c-Myc in RT with placebo compared to glucose [@mori_c-myc_2021; @poortinga_c-myc_2011; @west_acute_2016].
+In their study on recovery of skeletal muscle contractility after strength exercise, Raastad and Hallén (2000) found a rapid recovery during the initial 11hrs post-exercise, followed by either a levelling off or a drop until 22hrs post-exercise, returning to baseline values at 33hrs post-exercise.
+Herein, inflammation and phagocytic activity were proposed to be involved in the performance drop between 11-22hrs [@raastad_recovery_2000].
+Indeed, this pattern seems quite similar to what was observed in the present study, with a rapid recovery at 30min and 2hrs post-sixth RT session and a drop at 23hrs.
+Moreover, the strength measurements during the intervention and after five RT sessions were also conducted approximately 23hrs after RT.
+Thus, it is possible that these strength measurements reflect a biphasic recovery pattern.
+However, this does not explain the difference between treatments at post-fifth RT session.
+A possible argument could be that exercising without glucose may have caused more stress compared to exercising with glucose, as glucose is the preferred energy source during strenuous exercise [@mul_exercise_2015], thus increasing performance in RT with glucose compared to RT with placebo.
+This may also, at least partly, explain the significantly higher content of the stress-responsive c-Myc in RT with placebo compared to glucose [@mori_c-myc_2021; @poortinga_c-myc_2011; @west_acute_2016].
 However, training volume data do reveal that the total training session volume was equal on pairwise consecutive days, i.e.
-no difference between day 1-2, day 3-4 and so on.
+no difference between days 1-2, days 3-4 and so on.
 Thus, there were no differences in mechanical loading to induce higher stress in RT with placebo compared to RT with glucose.
-Arguably, the significantly higher plasma levels of glucose in RT with glucose may have increased energy availability and thus, ATP, resulting in a lower metabolic stress and less performance reduction, compared to RT with placebo [@westerblad_mechanisms_1998; @kent-braun_central_1999] Unfortunately, we did not conduct measurements of markers of metabolic stress; inorganic phosphate, H+, Mg2+ and the ADP/ATP ratio [@westerblad_mechanisms_1998; @kent-braun_central_1999], thus, discussing the potential effect of differences in metabolic stress between treatments would only be speculation, however possible.
+Arguably, the significantly higher plasma levels of glucose in RT with glucose may have increased energy availability and thus, ATP, resulting in lower metabolic stress and less performance reduction, compared to RT with placebo [@westerblad_mechanisms_1998; @kent-braun_central_1999] Unfortunately, we did not conduct measurements of markers of metabolic stress such ass inorganic phosphate, H+, Mg2+ and the ADP/ATP ratio [@westerblad_mechanisms_1998; @kent-braun_central_1999].
+Therefore, discussing the potential effect of differences in metabolic stress between treatments would only be speculation, however probable.
 Lastly, these inconclusive observations may also reflect our limited sample size, despite using a within-subjects design.
-Initially, we met our minimum limit of 16 participants however, three dropped out during the intervention, leaving us slightly underpowered according to the a priori calculation.
+Initially, we met our minimum limit of 16 participants.
+Unfortunately, three dropped out during the intervention, leaving our statistical power slightly underpowered according to the a priori calculation.
 
-In conclusion, ingestion of glucose immediately before and after resistance exercise training does not acutely augment ribosome biogenesis after two weeks of moderate volume exercise, in moderately trained young adults.
-Neither does it acutely enhance muscular performance during the exercise period nor after five sessions, or recovery within 23hrs of the last session.
+In conclusion, ingestion of glucose immediately before and after resistance exercise training does not acutely augment ribosome biogenesis after two weeks of heavy resistance exercise training, in moderately trained young adults.
+Neither does it acutely enhance muscular performance during the exercise period or after five sessions, nor recovery within 23hrs of the last session.
 The observations in total RNA, rRNA and rDNA transcription initiation associated proteins support the key role of UBF in ribosome biogenesis regulation in human skeletal muscle following resistance training.
-If future investigations are to be made on this topic, a greater sample size coupled with a negative control group and analyses of the intramuscular glycogen storage may provide higher resolution results.
+If future investigations are to be made on this topic, a greater sample size coupled with a negative control group and analyses of the intramuscular glycogen storage may provide higher-resolution results.
 

### 2023-03-31 ribose-paper.postfeedback.RMD
updating after feedback

--- 2023-03-30 ribose-paper.RMD
@@
 Such inability to benefit from RT may be explained by genetic and/or epigenetic disposition affecting the internal physiological milieu and subsequently affecting the ability to mount an anabolic response to RT [@figueiredo_genetic_2021;@thalacker-mercer_cluster_2013].<!-- Check these refs, do they correspond to genet epigen and internal mileu?
-Kristian: The way I read it, it looks like it corresponds --> Although RT program modifications such as altered exercise volume or intensity may induce more favourable adaptations in some individuals, such benefits are not uniform between individuals [@hammarstrom_benefits_2020;@krieger_single_2009;@schoenfeld_strength_2017].<!-- Add ref for more rt program modifications.
+Kristian: The way I read it, it looks like it corresponds --> Although RT program modifications such as altered exercise volume or intensity may induce more favourable adaptations in some individuals, these benefits are not uniform between individuals [@hammarstrom_benefits_2020;@krieger_single_2009;@schoenfeld_strength_2017].<!-- Add ref for more rt program modifications.
 Kristian: Added krieger et al.
@@
 This is surprising since glucose is the preferred energy substrate of the contracting skeletal muscle during strenuous exercise and a major supplier of energy to cells via ATP synthesis [@mul_exercise_2015;@sebata_control_2018].
-Furthermore, energy availability is a decisive factor in *de novo* synthesis of ribosomes [@kusnadi_regulation_2015;@moss_housekeeper_2007;@sebata_control_2018] <!--Add ref that indicate high energy requirement for ribo biog possibly (VONWALDEN19;TANAKA18?
+Furthermore, energy availability is a decisive factor in the*de novo* synthesis of ribosomes [@kusnadi_regulation_2015;@moss_housekeeper_2007;@sebata_control_2018] <!--Add ref that indicate high energy requirement for ribo biog possibly (VONWALDEN19;TANAKA18?
 MOSS.
@@
 Secondly, we aimed to describe the association between UBF and total RNA abundance in human skeletal muscle.
-There have been multiple studies in the last years suggesting translational capacity to be as important as, if not more than, translational efficiency in regard to long-term skeletal muscle adaptations to RT [@hammarstrom_benefits_2020;@hammarstrom_ribosome_2022;@figueiredo_revisiting_2019].
-Previously, *in vitro* studies have shown UBF and c-Myc to be important factors in the PIC, c-Myc as a general transcription factor and UBF as a specific transcription factor to rDNA transcription initiation [@mariappan_ribosomal_2011;@von_walden_ribosome_2019].
+There have been multiple studies in the last years suggesting translational capacity to be as important, if not more, as translational efficiency in regard to long-term skeletal muscle adaptations to RT [@hammarstrom_benefits_2020;@hammarstrom_ribosome_2022;@figueiredo_revisiting_2019].
+Previously, *in vitro* studies have shown UBF and c-Myc to be important factors in the PIC, c-Myc as a general transcription factor and UBF as a specific transcription factor for rDNA transcription initiation [@mariappan_ribosomal_2011;@von_walden_ribosome_2019].
 Coupled with the recent observations by Hammarström et al.
@@
 
-Sixteen healthy male and female participants (20-33yrs, Table 1) were recruited to the study through social media advertisement and word of mouth.
+Sixteen healthy male and female participants (20-33yrs, Tab 1) were recruited to the study through social media advertisement and word of mouth.
 The eligibility criteria were non-smokers and moderately trained (i.e.
@@
 Of the sixteen participants that commenced the intervention, three participants dropped out.
-One due to sickness and inability to resume, two participants experienced muscular discomfort connected to heavy resistance training.
+One due to sickness and inability to resume, two participants experienced muscular discomfort related to heavy resistance training.
 Lean mass (Tab 1) was measured using DXA (Prodigy Advance PA+302047, Lunar, San Francisco, CA, USA) on Day -1, the last day preceding the RT intervention.
@@
 The study was designed as a 12-day double-blinded randomized controlled trial, with an alternating unilateral RT protocol (Fig 1A).
-Participants were randomly allocated to exercise one leg with glucose and one leg with placebo (1A).
+Participants were randomly allocated to exercise one leg with glucose and one leg with a placebo (Fig 1A).
 <!--from the "old" methods: either their dominant or non-dominant leg with glucose or placebo before and after RT, and whether they started RT with glucose or placebo.
-Thus, participants exercised every day, alternating between e.g., one leg + glucose and the other leg + placebo - this might be TMI, as it's presented in the figure as well?--> One person was exclusively responsible for the randomization code and supplement distribution, blinding both investigators and participants in regard to which leg exercised with glucose/placebo.
-Glucose (Glucosum monohydricum, Merck KGaA, Darmstadt, Germany) and placebo (Steviosa, Soma Nordic AS, Oslo, Norway) were masked by mixing with 300ml Fun Light juice/saft (Orkla, Oslo, Norway).
+Thus, participants exercised every day, alternating between e.g., one leg + glucose and the other leg + placebo - this might be TMI, as it's presented in the figure as well?--> One person was exclusively responsible for the randomization code and supplement distribution, blinding both investigators and participants regarding which leg exercised with glucose/placebo.
+Glucose (Glucosum monohydricum, Merck KGaA, Darmstadt, Germany) and placebo (Steviosa, Soma Nordic AS, Oslo, Norway) were masked by mixing with 300ml Fun Light (Orkla, Oslo, Norway).
 A blinded taste test revealed that the participants were not able to disclose the contents of the provided boluses (30g glucosum monohydricum vs.
 ~0.3g Stevia rebaudiana extract).
-To ensure equal conditions during training sessions and strength testing, participants exercised and tested at the same time of day, +/- 1hr with the same supervisor on pairwise consecutive days.
+To ensure equal conditions during training sessions and strength testing, participants exercised and tested at the same time of day, +/- 1hr with the same supervisor on pairwise consecutive days (i.e.
+on days 1-2, 3-4, etc.).
 To further standardize this, participants also recorded and repeated their daily macronutrient intake (protein, fat, carbohydrate) and total calories on pairwise consecutive days.
@@
 0g glucose).
-Protein was ingested 2hrs before RT (0700hrs, 25g) and immediately after RT (~0930hrs, 25g).
+Whey Protein Isolate (Proteinfabrikken, Stokke, Norway) was ingested 2hrs before RT (0700hrs) and immediately after RT (~0930hrs), in boluses of 25 g mixed with 150 ml water.
 In the afternoon (1800hrs-1900hrs, period II) participants ingested glucose or placebo (3x30g vs.
@@
 Apart from this, participants ingested a self-chosen diet during period II, registered in MyFitnessPal or similar applications.
-The self-chosen diet was repeated on pairwise consecutive days (i.e.
-on days 1-2, 3-4, etc.), to ensure similar premises for resistance training responses between treatments.
+The self-chosen diet was repeated on pairwise consecutive days, to ensure similar premises for resistance training responses between treatments.
 During period III (2200-0700hrs), participants remained in an overnight fasted state.
 The daily onset of the dietary intervention (i.e., first ingestion of PRO supplement) varied between participants, from 0600hrs to 0900hrs to allow multiple participants to complete the protocol simultaneously.
-The protein supplement was ingested as 25g Whey Protein Isolate boluses (Proteinfabrikken, Stokke, Norway), diluted in 150 ml of water.
 During sessions, participants were free to ingest water ad libitum.
@@
 
-Strength tests were performed prior to (Days -7 and -5, and -1, both legs) and during the intervention (Days 4 and 8 leg 1, 5 and 9 leg two), after session 5 and after finalization of intervention (Day 11/12 leg 1 and 12/13 leg 2) (Fig 1A).
+Strength tests were performed prior to (Days -7 and -5, and -1, both legs) and during the intervention (Fig 1A, on days 4 and 8 for leg 1, and days 5 and 9 for leg two), after session 5 and after finalization of intervention (Fig 1A, on days 11/12 for leg 1 and days 12/13 for leg 2).
 Maximal isometric and isokinetic knee extension torque was measured with a Humac Norm Dynamometer (CSMi, Stoughton, Massachusetts, USA).
-Individual positions were recorded and standardized from pre-intervention tests (days -7 and -5).
+Individual positions were recorded and standardized from pre-intervention tests (Fig 1A, days -7 and -5).
 Isokinetic peak torque was measured at speeds of 60- and 240 d/s, 2x3 repetitions each, with the first set of each exercise as a sub-maximal warm-up.
-Isometric peak torque was measured at knee-angle 60, for a maximum of 10 seconds and two repetitions per test.
+Isometric peak torque was measured at a knee-angle of 60 degrees (extended knee was set to 0 degrees), for a maximum of 10 seconds and two repetitions per test.
+The isometric tests were ended when the participants reached a plateau or peak torque development decreased, which on average occurred between 2-4 seconds into the test.
 The highest peak torque values were used in the final analyses.
-During days 4, 5, 8 and 9 (Fig 1D, day 4 and 5 = Post 2RT, day 8 and 9 = Post 4RT), humac tests were conducted one hour before RT with the leg performing RT the previous day.
-During days 11 and 12, humac tests were performed four times: I) 45min before the last RT session (1D, Post 5RT), II) 30min after the last RT session (1D, 30 min post 6RT), III) 2hrs after the last RT session (1D, 2h post 6RT), and IV) 23hrs after the last RT session (1D, 23h post 6RT).
+During days 4, 5, 8 and 9 (Fig 1D, days 4 and 5 = Post 2RT, days 8 and 9 = Post 4RT), humac tests were conducted one hour before RT on the leg performing RT the previous day.
+During days 11 and 12, humac tests were performed four times: I) 45min before the last RT session (Fig 1D, Post 5RT), II) 30min after the last RT session (Fig 1D, 30 min post 6RT), III) 2hrs after the last RT session (Fig 1D, 2h post 6RT), and IV) 23hrs after the last RT session (Fig 1D, 23h post 6RT).
 Test I on day 12/13 included testing of both legs, representing 23hrs post-RT session test of one leg and post-session 5 test of the other leg.
@@
 Attempts with exaggerated hip movement or beneath full extension were not approved.
-Two-minute breaks were given during specific warm-up, three-minute breaks were given between 1RM attempts.
+Two-minute of rest were given during the specific warm-up, and three-minute of rest was given between 1RM attempts.
 
-### Resistance training protocol
+CONTINUE HERE
+### Resistance training protocoL
 
@@
 Before baseline testing, the participants were randomly allocated to exercise one leg with GLU (glucose) and the other with PLA (placebo), in a unilateral, alternating fashion.
-Further, non-dominant/dominant + GLU/PLA, and onset with GLU or PLA was also randomized, i.e.
-the figure illustrates an example where the participant was randomized to start RT with GLU.
+Further, non-dominant/dominant + GLU/PLA, and onset with GLU or PLA was also randomized, the figure illustrates an example where the participant was randomized to start RT with GLU.
 Biopsies were taken from m.
@@
 from gel 3 to 6, as shown by the duplicates (1 and 2).
-Change in protein levels were calculated as log-fold change scores normalized by pools (pool of all protein samples per gel).
+Changes in protein levels were calculated as log-fold change scores normalized by pools (pool of all protein samples per gel).
 Mean log-change scores of the duplicates were calculated before modelling and reverse-transformed for figure illustration.
@@
 C) Linear relationship between total RNA (ng x mg) and UBF levels (SD units), with time added as a covariate.
-Total RNA was normalized by wet muscle weight, and UBF was normalized by pools per gel, and total protein per lane factor.
+Total RNA was normalized by wet muscle weight, and UBF was normalized by pools per gel and total protein per lane factor.
 Values are presented as log-transformed means.

### 2023-04-11 ribose-paper.postfeedback.RMD
updates

@@
 Individual positions were recorded and standardized from pre-intervention tests (Fig 1A, days -7 and -5).
-Isokinetic peak torque was measured at speeds of 60- and 240 d/s, 2x3 repetitions each, with the first set of each exercise as a sub-maximal warm-up.
+Isokinetic peak torque was measured at extensions speeds of 60- and 240 d/s, 2x3 repetitions each, with the first set of each exercise as a sub-maximal warm-up.
 Isometric peak torque was measured at a knee-angle of 60 degrees (extended knee was set to 0 degrees), for a maximum of 10 seconds and two repetitions per test.
 The isometric tests were ended when the participants reached a plateau or peak torque development decreased, which on average occurred between 2-4 seconds into the test.
-The highest peak torque values were used in the final analyses.
 During days 4, 5, 8 and 9 (Fig 1D, days 4 and 5 = Post 2RT, days 8 and 9 = Post 4RT), humac tests were conducted one hour before RT on the leg performing RT the previous day.
@@
 Test I on day 12/13 included testing of both legs, representing 23hrs post-RT session test of one leg and post-session 5 test of the other leg.
+The highest peak torque values from the respective extension speeds and time points were summarized in an index.
+The index was calculated by dividing the average peak torque value by the highest observed peak torque value per extension speed, and summarizing this new index per extension speed to a mean strength index.
 
@@
 Attempts with exaggerated hip movement or beneath full extension were not approved.
-Two-minute of rest were given during the specific warm-up, and three-minute of rest was given between 1RM attempts.
+Two minutes of rest were given during the specific warm-up, and three minutes of rest was given between 1RM attempts.
 
-CONTINUE HERE
 ### Resistance training protocoL
@@
 In addition, before the respective exercises, two 10-repetition warm-up sets were completed at ~50% and ~70% of 10RM.
-Breaks between working sets were two minutes, and progressive loading was used to ensure adequate exercise stimulation throughout the intervention.
+To ensure an adequate exercise stimulation throughout the intervention, the exercise load was increased if the participants lifted more than 12 repetitions, as a progressive loading strategy.
+Rest between working sets was two minutes.
 For safety and standardization purposes, all sessions were monitored by trained personnel.
@@
 vastus lateralis using well-established procedures [@hammarstrom_benefits_2020].
-Briefly, muscle biopsy sampling was performed under local anaesthesia (Xylocaine, 10mg ml-1 with adrenaline 5μg ml−1, AstraZeneca AS, Oslo, Norway) using a 12-14-gauge needle (Universal Plus, Mermaid Medical AS, Stenløse, Denmark), operated with a spring-loaded biopsy gun.
+Briefly, muscle biopsy sampling was performed under local anaesthesia (Xylocaine, 10mg ml-1 with adrenaline 5μg ml−1, AstraZeneca AS, Oslo, Norway) using a 12-14-gauge needle (Universal Plus, Mermaid Medical AS, Stenløse, Denmark), operated with a spring-loaded biopsy gun (Bard Magnum, Bard, Rud, Norway).
 After the biopsy sampling, muscle tissue was divided into two aliquots for determination of total RNA/expression of rRNA and two aliquots for protein content measurement, snap frozen in isopentane (-80°C) and stored at -80°C until further analyses.
@@
 One finger draw was taken on day 1 (Fig 1A) to serve as a baseline.
-On days 11 and 12 (Fig 1A), finger draws were collected seven times: I) Immediately prior to protein ingestion (0700hrs) II) 45 minutes after protein ingestion (0745hrs) III) 1.5hrs after protein ingestion (0800hrs, i.e., immediately before GLU/PLAC intake), IIII) 2hrs after protein ingestion (0900hrs, i.e., immediately before training), IV) in the middle of RT (~0915hrs), V) immediately after training (~0930hrs), and VI) 2hrs after completion of training (~1130hrs).
+On days 11 and 12 (Fig 1A), finger draws were collected seven times: I) Immediately prior to protein ingestion (0700hrs) II) 45 minutes after protein ingestion (0745hrs) III) 1.5hrs after protein ingestion (0800hrs, i.e., immediately before GLU/PLAC intake), IV) 2hrs after protein ingestion (0900hrs, i.e., immediately before training), IV) in the middle of RT (~0915hrs), V) immediately after training (~0930hrs), and VI) 2hrs after completion of training (~1130hrs).
 Finger draws were analysed with in-house equipment (BIOSEN C-Line, EKF diagnostic GmbH, Barleben).
@@
 
-Two muscle biopsy samples were taken for total RNA extraction per leg per time point, and total RNA was extracted in two duplicates per muscle biopsy.
+Two muscle biopsy samples were collected for total RNA extraction per leg per time point, and total RNA was extracted in two duplicates per muscle biopsy.
 Thus, we had two duplicates per leg per time point, a total of eight RNA samples per participant.
@@
 An exogenous RNA control (λ polyA External Standard Kit, Takara Bio Inc., Shiga, Japan) was added at a fixed amount (0.04 ng ml-1 of Trizol reagent) to enable analysis of target gene expression per unit tissue weight [@ellefsen_improved_2008; @ellefsen_reliable_2014].
-Briefly, muscle tissue was homogenized in TRIzol using 0.5mm RNase-free Zirconium beads (~50 ul) (Next Advanced, Averill Park, NY, USA) and a bullet blender (bb).
+Briefly, muscle tissue was homogenized in TRIzol using 0.5mm RNase-free Zirconium beads (~50 ul) (Next Advanced, Averill Park, NY, USA) and a bullet blender (Bullet Blender, Next Advanced, Averill Park, NY, USA).
 Chloroform (Sigma-Aldrich, Missouri, USA) was used for phase separation, and an RNA pellet was precipitated with isopropanol (VWR International, Pennsylvania, USA).
@@
 RNA was reverse transcribed using Super Script IV Reverse Transcriptase (Invitrogen), according to the manufacturer’s instructions.
-Briefly, 1µl 10 mM dNTP mix, 0.5µl anchored oligo-dT, 0.5µl random hexamer primers (Thermo Scientific), a maximum of 9µl template RNA and 2 µl nuclease-free water were mixed, vortexed and briefly spun down.
+One µl 10 mM dNTP mix, 0.5µl anchored oligo-dT, 0.5µl random hexamer primers (Thermo Scientific), a maximum of 9µl template RNA and 2 µl nuclease-free water were mixed, vortexed and briefly spun down.
 Samples were then heated at 65°C for 5 minutes, followed by at least 1-minute incubation on ice.
@@
 Wet muscle was freeze-dried for 24hrs and dissected before extraction.
-The tissue was homogenized with a plastic rod in 80mg protein extraction powder (Invent Biotechnology) and 100ul iced cell lysis buffer (Denaturing Buffer, Invent Biotechnology), and centrifuged at top speed for 1 min.
+The tissue was homogenized with a plastic rod in 80mg protein extraction powder (Invent Biotechnology) and 100ul iced cell lysis buffer (Denaturing Buffer, Invent Biotechnology), and centrifuged at 19 000g for 1 min.
 The supernatant was divided into 2x25ul aliquots to run samples in duplicates, and total protein concentrations were determined in a 1:10 dilution (Pierce Detergent Compatible Bradford Assay Reagent, Thermo Fisher Scientific, Oslo, Norway).
@@
 Primary antibody incubation was performed at 4°C on a tipping board.
-Chemiluminescence signals were quantified using Image Studio Lite (LI-COR Biotechnology, Lincoln, NE, USA), and total protein content was quantified using ImageJ [@rueden_imagej2_2017], whereas total protein content was defined as mean grey value of the whole well with between-well values subtracted as background.
+Chemiluminescence signals were quantified using Image Studio Lite (LI-COR Biotechnology, Lincoln, NE, USA), and total protein content was quantified using ImageJ [@rueden_imagej2_2017], whereny total protein content was defined as mean grey value of the whole well with between-well values subtracted as background.
 
@@
 Before quantitative analyses, total RNA was normalized by wet muscle weight, qPCR values were normalized by wet muscle weight and an external reference gene (Lambda), while quantified protein signals from western blot were normalized by pool signals and total protein signal per gel.<!-- add explanations of normalization to extended methods --> Total RNA and qPCR values were analyzed by comparing post to pre log-fold change score differences.
-This was achieved by modelling change as response with baseline and supplement (glucose/placebo) as explanatory factors, and mean-centred baseline values as a covariate in a linear mixed effects model using, then calculating the estimated marginal means.
+This was achieved by modelling change as response with baseline and supplement (glucose/placebo) as explanatory factors, and mean-centred baseline values as a covariate in a linear mixed effects model, then calculating the estimated marginal means.
 Quantified western blot signals and the UBF/total RNA association were analyzed using a linear mixed effects model from the nlme package [@lindstrom_nonlinear_1990], with protein signal per time point as response, with time and supplement as explanatory factors.
@@
 The linear mixed effects model was designed with the lmer function (Satterthwaite's method) from the lme4 package [@bates_fitting_2014] in R studios, using the lmerTest function to procure p-values.
-Log-transformation and mean-centring baseline values were done to control for heteroscedasticity and regression to the mean, respectively.
-Log-transformed values were reverse-transformed for figure illustration.
-Values are presented as mean change + 95% CL unless otherwise stated, with the alpha-level set to α = 0.05.
+Log-transformation and mean-centring of baseline values were performed to control for heteroscedasticity and regression to the mean, respectively.
+Log-transformed values were reverse-transformed for figure illustrations.
+Values are presented as mean change + 95% confidence limits unless otherwise stated, with the alpha-level set to α = 0.05.
 
@@
 
-There was no difference in mean change of total training session volume between exercising with placebo or glucose, with an increase of 18 and 16% from session 1 to session 6, respectively (Tab 2; Mean change, 2: *p* = `r pvol.c2`, 3: *p* = `r pvol.c3`, 4: *p* = `r pvol.c4`, 5: *p* = `r pvol.c5`, 6: *p* = `r pvol.c6`).
+There was no difference in mean change of total training session volume between exercising with placebo or glucose, with an increase of 18 and 16% from session 1 to session 6, respectively (Tab 2; Mean change, Session 2: *p* = `r pvol.c2`, Session 3: *p* = `r pvol.c3`, Session 4: *p* = `r pvol.c4`, Session 5: *p* = `r pvol.c5`, Session 6: *p* = `r pvol.c6`).
 Both exercising with glucose and placebo led to significant increases in mean total training session volume at sessions four, five and six, compared to session 1 (Tab 2; Mean change, 2: *p* = `r pvol.s2`, 3: *p* = `r pvol.s3`, 4: *p* = `r pvol.s4`, 5: *p* = `r pvol.s5`, 6: *p* = 0.001).
-Regarding the dietary data, there were no differences in mean macronutrient intake (protein, fat, carbohydrates) or total calorie intake between glucose and placebo on pairwise consecutive days (*p* > 0.05).
+Regarding the dietary data, there were no differences in mean macronutrient intake (protein, fat, carbohydrates) or total calorie intake between glucose and placebo on pairwise consecutive days (*p* > 0.05 for all).
 
@@
 Levels of c-Myc increased by 42% more after RT with placebo compared to RT with glucose, a significant difference (Fig 3A: *p* = `r cdiff`).
-Levels of UBF and rpS6 increased by 19 and 23% more in placebo compared to glucose, however, this was not a significant difference (Fig 3A: UBF: *p* = `r udiff`, rpS6: *p* = `r rpdiff`).
+Levels of UBF and rpS6 increased by 19 and 23% more in placebo compared to glucose, with no significant differences between groups (Fig 3A: UBF: *p* = `r udiff`, rpS6: *p* = `r rpdiff`).
 The increase in both RT with glucose and placebo for all measured proteins was significant at post compared to baseline (*p* < 0.05).
@@
 
-The main findings of the present study were that heavy resistance training with glucose did not affect markers of ribosome biogenesis, measured as total RNA, rRNA and rDNA transcription initiation-associated proteins, compared to RT with placebo.
+The main findings of the present study were that heavy resistance training with glucose did not affect markers of ribosome biogenesis, measured as total RNA, rRNA, and rDNA transcription initiation-associated proteins, compared to RT with placebo.
 Similarly, RT with glucose in general did not affect markers of skeletal muscle functionality such as strength and recovery, or total training session volume.
-At one time point, RT with glucose led to less reduction in strength compared to placebo, and levels of plasma glucose and c-peptide were significantly higher before, during and after the RT sessions with glucose.
+At one time point, RT with glucose led to less reduction in strength compared to placebo.
+Levels of plasma glucose and c-peptide were significantly higher before, during and after the RT sessions with glucose.
 Importantly, there were no differences in daily macronutrient intake on pairwise consecutive days.
@@
 
-To our knowledge, no previous study has investigated the effects of glucose compared to placebo ingestion with RT on ribosome biogenesis in human skeletal muscle previously, making our hypotheses exploratory and comparisons few.
-Mariappan and colleagues [@mariappan_ribosomal_2011] investigated the effect of high glucose treatment on ribosome biogenesis in glomerular epithelial cells (GECs) from diabetic (type 1 and 2) mice, observing a UBF-dependent augmentation of ribosome biogenesis.
-In a study from 2015, Tanaka et al.
-[@tanaka_mild_2015] observed that mild glucose starvation reduced rRNA transcription in human breast cancer cells.
+To our knowledge, no previous study has investigated the effects of glucose compared to placebo ingestion with RT on ribosome biogenesis in human skeletal muscle previously, making our hypotheses exploratory and comparisons few.**In vitro** studies on mice glomerular epithelial cells [@mariappan_ribosomal_2011] and human breast cancer cells [@tanaka_mild_2015] have observed glucose-induced UBF-dependent augmentation of ribosome biogenesis [@mariappan_ribosomal_2011], and reduced rRNA transcription by glucose starvation [@tanaka_mild_2015].
 These findings do imply an important role for glucose in the initiation of rDNA transcription, the rate-limiting step of ribosome biogenesis [@mariappan_ribosomal_2011; @tanaka_mild_2015].
@@
 Moreover, robust increases in markers of ribosome biogenesis such as 47S pre-rRNA and mature rRNA can be expected after an acute bout of RT [@figueiredo_impact_2016] or a short period of RT [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022].
-Therefore, we hypothesized that if glucose can, directly or indirectly via insulin, stimulate initiation of rDNA transcription, five sessions of RT with glucose compared to placebo will elicit this effect.
-Further, glucose ingestion increases energy levels, thus, we also hypothesized that high glucose vs.
+Therefore, we hypothesized that if glucose can, directly or indirectly via insulin, stimulate the initiation of rDNA transcription, five sessions of RT with glucose compared to placebo will elicit this effect.
+Further, glucose ingestion increases energy levels, leading us to hypothesize that high glucose vs.
 placebo treatment would stimulate energy-sensitive pathways such as PIH1, mTORC1, ERK1/2, AMPK and SIRT1 [@kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012].
@@
 Due to the exploratory nature of our hypotheses, this was hardly a very surprising result.
-Indeed, previous studies have used cell cultures from yeast [@zhai_human_2012], rodent [@hoppe_amp-activated_2009; @mariappan_ribosomal_2011] or human cells [@tanaka_mild_2015] and are as such not directly comparable to human skeletal muscle cells.
+Indeed, previous studies have used cell cultures from yeast [@zhai_human_2012], rodent [@hoppe_amp-activated_2009; @mariappan_ribosomal_2011] or human cells [@tanaka_mild_2015] and are, as such, not directly comparable to human skeletal muscle cells.
 Importantly, previous studies investigated high vs.
@@
 
-In the present study, despite not measuring the activity in central pathways mediating anabolic signalling (mTORC1, ERK1/2), analyses of the downstream target upstream binding factor (UBF) and the ribosomal protein S6 (rpS6) further supported observations from the RNA data.
+In the present study, despite not measuring the activity in central pathways mediating anabolic signalling (mTORC1, ERK1/2), analyses of the downstream target UBF and the ribosomal protein S6 (rpS6) further supported observations from the RNA data.
 UBF has previously been described as a master regulator of rDNA transcription *in vitro* [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @russell_rna-polymerase-i-directed_2005], while rpS6 is proposed as a valid and reliable means to measure ribosome biogenesis [@chaillou_hypoxia_2012; @nakada_correlation_2016].
-However, content of c-Myc did differ between RT with glucose and RT with placebo, whereas RT with placebo led to significantly higher c-Myc content at the post measurement, compared to glucose.
-The general transcription factor c-Myc has previously been suggested as a potent regulator of ribosome biogenesis independent of mTORC1, and a direct regulator of UBF [@mori_c-myc_2021; @poortinga_c-myc_2011; @west_acute_2016].
-The independency of mTORC1 may explain why only c-Myc content differed between treatments, while UBF and rpS6 content changed equally.
+While content of UBF and rpS6 changed equally in RT with glucose compared to placebo, content of c-Myc was significanly higher after RT with placebo.
+This was unexpected, since the general transcription factor c-Myc previously has been described as a potent regulator of ribosome biogenesis, independent of mTORC1, and a direct regulator of UBF [@mori_c-myc_2021; @poortinga_c-myc_2011; @west_acute_2016].
+Along that notion, it would be reasonable to expect that differences in c-Myc content was reflected in total RNA and rRNA.
+However, this was not the case in the present study.
+c-Myc's independency from mTORC1 may explain why only c-Myc content differed between treatments, while UBF and rpS6 content changed equally.
 Moreover, the observation that total RNA and rRNA were unaffected by the significant differences in c-Myc content, does strengthen the importance and specificity of UBF content in human skeletal muscle ribosome biogenesis.
@@
 This could suggest that the RT protocol fatigued the participants, as both training frequency and intensity were quite high.
-Another promising explanation might be the biphasic recovery pattern, as described by Raastad & Hallén [@raastad_recovery_2000].
-In their study on recovery of skeletal muscle contractility after strength exercise, Raastad and Hallén (2000) found a rapid recovery during the initial 11hrs post-exercise, followed by either a levelling off or a drop until 22hrs post-exercise, returning to baseline values at 33hrs post-exercise.
+Another promising explanation might be the biphasic recovery pattern, as described by Raastad & Hallén [@raastad_recovery_2000], where the participants experienced a rapid recovery during the initial 11hrs post-exercise, followed by a levelling off or drop until 22hrs post-exercise.
 Herein, inflammation and phagocytic activity were proposed to be involved in the performance drop between 11-22hrs [@raastad_recovery_2000].
 Indeed, this pattern seems quite similar to what was observed in the present study, with a rapid recovery at 30min and 2hrs post-sixth RT session and a drop at 23hrs.
-Moreover, the strength measurements during the intervention and after five RT sessions were also conducted approximately 23hrs after RT.
-Thus, it is possible that these strength measurements reflect a biphasic recovery pattern.
-However, this does not explain the difference between treatments at post-fifth RT session.
-A possible argument could be that exercising without glucose may have caused more stress compared to exercising with glucose, as glucose is the preferred energy source during strenuous exercise [@mul_exercise_2015], thus increasing performance in RT with glucose compared to RT with placebo.
+Further, strength testing during the intervention was conducted 23hrs after RT, meaning that these tests also may have been influenced by the biphasic recovery.
+However, this does not explain the difference between treatments at post-fifth RT session, or the great drop in strength from post 4RT to post 5RT sessions.
+A possible argument could be that exercising without glucose may have caused more stress compared to exercising with glucose, as glucose is the preferred energy source during strenuous exercise [@mul_exercise_2015], thus increasing performance with glucose compared to placebo.
 This may also, at least partly, explain the significantly higher content of the stress-responsive c-Myc in RT with placebo compared to glucose [@mori_c-myc_2021; @poortinga_c-myc_2011; @west_acute_2016].
-However, training volume data do reveal that the total training session volume was equal on pairwise consecutive days, i.e.
+However, training volume showed that the total training session volume was equal on pairwise consecutive days, i.e.
 no difference between days 1-2, days 3-4 and so on.
-Thus, there were no differences in mechanical loading to induce higher stress in RT with placebo compared to RT with glucose.
-Arguably, the significantly higher plasma levels of glucose in RT with glucose may have increased energy availability and thus, ATP, resulting in lower metabolic stress and less performance reduction, compared to RT with placebo [@westerblad_mechanisms_1998; @kent-braun_central_1999] Unfortunately, we did not conduct measurements of markers of metabolic stress such ass inorganic phosphate, H+, Mg2+ and the ADP/ATP ratio [@westerblad_mechanisms_1998; @kent-braun_central_1999].
+Hence, there were no differences in mechanical loading to induce a higher stress between treatments.
+Arguably, the significantly higher plasma levels of glucose during RT with glucose may have increased energy availability in muscle, resulting in lower metabolic stress and less performance reduction, compared to placebo during RT [@westerblad_mechanisms_1998; @kent-braun_central_1999] Unfortunately, we did not conduct measurements of markers of metabolic stress such ass inorganic phosphate, H+, Mg2+ and the ADP/ATP ratio [@westerblad_mechanisms_1998; @kent-braun_central_1999].
 Therefore, discussing the potential effect of differences in metabolic stress between treatments would only be speculation, however probable.
@@
 Neither does it acutely enhance muscular performance during the exercise period or after five sessions, nor recovery within 23hrs of the last session.
-The observations in total RNA, rRNA and rDNA transcription initiation associated proteins support the key role of UBF in ribosome biogenesis regulation in human skeletal muscle following resistance training.
+The observations in total RNA, rRNA and rDNA transcription initiation-associated proteins support the key role of UBF in ribosome biogenesis regulation in human skeletal muscle following resistance training.
 If future investigations are to be made on this topic, a greater sample size coupled with a negative control group and analyses of the intramuscular glycogen storage may provide higher-resolution results.

### 2023-04-12 ext_method.RMD
updates

@@
 Notes for extended methods supplementary
-
-- Antiboy table?
-- Power calculation + rational?
 
@@
 
+## Total protein blots
+
+### Blot/Gel 1 & 4
+
+Gel 1 overview:
+
+* Ladder loaded in well 1 and 25.
+* Pools loaded in well 6, 12 and 20.
+* Well 2-5 = sample 1.1-1.4 (participant 101).
+* Well 7-10 = sample 2.1-2.4. (participant 102).
+* Well 11 + 13-15 = sample 3.1-3.4 (participant 103).
+* Well 16-19 = sample 5.1-5.4 (participant 105).
+* Well 21-24 = sample 6.1-6.4 (participant 106).
+
+![](figures/archive/total_protein_gel1.PNG)
+
+Gel 4 overview:
+
+* Ladder loaded in well 1 and 26.
+* Pools loaded in well 2, 7, 15 and 21.
+* Well 3-6 = sample 6.4-6.1 (participant 106).
+* Well 8-11 = sample 5.4-5.1 (participant 105).
+* Well 12-14 + 16 = sample 3.4-3.1 (participant 103).
+* Well 17-20 = sample 2.4-2.1 (participant 102).
+* Well 22-25 = sample 1.4-1.1 (participant 101).
+
+![](figures/archive/total_protein_gel4.PNG)
+
+Blot/Gel 2 & 5
+
+Gel 2 overview
+
+* Ladder loaded in well 1 and 26.
+* Pools loaded in well 6, 11, 19 and 25.
+* Well 2-5 = sample 7.1-7.4 (participant 107).
+* Well 7-10 = sample 8.1-8.4 (participant 108).
+* Well 12-15 = sample 9.1-9.4 (participant 109).
+* Well 16-18 + 20 = sample 12.1-12.4 (participant 112).
+* Well 21-24 = sample 13.1-13.4 (participant 113).
+
+![](figures/archive/total_protein_gel2.PNG)
+
+Gel 5 overview
+
+* Ladder loaded in well 1 and 26.
+* Pools loaded in well 2, 8, 16 and 21.
+* Well 3-6 = sample 13.4-13.1 (participant 113).
+* Well 7 + 9-11 = sample 12.4-12.1 (participant 112).
+* Well 12-15 = sample 9.4-9.1 (participant 109).
+* Well 17-20 = sample 8.4-8.1 (participant 108).
+* Well 22-25 = sample 7.4-7.1 (participant 107).
+
+![](figures/archive/total_protein_gel5.PNG)
+
+Blot/Gel 3 & 6
+
+Gel 3 overview
+
+* Ladder loaded in well 1 and 17.
+* Pools loaded in well 2, 7 and 13.
+* Well 3-6 = sample 14.1-14.4 (participant 114).
+* Well 8-11 = sample 15.1-15.4 (participant 115).
+* Well 12 + 14-16 = sample 16.1-16.4 (participant 116).
+
+![](figures/archive/total_protein_gel3.PNG)
+
+Gel 6 overview
+
+* Ladder loaded in well 1 and 17.
+* Pools loaded in well 5, 11 and 16.
+* Well 2-4 + 6 = sample 16.4-16.1 (participant 116).
+* Well 7-10 = sample 15.4-15.1 (participant 115).
+* Well 12-15 = sample 14.4-14.1 (participant 114).
+
+![](figures/archive/total_protein_gel6.PNG)
+
 ## Statistical analyses
@@
 RNA normalization:
+
  1. weight = mean(weight) 
@@
  4. RNA.weight = RNA / weight
- - weight = biopsy weight
+    + weight = biopsy weight
 
@@
 rRNA normalization:
+
  1. Lambda expression was first calculated and stored in the data.frame "nf"
@@
  6. nf.expr = log(expr / nf.w), nf.w = scale(nf.w) ?
- - expr = target expression 
- - weight = biopsy weight
+    + expr = target expression 
+    + weight = biopsy weight
 
@@
 Protein normalization:
+
  1. tpl = tp - bg
@@
  6. norm.sign = norm.sign / pool
- - tpl = total protein per lane
- - tp = total protein (mean gray pixels in lanes)
- - bg = background (mean gray pixels between lanes)
- - signal = pixels per band
+    + tpl = total protein per lane
+    + tp = total protein (mean gray pixels in lanes)
+    + bg = background (mean gray pixels between lanes)
+    + signal = pixels per band
 
@@
 So, after normalizing signal to tpl (norm.sign = signal / tpl), we summarized the average signal from the pools (pool = mean(pool)), and divided the tpl normalized signal by average pool signal (norm.sign = norm.sign / pool).
+
+### Statistical models
+
+RNA
+
+```{r total RNA model}
+totalrna.model <- readRDS("./data/data-gen/rna/totalrna.model.RDS")
+
+totalrna.model
+
+```
+
+Protein
+
+```{r protein model}
+protein.model <- readRDS("./data/data-gen/protein/protein.model.RDS")
+
+protein.model
+
+```
+
+Protein/RNA correlation
+
+```{r correlation model}
+correlation.model <- readRDS("./data/data-gen/protein/correlation.model.RDS")
+
+correlation.model
+
+```
+
+Strength, plasma glucose & c-peptide, nutrition and volume
+
+```{r strength, plasma, nutrition and volume model}
+strength.model <- readRDS("./data/data-gen/humac/strength.model.RDS")
+
+strength.model
+
+```

### 2023-04-12 R/rrna.analysis.R

@@ -186,6 +186,8 @@ ttest.47 <- t.test(base.47$GLUCOSE, base.47$PLACEBO, paired = TRUE)
 m1 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
                      data = change.18)
 
+saveRDS(m1, "./data/data-gen/rna/rrna.model.RDS")
+
 plot(m1)
 
 summary(m1)
@@ -200,7 +202,7 @@ summary(m2)
 
 # 5.8S
 m3 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
-                     data = change.5.8)
+                     data = change.58)
 
 plot(m3)
 

### 2023-04-12 R/strength.index.R

@@ -146,6 +146,8 @@ saveRDS(astr.lchange, "./data/data-gen/humac/astr.lchange.RDS")
 m.str <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
                          data = str.change)
 
+saveRDS(m.str, "./data/data-gen/humac/strength.model.RDS")
+
 m.lstr <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
                         data = str.lchange)
 plot(m.str)

### 2023-04-13 ribose-paper.postfeedback.RMD
misc changes

@@
 title: "Glucose ingestion before and after resistance training does not augment ribosome biogenesis in healthy moderately trained young adults"
-author: "Kristian Lian, Daniel Hammarström, Stian Ellefsen, Håvard Hamarsland"
+author: "Kristian Lian^1^, Sara<sup>1</sup>, Håvard Hamarsland<sup>1</sup>, Daniel Hammarström<sup>1</sup>, Stian Ellefsen<sup>1</sup>"
 bibliography: resources/ribose-paper.bib
@@
   word_document:
-    reference_docx: reference-template-02.docx
+    reference_docx: resources/reference-template-02.docx
 editor_options: 
   chunk_output_type: console
+knit: (function(input, ...) {
+    rmarkdown::render(
+      input,
+      output_dir = "../output"
+    )
+  })
 ---
@@
 ```
+
+## Abstract
+
+### Affiliations
+1: Inland Norway University of Applied Sciences, Section for , Adress
+
+### Correspondance
+Kristian Lian kristian.lian@inn.no
+
+Stian Ellefsen stian.ellefsen@inn.no
+
+### ORCID
+
+Kristian Lian
+<a href="https://orcid.org/0000-0002-5169-5960">
+<img alt="ORCID logo" src="https://info.orcid.org/wp-content/uploads/2019/11/orcid_16x16.png" width="16" height="16" /></a>,
+Daniel Hammarström
+<a href="https://orcid.org/0000-0001-8360-2100">
+<img alt="ORCID logo" src="https://info.orcid.org/wp-content/uploads/2019/11/orcid_16x16.png" width="16" height="16" /></a>
+
+<a href="https://orcid.org/0000-0003-0467-2018">
+Stian Ellefsen
+</a>,
+">
+Håvard Hamarsland: https://orcid.org/0000-0002-7343-2774
+
+</a>
+
+\newpage
 
@@
 
-Sixteen healthy male and female participants (20-33yrs, Tab 1) were recruited to the study through social media advertisement and word of mouth.
+Sixteen healthy male and female participants (20-33 years, Table 1) were recruited to the study through social media advertisement and word of mouth.
 The eligibility criteria were non-smokers and moderately trained (i.e.
@@
 One due to sickness and inability to resume, two participants experienced muscular discomfort related to heavy resistance training.
-Lean mass (Tab 1) was measured using DXA (Prodigy Advance PA+302047, Lunar, San Francisco, CA, USA) on Day -1, the last day preceding the RT intervention.
+Lean mass (Table 1) was measured using DXA (Prodigy Advance PA+302047, Lunar, San Francisco, CA, USA) on Day -1, the last day preceding the RT intervention.
 
@@
 
-The study was designed as a 12-day double-blinded randomized controlled trial, with an alternating unilateral RT protocol (Fig 1A).
+The study was designed as a 12-day double-blinded randomized controlled trial, with an alternating unilateral RT protocol (Figure 1A).
 Participants were randomly allocated to exercise one leg with glucose and one leg with a placebo (Fig 1A).
@@
 
-## Figure/table text
-
- **Table 1:**  Baseline characteristics.
-Age = mean age in years, Height = mean height in cm, Weight and Lean mass = mean in kg, Iso 60/240 = mean unilateral isokinetic knee extension peak torque in Nm at 60 d/s and 240 d/s, Isom = mean unilateral isometric knee extension peak torque in Nm.
-
-**Table 2:** Mean total session volume and mean change in total session volume from session 1 to session 6.
+## Figure captions
 
@@
 Values are presented as log-transformed means.
+
+## Tables
+
+```{r, echo = FALSE}
+
+###### Baseline tab
+#
+#
+## Author: KL
+#
+#
+## Purpose:
+
+## Packages
+library(readxl); library(tidyverse); library(flextable)
+
+## Data
+
+leg.dat <- read_excel("./data/dxa/ribose_dxaprleg.xlsx")
+
+dxa.dat <- read_excel("./data/dxa/ribose_dxa.xlsx")
+
+humac.clean2 <- readRDS("./data/data-gen/humac/humac.clean2.RDS")
+
+str.index <- readRDS("./data/data-gen/humac/str.index.RDS")
+
+code <- read_excel("./data/sup.xlsx")
+
+## DXA
+
+# Counts the "n" of females and males, added in the last step before 
+# making a flextable
+gendercount <- dxa.dat %>%
+  count(sex) %>%
+  mutate(sex = if_else(sex == "female",
+                       "Female",
+                       if_else(sex == "male",
+                               "Male", sex)))
+
+# Summarising and prepping dxa data
+dxa.tab <- dxa.dat %>%
+  select(subject, sex, age, height, weight, leanmassg) %>%
+  mutate(leanmass = leanmassg/1000) %>%
+  group_by(sex) %>%
+  summarise(mean.a = mean(age),
+            sd.a = sd(age),
+            mean.h = mean(height),
+            sd.h = sd(height),
+            mean.w = mean(weight),
+            sd.w = sd(weight),
+            mean.lm = mean(leanmass),
+            sd.lm = sd(leanmass)) %>%
+  mutate(age = paste0(sprintf("%.1f", mean.a), " (", sprintf("%.1f",sd.a), ")"),
+         height = paste0(sprintf("%.1f",mean.h), " (",sprintf("%.1f",sd.h), ")"),
+         weight = paste0(sprintf("%.1f",mean.w), " (", sprintf("%.1f",sd.w), ")"),
+         lm = paste0(sprintf("%.1f",mean.lm), " (", sprintf("%.1f",sd.lm), ")")) %>%
+  mutate(sex = if_else(sex == "female",
+                       "Female",
+                       if_else(sex == "male",
+                               "Male", sex))) %>%
+  select(sex, age, height, weight, lm) %>%
+  pivot_longer(values_to = "mean",
+               names_to = "measure",
+               cols = c(age:lm))
+
+# Gets gender per subject, to mark the subjects with gender in the humac 
+# data
+
+gender <- dxa.dat %>%
+  select(subject, sex)
+
+# Summarising and prepping humac data
+humac.tab <- humac.clean2 %>%
+  full_join(gender) %>%
+  select(subject, sex, time, acute, measure = test, peak.torque) %>%
+  filter(time == "baseline" & acute == "rest") %>%
+  group_by(sex, measure) %>%
+  summarise(mean.pt = mean(peak.torque),
+            sd.pt = sd(peak.torque)) %>%
+  mutate(mean = paste0(sprintf("%.1f",mean.pt), " (", sprintf("%.1f",sd.pt), ")")) %>%
+  select(sex, measure, mean) %>%
+  mutate(sex = if_else(sex == "female",
+                       "Female",
+                       if_else(sex == "male",
+                               "Male", sex)))
+
+# Joining the data frames to be used in the flextable
+
+dxa.tab %>%
+  full_join(humac.tab) %>%
+  pivot_wider(values_from = mean,
+              names_from = measure) %>%
+  full_join(gendercount) %>%
+  select(sex, n, age, height, weight, lm, isok.60, isok.240, isom) %>%
+
+  ## Make it to flextable
+  flextable() %>%
+  ## Change headers
+  compose(i = 1, j = c(1, 2, 3, 4, 5, 6, 7, 8, 9), part = "header", 
+          value = c(as_paragraph("Sex"),
+                    as_paragraph("n"), 
+                    as_paragraph("Age (yrs)"), 
+                    as_paragraph("Stature (cm)"), 
+                    as_paragraph("Body mass (kg)"),
+                    as_paragraph("Lean mass (kg)"),
+                    as_paragraph("60\U00BA sec", as_sup("-1")), 
+                    as_paragraph("240\U00BA sec", as_sup("-1")), 
+                    as_paragraph("0\U00BA sec", as_sup("-1")))) %>%
+  add_header_row(values = c(" ", "Knee-extension peak torque"), colwidths = c(6, 3)) %>%
+  add_body_row(top = FALSE, value = as_paragraph("Values are mean and (SD)"), colwidths = 9) %>%
+  set_caption("Table 1. Participant characteristics")
+
+```

### 2023-04-13 ribose-paper.postfeedback.RMD
misc changes

@@
   chunk_output_type: console
-knit: (function(input, ...) {
-    rmarkdown::render(
-      input,
-      output_dir = "../output"
-    )
-  })
 ---

### 2023-04-14 ribose-paper.postfeedback.RMD
updated fig/tab to figure/table

@@
 The study was designed as a 12-day double-blinded randomized controlled trial, with an alternating unilateral RT protocol (Figure 1A).
-Participants were randomly allocated to exercise one leg with glucose and one leg with a placebo (Fig 1A).
+Participants were randomly allocated to exercise one leg with glucose and one leg with a placebo (Figure 1A).
 <!--from the "old" methods: either their dominant or non-dominant leg with glucose or placebo before and after RT, and whether they started RT with glucose or placebo.
@@
 Data from the first five RT sessions was used to investigate main outcome measures (total RNA, rRNA and protein) and leg muscle strength, whereas data from the sixth RT session was used to explore secondary outcomes (muscular recovery, plasma glucose and -c-peptide levels).
-Participants were asked to avoid  resistance- or high-intensity training of the legs from Day -7 (Fig 1A) and onwards, until completion of the intervention and post-testing, to ensure the reliability of pre-intervention strength data and minimal interference from external exercise sources.
+Participants were asked to avoid  resistance- or high-intensity training of the legs from Day -7 (Figure 1A) and onwards, until completion of the intervention and post-testing, to ensure the reliability of pre-intervention strength data and minimal interference from external exercise sources.
 If participants did conduct exercise training outside of the prescribed protocol, they were asked to ensure equal loading on pairwise consecutive days.
@@
 
-Strength tests were performed prior to (Days -7 and -5, and -1, both legs) and during the intervention (Fig 1A, on days 4 and 8 for leg 1, and days 5 and 9 for leg two), after session 5 and after finalization of intervention (Fig 1A, on days 11/12 for leg 1 and days 12/13 for leg 2).
+Strength tests were performed prior to (Days -7 and -5, and -1, both legs) and during the intervention (Figure 1A, on days 4 and 8 for leg 1, and days 5 and 9 for leg two), after session 5 and after finalization of intervention (Figure 1A, on days 11/12 for leg 1 and days 12/13 for leg 2).
 Maximal isometric and isokinetic knee extension torque was measured with a Humac Norm Dynamometer (CSMi, Stoughton, Massachusetts, USA).
-Individual positions were recorded and standardized from pre-intervention tests (Fig 1A, days -7 and -5).
+Individual positions were recorded and standardized from pre-intervention tests (Figure 1A, days -7 and -5).
 Isokinetic peak torque was measured at extensions speeds of 60- and 240 d/s, 2x3 repetitions each, with the first set of each exercise as a sub-maximal warm-up.
@@
 The isometric tests were ended when the participants reached a plateau or peak torque development decreased, which on average occurred between 2-4 seconds into the test.
-During days 4, 5, 8 and 9 (Fig 1D, days 4 and 5 = Post 2RT, days 8 and 9 = Post 4RT), humac tests were conducted one hour before RT on the leg performing RT the previous day.
-During days 11 and 12, humac tests were performed four times: I) 45min before the last RT session (Fig 1D, Post 5RT), II) 30min after the last RT session (Fig 1D, 30 min post 6RT), III) 2hrs after the last RT session (Fig 1D, 2h post 6RT), and IV) 23hrs after the last RT session (Fig 1D, 23h post 6RT).
+During days 4, 5, 8 and 9 (Figure 1D, days 4 and 5 = Post 2RT, days 8 and 9 = Post 4RT), humac tests were conducted one hour before RT on the leg performing RT the previous day.
+During days 11 and 12, humac tests were performed four times: I) 45min before the last RT session (Figure 1D, Post 5RT), II) 30min after the last RT session (Figure 1D, 30 min post 6RT), III) 2hrs after the last RT session (Figure 1D, 2h post 6RT), and IV) 23hrs after the last RT session (Figure 1D, 23h post 6RT).
 Test I on day 12/13 included testing of both legs, representing 23hrs post-RT session test of one leg and post-session 5 test of the other leg.
@@
 
-Assessment of unilateral one repetition maximum (1RM) leg press and knee extension was conducted at pre-intervention testing during familiarization (days -7 and -5, Fig 1D).
+Assessment of unilateral one repetition maximum (1RM) leg press and knee extension was conducted at pre-intervention testing during familiarization (Figure 1, Ddays -7 and -5).
 The participants performed a general warm-up with 10min cycling on an indoor exercise bicycle.
-In addition, a protocol consisting of 1x10, 1x6 and 1x3 repetitions, was used as a specific warm-up before each of the tests.
+A protocol consisting of 1x10, 1x6 and 1x3 repetitions with a load equivalent to ~50-75% of assumed max repetitions, was used as a specific warm-up before each of the tests.
 All positions were controlled and recorded at the first 1RM test and repeated for the RT sessions.
@@
 After the biopsy sampling, muscle tissue was divided into two aliquots for determination of total RNA/expression of rRNA and two aliquots for protein content measurement, snap frozen in isopentane (-80°C) and stored at -80°C until further analyses.
-Muscle biopsies were sampled at four times: I/II) Pre-intervention (Fig 1A, 2hrs before training, Day 1 = leg 1, Day 2 = leg 2), and III/IV) after the fifth RT session, two hours before the sixth RT session (1A, Day 11 = leg 1, Day 12 = leg 2).
+Muscle biopsies were sampled at four times: I/II) Pre-intervention (Figure 1A, 2hrs before training, Day 1 = leg 1, Day 2 = leg 2), and III/IV) after the fifth RT session, two hours before the sixth RT session (Figure 1A, Day 11 = leg 1, Day 12 = leg 2).
 At each time point, two samples were taken from the same incision.
@@
 To measure blood glucose levels with and without glucose ingestion/training, blood was collected by means of finger draws on days with biopsy sampling.
-One finger draw was taken on day 1 (Fig 1A) to serve as a baseline.
-On days 11 and 12 (Fig 1A), finger draws were collected seven times: I) Immediately prior to protein ingestion (0700hrs) II) 45 minutes after protein ingestion (0745hrs) III) 1.5hrs after protein ingestion (0800hrs, i.e., immediately before GLU/PLAC intake), IV) 2hrs after protein ingestion (0900hrs, i.e., immediately before training), IV) in the middle of RT (~0915hrs), V) immediately after training (~0930hrs), and VI) 2hrs after completion of training (~1130hrs).
+One finger draw was taken on day 1 (Figure 1A) to serve as a baseline.
+On days 11 and 12 (Figure 1A), finger draws were collected seven times: I) Immediately prior to protein ingestion (0700hrs) II) 45 minutes after protein ingestion (0745hrs) III) 1.5hrs after protein ingestion (0800hrs, i.e., immediately before GLU/PLAC intake), IV) 2hrs after protein ingestion (0900hrs, i.e., immediately before training), IV) in the middle of RT (~0915hrs), V) immediately after training (~0930hrs), and VI) 2hrs after completion of training (~1130hrs).
 Finger draws were analysed with in-house equipment (BIOSEN C-Line, EKF diagnostic GmbH, Barleben).
@@
 
-There was no difference in mean change of total training session volume between exercising with placebo or glucose, with an increase of 18 and 16% from session 1 to session 6, respectively (Tab 2; Mean change, Session 2: *p* = `r pvol.c2`, Session 3: *p* = `r pvol.c3`, Session 4: *p* = `r pvol.c4`, Session 5: *p* = `r pvol.c5`, Session 6: *p* = `r pvol.c6`).
-Both exercising with glucose and placebo led to significant increases in mean total training session volume at sessions four, five and six, compared to session 1 (Tab 2; Mean change, 2: *p* = `r pvol.s2`, 3: *p* = `r pvol.s3`, 4: *p* = `r pvol.s4`, 5: *p* = `r pvol.s5`, 6: *p* = 0.001).
+There was no difference in mean change of total training session volume between exercising with placebo or glucose, with an increase of 18 and 16% from session 1 to session 6, respectively (Table 2; Mean change, S2 (session 2): *p* = `r pvol.c2`, S3: *p* = `r pvol.c3`, S4: *p* = `r pvol.c4`, S5: *p* = `r pvol.c5`, S6: *p* = `r pvol.c6`).
+Both exercising with glucose and placebo led to significant increases in mean total training session volume at sessions four, five and six, compared to session 1 (Table 2; Mean change, S2 (session 2): *p* = `r pvol.s2`, S3: *p* = `r pvol.s3`, S4: *p* = `r pvol.s4`, S5: *p* = `r pvol.s5`, S6: *p* = 0.001).
 Regarding the dietary data, there were no differences in mean macronutrient intake (protein, fat, carbohydrates) or total calorie intake between glucose and placebo on pairwise consecutive days (*p* > 0.05 for all).
@@
 
-Glucose ingestion before and after RT led to significant increases in plasma levels of glucose by 38% immediately before RT (Fig 1B, 0 min: *p* = 0.000), by 31% during RT (1B, 15 min: *p* = 0.000) and by 32% immediately after RT (1B, 30 min: *p* = 0.000), compared to placebo ingestion before and after RT.
-Two hours after the session, RT with glucose measured 8% lower plasma glucose levels compared to RT with placebo (1B, 270min: *p* = `r p.270`).
-Similarly, levels of c-peptide increased significantly more in exercises with glucose compared to placebo, by 95% immediately before (Fig 1C, 0 min: *p* = 0.000) and 87% after RT (1C, 30 min: *p* = 0.000).
+Glucose ingestion before and after RT led to significant increases in plasma levels of glucose by 38% immediately before RT (Figure 1B, 0 min: *p* = 0.000), by 31% during RT (Figure 1B, 15 min: *p* = 0.000) and by 32% immediately after RT (Figure 1B, 30 min: *p* = 0.000), compared to placebo ingestion before and after RT.
+Two hours after the session, RT with glucose measured 8% lower plasma glucose levels compared to RT with placebo (Figure 1B, 270min: *p* = `r p.270`).
+Similarly, levels of c-peptide increased significantly more in exercises with glucose compared to placebo, by 95% immediately before (Figure 1C, 0 min: *p* = 0.000) and 87% after RT (Figure 1C, 30 min: *p* = 0.000).
 
@@
 
-After 5 RT sessions (Fig 1D, Post 5RT), there was a significant difference in strength between glucose and placebo,  where placebo decreased strength by 7% more than glucose (1D, *p* = `r pdiff.c4`).
-At the other time points, there were no significant differences in mean change between exercising with glucose and placebo (1D, Post 2RT: *p* = `r pdiff.c2`, Post 4RT: *p* = `r pdiff.c3`, 30min post 6RT: *p* = `r pdiff.c5`, 2h post 6RT: *p* = `r pdiff.c6`, 23h post 6RT: *p* = `r pdiff.c7`).
-Both RT with glucose and placebo led to significantly reduced strength post-fifth session compared to baseline, by 11 and 18% respectively (1D, Post 5RT: *p* = 0.000).
-Comparisons of the acute data gathered from the post-fifth session until and including 23hrs post-sixth session showed a significant mean increase in strength of 5-9% after RT with glucose and placebo 30min after the sixth RT session (1D, 30min post 6RT: *p* = `r apc2`) and two hours after the sixth RT session (1D, 2h post 6RT: *p* = `r apc3`).
-Twenty-three hours after the last sixth RT session, strength was unchanged compared to after the fifth RT session (23h post 6RT: *p* = `r apc4`).
+After 5 RT sessions (Figure 1D, Post 5RT), there was a significant difference in strength between glucose and placebo,  where placebo decreased strength by 7% more than glucose (Figure 1D, *p* = `r pdiff.c4`).
+At the other time points, there were no significant differences in mean change between exercising with glucose and placebo (Figure 1D, Post 2RT: *p* = `r pdiff.c2`, Post 4RT: *p* = `r pdiff.c3`, 30min post 6RT: *p* = `r pdiff.c5`, 2h post 6RT: *p* = `r pdiff.c6`, 23h post 6RT: *p* = `r pdiff.c7`).
+Both RT with glucose and placebo led to significantly reduced strength post-fifth session compared to baseline, by 11 and 18% respectively (Figure 1D, Post 5RT: *p* = 0.000).
+Comparisons of the acute data gathered from the post-fifth session until and including 23hrs post-sixth session showed a significant mean increase in strength of 5-9% after RT with glucose and placebo 30min after the sixth RT session (Figure 1D, 30min post 6RT: *p* = `r apc2`) and two hours after the sixth RT session (Figure 1D, 2h post 6RT: *p* = `r apc3`).
+Twenty-three hours after the last sixth RT session, strength was unchanged compared to after the fifth RT session (Figure 1D, 23h post 6RT: *p* = `r apc4`).
 
@@
 
-Resistance training with glucose did not induce a higher accumulation of total RNA (fig 2A, *p* = `r trdiffp`) or rRNA  (fig 2B, 47S: *p* = `r rr47diffp`, 18S: *p* = `r rr18diffp`, 28S: *p* = `r rr28diffp`, 5.8S: *p* = `r rr58diffp`, 5S: *p* = `r rr5diffp`) compared to RT with placebo.
+Resistance training with glucose did not induce a higher accumulation of total RNA (Figure 2A, *p* = `r trdiffp`) or rRNA  (Figure 2B, 47S: *p* = `r rr47diffp`, 18S: *p* = `r rr18diffp`, 28S: *p* = `r rr28diffp`, 5.8S: *p* = `r rr58diffp`, 5S: *p* = `r rr5diffp`) compared to RT with placebo.
 From baseline to post-intervention, there was a mean increase in levels of total RNA by 26 and 22% after RT with glucose and placebo, respectively (*p* < 0.05, compared to baseline).
 A robust accumulation was also observed in ribosomal RNA expression in both RT with glucose and placebo, with mean increases between 34-43% (GLU) and 33-41% (PLAC) (*p* < 0.05, compared to baseline) in the four rRNAs.
-The expression of the 47S pre-rRNA increased by 37 and 59% with glucose and placebo respectively, where only RT with placebo increased significantly from baseline to post (fig 45, *p* = `r rr47tdiffp`).
+The expression of the 47S pre-rRNA increased by 37 and 59% with glucose and placebo respectively, where only RT with placebo increased significantly from baseline to post (Figure 2B, *p* = `r rr47tdiffp`).
 
@@
 
-Levels of c-Myc increased by 42% more after RT with placebo compared to RT with glucose, a significant difference (Fig 3A: *p* = `r cdiff`).
-Levels of UBF and rpS6 increased by 19 and 23% more in placebo compared to glucose, with no significant differences between groups (Fig 3A: UBF: *p* = `r udiff`, rpS6: *p* = `r rpdiff`).
+Levels of c-Myc increased by 42% more after RT with placebo compared to RT with glucose, a significant difference (Figure 3A: *p* = `r cdiff`).
+Levels of UBF and rpS6 increased by 19 and 23% more in placebo compared to glucose, with no significant differences between groups (Figure 3A: UBF: *p* = `r udiff`, rpS6: *p* = `r rpdiff`).
 The increase in both RT with glucose and placebo for all measured proteins was significant at post compared to baseline (*p* < 0.05).

### 2023-04-21 README.rmd
updated README
misc changes

--- 2023-01-05 README.rmd
@@
 
-# contratrain-muscle-paper
+# Ribose-paper
 
@@
 <img alt="ORCID logo" src="https://info.orcid.org/wp-content/uploads/2019/11/orcid_16x16.png" width="16" height="16" /></a> 
+Stian Ellefsen
+<a href="https://orcid.org/0000-0003-0467-2018">
+<img alt="ORCID logo" src="https://info.orcid.org/wp-content/uploads/2019/11/orcid_16x16.png" width="16" height="16" /></a>,
+Håvard Hamarsland
+<a href="https://orcid.org/0000-0002-7343-2774">
+<img alt="ORCID logo" src="https://info.orcid.org/wp-content/uploads/2019/11/orcid_16x16.png" width="16" height="16" /></a>,
+Sara Christine Moen
+
+## Abstract
 
@@
 
-- bibliography.bib
-- csl-files
+- Bibliography file
+- Csl-files
+- Reference template
+- Pptx file for the blot images in figure 3
+- RMD output files <!-- if you get this to work -->
 
@@
 
-- contains raw data not included in the data package
+- Contains raw data files
 
-`/data-gen/`
+`/data/data-gen/`
 
-- contains generated data stored for further analyses
+- Contains generated data stored for further analyses
 
-`contratrain-manuscript.qmd`
+`ribose-manuscript.rmd`
 
 - The manuscript source file
+
+`extended-methods.rmd`
+
+- Supplementary extended methods source file
 
@@
 - Figures source files and output
+
+`/figures/archive`
+
+- Png and tif files of total protein stains and antibody stains
+
+`/R`
+
+- Scripts for data handling/cleanup and analyses
+
+`/R/scraps`
+
+- Old/unused scripts

### 2023-04-26 ribose-paper.postfeedback.RMD
abstract

@@
 ## Abstract
+
+*Introduction:*Responses to resistance training (RT) are not uniform across the population, and individuals responding poorly to RT show a blunted ability to produce novel ribosomes [@hammarstrom2020;@stec2016].
+Recent research solidifies the relationship between increased rates of total RNA and the magnitude of RT-induced adaptations [@hammarstrom2020;@hammarstrom2022@figueiredo2019rev], with ribosome abundance being associated with UBF protein levels [@hammarstrom2022].
+However, 10-15% of individuals undertaking a period of RT showed considerable impaired growth [@alvarez2018;@mann2014;@thalacker2013].
+Since modifying RT alone does not convert low-responders to high-responders [@hammarstrom2020] other means than RT per se is necessary to circumvent this discrepancy.
+High glucose treatment has been observed to augment rDNA transcription in vitro[@tanaka2018], however, its effect on RT-induced adaptations remains quite unexplored.
+*Methods:*Sixteen healthy moderately trained participants were included and randomized to unilateral resistance training with glucose- and placebo supplement.
+Participants remained fasted overnight receiving glucose/placebo and protein only before and after RT.
+Resistance training consisted of 3 sets of 10 repetitions maximum unilateral leg press and knee extension.
+Micro biopsies were sampled pre (T1/T2) and post (T3/T4) the intervention, and maximal unilateral isometric and isokinetic knee extension torque were measured pre (T0), during (days 4, 5, 8, 9) and post (T3/T4, day 13) the intervention.
+*Results:*Despite robust increases in total RNA and rRNA in glucose and placebo (RNA: 26 and 22% respectively, rRNA: 33-43% and 33-41% respectively), there were no differences between supplements (p > 0.05).
+Content of c-Myc increased 42% more (*p* = 0.027) in placebo compared to glucose, but there were no differences in content of UBF (*p* = 0.185) and rpS6 (*p* = 0.178) between treatments.
+Further, there was a linear relationship between content of UBF and total RNA, where an increase of 1 SD unit in UBF equated to 13% increase in total RNA (*p* = 0.002).
+Lastly, there were no differences between supplements in maximal unilateral isometric knee extension force and isokinetic peak torque during the intervention (p > 0.05) and following the last session (p > 0.05), apart from the 60 d/s test before last RT session (p < 0.05).
+*Conclusion:* Glucose supplement before and after RT does not acutely augment ribosome biogenesis, nor does it enhance muscular performance and recovery after two weeks of heavy RT in moderately trained young individuals.
 

### 2023-04-26 ribose-paper.postfeedback.RMD
abstract

@@
 
-*Introduction:*Responses to resistance training (RT) are not uniform across the population, and individuals responding poorly to RT show a blunted ability to produce novel ribosomes [@hammarstrom2020;@stec2016].
-Recent research solidifies the relationship between increased rates of total RNA and the magnitude of RT-induced adaptations [@hammarstrom2020;@hammarstrom2022@figueiredo2019rev], with ribosome abundance being associated with UBF protein levels [@hammarstrom2022].
-However, 10-15% of individuals undertaking a period of RT showed considerable impaired growth [@alvarez2018;@mann2014;@thalacker2013].
-Since modifying RT alone does not convert low-responders to high-responders [@hammarstrom2020] other means than RT per se is necessary to circumvent this discrepancy.
-High glucose treatment has been observed to augment rDNA transcription in vitro[@tanaka2018], however, its effect on RT-induced adaptations remains quite unexplored.
+*Introduction:*Responses to resistance training (RT) are not uniform across the population, and individuals responding poorly to RT show a blunted ability to produce novel ribosomes [@hammarstrom_benefits_2020; @stec_ribosome_2016].
+Recent research solidifies the relationship between increased rates of total RNA and the magnitude of RT-induced adaptations [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022; @figueiredo_revisiting_2019], with ribosome abundance being associated with UBF protein levels [@hammarstrom_ribosome_2022].
+However, 10-15% of individuals undertaking a period of RT showed considerable impaired growth [@alvarez_interindividual_2018; @mann_high_2014; @thalacker-mercer_cluster_2013].
+Since modifying RT alone does not convert low-responders to high-responders [@hammarstrom_benefits_2020] other means than RT per se is necessary to circumvent this discrepancy.
+High glucose treatment has been observed to augment rDNA transcription in vitro[@sebata_control_2018], however, its effect on RT-induced adaptations remains quite unexplored.
 *Methods:*Sixteen healthy moderately trained participants were included and randomized to unilateral resistance training with glucose- and placebo supplement.

### 2023-06-01 ribose-paper.postfeedback.RMD
updates misc

--- 2023-04-14 ribose-paper.postfeedback.RMD
@@
 title: "Glucose ingestion before and after resistance training does not augment ribosome biogenesis in healthy moderately trained young adults"
-author: "Kristian Lian^1^, Sara<sup>1</sup>, Håvard Hamarsland<sup>1</sup>, Daniel Hammarström<sup>1</sup>, Stian Ellefsen<sup>1</sup>"
+author: "Kristian Lian^1^, Sara Christine Olsen Moen^1^, Håvard Hamarsland^1^, Daniel Hammarström^1^, Stian Ellefsen^1^"
 bibliography: resources/ribose-paper.bib
@@
 
+### Affiliations
+<div custom-style="correspondence_style">
+^1^ Section for Health and Exercise Physiology, Department of Public Health and Sport Sciences, Inland Norway University of Applied Sciences, Lillehammer, Norway.
+</div>
+
+<div custom-style="correspondence_style">
+### Correspondance
+Kristian Lian, Section for Health and Exercise Physiology, Department of Public Health and Sport Sciences, Inland Norway University of Applied Sciences, Lillehammer, Norway.
+Email: kristian.lian@inn.no
+
+Stian Ellefsen, Section for Health and Exercise Physiology, Department of Public Health and Sport Sciences, Inland Norway University of Applied Sciences, Lillehammer, Norway.
+Email: stian.ellefsen@inn.no
+</div>
+
+### ORCID
+<div custom-style="correspondence_style">
+Kristian Lian: https://orcid.org/0000-0002-5169-5960
+
+Håvard Hamarsland: https://orcid.org/0000-0002-7343-2774
+
+Daniel Hammarström https://orcid.org/0000-0001-8360-2100
+
+Stian Ellefsen https://orcid.org/0000-0003-0467-2018
+</div>
+
+\pagebreak
+<div custom-style="abstract_style">
 ## Abstract
 
-### Affiliations
-1: Inland Norway University of Applied Sciences, Section for , Adress
-
-### Correspondance
-Kristian Lian kristian.lian@inn.no
-
-Stian Ellefsen stian.ellefsen@inn.no
-
-### ORCID
-
-Kristian Lian
-<a href="https://orcid.org/0000-0002-5169-5960">
-<img alt="ORCID logo" src="https://info.orcid.org/wp-content/uploads/2019/11/orcid_16x16.png" width="16" height="16" /></a>,
-Daniel Hammarström
-<a href="https://orcid.org/0000-0001-8360-2100">
-<img alt="ORCID logo" src="https://info.orcid.org/wp-content/uploads/2019/11/orcid_16x16.png" width="16" height="16" /></a>
-
-<a href="https://orcid.org/0000-0003-0467-2018">
-Stian Ellefsen
-</a>,
-">
-Håvard Hamarsland: https://orcid.org/0000-0002-7343-2774
-
-</a>
-
-\newpage
-
+**Keywords: **
+</div>
+
+\pagebreak
 ## Introduction
@@
 vastus lateralis using well-established procedures [@hammarstrom_benefits_2020].
-Briefly, muscle biopsy sampling was performed under local anaesthesia (Xylocaine, 10mg ml-1 with adrenaline 5μg ml−1, AstraZeneca AS, Oslo, Norway) using a 12-14-gauge needle (Universal Plus, Mermaid Medical AS, Stenløse, Denmark), operated with a spring-loaded biopsy gun (Bard Magnum, Bard, Rud, Norway).
-After the biopsy sampling, muscle tissue was divided into two aliquots for determination of total RNA/expression of rRNA and two aliquots for protein content measurement, snap frozen in isopentane (-80°C) and stored at -80°C until further analyses.
+Briefly, muscle biopsy sampling was performed under local anaesthesia (Xylocaine, 10 mg ml^-1^ with adrenaline 5 μg ml^−1^, AstraZeneca AS, Oslo, Norway) using a 12-gauge needle (Universal Plus, Mermaid Medical AS, Stenløse, Denmark), operated with a spring-loaded biopsy gun (Bard Magnum, Bard, Rud, Norway).
+After the biopsy sampling, muscle tissue was divided into two aliquots for determination of total RNA/expression of rRNA and two aliquots for protein content measurement.
+Aliquots were snap frozen in isopentane (-80°C) and stored at -80°C until further analyses.
 Muscle biopsies were sampled at four times: I/II) Pre-intervention (Figure 1A, 2hrs before training, Day 1 = leg 1, Day 2 = leg 2), and III/IV) after the fifth RT session, two hours before the sixth RT session (Figure 1A, Day 11 = leg 1, Day 12 = leg 2).
@@
 
-To measure blood glucose levels with and without glucose ingestion/training, blood was collected by means of finger draws on days with biopsy sampling.
-One finger draw was taken on day 1 (Figure 1A) to serve as a baseline.
-On days 11 and 12 (Figure 1A), finger draws were collected seven times: I) Immediately prior to protein ingestion (0700hrs) II) 45 minutes after protein ingestion (0745hrs) III) 1.5hrs after protein ingestion (0800hrs, i.e., immediately before GLU/PLAC intake), IV) 2hrs after protein ingestion (0900hrs, i.e., immediately before training), IV) in the middle of RT (~0915hrs), V) immediately after training (~0930hrs), and VI) 2hrs after completion of training (~1130hrs).
-Finger draws were analysed with in-house equipment (BIOSEN C-Line, EKF diagnostic GmbH, Barleben).
-Venous blood samples were collected from the antecubital vein, coinciding with the finger draws except 45min after protein ingestion and in the middle of the RT session, to analyse endocrine variables.
+To measure blood glucose levels with and without glucose ingestion/training, capillary blood was collected from finger sticks on days with biopsy sampling.
+One capillary blood sample was taken on day 1 (Figure 1A) to serve as a baseline.
+On days 11 and 12 (Figure 1A), capillary blood samples were collected seven times: I) Immediately prior to protein ingestion (0700hrs) II) 45 minutes after protein ingestion (0745hrs) III) 1.5hrs after protein ingestion (0800hrs, i.e., immediately before GLU/PLAC intake), IV) 2hrs after protein ingestion (0900hrs, i.e., immediately before training), IV) in the middle of RT (~0915hrs), V) immediately after training (~0930hrs), and VI) 2hrs after completion of training (~1130hrs).
+Capillary blood samples were analysed with in-house equipment (BIOSEN C-Line, EKF diagnostic GmbH, Barleben).
+Venous blood samples were collected from the antecubital vein, coinciding with the capillary samples except 45min after protein ingestion and in the middle of the RT session, to analyse endocrine variables.
 
@@
 
-Two muscle biopsy samples were collected for total RNA extraction per leg per time point, and total RNA was extracted in two duplicates per muscle biopsy.
-Thus, we had two duplicates per leg per time point, a total of eight RNA samples per participant.
-Total RNA was extracted using TRIzol according to the manufacturer’s protocol.
-An exogenous RNA control (λ polyA External Standard Kit, Takara Bio Inc., Shiga, Japan) was added at a fixed amount (0.04 ng ml-1 of Trizol reagent) to enable analysis of target gene expression per unit tissue weight [@ellefsen_improved_2008; @ellefsen_reliable_2014].
-Briefly, muscle tissue was homogenized in TRIzol using 0.5mm RNase-free Zirconium beads (~50 ul) (Next Advanced, Averill Park, NY, USA) and a bullet blender (Bullet Blender, Next Advanced, Averill Park, NY, USA).
-Chloroform (Sigma-Aldrich, Missouri, USA) was used for phase separation, and an RNA pellet was precipitated with isopropanol (VWR International, Pennsylvania, USA).
-For assessment of RNA content and -purity 5µl RNA, from the RNA stock, was eluted in TE-buffer (1:2), for assessment via spectrophotometry.
-All samples had a 260nm to 280nm ratio > 1.9.
+Two muscle biopsy aliquots were used for total RNA extraction per leg and resulting in a total of eight RNA samples per participant.
+Total RNA was extracted using TRIzol with muscle tissue homogenized using 0.5mm RNase-free Zirconium beads (~50 ul; Next Advanced, Averill Park, NY, USA) and mechanical agitation (Bullet Blender, Next Advanced, Averill Park, NY, USA).
+Chloroform (Sigma-Aldrich, Oslo, Norway) was used for phase separation, and the RNA pellet was precipitated with isopropanol (VWR International, Oslo, Norway).
+To enable analysis of target gene expression per unit tissue weight [@ellefsen_improved_2008; @ellefsen_reliable_2014], an exogenous RNA control (λ polyA External Standard Kit, Takara Bio Inc., Shiga, Japan) was added at a fixed amount to each sample (0.04 ng ml^-1^ of Trizol reagent).
+For assessment of RNA content and purity, RNA was eluted in TE-buffer (1:2) and assessed via spectrophotometry.
+All samples had a 260 nm to 280 nm ratio > 1.9.
 The RNA stock was stored at -80°C until further analyses.
-RNA was reverse transcribed using Super Script IV Reverse Transcriptase (Invitrogen), according to the manufacturer’s instructions.
-One µl 10 mM dNTP mix, 0.5µl anchored oligo-dT, 0.5µl random hexamer primers (Thermo Scientific), a maximum of 9µl template RNA and 2 µl nuclease-free water were mixed, vortexed and briefly spun down.
-Samples were then heated at 65°C for 5 minutes, followed by at least 1-minute incubation on ice.
-For the next step, 4µl 5x SSIV Buffer, 1µl 100 mM DTT, 1 µl RNase OUT, and Super Script IV Reverse Transcriptase (Invitrogen) were mixed, vortexed and briefly spun down and added to the samples.
-Samples were incubated for 10 minutes at 23°C, 10 minutes at 50-55°C and 10 minutes at 80°C.
+
+Five-hundred ng of RNA was reverse transcribed using Super Script IV Reverse Transcriptase (Invitrogen, Oslo, Norway), according to the manufacturer's instructions using anchored oligo-dT and random hexamer primers (Thermo Scientific, Oslo, Norway).
 All samples were reverse transcribed and diluted to 1:50 before quantitative real-time polymerase chain reaction (qPCR).
-Lambda was used as an external reference gene, added in the RNA extraction (2µl per extraction).
-qPCR reactions were run on a fast-cycling real-time detection system (Applied Biosystems 7500 fast Real-Time PCR Systems, Life Technologies AS), with a total volume of 10 µl, containing 2 µl of cDNA, gene-specific primers (0.5 µM final concentration) and a commercial master mix (2X SYBR Select Master Mix, Applied Biosystems, Life Technologies AS) [@hammarstrom_benefits_2020].
-qPCR reactions consisted of 40 cycles (3 s 95°C denaturing and 30 s 60°C annealing) [@hammarstrom_benefits_2020].
-Raw fluorescence data were exported from the platform-specific software and amplification curves were modelled with a best-fit sigmoidal model using the qPCR package [@ritz_qpcr_2008] written for R [@hammarstrom_benefits_2020; @Rcore].
+qPCR reactions were run over 40 cycles (3 s 95°C denaturing and 30 s 60°C annealing) on a fast-cycling real-time detection system (Applied Biosystems 7500 fast Real-Time PCR Systems, Life Technologies AS), with a total reaction volume of 10 µl consisting of 2 µl of cDNA, gene-specific primers (0.5 µM final concentration) and a commercial master mix (2X SYBR Select Master Mix, Applied Biosystems, Life Technologies AS, Oslo, Norway).
+Raw fluorescence data was modelled with a best-fit sigmoidal model using the qPCR package [@ritz_qpcr_2008] written for R [@hammarstrom_benefits_2020; @Rcore].
 
@@
 Total protein was extracted using the Minute Total Protein Extraction Kit for Muscles (Invent Biotechnology), according to the manufacturer’s protocol, optimised for our lab.
-Wet muscle was freeze-dried for 24hrs and dissected before extraction.
-The tissue was homogenized with a plastic rod in 80mg protein extraction powder (Invent Biotechnology) and 100ul iced cell lysis buffer (Denaturing Buffer, Invent Biotechnology), and centrifuged at 19 000g for 1 min.
-The supernatant was divided into 2x25ul aliquots to run samples in duplicates, and total protein concentrations were determined in a 1:10 dilution (Pierce Detergent Compatible Bradford Assay Reagent, Thermo Fisher Scientific, Oslo, Norway).
-The protein samples were diluted to 20.25 µg µl−1 total protein with lysis buffer and 4X Laemmli sample buffer (Bio-Rad Laboratories AB, Oslo, Norway) containing 2-mercaptoethanol.
-Twenty microlitres from each protein sample were pooled as a normalization factor for gel separation.
+Wet muscle was freeze-dried for 24 hrs and dissected before extraction.
+The tissue was homogenized with a plastic rod in 80 mg protein extraction powder (Invent Biotechnology) and 100 ul ice cold cell lysis buffer (Denaturing Buffer, Invent Biotechnology), and centrifuged at 19 000 g for 1 min.
+The supernatant was divided into aliquots to run samples in duplicates, and total protein concentrations were determined in a 1:10 dilution (Pierce Detergent Compatible Bradford Assay Reagent, Thermo Fisher Scientific, Oslo, Norway).
+The protein samples were diluted to <!-- Check this, it should be 1.5?
+-->20.25 µg µl^−1^ total protein with lysis buffer and 4X Laemmli sample buffer (Bio-Rad Laboratories AB, Oslo, Norway) containing 2-mercaptoethanol.
 All protein samples were incubated at 95°C and stored at -20°C until further analysis.
-The protein samples (20.25µg total protein) were separated at 250V for 50min using 4-20% Tris-Glycine gels (Thermo Fisher Scientific), then transferred to PVDF membranes with wet transfer at 300mA for 3hrs.
+The protein samples (20.25 µg total protein) were separated at 250V on 4-20% Tris-Glycine gels (Bio-Rad Laboratories) for 50 min and then transferred to PVDF membranes with wet transfer at 300 mA for 3 hrs.
 Both gel electrophoresis and protein transfer were performed at 4°C.
-Following the wet transfer, membranes were stained with a MemCode reversible total protein stain (Thermo Fisher Scientific) and then blocked for 1hr at room temp with a blocking buffer of Tris-buffered saline (TBS; 20 mM Tris, 150 mM NaCl) with 5% non-fat dry milk and 0.1% Tween-20.
-Primary and secondary antibodies were purchased from Santa Cruz Biotechnology (Texas, USA) UBF:(UBF-9, sc-13125), rpS6:(Ribosomal protein S6 C-8, sc-74459), and Thermo Fisher Scientific (Oslo, Norway) c-Myc:(c-Myc 9E10), goat anti-mouse for c-Myc (goat anti-mouse IgG1 (y1) horseradish peroxidase conjugate), and anti-mouse (anti-mouse IgG1 horseradish peroxidase conjugate).
+Following the wet transfer, membranes were stained with a reversible total protein stain (Thermo Fisher Scientific) and then blocked for 1hr at room temp with a blocking buffer of Tris-buffered saline (TBS; 20 mM Tris, 150 mM NaCl) with 5% non-fat dry milk and 0.1% Tween-20.
+Primary and secondary antibodies were purchased from Santa Cruz Biotechnology (Texas, USA) UBF:(UBF-9, sc-13125), rpS6:(Ribosomal protein S6 C-8, sc-74459), and Thermo Fisher Scientific (Oslo, Norway) c-Myc:(c-Myc 9E10), goat anti-mouse for c-Myc (goat anti-mouse IgG1 (y1) horseradish peroxidase conjugate), and anti-mouse (anti-mouse IgG1 horseradish peroxidase conjugate)<!-- insert manufacturer -->.
 Antibodies were diluted in a 5% non-fat dry milk blocking buffer (TBS, 5% non-fat dry milk, 0.1% Tween-20).
-Membranes were incubated overnight with primary antibodies, and for 1hr with secondary antibodies.
-Between blocking and primary antibody staining, membranes were washed 1x5min, between primary and secondary staining, and after secondary staining, membranes were washed 3x5 with TBS-Tween (TBS; 20mM Tris, 150mM NaCl, 0.1% Tween).
-Following the last TBS-T wash, membranes were incubated 1x5min with enhanced chemiluminescent substrate (ECL, SuperSignal West Femto Maximum Sensitivity Substrate, Thermo Fisher Scientific).
-Membrane blocking, secondary antibody incubation, washing and ECL incubation were performed at room temp, on a tipping board.
-Primary antibody incubation was performed at 4°C on a tipping board.
-Chemiluminescence signals were quantified using Image Studio Lite (LI-COR Biotechnology, Lincoln, NE, USA), and total protein content was quantified using ImageJ [@rueden_imagej2_2017], whereny total protein content was defined as mean grey value of the whole well with between-well values subtracted as background.
-
-### Data handling and statistical analysis
-
-A priori power calculations showed that 20 participants would grant a statistical power of 80% (= 0.05), with an expected dropout of 20%.
-This power calculation was based on an assumption that the effects of glucose ingestion on ribosome biogenesis may equate to the effects of increasing RT volume from low to moderate [@hammarstrom_benefits_2020].
-<!-- maybe include the power calculation in extended methods?--> Raw data were imported and analyzed in R via R studios [@Rcore].
-Legs were categorized as either glucose or placebo in accordance with the randomization.
-Before quantitative analyses, total RNA was normalized by wet muscle weight, qPCR values were normalized by wet muscle weight and an external reference gene (Lambda), while quantified protein signals from western blot were normalized by pool signals and total protein signal per gel.<!-- add explanations of normalization to extended methods --> Total RNA and qPCR values were analyzed by comparing post to pre log-fold change score differences.
-This was achieved by modelling change as response with baseline and supplement (glucose/placebo) as explanatory factors, and mean-centred baseline values as a covariate in a linear mixed effects model, then calculating the estimated marginal means.
-Quantified western blot signals and the UBF/total RNA association were analyzed using a linear mixed effects model from the nlme package [@lindstrom_nonlinear_1990], with protein signal per time point as response, with time and supplement as explanatory factors.
-Plasma glucose and c-peptide, training volume, and the strength index were analyzed by multiple time-point log-fold change score comparisons, using the same linear mixed effects model with time added as explanatory factor.
-The linear mixed effects model was designed with the lmer function (Satterthwaite's method) from the lme4 package [@bates_fitting_2014] in R studios, using the lmerTest function to procure p-values.
-Log-transformation and mean-centring of baseline values were performed to control for heteroscedasticity and regression to the mean, respectively.
-Log-transformed values were reverse-transformed for figure illustrations.
-Values are presented as mean change + 95% confidence limits unless otherwise stated, with the alpha-level set to α = 0.05.
+Membranes were incubated overnight with primary antibodies, and for 1 hr with secondary antibodies.
+Between blocking and primary antibody staining, membranes were washed 5 min, between primary and secondary staining, and after secondary staining, membranes were washed 3 x 5 min with TBS-Tween (TBS; 20mM Tris, 150mM NaCl, 0.1% Tween).
+Following the last TBS-T wash, membranes were incubated 5 min with enhanced chemiluminescent substrate (ECL, SuperSignal West Femto Maximum Sensitivity Substrate, Thermo Fisher Scientific).
+Membrane blocking, secondary antibody incubation, washing and ECL incubation were performed at room temp.
+Primary antibody incubation was performed at 4°C.
+Chemiluminescence signals were quantified using Image Studio Lite (LI-COR Biotechnology, Lincoln, NE, USA), and total protein content was quantified using ImageJ [@rueden_imagej2_2017], where total protein content was defined as mean grey value of the whole well with between-well values subtracted as background.
+A pooled sample was used as a control on each gel to allow for between gel comparisons.
+
+### Statistics and data analysis
+
+A priori power calculations showed that 20 participants would grant a statistical power of 80% (alpha = 0.05), accounting for an expected dropout of 20%.
+This power calculation was based on an assumption that the effects of glucose ingestion on ribosome biogenesis (measured as within participant differences in total RNA) may equate to the effects of increasing RT volume from low to moderate [@hammarstrom_benefits_2020] corresponding to an effect size of d =  <!-- Add the effect size here-->.
+Before quantitative analyses, total RNA was normalized to wet muscle weight, qPCR values were normalized to wet muscle weight using the external reference gene (Lambda), while quantified protein signals from western blot were normalized to the pooled control sample and total protein.<!-- add explanations of normalization to extended methods -->
+Total RNA, protein and qPCR data were analyzed by comparing pre- to post-intervention log-transformed change-score differences between supplements.
+This was achieved by modelling change as the dependent variable with baseline-values and supplement (glucose/placebo) as independent variables in a linear mixed effects model giving each participant and leg a random intercept.
+Plasma glucose, c-peptide, training volume, and the strength index were analyzed by multiple time-point log-fold change score comparisons, using the same linear mixed effects model with time added as independent variable.
+The linear mixed effects models were fitted with the lmer function from the lme4 package using the lmerTest package to procure p-values (Satterthwaite's method for approximating degrees of freedom) [@bates_fitting_2014] or lme from the nlme package in R.
+Log-transformed values were expressed as fold-changes in presentations.
+Descriptive data are presented as mean and (SD).
+Inferential statistics are presented as means with 95% confidence intervals unless otherwise stated.
 
@@
 
-There was no difference in mean change of total training session volume between exercising with placebo or glucose, with an increase of 18 and 16% from session 1 to session 6, respectively (Table 2; Mean change, S2 (session 2): *p* = `r pvol.c2`, S3: *p* = `r pvol.c3`, S4: *p* = `r pvol.c4`, S5: *p* = `r pvol.c5`, S6: *p* = `r pvol.c6`).
-Both exercising with glucose and placebo led to significant increases in mean total training session volume at sessions four, five and six, compared to session 1 (Table 2; Mean change, S2 (session 2): *p* = `r pvol.s2`, S3: *p* = `r pvol.s3`, S4: *p* = `r pvol.s4`, S5: *p* = `r pvol.s5`, S6: *p* = 0.001).
-Regarding the dietary data, there were no differences in mean macronutrient intake (protein, fat, carbohydrates) or total calorie intake between glucose and placebo on pairwise consecutive days (*p* > 0.05 for all).
+There was no difference in mean change of total training session volume between exercising with placebo or glucose, with an increase of 18 and 16% from session 1 to session 6, respectively (Table 2).<!-- Add statistics to Table 2 -->
+There were no differences in mean macronutrient intake (protein, fat, carbohydrates) or total calorie intake between glucose and placebo on pairwise consecutive days (*p* > 0.05 for all).<!-- Should this also be in Table 2?
+-->
 
@@
 
-Glucose ingestion before and after RT led to significant increases in plasma levels of glucose by 38% immediately before RT (Figure 1B, 0 min: *p* = 0.000), by 31% during RT (Figure 1B, 15 min: *p* = 0.000) and by 32% immediately after RT (Figure 1B, 30 min: *p* = 0.000), compared to placebo ingestion before and after RT.
+Glucose ingestion before and after RT led to significant increases in plasma glucose levels by 38% immediately before RT (Figure 1B, 0 min), by 31% during RT (Figure 1B, 15 min) and by 32% immediately after RT (Figure 1B, 30 min; all : *p* < 0.001), compared to placebo.
 Two hours after the session, RT with glucose measured 8% lower plasma glucose levels compared to RT with placebo (Figure 1B, 270min: *p* = `r p.270`).
-Similarly, levels of c-peptide increased significantly more in exercises with glucose compared to placebo, by 95% immediately before (Figure 1C, 0 min: *p* = 0.000) and 87% after RT (Figure 1C, 30 min: *p* = 0.000).
+Similarly, levels of c-peptide increased with glucose ingestion over placebo, by 95% immediately before (Figure 1C, 0 min) and 87% after RT (Figure 1C, 30 min; both *p* < 0.001).
 
@@
 # Create model: 
-trna.lmod <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
+trna.lmod <- lmerTest::lmer(change ~ pre + supplement + (1|subject) , 
            data = trna.lchange)
@@
 
-# For inline reporting
-trcoef <- coef(summary(trna.lmod))
-trdiff <- round(trcoef[3,1], 3)
-trdiffp <- round(trcoef[3,5], 3)
-trdiffse <- round(trcoef[3,2], 2)
+### Function for inline reporting of total RNA and rRNA data
+
+stat_fun_rna <- function(model) {
+
+  coefs <- coef(summary(model))
+  ci <- confint(model)
+  meandiff <- round(100 * (exp(coefs[3,1]) -1), 1)
+  cilower <- round(100 * (exp(ci[5,1]) -1), 1)
+  ciupper <- round(100 * (exp(ci[5,2]) -1), 1)
+  pval <- round(coefs[3,5], 3)
+
+  stats <- paste0(meandiff, "%, [", cilower, ", ", ciupper,"], *p* = ", pval)
+
+  stats
+}
+
+total_rna_stats <- stat_fun_rna(trna.lmod)
 
@@
                      data = lchange.18)
-lres.18 <- plot(lmod.18)
-lsum.18 <- summary(lmod.18)
 
@@
                      data = lchange.28)
-lres.28 <- plot(lmod.28)
-lsum.28 <- summary(lmod.28)
 
@@
                      data = lchange.58)
-lres.58 <- plot(lmod.58)
-lsum.58 <- summary(lmod.58)
 
@@
                      data = lchange.5)
-lres.5 <- plot(lmod.5)
-lsum.5 <- summary(lmod.5)
 
@@
                      data = lchange.47)
-lres.47 <- plot(lmod.47)
-lsum.47 <- summary(lmod.47)
+
+# Visual checks of models
+
+plot(lmod.18)
+plot(lmod.28)
+plot(lmod.58)
+plot(lmod.5)
+plot(lmod.47)
 
@@
 
-# 18S
-lemm.18 <- confint(emmeans(lmod.18, specs = ~"supplement")) %>%
-        data.frame()
-
-# 28S
-lemm.28 <- confint(emmeans(lmod.28, specs = ~"supplement")) %>%
-        data.frame()
-
-# 5.8S
-lemm.58 <- confint(emmeans(lmod.58, specs = ~"supplement")) %>%
-        data.frame()
-
-# 5S
-lemm.5 <- confint(emmeans(lmod.5, specs = ~"supplement")) %>%
-        data.frame()
-
-# 47S
-lemm.47 <- confint(emmeans(lmod.47, specs = ~"supplement")) %>%
-        data.frame()
-
-## For inline reporting
-
-# 18S
-rr18coef <- coef(summary(lmod.18))
-rr18diffp <- round(rr18coef[3,5], 3)
-rr18diffse <- round(rr18coef[3,2], 3)
-
-# 28S
-rr28coef <- coef(summary(lmod.28))
-rr28diffp <- round(rr28coef[3,5], 3)
-rr28diffse <- round(rr28coef[3,2], 3)
-
-# 5.8S
-rr58coef <- coef(summary(lmod.58))
-rr58diffp <- round(rr58coef[3,5], 3)
-rr58diffse <- round(rr58coef[3,2], 3)
-
-# 5S
-rr5coef <- coef(summary(lmod.5))
-rr5diffp <- round(rr5coef[3,5], 3)
-rr5diffse <- round(rr5coef[3,2], 3)
-
-# 47S
-rr47coef <- coef(summary(lmod.47))
-rr47diffp <- round(rr47coef[3,5], 3)
-rr47tdiffp <- round(rr47coef[1,5], 3)
-rr47diffse <- round(rr47coef[3,2], 3)
-
-```
-
-Resistance training with glucose did not induce a higher accumulation of total RNA (Figure 2A, *p* = `r trdiffp`) or rRNA  (Figure 2B, 47S: *p* = `r rr47diffp`, 18S: *p* = `r rr18diffp`, 28S: *p* = `r rr28diffp`, 5.8S: *p* = `r rr58diffp`, 5S: *p* = `r rr5diffp`) compared to RT with placebo.
-From baseline to post-intervention, there was a mean increase in levels of total RNA by 26 and 22% after RT with glucose and placebo, respectively (*p* < 0.05, compared to baseline).
-A robust accumulation was also observed in ribosomal RNA expression in both RT with glucose and placebo, with mean increases between 34-43% (GLU) and 33-41% (PLAC) (*p* < 0.05, compared to baseline) in the four rRNAs.
-The expression of the 47S pre-rRNA increased by 37 and 59% with glucose and placebo respectively, where only RT with placebo increased significantly from baseline to post (Figure 2B, *p* = `r rr47tdiffp`).
+stat_rrna <- lapply(list(lmod.18, 
+                         lmod.28, 
+                         lmod.58, 
+                         lmod.5, 
+                         lmod.47), 
+       stat_fun_rna)
+
+```
+
+From baseline to post-intervention total RNA and rRNA increased on average by ~24% and ~33-43%, respectively (Figure 2).
+However, RT with glucose did not induce increased accumulation of total RNA (Figure 2A, mean difference `r total_rna_stats`) or rRNA  (Figure 2B; 47S, `r stat_rrna[[5]]`; 18S, `r stat_rrna[[1]]`; 28S, `r stat_rrna[[2]]`; 5.8S, `r stat_rrna[[3]]`; 5S, `r stat_rrna[[4]]`) compared to RT with placebo.
 

### 2023-06-01 R/totrna_cleanup.R

@@ -9,7 +9,7 @@
 ##
 ##-------------------------------------
 ## Notes:
-# Filter technical outliers by model log(rna) ~ muscle weight.
+# Filter outliers by model log(rna) ~ muscle weight.
 # Set threshold for filtering by adjusting threshold for standardized
 # residuals. A sensible threshold is 3 sd's from the modelled relationship
 # between RNA and muscle weight, taking training status into account.
@@ -24,7 +24,6 @@
 library(tidyverse); library(readxl); library(nlme)
 
 # A data frame with technical outliers
-
 
 outlier_tech <- data.frame(subject = 
                              c(101,
@@ -78,8 +77,7 @@ dat <- read_excel("./data/rna/RNA_raw.xlsx") %>%
 
 # Plot RNA vs. weight
 dat %>%
-  ggplot(aes(log(weight), log(RNA_tot), color = trained, 
-             shape = technical)) + geom_point()
+  ggplot(aes(log(weight), log(RNA_tot), color = trained)) + geom_point() + geom_smooth(method = "lm")
 
 
 # Model RNA to weight relationship using a linear model.
@@ -87,12 +85,6 @@ dat %>%
 
 m <- lm(log(RNA_tot) ~ log(weight) + trained,
         data = dat)
-
-
-m <- lme(log(RNA_tot) ~ log(weight) + trained,
-         random = list(subject = ~ 1,
-                       biopsy = ~ 1),
-         data = dat)
 
 
 summary(m)
@@ -104,7 +96,7 @@ dat$resid <- resid(m)
 # Plot the effect of outlier detection
 dat %>%
   mutate(resid = resid/sd(resid), 
-         outlier = if_else(resid < -2, "out", "in")) %>%
+         outlier = if_else(abs(resid) > 3, "out", "in")) %>%
   ggplot(aes(weight, log(RNA_tot), color = outlier)) +
   geom_point() +
   geom_smooth(method = "lm", aes(color = NULL))
@@ -112,7 +104,7 @@ dat %>%
 # Check the effect of outlier detection on included samples
 dat %>%
   mutate(resid = resid/sd(resid), 
-         outlier = if_else(resid < -2, "out", "in")) %>%
+         outlier = if_else(abs(resid) > 3, "out", "in")) %>%
   filter(outlier == "in") %>%
   group_by(subject) %>%
   summarise(n = n()/32)
@@ -121,7 +113,7 @@ dat %>%
 
[…]

### 2023-06-01 R/western_analysis.R

@@ -1,0 +1,324 @@
+##### Western analysis
+#
+#
+## Purpose: The purpose of this script is to analyse the cleaned up western data
+#
+# Author: KL/DH
+#
+# Packages
+library(tidyverse); library(nlme); library(emmeans); library(lme4)
+#
+# Data # from the script: western.cleanup.R
+prot.dat <- readRDS("./data/data-gen/protein/prot.datc.RDS") %>%
+  mutate(sample.id = paste0(gel.sorted, sample.id)) %>%
+  print()
+
+
+# Western data is analyzed as "maximal" random effects models  
+#
+
+m.cmyc.lmer <- lmer(log(norm.sign) ~ time * supplement + 
+                     (supplement*time|subject),
+                    
+              data = filter(prot.dat, target == "cmyc"))
+
+model <- m.cmyc.lmer
+summary(m.cmyc.lmer)
+
+
+### Updates model with new data ###
+m.rps6 <- update(m.cmyc.lmer, data = filter(prot.dat, target == "rps6"))
+m.ubf <- update(m.cmyc.lmer, data = filter(prot.dat, target == "ubf"))
+
+
+
+# Check assumptions
+
+
+check_cmyc <- mod_assume(m.cmyc.lmer) + cowplot::draw_label("c-myc", fontface = 'bold', x = 0.5, y = 1, hjust = 0.5, vjust = 1)
+check_rps6 <- mod_assume(m.rps6) + cowplot::draw_label("RPS6", fontface = 'bold', x = 0.5, y = 1, hjust = 0.5, vjust = 1)
+check_ubf <- mod_assume(m.ubf) + cowplot::draw_label("UBF", fontface = 'bold', x = 0.5, y = 1, hjust = 0.5, vjust = 1)
+
+### Save models 
+
+prot_models <- list(cmyc = m.cmyc.lmer, 
+                    ubf = m.ubf, 
+                    rps6 = m.rps6)
+
+saveRDS(prot_models, "./data/data-gen/protein/protein_lmer_models.RDS")
+
+
+
+
+
+
+
+
+
+# Western blot data analyzed as change scores -- 
+# This accounts for regression towards the mean in change-calculations
+# and a fully nested model can be fitted to account for the duplicate 
+# measures on the leg level.
+# Note that it is unclear how to deal with pairing of post/diff values and the covariate pre as
+# these are not linked. Aggregating throws away information that could be used as above in a 
[…]

### 2023-06-01 figures/figure2_alt.R

@@ -1,0 +1,153 @@
+### Ribose figure 2 ######
+#
+# This script creates figure 2, containing log-fold change data of total RNA and rRNA
+#
+# Packages 
+library(tidyverse); library(emmeans); library(cowplot); library(ggtext)
+   
+
+
+rrna_analysis_results <- readRDS("./data/data-gen/rna/rrna_analysis_results.RDS")
+
+
+rrna_foldchange <- rrna_analysis_results$fold_change %>%
+  mutate(target = gsub("trg_", "", target), 
+         target = paste0(toupper(target), " rRNA"), 
+         time = "post") %>%
+  add_row(time = "pre", supplement = "PLACEBO", estimate = 0, lower.CL = 0, upper.CL = 0, 
+          target = c("18S rRNA", 
+                       "28S rRNA" ,
+                       "47S rRNA" ,
+                       "5.8S rRNA",
+                       "5S rRNA", 
+                     "18S rRNA", 
+                     "28S rRNA" ,
+                     "47S rRNA" ,
+                     "5.8S rRNA",
+                     "5S rRNA")) %>%
+  add_row(time = "pre", supplement = "GLUCOSE", estimate = 0, lower.CL = 0, upper.CL = 0, 
+          target = c("18S rRNA", 
+                     "28S rRNA" ,
+                     "47S rRNA" ,
+                     "5.8S rRNA",
+                     "5S rRNA", 
+                     "18S rRNA", 
+                     "28S rRNA" ,
+                     "47S rRNA" ,
+                     "5.8S rRNA",
+                     "5S rRNA")) %>%
+  
+  mutate(estimate = (exp(estimate)), 
+         lower.CL =  (exp(lower.CL)),
+         upper.CL =  (exp(upper.CL))) %>%
+  
+  
+  print()
+  
+  
+
+# Designing the plot theme
+
+plot_theme <- theme(panel.grid.major = element_blank(),
+                    panel.grid.minor = element_blank(),
+                    panel.background = element_rect(fill = "lightblue", colour = NA),
+                    axis.line = element_line(colour = "black"))
+
+# Colours
+
+colors <- c("#d7191c",
+                     "#fdae61",
+                     "#abd9e9",
+                     "#2c7bb6")
+
+# Default textsizes
+labsize <- 8
+textsize <- 6
+htextsize <- 7
+atext <- 2
+labsize <- 8
+
[…]

### 2023-06-01 figures/figure3.R

@@ -5,7 +5,7 @@
 #
 # Packages
 library(emmeans); library(tidyverse); library(cowplot); library(magick); 
-library(ggpubr); library(nlme); library(ggtext)
+ library(nlme); library(ggtext)
 #
 # Data
 

### 2023-06-01 figures/figure3_alt.R

@@ -1,0 +1,506 @@
+### Ribose figure 3 ######
+#
+# This script produces the figures for cmyc, ubf and rps6
+#
+#
+# Packages
+library(emmeans); library(tidyverse); library(cowplot); library(magick); 
+
+library(nlme); library(ggtext)
+#
+# Data
+
+
+prot_models <- readRDS("./data/data-gen/protein/protein_lmer_models.RDS")
+
+
+cmyc <- readRDS("./data/data-gen/protein/cmyc.lemm.RDS")
+
+ubf <- readRDS("./data/data-gen/protein/ubf.lemm.RDS")
+
+rps6 <- readRDS("./data/data-gen/protein/rps6.lemm.RDS")
+
+joined.dat <- readRDS("./data/data-gen/rna.prot.RDS")
+
+# Images
+
+g3 <- cowplot::ggdraw() + cowplot::draw_image("./figures/archive/g3_115_annotated.png", scale = 0.9)
+g6 <- cowplot::ggdraw() + cowplot::draw_image("./figures/archive/g6_115_unannotated.png", scale = 0.9)
+
+ubf.x2.img <- cowplot::ggdraw() + cowplot::draw_image("./figures/archive/ubf_x2_115.png", scale = 0.8)
+cmyc.x2.img <- cowplot::ggdraw() + cowplot::draw_image("./figures/archive/cmyc_x2_115.png", scale = 0.8)
+rps6.x2.img <- cowplot::ggdraw() + cowplot::draw_image("./figures/archive/rps6_x2_115.png", scale = 0.8)
+
+# Colours
+
+colors <- c("#d7191c",
+            "#fdae61",
+            "#abd9e9",
+            "#2c7bb6")
+            
+# Textsizes
+
+labsize <- 8
+textsize <- 6
+htextsize <- 7
+legendti <- 6
+legendtex <- 5.25
+atext <- 2
+keysize <- 2
+
+# Designing the plot theme
+
+plot_theme <- theme(panel.grid.major = element_blank(),
+                    panel.grid.minor = element_blank(),
+                    panel.background = element_rect(fill = "lightblue", colour = NA),
+                    #plot.background = element_rect(fill = "lightblue", color = NA),
+                    axis.line = element_line(colour = "black"))
+
+## cmyc fig
+
+model <- prot_models$cmyc
+
+
+comp_plot <- function(model, method = "Wald") {
+  
+  ## Plot settings 
[…]

### 2023-06-02 manuscript.RMD
update manuscript

--- 2023-06-01 ribose-paper.postfeedback.RMD
@@
 ## Abstract
+
+**Introduction.** Muscle hypertrophy, induced by resistance training (RT), is determined by ribosomal biogenesis.
+Synthesis of ribosomes may augment the transcription of ribosomal RNA, as evident from in vitro studies.
+However, its effect on RT-induced adaptations remains unexplored.
+**Methods.** Sixteen healthy, moderately trained individuals (male/female, n = XX/XX; age, mean (sd)),<!-- Add male/female numbers, age--> participated in a within-participant trial with unilateral resistance training performed with glucose- and placebo supplement on alternate days over six sessions per condition.
+Participants remained fasted overnight, receiving glucose/placebo and protein supplementation only before and after RT sessions (3 sets of 10 repetitions maximum unilateral leg press and knee extension).
+Micro biopsies were sampled from m.
+vastus lateralis pre and post the intervention, and maximal unilateral isometric and isokinetic knee extension torque were measured before, during (days 4, 5, 8, 9) and after the intervention.
+**Results.** <!-- update statistics in this section -->RT led to robust increases in total RNA and rRNA in glucose and placebo (RNA: 26 and 22% respectively, rRNA: 33-43% and 33-41% respectively), with no differences between supplements (p > 0.05).
+There were no differences in content of c-Myc (*p* = 0.160), UBF (*p* = 0.154) and rpS6 (*p* = 0.386) between treatments in response to RT.
+A linear relationship was found between content of UBF and total RNA, where an increase of 1 SD unit in UBF equated to 6.8% increase in total RNA (*p* = 0.002).
+Lastly, there were no differences between supplements in maximal unilateral isometric knee extension force and isokinetic peak torque during the intervention (p > 0.05) and following the last session (p > 0.05), apart from the 60 d/s test before last RT session (p < 0.05).
+**Conclusion.** Glucose supplement before and after RT does not augment RT-induced ribosome biogenesis, nor does it enhance muscular performance and recovery after two weeks of heavy RT in moderately trained, young individuals.
 
@@
 ### Total RNA extraction and quantitative real-time reverse transcription polymerase chain reaction
+```{r totrna-extraction}
+
+totrna_complete <- readRDS("./data/data-gen/rna/rna_complete.RDS")
+
+samples_filtered <- totrna_complete %>%
+  group_by(subject, time, rep, technical, outlier) %>%
+  summarise(m = mean(RNA_tot)) %>%
+  group_by(technical, outlier) %>%
+  summarise(n = n())
+
+technical <- samples_filtered %>%
+  filter(technical == "out" & outlier == "in") %>%
+  pull(n)
+
+observed <- samples_filtered %>%
+  filter(technical == "in" & outlier == "out") %>%
+  pull(n)
+
+```
 
@@
 The RNA stock was stored at -80°C until further analyses.
+Before quantitative analyses of total RNA, samples with known loss of RNA during extraction (n = `r technical`) or a deviation from the observed RNA to muscle tissue weight relationship larger than 3 $\times$ residual SD while accounting for training status (n = `r observed`) were removed from the data set.
+Total RNA was normalized to wet muscle weight and log transformed before statistical analyses.
 
@@
 Raw fluorescence data was modelled with a best-fit sigmoidal model using the qPCR package [@ritz_qpcr_2008] written for R [@hammarstrom_benefits_2020; @Rcore].
+qPCR data was normalized to wet muscle weight using the external reference gene [@ellefsen_improved_2008; @ellefsen_reliable_2014] and analyzed on the log scale on a target-by-target basis.
+<!-- add explanations of normalization to extended methods -->
 
@@
 Following the wet transfer, membranes were stained with a reversible total protein stain (Thermo Fisher Scientific) and then blocked for 1hr at room temp with a blocking buffer of Tris-buffered saline (TBS; 20 mM Tris, 150 mM NaCl) with 5% non-fat dry milk and 0.1% Tween-20.
-Primary and secondary antibodies were purchased from Santa Cruz Biotechnology (Texas, USA) UBF:(UBF-9, sc-13125), rpS6:(Ribosomal protein S6 C-8, sc-74459), and Thermo Fisher Scientific (Oslo, Norway) c-Myc:(c-Myc 9E10), goat anti-mouse for c-Myc (goat anti-mouse IgG1 (y1) horseradish peroxidase conjugate), and anti-mouse (anti-mouse IgG1 horseradish peroxidase conjugate)<!-- insert manufacturer -->.
-Antibodies were diluted in a 5% non-fat dry milk blocking buffer (TBS, 5% non-fat dry milk, 0.1% Tween-20).
+Primary and secondary antibodies were purchased from Santa Cruz Biotechnology (Texas, USA): UBF, UBF F-9, sc-13125; rpS6, Ribosomal protein S6 C-8, sc-74459; and Thermo Fisher Scientific (Oslo, Norway): c-Myc, 9E10; goat anti-mouse (for c-Myc), goat anti-mouse IgG1 (y1) horseradish peroxidase conjugate; and anti-mouse (anti-mouse IgG1 horseradish peroxidase conjugate)<!-- insert manufacturer -->.
+Antibodies were diluted in blocking buffer to concentrations corresponding to <!-- Add final concentrations of antibodies -->.
 Membranes were incubated overnight with primary antibodies, and for 1 hr with secondary antibodies.
 Between blocking and primary antibody staining, membranes were washed 5 min, between primary and secondary staining, and after secondary staining, membranes were washed 3 x 5 min with TBS-Tween (TBS; 20mM Tris, 150mM NaCl, 0.1% Tween).
-Following the last TBS-T wash, membranes were incubated 5 min with enhanced chemiluminescent substrate (ECL, SuperSignal West Femto Maximum Sensitivity Substrate, Thermo Fisher Scientific).
+Following the last wash, membranes were incubated 5 min with enhanced chemiluminescent substrate (ECL, SuperSignal West Femto Maximum Sensitivity Substrate, Thermo Fisher Scientific).
 Membrane blocking, secondary antibody incubation, washing and ECL incubation were performed at room temp.
@@
 Chemiluminescence signals were quantified using Image Studio Lite (LI-COR Biotechnology, Lincoln, NE, USA), and total protein content was quantified using ImageJ [@rueden_imagej2_2017], where total protein content was defined as mean grey value of the whole well with between-well values subtracted as background.
-A pooled sample was used as a control on each gel to allow for between gel comparisons.
+A pooled sample was used as a control on each gel to allow for between gel comparisons and  quantified protein signals were subsequently normalized to the pooled control sample and total protein.
 
@@
 This power calculation was based on an assumption that the effects of glucose ingestion on ribosome biogenesis (measured as within participant differences in total RNA) may equate to the effects of increasing RT volume from low to moderate [@hammarstrom_benefits_2020] corresponding to an effect size of d =  <!-- Add the effect size here-->.
-Before quantitative analyses, total RNA was normalized to wet muscle weight, qPCR values were normalized to wet muscle weight using the external reference gene (Lambda), while quantified protein signals from western blot were normalized to the pooled control sample and total protein.<!-- add explanations of normalization to extended methods -->
-Total RNA, protein and qPCR data were analyzed by comparing pre- to post-intervention log-transformed change-score differences between supplements.
-This was achieved by modelling change as the dependent variable with baseline-values and supplement (glucose/placebo) as independent variables in a linear mixed effects model giving each participant and leg a random intercept.
-Plasma glucose, c-peptide, training volume, and the strength index were analyzed by multiple time-point log-fold change score comparisons, using the same linear mixed effects model with time added as independent variable.
-The linear mixed effects models were fitted with the lmer function from the lme4 package using the lmerTest package to procure p-values (Satterthwaite's method for approximating degrees of freedom) [@bates_fitting_2014] or lme from the nlme package in R.
+Total RNA, protein and qPCR data were analyzed by mixed-effects models with fixed effects included as *supplement $\times$ time*.
+To decrease the risk of Type I errors, random effects were selected from step-wise elimination of terms from the most complex structure (random slopes for time and supplement and their interaction) to less complex.
+The most complex random effects structure that converged was chosen as the final model[@Matuschek_2017].
+Plasma glucose, c-peptide, training volume, and the strength index were analyzed by multiple time-point log-fold change score comparisons, using mixed-effects model with baseline values, time and supplement, and the interaction between time and supplement as fixed effects.
+These data only supported a random intercept per participant.
+The linear mixed effects models were fitted with the lmer function from the lme4 package using the lmerTest package to procure p-values (Satterthwaite's method for approximating degrees of freedom)[@bates_fitting_2014; @lmerTest] written for R[@Rcore].
 Log-transformed values were expressed as fold-changes in presentations.
@@
 # Untransformed 
-vol.mod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
+vol.mod <- lmerTest::lmer(change ~ 0 + scale(baseline) + time + supplement:time + (1|subject),
                      data = vol.change)
@@
 # Log-transformed
-vol.lmod <- lmerTest::lmer(change ~ 0 + baseline + time + supplement:time + (1|subject),
+vol.lmod <- lmerTest::lmer(change ~ 0 + scale(baseline) + time + supplement:time + (1|subject),
                      data = vol.lchange)
@@
 
-# Data
-trna.lchange <- readRDS("./data/data-gen/rna/totrna.change.RDS")
-
-# Create model: 
-trna.lmod <- lmerTest::lmer(change ~ pre + supplement + (1|subject) , 
-           data = trna.lchange)
-
-trna.lres <- plot(trna.lmod)
-trna.lsum <- summary(trna.lmod)
-
-### Get estimated means from the model
-trna.emm <- confint(emmeans(trna.lmod, specs = ~"supplement")) %>%
-    data.frame()
+## Get results from totrna_analysis.R
+
+totrna_results <- readRDS("./data/data-gen/rna/tot_rna_coefs.RDS")
+
+# retrieve model
+trna.lmod <- totrna_results$model
 
@@
   ci <- confint(model)
-  meandiff <- round(100 * (exp(coefs[3,1]) -1), 1)
-  cilower <- round(100 * (exp(ci[5,1]) -1), 1)
-  ciupper <- round(100 * (exp(ci[5,2]) -1), 1)
-  pval <- round(coefs[3,5], 3)
+  meandiff <- sprintf("%.1f", 100 * (exp(coefs[4,1]) -1))
+  cilower <- sprintf("%.1f",100 * (exp(ci[6,1]) -1))
+  ciupper <- sprintf("%.1f",100 * (exp(ci[6,2]) -1))
+  pval <- sprintf("%.3f",coefs[4,5])
 
@@
 ## Scripts
-# - qpcr-import
-# - qpcr-analysis
-# - rrna.change
-# - rrnachange.figs
-
-# Log-transformed data
-lchange.18 <- readRDS("./data/data-gen/rna/lchange.18.RDS")
-lchange.28 <- readRDS("./data/data-gen/rna/lchange.28.RDS")
-lchange.58 <- readRDS("./data/data-gen/rna/lchange.58.RDS")
-lchange.5 <- readRDS("./data/data-gen/rna/lchange.5.RDS")
-lchange.47 <- readRDS("./data/data-gen/rna/lchange.47.RDS")
-
-# Create model: 
-# Needs to have an intercept per participant (mixed model)
-# Control for pre values.
-
-# 18S
-lmod.18 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
-                     data = lchange.18)
-
-# 28S
-lmod.28 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
-                     data = lchange.28)
-
-# 5.8S
-lmod.58 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
-                     data = lchange.58)
-
-# 5S
-lmod.5 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
-                     data = lchange.5)
-
-# 47S
-lmod.47 <- lmerTest::lmer(change ~ pre + supplement + (1|subject), 
-                     data = lchange.47)
-
-# Visual checks of models
-
-plot(lmod.18)
-plot(lmod.28)
-plot(lmod.58)
-plot(lmod.5)
-plot(lmod.47)
-
-### Get estimated means from the model, these are average increase at 
-# pre = 0 (the average pre value)
-# remember that these are originally log-fold change values, transformed back using exp(emmean) etc
-
-stat_rrna <- lapply(list(lmod.18, 
-                         lmod.28, 
-                         lmod.58, 
-                         lmod.5, 
-                         lmod.47), 
+# - qpcr_import
+# - qpcr_analysis
+
+# Retrieve models from qpcr_analysis.R
+
+rrna_analysis_results <- readRDS("./data/data-gen/rna/rrna_analysis_results.RDS")
+
+# Get estimated means from the model, these are average baseline-corrected log-fold differences 
+# at post.
+
+# Applying the stat_fun_rna function to models
+stat_rrna <- lapply(rrna_analysis_results$rrna_models, 
        stat_fun_rna)
@@
 
-From baseline to post-intervention total RNA and rRNA increased on average by ~24% and ~33-43%, respectively (Figure 2).
-However, RT with glucose did not induce increased accumulation of total RNA (Figure 2A, mean difference `r total_rna_stats`) or rRNA  (Figure 2B; 47S, `r stat_rrna[[5]]`; 18S, `r stat_rrna[[1]]`; 28S, `r stat_rrna[[2]]`; 5.8S, `r stat_rrna[[3]]`; 5S, `r stat_rrna[[4]]`) compared to RT with placebo.
+From baseline to post-intervention total RNA and rRNA per unit muscle weight increased on average by ~24% and ~33-43%, respectively (Figure 2).
+However, RT with glucose did not induce increased accumulation of total RNA (Figure 2A, mean difference `r total_rna_stats`) or rRNA  (Figure 2B; 47S, `r stat_rrna$trg_47s`; 18S, `r stat_rrna$trg_18s`; 28S, `r stat_rrna$trg_28s`; 5.8S, `r stat_rrna$trg_5.8s`; 5S, `r stat_rrna$trg_5s`) compared to RT with placebo.
 
@@
 
-# Data
-prot.dat <- readRDS("./data/data-gen/protein/prot.datc.RDS") # from the script: western.cleanup.R
-
-### Modelling target signal per supplement 
-# These models try to explain the normalised signal by time, supplement and time+supplement (?). The random = list command creates an individual 
-# intercept for the included (e.g., subject, gel.sample).
-
-## c-Myc model
-
-# Untransformed
-cmyc.mod <- lme(norm.sign ~ time + supplement + time:supplement, 
-                random = list(subject = ~ 1),
-                data = filter(prot.dat, target == "cmyc"))
-
-cmyc.sum <- summary(cmyc.mod)
-cmyc.res <- plot(cmyc.mod, resid(., type = "p") ~ fitted(.))
-
-# Log-transformed
-cmyc.lmod <- lme(log(norm.sign) ~ time + supplement + time:supplement, 
-                 random = list(subject = ~ 1),
-                 data = filter(prot.dat, target == "cmyc"))
-
-cmyc.lsum <- summary(cmyc.lmod)
-cmyc.lres <- plot(cmyc.lmod, resid(., type = "p") ~ fitted(.))
-
-## UBF model
-# Untransformed
-ubf.mod <- lme(norm.sign ~ time + supplement + time:supplement, 
-               random = list(subject = ~ 1),
-               data = filter(prot.dat, target == "ubf"))
-
-ubf.sum <- summary(ubf.mod)
-ubf.res <- plot(ubf.mod, resid(., type = "p") ~ fitted(.))
-
-# Log-transformed
-ubf.lmod <- lme(log(norm.sign) ~ time + supplement + time:supplement, 
-                random = list(subject = ~ 1),
-                data = filter(prot.dat, target == "ubf"))
-
-ubf.lsum <- summary(ubf.lmod)
-ubf.lres <- plot(ubf.lmod, resid(., type = "p") ~ fitted(.))
-
-# RPS6 model
-# Untransformed
-rps6.mod <- lme(norm.sign ~ time + supplement + time:supplement, 
-                 random = list(subject = ~ 1),
-                 data = filter(prot.dat, target == "rps6"))
-
-rps6.sum <- summary(rps6.mod)
-rps6.res <- plot(rps6.mod, resid(., type = "p") ~ fitted(.))
-
-# Log-transformed
-rps6.lmod <- lme(log(norm.sign) ~ time + supplement + time:supplement, 
-          random = list(subject = ~ 1),
-          data = filter(prot.dat, target == "rps6"))
-
-rps6.lsum <- summary(rps6.lmod)
-rps6.lres <- plot(rps6.lmod, resid(., type = "p") ~ fitted(.))
-
-### Getting emmeans from the models
-
-## c-Myc emmeans
-# Untransformed
-cmyc.emm <- emmeans(cmyc.mod, specs = ~ time|supplement) %>%
-  data.frame()
-
-# Log-transformed
-cmyc.lemm <- emmeans(cmyc.lmod, specs = ~ time|supplement) %>%
-  data.frame()
-
-## UBF emmeans
-# Untransformed
-ubf.emm <- emmeans(ubf.mod, specs = ~ time|supplement) %>%
-  data.frame()
-
-# Log-transformed
-ubf.lemm <- emmeans(ubf.lmod, specs = ~ time|supplement) %>%
-  data.frame()
-
-## RPS6
-# Untransformed
-rps6.emm <- emmeans(rps6.mod, specs = ~ time|supplement) %>%
-  data.frame()
-
-# Log-transformed
-rps6.lemm <- emmeans(rps6.lmod, specs = ~ time|supplement) %>%
-  data.frame()
-
-### For inline reporting
-## Diff cmyc
-
-ccoef <- coef(summary(cmyc.lmod))
-cdiff <- round(ccoef[4,5], 3)
-cpla <- round(ccoef[2,5], 20)
-
-## Diff UBF
-ucoef <- coef(summary(ubf.lmod))
-udiff <- round(ucoef[4,5], 3)
-
-## Diff rps6
-rpcoef <- coef(summary(rps6.lmod))
-rpdiff <- round(rpcoef[4,5], 3)
+# Data and models
+
+prot_models <- readRDS("./data/data-gen/protein/protein_lmer_models.RDS")
+
+model <- prot_models$prot_models$cmyc
+
+## extract estimated change
+west_estimate <- function(model, p = FALSE) {
+
+  # Return absolute values
+
+  if(isFALSE(p)){
+
+  est <- round(100 * (exp(coef(summary(model))[4,1])-1),0)
+
+  return(est)
+  }
+
+  if(p){
+
+   pval <- sprintf("%.3f", coef(summary(model))[4,5])
+
+    return(pval)
+  }
+
+}
+
+stat_west <- lapply(prot_models$prot_models, west_estimate)
+p_west <- unlist(lapply(prot_models$prot_models, west_estimate, p = TRUE))
+
+### Linear relationship between UBF and Total RNA
+
+joined.dat <- readRDS("./data/data-gen/rna.prot.RDS")
+
+mod <- lmerTest::lmer(log(mean.rna) ~ sd.ubf + time +  (1|subject), 
+            data = joined.dat)
+
+corr_incr <- round(100 * (exp(coef(summary(mod))[2,1])-1),1)
+corrp <- sprintf("%.3f", coef(summary(mod))[2,5])
 
@@
 rna.rdy <- rna.dat %>%
-  select(subject, time, outlier, supplement, RNA.weight) %>%
-  filter(outlier == "in") %>%
+  select(subject, time, supplement, RNA.weight) %>%
+
   mutate(subject = as.character(subject)) %>%
@@
 
-library(nlme)
-
-corr.mod <- lme(mean.rna ~ mean.sign + time , 
-         random = list(subject = ~ 1), 
+library(lme4)
+
+corr.mod <- lmerTest::lmer(mean.rna ~ mean.sign + time + (1|subject), 
           data = joined.dat, 
@@
 
-corr.lmod <- lme(log(mean.rna) ~ log(mean.sign) + time , 
-         random = list(subject = ~ 1), 
+corr.lmod <- lmerTest::lmer(log(mean.rna) ~ log(mean.sign) + time + (1|subject),
+
           data = joined.dat, 
@@
 corrp <- round(corrcoef[2,5], 3)
-```
-
-Levels of c-Myc increased by 42% more after RT with placebo compared to RT with glucose, a significant difference (Figure 3A: *p* = `r cdiff`).
-Levels of UBF and rpS6 increased by 19 and 23% more in placebo compared to glucose, with no significant differences between groups (Figure 3A: UBF: *p* = `r udiff`, rpS6: *p* = `r rpdiff`).
-The increase in both RT with glucose and placebo for all measured proteins was significant at post compared to baseline (*p* < 0.05).
-Further, we found a linear relationship between UBF and total RNA, where an increase of 1 SD unit of UBF equated to a 13% increase in total RNA (*p* = `r corrp`).
+
+## Percentage increase in RNA per 1 SD uinit increase in UBF using log-model to show relative increse
+corr_incr <- round(100 * (exp(coef(summary(corr.lmod))[2,1])-1), 0)
+
+```
+
+Overall, RT led to robust increases in all measured proteins (Figure 3A).
+RT with glucose resulted in estimated levels of c-Myc, UBF and RPS6 being `r stat_west$cmyc`, `r stat_west$ubf` and `r stat_west$rps6`% lower compared to placebo, respectively, without showing statistical significance (*p* = `r min(p_west)`-`r max(p_west)`; Figure 3A).
+There was a linear relationship between UBF and total RNA, where an increase of 1 SD unit of UBF equated to a $\sim$ `r corr_incr`% increase in total RNA (*p* = `r corrp`; Figure 3C).
 

### 2023-06-02 R/qpcr_analysis.R

@@ -7,8 +7,7 @@
 ## Project: Ribose
 #
 #
-## Purpose: This script analyses the cleaned up qpcr data by change-score
-# comparisons
+## Purpose: This script analyses the cleaned up qpcr data by a maximal mixed-model
 #
 #
 ## Asscociated scripts: qpcr.import.R
@@ -18,7 +17,9 @@
 
 library(tidyverse)
 library(readxl)
-library(nlme)
+library(lmerTest)
+library(buildmer)
+
 
 # Load data
 
@@ -109,42 +110,27 @@ models_final <- list()
 
 for(i in 1:length(targets)) {
   
-  m1 <- lmerTest::lmer(nf.expr ~ time * supplement + 
+  ## Using buildmer to get the maximal model
+  max_model <- buildmer(nf.expr ~ time * supplement + 
                   (time*supplement|subject),
                 
-                data = filter(qpcr_data2, target == targets[i]))
-  
-  m2 <- lmerTest::lmer(nf.expr ~ time * supplement + 
-               (time + supplement|subject),
-             
-             data = filter(qpcr_data2, target == targets[i]))
-  
-  m3 <- lmerTest::lmer(nf.expr ~ time * supplement + 
-               (supplement|subject),
-             
-             data = filter(qpcr_data2, target == targets[i]))
-  
-  m4 <- lmerTest::lmer(nf.expr ~ time * supplement + 
-               (time|subject),
-             
-             data = filter(qpcr_data2, target == targets[i]))
-  
-  m5 <- lmerTest::lmer(nf.expr ~ time * supplement + 
-               (1|subject),
-             
-             data = filter(qpcr_data2, target == targets[i]))
+                data = filter(qpcr_data2, target == targets[i]), 
+                buildmerControl=buildmerControl(direction='order',
+                                                args=list(control=lmerControl(optimizer='bobyqa')), 
+                                                include = ~ time * supplement +  (1|subject), 
+                                                
+                                                ddf = "Satterthwaite"))
   
   
-  models <- list(m1 = m1, m2 = m2, m3 = m3, m4 = m4, m5 = m5)
   
-
   
-  ## Using the custom function 
-  model <- return_best(models)
-
+  
   ## Get within condition confidence intervals
- 
-  models_final[[i]] <- model
+  models_final[[i]] <- max_model@model
+  
+  ## Convert to lmerTest model by extracting from the buildmer object
+  model <- max_model@model
+  
 
   names(models_final)[i] <- targets[i]
   

### 2023-06-02 R/totrna_cleanup.R

@@ -3,7 +3,7 @@
 ##
 ## Title: RNA cleanup
 ## Purpose: Filter away bad RNA estimates based on RNA/weight relationship
-## Author: DH
+## Author: DH/KL
 ##
 ##
 ##
@@ -11,7 +11,7 @@
 ## Notes:
 # Filter outliers by model log(rna) ~ muscle weight.
 # Set threshold for filtering by adjusting threshold for standardized
-# residuals. A sensible threshold is 3 sd's from the modelled relationship
+# residuals. A sensible threshold is 3 sd's from the modeled relationship
 # between RNA and muscle weight, taking training status into account.
 #
 #
@@ -90,7 +90,7 @@ m <- lm(log(RNA_tot) ~ log(weight) + trained,
 summary(m)
 
 # Calculate residuals
-# Thjese are raw residuals, they are standardized below.
+# These are raw residuals, they are standardized below.
 dat$resid <- resid(m)
 
 # Plot the effect of outlier detection
@@ -113,7 +113,7 @@ dat %>%
 
 dat_complete <- dat %>%
   mutate(resid = resid/sd(resid), 
-         outlier = if_else(abs(resid) > 2, "out", "in")) %>%
+         outlier = if_else(abs(resid) > 3, "out", "in")) %>%
   print()
 
 

### 2023-06-02 figures/figure3_alt.R

@@ -14,6 +14,9 @@ library(nlme); library(ggtext)
 prot_models <- readRDS("./data/data-gen/protein/protein_lmer_models.RDS")
 
 
+prot_models <- prot_models$prot_models
+
+
 cmyc <- readRDS("./data/data-gen/protein/cmyc.lemm.RDS")
 
 ubf <- readRDS("./data/data-gen/protein/ubf.lemm.RDS")
@@ -58,9 +61,6 @@ plot_theme <- theme(panel.grid.major = element_blank(),
 
 ## cmyc fig
 
-model <- prot_models$cmyc
-
-
 comp_plot <- function(model, method = "Wald") {
   
   ## Plot settings 
@@ -87,7 +87,7 @@ comp_plot <- function(model, method = "Wald") {
   
   ## Plot raw fold changes
   raw_fold <- model@frame %>%
-    pivot_wider(names_from = time, values_from = 'log(norm.sign)', values_fn = mean) %>%
+    pivot_wider(names_from = time, values_from = 'ln.norm.sign', values_fn = mean) %>%
   mutate(change = post - pre) %>%
     group_by(subject, supplement) %>%
     summarise(change = exp(mean(change))) 

### 2023-06-08 manuscript.RMD
almost done!

@@
 ## Abstract
+
+```{r}
+#| echo: false
+
+###
+
+```
 
@@
 **Keywords: **
+
 </div>
@@
 These data only supported a random intercept per participant.
-The linear mixed effects models were fitted with the lmer function from the lme4 package using the lmerTest package to procure p-values (Satterthwaite's method for approximating degrees of freedom)[@bates_fitting_2014; @lmerTest] written for R[@Rcore].
+The linear mixed effects models were fitted with the lmer function from the lme4 package using the lmerTest package to procure p-values (Satterthwaite's method for approximating degrees of freedom)[@bates_fitting_2014; @lmerTest], written for R[@Rcore].
 Log-transformed values were expressed as fold-changes in presentations.
 Descriptive data are presented as mean and (SD).
-Inferential statistics are presented as means with 95% confidence intervals unless otherwise stated.
+Inferential statistics are presented as means with 95% confidence intervals and *p*-values unless otherwise stated.
+*p* < 0.05 was considered statistically significant.
 
@@
 
+pval_report <- function(p) {
+
+  if(p < 0.001) {
+
+    "< 0.001"
+
+  } else {
+    sprintf("%.3f", p)
+
+  }
+
+}
+
 # For inline plotting
@@
 pscient.120 <- round(bcoef[8,5], 30)
-p.120 <- sprintf("%.5f", 1.700971e-16)
+p.120 <- pval_report(bcoef["timechange.120:supplementglucose", 5])
 pscient.135 <- round(bcoef[9,5], 30)
-p.135 <- sprintf("%.5f", 4.299568e-11)
+p.135 <- pval_report(bcoef["timechange.135:supplementglucose", 5])
 pscient.150 <- round(bcoef[10,5], 30)
-p.150 <- sprintf("%.5f", 1.445712e-11)
+p.150 <-  pval_report(bcoef["timechange.150:supplementglucose", 5])
 p.270 <- round(bcoef[11,5], 3)
@@
 ## Script: 
-# - western.import.R
-# - western.cleanup.R
-# - western.analysis.R
+# - western_import.R
+# - western_analysis.R
 
@@
 
-Overall, RT led to robust increases in all measured proteins (Figure 3A).
+Overall, RT led to robust increases from pre- to post-intervention in all measured proteins (Figure 3A).
 RT with glucose resulted in estimated levels of c-Myc, UBF and RPS6 being `r stat_west$cmyc`, `r stat_west$ubf` and `r stat_west$rps6`% lower compared to placebo, respectively, without showing statistical significance (*p* = `r min(p_west)`-`r max(p_west)`; Figure 3A).

### 2023-06-08 R/western_analysis.R

@@ -9,8 +9,9 @@
 library(tidyverse); library(emmeans); library(lme4); library(buildmer)
 
 #
-# Data # from the script: western.cleanup.R
-prot.dat <- readRDS("./data/data-gen/protein/prot.datc.RDS") %>%
+# Data 
+# from the script: western_import.R
+prot.dat <- readRDS("./data/data-gen/protein/prot.dat.RDS") %>%
   mutate(sample.id = paste0(gel.sorted, sample.id)) %>%
   print()
 

### 2023-06-08 README.rmd

@@
 
-# Ribose-paper
+# Repository organization and extended methods: *Glucose ingestion before and after resistance training does not augment ribosome biogenesis in healthy moderately trained young adults*
 
@@
 <a href="https://orcid.org/0000-0002-5169-5960">
+<img alt="ORCID logo" src="https://info.orcid.org/wp-content/uploads/2019/11/orcid_16x16.png" width="16" height="16" /></a>,
+Sara Christine Moen,
+Håvard Hamarsland
+<a href="https://orcid.org/0000-0002-7343-2774">
 <img alt="ORCID logo" src="https://info.orcid.org/wp-content/uploads/2019/11/orcid_16x16.png" width="16" height="16" /></a>,
@@
 <img alt="ORCID logo" src="https://info.orcid.org/wp-content/uploads/2019/11/orcid_16x16.png" width="16" height="16" /></a>,
-Håvard Hamarsland
-<a href="https://orcid.org/0000-0002-7343-2774">
-<img alt="ORCID logo" src="https://info.orcid.org/wp-content/uploads/2019/11/orcid_16x16.png" width="16" height="16" /></a>,
-Sara Christine Moen
+
+Section for Health and Exercise Physiology, Department of Public Health and Sport Sciences, Inland Norway University of Applied Sciences, Lillehammer, Norway.
 
@@
 - Old/unused scripts
+
+## Extended method descriptions
+
+### Tissue sampling
+
+We gathered muscle tissue from both legs at pre and post intervention, taking two samples from the same incision.
+Each piece of muscle tissue was split in two, i.e.
+we had two samples for RNA analyses and two samples for protein analyses.
+We analyzed the samples in duplicates, i.e.
+we had two samples per time point per leg, a total of eight samples per analysis.
+<!-- check if there is more tissue left --> These samples were then randomly labelled with participant number and a number between 1-8, to blind the investigators conducting the laboratory analyses.
+
+### Total RNA and rRNA
+
+#### Extraction of RNA from muscle tissue (protocol)
+
+##### Materials
+
+For extraction of RNA and protein per sample:
+
+- 1.5 ml Eppendorf tube with 0.5 mm RNase-free Zirconium beads (~50 ul) Keep out DNA RNAse contaminations (Eppendorf safe-lock for use in bullet blender.
+- 1.5 ml tube for RNA precipitation
+- 1.5 ml tube for RNA quantification (1:5 dilution, add 20 ul of DEPC-water)
+- 1.5 ml tubes for Protein precipitation
+- 1.5 ml tubes for Soluble protein fraction
+- 1.5 ml tubes for protein quantification (1:10 dilution, add 36 ul of ddH2O)
+
+##### Solutions and chemicals
+
+- Trizol (Life-technologies)
+- Chloroform
+- Isopropanol
+- External reference gene (Takara lambda)
+- 100% EtOH
+- 95% EtOH
+- 75% EtOH (kept on -20)
+
+##### Homogenization and phase separation
+
+1.
+Add 1 sample volume of beads and 300 ul Trizol to eppendorf 1.5 ml tubes.
+Keep on ice.
+OPTIONAL: If external reference gene is used, add 1 unit to every extraction with Trizol in first step.
+Lambda (0.04 ng/ul, 2 ul per extraction)
+2. Add frozen muscle samples (protocol optimized for 5-25 ug of wet weight muscle sample) without thawing and put tubes in Bullet blender
+3. Run Bullet Blender (BB) for 1 min at speed 10, put samples on ice for 1 min. Run samples for 1 min at speed 12. Check samples, if not dissolved run until fully homogenized.
+4. Add Trizol (700 ul) to total volume 1000 ul, run samples in BB for 1 min speed 3.
+5. Incubate @ RT for 5 min
+6. Add 200 ul chloroform, shake 15 sec, incubate for 2-3 min.
+7. Spin 12000g, 15min, 4°C
+8.
+Pipette 450 ul of the aqueous phase to a new tube.
+Be careful not to disrupt the interphase or organic phase.
+Add 500 ul of isopropanol to precipitate RNA, incubate over night at -20°C or 10 min at room temp and continue with RNA extraction.
+9. Completely remove the remaining aqueous phase, either store remaining phases at -20°C until protein extraction or continue right away
+
+##### Extraction
+
+Precipitation and wash
+
+1. Spin the isopropanol/water phase 12000g, 10min, 4°C. A pellet should form, decant supernatant, spin briefly and remove remaining liquids.
+2. Add 1000 ul 75% EtOH (-20°C) @ RT, vortex gently. Spin 7500g, 5min, 4°C.
+3. Repeat step 2 two more times.
+4. Remove all EtOH, air dry the pellet up to 10 min with tubes upside-down. Do not over-dry!
+
+Suspension
+1. Add 20-40 ul of 1x DEPC-treated water (55°C), incubate at 55°C for 10min.
+2. Aliquot:
+ 5 ul to 20 ul TE buffer for Nanodrop/Qubit2
+ x ul in stock solution
+3. Store RNA stock solution @ -80°C.
+
+#### cDNA synthesis using Superscript IV and general qPCR
+
+##### cDNA
+
+##### Materials
+
+- Heating block or thermal-cycler capable of 65, 50-55 and 80°C
+- RNase-free tubes, pipette tips
+- Superscript IV kit (SSIV Buffer, SSIV Reverese transcriptase, 100 nM DTT)
+- Oligo dT and/or Random Hexamer primers
+- 10 mM dNTP mix
+- Total-RNA or mRNA
+
+1. Combine the following components per reaction (consider creating a master mix from component a and b if you are running many samples, master mix 1):
+ 25 μM Oligo dT and 25 μM Random Hexamers - 0.5 μl Oligo dT (art nr. AB-1247) + 0.5 μl Random hexamers (SO142) per reaction
+ 10 mM dNTP mix (10 mM each) - 1 μl per reaction
+ Template RNA (0.1-5000 ng total-RNA) - up to 11 μl per reaction
+ Nuclease-free water - Add to total volume 13 μl per reaction
+
+2. Mix all components, vortex and briefly spin down.
+3. Heat the mix 5 min at 65°C followed by at least 1 min incubation on ice.
+4. Bring the 5x SSIV buffer to room-temp. Vortex and spin briefly.
+5. Combine the following into master-mix 2:
+ 5x SSIV Buffer - 4 μl per reaction
+ 100 mM DTT - 1 μl per reaction
+ RNase OUT - 1 μl per reaction
+ Super Script reverse transcriptase - 1 μl per reaction
+
+6. Vortex and briefly spin the components of master mix 2, add master-mix 2 to the reaction tube.
+7. Incubate the reaction tube at 23°C 10 min.
+8. Incubate the reaction mix at 50-55 °C 10 min followed by 10 min at 80°C
+9. Dilute the cDNA to desired concentration (e.g. 1:25)
+
+##### qPCR
+
+##### Materials
+
+- QuantStudio 5 real-time PCR machine
+- A qPCR reaction plate
+- Nuclease-free water and pipette tips
+- SYBR-green Master mix
+
+1. Combine master-mix:
+ Sybre-green 2X master-mix - 5 μl per reaction
+ Primermix, Forward and Reverse 5 μM each - 1 μl per reaction
+ H2O - 2 μl per reaction
+
+2. Load the plate with primer-specific master-mix
+3. Add 2 μl cDNA sample
+
+### Protein
+
+##### Total Protein Extraction via Minute Total Protein Extraction Kit for Muscles (Invent Biotechnology)
+
+Manufacturers protocol:
+1. Weight out 20-30 mg tissue (fresh/frozen) and cut it with a sharp blade or a pair of scissors into 
+small pieces (1 X 1 mm or smaller). Place tissue pieces in a filter cartridge with collection tube.
+
+2. Add 50 to 80 mg protein extraction powder on top of the tissue followed by addition of  75 ul 
+lysis buffer (Denaturing Buffer).
+
+3. Immediately grind the tissue with the plastic rod provided against the surface of the filter with 
+moderate twisting force for 2-3 min. Add another 25ul lysis buffer (Denaturing buffer) to the filter and continue to 
+grind for about 30 seconds to 1 min. The plastic rods are reusable after cleaning.
+
+4. Centrifuge the tube at top speed in a microcentrifuge for 1 min. Remove and discard the filter. 
+Transfer the supernatant to a fresh tube (this is extracted total protein). The white-grey pellet in 
+the bottom of collection tube is passing through protein extraction powder that should be 
+discarded.
+
+##### Measurement of total protein concentration
+
+Keep samples and ladder on ice as much as possible, minimize time off ice.
+
+1. Total protein samples from extraction are diluted 1:10 to a final volume of 40 ul (i.e. 4 units sample and 36 units water). 
+2. Use a 96 well plate, and pipette 3 wells per ladder (A-I) and sample (10ul). 
+3. After all ladders and samples have been added to the wells, pipette 250ul Bradford into each well.
+4. Run the plate using SkanIt software to examine the total protein concentration. (Fill in the running set-up and plate layout here)
+
+##### Gel electrophoresis
+Preparation
+
+1.
+Dilute with denaturing buffer to a protein concentration of i.e.
+3ug/ul (depending on measured total protein concentration).
+Standardize the start volume of the samples, i.e.
+to 50ul, then use C1V1 = C2V2 formula.
+Transfer 50ul total protein to a fresh tube, add dilution volume.
+Make sure to note the dilution volume of each sample, you will use it later on.
+2.
+Mix the loading buffer equal to 1/4 of the total sample volume (add startvolume and dilution volume from step 1, divide this by three = loading buffer volume).
+This should be fresh, mix it on the same day, see solutions.
+Add up the total volum of loading buffer needed and mix it under a vent.
+3. Add loading buffer to the samples, vortex briefly.
+4. Pool 20ul from each sample (or lower if necessary) into a fresh "pooling" tube.
+5. Incubate the samples and pool tube at 95 C for 5min. Lids may pop open during or after incubation, pay attention and close them if they do.
+6. Let the samples cool down briefly, store them at -20 C until loading.
+7. Prepare running buffer (solutions). This must be cold, prepare it the day prior to electrophoresis and store it at 4 C fridge.
+
+Setup:
+
+1. Unpack the gel, and make sure to flush it using water (blue tap/ddH2O). If you are running multiple gels, make sure to mark each gel (note something on the plastic case around the gel). 
+2. Remove the sticker from the gels, and insert the gels into the electrophoresis chamber. Keep on ice. 
+3. Fill up the main electrophoresis chamber and the small chamber on the gels surrounding plastic (with a green plastic placeholder) with running buffer, approximately 1L in total. Keep on ice.
+4. Carefully remove the green placeholder from each gel (use tweezers), then flush each well with running buffer from. Keep on ice.
+
+Loading:
+
+1.
+Make sure to draw or note which samples go into which wells.
+Always add Ladder in the first well (5ul, Precision plus Protein Dual Color standards), and the middle or last (3ul).
+Calculate loading volume.
+Our goal is to load as close as possible to 20.25 ug protein, i.e.
+divide 20.25 by standardised concentration (the concentration with loading buffer).
+If you diluted by three, the concentration will be down to 2.25 after adding loading buffer: 20.25/2.25 = 9.
+Therefore we load the gel with 9 ul sample per well in this case.
+Keep on ice.
+2.
+When all samples and ladders are added to the wells, fill the remaining running buffer into the small chamber where the wells are located.
+Do this carefully, and do not fill it on top of the wells or close to them.
+Keep on ice.
+
+Electrophoresis:
+1. Set the voltage to 250, constant, and mA to 500.
+2. Duration is set to 50min. Keep the electrophoresis unit on ice.
+
+##### Wet transfer
+
+Preperation
+1. Transfer buffer should be made prior to the transfer, and should be cold. 
+2. PVDF membran is cut from the blue scroll (8.5cm high, 14cm broad). DO NOT touch the white membrane. Cut of the top left corner. If you are transfering multiple membranes, mark them. 
+3. Soak foam pads in ddH2O and squeese out all bubbles.
+4. Incubate the membran for 5min in MeOH (make sure MeOH is discarded in designated bottle after). 
+5. Incubate membran 5-15min in cold transfer buffer.
+6. Remove the smallest plastic cover of the gel, and flush the gel with ddH2O, then put the gel in transfer buffer for approx 30sec.
+7. Soak filter paper in transfer buffer.
+
+Kit/Sandwich assembly
+1. Place a magnetic stirrer in the bottom of the Criterion Blotter tank, make sure it has room to spin.
+2. Place the Criterion Blotter tank in an ice tub to keep temperature low.
+3. Fill the tank to 50% of total fill volume with transfer buffer.
+4. Place iceblock in the back of the tank (designated area)
+5. Putt the sandwhich in its tray, pour transfer buffer over it, enoug to submerge it and the next components, then add the bottom foam pad
+6. Take 1 filter paper in one hand and place the gel on it
+7. While holding the filter with the gel on, place the membrane on top of the gel
+8. Put the filter/gel/membrane down on the bottom foam pad, with the filter side facing the foam pad
+9. Carefully use a roller to remove air bubbles, while holding the membrane down with two fingers
+10. Place the last filter paper on the membran, then carefully roll again
+11. Place the last foam pad on top, carefully roll again
+12. Put the sandwich in the Criterion Blotter, make sure red side faces the red electrode, and that the magnetic stirrer still can move. 
+13.
+Add remaining transfer buffer to the tank, fill it up to the marked level.
+Place the ice tub, with the blotter, upon a magnetic stirring plate and start stirring to maintain even buffer temperature and ion concetration during the transfer.
+
+Transfer
+1. Connect PowerPac HC power supply to Criterion Blotter.
+2. Set PowerPac to 400mA, constant A, 1t 45min, at 4 degrees celsius (ice).
+
+MemCode Staining
+NB: Keep the PVDF membrane side that was facing the Gel during transfer up at all times, and do not let the membrane dry!
+1. Wash the membran with water (ddH2O)
+2. Add MemCode Sensitizer. Carefully shake for 2min with the shaker. Sensitizer is reusable for a couple attempts
+3. Add MemCode Stain. Carefully shake for 1min with the shaker. Stain is reusable for a couple attempts
+4. Rinse the membrane 3 times with water (ddH2O).
+6. Wash with Destain/Methanol solution (1:1) on shaker for 5min.
+7. Manual capture in G:Box - make sure the black plate is removed, light is on and image is sharp
+8. Slice the membrane into desired strips using scalpel and a ruler
+9. Rinse 5 times with water (ddH2O)
+10. Wash with eraser/methanol solution (1:1) on a shaker for 10-20min
+11. Rinse 5 times with water (ddH2O)
+
+Antibody staining
+1. Block the membrane for 1h at roomtemp or 4C over night with "blocking buffer" (5% non-fat dry milk, see solutions)
+2. Wash once with TBS-T, 5min, on a tipping board
+3.
+Incubate membrane with suitable dilution of primary antibody.
+It is recommended to incubate at 4C over night.
+On a tipping board.
+Can be done using a polystyrene (isopor) box and ice to keep the temperature low.
+4. Wash the membrane 3 times with TBS-T, 5min per wash, on a tipping board. 
+5. Incubate membrane with recommended dilution of secondary antibodies in blocking buffer at roomtemp for 1h. On a tipping board
+6. Wash the membrane 3-5 times (5 times may yield nicer blots) with TBS-T, 5min per wash, on a tipping board. 
+7. Mix just sufficient ECL and incubate the membrane for 5min on a tipping board. 8ml should suffice if you have sliced the membrane in smaller strips
+8. Either store the membrane in TBS-T or proceed with visualisation. If you store it in TBST, make sure to incubate once more in ECL.
+
+Quantification with G:Box and Genesys
+1. Ready the G:Box and Genesys on the computer, use ECL protocol. 
+3. Sandwich the membrane between two pieces of transparency film, and use the roll to remove any airbubbles. Make sure you do not get any liquid on the black plate 
+3. Place the membrane within the G:Box and initiate protocol. Should not have to click anything but next, until "Capture". Make sure the membranes position is good, as close to centre as possible.
+
+###### **Table 3:** Primer sequence and performance. Average cycle thresholds (Ct) and primer efficiencies were calculated from all qPCR reactions.
+
+![](figures/archive/Primer.sequences.PNG)
+
+## Total protein blots
+
+### Blot/Gel 1 & 4
+
+Gel 1 overview:
+
+* Ladder loaded in well 1 and 25.
+* Pools loaded in well 6, 12 and 20.
+* Well 2-5 = sample 1.1-1.4 (participant 101).
+* Well 7-10 = sample 2.1-2.4. (participant 102).
+* Well 11 + 13-15 = sample 3.1-3.4 (participant 103).
+* Well 16-19 = sample 5.1-5.4 (participant 105).
+* Well 21-24 = sample 6.1-6.4 (participant 106).
+
+![](figures/archive/total_protein_gel1.PNG)
+
+Gel 4 overview:
+
+* Ladder loaded in well 1 and 26.
+* Pools loaded in well 2, 7, 15 and 21.
+* Well 3-6 = sample 6.4-6.1 (participant 106).
+* Well 8-11 = sample 5.4-5.1 (participant 105).
+* Well 12-14 + 16 = sample 3.4-3.1 (participant 103).
+* Well 17-20 = sample 2.4-2.1 (participant 102).
+* Well 22-25 = sample 1.4-1.1 (participant 101).
+
+![](figures/archive/total_protein_gel4.PNG)
+
+Blot/Gel 2 & 5
+
+Gel 2 overview
+
+* Ladder loaded in well 1 and 26.
+* Pools loaded in well 6, 11, 19 and 25.
+* Well 2-5 = sample 7.1-7.4 (participant 107).
+* Well 7-10 = sample 8.1-8.4 (participant 108).
+* Well 12-15 = sample 9.1-9.4 (participant 109).
+* Well 16-18 + 20 = sample 12.1-12.4 (participant 112).
+* Well 21-24 = sample 13.1-13.4 (participant 113).
+
+![](figures/archive/total_protein_gel2.PNG)
+
+Gel 5 overview
+
+* Ladder loaded in well 1 and 26.
+* Pools loaded in well 2, 8, 16 and 21.
+* Well 3-6 = sample 13.4-13.1 (participant 113).
+* Well 7 + 9-11 = sample 12.4-12.1 (participant 112).
+* Well 12-15 = sample 9.4-9.1 (participant 109).
+* Well 17-20 = sample 8.4-8.1 (participant 108).
+* Well 22-25 = sample 7.4-7.1 (participant 107).
+
+![](figures/archive/total_protein_gel5.PNG)
+
+Blot/Gel 3 & 6
+
+Gel 3 overview
+
+* Ladder loaded in well 1 and 17.
+* Pools loaded in well 2, 7 and 13.
+* Well 3-6 = sample 14.1-14.4 (participant 114).
+* Well 8-11 = sample 15.1-15.4 (participant 115).
+* Well 12 + 14-16 = sample 16.1-16.4 (participant 116).
+
+![](figures/archive/total_protein_gel3.PNG)
+
+Gel 6 overview
+
+* Ladder loaded in well 1 and 17.
+* Pools loaded in well 5, 11 and 16.
+* Well 2-4 + 6 = sample 16.4-16.1 (participant 116).
+* Well 7-10 = sample 15.4-15.1 (participant 115).
+* Well 12-15 = sample 14.4-14.1 (participant 114).
+
+![](figures/archive/total_protein_gel6.PNG)
+
+## Statistical analyses
+
+We chose to log-transformed all data, as visual inspection showed clear signs of heteroscedasticity in some of the data sets.
+Thus, data was log-transformed prior to modelling with the function log(), and then reverse-transformed for figure illustration via the function exp().
+Exceptions to this was the data used in figure 1 (B, C and D).
+In figure 1B and -C, the illustrated data are absolute changes in plasma glucose and c-peptide, as this demonstrated quite well the change that occured, and arguably, is quite more understandable compared to arbitrary units, such as the protein signal data.
+For figure 1D, we had already made in index of strength combining isometric and isokinetic peak torque, meaning the data was already put on a different scale than the original absolute change in peak torque.
+
+Further, we mean-centered the baseline values in each model to account for regression to the mean.
+This was simply done by subtracting the mean baseline value from each participants baseline value, to control for the possibility of extremely high or low baseline values to violate the assumption of equal variance.
+Estimated marginal means were calculated from the linear models, enabling acquisition of least-square means, showing the means for all involved groups adjusted to means of other variables in the model.
+This is a way to statistically model the means of the data, rather than describing them as they appear in the data.
+
+To enable cross nano Drop, qPCR plate and gel comparisons from the wet lab analyses, we normalized RNA and protein data.
+Total RNA was normalized to wet muscle weight, rRNA was normalized to wet muscle weight and the external reference gene Lambda, while protein data was normalized both to the total protein lane factor and to pools of all protein samples.
+
+RNA normalization:
+
+ 1. weight = mean(weight) 
+ 2. RNA = mean(RNA)
+ 3. weight.mc = weight/mean(weight)
+ 4. RNA.weight = RNA / weight
+    + weight = biopsy weight
+
+Weight and RNA was summarised, RNA was devided by mean centered weight. Thus, we could compare total RNA results analyzed on different nanoDrop plates.
+
+rRNA normalization:
+
+ 1. Lambda expression was first calculated and stored in the data.frame "nf"
+ 2. Calculation of target expression
+ 3. Joined the lambda and target expression data.frames
+ 4. nf.w = expr / weight
+ 5. nf.w = nf.w / max(nf.w)
+ 6. nf.expr = log(expr / nf.w), nf.w = scale(nf.w) ?
+    + expr = target expression 
+    + weight = biopsy weight
+
+To accurately measure changes in expression of rRNA, we added a fixed amount of Lambda to each total RNA sample during RNA extraction.
+Lambda expression and target (47S, 18S, 28S, 5.8S, 5S) were calculated in separate data frames, and then joined for normalization.
+In each data frame, we first created a normalization factor for weight (nf.w), a result of expression divided by weight.
+To make the nf.w comparable across the qPCR plates, we put it on a scale through the code nf.w = nf.w /max/(nf.w).
+This code divides the mean nf.w by the highest observed value in the variable nf.w.
+
+Protein normalization:
+
+ 1. tpl = tp - bg
+ 2. tpl = tpl/max(tpl)
+ 3. signal = signal / max(signal)
+ 4. norm.sign = signal / tpl
+ 5. pool = mean(pool)
+ 6. norm.sign = norm.sign / pool
+    + tpl = total protein per lane
+    + tp = total protein (mean gray pixels in lanes)
+    + bg = background (mean gray pixels between lanes)
+    + signal = pixels per band
+
+To create the tpl normalization factor, we measured the mean gray both within (tp) and between (bg) the lanes of each total protein stain blot with the ImageJ software.
+So, each lane and each space between lanes got a "mean gray" value.
+Subtracting background from total protein (tpl = tp-bg), and then using the resulting value divided by the maximum observed total protein in a lane per gel (tpl = tpl/max(tpl)), gave us a value normalized between gels.
+In addition, to account for variations in loading and unequal size in muscle tissue samples, we also normalized antibody signal by pools.
+The pools were mixes of total protein sample, 20ul from each sample, loaded into three lanes per gel.
+So, after normalizing signal to tpl (norm.sign = signal / tpl), we summarized the average signal from the pools (pool = mean(pool)), and divided the tpl normalized signal by average pool signal (norm.sign = norm.sign / pool).
+
+### Statistical models
+
+RNA
+
+```{r total RNA model}
+totalrna.model <- readRDS("./data/data-gen/rna/totalrna.model.RDS")
+
+totalrna.model
+
+```
+
+Protein
+
+```{r protein model}
+protein.model <- readRDS("./data/data-gen/protein/protein.model.RDS")
+
+protein.model
+
+```
+
+Protein/RNA correlation
+
+```{r correlation model}
+correlation.model <- readRDS("./data/data-gen/protein/correlation.model.RDS")
+
+correlation.model
+
+```
+
+Strength, plasma glucose & c-peptide, nutrition and volume
+
+```{r strength, plasma, nutrition and volume model}
+strength.model <- readRDS("./data/data-gen/humac/strength.model.RDS")
+
+strength.model
+
+```

### 2023-08-04 manuscript.RMD
fixed p-value format in manuscript

@@
   word_document:
-    reference_docx: resources/reference-template-02.docx
+    reference_docx: resources/jap-template.docx
 editor_options: 
@@
 
+<!-- To do: Fix tables, add outlier figures to the readme -->
 ```{r}
@@
 ###
+
+```
+
+```{r abstract in-line data}
+
+## total_rna_stats
+total_rna_stats <- readRDS("./data/data-gen/rna/total_rna_stats.RDS")
+
+## stat_rrna
+# stat_rrna$trg_47s, stat_rrna$trg_18s, stat_rrna$trg_28s, stat_rrna$trg_5.8s, stat_rrna$trg_5s
+stat_rrna <- readRDS("./data/data-gen/rna/stat_rrna.RDS")
+
+## stat_west
+#stat_west$cmyc, stat_west$ubf, stat_west$rps6
+stat_west <- readRDS("./data/data-gen/protein/stat_west.RDS")
+
+## p_west
+# min(p_west) - max(p_west)
+p_west <- readRDS("./data/data-gen/protein/p_west.RDS")
+
+## corr_incr
+# $\sim$ `r corr_incr`
+corr_incr <- readRDS("./data/data-gen/protein/corr_incr.RDS")
+corrp <- readRDS("./data/data-gen/protein/corrp.RDS")
 
@@
 However, its effect on RT-induced adaptations remains unexplored.
-**Methods.** Sixteen healthy, moderately trained individuals (male/female, n = XX/XX; age, mean (sd)),<!-- Add male/female numbers, age--> participated in a within-participant trial with unilateral resistance training performed with glucose- and placebo supplement on alternate days over six sessions per condition.
+
+**Methods.** Sixteen healthy, moderately trained individuals (male/female, n = 9/7; age, 23.7 (1.8)/24.6 (4.8)),<!-- Add male/female numbers, age.
+Kristian: Done--> participated in a within-participant trial with unilateral resistance training performed with glucose- and placebo supplement on alternate days over six sessions per condition.
 Participants remained fasted overnight, receiving glucose/placebo and protein supplementation only before and after RT sessions (3 sets of 10 repetitions maximum unilateral leg press and knee extension).
@@
 vastus lateralis pre and post the intervention, and maximal unilateral isometric and isokinetic knee extension torque were measured before, during (days 4, 5, 8, 9) and after the intervention.
-**Results.** <!-- update statistics in this section -->RT led to robust increases in total RNA and rRNA in glucose and placebo (RNA: 26 and 22% respectively, rRNA: 33-43% and 33-41% respectively), with no differences between supplements (p > 0.05).
-There were no differences in content of c-Myc (*p* = 0.160), UBF (*p* = 0.154) and rpS6 (*p* = 0.386) between treatments in response to RT.
-A linear relationship was found between content of UBF and total RNA, where an increase of 1 SD unit in UBF equated to 6.8% increase in total RNA (*p* = 0.002).
+
+**Results.** <!-- update statistics in this section -->From baseline to post-intervention total RNA and rRNA per unit muscle weight increased on average by ~24% and ~33-43%, respectively (Figure 2), with no differences between supplements (Figure 2A; mean difference `r total_rna_stats`, Figure 2B; 47S, `r stat_rrna$trg_47s`; 18S, `r stat_rrna$trg_18s`; 28S, `r stat_rrna$trg_28s`; 5.8S, `r stat_rrna$trg_5.8s`; 5S, `r stat_rrna$trg_5s`).
+RT with glucose resulted in estimated levels of c-Myc, UBF and RPS6 being `r stat_west$cmyc`, `r stat_west$ubf` and `r stat_west$rps6`% lower compared to placebo, respectively, without showing statistical significance (*p* = `r min(p_west)`-`r max(p_west)`; Figure 3A).
+There was a linear relationship between UBF and total RNA, where an increase of 1 SD unit of UBF equated to a $\sim$ `r corr_incr`% increase in total RNA (*p* = `r corrp`; Figure 3C).
 Lastly, there were no differences between supplements in maximal unilateral isometric knee extension force and isokinetic peak torque during the intervention (p > 0.05) and following the last session (p > 0.05), apart from the 60 d/s test before last RT session (p < 0.05).
+
 **Conclusion.** Glucose supplement before and after RT does not augment RT-induced ribosome biogenesis, nor does it enhance muscular performance and recovery after two weeks of heavy RT in moderately trained, young individuals.
+
+<!--*Introduction:*Responses to resistance training (RT) are not uniform across the population, and individuals responding poorly to RT show a blunted ability to produce novel ribosomes [@hammarstrom_benefits_2020; @stec_ribosome_2016].
+Recent research solidifies the relationship between increased rates of total RNA and the magnitude of RT-induced adaptations [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022; @figueiredo_revisiting_2019], with ribosome abundance being associated with UBF protein levels [@hammarstrom_ribosome_2022].
+However, 10-15% of individuals undertaking a period of RT showed considerable impaired growth [@alvarez_interindividual_2018; @mann_high_2014; @thalacker-mercer_cluster_2013].
+Since modifying RT alone does not convert low-responders to high-responders [@hammarstrom_benefits_2020] other means than RT *per se* is necessary to circumvent this discrepancy.
+High glucose treatment has been observed to augment rDNA transcription in vitro [@sebata_control_2018], however, its effect on RT-induced adaptations remains quite unexplored.
+*Methods:*Sixteen healthy moderately trained participants were included and randomized to unilateral resistance training with glucose- and placebo supplement.
+Participants remained fasted overnight receiving glucose/placebo and protein only before and after RT.
+Resistance training consisted of 3 sets of 10 repetitions maximum unilateral leg press and knee extension.
+Micro biopsies were sampled pre (T1/T2) and post (T3/T4) the intervention, and maximal unilateral isometric and isokinetic knee extension torque were measured pre (T0), during (days 4, 5, 8, 9) and post (T3/T4, day 13) the intervention.
+*Results:*Despite robust increases in total RNA and rRNA in glucose and placebo (RNA: 26 and 22% respectively, rRNA: 33-43% and 33-41% respectively), there were no differences between supplements (p > 0.05).
+There were no differences in change of c-Myc, UBF and rpS6 content between treatments (*p* > 0.05).
+There was indeed a linear relationship between content of UBF and total RNA, where an increase of 1 SD unit in UBF equated to 6.8% increase in total RNA (*p* = 0.002).
+Lastly, there were no differences between supplements in maximal unilateral isometric knee extension force and isokinetic peak torque during the intervention (p > 0.05) and following the last session (p > 0.05), apart from the 60 d/s test before last RT session (p < 0.05).
+*Conclusion:* Glucose supplement before and after RT does not acutely augment ribosome biogenesis, nor does it enhance muscular performance and recovery after two weeks of heavy RT in moderately trained young individuals.
+OLD ABSTRACT-->
 
@@
 Therefore, other means than modification of RT variables seem necessary to optimize individual responses.<!-- The above sentence might be a bit contradictory?
-Im not sure, are you thinking about "other means than modification of RT" being contradictory of "modifications such as altered exercise volume.." sentence?
+Kristian: Im not sure, are you thinking about "other means than modification of RT" being contradictory of "modifications such as altered exercise volume.." sentence?
 I think there is a point to establishing that yes, increasing exercise volume is a potent way to increase the RT stimulus in general.
@@
 This is surprising since glucose is the preferred energy substrate of the contracting skeletal muscle during strenuous exercise and a major supplier of energy to cells via ATP synthesis [@mul_exercise_2015;@sebata_control_2018].
-Furthermore, energy availability is a decisive factor in the*de novo* synthesis of ribosomes [@kusnadi_regulation_2015;@moss_housekeeper_2007;@sebata_control_2018] <!--Add ref that indicate high energy requirement for ribo biog possibly (VONWALDEN19;TANAKA18?
+Furthermore, energy availability is a decisive factor in the *de novo* synthesis of ribosomes [@kusnadi_regulation_2015;@moss_housekeeper_2007;@sebata_control_2018] <!--Add ref that indicate high energy requirement for ribo biog possibly (VONWALDEN19;TANAKA18?
 MOSS.
@@
 
-Transcription of ribosomal RNA (rRNA) by Pol I is considered the limiting step in synthesizing new ribosomes [@moss_promotion_1995].
+Transcription of ribosomal RNA (rRNA) by Pol I is considered the rate limiting step in synthesizing new ribosomes [@moss_promotion_1995].
 Multiple signalling pathways converge to regulate rRNA transcription, including the mammalian target of rapamycin (mTORC1) signal-transduction pathway and c-Myc [@kusnadi_regulation_2015;@von_walden_ribosome_2019;@sebata_control_2018; @west_acute_2016; @mori_c-myc_2021].
 mTOR receives input from growth factors, hormones, mechanical loading, and nutrients to balance protein synthesis through multiple mechanisms based on cellular energy levels [@hoppe_amp-activated_2009].
-Importantly, mTOR and its downstream target S6K1 serve as a mediator from insulin signalling in skeletal muscle [@hillier_physiological_2000].
+Importantly, mTOR and its downstream target S6K1 serve as a mediator for insulin signalling in skeletal muscle [@hillier_physiological_2000].
 The mTORC1 pathway contributes to ribosome biogenesis through the regulation of the translation of ribosomal proteins and to forming of the preinitiation complex (PIC) that marks the initiation of rRNA transcription [@figueiredo_regulation_2019;@von_walden_ribosome_2019].
@@
 Additionally, UBF phosphorylation, required for interaction with the rDNA promoter, is increased by high glucose in a mTORC1 dependent manner (rapamycin sensitive) in kidney glomerular epithelial cells[@mariappan_ribosomal_2011].
-Independently from UBF and mTORC1, high glucose was shown to lead to chromatin remodelling, which promotes rRNA transcription in cell cultures[@zhai_human_2012].
+Independently from UBF and mTORC1, high glucose was shown to lead to chromatin remodelling, which promoted rRNA transcription in cell cultures[@zhai_human_2012].
 Together these observations indicate a potential role of glucose in positively affecting muscle rRNA synthesis.
@@
 The study was designed as a 12-day double-blinded randomized controlled trial, with an alternating unilateral RT protocol (Figure 1A).
-Participants were randomly allocated to exercise one leg with glucose and one leg with a placebo (Figure 1A).
+Participants were randomly allocated to exercise one leg with a glucose condition and one leg with a placebo condition (Figure 1A).
 <!--from the "old" methods: either their dominant or non-dominant leg with glucose or placebo before and after RT, and whether they started RT with glucose or placebo.
@@
 Apart from this, participants ingested a self-chosen diet during period II, registered in MyFitnessPal or similar applications.
-The self-chosen diet was repeated on pairwise consecutive days, to ensure similar premises for resistance training responses between treatments.
+The self-chosen diet was repeated on pairwise consecutive days, to ensure similar premises for resistance training responses between conditions.
 During period III (2200-0700hrs), participants remained in an overnight fasted state.
@@
 Maximal leg press strength was defined as the maximal load lifted in a controlled fashion, with a knee angle of 45 degrees (start), within a 2.5kg range.
-Attempts, where participants did not reach 45 degrees during the eccentric phase, were not approved.
+Attempts where participants did not reach 45 degrees during the eccentric phase, were not approved.
 Maximal knee extension testing followed the same specific warm-up and pause protocol and was defined as maximal load lifted in a controlled fashion, reaching full extension of the knee joint, within a 1.25kg range.
@@
 
-### Resistance training protocoL
+### Resistance training protocol
 
@@
 
-To measure blood glucose levels with and without glucose ingestion/training, capillary blood was collected from finger sticks on days with biopsy sampling.
+To measure blood glucose levels with and without glucose ingestion/training, capillary blood was collected from finger draws on days with biopsy sampling.
 One capillary blood sample was taken on day 1 (Figure 1A) to serve as a baseline.
@@
 Comparisons of the acute data gathered from the post-fifth session until and including 23hrs post-sixth session showed a significant mean increase in strength of 5-9% after RT with glucose and placebo 30min after the sixth RT session (Figure 1D, 30min post 6RT: *p* = `r apc2`) and two hours after the sixth RT session (Figure 1D, 2h post 6RT: *p* = `r apc3`).
-Twenty-three hours after the last sixth RT session, strength was unchanged compared to after the fifth RT session (Figure 1D, 23h post 6RT: *p* = `r apc4`).
+Twenty-three hours after the last (sixth) RT session, strength was unchanged compared to after the fifth RT session (Figure 1D, 23h post 6RT: *p* = `r apc4`).
 
@@
 
+saveRDS(total_rna_stats, "./data/data-gen/rna/total_rna_stats.RDS")
+
 ```
@@
 
+saveRDS(stat_rrna, "./data/data-gen/rna/stat_rrna.RDS")
+
 ```
@@
 stat_west <- lapply(prot_models$prot_models, west_estimate)
+
+saveRDS(stat_west, "./data/data-gen/protein/stat_west.RDS")
+
 p_west <- unlist(lapply(prot_models$prot_models, west_estimate, p = TRUE))
+
+saveRDS(p_west, "./data/data-gen/protein/p_west.RDS")
 
@@
 corr_incr <- round(100 * (exp(coef(summary(mod))[2,1])-1),1)
+saveRDS(corr_incr, "./data/data-gen/protein/corr_incr.RDS")
+
 corrp <- sprintf("%.3f", coef(summary(mod))[2,5])
+saveRDS(corrp, "./data/data-gen/protein/corrp.RDS")
 
@@
 corrcoef <- coef(summary(corr.mod))
-corrp <- round(corrcoef[2,5], 3)
+corrp <- sprintf("%.4f", corrcoef[2,5])
 
@@
 These findings suggest that the effects of high vs.
-low glucose/glucose starvation treatment on rDNA transcription initiation observed in previous *in vitro* studies [@mariappan_ribosomal_2011; @tanaka_mild_2015] are not translatable to acute effects in human skeletal muscle *in vivo*.
+low glucose/glucose starvation condition on rDNA transcription initiation observed in previous *in vitro* studies [@mariappan_ribosomal_2011; @tanaka_mild_2015] are not translatable to acute effects in human skeletal muscle *in vivo*.
 Notably, the present study was specifically designed to investigate differences in pre-post changes within participants, and is as such also quite limited to this perspective.
-Hence, the observed effects of time serve as supplementary to the efficiency of the design, without a negative control group to compare to.
-
-To our knowledge, no previous study has investigated the effects of glucose compared to placebo ingestion with RT on ribosome biogenesis in human skeletal muscle previously, making our hypotheses exploratory and comparisons few.**In vitro** studies on mice glomerular epithelial cells [@mariappan_ribosomal_2011] and human breast cancer cells [@tanaka_mild_2015] have observed glucose-induced UBF-dependent augmentation of ribosome biogenesis [@mariappan_ribosomal_2011], and reduced rRNA transcription by glucose starvation [@tanaka_mild_2015].
+Hence, the observed effects of time serve as supplementary to the efficiency of the design, without a negative control group to compare with.
+
+To our knowledge, no previous study has investigated the effects of glucose compared to placebo ingestion with RT on ribosome biogenesis in human skeletal muscle previously, making our hypotheses exploratory and comparisons few.
+*In vitro* studies on mice glomerular epithelial cells [@mariappan_ribosomal_2011] and human breast cancer cells [@tanaka_mild_2015] have observed glucose-induced UBF-dependent augmentation of ribosome biogenesis [@mariappan_ribosomal_2011], and reduced rRNA transcription by glucose starvation [@tanaka_mild_2015].
 These findings do imply an important role for glucose in the initiation of rDNA transcription, the rate-limiting step of ribosome biogenesis [@mariappan_ribosomal_2011; @tanaka_mild_2015].
@@
 Further, glucose ingestion increases energy levels, leading us to hypothesize that high glucose vs.
-placebo treatment would stimulate energy-sensitive pathways such as PIH1, mTORC1, ERK1/2, AMPK and SIRT1 [@kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012].
+placebo condition would stimulate energy-sensitive pathways such as PIH1, mTORC1, ERK1/2, AMPK and SIRT1 [@kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012].
 Despite previously reported upregulation in PIC assembly due to high-glucose mediated mTORC1, ERK1/2 and PIH1 or low-glucose mediated AMPK and SIRT1 activation  [@hoppe_amp-activated_2009; @kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012], the present study showed no signs of such effects of glucose vs.
-placebo treatment.
-Resistance training, irrespective of treatment, yielded a robust accumulation of total RNA and expression of rRNA, in line with previous observations [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022].
+placebo conditions.
+Resistance training, irrespective of condition, yielded a robust accumulation of total RNA and expression of rRNA, in line with previous observations [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022].
 Due to the exploratory nature of our hypotheses, this was hardly a very surprising result.
@@
 Importantly, previous studies investigated high vs.
-low glucose treatment [@mariappan_ribosomal_2011], or high glucose vs.
-glucose starvation [@hoppe_amp-activated_2009; @tanaka_mild_2015], while the present study aimed to compare high glucose treatment to placebo, with a matched daily macronutrient and energy intake.
+low glucose conditions [@mariappan_ribosomal_2011], or high glucose vs.
+glucose starvation [@hoppe_amp-activated_2009; @tanaka_mild_2015], while the present study aimed to compare high glucose condition to placebo, with a matched daily macronutrient and energy intake.
 Therefore, the comparison made in the present study was high plasma glucose levels vs.
@@
 
-In the present study, despite not measuring the activity in central pathways mediating anabolic signalling (mTORC1, ERK1/2), analyses of the downstream target UBF and the ribosomal protein S6 (rpS6) further supported observations from the RNA data.
+In the present study, despite not measuring the activity in central pathways mediating anabolic signalling (mTORC1, ERK1/2), analyses of the downstream target UBF and the ribosomal protein S6 (rpS6), as well as the general transcription factor c-Myc, further supported observations from the RNA data.
 UBF has previously been described as a master regulator of rDNA transcription *in vitro* [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @russell_rna-polymerase-i-directed_2005], while rpS6 is proposed as a valid and reliable means to measure ribosome biogenesis [@chaillou_hypoxia_2012; @nakada_correlation_2016].
-While content of UBF and rpS6 changed equally in RT with glucose compared to placebo, content of c-Myc was significanly higher after RT with placebo.
+Lastly, c-Myc has previously been described as a potent regulator of ribosome biogenesis, independent of mTORC1, and a direct regulator of UBF [@mori_c-myc_2021; @poortinga_c-myc_2011; @west_acute_2016].
+Hence, it seems quite reasonable to observe equal changes in these three proteins.
+<!-- OLD: While content of UBF and rpS6 changed equally in RT with glucose compared to placebo, content of c-Myc was significanly higher after RT with placebo.
 This was unexpected, since the general transcription factor c-Myc previously has been described as a potent regulator of ribosome biogenesis, independent of mTORC1, and a direct regulator of UBF [@mori_c-myc_2021; @poortinga_c-myc_2011; @west_acute_2016].
@@
 Moreover, the observation that total RNA and rRNA were unaffected by the significant differences in c-Myc content, does strengthen the importance and specificity of UBF content in human skeletal muscle ribosome biogenesis.
-The linear relationship found exclusively between UBF content and total RNA accumulation, and not c-Myc or rpS6, found in the present study also support this notion.
+-->The linear relationship found exclusively between UBF content and total RNA accumulation, and neither c-Myc nor rpS6, support the importance and specificity of UBF content in human skeletal muscle ribosome biogenesis.
 While this is not a novel finding, it has only recently been observed in human skeletal muscle following RT [@hammarstrom_ribosome_2022].
@@
 
-As in the biological data, there was generally no difference in the change of strength as an index between the treatments.
+As in the biological data, there was generally no difference in the change of strength as an index between the conditions.
 Generally, skeletal muscle strength decreased from baseline to post-fifth and -sixth RT sessions, with a significantly less reduction in glucose compared to placebo at post-fifth session measurement.
@@
 Further, strength testing during the intervention was conducted 23hrs after RT, meaning that these tests also may have been influenced by the biphasic recovery.
-However, this does not explain the difference between treatments at post-fifth RT session, or the great drop in strength from post 4RT to post 5RT sessions.
+However, this does not explain the difference between conditions at post-fifth RT session, or the great drop in strength from post 4RT to post 5RT sessions.
 A possible argument could be that exercising without glucose may have caused more stress compared to exercising with glucose, as glucose is the preferred energy source during strenuous exercise [@mul_exercise_2015], thus increasing performance with glucose compared to placebo.
-This may also, at least partly, explain the significantly higher content of the stress-responsive c-Myc in RT with placebo compared to glucose [@mori_c-myc_2021; @poortinga_c-myc_2011; @west_acute_2016].
-However, training volume showed that the total training session volume was equal on pairwise consecutive days, i.e.
+Notably, training volume showed that the total training session volume was equal on pairwise consecutive days, i.e.
 no difference between days 1-2, days 3-4 and so on.
-Hence, there were no differences in mechanical loading to induce a higher stress between treatments.
-Arguably, the significantly higher plasma levels of glucose during RT with glucose may have increased energy availability in muscle, resulting in lower metabolic stress and less performance reduction, compared to placebo during RT [@westerblad_mechanisms_1998; @kent-braun_central_1999] Unfortunately, we did not conduct measurements of markers of metabolic stress such ass inorganic phosphate, H+, Mg2+ and the ADP/ATP ratio [@westerblad_mechanisms_1998; @kent-braun_central_1999].
-Therefore, discussing the potential effect of differences in metabolic stress between treatments would only be speculation, however probable.
+Hence, there were no differences in mechanical loading to induce a higher stress between conditions.
+Arguably, an increased energy availability via glucose ingestion during RT may induce less acute stress on the exercised skeletal muscle, therefore less fatigue, and perhaps less performance reduction, compared to placebo during RT <!-- this is possibly a stretch -->[@westerblad_mechanisms_1998; @kent-braun_central_1999] Unfortunately, we did not conduct measurements of markers of metabolic stress such ass inorganic phosphate, H+, Mg2+ and the ADP/ATP ratio [@westerblad_mechanisms_1998; @kent-braun_central_1999].
+Therefore, discussing the potential effect of differences in metabolic stress between conditions would only be speculation, however probable.
 Lastly, these inconclusive observations may also reflect our limited sample size, despite using a within-subjects design.
@@
 
-**Figure 3:** Changes in levels of protein.
-A) Mean UBF, c-Myc and rpS6 (AU, arbitrary units) at baseline and post + representative western blots of the respective proteins.
-Baseline = Day 1 leg 1, Day 2 leg 2, post = Day 11 leg 1, Day 12 leg 2.
-GLU = glucose, PLA = placebo.
+**Figure 3:** Changes in protein content.
+A) Fold-change pre-post in levels of protein in placebo and glucose + fold-change difference 95% CI, with representative western blots of the respective proteins.
 Protein samples were analyzed in two duplicates per biopsy per time point, loaded on separate gels in an inverted order, e.g.
@@
 Changes in protein levels were calculated as log-fold change scores normalized by pools (pool of all protein samples per gel).
-Mean log-change scores of the duplicates were calculated before modelling and reverse-transformed for figure illustration.
+Mean log fold-change scores of the duplicates were calculated before modelling and reverse-transformed for figure illustration.
 Values are estimated marginal means fold change per leg per supplement ± 95% CI.
-* = *p* < 0.05 between groups.
 Glucose n = 13, placebo n = 13.

### 2023-08-08 R/qpcr_analysis.R
manuscrit updates

@@ -164,7 +164,37 @@ rrna_analysis_results <- list(rrna_models = models_final,
 saveRDS(rrna_analysis_results, "./data/data-gen/rna/rrna_analysis_results.RDS")
 
 
+### Function for inline reporting of total RNA and rRNA data
 
+
+stat_fun_rna <- function(model) {
+  
+  coefs <- coef(summary(model))
+  ci <- confint(model)
+  meandiff <- sprintf("%.1f", 100 * (exp(coefs[4,1]) -1))
+  cilower <- sprintf("%.1f",100 * (exp(ci[6,1]) -1))
+  ciupper <- sprintf("%.1f",100 * (exp(ci[6,2]) -1))
+  pval <- sprintf("%.3f",coefs[4,5])
+  
+  
+  stats <- paste0(meandiff, "%, [", cilower, ", ", ciupper,"], *p* = ", pval)
+  
+  stats
+}
+
+# Retrieve models from rds
+
+rrna_analysis_results <- readRDS("./data/data-gen/rna/rrna_analysis_results.RDS")
+
+
+# Get estimated means from the model, these are average baseline-corrected log-fold differences 
+# at post.
+
+# Applying the stat_fun_rna function to models
+stat_rrna <- lapply(rrna_analysis_results$rrna_models, 
+                    stat_fun_rna)
+
+saveRDS(stat_rrna, "./data/data-gen/rna/stat_rrna.RDS")
 
 
 

### 2023-08-08 R/totrna_analysis.R

@@ -129,6 +129,35 @@ saveRDS(list(model = models_final,
              coefs = coefs), "./data/data-gen/rna/tot_rna_coefs.RDS")
 
 
+## Get results from totrna_analysis.R 
+
+totrna_results <- readRDS("./data/data-gen/rna/tot_rna_coefs.RDS")
+
+# retrieve model
+trna.lmod <- totrna_results$model
+
+### Function for inline reporting of total RNA and rRNA data
+
+stat_fun_rna <- function(model) {
+  
+  coefs <- coef(summary(model))
+  ci <- confint(model)
+  meandiff <- sprintf("%.1f", 100 * (exp(coefs[4,1]) -1))
+  cilower <- sprintf("%.1f",100 * (exp(ci[6,1]) -1))
+  ciupper <- sprintf("%.1f",100 * (exp(ci[6,2]) -1))
+  pval <- sprintf("%.3f",coefs[4,5])
+  
+  
+  stats <- paste0(meandiff, "%, [", cilower, ", ", ciupper,"], *p* = ", pval)
+  
+  stats
+}
+
+# Save these stats for inline reporting in abstract
+
+total_rna_stats <- stat_fun_rna(trna.lmod)
+
+saveRDS(total_rna_stats, "./data/data-gen/rna/total_rna_stats.RDS")
 
 
 # Saved for reference

### 2023-08-08 R/ubf_rna.correlation.R

@@ -8,35 +8,29 @@ library(tidyverse);library(ggpubr)
 
 # Data 
 west.dat <- readRDS("./data/data-gen/protein/prot.datc.RDS") # Quantified and cleaned western data,
-                                                               # (outliers marked) normalised by pool
+# (outliers marked) normalised by pool
 
 rna.dat <- readRDS("./data/data-gen/rna/rna.dat.RDS") # Quantified and cleaned (outliers marked), 
-                                                      # total RNA data, normalised by wet muscle weight
+# total RNA data, normalised by wet muscle weight
 
 
 ## Prepare data frames for joining
 
 ubf.rdy <- west.dat %>%  # One data frame without outliers
-  select(subject, time, outlier, supplement, target, norm.sign, tpl) %>%
+  select(subject, time, outlier, supplement, target, norm.sign) %>%
   filter(target == "ubf",
          outlier == "in") %>%
   select(-target) %>%
-  mutate(norm.sign = norm.sign / tpl) %>%
-  #print()
   group_by(subject, time, supplement) %>%
-  summarise(mean.sign = mean(norm.sign, na.rm = TRUE)) %>%
-  print()
+  summarise(mean.sign = mean(norm.sign, na.rm = TRUE)) 
 
-saveRDS(ubf.rdy, "./data/data-gen/protein/ubf.rdy.RDS")
 
 rna.rdy <- rna.dat %>%
   select(subject, time, supplement, RNA.weight) %>%
+  
   mutate(subject = as.character(subject)) %>%
   group_by(subject, time, supplement) %>%
-  summarise(mean.rna = mean(RNA.weight, na.rm = TRUE)) %>%
-  print()
-
-saveRDS(rna.rdy, "./data/data-gen/protein/rna.rdy.RDS")
+  summarise(mean.rna = mean(RNA.weight, na.rm = TRUE)) 
 
 ## Correlation analysis
 
@@ -44,140 +38,214 @@ joined.dat <- ubf.rdy %>%
   right_join(rna.rdy) %>%
   #filter(supplement != "GLUCOSE") %>%
   
-  mutate(sd.ubf = (mean.sign - mean(mean.sign, na.rm = TRUE )/sd(mean.sign, na.rm = TRUE))) %>%
-  print()
[…]

### 2023-08-08 figures/primer.tab.R

@@ -1,0 +1,27 @@
+### Primer sequence table
+
+## Packages
+library(readxl); library(tidyverse); library(flextable)
+
+## Data
+primer.dat <- read_excel("./data/primer.seq2.xlsx")
+
+tab.dat <- primer.dat %>%
+  select(Gene, fr, CT.mean, E) %>%
+  print()
+
+
+
+## Flextable
+tab.dat %>%
+  flextable() %>%
+  compose(i = 1, j = c(1, 2, 3, 4), part = "header",
+          value = c(as_paragraph("Gene"),
+                    as_paragraph("Sequence (forward - reverse)"),
+                    as_paragraph("CT mean (SD)"),
+                    as_paragraph("E"))) %>%
+  #add_header_row(values = c(" ", " ", " ", "Primer sequences"), 
+               #  colwidths = c(1, 1, 1, 1)) %>%
+  set_caption("Table 2. Primer Sequences") %>%
+  set_table_properties(layout = "autofit")
+

### 2023-08-08 manuscript.RMD
updates

@@
 <div custom-style="abstract_style">
-## Abstract
+# ABSTRACT
 
@@
 
-<!--*Introduction:*Responses to resistance training (RT) are not uniform across the population, and individuals responding poorly to RT show a blunted ability to produce novel ribosomes [@hammarstrom_benefits_2020; @stec_ribosome_2016].
-Recent research solidifies the relationship between increased rates of total RNA and the magnitude of RT-induced adaptations [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022; @figueiredo_revisiting_2019], with ribosome abundance being associated with UBF protein levels [@hammarstrom_ribosome_2022].
-However, 10-15% of individuals undertaking a period of RT showed considerable impaired growth [@alvarez_interindividual_2018; @mann_high_2014; @thalacker-mercer_cluster_2013].
-Since modifying RT alone does not convert low-responders to high-responders [@hammarstrom_benefits_2020] other means than RT *per se* is necessary to circumvent this discrepancy.
-High glucose treatment has been observed to augment rDNA transcription in vitro [@sebata_control_2018], however, its effect on RT-induced adaptations remains quite unexplored.
-*Methods:*Sixteen healthy moderately trained participants were included and randomized to unilateral resistance training with glucose- and placebo supplement.
-Participants remained fasted overnight receiving glucose/placebo and protein only before and after RT.
-Resistance training consisted of 3 sets of 10 repetitions maximum unilateral leg press and knee extension.
-Micro biopsies were sampled pre (T1/T2) and post (T3/T4) the intervention, and maximal unilateral isometric and isokinetic knee extension torque were measured pre (T0), during (days 4, 5, 8, 9) and post (T3/T4, day 13) the intervention.
-*Results:*Despite robust increases in total RNA and rRNA in glucose and placebo (RNA: 26 and 22% respectively, rRNA: 33-43% and 33-41% respectively), there were no differences between supplements (p > 0.05).
-There were no differences in change of c-Myc, UBF and rpS6 content between treatments (*p* > 0.05).
-There was indeed a linear relationship between content of UBF and total RNA, where an increase of 1 SD unit in UBF equated to 6.8% increase in total RNA (*p* = 0.002).
-Lastly, there were no differences between supplements in maximal unilateral isometric knee extension force and isokinetic peak torque during the intervention (p > 0.05) and following the last session (p > 0.05), apart from the 60 d/s test before last RT session (p < 0.05).
-*Conclusion:* Glucose supplement before and after RT does not acutely augment ribosome biogenesis, nor does it enhance muscular performance and recovery after two weeks of heavy RT in moderately trained young individuals.
-OLD ABSTRACT-->
-
-**Keywords: **
+## NEW AND NOTEWORTHY
+
+**Keywords:** Skeletal muscle; hypertrophy; resistance training; glucose; ribosome
 
@@
 \pagebreak
-## Introduction
+# INTRODUCTION
 
@@
 
-## Methods
-
-### Materials and methods
+# MATERIALS AND METHODS
 
@@
 
-### Participants
+## PARTICIPANTS
 
@@
 
-### Experimental design
+## EXPERIMENTAL DESIGN
 
@@
 
-### Dietary intervention
+## DIETARY INTERVENTION
 
@@
 
-### Assessment of muscular strength
+## ASSESSMENT OF MUSCULAR STRENGTH
 
@@
 
-### Resistance training protocol
+## RESISTANCE TRAINING PROTOCOL
 
@@
 
-### Sampling of muscle tissue and blood
+## SAMPLING OF MUSCLE TISSUE AND BLOOD
 
@@
 
-### Total RNA extraction and quantitative real-time reverse transcription polymerase chain reaction
+## TOTAL RNA EXTRACTION AND REAL-TIME REVERSE TRANSCRIPTION POLYMERASE CHAIN REACTION
 ```{r totrna-extraction}
@@
 
-### Protein extraction and immunoblotting
+## PROTEIN EXTRACTION AND IMMUNOBLOTTING
 
@@
 
-### Statistics and data analysis
+## STATISTICS AND DATA ANALYSIS
 
@@
 To decrease the risk of Type I errors, random effects were selected from step-wise elimination of terms from the most complex structure (random slopes for time and supplement and their interaction) to less complex.
-The most complex random effects structure that converged was chosen as the final model[@Matuschek_2017].
+The most complex random effects structure that converged was chosen as the final model[@matuschek_2017_balancing].
 Plasma glucose, c-peptide, training volume, and the strength index were analyzed by multiple time-point log-fold change score comparisons, using mixed-effects model with baseline values, time and supplement, and the interaction between time and supplement as fixed effects.
@@
 
-## Results
+# RESULTS
 
@@
 
-saveRDS(total_rna_stats, "./data/data-gen/rna/total_rna_stats.RDS")
-
 ```
@@
 
-saveRDS(stat_rrna, "./data/data-gen/rna/stat_rrna.RDS")
-
 ```
@@
 
-saveRDS(stat_west, "./data/data-gen/protein/stat_west.RDS")
-
 p_west <- unlist(lapply(prot_models$prot_models, west_estimate, p = TRUE))
-
-saveRDS(p_west, "./data/data-gen/protein/p_west.RDS")
 
@@
 corr_incr <- round(100 * (exp(coef(summary(mod))[2,1])-1),1)
-saveRDS(corr_incr, "./data/data-gen/protein/corr_incr.RDS")
 
 corrp <- sprintf("%.3f", coef(summary(mod))[2,5])
-saveRDS(corrp, "./data/data-gen/protein/corrp.RDS")
 
@@
 corrcoef <- coef(summary(corr.mod))
+
+corrp <- sprintf("%.3f", coef(summary(mod))[2,5])
+
 corrp <- sprintf("%.4f", corrcoef[2,5])
 
-## Percentage increase in RNA per 1 SD uinit increase in UBF using log-model to show relative increse
+## Percentage increase in RNA per 1 SD uinit increase in UBF using log-model to show relative increase
 corr_incr <- round(100 * (exp(coef(summary(corr.lmod))[2,1])-1), 0)
@@
 
-## Discussion
+# DISCUSSION
 
@@
 
-## References
+# APPENDIX
+
+IS THIS REQUIRED? IF NOT, IS IT NECESSARY?
+
+# DATA AVAILABILITY
+
+LINK TO GITHUB
+
+# GRANTS
+
+IS THIS NECESSARY?
+
+# DISCLOSURES
+
+# AUTHOR CONTRIBUTIONS
+
+# References
 
@@
 
-## Figure captions
+# FIGURE LEGENDS
 
@@
 
-## Tables
-
-```{r, echo = FALSE}
+# TABLES
+
+```{r baseline table, echo = FALSE}
 
@@
 ```
+
+```{r primer table, echo = FALSE}
+
+### Primer sequence table
+
+## Packages
+library(readxl); library(tidyverse); library(flextable)
+
+## Data
+primer.dat <- read_excel("./data/primer.seq2.xlsx")
+
+tab.dat <- primer.dat %>%
+  select(Gene, fr, CT.mean, E)
+
+## Flextable
+tab.dat %>%
+  flextable() %>%
+  compose(i = 1, j = c(1, 2, 3, 4), part = "header",
+          value = c(as_paragraph("Gene"),
+                    as_paragraph("Sequence (forward - reverse)"),
+                    as_paragraph("CT mean (SD)"),
+                    as_paragraph("E"))) %>%
+  #add_header_row(values = c(" ", " ", " ", "Primer sequences"), 
+               #  colwidths = c(1, 1, 1, 1)) %>%
+  set_caption("Table 2. Primer Sequences") %>%
+  width(j = 1, 1.5, unit = "in") %>%
+  width(j = 2, 3, unit = "in") %>%
+  width(j = 3, 1.5, unit = "in") %>%
+  width(j = 4, 0.5, unit = "in")
+  #set_table_properties(layout = "autofit")
+
+```

### 2023-08-08 jap.manuscript.RMD
manuscript work

--- 2023-08-08 manuscript.RMD
@@
 ---
-title: "Glucose ingestion before and after resistance training does not augment ribosome biogenesis in healthy moderately trained young adults"
-author: "Kristian Lian^1^, Sara Christine Olsen Moen^1^, Håvard Hamarsland^1^, Daniel Hammarström^1^, Stian Ellefsen^1^"
+title: ""
+author: ""
 bibliography: resources/ribose-paper.bib
@@
   word_document:
-    reference_docx: resources/jap-template.docx
+    reference_docx: resources/reference-template-02.docx
 editor_options: 
@@
 
-### Affiliations
+RESEARCH ARTICLE
+
+RUNNING HEAD: Glucose ingestion and resistance training in young adults
+
+Glucose ingestion before and after resistance training does not augment ribosome biogenesis in healthy moderately trained young adults
+
+Kristian Lian^1^, Sara Christine Olsen Moen^1^, Håvard Hamarsland^1^, Daniel Hammarström^1^, Stian Ellefsen^1^
+
 <div custom-style="correspondence_style">
-^1^ Section for Health and Exercise Physiology, Department of Public Health and Sport Sciences, Inland Norway University of Applied Sciences, Lillehammer, Norway.
+^1^Section for Health and Exercise Physiology, Department of Public Health and Sport Sciences, Inladn Norway University of Applied Sciences, Lillehammer, Norway
 </div>
@@
 <div custom-style="correspondence_style">
-### Correspondance
-Kristian Lian, Section for Health and Exercise Physiology, Department of Public Health and Sport Sciences, Inland Norway University of Applied Sciences, Lillehammer, Norway.
-Email: kristian.lian@inn.no
-
-Stian Ellefsen, Section for Health and Exercise Physiology, Department of Public Health and Sport Sciences, Inland Norway University of Applied Sciences, Lillehammer, Norway.
-Email: stian.ellefsen@inn.no
+Correspondance: Kristian Lian (kristian.lian@inn.no), Stian Ellefsen (stian.ellefsen@inn.no)
 </div>
 
-### ORCID
+ORCID:<!-- I does not say anything about ORCID in the journal template, do we leave it in?-->
 <div custom-style="correspondence_style">
@@
 
-**Introduction.** Muscle hypertrophy, induced by resistance training (RT), is determined by ribosomal biogenesis.
+Muscle hypertrophy, induced by resistance training (RT), is determined by ribosomal biogenesis.
 Synthesis of ribosomes may augment the transcription of ribosomal RNA, as evident from in vitro studies.
 However, its effect on RT-induced adaptations remains unexplored.
-
-**Methods.** Sixteen healthy, moderately trained individuals (male/female, n = 9/7; age, 23.7 (1.8)/24.6 (4.8)),<!-- Add male/female numbers, age.
-Kristian: Done--> participated in a within-participant trial with unilateral resistance training performed with glucose- and placebo supplement on alternate days over six sessions per condition.
+Sixteen healthy, moderately trained individuals (male/female, n = 9/7; age, 23.7 (1.8)/24.6 (4.8)), participated in a within-participant trial with unilateral resistance training with glucose- and placebo supplement on alternate days over six sessions per condition.
 Participants remained fasted overnight, receiving glucose/placebo and protein supplementation only before and after RT sessions (3 sets of 10 repetitions maximum unilateral leg press and knee extension).
@@
 vastus lateralis pre and post the intervention, and maximal unilateral isometric and isokinetic knee extension torque were measured before, during (days 4, 5, 8, 9) and after the intervention.
-
-**Results.** <!-- update statistics in this section -->From baseline to post-intervention total RNA and rRNA per unit muscle weight increased on average by ~24% and ~33-43%, respectively (Figure 2), with no differences between supplements (Figure 2A; mean difference `r total_rna_stats`, Figure 2B; 47S, `r stat_rrna$trg_47s`; 18S, `r stat_rrna$trg_18s`; 28S, `r stat_rrna$trg_28s`; 5.8S, `r stat_rrna$trg_5.8s`; 5S, `r stat_rrna$trg_5s`).
-RT with glucose resulted in estimated levels of c-Myc, UBF and RPS6 being `r stat_west$cmyc`, `r stat_west$ubf` and `r stat_west$rps6`% lower compared to placebo, respectively, without showing statistical significance (*p* = `r min(p_west)`-`r max(p_west)`; Figure 3A).
+Between conditions, there were no significant differences in increases of total RNA and rRNA (Figure 2A; mean difference `r total_rna_stats`, Figure 2B; 47S, `r stat_rrna$trg_47s`; 18S, `r stat_rrna$trg_18s`; 28S, `r stat_rrna$trg_28s`; 5.8S, `r stat_rrna$trg_5.8s`; 5S, `r stat_rrna$trg_5s`), or levels of c-Myc, UBF and RPS6 (*p* = `r min(p_west)`-`r max(p_west)`; Figure 3A).
 There was a linear relationship between UBF and total RNA, where an increase of 1 SD unit of UBF equated to a $\sim$ `r corr_incr`% increase in total RNA (*p* = `r corrp`; Figure 3C).
-Lastly, there were no differences between supplements in maximal unilateral isometric knee extension force and isokinetic peak torque during the intervention (p > 0.05) and following the last session (p > 0.05), apart from the 60 d/s test before last RT session (p < 0.05).
-
-**Conclusion.** Glucose supplement before and after RT does not augment RT-induced ribosome biogenesis, nor does it enhance muscular performance and recovery after two weeks of heavy RT in moderately trained, young individuals.
+In conclusion, glucose supplement before and after RT does not augment RT-induced ribosome biogenesis after two weeks of heavy RT in moderately trained young adults.
+
+</div>
 
 ## NEW AND NOTEWORTHY
+The effect of glucose ingestion on resistance training-induced ribosome biogenesis is quite unexplored in human skeletal muscle, and our results show no beneficial acute effect following six bouts of exercise.
+The linear relationship observed between UBF and total RNA do solidify the importance of UBF in human skeletal muscle ribosome biogenesis.
+
+</div>
 
@@
 Lastly, c-Myc has previously been described as a potent regulator of ribosome biogenesis, independent of mTORC1, and a direct regulator of UBF [@mori_c-myc_2021; @poortinga_c-myc_2011; @west_acute_2016].
-Hence, it seems quite reasonable to observe equal changes in these three proteins.
+Hence, it seems quite reasonable to observe similar changes in these three proteins.
 <!-- OLD: While content of UBF and rpS6 changed equally in RT with glucose compared to placebo, content of c-Myc was significanly higher after RT with placebo.
@@
 
+Perspectives and Significance
+
 In conclusion, ingestion of glucose immediately before and after resistance exercise training does not acutely augment ribosome biogenesis after two weeks of heavy resistance exercise training, in moderately trained young adults.
@@
 
-# APPENDIX
-
-IS THIS REQUIRED? IF NOT, IS IT NECESSARY?
-
 # DATA AVAILABILITY
 
-LINK TO GITHUB
-
-# GRANTS
-
-IS THIS NECESSARY?
-
-# DISCLOSURES
+Complete data sets and scripts are downloadable here; https://github.com/Kristianlian/ribose-paper
+
+# SUPPLEMENTAL MATERIAL
+
+Extended methods can be found here; https://github.com/Kristianlian/ribose-paper
 
 # AUTHOR CONTRIBUTIONS
+Conceived and designed research: Stian Ellefsen, Håvard Hamarsland, Daniel Hammarström.
+Performed experiments: Kristian Lian, Sara Christine Moen.
+Interpreted results of experiments: Kristian Lian, Daniel Hammarström.
+Prepared Figures: Kristian Lian, Daniel Hammarström.
+Drafted manuscripts: Kristian Lian, Daniel Hammarström.
+Edited and revised manuscript: Stian Ellefsen, Håvard Hamarsland, Daniel Hammarström.
+Approved final version of the manuscript: .
 

### 2023-08-09 jap.manuscript.RMD
table and template updates

@@
 
-RESEARCH ARTICLE
+##### RESEARCH ARTICLE
 
@@
 
-Glucose ingestion before and after resistance training does not augment ribosome biogenesis in healthy moderately trained young adults
-
-Kristian Lian^1^, Sara Christine Olsen Moen^1^, Håvard Hamarsland^1^, Daniel Hammarström^1^, Stian Ellefsen^1^
+#### Glucose ingestion before and after resistance training does not augment ribosome biogenesis in healthy moderately trained young adults
+
+###### Kristian Lian^1^, Sara Christine Olsen Moen^1^, Håvard Hamarsland^1^, Daniel Hammarström^1^, Stian Ellefsen^1^
 
@@
 
-**Keywords:** Skeletal muscle; hypertrophy; resistance training; glucose; ribosome
+**Keywords:** Glucose; hypertrophy; resistance training; ribosome; skeletal muscle
 
@@
 
-### Markers of ribosome biogenesis
-
-##### Total RNA and ribosomal RNA
+## Markers of ribosome biogenesis
+
+### Total RNA and ribosomal RNA
 
@@
 
-##### Protein
+### Protein
 
@@
 
-dxa.tab %>%
+dxa.flex <- dxa.tab %>%
   full_join(humac.tab) %>%
@@
   add_header_row(values = c(" ", "Knee-extension peak torque"), colwidths = c(6, 3)) %>%
-  add_body_row(top = FALSE, value = as_paragraph("Values are mean and (SD)"), colwidths = 9) %>%
   set_caption("Table 1. Participant characteristics")
 
-```
+dxa.flex <- border(dxa.flex, 
+                     i = NULL,
+                     j = NULL,
+                     border = NULL,
+                     border.top = fp_border_default(color = "grey"),
+                     border.bottom = fp_border_default(color = "grey"),
+                     border.left = fp_border_default(color = "grey"),
+                     border.right = fp_border_default(color = "grey"),
+                     part = "all")
+
+dxa.flex
+
+```
+Values are means ± SD.
+60º sec-1 = 60 degrees per second movement speed (isokinetic), 240º sec-1 = 240 degrees per second movement speed (isokinetic), 0º sec-1 = 0 degrees per second movement speed (isometric).
 
@@
 ## Flextable
-tab.dat %>%
+primer.tab <- tab.dat %>%
   flextable() %>%
@@
                     as_paragraph("Sequence (forward - reverse)"),
-                    as_paragraph("CT mean (SD)"),
+                    as_paragraph("Ct mean (SD)"),
                     as_paragraph("E"))) %>%
-  #add_header_row(values = c(" ", " ", " ", "Primer sequences"), 
-               #  colwidths = c(1, 1, 1, 1)) %>%
   set_caption("Table 2. Primer Sequences") %>%
-  width(j = 1, 1.5, unit = "in") %>%
-  width(j = 2, 3, unit = "in") %>%
-  width(j = 3, 1.5, unit = "in") %>%
-  width(j = 4, 0.5, unit = "in")
-  #set_table_properties(layout = "autofit")
-
-```
+  set_table_properties(layout = "autofit")
+
+primer.tab <- border(primer.tab, 
+                     i = NULL,
+                     j = NULL,
+                     border = NULL,
+                     border.top = fp_border_default(color = "grey"),
+                     border.bottom = fp_border_default(color = "grey"),
+                     border.left = fp_border_default(color = "grey"),
+                     border.right = fp_border_default(color = "grey"),
+                     part = "all")
+
+primer.tab
+
+```
+Values of CT are means ± SD. rRNA = ribosomal RNA, E = primer efficiency. Average cycle thresholds (Ct) and priming efficiencies were calculated from all qPCR reactions.

### 2023-08-09 figures/primer.tab.R

@@ -13,15 +13,25 @@ tab.dat <- primer.dat %>%
 
 
 ## Flextable
-tab.dat %>%
+primer.tab <- tab.dat %>%
   flextable() %>%
   compose(i = 1, j = c(1, 2, 3, 4), part = "header",
           value = c(as_paragraph("Gene"),
                     as_paragraph("Sequence (forward - reverse)"),
                     as_paragraph("CT mean (SD)"),
                     as_paragraph("E"))) %>%
-  #add_header_row(values = c(" ", " ", " ", "Primer sequences"), 
-               #  colwidths = c(1, 1, 1, 1)) %>%
-  set_caption("Table 2. Primer Sequences") %>%
+  set_caption("Table 2. Primer Sequences")
   set_table_properties(layout = "autofit")
+  
+
+
+primer.tab <- border(primer.tab, 
+                     i = NULL,
+                     j = NULL,
+                     border = NULL,
+                     border.top = fp_border_default(color = "grey"),
+                     border.bottom = fp_border_default(color = "grey"),
+                     border.left = fp_border_default(color = "grey"),
+                     border.right = fp_border_default(color = "grey"),
+                     part = "all")
 

### 2023-08-30 jap.manuscript.RMD
table and template updates
manuscript updates
updated jap.manuscript

@@
 Muscle hypertrophy, induced by resistance training (RT), is determined by ribosomal biogenesis.
-Synthesis of ribosomes may augment the transcription of ribosomal RNA, as evident from in vitro studies.
+Synthesis of ribosomes may augment the transcription of ribosomal RNA, as evident from in vitro studies<!-- what does this actually mean?-->.
 However, its effect on RT-induced adaptations remains unexplored.
-Sixteen healthy, moderately trained individuals (male/female, n = 9/7; age, 23.7 (1.8)/24.6 (4.8)), participated in a within-participant trial with unilateral resistance training with glucose- and placebo supplement on alternate days over six sessions per condition.
+Sixteen healthy, moderately trained individuals (male/female, n = 9/7; age, 23.7 (1.8)/24.6 (4.8)), participated in a within-participant trial with unilateral resistance training with glucose- and placebo supplementation on alternate days over six sessions per condition.
 Participants remained fasted overnight, receiving glucose/placebo and protein supplementation only before and after RT sessions (3 sets of 10 repetitions maximum unilateral leg press and knee extension).
 Micro biopsies were sampled from m.
-vastus lateralis pre and post the intervention, and maximal unilateral isometric and isokinetic knee extension torque were measured before, during (days 4, 5, 8, 9) and after the intervention.
-Between conditions, there were no significant differences in increases of total RNA and rRNA (Figure 2A; mean difference `r total_rna_stats`, Figure 2B; 47S, `r stat_rrna$trg_47s`; 18S, `r stat_rrna$trg_18s`; 28S, `r stat_rrna$trg_28s`; 5.8S, `r stat_rrna$trg_5.8s`; 5S, `r stat_rrna$trg_5s`), or levels of c-Myc, UBF and RPS6 (*p* = `r min(p_west)`-`r max(p_west)`; Figure 3A).
-There was a linear relationship between UBF and total RNA, where an increase of 1 SD unit of UBF equated to a $\sim$ `r corr_incr`% increase in total RNA (*p* = `r corrp`; Figure 3C).
-In conclusion, glucose supplement before and after RT does not augment RT-induced ribosome biogenesis after two weeks of heavy RT in moderately trained young adults.
+vastus lateralis before and after the intervention, and maximal unilateral isometric and isokinetic knee extension torque were measured before, during (days 4, 5, 8, 9) and after the intervention.
+Between conditions, there were no significant differences in increases of total RNA and rRNA (mean difference `r total_rna_stats`, 47S, `r stat_rrna$trg_47s`; 18S, `r stat_rrna$trg_18s`; 28S, `r stat_rrna$trg_28s`; 5.8S, `r stat_rrna$trg_5.8s`; 5S, `r stat_rrna$trg_5s`), or levels of c-Myc, UBF and RPS6 <!-- abbreviations ok here?--> (*p* = `r min(p_west)`-`r max(p_west)`).
+There was a linear relationship between UBF and total RNA, where an increase of 1 SD unit of UBF equated to a $\sim$ `r corr_incr`% increase in total RNA (*p* = `r corrp`).
+In conclusion, glucose supplementation before and after RT does not augment RT-induced ribosome biogenesis after two weeks of heavy RT in moderately trained young adults.
 
@@
 Responses to systematic resistance training (RT) vary widely between individuals, with as many as 10-15% showing impaired muscle growth in response to standardized training interventions [@alvarez_interindividual_2018; @mann_high_2014; @thalacker-mercer_cluster_2013].
-Such inability to benefit from RT may be explained by genetic and/or epigenetic disposition affecting the internal physiological milieu and subsequently affecting the ability to mount an anabolic response to RT [@figueiredo_genetic_2021;@thalacker-mercer_cluster_2013].<!-- Check these refs, do they correspond to genet epigen and internal mileu?
-Kristian: The way I read it, it looks like it corresponds --> Although RT program modifications such as altered exercise volume or intensity may induce more favourable adaptations in some individuals, these benefits are not uniform between individuals [@hammarstrom_benefits_2020;@krieger_single_2009;@schoenfeld_strength_2017].<!-- Add ref for more rt program modifications.
-Kristian: Added krieger et al.
-2009 and schoenfeld et al.
-2017 --> For instance, although changes in training volume generally induce greater benefits of RT, not all individuals benefit from such modifications [@hammarstrom_benefits_2020].
-Therefore, other means than modification of RT variables seem necessary to optimize individual responses.<!-- The above sentence might be a bit contradictory?
-Kristian: Im not sure, are you thinking about "other means than modification of RT" being contradictory of "modifications such as altered exercise volume.." sentence?
-I think there is a point to establishing that yes, increasing exercise volume is a potent way to increase the RT stimulus in general.
-However, in cases where this alone fails to elicit true responses, other means are necessary.
-Or am I missinterpreting this?
--->
-
-Nutritional supplements such as protein and creatine effectively optimize RT adaptations [@cermak_protein_2012;@lanhers_creatine_2015;@lanhers_creatine_2017;@morton_systematic_2018].
+Such inability to benefit from RT may be explained by genetic and/or epigenetic disposition affecting the internal physiological milieu and subsequently affecting the ability to mount an anabolic response to RT [@figueiredo_genetic_2021;@thalacker-mercer_cluster_2013].<!-- Rewrite this somewhat.
+Focus more on what can be done to alter the responsiveness to RT, rather than genetics.
+I.e.
+make the intervention seem more sensible --> Although RT program modifications such as altered exercise volume or intensity may induce more favourable adaptations in some individuals, these benefits are not uniform between individuals [@hammarstrom_benefits_2020;@krieger_single_2009;@schoenfeld_strength_2017].
+For instance, although increases in training volume generally induce greater benefits of RT, not all individuals benefit from such modifications [@hammarstrom_benefits_2020].
+Therefore, other means than modification of RT variables seem necessary to optimise individual responses.
+
+Nutritional supplements such as protein and creatine effectively optimise RT adaptations [@cermak_protein_2012;@lanhers_creatine_2015;@lanhers_creatine_2017;@morton_systematic_2018].
 However, it remains unknown if other nutritional adjuvants such as glucose can increase the efficacy of RT.
-This is surprising since glucose is the preferred energy substrate of the contracting skeletal muscle during strenuous exercise and a major supplier of energy to cells via ATP synthesis [@mul_exercise_2015;@sebata_control_2018].
-Furthermore, energy availability is a decisive factor in the *de novo* synthesis of ribosomes [@kusnadi_regulation_2015;@moss_housekeeper_2007;@sebata_control_2018] <!--Add ref that indicate high energy requirement for ribo biog possibly (VONWALDEN19;TANAKA18?
-MOSS.
-Kristian: These three should do it-->which in turn determines muscle growth by increasing the muscle's translational capacity [@figueiredo_regulation_2019;@hammarstrom_benefits_2020;@sebata_control_2018;@stec_ribosome_2016;@von_walden_ribosome_2019].
+This is surprising since glucose is the preferred energy substrate of the contracting skeletal muscle during strenuous exercise and a major energy supplier to cells via ATP synthesis [@mul_exercise_2015;@sebata_control_2018].
+Furthermore, energy availability is a decisive factor in the *de novo* synthesis of ribosomes [@kusnadi_regulation_2015;@moss_housekeeper_2007;@sebata_control_2018] which in turn determines muscle growth by increasing the muscle's translational capacity [@figueiredo_regulation_2019;@hammarstrom_benefits_2020;@sebata_control_2018;@stec_ribosome_2016;@von_walden_ribosome_2019].
 Therefore, investigating the effects of combined RT and glucose ingestion may provide valuable insight into the potential additive effect of glucose and RT on ribosome biogenesis.
 
-Transcription of ribosomal RNA (rRNA) by Pol I is considered the rate limiting step in synthesizing new ribosomes [@moss_promotion_1995].
+Transcription of ribosomal RNA (rRNA) by RNA Polymerase I is considered the rate-limiting step in synthesising new ribosomes [@moss_promotion_1995].
 Multiple signalling pathways converge to regulate rRNA transcription, including the mammalian target of rapamycin (mTORC1) signal-transduction pathway and c-Myc [@kusnadi_regulation_2015;@von_walden_ribosome_2019;@sebata_control_2018; @west_acute_2016; @mori_c-myc_2021].
@@
 
-The purpose of this investigation was to test the hypothesis that glucose supplementation before and after<!-- during?
-Kristian: before and after would be more precise, since they're not strictly ingesting anything during RT - I got a comment on that during the thesis defense --> RT increases markers of ribosomal abundance following five RT sessions.
+The purpose of this investigation was to test the hypothesis that glucose supplementation before and after RT increases markers of ribosomal abundance following five RT sessions.
 Secondly, we aimed to describe the association between UBF and total RNA abundance in human skeletal muscle.
-There have been multiple studies in the last years suggesting translational capacity to be as important, if not more, as translational efficiency in regard to long-term skeletal muscle adaptations to RT [@hammarstrom_benefits_2020;@hammarstrom_ribosome_2022;@figueiredo_revisiting_2019].
+There have been multiple studies in the last years suggesting translational capacity to be as important, if not more important than,  translational efficiency with regard to long-term skeletal muscle adaptations to RT [@hammarstrom_benefits_2020;@hammarstrom_ribosome_2022;@figueiredo_revisiting_2019].
 Previously, *in vitro* studies have shown UBF and c-Myc to be important factors in the PIC, c-Myc as a general transcription factor and UBF as a specific transcription factor for rDNA transcription initiation [@mariappan_ribosomal_2011;@von_walden_ribosome_2019].
 Coupled with the recent observations by Hammarström et al.
-[@hammarstrom_benefits_2020;@hammarstrom_ribosome_2022], understanding underlying factors regulating and affecting ribosome biogenesis seems key to furthering our understanding of the optimization of RT to individual phenotypes.
-<!-- We should maybe add a line or two about the rational here. Kristian: Tried to add some, what about this? to general? -->
+[@hammarstrom_benefits_2020;@hammarstrom_ribosome_2022], understanding underlying factors regulating and affecting ribosome biogenesis seems key to furthering our understanding of the optimisation of RT to individual phenotypes.
 
@@
 
-All participants gave their informed consent prior to study enrollment.
+All participants gave their written informed consent prior to study enrollment.
 The study was approved by the regional ethical committee (REK, ID nr.
@@
 2-8 RT sessions per 14 days for the last six months).
-Exclusion criteria were previous injury leading to impaired strength, inability to perform resistance exercise training, symptoms, and a medical record of metabolic disorders including hyperglycemia, i.e.
-fasting venous plasma glucose ≥6.1 mmol/L and/or 2-hour glucose tolerance ≥7.8 mmol/L, and/or HbA1c >42 mmol/mol.
+Exclusion criteria were previous injury leading to impaired strength, inability to perform resistance exercise training, symptoms, and a medical record of metabolic disorders including hyperglycemia.
 Of the sixteen participants that commenced the intervention, three participants dropped out.
 One due to sickness and inability to resume, two participants experienced muscular discomfort related to heavy resistance training.
-Lean mass (Table 1) was measured using DXA (Prodigy Advance PA+302047, Lunar, San Francisco, CA, USA) on Day -1, the last day preceding the RT intervention.
+Lean mass (Table 1) was measured using Dual-Energy X-ray Absorptiometry (Prodigy Advance PA+302047, Lunar, San Francisco, CA, USA) on Day -1, the last day preceding the RT intervention.
 
@@
 Participants were randomly allocated to exercise one leg with a glucose condition and one leg with a placebo condition (Figure 1A).
-<!--from the "old" methods: either their dominant or non-dominant leg with glucose or placebo before and after RT, and whether they started RT with glucose or placebo.
-Thus, participants exercised every day, alternating between e.g., one leg + glucose and the other leg + placebo - this might be TMI, as it's presented in the figure as well?--> One person was exclusively responsible for the randomization code and supplement distribution, blinding both investigators and participants regarding which leg exercised with glucose/placebo.
+One person was exclusively responsible for the randomisation code and supplement distribution, blinding both investigators and participants regarding which leg exercised with glucose/placebo conditions.
 Glucose (Glucosum monohydricum, Merck KGaA, Darmstadt, Germany) and placebo (Steviosa, Soma Nordic AS, Oslo, Norway) were masked by mixing with 300ml Fun Light (Orkla, Oslo, Norway).
@@
 on days 1-2, 3-4, etc.).
-To further standardize this, participants also recorded and repeated their daily macronutrient intake (protein, fat, carbohydrate) and total calories on pairwise consecutive days.
+To further standardise this, participants also recorded and repeated their daily macronutrient intake (protein, fat, carbohydrate) and total calories on pairwise consecutive days.
 
@@
 Participants were asked to avoid  resistance- or high-intensity training of the legs from Day -7 (Figure 1A) and onwards, until completion of the intervention and post-testing, to ensure the reliability of pre-intervention strength data and minimal interference from external exercise sources.
-If participants did conduct exercise training outside of the prescribed protocol, they were asked to ensure equal loading on pairwise consecutive days.
+If participants did conduct exercise training outside of the prescribed protocol, they were asked to ensure equal loading on pairwise consecutive days <!-- we do not have any data on whether this was done ore not, do we leave it out?-->.
 
@@
 Individual positions were recorded and standardized from pre-intervention tests (Figure 1A, days -7 and -5).
-Isokinetic peak torque was measured at extensions speeds of 60- and 240 d/s, 2x3 repetitions each, with the first set of each exercise as a sub-maximal warm-up.
-Isometric peak torque was measured at a knee-angle of 60 degrees (extended knee was set to 0 degrees), for a maximum of 10 seconds and two repetitions per test.
+Isokinetic peak torque was measured concentrically from 90 to 0 degrees knee angle (extended knee was set to 0 degrees) at angular velocities of 60- and 240 degrees per second, 2x3 repetitions each, with the first set of each exercise as a sub-maximal warm-up.
+Isometric knee extensor peak torque was measured at a knee-angle of 60 degrees, for a maximum og 10 seconds and two repetitions per test.
 The isometric tests were ended when the participants reached a plateau or peak torque development decreased, which on average occurred between 2-4 seconds into the test.
@@
 Test I on day 12/13 included testing of both legs, representing 23hrs post-RT session test of one leg and post-session 5 test of the other leg.
-The highest peak torque values from the respective extension speeds and time points were summarized in an index.
-The index was calculated by dividing the average peak torque value by the highest observed peak torque value per extension speed, and summarizing this new index per extension speed to a mean strength index.
+The highest peak torque values from the respective angular velocities and time points were summarized in an index.
+The index was calculated by dividing the average peak torque value by the highest observed peak torque value per angular velocity, and summarizing this new index per angular velocity to a mean strength index.
 
@@
 All positions were controlled and recorded at the first 1RM test and repeated for the RT sessions.
-Maximal leg press strength was defined as the maximal load lifted in a controlled fashion, with a knee angle of 45 degrees (start), within a 2.5kg range.
-Attempts where participants did not reach 45 degrees during the eccentric phase, were not approved.
-Maximal knee extension testing followed the same specific warm-up and pause protocol and was defined as maximal load lifted in a controlled fashion, reaching full extension of the knee joint, within a 1.25kg range.
+Maximal leg press strength was defined as the maximal load lifted in a controlled fashion, starting from a knee angle of 90 degrees.
+To find a reproducible 90 degree knee angle for each participant, centimeter markings on the side panels of the leg press machine was used to record were to find 90 degrees for each separate leg and participant.
+Attempts where participants did not reach 90 degrees during the eccentric phase, were not approved.
+Maximal knee extension testing followed the same specific warm-up was defined as maximal load lifted in a controlled fashion, reaching full extension of the knee joint.
 Attempts with exaggerated hip movement or beneath full extension were not approved.
-Two minutes of rest were given during the specific warm-up, and three minutes of rest was given between 1RM attempts.
+Two minutes of rest were given during the specific warm-up, and three minutes of rest were given between 1RM attempts.
 
@@
 
-Resistance training consisted of three sets of unilateral leg presses and three sets of knee extensions, with an intensity of 10 repetition maximum (10RM).
+Resistance training consisted of three sets of unilateral leg presses and three sets of knee extensions, with an intensity of 10 repetitions maximum (10RM).
 As a general warm-up, the participants cycled on an indoor exercise bicycle for 5-10min.
 In addition, before the respective exercises, two 10-repetition warm-up sets were completed at ~50% and ~70% of 10RM.
-To ensure an adequate exercise stimulation throughout the intervention, the exercise load was increased if the participants lifted more than 12 repetitions, as a progressive loading strategy.
+To ensure an adequate exercise stimulation throughout the intervention, the exercise load was increased the following set if the participants lifted more than 12 repetitions, as a progressive loading strategy.
+If the participants lifted fewer than 8 repetitions per set, the load was reduced in the following set.
 Rest between working sets was two minutes.
-For safety and standardization purposes, all sessions were monitored by trained personnel.
-<!-- From the "old" methods: The rate of perceived exertion (RPE, 0–10-point scale) was logged before every session, and the session score was logged 15min after each session - is this necessary info, as we are not reporting these data?
---> Lastly, training volume (load and repetitions) was logged for every session.
+For safety and standardisation purposes, all sessions were monitored by trained personnel.
+Lastly, training volume (load and repetitions) was logged for every session.
 
@@
 After the biopsy sampling, muscle tissue was divided into two aliquots for determination of total RNA/expression of rRNA and two aliquots for protein content measurement.
-Aliquots were snap frozen in isopentane (-80°C) and stored at -80°C until further analyses.
-Muscle biopsies were sampled at four times: I/II) Pre-intervention (Figure 1A, 2hrs before training, Day 1 = leg 1, Day 2 = leg 2), and III/IV) after the fifth RT session, two hours before the sixth RT session (Figure 1A, Day 11 = leg 1, Day 12 = leg 2).
+Aliquots were snap-frozen in isopentane (-80°C) and stored at -80°C until further analyses.
+Muscle biopsies were collected at four time points: I/II) Pre-intervention (Figure 1A, 2hrs before training, Day 1 = leg 1, Day 2 = leg 2), and III/IV) after the fifth RT session, two hours before the sixth RT session (Figure 1A, Day 11 = leg 1, Day 12 = leg 2).
 At each time point, two samples were taken from the same incision.
@@
 To measure blood glucose levels with and without glucose ingestion/training, capillary blood was collected from finger draws on days with biopsy sampling.
-One capillary blood sample was taken on day 1 (Figure 1A) to serve as a baseline.
-On days 11 and 12 (Figure 1A), capillary blood samples were collected seven times: I) Immediately prior to protein ingestion (0700hrs) II) 45 minutes after protein ingestion (0745hrs) III) 1.5hrs after protein ingestion (0800hrs, i.e., immediately before GLU/PLAC intake), IV) 2hrs after protein ingestion (0900hrs, i.e., immediately before training), IV) in the middle of RT (~0915hrs), V) immediately after training (~0930hrs), and VI) 2hrs after completion of training (~1130hrs).
+One capillary blood sample was collected on day 1 (Figure 1A) to serve as a baseline.
+On days 11 and 12 (Figure 1A), capillary blood samples were collected seven times: I) Immediately prior to protein ingestion (0700hrs) II) 45 minutes after protein ingestion (0745hrs) III) 1.5hrs after protein ingestion (0830hrs, i.e., immediately before GLU/PLAC intake), IV) 2hrs after protein ingestion (0900hrs, i.e., immediately before training), IV) in the middle of RT (~0915hrs), V) immediately after training (~0930hrs), and VI) 2hrs after completion of training (~1130hrs).
 Capillary blood samples were analysed with in-house equipment (BIOSEN C-Line, EKF diagnostic GmbH, Barleben).
@@
 qPCR reactions were run over 40 cycles (3 s 95°C denaturing and 30 s 60°C annealing) on a fast-cycling real-time detection system (Applied Biosystems 7500 fast Real-Time PCR Systems, Life Technologies AS), with a total reaction volume of 10 µl consisting of 2 µl of cDNA, gene-specific primers (0.5 µM final concentration) and a commercial master mix (2X SYBR Select Master Mix, Applied Biosystems, Life Technologies AS, Oslo, Norway).
+An overview of the primers may be found in table 2.
 Raw fluorescence data was modelled with a best-fit sigmoidal model using the qPCR package [@ritz_qpcr_2008] written for R [@hammarstrom_benefits_2020; @Rcore].
@@
 Following the wet transfer, membranes were stained with a reversible total protein stain (Thermo Fisher Scientific) and then blocked for 1hr at room temp with a blocking buffer of Tris-buffered saline (TBS; 20 mM Tris, 150 mM NaCl) with 5% non-fat dry milk and 0.1% Tween-20.
-Primary and secondary antibodies were purchased from Santa Cruz Biotechnology (Texas, USA): UBF, UBF F-9, sc-13125; rpS6, Ribosomal protein S6 C-8, sc-74459; and Thermo Fisher Scientific (Oslo, Norway): c-Myc, 9E10; goat anti-mouse (for c-Myc), goat anti-mouse IgG1 (y1) horseradish peroxidase conjugate; and anti-mouse (anti-mouse IgG1 horseradish peroxidase conjugate)<!-- insert manufacturer -->.
-Antibodies were diluted in blocking buffer to concentrations corresponding to <!-- Add final concentrations of antibodies -->.
+Primary and secondary antibodies were purchased from Santa Cruz Biotechnology (Texas, USA): UBF, UBF F-9, sc-13125; rpS6, Ribosomal protein S6 C-8, sc-74459; and Thermo Fisher Scientific (Oslo, Norway): c-Myc, 9E10; goat anti-mouse (for c-Myc), goat anti-mouse IgG1 (y1) horseradish peroxidase conjugate; and anti-mouse (anti-mouse IgG1 horseradish peroxidase conjugate).
+Antibodies were diluted in blocking buffer to concentrations corresponding to 1:25 000 (UBF, rpS6) and 1:5000 (c-Myc).
+
 Membranes were incubated overnight with primary antibodies, and for 1 hr with secondary antibodies.
@@
 A priori power calculations showed that 20 participants would grant a statistical power of 80% (alpha = 0.05), accounting for an expected dropout of 20%.
-This power calculation was based on an assumption that the effects of glucose ingestion on ribosome biogenesis (measured as within participant differences in total RNA) may equate to the effects of increasing RT volume from low to moderate [@hammarstrom_benefits_2020] corresponding to an effect size of d =  <!-- Add the effect size here-->.
+This power calculation was based on an assumption that the effects of glucose ingestion on ribosome biogenesis (measured as within-participant differences in total RNA) may equate to the effects of increasing RT volume from low to moderate [@hammarstrom_benefits_2020].
 Total RNA, protein and qPCR data were analyzed by mixed-effects models with fixed effects included as *supplement $\times$ time*.
@@
 These data only supported a random intercept per participant.
-The linear mixed effects models were fitted with the lmer function from the lme4 package using the lmerTest package to procure p-values (Satterthwaite's method for approximating degrees of freedom)[@bates_fitting_2014; @lmerTest], written for R[@Rcore].
-Log-transformed values were expressed as fold-changes in presentations.
+The linear mixed-effects models were fitted with the lmer function from the lme4 package using the lmerTest package to procure p-values (Satterthwaite's method for approximating degrees of freedom)[@bates_fitting_2014; @lmerTest], written for R[@Rcore].
+Log-transformed values were expressed as fold changes in visualisations.
 Descriptive data are presented as mean and (SD).
@@
 
-From baseline to post-intervention total RNA and rRNA per unit muscle weight increased on average by ~24% and ~33-43%, respectively (Figure 2).
+From baseline to post-intervention total RNA and rRNA per unit muscle weight increased on average by ~20-27% and ~25-57%, respectively (Figure 2).
 However, RT with glucose did not induce increased accumulation of total RNA (Figure 2A, mean difference `r total_rna_stats`) or rRNA  (Figure 2B; 47S, `r stat_rrna$trg_47s`; 18S, `r stat_rrna$trg_18s`; 28S, `r stat_rrna$trg_28s`; 5.8S, `r stat_rrna$trg_5.8s`; 5S, `r stat_rrna$trg_5s`) compared to RT with placebo.
@@
 low glucose/glucose starvation condition on rDNA transcription initiation observed in previous *in vitro* studies [@mariappan_ribosomal_2011; @tanaka_mild_2015] are not translatable to acute effects in human skeletal muscle *in vivo*.
-Notably, the present study was specifically designed to investigate differences in pre-post changes within participants, and is as such also quite limited to this perspective.
+<!-- Is there a point to discuss the fact that both legs may have exercised with sufficient glucose levels here, as the probably had full stores from the morning?
+-->Notably, the present study was specifically designed to investigate differences in pre to post changes within participants, and is as such also quite limited to this perspective.
 Hence, the observed effects of time serve as supplementary to the efficiency of the design, without a negative control group to compare with.
@@
 Hence, there were no differences in mechanical loading to induce a higher stress between conditions.
-Arguably, an increased energy availability via glucose ingestion during RT may induce less acute stress on the exercised skeletal muscle, therefore less fatigue, and perhaps less performance reduction, compared to placebo during RT <!-- this is possibly a stretch -->[@westerblad_mechanisms_1998; @kent-braun_central_1999] Unfortunately, we did not conduct measurements of markers of metabolic stress such ass inorganic phosphate, H+, Mg2+ and the ADP/ATP ratio [@westerblad_mechanisms_1998; @kent-braun_central_1999].
+Arguably, an increased energy availability via glucose ingestion during RT may induce less acute stress on the exercised skeletal muscle, therefore less fatigue, and perhaps less performance reduction, compared to placebo during RT <!-- this is possibly a stretch -->[@westerblad_mechanisms_1998; @kent-braun_central_1999] Unfortunately, we did not conduct measurements of markers of metabolic stress such as inorganic phosphate, H+, Mg2+ and the ADP/ATP ratio [@westerblad_mechanisms_1998; @kent-braun_central_1999].
 Therefore, discussing the potential effect of differences in metabolic stress between conditions would only be speculation, however probable.
@@
 
-Perspectives and Significance
+### Perspectives and Significance
 
@@
 Conceived and designed research: Stian Ellefsen, Håvard Hamarsland, Daniel Hammarström.
-Performed experiments: Kristian Lian, Sara Christine Moen.
+Performed experiments: Kristian Lian, Daniel Hammarström, Håvard Hamarsland, Håvard Nygaard, Sara Christine Moen.
 Interpreted results of experiments: Kristian Lian, Daniel Hammarström.

### 2023-09-01 jap.manuscript.RMD
updates

@@
 
-##### RESEARCH ARTICLE
-
-RUNNING HEAD: Glucose ingestion and resistance training in young adults
-
 #### Glucose ingestion before and after resistance training does not augment ribosome biogenesis in healthy moderately trained young adults
 
-###### Kristian Lian^1^, Sara Christine Olsen Moen^1^, Håvard Hamarsland^1^, Daniel Hammarström^1^, Stian Ellefsen^1^
+###### Kristian Lian^1^, Daniel Hammarström^1^, Håvard Hamarsland^1^, Sara Christine Olsen Moen^1^, Stian Ellefsen^1^
 
 <div custom-style="correspondence_style">
-^1^Section for Health and Exercise Physiology, Department of Public Health and Sport Sciences, Inladn Norway University of Applied Sciences, Lillehammer, Norway
+^1^Section for Health and Exercise Physiology, Department of Public Health and Sport Sciences, Inland Norway University of Applied Sciences, Lillehammer, Norway
 </div>
@@
 
-Muscle hypertrophy, induced by resistance training (RT), is determined by ribosomal biogenesis.
-Synthesis of ribosomes may augment the transcription of ribosomal RNA, as evident from in vitro studies<!-- what does this actually mean?-->.
-However, its effect on RT-induced adaptations remains unexplored.
-Sixteen healthy, moderately trained individuals (male/female, n = 9/7; age, 23.7 (1.8)/24.6 (4.8)), participated in a within-participant trial with unilateral resistance training with glucose- and placebo supplementation on alternate days over six sessions per condition.
+*Purpose:* Muscle hypertrophy, induced by resistance training, is determined by ribosomal biogenesis.
+High-glucose treatment may augment the transcription of ribosomal RNA, as evident from in vitro studies.
+However, its effect on resistance training-induced adaptations remains unexplored.
+
+*Methods:* Sixteen healthy, moderately trained individuals (male/female, n = 9/7; age, 23.7 (1.8)/24.6 (4.8)), participated in a within-participant trial with unilateral resistance training with glucose- and placebo supplementation on alternate days over six sessions per condition.
 Participants remained fasted overnight, receiving glucose/placebo and protein supplementation only before and after RT sessions (3 sets of 10 repetitions maximum unilateral leg press and knee extension).
 Micro biopsies were sampled from m.
-vastus lateralis before and after the intervention, and maximal unilateral isometric and isokinetic knee extension torque were measured before, during (days 4, 5, 8, 9) and after the intervention.
-Between conditions, there were no significant differences in increases of total RNA and rRNA (mean difference `r total_rna_stats`, 47S, `r stat_rrna$trg_47s`; 18S, `r stat_rrna$trg_18s`; 28S, `r stat_rrna$trg_28s`; 5.8S, `r stat_rrna$trg_5.8s`; 5S, `r stat_rrna$trg_5s`), or levels of c-Myc, UBF and RPS6 <!-- abbreviations ok here?--> (*p* = `r min(p_west)`-`r max(p_west)`).
-There was a linear relationship between UBF and total RNA, where an increase of 1 SD unit of UBF equated to a $\sim$ `r corr_incr`% increase in total RNA (*p* = `r corrp`).
-In conclusion, glucose supplementation before and after RT does not augment RT-induced ribosome biogenesis after two weeks of heavy RT in moderately trained young adults.
+vastus lateralis before and after the intervention.
+
+*Results:* Between conditions, there were no significant differences in increases of total RNA and ribosomal RNA (mean difference `r total_rna_stats`, 47S, `r stat_rrna$trg_47s`; 18S, `r stat_rrna$trg_18s`; 28S, `r stat_rrna$trg_28s`; 5.8S, `r stat_rrna$trg_5.8s`; 5S, `r stat_rrna$trg_5s`), or levels of c-Myc, upstream binding factor and ribosomal protein S6 (*p* = `r min(p_west)`-`r max(p_west)`).
+There was a linear relationship between upstream binding factor and total RNA, where an increase of 1 standard deviance unit of upstream binding factor equated to a $\sim$ `r corr_incr`% increase in total RNA (*p* = `r corrp`).
+
+*Conclusion:* Glucose supplementation before and after RT does not augment RT-induced ribosome biogenesis after two weeks of heavy RT in moderately trained young adults.
 
@@
 
-## NEW AND NOTEWORTHY
-The effect of glucose ingestion on resistance training-induced ribosome biogenesis is quite unexplored in human skeletal muscle, and our results show no beneficial acute effect following six bouts of exercise.
-The linear relationship observed between UBF and total RNA do solidify the importance of UBF in human skeletal muscle ribosome biogenesis.
+*Keywords:* Glucose; hypertrophy; resistance training; ribosome; skeletal muscle
 
@@
 
-**Keywords:** Glucose; hypertrophy; resistance training; ribosome; skeletal muscle
-
-</div>
+# Statements and Declarations
+????
+
+# Abbreviations
+Make a table for this
 
@@
 
-Responses to systematic resistance training (RT) vary widely between individuals, with as many as 10-15% showing impaired muscle growth in response to standardized training interventions [@alvarez_interindividual_2018; @mann_high_2014; @thalacker-mercer_cluster_2013].
-Such inability to benefit from RT may be explained by genetic and/or epigenetic disposition affecting the internal physiological milieu and subsequently affecting the ability to mount an anabolic response to RT [@figueiredo_genetic_2021;@thalacker-mercer_cluster_2013].<!-- Rewrite this somewhat.
-Focus more on what can be done to alter the responsiveness to RT, rather than genetics.
-I.e.
-make the intervention seem more sensible --> Although RT program modifications such as altered exercise volume or intensity may induce more favourable adaptations in some individuals, these benefits are not uniform between individuals [@hammarstrom_benefits_2020;@krieger_single_2009;@schoenfeld_strength_2017].
-For instance, although increases in training volume generally induce greater benefits of RT, not all individuals benefit from such modifications [@hammarstrom_benefits_2020].
-Therefore, other means than modification of RT variables seem necessary to optimise individual responses.
+Responses to systematic resistance training (RT) vary widely between individuals, with as much as 10-15% showing impaired skeletal muscle growth in response to standardized training interventions [@alvarez_interindividual_2018; @mann_high_2014; @thalacker-mercer_cluster_2013].
+While there certainly may be differences in the genetic predisposition in terms of readiness to adapt to exercise [@thalacker-mercer_cluster_2013], there are also recent studies showing that epigenetic factors such as ribosomal DNA (rDNA) availability may also impact the ability to mount an anabolic response to RT [@figueiredo_genetic_2021]<!--check this source -->.
+Indeed, the ability to synthesise ribosomes following a period of RT is connected to the magnitude of RT-induced responses in an individual [@hammarstrom_benefits_2020].
+For instance, increasing training volume generally induce greater ribosome biogenesis and thus, greater benefits of RT [@hammarstrom_benefits_2020;@krieger_single_2009;@schoenfeld_strength_2017].
+However, as evident from Hammarström et al.
+(2020), not all participants experienced increased muscle mass and -strength despite increasing training volume.
+Therefore, other means than modification of RT variables alone seem necessary to optimise individual responses to RT.
 
@@
 However, it remains unknown if other nutritional adjuvants such as glucose can increase the efficacy of RT.
-This is surprising since glucose is the preferred energy substrate of the contracting skeletal muscle during strenuous exercise and a major energy supplier to cells via ATP synthesis [@mul_exercise_2015;@sebata_control_2018].
+This is surprising since glucose is the preferred energy substrate of the contracting skeletal muscle during strenuous exercise and a major energy supplier to cells via adenosine triphosphate (ATP) synthesis [@mul_exercise_2015;@sebata_control_2018].
 Furthermore, energy availability is a decisive factor in the *de novo* synthesis of ribosomes [@kusnadi_regulation_2015;@moss_housekeeper_2007;@sebata_control_2018] which in turn determines muscle growth by increasing the muscle's translational capacity [@figueiredo_regulation_2019;@hammarstrom_benefits_2020;@sebata_control_2018;@stec_ribosome_2016;@von_walden_ribosome_2019].
@@
 mTOR receives input from growth factors, hormones, mechanical loading, and nutrients to balance protein synthesis through multiple mechanisms based on cellular energy levels [@hoppe_amp-activated_2009].
-Importantly, mTOR and its downstream target S6K1 serve as a mediator for insulin signalling in skeletal muscle [@hillier_physiological_2000].
+Importantly, mTOR and its downstream target ribosomal protein S6 kinase beta-1 (S6K1) serve as a mediator for insulin signalling in skeletal muscle [@hillier_physiological_2000].
 The mTORC1 pathway contributes to ribosome biogenesis through the regulation of the translation of ribosomal proteins and to forming of the preinitiation complex (PIC) that marks the initiation of rRNA transcription [@figueiredo_regulation_2019;@von_walden_ribosome_2019].
@@
 Secondly, we aimed to describe the association between UBF and total RNA abundance in human skeletal muscle.
-There have been multiple studies in the last years suggesting translational capacity to be as important, if not more important than,  translational efficiency with regard to long-term skeletal muscle adaptations to RT [@hammarstrom_benefits_2020;@hammarstrom_ribosome_2022;@figueiredo_revisiting_2019].
+There have been multiple studies recently suggesting translational capacity to be as important, if not more important than, translational efficiency with regard to long-term skeletal muscle adaptations to RT [@hammarstrom_benefits_2020;@hammarstrom_ribosome_2022;@figueiredo_revisiting_2019].
 Previously, *in vitro* studies have shown UBF and c-Myc to be important factors in the PIC, c-Myc as a general transcription factor and UBF as a specific transcription factor for rDNA transcription initiation [@mariappan_ribosomal_2011;@von_walden_ribosome_2019].
@@
 One due to sickness and inability to resume, two participants experienced muscular discomfort related to heavy resistance training.
-Lean mass (Table 1) was measured using Dual-Energy X-ray Absorptiometry (Prodigy Advance PA+302047, Lunar, San Francisco, CA, USA) on Day -1, the last day preceding the RT intervention.
+Lean mass (Table 1) was measured using Dual-Energy X-ray Absorptiometry (DXA, Prodigy Advance PA+302047, Lunar, San Francisco, CA, USA) on Day -1, the last day preceding the RT intervention.
 
@@
 Participants were asked to avoid  resistance- or high-intensity training of the legs from Day -7 (Figure 1A) and onwards, until completion of the intervention and post-testing, to ensure the reliability of pre-intervention strength data and minimal interference from external exercise sources.
-If participants did conduct exercise training outside of the prescribed protocol, they were asked to ensure equal loading on pairwise consecutive days <!-- we do not have any data on whether this was done ore not, do we leave it out?-->.
 
@@
 
-The dietary intervention spanned the whole day, divided into three periods: I) From awakening until 2.5 hours after RT, II) from 2.5 hours after RT until 2200hrs, and III) from 2200hrs until awakening.
-During period I, participants ingested protein and glucose/placebo only.
+The dietary intervention spanned the whole day, divided into three periods: I) From awakening until 2.5 hours (hrs) after RT, II) from 2.5hrs after RT until 2200hrs, and III) from 2200hrs until awakening.
+During period I, participants ingested protein and glucose **or** protein and placebo only.
 Glucose/placebo was ingested three times in period I: 30 minutes before RT (0830hrs, 30g vs 0g glucose), immediately before RT (0900hrs, 30g vs.
@@
 During period III (2200-0700hrs), participants remained in an overnight fasted state.
-The daily onset of the dietary intervention (i.e., first ingestion of PRO supplement) varied between participants, from 0600hrs to 0900hrs to allow multiple participants to complete the protocol simultaneously.
+The daily onset of the dietary intervention (i.e., first ingestion of protein supplement) varied between participants, from 0600hrs to 0900hrs to allow multiple participants to complete the protocol simultaneously.
 During sessions, participants were free to ingest water ad libitum.

### 2023-09-05 ejap.manuscript.RMD
updates

--- 2023-09-01 jap.manuscript.RMD
@@
 
-*Results:* Between conditions, there were no significant differences in increases of total RNA and ribosomal RNA (mean difference `r total_rna_stats`, 47S, `r stat_rrna$trg_47s`; 18S, `r stat_rrna$trg_18s`; 28S, `r stat_rrna$trg_28s`; 5.8S, `r stat_rrna$trg_5.8s`; 5S, `r stat_rrna$trg_5s`), or levels of c-Myc, upstream binding factor and ribosomal protein S6 (*p* = `r min(p_west)`-`r max(p_west)`).
+*Results:* Between conditions, there were no significant differences in increases of total RNA and ribosomal RNA (mean difference `r total_rna_stats`, 47S, `r stat_rrna$trg_47s`; 18S, `r stat_rrna$trg_18s`; 28S, `r stat_rrna$trg_28s`; 5.8S, `r stat_rrna$trg_5.8s`; 5S, `r stat_rrna$trg_5s`), or levels of cellular myelocutmatosis oncogene, upstream binding factor and ribosomal protein S6 (*p* = `r min(p_west)`-`r max(p_west)`).
 There was a linear relationship between upstream binding factor and total RNA, where an increase of 1 standard deviance unit of upstream binding factor equated to a $\sim$ `r corr_incr`% increase in total RNA (*p* = `r corrp`).
@@
 Resistance training consisted of three sets of unilateral leg presses and three sets of knee extensions, with an intensity of 10 repetitions maximum (10RM).
-As a general warm-up, the participants cycled on an indoor exercise bicycle for 5-10min.
+As a general warm-up, the participants cycled on an indoor exercise bicycle for 5-10 minutes (min).
 In addition, before the respective exercises, two 10-repetition warm-up sets were completed at ~50% and ~70% of 10RM.
@@
 All samples were reverse transcribed and diluted to 1:50 before quantitative real-time polymerase chain reaction (qPCR).
-qPCR reactions were run over 40 cycles (3 s 95°C denaturing and 30 s 60°C annealing) on a fast-cycling real-time detection system (Applied Biosystems 7500 fast Real-Time PCR Systems, Life Technologies AS), with a total reaction volume of 10 µl consisting of 2 µl of cDNA, gene-specific primers (0.5 µM final concentration) and a commercial master mix (2X SYBR Select Master Mix, Applied Biosystems, Life Technologies AS, Oslo, Norway).
+qPCR reactions were run over 40 cycles (3 s 95°C denaturing and 30 s 60°C annealing) on a fast-cycling real-time detection system (Applied Biosystems 7500 fast Real-Time PCR Systems, Life Technologies AS), with a total reaction volume of 10 µl consisting of 2 µl of complementary DNA (cDNA), gene-specific primers (0.5 µM final concentration) and a commercial master mix (2X SYBR Select Master Mix, Applied Biosystems, Life Technologies AS, Oslo, Norway).
 An overview of the primers may be found in table 2.
@@
 
-There was no difference in mean change of total training session volume between exercising with placebo or glucose, with an increase of 18 and 16% from session 1 to session 6, respectively (Table 2).<!-- Add statistics to Table 2 -->
-There were no differences in mean macronutrient intake (protein, fat, carbohydrates) or total calorie intake between glucose and placebo on pairwise consecutive days (*p* > 0.05 for all).<!-- Should this also be in Table 2?
--->
+There was no difference in mean change of total training session volume between exercising with placebo or glucose, with an increase of 18 and 16% from session 1 to session 6, respectively (*p* > 0.05).
+There were no differences in mean macronutrient intake (protein, fat, carbohydrates) or total calorie intake between glucose and placebo on pairwise consecutive days (*p* > 0.05 for all).
 
@@
 Further, glucose ingestion increases energy levels, leading us to hypothesize that high glucose vs.
-placebo condition would stimulate energy-sensitive pathways such as PIH1, mTORC1, ERK1/2, AMPK and SIRT1 [@kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012].
+placebo condition would stimulate energy-sensitive pathways such as mTORC1, PIH1, extracellular signal-regulated kinase 1/2 (ERK1/2), AMP-dependent protein kinase (AMPK) and Sirtuin 1 (SIRT1) [@kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012].
 Despite previously reported upregulation in PIC assembly due to high-glucose mediated mTORC1, ERK1/2 and PIH1 or low-glucose mediated AMPK and SIRT1 activation  [@hoppe_amp-activated_2009; @kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012], the present study showed no signs of such effects of glucose vs.
@@
 Between days -7 and -1, participants were familiarized to the RT exercises via 1RM leg press and knee extension testing, and to the strength tests via Humac Norm dynamometer (days -7 and -5).
-Before baseline testing, the participants were randomly allocated to exercise one leg with GLU (glucose) and the other with PLA (placebo), in a unilateral, alternating fashion.
+Before baseline testing, the participants were randomly allocated to exercise one leg with glucose (GLU) and the other with placebo (PLA), in a unilateral, alternating fashion.
 Further, non-dominant/dominant + GLU/PLA, and onset with GLU or PLA was also randomized, the figure illustrates an example where the participant was randomized to start RT with GLU.
@@
 Blood for measurement of plasma glucose and -c-peptide was sampled at baseline (Day 1), and during post testing (Day 11 leg 1, Day 12 leg 2), via finger draws and venous blood samples.
-Skeletal muscle strength was measured as peak torque in unilateral isometric and isokinetic (at 60 and 240 d/s) knee extension before, multiple times during, and after five and six session.
+Skeletal muscle strength was measured as peak torque in unilateral isometric and isokinetic (at 60 and 240 degrees per second) knee extension before, multiple times during, and after five and six session.
 <!-- A total of three participants dropped out of the intervention, either due to reported muscular discomforts of heavy RT, or sickness in their family - necessary to include this here?
 Its already reported under "experimental design" -->.
-B and C) Changes in plasma glucose (B, mmol/L) and c-peptide levels (C, pmol/L).
+B and C) Changes in plasma glucose (B, mmol/L) and serum c-peptide levels (C, pmol/L).
 Glucose levels in blood was measured via finger draws 120 (-120), 90 (-90), and 30min (-30) before RT, immediately before RT (0), during RT (15), immediately after RT (30) and 2hrs after RT (120).
@@
 The index was calculated by normalizing peak torque values to the highest peak torque value at each respective speed, and then summarized and used in change score calculations.
-Values are presented as changes in estimated marginal means ± 95% CI.
+Values are presented as changes in estimated marginal means ± 95% confidence intervals (CI).
 * = *p* < 0.05 between groups.

### 2023-09-12 ejap.manuscript.RMD
discussion updates

@@
 
-ORCID:<!-- I does not say anything about ORCID in the journal template, do we leave it in?-->
+ORCID:
 <div custom-style="correspondence_style">
@@
 Responses to systematic resistance training (RT) vary widely between individuals, with as much as 10-15% showing impaired skeletal muscle growth in response to standardized training interventions [@alvarez_interindividual_2018; @mann_high_2014; @thalacker-mercer_cluster_2013].
-While there certainly may be differences in the genetic predisposition in terms of readiness to adapt to exercise [@thalacker-mercer_cluster_2013], there are also recent studies showing that epigenetic factors such as ribosomal DNA (rDNA) availability may also impact the ability to mount an anabolic response to RT [@figueiredo_genetic_2021]<!--check this source -->.
+While there certainly may be differences in the genetic predisposition in terms of readiness to adapt to exercise [@thalacker-mercer_cluster_2013], there are also recent studies showing that epigenetic factors such as ribosomal DNA (rDNA) availability may also impact the ability to mount an anabolic response to RT [@figueiredo_genetic_2021].
 Indeed, the ability to synthesise ribosomes following a period of RT is connected to the magnitude of RT-induced responses in an individual [@hammarstrom_benefits_2020].
@@
 low glucose/glucose starvation condition on rDNA transcription initiation observed in previous *in vitro* studies [@mariappan_ribosomal_2011; @tanaka_mild_2015] are not translatable to acute effects in human skeletal muscle *in vivo*.
-<!-- Is there a point to discuss the fact that both legs may have exercised with sufficient glucose levels here, as the probably had full stores from the morning?
--->Notably, the present study was specifically designed to investigate differences in pre to post changes within participants, and is as such also quite limited to this perspective.
+Notably, the present study was specifically designed to investigate differences in pre to post changes within participants, and is as such also quite limited to this perspective.
 Hence, the observed effects of time serve as supplementary to the efficiency of the design, without a negative control group to compare with.
@@
 Unfortunately, three dropped out during the intervention, leaving our statistical power slightly underpowered according to the a priori calculation.
+Furhter, we did not measure the intramuscular glycogen storage, representing another limitation to our design.
+Hence, we cannot know for certain whether there were actual differences within the exercising skeletal muscles between the glucose and placebo condition.
+The participants did indeed match their macro nutrients on pairwise consecutive days, and the plasma glucose measurements did not show any differences in blood sugar until ingestion of placebo/glucose.
+Thus, it may stand to reason that the participants had similar levels of intramuscular levels before ingesting any supplement and that the glucose supplement led to higher glucose availability during exercise, as by design.
+However, even if the leg exercising with glucose supplement did in fact take up the extra glucose, it did not seem to affect any of our outcome measures.
 

### 2023-09-12 jap.manuscript.RMD
discussion updates

@@
 Responses to systematic resistance training (RT) vary widely between individuals, with as much as 10-15% showing impaired skeletal muscle growth in response to standardized training interventions [@alvarez_interindividual_2018; @mann_high_2014; @thalacker-mercer_cluster_2013].
-While there certainly may be differences in the genetic predisposition in terms of readiness to adapt to exercise [@thalacker-mercer_cluster_2013], there are also recent studies showing that epigenetic factors such as ribosomal DNA (rDNA) availability may also impact the ability to mount an anabolic response to RT [@figueiredo_genetic_2021]<!--check this source -->.
+While there certainly may be differences in the genetic predisposition in terms of readiness to adapt to exercise [@thalacker-mercer_cluster_2013], there are also recent studies showing that epigenetic factors such as ribosomal DNA (rDNA) availability may also impact the ability to mount an anabolic response to RT, given the importance of translational capacity in this regard [@figueiredo_genetic_2021].
 Indeed, the ability to synthesise ribosomes following a period of RT is connected to the magnitude of RT-induced responses in an individual [@hammarstrom_benefits_2020].

### 2023-09-26 ejap.manuscript.RMD
added abbreviation table

@@
 
+```{r abbreviation table}
+### Abbreviations table
+
+## Packages
+library(readxl); library(tidyverse); library(flextable)
+
+## Data
+abbreviations.dat <- read_excel("./data/abbreviations.xlsx")
+
+tab.dat <- abbreviations.dat
+
+## Flextable
+abb.tab <- tab.dat %>%
+  flextable() %>%
+  set_caption(caption = "Abbreviations",
+              align_with_table = FALSE) %>%
+  set_table_properties(layout = "autofit")
+
+abb.tab <- delete_part(x = abb.tab, part = "header")
+
+abb.tab <- border(abb.tab, 
+                     i = NULL,
+                     j = NULL,
+                     border = NULL,
+                     border.top = fp_border_default(color = "black"),
+                     border.bottom = fp_border_default(color = "black"),
+                     border.left = fp_border_default(color = "black"),
+                     border.right = fp_border_default(color = "black"),
+                     part = "all")
+
+abb.tab
+```
+
 </div>
@@
 # Statements and Declarations
-????
-
-# Abbreviations
-Make a table for this
+This study was financed through research funds from the Inland Norway University of Applied Sciences.
+There were no economical vested interest and everyone participating as researchers and associates, received only a regular salary during this study.
 
@@
 # AUTHOR CONTRIBUTIONS
-Conceived and designed research: Stian Ellefsen, Håvard Hamarsland, Daniel Hammarström.
-Performed experiments: Kristian Lian, Daniel Hammarström, Håvard Hamarsland, Håvard Nygaard, Sara Christine Moen.
-Interpreted results of experiments: Kristian Lian, Daniel Hammarström.
-Prepared Figures: Kristian Lian, Daniel Hammarström.
-Drafted manuscripts: Kristian Lian, Daniel Hammarström.
-Edited and revised manuscript: Stian Ellefsen, Håvard Hamarsland, Daniel Hammarström.
+Conceived and designed research: SE, HH, DH.
+Performed experiments: KL, DH, HH, HN, SCM.
+Interpreted results of experiments: KL, DH.
+Prepared Figures: KL, DH.
+Drafted manuscripts: KL, DH.
+Edited and revised manuscript: SE, HH, DH.
 Approved final version of the manuscript: .
@@
   add_header_row(values = c(" ", "Knee-extension peak torque"), colwidths = c(6, 3)) %>%
-  set_caption("Table 1. Participant characteristics")
+  set_caption(caption = "Table 1. Participant characteristics",
+              align_with_table = FALSE)
 
@@
                      border = NULL,
-                     border.top = fp_border_default(color = "grey"),
-                     border.bottom = fp_border_default(color = "grey"),
-                     border.left = fp_border_default(color = "grey"),
-                     border.right = fp_border_default(color = "grey"),
+                     border.top = fp_border_default(color = "black"),
+                     border.bottom = fp_border_default(color = "black"),
+                     border.left = fp_border_default(color = "black"),
+                     border.right = fp_border_default(color = "black"),
                      part = "all")
@@
                     as_paragraph("E"))) %>%
-  set_caption("Table 2. Primer Sequences") %>%
+  set_caption(caption = "Table 2. Primer Sequences",
+              align_with_table = FALSE) %>%
   set_table_properties(layout = "autofit")

### 2023-09-26 figures/abbreviations.table.R

@@ -1,0 +1,37 @@
+### Primer sequence table
+
+## Packages
+library(readxl); library(tidyverse); library(flextable)
+
+## Data
+abbreviations.dat <- read_excel("./data/abbreviations.xlsx")
+
+tab.dat <- abbreviations.dat
+
+
+## Flextable
+abb.tab <- tab.dat %>%
+  flextable() %>%
+  #print()
+  #compose(i = 1, j = c(1, 2), part = "header",
+   #       value = c(as_paragraph(""),
+    #                as_paragraph(""))) %>%
+  #set_caption("") %>%
+  set_table_properties(layout = "autofit")
+
+abb.tab <- delete_part(x = abb.tab, part = "header")
+  
+
+
+
+abb.tab <- border(abb.tab, 
+                     i = NULL,
+                     j = NULL,
+                     border = NULL,
+                     border.top = fp_border_default(color = "black"),
+                     border.bottom = fp_border_default(color = "black"),
+                     border.left = fp_border_default(color = "black"),
+                     border.right = fp_border_default(color = "black"),
+                     part = "all") 
+
+abb.tab

### 2023-09-27 jap.manuscript.RMD
updated results

@@
 Both RT with glucose and placebo led to significantly reduced strength post-fifth session compared to baseline, by 11 and 18% respectively (Figure 1D, Post 5RT: *p* = 0.000).
-Comparisons of the acute data gathered from the post-fifth session until and including 23hrs post-sixth session showed a significant mean increase in strength of 5-9% after RT with glucose and placebo 30min after the sixth RT session (Figure 1D, 30min post 6RT: *p* = `r apc2`) and two hours after the sixth RT session (Figure 1D, 2h post 6RT: *p* = `r apc3`).
+Comparisons of the acute data gathered from the post-fifth session until and including 23hrs post-sixth session showed a significant mean increase in strength of 5-9% after RT with glucose and placebo, 30min after the sixth RT session (Figure 1D, 30min post 6RT: *p* = `r apc2`) and two hours after the sixth RT session (Figure 1D, 2h post 6RT: *p* = `r apc3`).
 Twenty-three hours after the last (sixth) RT session, strength was unchanged compared to after the fifth RT session (Figure 1D, 23h post 6RT: *p* = `r apc4`).
@@
 
-From baseline to post-intervention total RNA and rRNA per unit muscle weight increased on average by ~20-27% and ~25-57%, respectively (Figure 2).
+From baseline to after the intervention, total RNA and rRNA per unit muscle weight increased on average by ~20-27% and ~25-57%, respectively (Figure 2).
 However, RT with glucose did not induce increased accumulation of total RNA (Figure 2A, mean difference `r total_rna_stats`) or rRNA  (Figure 2B; 47S, `r stat_rrna$trg_47s`; 18S, `r stat_rrna$trg_18s`; 28S, `r stat_rrna$trg_28s`; 5.8S, `r stat_rrna$trg_5.8s`; 5S, `r stat_rrna$trg_5s`) compared to RT with placebo.
@@
 
-Overall, RT led to robust increases from pre- to post-intervention in all measured proteins (Figure 3A).
-RT with glucose resulted in estimated levels of c-Myc, UBF and RPS6 being `r stat_west$cmyc`, `r stat_west$ubf` and `r stat_west$rps6`% lower compared to placebo, respectively, without showing statistical significance (*p* = `r min(p_west)`-`r max(p_west)`; Figure 3A).
+Overall, RT led to robust increases in all measured proteins from before to after the intervention (Figure 3A).
+RT with glucose resulted in estimated levels of c-Myc, UBF and RPS6 being `r stat_west$cmyc`, `r stat_west$ubf` and `r stat_west$rps6`% lower compared to placebo, respectively, without showing a statistically significant difference between conditions (*p* = `r min(p_west)`-`r max(p_west)`; Figure 3A).
 There was a linear relationship between UBF and total RNA, where an increase of 1 SD unit of UBF equated to a $\sim$ `r corr_incr`% increase in total RNA (*p* = `r corrp`; Figure 3C).

### 2023-09-27 ejap.manuscript.RMD
small updates

@@
 There were no economical vested interest and everyone participating as researchers and associates, received only a regular salary during this study.
+
+## Acknowledgements
+Great thanks are in order to Henriette Spilhaug Bollandsås, Nora Lerdalen Bonsak and Thea Bøhn Nyløkken for their impressive commitment and dedication in the training intervention
+the Inland Norway University of Applied Sciences, for funding this research.
+Gratitude is also due to Daniel Hammarström, Håvard Hamarsland and Prof.
+Stian Ellefsen for the guidance and <!-- needs more work -->
 
@@
 
-### Perspectives and Significance
+### Conclusion
 
@@
 
-# DATA AVAILABILITY
+## DATA AVAILABILITY
 
@@
 
-# SUPPLEMENTAL MATERIAL
+## SUPPLEMENTAL MATERIAL
 
@@
 
-# AUTHOR CONTRIBUTIONS
+## AUTHOR CONTRIBUTIONS
 Conceived and designed research: SE, HH, DH.
@@
 
-# References
+## References
 

### 2023-09-28 ejap.manuscript.RMD
working on references

--- 2023-09-26 ejap.manuscript.RMD
@@
 bibliography: resources/ribose-paper.bib
-csl: resources/jap.csl
+csl: resources/ejap.csl
 output:
@@
 There were no economical vested interest and everyone participating as researchers and associates, received only a regular salary during this study.
+
+## AUTHOR CONTRIBUTIONS
+
+Data was collected at the sports science lab at the Inland Norway University of Applied Sciences, with all analyses except for serum blood samples were performed at the university.
+Serum blood samples were performed at Innlandet Hospital Trust.
+S.E., H.H.
+and D.H.
+conceived and designed the study; K.L., S.C.M., D.H.
+and H.H.
+performed experiements; K.L.
+and D.H.
+interpreted the results; K.L.
+and D.H.
+prepared figures; K.L.
+drafted the manuscript; D.H., H.H.
+and S.E.
+edited and revised the manuscript; D.H., H.H.
+and S.E.
+approved the final version of the manuscript.
+
+## Acknowledgements
+Gratitude are in order to Henriette Spilhaug Bollandsås, Nora Lerdalen Bonsak and Thea Bøhn Nyløkken for their dedication and hard work in the training intervention.
+The authors would also like to express their gratitude to all the participants, staying commited to this demanding protocol in an impressive way.
+We also thank Håvard Nygaard for lending your expertise and technical skills to improve this project.
 
@@
 
-# AUTHOR CONTRIBUTIONS
-Conceived and designed research: SE, HH, DH.
-Performed experiments: KL, DH, HH, HN, SCM.
-Interpreted results of experiments: KL, DH.
-Prepared Figures: KL, DH.
-Drafted manuscripts: KL, DH.
-Edited and revised manuscript: SE, HH, DH.
-Approved final version of the manuscript: .
-
-# References
+## References
 

### 2023-09-28 ejap.manuscript.RMD
ejap format

@@
   word_document:
-    reference_docx: resources/reference-template-02.docx
+    reference_docx: resources/ejap-template.docx
 editor_options: 
@@
 
-#### Glucose ingestion before and after resistance training does not augment ribosome biogenesis in healthy moderately trained young adults
+# Glucose ingestion before and after resistance training does not augment ribosome biogenesis in healthy moderately trained young adults
 
@@
 <div custom-style="abstract_style">
-# ABSTRACT
+## Abstract
 
@@
 
-*Purpose:* Muscle hypertrophy, induced by resistance training, is determined by ribosomal biogenesis.
+**Purpose:** Muscle hypertrophy, induced by resistance training, is determined by ribosomal biogenesis.
 High-glucose treatment may augment the transcription of ribosomal RNA, as evident from in vitro studies.
@@
 
-*Methods:* Sixteen healthy, moderately trained individuals (male/female, n = 9/7; age, 23.7 (1.8)/24.6 (4.8)), participated in a within-participant trial with unilateral resistance training with glucose- and placebo supplementation on alternate days over six sessions per condition.
+**Methods:** Sixteen healthy, moderately trained individuals (male/female, n = 9/7; age, 23.7 (1.8)/24.6 (4.8)), participated in a within-participant trial with unilateral resistance training with glucose- and placebo supplementation on alternate days over six sessions per condition.
 Participants remained fasted overnight, receiving glucose/placebo and protein supplementation only before and after RT sessions (3 sets of 10 repetitions maximum unilateral leg press and knee extension).
@@
 
-*Results:* Between conditions, there were no significant differences in increases of total RNA and ribosomal RNA (mean difference `r total_rna_stats`, 47S, `r stat_rrna$trg_47s`; 18S, `r stat_rrna$trg_18s`; 28S, `r stat_rrna$trg_28s`; 5.8S, `r stat_rrna$trg_5.8s`; 5S, `r stat_rrna$trg_5s`), or levels of cellular myelocutmatosis oncogene, upstream binding factor and ribosomal protein S6 (*p* = `r min(p_west)`-`r max(p_west)`).
+**Results:** Between conditions, there were no significant differences in increases of total RNA and ribosomal RNA (mean difference `r total_rna_stats`, 47S, `r stat_rrna$trg_47s`; 18S, `r stat_rrna$trg_18s`; 28S, `r stat_rrna$trg_28s`; 5.8S, `r stat_rrna$trg_5.8s`; 5S, `r stat_rrna$trg_5s`), or levels of cellular myelocutmatosis oncogene, upstream binding factor and ribosomal protein S6 (*p* = `r min(p_west)`-`r max(p_west)`).
 There was a linear relationship between upstream binding factor and total RNA, where an increase of 1 standard deviance unit of upstream binding factor equated to a $\sim$ `r corr_incr`% increase in total RNA (*p* = `r corrp`).
 
-*Conclusion:* Glucose supplementation before and after RT does not augment RT-induced ribosome biogenesis after two weeks of heavy RT in moderately trained young adults.
-
-</div>
-
-*Keywords:* Glucose; hypertrophy; resistance training; ribosome; skeletal muscle
-
+**Conclusion:** Glucose supplementation before and after RT does not augment RT-induced ribosome biogenesis after two weeks of heavy RT in moderately trained young adults.
+
+**Keywords:** Glucose; hypertrophy; resistance training; ribosome; skeletal muscle
+
+**Abbreviations**
 ```{r abbreviation table}
@@
   flextable() %>%
-  set_caption(caption = "Abbreviations",
-              align_with_table = FALSE) %>%
   set_table_properties(layout = "autofit")
@@
 
-</div>
-
-# Statements and Declarations
+### Statements and Declarations
 This study was financed through research funds from the Inland Norway University of Applied Sciences.
@@
 
-## AUTHOR CONTRIBUTIONS
+### Author contributions
 
@@
 
-## Acknowledgements
+### Acknowledgements
 Gratitude are in order to Henriette Spilhaug Bollandsås, Nora Lerdalen Bonsak and Thea Bøhn Nyløkken for their dedication and hard work in the training intervention.
@@
 
+</div>
+
 \pagebreak
-# INTRODUCTION
+## Introduction
 
@@
 
-# MATERIALS AND METHODS
+## Materials and methods
 
@@
 
-## PARTICIPANTS
+### Participants
 
@@
 
-## EXPERIMENTAL DESIGN
+### Experimental design
 
@@
 
-## DIETARY INTERVENTION
+### Dietary intervention
 
@@
 
-## ASSESSMENT OF MUSCULAR STRENGTH
+### Assessment of muscular strength
 
@@
 
-## RESISTANCE TRAINING PROTOCOL
+### Resistance training protocol
 
@@
 
-## SAMPLING OF MUSCLE TISSUE AND BLOOD
+### Sampling of muscle tissue and blood
 
@@
 
-## TOTAL RNA EXTRACTION AND REAL-TIME REVERSE TRANSCRIPTION POLYMERASE CHAIN REACTION
+### Total RNA extraction and real-time reverse transcription polymerase chain reaction
 ```{r totrna-extraction}
@@
 
-## PROTEIN EXTRACTION AND IMMUNOBLOTTING
+### Protein extraction and immunoblotting
 
@@
 
-## STATISTICS AND DATA ANALYSIS
+### Statistics and data analysis
 
@@
 
-# RESULTS
+## Results
 
@@
 
-## Markers of ribosome biogenesis
-
-### Total RNA and ribosomal RNA
+### Markers of ribosome biogenesis
+
+*Total RNA and ribosomal RNA*
 
@@
 
-### Protein
+*Protein*
 
@@
 
-# DISCUSSION
+## Discussion
 
@@
 
-### Conclusion
+*Conclusion*
 
@@
 
-## DATA AVAILABILITY
+### DATA AVAILABILITY
 
@@
 
-## SUPPLEMENTAL MATERIAL
+### SUPPLEMENt MATERIAL
 
@@
 
-# FIGURE LEGENDS
+# Figure/table legends
 
@@
 
-# TABLES
-
+# Tables
+
+**Table 1.
+Participant characteristics:** Values are means ± SD.
+60º sec-1 = 60 degrees per second movement speed (isokinetic), 240º sec-1 = 240 degrees per second movement speed (isokinetic), 0º sec-1 = 0 degrees per second movement speed (isometric).
 ```{r baseline table, echo = FALSE}
@@
                     as_paragraph("0\U00BA sec", as_sup("-1")))) %>%
-  add_header_row(values = c(" ", "Knee-extension peak torque"), colwidths = c(6, 3)) %>%
-  set_caption(caption = "Table 1. Participant characteristics",
-              align_with_table = FALSE)
+  add_header_row(values = c(" ", "Knee-extension peak torque"), colwidths = c(6, 3))
 
@@
 ```
-Values are means ± SD.
-60º sec-1 = 60 degrees per second movement speed (isokinetic), 240º sec-1 = 240 degrees per second movement speed (isokinetic), 0º sec-1 = 0 degrees per second movement speed (isometric).
-
+
+**Table 2.
+Primer sequences:** Values of Ct are means ± SD.
+rRNA = ribosomal RNA, E = primer efficiency.
+Average cycle thresholds (Ct) and priming efficiencies were calculated from all qPCR reactions.
 ```{r primer table, echo = FALSE}
@@
                     as_paragraph("E"))) %>%
-  set_caption(caption = "Table 2. Primer Sequences",
-              align_with_table = FALSE) %>%
   set_table_properties(layout = "autofit")
@@
 ```
-Values of CT are means ± SD. rRNA = ribosomal RNA, E = primer efficiency. Average cycle thresholds (Ct) and priming efficiencies were calculated from all qPCR reactions.

### 2023-09-28 figures/ejap_figures/ejap_figure1.R

@@ -1,0 +1,333 @@
+### Ribose figure 1 in ejap format######
+#
+# Study design, strength (1RM, Humac), c-peptide and blood glucose
+#
+#
+#
+# Packages 
+library(tidyverse); library(emmeans); library(cowplot); library(grid); library(gridExtra); library(readxl); 
+library(elementalist); library(ggtext)
+#
+## Data
+
+# Study design
+d.dat <- read_excel("./data/design.dat.xlsx", na = "NA")
+
+# C-peptide
+cpep.change <- readRDS("./data/data-gen/glucose/insulin.figchange.RDS") # Untransformed data
+
+# Blood glucose
+glu.change <- readRDS("./data/data-gen/glucose/gluc.change.RDS") # Untransformed data
+
+# Humac data
+str.change <- readRDS("./data/data-gen/humac/str.change.RDS")
+str.emm <- readRDS("./data/data-gen/humac/emm.str.RDS")
+
+# Training volume data
+vol.lemm <- readRDS("./data/data-gen/training/vol.lemm.RDS")
+
+
+### Plots
+pos <- position_dodge(width = .2)
+
+# Designing the plot theme
+
+plot_theme <- theme(panel.grid.major = element_blank(),
+                    panel.grid.minor = element_blank(),
+                    panel.background = element_rect(fill = "white", colour = NA),
+                    axis.line = element_line(colour = "black"))
+
+# Colours
+# Colours
+
+colors <- c("black","white")
+                     
+
+
+## C-peptide
+
+cpep.fig <- cpep.change %>%
+  data.frame() %>%
+  add_row(supplement = "PLACEBO", time = "change.1", emmean = 0, SE = 0, df = 0, lower.CL = 0, upper.CL = 0, .before =1) %>%
+  add_row(supplement = "GLUCOSE", time = "change.1", emmean = 0, SE = 0, df = 0, lower.CL = 0, upper.CL = 0, .before = 2) %>%
+  mutate(time.c = gsub("change.", "", time), 
+         time.c = as.numeric(time.c)) %>%
+  ggplot(aes(time.c, emmean, group = supplement, fill = supplement)) +
+  annotate("text", x = c(120, 150), y = c(975, 950), label = "*") +
+  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
+                size = 0.5,
+                position = pos) +
+  geom_line(position = pos, lty = 2) +
[…]

### 2023-09-28 figures/ejap_figures/ejap_figure2.R

@@ -1,0 +1,145 @@
+### Ribose figure 2 in ejap format ######
+#
+# This script creates figure 2, containing log-fold change data of total RNA and rRNA
+#
+# Packages 
+library(tidyverse); library(emmeans); library(cowplot); library(ggtext)
+
+#
+### Data
+#
+## Total RNA
+emm.tot <- readRDS("./data/data-gen/rna/emm.tot.RDS")
+
+## rRNA
+# 18S
+emm.18 <- readRDS("./data/data-gen/rna/emm.18.RDS") %>%
+  mutate(target = "18S rRNA",
+         time = "post") %>%
+  add_row(supplement = c("PLACEBO", "GLUCOSE"), time = "pre", emmean = 0, lower.CL = 0, upper.CL = 0,
+          target = "18S rRNA") %>%
+  print()
+
+# 28S
+emm.28 <- readRDS("./data/data-gen/rna/emm.28.RDS")  %>%
+  mutate(target = "28S rRNA",
+         time = "post") %>%
+  add_row(supplement = c("PLACEBO", "GLUCOSE"), time = "pre", emmean = 0, lower.CL = 0, upper.CL = 0,
+          target = "28S rRNA") %>%
+  print()
+
+# 5.8S
+emm.5.8 <- readRDS("./data/data-gen/rna/emm.5.8.RDS") %>%
+  mutate(target = "5.8S rRNA",
+         time = "post") %>%
+  add_row(supplement = c("PLACEBO", "GLUCOSE"), time = "pre", emmean = 0, lower.CL = 0, upper.CL = 0,
+          target = "5.8S rRNA") %>%
+  print()
+
+# 5S
+emm.5 <- readRDS("./data/data-gen/rna/emm.5.RDS")  %>%
+  mutate(target = "5S rRNA",
+         time = "post") %>%
+  add_row(supplement = c("PLACEBO", "GLUCOSE"), time = "pre", emmean = 0, lower.CL = 0, upper.CL = 0,
+          target = "5S rRNA") %>%
+  print()
+
+# 47S
+emm.47 <- readRDS("./data/data-gen/rna/emm.47.RDS") %>%
+  mutate(target = "47S rRNA",
+         time = "post") %>%
+  add_row(supplement = c("PLACEBO", "GLUCOSE"), time = "pre", emmean = 0, lower.CL = 0, upper.CL = 0,
+          target = "47S rRNA") %>%
+  print() 
+
+# Designing the plot theme
+
[…]

### 2023-09-29 README.rmd
fixe on ejap figs

@@
 
-# Repository organization and extended methods: *Glucose ingestion before and after resistance training does not augment ribosome biogenesis in healthy moderately trained young adults*
+# Extended methods: *Glucose ingestion before and after resistance training does not augment ribosome biogenesis in healthy moderately trained young adults*
 

### 2023-09-29 figures/ejap_figures/ejap_figure1.R

@@ -6,7 +6,7 @@
 #
 # Packages 
 library(tidyverse); library(emmeans); library(cowplot); library(grid); library(gridExtra); library(readxl); 
-library(elementalist); library(ggtext)
+library(ggtext)
 #
 ## Data
 
@@ -28,14 +28,14 @@ vol.lemm <- readRDS("./data/data-gen/training/vol.lemm.RDS")
 
 
 ### Plots
-pos <- position_dodge(width = .2)
+pos <- position_dodge(width = 10)
 
 # Designing the plot theme
 
 plot_theme <- theme(panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(),
                     panel.background = element_rect(fill = "white", colour = NA),
-                    axis.line = element_line(colour = "black"))
+                    axis.line = element_line(colour = "black", linewidth = 0.4))
 
 # Colours
 # Colours
@@ -56,9 +56,10 @@ cpep.fig <- cpep.change %>%
   annotate("text", x = c(120, 150), y = c(975, 950), label = "*") +
   geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
                 size = 0.5,
-                position = pos) +
-  geom_line(position = pos, lty = 2) +
-  geom_point(shape = 21, position = pos, size = 2) +
+                width = 0,
+                position = position_dodge(width = 10)) +
+  geom_line(position = position_dodge(width = 10), lty = 2) +
+  geom_point(shape = 21, position = position_dodge(width = 10), size = 2) +
   scale_fill_manual(values = colors[c(1,2)]) +
   scale_x_continuous(limits = c(0, 300), breaks = c(0, 90, 120, 150, 270),
                      expand = expansion(0), labels = c("change.1" = "-120 min", "change.90" = "-30 min", "change.120" = "0 min", 
@@ -87,6 +88,7 @@ glu.fig <- glu.change %>%
   annotate("text", x = c(270), y = c(0.23), label = "*") +
   geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
                 size = 0.5,
+                width = 0,
                 position = pos) +
   geom_line(position = pos, lty = 2) +
   geom_point(shape = 21, position = pos, size = 2) +
@@ -134,14 +136,19 @@ str.fig2 <- str.emm2 %>%
   data.frame() %>%
   ggplot(aes(timeh, emmean, fill = supplement)) +
[…]

### 2023-09-29 figures/ejap_figures/ejap_figure2.R

@@ -57,7 +57,7 @@ emm.47 <- readRDS("./data/data-gen/rna/emm.47.RDS") %>%
 plot_theme <- theme(panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(),
                     panel.background = element_rect(fill = "white", colour = NA),
-                    axis.line = element_line(colour = "black"))
+                    axis.line = element_line(colour = "black", linewidth = 0.4))
 
 # Colours
 
@@ -77,7 +77,7 @@ totrna.plot <- emm.tot %>%
   mutate(time = factor(time, levels = c("pre", "post"), labels = c("Baseline", "Post"))) %>%
   ggplot(aes(time, exp(emmean), group = supplement, fill = supplement)) +
   geom_errorbar(aes(ymin = exp(lower.CL), ymax = exp(upper.CL)),
-                width = 0.1,
+                width = 0,
                 position = position_dodge(width = 0.2)) +
   geom_line(position = position_dodge(width = 0.2), lty = 2) +
   geom_point(position = position_dodge(width = 0.2), shape = 21, size = 2) +
@@ -87,8 +87,8 @@ totrna.plot <- emm.tot %>%
   theme_classic() +
   theme(axis.title.x = element_blank(), 
         axis.title.y = element_markdown(size = htextsize),
-        axis.text.x = element_text(size = textsize),
-        axis.text.y = element_text(size = textsize),
+        axis.text.x = element_text(size = textsize, color = "black"),
+        axis.text.y = element_text(size = textsize, color = "black"),
         legend.title = element_text(size = htextsize),
         legend.text = element_text(size = textsize),
         legend.key.size = unit(0.3, "cm")) +
@@ -112,7 +112,7 @@ rrna.fig <- bind_rows(emm.47, emm.18, emm.28, emm.5.8, emm.5) %>%
   labs(x = "Time", y = "AU fold change", fill = "Supplement") +
   facet_wrap(~target, nrow = 1) +
   theme(axis.title = element_text(size = htextsize),
-        axis.text = element_text(size = textsize),
+        axis.text = element_text(size = textsize, color = "black"),
         legend.title = element_text(size = textsize),
         legend.text = element_text(size = textsize),
         strip.text = element_text(hjust = 0, size = htextsize),
@@ -129,7 +129,7 @@ fig2 <- plot_grid(plot_grid(NULL,
[…]

### 2023-09-29 figures/ejap_figures/ejap_figure3.R

@@ -1,0 +1,508 @@
+### Ribose figure 3 ######
+#
+# This script produces the figures for cmyc, ubf and rps6
+#
+#
+# Packages
+library(emmeans); library(tidyverse); library(cowplot); library(magick); 
+
+library(nlme); library(ggtext); library(lme4)
+#
+# Data
+
+
+prot_models <- readRDS("./data/data-gen/protein/protein_lmer_models.RDS")
+
+
+prot_models <- prot_models$prot_models
+
+
+cmyc <- readRDS("./data/data-gen/protein/cmyc.lemm.RDS")
+
+ubf <- readRDS("./data/data-gen/protein/ubf.lemm.RDS")
+
+rps6 <- readRDS("./data/data-gen/protein/rps6.lemm.RDS")
+
+joined.dat <- readRDS("./data/data-gen/rna.prot.RDS")
+
+# Images
+
+g3 <- cowplot::ggdraw() + cowplot::draw_image("./figures/archive/g3_115_annotated.png", scale = 0.9)
+g6 <- cowplot::ggdraw() + cowplot::draw_image("./figures/archive/g6_115_unannotated.png", scale = 0.9)
+
+ubf.x2.img <- cowplot::ggdraw() + cowplot::draw_image("./figures/archive/ubf_x2_115.png", scale = 0.8)
+cmyc.x2.img <- cowplot::ggdraw() + cowplot::draw_image("./figures/archive/cmyc_x2_115.png", scale = 0.8)
+rps6.x2.img <- cowplot::ggdraw() + cowplot::draw_image("./figures/archive/rps6_x2_115.png", scale = 0.8)
+
+# Colours
+
+colors <- c("#d7191c",
+            "gray80",
+            "gray40",
+            "#2c7bb6")
+            
+# Textsizes
+
+labsize <- 8
+textsize <- 6
+htextsize <- 7
+legendti <- 6
+legendtex <- 5.25
+atext <- 2
+keysize <- 2
+
+# Designing the plot theme
+
+plot_theme <- theme(panel.grid.major = element_blank(),
+                    panel.grid.minor = element_blank(),
+                    panel.background = element_rect(fill = "white", colour = NA),
+                    #plot.background = element_rect(fill = "lightblue", color = NA),
+                    axis.line = element_line(colour = "black", linewidth = 1))
+
+## cmyc fig
+
+comp_plot <- function(model, method = "Wald") {
+  
[…]

### 2023-09-29 README.rmd
proofreading

@@
 <img alt="ORCID logo" src="https://info.orcid.org/wp-content/uploads/2019/11/orcid_16x16.png" width="16" height="16" /></a>,
-Sara Christine Moen,
+Daniel Hammarström
+<a href="https://orcid.org/0000-0001-8360-2100">
+<img alt="ORCID logo" src="https://info.orcid.org/wp-content/uploads/2019/11/orcid_16x16.png" width="16" height="16" /></a>,
 Håvard Hamarsland
@@
 <img alt="ORCID logo" src="https://info.orcid.org/wp-content/uploads/2019/11/orcid_16x16.png" width="16" height="16" /></a>,
-Daniel Hammarström
-<a href="https://orcid.org/0000-0001-8360-2100">
-<img alt="ORCID logo" src="https://info.orcid.org/wp-content/uploads/2019/11/orcid_16x16.png" width="16" height="16" /></a> 
+Sara Christine Moen,
 Stian Ellefsen

### 2023-09-29 ejap.manuscript.RMD
proofreading

@@
 
-ORCID:
-<div custom-style="correspondence_style">
-Kristian Lian: https://orcid.org/0000-0002-5169-5960
-
-Håvard Hamarsland: https://orcid.org/0000-0002-7343-2774
-
-Daniel Hammarström https://orcid.org/0000-0001-8360-2100
-
-Stian Ellefsen https://orcid.org/0000-0003-0467-2018
+ORCID: Kristian Lian https://orcid.org/0000-0002-5169-5960, Daniel Hammarström https://orcid.org/0000-0001-8360-2100, Håvard Hamarsland https://orcid.org/0000-0002-7343-2774, Stian Ellefsen https://orcid.org/0000-0003-0467-2018
 </div>
 
-\pagebreak
 <div custom-style="abstract_style">
@@
 
-**Results:** Between conditions, there were no significant differences in increases of total RNA and ribosomal RNA (mean difference `r total_rna_stats`, 47S, `r stat_rrna$trg_47s`; 18S, `r stat_rrna$trg_18s`; 28S, `r stat_rrna$trg_28s`; 5.8S, `r stat_rrna$trg_5.8s`; 5S, `r stat_rrna$trg_5s`), or levels of cellular myelocutmatosis oncogene, upstream binding factor and ribosomal protein S6 (*p* = `r min(p_west)`-`r max(p_west)`).
+**Results:** Between conditions, there were no significant differences in increases of total RNA and ribosomal RNA (mean difference `r total_rna_stats`, 47S, `r stat_rrna$trg_47s`; 18S, `r stat_rrna$trg_18s`; 28S, `r stat_rrna$trg_28s`; 5.8S, `r stat_rrna$trg_5.8s`; 5S, `r stat_rrna$trg_5s`), or levels of cellular myelocytomatosis oncogene, upstream binding factor and ribosomal protein S6 (*p* = `r min(p_west)`-`r max(p_west)`).
 There was a linear relationship between upstream binding factor and total RNA, where an increase of 1 standard deviance unit of upstream binding factor equated to a $\sim$ `r corr_incr`% increase in total RNA (*p* = `r corrp`).
@@
 
-**Keywords:** Glucose; hypertrophy; resistance training; ribosome; skeletal muscle
+**Keywords:** Glucose; hypertrophy; resistance training; ribosome; skeletal muscle.
+
+### Statements and Declarations
+This study was financed through research funds from the Inland Norway University of Applied Sciences.
+There was no economic vested interest and everyone participating as researchers and associates received only a regular salary during this study.
+
+### Author contributions
+
+Data was collected at the sports science lab at the Inland Norway University of Applied Sciences, with all analyses except for serum blood samples were performed at the sports science lab.
+Serum blood samples were performed at Innlandet Hospital Trust.
+S.E., H.H.
+and D.H.
+conceived and designed the study; K.L., S.C.M., D.H.
+and H.H.
+performed experiments; K.L.
+and D.H.
+interpreted the results; K.L.
+and D.H.
+prepared figures; K.L.
+drafted the manuscript; D.H., H.H.
+and S.E.
+edited and revised the manuscript; All authors approved the final version of the manuscript.
+
+### Acknowledgements
+Gratitude is due, to Henriette Spilhaug Bollandsås, Nora Lerdalen Bonsak and Thea Bøhn Nyløkken for their dedication and hard work in the training intervention.
+The authors would also like to express their gratitude to all the participants for staying committed to this demanding protocol in an impressive way.
+We also thank Håvard Nygaard for lending your expertise and technical skills to improve this project.
 
@@
 
-### Statements and Declarations
-This study was financed through research funds from the Inland Norway University of Applied Sciences.
-There were no economical vested interest and everyone participating as researchers and associates, received only a regular salary during this study.
-
-### Author contributions
-
-Data was collected at the sports science lab at the Inland Norway University of Applied Sciences, with all analyses except for serum blood samples were performed at the university.
-Serum blood samples were performed at Innlandet Hospital Trust.
-S.E., H.H.
-and D.H.
-conceived and designed the study; K.L., S.C.M., D.H.
-and H.H.
-performed experiements; K.L.
-and D.H.
-interpreted the results; K.L.
-and D.H.
-prepared figures; K.L.
-drafted the manuscript; D.H., H.H.
-and S.E.
-edited and revised the manuscript; D.H., H.H.
-and S.E.
-approved the final version of the manuscript.
-
-### Acknowledgements
-Gratitude are in order to Henriette Spilhaug Bollandsås, Nora Lerdalen Bonsak and Thea Bøhn Nyløkken for their dedication and hard work in the training intervention.
-The authors would also like to express their gratitude to all the participants, staying commited to this demanding protocol in an impressive way.
-We also thank Håvard Nygaard for lending your expertise and technical skills to improve this project.
-
 </div>
@@
 Indeed, the ability to synthesise ribosomes following a period of RT is connected to the magnitude of RT-induced responses in an individual [@hammarstrom_benefits_2020].
-For instance, increasing training volume generally induce greater ribosome biogenesis and thus, greater benefits of RT [@hammarstrom_benefits_2020;@krieger_single_2009;@schoenfeld_strength_2017].
+For instance, increasing training volume generally induces greater ribosome biogenesis and thus, greater benefits of RT [@hammarstrom_benefits_2020;@krieger_single_2009;@schoenfeld_strength_2017].
 However, as evident from Hammarström et al.
@@
 Nutritional supplements such as protein and creatine effectively optimise RT adaptations [@cermak_protein_2012;@lanhers_creatine_2015;@lanhers_creatine_2017;@morton_systematic_2018].
-However, it remains unknown if other nutritional adjuvants such as glucose can increase the efficacy of RT.
+However, it remains equivocal if other nutritional adjuvants such as glucose can increase the efficacy of RT [@tezze_anabolic_2023].
 This is surprising since glucose is the preferred energy substrate of the contracting skeletal muscle during strenuous exercise and a major energy supplier to cells via adenosine triphosphate (ATP) synthesis [@mul_exercise_2015;@sebata_control_2018].
@@
 mTOR receives input from growth factors, hormones, mechanical loading, and nutrients to balance protein synthesis through multiple mechanisms based on cellular energy levels [@hoppe_amp-activated_2009].
-Importantly, mTOR and its downstream target ribosomal protein S6 kinase beta-1 (S6K1) serve as a mediator for insulin signalling in skeletal muscle [@hillier_physiological_2000].
+Importantly, mTOR and its downstream target ribosomal protein S6 kinase beta-1 (S6K1) serves as a mediator for insulin signalling in skeletal muscle [@hillier_physiological_2000].
 The mTORC1 pathway contributes to ribosome biogenesis through the regulation of the translation of ribosomal proteins and to forming of the preinitiation complex (PIC) that marks the initiation of rRNA transcription [@figueiredo_regulation_2019;@von_walden_ribosome_2019].
 Parallel to mTORC1 [@mori_c-myc_2021; @west_acute_2016], c-Myc increases ribosomal biogenesis directly and through transcriptional control of the upstream binding factor (UBF) [@sanij_ubf_2008; @poortinga_c-myc_2011].
-Additionally, UBF phosphorylation, required for interaction with the rDNA promoter, is increased by high glucose in a mTORC1 dependent manner (rapamycin sensitive) in kidney glomerular epithelial cells[@mariappan_ribosomal_2011].
-Independently from UBF and mTORC1, high glucose was shown to lead to chromatin remodelling, which promoted rRNA transcription in cell cultures[@zhai_human_2012].
+Additionally, UBF phosphorylation, required for interaction with the ribosomal deoxyribonucleic acid (rDNA) promoter, is increased by high glucose in a mTORC1 dependent manner (rapamycin sensitive) in kidney glomerular epithelial cells [@mariappan_ribosomal_2011].
+Independently from UBF and mTORC1, high glucose was shown to lead to chromatin remodelling, which promoted rRNA transcription in cell cultures [@zhai_human_2012].
 Together these observations indicate a potential role of glucose in positively affecting muscle rRNA synthesis.
@@
 Secondly, we aimed to describe the association between UBF and total RNA abundance in human skeletal muscle.
-There have been multiple studies recently suggesting translational capacity to be as important, if not more important than, translational efficiency with regard to long-term skeletal muscle adaptations to RT [@hammarstrom_benefits_2020;@hammarstrom_ribosome_2022;@figueiredo_revisiting_2019].
+There have been multiple studies recently suggesting translational capacity to be as important, if not more important than, translational efficiency concerning long-term skeletal muscle adaptations to RT [@hammarstrom_benefits_2020;@hammarstrom_ribosome_2022;@figueiredo_revisiting_2019].
 Previously, *in vitro* studies have shown UBF and c-Myc to be important factors in the PIC, c-Myc as a general transcription factor and UBF as a specific transcription factor for rDNA transcription initiation [@mariappan_ribosomal_2011;@von_walden_ribosome_2019].
@@
 
-All participants gave their written informed consent prior to study enrollment.
+All participants gave their written informed consent before study enrollment.
 The study was approved by the regional ethical committee (REK, ID nr.
-153628), pre-registered at clinicaltrials.gov (Identifier: NCT04545190), and conducted according to the Helsinki declaration.
+153628), pre-registered at clinicaltrials.gov (Identifier: NCT04545190), and conducted according to the Helsinki Declaration.
 
@@
 
-The dietary intervention spanned the whole day, divided into three periods: I) From awakening until 2.5 hours (hrs) after RT, II) from 2.5hrs after RT until 2200hrs, and III) from 2200hrs until awakening.
-During period I, participants ingested protein and glucose **or** protein and placebo only.
-Glucose/placebo was ingested three times in period I: 30 minutes before RT (0830hrs, 30g vs 0g glucose), immediately before RT (0900hrs, 30g vs.
+The dietary intervention spanned the whole day, divided into three periods: I) From awakening until 2.5 hours (hrs) after RT, II) from 2.5 hours after RT until 2200hrs, and III) from 2200hrs until awakening.
+During period I, participants ingested protein and glucose *or* protein and placebo only.
+Glucose/placebo was ingested three times in period I: 30 minutes before RT (0830hrs, 30g vs.
+0g glucose), immediately before RT (0900hrs, 30g vs.
 0g glucose), and immediately after RT (~0930hrs, 30g vs.
@@
 During sessions, participants were free to ingest water ad libitum.
+
+CONTINUE PROOFREADING FROM HERE!!!!!!!!!!!!!!!!!!!!
 

### 2023-09-29 README.rmd
cleanup

@@
 ## Abstract
+
+**Purpose:** Muscle hypertrophy, induced by resistance training, is determined by ribosomal biogenesis.
+High-glucose treatment may augment the transcription of ribosomal RNA, as evident from in vitro studies.
+However, its effect on resistance training-induced adaptations remains unexplored.
+
+**Methods:** Sixteen healthy, moderately trained individuals (male/female, n = 9/7; age, 23.7 (1.8)/24.6 (4.8)), participated in a within-participant trial with unilateral resistance training with glucose- and placebo supplementation on alternate days over six sessions per condition.
+Participants remained fasted overnight, receiving glucose/placebo and protein supplementation only before and after RT sessions (3 sets of 10 repetitions maximum unilateral leg press and knee extension).
+Micro biopsies were sampled from m.
+vastus lateralis before and after the intervention.
+
+**Results:** Between conditions, there were no significant differences in increases of total RNA and ribosomal RNA (mean difference 7.6%, [-7.2, 24.9], p = 0.337, 47S, 37.9%, [-28.4, 131.6], p = 0.400; 18S, -7.6%, [-34.0, 29.8], p = 0.652; 28S, -2.5%, [-37.7, 53.2], p = 0.915; 5.8S, -7.7%, [9.8, 98.0], p = 0.644; 5S, -0.4%, [-31.1, 44.2], p = 0.982), or levels of cellular myelocytomatosis oncogene, upstream binding factor and ribosomal protein S6 (p = 0.094-0.292).
+There was a linear relationship between upstream binding factor and total RNA, where an increase of 1 standard deviance unit of upstream binding factor equated to a ∼ 14% increase in total RNA (p = 0.003).
+
+**Conclusion:** Glucose supplementation before and after RT does not augment RT-induced ribosome biogenesis after two weeks of heavy RT in moderately trained young adults.
 

### 2023-09-29 manuscript.RMD
cleanup

@@
 ---
-title: "Glucose ingestion before and after resistance training does not augment ribosome biogenesis in healthy moderately trained young adults"
-author: "Kristian Lian^1^, Sara Christine Olsen Moen^1^, Håvard Hamarsland^1^, Daniel Hammarström^1^, Stian Ellefsen^1^"
+title: ""
+author: ""
 bibliography: resources/ribose-paper.bib
-csl: resources/jap.csl
+csl: resources/ejap.csl
 output:
   word_document:
-    reference_docx: resources/jap-template.docx
+    reference_docx: resources/ejap-template.docx
 editor_options: 
@@
 
-### Affiliations
+# Glucose ingestion before and after resistance training does not augment ribosome biogenesis in healthy moderately trained young adults
+
+###### Kristian Lian^1^, Daniel Hammarström^1^, Håvard Hamarsland^1^, Sara Christine Olsen Moen^1^, Stian Ellefsen^1^
+
 <div custom-style="correspondence_style">
-^1^ Section for Health and Exercise Physiology, Department of Public Health and Sport Sciences, Inland Norway University of Applied Sciences, Lillehammer, Norway.
+^1^Section for Health and Exercise Physiology, Department of Public Health and Sport Sciences, Inland Norway University of Applied Sciences, Lillehammer, Norway
 </div>
@@
 <div custom-style="correspondence_style">
-### Correspondance
-Kristian Lian, Section for Health and Exercise Physiology, Department of Public Health and Sport Sciences, Inland Norway University of Applied Sciences, Lillehammer, Norway.
-Email: kristian.lian@inn.no
-
-Stian Ellefsen, Section for Health and Exercise Physiology, Department of Public Health and Sport Sciences, Inland Norway University of Applied Sciences, Lillehammer, Norway.
-Email: stian.ellefsen@inn.no
+Correspondance: Kristian Lian (kristian.lian@inn.no), Stian Ellefsen (stian.ellefsen@inn.no)
 </div>
 
-### ORCID
-<div custom-style="correspondence_style">
-Kristian Lian: https://orcid.org/0000-0002-5169-5960
-
-Håvard Hamarsland: https://orcid.org/0000-0002-7343-2774
-
-Daniel Hammarström https://orcid.org/0000-0001-8360-2100
-
-Stian Ellefsen https://orcid.org/0000-0003-0467-2018
+ORCID: Kristian Lian https://orcid.org/0000-0002-5169-5960, Daniel Hammarström https://orcid.org/0000-0001-8360-2100, Håvard Hamarsland https://orcid.org/0000-0002-7343-2774, Stian Ellefsen https://orcid.org/0000-0003-0467-2018
 </div>
 
-\pagebreak
 <div custom-style="abstract_style">
-# ABSTRACT
+## Abstract
 
@@
 
-**Introduction.** Muscle hypertrophy, induced by resistance training (RT), is determined by ribosomal biogenesis.
-Synthesis of ribosomes may augment the transcription of ribosomal RNA, as evident from in vitro studies.
-However, its effect on RT-induced adaptations remains unexplored.
-
-**Methods.** Sixteen healthy, moderately trained individuals (male/female, n = 9/7; age, 23.7 (1.8)/24.6 (4.8)),<!-- Add male/female numbers, age.
-Kristian: Done--> participated in a within-participant trial with unilateral resistance training performed with glucose- and placebo supplement on alternate days over six sessions per condition.
+**Purpose:** Muscle hypertrophy, induced by resistance training, is determined by ribosomal biogenesis.
+High-glucose treatment may augment the transcription of ribosomal RNA, as evident from in vitro studies.
+However, its effect on resistance training-induced adaptations remains unexplored.
+
+**Methods:** Sixteen healthy, moderately trained individuals (male/female, n = 9/7; age, 23.7 (1.8)/24.6 (4.8)), participated in a within-participant trial with unilateral resistance training with glucose- and placebo supplementation on alternate days over six sessions per condition.
 Participants remained fasted overnight, receiving glucose/placebo and protein supplementation only before and after RT sessions (3 sets of 10 repetitions maximum unilateral leg press and knee extension).
 Micro biopsies were sampled from m.
-vastus lateralis pre and post the intervention, and maximal unilateral isometric and isokinetic knee extension torque were measured before, during (days 4, 5, 8, 9) and after the intervention.
-
-**Results.** <!-- update statistics in this section -->From baseline to post-intervention total RNA and rRNA per unit muscle weight increased on average by ~24% and ~33-43%, respectively (Figure 2), with no differences between supplements (Figure 2A; mean difference `r total_rna_stats`, Figure 2B; 47S, `r stat_rrna$trg_47s`; 18S, `r stat_rrna$trg_18s`; 28S, `r stat_rrna$trg_28s`; 5.8S, `r stat_rrna$trg_5.8s`; 5S, `r stat_rrna$trg_5s`).
-RT with glucose resulted in estimated levels of c-Myc, UBF and RPS6 being `r stat_west$cmyc`, `r stat_west$ubf` and `r stat_west$rps6`% lower compared to placebo, respectively, without showing statistical significance (*p* = `r min(p_west)`-`r max(p_west)`; Figure 3A).
-There was a linear relationship between UBF and total RNA, where an increase of 1 SD unit of UBF equated to a $\sim$ `r corr_incr`% increase in total RNA (*p* = `r corrp`; Figure 3C).
-Lastly, there were no differences between supplements in maximal unilateral isometric knee extension force and isokinetic peak torque during the intervention (p > 0.05) and following the last session (p > 0.05), apart from the 60 d/s test before last RT session (p < 0.05).
-
-**Conclusion.** Glucose supplement before and after RT does not augment RT-induced ribosome biogenesis, nor does it enhance muscular performance and recovery after two weeks of heavy RT in moderately trained, young individuals.
-
-## NEW AND NOTEWORTHY
-
-**Keywords:** Skeletal muscle; hypertrophy; resistance training; glucose; ribosome
+vastus lateralis before and after the intervention.
+
+**Results:** Between conditions, there were no significant differences in increases of total RNA and ribosomal RNA (mean difference `r total_rna_stats`, 47S, `r stat_rrna$trg_47s`; 18S, `r stat_rrna$trg_18s`; 28S, `r stat_rrna$trg_28s`; 5.8S, `r stat_rrna$trg_5.8s`; 5S, `r stat_rrna$trg_5s`), or levels of cellular myelocytomatosis oncogene, upstream binding factor and ribosomal protein S6 (*p* = `r min(p_west)`-`r max(p_west)`).
+There was a linear relationship between upstream binding factor and total RNA, where an increase of 1 standard deviance unit of upstream binding factor equated to a $\sim$ `r corr_incr`% increase in total RNA (*p* = `r corrp`).
+
+**Conclusion:** Glucose supplementation before and after RT does not augment RT-induced ribosome biogenesis after two weeks of heavy RT in moderately trained young adults.
+
+**Keywords:** Glucose; hypertrophy; resistance training; ribosome; skeletal muscle.
+
+### Statements and Declarations
+This study was financed through research funds from the Inland Norway University of Applied Sciences.
+There was no economic vested interest and everyone participating as researchers and associates received only a regular salary during this study.
+
+### Author contributions
+
+Data was collected at the sports science lab at the Inland Norway University of Applied Sciences, with all analyses except for serum blood samples performed at the sports science lab.
+Serum blood samples were performed at Innlandet Hospital Trust.
+S.E., H.H.
+and D.H.
+conceived and designed the study; K.L., S.C.M., D.H.
+and H.H.
+performed experiments; K.L.
+and D.H.
+interpreted the results; K.L.
+and D.H.
+prepared figures; K.L.
+drafted the manuscript; D.H., H.H.
+and S.E.
+edited and revised the manuscript; All authors approved the final version of the manuscript.
+
+### Acknowledgements
+Gratitude is due, to Henriette Spilhaug Bollandsås, Nora Lerdalen Bonsak and Thea Bøhn Nyløkken for their dedication and hard work in the training intervention.
+The authors would also like to express their gratitude to all the participants for staying committed to this demanding protocol in an impressive way.
+We also thank Håvard Nygaard for lending your expertise and technical skills to improve this project.
+
+**Abbreviations**
+```{r abbreviation table}
+### Abbreviations table
+
+## Packages
+library(readxl); library(tidyverse); library(flextable)
+
+## Data
+abbreviations.dat <- read_excel("./data/abbreviations.xlsx")
+
+tab.dat <- abbreviations.dat
+
+## Flextable
+abb.tab <- tab.dat %>%
+  flextable() %>%
+  set_table_properties(layout = "autofit")
+
+abb.tab <- delete_part(x = abb.tab, part = "header")
+
+abb.tab <- border(abb.tab, 
+                     i = NULL,
+                     j = NULL,
+                     border = NULL,
+                     border.top = fp_border_default(color = "black"),
+                     border.bottom = fp_border_default(color = "black"),
+                     border.left = fp_border_default(color = "black"),
+                     border.right = fp_border_default(color = "black"),
+                     part = "all")
+
+abb.tab
+```
 
@@
 \pagebreak
-# INTRODUCTION
-
-Responses to systematic resistance training (RT) vary widely between individuals, with as many as 10-15% showing impaired muscle growth in response to standardized training interventions [@alvarez_interindividual_2018; @mann_high_2014; @thalacker-mercer_cluster_2013].
-Such inability to benefit from RT may be explained by genetic and/or epigenetic disposition affecting the internal physiological milieu and subsequently affecting the ability to mount an anabolic response to RT [@figueiredo_genetic_2021;@thalacker-mercer_cluster_2013].<!-- Check these refs, do they correspond to genet epigen and internal mileu?
-Kristian: The way I read it, it looks like it corresponds --> Although RT program modifications such as altered exercise volume or intensity may induce more favourable adaptations in some individuals, these benefits are not uniform between individuals [@hammarstrom_benefits_2020;@krieger_single_2009;@schoenfeld_strength_2017].<!-- Add ref for more rt program modifications.
-Kristian: Added krieger et al.
-2009 and schoenfeld et al.
-2017 --> For instance, although changes in training volume generally induce greater benefits of RT, not all individuals benefit from such modifications [@hammarstrom_benefits_2020].
-Therefore, other means than modification of RT variables seem necessary to optimize individual responses.<!-- The above sentence might be a bit contradictory?
-Kristian: Im not sure, are you thinking about "other means than modification of RT" being contradictory of "modifications such as altered exercise volume.." sentence?
-I think there is a point to establishing that yes, increasing exercise volume is a potent way to increase the RT stimulus in general.
-However, in cases where this alone fails to elicit true responses, other means are necessary.
-Or am I missinterpreting this?
--->
-
-Nutritional supplements such as protein and creatine effectively optimize RT adaptations [@cermak_protein_2012;@lanhers_creatine_2015;@lanhers_creatine_2017;@morton_systematic_2018].
-However, it remains unknown if other nutritional adjuvants such as glucose can increase the efficacy of RT.
-This is surprising since glucose is the preferred energy substrate of the contracting skeletal muscle during strenuous exercise and a major supplier of energy to cells via ATP synthesis [@mul_exercise_2015;@sebata_control_2018].
-Furthermore, energy availability is a decisive factor in the *de novo* synthesis of ribosomes [@kusnadi_regulation_2015;@moss_housekeeper_2007;@sebata_control_2018] <!--Add ref that indicate high energy requirement for ribo biog possibly (VONWALDEN19;TANAKA18?
-MOSS.
-Kristian: These three should do it-->which in turn determines muscle growth by increasing the muscle's translational capacity [@figueiredo_regulation_2019;@hammarstrom_benefits_2020;@sebata_control_2018;@stec_ribosome_2016;@von_walden_ribosome_2019].
+## Introduction
+
+Responses to systematic resistance training (RT) vary widely between individuals, with as much as 10-15% showing impaired skeletal muscle growth in response to standardized training interventions [@alvarez_interindividual_2018; @mann_high_2014; @thalacker-mercer_cluster_2013].
+While there certainly may be differences in the genetic predisposition in terms of readiness to adapt to exercise [@thalacker-mercer_cluster_2013], there are also recent studies showing that epigenetic factors such as ribosomal DNA (rDNA) availability may also impact the ability to mount an anabolic response to RT [@figueiredo_genetic_2021].
+Indeed, the ability to synthesise ribosomes following a period of RT is connected to the magnitude of RT-induced responses in an individual [@hammarstrom_benefits_2020].
+For instance, increasing training volume generally induces greater ribosome biogenesis and thus, greater benefits of RT [@hammarstrom_benefits_2020;@krieger_single_2009;@schoenfeld_strength_2017].
+However, as evident from Hammarström et al.
+(2020), not all participants experienced increased muscle mass and -strength despite increasing training volume.
+Therefore, other means than modification of RT variables alone seem necessary to optimise individual responses to RT.
+
+Nutritional supplements such as protein and creatine effectively optimise RT adaptations [@cermak_protein_2012;@lanhers_creatine_2015;@lanhers_creatine_2017;@morton_systematic_2018].
+However, it remains equivocal if other nutritional adjuvants such as glucose can increase the efficacy of RT [@tezze_anabolic_2023].
+This is surprising since glucose is the preferred energy substrate of the contracting skeletal muscle during strenuous exercise and a major energy supplier to cells via adenosine triphosphate (ATP) synthesis [@mul_exercise_2015;@sebata_control_2018].
+Furthermore, energy availability is a decisive factor in the *de novo* synthesis of ribosomes [@kusnadi_regulation_2015;@moss_housekeeper_2007;@sebata_control_2018] which in turn determines muscle growth by increasing the muscle's translational capacity [@figueiredo_regulation_2019;@hammarstrom_benefits_2020;@sebata_control_2018;@stec_ribosome_2016;@von_walden_ribosome_2019].
 Therefore, investigating the effects of combined RT and glucose ingestion may provide valuable insight into the potential additive effect of glucose and RT on ribosome biogenesis.
 
-Transcription of ribosomal RNA (rRNA) by Pol I is considered the rate limiting step in synthesizing new ribosomes [@moss_promotion_1995].
+Transcription of ribosomal RNA (rRNA) by RNA Polymerase I is considered the rate-limiting step in synthesising new ribosomes [@moss_promotion_1995].
 Multiple signalling pathways converge to regulate rRNA transcription, including the mammalian target of rapamycin (mTORC1) signal-transduction pathway and c-Myc [@kusnadi_regulation_2015;@von_walden_ribosome_2019;@sebata_control_2018; @west_acute_2016; @mori_c-myc_2021].
 mTOR receives input from growth factors, hormones, mechanical loading, and nutrients to balance protein synthesis through multiple mechanisms based on cellular energy levels [@hoppe_amp-activated_2009].
-Importantly, mTOR and its downstream target S6K1 serve as a mediator for insulin signalling in skeletal muscle [@hillier_physiological_2000].
+Importantly, mTOR and its downstream target ribosomal protein S6 kinase beta-1 (S6K1) serves as a mediator for insulin signalling in skeletal muscle [@hillier_physiological_2000].
 The mTORC1 pathway contributes to ribosome biogenesis through the regulation of the translation of ribosomal proteins and to forming of the preinitiation complex (PIC) that marks the initiation of rRNA transcription [@figueiredo_regulation_2019;@von_walden_ribosome_2019].
 Parallel to mTORC1 [@mori_c-myc_2021; @west_acute_2016], c-Myc increases ribosomal biogenesis directly and through transcriptional control of the upstream binding factor (UBF) [@sanij_ubf_2008; @poortinga_c-myc_2011].
-Additionally, UBF phosphorylation, required for interaction with the rDNA promoter, is increased by high glucose in a mTORC1 dependent manner (rapamycin sensitive) in kidney glomerular epithelial cells[@mariappan_ribosomal_2011].
-Independently from UBF and mTORC1, high glucose was shown to lead to chromatin remodelling, which promoted rRNA transcription in cell cultures[@zhai_human_2012].
+Additionally, UBF phosphorylation, required for interaction with the ribosomal deoxyribonucleic acid (rDNA) promoter, is increased by high glucose in a mTORC1 dependent manner (rapamycin sensitive) in kidney glomerular epithelial cells [@mariappan_ribosomal_2011].
+Independently from UBF and mTORC1, high glucose was shown to lead to chromatin remodelling, which promoted rRNA transcription in cell cultures [@zhai_human_2012].
 Together these observations indicate a potential role of glucose in positively affecting muscle rRNA synthesis.
 
-The purpose of this investigation was to test the hypothesis that glucose supplementation before and after<!-- during?
-Kristian: before and after would be more precise, since they're not strictly ingesting anything during RT - I got a comment on that during the thesis defense --> RT increases markers of ribosomal abundance following five RT sessions.
+The purpose of this investigation was to test the hypothesis that glucose supplementation before and after RT increases markers of ribosomal abundance following five RT sessions.
 Secondly, we aimed to describe the association between UBF and total RNA abundance in human skeletal muscle.
-There have been multiple studies in the last years suggesting translational capacity to be as important, if not more, as translational efficiency in regard to long-term skeletal muscle adaptations to RT [@hammarstrom_benefits_2020;@hammarstrom_ribosome_2022;@figueiredo_revisiting_2019].
+There have been multiple studies recently suggesting translational capacity to be as important, if not more important than, translational efficiency concerning long-term skeletal muscle adaptations to RT [@hammarstrom_benefits_2020;@hammarstrom_ribosome_2022;@figueiredo_revisiting_2019].
 Previously, *in vitro* studies have shown UBF and c-Myc to be important factors in the PIC, c-Myc as a general transcription factor and UBF as a specific transcription factor for rDNA transcription initiation [@mariappan_ribosomal_2011;@von_walden_ribosome_2019].
 Coupled with the recent observations by Hammarström et al.
-[@hammarstrom_benefits_2020;@hammarstrom_ribosome_2022], understanding underlying factors regulating and affecting ribosome biogenesis seems key to furthering our understanding of the optimization of RT to individual phenotypes.
-<!-- We should maybe add a line or two about the rational here. Kristian: Tried to add some, what about this? to general? -->
-
-# MATERIALS AND METHODS
-
-All participants gave their informed consent prior to study enrollment.
+[@hammarstrom_benefits_2020;@hammarstrom_ribosome_2022], understanding underlying factors regulating and affecting ribosome biogenesis seems key to furthering our understanding of the optimisation of RT to individual phenotypes.
+
+## Materials and methods
+
+All participants gave their written informed consent before study enrollment.
 The study was approved by the regional ethical committee (REK, ID nr.
-153628), pre-registered at clinicaltrials.gov (Identifier: NCT04545190), and conducted according to the Helsinki declaration.
-
-## PARTICIPANTS
+153628), pre-registered at clinicaltrials.gov (Identifier: NCT04545190), and conducted according to the Helsinki Declaration.
+
+### Participants
 
@@
 2-8 RT sessions per 14 days for the last six months).
-Exclusion criteria were previous injury leading to impaired strength, inability to perform resistance exercise training, symptoms, and a medical record of metabolic disorders including hyperglycemia, i.e.
-fasting venous plasma glucose ≥6.1 mmol/L and/or 2-hour glucose tolerance ≥7.8 mmol/L, and/or HbA1c >42 mmol/mol.
+Exclusion criteria were previous injury leading to impaired strength, inability to perform resistance exercise training, symptoms, and a medical record of metabolic disorders including hyperglycemia.
 Of the sixteen participants that commenced the intervention, three participants dropped out.
 One due to sickness and inability to resume, two participants experienced muscular discomfort related to heavy resistance training.
-Lean mass (Table 1) was measured using DXA (Prodigy Advance PA+302047, Lunar, San Francisco, CA, USA) on Day -1, the last day preceding the RT intervention.
-
-## EXPERIMENTAL DESIGN
+Lean mass (Table 1) was measured using Dual-Energy X-ray Absorptiometry (DXA, Prodigy Advance PA+302047, Lunar, San Francisco, CA, USA) on Day -1, the last day preceding the RT intervention.
+
+### Experimental design
 
@@
 Participants were asked to avoid  resistance- or high-intensity training of the legs from Day -7 (Figure 1A) and onwards, until completion of the intervention and post-testing, to ensure the reliability of pre-intervention strength data and minimal interference from external exercise sources.
-If participants did conduct exercise training outside of the prescribed protocol, they were asked to ensure equal loading on pairwise consecutive days.
-
-## DIETARY INTERVENTION
-
-The dietary intervention spanned the whole day, divided into three periods: I) From awakening until 2.5 hours after RT, II) from 2.5 hours after RT until 2200hrs, and III) from 2200hrs until awakening.
-During period I, participants ingested protein and glucose/placebo only.
-Glucose/placebo was ingested three times in period I: 30 minutes before RT (0830hrs, 30g vs 0g glucose), immediately before RT (0900hrs, 30g vs.
+
+### Dietary intervention
+
+The dietary intervention spanned the whole day, divided into three periods: I) From awakening until 2.5 hours (hrs) after RT, II) from 2.5 hours after RT until 2200hrs, and III) from 2200hrs until awakening.
+During period I, participants ingested protein and glucose *or* protein and placebo only.
+Glucose/placebo was ingested three times in period I: 30 minutes before RT (0830hrs, 30g vs.
+0g glucose), immediately before RT (0900hrs, 30g vs.
 0g glucose), and immediately after RT (~0930hrs, 30g vs.
@@
 During period III (2200-0700hrs), participants remained in an overnight fasted state.
-The daily onset of the dietary intervention (i.e., first ingestion of PRO supplement) varied between participants, from 0600hrs to 0900hrs to allow multiple participants to complete the protocol simultaneously.
+The daily onset of the dietary intervention (i.e., first ingestion of protein supplement) varied between participants, from 0600hrs to 0900hrs to allow multiple participants to complete the protocol simultaneously.
 During sessions, participants were free to ingest water ad libitum.
 
-## ASSESSMENT OF MUSCULAR STRENGTH
-
-Strength tests were performed prior to (Days -7 and -5, and -1, both legs) and during the intervention (Figure 1A, on days 4 and 8 for leg 1, and days 5 and 9 for leg two), after session 5 and after finalization of intervention (Figure 1A, on days 11/12 for leg 1 and days 12/13 for leg 2).
+### Assessment of muscular strength
+
+Strength tests were performed before (Figure 1A, Days -7 and -5, and -1, both legs) and during the intervention (Figure 1A, on days 4 and 8 for leg 1, and days 5 and 9 for leg two), after session 5 and after finalization of the intervention (Figure 1A, on days 11/12 for leg 1 and days 12/13 for leg 2).
 Maximal isometric and isokinetic knee extension torque was measured with a Humac Norm Dynamometer (CSMi, Stoughton, Massachusetts, USA).
 Individual positions were recorded and standardized from pre-intervention tests (Figure 1A, days -7 and -5).
-Isokinetic peak torque was measured at extensions speeds of 60- and 240 d/s, 2x3 repetitions each, with the first set of each exercise as a sub-maximal warm-up.
-Isometric peak torque was measured at a knee-angle of 60 degrees (extended knee was set to 0 degrees), for a maximum of 10 seconds and two repetitions per test.
+Isokinetic peak torque was measured concentrically from 90 to 0 degrees knee angle (extended knee was set to 0 degrees) at angular velocities of 60- and 240 degrees per second, 2x3 repetitions each, with the first set of each exercise as a sub-maximal warm-up.
+Isometric knee extensor peak torque was measured at a kneen angle of 60 degrees, for a maximum of 10 seconds and two repetitions per test.
 The isometric tests were ended when the participants reached a plateau or peak torque development decreased, which on average occurred between 2-4 seconds into the test.
@@
 Test I on day 12/13 included testing of both legs, representing 23hrs post-RT session test of one leg and post-session 5 test of the other leg.
-The highest peak torque values from the respective extension speeds and time points were summarized in an index.
-The index was calculated by dividing the average peak torque value by the highest observed peak torque value per extension speed, and summarizing this new index per extension speed to a mean strength index.
-
-Assessment of unilateral one repetition maximum (1RM) leg press and knee extension was conducted at pre-intervention testing during familiarization (Figure 1, Ddays -7 and -5).
-The participants performed a general warm-up with 10min cycling on an indoor exercise bicycle.
+The highest peak torque values from the respective angular velocities and time points were summarized in an index.
+The index was calculated by dividing the average peak torque value by the highest observed peak torque value per angular velocity and summarizing this new index per angular velocity to a mean strength index.
+
+Assessment of unilateral one repetition maximum (1RM) leg press and knee extension was conducted at before the intervention during familiarization (Figure 1, Days -7 and -5).
+The participants performed a general warm-up with 10 minutes of cycling on an indoor exercise bicycle.
 A protocol consisting of 1x10, 1x6 and 1x3 repetitions with a load equivalent to ~50-75% of assumed max repetitions, was used as a specific warm-up before each of the tests.
 All positions were controlled and recorded at the first 1RM test and repeated for the RT sessions.
-Maximal leg press strength was defined as the maximal load lifted in a controlled fashion, with a knee angle of 45 degrees (start), within a 2.5kg range.
-Attempts where participants did not reach 45 degrees during the eccentric phase, were not approved.
-Maximal knee extension testing followed the same specific warm-up and pause protocol and was defined as maximal load lifted in a controlled fashion, reaching full extension of the knee joint, within a 1.25kg range.
+Maximal leg press strength was defined as the maximal load lifted in a controlled fashion, starting from a knee angle of 90 degrees.
+To find a reproducible 90-degree knee angle for each participant, centimetre markings on the side panels of the leg press machine were used to record where to find 90 degrees for each separate leg and participant.
+Attempts where participants did not reach 90 degrees during the eccentric phase, were not approved.
+Maximal knee extension testing followed the same specific warm-up as the maximal leg press test and was defined as the maximal load lifted in a controlled fashion, reaching full extension of the knee joint.
 Attempts with exaggerated hip movement or beneath full extension were not approved.
-Two minutes of rest were given during the specific warm-up, and three minutes of rest was given between 1RM attempts.
-
-## RESISTANCE TRAINING PROTOCOL
-
-Resistance training consisted of three sets of unilateral leg presses and three sets of knee extensions, with an intensity of 10 repetition maximum (10RM).
-As a general warm-up, the participants cycled on an indoor exercise bicycle for 5-10min.
+Two minutes of rest were given during the specific warm-up, and three minutes of rest were given between 1RM attempts.
+
+### Resistance training protocol
+
+Resistance training consisted of three sets of unilateral leg presses and three sets of unilateral knee extensions, with an intensity of 10 repetitions maximum (10RM).
+As a general warm-up, the participants cycled on an indoor exercise bicycle for 5-10 minutes.
 In addition, before the respective exercises, two 10-repetition warm-up sets were completed at ~50% and ~70% of 10RM.
-To ensure an adequate exercise stimulation throughout the intervention, the exercise load was increased if the participants lifted more than 12 repetitions, as a progressive loading strategy.
-Rest between working sets was two minutes.
-For safety and standardization purposes, all sessions were monitored by trained personnel.
-<!-- From the "old" methods: The rate of perceived exertion (RPE, 0–10-point scale) was logged before every session, and the session score was logged 15min after each session - is this necessary info, as we are not reporting these data?
---> Lastly, training volume (load and repetitions) was logged for every session.
-
-## SAMPLING OF MUSCLE TISSUE AND BLOOD
+To ensure adequate exercise stimulation throughout the intervention, the exercise load was increased the following set if the participants lifted more than 12 repetitions, as a progressive loading strategy.
+If the participants lifted fewer than 8 repetitions per set, the load was reduced in the following set.
+The resting time between working sets was two minutes.
+For safety and standardisation purposes, all sessions were monitored by trained personnel.
+Lastly, training volume (load and repetitions) was logged for every session.
+
+### Sampling of muscle tissue and blood
 
@@
 After the biopsy sampling, muscle tissue was divided into two aliquots for determination of total RNA/expression of rRNA and two aliquots for protein content measurement.
-Aliquots were snap frozen in isopentane (-80°C) and stored at -80°C until further analyses.
-Muscle biopsies were sampled at four times: I/II) Pre-intervention (Figure 1A, 2hrs before training, Day 1 = leg 1, Day 2 = leg 2), and III/IV) after the fifth RT session, two hours before the sixth RT session (Figure 1A, Day 11 = leg 1, Day 12 = leg 2).
+Aliquots were snap-frozen in isopentane (-80°C) and stored at -80°C until further analyses.
+Muscle biopsies were collected at four time points: I/II) Before the intervention (Figure 1A, 2hrs before training, Day 1 = leg 1, Day 2 = leg 2), and III/IV) after the fifth RT session, two hours before the sixth RT session (Figure 1A, Day 11 = leg 1, Day 12 = leg 2).
 At each time point, two samples were taken from the same incision.
@@
 To measure blood glucose levels with and without glucose ingestion/training, capillary blood was collected from finger draws on days with biopsy sampling.
-One capillary blood sample was taken on day 1 (Figure 1A) to serve as a baseline.
-On days 11 and 12 (Figure 1A), capillary blood samples were collected seven times: I) Immediately prior to protein ingestion (0700hrs) II) 45 minutes after protein ingestion (0745hrs) III) 1.5hrs after protein ingestion (0800hrs, i.e., immediately before GLU/PLAC intake), IV) 2hrs after protein ingestion (0900hrs, i.e., immediately before training), IV) in the middle of RT (~0915hrs), V) immediately after training (~0930hrs), and VI) 2hrs after completion of training (~1130hrs).
+One capillary blood sample was collected on day 1 (Figure 1A) to serve as a baseline.
+On days 11 and 12 (Figure 1A), capillary blood samples were collected seven times: I) Immediately before protein ingestion (0700hrs) II) 45 minutes after protein ingestion (0745hrs) III) 1.5 hours after protein ingestion (0830hrs, i.e., immediately before GLU/PLAC intake), IV) 2 hours after protein ingestion (0900hrs, i.e., immediately before training), IV) in the middle of RT (~0915hrs), V) immediately after training (~0930hrs), and VI) 2 hours after completion of training (~1130hrs).
 Capillary blood samples were analysed with in-house equipment (BIOSEN C-Line, EKF diagnostic GmbH, Barleben).
-Venous blood samples were collected from the antecubital vein, coinciding with the capillary samples except 45min after protein ingestion and in the middle of the RT session, to analyse endocrine variables.
-
-## TOTAL RNA EXTRACTION AND REAL-TIME REVERSE TRANSCRIPTION POLYMERASE CHAIN REACTION
+Venous blood samples were collected from the antecubital vein, coinciding with the capillary samples except 45 minutes after protein ingestion and in the middle of the RT session, to analyse endocrine variables.
+
+### Total RNA extraction and real-time reverse transcription polymerase chain reaction
 ```{r totrna-extraction}
@@
 
-Two muscle biopsy aliquots were used for total RNA extraction per leg and resulting in a total of eight RNA samples per participant.
-Total RNA was extracted using TRIzol with muscle tissue homogenized using 0.5mm RNase-free Zirconium beads (~50 ul; Next Advanced, Averill Park, NY, USA) and mechanical agitation (Bullet Blender, Next Advanced, Averill Park, NY, USA).
+Two muscle biopsy aliquots were used for total RNA extraction per leg, resulting in a total of eight RNA samples per participant.
+Total RNA was extracted using TRIzol with muscle tissue homogenised using 0.5mm RNase-free Zirconium beads (~50 ul; Next Advanced, Averill Park, NY, USA) and mechanical agitation (Bullet Blender, Next Advanced, Averill Park, NY, USA).
 Chloroform (Sigma-Aldrich, Oslo, Norway) was used for phase separation, and the RNA pellet was precipitated with isopropanol (VWR International, Oslo, Norway).
-To enable analysis of target gene expression per unit tissue weight [@ellefsen_improved_2008; @ellefsen_reliable_2014], an exogenous RNA control (λ polyA External Standard Kit, Takara Bio Inc., Shiga, Japan) was added at a fixed amount to each sample (0.04 ng ml^-1^ of Trizol reagent).
+To enable analysis of target gene expression per unit tissue weight [@ellefsen_improved_2008; @ellefsen_reliable_2014], an exogenous RNA control (Lambda, λ polyA External Standard Kit, Takara Bio Inc., Shiga, Japan) was added at a fixed amount to each sample (0.04 ng ml^-1^ of TRIzol reagent).
 For assessment of RNA content and purity, RNA was eluted in TE-buffer (1:2) and assessed via spectrophotometry.
@@
 Before quantitative analyses of total RNA, samples with known loss of RNA during extraction (n = `r technical`) or a deviation from the observed RNA to muscle tissue weight relationship larger than 3 $\times$ residual SD while accounting for training status (n = `r observed`) were removed from the data set.
-Total RNA was normalized to wet muscle weight and log transformed before statistical analyses.
-
-Five-hundred ng of RNA was reverse transcribed using Super Script IV Reverse Transcriptase (Invitrogen, Oslo, Norway), according to the manufacturer's instructions using anchored oligo-dT and random hexamer primers (Thermo Scientific, Oslo, Norway).
+Total RNA was normalised to wet muscle weight and log transformed before statistical analyses.
+
+Five hundred ng of RNA was reverse transcribed using Super Script IV Reverse Transcriptase (Invitrogen, Oslo, Norway), according to the manufacturer's instructions using anchored oligo-dT and random hexamer primers (Thermo Scientific, Oslo, Norway).
 All samples were reverse transcribed and diluted to 1:50 before quantitative real-time polymerase chain reaction (qPCR).
-qPCR reactions were run over 40 cycles (3 s 95°C denaturing and 30 s 60°C annealing) on a fast-cycling real-time detection system (Applied Biosystems 7500 fast Real-Time PCR Systems, Life Technologies AS), with a total reaction volume of 10 µl consisting of 2 µl of cDNA, gene-specific primers (0.5 µM final concentration) and a commercial master mix (2X SYBR Select Master Mix, Applied Biosystems, Life Technologies AS, Oslo, Norway).
+qPCR reactions were run over 40 cycles (3 s 95°C denaturing and 30 s 60°C annealing) on a fast-cycling real-time detection system (Applied Biosystems 7500 fast Real-Time PCR Systems, Life Technologies AS), with a total reaction volume of 10 µl consisting of 2 µl of complementary DNA (cDNA), gene-specific primers (0.5 µM final concentration) and a commercial master mix (2X SYBR Select Master Mix, Applied Biosystems, Life Technologies AS, Oslo, Norway).
+An overview of the primers may be found in Table 2.
 Raw fluorescence data was modelled with a best-fit sigmoidal model using the qPCR package [@ritz_qpcr_2008] written for R [@hammarstrom_benefits_2020; @Rcore].
-qPCR data was normalized to wet muscle weight using the external reference gene [@ellefsen_improved_2008; @ellefsen_reliable_2014] and analyzed on the log scale on a target-by-target basis.
+qPCR data was normalised to wet muscle weight using the external reference gene Lambda [@ellefsen_improved_2008; @ellefsen_reliable_2014] and analysed on the log scale on a target-by-target basis.
 <!-- add explanations of normalization to extended methods -->
 
-## PROTEIN EXTRACTION AND IMMUNOBLOTTING
+### Protein extraction and immunoblotting
 
 Total protein was extracted using the Minute Total Protein Extraction Kit for Muscles (Invent Biotechnology), according to the manufacturer’s protocol, optimised for our lab.
-Wet muscle was freeze-dried for 24 hrs and dissected before extraction.
-The tissue was homogenized with a plastic rod in 80 mg protein extraction powder (Invent Biotechnology) and 100 ul ice cold cell lysis buffer (Denaturing Buffer, Invent Biotechnology), and centrifuged at 19 000 g for 1 min.
+Wet muscle was freeze-dried for 24 hours and dissected before extraction.
+The tissue was homogenised with a plastic rod in 80 mg protein extraction powder (Invent Biotechnology) and 100 ul ice-cold cell lysis buffer (Denaturing Buffer, Invent Biotechnology), and centrifuged at 19 000 g for 1 minute.
 The supernatant was divided into aliquots to run samples in duplicates, and total protein concentrations were determined in a 1:10 dilution (Pierce Detergent Compatible Bradford Assay Reagent, Thermo Fisher Scientific, Oslo, Norway).
-The protein samples were diluted to <!-- Check this, it should be 1.5?
--->20.25 µg µl^−1^ total protein with lysis buffer and 4X Laemmli sample buffer (Bio-Rad Laboratories AB, Oslo, Norway) containing 2-mercaptoethanol.
+The protein samples were diluted to a concentration of 1.5 µg µl^−1^ with lysis buffer and 4X Laemmli sample buffer (Bio-Rad Laboratories AB, Oslo, Norway) containing 2-mercaptoethanol..
 All protein samples were incubated at 95°C and stored at -20°C until further analysis.
-The protein samples (20.25 µg total protein) were separated at 250V on 4-20% Tris-Glycine gels (Bio-Rad Laboratories) for 50 min and then transferred to PVDF membranes with wet transfer at 300 mA for 3 hrs.
+The protein samples (20.25 µg total protein) were separated at 250V on 4-20% Tris-Glycine gels (Bio-Rad Laboratories) for 50 minutes and then transferred to PVDF membranes with wet transfer at 300 mA for 3 hours.
 Both gel electrophoresis and protein transfer were performed at 4°C.
-Following the wet transfer, membranes were stained with a reversible total protein stain (Thermo Fisher Scientific) and then blocked for 1hr at room temp with a blocking buffer of Tris-buffered saline (TBS; 20 mM Tris, 150 mM NaCl) with 5% non-fat dry milk and 0.1% Tween-20.
-Primary and secondary antibodies were purchased from Santa Cruz Biotechnology (Texas, USA): UBF, UBF F-9, sc-13125; rpS6, Ribosomal protein S6 C-8, sc-74459; and Thermo Fisher Scientific (Oslo, Norway): c-Myc, 9E10; goat anti-mouse (for c-Myc), goat anti-mouse IgG1 (y1) horseradish peroxidase conjugate; and anti-mouse (anti-mouse IgG1 horseradish peroxidase conjugate)<!-- insert manufacturer -->.
-Antibodies were diluted in blocking buffer to concentrations corresponding to <!-- Add final concentrations of antibodies -->.
-Membranes were incubated overnight with primary antibodies, and for 1 hr with secondary antibodies.
-Between blocking and primary antibody staining, membranes were washed 5 min, between primary and secondary staining, and after secondary staining, membranes were washed 3 x 5 min with TBS-Tween (TBS; 20mM Tris, 150mM NaCl, 0.1% Tween).
-Following the last wash, membranes were incubated 5 min with enhanced chemiluminescent substrate (ECL, SuperSignal West Femto Maximum Sensitivity Substrate, Thermo Fisher Scientific).
+Following the wet transfer, membranes were stained with a reversible total protein stain (Thermo Fisher Scientific) and then blocked for 1 hour at room temp with a blocking buffer of Tris-buffered saline (TBS; 20 mM Tris, 150 mM NaCl) with 5% non-fat dry milk and 0.1% Tween-20.
+Primary and secondary antibodies were purchased from Santa Cruz Biotechnology (Texas, USA): UBF, UBF F-9, sc-13125; rpS6, Ribosomal protein S6 C-8, sc-74459; and Thermo Fisher Scientific (Oslo, Norway): c-Myc, 9E10; goat anti-mouse (for c-Myc), goat anti-mouse IgG1 (y1) horseradish peroxidase conjugate; and anti-mouse (anti-mouse IgG1 horseradish peroxidase conjugate).
+Antibodies were diluted in blocking buffer to concentrations corresponding to 1:25 000 (UBF, rpS6) and 1:5000 (c-Myc).
+
+Membranes were incubated overnight with primary antibodies and for 1 hour with secondary antibodies.
+Between blocking and primary antibody staining, membranes were washed for 5 minutes, between primary and secondary staining, and after secondary staining, membranes were washed for 3 x 5 minutes with TBS-Tween (TBS; 20mM Tris, 150mM NaCl, 0.1% Tween).
+Following the last wash, membranes were incubated for 5 minutes with enhanced chemiluminescent substrate (ECL, SuperSignal West Femto Maximum Sensitivity Substrate, Thermo Fisher Scientific).
 Membrane blocking, secondary antibody incubation, washing and ECL incubation were performed at room temp.
@@
 Chemiluminescence signals were quantified using Image Studio Lite (LI-COR Biotechnology, Lincoln, NE, USA), and total protein content was quantified using ImageJ [@rueden_imagej2_2017], where total protein content was defined as mean grey value of the whole well with between-well values subtracted as background.
-A pooled sample was used as a control on each gel to allow for between gel comparisons and  quantified protein signals were subsequently normalized to the pooled control sample and total protein.
-
-## STATISTICS AND DATA ANALYSIS
+A pooled sample was used as a control on each gel to allow for between-gel comparisons and  quantified protein signals were subsequently normalized to the pooled control sample and total protein.
+
+CONTINUE PROOFREADING HERE!!!!!
+
+### Statistics and data analysis
 
 A priori power calculations showed that 20 participants would grant a statistical power of 80% (alpha = 0.05), accounting for an expected dropout of 20%.
-This power calculation was based on an assumption that the effects of glucose ingestion on ribosome biogenesis (measured as within participant differences in total RNA) may equate to the effects of increasing RT volume from low to moderate [@hammarstrom_benefits_2020] corresponding to an effect size of d =  <!-- Add the effect size here-->.
+This power calculation was based on an assumption that the effects of glucose ingestion on ribosome biogenesis (measured as within-participant differences in total RNA) may equate to the effects of increasing RT volume from low to moderate [@hammarstrom_benefits_2020].
 Total RNA, protein and qPCR data were analyzed by mixed-effects models with fixed effects included as *supplement $\times$ time*.
@@
 
-##### Total RNA and ribosomal RNA
+*Total RNA and ribosomal RNA*
 
@@
 
-From baseline to post-intervention total RNA and rRNA per unit muscle weight increased on average by ~24% and ~33-43%, respectively (Figure 2).
+From baseline to post-intervention total RNA and rRNA per unit muscle weight increased on average by ~20-27% and ~25-57%, respectively (Figure 2).
 However, RT with glucose did not induce increased accumulation of total RNA (Figure 2A, mean difference `r total_rna_stats`) or rRNA  (Figure 2B; 47S, `r stat_rrna$trg_47s`; 18S, `r stat_rrna$trg_18s`; 28S, `r stat_rrna$trg_28s`; 5.8S, `r stat_rrna$trg_5.8s`; 5S, `r stat_rrna$trg_5s`) compared to RT with placebo.
 
-##### Protein
+*Protein*
 
@@
 low glucose/glucose starvation condition on rDNA transcription initiation observed in previous *in vitro* studies [@mariappan_ribosomal_2011; @tanaka_mild_2015] are not translatable to acute effects in human skeletal muscle *in vivo*.
-Notably, the present study was specifically designed to investigate differences in pre-post changes within participants, and is as such also quite limited to this perspective.
+Notably, the present study was specifically designed to investigate differences in pre to post changes within participants, and is as such also quite limited to this perspective.
 Hence, the observed effects of time serve as supplementary to the efficiency of the design, without a negative control group to compare with.
@@
 Unfortunately, three dropped out during the intervention, leaving our statistical power slightly underpowered according to the a priori calculation.
+Furhter, we did not measure the intramuscular glycogen storage, representing another limitation to our design.
+Hence, we cannot know for certain whether there were actual differences within the exercising skeletal muscles between the glucose and placebo condition.
+The participants did indeed match their macro nutrients on pairwise consecutive days, and the plasma glucose measurements did not show any differences in blood sugar until ingestion of placebo/glucose.
+Thus, it may stand to reason that the participants had similar levels of intramuscular levels before ingesting any supplement and that the glucose supplement led to higher glucose availability during exercise, as by design.
+However, even if the leg exercising with glucose supplement did in fact take up the extra glucose, it did not seem to affect any of our outcome measures.
+
+*Conclusion*
 
@@
 
-# APPENDIX
-
-IS THIS REQUIRED? IF NOT, IS IT NECESSARY?
-
-# DATA AVAILABILITY
-
-LINK TO GITHUB
-
-# GRANTS
-
-IS THIS NECESSARY?
-
-# DISCLOSURES
-
-# AUTHOR CONTRIBUTIONS
-
-# References
+### DATA AVAILABILITY
+
+Complete data sets and scripts are downloadable here; https://github.com/Kristianlian/ribose-paper
+
+### SUPPLEMENt MATERIAL
+
+Extended methods can be found here; https://github.com/Kristianlian/ribose-paper
+
+## References
 
@@
                     as_paragraph("0\U00BA sec", as_sup("-1")))) %>%
-  add_header_row(values = c(" ", "Knee-extension peak torque"), colwidths = c(6, 3)) %>%
-  add_body_row(top = FALSE, value = as_paragraph("Values are mean and (SD)"), colwidths = 9) %>%
-  set_caption("Table 1. Participant characteristics")
-
-```
-
+  add_header_row(values = c(" ", "Knee-extension peak torque"), colwidths = c(6, 3))
+
+dxa.flex <- border(dxa.flex, 
+                     i = NULL,
+                     j = NULL,
+                     border = NULL,
+                     border.top = fp_border_default(color = "black"),
+                     border.bottom = fp_border_default(color = "black"),
+                     border.left = fp_border_default(color = "black"),
+                     border.right = fp_border_default(color = "black"),
+                     part = "all")
+
+dxa.flex
+
+```
+
+**Table 2.
+Primer sequences:** Values of Ct are means ± SD.
+rRNA = ribosomal RNA, E = primer efficiency.
+Average cycle thresholds (Ct) and priming efficiencies were calculated from all qPCR reactions.
 ```{r primer table, echo = FALSE}
@@
                     as_paragraph("Sequence (forward - reverse)"),
-                    as_paragraph("CT mean (SD)"),
+                    as_paragraph("Ct mean (SD)"),
                     as_paragraph("E"))) %>%
-  #add_header_row(values = c(" ", " ", " ", "Primer sequences"), 
-               #  colwidths = c(1, 1, 1, 1)) %>%
-  set_caption("Table 2. Primer Sequences") %>%
-  width(j = 1, 1.5, unit = "in") %>%
-  width(j = 2, 3, unit = "in") %>%
-  width(j = 3, 1.5, unit = "in") %>%
-  width(j = 4, 0.5, unit = "in")
-  #set_table_properties(layout = "autofit")
-
-```
+  set_table_properties(layout = "autofit")
+
+primer.tab <- border(primer.tab, 
+                     i = NULL,
+                     j = NULL,
+                     border = NULL,
+                     border.top = fp_border_default(color = "black"),
+                     border.bottom = fp_border_default(color = "black"),
+                     border.left = fp_border_default(color = "black"),
+                     border.right = fp_border_default(color = "black"),
+                     part = "all")
+
+primer.tab
+
+```

### 2023-09-29 figures/abbreviations.table.R

@@ -12,23 +12,21 @@ tab.dat <- abbreviations.dat
 ## Flextable
 abb.tab <- tab.dat %>%
   flextable() %>%
-  set_caption(caption = "Abbreviations",
-              align_with_table = FALSE) %>%
   set_table_properties(layout = "autofit")
 
 abb.tab <- delete_part(x = abb.tab, part = "header")
-  
+
 
 
 
 abb.tab <- border(abb.tab, 
-                     i = NULL,
-                     j = NULL,
-                     border = NULL,
-                     border.top = fp_border_default(color = "black"),
-                     border.bottom = fp_border_default(color = "black"),
-                     border.left = fp_border_default(color = "black"),
-                     border.right = fp_border_default(color = "black"),
-                     part = "all") 
+                  i = NULL,
+                  j = NULL,
+                  border = NULL,
+                  border.top = fp_border_default(color = "black"),
+                  border.bottom = fp_border_default(color = "black"),
+                  border.left = fp_border_default(color = "black"),
+                  border.right = fp_border_default(color = "black"),
+                  part = "all") 
 
 abb.tab

### 2023-09-29 figures/archive/figure1.R

@@ -1,4 +1,4 @@
-### Ribose figure 1 in ejap format######
+### Ribose figure 1 ######
 #
 # Study design, strength (1RM, Humac), c-peptide and blood glucose
 #
@@ -6,7 +6,7 @@
 #
 # Packages 
 library(tidyverse); library(emmeans); library(cowplot); library(grid); library(gridExtra); library(readxl); 
-library(ggtext)
+library(elementalist); library(ggtext)
 #
 ## Data
 
@@ -28,20 +28,23 @@ vol.lemm <- readRDS("./data/data-gen/training/vol.lemm.RDS")
 
 
 ### Plots
-pos <- position_dodge(width = 10)
+pos <- position_dodge(width = .2)
 
 # Designing the plot theme
 
 plot_theme <- theme(panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(),
-                    panel.background = element_rect(fill = "white", colour = NA),
-                    axis.line = element_line(colour = "black", linewidth = 0.4))
+                    panel.background = element_rect(fill = "lightblue", colour = NA),
+                    axis.line = element_line(colour = "black"))
 
 # Colours
 # Colours
 
-colors <- c("black","white")
-                     
+colors <- c("#d7191c",
+            "#fdae61",
+            "#abd9e9",
+            "#2c7bb6")
+
 
 
 ## C-peptide
@@ -56,11 +59,10 @@ cpep.fig <- cpep.change %>%
   annotate("text", x = c(120, 150), y = c(975, 950), label = "*") +
   geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
                 size = 0.5,
-                width = 0,
-                position = position_dodge(width = 10)) +
-  geom_line(position = position_dodge(width = 10), lty = 2) +
-  geom_point(shape = 21, position = position_dodge(width = 10), size = 2) +
-  scale_fill_manual(values = colors[c(1,2)]) +
+                position = pos) +
+  geom_line(position = pos, lty = 2) +
+  geom_point(shape = 21, position = pos, size = 2) +
+  scale_fill_manual(values = colors[c(1,4)]) +
   scale_x_continuous(limits = c(0, 300), breaks = c(0, 90, 120, 150, 270),
[…]

### 2023-09-29 figures/archive/figure3.R

@@ -1,0 +1,352 @@
+### Ribose figure 3 ######
+#
+# This script produces the figures for cmyc, ubf and rps6
+#
+#
+# Packages
+library(emmeans); library(tidyverse); library(cowplot); library(magick); 
+library(nlme); library(ggtext)
+#
+# Data
+
+cmyc <- readRDS("./data/data-gen/protein/cmyc.lemm.RDS")
+
+ubf <- readRDS("./data/data-gen/protein/ubf.lemm.RDS")
+
+rps6 <- readRDS("./data/data-gen/protein/rps6.lemm.RDS")
+
+joined.dat <- readRDS("./data/data-gen/rna.prot.RDS")
+
+# Images
+
+g3 <- cowplot::ggdraw() + cowplot::draw_image("./figures/archive/g3_115_annotated.png", scale = 0.9)
+g6 <- cowplot::ggdraw() + cowplot::draw_image("./figures/archive/g6_115_unannotated.png", scale = 0.9)
+
+ubf.x2.img <- cowplot::ggdraw() + cowplot::draw_image("./figures/archive/ubf_x2_115.png", scale = 0.8)
+cmyc.x2.img <- cowplot::ggdraw() + cowplot::draw_image("./figures/archive/cmyc_x2_115.png", scale = 0.8)
+rps6.x2.img <- cowplot::ggdraw() + cowplot::draw_image("./figures/archive/rps6_x2_115.png", scale = 0.8)
+
+# Colours
+
+colors <- c("#d7191c",
+                     "#fdae61",
+                     "#abd9e9",
+                     "#2c7bb6")
+                     
+# Textsizes
+
+labsize <- 8
+textsize <- 6
+htextsize <- 7
+legendti <- 6
+legendtex <- 5.25
+atext <- 2
+keysize <- 2
+
+# Designing the plot theme
+
+plot_theme <- theme(panel.grid.major = element_blank(),
+                    panel.grid.minor = element_blank(),
+                    panel.background = element_rect(fill = "lightblue", colour = NA),
+                    #plot.background = element_rect(fill = "lightblue", color = NA),
+                    axis.line = element_line(colour = "black"))
+
+## cmyc fig
+
+cmyc.plot <- cmyc %>%
+  # mutate(time = "post") %>%
+  # add_row(supplement = c("PLACEBO", "GLUCOSE"), time = "pre", emmean = 0, lower.CL = 0, upper.CL = 0) %>%
[…]

### 2023-09-29 figures/archive/tab1.R

@@ -48,10 +48,10 @@ dxa.tab <- dxa.dat %>%
             sd.w = sd(weight),
             mean.lm = mean(leanmass),
             sd.lm = sd(leanmass)) %>%
-  mutate(age = paste0(sprintf("%.1f", mean.a), " (", sprintf("%.1f",sd.a), ")"),
-         height = paste0(sprintf("%.1f",mean.h), " (",sprintf("%.1f",sd.h), ")"),
-         weight = paste0(sprintf("%.1f",mean.w), " (", sprintf("%.1f",sd.w), ")"),
-         lm = paste0(sprintf("%.1f",mean.lm), " (", sprintf("%.1f",sd.lm), ")")) %>%
+  mutate(age = paste0(round(mean.a, 2), " (", round(sd.a, 2), ")"),
+         height = paste0(round(mean.h, 2), " (", round(sd.h, 2), ")"),
+         weight = paste0(round(mean.w, 2), " (", round(sd.w, 2), ")"),
+         lm = paste0(round(mean.lm, 2), " (", round(sd.lm, 2), ")")) %>%
   mutate(sex = if_else(sex == "female",
                        "Female",
                        if_else(sex == "male",
@@ -66,7 +66,8 @@ dxa.tab <- dxa.dat %>%
 # data
 
 gender <- dxa.dat %>%
-  select(subject, sex) 
+  select(subject, sex) %>%
+  print()
 
 # Summarising and prepping humac data
 humac.tab <- humac.clean2 %>%
@@ -76,47 +77,60 @@ humac.tab <- humac.clean2 %>%
   group_by(sex, measure) %>%
   summarise(mean.pt = mean(peak.torque),
             sd.pt = sd(peak.torque)) %>%
-  mutate(mean = paste0(sprintf("%.1f",mean.pt), " (", sprintf("%.1f",sd.pt), ")")) %>%
+  mutate(mean = paste0(round(mean.pt, 2), " (", round(sd.pt, 2), ")")) %>%
   select(sex, measure, mean) %>%
   mutate(sex = if_else(sex == "female",
                        "Female",
                        if_else(sex == "male",
                                "Male", sex))) 
-
+ 
 
 # Joining the data frames to be used in the flextable
-
-dxa.flex <- dxa.tab %>%
+desc.dat <- dxa.tab %>%
   full_join(humac.tab) %>%
   pivot_wider(values_from = mean,
               names_from = measure) %>%
[…]

### 2023-09-29 figures/ejap_figures/ejap_figure3.R

@@ -1,508 +1,0 @@
-### Ribose figure 3 ######
-#
-# This script produces the figures for cmyc, ubf and rps6
-#
-#
-# Packages
-library(emmeans); library(tidyverse); library(cowplot); library(magick); 
-
-library(nlme); library(ggtext); library(lme4)
-#
-# Data
-
-
-prot_models <- readRDS("./data/data-gen/protein/protein_lmer_models.RDS")
-
-
-prot_models <- prot_models$prot_models
-
-
-cmyc <- readRDS("./data/data-gen/protein/cmyc.lemm.RDS")
-
-ubf <- readRDS("./data/data-gen/protein/ubf.lemm.RDS")
-
-rps6 <- readRDS("./data/data-gen/protein/rps6.lemm.RDS")
-
-joined.dat <- readRDS("./data/data-gen/rna.prot.RDS")
-
-# Images
-
-g3 <- cowplot::ggdraw() + cowplot::draw_image("./figures/archive/g3_115_annotated.png", scale = 0.9)
-g6 <- cowplot::ggdraw() + cowplot::draw_image("./figures/archive/g6_115_unannotated.png", scale = 0.9)
-
-ubf.x2.img <- cowplot::ggdraw() + cowplot::draw_image("./figures/archive/ubf_x2_115.png", scale = 0.8)
-cmyc.x2.img <- cowplot::ggdraw() + cowplot::draw_image("./figures/archive/cmyc_x2_115.png", scale = 0.8)
-rps6.x2.img <- cowplot::ggdraw() + cowplot::draw_image("./figures/archive/rps6_x2_115.png", scale = 0.8)
-
-# Colours
-
-colors <- c("#d7191c",
-            "gray80",
-            "gray40",
-            "#2c7bb6")
-            
-# Textsizes
-
-labsize <- 8
-textsize <- 6
-htextsize <- 7
-legendti <- 6
-legendtex <- 5.25
-atext <- 2
-keysize <- 2
-
-# Designing the plot theme
-
-plot_theme <- theme(panel.grid.major = element_blank(),
-                    panel.grid.minor = element_blank(),
-                    panel.background = element_rect(fill = "white", colour = NA),
-                    #plot.background = element_rect(fill = "lightblue", color = NA),
-                    axis.line = element_line(colour = "black", linewidth = 1))
-
-## cmyc fig
-
-comp_plot <- function(model, method = "Wald") {
-  
[…]

### 2023-09-29 figures/figure3.R

@@ -5,9 +5,17 @@
 #
 # Packages
 library(emmeans); library(tidyverse); library(cowplot); library(magick); 
-library(nlme); library(ggtext)
+
+library(nlme); library(ggtext); library(lme4)
 #
 # Data
+
+
+prot_models <- readRDS("./data/data-gen/protein/protein_lmer_models.RDS")
+
+
+prot_models <- prot_models$prot_models
+
 
 cmyc <- readRDS("./data/data-gen/protein/cmyc.lemm.RDS")
 
@@ -29,10 +37,10 @@ rps6.x2.img <- cowplot::ggdraw() + cowplot::draw_image("./figures/archive/rps6_x
 # Colours
 
 colors <- c("#d7191c",
-                     "#fdae61",
-                     "#abd9e9",
-                     "#2c7bb6")
-                     
+            "gray80",
+            "gray40",
+            "#2c7bb6")
+            
 # Textsizes
 
 labsize <- 8
@@ -47,81 +55,238 @@ keysize <- 2
 
 plot_theme <- theme(panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(),
-                    panel.background = element_rect(fill = "lightblue", colour = NA),
+                    panel.background = element_rect(fill = "white", colour = NA),
                     #plot.background = element_rect(fill = "lightblue", color = NA),
-                    axis.line = element_line(colour = "black"))
+                    axis.line = element_line(colour = "black", linewidth = 1))
 
 ## cmyc fig
 
-cmyc.plot <- cmyc %>%
-  # mutate(time = "post") %>%
-  # add_row(supplement = c("PLACEBO", "GLUCOSE"), time = "pre", emmean = 0, lower.CL = 0, upper.CL = 0) %>%
-  mutate(time = factor(time, levels = c("pre", "post"), labels = c("Baseline", "Post"))) %>%
-  ggplot(aes(time, exp(emmean), group = supplement, fill = supplement)) +
-  annotate("text", x = "Post", y = 1.6, label = "*", size = 3) +
-  geom_errorbar(aes(ymin = exp(lower.CL), ymax = exp(upper.CL)),
-                width = 0.1,
-                position = position_dodge(width = 0.2)) +
-  geom_line(position = position_dodge(width = 0.2), lty = 2) +
-  geom_point(shape = 21, size = 2, position = position_dodge(width = 0.2)) +
-  scale_fill_manual(values = colors[c(4,1)]) +
[…]

### 2023-09-29 figures/primer.tab.R

@@ -1,37 +1,0 @@
-### Primer sequence table
-
-## Packages
-library(readxl); library(tidyverse); library(flextable)
-
-## Data
-primer.dat <- read_excel("./data/primer.seq2.xlsx")
-
-tab.dat <- primer.dat %>%
-  select(Gene, fr, CT.mean, E) %>%
-  print()
-
-
-
-## Flextable
-primer.tab <- tab.dat %>%
-  flextable() %>%
-  compose(i = 1, j = c(1, 2, 3, 4), part = "header",
-          value = c(as_paragraph("Gene"),
-                    as_paragraph("Sequence (forward - reverse)"),
-                    as_paragraph("CT mean (SD)"),
-                    as_paragraph("E"))) %>%
-  set_caption("Table 2. Primer Sequences")
-  set_table_properties(layout = "autofit")
-  
-
-
-primer.tab <- border(primer.tab, 
-                     i = NULL,
-                     j = NULL,
-                     border = NULL,
-                     border.top = fp_border_default(color = "grey"),
-                     border.bottom = fp_border_default(color = "grey"),
-                     border.left = fp_border_default(color = "grey"),
-                     border.right = fp_border_default(color = "grey"),
-                     part = "all")
-

### 2023-09-29 figures/tab2.R

@@ -1,140 +1,35 @@
-##### Training volume table
-#
-#
-## Author: KL
-#
-#
-## Purpose:
-# The purpose of this script is to produce table 2, presenting average total 
-# session volume and average change per RT session per supplement.  
+### Primer sequence table
 
 ## Packages
-library(tidyverse); library(flextable)
+library(readxl); library(tidyverse); library(flextable)
 
 ## Data
+primer.dat <- read_excel("./data/primer.seq2.xlsx")
 
-vol.clean <- readRDS("./data/data-gen/training/vol.clean.RDS") 
-# Cleaned up data, from the script "volume.change.R#
-
-vol.change <- readRDS("./data/data-gen/training/vol.change.RDS")
-# Total session volum change data, from the script "volume.change.R#
-  
-## Summarising
-
-# First, summarizing total session volume per time point and supplement
-vol.summarised <- vol.clean  %>%
-  # Groups the data by time point and supplement, i.e. session 1, 2, 3.. and placebo or glucose
-  group_by(time, supplement) %>%
-  # Summarises mean and SD of total session volume
-  summarise(mean.vol = mean(tot.volume, na.rm = TRUE),
-            sd.vol = sd(tot.volume, na.rm = TRUE)) %>%
-  mutate(tot.stat = paste0(round(mean.vol, 2), " ± ", round(sd.vol, 2))) %>%
-  select(time, supplement, tot.stat) %>%
-  print()
-  
+tab.dat <- primer.dat %>%
+  select(Gene, fr, CT.mean, E)
 
 
-# Then, summarizing the change in total session volume per time point and supplement
-vol.ch <- vol.change %>%
-  group_by(time, supplement) %>%
-  summarise(mean.change = mean(change, na.rm = TRUE),
-            sd.change = sd(change, na.rm = TRUE)) %>%
-  ungroup() %>%
-  # Adds baseline rows set to 0
-  add_row(time = "baseline", supplement = "placebo", mean.change = 0, sd.change = 0, .before = 1) %>%
-  add_row(time = "baseline", supplement = "glucose", mean.change = 0, sd.change = 0, .before = 2) %>%
-  # Changes the name of time points to correspond with "vol.summarised"
-  mutate(time = if_else(time == "baseline",
-                         "baseline",
-                         if_else(time == "change.2",
-                                 "session2",
-                                 if_else(time == "change.3",
-                                         "session3",
-                                         if_else(time == "change.4",
-                                                 "session4",
-                                                 if_else(time == "change.5",
-                                                         "session5",
[…]

### 2023-09-29 manuscript.RMD
proofreading

@@
 The supernatant was divided into aliquots to run samples in duplicates, and total protein concentrations were determined in a 1:10 dilution (Pierce Detergent Compatible Bradford Assay Reagent, Thermo Fisher Scientific, Oslo, Norway).
-The protein samples were diluted to a concentration of 1.5 µg µl^−1^ with lysis buffer and 4X Laemmli sample buffer (Bio-Rad Laboratories AB, Oslo, Norway) containing 2-mercaptoethanol..
+The protein samples were diluted to a concentration of 1.5 µg µl^−1^ with lysis buffer and 4X Laemmli sample buffer (Bio-Rad Laboratories AB, Oslo, Norway) containing 2-mercaptoethanol.
 All protein samples were incubated at 95°C and stored at -20°C until further analysis.
@@
 
-CONTINUE PROOFREADING HERE!!!!!
-
 ### Statistics and data analysis
@@
 To decrease the risk of Type I errors, random effects were selected from step-wise elimination of terms from the most complex structure (random slopes for time and supplement and their interaction) to less complex.
-The most complex random effects structure that converged was chosen as the final model[@matuschek_2017_balancing].
-Plasma glucose, c-peptide, training volume, and the strength index were analyzed by multiple time-point log-fold change score comparisons, using mixed-effects model with baseline values, time and supplement, and the interaction between time and supplement as fixed effects.
+The most complex random effects structure that converged was chosen as the final model[@matuschek_balancing_2017].
+Plasma glucose, c-peptide, training volume, and the strength index were analyzed by multiple time-point log-fold change score comparisons, using a mixed-effects model with baseline values, time and supplement, and the interaction between time and supplement as fixed effects.
 These data only supported a random intercept per participant.
-The linear mixed-effects models were fitted with the lmer function from the lme4 package using the lmerTest package to procure p-values (Satterthwaite's method for approximating degrees of freedom)[@bates_fitting_2014; @lmerTest], written for R[@Rcore].
+The linear mixed-effects models were fitted with the lmer function from the lme4 package using the lmerTest package to procure p-values (Satterthwaite's method for approximating degrees of freedom)[@bates_fitting_2014], written for R[@Rcore].
 Log-transformed values were expressed as fold changes in visualisations.
@@
 *p* < 0.05 was considered statistically significant.
+
+CONTINUE PROOFREADING HERE!!!!!
 

### 2023-09-29 manuscript.RMD
completed proofreading

@@
 
-CONTINUE PROOFREADING HERE!!!!!
-
 ## Results
@@
 At the other time points, there were no significant differences in mean change between exercising with glucose and placebo (Figure 1D, Post 2RT: *p* = `r pdiff.c2`, Post 4RT: *p* = `r pdiff.c3`, 30min post 6RT: *p* = `r pdiff.c5`, 2h post 6RT: *p* = `r pdiff.c6`, 23h post 6RT: *p* = `r pdiff.c7`).
-Both RT with glucose and placebo led to significantly reduced strength post-fifth session compared to baseline, by 11 and 18% respectively (Figure 1D, Post 5RT: *p* = 0.000).
-Comparisons of the acute data gathered from the post-fifth session until and including 23hrs post-sixth session showed a significant mean increase in strength of 5-9% after RT with glucose and placebo 30min after the sixth RT session (Figure 1D, 30min post 6RT: *p* = `r apc2`) and two hours after the sixth RT session (Figure 1D, 2h post 6RT: *p* = `r apc3`).
+Both RT with glucose and placebo led to significantly reduced strength after the fifth session compared to baseline, by 11 and 18% respectively (Figure 1D, Post 5RT: *p* = 0.000).
+Comparisons of the acute data gathered from after five sessions until and including 23 hours after the sixth session showed a significant mean increase in strength of 5-9% from RT with glucose and placebo 30 minutes after the sixth RT session (Figure 1D, 30min post 6RT: *p* = `r apc2`) and two hours after the sixth RT session (Figure 1D, 2h post 6RT: *p* = `r apc3`).
 Twenty-three hours after the last (sixth) RT session, strength was unchanged compared to after the fifth RT session (Figure 1D, 23h post 6RT: *p* = `r apc4`).
@@
 
-From baseline to post-intervention total RNA and rRNA per unit muscle weight increased on average by ~20-27% and ~25-57%, respectively (Figure 2).
+From baseline to after the intervention total RNA and rRNA per unit muscle weight increased on average by ~20-27% and ~25-57%, respectively (Figure 2).
 However, RT with glucose did not induce increased accumulation of total RNA (Figure 2A, mean difference `r total_rna_stats`) or rRNA  (Figure 2B; 47S, `r stat_rrna$trg_47s`; 18S, `r stat_rrna$trg_18s`; 28S, `r stat_rrna$trg_28s`; 5.8S, `r stat_rrna$trg_5.8s`; 5S, `r stat_rrna$trg_5s`) compared to RT with placebo.
@@
 
-Overall, RT led to robust increases from pre- to post-intervention in all measured proteins (Figure 3A).
+Overall, RT led to robust increases from before to after the intervention in all measured proteins (Figure 3A).
 RT with glucose resulted in estimated levels of c-Myc, UBF and RPS6 being `r stat_west$cmyc`, `r stat_west$ubf` and `r stat_west$rps6`% lower compared to placebo, respectively, without showing statistical significance (*p* = `r min(p_west)`-`r max(p_west)`; Figure 3A).
@@
 These findings suggest that the effects of high vs.
-low glucose/glucose starvation condition on rDNA transcription initiation observed in previous *in vitro* studies [@mariappan_ribosomal_2011; @tanaka_mild_2015] are not translatable to acute effects in human skeletal muscle *in vivo*.
-Notably, the present study was specifically designed to investigate differences in pre to post changes within participants, and is as such also quite limited to this perspective.
+low glucose/glucose starvation conditions on rDNA transcription initiation observed in previous *in vitro* studies [@mariappan_ribosomal_2011; @tanaka_mild_2015] are not translatable to acute effects in human skeletal muscle *in vivo*.
+Notably, the present study was specifically designed to investigate differences in pre to post changes within participants and is as such also quite limited to this perspective.
 Hence, the observed effects of time serve as supplementary to the efficiency of the design, without a negative control group to compare with.
@@
 To our knowledge, no previous study has investigated the effects of glucose compared to placebo ingestion with RT on ribosome biogenesis in human skeletal muscle previously, making our hypotheses exploratory and comparisons few.
-*In vitro* studies on mice glomerular epithelial cells [@mariappan_ribosomal_2011] and human breast cancer cells [@tanaka_mild_2015] have observed glucose-induced UBF-dependent augmentation of ribosome biogenesis [@mariappan_ribosomal_2011], and reduced rRNA transcription by glucose starvation [@tanaka_mild_2015].
+*In vitro* studies on mice glomerular epithelial cells [@mariappan_ribosomal_2011] and human breast cancer cells [@tanaka_mild_2015] have observed glucose-induced UBF-dependent augmentation of ribosome biogenesis [@mariappan_ribosomal_2011] and reduced rRNA transcription by glucose starvation [@tanaka_mild_2015].
 These findings do imply an important role for glucose in the initiation of rDNA transcription, the rate-limiting step of ribosome biogenesis [@mariappan_ribosomal_2011; @tanaka_mild_2015].
@@
 Moreover, robust increases in markers of ribosome biogenesis such as 47S pre-rRNA and mature rRNA can be expected after an acute bout of RT [@figueiredo_impact_2016] or a short period of RT [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022].
-Therefore, we hypothesized that if glucose can, directly or indirectly via insulin, stimulate the initiation of rDNA transcription, five sessions of RT with glucose compared to placebo will elicit this effect.
+Therefore, we hypothesized that if glucose can, directly or indirectly via insulin, stimulate the initiation of rDNA transcription, five sessions of RT with glucose compared to placebo would elicit this effect.
 Further, glucose ingestion increases energy levels, leading us to hypothesize that high glucose vs.
@@
 Due to the exploratory nature of our hypotheses, this was hardly a very surprising result.
-Indeed, previous studies have used cell cultures from yeast [@zhai_human_2012], rodent [@hoppe_amp-activated_2009; @mariappan_ribosomal_2011] or human cells [@tanaka_mild_2015] and are, as such, not directly comparable to human skeletal muscle cells.
+Indeed, previous studies have used cell cultures from yeast [@zhai_human_2012], rodents [@hoppe_amp-activated_2009; @mariappan_ribosomal_2011] or human cells [@tanaka_mild_2015] and are, as such, not directly comparable to human skeletal muscle cells.
 Importantly, previous studies investigated high vs.
 low glucose conditions [@mariappan_ribosomal_2011], or high glucose vs.
-glucose starvation [@hoppe_amp-activated_2009; @tanaka_mild_2015], while the present study aimed to compare high glucose condition to placebo, with a matched daily macronutrient and energy intake.
+glucose starvation [@hoppe_amp-activated_2009; @tanaka_mild_2015], while the present study aimed to compare the high glucose condition to a placebo condition, with a matched daily macronutrient and energy intake.
 Therefore, the comparison made in the present study was high plasma glucose levels vs.
@@
 Hence, it seems quite reasonable to observe similar changes in these three proteins.
-<!-- OLD: While content of UBF and rpS6 changed equally in RT with glucose compared to placebo, content of c-Myc was significanly higher after RT with placebo.
-This was unexpected, since the general transcription factor c-Myc previously has been described as a potent regulator of ribosome biogenesis, independent of mTORC1, and a direct regulator of UBF [@mori_c-myc_2021; @poortinga_c-myc_2011; @west_acute_2016].
-Along that notion, it would be reasonable to expect that differences in c-Myc content was reflected in total RNA and rRNA.
-However, this was not the case in the present study.
-c-Myc's independency from mTORC1 may explain why only c-Myc content differed between treatments, while UBF and rpS6 content changed equally.
-Moreover, the observation that total RNA and rRNA were unaffected by the significant differences in c-Myc content, does strengthen the importance and specificity of UBF content in human skeletal muscle ribosome biogenesis.
--->The linear relationship found exclusively between UBF content and total RNA accumulation, and neither c-Myc nor rpS6, support the importance and specificity of UBF content in human skeletal muscle ribosome biogenesis.
+The linear relationship found exclusively between UBF content and total RNA accumulation, and neither c-Myc nor rpS6, supports the importance and specificity of UBF content in human skeletal muscle ribosome biogenesis.
 While this is not a novel finding, it has only recently been observed in human skeletal muscle following RT [@hammarstrom_ribosome_2022].
@@
 As in the biological data, there was generally no difference in the change of strength as an index between the conditions.
-Generally, skeletal muscle strength decreased from baseline to post-fifth and -sixth RT sessions, with a significantly less reduction in glucose compared to placebo at post-fifth session measurement.
+Generally, skeletal muscle strength decreased from baseline to after the fifth and -sixth RT sessions, with a significantly less reduction in glucose compared to placebo at the measurement after five sessions.
 This could suggest that the RT protocol fatigued the participants, as both training frequency and intensity were quite high.
-Another promising explanation might be the biphasic recovery pattern, as described by Raastad & Hallén [@raastad_recovery_2000], where the participants experienced a rapid recovery during the initial 11hrs post-exercise, followed by a levelling off or drop until 22hrs post-exercise.
-Herein, inflammation and phagocytic activity were proposed to be involved in the performance drop between 11-22hrs [@raastad_recovery_2000].
-Indeed, this pattern seems quite similar to what was observed in the present study, with a rapid recovery at 30min and 2hrs post-sixth RT session and a drop at 23hrs.
-Further, strength testing during the intervention was conducted 23hrs after RT, meaning that these tests also may have been influenced by the biphasic recovery.
-However, this does not explain the difference between conditions at post-fifth RT session, or the great drop in strength from post 4RT to post 5RT sessions.
+Another promising explanation might be the biphasic recovery pattern, as described by Raastad & Hallén [@raastad_recovery_2000], where the participants experienced a rapid recovery during the initial 11 hours after exercise, followed by a levelling off or drop until 22 hours after exercise.
+Herein, inflammation and phagocytic activity were proposed to be involved in the performance drop between 11-22 hours [@raastad_recovery_2000].
+Indeed, this pattern seems quite similar to what was observed in the present study, with a rapid recovery at 30 minutes and 2 hours after the sixth RT session and a drop at 23 hours after the sixth session.
+Further, strength testing during the intervention was conducted 23 hours after RT, meaning that the biphasic recovery may have also influenced these tests.
+However, this does not explain the difference between conditions at observed after five RT session, or the drop in strength from after four RT sessions to after five RT sessions.
 A possible argument could be that exercising without glucose may have caused more stress compared to exercising with glucose, as glucose is the preferred energy source during strenuous exercise [@mul_exercise_2015], thus increasing performance with glucose compared to placebo.
@@
 Hence, there were no differences in mechanical loading to induce a higher stress between conditions.
-Arguably, an increased energy availability via glucose ingestion during RT may induce less acute stress on the exercised skeletal muscle, therefore less fatigue, and perhaps less performance reduction, compared to placebo during RT <!-- this is possibly a stretch -->[@westerblad_mechanisms_1998; @kent-braun_central_1999] Unfortunately, we did not conduct measurements of markers of metabolic stress such as inorganic phosphate, H+, Mg2+ and the ADP/ATP ratio [@westerblad_mechanisms_1998; @kent-braun_central_1999].
+Arguably, an increased energy availability via glucose ingestion during RT may induce less acute stress on the exercised skeletal muscle, therefore less fatigue, and perhaps less performance reduction, compared to placebo during RT [@westerblad_mechanisms_1998; @kent-braun_central_1999].
+Unfortunately, we did not conduct measurements of markers of metabolic stress such as inorganic phosphate, H+, Mg2+ and the ADP/ATP ratio [@westerblad_mechanisms_1998; @kent-braun_central_1999].
 Therefore, discussing the potential effect of differences in metabolic stress between conditions would only be speculation, however probable.
@@
 Initially, we met our minimum limit of 16 participants.
-Unfortunately, three dropped out during the intervention, leaving our statistical power slightly underpowered according to the a priori calculation.
-Furhter, we did not measure the intramuscular glycogen storage, representing another limitation to our design.
-Hence, we cannot know for certain whether there were actual differences within the exercising skeletal muscles between the glucose and placebo condition.
-The participants did indeed match their macro nutrients on pairwise consecutive days, and the plasma glucose measurements did not show any differences in blood sugar until ingestion of placebo/glucose.
+Unfortunately, three dropped out during the intervention, leaving our statistical analyses slightly underpowered according to the a priori calculation.
+Moreover, we did not measure the intramuscular glycogen storage, representing another limitation to our design.
+Hence, we cannot know for certain whether there were actual differences within the exercising skeletal muscles between the glucose and placebo conditions.
+The participants did indeed match their macronutrients on pairwise consecutive days, and the plasma glucose measurements did not show any differences in blood sugar until ingestion of placebo/glucose.
 Thus, it may stand to reason that the participants had similar levels of intramuscular levels before ingesting any supplement and that the glucose supplement led to higher glucose availability during exercise, as by design.
-However, even if the leg exercising with glucose supplement did in fact take up the extra glucose, it did not seem to affect any of our outcome measures.
+However, even if the leg exercising with glucose supplement did in fact take up the extra glucose, it did not seem to affect any of our main outcome measures.
 
@@
 In conclusion, ingestion of glucose immediately before and after resistance exercise training does not acutely augment ribosome biogenesis after two weeks of heavy resistance exercise training, in moderately trained young adults.
-Neither does it acutely enhance muscular performance during the exercise period or after five sessions, nor recovery within 23hrs of the last session.
+Neither does it acutely enhance muscular performance during the exercise period or after five sessions, nor recovery within 23 hours of the last session.
 The observations in total RNA, rRNA and rDNA transcription initiation-associated proteins support the key role of UBF in ribosome biogenesis regulation in human skeletal muscle following resistance training.
@@
 
-### DATA AVAILABILITY
+### Data availability
 
@@
 
-### SUPPLEMENt MATERIAL
+### Supplementary material
 
@@
 
+\pagebreak
 ## References
@@
 Skeletal muscle strength was measured as peak torque in unilateral isometric and isokinetic (at 60 and 240 degrees per second) knee extension before, multiple times during, and after five and six session.
-<!-- A total of three participants dropped out of the intervention, either due to reported muscular discomforts of heavy RT, or sickness in their family - necessary to include this here?
-Its already reported under "experimental design" -->.
 B and C) Changes in plasma glucose (B, mmol/L) and serum c-peptide levels (C, pmol/L).
@@
 Total RNA and rRNA changes were calculated as log-fold change score per mg wet muscle weight.
-Mean change scores of the duplicates were calculated and transformed to the log-scale before modelling, then reverse-transformed for figure illustration.
+Mean change scores of the duplicates were calculated and transformed to the log-scale before modelling, then reverse-transformed for figure visualisation.
 Values are estimated marginal means fold change per leg per supplement ± 95% CI.
@@
 Changes in protein levels were calculated as log-fold change scores normalized by pools (pool of all protein samples per gel).
-Mean log fold-change scores of the duplicates were calculated before modelling and reverse-transformed for figure illustration.
+Mean log fold-change scores of the duplicates were calculated before modelling and reverse-transformed for figure visualisation.
 Values are estimated marginal means fold change per leg per supplement ± 95% CI.

### 2023-09-29 manuscript.RMD
update of figure legend for fig 3

@@
 
-**Figure 3:** Changes in protein content.
-A) Fold-change pre-post in levels of protein in placebo and glucose + fold-change difference 95% CI, with representative western blots of the respective proteins.
-Protein samples were analyzed in two duplicates per biopsy per time point, loaded on separate gels in an inverted order, e.g.
-from gel 3 to 6, as shown by the duplicates (1 and 2).
-Changes in protein levels were calculated as log-fold change scores normalized by pools (pool of all protein samples per gel).
-Mean log fold-change scores of the duplicates were calculated before modelling and reverse-transformed for figure visualisation.
-Values are estimated marginal means fold change per leg per supplement ± 95% CI.
-Glucose n = 13, placebo n = 13.
-B) Representative total protein stain blot.
-C) Linear relationship between total RNA (ng x mg) and UBF levels (SD units), with time added as a covariate.
-Total RNA was normalized by wet muscle weight, and UBF was normalized by pools per gel and total protein per lane factor.
-Values are presented as log-transformed means.
+**Figure 3:** Changes in c-Myc, UBF and RPS6 protein content from pre- to post-training in Placebo and Glucose conditions together with differences between conditions (second axis in a).
+Representative western blots of the respective proteins are shown under each panel together with total protein stains in (b).
+Protein samples were analyzed in two duplicates per biopsy per time point, loaded on separate gels in an inverted order as exemplified by the duplicates (1 and 2 in a and b).
+Values are estimated fold change per condition with 95% CI, Glucose n = 13 and placebo n = 13.
+A linear relationship was shown between between total RNA (ng x mg) and UBF levels (SD units) while controlling for time.
+Total RNA was normalized by wet muscle weight, and UBF was normalized by a pooled sample used on each gel.
 

### 2023-09-30 manuscript.RMD
final updates

@@
 
-**Figure 1:** A) An overview of the experimental design with 12 days of concomitant dietary intervention and resistance training (RT), preceded by 7 days involving familiarization.
+**Figure 1:** a) An overview of the experimental design with 12 days of concomitant dietary intervention and resistance training (RT), preceded by 7 days involving familiarization.
 Between days -7 and -1, participants were familiarized to the RT exercises via 1RM leg press and knee extension testing, and to the strength tests via Humac Norm dynamometer (days -7 and -5).
@@
 Blood for measurement of plasma glucose and -c-peptide was sampled at baseline (Day 1), and during post testing (Day 11 leg 1, Day 12 leg 2), via finger draws and venous blood samples.
-Skeletal muscle strength was measured as peak torque in unilateral isometric and isokinetic (at 60 and 240 degrees per second) knee extension before, multiple times during, and after five and six session.
-B and C) Changes in plasma glucose (B, mmol/L) and serum c-peptide levels (C, pmol/L).
+Skeletal muscle strength was measured as peak torque in unilateral isometric and isokinetic (at 60 and 240 degrees per second) knee extension before, multiple times during, and after five and six sessions.
+B and c) Changes in plasma glucose (b, mmol/L) and serum c-peptide levels (c, pmol/L).
 Glucose levels in blood was measured via finger draws 120 (-120), 90 (-90), and 30min (-30) before RT, immediately before RT (0), during RT (15), immediately after RT (30) and 2hrs after RT (120).
 C-peptide levels were measured simultaneously to these finger draws, except for 90min before and during RT.
-D) Changes in muscular strength measured as isometric and isokinetic peak torque (60 and 240 d/s) via Humac Norm Dynamometer, conducted at baseline (A: Day -1), after two and four RT sessions (A: Day 4 and 8 leg 1, Day 5 and 9 leg 2), after five RT sessions/before the 6th session (Day 11 leg 1, Day 12 leg 2), as well as 30min, 2hrs and 23hrs after the 6th RT session (A: Day 11/12 leg 1, Day 12/13 leg 2).
+d) Changes in muscular strength measured as isometric and isokinetic peak torque (60 and 240 d/s) via Humac Norm Dynamometer, conducted at baseline (a: Day -1), after two and four RT sessions (a: Day 4 and 8 leg 1, Day 5 and 9 leg 2), after five RT sessions/before the 6th session (Day 11 leg 1, Day 12 leg 2), as well as 30min, 2hrs and 23hrs after the 6th RT session (a: Day 11/12 leg 1, Day 12/13 leg 2).
 The index was calculated by normalizing peak torque values to the highest peak torque value at each respective speed, and then summarized and used in change score calculations.
@@
 
-**Figure 2:** Changes in total RNA and ribosomal RNA.
-A) Total RNA, B) 47S pre-rRNA, 18S rRNA, 28S rRNA, 5.8S rRNA, 5S rRNA.
+**Figure 2:** Changes in total RNA and ribosomal RNA with Glucose and Placebo conditions.
+a) Total RNA, b) 47S pre-rRNA, 18S rRNA, 28S rRNA, 5.8S rRNA, 5S rRNA.
 Baseline = Day 1 leg 1/ Day 2 leg 2, Post = Day 11 leg 1, Day 12 leg 2.

### 2023-10-06 manuscript2.RMD
working on feedback

--- 2023-09-30 manuscript.RMD
@@
 
-**Purpose:** Muscle hypertrophy, induced by resistance training, is determined by ribosomal biogenesis.
-High-glucose treatment may augment the transcription of ribosomal RNA, as evident from in vitro studies.
-However, its effect on resistance training-induced adaptations remains unexplored.
-
-**Methods:** Sixteen healthy, moderately trained individuals (male/female, n = 9/7; age, 23.7 (1.8)/24.6 (4.8)), participated in a within-participant trial with unilateral resistance training with glucose- and placebo supplementation on alternate days over six sessions per condition.
-Participants remained fasted overnight, receiving glucose/placebo and protein supplementation only before and after RT sessions (3 sets of 10 repetitions maximum unilateral leg press and knee extension).
+**Purpose:** Resistance training-induced skeletal muscle hypertrophy seems to depend on ribosome biogenesis and content.
+High glucose treatment may augment ribosomal biogenesis, as evident from *in vitro* studies, and may thus act to potentiate resistance training-induced adaptations *in vivo*.
+Hoever, this remains largely unexplored.
+
+**Methods:** Sixteen healthy, moderately trained individuals (male/female, n = 9/7; age, 24.1 (3.3)), participated in a within-participant crossover trial, conducting unilateral resistance training (leg press and knee extension, 3 sets of 10 repetitions maximum) with pre- and post-exercise ingestion of either glucose (3x30g, 90g total) or placebo supplement (Stevia rebaudiana extract, 3x0.3g, 0.9g total), together with protein (2x25g, 50g total), on alternating days over a period of twelve days; six sessions per condition, resistance training was performed in an otherwise fasted state, as morning sessions.
 Micro biopsies were sampled from m.
-vastus lateralis before and after the intervention.
-
-**Results:** Between conditions, there were no significant differences in increases of total RNA and ribosomal RNA (mean difference `r total_rna_stats`, 47S, `r stat_rrna$trg_47s`; 18S, `r stat_rrna$trg_18s`; 28S, `r stat_rrna$trg_28s`; 5.8S, `r stat_rrna$trg_5.8s`; 5S, `r stat_rrna$trg_5s`), or levels of cellular myelocytomatosis oncogene, upstream binding factor and ribosomal protein S6 (*p* = `r min(p_west)`-`r max(p_west)`).
-There was a linear relationship between upstream binding factor and total RNA, where an increase of 1 standard deviance unit of upstream binding factor equated to a $\sim$ `r corr_incr`% increase in total RNA (*p* = `r corrp`).
-
-**Conclusion:** Glucose supplementation before and after RT does not augment RT-induced ribosome biogenesis after two weeks of heavy RT in moderately trained young adults.
+vastus lateralis before and after the intervention, and maximal unilateral isometric and isokinetic knee extension torque were measured before, during (days 4/5, 8/9) and after the intervention.
+
+**Results:** Glucose ingestion did not have beneficial effects on resistance training-induced increases of ribosomal content (mean difference 7.6% [-7.2, 24.9], *p* = 0.34; ribosomal RNA, 47S/18S/28S/5.8S/5S, range 7.6%-37.9%, *p* = 0.40-0.98) or levels of relevant transcriptional or translational regulators (c-MYK/UBF/rpS6, *p* = 0.094-0.292).
+Of note, both baseline and trained state data of total RNA showed a linear relationship with UBF; ∼14% increase in total RNA corresponded to 1 standard deviation unit increase in UBF (*p* = 0.003).
+
+**Conclusion:** Glucose supplementation before and after resistance training did not augment ribosomal accumulation during two weeks of heavy-load resistance training in moderately trained young adults.
 
@@
 ### Statements and Declarations
+
 This study was financed through research funds from the Inland Norway University of Applied Sciences.
-There was no economic vested interest and everyone participating as researchers and associates received only a regular salary during this study.
+The authors have no competing interests to declare, and researchers and associates participating in the study conduct received only a regular salary.
 
@@
 
-Data was collected at the sports science lab at the Inland Norway University of Applied Sciences, with all analyses except for serum blood samples performed at the sports science lab.
-Serum blood samples were performed at Innlandet Hospital Trust.
+Data was collected at the sports science lab at the Inland Norway University of Applied Sciences, with all analyses performed on site, except serum blood analyses, which were performed at Innlandet Hospital Trust.
 S.E., H.H.
@@
 ### Acknowledgements
-Gratitude is due, to Henriette Spilhaug Bollandsås, Nora Lerdalen Bonsak and Thea Bøhn Nyløkken for their dedication and hard work in the training intervention.
+Gratitude is due, to Henriette Spilhaug Bollandsås, Nora Lerdalen Bonsak and Thea Bøhn Nyløkken for their dedication and hard work during the training intervention.
 The authors would also like to express their gratitude to all the participants for staying committed to this demanding protocol in an impressive way.
@@
 While there certainly may be differences in the genetic predisposition in terms of readiness to adapt to exercise [@thalacker-mercer_cluster_2013], there are also recent studies showing that epigenetic factors such as ribosomal DNA (rDNA) availability may also impact the ability to mount an anabolic response to RT [@figueiredo_genetic_2021].
-Indeed, the ability to synthesise ribosomes following a period of RT is connected to the magnitude of RT-induced responses in an individual [@hammarstrom_benefits_2020].
+Indeed, the ability to synthesise ribosomes during a period of RT is connected to the magnitude of RT-induced responses in an individual [@hammarstrom_benefits_2020].
 For instance, increasing training volume generally induces greater ribosome biogenesis and thus, greater benefits of RT [@hammarstrom_benefits_2020;@krieger_single_2009;@schoenfeld_strength_2017].
@@
 
-Nutritional supplements such as protein and creatine effectively optimise RT adaptations [@cermak_protein_2012;@lanhers_creatine_2015;@lanhers_creatine_2017;@morton_systematic_2018].
+Nutritional supplements such as protein and creatine is often used to optimise RT adaptations [@cermak_protein_2012;@lanhers_creatine_2015;@lanhers_creatine_2017;@morton_systematic_2018].
 However, it remains equivocal if other nutritional adjuvants such as glucose can increase the efficacy of RT [@tezze_anabolic_2023].
@@
 Furthermore, energy availability is a decisive factor in the *de novo* synthesis of ribosomes [@kusnadi_regulation_2015;@moss_housekeeper_2007;@sebata_control_2018] which in turn determines muscle growth by increasing the muscle's translational capacity [@figueiredo_regulation_2019;@hammarstrom_benefits_2020;@sebata_control_2018;@stec_ribosome_2016;@von_walden_ribosome_2019].
+In addition, insulin may have anabolic effects with elevated amino acids, and has a clear role in reduction of muscle protein breakdown independent of amino acid availability [@abdulla_role_2016].
 Therefore, investigating the effects of combined RT and glucose ingestion may provide valuable insight into the potential additive effect of glucose and RT on ribosome biogenesis.
 
-Transcription of ribosomal RNA (rRNA) by RNA Polymerase I is considered the rate-limiting step in synthesising new ribosomes [@moss_promotion_1995].
+Ribosomal biogenesis and content seem to be a prerequisite for skeletal muscle growth, and transcription of ribosomal RNA (rRNA) by RNA Polymerase I is considered the rate-limiting step in synthesising new ribosomes [@moss_promotion_1995].
 Multiple signalling pathways converge to regulate rRNA transcription, including the mammalian target of rapamycin (mTORC1) signal-transduction pathway and c-Myc [@kusnadi_regulation_2015;@von_walden_ribosome_2019;@sebata_control_2018; @west_acute_2016; @mori_c-myc_2021].
 mTOR receives input from growth factors, hormones, mechanical loading, and nutrients to balance protein synthesis through multiple mechanisms based on cellular energy levels [@hoppe_amp-activated_2009].
-Importantly, mTOR and its downstream target ribosomal protein S6 kinase beta-1 (S6K1) serves as a mediator for insulin signalling in skeletal muscle [@hillier_physiological_2000].
-The mTORC1 pathway contributes to ribosome biogenesis through the regulation of the translation of ribosomal proteins and to forming of the preinitiation complex (PIC) that marks the initiation of rRNA transcription [@figueiredo_regulation_2019;@von_walden_ribosome_2019].
-Parallel to mTORC1 [@mori_c-myc_2021; @west_acute_2016], c-Myc increases ribosomal biogenesis directly and through transcriptional control of the upstream binding factor (UBF) [@sanij_ubf_2008; @poortinga_c-myc_2011].
-Additionally, UBF phosphorylation, required for interaction with the ribosomal deoxyribonucleic acid (rDNA) promoter, is increased by high glucose in a mTORC1 dependent manner (rapamycin sensitive) in kidney glomerular epithelial cells [@mariappan_ribosomal_2011].
-Independently from UBF and mTORC1, high glucose was shown to lead to chromatin remodelling, which promoted rRNA transcription in cell cultures [@zhai_human_2012].
+Importantly, mTOR and its downstream target ribosomal protein S6 kinase beta-1 (S6K1) serves as a direct mediator of insulin signalling in skeletal muscle [@hillier_physiological_2000].
+The mTORC1 pathway contributes to ribosome biogenesis through the regulation of the translation of ribosomal proteins and to forming of the preinitiation complex (PIC) that marks the initiation of rRNA transcription [@figueiredo_regulation_2019; @von_walden_ribosome_2019].
+Parallel to mTORC1 [@mori_c-myc_2021; @west_acute_2016], c-Myc increases ribosomal biogenesis directly through transcriptional control of the upstream binding factor (UBF) [@sanij_ubf_2008; @poortinga_c-myc_2011].
+Indeed, UBF phosphorylation, which is required for interaction with the ribosomal deoxyribonucleic acid (rDNA) promoter, seems to be increased by high-glucose treatment in a mTORC1-dependent manner (rapamycin sensitive) in kidney glomerular epithelial cells [@mariappan_ribosomal_2011].
+Independent of UBF and mTORC1, high glucose was shown to lead to chromatin remodelling, which promoted rRNA transcription in cell cultures [@zhai_human_2012].
 Together these observations indicate a potential role of glucose in positively affecting muscle rRNA synthesis.
-
-The purpose of this investigation was to test the hypothesis that glucose supplementation before and after RT increases markers of ribosomal abundance following five RT sessions.
-Secondly, we aimed to describe the association between UBF and total RNA abundance in human skeletal muscle.
+<!-- sørg for at disse to avsnittene henger godt sammen!-->
 There have been multiple studies recently suggesting translational capacity to be as important, if not more important than, translational efficiency concerning long-term skeletal muscle adaptations to RT [@hammarstrom_benefits_2020;@hammarstrom_ribosome_2022;@figueiredo_revisiting_2019].
@@
 [@hammarstrom_benefits_2020;@hammarstrom_ribosome_2022], understanding underlying factors regulating and affecting ribosome biogenesis seems key to furthering our understanding of the optimisation of RT to individual phenotypes.
+
+The purpose of this investigation was to test the hypothesis that glucose supplementation before and after five RT-sessions over a period of 12 days will potentiate RT-associated accumulation of markers of ribosomal abundance following five RT sessions.
+Secondly, we aimed to describe the association between changes in total RNA abundance and UBF in human skeletal muscle.
 
@@
 
-There was no difference in mean change of total training session volume between exercising with placebo or glucose, with an increase of 18 and 16% from session 1 to session 6, respectively (*p* > 0.05).
+Glucose ingestion compared to placebo before and after RT-sessions did not lead to a difference in mean change of total training session volume, with an average increase of 18 and 16% from session 1 to session 6, respectively (*p* > 0.05).
 There were no differences in mean macronutrient intake (protein, fat, carbohydrates) or total calorie intake between glucose and placebo on pairwise consecutive days (*p* > 0.05 for all).
@@
 
-Glucose ingestion before and after RT led to significant increases in plasma glucose levels by 38% immediately before RT (Figure 1B, 0 min), by 31% during RT (Figure 1B, 15 min) and by 32% immediately after RT (Figure 1B, 30 min; all : *p* < 0.001), compared to placebo.
-Two hours after the session, RT with glucose measured 8% lower plasma glucose levels compared to RT with placebo (Figure 1B, 270min: *p* = `r p.270`).
-Similarly, levels of c-peptide increased with glucose ingestion over placebo, by 95% immediately before (Figure 1C, 0 min) and 87% after RT (Figure 1C, 30 min; both *p* < 0.001).
+Glucose ingestion before and after RT led to increases in plasma glucose levels compare to baseline by 38% immediately before RT (Figure 1B, 0 min), by 31% during RT (Figure 1B, 15 min) and by 32% immediately after RT (Figure 1B, 30 min; all : *p* < 0.001), with no changes observed in the placebo condition (*p* > 0.05).
+Compared to the placebo condition, ingestion of glucose increased plasma glucose levels by 36% immediately before RT (Figure 1B, 0 min), by 27% during RT (Figure 1B, 15 min) and by 28% immediately after RT (Figure 1B, 30 min; all : *p* < 0.001).
+Two hours after the RT session, glucose ingestion was associated with 12% lower plasma glucose levels compared to baseline, and 8% lower compared to placebo (Figure 1B, 270min: *p* = `r p.270`).
+
+Glucose ingestion before and after RT led to increases in levels of c-peptide compared to baseline, by 95% immediately before (Figure 1C, 0 min) and 87% after RT (Figure 1C, 30 min; both *p* < 0.001), with no changes observed with the placebo condition (*p* > 0.05).
+Compared to the placebo condition, ingestion of glucose increased levels of c-peptide by 85% immediately before (Figure 1C, 0 min) and 85% after RT (Figure 1C, 30 min; both *p* < 0.001).
 
@@
 
-After 5 RT sessions (Figure 1D, Post 5RT), there was a significant difference in strength between glucose and placebo,  where placebo decreased strength by 7% more than glucose (Figure 1D, *p* = `r pdiff.c4`).
-At the other time points, there were no significant differences in mean change between exercising with glucose and placebo (Figure 1D, Post 2RT: *p* = `r pdiff.c2`, Post 4RT: *p* = `r pdiff.c3`, 30min post 6RT: *p* = `r pdiff.c5`, 2h post 6RT: *p* = `r pdiff.c6`, 23h post 6RT: *p* = `r pdiff.c7`).
+Glucose ingestion before and after RT sessions did not generally enhance skeletal muscle recovery throughout the intervention, neither during the intervention in a rested state (Figure 1D, 23hrs after exercise; Post 2RT: *p* = `r pdiff.c2`, Post 4RT: *p* = `r pdiff.c3`), nor acutely after the sixth and final RT-session (30min post 6RT: *p* = `r pdiff.c5`, 2h post 6RT: *p* = `r pdiff.c6`) or in a rested state after the sixth RT session (23h post 6RT: *p* = `r pdiff.c7`).
+Glucose ingestion before and after RT compared to placebo was associated with a 7% less reduction in strength after five RT sessions (Figure 1D, *p* = `r pdiff.c4`).
+
 Both RT with glucose and placebo led to significantly reduced strength after the fifth session compared to baseline, by 11 and 18% respectively (Figure 1D, Post 5RT: *p* = 0.000).
-Comparisons of the acute data gathered from after five sessions until and including 23 hours after the sixth session showed a significant mean increase in strength of 5-9% from RT with glucose and placebo 30 minutes after the sixth RT session (Figure 1D, 30min post 6RT: *p* = `r apc2`) and two hours after the sixth RT session (Figure 1D, 2h post 6RT: *p* = `r apc3`).
+Comparisons of the acute data gathered from after five sessions until and including 23 hours after the sixth session showed an average increase in strength of 5-9% from RT with glucose and placebo 30 minutes after the sixth RT session (Figure 1D, 30min post 6RT: *p* = `r apc2`) and two hours after the sixth RT session (Figure 1D, 2h post 6RT: *p* = `r apc3`).
 Twenty-three hours after the last (sixth) RT session, strength was unchanged compared to after the fifth RT session (Figure 1D, 23h post 6RT: *p* = `r apc4`).
@@
 
-From baseline to after the intervention total RNA and rRNA per unit muscle weight increased on average by ~20-27% and ~25-57%, respectively (Figure 2).
+The five-session-RT intervention led to increases in total RNA and rRNA per unit muscle weight in both glucose and placebo, on average by ~20-27% and ~25-57%, respectively (Figure 2).
 However, RT with glucose did not induce increased accumulation of total RNA (Figure 2A, mean difference `r total_rna_stats`) or rRNA  (Figure 2B; 47S, `r stat_rrna$trg_47s`; 18S, `r stat_rrna$trg_18s`; 28S, `r stat_rrna$trg_28s`; 5.8S, `r stat_rrna$trg_5.8s`; 5S, `r stat_rrna$trg_5s`) compared to RT with placebo.
@@
 
-Overall, RT led to robust increases from before to after the intervention in all measured proteins (Figure 3A).
+The five-session-RT intervention led to increases in all measured proteins in both glucose and placebo (Figure 3A).
 RT with glucose resulted in estimated levels of c-Myc, UBF and RPS6 being `r stat_west$cmyc`, `r stat_west$ubf` and `r stat_west$rps6`% lower compared to placebo, respectively, without showing statistical significance (*p* = `r min(p_west)`-`r max(p_west)`; Figure 3A).
-There was a linear relationship between UBF and total RNA, where an increase of 1 SD unit of UBF equated to a $\sim$ `r corr_incr`% increase in total RNA (*p* = `r corrp`; Figure 3C).
+
+Both baseline and trained state data of total RNA showed a linear relationship with UBF, where $\sim$ `r corr_incr`% increase in total RNA corresponded to 1 standard deviation unit increase in UBF (*p* = `r corrp`; Figure 3C).
 
@@
 Similarly, RT with glucose in general did not affect markers of skeletal muscle functionality such as strength and recovery, or total training session volume.
-At one time point, RT with glucose led to less reduction in strength compared to placebo.
+Towards the end of the intervention, RT with glucose led to less reduction in strength compared to placebo.
 Levels of plasma glucose and c-peptide were significantly higher before, during and after the RT sessions with glucose.

### 2023-10-09 manuscript2.RMD
updates

@@
 Similarly, RT with glucose in general did not affect markers of skeletal muscle functionality such as strength and recovery, or total training session volume.
-Towards the end of the intervention, RT with glucose led to less reduction in strength compared to placebo.
-Levels of plasma glucose and c-peptide were significantly higher before, during and after the RT sessions with glucose.
-Importantly, there were no differences in daily macronutrient intake on pairwise consecutive days.
+Towards the end of the intervention, RT with glucose led to less reduction in strength compared to placebo, which may point to an accumulated effect of the glucose condition.
+As by design the levels of plasma glucose and c-peptide were significantly higher before, during and after the RT sessions with the glucose condition compared to placebo, and there were no differences in the daily macro nutrient intake between conditions on consecutive days, thus enabling an effective design to study the biological and functional effects of the glucose condition.
 These findings suggest that the effects of high vs.
-low glucose/glucose starvation conditions on rDNA transcription initiation observed in previous *in vitro* studies [@mariappan_ribosomal_2011; @tanaka_mild_2015] are not translatable to acute effects in human skeletal muscle *in vivo*.
+low glucose/glucose starvation conditions on rDNA transcription initiation observed in previous *in vitro* studies [@mariappan_ribosomal_2011; @tanaka_mild_2015] are not translatable to acute effects in human skeletal muscle *in vivo*, with a design as the present study.
 Notably, the present study was specifically designed to investigate differences in pre to post changes within participants and is as such also quite limited to this perspective.
@@
 Hence, it seems quite reasonable to observe similar changes in these three proteins.
-The linear relationship found exclusively between UBF content and total RNA accumulation, and neither c-Myc nor rpS6, supports the importance and specificity of UBF content in human skeletal muscle ribosome biogenesis.
+The linear relationship found exclusively between UBF content and total RNA levels, and neither c-Myc nor rpS6, supports the importance and specificity of UBF content in human skeletal muscle ribosome biogenesis.
 While this is not a novel finding, it has only recently been observed in human skeletal muscle following RT [@hammarstrom_ribosome_2022].

### 2023-10-10 manuscript2.RMD
updates

--- 2023-10-06 manuscript2.RMD
@@
 Responses to systematic resistance training (RT) vary widely between individuals, with as much as 10-15% showing impaired skeletal muscle growth in response to standardized training interventions [@alvarez_interindividual_2018; @mann_high_2014; @thalacker-mercer_cluster_2013].
-While there certainly may be differences in the genetic predisposition in terms of readiness to adapt to exercise [@thalacker-mercer_cluster_2013], there are also recent studies showing that epigenetic factors such as ribosomal DNA (rDNA) availability may also impact the ability to mount an anabolic response to RT [@figueiredo_genetic_2021].
-Indeed, the ability to synthesise ribosomes during a period of RT is connected to the magnitude of RT-induced responses in an individual [@hammarstrom_benefits_2020].
+There may be differences in the genetic predisposition in terms of readiness to adapt to exercise [@thalacker-mercer_cluster_2013], however, the internal physiological milieu seem favourably shaped for muscle growth by factors such as different types of training stress and nutrients [@figueiredo_genetic_2021; @sebata_control_2018; @thalacker-mercer_cluster_2013].
+Indeed, training stress and nutrients impact the ability to synthesis ribosomes, which in turn is connected to the magnitude of RT-induced responses in an individual [@hammarstrom_benefits_2020; @kusnadi_regulation_2015; @sebata_control_2018].
 For instance, increasing training volume generally induces greater ribosome biogenesis and thus, greater benefits of RT [@hammarstrom_benefits_2020;@krieger_single_2009;@schoenfeld_strength_2017].
-However, as evident from Hammarström et al.
-(2020), not all participants experienced increased muscle mass and -strength despite increasing training volume.
+Still, as evident from Hammarström et al.
+(2020), not all participants experience increased muscle mass and -strength despite increasing training volume.
 Therefore, other means than modification of RT variables alone seem necessary to optimise individual responses to RT.

### 2023-10-10 manuscript2.RMD
updates

@@
 
-**Conclusion:** Glucose supplementation before and after resistance training did not augment ribosomal accumulation during two weeks of heavy-load resistance training in moderately trained young adults.
+**Conclusion:** Glucose ingestion before and after resistance training did not augment ribosomal RNA accumulation during two weeks of heavy-load resistance training in moderately trained young adults.
 
@@
 
-The purpose of this investigation was to test the hypothesis that glucose supplementation before and after five RT-sessions over a period of 12 days will potentiate RT-associated accumulation of markers of ribosomal abundance following five RT sessions.
+Therefore, the purpose of this investigation was to test the hypothesis that glucose supplementation before and after five RT-sessions over a period of 12 days will potentiate RT-associated accumulation of markers of ribosomal abundance following five RT sessions.
 Secondly, we aimed to describe the association between changes in total RNA abundance and UBF in human skeletal muscle.
@@
 RT with glucose resulted in estimated levels of c-Myc, UBF and RPS6 being `r stat_west$cmyc`, `r stat_west$ubf` and `r stat_west$rps6`% lower compared to placebo, respectively, without showing statistical significance (*p* = `r min(p_west)`-`r max(p_west)`; Figure 3A).
-
 Both baseline and trained state data of total RNA showed a linear relationship with UBF, where $\sim$ `r corr_incr`% increase in total RNA corresponded to 1 standard deviation unit increase in UBF (*p* = `r corrp`; Figure 3C).
@@
 These findings suggest that the effects of high vs.
-low glucose/glucose starvation conditions on rDNA transcription initiation observed in previous *in vitro* studies [@mariappan_ribosomal_2011; @tanaka_mild_2015] are not translatable to acute effects in human skeletal muscle *in vivo*, with a design as the present study.
-Notably, the present study was specifically designed to investigate differences in pre to post changes within participants and is as such also quite limited to this perspective.
-Hence, the observed effects of time serve as supplementary to the efficiency of the design, without a negative control group to compare with.
-
-To our knowledge, no previous study has investigated the effects of glucose compared to placebo ingestion with RT on ribosome biogenesis in human skeletal muscle previously, making our hypotheses exploratory and comparisons few.
+low glucose/glucose starvation conditions on rDNA transcription initiation observed in previous *in vitro* studies [@mariappan_ribosomal_2011; @tanaka_mild_2015] are not translatable to acute effects in human skeletal muscle *in vivo*, with a design like the present study.
+
+To our knowledge, no previous study has investigated the effects of glucose compared to placebo ingestion with RT on ribosome biogenesis in human skeletal muscle, making our hypotheses exploratory and comparisons few.
 *In vitro* studies on mice glomerular epithelial cells [@mariappan_ribosomal_2011] and human breast cancer cells [@tanaka_mild_2015] have observed glucose-induced UBF-dependent augmentation of ribosome biogenesis [@mariappan_ribosomal_2011] and reduced rRNA transcription by glucose starvation [@tanaka_mild_2015].
@@
 normal plasma glucose levels, to investigate the effect of glucose *per se* and not intracellular energy status.
-Thus, while glucose ingestion presumably is important in supplying energy for growth-inducing processes such as ribosome biogenesis [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @sebata_control_2018] there is no apparent effect of ingesting glucose *per se* on markers of ribosome biogenesis.
-
-In the present study, despite not measuring the activity in central pathways mediating anabolic signalling (mTORC1, ERK1/2), analyses of the downstream target UBF and the ribosomal protein S6 (rpS6), as well as the general transcription factor c-Myc, further supported observations from the RNA data.
+Thus, while glucose ingestion presumably is important for supplying energy for growth-inducing processes such as ribosome biogenesis [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @sebata_control_2018] there is no apparent effect of ingesting glucose *per se* on markers of ribosome biogenesis.
+
+In the present study, despite not measuring the activity in central pathways mediating anabolic signalling (mTORC1, ERK1/2), analyses of the downstream target UBF and the ribosomal protein S6 (rpS6), as well as the general transcription factor c-Myc, further supported observations from the RNA data as there was no difference between the glucose and placebo conditions.
 UBF has previously been described as a master regulator of rDNA transcription *in vitro* [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @russell_rna-polymerase-i-directed_2005], while rpS6 is proposed as a valid and reliable means to measure ribosome biogenesis [@chaillou_hypoxia_2012; @nakada_correlation_2016].
@@
 
+*Limitations and strengths*
+Notably, the present study was specifically designed to investigate differences in pre to post changes within participants and is as such also quite limited to this perspective.
+Hence, the observed effects of time serve as supplementary to the efficiency of the design, without a negative control group to compare with.
+
 *Conclusion*
 
-In conclusion, ingestion of glucose immediately before and after resistance exercise training does not acutely augment ribosome biogenesis after two weeks of heavy resistance exercise training, in moderately trained young adults.
-Neither does it acutely enhance muscular performance during the exercise period or after five sessions, nor recovery within 23 hours of the last session.
-The observations in total RNA, rRNA and rDNA transcription initiation-associated proteins support the key role of UBF in ribosome biogenesis regulation in human skeletal muscle following resistance training.
-If future investigations are to be made on this topic, a greater sample size coupled with a negative control group and analyses of the intramuscular glycogen storage may provide higher-resolution results.
+In conclusion, ingestion of glucose immediately before and after five heavy-load resistance training sessions conducted over a period of 12 days did not augment accumulation of ribosomal RNA, in moderately trained young adults compared to ingestion of placebo.
+Glucose ingestion did not affect muscular performance over the course of the study, nor did it affect muscular performance measured 30 minutes, 2 hours or 23 hours after the last session.
+Glucose ingestion was associated with less reduction in muscular performance 23 hours after the fifth training session, thus we cannot rule out a possible accumulated effect of ingestion glucose compared to placebo.
+The relationship between baseline and trained state data of total RNA and UBF levels do support the key role of UBF in ribosomal biogenesis regulation in human skeletal muscle following resistance training.
+Future investigations should aim for a greater sample size over a longer time period, to investigate the potential accumulated effects of glucose ingestion, as well as analyses of the intramuscular glycogen storage to provide higher-resolution results.
 

### 2023-10-11 manuscript2.RMD
updates

@@
 The linear relationship found exclusively between UBF content and total RNA levels, and neither c-Myc nor rpS6, supports the importance and specificity of UBF content in human skeletal muscle ribosome biogenesis.
-While this is not a novel finding, it has only recently been observed in human skeletal muscle following RT [@hammarstrom_ribosome_2022].
+While this is not a novel finding, it has only recently been observed in human skeletal muscle following a period RT [@hammarstrom_ribosome_2022].
 As such, UBF also seems to respond to mechanical loading in human muscle cells in line with responses seen in cell cultures and synergist ablation models [@mariappan_ribosomal_2011; @von_walden_mechanical_2012], in addition to the previously reported response to hormonal, nutritional, and cellular energy signals [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @russell_rna-polymerase-i-directed_2005].
 
-As in the biological data, there was generally no difference in the change of strength as an index between the conditions.
-Generally, skeletal muscle strength decreased from baseline to after the fifth and -sixth RT sessions, with a significantly less reduction in glucose compared to placebo at the measurement after five sessions.
-This could suggest that the RT protocol fatigued the participants, as both training frequency and intensity were quite high.
-Another promising explanation might be the biphasic recovery pattern, as described by Raastad & Hallén [@raastad_recovery_2000], where the participants experienced a rapid recovery during the initial 11 hours after exercise, followed by a levelling off or drop until 22 hours after exercise.
+As in the biological data, glucose ingestion had no effect throughout the intervention on muscular strength compared to placebo.
+Generally, skeletal muscle performance measured by a strength index and used as a proxy marker for muscular recovery, decreased from baseline to after the condition in both conditions.
+Glucose ingestion was associated with a less reduction in muscular strength after five RT sessions compared to placebo, which may point towards a beneficial accumulated effect where the heavy-load RT fatigued the participants but glucose ingestion may have improved muscular recovery [@mul_exercise_2015; @sebata_control_2018].
+However, glucose did not improve muscular performance/recovery acutely following one RT session compared to placebo, measure 30 minutes, 2 hours and 23 hours after the sixth and final training session.
+As such, the potential accumulated effect observed after five RT sessions did not extend to after session six.
+Therefore, we cant rule out that there may be a long-term effect of ingesting glucose when exercising heavy-load resistance training, which would require a longer intervention period than that of the present study.
+Another possible explanation to the decrease in muscular performance during the intervention might be the biphasic recovery pattern, as described by Raastad & Hallén [@raastad_recovery_2000], where the participants experienced a rapid recovery during the initial 11 hours after exercise, followed by a levelling off or drop until 22 hours after exercise.
 Herein, inflammation and phagocytic activity were proposed to be involved in the performance drop between 11-22 hours [@raastad_recovery_2000].
@@
 Further, strength testing during the intervention was conducted 23 hours after RT, meaning that the biphasic recovery may have also influenced these tests.
-However, this does not explain the difference between conditions at observed after five RT session, or the drop in strength from after four RT sessions to after five RT sessions.
+However, this does not explain the difference between conditions observed after five RT session, or the drop in strength from after four RT sessions to after five RT sessions.
 A possible argument could be that exercising without glucose may have caused more stress compared to exercising with glucose, as glucose is the preferred energy source during strenuous exercise [@mul_exercise_2015], thus increasing performance with glucose compared to placebo.
@@
 Therefore, discussing the potential effect of differences in metabolic stress between conditions would only be speculation, however probable.
+
+*Limitations and Strengths*
+The present study was designed specifically to investigate the acute biological and functional effects of ingesting glucose compared to placebo, with unilateral training and testing to reduce biological variation between participants in our analyses [KILDE].
+Further, to ensure that the legs exercised under the same conditions apart from glucose/placebo during exercise, macronutrients, the time of day and test/training personell were standardized for each participant on consequtive pairwise days.
+
 Lastly, these inconclusive observations may also reflect our limited sample size, despite using a within-subjects design.

### 2023-10-11 manuscript2.RMD
updates

@@
 High glucose treatment may augment ribosomal biogenesis, as evident from *in vitro* studies, and may thus act to potentiate resistance training-induced adaptations *in vivo*.
-Hoever, this remains largely unexplored.
+However, this remains largely unexplored.
 
@@
 The present study was designed specifically to investigate the acute biological and functional effects of ingesting glucose compared to placebo, with unilateral training and testing to reduce biological variation between participants in our analyses [KILDE].
-Further, to ensure that the legs exercised under the same conditions apart from glucose/placebo during exercise, macronutrients, the time of day and test/training personell were standardized for each participant on consequtive pairwise days.
-
-Lastly, these inconclusive observations may also reflect our limited sample size, despite using a within-subjects design.
-Initially, we met our minimum limit of 16 participants.
-Unfortunately, three dropped out during the intervention, leaving our statistical analyses slightly underpowered according to the a priori calculation.
-Moreover, we did not measure the intramuscular glycogen storage, representing another limitation to our design.
-Hence, we cannot know for certain whether there were actual differences within the exercising skeletal muscles between the glucose and placebo conditions.
-The participants did indeed match their macronutrients on pairwise consecutive days, and the plasma glucose measurements did not show any differences in blood sugar until ingestion of placebo/glucose.
-Thus, it may stand to reason that the participants had similar levels of intramuscular levels before ingesting any supplement and that the glucose supplement led to higher glucose availability during exercise, as by design.
-However, even if the leg exercising with glucose supplement did in fact take up the extra glucose, it did not seem to affect any of our main outcome measures.
+Further, to ensure that the legs exercised under the same conditions apart from glucose/placebo during exercise, macro nutrients, the time of day and test/training personnel were standardized for each participant on pairwise consecutive days.
+Every day, the participants showed up in an overnight fasted state, and ingested only either protein and glucose or protein and placebo before and during exercise.
+As evident from the blood data, glucose ingestion thus led to higher levels both of plasma glucose and serum c-peptide compared to placebo, together with the aforementioned standardisations enabling analyses of the effect of ingesting glucose on total- and specific RNA levels and proteins, as well as muscular performance, within-participant.
+The present study also had a few limitations; firstly, the sample size was smaller than expected and planned for.
+Initially the minimum limit of 16 participants, according to the a priori power calculation, was met.
+However, three dropped out during the intervention, possibly leaving our statistical analyses slightly underpowered.
+Moreover, the intramuscular glycogen stores were not measured, hence it cannot be determined whether the glucose ingestion in fact increased intramuscular glucose.
+Though it may be a reasonable assumption that the glucose ingestion in this design did lead to increased intramuscular glucose and thus energy levels, it cannot be excluded that the participants intramuscular glycogen stores were topped up from the previous day, and as such, ingesting more glucose had no further benefit.
+Nevertheless, even if the leg exercising with glucose supplement did in fact take up the extra glucose, it did not seem to affect any of our main outcome measures.
 

### 2023-10-12 manuscript2.RMD
updates

@@
 Furthermore, energy availability is a decisive factor in the *de novo* synthesis of ribosomes [@kusnadi_regulation_2015;@moss_housekeeper_2007;@sebata_control_2018] which in turn determines muscle growth by increasing the muscle's translational capacity [@figueiredo_regulation_2019;@hammarstrom_benefits_2020;@sebata_control_2018;@stec_ribosome_2016;@von_walden_ribosome_2019].
-In addition, insulin may have anabolic effects with elevated amino acids, and has a clear role in reduction of muscle protein breakdown independent of amino acid availability [@abdulla_role_2016].
+In addition, insulin may have anabolic effects with elevated amino acids, and has a clear role in reduction of muscle protein breakdown independent of amino acid availability [@abdulla_role_2016; @hillier_physiological_2000].
 Therefore, investigating the effects of combined RT and glucose ingestion may provide valuable insight into the potential additive effect of glucose and RT on ribosome biogenesis.
@@
 
-To our knowledge, no previous study has investigated the effects of glucose compared to placebo ingestion with RT on ribosome biogenesis in human skeletal muscle, making our hypotheses exploratory and comparisons few.
-*In vitro* studies on mice glomerular epithelial cells [@mariappan_ribosomal_2011] and human breast cancer cells [@tanaka_mild_2015] have observed glucose-induced UBF-dependent augmentation of ribosome biogenesis [@mariappan_ribosomal_2011] and reduced rRNA transcription by glucose starvation [@tanaka_mild_2015].
-These findings do imply an important role for glucose in the initiation of rDNA transcription, the rate-limiting step of ribosome biogenesis [@mariappan_ribosomal_2011; @tanaka_mild_2015].
-In addition, Hillier et al.
-found physiological hyperinsulinemia to stimulate p70S6K phosphorylation in human skeletal muscle [@hillier_physiological_2000].
-Moreover, robust increases in markers of ribosome biogenesis such as 47S pre-rRNA and mature rRNA can be expected after an acute bout of RT [@figueiredo_impact_2016] or a short period of RT [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022].
-Therefore, we hypothesized that if glucose can, directly or indirectly via insulin, stimulate the initiation of rDNA transcription, five sessions of RT with glucose compared to placebo would elicit this effect.
-Further, glucose ingestion increases energy levels, leading us to hypothesize that high glucose vs.
-placebo condition would stimulate energy-sensitive pathways such as mTORC1, PIH1, extracellular signal-regulated kinase 1/2 (ERK1/2), AMP-dependent protein kinase (AMPK) and Sirtuin 1 (SIRT1) [@kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012].
+The observations of the present study, from human skeletal muscle, do not support previous indications from non-human skeletal muscle [@mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012].
+Allthough we observed a linear relationship between baseline and trained state UBF and total RNA levels, a UBF-dependent augmentation of ribosome biogenesis did not seem to be induced by glucose ingestion in the present study, as compared to in mice glomerular epithelial cells [@mariappan_ribosomal_2011] and in human breast cancer cells [@tanaka_mild_2015].
+Neither do our findings indicate that insulin *per se*, at least in physiological levels, potentiates accumulation of total RNA through p70S6K stimulation as observed with hyperinsulinemia in human skeletal muscle [@hillier_physiological_2000].
+Robust increases in markers of ribosome biogenesis such as 47S pre-rRNA and mature rRNA are expected to occur after an acute bout of RT [@figueiredo_impact_2016], as well as after a short period of RT [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022].
+Therefore, in the present study, it was expected that eventual benefits of ingesting glucose compared to placebo with RT would be measurable after either 5 training sessions, either due to glucose-induced stimulation of energy-sensitive pathways such as mTORC1, PIH1, extracellular signal-regulated kinase 1/2 (ERK1/2), AMP-dependent protein kinase (AMPK) or Sirtuin 1 (SIRT1) [@kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012].
 Despite previously reported upregulation in PIC assembly due to high-glucose mediated mTORC1, ERK1/2 and PIH1 or low-glucose mediated AMPK and SIRT1 activation  [@hoppe_amp-activated_2009; @kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012], the present study showed no signs of such effects of glucose vs.
 placebo conditions.
-Resistance training, irrespective of condition, yielded a robust accumulation of total RNA and expression of rRNA, in line with previous observations [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022].
-Due to the exploratory nature of our hypotheses, this was hardly a very surprising result.
-Indeed, previous studies have used cell cultures from yeast [@zhai_human_2012], rodents [@hoppe_amp-activated_2009; @mariappan_ribosomal_2011] or human cells [@tanaka_mild_2015] and are, as such, not directly comparable to human skeletal muscle cells.
 Importantly, previous studies investigated high vs.
@@
 normal plasma glucose levels, to investigate the effect of glucose *per se* and not intracellular energy status.
-Thus, while glucose ingestion presumably is important for supplying energy for growth-inducing processes such as ribosome biogenesis [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @sebata_control_2018] there is no apparent effect of ingesting glucose *per se* on markers of ribosome biogenesis.
+Thus, while glucose ingestion presumably is important for supplying energy for growth-inducing processes such as ribosome biogenesis [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @sebata_control_2018] there is no apparent effect of ingesting glucose *per se* on markers of ribosome biogenesis, during 12 days of heavy-load RT.
+Further, previous studies have used cell cultures from yeast [@zhai_human_2012], rodents [@hoppe_amp-activated_2009; @mariappan_ribosomal_2011] or human breast cancer cells [@tanaka_mild_2015] and are, as such, not directly comparable to human skeletal muscle cells.
+Nevertheless, resistance training irrespective of condition yielded a robust accumulation of total RNA and expression of rRNA, in line with previous observation for heavy-load RT [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022].
 
@@
 The linear relationship found exclusively between UBF content and total RNA levels, and neither c-Myc nor rpS6, supports the importance and specificity of UBF content in human skeletal muscle ribosome biogenesis.
-While this is not a novel finding, it has only recently been observed in human skeletal muscle following a period RT [@hammarstrom_ribosome_2022].
+While this is not a novel finding, it was only recently observed in human skeletal muscle following a period of RT [@hammarstrom_ribosome_2022].
 As such, UBF also seems to respond to mechanical loading in human muscle cells in line with responses seen in cell cultures and synergist ablation models [@mariappan_ribosomal_2011; @von_walden_mechanical_2012], in addition to the previously reported response to hormonal, nutritional, and cellular energy signals [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @russell_rna-polymerase-i-directed_2005].

### 2023-10-12 manuscript2.RMD
proofreading

@@
 
-**Methods:** Sixteen healthy, moderately trained individuals (male/female, n = 9/7; age, 24.1 (3.3)), participated in a within-participant crossover trial, conducting unilateral resistance training (leg press and knee extension, 3 sets of 10 repetitions maximum) with pre- and post-exercise ingestion of either glucose (3x30g, 90g total) or placebo supplement (Stevia rebaudiana extract, 3x0.3g, 0.9g total), together with protein (2x25g, 50g total), on alternating days over a period of twelve days; six sessions per condition, resistance training was performed in an otherwise fasted state, as morning sessions.
+**Methods:** Sixteen healthy, moderately trained individuals (male/female, n = 9/7; age, 24.1 (3.3)), participated in a within-participant crossover trial, conducting unilateral resistance training (leg press and knee extension, 3 sets of 10 repetitions maximum) with pre- and post-exercise ingestion of either glucose (3x30g, 90g total) or placebo supplement (Stevia rebaudiana extract, 3x0.3g, 0.9g total), together with protein (2x25g, 50g total), on alternating days for twelve days; six sessions per condition, resistance training was performed in an otherwise fasted state, as morning sessions.
 Micro biopsies were sampled from m.
@@
 **Results:** Glucose ingestion did not have beneficial effects on resistance training-induced increases of ribosomal content (mean difference 7.6% [-7.2, 24.9], *p* = 0.34; ribosomal RNA, 47S/18S/28S/5.8S/5S, range 7.6%-37.9%, *p* = 0.40-0.98) or levels of relevant transcriptional or translational regulators (c-MYK/UBF/rpS6, *p* = 0.094-0.292).
-Of note, both baseline and trained state data of total RNA showed a linear relationship with UBF; ∼14% increase in total RNA corresponded to 1 standard deviation unit increase in UBF (*p* = 0.003).
+Of note, both baseline and trained state data of total RNA showed a linear relationship with UBF; a ∼14% increase in total RNA corresponded to 1 standard deviation unit increase in UBF (*p* = 0.003).
 
@@
 
-Data was collected at the sports science lab at the Inland Norway University of Applied Sciences, with all analyses performed on site, except serum blood analyses, which were performed at Innlandet Hospital Trust.
+Data was collected at the sports science lab at the Inland Norway University of Applied Sciences, with all analyses performed on-site, except serum blood analyses, which were performed at Innlandet Hospital Trust.
 S.E., H.H.
@@
 Responses to systematic resistance training (RT) vary widely between individuals, with as much as 10-15% showing impaired skeletal muscle growth in response to standardized training interventions [@alvarez_interindividual_2018; @mann_high_2014; @thalacker-mercer_cluster_2013].
-There may be differences in the genetic predisposition in terms of readiness to adapt to exercise [@thalacker-mercer_cluster_2013], however, the internal physiological milieu seem favourably shaped for muscle growth by factors such as different types of training stress and nutrients [@figueiredo_genetic_2021; @sebata_control_2018; @thalacker-mercer_cluster_2013].
-Indeed, training stress and nutrients impact the ability to synthesis ribosomes, which in turn is connected to the magnitude of RT-induced responses in an individual [@hammarstrom_benefits_2020; @kusnadi_regulation_2015; @sebata_control_2018].
+There may be differences in the genetic predisposition in terms of readiness to adapt to exercise [@thalacker-mercer_cluster_2013], however, the internal physiological milieu seems favourably shaped for muscle growth by factors such as different types of training stress and nutrients [@figueiredo_genetic_2021; @sebata_control_2018; @thalacker-mercer_cluster_2013].
+Indeed, training stress and nutrients impact the ability to synthesise ribosomes, which in turn is connected to the magnitude of RT-induced responses in an individual [@hammarstrom_benefits_2020; @kusnadi_regulation_2015; @sebata_control_2018].
 For instance, increasing training volume generally induces greater ribosome biogenesis and thus, greater benefits of RT [@hammarstrom_benefits_2020;@krieger_single_2009;@schoenfeld_strength_2017].
@@
 
-Nutritional supplements such as protein and creatine is often used to optimise RT adaptations [@cermak_protein_2012;@lanhers_creatine_2015;@lanhers_creatine_2017;@morton_systematic_2018].
+Nutritional supplements such as protein and creatine are often used to optimise RT adaptations [@cermak_protein_2012;@lanhers_creatine_2015;@lanhers_creatine_2017;@morton_systematic_2018].
 However, it remains equivocal if other nutritional adjuvants such as glucose can increase the efficacy of RT [@tezze_anabolic_2023].
@@
 Furthermore, energy availability is a decisive factor in the *de novo* synthesis of ribosomes [@kusnadi_regulation_2015;@moss_housekeeper_2007;@sebata_control_2018] which in turn determines muscle growth by increasing the muscle's translational capacity [@figueiredo_regulation_2019;@hammarstrom_benefits_2020;@sebata_control_2018;@stec_ribosome_2016;@von_walden_ribosome_2019].
-In addition, insulin may have anabolic effects with elevated amino acids, and has a clear role in reduction of muscle protein breakdown independent of amino acid availability [@abdulla_role_2016; @hillier_physiological_2000].
+In addition, insulin may have anabolic effects with elevated amino acids, and has a clear role in the reduction of muscle protein breakdown independent of amino acid availability [@abdulla_role_2016; @hillier_physiological_2000].
 Therefore, investigating the effects of combined RT and glucose ingestion may provide valuable insight into the potential additive effect of glucose and RT on ribosome biogenesis.
@@
 
-Therefore, the purpose of this investigation was to test the hypothesis that glucose supplementation before and after five RT-sessions over a period of 12 days will potentiate RT-associated accumulation of markers of ribosomal abundance following five RT sessions.
+Therefore, the purpose of this investigation was to test the hypothesis that glucose supplementation before and after five RT sessions over a period of 12 days will potentiate RT-associated accumulation of markers of ribosomal abundance following five RT sessions.
 Secondly, we aimed to describe the association between changes in total RNA abundance and UBF in human skeletal muscle.
@@
 Glucose (Glucosum monohydricum, Merck KGaA, Darmstadt, Germany) and placebo (Steviosa, Soma Nordic AS, Oslo, Norway) were masked by mixing with 300ml Fun Light (Orkla, Oslo, Norway).
-A blinded taste test revealed that the participants were not able to disclose the contents of the provided boluses (30g glucosum monohydricum vs.
+A blinded taste test revealed that the participants were not able to disclose the contents of the provided boluses (30g glucoseum monohydricum vs.
 ~0.3g Stevia rebaudiana extract).
@@
 To enable analysis of target gene expression per unit tissue weight [@ellefsen_improved_2008; @ellefsen_reliable_2014], an exogenous RNA control (Lambda, λ polyA External Standard Kit, Takara Bio Inc., Shiga, Japan) was added at a fixed amount to each sample (0.04 ng ml^-1^ of TRIzol reagent).
-For assessment of RNA content and purity, RNA was eluted in TE-buffer (1:2) and assessed via spectrophotometry.
+For assessment of RNA content and purity, RNA was eluted in TE buffer (1:2) and assessed via spectrophotometry.
 All samples had a 260 nm to 280 nm ratio > 1.9.
@@
 
-Glucose ingestion compared to placebo before and after RT-sessions did not lead to a difference in mean change of total training session volume, with an average increase of 18 and 16% from session 1 to session 6, respectively (*p* > 0.05).
+Glucose ingestion compared to placebo before and after RT sessions did not lead to a difference in mean change of total training session volume, with an average increase of 18 and 16% from session 1 to session 6, respectively (*p* > 0.05).
 There were no differences in mean macronutrient intake (protein, fat, carbohydrates) or total calorie intake between glucose and placebo on pairwise consecutive days (*p* > 0.05 for all).
@@
 
-Glucose ingestion before and after RT led to increases in plasma glucose levels compare to baseline by 38% immediately before RT (Figure 1B, 0 min), by 31% during RT (Figure 1B, 15 min) and by 32% immediately after RT (Figure 1B, 30 min; all : *p* < 0.001), with no changes observed in the placebo condition (*p* > 0.05).
+Glucose ingestion before and after RT led to increases in plasma glucose levels compared to baseline by 38% immediately before RT (Figure 1B, 0 min), by 31% during RT (Figure 1B, 15 min) and by 32% immediately after RT (Figure 1B, 30 min; all : *p* < 0.001), with no changes observed in the placebo condition (*p* > 0.05).
 Compared to the placebo condition, ingestion of glucose increased plasma glucose levels by 36% immediately before RT (Figure 1B, 0 min), by 27% during RT (Figure 1B, 15 min) and by 28% immediately after RT (Figure 1B, 30 min; all : *p* < 0.001).
@@
 Towards the end of the intervention, RT with glucose led to less reduction in strength compared to placebo, which may point to an accumulated effect of the glucose condition.
-As by design the levels of plasma glucose and c-peptide were significantly higher before, during and after the RT sessions with the glucose condition compared to placebo, and there were no differences in the daily macro nutrient intake between conditions on consecutive days, thus enabling an effective design to study the biological and functional effects of the glucose condition.
+As by design, the levels of plasma glucose and c-peptide were significantly higher before, during and after the RT sessions with the glucose condition compared to placebo, and there were no differences in the daily macronutrient intake between conditions on consecutive days, thus enabling an effective design to study the biological and functional effects of the glucose condition.
 These findings suggest that the effects of high vs.
@@
 The observations of the present study, from human skeletal muscle, do not support previous indications from non-human skeletal muscle [@mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012].
-Allthough we observed a linear relationship between baseline and trained state UBF and total RNA levels, a UBF-dependent augmentation of ribosome biogenesis did not seem to be induced by glucose ingestion in the present study, as compared to in mice glomerular epithelial cells [@mariappan_ribosomal_2011] and in human breast cancer cells [@tanaka_mild_2015].
-Neither do our findings indicate that insulin *per se*, at least in physiological levels, potentiates accumulation of total RNA through p70S6K stimulation as observed with hyperinsulinemia in human skeletal muscle [@hillier_physiological_2000].
+Although we observed a linear relationship between baseline and trained state UBF and total RNA levels, a UBF-dependent augmentation of ribosome biogenesis did not seem to be induced by glucose ingestion in the present study, as compared to in mice glomerular epithelial cells [@mariappan_ribosomal_2011] and in human breast cancer cells [@tanaka_mild_2015].
+Neither do our findings indicate that insulin *per se*, at least at physiological levels, potentiates accumulation of total RNA through p70S6K stimulation as observed with hyperinsulinemia in human skeletal muscle [@hillier_physiological_2000].
 Robust increases in markers of ribosome biogenesis such as 47S pre-rRNA and mature rRNA are expected to occur after an acute bout of RT [@figueiredo_impact_2016], as well as after a short period of RT [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022].
@@
 Further, previous studies have used cell cultures from yeast [@zhai_human_2012], rodents [@hoppe_amp-activated_2009; @mariappan_ribosomal_2011] or human breast cancer cells [@tanaka_mild_2015] and are, as such, not directly comparable to human skeletal muscle cells.
-Nevertheless, resistance training irrespective of condition yielded a robust accumulation of total RNA and expression of rRNA, in line with previous observation for heavy-load RT [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022].
+Nevertheless, resistance training irrespective of condition yielded a robust accumulation of total RNA and expression of rRNA, in line with previous observations for heavy-load RT [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022].
 
@@
 As in the biological data, glucose ingestion had no effect throughout the intervention on muscular strength compared to placebo.
-Generally, skeletal muscle performance measured by a strength index and used as a proxy marker for muscular recovery, decreased from baseline to after the condition in both conditions.
+Generally, skeletal muscle performance measured by a strength index, used as a proxy marker for muscular recovery, decreased from baseline to after the condition in both conditions.
 Glucose ingestion was associated with a less reduction in muscular strength after five RT sessions compared to placebo, which may point towards a beneficial accumulated effect where the heavy-load RT fatigued the participants but glucose ingestion may have improved muscular recovery [@mul_exercise_2015; @sebata_control_2018].
-However, glucose did not improve muscular performance/recovery acutely following one RT session compared to placebo, measure 30 minutes, 2 hours and 23 hours after the sixth and final training session.
+However, glucose did not improve muscular performance/recovery acutely following one RT session compared to placebo, measured 30 minutes, 2 hours and 23 hours after the sixth and final training session.
 As such, the potential accumulated effect observed after five RT sessions did not extend to after session six.
-Therefore, we cant rule out that there may be a long-term effect of ingesting glucose when exercising heavy-load resistance training, which would require a longer intervention period than that of the present study.
-Another possible explanation to the decrease in muscular performance during the intervention might be the biphasic recovery pattern, as described by Raastad & Hallén [@raastad_recovery_2000], where the participants experienced a rapid recovery during the initial 11 hours after exercise, followed by a levelling off or drop until 22 hours after exercise.
+Therefore, we cant rule out a possible long-term effect of ingesting glucose when exercising heavy-load resistance training, which would require a longer intervention period than that of the present study.
+Another possible explanation for the decrease in muscular performance during the intervention might be the biphasic recovery pattern, as described by Raastad & Hallén [@raastad_recovery_2000], where the participants experienced a rapid recovery during the initial 11 hours after exercise, followed by a levelling off or drop until 22 hours after exercise.
 Herein, inflammation and phagocytic activity were proposed to be involved in the performance drop between 11-22 hours [@raastad_recovery_2000].
@@
 Further, strength testing during the intervention was conducted 23 hours after RT, meaning that the biphasic recovery may have also influenced these tests.
-However, this does not explain the difference between conditions observed after five RT session, or the drop in strength from after four RT sessions to after five RT sessions.
+However, this does not explain the difference between conditions observed after five RT sessions, or the drop in strength from after four RT sessions to after five RT sessions.
 A possible argument could be that exercising without glucose may have caused more stress compared to exercising with glucose, as glucose is the preferred energy source during strenuous exercise [@mul_exercise_2015], thus increasing performance with glucose compared to placebo.
@@
 *Limitations and Strengths*
+
 The present study was designed specifically to investigate the acute biological and functional effects of ingesting glucose compared to placebo, with unilateral training and testing to reduce biological variation between participants in our analyses [KILDE].
-Further, to ensure that the legs exercised under the same conditions apart from glucose/placebo during exercise, macro nutrients, the time of day and test/training personnel were standardized for each participant on pairwise consecutive days.
-Every day, the participants showed up in an overnight fasted state, and ingested only either protein and glucose or protein and placebo before and during exercise.
+Further, to ensure that the legs were exercised under the same conditions apart from glucose/placebo during exercise, macronutrients, the time of day and test/training personnel were standardized for each participant on pairwise consecutive days.
+Every day, the participants showed up in an overnight fasted state and ingested only either protein and glucose or protein and placebo before and during exercise.
 As evident from the blood data, glucose ingestion thus led to higher levels both of plasma glucose and serum c-peptide compared to placebo, together with the aforementioned standardisations enabling analyses of the effect of ingesting glucose on total- and specific RNA levels and proteins, as well as muscular performance, within-participant.
@@
 However, three dropped out during the intervention, possibly leaving our statistical analyses slightly underpowered.
-Moreover, the intramuscular glycogen stores were not measured, hence it cannot be determined whether the glucose ingestion in fact increased intramuscular glucose.
-Though it may be a reasonable assumption that the glucose ingestion in this design did lead to increased intramuscular glucose and thus energy levels, it cannot be excluded that the participants intramuscular glycogen stores were topped up from the previous day, and as such, ingesting more glucose had no further benefit.
-Nevertheless, even if the leg exercising with glucose supplement did in fact take up the extra glucose, it did not seem to affect any of our main outcome measures.
-
-*Limitations and strengths*
-Notably, the present study was specifically designed to investigate differences in pre to post changes within participants and is as such also quite limited to this perspective.
-Hence, the observed effects of time serve as supplementary to the efficiency of the design, without a negative control group to compare with.
+Moreover, the intramuscular glycogen stores were not measured, hence it can't be determined whether the glucose ingestion increased intramuscular glucose.
+Though it may be a reasonable assumption that the glucose ingestion in this design did lead to increased intramuscular glucose and thus energy levels, it cannot be excluded that the participants' intramuscular glycogen stores were topped up from the previous day, and as such, ingesting more glucose had no further benefit.
+Nevertheless, even if the skeletal muscle of the leg exercising with glucose supplement took up the extra glucose, it did not seem to affect any of our main outcome measures.
 
@@
 In conclusion, ingestion of glucose immediately before and after five heavy-load resistance training sessions conducted over a period of 12 days did not augment accumulation of ribosomal RNA, in moderately trained young adults compared to ingestion of placebo.
-Glucose ingestion did not affect muscular performance over the course of the study, nor did it affect muscular performance measured 30 minutes, 2 hours or 23 hours after the last session.
-Glucose ingestion was associated with less reduction in muscular performance 23 hours after the fifth training session, thus we cannot rule out a possible accumulated effect of ingestion glucose compared to placebo.
+Glucose ingestion did not affect muscular performance throughout the study, nor did it affect muscular performance measured 30 minutes, 2 hours or 23 hours after the last session.
+Glucose ingestion was associated with less reduction in muscular performance 23 hours after the fifth training session, thus we can't rule out a possible accumulated effect of ingestion glucose compared to placebo.
 The relationship between baseline and trained state data of total RNA and UBF levels do support the key role of UBF in ribosomal biogenesis regulation in human skeletal muscle following resistance training.
@@
 vastus lateralis at baseline (Day 1 leg 1, Day 2 leg 2), and after five RT sessions (Day 11 leg 1, Day 12 leg 2).
-Blood for measurement of plasma glucose and -c-peptide was sampled at baseline (Day 1), and during post testing (Day 11 leg 1, Day 12 leg 2), via finger draws and venous blood samples.
+Blood for measuring plasma glucose and -c-peptide was sampled at baseline (Day 1), and during post-testing (Day 11 leg 1, Day 12 leg 2), via finger draws and venous blood samples.
 Skeletal muscle strength was measured as peak torque in unilateral isometric and isokinetic (at 60 and 240 degrees per second) knee extension before, multiple times during, and after five and six sessions.
 B and c) Changes in plasma glucose (b, mmol/L) and serum c-peptide levels (c, pmol/L).
-Glucose levels in blood was measured via finger draws 120 (-120), 90 (-90), and 30min (-30) before RT, immediately before RT (0), during RT (15), immediately after RT (30) and 2hrs after RT (120).
+Glucose levels in blood were measured via finger draws 120 (-120), 90 (-90), and 30min (-30) before RT, immediately before RT (0), during RT (15), immediately after RT (30) and 2hrs after RT (120).
 C-peptide levels were measured simultaneously to these finger draws, except for 90min before and during RT.
@@
 Total RNA and rRNA changes were calculated as log-fold change score per mg wet muscle weight.
-Mean change scores of the duplicates were calculated and transformed to the log-scale before modelling, then reverse-transformed for figure visualisation.
+Mean change scores of the duplicates were calculated and transformed to the log scale before modelling, then reverse-transformed for figure visualisation.
 Values are estimated marginal means fold change per leg per supplement ± 95% CI.
@@
 Values are estimated fold change per condition with 95% CI, Glucose n = 13 and placebo n = 13.
-A linear relationship was shown between between total RNA (ng x mg) and UBF levels (SD units) while controlling for time.
+A linear relationship was shown between total RNA (ng x mg) and UBF levels (SD units) while controlling for time.
 Total RNA was normalized by wet muscle weight, and UBF was normalized by a pooled sample used on each gel.

### 2023-10-12 manuscript2.RMD
proofreading

@@
 
-The present study was designed specifically to investigate the acute biological and functional effects of ingesting glucose compared to placebo, with unilateral training and testing to reduce biological variation between participants in our analyses [KILDE].
+The present study was designed specifically to investigate the acute biological and functional effects of ingesting glucose compared to placebo, with unilateral training and testing to reduce biological variation between participants in our analyses [@macinnis_investigating_2017].
 Further, to ensure that the legs were exercised under the same conditions apart from glucose/placebo during exercise, macronutrients, the time of day and test/training personnel were standardized for each participant on pairwise consecutive days.

### 2023-10-13 manuscript2.RMD
proofreading

@@
 
-Therefore, the purpose of this investigation was to test the hypothesis that glucose supplementation before and after five RT sessions over a period of 12 days will potentiate RT-associated accumulation of markers of ribosomal abundance following five RT sessions.
+Therefore, the purpose of this investigation was to test the hypothesis that glucose supplementation before and after five RT sessions over 12 days will potentiate RT-associated accumulation of markers of ribosomal abundance following five RT sessions.
 Secondly, we aimed to describe the association between changes in total RNA abundance and UBF in human skeletal muscle.
@@
 Neither do our findings indicate that insulin *per se*, at least at physiological levels, potentiates accumulation of total RNA through p70S6K stimulation as observed with hyperinsulinemia in human skeletal muscle [@hillier_physiological_2000].
-Robust increases in markers of ribosome biogenesis such as 47S pre-rRNA and mature rRNA are expected to occur after an acute bout of RT [@figueiredo_impact_2016], as well as after a short period of RT [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022].
+Increases in markers of ribosome biogenesis such as 47S pre-rRNA and mature rRNA are expected to occur after a single session of RT [@figueiredo_impact_2016], as well as after a short period of RT [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022].
 Therefore, in the present study, it was expected that eventual benefits of ingesting glucose compared to placebo with RT would be measurable after either 5 training sessions, either due to glucose-induced stimulation of energy-sensitive pathways such as mTORC1, PIH1, extracellular signal-regulated kinase 1/2 (ERK1/2), AMP-dependent protein kinase (AMPK) or Sirtuin 1 (SIRT1) [@kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012].
@@
 As such, the potential accumulated effect observed after five RT sessions did not extend to after session six.
-Therefore, we cant rule out a possible long-term effect of ingesting glucose when exercising heavy-load resistance training, which would require a longer intervention period than that of the present study.
+Therefore, we can't rule out a possible long-term effect of ingesting glucose when exercising heavy-load resistance training, which would require a longer intervention period than that of the present study.
 Another possible explanation for the decrease in muscular performance during the intervention might be the biphasic recovery pattern, as described by Raastad & Hallén [@raastad_recovery_2000], where the participants experienced a rapid recovery during the initial 11 hours after exercise, followed by a levelling off or drop until 22 hours after exercise.
@@
 
-In conclusion, ingestion of glucose immediately before and after five heavy-load resistance training sessions conducted over a period of 12 days did not augment accumulation of ribosomal RNA, in moderately trained young adults compared to ingestion of placebo.
+In conclusion, ingestion of glucose immediately before and after five heavy-load resistance training sessions conducted over 12 days did not augment accumulation of ribosomal RNA, in moderately trained young adults compared to ingestion of placebo.
 Glucose ingestion did not affect muscular performance throughout the study, nor did it affect muscular performance measured 30 minutes, 2 hours or 23 hours after the last session.
-Glucose ingestion was associated with less reduction in muscular performance 23 hours after the fifth training session, thus we can't rule out a possible accumulated effect of ingestion glucose compared to placebo.
+Glucose ingestion was associated with less reduction in muscular performance 23 hours after the fifth training session, thus we can't rule out a possible accumulated effect of ingesting glucose compared to placebo.
 The relationship between baseline and trained state data of total RNA and UBF levels do support the key role of UBF in ribosomal biogenesis regulation in human skeletal muscle following resistance training.

### 2023-10-13 manuscript2.RMD
proofreading

@@
 
-<div custom-style="correspondence_style">
+::: {custom-style="correspondence_style"}
 ^1^Section for Health and Exercise Physiology, Department of Public Health and Sport Sciences, Inland Norway University of Applied Sciences, Lillehammer, Norway
+:::
+
+::: {custom-style="correspondence_style"}
+Correspondance: Kristian Lian ([kristian.lian\@inn.no](mailto:kristian.lian@inn.no){.email}), Stian Ellefsen ([stian.ellefsen\@inn.no](mailto:stian.ellefsen@inn.no){.email})
+:::
+
+ORCID: Kristian Lian <https://orcid.org/0000-0002-5169-5960>, Daniel Hammarström <https://orcid.org/0000-0001-8360-2100>, Håvard Hamarsland <https://orcid.org/0000-0002-7343-2774>, Stian Ellefsen <https://orcid.org/0000-0003-0467-2018>
+
 </div>
 
-<div custom-style="correspondence_style">
-Correspondance: Kristian Lian (kristian.lian@inn.no), Stian Ellefsen (stian.ellefsen@inn.no)
-</div>
-
-ORCID: Kristian Lian https://orcid.org/0000-0002-5169-5960, Daniel Hammarström https://orcid.org/0000-0001-8360-2100, Håvard Hamarsland https://orcid.org/0000-0002-7343-2774, Stian Ellefsen https://orcid.org/0000-0003-0467-2018
-</div>
-
-<div custom-style="abstract_style">
+::: {custom-style="abstract_style"}
 ## Abstract
@@
 <!-- To do: Fix tables, add outlier figures to the readme -->
+
 ```{r}
@@
 
-**Conclusion:** Glucose ingestion before and after resistance training did not augment ribosomal RNA accumulation during two weeks of heavy-load resistance training in moderately trained young adults.
+**Conclusion:** Glucose ingestion before and after resistance training did not augment ribosomal RNA accumulation during twelve days of heavy-load resistance training in moderately trained young adults.
 
@@
 ### Acknowledgements
+
 Gratitude is due, to Henriette Spilhaug Bollandsås, Nora Lerdalen Bonsak and Thea Bøhn Nyløkken for their dedication and hard work during the training intervention.
@@
 **Abbreviations**
+
 ```{r abbreviation table}
@@
 ```
-
-</div>
+:::
 
 \pagebreak
+
 ## Introduction
@@
 Indeed, training stress and nutrients impact the ability to synthesise ribosomes, which in turn is connected to the magnitude of RT-induced responses in an individual [@hammarstrom_benefits_2020; @kusnadi_regulation_2015; @sebata_control_2018].
-For instance, increasing training volume generally induces greater ribosome biogenesis and thus, greater benefits of RT [@hammarstrom_benefits_2020;@krieger_single_2009;@schoenfeld_strength_2017].
+For instance, increasing training volume generally induces greater ribosome biogenesis and thus, greater benefits of RT [@hammarstrom_benefits_2020; @krieger_single_2009; @schoenfeld_strength_2017].
 Still, as evident from Hammarström et al.
@@
 
-Nutritional supplements such as protein and creatine are often used to optimise RT adaptations [@cermak_protein_2012;@lanhers_creatine_2015;@lanhers_creatine_2017;@morton_systematic_2018].
+Nutritional supplements such as protein and creatine are often used to optimise RT adaptations [@cermak_protein_2012; @lanhers_creatine_2015; @lanhers_creatine_2017; @morton_systematic_2018].
 However, it remains equivocal if other nutritional adjuvants such as glucose can increase the efficacy of RT [@tezze_anabolic_2023].
-This is surprising since glucose is the preferred energy substrate of the contracting skeletal muscle during strenuous exercise and a major energy supplier to cells via adenosine triphosphate (ATP) synthesis [@mul_exercise_2015;@sebata_control_2018].
-Furthermore, energy availability is a decisive factor in the *de novo* synthesis of ribosomes [@kusnadi_regulation_2015;@moss_housekeeper_2007;@sebata_control_2018] which in turn determines muscle growth by increasing the muscle's translational capacity [@figueiredo_regulation_2019;@hammarstrom_benefits_2020;@sebata_control_2018;@stec_ribosome_2016;@von_walden_ribosome_2019].
+This is surprising since glucose is the preferred energy substrate of the contracting skeletal muscle during strenuous exercise and a major energy supplier to cells via adenosine triphosphate (ATP) synthesis [@mul_exercise_2015; @sebata_control_2018].
+Furthermore, energy availability is a decisive factor in the *de novo* synthesis of ribosomes [@kusnadi_regulation_2015; @moss_housekeeper_2007; @sebata_control_2018] which in turn determines muscle growth by increasing the muscle's translational capacity [@figueiredo_regulation_2019; @hammarstrom_benefits_2020; @sebata_control_2018; @stec_ribosome_2016; @von_walden_ribosome_2019].
 In addition, insulin may have anabolic effects with elevated amino acids, and has a clear role in the reduction of muscle protein breakdown independent of amino acid availability [@abdulla_role_2016; @hillier_physiological_2000].
@@
 
-Ribosomal biogenesis and content seem to be a prerequisite for skeletal muscle growth, and transcription of ribosomal RNA (rRNA) by RNA Polymerase I is considered the rate-limiting step in synthesising new ribosomes [@moss_promotion_1995].
-Multiple signalling pathways converge to regulate rRNA transcription, including the mammalian target of rapamycin (mTORC1) signal-transduction pathway and c-Myc [@kusnadi_regulation_2015;@von_walden_ribosome_2019;@sebata_control_2018; @west_acute_2016; @mori_c-myc_2021].
-mTOR receives input from growth factors, hormones, mechanical loading, and nutrients to balance protein synthesis through multiple mechanisms based on cellular energy levels [@hoppe_amp-activated_2009].
-Importantly, mTOR and its downstream target ribosomal protein S6 kinase beta-1 (S6K1) serves as a direct mediator of insulin signalling in skeletal muscle [@hillier_physiological_2000].
-The mTORC1 pathway contributes to ribosome biogenesis through the regulation of the translation of ribosomal proteins and to forming of the preinitiation complex (PIC) that marks the initiation of rRNA transcription [@figueiredo_regulation_2019; @von_walden_ribosome_2019].
-Parallel to mTORC1 [@mori_c-myc_2021; @west_acute_2016], c-Myc increases ribosomal biogenesis directly through transcriptional control of the upstream binding factor (UBF) [@sanij_ubf_2008; @poortinga_c-myc_2011].
+Ribosomal biogenesis and content seem to be a prerequisite for skeletal muscle growth, and transcription of ribosomal ribonucleic acid (rRNA) by RNA Polymerase I is considered the rate-limiting step in synthesising new ribosomes [@moss_promotion_1995].
+Multiple signalling pathways converge to regulate rRNA transcription, including c-Myc and the mammalian target of rapamycin complex 1 (mTORC1) signal-transduction pathway [@kusnadi_regulation_2015; @von_walden_ribosome_2019; @sebata_control_2018; @west_acute_2016; @mori_c-myc_2021].
+The general transcription factor c-Myc increases ribosomal biogenesis directly through transcriptional control of the upstream binding factor (UBF) [@sanij_ubf_2008; @poortinga_c-myc_2011; @mori_c-myc_2021; @west_acute_2016].
 Indeed, UBF phosphorylation, which is required for interaction with the ribosomal deoxyribonucleic acid (rDNA) promoter, seems to be increased by high-glucose treatment in a mTORC1-dependent manner (rapamycin sensitive) in kidney glomerular epithelial cells [@mariappan_ribosomal_2011].
 Independent of UBF and mTORC1, high glucose was shown to lead to chromatin remodelling, which promoted rRNA transcription in cell cultures [@zhai_human_2012].
-Together these observations indicate a potential role of glucose in positively affecting muscle rRNA synthesis.
-<!-- sørg for at disse to avsnittene henger godt sammen!-->
-There have been multiple studies recently suggesting translational capacity to be as important, if not more important than, translational efficiency concerning long-term skeletal muscle adaptations to RT [@hammarstrom_benefits_2020;@hammarstrom_ribosome_2022;@figueiredo_revisiting_2019].
-Previously, *in vitro* studies have shown UBF and c-Myc to be important factors in the PIC, c-Myc as a general transcription factor and UBF as a specific transcription factor for rDNA transcription initiation [@mariappan_ribosomal_2011;@von_walden_ribosome_2019].
+The mTORC1 pathway receives input from growth factors, hormones, mechanical loading, and nutrients to balance protein synthesis through multiple mechanisms based on cellular energy levels [@hoppe_amp-activated_2009] and contributes to ribosome biogenesis through the forming of the preinitiation complex (PIC) that marks the initiation of rRNA transcription and through the regulation of ribosomal protein translation [@figueiredo_regulation_2019; @von_walden_ribosome_2019].
+Moreover, mTORC1 and its downstream target ribosomal protein S6 kinase beta-1 (S6K1) serves as direct mediators of insulin signalling in skeletal muscle [@hillier_physiological_2000].
+Together these observations indicate a potential role of glucose in positively affecting human skeletal muscle rRNA synthesis, through potentiating transcription or translation of ribosomal RNA and increasing the translational capacity of skeletal muscle cells [@sebata_control_2018; @zhai_human_2012; @hillier_physiological_2000; @hoppe_amp-activated_2009].
+
+There have been multiple studies recently suggesting translational capacity to be as important, if not more important than, translational efficiency concerning long-term skeletal muscle adaptations to RT [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022; @figueiredo_revisiting_2019].
+Previously, *in vitro* studies have shown UBF and c-Myc to be important factors in the PIC, c-Myc as a general transcription factor and UBF as a specific transcription factor for rDNA transcription initiation [@mariappan_ribosomal_2011; @von_walden_ribosome_2019].
 Coupled with the recent observations by Hammarström et al.
-[@hammarstrom_benefits_2020;@hammarstrom_ribosome_2022], understanding underlying factors regulating and affecting ribosome biogenesis seems key to furthering our understanding of the optimisation of RT to individual phenotypes.
+[@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022], understanding underlying factors regulating and affecting ribosome biogenesis seems key to furthering our understanding of the optimisation of RT to individual phenotypes.
 
@@
 
-The study was designed as a 12-day double-blinded randomized controlled trial, with an alternating unilateral RT protocol (Figure 1A).
+The study was designed as a 12-day double-blinded placebo-controlled simultaneous crossover trial, with an alternating unilateral RT protocol (Figure 1A).
 Participants were randomly allocated to exercise one leg with a glucose condition and one leg with a placebo condition (Figure 1A).
@@
 A blinded taste test revealed that the participants were not able to disclose the contents of the provided boluses (30g glucoseum monohydricum vs.
-~0.3g Stevia rebaudiana extract).
+\~0.3g Stevia rebaudiana extract).
 To ensure equal conditions during training sessions and strength testing, participants exercised and tested at the same time of day, +/- 1hr with the same supervisor on pairwise consecutive days (i.e.
@@
 All participants completed six RT sessions with glucose and six with placebo, allowing a within-subjects analysis of the effects of glucose ingestion before and after RT.
-Data from the first five RT sessions was used to investigate main outcome measures (total RNA, rRNA and protein) and leg muscle strength, whereas data from the sixth RT session was used to explore secondary outcomes (muscular recovery, plasma glucose and -c-peptide levels).
+Data from the first five RT sessions was used to investigate main outcome measures (total RNA, rRNA and protein) and leg muscle strength, whereas data from the sixth RT session was used to explore secondary outcomes (muscular recovery, plasma glucose and serum c-peptide levels).
 Participants were asked to avoid resistance- or high-intensity training of the legs from Day -7 (Figure 1A) and onwards, until completion of the intervention and post-testing, to ensure the reliability of pre-intervention strength data and minimal interference from external exercise sources.
@@
 0g glucose), immediately before RT (0900hrs, 30g vs.
-0g glucose), and immediately after RT (~0930hrs, 30g vs.
+0g glucose), and immediately after RT (\~0930hrs, 30g vs.
 0g glucose).
-Whey Protein Isolate (Proteinfabrikken, Stokke, Norway) was ingested 2hrs before RT (0700hrs) and immediately after RT (~0930hrs), in boluses of 25 g mixed with 150 ml water.
+Whey Protein Isolate (Proteinfabrikken, Stokke, Norway) was ingested 2hrs before RT (0700hrs) and immediately after RT (\~0930hrs), in boluses of 25 g mixed with 150 ml water.
 In the afternoon (1800hrs-1900hrs, period II) participants ingested glucose or placebo (3x30g vs.
@@
 Isokinetic peak torque was measured concentrically from 90 to 0 degrees knee angle (extended knee was set to 0 degrees) at angular velocities of 60- and 240 degrees per second, 2x3 repetitions each, with the first set of each exercise as a sub-maximal warm-up.
-Isometric knee extensor peak torque was measured at a kneen angle of 60 degrees, for a maximum of 10 seconds and two repetitions per test.
+Isometric knee extensor peak torque was measured at a knee angle of 60 degrees, for a maximum of 10 seconds and two repetitions per test.
 The isometric tests were ended when the participants reached a plateau or peak torque development decreased, which on average occurred between 2-4 seconds into the test.
@@
 The participants performed a general warm-up with 10 minutes of cycling on an indoor exercise bicycle.
-A protocol consisting of 1x10, 1x6 and 1x3 repetitions with a load equivalent to ~50-75% of assumed max repetitions, was used as a specific warm-up before each of the tests.
+A protocol consisting of 1x10, 1x6 and 1x3 repetitions with a load equivalent to \~50-75% of assumed max repetitions, was used as a specific warm-up before each of the tests.
 All positions were controlled and recorded at the first 1RM test and repeated for the RT sessions.
@@
 As a general warm-up, the participants cycled on an indoor exercise bicycle for 5-10 minutes.
-In addition, before the respective exercises, two 10-repetition warm-up sets were completed at ~50% and ~70% of 10RM.
+In addition, before the respective exercises, two 10-repetition warm-up sets were completed at \~50% and \~70% of 10RM.
 To ensure adequate exercise stimulation throughout the intervention, the exercise load was increased the following set if the participants lifted more than 12 repetitions, as a progressive loading strategy.
@@
 Aliquots were snap-frozen in isopentane (-80°C) and stored at -80°C until further analyses.
-Muscle biopsies were collected at four time points: I/II) Before the intervention (Figure 1A, 2hrs before training, Day 1 = leg 1, Day 2 = leg 2), and III/IV) after the fifth RT session, two hours before the sixth RT session (Figure 1A, Day 11 = leg 1, Day 12 = leg 2).
+Muscle biopsies were collected at four time points: I/II) Before the intervention (Figure 1A, 2hrs before training, Day 1 = leg 1, Day 2 = leg 2), and III/IV) approximately 22 hours after the fifth RT session, two hours before the sixth RT session (Figure 1A, Day 11 = leg 1, Day 12 = leg 2).
 At each time point, two samples were taken from the same incision.
@@
 One capillary blood sample was collected on day 1 (Figure 1A) to serve as a baseline.
-On days 11 and 12 (Figure 1A), capillary blood samples were collected seven times: I) Immediately before protein ingestion (0700hrs) II) 45 minutes after protein ingestion (0745hrs) III) 1.5 hours after protein ingestion (0830hrs, i.e., immediately before GLU/PLAC intake), IV) 2 hours after protein ingestion (0900hrs, i.e., immediately before training), IV) in the middle of RT (~0915hrs), V) immediately after training (~0930hrs), and VI) 2 hours after completion of training (~1130hrs).
+On days 11 and 12 (Figure 1A), capillary blood samples were collected seven times: I) Immediately before protein ingestion (0700hrs) II) 45 minutes after protein ingestion (0745hrs) III) 1.5 hours after protein ingestion (0830hrs, i.e., immediately before GLU/PLAC intake), IV) 2 hours after protein ingestion (0900hrs, i.e., immediately before training), IV) in the middle of RT (\~0915hrs), V) immediately after training (\~0930hrs), and VI) 2 hours after completion of training (\~1130hrs).
 Capillary blood samples were analysed with in-house equipment (BIOSEN C-Line, EKF diagnostic GmbH, Barleben).
@@
 ### Total RNA extraction and real-time reverse transcription polymerase chain reaction
+
 ```{r totrna-extraction}
@@
 Two muscle biopsy aliquots were used for total RNA extraction per leg, resulting in a total of eight RNA samples per participant.
-Total RNA was extracted using TRIzol with muscle tissue homogenised using 0.5mm RNase-free Zirconium beads (~50 ul; Next Advanced, Averill Park, NY, USA) and mechanical agitation (Bullet Blender, Next Advanced, Averill Park, NY, USA).
+Total RNA was extracted using TRIzol with muscle tissue homogenised using 0.5mm RNase-free Zirconium beads (\~50 ul; Next Advanced, Averill Park, NY, USA) and mechanical agitation (Bullet Blender, Next Advanced, Averill Park, NY, USA).
 Chloroform (Sigma-Aldrich, Oslo, Norway) was used for phase separation, and the RNA pellet was precipitated with isopropanol (VWR International, Oslo, Norway).
@@
 For assessment of RNA content and purity, RNA was eluted in TE buffer (1:2) and assessed via spectrophotometry.
-All samples had a 260 nm to 280 nm ratio > 1.9.
+All samples had a 260 nm to 280 nm ratio \> 1.9.
 The RNA stock was stored at -80°C until further analyses.
@@
 
-Total protein was extracted using the Minute Total Protein Extraction Kit for Muscles (Invent Biotechnology), according to the manufacturer’s protocol, optimised for our lab.
+Total protein was extracted using the Minute Total Protein Extraction Kit for Muscles (Invent Biotechnology), according to the manufacturer's protocol, optimised for our lab.
 Wet muscle was freeze-dried for 24 hours and dissected before extraction.
@@
 A priori power calculations showed that 20 participants would grant a statistical power of 80% (alpha = 0.05), accounting for an expected dropout of 20%.
-This power calculation was based on an assumption that the effects of glucose ingestion on ribosome biogenesis (measured as within-participant differences in total RNA) may equate to the effects of increasing RT volume from low to moderate [@hammarstrom_benefits_2020].
-Total RNA, protein and qPCR data were analyzed by mixed-effects models with fixed effects included as *supplement $\times$ time*.
+This power calculation was based on an assumption that the effects of glucose ingestion on total RNA accumulation and rRNA expression may equate to the effects of increasing RT volume from low to moderate [@hammarstrom_benefits_2020].
+Total RNA, protein and qPCR data were analyzed by mixed-effects models with fixed effects included as *supplement* $\times$ time.
 To decrease the risk of Type I errors, random effects were selected from step-wise elimination of terms from the most complex structure (random slopes for time and supplement and their interaction) to less complex.
-The most complex random effects structure that converged was chosen as the final model[@matuschek_balancing_2017].
-Plasma glucose, c-peptide, training volume, and the strength index were analyzed by multiple time-point log-fold change score comparisons, using a mixed-effects model with baseline values, time and supplement, and the interaction between time and supplement as fixed effects.
+The most complex random effect structure that converged was chosen as the final model [@matuschek_balancing_2017].
+Plasma glucose, serum c-peptide, training volume, and the strength index were analyzed by multiple time-point log-fold change score comparisons, using a mixed-effects model with baseline values, time and supplement, and the interaction between time and supplement as fixed effects.
 These data only supported a random intercept per participant.
-The linear mixed-effects models were fitted with the lmer function from the lme4 package using the lmerTest package to procure p-values (Satterthwaite's method for approximating degrees of freedom)[@bates_fitting_2014], written for R[@Rcore].
+The linear mixed-effects models were fitted with the lmer function from the lme4 package using the lmerTest package to procure p-values (Satterthwaite's method for approximating degrees of freedom) [@bates_fitting_2014], written for R [@Rcore].
 Log-transformed values were expressed as fold changes in visualisations.
-Descriptive data are presented as mean and (SD).
+Descriptive data are presented as mean and standard deviation (SD).
 Inferential statistics are presented as means with 95% confidence intervals and *p*-values unless otherwise stated.
-*p* < 0.05 was considered statistically significant.
+*p* \< 0.05 was considered statistically significant.
 
@@
 
-Glucose ingestion compared to placebo before and after RT sessions did not lead to a difference in mean change of total training session volume, with an average increase of 18 and 16% from session 1 to session 6, respectively (*p* > 0.05).
-There were no differences in mean macronutrient intake (protein, fat, carbohydrates) or total calorie intake between glucose and placebo on pairwise consecutive days (*p* > 0.05 for all).
+Glucose ingestion before and after RT sessions did not lead to a higher mean change of total training session volume, with a mean increase of 17% in both the glucose condition (pre: 5262 ± 1799kg, post: 6319 ± 2256kg, *p* > 0.05) and the placebo condition (pre: 5351 ± 1615kg, post: 6438 ± 2092, *p* > 0.05) from session 1 to session 6, respectively.
+There were no differences in mean macronutrient intake (protein, fat, carbohydrates) or total calorie intake between glucose and placebo on pairwise consecutive days (*p* \> 0.05 for all, raw data and table available at GitHub repository).
 
@@
 
-Glucose ingestion before and after RT led to increases in plasma glucose levels compared to baseline by 38% immediately before RT (Figure 1B, 0 min), by 31% during RT (Figure 1B, 15 min) and by 32% immediately after RT (Figure 1B, 30 min; all : *p* < 0.001), with no changes observed in the placebo condition (*p* > 0.05).
-Compared to the placebo condition, ingestion of glucose increased plasma glucose levels by 36% immediately before RT (Figure 1B, 0 min), by 27% during RT (Figure 1B, 15 min) and by 28% immediately after RT (Figure 1B, 30 min; all : *p* < 0.001).
+Glucose ingestion before and after RT led to increases in plasma glucose levels compared to baseline by 38% immediately before RT (Figure 1B, 0 min), by 31% during RT (Figure 1B, 15 min) and by 32% immediately after RT (Figure 1B, 30 min; all : *p* \< 0.001), with no changes observed in the placebo condition (*p* \> 0.05).
+Compared to the placebo condition, ingestion of glucose increased plasma glucose levels by 36% immediately before RT (Figure 1B, 0 min), by 27% during RT (Figure 1B, 15 min) and by 28% immediately after RT (Figure 1B, 30 min; all : *p* \< 0.001).
 Two hours after the RT session, glucose ingestion was associated with 12% lower plasma glucose levels compared to baseline, and 8% lower compared to placebo (Figure 1B, 270min: *p* = `r p.270`).
 
-Glucose ingestion before and after RT led to increases in levels of c-peptide compared to baseline, by 95% immediately before (Figure 1C, 0 min) and 87% after RT (Figure 1C, 30 min; both *p* < 0.001), with no changes observed with the placebo condition (*p* > 0.05).
-Compared to the placebo condition, ingestion of glucose increased levels of c-peptide by 85% immediately before (Figure 1C, 0 min) and 85% after RT (Figure 1C, 30 min; both *p* < 0.001).
+Glucose ingestion before and after RT led to increases in levels of c-peptide compared to baseline, by 95% immediately before (Figure 1C, 0 min) and 87% after RT (Figure 1C, 30 min; both *p* \< 0.001), with no changes observed with the placebo condition (*p* \> 0.05).
+Compared to the placebo condition, ingestion of glucose increased levels of c-peptide by 85% immediately before (Figure 1C, 0 min) and 85% after RT (Figure 1C, 30 min; both *p* \< 0.001).
 
@@
 
-Glucose ingestion before and after RT sessions did not generally enhance skeletal muscle recovery throughout the intervention, neither during the intervention in a rested state (Figure 1D, 23hrs after exercise; Post 2RT: *p* = `r pdiff.c2`, Post 4RT: *p* = `r pdiff.c3`), nor acutely after the sixth and final RT-session (30min post 6RT: *p* = `r pdiff.c5`, 2h post 6RT: *p* = `r pdiff.c6`) or in a rested state after the sixth RT session (23h post 6RT: *p* = `r pdiff.c7`).
+Glucose ingestion before and after RT sessions did not generally improve skeletal muscle recovery throughout the intervention, neither during the intervention in a rested state (Figure 1D, 23hrs after exercise; Post 2RT: *p* = `r pdiff.c2`, Post 4RT: *p* = `r pdiff.c3`), nor acutely after the sixth and final RT-session (30min post 6RT: *p* = `r pdiff.c5`, 2h post 6RT: *p* = `r pdiff.c6`) or in a rested state after the sixth RT session (23h post 6RT: *p* = `r pdiff.c7`).
 Glucose ingestion before and after RT compared to placebo was associated with a 7% less reduction in strength after five RT sessions (Figure 1D, *p* = `r pdiff.c4`).
@@
 
-The five-session-RT intervention led to increases in total RNA and rRNA per unit muscle weight in both glucose and placebo, on average by ~20-27% and ~25-57%, respectively (Figure 2).
+The five-session-RT intervention led to increases in total RNA and rRNA per unit muscle weight both in the glucose and the placebo condition, on average by \~20-27% and \~25-57%, respectively (Figure 2).
 However, RT with glucose did not induce increased accumulation of total RNA (Figure 2A, mean difference `r total_rna_stats`) or rRNA (Figure 2B; 47S, `r stat_rrna$trg_47s`; 18S, `r stat_rrna$trg_18s`; 28S, `r stat_rrna$trg_28s`; 5.8S, `r stat_rrna$trg_5.8s`; 5S, `r stat_rrna$trg_5s`) compared to RT with placebo.
@@
 
-The five-session-RT intervention led to increases in all measured proteins in both glucose and placebo (Figure 3A).
+The five-session-RT intervention led to increases in all measured proteins both in the glucose and the placebo condition (Figure 3A).
 RT with glucose resulted in estimated levels of c-Myc, UBF and RPS6 being `r stat_west$cmyc`, `r stat_west$ubf` and `r stat_west$rps6`% lower compared to placebo, respectively, without showing statistical significance (*p* = `r min(p_west)`-`r max(p_west)`; Figure 3A).
@@
 
-Complete data sets and scripts are downloadable here; https://github.com/Kristianlian/ribose-paper
+Complete data sets and scripts are downloadable here; <https://github.com/Kristianlian/ribose-paper>
 
@@
 
-Extended methods can be found here; https://github.com/Kristianlian/ribose-paper
+Extended methods can be found here; <https://github.com/Kristianlian/ribose-paper>
 
 \pagebreak
+
 ## References
 
-<div id="refs"></div>
+::: {#refs}
+:::
 
@@
 Values are presented as changes in estimated marginal means ± 95% confidence intervals (CI).
-* = *p* < 0.05 between groups.
+\* = *p* \< 0.05 between groups.
 Glucose n = 13, placebo n = 13.
@@
 60º sec-1 = 60 degrees per second movement speed (isokinetic), 240º sec-1 = 240 degrees per second movement speed (isokinetic), 0º sec-1 = 0 degrees per second movement speed (isometric).
+
 ```{r baseline table, echo = FALSE}
@@
 
-**Table 2.
-Primer sequences:** Values of Ct are means ± SD.
-rRNA = ribosomal RNA, E = primer efficiency.
-Average cycle thresholds (Ct) and priming efficiencies were calculated from all qPCR reactions.
+**Table 2. Primer sequences:** Values of Ct are means ± SD. rRNA = ribosomal RNA, E = primer efficiency. Average cycle thresholds (Ct) and priming efficiencies were calculated from all qPCR reactions.
+
 ```{r primer table, echo = FALSE}

### 2023-10-13 manuscript2.RMD
final push

@@
 
-The observations of the present study, from human skeletal muscle, do not support previous indications from non-human skeletal muscle [@mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012].
+The observations of the present study from human skeletal muscle do not support previous indications from non-human skeletal muscle [@mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012].
 Although we observed a linear relationship between baseline and trained state UBF and total RNA levels, a UBF-dependent augmentation of ribosome biogenesis did not seem to be induced by glucose ingestion in the present study, as compared to in mice glomerular epithelial cells [@mariappan_ribosomal_2011] and in human breast cancer cells [@tanaka_mild_2015].
@@
 Increases in markers of ribosome biogenesis such as 47S pre-rRNA and mature rRNA are expected to occur after a single session of RT [@figueiredo_impact_2016], as well as after a short period of RT [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022].
-Therefore, in the present study, it was expected that eventual benefits of ingesting glucose compared to placebo with RT would be measurable after either 5 training sessions, either due to glucose-induced stimulation of energy-sensitive pathways such as mTORC1, PIH1, extracellular signal-regulated kinase 1/2 (ERK1/2), AMP-dependent protein kinase (AMPK) or Sirtuin 1 (SIRT1) [@kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012].
+Therefore, in the present study, it was expected that eventual benefits of ingesting glucose compared to placebo with RT would be measurable after 5 training sessions, either due to glucose-induced stimulation of energy-sensitive pathways such as mTORC1, PIH1, extracellular signal-regulated kinase 1/2 (ERK1/2), AMP-dependent protein kinase (AMPK) or Sirtuin 1 (SIRT1) [@kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012].
 Despite previously reported upregulation in PIC assembly due to high-glucose mediated mTORC1, ERK1/2 and PIH1 or low-glucose mediated AMPK and SIRT1 activation [@hoppe_amp-activated_2009; @kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012], the present study showed no signs of such effects of glucose vs.
@@
 The present study was designed specifically to investigate the acute biological and functional effects of ingesting glucose compared to placebo, with unilateral training and testing to reduce biological variation between participants in our analyses [@macinnis_investigating_2017].
-Further, to ensure that the legs were exercised under the same conditions apart from glucose/placebo during exercise, macronutrients, the time of day and test/training personnel were standardized for each participant on pairwise consecutive days.
+Further, to ensure that the legs were exercised under the same conditions apart from glucose/placebo during exercise, macronutrients, the time of day and test/training personnel were standardized for each participant on pairwise consecutive days[@halperin_threats_2015].
 Every day, the participants showed up in an overnight fasted state and ingested only either protein and glucose or protein and placebo before and during exercise.

### 2023-10-13 resources/older manuscript versions/first_draft.RMD
cleanup

--- 2023-09-01 jap.manuscript.RMD
@@
 ---
-title: ""
-author: ""
+title: "Glucose ingestion before and after resistance training does not augment ribosome biogenesis in healthy moderately trained young adults"
+author: "Kristian Lian^1^, Sara Christine Olsen Moen^1^, Håvard Hamarsland^1^, Daniel Hammarström^1^, Stian Ellefsen^1^"
 bibliography: resources/ribose-paper.bib
@@
 
-#### Glucose ingestion before and after resistance training does not augment ribosome biogenesis in healthy moderately trained young adults
-
-###### Kristian Lian^1^, Daniel Hammarström^1^, Håvard Hamarsland^1^, Sara Christine Olsen Moen^1^, Stian Ellefsen^1^
-
+### Affiliations
 <div custom-style="correspondence_style">
-^1^Section for Health and Exercise Physiology, Department of Public Health and Sport Sciences, Inland Norway University of Applied Sciences, Lillehammer, Norway
+^1^ Section for Health and Exercise Physiology, Department of Public Health and Sport Sciences, Inland Norway University of Applied Sciences, Lillehammer, Norway.
 </div>
@@
 <div custom-style="correspondence_style">
-Correspondance: Kristian Lian (kristian.lian@inn.no), Stian Ellefsen (stian.ellefsen@inn.no)
+### Correspondance
+Kristian Lian, Section for Health and Exercise Physiology, Department of Public Health and Sport Sciences, Inland Norway University of Applied Sciences, Lillehammer, Norway.
+Email: kristian.lian@inn.no
+
+Stian Ellefsen, Section for Health and Exercise Physiology, Department of Public Health and Sport Sciences, Inland Norway University of Applied Sciences, Lillehammer, Norway.
+Email: stian.ellefsen@inn.no
 </div>
 
-ORCID:<!-- I does not say anything about ORCID in the journal template, do we leave it in?-->
+### ORCID
 <div custom-style="correspondence_style">
@@
 
-*Purpose:* Muscle hypertrophy, induced by resistance training, is determined by ribosomal biogenesis.
+**Introduction.** Muscle hypertrophy, induced by resistance training (RT), is determined by ribosomal biogenesis.
 High-glucose treatment may augment the transcription of ribosomal RNA, as evident from in vitro studies.
-However, its effect on resistance training-induced adaptations remains unexplored.
-
-*Methods:* Sixteen healthy, moderately trained individuals (male/female, n = 9/7; age, 23.7 (1.8)/24.6 (4.8)), participated in a within-participant trial with unilateral resistance training with glucose- and placebo supplementation on alternate days over six sessions per condition.
+However, its effect on RT-induced adaptations remains unexplored.
+
+**Methods.** Sixteen healthy, moderately trained individuals (male/female, n = 9/7; age, 23.7 (1.8)/24.6 (4.8)),<!-- Add male/female numbers, age.
+Kristian: Done--> participated in a within-participant trial with unilateral resistance training performed with glucose- and placebo supplement on alternate days over six sessions per condition.
 Participants remained fasted overnight, receiving glucose/placebo and protein supplementation only before and after RT sessions (3 sets of 10 repetitions maximum unilateral leg press and knee extension).
 Micro biopsies were sampled from m.
-vastus lateralis before and after the intervention.
-
-*Results:* Between conditions, there were no significant differences in increases of total RNA and ribosomal RNA (mean difference `r total_rna_stats`, 47S, `r stat_rrna$trg_47s`; 18S, `r stat_rrna$trg_18s`; 28S, `r stat_rrna$trg_28s`; 5.8S, `r stat_rrna$trg_5.8s`; 5S, `r stat_rrna$trg_5s`), or levels of c-Myc, upstream binding factor and ribosomal protein S6 (*p* = `r min(p_west)`-`r max(p_west)`).
-There was a linear relationship between upstream binding factor and total RNA, where an increase of 1 standard deviance unit of upstream binding factor equated to a $\sim$ `r corr_incr`% increase in total RNA (*p* = `r corrp`).
-
-*Conclusion:* Glucose supplementation before and after RT does not augment RT-induced ribosome biogenesis after two weeks of heavy RT in moderately trained young adults.
+vastus lateralis pre and post the intervention, and maximal unilateral isometric and isokinetic knee extension torque were measured before, during (days 4, 5, 8, 9) and after the intervention.
+
+**Results.** <!-- update statistics in this section -->From baseline to post-intervention total RNA and rRNA per unit muscle weight increased on average by ~24% and ~33-43%, respectively (Figure 2), with no differences between supplements (Figure 2A; mean difference `r total_rna_stats`, Figure 2B; 47S, `r stat_rrna$trg_47s`; 18S, `r stat_rrna$trg_18s`; 28S, `r stat_rrna$trg_28s`; 5.8S, `r stat_rrna$trg_5.8s`; 5S, `r stat_rrna$trg_5s`).
+RT with glucose resulted in estimated levels of c-Myc, UBF and RPS6 being `r stat_west$cmyc`, `r stat_west$ubf` and `r stat_west$rps6`% lower compared to placebo, respectively, without showing statistical significance (*p* = `r min(p_west)`-`r max(p_west)`; Figure 3A).
+There was a linear relationship between UBF and total RNA, where an increase of 1 SD unit of UBF equated to a $\sim$ `r corr_incr`% increase in total RNA (*p* = `r corrp`; Figure 3C).
+Lastly, there were no differences between supplements in maximal unilateral isometric knee extension force and isokinetic peak torque during the intervention (p > 0.05) and following the last session (p > 0.05), apart from the 60 d/s test before last RT session (p < 0.05).
+
+**Conclusion.** Glucose supplement before and after RT does not augment RT-induced ribosome biogenesis, nor does it enhance muscular performance and recovery after two weeks of heavy RT in moderately trained, young individuals.
+
+## NEW AND NOTEWORTHY
+
+**Keywords:** Skeletal muscle; hypertrophy; resistance training; glucose; ribosome
 
 </div>
-
-*Keywords:* Glucose; hypertrophy; resistance training; ribosome; skeletal muscle
-
-</div>
-
-# Statements and Declarations
-????
-
-# Abbreviations
-Make a table for this
 
@@
 
-Responses to systematic resistance training (RT) vary widely between individuals, with as much as 10-15% showing impaired skeletal muscle growth in response to standardized training interventions [@alvarez_interindividual_2018; @mann_high_2014; @thalacker-mercer_cluster_2013].
-While there certainly may be differences in the genetic predisposition in terms of readiness to adapt to exercise [@thalacker-mercer_cluster_2013], there are also recent studies showing that epigenetic factors such as ribosomal DNA (rDNA) availability may also impact the ability to mount an anabolic response to RT [@figueiredo_genetic_2021]<!--check this source -->.
-Indeed, the ability to synthesise ribosomes following a period of RT is connected to the magnitude of RT-induced responses in an individual [@hammarstrom_benefits_2020].
-For instance, increasing training volume generally induce greater ribosome biogenesis and thus, greater benefits of RT [@hammarstrom_benefits_2020;@krieger_single_2009;@schoenfeld_strength_2017].
-However, as evident from Hammarström et al.
-(2020), not all participants experienced increased muscle mass and -strength despite increasing training volume.
-Therefore, other means than modification of RT variables alone seem necessary to optimise individual responses to RT.
-
-Nutritional supplements such as protein and creatine effectively optimise RT adaptations [@cermak_protein_2012;@lanhers_creatine_2015;@lanhers_creatine_2017;@morton_systematic_2018].
+Responses to systematic resistance training (RT) vary widely between individuals, with as many as 10-15% showing impaired muscle growth in response to standardized training interventions [@alvarez_interindividual_2018; @mann_high_2014; @thalacker-mercer_cluster_2013].
+Such inability to benefit from RT may be explained by genetic and/or epigenetic disposition affecting the internal physiological milieu and subsequently affecting the ability to mount an anabolic response to RT [@figueiredo_genetic_2021;@thalacker-mercer_cluster_2013].<!-- Check these refs, do they correspond to genet epigen and internal mileu?
+Kristian: The way I read it, it looks like it corresponds --> Although RT program modifications such as altered exercise volume or intensity may induce more favourable adaptations in some individuals, these benefits are not uniform between individuals [@hammarstrom_benefits_2020;@krieger_single_2009;@schoenfeld_strength_2017].<!-- Add ref for more rt program modifications.
+Kristian: Added krieger et al.
+2009 and schoenfeld et al.
+2017 --> For instance, although changes in training volume generally induce greater benefits of RT, not all individuals benefit from such modifications [@hammarstrom_benefits_2020].
+Therefore, other means than modification of RT variables seem necessary to optimize individual responses.<!-- The above sentence might be a bit contradictory?
+Kristian: Im not sure, are you thinking about "other means than modification of RT" being contradictory of "modifications such as altered exercise volume.." sentence?
+I think there is a point to establishing that yes, increasing exercise volume is a potent way to increase the RT stimulus in general.
+However, in cases where this alone fails to elicit true responses, other means are necessary.
+Or am I missinterpreting this?
+-->
+
+Nutritional supplements such as protein and creatine effectively optimize RT adaptations [@cermak_protein_2012;@lanhers_creatine_2015;@lanhers_creatine_2017;@morton_systematic_2018].
 However, it remains unknown if other nutritional adjuvants such as glucose can increase the efficacy of RT.
-This is surprising since glucose is the preferred energy substrate of the contracting skeletal muscle during strenuous exercise and a major energy supplier to cells via adenosine triphosphate (ATP) synthesis [@mul_exercise_2015;@sebata_control_2018].
-Furthermore, energy availability is a decisive factor in the *de novo* synthesis of ribosomes [@kusnadi_regulation_2015;@moss_housekeeper_2007;@sebata_control_2018] which in turn determines muscle growth by increasing the muscle's translational capacity [@figueiredo_regulation_2019;@hammarstrom_benefits_2020;@sebata_control_2018;@stec_ribosome_2016;@von_walden_ribosome_2019].
+This is surprising since glucose is the preferred energy substrate of the contracting skeletal muscle during strenuous exercise and a major supplier of energy to cells via ATP synthesis [@mul_exercise_2015;@sebata_control_2018].
+Furthermore, energy availability is a decisive factor in the *de novo* synthesis of ribosomes [@kusnadi_regulation_2015;@moss_housekeeper_2007;@sebata_control_2018] <!--Add ref that indicate high energy requirement for ribo biog possibly (VONWALDEN19;TANAKA18?
+MOSS.
+Kristian: These three should do it-->which in turn determines muscle growth by increasing the muscle's translational capacity [@figueiredo_regulation_2019;@hammarstrom_benefits_2020;@sebata_control_2018;@stec_ribosome_2016;@von_walden_ribosome_2019].
 Therefore, investigating the effects of combined RT and glucose ingestion may provide valuable insight into the potential additive effect of glucose and RT on ribosome biogenesis.
 
-Transcription of ribosomal RNA (rRNA) by RNA Polymerase I is considered the rate-limiting step in synthesising new ribosomes [@moss_promotion_1995].
+Transcription of ribosomal RNA (rRNA) by Pol I is considered the rate limiting step in synthesizing new ribosomes [@moss_promotion_1995].
 Multiple signalling pathways converge to regulate rRNA transcription, including the mammalian target of rapamycin (mTORC1) signal-transduction pathway and c-Myc [@kusnadi_regulation_2015;@von_walden_ribosome_2019;@sebata_control_2018; @west_acute_2016; @mori_c-myc_2021].
 mTOR receives input from growth factors, hormones, mechanical loading, and nutrients to balance protein synthesis through multiple mechanisms based on cellular energy levels [@hoppe_amp-activated_2009].
-Importantly, mTOR and its downstream target ribosomal protein S6 kinase beta-1 (S6K1) serve as a mediator for insulin signalling in skeletal muscle [@hillier_physiological_2000].
+Importantly, mTOR and its downstream target S6K1 serve as a mediator for insulin signalling in skeletal muscle [@hillier_physiological_2000].
 The mTORC1 pathway contributes to ribosome biogenesis through the regulation of the translation of ribosomal proteins and to forming of the preinitiation complex (PIC) that marks the initiation of rRNA transcription [@figueiredo_regulation_2019;@von_walden_ribosome_2019].
@@
 
-The purpose of this investigation was to test the hypothesis that glucose supplementation before and after RT increases markers of ribosomal abundance following five RT sessions.
+The purpose of this investigation was to test the hypothesis that glucose supplementation before and after<!-- during?
+Kristian: before and after would be more precise, since they're not strictly ingesting anything during RT - I got a comment on that during the thesis defense --> RT increases markers of ribosomal abundance following five RT sessions.
 Secondly, we aimed to describe the association between UBF and total RNA abundance in human skeletal muscle.
-There have been multiple studies recently suggesting translational capacity to be as important, if not more important than, translational efficiency with regard to long-term skeletal muscle adaptations to RT [@hammarstrom_benefits_2020;@hammarstrom_ribosome_2022;@figueiredo_revisiting_2019].
+There have been multiple studies in the last years suggesting translational capacity to be as important, if not more, as translational efficiency in regard to long-term skeletal muscle adaptations to RT [@hammarstrom_benefits_2020;@hammarstrom_ribosome_2022;@figueiredo_revisiting_2019].
 Previously, *in vitro* studies have shown UBF and c-Myc to be important factors in the PIC, c-Myc as a general transcription factor and UBF as a specific transcription factor for rDNA transcription initiation [@mariappan_ribosomal_2011;@von_walden_ribosome_2019].
 Coupled with the recent observations by Hammarström et al.
-[@hammarstrom_benefits_2020;@hammarstrom_ribosome_2022], understanding underlying factors regulating and affecting ribosome biogenesis seems key to furthering our understanding of the optimisation of RT to individual phenotypes.
+[@hammarstrom_benefits_2020;@hammarstrom_ribosome_2022], understanding underlying factors regulating and affecting ribosome biogenesis seems key to furthering our understanding of the optimization of RT to individual phenotypes.
+<!-- We should maybe add a line or two about the rational here. Kristian: Tried to add some, what about this? to general? -->
 
@@
 
-All participants gave their written informed consent prior to study enrollment.
+All participants gave their informed consent prior to study enrollment.
 The study was approved by the regional ethical committee (REK, ID nr.
@@
 2-8 RT sessions per 14 days for the last six months).
-Exclusion criteria were previous injury leading to impaired strength, inability to perform resistance exercise training, symptoms, and a medical record of metabolic disorders including hyperglycemia.
+Exclusion criteria were previous injury leading to impaired strength, inability to perform resistance exercise training, symptoms, and a medical record of metabolic disorders including hyperglycemia, i.e.
+fasting venous plasma glucose ≥6.1 mmol/L and/or 2-hour glucose tolerance ≥7.8 mmol/L, and/or HbA1c >42 mmol/mol.
 Of the sixteen participants that commenced the intervention, three participants dropped out.
 One due to sickness and inability to resume, two participants experienced muscular discomfort related to heavy resistance training.
-Lean mass (Table 1) was measured using Dual-Energy X-ray Absorptiometry (DXA, Prodigy Advance PA+302047, Lunar, San Francisco, CA, USA) on Day -1, the last day preceding the RT intervention.
+Lean mass (Table 1) was measured using DXA (Prodigy Advance PA+302047, Lunar, San Francisco, CA, USA) on Day -1, the last day preceding the RT intervention.
 
@@
 Participants were randomly allocated to exercise one leg with a glucose condition and one leg with a placebo condition (Figure 1A).
-One person was exclusively responsible for the randomisation code and supplement distribution, blinding both investigators and participants regarding which leg exercised with glucose/placebo conditions.
+<!--from the "old" methods: either their dominant or non-dominant leg with glucose or placebo before and after RT, and whether they started RT with glucose or placebo.
+Thus, participants exercised every day, alternating between e.g., one leg + glucose and the other leg + placebo - this might be TMI, as it's presented in the figure as well?--> One person was exclusively responsible for the randomization code and supplement distribution, blinding both investigators and participants regarding which leg exercised with glucose/placebo.
 Glucose (Glucosum monohydricum, Merck KGaA, Darmstadt, Germany) and placebo (Steviosa, Soma Nordic AS, Oslo, Norway) were masked by mixing with 300ml Fun Light (Orkla, Oslo, Norway).
@@
 on days 1-2, 3-4, etc.).
-To further standardise this, participants also recorded and repeated their daily macronutrient intake (protein, fat, carbohydrate) and total calories on pairwise consecutive days.
+To further standardize this, participants also recorded and repeated their daily macronutrient intake (protein, fat, carbohydrate) and total calories on pairwise consecutive days.
 
@@
 Participants were asked to avoid  resistance- or high-intensity training of the legs from Day -7 (Figure 1A) and onwards, until completion of the intervention and post-testing, to ensure the reliability of pre-intervention strength data and minimal interference from external exercise sources.
+If participants did conduct exercise training outside of the prescribed protocol, they were asked to ensure equal loading on pairwise consecutive days.
 
@@
 
-The dietary intervention spanned the whole day, divided into three periods: I) From awakening until 2.5 hours (hrs) after RT, II) from 2.5hrs after RT until 2200hrs, and III) from 2200hrs until awakening.
-During period I, participants ingested protein and glucose **or** protein and placebo only.
+The dietary intervention spanned the whole day, divided into three periods: I) From awakening until 2.5 hours after RT, II) from 2.5 hours after RT until 2200hrs, and III) from 2200hrs until awakening.
+During period I, participants ingested protein and glucose/placebo only.
 Glucose/placebo was ingested three times in period I: 30 minutes before RT (0830hrs, 30g vs 0g glucose), immediately before RT (0900hrs, 30g vs.
@@
 During period III (2200-0700hrs), participants remained in an overnight fasted state.
-The daily onset of the dietary intervention (i.e., first ingestion of protein supplement) varied between participants, from 0600hrs to 0900hrs to allow multiple participants to complete the protocol simultaneously.
+The daily onset of the dietary intervention (i.e., first ingestion of PRO supplement) varied between participants, from 0600hrs to 0900hrs to allow multiple participants to complete the protocol simultaneously.
 During sessions, participants were free to ingest water ad libitum.
@@
 Individual positions were recorded and standardized from pre-intervention tests (Figure 1A, days -7 and -5).
-Isokinetic peak torque was measured concentrically from 90 to 0 degrees knee angle (extended knee was set to 0 degrees) at angular velocities of 60- and 240 degrees per second, 2x3 repetitions each, with the first set of each exercise as a sub-maximal warm-up.
-Isometric knee extensor peak torque was measured at a knee-angle of 60 degrees, for a maximum og 10 seconds and two repetitions per test.
+Isokinetic peak torque was measured at extensions speeds of 60- and 240 d/s, 2x3 repetitions each, with the first set of each exercise as a sub-maximal warm-up.
+Isometric peak torque was measured at a knee-angle of 60 degrees (extended knee was set to 0 degrees), for a maximum of 10 seconds and two repetitions per test.
 The isometric tests were ended when the participants reached a plateau or peak torque development decreased, which on average occurred between 2-4 seconds into the test.
@@
 Test I on day 12/13 included testing of both legs, representing 23hrs post-RT session test of one leg and post-session 5 test of the other leg.
-The highest peak torque values from the respective angular velocities and time points were summarized in an index.
-The index was calculated by dividing the average peak torque value by the highest observed peak torque value per angular velocity, and summarizing this new index per angular velocity to a mean strength index.
+The highest peak torque values from the respective extension speeds and time points were summarized in an index.
+The index was calculated by dividing the average peak torque value by the highest observed peak torque value per extension speed, and summarizing this new index per extension speed to a mean strength index.
 
@@
 All positions were controlled and recorded at the first 1RM test and repeated for the RT sessions.
-Maximal leg press strength was defined as the maximal load lifted in a controlled fashion, starting from a knee angle of 90 degrees.
-To find a reproducible 90 degree knee angle for each participant, centimeter markings on the side panels of the leg press machine was used to record were to find 90 degrees for each separate leg and participant.
-Attempts where participants did not reach 90 degrees during the eccentric phase, were not approved.
-Maximal knee extension testing followed the same specific warm-up was defined as maximal load lifted in a controlled fashion, reaching full extension of the knee joint.
+Maximal leg press strength was defined as the maximal load lifted in a controlled fashion, with a knee angle of 45 degrees (start), within a 2.5kg range.
+Attempts where participants did not reach 45 degrees during the eccentric phase, were not approved.
+Maximal knee extension testing followed the same specific warm-up and pause protocol and was defined as maximal load lifted in a controlled fashion, reaching full extension of the knee joint, within a 1.25kg range.
 Attempts with exaggerated hip movement or beneath full extension were not approved.
-Two minutes of rest were given during the specific warm-up, and three minutes of rest were given between 1RM attempts.
+Two minutes of rest were given during the specific warm-up, and three minutes of rest was given between 1RM attempts.
 
@@
 
-Resistance training consisted of three sets of unilateral leg presses and three sets of knee extensions, with an intensity of 10 repetitions maximum (10RM).
+Resistance training consisted of three sets of unilateral leg presses and three sets of knee extensions, with an intensity of 10 repetition maximum (10RM).
 As a general warm-up, the participants cycled on an indoor exercise bicycle for 5-10min.
 In addition, before the respective exercises, two 10-repetition warm-up sets were completed at ~50% and ~70% of 10RM.
-To ensure an adequate exercise stimulation throughout the intervention, the exercise load was increased the following set if the participants lifted more than 12 repetitions, as a progressive loading strategy.
-If the participants lifted fewer than 8 repetitions per set, the load was reduced in the following set.
+To ensure an adequate exercise stimulation throughout the intervention, the exercise load was increased if the participants lifted more than 12 repetitions, as a progressive loading strategy.
 Rest between working sets was two minutes.
-For safety and standardisation purposes, all sessions were monitored by trained personnel.
-Lastly, training volume (load and repetitions) was logged for every session.
+For safety and standardization purposes, all sessions were monitored by trained personnel.
+<!-- From the "old" methods: The rate of perceived exertion (RPE, 0–10-point scale) was logged before every session, and the session score was logged 15min after each session - is this necessary info, as we are not reporting these data?
+--> Lastly, training volume (load and repetitions) was logged for every session.
 
@@
 After the biopsy sampling, muscle tissue was divided into two aliquots for determination of total RNA/expression of rRNA and two aliquots for protein content measurement.
-Aliquots were snap-frozen in isopentane (-80°C) and stored at -80°C until further analyses.
-Muscle biopsies were collected at four time points: I/II) Pre-intervention (Figure 1A, 2hrs before training, Day 1 = leg 1, Day 2 = leg 2), and III/IV) after the fifth RT session, two hours before the sixth RT session (Figure 1A, Day 11 = leg 1, Day 12 = leg 2).
+Aliquots were snap frozen in isopentane (-80°C) and stored at -80°C until further analyses.
+Muscle biopsies were sampled at four times: I/II) Pre-intervention (Figure 1A, 2hrs before training, Day 1 = leg 1, Day 2 = leg 2), and III/IV) after the fifth RT session, two hours before the sixth RT session (Figure 1A, Day 11 = leg 1, Day 12 = leg 2).
 At each time point, two samples were taken from the same incision.
@@
 To measure blood glucose levels with and without glucose ingestion/training, capillary blood was collected from finger draws on days with biopsy sampling.
-One capillary blood sample was collected on day 1 (Figure 1A) to serve as a baseline.
-On days 11 and 12 (Figure 1A), capillary blood samples were collected seven times: I) Immediately prior to protein ingestion (0700hrs) II) 45 minutes after protein ingestion (0745hrs) III) 1.5hrs after protein ingestion (0830hrs, i.e., immediately before GLU/PLAC intake), IV) 2hrs after protein ingestion (0900hrs, i.e., immediately before training), IV) in the middle of RT (~0915hrs), V) immediately after training (~0930hrs), and VI) 2hrs after completion of training (~1130hrs).
+One capillary blood sample was taken on day 1 (Figure 1A) to serve as a baseline.
+On days 11 and 12 (Figure 1A), capillary blood samples were collected seven times: I) Immediately prior to protein ingestion (0700hrs) II) 45 minutes after protein ingestion (0745hrs) III) 1.5hrs after protein ingestion (0800hrs, i.e., immediately before GLU/PLAC intake), IV) 2hrs after protein ingestion (0900hrs, i.e., immediately before training), IV) in the middle of RT (~0915hrs), V) immediately after training (~0930hrs), and VI) 2hrs after completion of training (~1130hrs).
 Capillary blood samples were analysed with in-house equipment (BIOSEN C-Line, EKF diagnostic GmbH, Barleben).
@@
 qPCR reactions were run over 40 cycles (3 s 95°C denaturing and 30 s 60°C annealing) on a fast-cycling real-time detection system (Applied Biosystems 7500 fast Real-Time PCR Systems, Life Technologies AS), with a total reaction volume of 10 µl consisting of 2 µl of cDNA, gene-specific primers (0.5 µM final concentration) and a commercial master mix (2X SYBR Select Master Mix, Applied Biosystems, Life Technologies AS, Oslo, Norway).
-An overview of the primers may be found in table 2.
 Raw fluorescence data was modelled with a best-fit sigmoidal model using the qPCR package [@ritz_qpcr_2008] written for R [@hammarstrom_benefits_2020; @Rcore].
@@
 Following the wet transfer, membranes were stained with a reversible total protein stain (Thermo Fisher Scientific) and then blocked for 1hr at room temp with a blocking buffer of Tris-buffered saline (TBS; 20 mM Tris, 150 mM NaCl) with 5% non-fat dry milk and 0.1% Tween-20.
-Primary and secondary antibodies were purchased from Santa Cruz Biotechnology (Texas, USA): UBF, UBF F-9, sc-13125; rpS6, Ribosomal protein S6 C-8, sc-74459; and Thermo Fisher Scientific (Oslo, Norway): c-Myc, 9E10; goat anti-mouse (for c-Myc), goat anti-mouse IgG1 (y1) horseradish peroxidase conjugate; and anti-mouse (anti-mouse IgG1 horseradish peroxidase conjugate).
-Antibodies were diluted in blocking buffer to concentrations corresponding to 1:25 000 (UBF, rpS6) and 1:5000 (c-Myc).
-
+Primary and secondary antibodies were purchased from Santa Cruz Biotechnology (Texas, USA): UBF, UBF F-9, sc-13125; rpS6, Ribosomal protein S6 C-8, sc-74459; and Thermo Fisher Scientific (Oslo, Norway): c-Myc, 9E10; goat anti-mouse (for c-Myc), goat anti-mouse IgG1 (y1) horseradish peroxidase conjugate; and anti-mouse (anti-mouse IgG1 horseradish peroxidase conjugate)<!-- insert manufacturer -->.
+Antibodies were diluted in blocking buffer to concentrations corresponding to <!-- Add final concentrations of antibodies -->.
 Membranes were incubated overnight with primary antibodies, and for 1 hr with secondary antibodies.
@@
 A priori power calculations showed that 20 participants would grant a statistical power of 80% (alpha = 0.05), accounting for an expected dropout of 20%.
-This power calculation was based on an assumption that the effects of glucose ingestion on ribosome biogenesis (measured as within-participant differences in total RNA) may equate to the effects of increasing RT volume from low to moderate [@hammarstrom_benefits_2020].
+This power calculation was based on an assumption that the effects of glucose ingestion on ribosome biogenesis (measured as within participant differences in total RNA) may equate to the effects of increasing RT volume from low to moderate [@hammarstrom_benefits_2020] corresponding to an effect size of d =  <!-- Add the effect size here-->.
 Total RNA, protein and qPCR data were analyzed by mixed-effects models with fixed effects included as *supplement $\times$ time*.
@@
 These data only supported a random intercept per participant.
-The linear mixed-effects models were fitted with the lmer function from the lme4 package using the lmerTest package to procure p-values (Satterthwaite's method for approximating degrees of freedom)[@bates_fitting_2014; @lmerTest], written for R[@Rcore].
-Log-transformed values were expressed as fold changes in visualisations.
+The linear mixed effects models were fitted with the lmer function from the lme4 package using the lmerTest package to procure p-values (Satterthwaite's method for approximating degrees of freedom)[@bates_fitting_2014; @lmerTest], written for R[@Rcore].
+Log-transformed values were expressed as fold-changes in presentations.
 Descriptive data are presented as mean and (SD).
@@
 
-## Markers of ribosome biogenesis
-
-### Total RNA and ribosomal RNA
+### Markers of ribosome biogenesis
+
+##### Total RNA and ribosomal RNA
 
@@
 
-From baseline to post-intervention total RNA and rRNA per unit muscle weight increased on average by ~20-27% and ~25-57%, respectively (Figure 2).
+From baseline to post-intervention total RNA and rRNA per unit muscle weight increased on average by ~24% and ~33-43%, respectively (Figure 2).
 However, RT with glucose did not induce increased accumulation of total RNA (Figure 2A, mean difference `r total_rna_stats`) or rRNA  (Figure 2B; 47S, `r stat_rrna$trg_47s`; 18S, `r stat_rrna$trg_18s`; 28S, `r stat_rrna$trg_28s`; 5.8S, `r stat_rrna$trg_5.8s`; 5S, `r stat_rrna$trg_5s`) compared to RT with placebo.
 
-### Protein
+##### Protein
 
@@
 low glucose/glucose starvation condition on rDNA transcription initiation observed in previous *in vitro* studies [@mariappan_ribosomal_2011; @tanaka_mild_2015] are not translatable to acute effects in human skeletal muscle *in vivo*.
-<!-- Is there a point to discuss the fact that both legs may have exercised with sufficient glucose levels here, as the probably had full stores from the morning?
--->Notably, the present study was specifically designed to investigate differences in pre to post changes within participants, and is as such also quite limited to this perspective.
+Notably, the present study was specifically designed to investigate differences in pre-post changes within participants, and is as such also quite limited to this perspective.
 Hence, the observed effects of time serve as supplementary to the efficiency of the design, without a negative control group to compare with.
@@
 Lastly, c-Myc has previously been described as a potent regulator of ribosome biogenesis, independent of mTORC1, and a direct regulator of UBF [@mori_c-myc_2021; @poortinga_c-myc_2011; @west_acute_2016].
-Hence, it seems quite reasonable to observe similar changes in these three proteins.
+Hence, it seems quite reasonable to observe equal changes in these three proteins.
 <!-- OLD: While content of UBF and rpS6 changed equally in RT with glucose compared to placebo, content of c-Myc was significanly higher after RT with placebo.
@@
 Hence, there were no differences in mechanical loading to induce a higher stress between conditions.
-Arguably, an increased energy availability via glucose ingestion during RT may induce less acute stress on the exercised skeletal muscle, therefore less fatigue, and perhaps less performance reduction, compared to placebo during RT <!-- this is possibly a stretch -->[@westerblad_mechanisms_1998; @kent-braun_central_1999] Unfortunately, we did not conduct measurements of markers of metabolic stress such as inorganic phosphate, H+, Mg2+ and the ADP/ATP ratio [@westerblad_mechanisms_1998; @kent-braun_central_1999].
+Arguably, an increased energy availability via glucose ingestion during RT may induce less acute stress on the exercised skeletal muscle, therefore less fatigue, and perhaps less performance reduction, compared to placebo during RT <!-- this is possibly a stretch -->[@westerblad_mechanisms_1998; @kent-braun_central_1999] Unfortunately, we did not conduct measurements of markers of metabolic stress such ass inorganic phosphate, H+, Mg2+ and the ADP/ATP ratio [@westerblad_mechanisms_1998; @kent-braun_central_1999].
 Therefore, discussing the potential effect of differences in metabolic stress between conditions would only be speculation, however probable.
@@
 
-### Perspectives and Significance
-
 In conclusion, ingestion of glucose immediately before and after resistance exercise training does not acutely augment ribosome biogenesis after two weeks of heavy resistance exercise training, in moderately trained young adults.
@@
 
+# APPENDIX
+
+IS THIS REQUIRED? IF NOT, IS IT NECESSARY?
+
 # DATA AVAILABILITY
 
-Complete data sets and scripts are downloadable here; https://github.com/Kristianlian/ribose-paper
-
-# SUPPLEMENTAL MATERIAL
-
-Extended methods can be found here; https://github.com/Kristianlian/ribose-paper
+LINK TO GITHUB
+
+# GRANTS
+
+IS THIS NECESSARY?
+
+# DISCLOSURES
 
 # AUTHOR CONTRIBUTIONS
-Conceived and designed research: Stian Ellefsen, Håvard Hamarsland, Daniel Hammarström.
-Performed experiments: Kristian Lian, Daniel Hammarström, Håvard Hamarsland, Håvard Nygaard, Sara Christine Moen.
-Interpreted results of experiments: Kristian Lian, Daniel Hammarström.
-Prepared Figures: Kristian Lian, Daniel Hammarström.
-Drafted manuscripts: Kristian Lian, Daniel Hammarström.
-Edited and revised manuscript: Stian Ellefsen, Håvard Hamarsland, Daniel Hammarström.
-Approved final version of the manuscript: .
 
@@
 
-dxa.flex <- dxa.tab %>%
+dxa.tab %>%
   full_join(humac.tab) %>%
@@
   add_header_row(values = c(" ", "Knee-extension peak torque"), colwidths = c(6, 3)) %>%
+  add_body_row(top = FALSE, value = as_paragraph("Values are mean and (SD)"), colwidths = 9) %>%
   set_caption("Table 1. Participant characteristics")
 
-dxa.flex <- border(dxa.flex, 
-                     i = NULL,
-                     j = NULL,
-                     border = NULL,
-                     border.top = fp_border_default(color = "grey"),
-                     border.bottom = fp_border_default(color = "grey"),
-                     border.left = fp_border_default(color = "grey"),
-                     border.right = fp_border_default(color = "grey"),
-                     part = "all")
-
-dxa.flex
-
-```
-Values are means ± SD.
-60º sec-1 = 60 degrees per second movement speed (isokinetic), 240º sec-1 = 240 degrees per second movement speed (isokinetic), 0º sec-1 = 0 degrees per second movement speed (isometric).
+```
 
@@
 ## Flextable
-primer.tab <- tab.dat %>%
+tab.dat %>%
   flextable() %>%
@@
                     as_paragraph("Sequence (forward - reverse)"),
-                    as_paragraph("Ct mean (SD)"),
+                    as_paragraph("CT mean (SD)"),
                     as_paragraph("E"))) %>%
+  #add_header_row(values = c(" ", " ", " ", "Primer sequences"), 
+               #  colwidths = c(1, 1, 1, 1)) %>%
   set_caption("Table 2. Primer Sequences") %>%
-  set_table_properties(layout = "autofit")
-
-primer.tab <- border(primer.tab, 
-                     i = NULL,
-                     j = NULL,
-                     border = NULL,
-                     border.top = fp_border_default(color = "grey"),
-                     border.bottom = fp_border_default(color = "grey"),
-                     border.left = fp_border_default(color = "grey"),
-                     border.right = fp_border_default(color = "grey"),
-                     part = "all")
-
-primer.tab
-
-```
-Values of CT are means ± SD. rRNA = ribosomal RNA, E = primer efficiency. Average cycle thresholds (Ct) and priming efficiencies were calculated from all qPCR reactions.
+  width(j = 1, 1.5, unit = "in") %>%
+  width(j = 2, 3, unit = "in") %>%
+  width(j = 3, 1.5, unit = "in") %>%
+  width(j = 4, 0.5, unit = "in")
+  #set_table_properties(layout = "autofit")
+
+```

### 2023-10-13 resources/older manuscript versions/second_draft.RMD
cleanup

--- 2023-09-30 manuscript.RMD
@@
 found physiological hyperinsulinemia to stimulate p70S6K phosphorylation in human skeletal muscle [@hillier_physiological_2000].
-Moreover, robust increases in markers of ribosome biogenesis such as 47S pre-rRNA and mature rRNA can be expected after an acute bout of RT [@figueiredo_impact_2016] or a short period of RT [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022].
+Moreover, robust increases in markers of ribosome biogenesis such as 47S pre-rRNA and mature rRNA can be expected acutely after a session of RT [@figueiredo_impact_2016] or a short period of RT [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022].
 Therefore, we hypothesized that if glucose can, directly or indirectly via insulin, stimulate the initiation of rDNA transcription, five sessions of RT with glucose compared to placebo would elicit this effect.

### 2023-10-18 README.rmd
readme updates

@@
 ```
+
+### Supplementary figures
+
+```{r macro nutrient figure, echo=FALSE, message=FALSE, warning=FALSE}
+#### Nutrition change figs
+
+# Author: KL
+# Project: Ribose
+
+# Purpose: This script plots figures from protein, fat, carbs, calories and propr kg emmeans
+
+# Packages
+library(tidyverse); library(emmeans); library(cowplot)
+
+# Data
+
+pro <- readRDS("./data/data-gen/nutrition/prot.emm.RDS")
+fat <- readRDS("./data/data-gen/nutrition/fat.emm.RDS")
+cho <- readRDS("./data/data-gen/nutrition/cho.emm.RDS")
+cal <- readRDS("./data/data-gen/nutrition/cal.emm.RDS")
+propr <- readRDS("./data/data-gen/nutrition/propr.emm.RDS")
+
+## Figures
+
+pos <- position_dodge(width = 0.2)
+
+# Protein
+
+pro.fig <- pro %>%
+  data.frame() %>%
+  add_row(supplement = "placebo", time = "change.d1", emmean = 0, SE = 0, df = 0, lower.CL = 0, upper.CL = 0, .before =1) %>%
+  add_row(supplement = "glucose", time = "change.d1", emmean = 0, SE = 0, df = 0, lower.CL = 0, upper.CL = 0, .before =2) %>%
+  ggplot(aes(time, emmean, group = supplement, fill = supplement)) +
+  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
+                position = pos,
+                width = 0.2) +
+  geom_line(position = pos) +
+  geom_point(shape = 21, position = pos, size = 3) +
+  scale_x_discrete(labels=c("change.d1" = "Day 1", "change.d2" = "Day 2", "change.d3" = "Day 3",
+                            "change.d4" = "Day 4", "change.d5" = "Day 5", 
+                            "change.d6" = "Day 6")) +
+  labs(x = "", y = "Protein (g)", fill = "Supplement") +
+  theme_classic() +
+  theme(axis.text.x = element_text(size=8))
+
+# Fat
+fat.fig <- fat %>%
+  data.frame() %>%
+  add_row(supplement = "placebo", time = "change.d1", emmean = 0, SE = 0, df = 0, lower.CL = 0, upper.CL = 0, .before =1) %>%
+  add_row(supplement = "glucose", time = "change.d1", emmean = 0, SE = 0, df = 0, lower.CL = 0, upper.CL = 0, .before =2) %>%
+  ggplot(aes(time, emmean, group = supplement, fill = supplement)) +
+  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
+                position = pos,
+                width = 0.2) +
+  geom_line(position = pos) +
+  geom_point(shape = 21, position = pos, size = 3) +
+  scale_x_discrete(labels=c("change.d1" = "Day 1", "change.d2" = "Day 2", "change.d3" = "Day 3",
+                            "change.d4" = "Day 4", "change.d5" = "Day 5", "change.d6" = "Day 6")) +
+  labs(x = "", y = "Fat (g)", fill = "Supplement") +
+  theme_classic() +
+  theme(axis.text.x = element_text(size=8))
+
+# Carbs
+cho.fig <- cho %>%
+  data.frame() %>%
+  add_row(supplement = "placebo", time = "change.d1", emmean = 0, SE = 0, df = 0, lower.CL = 0, upper.CL = 0, .before =1) %>%
+  add_row(supplement = "glucose", time = "change.d1", emmean = 0, SE = 0, df = 0, lower.CL = 0, upper.CL = 0, .before =2) %>%
+  ggplot(aes(time, emmean, group = supplement, fill = supplement)) +
+  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
+                position = pos,
+                width = 0.2) +
+  geom_line(position = pos) +
+  geom_point(shape = 21, position = pos, size = 3) +
+  scale_x_discrete(labels=c("change.d1" = "Day 1", "change.d2" = "Day 2", "change.d3" = "Day 3",
+                            "change.d4" = "Day 4", "change.d5" = "Day 5", "change.d6" = "Day 6")) +
+  labs(x = "", y = "Carbohydrates (g)", fill = "Supplement") +
+  theme_classic() +
+  theme(axis.text.x = element_text(size=8))
+
+# Calories
+cal.fig <- cal %>%
+  data.frame() %>%
+  add_row(supplement = "placebo", time = "change.d1", emmean = 0, SE = 0, df = 0, lower.CL = 0, upper.CL = 0, .before =1) %>%
+  add_row(supplement = "glucose", time = "change.d1", emmean = 0, SE = 0, df = 0, lower.CL = 0, upper.CL = 0, .before =2) %>%
+  ggplot(aes(time, emmean, group = supplement, fill = supplement)) +
+  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
+                position = pos,
+                width = 0.2) +
+  geom_line(position = pos) +
+  geom_point(shape = 21, position = pos, size = 3) +
+  scale_x_discrete(labels=c("change.d1" = "Day 1", "change.d2" = "Day 2", "change.d3" = "Day 3",
+                            "change.d4" = "Day 4", "change.d5" = "Day 5", "change.d6" = "Day 6")) +
+  labs(x = "", y = "Calories", fill = "Supplement") +
+  theme_classic() +
+  theme(axis.text.x = element_text(size=8))
+
+# Protein pr kg
+propr.fig <- propr %>%
+  data.frame() %>%
+  add_row(supplement = "placebo", time = "change.d1", emmean = 0, SE = 0, df = 0, lower.CL = 0, upper.CL = 0, .before =1) %>%
+  add_row(supplement = "glucose", time = "change.d1", emmean = 0, SE = 0, df = 0, lower.CL = 0, upper.CL = 0, .before =2) %>%
+  ggplot(aes(time, emmean, group = supplement, fill = supplement)) +
+  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
+                position = pos,
+                width = 0.2) +
+  geom_line(position = pos) +
+  geom_point(shape = 21, position = pos, size = 3) +
+  scale_x_discrete(labels=c("change.d1" = "Day 1", "change.d2" = "Day 2", "change.d3" = "Day 3",
+                            "change.d4" = "Day 4", "change.d5" = "Day 5", "change.d6" = "Day 6")) +
+  labs(x = "", y = "Protein (g) \nper kg body weight\n", fill = "Supplement") +
+  theme_classic() +
+  theme(axis.text.x = element_text(size=8))
+
+# Cowplot for gathering figures
+
+legend <- get_legend(pro.fig + theme(legend.box.margin = margin(0, 0, 0,12)))
+
+nut.fig <- plot_grid(pro.fig + theme(legend.position = "none"),
+                     propr.fig + theme(legend.position = "none"),
+                     cho.fig + theme(legend.position = "none"),
+                     fat.fig + theme(legend.position = "none"),
+                     cal.fig + theme(legend.position = "none"), 
+                     legend, ncol = 2, nrow = 3)
+nut.fig
+
+```
+
+**Figure 4: Daily change in macronutrient intake per condition.** For the purpose of illustration the first day with the glucose condition and the first day with the placebo condition are combined into "Day 1", though they were completed on separate days (i.e., day 1 and 2).
+The same goes for Day 2, 3, 4, 5 and 6.
+Macronutrient intake data was gathered from the MyFitnessPal application, where participants registered their daily intake.
+Values are estimated marginal means with 95% CI.
+
+**Table 3: Daily mean dietary intake of macro nutrients during each training day per supplement leg.** Glu = glucose,Plac = placebo.
+Training day 1-training day 12 are combined pairwise to day 1-6: Day 1 = training day 1/2, Day 2 = training day 3/4, Day 3 = training day 5/6, Day 4 = training day 7/8, Day 5 = training day 9/10, Day 6 = training day 11/12.
+Values are mean ± standard deviation.
+P = difference between Glu and Plac.
+Glucose n = 12, placebo n = 12.
+Day 6: Glu = 11, Plac = 11.
+```{r nutrition table, echo=FALSE, message=FALSE, warning=FALSE}
+### Nutrition data table
+
+# This script produces a table of the daily nutritional intake, calculated to mean and SD per supplement.
+
+library(knitr)
+library(rmarkdown)
+library(tidyverse)
+library(readxl)
+library(readr)
+library(tinytex)
+library(tidyr)
+library(knitr)
+library(kableExtra)
+
+#nutritiontable with pro.pr kg bodyweight pr. group
+
+#laod dxa data and take out weight and subject to join with nutrition data
+fp.weight <-read_excel("data/dxa/ribose_dxa.xlsx") %>%
+  select(weight, subject)
+
+#nutrition data.
+# Data are presented as grams, supplemented protein (25g whey protein isolate = 42g protein) and glucose (90g) has been added by investigators,
+# and calculated into the total amount of ingested per day.
+
+# Load nutrition data. Sorts data and summarises into mean per supplement per day
+
+nut.weight <- read_excel("data/nutrition/ribose_nutrition.xlsx")%>%
+  select(timepoint, meal, subject, protein, fat, calories,
+         carbohydrates, group, sup_pro, sup_gluc, kcal_glu) %>%
+  #changed names to compare groups each paired days
+  mutate(timepoint = if_else(timepoint %in% c("T1","T2"), 
+                             "Day 1",
+                             if_else(timepoint %in% c("3", "4"),
+                                     "Day 2",
+                                     if_else(timepoint %in% c("5", "6"),
+                                             "Day 3",
+                                             if_else(timepoint %in% c("7", "8"),
+                                                     "Day 4",
+                                                     if_else(timepoint %in% c("9", "10"),
+                                                             "Day 5",
+                                                             if_else(timepoint %in% c("T3", "T4"),
+                                                                     "Day 6",
+                                                                     "na"))))))) %>%
+  group_by(timepoint, subject, group) %>%
+  #take in weight
+  inner_join(fp.weight) %>%
+  #summarise each variable to get total pr timepoint.
+  #Whey protein supplement is added to total protein ingestion
+  summarise(protein = sum(protein + sup_pro),
+            fat = sum(fat),
+            calories = sum(calories + kcal_glu),
+            carbohydrates = sum(carbohydrates + sup_gluc),
+            weight = sum(weight)) %>%
+  #summarise to get protein pr weight.
+  mutate(proprkg = (protein/weight)) %>%
+  pivot_wider(names_from = group,
+              values_from = c(fat, protein, calories, carbohydrates, weight, proprkg))
+
+# Nutrition table with pro/weight. The table shows mean and SD of total intake 
+# of grams per supplement per time point.
+
+nuttable <- nut.weight %>%
+  pivot_longer(names_to = "variable",
+               values_to = "values", col= fat_glucose:proprkg_placebo) %>%
+  group_by(timepoint, variable) %>%
+  summarise(m = mean(values, na.rm = TRUE),
+            s= sd(values, na.rm = TRUE)) %>%
+  ungroup() %>%
+  separate(variable, into = c("variable", "group")) %>%
+  mutate(stat = paste0(round(m, 1), " (", round(s, 1), ")")) %>%
+  select(timepoint, variable, group, stat,) %>%
+  pivot_wider(names_from = variable,
+              values_from = stat) %>%
+  select(timepoint, group, calories, carbohydrates, fat, protein,
+         proprkg) %>%
+  #changed names of group for simplicity 
+  mutate(group = if_else(group %in% c("glucose"),
+                         "G",
+                         if_else(group %in% c("placebo"),
+                                 "P",
+                                 "")))
+
+nuttable %>%
+  kable()
+
+```

### 2023-10-18 manuscript.RMD
readme updates

@@
 
-**Purpose:** Muscle hypertrophy, induced by resistance training, is determined by ribosomal biogenesis.
-High-glucose treatment may augment the transcription of ribosomal RNA, as evident from in vitro studies.
-However, its effect on resistance training-induced adaptations remains unexplored.
-
-**Methods:** Sixteen healthy, moderately trained individuals (male/female, n = 9/7; age, 23.7 (1.8)/24.6 (4.8)), participated in a within-participant trial with unilateral resistance training with glucose- and placebo supplementation on alternate days over six sessions per condition.
-Participants remained fasted overnight, receiving glucose/placebo and protein supplementation only before and after RT sessions (3 sets of 10 repetitions maximum unilateral leg press and knee extension).
+**Purpose:** Resistance training-induced skeletal muscle hypertrophy seems to depend on ribosome biogenesis and content.
+High glucose treatment may augment ribosomal biogenesis, as evident from *in vitro* studies, and may thus act to potentiate resistance training-induced adaptations *in vivo*.
+However, this remains largely unexplored.
+
+**Methods:** Sixteen healthy, moderately trained individuals (male/female, n = 9/7; age, 24.1 (3.3)), participated in a within-participant crossover trial, conducting unilateral resistance training (leg press and knee extension, 3 sets of 10 repetitions maximum) with pre- and post-exercise ingestion of either glucose (3x30g, 90g total) or placebo supplement (Stevia rebaudiana extract, 3x0.3g, 0.9g total), together with protein (2x25g, 50g total), on alternating days for twelve days; six sessions per condition, resistance training was performed in an otherwise fasted state, as morning sessions.
 Micro biopsies were sampled from m.
-vastus lateralis before and after the intervention.
-
-**Results:** Between conditions, there were no significant differences in increases of total RNA and ribosomal RNA (mean difference `r total_rna_stats`, 47S, `r stat_rrna$trg_47s`; 18S, `r stat_rrna$trg_18s`; 28S, `r stat_rrna$trg_28s`; 5.8S, `r stat_rrna$trg_5.8s`; 5S, `r stat_rrna$trg_5s`), or levels of cellular myelocytomatosis oncogene, upstream binding factor and ribosomal protein S6 (*p* = `r min(p_west)`-`r max(p_west)`).
-There was a linear relationship between upstream binding factor and total RNA, where an increase of 1 standard deviance unit of upstream binding factor equated to a $\sim$ `r corr_incr`% increase in total RNA (*p* = `r corrp`).
-
-**Conclusion:** Glucose supplementation before and after RT does not augment RT-induced ribosome biogenesis after two weeks of heavy RT in moderately trained young adults.
+vastus lateralis before and after the intervention, and maximal unilateral isometric and isokinetic knee extension torque were measured before, during (days 4/5, 8/9) and after the intervention.
+
+**Results:** Glucose ingestion did not have beneficial effects on resistance training-induced increases of ribosomal content (mean difference 7.6% [-7.2, 24.9], *p* = 0.34; ribosomal RNA, 47S/18S/28S/5.8S/5S, range 7.6%-37.9%, *p* = 0.40-0.98) or levels of relevant transcriptional or translational regulators (c-MYK/UBF/rpS6, *p* = 0.094-0.292).
+Of note, both baseline and trained state data of total RNA showed a linear relationship with UBF; a ∼14% increase in total RNA corresponded to 1 standard deviation unit increase in UBF (*p* = 0.003).
+
+**Conclusion:** Glucose ingestion before and after resistance training did not augment ribosomal RNA accumulation during twelve days of heavy-load resistance training in moderately trained young adults.
 
@@
 
-Data was collected at the sports science lab at the Inland Norway University of Applied Sciences, with all analyses except for serum blood samples performed at the sports science lab.
-Serum blood samples were performed at Innlandet Hospital Trust.
+Data was collected at the sports science lab at the Inland Norway University of Applied Sciences, with all analyses performed on-site, except serum blood analyses, which were performed at Innlandet Hospital Trust.
 S.E., H.H.
@@
 ### Acknowledgements
-Gratitude is due, to Henriette Spilhaug Bollandsås, Nora Lerdalen Bonsak and Thea Bøhn Nyløkken for their dedication and hard work in the training intervention.
+
+Gratitude is due, to Henriette Spilhaug Bollandsås, Nora Lerdalen Bonsak and Thea Bøhn Nyløkken for their dedication and hard work during the training intervention.
 The authors would also like to express their gratitude to all the participants for staying committed to this demanding protocol in an impressive way.
@@
 Responses to systematic resistance training (RT) vary widely between individuals, with as much as 10-15% showing impaired skeletal muscle growth in response to standardized training interventions [@alvarez_interindividual_2018; @mann_high_2014; @thalacker-mercer_cluster_2013].
-While there certainly may be differences in the genetic predisposition in terms of readiness to adapt to exercise [@thalacker-mercer_cluster_2013], there are also recent studies showing that epigenetic factors such as ribosomal DNA (rDNA) availability may also impact the ability to mount an anabolic response to RT [@figueiredo_genetic_2021].
-Indeed, the ability to synthesise ribosomes following a period of RT is connected to the magnitude of RT-induced responses in an individual [@hammarstrom_benefits_2020].
-For instance, increasing training volume generally induces greater ribosome biogenesis and thus, greater benefits of RT [@hammarstrom_benefits_2020;@krieger_single_2009;@schoenfeld_strength_2017].
-However, as evident from Hammarström et al.
-(2020), not all participants experienced increased muscle mass and -strength despite increasing training volume.
+There may be differences in the genetic predisposition in terms of readiness to adapt to exercise [@thalacker-mercer_cluster_2013], however, the internal physiological milieu seems favourably shaped for muscle growth by factors such as different types of training stress and nutrients [@figueiredo_genetic_2021; @sebata_control_2018; @thalacker-mercer_cluster_2013].
+Indeed, training stress and nutrients impact the ability to synthesise ribosomes, which in turn is connected to the magnitude of RT-induced responses in an individual [@hammarstrom_benefits_2020; @kusnadi_regulation_2015; @sebata_control_2018].
+For instance, increasing training volume generally induces greater ribosome biogenesis and thus, greater benefits of RT [@hammarstrom_benefits_2020; @krieger_single_2009; @schoenfeld_strength_2017].
+Still, as evident from Hammarström et al.
+(2020), not all participants experience increased muscle mass and -strength despite increasing training volume.
 Therefore, other means than modification of RT variables alone seem necessary to optimise individual responses to RT.
 
-Nutritional supplements such as protein and creatine effectively optimise RT adaptations [@cermak_protein_2012;@lanhers_creatine_2015;@lanhers_creatine_2017;@morton_systematic_2018].
+Nutritional supplements such as protein and creatine are often used to optimise RT adaptations [@cermak_protein_2012; @lanhers_creatine_2015; @lanhers_creatine_2017; @morton_systematic_2018].
 However, it remains equivocal if other nutritional adjuvants such as glucose can increase the efficacy of RT [@tezze_anabolic_2023].
-This is surprising since glucose is the preferred energy substrate of the contracting skeletal muscle during strenuous exercise and a major energy supplier to cells via adenosine triphosphate (ATP) synthesis [@mul_exercise_2015;@sebata_control_2018].
-Furthermore, energy availability is a decisive factor in the *de novo* synthesis of ribosomes [@kusnadi_regulation_2015;@moss_housekeeper_2007;@sebata_control_2018] which in turn determines muscle growth by increasing the muscle's translational capacity [@figueiredo_regulation_2019;@hammarstrom_benefits_2020;@sebata_control_2018;@stec_ribosome_2016;@von_walden_ribosome_2019].
+This is surprising since glucose is the preferred energy substrate of the contracting skeletal muscle during strenuous exercise and a major energy supplier to cells via adenosine triphosphate (ATP) synthesis [@mul_exercise_2015; @sebata_control_2018].
+Furthermore, energy availability is a decisive factor in the *de novo* synthesis of ribosomes [@kusnadi_regulation_2015; @moss_housekeeper_2007; @sebata_control_2018] which in turn determines muscle growth by increasing the muscle's translational capacity [@figueiredo_regulation_2019; @hammarstrom_benefits_2020; @sebata_control_2018; @stec_ribosome_2016; @von_walden_ribosome_2019].
+In addition, insulin may have anabolic effects with elevated amino acids, and has a clear role in the reduction of muscle protein breakdown independent of amino acid availability [@abdulla_role_2016; @hillier_physiological_2000].
 Therefore, investigating the effects of combined RT and glucose ingestion may provide valuable insight into the potential additive effect of glucose and RT on ribosome biogenesis.
 
-Transcription of ribosomal RNA (rRNA) by RNA Polymerase I is considered the rate-limiting step in synthesising new ribosomes [@moss_promotion_1995].
-Multiple signalling pathways converge to regulate rRNA transcription, including the mammalian target of rapamycin (mTORC1) signal-transduction pathway and c-Myc [@kusnadi_regulation_2015;@von_walden_ribosome_2019;@sebata_control_2018; @west_acute_2016; @mori_c-myc_2021].
-mTOR receives input from growth factors, hormones, mechanical loading, and nutrients to balance protein synthesis through multiple mechanisms based on cellular energy levels [@hoppe_amp-activated_2009].
-Importantly, mTOR and its downstream target ribosomal protein S6 kinase beta-1 (S6K1) serves as a mediator for insulin signalling in skeletal muscle [@hillier_physiological_2000].
-The mTORC1 pathway contributes to ribosome biogenesis through the regulation of the translation of ribosomal proteins and to forming of the preinitiation complex (PIC) that marks the initiation of rRNA transcription [@figueiredo_regulation_2019;@von_walden_ribosome_2019].
-Parallel to mTORC1 [@mori_c-myc_2021; @west_acute_2016], c-Myc increases ribosomal biogenesis directly and through transcriptional control of the upstream binding factor (UBF) [@sanij_ubf_2008; @poortinga_c-myc_2011].
-Additionally, UBF phosphorylation, required for interaction with the ribosomal deoxyribonucleic acid (rDNA) promoter, is increased by high glucose in a mTORC1 dependent manner (rapamycin sensitive) in kidney glomerular epithelial cells [@mariappan_ribosomal_2011].
-Independently from UBF and mTORC1, high glucose was shown to lead to chromatin remodelling, which promoted rRNA transcription in cell cultures [@zhai_human_2012].
-Together these observations indicate a potential role of glucose in positively affecting muscle rRNA synthesis.
-
-The purpose of this investigation was to test the hypothesis that glucose supplementation before and after RT increases markers of ribosomal abundance following five RT sessions.
-Secondly, we aimed to describe the association between UBF and total RNA abundance in human skeletal muscle.
-There have been multiple studies recently suggesting translational capacity to be as important, if not more important than, translational efficiency concerning long-term skeletal muscle adaptations to RT [@hammarstrom_benefits_2020;@hammarstrom_ribosome_2022;@figueiredo_revisiting_2019].
-Previously, *in vitro* studies have shown UBF and c-Myc to be important factors in the PIC, c-Myc as a general transcription factor and UBF as a specific transcription factor for rDNA transcription initiation [@mariappan_ribosomal_2011;@von_walden_ribosome_2019].
+Ribosomal biogenesis and content seem to be a prerequisite for skeletal muscle growth, and transcription of ribosomal ribonucleic acid (rRNA) by RNA Polymerase I is considered the rate-limiting step in synthesising new ribosomes [@moss_promotion_1995].
+Multiple signalling pathways converge to regulate rRNA transcription, including c-Myc and the mammalian target of rapamycin complex 1 (mTORC1) signal-transduction pathway [@kusnadi_regulation_2015; @von_walden_ribosome_2019; @sebata_control_2018; @west_acute_2016; @mori_c-myc_2021].
+The general transcription factor c-Myc increases ribosomal biogenesis directly through transcriptional control of the upstream binding factor (UBF) [@sanij_ubf_2008; @poortinga_c-myc_2011; @mori_c-myc_2021; @west_acute_2016].
+Indeed, UBF phosphorylation, which is required for interaction with the ribosomal deoxyribonucleic acid (rDNA) promoter, seems to be increased by high-glucose treatment in a mTORC1-dependent manner (rapamycin sensitive) in kidney glomerular epithelial cells [@mariappan_ribosomal_2011].
+Independent of UBF and mTORC1, high glucose was shown to lead to chromatin remodelling, which promoted rRNA transcription in cell cultures [@zhai_human_2012].
+The mTORC1 pathway receives input from growth factors, hormones, mechanical loading, and nutrients to balance protein synthesis through multiple mechanisms based on cellular energy levels [@hoppe_amp-activated_2009] and contributes to ribosome biogenesis through the forming of the preinitiation complex (PIC) that marks the initiation of rRNA transcription and through the regulation of ribosomal protein translation [@figueiredo_regulation_2019; @von_walden_ribosome_2019].
+Moreover, mTORC1 and its downstream target ribosomal protein S6 kinase beta-1 (S6K1) serves as direct mediators of insulin signalling in skeletal muscle [@hillier_physiological_2000].
+Together these observations indicate a potential role of glucose in positively affecting human skeletal muscle rRNA synthesis, through potentiating transcription or translation of ribosomal RNA and increasing the translational capacity of skeletal muscle cells [@sebata_control_2018; @zhai_human_2012; @hillier_physiological_2000; @hoppe_amp-activated_2009].
+
+There have been multiple studies recently suggesting translational capacity to be as important, if not more important than, translational efficiency concerning long-term skeletal muscle adaptations to RT [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022; @figueiredo_revisiting_2019].
+Previously, *in vitro* studies have shown UBF and c-Myc to be important factors in the PIC, c-Myc as a general transcription factor and UBF as a specific transcription factor for rDNA transcription initiation [@mariappan_ribosomal_2011; @von_walden_ribosome_2019].
 Coupled with the recent observations by Hammarström et al.
-[@hammarstrom_benefits_2020;@hammarstrom_ribosome_2022], understanding underlying factors regulating and affecting ribosome biogenesis seems key to furthering our understanding of the optimisation of RT to individual phenotypes.
+[@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022], understanding underlying factors regulating and affecting ribosome biogenesis seems key to furthering our understanding of the optimisation of RT to individual phenotypes.
+
+Therefore, the purpose of this investigation was to test the hypothesis that glucose supplementation before and after five RT sessions over 12 days will potentiate RT-associated accumulation of markers of ribosomal abundance following five RT sessions.
+Secondly, we aimed to describe the association between changes in total RNA abundance and UBF in human skeletal muscle.
 
@@
 Glucose (Glucosum monohydricum, Merck KGaA, Darmstadt, Germany) and placebo (Steviosa, Soma Nordic AS, Oslo, Norway) were masked by mixing with 300ml Fun Light (Orkla, Oslo, Norway).
-A blinded taste test revealed that the participants were not able to disclose the contents of the provided boluses (30g glucosum monohydricum vs.
-~0.3g Stevia rebaudiana extract).
+A blinded taste test revealed that the participants were not able to disclose the contents of the provided boluses (30g glucoseum monohydricum vs.
+\~0.3g Stevia rebaudiana extract).
 To ensure equal conditions during training sessions and strength testing, participants exercised and tested at the same time of day, +/- 1hr with the same supervisor on pairwise consecutive days (i.e.
@@
 Two muscle biopsy aliquots were used for total RNA extraction per leg, resulting in a total of eight RNA samples per participant.
-Total RNA was extracted using TRIzol with muscle tissue homogenised using 0.5mm RNase-free Zirconium beads (~50 ul; Next Advanced, Averill Park, NY, USA) and mechanical agitation (Bullet Blender, Next Advanced, Averill Park, NY, USA).
+Total RNA was extracted using TRIzol with muscle tissue homogenised using 0.5mm RNase-free Zirconium beads (\~50 ul; Next Advanced, Averill Park, NY, USA) and mechanical agitation (Bullet Blender, Next Advanced, Averill Park, NY, USA).
 Chloroform (Sigma-Aldrich, Oslo, Norway) was used for phase separation, and the RNA pellet was precipitated with isopropanol (VWR International, Oslo, Norway).
 To enable analysis of target gene expression per unit tissue weight [@ellefsen_improved_2008; @ellefsen_reliable_2014], an exogenous RNA control (Lambda, λ polyA External Standard Kit, Takara Bio Inc., Shiga, Japan) was added at a fixed amount to each sample (0.04 ng ml^-1^ of TRIzol reagent).
-For assessment of RNA content and purity, RNA was eluted in TE-buffer (1:2) and assessed via spectrophotometry.
-All samples had a 260 nm to 280 nm ratio > 1.9.
+For assessment of RNA content and purity, RNA was eluted in TE buffer (1:2) and assessed via spectrophotometry.
+All samples had a 260 nm to 280 nm ratio \> 1.9.
 The RNA stock was stored at -80°C until further analyses.
@@
 
-There was no difference in mean change of total training session volume between exercising with placebo or glucose, with an increase of 18 and 16% from session 1 to session 6, respectively (*p* > 0.05).
-There were no differences in mean macronutrient intake (protein, fat, carbohydrates) or total calorie intake between glucose and placebo on pairwise consecutive days (*p* > 0.05 for all).
+Glucose ingestion before and after RT sessions did not lead to a higher mean change of total training session volume, with a mean increase of 17% in both the glucose condition (pre: 5262 ± 1799kg, post: 6319 ± 2256kg, *p* > 0.05) and the placebo condition (pre: 5351 ± 1615kg, post: 6438 ± 2092, *p* > 0.05) from session 1 to session 6, respectively.
+There were no differences in mean macronutrient intake (protein, fat, carbohydrates) or total calorie intake between glucose and placebo on pairwise consecutive days (*p* \> 0.05 for all, raw data and table available at GitHub repository).
 
@@
 
-Glucose ingestion before and after RT led to significant increases in plasma glucose levels by 38% immediately before RT (Figure 1B, 0 min), by 31% during RT (Figure 1B, 15 min) and by 32% immediately after RT (Figure 1B, 30 min; all : *p* < 0.001), compared to placebo.
-Two hours after the session, RT with glucose measured 8% lower plasma glucose levels compared to RT with placebo (Figure 1B, 270min: *p* = `r p.270`).
-Similarly, levels of c-peptide increased with glucose ingestion over placebo, by 95% immediately before (Figure 1C, 0 min) and 87% after RT (Figure 1C, 30 min; both *p* < 0.001).
+Glucose ingestion before and after RT led to increases in plasma glucose levels compared to baseline by 38% immediately before RT (Figure 1B, 0 min), by 31% during RT (Figure 1B, 15 min) and by 32% immediately after RT (Figure 1B, 30 min; all : *p* \< 0.001), with no changes observed in the placebo condition (*p* \> 0.05).
+Compared to the placebo condition, ingestion of glucose increased plasma glucose levels by 36% immediately before RT (Figure 1B, 0 min), by 27% during RT (Figure 1B, 15 min) and by 28% immediately after RT (Figure 1B, 30 min; all : *p* \< 0.001).
+Two hours after the RT session, glucose ingestion was associated with 12% lower plasma glucose levels compared to baseline, and 8% lower compared to placebo (Figure 1B, 270min: *p* = `r p.270`).
+
+Glucose ingestion before and after RT led to increases in levels of c-peptide compared to baseline, by 95% immediately before (Figure 1C, 0 min) and 87% after RT (Figure 1C, 30 min; both *p* \< 0.001), with no changes observed with the placebo condition (*p* \> 0.05).
+Compared to the placebo condition, ingestion of glucose increased levels of c-peptide by 85% immediately before (Figure 1C, 0 min) and 85% after RT (Figure 1C, 30 min; both *p* \< 0.001).
 
@@
 
-After 5 RT sessions (Figure 1D, Post 5RT), there was a significant difference in strength between glucose and placebo,  where placebo decreased strength by 7% more than glucose (Figure 1D, *p* = `r pdiff.c4`).
-At the other time points, there were no significant differences in mean change between exercising with glucose and placebo (Figure 1D, Post 2RT: *p* = `r pdiff.c2`, Post 4RT: *p* = `r pdiff.c3`, 30min post 6RT: *p* = `r pdiff.c5`, 2h post 6RT: *p* = `r pdiff.c6`, 23h post 6RT: *p* = `r pdiff.c7`).
+Glucose ingestion before and after RT sessions did not generally improve skeletal muscle recovery throughout the intervention, neither during the intervention in a rested state (Figure 1D, 23hrs after exercise; Post 2RT: *p* = `r pdiff.c2`, Post 4RT: *p* = `r pdiff.c3`), nor acutely after the sixth and final RT-session (30min post 6RT: *p* = `r pdiff.c5`, 2h post 6RT: *p* = `r pdiff.c6`) or in a rested state after the sixth RT session (23h post 6RT: *p* = `r pdiff.c7`).
+Glucose ingestion before and after RT compared to placebo was associated with a 7% less reduction in strength after five RT sessions (Figure 1D, *p* = `r pdiff.c4`).
+
 Both RT with glucose and placebo led to significantly reduced strength after the fifth session compared to baseline, by 11 and 18% respectively (Figure 1D, Post 5RT: *p* = 0.000).
-Comparisons of the acute data gathered from after five sessions until and including 23 hours after the sixth session showed a significant mean increase in strength of 5-9% from RT with glucose and placebo 30 minutes after the sixth RT session (Figure 1D, 30min post 6RT: *p* = `r apc2`) and two hours after the sixth RT session (Figure 1D, 2h post 6RT: *p* = `r apc3`).
+Comparisons of the acute data gathered from after five sessions until and including 23 hours after the sixth session showed an average increase in strength of 5-9% from RT with glucose and placebo 30 minutes after the sixth RT session (Figure 1D, 30min post 6RT: *p* = `r apc2`) and two hours after the sixth RT session (Figure 1D, 2h post 6RT: *p* = `r apc3`).
 Twenty-three hours after the last (sixth) RT session, strength was unchanged compared to after the fifth RT session (Figure 1D, 23h post 6RT: *p* = `r apc4`).
@@
 
-From baseline to after the intervention total RNA and rRNA per unit muscle weight increased on average by ~20-27% and ~25-57%, respectively (Figure 2).
+The five-session-RT intervention led to increases in total RNA and rRNA per unit muscle weight both in the glucose and the placebo condition, on average by \~20-27% and \~25-57%, respectively (Figure 2).
 However, RT with glucose did not induce increased accumulation of total RNA (Figure 2A, mean difference `r total_rna_stats`) or rRNA (Figure 2B; 47S, `r stat_rrna$trg_47s`; 18S, `r stat_rrna$trg_18s`; 28S, `r stat_rrna$trg_28s`; 5.8S, `r stat_rrna$trg_5.8s`; 5S, `r stat_rrna$trg_5s`) compared to RT with placebo.
@@
 
-Overall, RT led to robust increases from before to after the intervention in all measured proteins (Figure 3A).
+The five-session-RT intervention led to increases in all measured proteins both in the glucose and the placebo condition (Figure 3A).
 RT with glucose resulted in estimated levels of c-Myc, UBF and RPS6 being `r stat_west$cmyc`, `r stat_west$ubf` and `r stat_west$rps6`% lower compared to placebo, respectively, without showing statistical significance (*p* = `r min(p_west)`-`r max(p_west)`; Figure 3A).
-There was a linear relationship between UBF and total RNA, where an increase of 1 SD unit of UBF equated to a $\sim$ `r corr_incr`% increase in total RNA (*p* = `r corrp`; Figure 3C).
+Both baseline and trained state data of total RNA showed a linear relationship with UBF, where $\sim$ `r corr_incr`% increase in total RNA corresponded to 1 standard deviation unit increase in UBF (*p* = `r corrp`; Figure 3C).
 
@@
 Similarly, RT with glucose in general did not affect markers of skeletal muscle functionality such as strength and recovery, or total training session volume.
-At one time point, RT with glucose led to less reduction in strength compared to placebo.
-Levels of plasma glucose and c-peptide were significantly higher before, during and after the RT sessions with glucose.
-Importantly, there were no differences in daily macronutrient intake on pairwise consecutive days.
+Towards the end of the intervention, RT with glucose led to less reduction in strength compared to placebo, which may point to an accumulated effect of the glucose condition.
+As by design, the levels of plasma glucose and c-peptide were significantly higher before, during and after the RT sessions with the glucose condition compared to placebo, and there were no differences in the daily macronutrient intake between conditions on consecutive days, thus enabling an effective design to study the biological and functional effects of the glucose condition.
 These findings suggest that the effects of high vs.
-low glucose/glucose starvation conditions on rDNA transcription initiation observed in previous *in vitro* studies [@mariappan_ribosomal_2011; @tanaka_mild_2015] are not translatable to acute effects in human skeletal muscle *in vivo*.
-Notably, the present study was specifically designed to investigate differences in pre to post changes within participants and is as such also quite limited to this perspective.
-Hence, the observed effects of time serve as supplementary to the efficiency of the design, without a negative control group to compare with.
-
-To our knowledge, no previous study has investigated the effects of glucose compared to placebo ingestion with RT on ribosome biogenesis in human skeletal muscle previously, making our hypotheses exploratory and comparisons few.
-*In vitro* studies on mice glomerular epithelial cells [@mariappan_ribosomal_2011] and human breast cancer cells [@tanaka_mild_2015] have observed glucose-induced UBF-dependent augmentation of ribosome biogenesis [@mariappan_ribosomal_2011] and reduced rRNA transcription by glucose starvation [@tanaka_mild_2015].
-These findings do imply an important role for glucose in the initiation of rDNA transcription, the rate-limiting step of ribosome biogenesis [@mariappan_ribosomal_2011; @tanaka_mild_2015].
-In addition, Hillier et al.
-found physiological hyperinsulinemia to stimulate p70S6K phosphorylation in human skeletal muscle [@hillier_physiological_2000].
-Moreover, robust increases in markers of ribosome biogenesis such as 47S pre-rRNA and mature rRNA can be expected after an acute bout of RT [@figueiredo_impact_2016] or a short period of RT [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022].
-Therefore, we hypothesized that if glucose can, directly or indirectly via insulin, stimulate the initiation of rDNA transcription, five sessions of RT with glucose compared to placebo would elicit this effect.
-Further, glucose ingestion increases energy levels, leading us to hypothesize that high glucose vs.
-placebo condition would stimulate energy-sensitive pathways such as mTORC1, PIH1, extracellular signal-regulated kinase 1/2 (ERK1/2), AMP-dependent protein kinase (AMPK) and Sirtuin 1 (SIRT1) [@kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012].
+low glucose/glucose starvation conditions on rDNA transcription initiation observed in previous *in vitro* studies [@mariappan_ribosomal_2011; @tanaka_mild_2015] are not translatable to acute effects in human skeletal muscle *in vivo*, with a design like the present study.
+
+The observations of the present study from human skeletal muscle do not support previous indications from non-human skeletal muscle [@mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012].
+Although we observed a linear relationship between baseline and trained state UBF and total RNA levels, a UBF-dependent augmentation of ribosome biogenesis did not seem to be induced by glucose ingestion in the present study, as compared to in mice glomerular epithelial cells [@mariappan_ribosomal_2011] and in human breast cancer cells [@tanaka_mild_2015].
+Neither do our findings indicate that insulin *per se*, at least at physiological levels, potentiates accumulation of total RNA through p70S6K stimulation as observed with hyperinsulinemia in human skeletal muscle [@hillier_physiological_2000].
+Increases in markers of ribosome biogenesis such as 47S pre-rRNA and mature rRNA are expected to occur after a single session of RT [@figueiredo_impact_2016], as well as after a short period of RT [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022].
+Therefore, in the present study, it was expected that eventual benefits of ingesting glucose compared to placebo with RT would be measurable after 5 training sessions, either due to glucose-induced stimulation of energy-sensitive pathways such as mTORC1, PIH1, extracellular signal-regulated kinase 1/2 (ERK1/2), AMP-dependent protein kinase (AMPK) or Sirtuin 1 (SIRT1) [@kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012].
 Despite previously reported upregulation in PIC assembly due to high-glucose mediated mTORC1, ERK1/2 and PIH1 or low-glucose mediated AMPK and SIRT1 activation [@hoppe_amp-activated_2009; @kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012], the present study showed no signs of such effects of glucose vs.
 placebo conditions.
-Resistance training, irrespective of condition, yielded a robust accumulation of total RNA and expression of rRNA, in line with previous observations [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022].
-Due to the exploratory nature of our hypotheses, this was hardly a very surprising result.
-Indeed, previous studies have used cell cultures from yeast [@zhai_human_2012], rodents [@hoppe_amp-activated_2009; @mariappan_ribosomal_2011] or human cells [@tanaka_mild_2015] and are, as such, not directly comparable to human skeletal muscle cells.
 Importantly, previous studies investigated high vs.
@@
 normal plasma glucose levels, to investigate the effect of glucose *per se* and not intracellular energy status.
-Thus, while glucose ingestion presumably is important in supplying energy for growth-inducing processes such as ribosome biogenesis [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @sebata_control_2018] there is no apparent effect of ingesting glucose *per se* on markers of ribosome biogenesis.
-
-In the present study, despite not measuring the activity in central pathways mediating anabolic signalling (mTORC1, ERK1/2), analyses of the downstream target UBF and the ribosomal protein S6 (rpS6), as well as the general transcription factor c-Myc, further supported observations from the RNA data.
+Thus, while glucose ingestion presumably is important for supplying energy for growth-inducing processes such as ribosome biogenesis [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @sebata_control_2018] there is no apparent effect of ingesting glucose *per se* on markers of ribosome biogenesis, during 12 days of heavy-load RT.
+Further, previous studies have used cell cultures from yeast [@zhai_human_2012], rodents [@hoppe_amp-activated_2009; @mariappan_ribosomal_2011] or human breast cancer cells [@tanaka_mild_2015] and are, as such, not directly comparable to human skeletal muscle cells.
+Nevertheless, resistance training irrespective of condition yielded a robust accumulation of total RNA and expression of rRNA, in line with previous observations for heavy-load RT [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022].
+
+In the present study, despite not measuring the activity in central pathways mediating anabolic signalling (mTORC1, ERK1/2), analyses of the downstream target UBF and the ribosomal protein S6 (rpS6), as well as the general transcription factor c-Myc, further supported observations from the RNA data as there was no difference between the glucose and placebo conditions.
 UBF has previously been described as a master regulator of rDNA transcription *in vitro* [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @russell_rna-polymerase-i-directed_2005], while rpS6 is proposed as a valid and reliable means to measure ribosome biogenesis [@chaillou_hypoxia_2012; @nakada_correlation_2016].
@@
 Hence, it seems quite reasonable to observe similar changes in these three proteins.
-The linear relationship found exclusively between UBF content and total RNA accumulation, and neither c-Myc nor rpS6, supports the importance and specificity of UBF content in human skeletal muscle ribosome biogenesis.
-While this is not a novel finding, it has only recently been observed in human skeletal muscle following RT [@hammarstrom_ribosome_2022].
+The linear relationship found exclusively between UBF content and total RNA levels, and neither c-Myc nor rpS6, supports the importance and specificity of UBF content in human skeletal muscle ribosome biogenesis.
+While this is not a novel finding, it was only recently observed in human skeletal muscle following a period of RT [@hammarstrom_ribosome_2022].
 As such, UBF also seems to respond to mechanical loading in human muscle cells in line with responses seen in cell cultures and synergist ablation models [@mariappan_ribosomal_2011; @von_walden_mechanical_2012], in addition to the previously reported response to hormonal, nutritional, and cellular energy signals [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @russell_rna-polymerase-i-directed_2005].
 
-As in the biological data, there was generally no difference in the change of strength as an index between the conditions.
-Generally, skeletal muscle strength decreased from baseline to after the fifth and -sixth RT sessions, with a significantly less reduction in glucose compared to placebo at the measurement after five sessions.
-This could suggest that the RT protocol fatigued the participants, as both training frequency and intensity were quite high.
-Another promising explanation might be the biphasic recovery pattern, as described by Raastad & Hallén [@raastad_recovery_2000], where the participants experienced a rapid recovery during the initial 11 hours after exercise, followed by a levelling off or drop until 22 hours after exercise.
+As in the biological data, glucose ingestion had no effect throughout the intervention on muscular strength compared to placebo.
+Generally, skeletal muscle performance measured by a strength index, used as a proxy marker for muscular recovery, decreased from baseline to after the condition in both conditions.
+Glucose ingestion was associated with a less reduction in muscular strength after five RT sessions compared to placebo, which may point towards a beneficial accumulated effect where the heavy-load RT fatigued the participants but glucose ingestion may have improved muscular recovery [@mul_exercise_2015; @sebata_control_2018].
+However, glucose did not improve muscular performance/recovery acutely following one RT session compared to placebo, measured 30 minutes, 2 hours and 23 hours after the sixth and final training session.
+As such, the potential accumulated effect observed after five RT sessions did not extend to after session six.
+Therefore, we can't rule out a possible long-term effect of ingesting glucose when exercising heavy-load resistance training, which would require a longer intervention period than that of the present study.
+Another possible explanation for the decrease in muscular performance during the intervention might be the biphasic recovery pattern, as described by Raastad & Hallén [@raastad_recovery_2000], where the participants experienced a rapid recovery during the initial 11 hours after exercise, followed by a levelling off or drop until 22 hours after exercise.
 Herein, inflammation and phagocytic activity were proposed to be involved in the performance drop between 11-22 hours [@raastad_recovery_2000].
@@
 Further, strength testing during the intervention was conducted 23 hours after RT, meaning that the biphasic recovery may have also influenced these tests.
-However, this does not explain the difference between conditions at observed after five RT session, or the drop in strength from after four RT sessions to after five RT sessions.
+However, this does not explain the difference between conditions observed after five RT sessions, or the drop in strength from after four RT sessions to after five RT sessions.
 A possible argument could be that exercising without glucose may have caused more stress compared to exercising with glucose, as glucose is the preferred energy source during strenuous exercise [@mul_exercise_2015], thus increasing performance with glucose compared to placebo.
@@
 Therefore, discussing the potential effect of differences in metabolic stress between conditions would only be speculation, however probable.
-Lastly, these inconclusive observations may also reflect our limited sample size, despite using a within-subjects design.
-Initially, we met our minimum limit of 16 participants.
-Unfortunately, three dropped out during the intervention, leaving our statistical analyses slightly underpowered according to the a priori calculation.
-Moreover, we did not measure the intramuscular glycogen storage, representing another limitation to our design.
-Hence, we cannot know for certain whether there were actual differences within the exercising skeletal muscles between the glucose and placebo conditions.
-The participants did indeed match their macronutrients on pairwise consecutive days, and the plasma glucose measurements did not show any differences in blood sugar until ingestion of placebo/glucose.
-Thus, it may stand to reason that the participants had similar levels of intramuscular levels before ingesting any supplement and that the glucose supplement led to higher glucose availability during exercise, as by design.
-However, even if the leg exercising with glucose supplement did in fact take up the extra glucose, it did not seem to affect any of our main outcome measures.
+
+*Limitations and Strengths*
+
+The present study was designed specifically to investigate the acute biological and functional effects of ingesting glucose compared to placebo, with unilateral training and testing to reduce biological variation between participants in our analyses [@macinnis_investigating_2017].
+Further, to ensure that the legs were exercised under the same conditions apart from glucose/placebo during exercise, macronutrients, the time of day and test/training personnel were standardized for each participant on pairwise consecutive days[@halperin_threats_2015].
+Every day, the participants showed up in an overnight fasted state and ingested only either protein and glucose or protein and placebo before and during exercise.
+As evident from the blood data, glucose ingestion thus led to higher levels both of plasma glucose and serum c-peptide compared to placebo, together with the aforementioned standardisations enabling analyses of the effect of ingesting glucose on total- and specific RNA levels and proteins, as well as muscular performance, within-participant.
+The present study also had a few limitations; firstly, the sample size was smaller than expected and planned for.
+Initially the minimum limit of 16 participants, according to the a priori power calculation, was met.
+However, three dropped out during the intervention, possibly leaving our statistical analyses slightly underpowered.
+Moreover, the intramuscular glycogen stores were not measured, hence it can't be determined whether the glucose ingestion increased intramuscular glucose.
+Though it may be a reasonable assumption that the glucose ingestion in this design did lead to increased intramuscular glucose and thus energy levels, it cannot be excluded that the participants' intramuscular glycogen stores were topped up from the previous day, and as such, ingesting more glucose had no further benefit.
+Nevertheless, even if the skeletal muscle of the leg exercising with glucose supplement took up the extra glucose, it did not seem to affect any of our main outcome measures.
 
@@
 
-In conclusion, ingestion of glucose immediately before and after resistance exercise training does not acutely augment ribosome biogenesis after two weeks of heavy resistance exercise training, in moderately trained young adults.
-Neither does it acutely enhance muscular performance during the exercise period or after five sessions, nor recovery within 23 hours of the last session.
-The observations in total RNA, rRNA and rDNA transcription initiation-associated proteins support the key role of UBF in ribosome biogenesis regulation in human skeletal muscle following resistance training.
-If future investigations are to be made on this topic, a greater sample size coupled with a negative control group and analyses of the intramuscular glycogen storage may provide higher-resolution results.
+In conclusion, ingestion of glucose immediately before and after five heavy-load resistance training sessions conducted over 12 days did not augment accumulation of ribosomal RNA, in moderately trained young adults compared to ingestion of placebo.
+Glucose ingestion did not affect muscular performance throughout the study, nor did it affect muscular performance measured 30 minutes, 2 hours or 23 hours after the last session.
+Glucose ingestion was associated with less reduction in muscular performance 23 hours after the fifth training session, thus we can't rule out a possible accumulated effect of ingesting glucose compared to placebo.
+The relationship between baseline and trained state data of total RNA and UBF levels do support the key role of UBF in ribosomal biogenesis regulation in human skeletal muscle following resistance training.
+Future investigations should aim for a greater sample size over a longer time period, to investigate the potential accumulated effects of glucose ingestion, as well as analyses of the intramuscular glycogen storage to provide higher-resolution results.
 
@@
 
-Complete data sets and scripts are downloadable here; https://github.com/Kristianlian/ribose-paper
+The datasets generated during and analysed during the current study are available in the "ribose-paper" repository; <https://github.com/Kristianlian/ribose-paper>
 
@@
 
-Extended methods can be found here; https://github.com/Kristianlian/ribose-paper
+Supplementary material can be found in the readme file of the "ribose-paper" repository, available here; <https://github.com/Kristianlian/ribose-paper>
 
 \pagebreak
+
 ## References
 
-<div id="refs"></div>
+::: {#refs}
+:::
 

### 2023-10-18 README.rmd
repository organization

@@
 
-### Supplementary figures
+## Supplementary figures
 
@@
 
+#### Macronutrients
+
 **Figure 4: Daily change in macronutrient intake per condition.** For the purpose of illustration the first day with the glucose condition and the first day with the placebo condition are combined into "Day 1", though they were completed on separate days (i.e., day 1 and 2).
@@
 Day 6: Glu = 11, Plac = 11.
+
 ```{r nutrition table, echo=FALSE, message=FALSE, warning=FALSE}
@@
 ```
+
+#### Training volume
+
+```{r total session volume figure, echo=FALSE, message=FALSE, warning=FALSE}
+
+### Total session volume change
+
+# Absolute changes are presented as mean change ± SD per supplement
+
+# Packages
+library(readxl);library(tidyverse);library(knitr);library(lme4);library(broom);library(emmeans)
+
+# Data
+tot.vol <- read_excel("./data/training/ribose_volume.xlsx", na = "NA") %>%
+  select(subject, timepoint, tot.volume, supplement)
+
+## Handling the data by creating a new factor called time from timepoint. This factor combines any observation at T1 and T2 to baseline, etc. 
+# The code also sorts the order of the factor time, from baseline to session 6, using time = factor(time, levels c()), and sets placebo to be compared to 
+# glucose via supplement = factor(supplement, levels = c()).
+
+tot.volh <- tot.vol %>%
+  mutate(time = if_else(timepoint %in% c("T1", "T2"),
+                        "baseline",
+                        if_else(timepoint %in% c("D3", "D4"),
+                                "session2",
+                                if_else(timepoint %in% c("D5", "D6"),
+                                        "session3",
+                                        if_else(timepoint %in% c("D7", "D8"),
+                                                "session4",
+                                                if_else(timepoint %in% c("D9", "D10"),
+                                                        "session5",
+                                                        if_else(timepoint %in% c("T3", "T4"),
+                                                                "session6", timepoint))))))) %>%
+  mutate(time = factor(time, levels = c("baseline", "session1", "session2", "session3", "session4", "session5", "session6")),
+         supplement = factor(supplement, levels = c("placebo", "glucose")))
+
+## Baseline analysis - comparison of the two legs
+# A baseline analysis comparing training volume at baseline sessions between the two legs via a paired t.test, and providing a summary of mean training
+# volume and sd
+
+base.vol <- tot.volh %>%
+  filter(time == "baseline") %>%
+  select(subject, time, supplement, tot.volume) %>%
+  group_by(supplement) %>%
+  pivot_wider(names_from = supplement,
+              values_from = tot.volume)
+
+vol.ttest <- t.test(base.vol$glucose, base.vol$placebo, paired = TRUE)
+
+vol.summary <- tot.volh %>%
+  filter(time == "baseline") %>%
+  select(subject, time, supplement, tot.volume) %>%
+  group_by(supplement) %>%
+  mutate(m = mean(tot.volume),
+         s = sd(tot.volume))
+
+## Absolute data - summarising mean and SD for barplot
+# Mean and SD are calculated from absolute (kg) total volume, creating a basis for a barplot.
+
+vol.exp <- tot.volh %>%
+  select(subject, supplement, time, tot.volume) %>% 
+  group_by(supplement, time) %>%
+  summarise(mean.vol = mean(tot.volume),
+            sd.vol = sd(tot.volume))
+
+totvol.barplot <- ggplot(vol.exp, aes(fill = supplement, y = mean.vol, x = time)) +
+  #annotate("text", x = c("session4", "session5", "session6"),
+   #        y = c(8500, 8500, 8700), label = "†") +
+  geom_bar(position = "dodge", stat = "identity") +
+  geom_errorbar(aes( ymin = mean.vol - sd.vol, ymax = mean.vol + sd.vol),
+                width = 0.2,
+                position = position_dodge(width = 1)) +
+  labs(x = "", y = "Training volume \n(kg)\n", fill = "Supplement") +
+  scale_x_discrete(labels=c("baseline" = "Baseline", "session2" = "Session 2", "session3" = "Session 3",
+                            "session4" = "Session 4", "session5" = "Session 5", 
+                            "session6" = "Session")) +
+  #scale_y_continuous(limits = c(0, 10), breaks = c(0, 1, 2, 3, 4, 5, 6,
+  #                                                7, 8, 9, 10),
+  #                 expand = expand_scale(0)) +
+  theme_classic() +
+  theme(axis.text.x = element_text(size = 7, angle = 45, hjust = 1))
+
+totvol.barplot
+
+```
+
+**Figure 5:** Changes in total training session volume from baseline until resistance exercise session 6 (training day 11 and 12).
+Baseline = Training day 1/2, Session 2 = Training day 3/4, Session 3 = Training day 5/6, Session 4 = Training day 7/8, Session 5 = Training day 9/10, Session 6 = Training day 11/12.
+Values are mean kg ± standard deviation.

### 2023-10-18 R/scraps/nutrition.table.R

@@ -1,0 +1,91 @@
+### Nutrition data table
+
+# This script produces a table of the daily nutritional intake, calculated to mean and SD per supplement.
+
+library(knitr)
+library(rmarkdown)
+library(tidyverse)
+library(readxl)
+library(readr)
+library(tinytex)
+library(tidyr)
+library(knitr)
+library(kableExtra)
+
+#nutritiontable with pro.pr kg bodyweight pr. group
+
+#laod dxa data and take out weight and subject to join with nutrition data
+fp.weight <-read_excel("data/dxa/ribose_dxa.xlsx") %>%
+  select(weight, subject)
+
+#nutrition data.
+# Data are presented as grams, supplemented protein (25g whey protein isolate = 42g protein) and glucose (90g) has been added by investigators,
+# and calculated into the total amount of ingested per day.
+
+# Load nutrition data. Sorts data and summarises into mean per supplement per day
+
+nut.weight <- read_excel("data/nutrition/ribose_nutrition.xlsx")%>%
+  select(timepoint, meal, subject, protein, fat, calories,
+         carbohydrates, group, sup_pro, sup_gluc, kcal_glu) %>%
+  #changed names to compare groups each paired days
+  mutate(timepoint = if_else(timepoint %in% c("T1","T2"), 
+                             "Day 1",
+                             if_else(timepoint %in% c("3", "4"),
+                                     "Day 2",
+                                     if_else(timepoint %in% c("5", "6"),
+                                             "Day 3",
+                                             if_else(timepoint %in% c("7", "8"),
+                                                     "Day 4",
+                                                     if_else(timepoint %in% c("9", "10"),
+                                                             "Day 5",
+                                                             if_else(timepoint %in% c("T3", "T4"),
+                                                                     "Day 6",
+                                                                     "na"))))))) %>%
+  group_by(timepoint, subject, group) %>%
+  #take in weight
+  inner_join(fp.weight) %>%
+  #summarise each variable to get total pr timepoint.
+  #Whey protein supplement is added to total protein ingestion
+  summarise(protein = sum(protein + sup_pro),
+            fat = sum(fat),
+            calories = sum(calories + kcal_glu),
+            carbohydrates = sum(carbohydrates + sup_gluc),
+            weight = sum(weight)) %>%
[…]

### 2023-10-18 R/scraps/training.volume.R

@@ -1,0 +1,158 @@
+### Total volume change, both absolute values and fold-change
+
+# This script provides log-fold change score calculation of training volume, and illustration of absolute and fold change in training volume. 
+# Absolute changes are presented as mean change ± SD per supplement, fold changes are represented as mean change ± CI. 
+
+# Packages
+library(readxl);library(tidyverse);library(knitr);library(lme4);library(broom);library(emmeans)
+
+# Data
+tot.vol <- read_excel("./data/training/ribose_volume.xlsx", na = "NA") %>%
+  select(subject, timepoint, tot.volume, supplement) 
+
+## Handling the data by creating a new factor called time from timepoint. This factor combines any observation at T1 and T2 to baseline, etc. 
+# The code also sorts the order of the factor time, from baseline to session 6, using time = factor(time, levels c()), and sets placebo to be compared to 
+# glucose via supplement = factor(supplement, levels = c()).
+
+tot.volh <- tot.vol %>%
+  mutate(time = if_else(timepoint %in% c("T1", "T2"),
+                        "baseline",
+                        if_else(timepoint %in% c("D3", "D4"),
+                                "session2",
+                                if_else(timepoint %in% c("D5", "D6"),
+                                        "session3",
+                                        if_else(timepoint %in% c("D7", "D8"),
+                                                "session4",
+                                                if_else(timepoint %in% c("D9", "D10"),
+                                                        "session5",
+                                                        if_else(timepoint %in% c("T3", "T4"),
+                                                                "session6", timepoint))))))) %>%
+  mutate(time = factor(time, levels = c("baseline", "session1", "session2", "session3", "session4", "session5", "session6")),
+         supplement = factor(supplement, levels = c("placebo", "glucose")))
+
+## Baseline analysis - comparison of the two legs
+# A baseline analysis comparing training volume at baseline sessions between the two legs via a paired t.test, and providing a summary of mean training
+# volume and sd
+
+base.vol <- tot.volh %>%
+  filter(time == "baseline") %>%
+  select(subject, time, supplement, tot.volume) %>%
[…]

### 2024-01-03 manuscript_revised.RMD
working on revision

--- 2023-10-18 manuscript.RMD
@@
 High glucose treatment may augment ribosomal biogenesis, as evident from *in vitro* studies, and may thus act to potentiate resistance training-induced adaptations *in vivo*.
-However, this remains largely unexplored.
+This was investigated with the main outcomes being total RNA and ribosomal RNA abundances, as well as relevant transcriptional or translational regulators (c-Myc/UBF/rpS6) to further describe the relationship between these regulators and ribosome biogenesis in human skeletal muscle tissue.
 
@@
 Primary and secondary antibodies were purchased from Santa Cruz Biotechnology (Texas, USA): UBF, UBF F-9, sc-13125; rpS6, Ribosomal protein S6 C-8, sc-74459; and Thermo Fisher Scientific (Oslo, Norway): c-Myc, 9E10; goat anti-mouse (for c-Myc), goat anti-mouse IgG1 (y1) horseradish peroxidase conjugate; and anti-mouse (anti-mouse IgG1 horseradish peroxidase conjugate).
-Antibodies were diluted in blocking buffer to concentrations corresponding to 1:25 000 (UBF, rpS6) and 1:5000 (c-Myc).
+<!-- Reviewer comment: 25,000 dilution for primary seems quite diluted.
+Is the dilution is correct?
+Revision: Antibodies were diluted in blocking buffer to concentrations corresponding to 1:500 (UBF and rpS6, primary), 1:2000 (c-Myc, primary), 1:5000 (c-Myc, secondary) and 1:25 000 (UBF, rpS6, secondary).-->
 

### 2024-01-03 resources/older manuscript versions/revision_notes.RMD
working on revision

---
title: "Untitled"
format: docx
editor: visual
---

### Revision

Reviewer 1, comment 1: 25,000 dilution for primary seems quite diluted. Is the dilution is correct? Page 9, line 257-258 (line 211 in quarto)
 - Original draft: Antibodies were diluted in blocking buffer to concentrations corresponding to 1:25 000 (UBF, rpS6) and 1:5000 (c-Myc).
 - Authors comment: By mistake, only the dilution volumes of the secondary antibodies were included in the original description. This section is now updated with dilution volumes of both primary and secondary antibodies. 
 - Revision: Antibodies were diluted in blocking buffer to concentrations corresponding to 1:500 (UBF and rpS6, primary), 1:2000 (c-Myc, primary), 1:5000 (c-Myc, secondary) and 1:25 000 (UBF, rpS6, secondary).

Reviewer 1, comment 2: Most of those antibodies have been discontinued by the commercial provided (Santa Cruz). Did the authors try to validate these antibodies?
 - Authors answer: We did several rounds of antibody testing with these specific primers before analysing the muscle tissue from the present study. We compared the blots from said test runs to reviews on the providers own web page as a form of validation. 
 - Primary antibodies are available at:
   - UBF F9: https://www.scbt.com/p/ubf-antibody-f-9
   - rpS6 C-8: https://www.scbt.com/p/ribosomal-protein-s6-antibody-c-8
   - c-Myc 9E10: https://www.thermofisher.com/antibody/product/c-Myc-Antibody-clone-9E10-Monoclonal/MA1-980?imageId=1024462
   - Finner ikke eksakte matcher for sekundære, må sjekke navnene på laben.

Reviewer 1, comment 3:Given that whey protein was also given to the participants following the resistance exercise session, could the increase insulin from whey protein ingestion already be enough to evoke the anabolic effects related to ribosome biogenesis? Did the authors measure plasma insulin levels? It would be interesting to this story to check insulin levels.
 - We did expect protein ingestion to possible impact plasma insulin levels, hence the participants waited for 2hrs from ingesting protein before exercising, thus "washing out" the potential effect of protein on plasma insulin levels. Further, 30 minutes before exercise there was no significant change in c-peptide levels, indicating no significant change in plasma insuline at this time. What occured between protein ingestion and the 30 min pre exericise mark, and/or if anything, this impacted ribosome biogenesis prior to glucose ingestion and exercise. 
 - We decided to measure plasma c-peptide rather than insulin. Indeed this provides an indirect measure as opposed to directly measuring plasma insulin, however, c-peptide stays in the blood longer than insulin, making c-peptide easier to accurately measure over several time points [LEIGHTON17]. Further, c-peptide is widely used as an insulin secretion measure [LEIGHTON17].
 - Er det noe argument for at ikke-signifikante endringer i plasma glukose 30 min etter protein inntak også underbygger lite endring i plasma insulin? Altså, dersom det hadde vært en stor utskillelse i plasma insulin, ville ikke dette ha påvirket plasma glukose?

Reviewer 2, comment 1: Please, also describe in the abstract the second objective of the study (lines 98-99). The abstract does not state, and, therefore is not clear, what are the main outcomes (parameters) measured in the clinical trial.
 - Authors comment: The abstract has now been updated with this information.
 - Revision:
   - Removed "However, this remains largely unexplored."
   - Added "This was investigated with the main outcomes being total RNA and ribosomal RNA abundances, as well as relevant transcriptional or translational regulators (c-Myc/UBF/rpS6) to further described relationship between these regulators and ribosome biogenesis in human skeletal muscle tissue."
   - Abstract trenger nå mer jobb, for å kutte ned til akseptert lengde/antall ord.

Reviewer 2, comment 2: In the results the data were described in relation to % ou fold of changes. It is important to report the data also in absolute values.

Reviewer 2, comment 3: There is a lack of information regarding the use of supplements by participants. This aspect is important for characterizing the sample and as stated by the authors on page 3, line 57.

Reviewer 2, comment 4: Figure 3 shows the result for RPS6. However, only in the discussion section there is some description about this protein (lines 377 - 381). I suggest that this be addressed in the introduction for better understanding of the reader.

Reviewer 2, comment 5: Figure 1d, use only the word "post" or "after".

Reviewer 2, comment 6: In Table 1, provide body fat percentage data, together with the absolute lean mass data (already shown), they become more informative to the reader.

Reviewer 2, comment 7: Please specify how the "blinded taste test" (described on page 6, line 120) was carried out.

Reviewer 2, comment 8: Mention more characteristics related to the training of the participants at baseline, such as number of weekly or 14-day RT sessions (which was an inclusion criteria).

### 2024-01-04 manuscript_revised.RMD
further revision

@@
 High glucose treatment may augment ribosomal biogenesis, as evident from *in vitro* studies, and may thus act to potentiate resistance training-induced adaptations *in vivo*.
-This was investigated with the main outcomes being total RNA and ribosomal RNA abundances, as well as relevant transcriptional or translational regulators (c-Myc/UBF/rpS6) to further describe the relationship between these regulators and ribosome biogenesis in human skeletal muscle tissue.
+This was investigated with total RNA and ribosomal RNA abundances as main outcomes, with relevant transcriptional or translational regulators (c-Myc/UBF/rpS6) as a secondary outcome.
 
@@
 3x0g glucose) opposite to the supplement they received during RT, to ensure a balanced daily intake of glucose.
-Apart from this, participants ingested a self-chosen diet during period II, registered in MyFitnessPal or similar applications.
+Apart from this, participants ingested a self-chosen diet during period II.
+Further, participants were asked not to use any other supplements such as additional protein and/or creatine, and to register all food/drink consumption in MyFitnessPal or similar applications.
 The self-chosen diet was repeated on pairwise consecutive days, to ensure similar premises for resistance training responses between conditions.

### 2024-01-04 resources/older manuscript versions/revision_notes.RMD
further revision

@@
 - Original draft: Antibodies were diluted in blocking buffer to concentrations corresponding to 1:25 000 (UBF, rpS6) and 1:5000 (c-Myc). 
- - Authors comment: By mistake, only the dilution volumes of the secondary antibodies were included in the original description.
+- Authors answer: By mistake, only the dilution volumes of the secondary antibodies were included in the original description.
 This section is now updated with dilution volumes of both primary and secondary antibodies.
@@
   - Removed "However, this remains largely unexplored." 
-   - Added "This was investigated with the main outcomes being total RNA and ribosomal RNA abundances, as well as relevant transcriptional or translational regulators (c-Myc/UBF/rpS6) to further described relationship between these regulators and ribosome biogenesis in human skeletal muscle tissue."
-   - Abstract trenger nå mer jobb, for å kutte ned til akseptert lengde/antall ord.
+  - Added "This was investigated with total RNA and ribosomal RNA abundances as main outcomes, with relevant transcriptional or translational regulators (c-Myc/UBF/rpS6) as a secondary outcome." 
+- Abstract trenger nå mer jobb, for å kutte ned til akseptert lengde/antall ord. Er det ok med 30 ord for mye?
 
@@
 This aspect is important for characterizing the sample and as stated by the authors on page 3, line 57.
+ - Authors comment: The participants were asked not to use any other form of supplements during the course of the study.
+However, I can see that this is not clearly stated in the methods chapter, and will now be added.
+ - Revision: Page 5, lines 141-143.
+   - "Apart from this, participants ingested a self-chosen diet during period II, registered in MyFitnessPal or similar applications" changed to "Apart from this, participants ingested a self-chosen diet during period II.
+Further, participants were asked not to use any other supplements such as additional protein and/or creatine, and to register all food/drink consumption in MyFitnessPal or similar applications"
 

### 2024-01-05 manuscript_revised.RMD
further revision

@@
 \~0.3g Stevia rebaudiana extract).
+In this blinded taste test, the participants were given two boluses of placebo (75ml) and two boluses of glucose (75ml), consumed in a randomised order.
+The participants were instructed to finish one bolus, guess the content, and move on to the next bolus.
 To ensure equal conditions during training sessions and strength testing, participants exercised and tested at the same time of day, +/- 1hr with the same supervisor on pairwise consecutive days (i.e.
@@
 
-Glucose ingestion before and after RT led to increases in plasma glucose levels compared to baseline by 38% immediately before RT (Figure 1B, 0 min), by 31% during RT (Figure 1B, 15 min) and by 32% immediately after RT (Figure 1B, 30 min; all : *p* \< 0.001), with no changes observed in the placebo condition (*p* \> 0.05).
-Compared to the placebo condition, ingestion of glucose increased plasma glucose levels by 36% immediately before RT (Figure 1B, 0 min), by 27% during RT (Figure 1B, 15 min) and by 28% immediately after RT (Figure 1B, 30 min; all : *p* \< 0.001).
+Glucose ingestion before and after RT led to increases in plasma glucose levels compared to baseline by 38% immediately before RT (Figure 1B, 0 min: 2.05 ± 0.73 mmol/L), by 31% during RT (Figure 1B, 15 min: 1.75 ± 1.44 mmol/L) and by 32% immediately after RT (Figure 1B, 30 min: 1.62 ± 1.10 mmol/L, all : *p* \< 0.001), with no changes observed in the placebo condition (Figure 1B, 0 min: 0.09 ± 0.3 mmol/L, 15 min: 0.16 ± 0.35 mmol/L, 30 min: 0.18 ± 0.39 mmol/L, *p* \> 0.05).
+Compared to the placebo condition, ingestion of glucose increased plasma glucose levels by 36% immediately before RT (Figure 1B, 0 min), by 27% during RT (Figure 1B, 15 min) and by 28% immediately after RT (Figure 1B, 30 min, all : *p* \< 0.001).
 Two hours after the RT session, glucose ingestion was associated with 12% lower plasma glucose levels compared to baseline, and 8% lower compared to placebo (Figure 1B, 270min: *p* = `r p.270`).
 
-Glucose ingestion before and after RT led to increases in levels of c-peptide compared to baseline, by 95% immediately before (Figure 1C, 0 min) and 87% after RT (Figure 1C, 30 min; both *p* \< 0.001), with no changes observed with the placebo condition (*p* \> 0.05).
+Glucose ingestion before and after RT led to increases in levels of c-peptide compared to baseline, by 95% immediately before (Figure 1C, 0 min: 796 ± 376.0 pmol/L) and 87% after RT (Figure 1C, 30 min: 793 ± 581.0 pmol/L, both *p* \< 0.001), with no changes observed with the placebo condition (Figure 1C, 0 min: 63.7 ± 71.0 pmol/L, 30 min: 53.9 ± 134.0 pmol/L, both *p* \> 0.05).
 Compared to the placebo condition, ingestion of glucose increased levels of c-peptide by 85% immediately before (Figure 1C, 0 min) and 85% after RT (Figure 1C, 30 min; both *p* \< 0.001).
@@
 
-Glucose ingestion before and after RT sessions did not generally improve skeletal muscle recovery throughout the intervention, neither during the intervention in a rested state (Figure 1D, 23hrs after exercise; Post 2RT: *p* = `r pdiff.c2`, Post 4RT: *p* = `r pdiff.c3`), nor acutely after the sixth and final RT-session (30min post 6RT: *p* = `r pdiff.c5`, 2h post 6RT: *p* = `r pdiff.c6`) or in a rested state after the sixth RT session (23h post 6RT: *p* = `r pdiff.c7`).
-Glucose ingestion before and after RT compared to placebo was associated with a 7% less reduction in strength after five RT sessions (Figure 1D, *p* = `r pdiff.c4`).
+Glucose ingestion before and after RT sessions did not generally improve skeletal muscle recovery throughout the intervention, neither during the intervention in a rested state (Figure 1D, 23hrs after exercise; Post 2RT: -0.02 ± 0.05 vs.
+-0.04 ± 0.03 *p* = `r pdiff.c2`, Post 4RT: -0.02 ± 0.05 vs.
+-0.06 ± 0.03 *p* = `r pdiff.c3`, glucose vs.
+placebo), nor acutely after the sixth and final RT-session (30min post 6RT: -0.03 ± 0.05 vs.
+-0.09 ± 0.06 *p* = `r pdiff.c5`, 2h post 6RT: -0.02 ± 0.07 vs.
+-0.08 ± 0.07 *p* = `r pdiff.c6`, glucose vs.
+placebo) or in a rested state after the sixth RT session (23h post 6RT: -0.05 ± 0.06 vs.
+-0.09 vs.
+0.05 *p* = `r pdiff.c7`, glucose vs.
+placebo).
+Glucose ingestion before and after RT compared to placebo was associated with a 7% less reduction in strength after five RT sessions (Figure 1D, -0.06 ± 0.06 vs -0.14 ± 0.12 *p* = `r pdiff.c4`, glucose vs.
+placebo).
 

### 2024-01-05 resources/older manuscript versions/revision_notes.RMD
further revision

@@
 Reviewer 1, comment 1: 25,000 dilution for primary seems quite diluted. Is the dilution is correct? Page 9, line 257-258 (line 211 in quarto) 
+- Authors reply: By mistake, only the dilution volumes for the secondary antibodies were included in the original description.
+This section is now updated with dilution volumes of both primary and secondary antibodies.
+- Revision: 
    - Original draft: Antibodies were diluted in blocking buffer to concentrations corresponding to 1:25 000 (UBF, rpS6) and 1:5000 (c-Myc).
-- Authors answer: By mistake, only the dilution volumes of the secondary antibodies were included in the original description.
-This section is now updated with dilution volumes of both primary and secondary antibodies.
    - Revision: Antibodies were diluted in blocking buffer to concentrations corresponding to 1:500 (UBF and rpS6, primary), 1:2000 (c-Myc, primary), 1:5000 (c-Myc, secondary) and 1:25 000 (UBF, rpS6, secondary).
@@
 Reviewer 1, comment 2: Most of those antibodies have been discontinued by the commercial provided (Santa Cruz). Did the authors try to validate these antibodies? 
-- Authors answer: We did several rounds of antibody testing with these specific primers before analysing the muscle tissue from the present study.
+- Authors reply: We did several rounds of antibody testing with these specific primers before analysing the muscle tissue from the present study.
 We compared the blots from said test runs to reviews on the providers own web page as a form of validation.
@@
 - c-Myc 9E10: https://www.thermofisher.com/antibody/product/c-Myc-Antibody-clone-9E10-Monoclonal/MA1-980?imageId=1024462 
-- Finner ikke eksakte matcher for sekundære, må sjekke navnene på laben.
+- Finner ikke eksakte matcher for sekundære, må sjekke navnene på laben. Finner allikevel ikke noen info om at de primære antistoffene er avviklet.
 
@@
 It would be interesting to this story to check insulin levels.
-- We did expect protein ingestion to possible impact plasma insulin levels, hence the participants waited for 2hrs from ingesting protein before exercising, thus "washing out" the potential effect of protein on plasma insulin levels.
+- Authors reply: We did expect protein ingestion to possible impact plasma insulin levels, hence the participants waited for 2hrs from ingesting protein before exercising, thus "washing out" the potential effect of protein on plasma insulin levels.
 Further, 30 minutes before exercise there was no significant change in c-peptide levels, indicating no significant change in plasma insuline at this time.
 What occured between protein ingestion and the 30 min pre exericise mark, and/or if anything, this impacted ribosome biogenesis prior to glucose ingestion and exercise.
-- We decided to measure plasma c-peptide rather than insulin.
+We decided to measure plasma c-peptide rather than insulin.
 Indeed this provides an indirect measure as opposed to directly measuring plasma insulin, however, c-peptide stays in the blood longer than insulin, making c-peptide easier to accurately measure over several time points [LEIGHTON17].
@@
 The abstract does not state, and, therefore is not clear, what are the main outcomes (parameters) measured in the clinical trial.
-- Authors comment: The abstract has now been updated with this information. 
+- Authors reply: The abstract has now been updated with this information. 
 - Revision: 
@@
 Reviewer 2, comment 2: In the results the data were described in relation to % ou fold of changes. It is important to report the data also in absolute values.
+ - Authors reply: Results chapter now updated with absolute values of change from baseline to post in both conditions
+   - Glucose revision: Glucose ingestion before and after RT led to increases in plasma glucose levels compared to baseline by 38% immediately before RT (Figure 1B, 0 min: 2.05 ± 0.73 mmol/L), by 31% during RT (Figure 1B, 15 min: 1.75 ± 1.44 mmol/L) and by 32% immediately after RT (Figure 1B, 30 min: 1.62 ± 1.10 mmol/L, all : *p* \< 0.001), with no changes observed in the placebo condition (Figure 1B, 0 min: 0.09 ± 0.3 mmol/L, 15 min: 0.16 ± 0.35 mmol/L, 30 min: 0.18 ± 0.39 mmol/L, *p* \> 0.05).
+   - C-peptide revision: Glucose ingestion before and after RT led to increases in levels of c-peptide compared to baseline, by 95% immediately before (Figure 1C, 0 min: 796 ± 376.0 pmol/L) and 87% after RT (Figure 1C, 30 min: 793 ± 581.0 pmol/L, both *p* \< 0.001), with no changes observed with the placebo condition (Figure 1C, 0 min: 63.7 ± 71.0 pmol/L, 30 min: 53.9 ± 134.0 pmol/L, both *p* \> 0.05).
+   - Strength index revision: Glucose ingestion before and after RT sessions did not generally improve skeletal muscle recovery throughout the intervention, neither during the intervention in a rested state (Figure 1D, 23hrs after exercise; Post 2RT: -0.02 ± 0.05 vs.
+-0.04 ± 0.03 *p* = `r pdiff.c2`, Post 4RT: -0.02 ± 0.05 vs.
+-0.06 ± 0.03 *p* = `r pdiff.c3`, glucose vs.
+placebo), nor acutely after the sixth and final RT-session (30min post 6RT: -0.03 ± 0.05 vs.
+-0.09 ± 0.06 *p* = `r pdiff.c5`, 2h post 6RT: -0.02 ± 0.07 vs.
+-0.08 ± 0.07 *p* = `r pdiff.c6`, glucose vs.
+placebo) or in a rested state after the sixth RT session (23h post 6RT: -0.05 ± 0.06 vs.
+-0.09 vs.
+0.05 *p* = `r pdiff.c7`, glucose vs.
+placebo).
+Glucose ingestion before and after RT compared to placebo was associated with a 7% less reduction in strength after five RT sessions (Figure 1D, -0.06 ± 0.06 vs -0.14 ± 0.12 *p* = `r pdiff.c4`, glucose vs.
+placebo).
+   - Total RNA and rRNA revision:
 
@@
 Reviewer 2, comment 5: Figure 1d, use only the word "post" or "after".
+ - Authors answer: Figure 1d updated only to use the word "after" instead of the mix of "after" and "post".
+ - Men, er ikke helt sikker på om jeg egentlig vil bruke "after" eller "post" her.
+Jeg synes "after" leses bedre på denne figuren, men "post" passer bedre i figur 2.
+Er det uryddig å ha forskjellig på disse to figurene?
 
 Reviewer 2, comment 6: In Table 1, provide body fat percentage data, together with the absolute lean mass data (already shown), they become more informative to the reader.
+ - Må hente ut denne dataen
 
 Reviewer 2, comment 7: Please specify how the "blinded taste test" (described on page 6, line 120) was carried out.
+ - Authors answer: Added following description of the blinded taste test (page 5, line 120-121).
+   - "In this blinded taste test, the participants were given two boluses of placebo (75ml) and two boluses of glucose (75ml), consumed in a randomised order.
+The participants were instructed to finish one bolus, guess the content, and move on to the next bolus."
 
 Reviewer 2, comment 8: Mention more characteristics related to the training of the participants at baseline, such as number of weekly or 14-day RT sessions (which was an inclusion criteria).
+ - Authors answer: Sadly, we do not have detailed characteristics on individual training sessions per week prior to enrollment in the study, other than that they did indeed have a weekly resistance training session volume between two and eight sessions per 14 days for the last six months.
+The authors realise that this leaves room for variation between participants in terms of training status a baseline, however variations such as these should be controlled for by the within-participants analyses.

### 2024-01-05 R/strength.index.R

@@ -88,6 +88,12 @@ str.change <- strength.index %>%
                cols = c(change.2:change.7)) %>%
   print()
 
+# Index change in absolute value
+str.change |>
+  group_by(time, supplement) |>
+  summarise(mc = mean(change),
+            sd = sd(change))
+
 saveRDS(str.change, "./data/data-gen/humac/str.change.RDS")
 
 str.lchange <- strength.index %>%

### 2024-01-05 figures/figure1.R

@@ -151,8 +151,8 @@ str.fig2 <- str.emm2 %>%
   scale_x_continuous(limits = c(-1,25), breaks = c(0, 8, 16, 20, 21, 22, 23.5),
                      expand = expansion(0),
                      labels = c("0" = "Baseline", "8" = "23hrs after 2RT", "16" = "23hrs after 4RT", "21" = "23hrs after 5RT",
-                                "22.7" = "30min post 6RT", "23.4" = "2h post 6RT",
-                                "24.4" = "23h post 6RT")) +
+                                "22.7" = "30min after 6RT", "23.4" = "2h after 6RT",
+                                "24.4" = "23h after 6RT")) +
   labs(x = "Time", y = "Strength index change", fill = "Supplement") +
   theme(axis.text.x = element_text(angle = 50, hjust = 1, vjust = 1, size = 6),
         axis.title = element_text(size = 7),

### 2024-01-08 R/humac.main.analysis.R
revision updates

@@ -54,7 +54,29 @@ change.isom <- isom.dat %>%
                cols = (change.2:change.4)) #%>%
 #  print()
 
-saveRDS(change.isom, "./data/data-gen/humac/isom.lchange.RDS")
+saveRDS(change.isom, "./data/data-gen/humac/isom.lchange.RDS") # This was used for analyses of strength index
+
+# Absolute change isometric
+
+absolutechange.isom <- isom.dat %>%
+  dplyr::select(subject, time, supplement, peak.torque) %>%
+  group_by(subject, time, supplement) %>%
+  summarise(peak.torque = mean(peak.torque, na.rm = TRUE)) %>%
+  pivot_wider(names_from = time, 
+              values_from = peak.torque) %>%
+  
+  ungroup() %>%
+  mutate(change.2 = test1-baseline,
+         change.3 = test2-baseline,
+         change.4 = test3-baseline,
+         baseline = baseline - mean(baseline, na.rm = TRUE),
+         supplement = factor(supplement, levels = c("placebo", "glucose"))) %>%
+  select(subject, supplement, baseline, change.2, change.3, change.4) %>%
+  pivot_longer(names_to = "time",
+               values_to = "change",
+               cols = (change.2:change.4))
+
+saveRDS(absolutechange.isom, "./data/data-gen/humac/isom.absolutechange.RDS") # this was used for the humac table
 
 # Isok 60
 
@@ -81,7 +103,29 @@ change.60 <- isok60.dat %>%
                cols = (change.2:change.4)) #%>%
   #print()
 
-saveRDS(change.60, "./data/data-gen/humac/isok60.lchange.RDS")
+saveRDS(change.60, "./data/data-gen/humac/isok60.lchange.RDS") # This was used for analyses of strength index
+
+# Absolute change isokinetic 60
+
+absolutechange.60 <- isok60.dat %>%
+  dplyr::select(subject, time, supplement, peak.torque) %>%
+  group_by(subject, time, supplement) %>%
+  summarise(peak.torque = mean(peak.torque, na.rm = TRUE)) %>%
+  pivot_wider(names_from = time, 
+              values_from = peak.torque) %>%
+  
+  ungroup() %>%
+  mutate(change.2 = test1-baseline,
+         change.3 = test2-baseline,
+         change.4 = test3-baseline,
+         baseline = baseline - mean(baseline, na.rm = TRUE),
+         supplement = factor(supplement, levels = c("placebo", "glucose"))) %>%
[…]

### 2024-01-09 manuscript_revised.RMD
added reference to tab 3

@@
 
-Glucose ingestion before and after RT sessions did not generally improve skeletal muscle recovery throughout the intervention, neither during the intervention in a rested state (Figure 1D, 23hrs after exercise; Post 2RT: -0.02 ± 0.05 vs.
--0.04 ± 0.03 *p* = `r pdiff.c2`, Post 4RT: -0.02 ± 0.05 vs.
--0.06 ± 0.03 *p* = `r pdiff.c3`, glucose vs.
-placebo), nor acutely after the sixth and final RT-session (30min post 6RT: -0.03 ± 0.05 vs.
--0.09 ± 0.06 *p* = `r pdiff.c5`, 2h post 6RT: -0.02 ± 0.07 vs.
--0.08 ± 0.07 *p* = `r pdiff.c6`, glucose vs.
-placebo) or in a rested state after the sixth RT session (23h post 6RT: -0.05 ± 0.06 vs.
--0.09 vs.
-0.05 *p* = `r pdiff.c7`, glucose vs.
-placebo).
-Glucose ingestion before and after RT compared to placebo was associated with a 7% less reduction in strength after five RT sessions (Figure 1D, -0.06 ± 0.06 vs -0.14 ± 0.12 *p* = `r pdiff.c4`, glucose vs.
-placebo).
+Glucose ingestion before and after RT sessions did not generally improve skeletal muscle recovery throughout the intervention, neither during the intervention in a rested state (Figure 1D, 23hrs after exercise; Post 2RT: *p* = `r pdiff.c2`, Post 4RT: *p* = `r pdiff.c3`), nor acutely after the sixth and final RT-session (30min post 6RT: *p* = `r pdiff.c5`, 2h post 6RT: *p* = `r pdiff.c6`) or in a rested state after the sixth RT session (23h post 6RT: *p* = `r pdiff.c7`).
+Glucose ingestion before and after RT compared to placebo was associated with a 7% less reduction in strength after five RT sessions (Figure 1D, *p* = `r pdiff.c4`).
 
@@
 Twenty-three hours after the last (sixth) RT session, strength was unchanged compared to after the fifth RT session (Figure 1D, 23h post 6RT: *p* = `r apc4`).
+Table 3 shows the mean change in absolute peak torque values per condition and angular velocity.
 
@@
 ```{r baseline table, echo = FALSE}
-
-###### Baseline tab
-#
-#
-## Author: KL
-#
-#
-## Purpose:
 
@@
 
-### Primer sequence table
-
-## Packages
-library(readxl); library(tidyverse); library(flextable)
-
 ## Data
@@
 ```
+
+**Table 3.
+Peak Torque** Mean peak torque per condition and angular velocity measured at the different times during the intervention.
+Baseline = before exercise, After 2RT = 23 hours after two training sessions, After 4RT = 23 hours after four training sessions, After RT5 = 23 hours after five training sessions, After 6RT#1 = 30min after the sixth training session, After 6RT#2 = 2 hours after the sixth training session, After 6RT#3 = 23 hours after the sixth training session.
+Values are reported as mean ± SD.
+* = significant difference between glucose and placebo.
+
+```{r humac table, echo = FALSE}
+
+## Data
+
+# Baseline data
+humac.clean2 <- readRDS("./data/data-gen/humac/humac.clean2.RDS")
+
+# Pre-post five sessions data
+
+isometric <- readRDS("./data/data-gen/humac/isom.absolutechange.RDS")
+iso60 <- readRDS("./data/data-gen/humac/iso60.absolutechange.RDS")
+iso240 <- readRDS("./data/data-gen/humac/iso240.absolutechange.RDS")
+
+# Pre-post session six data (acute data)
+isometric.acute <- readRDS("./data/data-gen/humac/isom.absolutechange.ac.RDS")
+iso60.acute <- readRDS("./data/data-gen/humac/iso60.absolutechange.ac.RDS")
+iso240.acute <- readRDS("./data/data-gen/humac/iso240.absolutechange.ac.RDS")
+
+## Summarising
+
+# Wide format
+
+tabdat <- humac.clean2 |>
+  select(subject, supplement, time, acute, measure = test, peak.torque) %>%
+#  filter(time == "baseline" & acute == "rest") |>
+  group_by(measure, supplement, time) |>
+  summarise(mean.pt = mean(peak.torque),
+            sd.pt = sd(peak.torque)) |>
+  mutate(mean = paste0(sprintf("%.1f",mean.pt), " (", sprintf("%.1f",sd.pt), ")")) |>
+  select(supplement, time, measure, mean) |>
+  mutate(supplement = if_else(supplement == "glucose",
+                       "Glucose",
+                       if_else(supplement == "placebo",
+                               "Placebo", supplement))) |>
+  mutate(measure = if_else(measure == "isom",
+                           "0\U00BA sec",
+                           if_else(measure == "isok.60",
+                                   "60\U00BA sec",
+                                   if_else(measure == "isok.240",
+                                           "240\U00BA sec", measure)))) |>
+  pivot_wider(names_from = time,
+              values_from = mean)
+
+## Making the table
+
+humac.flex <- tabdat |>
+  flextable() |>
+  compose(i = 1, j = c(1, 2, 3, 4, 5, 6, 7, 8, 9), part = "header",
+          value = c(as_paragraph("Condition"),
+                    as_paragraph("Velocity"),
+                    as_paragraph("Baseline"),
+                    as_paragraph("After 2RT"),
+                    as_paragraph("After 4RT"),
+                    as_paragraph("After 5RT"),
+                    as_paragraph("After 6RT#1"),
+                    as_paragraph("After 6RT#2"),
+                    as_paragraph("After 6RT#3"))) |>
+  add_body_row(values = c(" ", "Knee-extension peak torque (Nm)"), colwidths = c(2, 7))
+
+humac.flex <- border(humac.flex, 
+                     i = NULL,
+                     j = NULL,
+                     border = NULL,
+                     border.top = fp_border_default(color = "black"),
+                     border.bottom = fp_border_default(color = "black"),
+                     border.left = fp_border_default(color = "black"),
+                     border.right = fp_border_default(color = "black"),
+                     part = "all")
+
+humac.flex <- footnote(humac.flex,
+                       i = 4,
+                       j = 6,
+                       as_paragraph("Sig. difference"),
+                       ref_symbols = "*",
+                       part = "body", inline = TRUE)
+
+humac.flex
+
+```

### 2024-01-09 resources/older manuscript versions/revision_notes.RMD
revision updates
finished tab3

@@
    - C-peptide revision: Glucose ingestion before and after RT led to increases in levels of c-peptide compared to baseline, by 95% immediately before (Figure 1C, 0 min: 796 ± 376.0 pmol/L) and 87% after RT (Figure 1C, 30 min: 793 ± 581.0 pmol/L, both *p* \< 0.001), with no changes observed with the placebo condition (Figure 1C, 0 min: 63.7 ± 71.0 pmol/L, 30 min: 53.9 ± 134.0 pmol/L, both *p* \> 0.05).
-   - Strength index revision: Glucose ingestion before and after RT sessions did not generally improve skeletal muscle recovery throughout the intervention, neither during the intervention in a rested state (Figure 1D, 23hrs after exercise; Post 2RT: -0.02 ± 0.05 vs.
--0.04 ± 0.03 *p* = `r pdiff.c2`, Post 4RT: -0.02 ± 0.05 vs.
--0.06 ± 0.03 *p* = `r pdiff.c3`, glucose vs.
-placebo), nor acutely after the sixth and final RT-session (30min post 6RT: -0.03 ± 0.05 vs.
--0.09 ± 0.06 *p* = `r pdiff.c5`, 2h post 6RT: -0.02 ± 0.07 vs.
--0.08 ± 0.07 *p* = `r pdiff.c6`, glucose vs.
-placebo) or in a rested state after the sixth RT session (23h post 6RT: -0.05 ± 0.06 vs.
--0.09 vs.
-0.05 *p* = `r pdiff.c7`, glucose vs.
-placebo).
-Glucose ingestion before and after RT compared to placebo was associated with a 7% less reduction in strength after five RT sessions (Figure 1D, -0.06 ± 0.06 vs -0.14 ± 0.12 *p* = `r pdiff.c4`, glucose vs.
-placebo).
-   - Total RNA and rRNA revision:
+   - Absolute values for humac added in a table (tab3), reference by "Table 3 shows the mean change in absolute peak torque values per condition and angular velocity." at the end of the strength index results.
+   - western/rna??Prøve å inkludere rådata i figurene?
 
@@
 Reviewer 2, comment 6: In Table 1, provide body fat percentage data, together with the absolute lean mass data (already shown), they become more informative to the reader.
- - Må hente ut denne dataen
+ - Må hente ut denne dataen. Tabellen er oppdatert med tilfeldige tall for kroppsfett som "placeholder". Trenger bare å plotte inn rett data i dxa arket.
 

### 2024-01-09 figures/tab3.R

@@ -53,7 +53,7 @@ tabdat <- humac.clean2 |>
 
 ## Making the table
 
-humac.flex <- tabdat.wide |>
+humac.flex <- tabdat |>
   flextable() |>
   compose(i = 1, j = c(1, 2, 3, 4, 5, 6, 7, 8, 9), part = "header",
           value = c(as_paragraph("Condition"),

### 2024-01-09 manuscript_revised.RMD
added Knut Sindre as author
updating intro

@@
 
-###### Kristian Lian^1^, Daniel Hammarström^1^, Håvard Hamarsland^1^, Sara Christine Olsen Moen^1^, Stian Ellefsen^1^
+###### Kristian Lian^1^, Daniel Hammarström^1^, Håvard Hamarsland^1^, Knut Sindre Mølmen^1^, Sara Christine Olsen Moen^1^, Stian Ellefsen^1^
 
@@
 Moreover, mTORC1 and its downstream target ribosomal protein S6 kinase beta-1 (S6K1) serves as direct mediators of insulin signalling in skeletal muscle [@hillier_physiological_2000].
-Together these observations indicate a potential role of glucose in positively affecting human skeletal muscle rRNA synthesis, through potentiating transcription or translation of ribosomal RNA and increasing the translational capacity of skeletal muscle cells [@sebata_control_2018; @zhai_human_2012; @hillier_physiological_2000; @hoppe_amp-activated_2009].
+Previously, increases in rpS6 coincided with increases in rRNA and total RNA in the early phase resistance training in humans [@hammarstrom_ribosome_2022].<!-- FORTSETT HER --> Together these observations indicate a potential role of glucose in positively affecting human skeletal muscle rRNA synthesis, through potentiating transcription or translation of ribosomal RNA and increasing the translational capacity of skeletal muscle cells [@sebata_control_2018; @zhai_human_2012; @hillier_physiological_2000; @hoppe_amp-activated_2009].
 

### 2024-01-09 resources/older manuscript versions/revision_notes.RMD
updating intro

@@
 We decided to measure plasma c-peptide rather than insulin.
-Indeed this provides an indirect measure as opposed to directly measuring plasma insulin, however, c-peptide stays in the blood longer than insulin, making c-peptide easier to accurately measure over several time points [LEIGHTON17].
-Further, c-peptide is widely used as an insulin secretion measure [LEIGHTON17].
+Indeed this provides an indirect measure as opposed to directly measuring plasma insulin, however, c-peptide stays in the blood longer than insulin, making c-peptide easier to accurately measure over several time points [@leighton2017practical].
+Further, c-peptide is widely used as an insulin secretion measure [@leighton2017practical].
 - Er det noe argument for at ikke-signifikante endringer i plasma glukose 30 min etter protein inntak også underbygger lite endring i plasma insulin?

### 2024-01-10 manuscript_revised.RMD
tab 3 added, tab 1 modified accordingly

@@
 
-**Table 1.
-Participant characteristics:** Values are means ± SD.
-60º sec-1 = 60 degrees per second movement speed (isokinetic), 240º sec-1 = 240 degrees per second movement speed (isokinetic), 0º sec-1 = 0 degrees per second movement speed (isometric).
+**Table 1. Participant characteristics:** Values are means ± SD.
 
@@
 dxa.tab <- dxa.dat %>%
-  select(subject, sex, age, height, weight, leanmassg) %>%
+  select(subject, sex, age, height, weight, leanmassg, bodyfat) %>%
   mutate(leanmass = leanmassg/1000) %>%
@@
             mean.lm = mean(leanmass),
-            sd.lm = sd(leanmass)) %>%
+            sd.lm = sd(leanmass),
+            mean.bf = mean(bodyfat),
+            sd.bf = sd(bodyfat)) %>%
   mutate(age = paste0(sprintf("%.1f", mean.a), " (", sprintf("%.1f",sd.a), ")"),
@@
          weight = paste0(sprintf("%.1f",mean.w), " (", sprintf("%.1f",sd.w), ")"),
-         lm = paste0(sprintf("%.1f",mean.lm), " (", sprintf("%.1f",sd.lm), ")")) %>%
+         lm = paste0(sprintf("%.1f",mean.lm), " (", sprintf("%.1f",sd.lm), ")"),
+         bf = paste0(sprintf("%.1f",mean.bf), " (", sprintf("%.1f",sd.bf), ")")) %>%
   mutate(sex = if_else(sex == "female",
@@
                                "Male", sex))) %>%
-  select(sex, age, height, weight, lm) %>%
+  select(sex, age, height, weight, lm, bf) %>%
   pivot_longer(values_to = "mean",
                names_to = "measure",
-               cols = c(age:lm))
+               cols = c(age:bf))
 
@@
 # Summarising and prepping humac data
-humac.tab <- humac.clean2 %>%
-  full_join(gender) %>%
-  select(subject, sex, time, acute, measure = test, peak.torque) %>%
-  filter(time == "baseline" & acute == "rest") %>%
-  group_by(sex, measure) %>%
-  summarise(mean.pt = mean(peak.torque),
-            sd.pt = sd(peak.torque)) %>%
-  mutate(mean = paste0(sprintf("%.1f",mean.pt), " (", sprintf("%.1f",sd.pt), ")")) %>%
-  select(sex, measure, mean) %>%
-  mutate(sex = if_else(sex == "female",
-                       "Female",
-                       if_else(sex == "male",
-                               "Male", sex)))
+#humac.tab <- humac.clean2 %>%
+#  full_join(gender) %>%
+#  select(subject, sex, time, acute, measure = test, peak.torque) %>%
+#  filter(time == "baseline" & acute == "rest") %>%
+#  group_by(sex, measure) %>%
+#  summarise(mean.pt = mean(peak.torque),
+#            sd.pt = sd(peak.torque)) %>%
+#  mutate(mean = paste0(sprintf("%.1f",mean.pt), " (", sprintf("%.1f",sd.pt), ")")) %>%
+#  select(sex, measure, mean) %>%
+#  mutate(sex = if_else(sex == "female",
+#                       "Female",
+#                       if_else(sex == "male",
+#                               "Male", sex)))
 
@@
 dxa.flex <- dxa.tab %>%
-  full_join(humac.tab) %>%
+  #full_join(humac.tab) %>%
   pivot_wider(values_from = mean,
@@
   full_join(gendercount) %>%
-  select(sex, n, age, height, weight, lm, isok.60, isok.240, isom) %>%
+  select(sex, n, age, height, weight, lm, bf) %>%
 
@@
   ## Change headers
-  compose(i = 1, j = c(1, 2, 3, 4, 5, 6, 7, 8, 9), part = "header", 
+  compose(i = 1, j = c(1, 2, 3, 4, 5, 6, 7), part = "header", 
           value = c(as_paragraph("Sex"),
@@
                     as_paragraph("Lean mass (kg)"),
-                    as_paragraph("60\U00BA sec", as_sup("-1")), 
-                    as_paragraph("240\U00BA sec", as_sup("-1")), 
-                    as_paragraph("0\U00BA sec", as_sup("-1")))) %>%
-  add_header_row(values = c(" ", "Knee-extension peak torque"), colwidths = c(6, 3))
+                    as_paragraph("Body Fat (%)")))
+                  #  as_paragraph("60\U00BA sec", as_sup("-1")), 
+                   # as_paragraph("240\U00BA sec", as_sup("-1")), 
+                   # as_paragraph("0\U00BA sec", as_sup("-1")))) %>%
+ # add_header_row(values = c(" ", "Knee-extension peak torque"), colwidths = c(7, 3))
 
@@
 Baseline = before exercise, After 2RT = 23 hours after two training sessions, After 4RT = 23 hours after four training sessions, After RT5 = 23 hours after five training sessions, After 6RT#1 = 30min after the sixth training session, After 6RT#2 = 2 hours after the sixth training session, After 6RT#3 = 23 hours after the sixth training session.
+240º sec-1 = 240 degrees per second movement speed (isokinetic), 60º sec-1 = 60 degrees per second movement speed (isokinetic), 0º sec-1 = 0 degrees per second movement speed (isometric).
 Values are reported as mean ± SD.
@@
   mutate(measure = if_else(measure == "isom",
-                           "0\U00BA sec",
+                           "0",
                            if_else(measure == "isok.60",
-                                   "60\U00BA sec",
+                                   "60",
                                    if_else(measure == "isok.240",
-                                           "240\U00BA sec", measure)))) |>
+                                           "240", measure)))) |>
   pivot_wider(names_from = time,
@@
                     as_paragraph("After 6RT#3"))) |>
-  add_body_row(values = c(" ", "Knee-extension peak torque (Nm)"), colwidths = c(2, 7))
+  add_body_row(values = c(" ", "Knee-extension peak torque (Nm)"), colwidths = c(2, 7)) |>
+  compose(i = c(2, 3, 4, 5, 6, 7), j = 2, part = "body",
+          value = c(as_paragraph("240\U00BA sec", as_sup("-1")),
+                    as_paragraph("240\U00BA sec", as_sup("-1")),
+                    as_paragraph("60\U00BA sec", as_sup("-1")),
+                    as_paragraph("60\U00BA sec", as_sup("-1")),
+                    as_paragraph("0\U00BA sec", as_sup("-1")),
+                    as_paragraph("0\U00BA sec", as_sup("-1"))))
 

### 2024-01-10 resources/older manuscript versions/revision_notes.RMD
revision updates
tab 3 added, tab 1 modified accordingly

@@
 
-Reviewer 1, comment 1: 25,000 dilution for primary seems quite diluted. Is the dilution is correct? Page 9, line 257-258 (line 211 in quarto) 
+<!--Reviewer 1, comment 1: 25,000 dilution for primary seems quite diluted. Is the dilution is correct? Page 9, line 257-258 (line 211 in quarto) 
 - Authors reply: By mistake, only the dilution volumes for the secondary antibodies were included in the original description.
@@
    - Original draft: Antibodies were diluted in blocking buffer to concentrations corresponding to 1:25 000 (UBF, rpS6) and 1:5000 (c-Myc).
-   - Revision: Antibodies were diluted in blocking buffer to concentrations corresponding to 1:500 (UBF and rpS6, primary), 1:2000 (c-Myc, primary), 1:5000 (c-Myc, secondary) and 1:25 000 (UBF, rpS6, secondary).
+   - Revision: Antibodies were diluted in blocking buffer to concentrations corresponding to 1:500 (UBF and rpS6, primary), 1:2000 (c-Myc, primary), 1:5000 (c-Myc, secondary) and 1:25 000 (UBF, rpS6, secondary).-->
 
-Reviewer 1, comment 2: Most of those antibodies have been discontinued by the commercial provided (Santa Cruz). Did the authors try to validate these antibodies? 
+<!--Reviewer 1, comment 2: Most of those antibodies have been discontinued by the commercial provided (Santa Cruz). Did the authors try to validate these antibodies? 
 - Authors reply: We did several rounds of antibody testing with these specific primers before analysing the muscle tissue from the present study.
@@
 - c-Myc 9E10: https://www.thermofisher.com/antibody/product/c-Myc-Antibody-clone-9E10-Monoclonal/MA1-980?imageId=1024462 
-- Finner ikke eksakte matcher for sekundære, må sjekke navnene på laben. Finner allikevel ikke noen info om at de primære antistoffene er avviklet.
+- Finner ikke eksakte matcher for sekundære, må sjekke navnene på laben. Finner allikevel ikke noen info om at de primære antistoffene er avviklet.-->
 
-Reviewer 1, comment 3:Given that whey protein was also given to the participants following the resistance exercise session, could the increase insulin from whey protein ingestion already be enough to evoke the anabolic effects related to ribosome biogenesis?
+<!--Reviewer 1, comment 3:Given that whey protein was also given to the participants following the resistance exercise session, could the increase insulin from whey protein ingestion already be enough to evoke the anabolic effects related to ribosome biogenesis?
 Did the authors measure plasma insulin levels?
@@
 - Er det noe argument for at ikke-signifikante endringer i plasma glukose 30 min etter protein inntak også underbygger lite endring i plasma insulin?
-Altså, dersom det hadde vært en stor utskillelse i plasma insulin, ville ikke dette ha påvirket plasma glukose?
+Altså, dersom det hadde vært en stor utskillelse i plasma insulin, ville ikke dette ha påvirket plasma glukose?-->
 
-Reviewer 2, comment 1: Please, also describe in the abstract the second objective of the study (lines 98-99).
+<!--Reviewer 2, comment 1: Please, also describe in the abstract the second objective of the study (lines 98-99).
 The abstract does not state, and, therefore is not clear, what are the main outcomes (parameters) measured in the clinical trial.
@@
   - Added "This was investigated with total RNA and ribosomal RNA abundances as main outcomes, with relevant transcriptional or translational regulators (c-Myc/UBF/rpS6) as a secondary outcome." 
-- Abstract trenger nå mer jobb, for å kutte ned til akseptert lengde/antall ord. Er det ok med 30 ord for mye?
+- Abstract trenger nå mer jobb, for å kutte ned til akseptert lengde/antall ord. Er det ok med 30 ord for mye?-->
 
@@
    - Absolute values for humac added in a table (tab3), reference by "Table 3 shows the mean change in absolute peak torque values per condition and angular velocity." at the end of the strength index results.
-   - western/rna??Prøve å inkludere rådata i figurene?
+   - Baseline strength values removed from tab 1, since a detailed table of humac measurements has been added (table 3).
+   - Mangler å plotte rådata i total RNA, så burde denne kommentaren være besvart.
 
-Reviewer 2, comment 3: There is a lack of information regarding the use of supplements by participants.
+<!--Reviewer 2, comment 3: There is a lack of information regarding the use of supplements by participants.
 This aspect is important for characterizing the sample and as stated by the authors on page 3, line 57.
@@
    - "Apart from this, participants ingested a self-chosen diet during period II, registered in MyFitnessPal or similar applications" changed to "Apart from this, participants ingested a self-chosen diet during period II.
-Further, participants were asked not to use any other supplements such as additional protein and/or creatine, and to register all food/drink consumption in MyFitnessPal or similar applications"
+Further, participants were asked not to use any other supplements such as additional protein and/or creatine, and to register all food/drink consumption in MyFitnessPal or similar applications"-->
 
@@
 
-Reviewer 2, comment 5: Figure 1d, use only the word "post" or "after".
+<!--Reviewer 2, comment 5: Figure 1d, use only the word "post" or "after".
  - Authors answer: Figure 1d updated only to use the word "after" instead of the mix of "after" and "post".
@@
 Jeg synes "after" leses bedre på denne figuren, men "post" passer bedre i figur 2.
-Er det uryddig å ha forskjellig på disse to figurene?
+Er det uryddig å ha forskjellig på disse to figurene?-->
 
-Reviewer 2, comment 6: In Table 1, provide body fat percentage data, together with the absolute lean mass data (already shown), they become more informative to the reader.
- - Må hente ut denne dataen. Tabellen er oppdatert med tilfeldige tall for kroppsfett som "placeholder". Trenger bare å plotte inn rett data i dxa arket.
+<!--Reviewer 2, comment 6: In Table 1, provide body fat percentage data, together with the absolute lean mass data (already shown), they become more informative to the reader.
+ - Authors answer: Updated figure one with body fat %-->
 
-Reviewer 2, comment 7: Please specify how the "blinded taste test" (described on page 6, line 120) was carried out.
+<!--Reviewer 2, comment 7: Please specify how the "blinded taste test" (described on page 6, line 120) was carried out.
  - Authors answer: Added following description of the blinded taste test (page 5, line 120-121).
    - "In this blinded taste test, the participants were given two boluses of placebo (75ml) and two boluses of glucose (75ml), consumed in a randomised order.
-The participants were instructed to finish one bolus, guess the content, and move on to the next bolus."
+The participants were instructed to finish one bolus, guess the content, and move on to the next bolus."-->
 
-Reviewer 2, comment 8: Mention more characteristics related to the training of the participants at baseline, such as number of weekly or 14-day RT sessions (which was an inclusion criteria).
+<!--Reviewer 2, comment 8: Mention more characteristics related to the training of the participants at baseline, such as number of weekly or 14-day RT sessions (which was an inclusion criteria).
  - Authors answer: Sadly, we do not have detailed characteristics on individual training sessions per week prior to enrollment in the study, other than that they did indeed have a weekly resistance training session volume between two and eight sessions per 14 days for the last six months.
-The authors realise that this leaves room for variation between participants in terms of training status a baseline, however variations such as these should be controlled for by the within-participants analyses.
+The authors realise that this leaves room for variation between participants in terms of training status a baseline, however variations such as these should be controlled for by the within-participants analyses.-->

### 2024-01-10 figures/tab1.R

@@ -72,45 +72,45 @@ gender <- dxa.dat %>%
   select(subject, sex) 
 
 # Summarising and prepping humac data
-humac.tab <- humac.clean2 %>%
-  full_join(gender) %>%
-  select(subject, sex, time, acute, measure = test, peak.torque) %>%
-  filter(time == "baseline" & acute == "rest") %>%
-  group_by(sex, measure) %>%
-  summarise(mean.pt = mean(peak.torque),
-            sd.pt = sd(peak.torque)) %>%
-  mutate(mean = paste0(sprintf("%.1f",mean.pt), " (", sprintf("%.1f",sd.pt), ")")) %>%
-  select(sex, measure, mean) %>%
-  mutate(sex = if_else(sex == "female",
-                       "Female",
-                       if_else(sex == "male",
-                               "Male", sex))) 
+#humac.tab <- humac.clean2 %>%
+#  full_join(gender) %>%
+#  select(subject, sex, time, acute, measure = test, peak.torque) %>%
+#  filter(time == "baseline" & acute == "rest") %>%
+#  group_by(sex, measure) %>%
+#  summarise(mean.pt = mean(peak.torque),
+#            sd.pt = sd(peak.torque)) %>%
+#  mutate(mean = paste0(sprintf("%.1f",mean.pt), " (", sprintf("%.1f",sd.pt), ")")) %>%
+#  select(sex, measure, mean) %>%
+#  mutate(sex = if_else(sex == "female",
+#                       "Female",
+#                       if_else(sex == "male",
+#                               "Male", sex))) 
 
 
 # Joining the data frames to be used in the flextable
 
 dxa.flex <- dxa.tab %>%
-  full_join(humac.tab) %>%
+  #full_join(humac.tab) %>%
   pivot_wider(values_from = mean,
               names_from = measure) %>%
   full_join(gendercount) %>%
-  select(sex, n, age, height, weight, lm, bf, isok.60, isok.240, isom) %>%
+  select(sex, n, age, height, weight, lm, bf) %>%
   
   ## Make it to flextable
   flextable() %>%
   ## Change headers
-  compose(i = 1, j = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10), part = "header", 
+  compose(i = 1, j = c(1, 2, 3, 4, 5, 6, 7), part = "header", 
           value = c(as_paragraph("Sex"),
                     as_paragraph("n"), 
                     as_paragraph("Age (yrs)"), 
[…]

### 2024-01-11 R/qpcr_analysis.R
getting absolute values RNA

@@ -81,6 +81,16 @@ qpcr_data2 <- qpcr_data %>%
 ### Save raw data 
 saveRDS(qpcr_data2, "./data/data-gen/rna/qpcr_data2.RDS")
 
+## Summarising absolute values
+abs.sum <- qpcr_data2 |>
+  select(subject, time, target, supplement, nf.w) |>
+  filter(target %in% c("trg_18s", "trg_47s", "trg_5s", "trg_5.8s", "trg_28s")) |>
+  group_by(time, target, supplement) |>
+  summarise(mean.nf = mean(nf.w),
+            sd.nf = sd(nf.w)) |>
+  pivot_wider(names_from = time,
+              values_from = mean.nf)
+  mutate(change = )
 
 
 qpcr_data2 %>%

### 2024-01-11 R/totrna_analysis.R

@@ -58,6 +58,22 @@ rna_dat2 <- rna_dat %>%
          ln.RNA.weight = log(RNA.weight)) %>%
   print()
 
+## Summarising for absolute values
+# RNA normalized to weight
+rna_dat2 |>
+  select(subject, time, supplement, RNA.weight) |>
+  group_by(supplement, time) |>
+  summarise(mean.rna = mean(RNA.weight),
+            sd.rna = sd(RNA.weight)) |>
+  
+
+# Raw RNA concentration
+rna_dat2 |>
+  select(subject, time, supplement, RNA) |>
+  group_by(supplement, time) |>
+  summarise(mean.rna = mean(RNA),
+            sd.rna = sd(RNA))
+
 ## Save for plotting/stats in UBF-RNA correlation
 saveRDS(rna_dat2, "./data/data-gen/rna/rna.dat.RDS")
 

### 2024-01-12 manuscript_revised.RMD
getting absolute values RNA
revision work

@@
 Data was collected at the sports science lab at the Inland Norway University of Applied Sciences, with all analyses performed on-site, except serum blood analyses, which were performed at Innlandet Hospital Trust.
-S.E., H.H.
-and D.H.
-conceived and designed the study; K.L., S.C.M., D.H.
-and H.H.
+D.H., H.H, K.S.M and S.E.
+conceived and designed the study; D.H., H.H., K.L., K.S.M, and S.C.M.
 performed experiments; K.L.
@@
 prepared figures; K.L.
-drafted the manuscript; D.H., H.H.
+drafted the manuscript; D.H., H.H., K.S.M.
 and S.E.
@@
 Indeed, UBF phosphorylation, which is required for interaction with the ribosomal deoxyribonucleic acid (rDNA) promoter, seems to be increased by high-glucose treatment in a mTORC1-dependent manner (rapamycin sensitive) in kidney glomerular epithelial cells [@mariappan_ribosomal_2011].
+In addition, mTORC1 regulation of ribosomal gene transcription is dependent on ribosomal protein S6 kinase beta-1 (S6K1) [@hannan2003mtor] with its target substrate ribosomal protein S6 (rpS6) coinciding with increases in rRNA and total RNA during early phase RT in humans [@hammarstrom_ribosome_2022].
 Independent of UBF and mTORC1, high glucose was shown to lead to chromatin remodelling, which promoted rRNA transcription in cell cultures [@zhai_human_2012].
 The mTORC1 pathway receives input from growth factors, hormones, mechanical loading, and nutrients to balance protein synthesis through multiple mechanisms based on cellular energy levels [@hoppe_amp-activated_2009] and contributes to ribosome biogenesis through the forming of the preinitiation complex (PIC) that marks the initiation of rRNA transcription and through the regulation of ribosomal protein translation [@figueiredo_regulation_2019; @von_walden_ribosome_2019].
-Moreover, mTORC1 and its downstream target ribosomal protein S6 kinase beta-1 (S6K1) serves as direct mediators of insulin signalling in skeletal muscle [@hillier_physiological_2000].
-Previously, increases in rpS6 coincided with increases in rRNA and total RNA in the early phase resistance training in humans [@hammarstrom_ribosome_2022].<!-- FORTSETT HER --> Together these observations indicate a potential role of glucose in positively affecting human skeletal muscle rRNA synthesis, through potentiating transcription or translation of ribosomal RNA and increasing the translational capacity of skeletal muscle cells [@sebata_control_2018; @zhai_human_2012; @hillier_physiological_2000; @hoppe_amp-activated_2009].
+Moreover, mTORC1 and S6K1 serves as direct mediators of insulin signalling in skeletal muscle [@hillier_physiological_2000].
+Together these observations indicate a potential role of glucose in positively affecting human skeletal muscle rRNA synthesis, through potentiating transcription or translation of ribosomal RNA and increasing the translational capacity of skeletal muscle cells [@sebata_control_2018; @zhai_human_2012; @hillier_physiological_2000; @hoppe_amp-activated_2009].
 
@@
 
-The five-session-RT intervention led to increases in total RNA and rRNA per unit muscle weight both in the glucose and the placebo condition, on average by \~20-27% and \~25-57%, respectively (Figure 2).
+The five-session-RT intervention led to increases in total RNA and rRNA per unit muscle weight both in the glucose and the placebo condition, on average by \~20-27% <!--(glucose: 263 ± 50 mg/ng-1, placebo: 210 ± 121 mg/ng-1)--> and \~25-57% <!--(47S; 0.253 ± 1.27 and 0.576 ± 0.677, 18S; 0.336 ± 0.460 and 0.271 ± 0.470, 28S; 0.314 ± 0.504 and 0.311 ± 0.582, 5.8S; 0.388 ± 0.576 and 0.322 ± 0.520, 5S; 0.305 ± 0.608 and 0.292 ± 0.432, arbitrary units for glucose and placebo respectively)--> , respectively (Figure 2).
 However, RT with glucose did not induce increased accumulation of total RNA (Figure 2A, mean difference `r total_rna_stats`) or rRNA (Figure 2B; 47S, `r stat_rrna$trg_47s`; 18S, `r stat_rrna$trg_18s`; 28S, `r stat_rrna$trg_28s`; 5.8S, `r stat_rrna$trg_5.8s`; 5S, `r stat_rrna$trg_5s`) compared to RT with placebo.

### 2024-01-12 resources/older manuscript versions/revision_notes.RMD
getting absolute values RNA
revision work

@@
    - Baseline strength values removed from tab 1, since a detailed table of humac measurements has been added (table 3).
-   - Mangler å plotte rådata i total RNA, så burde denne kommentaren være besvart.
+   - Added the following for total RNA and rRNA: (glucose: 263 ± 50 mg/ng-1, placebo: 210 ± 121 mg/ng-1), and; (47S; 0.253 ± 1.27 and 0.576 ± 0.677, 18S; 0.336 ± 0.460 and 0.271 ± 0.470, 28S; 0.314 ± 0.504 and 0.311 ± 0.582, 5.8S; 0.388 ± 0.576 and 0.322 ± 0.520, 5S; 0.305 ± 0.608 and 0.292 ± 0.432, arbitrary units for glucose and placebo respectively)
 
@@
 I suggest that this be addressed in the introduction for better understanding of the reader.
+ - Authors answer: Added the following;
+   - Introduction: "In addition, mTORC1 regulation of ribosomal gene transcription is dependent on ribosomal protein S6 kinase beta-1 (S6K1) [@hannan2003mtor] with its target substrate ribosomal protein S6 (rpS6) coinciding with increases in rRNA and total RNA during early phase RT in humans [@hammarstrom_ribosome_2022]"
+   - Discussion:
 
@@
 The authors realise that this leaves room for variation between participants in terms of training status a baseline, however variations such as these should be controlled for by the within-participants analyses.-->
+
+<!--Also, added Knut Sindre Mølmen to authors, and author contributions (conception of idea, experiments and edition of manuscript)-->

### 2024-01-12 R/qpcr_analysis.R

@@ -19,6 +19,7 @@ library(tidyverse)
 library(readxl)
 library(lmerTest)
 library(buildmer)
+library(emmeans)
 
 
 # Load data
@@ -81,16 +82,6 @@ qpcr_data2 <- qpcr_data %>%
 ### Save raw data 
 saveRDS(qpcr_data2, "./data/data-gen/rna/qpcr_data2.RDS")
 
-## Summarising absolute values
-abs.sum <- qpcr_data2 |>
-  select(subject, time, target, supplement, nf.w) |>
-  filter(target %in% c("trg_18s", "trg_47s", "trg_5s", "trg_5.8s", "trg_28s")) |>
-  group_by(time, target, supplement) |>
-  summarise(mean.nf = mean(nf.w),
-            sd.nf = sd(nf.w)) |>
-  pivot_wider(names_from = time,
-              values_from = mean.nf)
-  mutate(change = )
 
 
 qpcr_data2 %>%
@@ -207,77 +198,148 @@ stat_rrna <- lapply(rrna_analysis_results$rrna_models,
 saveRDS(stat_rrna, "./data/data-gen/rna/stat_rrna.RDS")
 
 
+## "Absolute change"
+ qdat <- qpcr_data2 %>%
+   mutate(target = gsub("rRNA", "", target),
+          target = gsub("  ", " ", target),
+          target = paste0("trg_", target)) %>%
+   separate(target, into = c("target", "primer"), sep = " ") %>%
+   
+   # Create a weight-normalized variable
+   mutate(nf.expr = log(expr / nf.w), 
+          nf.w = scale(nf.w),
+          # Technical random effect
+          technical = paste(subject, time, supplement, rep, sep = "_"),
+          biological = paste0("S", sample)) %>%
+   filter(!(target %in% c("trg_MHC1", "trg_MHC2A", "trg_MHC2X"))) %>%
+ 
+   
+   print()
+ 
+ saveRDS(qdat, "./data/data-gen/rna/qclean.RDS")
+ 
+ ## Change data
+ rrna18 <- qdat %>%
+   filter(target == "trg_trg_18s") %>%
+   print()
+ 
+ rrna28 <- qdat %>%
+   filter(target == "trg_trg_28s") %>%
+   print()
+ 
+ rrna5.8 <- qdat %>%
+   filter(target == "trg_trg_5.8s") %>%
+   print()
+ 
+ rrna5 <- qdat %>%
[…]

### 2024-01-15 manuscript_revised.RMD
revision work

@@
 
-ORCID: Kristian Lian <https://orcid.org/0000-0002-5169-5960>, Daniel Hammarström <https://orcid.org/0000-0001-8360-2100>, Håvard Hamarsland <https://orcid.org/0000-0002-7343-2774>, Stian Ellefsen <https://orcid.org/0000-0003-0467-2018>
+ORCID: Kristian Lian <https://orcid.org/0000-0002-5169-5960>, Daniel Hammarström <https://orcid.org/0000-0001-8360-2100>, Håvard Hamarsland <https://orcid.org/0000-0002-7343-2774>, Knut Sindre Mølmen<https://orcid.org/0000-0001-8924-6848>, Stian Ellefsen <https://orcid.org/0000-0003-0467-2018>
 
@@
 Indeed, UBF phosphorylation, which is required for interaction with the ribosomal deoxyribonucleic acid (rDNA) promoter, seems to be increased by high-glucose treatment in a mTORC1-dependent manner (rapamycin sensitive) in kidney glomerular epithelial cells [@mariappan_ribosomal_2011].
-In addition, mTORC1 regulation of ribosomal gene transcription is dependent on ribosomal protein S6 kinase beta-1 (S6K1) [@hannan2003mtor] with its target substrate ribosomal protein S6 (rpS6) coinciding with increases in rRNA and total RNA during early phase RT in humans [@hammarstrom_ribosome_2022].
+In addition, mTORC1 regulation of ribosomal gene transcription is dependent on ribosomal protein S6 kinase beta-1 (S6K1) [@hannan2003mtor] with its target substrate ribosomal protein S6 (rpS6) also increasing during early phase RT in humans [@hammarstrom_ribosome_2022].
 Independent of UBF and mTORC1, high glucose was shown to lead to chromatin remodelling, which promoted rRNA transcription in cell cultures [@zhai_human_2012].
@@
 
-The five-session-RT intervention led to increases in total RNA and rRNA per unit muscle weight both in the glucose and the placebo condition, on average by \~20-27% <!--(glucose: 263 ± 50 mg/ng-1, placebo: 210 ± 121 mg/ng-1)--> and \~25-57% <!--(47S; 0.253 ± 1.27 and 0.576 ± 0.677, 18S; 0.336 ± 0.460 and 0.271 ± 0.470, 28S; 0.314 ± 0.504 and 0.311 ± 0.582, 5.8S; 0.388 ± 0.576 and 0.322 ± 0.520, 5S; 0.305 ± 0.608 and 0.292 ± 0.432, arbitrary units for glucose and placebo respectively)--> , respectively (Figure 2).
+The five-session-RT intervention led to increases in total RNA and rRNA per unit muscle weight both in the glucose and the placebo condition, on average by \~20-27% (glucose: 263 ± 50 mg/ng-1, placebo: 210 ± 121 mg/ng-1) and \~25-57% (47S; 0.253 ± 1.27 and 0.576 ± 0.677, 18S; 0.336 ± 0.460 and 0.271 ± 0.470, 28S; 0.314 ± 0.504 and 0.311 ± 0.582, 5.8S; 0.388 ± 0.576 and 0.322 ± 0.520, 5S; 0.305 ± 0.608 and 0.292 ± 0.432, arbitrary units for glucose and placebo respectively), respectively (Figure 2).
 However, RT with glucose did not induce increased accumulation of total RNA (Figure 2A, mean difference `r total_rna_stats`) or rRNA (Figure 2B; 47S, `r stat_rrna$trg_47s`; 18S, `r stat_rrna$trg_18s`; 28S, `r stat_rrna$trg_28s`; 5.8S, `r stat_rrna$trg_5.8s`; 5S, `r stat_rrna$trg_5s`) compared to RT with placebo.

### 2024-01-15 resources/older manuscript versions/revision_notes.RMD
revision work

--- 2024-01-09 resources/older manuscript versions/revision_notes.RMD
@@
 - Authors reply: We did expect protein ingestion to possible impact plasma insulin levels, hence the participants waited for 2hrs from ingesting protein before exercising, thus "washing out" the potential effect of protein on plasma insulin levels.
-Further, 30 minutes before exercise there was no significant change in c-peptide levels, indicating no significant change in plasma insuline at this time.
-What occured between protein ingestion and the 30 min pre exericise mark, and/or if anything, this impacted ribosome biogenesis prior to glucose ingestion and exercise.
-We decided to measure plasma c-peptide rather than insulin.
+Further, 30 minutes before exercise there was no significant change in c-peptide levels, indicating no significant change in plasma insulin at this time.
+What occured between protein ingestion and the 30 min pre exericise mark, and if anything, this impacted ribosome biogenesis prior to glucose ingestion and exercise we cannot say for certain.
+- We decided to measure plasma c-peptide rather than insulin.
 Indeed this provides an indirect measure as opposed to directly measuring plasma insulin, however, c-peptide stays in the blood longer than insulin, making c-peptide easier to accurately measure over several time points [@leighton2017practical].
-Further, c-peptide is widely used as an insulin secretion measure [@leighton2017practical].
+Further, c-peptide is widely used as an insulin secretion measure, and should provide insight into changes in plasma insulin levels [@leighton2017practical].
 - Er det noe argument for at ikke-signifikante endringer i plasma glukose 30 min etter protein inntak også underbygger lite endring i plasma insulin?
@@
  - Authors reply: Results chapter now updated with absolute values of change from baseline to post in both conditions
-   - Glucose revision: Glucose ingestion before and after RT led to increases in plasma glucose levels compared to baseline by 38% immediately before RT (Figure 1B, 0 min: 2.05 ± 0.73 mmol/L), by 31% during RT (Figure 1B, 15 min: 1.75 ± 1.44 mmol/L) and by 32% immediately after RT (Figure 1B, 30 min: 1.62 ± 1.10 mmol/L, all : *p* \< 0.001), with no changes observed in the placebo condition (Figure 1B, 0 min: 0.09 ± 0.3 mmol/L, 15 min: 0.16 ± 0.35 mmol/L, 30 min: 0.18 ± 0.39 mmol/L, *p* \> 0.05).
-   - C-peptide revision: Glucose ingestion before and after RT led to increases in levels of c-peptide compared to baseline, by 95% immediately before (Figure 1C, 0 min: 796 ± 376.0 pmol/L) and 87% after RT (Figure 1C, 30 min: 793 ± 581.0 pmol/L, both *p* \< 0.001), with no changes observed with the placebo condition (Figure 1C, 0 min: 63.7 ± 71.0 pmol/L, 30 min: 53.9 ± 134.0 pmol/L, both *p* \> 0.05).
+   - Glucose results revision: Glucose ingestion before and after RT led to increases in plasma glucose levels compared to baseline by 38% immediately before RT (Figure 1B, 0 min: 2.05 ± 0.73 mmol/L), by 31% during RT (Figure 1B, 15 min: 1.75 ± 1.44 mmol/L) and by 32% immediately after RT (Figure 1B, 30 min: 1.62 ± 1.10 mmol/L, all : *p* \< 0.001), with no changes observed in the placebo condition (Figure 1B, 0 min: 0.09 ± 0.3 mmol/L, 15 min: 0.16 ± 0.35 mmol/L, 30 min: 0.18 ± 0.39 mmol/L, *p* \> 0.05).
+   - C-peptide results revision: Glucose ingestion before and after RT led to increases in levels of c-peptide compared to baseline, by 95% immediately before (Figure 1C, 0 min: 796 ± 376.0 pmol/L) and 87% after RT (Figure 1C, 30 min: 793 ± 581.0 pmol/L, both *p* \< 0.001), with no changes observed with the placebo condition (Figure 1C, 0 min: 63.7 ± 71.0 pmol/L, 30 min: 53.9 ± 134.0 pmol/L, both *p* \> 0.05).
    - Absolute values for humac added in a table (tab3), reference by "Table 3 shows the mean change in absolute peak torque values per condition and angular velocity." at the end of the strength index results.
-   - western/rna??Prøve å inkludere rådata i figurene?
+   - Baseline strength values removed from tab 1, since a detailed table of humac measurements has been added (table 3).
+   - Total total RNA and rRNA revision: Addition of: (glucose: 263 ± 50 mg/ng-1, placebo: 210 ± 121 mg/ng-1), and; (47S; 0.253 ± 1.27 and 0.576 ± 0.677, 18S; 0.336 ± 0.460 and 0.271 ± 0.470, 28S; 0.314 ± 0.504 and 0.311 ± 0.582, 5.8S; 0.388 ± 0.576 and 0.322 ± 0.520, 5S; 0.305 ± 0.608 and 0.292 ± 0.432, arbitrary units for glucose and placebo respectively)
 
@@
 This aspect is important for characterizing the sample and as stated by the authors on page 3, line 57.
- - Authors comment: The participants were asked not to use any other form of supplements during the course of the study.
+ - Authors reply: The participants were asked not to use any other form of supplements during the course of the study.
 However, I can see that this is not clearly stated in the methods chapter, and will now be added.
@@
 I suggest that this be addressed in the introduction for better understanding of the reader.
+ - Authors reply: Added the following;
+   - Introduction: "In addition, mTORC1 regulation of ribosomal gene transcription is dependent on ribosomal protein S6 kinase beta-1 (S6K1) [@hannan2003mtor] with its target substrate ribosomal protein S6 (rpS6) also increasing during early phase RT in humans [@hammarstrom_ribosome_2022]"
+   - Denne føler jeg muligens er litt svak, men slet litt med å finne en god måte å bake inn rpS6 på.
 
-Reviewer 2, comment 5: Figure 1d, use only the word "post" or "after".
- - Authors answer: Figure 1d updated only to use the word "after" instead of the mix of "after" and "post".
+<Reviewer 2, comment 5: Figure 1d, use only the word "post" or "after".
+ - Authors reply: Figure 1d updated only to use the word "after" instead of the mix of "after" and "post".
  - Men, er ikke helt sikker på om jeg egentlig vil bruke "after" eller "post" her.
@@
 Reviewer 2, comment 6: In Table 1, provide body fat percentage data, together with the absolute lean mass data (already shown), they become more informative to the reader.
- - Må hente ut denne dataen. Tabellen er oppdatert med tilfeldige tall for kroppsfett som "placeholder". Trenger bare å plotte inn rett data i dxa arket.
+ - Authors reply: Updated table 1 with body fat %
 
 Reviewer 2, comment 7: Please specify how the "blinded taste test" (described on page 6, line 120) was carried out.
- - Authors answer: Added following description of the blinded taste test (page 5, line 120-121).
+ - Authors reply: Added following description of the blinded taste test:
    - "In this blinded taste test, the participants were given two boluses of placebo (75ml) and two boluses of glucose (75ml), consumed in a randomised order.
@@
 Reviewer 2, comment 8: Mention more characteristics related to the training of the participants at baseline, such as number of weekly or 14-day RT sessions (which was an inclusion criteria).
- - Authors answer: Sadly, we do not have detailed characteristics on individual training sessions per week prior to enrollment in the study, other than that they did indeed have a weekly resistance training session volume between two and eight sessions per 14 days for the last six months.
-The authors realise that this leaves room for variation between participants in terms of training status a baseline, however variations such as these should be controlled for by the within-participants analyses.
+ - Authors reply: Sadly, we do not have detailed characteristics on individual training sessions per week prior to enrollment in the study, other than that they did indeed have a weekly resistance training session volume between two and eight sessions per 14 days for the last six months.
+The authors realise that this leaves room for variation between participants in terms of training status at baseline, however variations such as these should be controlled for by the within-participants analyses.
+
+Other changes: Added Knut Sindre Mølmen to authors, and author contributions (conception of idea, experiments and edition of manuscript)

### 2024-01-29 manuscript_revised.RMD
updating revised manuscript

@@
 
-# Glucose ingestion before and after resistance training does not augment ribosome biogenesis in healthy moderately trained young adults
+# Glucose ingestion before and after resistance training sessions does not augment ribosome biogenesis in healthy moderately trained young adults
 
@@
 ::: {custom-style="correspondence_style"}
-Correspondance: Kristian Lian ([kristian.lian\@inn.no](mailto:kristian.lian@inn.no){.email}), Stian Ellefsen ([stian.ellefsen\@inn.no](mailto:stian.ellefsen@inn.no){.email})
+Correspondance: Kristian Lian ([kristian.lian\@inn.no](mailto:kristian.lian@inn.no){.email})
 :::
@@
 ## Abstract
-
-<!-- To do: Fix tables, add outlier figures to the readme -->
 
@@
 
-**Purpose:** Resistance training-induced skeletal muscle hypertrophy seems to depend on ribosome biogenesis and content.
-High glucose treatment may augment ribosomal biogenesis, as evident from *in vitro* studies, and may thus act to potentiate resistance training-induced adaptations *in vivo*.
-This was investigated with total RNA and ribosomal RNA abundances as main outcomes, with relevant transcriptional or translational regulators (c-Myc/UBF/rpS6) as a secondary outcome.
-
-**Methods:** Sixteen healthy, moderately trained individuals (male/female, n = 9/7; age, 24.1 (3.3)), participated in a within-participant crossover trial, conducting unilateral resistance training (leg press and knee extension, 3 sets of 10 repetitions maximum) with pre- and post-exercise ingestion of either glucose (3x30g, 90g total) or placebo supplement (Stevia rebaudiana extract, 3x0.3g, 0.9g total), together with protein (2x25g, 50g total), on alternating days for twelve days; six sessions per condition, resistance training was performed in an otherwise fasted state, as morning sessions.
+**Purpose:** Resistance training-induced skeletal muscle hypertrophy seems to depend on ribosome biogenesis and -content.
+High glucose treatment may augment ribosome biogenesis, as evident from *in vitro* studies, and may thus act to potentiate resistance training-induced adaptations *in vivo*.
+This was investigated with total RNA and ribosomal RNA as main outcomes, with relevant transcriptional or translational regulaters (c-Myc/UBF/rpS6) as a secondary outcome.
+
+**Methods:** Sixteen healthy, moderately trained individuals (male/female, n = 9/7; age, 24.1 (3.3)) participated in a within-participant crossover trial where they conducted unilateral resistance training (leg press and knee extension, 3 sets of 10 repetitions maximum) with pre- and post-exercise ingestion of either glucose (3x30g, 90g total) or placebo supplements (Stevia rebaudiana, 3x0.3g, 0.9g total), together with protein (2x25g, 50g total), on alternating days for twelve days.
+Six morning resistance exercise sessions were conducted per condition, and the sessions were performed in an otherwise fasted state.
 Micro biopsies were sampled from m.
@@
 **Results:** Glucose ingestion did not have beneficial effects on resistance training-induced increases of ribosomal content (mean difference 7.6% [-7.2, 24.9], *p* = 0.34; ribosomal RNA, 47S/18S/28S/5.8S/5S, range 7.6%-37.9%, *p* = 0.40-0.98) or levels of relevant transcriptional or translational regulators (c-MYK/UBF/rpS6, *p* = 0.094-0.292).
-Of note, both baseline and trained state data of total RNA showed a linear relationship with UBF; a ∼14% increase in total RNA corresponded to 1 standard deviation unit increase in UBF (*p* = 0.003).
-
-**Conclusion:** Glucose ingestion before and after resistance training did not augment ribosomal RNA accumulation during twelve days of heavy-load resistance training in moderately trained young adults.
+Of note, both baseline and trained state data of total RNA showed a linear relationship with UBF; a ∼14% increase in total RNA corresponded to 1 SD unit increase in UBF (*p* = 0.003).
+
+**Conclusion:** Glucose ingestion before and after resistance training sessions did not augment ribosomal RNA accumulation during twelve days of heavy-load resistance training in moderately trained young adults.
 
@@
 
-Gratitude is due, to Henriette Spilhaug Bollandsås, Nora Lerdalen Bonsak and Thea Bøhn Nyløkken for their dedication and hard work during the training intervention.
+Gratitude is due, to Henriette Spilhaug Bollandsås, Nora Lerdalen Bonsak, Thea Bøhn Nyløkken, Max Ullrich and Chris Sylstad for their dedication and hard work during the training intervention.
 The authors would also like to express their gratitude to all the participants for staying committed to this demanding protocol in an impressive way.
@@
 
-Responses to systematic resistance training (RT) vary widely between individuals, with as much as 10-15% showing impaired skeletal muscle growth in response to standardized training interventions [@alvarez_interindividual_2018; @mann_high_2014; @thalacker-mercer_cluster_2013].
-There may be differences in the genetic predisposition in terms of readiness to adapt to exercise [@thalacker-mercer_cluster_2013], however, the internal physiological milieu seems favourably shaped for muscle growth by factors such as different types of training stress and nutrients [@figueiredo_genetic_2021; @sebata_control_2018; @thalacker-mercer_cluster_2013].
-Indeed, training stress and nutrients impact the ability to synthesise ribosomes, which in turn is connected to the magnitude of RT-induced responses in an individual [@hammarstrom_benefits_2020; @kusnadi_regulation_2015; @sebata_control_2018].
-For instance, increasing training volume generally induces greater ribosome biogenesis and thus, greater benefits of RT [@hammarstrom_benefits_2020; @krieger_single_2009; @schoenfeld_strength_2017].
+Responses to systematic resistance training (RT) vary widely between individuals, with as much as 10-15% showing impaired skeletal muscle growth in response to standardised training interventions [@alvarez_interindividual_2018; @mann_high_2014; @thalacker-mercer_cluster_2013].
+Genetic predisposition may explain some of this variation [@thalacker-mercer_cluster_2013], but in general the internal physiological milieu seems to be favourably shaped for muscle growth in response to different types of nutrient intake and exercise training [@figueiredo_genetic_2021; @sebata_control_2018; @thalacker-mercer_cluster_2013].
+Indeed, exercise training and nutrients impact the ability to synthesise ribosomes, which in turn is demonstrated to be connected to the magnitude of RT-induced responses [@hammarstrom_benefits_2020; @kusnadi_regulation_2015; @sebata_control_2018].
+For instance, increasing training volume generally induces greater ribosome biogenesis and is associated with greater benefits of RT in terms of gain in skeletal muscle mass and -strength [@hammarstrom_benefits_2020; @krieger_single_2009; @schoenfeld_strength_2017].
 Still, as evident from Hammarström et al.
-(2020), not all participants experience increased muscle mass and -strength despite increasing training volume.
-Therefore, other means than modification of RT variables alone seem necessary to optimise individual responses to RT.
-
-Nutritional supplements such as protein and creatine are often used to optimise RT adaptations [@cermak_protein_2012; @lanhers_creatine_2015; @lanhers_creatine_2017; @morton_systematic_2018].
-However, it remains equivocal if other nutritional adjuvants such as glucose can increase the efficacy of RT [@tezze_anabolic_2023].
-This is surprising since glucose is the preferred energy substrate of the contracting skeletal muscle during strenuous exercise and a major energy supplier to cells via adenosine triphosphate (ATP) synthesis [@mul_exercise_2015; @sebata_control_2018].
-Furthermore, energy availability is a decisive factor in the *de novo* synthesis of ribosomes [@kusnadi_regulation_2015; @moss_housekeeper_2007; @sebata_control_2018] which in turn determines muscle growth by increasing the muscle's translational capacity [@figueiredo_regulation_2019; @hammarstrom_benefits_2020; @sebata_control_2018; @stec_ribosome_2016; @von_walden_ribosome_2019].
-In addition, insulin may have anabolic effects with elevated amino acids, and has a clear role in the reduction of muscle protein breakdown independent of amino acid availability [@abdulla_role_2016; @hillier_physiological_2000].
-Therefore, investigating the effects of combined RT and glucose ingestion may provide valuable insight into the potential additive effect of glucose and RT on ribosome biogenesis.
-
-Ribosomal biogenesis and content seem to be a prerequisite for skeletal muscle growth, and transcription of ribosomal ribonucleic acid (rRNA) by RNA Polymerase I is considered the rate-limiting step in synthesising new ribosomes [@moss_promotion_1995].
-Multiple signalling pathways converge to regulate rRNA transcription, including c-Myc and the mammalian target of rapamycin complex 1 (mTORC1) signal-transduction pathway [@kusnadi_regulation_2015; @von_walden_ribosome_2019; @sebata_control_2018; @west_acute_2016; @mori_c-myc_2021].
-The general transcription factor c-Myc increases ribosomal biogenesis directly through transcriptional control of the upstream binding factor (UBF) [@sanij_ubf_2008; @poortinga_c-myc_2011; @mori_c-myc_2021; @west_acute_2016].
-Indeed, UBF phosphorylation, which is required for interaction with the ribosomal deoxyribonucleic acid (rDNA) promoter, seems to be increased by high-glucose treatment in a mTORC1-dependent manner (rapamycin sensitive) in kidney glomerular epithelial cells [@mariappan_ribosomal_2011].
-In addition, mTORC1 regulation of ribosomal gene transcription is dependent on ribosomal protein S6 kinase beta-1 (S6K1) [@hannan2003mtor] with its target substrate ribosomal protein S6 (rpS6) also increasing during early phase RT in humans [@hammarstrom_ribosome_2022].
-Independent of UBF and mTORC1, high glucose was shown to lead to chromatin remodelling, which promoted rRNA transcription in cell cultures [@zhai_human_2012].
-The mTORC1 pathway receives input from growth factors, hormones, mechanical loading, and nutrients to balance protein synthesis through multiple mechanisms based on cellular energy levels [@hoppe_amp-activated_2009] and contributes to ribosome biogenesis through the forming of the preinitiation complex (PIC) that marks the initiation of rRNA transcription and through the regulation of ribosomal protein translation [@figueiredo_regulation_2019; @von_walden_ribosome_2019].
-Moreover, mTORC1 and S6K1 serves as direct mediators of insulin signalling in skeletal muscle [@hillier_physiological_2000].
-Together these observations indicate a potential role of glucose in positively affecting human skeletal muscle rRNA synthesis, through potentiating transcription or translation of ribosomal RNA and increasing the translational capacity of skeletal muscle cells [@sebata_control_2018; @zhai_human_2012; @hillier_physiological_2000; @hoppe_amp-activated_2009].
-
-There have been multiple studies recently suggesting translational capacity to be as important, if not more important than, translational efficiency concerning long-term skeletal muscle adaptations to RT [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022; @figueiredo_revisiting_2019].
-Previously, *in vitro* studies have shown UBF and c-Myc to be important factors in the PIC, c-Myc as a general transcription factor and UBF as a specific transcription factor for rDNA transcription initiation [@mariappan_ribosomal_2011; @von_walden_ribosome_2019].
-Coupled with the recent observations by Hammarström et al.
-[@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022], understanding underlying factors regulating and affecting ribosome biogenesis seems key to furthering our understanding of the optimisation of RT to individual phenotypes.
-
-Therefore, the purpose of this investigation was to test the hypothesis that glucose supplementation before and after five RT sessions over 12 days will potentiate RT-associated accumulation of markers of ribosomal abundance following five RT sessions.
+(2020), not all participants experience increased muscle mass and improved muscle strength with increasing training volume.
+Therefore, means other than modification of RT variables seem necessary to optimise individual responses to RT, for instance nutritional adjuvants.
+
+Nutritional supplements such as protein and creatine are frequently advocated as means to optimise RT adaptations [@cermak_protein_2012; @lanhers_creatine_2015; @lanhers_creatine_2017; @morton_systematic_2018].
+The efficacy of other nutritional adjuvants such as glucose remains equivocal [@tezze_anabolic_2023].
+This is surprising given that glucose is the preferred energy substrate of the contracting skeletal muscle during strenuous exercise and a major energy supplier to cells via adenosine triphosphate (ATP) synthesis [@mul_exercise_2015; @sebata_control_2018].
+Furthermore, energy availability is a decisive factor in the *de novo* synthesis of ribosomes [@kusnadi_regulation_2015; @moss_housekeeper_2007; @sebata_control_2018] which in turn seems to determine muscle growth by increasing the muscle's translational capacity [@figueiredo_regulation_2019; @hammarstrom_benefits_2020; @sebata_control_2018; @stec_ribosome_2016; @von_walden_ribosome_2019].
+In addition, insulin, which is secreted from the beta cells of pancreas in response to rising blood glucose levels, may itself exert anabolic effects irrespective of muscle contraction, e.g.
+by elevating levels of amino acids, an also plays a role in reducing muscle protein breakdown independent of amino acid availability [@abdulla_role_2016; @hillier_physiological_2000].
+It seems plausible that combined RT and glucose ingestion provides effects on ribosome biogenesis compared to RT alone.
+
+Ribosome biogenesis and content seem to be a prerequisite for skeletal muscle growth, and transcription of ribosomal ribonucleic acid (rRNA) by RNA Polymerase I is considered the rate-limiting step in *de novo* ribosome biogenesis [@moss_promotion_1995].
+Multiple proteins and signalling pathways converge to regulate rRNA transcription, including c-Myc and the mammalian target of rapamycin complex 1 (mTORC1) signal-transduction pathway [@kusnadi_regulation_2015; @von_walden_ribosome_2019; @sebata_control_2018; @west_acute_2016; @mori_c-myc_2021].
+First, the general transcription factor c-Myc increases ribosome biogenesis directly through transcriptional control of the upstream binding factor (UBF) [@sanij_ubf_2008; @poortinga_c-myc_2011; @mori_c-myc_2021; @west_acute_2016].
+Indeed, UBF phosphorylation, which is required for interaction with the ribosomal deoxyribonucleic acid (rDNA) promoter, is increased by RT alone in muscle and also seems to be increased by high-glucose treatment in a mTORC1-dependent manner (rapamycin sensitive) in kidney glomerular epithelial cells [@mariappan_ribosomal_2011].
+
+Second, the mTORC1 pathway receives input from growth factors, hormones, mechanical loading, and nutrients to balance protein synthesis through multiple mechanisms based on cellular energy levels [@hoppe_amp-activated_2009].
+This contributes to ribosome biogenesis by forming the preinitiation complex (PIC) that marks the initiation of rRNA transcription, as well as through regulation of ribosomal protein translation [@figueiredo_regulation_2019; @von_walden_ribosome_2019].
+In addition, mTORC1 and ribosomal protein S6 kinase beta-1 (S6K1) are direct mediators of insulin signalling in skeletal muscle [@hillier_physiological_2000].
+Third, high glucose was shown to lead to chromatin remodelling independent of UBF and mTORC1, which in turn promotes rRNA transcription in cell cultures [@zhai_human_2012].
+Together, these mechanistic observations underscore a potential role for glucose in muscle ribosome biogenesis and function in human skeletal muscle, acting in concert with RT to potentiate transcription of ribosomal RNA and increasing translational capacity [@sebata_control_2018; @zhai_human_2012; @hillier_physiological_2000; @hoppe_amp-activated_2009].
+
+Multiple studies have suggested translational capacity to be as important, if not more important, than translational efficiency for promoting long-term skeletal muscle adaptations to RT [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022; @figueiredo_revisiting_2019].
+While the regulation of translational capacity itself involves activation of c-Myc and UBF, acting to stimulate formation of the PIC through the general transcription factor, as well as through a specific transcription factor facilitating rDNA transcription initiation [@mariappan_ribosomal_2011; @von_walden_ribosome_2019], the content of ribosomes, c-Myc and UBF is increased with RT-induced muscle accretion [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022].
+Furthermore, Nakada et al.
+[@nakada2016correlation] found a correlation between rRNA content and rpS6 content in their synergist ablation model on rats.
+This makes knowledge about factors that regulate and affect ribosome biogenesis key for optimising RT at the individual level.
+
+Therefore, the main purpose of this investigation was to test the hypothesis that glucose supplementation given before and after six RT sessions, conducted over a period of 12 days will potentiate RT-associated accumulation of markers of ribosomal abundance.
 Secondly, we aimed to describe the association between changes in total RNA abundance and UBF in human skeletal muscle.
@@
 
-All participants gave their written informed consent before study enrollment.
-The study was approved by the regional ethical committee (REK, ID nr.
+All participants gave their written informed consent before study enrolment.
+The study was approved by the Regional Committee for Medical and Health Research Ethics - South-East Norway (ID nr.
 153628), pre-registered at clinicaltrials.gov (Identifier: NCT04545190), and conducted according to the Helsinki Declaration.
@@
 Sixteen healthy male and female participants (20-33 years, Table 1) were recruited to the study through social media advertisement and word of mouth.
-The eligibility criteria were non-smokers and moderately trained (i.e.
+The eligibility criteria were non-smoking and moderately trained (i.e.
 2-8 RT sessions per 14 days for the last six months).
-Exclusion criteria were previous injury leading to impaired strength, inability to perform resistance exercise training, symptoms, and a medical record of metabolic disorders including hyperglycemia.
+Exclusion criteria were previous injury leading to impaired muscle strength, inability to perform resistance exercise training, symptoms, and a medical record of metabolic disorders including hyperglycemia.
 Of the sixteen participants that commenced the intervention, three participants dropped out.
-One due to sickness and inability to resume, two participants experienced muscular discomfort related to heavy resistance training.
-Lean mass (Table 1) was measured using Dual-Energy X-ray Absorptiometry (DXA, Prodigy Advance PA+302047, Lunar, San Francisco, CA, USA) on Day -1, the last day preceding the RT intervention.
+One due to sickness and inability to resume, while two participants experienced muscular discomfort related to heavy resistance training.
+Lean mass and body fat % (Table 1) was measured using Dual-Energy X-ray Absorptiometry (DXA, Prodigy Advance PA+302047, Lunar, San Francisco, CA, USA) on Day -1, the last day preceding the RT intervention.
+
+Table 1 could be placed here.
 
@@
 One person was exclusively responsible for the randomisation code and supplement distribution, blinding both investigators and participants regarding which leg exercised with glucose/placebo conditions.
-Glucose (Glucosum monohydricum, Merck KGaA, Darmstadt, Germany) and placebo (Steviosa, Soma Nordic AS, Oslo, Norway) were masked by mixing with 300ml Fun Light (Orkla, Oslo, Norway).
-A blinded taste test revealed that the participants were not able to disclose the contents of the provided boluses (30g glucoseum monohydricum vs.
-\~0.3g Stevia rebaudiana extract).
-In this blinded taste test, the participants were given two boluses of placebo (75ml) and two boluses of glucose (75ml), consumed in a randomised order.
-The participants were instructed to finish one bolus, guess the content, and move on to the next bolus.
+One bolus of glucose was ingested as 30g of glucose (Glucosum monohydricum, Merck KGaA, Darmstadt, Germany) and one bolus of placebo was ingested as 100mg Stevia powder (Steviosa, Soma Nordic AS, Oslo, Norway), containing the natural sweetener erythritol in amounts equivalent to the sweetness of 30g glucose, mixed with 300ml Fun Light (Orkla, Oslo, Norway).
+Hence, the glucose and placebo supplements had identical chemical composition, except for their content of glucose/Stevia.
+To test whether the boluses truly were masked sufficiently to avoid detection, a blinded taste test was carried out.
+In this blinded taste test, the participants were given two glasses of glucose mix (75ml per) and two glasses of the placebo mix (75ml per), consumed in a randomised order per participant.
+The participants were instructed to finish one bolus, note their guess for its content, and move on to the next glass.
+On average, the participants had a score of 2 points (2.08 ± 1.24) out of 4 possible.
 To ensure equal conditions during training sessions and strength testing, participants exercised and tested at the same time of day, +/- 1hr with the same supervisor on pairwise consecutive days (i.e.
@@
 All participants completed six RT sessions with glucose and six with placebo, allowing a within-subjects analysis of the effects of glucose ingestion before and after RT.
-Data from the first five RT sessions was used to investigate main outcome measures (total RNA, rRNA and protein) and leg muscle strength, whereas data from the sixth RT session was used to explore secondary outcomes (muscular recovery, plasma glucose and serum c-peptide levels).
+Data from the first five RT sessions was used to investigate main outcome measures (total RNA, rRNA and protein) and leg muscle strength, whereas data from the sixth RT session was used to explore secondary outcomes (muscular recovery, plasma glucose and plasma c-peptide levels).
 Participants were asked to avoid resistance- or high-intensity training of the legs from Day -7 (Figure 1A) and onwards, until completion of the intervention and post-testing, to ensure the reliability of pre-intervention strength data and minimal interference from external exercise sources.
@@
 During period I, participants ingested protein and glucose *or* protein and placebo only.
-Glucose/placebo was ingested three times in period I: 30 minutes before RT (0830hrs, 30g vs.
-0g glucose), immediately before RT (0900hrs, 30g vs.
-0g glucose), and immediately after RT (\~0930hrs, 30g vs.
+Glucose/placebo was ingested three times in period I: 30 minutes before RT (30g vs.
+0g glucose), immediately before RT (30g vs.
+0g glucose), and immediately after RT (30g vs.
 0g glucose).
-Whey Protein Isolate (Proteinfabrikken, Stokke, Norway) was ingested 2hrs before RT (0700hrs) and immediately after RT (\~0930hrs), in boluses of 25 g mixed with 150 ml water.
-In the afternoon (1800hrs-1900hrs, period II) participants ingested glucose or placebo (3x30g vs.
+Whey Protein Isolate (Proteinfabrikken, Stokke, Norway) was ingested 2hrs before RT and immediately after RT, in boluses of 25 g mixed with 150 ml water.
+In the afternoon (1800-1900hrs, period II) participants ingested glucose or placebo (3x30g vs.
 3x0g glucose) opposite to the supplement they received during RT, to ensure a balanced daily intake of glucose.
@@
 The daily onset of the dietary intervention (i.e., first ingestion of protein supplement) varied between participants, from 0600hrs to 0900hrs to allow multiple participants to complete the protocol simultaneously.
-During sessions, participants were free to ingest water ad libitum.
+During resistance training sessions, participants were free to ingest water ad libitum.
 
@@
 
-Strength tests were performed before (Figure 1A, Days -7 and -5, and -1, both legs) and during the intervention (Figure 1A, on days 4 and 8 for leg 1, and days 5 and 9 for leg two), after session 5 and after finalization of the intervention (Figure 1A, on days 11/12 for leg 1 and days 12/13 for leg 2).
+Muscle strength tests were performed before (Figure 1A, Days -7 and -5, and -1, both legs) and during the intervention (Figure 1A, on days 4 and 8 for leg 1, and days 5 and 9 for leg two), after session 5 and after finalization of the intervention (Figure 1A, on days 11/12 for leg 1 and days 12/13 for leg 2).
 Maximal isometric and isokinetic knee extension torque was measured with a Humac Norm Dynamometer (CSMi, Stoughton, Massachusetts, USA).
@@
 The isometric tests were ended when the participants reached a plateau or peak torque development decreased, which on average occurred between 2-4 seconds into the test.
-During days 4, 5, 8 and 9 (Figure 1D, days 4 and 5 = Post 2RT, days 8 and 9 = Post 4RT), humac tests were conducted one hour before RT on the leg performing RT the previous day.
-During days 11 and 12, humac tests were performed four times: I) 45min before the last RT session (Figure 1D, Post 5RT), II) 30min after the last RT session (Figure 1D, 30 min post 6RT), III) 2hrs after the last RT session (Figure 1D, 2h post 6RT), and IV) 23hrs after the last RT session (Figure 1D, 23h post 6RT).
+During days 4, 5, 8 and 9 (Figure 1D, days 4 and 5 = Post 2RT, days 8 and 9 = Post 4RT), knee extension torque tests were conducted one hour before RT on the leg performing RT the previous day.
+During days 11 and 12, these tests were performed four times: I) 45min before the last RT session (Figure 1D, Post 5RT), II) 30min after the last RT session (Figure 1D, 30 min post 6RT), III) 2hrs after the last RT session (Figure 1D, 2h post 6RT), and IV) 23hrs after the last RT session (Figure 1D, 23h post 6RT).
 Test I on day 12/13 included testing of both legs, representing 23hrs post-RT session test of one leg and post-session 5 test of the other leg.
@@
 
-Assessment of unilateral one repetition maximum (1RM) leg press and knee extension was conducted at before the intervention during familiarization (Figure 1, Days -7 and -5).
+Assessment of unilateral one repetition maximum (1RM) leg press and knee extension was conducted in the familiarisation phase prior to the intervention (Figure 1, Days -7 and -5).
 The participants performed a general warm-up with 10 minutes of cycling on an indoor exercise bicycle.
-A protocol consisting of 1x10, 1x6 and 1x3 repetitions with a load equivalent to \~50-75% of assumed max repetitions, was used as a specific warm-up before each of the tests.
-All positions were controlled and recorded at the first 1RM test and repeated for the RT sessions.
+A protocol consisting of 1x10, 1x6 and 1x3 repetitions with a load equivalent to \~50-75% of assumed maximal repetitions, was used as a specific warm-up before each of the tests.
+All positions were controlled and recorded at the first 1RM test and reproduced during the RT sessions.
 Maximal leg press strength was defined as the maximal load lifted in a controlled fashion, starting from a knee angle of 90 degrees.
@@
 Attempts where participants did not reach 90 degrees during the eccentric phase, were not approved.
-Maximal knee extension testing followed the same specific warm-up as the maximal leg press test and was defined as the maximal load lifted in a controlled fashion, reaching full extension of the knee joint.
+Maximal knee extension strength testing followed the same specific warm-up as the maximal leg press test and was defined as the maximal load lifted in a controlled fashion, reaching full extension of the knee joint.
 Attempts with exaggerated hip movement or beneath full extension were not approved.
@@
 
-Resistance training consisted of three sets of unilateral leg presses and three sets of unilateral knee extensions, with an intensity of 10 repetitions maximum (10RM).
+Resistance training consisted of three sets of unilateral leg presses and three sets of unilateral knee extensions, with an exercise intensity of 10 repetitions maximum (10RM).
 As a general warm-up, the participants cycled on an indoor exercise bicycle for 5-10 minutes.
@@
 One capillary blood sample was collected on day 1 (Figure 1A) to serve as a baseline.
-On days 11 and 12 (Figure 1A), capillary blood samples were collected seven times: I) Immediately before protein ingestion (0700hrs) II) 45 minutes after protein ingestion (0745hrs) III) 1.5 hours after protein ingestion (0830hrs, i.e., immediately before GLU/PLAC intake), IV) 2 hours after protein ingestion (0900hrs, i.e., immediately before training), IV) in the middle of RT (\~0915hrs), V) immediately after training (\~0930hrs), and VI) 2 hours after completion of training (\~1130hrs).
+On days 11 and 12 (Figure 1A), capillary blood samples were collected seven times: I) Immediately before protein ingestion (0700hrs) II) 45 minutes after protein ingestion (0745hrs) III) 1.5 hours after protein ingestion (0830hrs, i.e., immediately before glucose/placebo intake), IV) 2 hours after protein ingestion (0900hrs, i.e., immediately before training), IV) in the middle of RT (\~0915hrs), V) immediately after training (\~0930hrs), and VI) 2 hours after completion of training (\~1130hrs).
 Capillary blood samples were analysed with in-house equipment (BIOSEN C-Line, EKF diagnostic GmbH, Barleben).
@@
 qPCR reactions were run over 40 cycles (3 s 95°C denaturing and 30 s 60°C annealing) on a fast-cycling real-time detection system (Applied Biosystems 7500 fast Real-Time PCR Systems, Life Technologies AS), with a total reaction volume of 10 µl consisting of 2 µl of complementary DNA (cDNA), gene-specific primers (0.5 µM final concentration) and a commercial master mix (2X SYBR Select Master Mix, Applied Biosystems, Life Technologies AS, Oslo, Norway).
-An overview of the primers may be found in Table 2.
+An overview of the primers can be found in Table 2.
 Raw fluorescence data was modelled with a best-fit sigmoidal model using the qPCR package [@ritz_qpcr_2008] written for R [@hammarstrom_benefits_2020; @Rcore].
 qPCR data was normalised to wet muscle weight using the external reference gene Lambda [@ellefsen_improved_2008; @ellefsen_reliable_2014] and analysed on the log scale on a target-by-target basis.
-<!-- add explanations of normalization to extended methods -->
+
+Table 2 can be placed here.
 
@@
 Both gel electrophoresis and protein transfer were performed at 4°C.
-Following the wet transfer, membranes were stained with a reversible total protein stain (Thermo Fisher Scientific) and then blocked for 1 hour at room temp with a blocking buffer of Tris-buffered saline (TBS; 20 mM Tris, 150 mM NaCl) with 5% non-fat dry milk and 0.1% Tween-20.
+Following the wet transfer, membranes were stained with a reversible total protein stain (Thermo Fisher Scientific) and then blocked for 1 hour at room temperature with a blocking buffer of Tris-buffered saline (TBS; 20 mM Tris, 150 mM NaCl) with 5% non-fat dry milk and 0.1% Tween-20.
 Primary and secondary antibodies were purchased from Santa Cruz Biotechnology (Texas, USA): UBF, UBF F-9, sc-13125; rpS6, Ribosomal protein S6 C-8, sc-74459; and Thermo Fisher Scientific (Oslo, Norway): c-Myc, 9E10; goat anti-mouse (for c-Myc), goat anti-mouse IgG1 (y1) horseradish peroxidase conjugate; and anti-mouse (anti-mouse IgG1 horseradish peroxidase conjugate).
-<!-- Reviewer comment: 25,000 dilution for primary seems quite diluted.
-Is the dilution is correct?
-Revision: Antibodies were diluted in blocking buffer to concentrations corresponding to 1:500 (UBF and rpS6, primary), 1:2000 (c-Myc, primary), 1:5000 (c-Myc, secondary) and 1:25 000 (UBF, rpS6, secondary).-->
+Antibodies were diluted in blocking buffer to concentrations corresponding to 1:500 (UBF and rpS6, primary), 1:2000 (c-Myc, primary), 1:5000 (c-Myc, secondary), and 1:25 000 (UBF and rpS6, secondary).
 
@@
 Following the last wash, membranes were incubated for 5 minutes with enhanced chemiluminescent substrate (ECL, SuperSignal West Femto Maximum Sensitivity Substrate, Thermo Fisher Scientific).
-Membrane blocking, secondary antibody incubation, washing and ECL incubation were performed at room temp.
+Membrane blocking, secondary antibody incubation, washing and ECL incubation were performed at room temperature.
 Primary antibody incubation was performed at 4°C.
@@
 This power calculation was based on an assumption that the effects of glucose ingestion on total RNA accumulation and rRNA expression may equate to the effects of increasing RT volume from low to moderate [@hammarstrom_benefits_2020].
-Total RNA, protein and qPCR data were analyzed by mixed-effects models with fixed effects included as *supplement* $\times$ time.
+Total RNA, protein and qPCR data were analysed by mixed-effects models with fixed effects included as *supplement* $\times$ time.
 To decrease the risk of Type I errors, random effects were selected from step-wise elimination of terms from the most complex structure (random slopes for time and supplement and their interaction) to less complex.

### 2024-01-29 resources/older manuscript versions/revision_notes.RMD
updating revised manuscript

--- 2024-01-10 resources/older manuscript versions/revision_notes.RMD
@@
 editor: visual
+bibliography: resources/ribose-paper.bib
 ---
 
-### Revision
+Dear Editor, Michalis G. Noklaidis, and reviewers
 
-<!--Reviewer 1, comment 1: 25,000 dilution for primary seems quite diluted. Is the dilution is correct? Page 9, line 257-258 (line 211 in quarto) 
-- Authors reply: By mistake, only the dilution volumes for the secondary antibodies were included in the original description.
+We are grateful for all the constructive comments on our manuscript.
+We have done our best to direct all comments and suggestions in the revised manuscript.
+Attached you will find a clean version of the revised manuscript and a version that highlights (in red, with "track changes") all changes made during the revision process.
+We also provide a point-by-point response explaining how we have adressed each comment.
+Please let us know if something is still unclear.
+
+On behalf of the authors,
+
+Kristian Lian
+
+(The part above is written in the email, copy and cut it from here.)
+
+## Revision comments and answers
+
+#### Reviewer 1
+
+"Lian et al.
+investigated the effects of glucose ingestion on the ribosome biogenesis responses post resistance exercise.
+This is a very interesting research question, which was well performed by this research group.
+I overall enjoyed reading this well-written manuscript and have very little to suggest.
+Please see my comments below."
+
+**Authors reply:** Thank you very much, we are glad the reviewer enjoyed the read and found our research question interesting.
+
+*Comment 1: 25,000 dilution for primary seems quite diluted. Is the dilution is correct? Page 9, line 257-258.*
+
+**Authors reply:** Thank you for bringing this to our attention.
+By mistake, only the dilution volumes for the secondary antibodies were included in the original description.
 This section is now updated with dilution volumes of both primary and secondary antibodies.
-- Revision: 
-   - Original draft: Antibodies were diluted in blocking buffer to concentrations corresponding to 1:25 000 (UBF, rpS6) and 1:5000 (c-Myc).
-   - Revision: Antibodies were diluted in blocking buffer to concentrations corresponding to 1:500 (UBF and rpS6, primary), 1:2000 (c-Myc, primary), 1:5000 (c-Myc, secondary) and 1:25 000 (UBF, rpS6, secondary).-->
+Original draft: Antibodies were diluted in blocking buffer to concentrations corresponding to 1:25 000 (UBF, rpS6) and 1:5000 (c-Myc).
+Revision: Antibodies were diluted in blocking buffer to concentrations corresponding to 1:500 (UBF and rpS6, primary), 1:2000 (c-Myc, primary), 1:5000 (c-Myc, secondary) and 1:25 000 (UBF and rpS6, secondary).
+Added on page 9, line 293.
 
-<!--Reviewer 1, comment 2: Most of those antibodies have been discontinued by the commercial provided (Santa Cruz). Did the authors try to validate these antibodies? 
-- Authors reply: We did several rounds of antibody testing with these specific primers before analysing the muscle tissue from the present study.
-We compared the blots from said test runs to reviews on the providers own web page as a form of validation.
-- Primary antibodies are available at: 
-- UBF F9: https://www.scbt.com/p/ubf-antibody-f-9 
-- rpS6 C-8: https://www.scbt.com/p/ribosomal-protein-s6-antibody-c-8 
-- c-Myc 9E10: https://www.thermofisher.com/antibody/product/c-Myc-Antibody-clone-9E10-Monoclonal/MA1-980?imageId=1024462 
-- Finner ikke eksakte matcher for sekundære, må sjekke navnene på laben. Finner allikevel ikke noen info om at de primære antistoffene er avviklet.-->
+*Comment 2: Most of those antibodies have been discontinued by the commercial provided (Santa Cruz). Did the authors try to validate these antibodies?*
 
-<!--Reviewer 1, comment 3:Given that whey protein was also given to the participants following the resistance exercise session, could the increase insulin from whey protein ingestion already be enough to evoke the anabolic effects related to ribosome biogenesis?
+**Authors reply:** Thank you for this input.
+We did several runs of antibody testing with these specific antibodies before analysing the muscle tissue from the present study.
+We compared the blots from our test runs to expected molecular sizes a provided by the manufacturer as a form of validation.
+Additionally, no secondary bands were detected.
+Unfortunately, we are not in possession of any knock-out tissue for a complete validation.
+We believe all of the primary antibodies are still provided by the vendor:
+- UBF: https://www.scbt.com/p/ubf-antibody-f-9 
+- rpS6: https://www.scbt.com/p/ribosomal-protein-s6-antibody-c-8 
+- c-Myc: https://www.thermofisher.com/antibody/product/c-Myc-Antibody-clone-9E10-Monoclonal/MA1-980?imageId=1024462
+
+*Comment 3: Given that whey protein was also given to the participants following the resistance exercise session, could the increase insulin from whey protein ingestion already be enough to evoke the anabolic effects related to ribosome biogenesis?
 Did the authors measure plasma insulin levels?
-It would be interesting to this story to check insulin levels.
-- Authors reply: We did expect protein ingestion to possible impact plasma insulin levels, hence the participants waited for 2hrs from ingesting protein before exercising, thus "washing out" the potential effect of protein on plasma insulin levels.
-Further, 30 minutes before exercise there was no significant change in c-peptide levels, indicating no significant change in plasma insuline at this time.
-What occured between protein ingestion and the 30 min pre exericise mark, and/or if anything, this impacted ribosome biogenesis prior to glucose ingestion and exercise.
+It would be interesting to this story to check insulin levels.*
+
+**Authors reply:** Thank you for bringing up this subject.
+We did assume tat protein ingestiuon cold impact plasma insulin levels.
+Therefore, the participants waited for 2hrs from ingesting the protein supplement before starting the training sessions, thus "washing out" the potential effect of protein on plasma insulin levels before the training session.
+However, as noted by the reviewer, the participants did not wait for 2hrs after the sessions to ingest the second/last protein supplement.
+In the glucose condition sessions, plasa c-peptide levels were significantly higher at onset of training (Figure 1C, 0min) and at the end of training (Figure 1C, 30min) compared to placebo, and returned to baseline values 2hrs after training (Figure 1C, 120min).
+So, based on these data, glucose ingestion did significantly increase plasma c-peptide compared to placebo during exercise, with no measured effect of protein ingestion on plasma c-peptid levels from 2hrs pre until 2hrs post exercise.
+However, we cannot exclude that there were peaks/fluctuations in plasma c-peptide as a response to ingesting protein that we did not measure in this time frame, especially after exercise with a 2 hour-long period without measurements.
+One could arguer that if the protein ingestion *per se* stimulated a large insulin secretion, this would trigger increased glucose secretion from the liver to the bloodstream as a response to the insulin, to stabilise the plasma glucose levels.
+Our 30min post-protein ingestion measurement of plasma glucose showed no such changes at this time.
+Nevertheless, without having measured plasma c-peptide/insulin levels more frequently following the training sessions, we have no way of knowing what truly occurred in this time frame.
+
 We decided to measure plasma c-peptide rather than insulin.
-Indeed this provides an indirect measure as opposed to directly measuring plasma insulin, however, c-peptide stays in the blood longer than insulin, making c-peptide easier to accurately measure over several time points [@leighton2017practical].
-Further, c-peptide is widely used as an insulin secretion measure [@leighton2017practical].
-- Er det noe argument for at ikke-signifikante endringer i plasma glukose 30 min etter protein inntak også underbygger lite endring i plasma insulin?
-Altså, dersom det hadde vært en stor utskillelse i plasma insulin, ville ikke dette ha påvirket plasma glukose?-->
+Indeed, this provides indirect measure as opposed to direclty measuring plasma insulin.
+However, c-peptide stays in the blood for a longer time than insulin, making c-peptide easier to accurately measure over several time points [@leighton2017practical].
+Further, c-peptide is widely used as an insulin secretion measure, and is demonstrated to provide insight into changes in plasma insulin levels [@leighton2017practical].
 
-<!--Reviewer 2, comment 1: Please, also describe in the abstract the second objective of the study (lines 98-99).
-The abstract does not state, and, therefore is not clear, what are the main outcomes (parameters) measured in the clinical trial.
-- Authors reply: The abstract has now been updated with this information. 
-- Revision: 
+#### Reviewer 2
+
+"This manuscript is an original article that performed double-blinded placebo-controlled crossover trial with the main purpose to evaluate if the glucose supplementation before and after five RT sessions over 12 days will potentiate RT-associated accumulation of markers of ribosomal abundance following five RT sessions in healthy moderately trained young adults.
+The article is well written and the clinical trial well designed.
+The study presents novelty.
+Although, this clinical trial has important limitations: not reaching the calculated sample; intramuscular glycogen and circulating insulin levels were not performed."
+
+**Authors reply:** Thanks for these kind words. We are glad the reviewer appreciates the manuscript.
+
+*Comment 1: Please, also describe in the abstract the second objective of the study (lines 98-99).
+The abstract does not state, and, therefore is not clear, what are the main outcomes (parameters) measured in the clinical trial.*
+
+**Authors reply:** Thank you for pointing this out, the abstract has now been updated with this information.
   - Removed "However, this remains largely unexplored." 
-  - Added "This was investigated with total RNA and ribosomal RNA abundances as main outcomes, with relevant transcriptional or translational regulators (c-Myc/UBF/rpS6) as a secondary outcome." 
-- Abstract trenger nå mer jobb, for å kutte ned til akseptert lengde/antall ord. Er det ok med 30 ord for mye?-->
+  - Added "This was investigated with total RNA and ribosomal RNA abundances as main outcomes, with relevant transcriptional or translational regulators (c-Myc/UBF/rpS6) as a secondary outcome." On page 1, line 14-16
+The abstract is now approximately 35 words over the limit of 250 words to accommodate these changes, we hope that this does not cause any inconvenience. If necessary, we will adjust this too.
 
-Reviewer 2, comment 2: In the results the data were described in relation to % ou fold of changes. It is important to report the data also in absolute values.
- - Authors reply: Results chapter now updated with absolute values of change from baseline to post in both conditions
-   - Glucose revision: Glucose ingestion before and after RT led to increases in plasma glucose levels compared to baseline by 38% immediately before RT (Figure 1B, 0 min: 2.05 ± 0.73 mmol/L), by 31% during RT (Figure 1B, 15 min: 1.75 ± 1.44 mmol/L) and by 32% immediately after RT (Figure 1B, 30 min: 1.62 ± 1.10 mmol/L, all : *p* \< 0.001), with no changes observed in the placebo condition (Figure 1B, 0 min: 0.09 ± 0.3 mmol/L, 15 min: 0.16 ± 0.35 mmol/L, 30 min: 0.18 ± 0.39 mmol/L, *p* \> 0.05).
-   - C-peptide revision: Glucose ingestion before and after RT led to increases in levels of c-peptide compared to baseline, by 95% immediately before (Figure 1C, 0 min: 796 ± 376.0 pmol/L) and 87% after RT (Figure 1C, 30 min: 793 ± 581.0 pmol/L, both *p* \< 0.001), with no changes observed with the placebo condition (Figure 1C, 0 min: 63.7 ± 71.0 pmol/L, 30 min: 53.9 ± 134.0 pmol/L, both *p* \> 0.05).
-   - Absolute values for humac added in a table (tab3), reference by "Table 3 shows the mean change in absolute peak torque values per condition and angular velocity." at the end of the strength index results.
-   - Baseline strength values removed from tab 1, since a detailed table of humac measurements has been added (table 3).
-   - Mangler å plotte rådata i total RNA, så burde denne kommentaren være besvart.
+*Comment 2: In the results the data were described in relation to % ou fold of changes. It is important to report the data also in absolute values.*
 
-<!--Reviewer 2, comment 3: There is a lack of information regarding the use of supplements by participants.
-This aspect is important for characterizing the sample and as stated by the authors on page 3, line 57.
- - Authors comment: The participants were asked not to use any other form of supplements during the course of the study.
-However, I can see that this is not clearly stated in the methods chapter, and will now be added.
- - Revision: Page 5, lines 141-143.
-   - "Apart from this, participants ingested a self-chosen diet during period II, registered in MyFitnessPal or similar applications" changed to "Apart from this, participants ingested a self-chosen diet during period II.
-Further, participants were asked not to use any other supplements such as additional protein and/or creatine, and to register all food/drink consumption in MyFitnessPal or similar applications"-->
+**Authors reply:** Thank you for your valuable input. The results are now updated with absolute values of change from baselin to post in both conditions:
 
-Reviewer 2, comment 4: Figure 3 shows the result for RPS6.
+Glucose results revision: Glucose ingestion before and after RT led to increases in plasma glucose levels compared to baseline by 38% immediately before RT (Figure 1B, 0 min: 2.05 ± 0.73 mmol/L), by 31% during RT (Figure 1B, 15 min: 1.75 ± 1.44 mmol/L) and by 32% immediately after RT (Figure 1B, 30 min: 1.62 ± 1.10 mmol/L, all : p < 0.001), with no changes being observed in the placebo condition (Figure 1B, 0 min: 0.09 ± 0.3 mmol/L, 15 min: 0.16 ± 0.35 mmol/L, 30 min: 0.18 ± 0.39 mmol/L, p > 0.05).
+Added on page 11, line 335
+
+C-peptide results revision: Glucose ingestion before and after RT led to increases in levels of c-peptide compared to baseline, by 95% immediately before (Figure 1C, 0 min: 796 ± 376.0 pmol/L) and 87% after RT (Figure 1C, 30 min: 793 ± 581.0 pmol/L, both p < 0.001), with no changes observed with the placebo condition (Figure 1C, 0 min: 63.7 ± 71.0 pmol/L, 30 min: 53.9 ± 134.0 pmol/L, both p > 0.05).
+Addded on page 11 line 347
+
+Absolute values for knee extension torque are added to Table 3, referenced by "Table 3 shows the mean change in absolute peak torque per condition and angular velocity." on page 12 line 367.
+Baseline muscle strength values are now removed from table 1, since the detailed table of knee extension torque measurements are added (Table 3).
+
+Revision of total RNA and rRNA: (glucose: 263 ± 50 ng/mg-1 , placebo: 210 ± 121 ng/mg-1) added on page 12, line 371, and; (47S; 0.253 ± 1.27 and 0.576 ± 0.677, 18S; 0.336 ± 0.460 and 0.271 ± 0.470, 28S; 0.314 ± 0.504 and 0.311 ± 0.582, 5.8S; 0.388 ± 0.576 and 0.322 ± 0.520, 5S; 0.305 ± 0.608 and 0.292 ± 0.432, arbitrary units for glucose and placebo respectively) added on page 12 line 374.
+
+*Comment 3: There is a lack of information regarding the use of supplements by participants. This aspect is important for characterizing the sample and as stated by the authors on page 3, line 57.*
+
+*Authors reply:* Thank you for this important remark.
+The participants were asked not to use any other form of supplements during the course of the study.
+However, the authors can see that this was not clearly stated in the methods chapter, but this has been clarified in the revised manuscript.
+"Apart from this, participants ingested a self-chosen diet during period II, registered in MyFitnessPal or similar applications" has been changed to "Apart from this, participants ingested a self-chosen diet during period II.
+Further, participants were asked not to use any other supplements such as additional protein and/or creatine, and to register all food/drink consumption in MyFitnessPal or similar applications." Page 6, line 175.
+
+*Comment 4: Figure 3 shows the result for RPS6.
 However, only in the discussion section there is some description about this protein (lines 377 - 381).
-I suggest that this be addressed in the introduction for better understanding of the reader.
+I suggest that this be addressed in the introduction for better understanding of the reader.*
 
-<!--Reviewer 2, comment 5: Figure 1d, use only the word "post" or "after".
- - Authors answer: Figure 1d updated only to use the word "after" instead of the mix of "after" and "post".
- - Men, er ikke helt sikker på om jeg egentlig vil bruke "after" eller "post" her.
-Jeg synes "after" leses bedre på denne figuren, men "post" passer bedre i figur 2.
-Er det uryddig å ha forskjellig på disse to figurene?-->
+**Authors reply:** Thank you for pointing this out.
+The following has been added about rpS6 in the introduction: "Furthermore, Nakada et al.
+(Nakada, Ogasawara, Kawada, Maekawa and Ishii, 2016) found a correlation between rRNA content and rpS6 content in their synergist ablation model on rats." Page 4, line 114.
 
-<!--Reviewer 2, comment 6: In Table 1, provide body fat percentage data, together with the absolute lean mass data (already shown), they become more informative to the reader.
- - Authors answer: Updated figure one with body fat %-->
+*Comment 5: Figure 1d, use only the word "post" or "after".*
 
-<!--Reviewer 2, comment 7: Please specify how the "blinded taste test" (described on page 6, line 120) was carried out.
- - Authors answer: Added following description of the blinded taste test (page 5, line 120-121).
-   - "In this blinded taste test, the participants were given two boluses of placebo (75ml) and two boluses of glucose (75ml), consumed in a randomised order.
-The participants were instructed to finish one bolus, guess the content, and move on to the next bolus."-->
+**Authors reply:** Thank you for bringing this to our attention, Figure 1D is now updated with only "post" instead of the mix of "after" and "post".
 
-<!--Reviewer 2, comment 8: Mention more characteristics related to the training of the participants at baseline, such as number of weekly or 14-day RT sessions (which was an inclusion criteria).
- - Authors answer: Sadly, we do not have detailed characteristics on individual training sessions per week prior to enrollment in the study, other than that they did indeed have a weekly resistance training session volume between two and eight sessions per 14 days for the last six months.
-The authors realise that this leaves room for variation between participants in terms of training status a baseline, however variations such as these should be controlled for by the within-participants analyses.-->
+*Comment 6: In Table 1, provide body fat percentage data, together with the absolute lean mass data (already shown), they become more informative to the reader.*
+
+**Authors reply:** Thank you for that suggestion, Table one is now updated with body fat %.
+
+*Comment 7: Please specify how the "blinded taste test" (described on page 6, line 120) was carried out.*
+**Authors reply:** Thank you for this valuable input. We have added the following description of the blinded taste test:
+   - "In this blinded taste test, the participants were given two glasses of glucose mix (75ml per) and two glasses of the placebo mix (75ml per), consumed in a randomised order per participant.
+The participants were instructed to finish one bolus, note their guess for its content, and move on to the next glass.
+On average, the participants had a score of 2 points (2.08 ± 1.24) out of 4 possible.
+" Page 5-6, line 151
+
+*Comment 8: Mention more characteristics related to the training of the participants at baseline, such as number of weekly or 14-day RT sessions (which was an inclusion criteria).*
+
+**Authors reply:** Thank you for bringing this up.
+Sadly, we do not have detailed characteristics on individual training sessions per week prior to enrollment in the study, other than that they did indeed have a weekly resistance training session volume between two and eight sessions per 14 days for the last six months.
+The authors realise that this leaves room for variation between participants in terms of training status at baseline, however variations such as these should be controlled for by the within-participants analyses.
+This is now addressed in "Limitations and strengths" Page 16, line 511.
+
+**Additional changes to the manuscript:* Knut Sindre Mølmen has been added to the author list.
+He contributed with *conception of idea, performed experiments and edition of manuscript* and was accidentally omitted from the list of authors in the original submission.
+We have also taken steps to improve the language and flow of the manuscript at multiple points, as highlighted in red and with "track changes".
+Lastly, we added suggestions for where tables and figures may be placed.

### 2024-01-30 manuscript_revised.RMD
finished revision

@@
 
-ORCID: Kristian Lian <https://orcid.org/0000-0002-5169-5960>, Daniel Hammarström <https://orcid.org/0000-0001-8360-2100>, Håvard Hamarsland <https://orcid.org/0000-0002-7343-2774>, Knut Sindre Mølmen<https://orcid.org/0000-0001-8924-6848>, Stian Ellefsen <https://orcid.org/0000-0003-0467-2018>
+ORCID: Kristian Lian <https://orcid.org/0000-0002-5169-5960>, Daniel Hammarström <https://orcid.org/0000-0001-8360-2100>, Håvard Hamarsland <https://orcid.org/0000-0002-7343-2774>, Knut Sindre Mølmen <https://orcid.org/0000-0001-8924-6848>, Stian Ellefsen <https://orcid.org/0000-0003-0467-2018>
 
@@
 High glucose treatment may augment ribosome biogenesis, as evident from *in vitro* studies, and may thus act to potentiate resistance training-induced adaptations *in vivo*.
-This was investigated with total RNA and ribosomal RNA as main outcomes, with relevant transcriptional or translational regulaters (c-Myc/UBF/rpS6) as a secondary outcome.
+This was investigated with total RNA and ribosomal RNA abundances as main outcomes, with relevant transcriptional or translational regulators (c-Myc/UBF/rpS6) as a secondary outcome.
 
@@
 Responses to systematic resistance training (RT) vary widely between individuals, with as much as 10-15% showing impaired skeletal muscle growth in response to standardised training interventions [@alvarez_interindividual_2018; @mann_high_2014; @thalacker-mercer_cluster_2013].
-Genetic predisposition may explain some of this variation [@thalacker-mercer_cluster_2013], but in general the internal physiological milieu seems to be favourably shaped for muscle growth in response to different types of nutrient intake and exercise training [@figueiredo_genetic_2021; @sebata_control_2018; @thalacker-mercer_cluster_2013].
+Genetic predisposition may explain some of this variation [@thalacker-mercer_cluster_2013], but in general, the internal physiological milieu seems to be favourably shaped for muscle growth in response to different types of nutrient intake and exercise training [@figueiredo_genetic_2021; @sebata_control_2018; @thalacker-mercer_cluster_2013].
 Indeed, exercise training and nutrients impact the ability to synthesise ribosomes, which in turn is demonstrated to be connected to the magnitude of RT-induced responses [@hammarstrom_benefits_2020; @kusnadi_regulation_2015; @sebata_control_2018].
@@
 (2020), not all participants experience increased muscle mass and improved muscle strength with increasing training volume.
-Therefore, means other than modification of RT variables seem necessary to optimise individual responses to RT, for instance nutritional adjuvants.
+Therefore, means other than modification of RT variables seem necessary to optimise individual responses to RT, for instance, nutritional adjuvants.
 
@@
 Furthermore, energy availability is a decisive factor in the *de novo* synthesis of ribosomes [@kusnadi_regulation_2015; @moss_housekeeper_2007; @sebata_control_2018] which in turn seems to determine muscle growth by increasing the muscle's translational capacity [@figueiredo_regulation_2019; @hammarstrom_benefits_2020; @sebata_control_2018; @stec_ribosome_2016; @von_walden_ribosome_2019].
-In addition, insulin, which is secreted from the beta cells of pancreas in response to rising blood glucose levels, may itself exert anabolic effects irrespective of muscle contraction, e.g.
-by elevating levels of amino acids, an also plays a role in reducing muscle protein breakdown independent of amino acid availability [@abdulla_role_2016; @hillier_physiological_2000].
-It seems plausible that combined RT and glucose ingestion provides effects on ribosome biogenesis compared to RT alone.
+In addition, insulin, which is secreted from the beta cells of the pancreas in response to rising blood glucose levels, may itself exert anabolic effects irrespective of muscle contraction, e.g.
+by elevating levels of amino acids, and also plays a role in reducing muscle protein breakdown independent of amino acid availability [@abdulla_role_2016; @hillier_physiological_2000].
+It seems plausible that combined RT and glucose ingestion provide additive effects on ribosome biogenesis compared to RT alone.
 
@@
 One due to sickness and inability to resume, while two participants experienced muscular discomfort related to heavy resistance training.
-Lean mass and body fat % (Table 1) was measured using Dual-Energy X-ray Absorptiometry (DXA, Prodigy Advance PA+302047, Lunar, San Francisco, CA, USA) on Day -1, the last day preceding the RT intervention.
-
-Table 1 could be placed here.
+Lean mass and body fat % (Table 1) were measured using Dual-Energy X-ray Absorptiometry (DXA, Prodigy Advance PA+302047, Lunar, San Francisco, CA, USA) on Day -1, the last day preceding the RT intervention.
+
+*Table 1 could be placed around here.*
 
@@
 All participants completed six RT sessions with glucose and six with placebo, allowing a within-subjects analysis of the effects of glucose ingestion before and after RT.
-Data from the first five RT sessions was used to investigate main outcome measures (total RNA, rRNA and protein) and leg muscle strength, whereas data from the sixth RT session was used to explore secondary outcomes (muscular recovery, plasma glucose and plasma c-peptide levels).
-Participants were asked to avoid resistance- or high-intensity training of the legs from Day -7 (Figure 1A) and onwards, until completion of the intervention and post-testing, to ensure the reliability of pre-intervention strength data and minimal interference from external exercise sources.
+Data from the first five RT sessions was used to investigate main outcome measures (total RNA, rRNA and protein) and leg muscle strength, whereas data from the sixth RT session was used to explore secondary outcomes (muscular recovery, plasma glucose and serum c-peptide levels).
+Participants were asked to avoid resistance- or high-intensity training of the legs from Day -7 (Figure 1A) and onwards, until completion of the intervention and post-testing, to ensure the reliability of pre-intervention strength data and minimise interference from external exercise sources.
 
@@
 
-Table 2 can be placed here.
+*Table 2 can be placed here.*
 
@@
 
-Glucose ingestion before and after RT sessions did not lead to a higher mean change of total training session volume, with a mean increase of 17% in both the glucose condition (pre: 5262 ± 1799kg, post: 6319 ± 2256kg, *p* > 0.05) and the placebo condition (pre: 5351 ± 1615kg, post: 6438 ± 2092, *p* > 0.05) from session 1 to session 6, respectively.
-There were no differences in mean macronutrient intake (protein, fat, carbohydrates) or total calorie intake between glucose and placebo on pairwise consecutive days (*p* \> 0.05 for all, raw data and table available at GitHub repository).
+Glucose ingestion before and after RT sessions did not lead to a higher mean change of total training session volume, with a mean increase of 17% in both the glucose condition (pre, 5262 ± 1799kg; post, 6319 ± 2256kg, *p* > 0.05) and the placebo condition (pre, 5351 ± 1615kg; post, 6438 ± 2092, *p* > 0.05) from session 1 to session 6, respectively.
+There were no differences in mean macronutrient intake (protein, fat, carbohydrates) or total calorie intake between glucose and placebo on pairwise consecutive days (*p* \> 0.05 for all, raw data and table available in GitHub repository).
 
@@
 
-Glucose ingestion before and after RT led to increases in plasma glucose levels compared to baseline by 38% immediately before RT (Figure 1B, 0 min: 2.05 ± 0.73 mmol/L), by 31% during RT (Figure 1B, 15 min: 1.75 ± 1.44 mmol/L) and by 32% immediately after RT (Figure 1B, 30 min: 1.62 ± 1.10 mmol/L, all : *p* \< 0.001), with no changes observed in the placebo condition (Figure 1B, 0 min: 0.09 ± 0.3 mmol/L, 15 min: 0.16 ± 0.35 mmol/L, 30 min: 0.18 ± 0.39 mmol/L, *p* \> 0.05).
-Compared to the placebo condition, ingestion of glucose increased plasma glucose levels by 36% immediately before RT (Figure 1B, 0 min), by 27% during RT (Figure 1B, 15 min) and by 28% immediately after RT (Figure 1B, 30 min, all : *p* \< 0.001).
-Two hours after the RT session, glucose ingestion was associated with 12% lower plasma glucose levels compared to baseline, and 8% lower compared to placebo (Figure 1B, 270min: *p* = `r p.270`).
-
-Glucose ingestion before and after RT led to increases in levels of c-peptide compared to baseline, by 95% immediately before (Figure 1C, 0 min: 796 ± 376.0 pmol/L) and 87% after RT (Figure 1C, 30 min: 793 ± 581.0 pmol/L, both *p* \< 0.001), with no changes observed with the placebo condition (Figure 1C, 0 min: 63.7 ± 71.0 pmol/L, 30 min: 53.9 ± 134.0 pmol/L, both *p* \> 0.05).
+Glucose ingestion before and after RT led to increases in plasma glucose levels compared to baseline by 38% immediately before RT (Figure 1B, 0min, 2.05 ± 0.73 mmol/L), by 31% during RT (Figure 1B, 15 min, 1.75 ± 1.44 mmol/L) and by 32% immediately after RT (Figure 1B, 30 min, 1.62 ± 1.10 mmol/L, all *p* < 0.001), with no changes being observed in the placebo condition (Figure 1B, 0 min, 0.09 ± 0.3 mmol/L; 15 min, 0.16 ± 0.35 mmol/L; 30 min, 0.18 ± 0.39 mmol/L, all *p* > 0.05).
+Compared to the placebo condition, ingestion of glucose increased plasma glucose levels by 36% immediately before RT (Figure 1B, 0 min), by 27% during RT (Figure 1B, 15 min) and by 28% immediately after RT (Figure 1B, 30 min, all *p* \< 0.001).
+Two hours after the RT session, glucose ingestion was associated with 12% lower plasma glucose levels compared to baseline, and 8% lower compared to placebo (Figure 1B, 270min, *p* = `r p.270`).
+
+Glucose ingestion before and after RT led to increases in levels of c-peptide compared to baseline, by 95% immediately before (Figure 1C, 0 min, 796 ± 376.0 pmol/L) and 87% after RT (Figure 1C, 30 min, 793 ± 581.0 pmol/L, both *p* < 0.001), with no changes observed with the placebo condition (Figure 1C, 0 min, 63.7 ± 71.0 pmol/L; 30 min, 53.9 ± 134.0 pmol/L; both *p* > 0.05).
 Compared to the placebo condition, ingestion of glucose increased levels of c-peptide by 85% immediately before (Figure 1C, 0 min) and 85% after RT (Figure 1C, 30 min; both *p* \< 0.001).
+
+*Figure 1 can be placed here.*
 
@@
 
-Glucose ingestion before and after RT sessions did not generally improve skeletal muscle recovery throughout the intervention, neither during the intervention in a rested state (Figure 1D, 23hrs after exercise; Post 2RT: *p* = `r pdiff.c2`, Post 4RT: *p* = `r pdiff.c3`), nor acutely after the sixth and final RT-session (30min post 6RT: *p* = `r pdiff.c5`, 2h post 6RT: *p* = `r pdiff.c6`) or in a rested state after the sixth RT session (23h post 6RT: *p* = `r pdiff.c7`).
-Glucose ingestion before and after RT compared to placebo was associated with a 7% less reduction in strength after five RT sessions (Figure 1D, *p* = `r pdiff.c4`).
-
-Both RT with glucose and placebo led to significantly reduced strength after the fifth session compared to baseline, by 11 and 18% respectively (Figure 1D, Post 5RT: *p* = 0.000).
-Comparisons of the acute data gathered from after five sessions until and including 23 hours after the sixth session showed an average increase in strength of 5-9% from RT with glucose and placebo 30 minutes after the sixth RT session (Figure 1D, 30min post 6RT: *p* = `r apc2`) and two hours after the sixth RT session (Figure 1D, 2h post 6RT: *p* = `r apc3`).
-Twenty-three hours after the last (sixth) RT session, strength was unchanged compared to after the fifth RT session (Figure 1D, 23h post 6RT: *p* = `r apc4`).
+In general, glucose ingestion before and after RT sessions did not improve skeletal muscle recovery compared to placebo throughout the intervention, neither the rested state (Figure 1D, 23hrs after exercise, Post 2RT, *p* = `r pdiff.c2`; Post 4RT, *p* = `r pdiff.c3`), nor acutely after the sixth/final RT-session (30min post 6RT, *p* = `r pdiff.c5`; 2h post 6RT, *p* = `r pdiff.c6`) or in the rested state after the sixth RT session (23h post 6RT, *p* = `r pdiff.c7`).
+In contrast to this, glucose ingestion was associated with a 7% less reduction in muscle strength after the fifth session compared to placebo (Figure 1D, *p* = `r pdiff.c4`).
+
+In knee extension torque, both RT with glucose and placebo led to significantly reduced muscle strength after the fifth session compared to baseline, by 11 and 18% respectively (Figure 1D, Post 5RT, *p* = 0.000).
+Comparisons of the acute data gathered from after five sessions until and including 23 hours after the sixth session showed an average increase in muscle strength of 5-9% from RT with glucose and placebo 30 minutes after the sixth RT session (Figure 1D, 30min post 6RT, *p* = `r apc2`) and two hours after the sixth RT session (Figure 1D, 2h post 6RT, *p* = `r apc3`).
+Twenty-three hours after the last (sixth) RT session, muscle strength was unchanged compared to after the fifth RT session (Figure 1D, 23h post 6RT, *p* = `r apc4`).
 Table 3 shows the mean change in absolute peak torque values per condition and angular velocity.
+
+*Table 3 can be placed here.*
 
@@
 
-The five-session-RT intervention led to increases in total RNA and rRNA per unit muscle weight both in the glucose and the placebo condition, on average by \~20-27% (glucose: 263 ± 50 mg/ng-1, placebo: 210 ± 121 mg/ng-1) and \~25-57% (47S; 0.253 ± 1.27 and 0.576 ± 0.677, 18S; 0.336 ± 0.460 and 0.271 ± 0.470, 28S; 0.314 ± 0.504 and 0.311 ± 0.582, 5.8S; 0.388 ± 0.576 and 0.322 ± 0.520, 5S; 0.305 ± 0.608 and 0.292 ± 0.432, arbitrary units for glucose and placebo respectively), respectively (Figure 2).
-However, RT with glucose did not induce increased accumulation of total RNA (Figure 2A, mean difference `r total_rna_stats`) or rRNA (Figure 2B; 47S, `r stat_rrna$trg_47s`; 18S, `r stat_rrna$trg_18s`; 28S, `r stat_rrna$trg_28s`; 5.8S, `r stat_rrna$trg_5.8s`; 5S, `r stat_rrna$trg_5s`) compared to RT with placebo.
+The five-session-RT intervention led to on average ~20-27% increases in total RNA (glucose, 263 ± 50 ng/mg^-1^; placebo, 210 ± 121 ng/mg^-1^) and ~25-57% increases in rRNA per unit muscle weight (47S, 0.253 ± 1.27 and 0.576 ± 0.677; 18S, 0.336 ± 0.460 and 0.271 ± 0.470; 28S, 0.314 ± 0.504 and 0.311 ± 0.582; 5.8S, 0.388 ± 0.576 and 0.322 ± 0.520; 5S, 0.305 ± 0.608 and 0.292 ± 0.432; arbitrary units for glucose and placebo respectively) (Figure 2).
+However, RT with glucose did not induce more pronounced accumulation of total RNA compared to RT with placebo (Figure 2A, mean difference `r total_rna_stats`) or rRNA (Figure 2B, 47S, `r stat_rrna$trg_47s`; 18S, `r stat_rrna$trg_18s`; 28S, `r stat_rrna$trg_28s`; 5.8S, `r stat_rrna$trg_5.8s`; 5S, `r stat_rrna$trg_5s`).
+
+*Figure 2 can be placed here. *
 
@@
 
-The five-session-RT intervention led to increases in all measured proteins both in the glucose and the placebo condition (Figure 3A).
-RT with glucose resulted in estimated levels of c-Myc, UBF and RPS6 being `r stat_west$cmyc`, `r stat_west$ubf` and `r stat_west$rps6`% lower compared to placebo, respectively, without showing statistical significance (*p* = `r min(p_west)`-`r max(p_west)`; Figure 3A).
-Both baseline and trained state data of total RNA showed a linear relationship with UBF, where $\sim$ `r corr_incr`% increase in total RNA corresponded to 1 standard deviation unit increase in UBF (*p* = `r corrp`; Figure 3C).
+The five-session-RT intervention led to increased abundances of all measured proteins, both in the glucose and in the placebo condition (Figure 3A).
+RT with glucose resulted in lowered estimates of c-Myc, UBF and rpS6 levels compared to placebo (being `r stat_west$cmyc`, `r stat_west$ubf` and `r stat_west$rps6`% lower compared to placebo, respectively), without reaching statistical significance (Figure 3A, *p* = `r min(p_west)`-`r max(p_west)`).
+Baseline and trained-state total RNA levels showed a linear relationship with UBF abundances; a $\sim$ `r corr_incr`% increase in total RNA corresponded to 1 SD unit increase in UBF (Figure 3C, *p* = `r corrp`).
+
+*Figure 3 can be placed here.*
 
@@
 
-The main findings of the present study were that heavy resistance training with glucose did not affect markers of ribosome biogenesis, measured as total RNA, rRNA, and rDNA transcription initiation-associated proteins, compared to RT with placebo.
-Similarly, RT with glucose in general did not affect markers of skeletal muscle functionality such as strength and recovery, or total training session volume.
-Towards the end of the intervention, RT with glucose led to less reduction in strength compared to placebo, which may point to an accumulated effect of the glucose condition.
-As by design, the levels of plasma glucose and c-peptide were significantly higher before, during and after the RT sessions with the glucose condition compared to placebo, and there were no differences in the daily macronutrient intake between conditions on consecutive days, thus enabling an effective design to study the biological and functional effects of the glucose condition.
-These findings suggest that the effects of high vs.
-low glucose/glucose starvation conditions on rDNA transcription initiation observed in previous *in vitro* studies [@mariappan_ribosomal_2011; @tanaka_mild_2015] are not translatable to acute effects in human skeletal muscle *in vivo*, with a design like the present study.
-
-The observations of the present study from human skeletal muscle do not support previous indications from non-human skeletal muscle [@mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012].
-Although we observed a linear relationship between baseline and trained state UBF and total RNA levels, a UBF-dependent augmentation of ribosome biogenesis did not seem to be induced by glucose ingestion in the present study, as compared to in mice glomerular epithelial cells [@mariappan_ribosomal_2011] and in human breast cancer cells [@tanaka_mild_2015].
+The main findings of the present study were that heavy resistance training with glucose did not affect markers of ribosome biogenesis compared to RT with placebo, measured as total RNA, rRNA, and protein involved in rDNA transcription initiation.
+Similarly, RT with glucose  did not affect markers of skeletal muscle functionality compared to placebo, such as muscle strength and recovery, and total training session volume.
+Towards the end of the intervention, RT with glucose led to less reduction in muscle strength compared to placebo, which may point to an accumulated effect of the glucose condition.
+As by design, the levels of plasma glucose and serum c-peptide were significantly higher before, during and after the glucose condition RT sessions compared to placebo, and there were no differences in daily macronutrient intake between conditions on consecutive days, suggesting that the study was adequately designed to elucidate the biological and functional effects of the glucose condition.
+These findings suggest that the previously observed significance of glucose exposure for rDNA transcription initiation in *in vitro* studies [@mariappan_ribosomal_2011; @tanaka_mild_2015] is not translatable to acute effects in resistance exercised human skeletal muscle *in vivo*, with a design like the present study.
+
+The observations of the present study from human skeletal muscle cells do not support previous indications of high glucose exposure on rDNA transcription initiation in non-human and/or non-skeletal muscle cells [@mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012].
+A UBF-dependent augmentation of ribosome biogenesis did not seem to be induced by glucose ingestion in the present study, as compared to in mice glomerular epithelial cells [@mariappan_ribosomal_2011] and in human breast cancer cells [@tanaka_mild_2015].
 Neither do our findings indicate that insulin *per se*, at least at physiological levels, potentiates accumulation of total RNA through p70S6K stimulation as observed with hyperinsulinemia in human skeletal muscle [@hillier_physiological_2000].
 Increases in markers of ribosome biogenesis such as 47S pre-rRNA and mature rRNA are expected to occur after a single session of RT [@figueiredo_impact_2016], as well as after a short period of RT [@hammarstrom_benefits_2020; @hammarstrom_ribosome_2022].
-Therefore, in the present study, it was expected that eventual benefits of ingesting glucose compared to placebo with RT would be measurable after 5 training sessions, either due to glucose-induced stimulation of energy-sensitive pathways such as mTORC1, PIH1, extracellular signal-regulated kinase 1/2 (ERK1/2), AMP-dependent protein kinase (AMPK) or Sirtuin 1 (SIRT1) [@kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012].
-Despite previously reported upregulation in PIC assembly due to high-glucose mediated mTORC1, ERK1/2 and PIH1 or low-glucose mediated AMPK and SIRT1 activation [@hoppe_amp-activated_2009; @kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012], the present study showed no signs of such effects of glucose vs.
+Therefore, in the present study, it was expected that any benefits of ingesting glucose compared to placebo with RT would be measurable after five training sessions, either due to glucose-induced stimulation of energy-sensitive pathways such as mTORC1, PIH1, extracellular signal-regulated kinase 1/2 (ERK1/2), AMP-dependent protein kinase (AMPK) or Sirtuin 1 (SIRT1) [@kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012].
+Despite previously reported upregulation in PIC assembly due to high-glucose mediated mTORC1, ERK1/2 and PIH1 or low-glucose mediated AMPK and SIRT1 activation [@hoppe_amp-activated_2009; @kim_nutrient_2013; @mariappan_ribosomal_2011; @tanaka_mild_2015; @zhai_human_2012], the present study displayed no signs of such effects of glucose vs.
 placebo conditions.
+
 Importantly, previous studies investigated high vs.
 low glucose conditions [@mariappan_ribosomal_2011], or high glucose vs.
-glucose starvation [@hoppe_amp-activated_2009; @tanaka_mild_2015], while the present study aimed to compare the high glucose condition to a placebo condition, with a matched daily macronutrient and energy intake.
+glucose starvation [@hoppe_amp-activated_2009; @tanaka_mild_2015], while the present study aimed to compare the high glucose condition to a placebo condition (stevia), with a matched daily macronutrient and energy intake.
 Therefore, the comparison made in the present study was high plasma glucose levels vs.
 normal plasma glucose levels, to investigate the effect of glucose *per se* and not intracellular energy status.
-Thus, while glucose ingestion presumably is important for supplying energy for growth-inducing processes such as ribosome biogenesis [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @sebata_control_2018] there is no apparent effect of ingesting glucose *per se* on markers of ribosome biogenesis, during 12 days of heavy-load RT.
+Thus, while glucose ingestion presumably is important for supplying energy for growth-inducing processes such as ribosome biogenesis [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @sebata_control_2018] there is no apparent effect of ingesting added glucose *per se* on markers of ribosome biogenesis, during 12 days of heavy-load RT.
 Further, previous studies have used cell cultures from yeast [@zhai_human_2012], rodents [@hoppe_amp-activated_2009; @mariappan_ribosomal_2011] or human breast cancer cells [@tanaka_mild_2015] and are, as such, not directly comparable to human skeletal muscle cells.
@@
 
-In the present study, despite not measuring the activity in central pathways mediating anabolic signalling (mTORC1, ERK1/2), analyses of the downstream target UBF and the ribosomal protein S6 (rpS6), as well as the general transcription factor c-Myc, further supported observations from the RNA data as there was no difference between the glucose and placebo conditions.
-UBF has previously been described as a master regulator of rDNA transcription *in vitro* [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @russell_rna-polymerase-i-directed_2005], while rpS6 is proposed as a valid and reliable means to measure ribosome biogenesis [@chaillou_hypoxia_2012; @nakada_correlation_2016].
+In the present study, despite not measuring the activity in central pathways mediating anabolic signalling (mTORC1, ERK1/2), analyses of the downstream target UBF and the ribosomal protein S6 (rpS6), as well as the general transcription factor c-Myc were performed.
+These analyses further supported observations from the RNA data as there was no difference between the glucose and placebo conditions.
+UBF has previously been described as a master regulator of rDNA transcription *in vitro* [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @russell_rna-polymerase-i-directed_2005], while rpS6 previously correlated with 18S and 28S rRNA content and is proposed as a valid and reliable means to measure ribosome biogenesis [@chaillou_hypoxia_2012; @nakada_correlation_2016].
 Lastly, c-Myc has previously been described as a potent regulator of ribosome biogenesis, independent of mTORC1, and a direct regulator of UBF [@mori_c-myc_2021; @poortinga_c-myc_2011; @west_acute_2016].
 Hence, it seems quite reasonable to observe similar changes in these three proteins.
-The linear relationship found exclusively between UBF content and total RNA levels, and neither c-Myc nor rpS6, supports the importance and specificity of UBF content in human skeletal muscle ribosome biogenesis.
-While this is not a novel finding, it was only recently observed in human skeletal muscle following a period of RT [@hammarstrom_ribosome_2022].
-As such, UBF also seems to respond to mechanical loading in human muscle cells in line with responses seen in cell cultures and synergist ablation models [@mariappan_ribosomal_2011; @von_walden_mechanical_2012], in addition to the previously reported response to hormonal, nutritional, and cellular energy signals [@figueiredo_regulation_2019; @kusnadi_regulation_2015; @russell_rna-polymerase-i-directed_2005].
-
-As in the biological data, glucose ingestion had no effect throughout the intervention on muscular strength compared to placebo.
-Generally, skeletal muscle performance measured by a strength index, used as a proxy marker for muscular recovery, decreased from baseline to after the condition in both conditions.
-Glucose ingestion was associated with a less reduction in muscular strength after five RT sessions compared to placebo, which may point towards a beneficial accumulated effect where the heavy-load RT fatigued the participants but glucose ingestion may have improved muscular recovery [@mul_exercise_2015; @sebata_control_2018].
-However, glucose did not improve muscular performance/recovery acutely following one RT session compared to placebo, measured 30 minutes, 2 hours and 23 hours after the sixth and final training session.
-As such, the potential accumulated effect observed after five RT sessions did not extend to after session six.
-Therefore, we can't rule out a possible long-term effect of ingesting glucose when exercising heavy-load resistance training, which would require a longer intervention period than that of the present study.
-Another possible explanation for the decrease in muscular performance during the intervention might be the biphasic recovery pattern, as described by Raastad & Hallén [@raastad_recovery_2000], where the participants experienced a rapid recovery during the initial 11 hours after exercise, followed by a levelling off or drop until 22 hours after exercise.
+The linear relationship found exclusively between UBF content and total RNA levels, and not between total RNA and c-Myc or rpS6, supports a specific role for UBF in regulating ribosomal content in human skeletal muscle.
+This is in line with recent observations in human skeletal muscle following a period of RT [@hammarstrom_ribosome_2022].
+The observed RT-induced increase in levels of UBF, c-Myc and rpS6 is itself in line with responses seen in cell cultures and synergist ablation models [@mariappan_ribosomal_2011; @von_walden_mechanical_2012].
+
+As in the biological data, combined glucose ingestion and RT did not exert measurable effects on muscular strength throughout the intervention compared to placebo.
+In general, the skeletal muscle performance index, which was used as a proxy marker for muscular recovery, decreased similarly from baseline to after the intervention in both conditions.
+There was one exception to this however, as glucose ingestion was associated with a lower reduction in muscular strength after five RT sessions compared to placebo, which may point towards a beneficial accumulated effect where the heavy-load RT gradually fatigued the participants but glucose ingestion counteracted this response [@mul_exercise_2015; @sebata_control_2018].
+Therefore, we cannot rule out potential long-term benefits of ingesting glucose in connection with heavy-load resistance training, which would require a longer intervention period than that of the present study.
+Having noted this, glucose did not improve muscular performance/recovery acutely following one RT session compared to placebo, measured 30 minutes, 2 hours and 23 hours after the sixth/final training session.
+As such, the potential accumulated effect observed after five RT sessions did not extend to acute effects measured after RT session six.
+
+The decrease in muscular performance observed over the course of the intervention might be explained by the biphasic recovery pattern, as described by Raastad & Hallén [@raastad_recovery_2000], where the participants experienced a rapid recovery during the initial 11 hours after exercise, followed by a levelling off or drop until 22 hours after exercise.
 Herein, inflammation and phagocytic activity were proposed to be involved in the performance drop between 11-22 hours [@raastad_recovery_2000].
 Indeed, this pattern seems quite similar to what was observed in the present study, with a rapid recovery at 30 minutes and 2 hours after the sixth RT session and a drop at 23 hours after the sixth session.
-Further, strength testing during the intervention was conducted 23 hours after RT, meaning that the biphasic recovery may have also influenced these tests.
-However, this does not explain the difference between conditions observed after five RT sessions, or the drop in strength from after four RT sessions to after five RT sessions.
-A possible argument could be that exercising without glucose may have caused more stress compared to exercising with glucose, as glucose is the preferred energy source during strenuous exercise [@mul_exercise_2015], thus increasing performance with glucose compared to placebo.
-Notably, training volume showed that the total training session volume was equal on pairwise consecutive days, i.e.
+Further, muscle strength testing during the intervention was conducted 23 hours after RT, meaning that the biphasic recovery may have also influenced these tests.
+However, this neither explains the difference observed after five RT sessions between conditions, nor the drop in muscle strength from after the fourth RT session to after the fifth RT session.
+A possible explanation could be that exercising without glucose ingestion may have caused more lower cellular energy substrate availability compared to exercising with glucose, as glucose is the preferred energy source during strenuous exercise [@mul_exercise_2015], thus increasing performance with glucose compared to placebo.
+Notably, training volume data displayed that the total training session volume was not different on pairwise consecutive days, i.e.
 no difference between days 1-2, days 3-4 and so on.
-Hence, there were no differences in mechanical loading to induce a higher stress between conditions.
-Arguably, an increased energy availability via glucose ingestion during RT may induce less acute stress on the exercised skeletal muscle, therefore less fatigue, and perhaps less performance reduction, compared to placebo during RT [@westerblad_mechanisms_1998; @kent-braun_central_1999].
+Hence, there were no differences in mechanical loading to induce greater fatigue between conditions.
+Arguably, an increased energy availability via glucose ingestion during RT may induce less acute fatigue on the exercised skeletal muscle, and perhaps less performance reduction, compared to placebo during RT [@westerblad_mechanisms_1998; @kent-braun_central_1999].
 Unfortunately, we did not conduct measurements of markers of metabolic stress such as inorganic phosphate, H+, Mg2+ and the ADP/ATP ratio [@westerblad_mechanisms_1998; @kent-braun_central_1999].
@@
 
-The present study was designed specifically to investigate the acute biological and functional effects of ingesting glucose compared to placebo, with unilateral training and testing to reduce biological variation between participants in our analyses [@macinnis_investigating_2017].
-Further, to ensure that the legs were exercised under the same conditions apart from glucose/placebo during exercise, macronutrients, the time of day and test/training personnel were standardized for each participant on pairwise consecutive days[@halperin_threats_2015].
-Every day, the participants showed up in an overnight fasted state and ingested only either protein and glucose or protein and placebo before and during exercise.
-As evident from the blood data, glucose ingestion thus led to higher levels both of plasma glucose and serum c-peptide compared to placebo, together with the aforementioned standardisations enabling analyses of the effect of ingesting glucose on total- and specific RNA levels and proteins, as well as muscular performance, within-participant.
-The present study also had a few limitations; firstly, the sample size was smaller than expected and planned for.
-Initially the minimum limit of 16 participants, according to the a priori power calculation, was met.
-However, three dropped out during the intervention, possibly leaving our statistical analyses slightly underpowered.
-Moreover, the intramuscular glycogen stores were not measured, hence it can't be determined whether the glucose ingestion increased intramuscular glucose.
+The present study was designed specifically to investigate the acute biological and functional effects of ingesting glucose compared to placebo, with unilateral training and testing in a crossover design.
+This design allowed for high-resolution analyses of within-participant comparisons of the two treatments and hence removing biological diversity between individuals as a confounding factor [@macinnis_investigating_2017].
+Further, to ensure that the legs were exercised under the same conditions (apart from glucose/placebo during exercise), macronutrients, the time of day and the test/training personnel were standardised for each participant on pairwise consecutive days [@halperin_threats_2015].
+Every day, the participants showed up in an overnight fasted state and ingested either protein and glucose, or protein and placebo before and after exercise.
+Taken together, the clear difference between the glucose and placebo conditions in plasma glucose and serum c-peptide, along with the aforementioned standardisations enabled high-resolution analyses of the effect of ingestion glucose on total- and specific RNA levels and proteins, as well as muscular performance, within-participant.
+
+The present study also had a few limitations.
+Firstly, the sample size was smaller than expected and planned for.
+Initially the minimium limit of 16 participants, according to the a priori power calculation, was met.
+However, three dropped out during the intervention.
+The crossover design (along with all its standardisations) likely still contributed to the validity and reliability of the analyses, though these dropouts possibly left our statistical analyses slightly underpowered.
+In addition, we did not keep detailed information on each participant's training history prior to enrolment in the study, other than being within our definition (between two and eight RT sessions per 14 days for the last six months).
+Indeed, this leaves room for variation between participants in terms of training status at baseline, which should be controlled for with the within-participants design [@macinnis_investigating_2017].
+Furthermore, the intramuscular glycogen stores were not measured, hence it cannot be determined whether the glucose ingestion increased intramuscular glucose.
 Though it may be a reasonable assumption that the glucose ingestion in this design did lead to increased intramuscular glucose and thus energy levels, it cannot be excluded that the participants' intramuscular glycogen stores were topped up from the previous day, and as such, ingesting more glucose had no further benefit.
-Nevertheless, even if the skeletal muscle of the leg exercising with glucose supplement took up the extra glucose, it did not seem to affect any of our main outcome measures.
+Importantly though, our main hypothesis was centered around extracellular signalling and the effects of elevated plasma glucose/insulin on rRNA synthesis, and not intracellular energy stores.
+Nevertheless, even if the skeletal muscle of the leg exercising with the glucose condition took up the extra glucose, it did not seem to affect any of our main outcome measures.
 
@@
 Glucose ingestion did not affect muscular performance throughout the study, nor did it affect muscular performance measured 30 minutes, 2 hours or 23 hours after the last session.
-Glucose ingestion was associated with less reduction in muscular performance 23 hours after the fifth training session, thus we can't rule out a possible accumulated effect of ingesting glucose compared to placebo.
-The relationship between baseline and trained state data of total RNA and UBF levels do support the key role of UBF in ribosomal biogenesis regulation in human skeletal muscle following resistance training.
-Future investigations should aim for a greater sample size over a longer time period, to investigate the potential accumulated effects of glucose ingestion, as well as analyses of the intramuscular glycogen storage to provide higher-resolution results.
+Glucose ingestion was associated with a lower reduction in muscular performance 23 hours after the fifth training session, and we can therefore not rule out a possible accumulated effect of ingesting glucose compared to placebo on recovery.
+There was a relationship between baseline and trained state data of total RNA and UBF levels.
+This supports a key role for UBF in ribosome biogenesis in human skeletal muscle following resistance training.
+Future investigations should focus on the accumulated long-term effects of simultaneous glucose ingestion and RT on RT-related muscular adaptations, as well as include analysis of intramuscular glycogen storage.
 
@@
 vastus lateralis at baseline (Day 1 leg 1, Day 2 leg 2), and after five RT sessions (Day 11 leg 1, Day 12 leg 2).
-Blood for measuring plasma glucose and -c-peptide was sampled at baseline (Day 1), and during post-testing (Day 11 leg 1, Day 12 leg 2), via finger draws and venous blood samples.
+Blood for measuring plasma glucose and serum c-peptide was sampled at baseline (Day 1), and during post-testing (Day 11 leg 1, Day 12 leg 2), via finger draws and venous blood samples.
 Skeletal muscle strength was measured as peak torque in unilateral isometric and isokinetic (at 60 and 240 degrees per second) knee extension before, multiple times during, and after five and six sessions.
@@
 d) Changes in muscular strength measured as isometric and isokinetic peak torque (60 and 240 d/s) via Humac Norm Dynamometer, conducted at baseline (a: Day -1), after two and four RT sessions (a: Day 4 and 8 leg 1, Day 5 and 9 leg 2), after five RT sessions/before the 6th session (Day 11 leg 1, Day 12 leg 2), as well as 30min, 2hrs and 23hrs after the 6th RT session (a: Day 11/12 leg 1, Day 12/13 leg 2).
-The index was calculated by normalizing peak torque values to the highest peak torque value at each respective speed, and then summarized and used in change score calculations.
+The index was calculated by normalizing peak torque values to the highest peak torque value at each respective angular velocity, and then summarized and used in change score calculations.
 Values are presented as changes in estimated marginal means ± 95% confidence intervals (CI).
@@
 **Table 3.
-Peak Torque** Mean peak torque per condition and angular velocity measured at the different times during the intervention.
-Baseline = before exercise, After 2RT = 23 hours after two training sessions, After 4RT = 23 hours after four training sessions, After RT5 = 23 hours after five training sessions, After 6RT#1 = 30min after the sixth training session, After 6RT#2 = 2 hours after the sixth training session, After 6RT#3 = 23 hours after the sixth training session.
-240º sec-1 = 240 degrees per second movement speed (isokinetic), 60º sec-1 = 60 degrees per second movement speed (isokinetic), 0º sec-1 = 0 degrees per second movement speed (isometric).
+Peak Torque:** Mean peak torque per condition and angular velocity measured at the different times during the intervention.
+Baseline = before exercise, Post 2RT = 23 hours after two training sessions, Post 4RT = 23 hours after four training sessions, Post RT5 = 23 hours after five training sessions, Post 6RT#1 = 30min after the sixth training session, Post 6RT#2 = 2 hours after the sixth training session, Post 6RT#3 = 23 hours after the sixth training session.
+240º sec-1 = 240 degrees per second angular velocity (isokinetic), 60º sec-1 = 60 degrees per second angular velocity (isokinetic), 0º sec-1 = 0 degrees per second angular velocity (isometric).
 Values are reported as mean ± SD.

### 2024-01-30 revision_notes.RMD
finished revision

--- 2024-01-29 resources/older manuscript versions/revision_notes.RMD
@@
 ---
-title: "Untitled"
-format: docx
-editor: visual
+title: ""
+author: ""
 bibliography: resources/ribose-paper.bib
+output:
+  word_document:
+editor_options: 
+  chunk_output_type: console
 ---
@@
 Revision: Antibodies were diluted in blocking buffer to concentrations corresponding to 1:500 (UBF and rpS6, primary), 1:2000 (c-Myc, primary), 1:5000 (c-Myc, secondary) and 1:25 000 (UBF and rpS6, secondary).
-Added on page 9, line 293.
+Added on page 11, line 302 (page 10, line 279 in the clean manuscript).
 
@@
 **Authors reply:** Thank you for bringing up this subject.
-We did assume tat protein ingestiuon cold impact plasma insulin levels.
+We did assume tat protein ingestion cold impact plasma insulin levels.
 Therefore, the participants waited for 2hrs from ingesting the protein supplement before starting the training sessions, thus "washing out" the potential effect of protein on plasma insulin levels before the training session.
 However, as noted by the reviewer, the participants did not wait for 2hrs after the sessions to ingest the second/last protein supplement.
-In the glucose condition sessions, plasa c-peptide levels were significantly higher at onset of training (Figure 1C, 0min) and at the end of training (Figure 1C, 30min) compared to placebo, and returned to baseline values 2hrs after training (Figure 1C, 120min).
-So, based on these data, glucose ingestion did significantly increase plasma c-peptide compared to placebo during exercise, with no measured effect of protein ingestion on plasma c-peptid levels from 2hrs pre until 2hrs post exercise.
-However, we cannot exclude that there were peaks/fluctuations in plasma c-peptide as a response to ingesting protein that we did not measure in this time frame, especially after exercise with a 2 hour-long period without measurements.
+In the glucose condition sessions, c-peptide levels were significantly higher at onset of training (Figure 1C, 0min) and at the end of training (Figure 1C, 30min) compared to placebo, and returned to baseline values 2hrs after training (Figure 1C, 120min).
+So, based on these data, glucose ingestion did significantly increase c-peptide compared to placebo during exercise, with no measured effect of protein ingestion on c-peptid levels from 2hrs pre until 2hrs post exercise.
+However, we cannot exclude that there were peaks/fluctuations in c-peptide as a response to ingesting protein that we did not measure in this time frame, especially after exercise with a 2 hour-long period without measurements.
 One could arguer that if the protein ingestion *per se* stimulated a large insulin secretion, this would trigger increased glucose secretion from the liver to the bloodstream as a response to the insulin, to stabilise the plasma glucose levels.
 Our 30min post-protein ingestion measurement of plasma glucose showed no such changes at this time.
-Nevertheless, without having measured plasma c-peptide/insulin levels more frequently following the training sessions, we have no way of knowing what truly occurred in this time frame.
+Nevertheless, without having measured c-peptide/insulin levels more frequently following the training sessions, we have no way of knowing what truly occurred in this time frame.
 
-We decided to measure plasma c-peptide rather than insulin.
-Indeed, this provides indirect measure as opposed to direclty measuring plasma insulin.
+We decided to measure c-peptide from serum samples rather than insulin.
+Indeed, this provides an indirect measure as opposed to direclty measuring plasma insulin.
 However, c-peptide stays in the blood for a longer time than insulin, making c-peptide easier to accurately measure over several time points [@leighton2017practical].
@@
 Glucose results revision: Glucose ingestion before and after RT led to increases in plasma glucose levels compared to baseline by 38% immediately before RT (Figure 1B, 0 min: 2.05 ± 0.73 mmol/L), by 31% during RT (Figure 1B, 15 min: 1.75 ± 1.44 mmol/L) and by 32% immediately after RT (Figure 1B, 30 min: 1.62 ± 1.10 mmol/L, all : p < 0.001), with no changes being observed in the placebo condition (Figure 1B, 0 min: 0.09 ± 0.3 mmol/L, 15 min: 0.16 ± 0.35 mmol/L, 30 min: 0.18 ± 0.39 mmol/L, p > 0.05).
-Added on page 11, line 335
+Added on page 12, line 344 (page 11, line 319 in the clean manuscript)
 
 C-peptide results revision: Glucose ingestion before and after RT led to increases in levels of c-peptide compared to baseline, by 95% immediately before (Figure 1C, 0 min: 796 ± 376.0 pmol/L) and 87% after RT (Figure 1C, 30 min: 793 ± 581.0 pmol/L, both p < 0.001), with no changes observed with the placebo condition (Figure 1C, 0 min: 63.7 ± 71.0 pmol/L, 30 min: 53.9 ± 134.0 pmol/L, both p > 0.05).
-Addded on page 11 line 347
+Addded on page 12 line 356 (page 11, line 328 in the clean manuscript)
 
-Absolute values for knee extension torque are added to Table 3, referenced by "Table 3 shows the mean change in absolute peak torque per condition and angular velocity." on page 12 line 367.
+Absolute values for knee extension torque are added to Table 3, referenced by "Table 3 shows the mean change in absolute peak torque per condition and angular velocity." on page 13 line 376 (page 12, line 347 in the clean manuscript).
 Baseline muscle strength values are now removed from table 1, since the detailed table of knee extension torque measurements are added (Table 3).
 
-Revision of total RNA and rRNA: (glucose: 263 ± 50 ng/mg-1 , placebo: 210 ± 121 ng/mg-1) added on page 12, line 371, and; (47S; 0.253 ± 1.27 and 0.576 ± 0.677, 18S; 0.336 ± 0.460 and 0.271 ± 0.470, 28S; 0.314 ± 0.504 and 0.311 ± 0.582, 5.8S; 0.388 ± 0.576 and 0.322 ± 0.520, 5S; 0.305 ± 0.608 and 0.292 ± 0.432, arbitrary units for glucose and placebo respectively) added on page 12 line 374.
+Revision of total RNA and rRNA: (glucose: 263 ± 50 ng/mg-1 , placebo: 210 ± 121 ng/mg-1) and; (47S; 0.253 ± 1.27 and 0.576 ± 0.677, 18S; 0.336 ± 0.460 and 0.271 ± 0.470, 28S; 0.314 ± 0.504 and 0.311 ± 0.582, 5.8S; 0.388 ± 0.576 and 0.322 ± 0.520, 5S; 0.305 ± 0.608 and 0.292 ± 0.432, arbitrary units for glucose and placebo respectively) added on page 13 line 383 (page 12, line 352 in the clean manuscript).
 
@@
 "Apart from this, participants ingested a self-chosen diet during period II, registered in MyFitnessPal or similar applications" has been changed to "Apart from this, participants ingested a self-chosen diet during period II.
-Further, participants were asked not to use any other supplements such as additional protein and/or creatine, and to register all food/drink consumption in MyFitnessPal or similar applications." Page 6, line 175.
+Further, participants were asked not to use any other supplements such as additional protein and/or creatine, and to register all food/drink consumption in MyFitnessPal or similar applications." Page 7, line 183 (page 7, line 163 in the clean manuscript).
 
@@
 The following has been added about rpS6 in the introduction: "Furthermore, Nakada et al.
-(Nakada, Ogasawara, Kawada, Maekawa and Ishii, 2016) found a correlation between rRNA content and rpS6 content in their synergist ablation model on rats." Page 4, line 114.
+(Nakada, Ogasawara, Kawada, Maekawa and Ishii, 2016) found a correlation between rRNA content and rpS6 content in their synergist ablation model on rats." Page 5, line 123 (page 5, line 106 in the clean manuscript).
 
@@
 
-**Authors reply:** Thank you for that suggestion, Table one is now updated with body fat %.
+**Authors reply:** Thank you for that suggestion, Table 1 is now updated with body fat %.
 
@@
 On average, the participants had a score of 2 points (2.08 ± 1.24) out of 4 possible.
-" Page 5-6, line 151
+" Page 5-6, line 157 (page 6, line 138 in the clean manuscript)
 
@@
 **Authors reply:** Thank you for bringing this up.
-Sadly, we do not have detailed characteristics on individual training sessions per week prior to enrollment in the study, other than that they did indeed have a weekly resistance training session volume between two and eight sessions per 14 days for the last six months.
+Sadly, we do not have detailed characteristics on individual training sessions per week prior to enrollment in the study, other than that they did indeed confirm to have a weekly resistance training session volume between two and eight sessions per 14 days for the last six months.
 The authors realise that this leaves room for variation between participants in terms of training status at baseline, however variations such as these should be controlled for by the within-participants analyses.
-This is now addressed in "Limitations and strengths" Page 16, line 511.
+This is now addressed in "Limitations and strengths" Page 17, line 519 (page 16, line 471 in the clean manuscript).
 
-**Additional changes to the manuscript:* Knut Sindre Mølmen has been added to the author list.
+**Additional changes to the manuscript:** Knut Sindre Mølmen has been added to the author list.
 He contributed with *conception of idea, performed experiments and edition of manuscript* and was accidentally omitted from the list of authors in the original submission.
+Bachelor students Max Ullrich and Chris Sylstad were also accidentally omitted from the acknowledgements in the original submission, but have now been added.
 We have also taken steps to improve the language and flow of the manuscript at multiple points, as highlighted in red and with "track changes".

### 2024-01-30 figures/figure1.R

@@ -64,7 +64,7 @@ cpep.fig <- cpep.change %>%
   scale_x_continuous(limits = c(0, 300), breaks = c(0, 90, 120, 150, 270),
                      expand = expansion(0), labels = c("change.1" = "-120 min", "change.90" = "-30 min", "change.120" = "0 min", 
                                                        "change.150" = "30 min", "change.270" = "120 min")) +
-  labs(x = "", y = "Plasma c-peptide levels <br>(pmol &times; L<sup>-1</sup> change)", fill = "Supplement") +
+  labs(x = "", y = "Serum c-peptide levels <br>(pmol &times; L<sup>-1</sup> change)", fill = "Supplement") +
   theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6),
         axis.title.y = element_markdown(size = 7),
         legend.title = element_text(size = 7),
